### 2019-12-30 README.Rmd
Add 2018 phenotypic data, metadata, and Taxa

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# pvdiv-phenology-gxe

<!-- badges: start -->
<!-- badges: end -->

The goal of pvdiv-phenology-gxe is to analyze the genetics of GxE in the phenology data collected for the switchgrass diversity panel in 2018 and 2019, including tiller count in 2018, survival to 2019, and greenup to tiller count & biomass in 2019. This repository will be private until the associated dataset is published.

As of December 2019, this analysis is on data Jason Bonette compiled 2019-12-25. I create downstream data objects from data collected on this date. At the beginning of every month, and whenever I hear from him, I will check his dataset for any changes.

R Files:

In data-raw/:

    *metadata.R creates the metadata rda file used for subsequent analyses.
    *phenotypes_2018.R creates the phenotypic data rda file used for analyses of 2018 data
    *phenotypes_2019.Rmd creates the phenotypic data rda file used for analyses of 2019 data. These plants differ from those in phenotypes_2018 - many plants that died were replaced by Blackwell cultivar clones, and any phenotypes measured on these clones were not included in downstream analyses.
    *Taxa.R creates the Taxa rda file used to tie sequence data to the 785 sequenced PLANT_IDs.

### 2019-12-30 data-raw/phenotypes_2019.Rmd
Add 2018 phenotypic data, metadata, and Taxa

---
title: "Phenotypes_Planting_2019"
author: "Alice MacQueen"
date: "12/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(XLConnect)
library(rlang)
library(switchgrassGWAS)
```

## Load all phenotypic data from 2019

```{r}
wb_pheno <- loadWorkbook(file.path("C:", "Users", "alice", "Dropbox",
                                   paste0("GWAS DOE Switchgrass, ",
                                          "Consolidated Data"),
                                   paste0("DOE_GWAS_2019_Consolidated ",
                                          "Pheno Data_12-25-19.xlsx")))
lst_pheno <- readWorksheet(wb_pheno, sheet = getSheets(wb_pheno))
phenos_planting1 <- lst_pheno$`GWAS 2019 Greenup Data`
phenos_planting2 <- lst_pheno$`GWAS 2019 Core Phenotype Data`
phenos_key <- lst_pheno$`Column Key and Notes`

wb_KING2019 <- loadWorkbook(file.path("C:", "Users", "alice", "Dropbox",
                                      paste0("GWAS DOE Switchgrass, ",
                                             "Consolidated Data"),
                                      "Individual Site Files",
                                      "KING_GWAS_2019_Chlorosis Data.xlsx"))
lst_KING2019 <- readWorksheet(wb_KING2019, sheet = getSheets(wb_KING2019))
KING2019 <- as_tibble(lst_KING2019$`KING GWAS 2019 Chlorosis`)
key_KING19 <- as_tibble(lst_KING2019$`COLUMN KEY`)

all_plants_phenos <- phenos_planting1 %>%
  left_join(PKLE_2018, by = c("SITE", "PLOT_GL", "PLOT_LC", "ACC", "INDV", "PLANT_ID")) %>%
  left_join(KBSM2018, by = c("PLOT_GL")) %>%
  left_join(TMPL2018, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
  mutate(DAM = DAM/10) %>%
  left_join(KING2018, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
  left_join(CLMB2018, by = "PLOT_GL") %>%
  left_join(KING2019, by = "PLOT_GL") %>%
  left_join(LOST2018, by = c("SITE", "PLOT_GL", "PLOT_LC", "PLANT_ID",
                             "PLANT_.ID_GL")) %>%
  mutate(PLANT_ID = ifelse(ACC == "AP13",
                           "AP13",
                           PLANT_ID),
         DEAD_2018 = ifelse(is.na(DEAD_2018), 0, DEAD_2018))
```

# Look at 2019 data

Remove 2019 phenotype scores for plants that didn't survive the winter of 2018.
```{r}
all_plants_phenos %>%
  left_join(phenos_planting2, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
  group_by(SRV, DEAD_2018) %>% tally()

all_plants_phenos %>%
  left_join(phenos_planting2, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
  filter(SRV == "W" & DEAD_2018 == 1)

all_plants_phenos %>%
  left_join(phenos_planting2, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
  filter(SRV == "Y" & DEAD_2018 == 1)

```
12 called as "N" but I called previously as survived (DEAD_2018 = 0) - I'm willing to call these dead. Little data was collected.

13 called as "W" but I called previously as dead (and were noted (in 'NOTES') as "Dead" or "dead".) 8 have phenotype data collected in 2019.
CLMB	C6920	6920	J216	A	J216.A	J216.A.CL	
FRMI	F1002	1002	J462	A	J462.A	J462.A.FR	
FRMI	F1714	1714	J458	B	J458.B	J458.B.FR	
FRMI	F2716	2716	J498	C	J498.C	J498.C.FR	
FRMI	F2724	2724	J235	A	J235.A	J235.A.FR	
KING	K1909	1909	J465	C	J465.C	J465.C.KG	
KING	K2008	2008	J597	A	J597.A	J597.A.KG	
PKLE	P5601	5601	J036	A	J036.A	J036.A.PK	
TMPL	T1303	1303	J597	A	J597.A	J597.A.TP	
TMPL	T1609	1609	J538	A	J538.A	J538.A.TP
TMPL	T2308	2308	J504	A	J504.A	J504.A.TP	
TMPL	T2721	2721	J477	B	J477.B	J477.B.TP	
TMPL	T2803	2803	J592	C	J592.C	J592.C.TP

6 with issues on whether or not the plant died:
For these, SRV was called as "Y", but were noted (in 'NOTES') as "Dead". All 6 have some phenotypic data collected in 2019.

FRMI	F2519	2519	J482	A	J482.A	J482.A.FR	
FRMI	F2523	2523	J577	B	J577.B	J577.B.FR	
FRMI	F2720	2720	J465	A	J465.A	J465.A.FR	
FRMI	F3302	3302	J340	A	J340.A	J340.A.FR	
TMPL	T2106	2106	J019	A	J019.A	J019.A.TP	
OVTN	V1319	1319	J580	A	J580.A	J580.A.VN

Ask Jason about 19 of these. Drop all 31 for now.

```{r}
phenotypes_2019 <- all_plants_phenos %>%
  left_join(phenos_planting2, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
  filter(!(SRV == "W" & DEAD_2018 == 1) & !(SRV == "Y" & DEAD_2018 == 1) & !(SRV == "N")) %>%
  dplyr::select(SITE:PLANT_ID, GR1:GR100, EM1:NOTE_HARVEST)
```

# 2019 Phenotype Cleanup

Ensure all of the columns have the correct data type - mostly, numeric.

```{r}
phenotypes_2019 %>%
  mutate(GR1 = as.numeric(GR1),
         GR50 = as.numeric(GR50),
         GR100 = as.numeric(GR100),
         EM1 = as.numeric(EM1),
         EM50 = as.numeric(EM50),
         FL1 = as.numeric(FL1),
         FL50 = as.numeric(FL50),
         TC_EOS = as.numeric(TC_EOS),
         HT_PAN_EOS = as.numeric(HT_PAN_EOS),
         LD_EOS = as.numeric(LD_EOS),
         REGR = as.numeric(REGR),
         MOISTURE = as.numeric(MOISTURE),
         BIOMASS = as.numeric(BIOMASS))

```

## Look for outlier values by site

and maybe by genetic subgroup, or ACC, or PLANT_ID

### 2019-12-30 data-raw/Taxa.R

@@ -1,0 +1,13 @@
+## code to prepare `Taxa` dataset goes here
+
+library(readr)
+library(magrittr)
+library(dplyr)
+
+SNP_ID <- read_delim(file = file.path("data-raw",
+                                      "Pvirgatum_4x_784g_imp_phased_maf0.019.012.indv"),
+                     delim = "\n", col_names = FALSE)
+Taxa <- SNP_ID %>%
+  rename(PLANT_ID = X1)
+
+save(Taxa, file = "data/Taxa.rda")

### 2019-12-30 data-raw/metadata.R

@@ -1,0 +1,43 @@
+## code to prepare `metadata` dataset goes here
+library(tidyverse)
+library(XLConnect)
+wb_metadata <- loadWorkbook(file.path("data-raw",
+                                      "PVDIV_Metadata_2019-09-03.xlsx"))
+lst_metadata <- readWorksheet(wb_metadata, sheet = getSheets(wb_metadata))
+
+pvdiv <- as_tibble(lst_metadata$`PVDIV MASTER METADATA`)
+
+kof5 <- read_csv("data-raw/k3.nnet.gen.strict.ecotypeClassification.csv")
+
+eco_geno_subgroups <- kof5 %>%
+  dplyr::select(PLANT_ID:pc1_53.7perc, main.subpop:NEC) %>%
+  dplyr::select(-train.id) %>%
+  dplyr::rename(UP_ecotype = Upland, TX_ecotype = TX.lowland,
+                CO_ecotype = Coast.lowland, Ecotype = eco.class,
+                Ecotype_PC1 = pc1_53.7perc, JL_genetic_subpop = main.subpop,
+                JL_detail_genetic = detail.subpop, TX_genetic = TX,
+                GC_genetic = GC, MW_genetic = MW, SEC_genetic = SEC,
+                NEC_genetic = NEC)
+
+pvdiv <- pvdiv %>%
+  select(PLANT_ID:LIB_NOTES, PLOIDY, COLLECTION_TYPE:COLLECTION_METHOD, COLLECTOR, STATE:COLL_DATE) %>%
+  mutate(LATITUDE = as.numeric(LATITUDE),
+         LONGITUDE = as.numeric(LONGITUDE),
+         ELEVATION = as.numeric(ELEVATION),
+         LIBRARY = ifelse(LIBRARY == "NA", NA, LIBRARY),
+         LIB_CLIMATE = ifelse(LIB_CLIMATE == "NA", NA, LIB_CLIMATE),
+         LIB_PHENO = ifelse(LIB_PHENO == "NA", NA, LIB_PHENO),
+         LIB_ACC = ifelse(LIB_ACC == "NA", NA, LIB_ACC),
+         COLL_DATE = ifelse(COLL_DATE == "NA", NA, COLL_DATE),
+         HABITAT = ifelse(HABITAT == "NA", NA, HABITAT),
+         LOCALITY = ifelse(LOCALITY == "NA", NA, LOCALITY),
[…]

### 2019-12-30 data-raw/phenotypes_2018.R

@@ -1,0 +1,233 @@
+## code to prepare `phenotypes_2018` dataset goes here
+library(tidyverse)
+library(XLConnect)
+library(rlang)
+library(switchgrassGWAS)
+
+## Load all the phenotypic data from 2018 ## --------------
+## -------------------------------------------
+# Working Phenotypes, most phenotypes should be in here
+wb_pheno <- loadWorkbook(file.path("C:", "Users", "alice", "Dropbox",
+                                   paste0("GWAS DOE Switchgrass, ",
+                                          "Consolidated Data"),
+                                   paste0("DOE_GWAS_2019_Consolidated ",
+                                          "Pheno Data_12-25-19.xlsx")))
+lst_pheno <- readWorksheet(wb_pheno, sheet = getSheets(wb_pheno))
+phenos_planting1 <- lst_pheno$`GWAS 2019 Greenup Data`
+phenos_planting2 <- lst_pheno$`GWAS 2019 Core Phenotype Data`
+phenos_key <- lst_pheno$`Column Key and Notes`
+
+# Individual site phenotypes
+wb_CLMB2018 <- loadWorkbook(file.path("C:", "Users", "alice", "Dropbox",
+                                      paste0("GWAS DOE Switchgrass, ",
+                                             "Consolidated Data"),
+                                      "Individual Site Files",
+                                      "CLMB_GWAS_2018_Rust Data_Final.xlsx"))
+CLMB2018 <- readWorksheet(wb_CLMB2018, sheet = getSheets(wb_CLMB2018))
+
+KBSM2018 <- read_csv(file.path("C:", "Users", "alice", "Dropbox",
+                               paste0("GWAS DOE Switchgrass, ",
+                                      "Consolidated Data"),
+                               "Individual Site Files",
+                               "KBSM_GWAS_2018_Rust Data_Final.csv"),
+                     col_names = TRUE)
+KING2018 <- read_csv(file = file.path("C:", "Users", "alice", "OneDrive",
+                                      "Juenger Projects", "SwitchgrassGWAS",
[…]

### 2019-12-31 data-raw/phenotypes_2019.Rmd
Plot Phenology data by site and subpopulation

@@
 library(switchgrassGWAS)
+library(cowplot)
+library(viridis)
+library(bigsnpr)
+library(scales)
+
+source("C:/Users/alice/Documents/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+theme_set(theme_poster)
+
+load("../data/sites.rda")
+load("../data/metadata.rda")
+load("../data/Taxa.rda")
+
+usa <- map_data("state")
 ```
@@
 
-all_plants_phenos <- phenos_planting1 %>%
-  left_join(PKLE_2018, by = c("SITE", "PLOT_GL", "PLOT_LC", "ACC", "INDV", "PLANT_ID")) %>%
-  left_join(KBSM2018, by = c("PLOT_GL")) %>%
-  left_join(TMPL2018, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
-  mutate(DAM = DAM/10) %>%
-  left_join(KING2018, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
-  left_join(CLMB2018, by = "PLOT_GL") %>%
-  left_join(KING2019, by = "PLOT_GL") %>%
+wb_DEAD2018 <- loadWorkbook(file.path("C:", "Users", "alice", "Dropbox",
+                                      paste0("GWAS DOE Switchgrass, ",
+                                             "Consolidated Data"),
+                                      "Forms and Miscellaneous",
+                                      paste0("GWAS_2019_Mortality, Weak and ",
+                                             "Not P. Virgatum Plants_ARCHIVE.xlsx")))
+lst_DEAD2018 <- readWorksheet(wb_DEAD2018, sheet = getSheets(wb_DEAD2018))
+DEAD2018 <- as_tibble(lst_DEAD2018$`GWAS 2019 Dead Only`)
+LOST2018 <- as_tibble(lst_DEAD2018$`GWAS 2019 Dead, Weak, Not PV`)
+
+LOST2018 <- LOST2018 %>%
+  mutate(DEAD_2018 = case_when(
+    NOTES %in% c("dead", "Dead", "Dead?", "DEAD") ~ 1,
+    NOTES %in% c("1 tiller", "2 tillers", "3 tillers", "4 tillers",
+                 "5 tillers", "Very Weak", "small crown", "Partial crown death",
+                 "small crown/2 tillers", "small crown/3 tillers",
+                 "small crown/4 tillers", "small crown/5 tillers",
+                 "small crown/7 tillers", "small crown/8 tillers") ~ 0,
+    NOTES %in% c("Not Virgatum", "switchgrass?", "Not Virgatum, Replant") ~ 2,
+    TRUE ~ 0
+  ))
+
+phenos_2019 <- phenos_planting1 %>%
   left_join(LOST2018, by = c("SITE", "PLOT_GL", "PLOT_LC", "PLANT_ID",
                              "PLANT_.ID_GL")) %>%
+  left_join(phenos_planting2, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
+  left_join(KING2019, by = "PLOT_GL") %>%
   mutate(PLANT_ID = ifelse(ACC == "AP13",
                            "AP13",
-                           PLANT_ID),
-         DEAD_2018 = ifelse(is.na(DEAD_2018), 0, DEAD_2018))
+                           PLANT_ID)) %>%
+  dplyr::select(-TC_EOS_2018, -`PLANT_.ID_GL`, -INDV, -(LD_EOS:BIOMASS))
 ```
@@
 ```{r}
-all_plants_phenos %>%
-  left_join(phenos_planting2, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
+phenos_2019 %>%
   group_by(SRV, DEAD_2018) %>% tally()
 
-all_plants_phenos %>%
-  left_join(phenos_planting2, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
+phenos_2019 %>%
   filter(SRV == "W" & DEAD_2018 == 1)
 
-all_plants_phenos %>%
-  left_join(phenos_planting2, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
+phenos_2019 %>%
   filter(SRV == "Y" & DEAD_2018 == 1)
@@
 ```
+Anything where DEAD_2018 = 2 is called as "Not virgatum" in the past, so should be dropped. Does Jason agree?
+
 12 called as "N" but I called previously as survived (DEAD_2018 = 0) - I'm willing to call these dead. Little data was collected.
@@
 ```{r}
-phenotypes_2019 <- all_plants_phenos %>%
-  left_join(phenos_planting2, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
-  filter(!(SRV == "W" & DEAD_2018 == 1) & !(SRV == "Y" & DEAD_2018 == 1) & !(SRV == "N")) %>%
-  dplyr::select(SITE:PLANT_ID, GR1:GR100, EM1:NOTE_HARVEST)
+phenotypes_2019 <- phenos_2019 %>%
+  filter(!(DEAD_2018 %in% c(1,2)) | is.na(DEAD_2018)) %>% # 1 = dead in 2018, and 2 = not a virgatum plant. Drop 1's and 2's. Keep 0's and NA's.
+  filter(SRV != "N") %>% # Drop individuals that did not survive
+  dplyr::select(-VERIFY, -PLOT_GL_CHK)
 ```
@@
 ```{r}
-phenotypes_2019 %>%
-  mutate(GR1 = as.numeric(GR1),
+phenotypes_2019 <- phenotypes_2019 %>%
+  mutate(FRZ_DMG = as.numeric(FRZ_DMG),
+         GR1 = as.numeric(GR1),
          GR50 = as.numeric(GR50),
@@
          HT_PAN_EOS = as.numeric(HT_PAN_EOS),
-         LD_EOS = as.numeric(LD_EOS),
-         REGR = as.numeric(REGR),
-         MOISTURE = as.numeric(MOISTURE),
-         BIOMASS = as.numeric(BIOMASS))
+         DEAD_2018 = ifelse(is.na(DEAD_2018), 0, DEAD_2018),
+         CHLR_101 = as.numeric(CHLR_101),
+         CHLR_121 = as.numeric(CHLR_121),
+         CHLR_136 = as.numeric(CHLR_136),
+         CHLR_JASON_138 = as.numeric(CHLR_JASON_138),
+         CHLR_150. = as.numeric(CHLR_150.),
+         SPAD_158 = as.numeric(SPAD_158),
+         SPADCOR_158 = as.numeric(SPADCOR_158)
+         )
+```
 
+# Put the phenotypes in wider format
+
+so each column has one phenotype at one site.
+
+```{r}
+# Get the mean phenotype value for each PLANT_ID at each SITE as the GWAS value
+phemean_2019 <- phenotypes_2019 %>%
+  group_by(PLANT_ID, SITE) %>%
+  summarise(FRZ_DMG = mean(FRZ_DMG, na.rm = TRUE),
+            GR1 = mean(GR1, na.rm = TRUE),
+            GR50 = mean(GR50, na.rm = TRUE),
+            GR100 = mean(GR100, na.rm = TRUE),
+            EM1 = mean(EM1, na.rm = TRUE),
+            EM50 = mean(EM50, na.rm = TRUE),
+            FL1 = mean(FL1, na.rm = TRUE),
+            FL50 = mean(FL50, na.rm = TRUE),
+            TC_EOS = mean(TC_EOS, na.rm = TRUE),
+            HT_PAN_EOS = mean(HT_PAN_EOS, na.rm = TRUE),
+            CHLR_101 = mean(CHLR_101, na.rm = TRUE),
+            CHLR_121 = mean(CHLR_121, na.rm = TRUE),
+            CHLR_136 = mean(CHLR_136, na.rm = TRUE),
+            CHLR_JASON_138 = mean(CHLR_JASON_138, na.rm = TRUE),
+            CHLR_150. = mean(CHLR_150., na.rm = TRUE),
+            SPAD_158 = mean(SPAD_158, na.rm = TRUE),
+            SPADCOR_158 = mean(SPADCOR_158, na.rm = TRUE)
+  ) %>%
+  mutate_all( ~ case_when(!is.nan(.x) ~ .x)) %>% # if the column is not NA, keep the value; else replace with NA as nothing is provided as another choice in case_when.
+  filter(!is.na(SITE) & !is.na(PLANT_ID))
+```
+
+```{r}
+load("../data/Taxa.rda")
+
+phenotypes_wide_2019 <- phemean_2019 %>%
+  gather(FRZ_DMG:SPADCOR_158, key = "Phenotype", value = value) %>%
+  filter(!is.na(value)) %>%
+  # saveRDS(file = "Cleaned pvdiv Phenotypes from 2018 for 784g 2019-05-16.rds")
+  unite(col = "SITE_PHE", SITE, Phenotype) %>%
+  spread(key = SITE_PHE, value = value) %>%
+  right_join(Taxa)
 ```
@@
 
+Use long format data to look for outlier values.
+This data has associated plot positions (PLOT_GL and PLOT_LC), as well as all associated notes (on greenup, flowering, harvest, chlr and spad).
+The notes might be relevant for outlier analysis.
+
+```{r}
+phenotypes_long_2019 <- phenotypes_2019 %>%
+  dplyr::select(-SRV, -DEAD_2018) %>%
+  dplyr::select(SITE:PLANT_ID, starts_with("NOTE"), everything()) %>%
+  pivot_longer(cols = FRZ_DMG:SPADCOR_158, names_to = "PHE", values_to = "Value") %>%
+  filter(!is.na(Value)) %>%
+  dplyr::select(SITE:PLANT_ID, PHE, Value, everything())
+
+sites %>%
+  arrange(Latitude)
+
+sites$Factor <- factor(sites$Factor, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
+sites$Factor <- factor(sites$Factor, levels = c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING"))
+levels(sites$Factor)
+
+phenotypes_long_2019$SITE <- factor(phenotypes_long_2019$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
+phenotypes_long_2019$SITE <- factor(phenotypes_long_2019$SITE, levels = c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING"))
+phenotypes_long_2019$PHE <- factor(phenotypes_long_2019$PHE, levels = c("GR1", "GR50", "GR100", "EM1", "EM50", "FL1", "FL50", "TC_EOS", "HT_PAN_EOS", "FRZ_DMG", "CHLR_101", "CHLR_121", "CHLR_136", "CHLR_JASON_138", "CHLR_150.", "SPAD_158", "SPADCOR_158"))
+
+```
+
+```{r}
+
+phenotypes_long_2019 %>%
+  filter(!(grepl("CHLR", PHE)) & !(grepl("SPAD", PHE))) %>%
+  ggplot(aes(x = Value)) + 
+  geom_histogram(aes(fill = SITE)) + 
+  facet_wrap(~PHE, scales = "free")
+
+phenology_long_2019 <- phenotypes_long_2019 %>%
+  filter(!(grepl("CHLR", PHE)) & !(grepl("SPAD", PHE)) & !grepl("FRZ", PHE) & !is.na(PHE) & !grepl("TC_EOS", PHE) & !grepl("HT_PAN_EOS", PHE))
+
+library(lubridate)
+phenology_long_2019 %>%
+  mutate(Days = as_date(Value)) %>%
+  ggplot(aes(x = Days)) + 
+  geom_histogram(aes(fill = SITE)) + 
+  facet_grid(SITE~PHE, scales = "free") +
+  scale_fill_viridis(discrete = TRUE, end = 0.95, direction = -1) +
+  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) + xlab("Date") + ylab("Number of Individuals") +
+  geom_vline(xintercept = c(120), linetype = 2) +
+  geom_vline(xintercept = c(181), linetype = 3, color = "grey20") +
+  geom_vline(xintercept = c(243), linetype = 4, color = "grey80")
+
+save_plot("Phenology_Data_Only_by_Site_2019-12-30.png", plot = last_plot(), base_height = 4*2.5, base_width = 5*2.5)
+
+phenology_long_2019 %>%
+  mutate(Days = as_date(Value)) %>%
+  ggplot(aes(x = Days)) + 
+  geom_histogram(aes(fill = SITE)) + 
+  facet_grid(PHE~SITE, scales = "free") +
+  scale_fill_viridis(discrete = TRUE, end = 0.95, direction = -1) +
+  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) + xlab("Date") + ylab("Number of Individuals") +
+  geom_vline(xintercept = c(120), linetype = 2) +
+  geom_vline(xintercept = c(181), linetype = 3, color = "grey20") +
+  geom_vline(xintercept = c(243), linetype = 4, color = "grey80")
+
+save_plot("Phenology_Data_Only_by_Site_CoordFlip_2019-12-30.png", plot = last_plot(), base_height = 4*2, base_width = 8*2)
+
+as_date(120) # May 1st
+as_date(181) # July 1st
+as_date(243) # September 1st
+
+212-181
+181-120
+```
+
+```{r}
+phenotypes_long_2019 %>%
+  filter(grepl("TC_EOS", PHE) | grepl("HT_PAN_EOS", PHE) | grepl("FRZ", PHE)) %>%
+  ggplot(aes(x = Value)) +
+  geom_histogram(aes(fill = SITE)) + 
+  facet_grid(SITE~PHE, scales = "free") +
+  scale_fill_viridis(discrete = TRUE, end = 0.95, direction = -1) +
+  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) + xlab("Date") + ylab("Number of Individuals")
+
+save_plot("2019_Phenotypic_Data_by_Site_2019-12-30.png", plot = last_plot(), base_height = 10, base_width = 6)
+```
+
 and maybe by genetic subgroup, or ACC, or PLANT_ID
+
+```{r}
+
+load("../data/metadata.rda")
+
+metadata <- metadata %>%
+  mutate(Genetic_subpop = case_when(
+           is.na(JL_genetic_subpop) ~ "Unknown",
+           JL_genetic_subpop == "other" ~ "Admixed",
+           TRUE ~ JL_genetic_subpop
+           ))
+metadata$Genetic_subpop <- factor(metadata$Genetic_subpop, levels = c("TX", "GC,TX", "GC", "SEC", "NEC,SEC", "NEC", "MW", "Admixed", "Unknown"))
+metadata$Ecotype <- factor(metadata$Ecotype, levels = c("TX", "Coast", "Up"))
+
+metadata <- add_row(metadata, PLANT_ID = "AP13", LIB_CLIMATE = "Y", LIB_PHENO = "Y", LIB_ACC = "Y", Ecotype = "TX", Genetic_subpop = "TX")
+```
+
+```{r}
+phenology_long_2019 %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  filter(Genetic_subpop != "Unknown") %>%
+  mutate(Days = as_date(Value)) %>%
+  ggplot(aes(x = Days)) + 
+  geom_histogram(aes(fill = Genetic_subpop)) + 
+  facet_grid(SITE~PHE, scales = "free") +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = -1) +
+  theme(legend.position = "right", axis.text.x = element_text(angle = 45, hjust = 1)) + xlab("Date") + ylab("Number of Individuals") +
+  geom_vline(xintercept = c(120), linetype = 2) +
+  geom_vline(xintercept = c(181), linetype = 3, color = "grey20") +
+  geom_vline(xintercept = c(243), linetype = 4, color = "grey80")
+
+save_plot("Phenology_Data_by_Site_Genetic_Subpop_2019-12-30.png", plot = last_plot(), base_height = 7, base_width = 11)
+
+phenology_long_2019 %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  filter(Genetic_subpop != "Unknown" & PLANT_ID != "AP13") %>%
+  mutate(Days = as_date(Value)) %>%
+  ggplot(aes(x = Days)) + 
+  geom_histogram(aes(fill = Genetic_subpop)) + 
+  facet_grid(SITE~PHE, scales = "free") +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = -1) +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1)) +#,
+        #panel.spacing = unit(2, "lines")) +
+  xlab("Date") + 
+  ylab("Number of Individuals") +
+  geom_vline(xintercept = c(120), linetype = 2) +
+  geom_vline(xintercept = c(181), linetype = 3, color = "grey20") +
+  geom_vline(xintercept = c(243), linetype = 4, color = "grey80")
+
+save_plot("Phenology_Data_by_Site_Genetic_Subpop_noAP13_2019-12-30.png", plot = last_plot(), base_height = 8, base_width = 14)
+
+phenology_long_2019 %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  filter(Genetic_subpop != "Unknown" & PLANT_ID != "AP13") %>%
+  mutate(Days = as_date(Value)) %>%
+  ggplot(aes(x = Days)) + 
+  geom_histogram(aes(fill = Genetic_subpop)) + 
+  facet_grid(PHE~SITE, scales = "free") +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = -1) +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1)) +#,
+        #panel.spacing = unit(2, "lines")) +
+  xlab("Date") + 
+  ylab("Number of Individuals") +
+  geom_vline(xintercept = c(120), linetype = 2) +
+  geom_vline(xintercept = c(181), linetype = 3, color = "grey20") +
+  geom_vline(xintercept = c(243), linetype = 4, color = "grey80")
+
+save_plot("Phenology_Data_by_Site_Genetic_Subpop_noAP13_Coord_Flip_2019-12-30.png", plot = last_plot(), base_height = 8, base_width = 14)
+
+```
+
+## Phenology plots by Ecotype
+
+```{r}
+phenology_long_2019 %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  filter(Genetic_subpop != "Unknown" & PLANT_ID != "AP13") %>%
+  mutate(Days = as_date(Value)) %>%
+  ggplot(aes(x = Days)) + 
+  geom_histogram(aes(fill = Ecotype)) + 
+  facet_grid(SITE~PHE, scales = "free") +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = -1) +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1)) +#,
+        #panel.spacing = unit(2, "lines")) +
+  xlab("Date") + 
+  ylab("Number of Individuals") +
+  geom_vline(xintercept = c(120), linetype = 2) +
+  geom_vline(xintercept = c(181), linetype = 3, color = "grey20") +
+  geom_vline(xintercept = c(243), linetype = 4, color = "grey80")
+
+save_plot("Phenology_Data_by_Site_Ecotype_noAP13_2019-12-30.png", plot = last_plot(), base_height = 8, base_width = 14)
+
+phenology_long_2019 %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  filter(Genetic_subpop != "Unknown" & PLANT_ID != "AP13") %>%
+  mutate(Days = as_date(Value)) %>%
+  ggplot(aes(x = Days)) + 
+  geom_histogram(aes(fill = Ecotype)) + 
+  facet_grid(PHE~SITE, scales = "free") +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = -1) +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1)) +#,
+        #panel.spacing = unit(2, "lines")) +
+  xlab("Date") + 
+  ylab("Number of Individuals") +
+  geom_vline(xintercept = c(120), linetype = 2) +
+  geom_vline(xintercept = c(181), linetype = 3, color = "grey20") +
+  geom_vline(xintercept = c(243), linetype = 4, color = "grey80")
+
+save_plot("Phenology_Data_by_Site_Ecotype_noAP13_Coord_Flip_2019-12-30.png", plot = last_plot(), base_height = 8, base_width = 14)
+
+```

### 2019-12-31 data-raw/sites.R

@@ -1,0 +1,16 @@
+library(tidyverse)
+
+sites <- tribble(~Site, ~Latitude, ~Longitude,
+                 "PKLE", 30.383979, -97.729383,
+                 "FRMI", 41.837133, -88.239667,
+                 "OVTN", 32.3029, -94.9794,
+                 "KING", 27.549863, -97.881012,
+                 "BRKG", 44.3068, -96.6705,
+                 "CLMB", 38.8969, -92.2178,
+                 "KBSM", 42.419618, -85.371266,
+                 "LINC", 41.1543, -96.4153,
+                 "STIL", 35.991148, -97.046489,
+                 "TMPL", 31.043383, -97.349498
+)
+
+save(sites, file = "data/sites.rda")

### 2020-01-06 data-raw/phenotypes_2019.Rmd
Plots of 2019 phenotyping errors and 2019 phenotypes

@@
 library(scales)
+library(lubridate)
 
@@
 phenos_2019 %>%
+  filter(SRV == "W" & DEAD_2018 == 2)
+
+phenos_2019 %>%
   filter(SRV == "W" & DEAD_2018 == 1)
@@
 Anything where DEAD_2018 = 2 is called as "Not virgatum" in the past, so should be dropped. Does Jason agree?
+FRMI	F2610	2610	J203	J203.A	NA	156	NA	NA	W	Likely not virgatum, Weak	switchgrass?
+FRMI	F2707	2707	J421	J421.A	NA	149	176	NA 	W	Likely not virgatum, Weak	switchgrass?
 
@@
 ```{r}
-load("../data/Taxa.rda")
-
 phenotypes_wide_2019 <- phemean_2019 %>%
@@
 
-sites$Factor <- factor(sites$Factor, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
-sites$Factor <- factor(sites$Factor, levels = c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING"))
-levels(sites$Factor)
+#sites$Factor <- factor(sites$Factor, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
+#sites$Factor <- factor(sites$Factor, levels = c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING"))
+#levels(sites$Factor)
 
-phenotypes_long_2019$SITE <- factor(phenotypes_long_2019$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
+# phenotypes_long_2019$SITE <- factor(phenotypes_long_2019$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
 phenotypes_long_2019$SITE <- factor(phenotypes_long_2019$SITE, levels = c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING"))
 phenotypes_long_2019$PHE <- factor(phenotypes_long_2019$PHE, levels = c("GR1", "GR50", "GR100", "EM1", "EM50", "FL1", "FL50", "TC_EOS", "HT_PAN_EOS", "FRZ_DMG", "CHLR_101", "CHLR_121", "CHLR_136", "CHLR_JASON_138", "CHLR_150.", "SPAD_158", "SPADCOR_158"))
+
+phenology_long_2019 <- phenotypes_long_2019 %>%
+  filter(!(grepl("CHLR", PHE)) & !(grepl("SPAD", PHE)) & !grepl("FRZ", PHE) & !is.na(PHE) & !grepl("TC_EOS", PHE) & !grepl("HT_PAN_EOS", PHE))
 
@@
 ```{r}
-
 phenotypes_long_2019 %>%
@@
 
-phenology_long_2019 <- phenotypes_long_2019 %>%
-  filter(!(grepl("CHLR", PHE)) & !(grepl("SPAD", PHE)) & !grepl("FRZ", PHE) & !is.na(PHE) & !grepl("TC_EOS", PHE) & !grepl("HT_PAN_EOS", PHE))
-
-library(lubridate)
 phenology_long_2019 %>%
@@
 ```{r}
-
-load("../data/metadata.rda")
-
 metadata <- metadata %>%
@@
            TRUE ~ JL_genetic_subpop
+           ),
+        Ecotype = case_when(
+          Ecotype == "TX" ~ "Texas",
+          Ecotype == "Coast" ~ "Coastal",
+          Ecotype == "Up" ~ "Upland"
           ))
 metadata$Genetic_subpop <- factor(metadata$Genetic_subpop, levels = c("TX", "GC,TX", "GC", "SEC", "NEC,SEC", "NEC", "MW", "Admixed", "Unknown"))
-metadata$Ecotype <- factor(metadata$Ecotype, levels = c("TX", "Coast", "Up"))
+metadata$Ecotype <- factor(metadata$Ecotype, levels = c("Texas", "Coastal", "Upland"))
 
@@
 ```
+
+## Other phenotypes by Ecotype/Subpop
+
+```{r}
+phenotypes_long_2019 %>%
+  filter(grepl("TC_EOS", PHE) | grepl("HT_PAN_EOS", PHE) | grepl("FRZ", PHE)) %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  filter(!(Genetic_subpop %in% c("Admixed", "Unknown")) & 
+           PLANT_ID != "AP13") %>%
+  ggplot(aes(x = Value)) +
+  geom_histogram(aes(fill = Genetic_subpop)) + 
+  facet_grid(SITE~PHE, scales = "free") +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = -1) +
+  theme(legend.position = "right", axis.text.x = element_text(angle = 45, hjust = 1)) + xlab("Value") + ylab("Count of Individuals")
+
+save_plot("2019_Phenotypic_Data_by_Site_and_Genetic_Subpop_2019-12-30.png", plot = last_plot(), base_height = 10, base_width = 9)
+
+phenotypes_long_2019 %>%
+  filter(grepl("TC_EOS", PHE) | grepl("HT_PAN_EOS", PHE) | grepl("FRZ", PHE)) %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  filter(!(Genetic_subpop %in% c("Unknown")) & 
+           PLANT_ID != "AP13") %>%
+  ggplot(aes(x = Value)) +
+  geom_histogram(aes(fill = Ecotype)) + 
+  facet_grid(SITE~PHE, scales = "free") +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = -1) +
+  theme(legend.position = "right", axis.text.x = element_text(angle = 45, hjust = 1)) + xlab("Value") + ylab("Count of Individuals")
+
+save_plot("2019_Phenotypic_Data_by_Site_and_Ecotype_2019-12-30.png", plot = last_plot(), base_height = 10, base_width = 9)
+```
+
+# Do dates make sense for each SITE*PLANT_ID?
+
+In the aggregate these values seem to make sense, check them for individual cases.
+
+```{r}
+phenotypes_long_2019 %>%
+  filter(grepl("TC_EOS", PHE) | grepl("HT_PAN_EOS", PHE)) %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  filter(!(Genetic_subpop %in% c("Unknown")) & 
+           PLANT_ID != "AP13") %>%
+  pivot_wider(names_from = PHE, values_from = Value) %>%
+  ggplot(aes(x = TC_EOS, y = HT_PAN_EOS)) +
+  geom_point(aes(color = Ecotype), alpha = 0.2) +
+  facet_wrap(~SITE) +
+  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1) +
+  theme(legend.position = "right")
+
+phe_comp <- phenology_long_2019 %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  filter(!(Genetic_subpop %in% c("Unknown")) & 
+           PLANT_ID != "AP13") %>%
+  pivot_wider(names_from = PHE, values_from = Value)
+```
+
+```{r}
+phe_comp %>%
+  ggplot(aes(x = GR50, y = EM50)) +
+  geom_point(aes(color = Ecotype), alpha = 0.2) +
+  facet_wrap(~SITE) +
+  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1) +
+  theme(legend.position = "right") + 
+  geom_abline(slope = 1, intercept = 0)
+save_plot("Greenup_vs_Emergence_50percent_by_Site_2019-12-31.png", plot = last_plot(), base_height = 4)
+phe_comp %>%
+  ggplot(aes(x = GR50, y = FL50)) +
+  geom_point(aes(color = Ecotype), alpha = 0.2) +
+  facet_wrap(~SITE) +
+  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1) +
+  theme(legend.position = "right") + 
+  geom_abline(slope = 1, intercept = 0)
+save_plot("Greenup_vs_Flowering_50percent_by_Site_2019-12-31.png", plot = last_plot(), base_height = 4)
+phe_comp %>%
+  ggplot(aes(x = EM50, y = FL50)) +
+  geom_point(aes(color = Ecotype), alpha = 0.2) +
+  facet_wrap(~SITE) +
+  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1) +
+  theme(legend.position = "right") + 
+  geom_abline(slope = 1, intercept = 0)
+save_plot("Emergence_vs_Flowering_50percent_by_Site_2019-12-31.png", plot = last_plot(), base_height = 4)
+```
+
+TMPL has some EM1 and EM50 issues. not collected very much at later timepoints.
+OVTN: I don't have EM or FL data. Was this data not collected?
+
+TMPL and PKLE (and KING, and STIL less so) have a huge spread of FL1 vs GR1. Also true for FL1 vs GR100 and FL50 vs GR100. GxE?!
+
+```{r}
+
+phe_comp %>%
+  ggplot(aes(x = FL1, y = FL50)) +
+  geom_point(aes(color = Ecotype), alpha = 0.5) +
+  facet_wrap(~SITE) +
+  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1) +
+  theme(legend.position = "right") + 
+  geom_abline(slope = 1, intercept = 0)
+save_plot("Possible_Flowering_Errors_below_the_line_by_Site_2019-12-30.png", plot = last_plot(), base_height = 4)
+
+phe_comp %>%
+  ggplot(aes(x = GR1, y = GR100)) +
+  geom_point(aes(color = Ecotype), alpha = 0.2) +
+  #facet_wrap(~SITE) +
+  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1) +
+  theme(legend.position = "right") + 
+  geom_abline(slope = 1, intercept = 0)
+
+save_plot("Possible_Greenup_Errors_below_the_line_by_Site_2019-12-30.png", plot = last_plot(), base_height = 4)
+
+phe_comp %>%
+  ggplot(aes(x = EM1, y = EM50)) +
+  geom_point(aes(color = Ecotype), alpha = 0.4) +
+  facet_wrap(~SITE) +
+  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1) +
+  theme(legend.position = "right") + 
+  geom_abline(slope = 1, intercept = 0)
+save_plot("Possible_Emergence_Errors_below_the_line_by_Site_2019-12-30.png", plot = last_plot(), base_height = 4)
+```
+Generally, EM50 should not be less than EM1, and it certainly shouldn't be much less (like where EM1 is 200 and EM50 is 50, must be an error)
+```{r}
+phe_comp %>%
+  mutate(EMdiff = EM50 - EM1) %>%
+  filter(EMdiff < 0) %>%
+  dplyr::select(SITE:PLANT_ID, EM1, EM50, EMdiff)
+
+phe_comp %>%
+  mutate(FLdiff = FL50 - FL1) %>%
+  filter(FLdiff < 0) %>%
+  dplyr::select(SITE:PLANT_ID, FL1, FL50, FLdiff)
+```
+SITE PLOT_GL PLOT_LC ACC PLANT_ID EM1 EM50 EMdiff
+
+STIL	S1110	1110	J587	J587.A	186	185	-1
+STIL	S2907	2907	J073	J073.A	190	184	-6
+STIL	S3603	3603	J210	J210.A	188	186	-2
+TMPL	T1417	1417	J321	J321.A	182	180	-2
+TMPL	T2101	2101	J586	J586.B	199	182	-17
+TMPL	T2103	2103	J008	J008.A	182	126	-56
+TMPL	T2111	2111	J326	J326.B	199	19	-180
+TMPL	T2203	2203	J614	J614.B	142	141	-1
+TMPL	T2411	2411	J270	J270.A	199	99	-100
+
+For GxE study:
+
+Greenup date (50%)
+time to bolting (EM50 - GR50)
+Panicle emergence date (50%)
+time to flowering (FL50 - GR50)
+Flowering date (50%)
+Tiller count EOS 2019
+Panicle height EOS 2019
+
+Also GR1, GR100, EM1, FL1, CHLR/SPAD in KING, FRZ_DAM, but the 7 above seem more interesting.
+
+# Remove outliers
+
+```{r}
+phe_2019_fil <- phenotypes_long_2019 %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  filter(!is.na(JL_genetic_subpop)) %>%
+  pivot_wider(names_from = PHE, values_from = Value) %>%
+  dplyr::select(SITE:PLANT_ID, LIB_PHENO, Ecotype, GR1:FRZ_DMG) %>%
+  filter(LIB_PHENO == "Y") %>%
+  mutate(EM50 = ifelse(EM50 - EM1 < 0, NA, EM50),
+         FL50 = ifelse(FL50 - FL1 < 0, NA, FL50),
+         TBOLT = EM50 - GR50,
+         TFLOW = FL50 - GR50)
+
+phemean_2019 <- phe_2019_fil %>%
+  group_by(PLANT_ID, SITE) %>%
+  summarise(FRZ_DMG = mean(FRZ_DMG, na.rm = TRUE),
+            GR1 = mean(GR1, na.rm = TRUE),
+            GR50 = mean(GR50, na.rm = TRUE),
+            GR100 = mean(GR100, na.rm = TRUE),
+            EM1 = mean(EM1, na.rm = TRUE),
+            EM50 = mean(EM50, na.rm = TRUE),
+            FL1 = mean(FL1, na.rm = TRUE),
+            FL50 = mean(FL50, na.rm = TRUE),
+            TC_EOS = mean(TC_EOS, na.rm = TRUE),
+            HT_PAN_EOS = mean(HT_PAN_EOS, na.rm = TRUE),
+            CHLR_101 = mean(CHLR_101, na.rm = TRUE),
+            CHLR_121 = mean(CHLR_121, na.rm = TRUE),
+            CHLR_136 = mean(CHLR_136, na.rm = TRUE),
+            CHLR_JASON_138 = mean(CHLR_JASON_138, na.rm = TRUE),
+            CHLR_150. = mean(CHLR_150., na.rm = TRUE),
+            SPAD_158 = mean(SPAD_158, na.rm = TRUE),
+            SPADCOR_158 = mean(SPADCOR_158, na.rm = TRUE),
+            TBOLT = mean(TBOLT, na.rm = TRUE),
+            TFLOW = mean(TFLOW, na.rm = TRUE)
+  ) %>%
+  mutate_all( ~ case_when(!is.nan(.x) ~ .x)) %>% # if the column is not NA, keep the value; else replace with NA as nothing is provided as another choice in case_when.
+  filter(!is.na(SITE) & !is.na(PLANT_ID))
+```
+
+```{r}
+phenotypes_wide_2019 <- phemean_2019 %>%
+  dplyr::select(PLANT_ID:SITE, GR1:HT_PAN_EOS, TBOLT, TFLOW) %>%
+  gather(c(GR1:TFLOW), key = "Phenotype", value = value) %>%
+  filter(!is.na(value)) %>%
+  # saveRDS(file = "Cleaned pvdiv Phenotypes from 2018 for 784g 2019-05-16.rds")
+  unite(col = "SITE_PHE", SITE, Phenotype) %>%
+  spread(key = SITE_PHE, value = value) %>%
+  right_join(Taxa)
+
+save(phenotypes_wide_2019, file = "../data/phenotypes_2019.rda")
+```
+
+# How many plants are phenotyped at each site?
+
+```{r}
+colSums(!is.na(phenotypes_wide_2019))
+
+phenotypes_wide_2019 %>%
+  left_join(metadata) %>%
+  group_by(Ecotype, is.na(BRKG_GR50)) %>%
+  tally()
+
+phenotypes_long_2019 %>%
+  left_join(metadata) %>%
+  group_by(SITE, PHE, Ecotype) %>%
+  tally()
+phemean_2019 %>%
+  left_join(metadata) %>%
+  dplyr::select(PLANT_ID, SITE, GR1:HT_PAN_EOS) %>%
+  ggplot(aes(x = PLANT_ID, y = SITE)) +
+  geom_heatm
+  pivot_longer(cols = GR1:HT_PAN_EOS, names_to = "PHE", values_to = "Value") %>%
+  group_by(SITE, PHE) %>%
+  tally()
+
+tibble(PLANT_ID = Taxa$PLANT_ID, PHE = rowSums(!is.na(phenotypes_wide_2019)) - 1) %>%
+  filter(PHE > 0)
+
+colremove <- names(which(colSums(!is.na(phenotypes_wide_2019)) < 200))
+
+phe_gwas_2019 <- phenotypes_wide_2019 %>%
+  dplyr::select(-colremove, -ends_with("1"), -ends_with("100"))
+
+plotphe <- phe_gwas_2019 %>%
+  #dplyr::select(PLANT_ID, ends_with("TBOLT")) %>%
+  pivot_longer(cols = BRKG_EM50:TMPL_TFLOW, names_to = "PHE", values_to = "VALUE") %>%
+  left_join(metadata) %>%
+  filter(!is.na(VALUE)) %>%
+  separate(PHE, into = c("SITE", "_", "PHE"), sep = c(4,5))
+
+plotphe$SITE <- factor(plotphe$SITE, levels = c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING"))
+plotphe$PHE <- factor(plotphe$PHE, levels = c("GR50", "EM50", "FL50", "TBOLT", "TFLOW", "TC_EOS", "HT_PAN_EOS"))
+
+plotphe %>%
+  group_by(PHE, Ecotype) %>% 
+  ggplot(aes(x = SITE)) +
+  geom_bar(aes( fill = Ecotype)) +
+  facet_wrap(~PHE) +
+  scale_fill_viridis(discrete = TRUE, end = 0.95, direction = -1) +
+  theme(legend.position = "right", axis.text.x = element_text(angle = 45, 
+                                                              hjust = 1)) +
+  xlab("Site") + ylab("Number of Individuals")
+
+save_plot("Number_of_Individuals_per_Phenotype_2_2020-01-04.png", plot = last_plot(), base_height = 4*2.5, base_width = 5*2.5)
+
+summary(metadata$Ecotype)
+```

### 2020-01-06 data-raw/phenotypes_2019.Rmd
Modify phenotypes_2019 and metadata to have AP13 values

@@
 
-source("C:/Users/alice/Documents/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+source("../../../Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
 theme_set(theme_poster)
@@
 ```{r}
-wb_pheno <- loadWorkbook(file.path("C:", "Users", "alice", "Dropbox",
+wb_pheno <- loadWorkbook(file.path("C:", "Users", "ahm543", "Dropbox",
                                    paste0("GWAS DOE Switchgrass, ",
@@
 
-wb_KING2019 <- loadWorkbook(file.path("C:", "Users", "alice", "Dropbox",
+wb_KING2019 <- loadWorkbook(file.path("C:", "Users", "ahm543", "Dropbox",
                                       paste0("GWAS DOE Switchgrass, ",
@@
 
-wb_DEAD2018 <- loadWorkbook(file.path("C:", "Users", "alice", "Dropbox",
+wb_DEAD2018 <- loadWorkbook(file.path("C:", "Users", "ahm543", "Dropbox",
                                       paste0("GWAS DOE Switchgrass, ",
@@
 
-metadata <- add_row(metadata, PLANT_ID = "AP13", LIB_CLIMATE = "Y", LIB_PHENO = "Y", LIB_ACC = "Y", Ecotype = "TX", Genetic_subpop = "TX")
+# metadata <- metadata %>%
+#   filter(PLANT_ID != "AP13")
+
+metadata <- add_row(metadata, PLANT_ID = "AP13", LIBRARY = "REF", LIB_CLIMATE = "Y", LIB_PHENO = "Y", LIB_ACC = "Y", JL_genetic_subpop = "TX", JL_detail_genetic = "TX", PLOIDY = "4X", Ecotype = "TX", Genetic_subpop = "TX")
+
+save(metadata, file = "../data/metadata.rda")
 ```

### 2020-01-06 data-raw/phenotypes_2019.R

@@ -1,0 +1,151 @@
+# code to generate phenotypes_wide_2019 goes here
+
+library(tidyverse)
+library(XLConnect)
+library(rlang)
+library(switchgrassGWAS)
+library(cowplot)
+library(viridis)
+library(bigsnpr)
+library(scales)
+library(lubridate)
+
+source("../../Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+theme_set(theme_poster)
+
+# Load all data
+
+load("data/sites.rda")
+load("data/metadata.rda")
+load("data/Taxa.rda")
+
+usa <- map_data("state")
+wb_pheno <- loadWorkbook(file.path("C:", "Users", "ahm543", "Dropbox",
+                                   paste0("GWAS DOE Switchgrass, ",
+                                          "Consolidated Data"),
+                                   paste0("DOE_GWAS_2019_Consolidated ",
+                                          "Pheno Data_12-25-19.xlsx")))
+lst_pheno <- readWorksheet(wb_pheno, sheet = getSheets(wb_pheno))
+phenos_planting1 <- lst_pheno$`GWAS 2019 Greenup Data`
+phenos_planting2 <- lst_pheno$`GWAS 2019 Core Phenotype Data`
+phenos_key <- lst_pheno$`Column Key and Notes`
+
+wb_KING2019 <- loadWorkbook(file.path("C:", "Users", "ahm543", "Dropbox",
+                                      paste0("GWAS DOE Switchgrass, ",
+                                             "Consolidated Data"),
+                                      "Individual Site Files",
+                                      "KING_GWAS_2019_Chlorosis Data.xlsx"))
+lst_KING2019 <- readWorksheet(wb_KING2019, sheet = getSheets(wb_KING2019))
+KING2019 <- as_tibble(lst_KING2019$`KING GWAS 2019 Chlorosis`)
+key_KING19 <- as_tibble(lst_KING2019$`COLUMN KEY`)
+
+wb_DEAD2018 <- loadWorkbook(file.path("C:", "Users", "ahm543", "Dropbox",
+                                      paste0("GWAS DOE Switchgrass, ",
[…]

### 2020-01-06 analysis/1_bigsnpr_gwas_prep.Rmd
PAG talk draft from Tom's slides

---
title: "1_bigsnpr_gwas_prep"
author: "Alice MacQueen"
date: "1/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(XLConnect)
library(rlang)
library(bigsnpr)
library(switchgrassGWAS)
library(AnnotationDbi)
library(cowplot)
library(viridis)
library(bigsnpr)
library(scales)
library(lubridate)

source("../../Functions_ggplot-theme-adjustments_2018-01-03.R")
theme_set(theme_poster)

load("../data/sites.rda")
load("../data/metadata.rda")
load("../data/Taxa.rda")
load("../data/phenotypes_2019.rda")

usa <- map_data("state")
```

## prepare phenotypes for GWAS

```{r}
colremove <- names(which(colSums(!is.na(phenotypes_wide_2019)) < 200))

phe_gwas_4big <- phenotypes_wide_2019 %>%
  dplyr::select(-colremove, -ends_with("1"), -ends_with("100"), 
                -starts_with("BRKG"), -starts_with("FRMI"), 
                -starts_with("LINC"), -starts_with("OVTN"), 
                -starts_with("TMPL"), -starts_with("KING"))

```

# gwas with 9 PCs for 509 individuals

```{r}

# Load files used in creating the annotation tables.

txdb <- loadDb(file = file.path("/", "home", "alice", "pvdiv", "switchgrassGWAS", 
                                "data",
                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))

anno_info <-
  read_delim(file = file.path("/", "home", "alice",  "pvdiv", "switchgrassGWAS", 
                              "data",
                              "Pvirgatum_516_v5.1.annotation_info.txt"),
             col_names = TRUE, delim = "\t")
gene_anno_info <- tbl_df(anno_info) %>%
  distinct(locusName, .keep_all = TRUE)

svd <- readRDS(file.path("/", "home", "alice", "Github", "pvdiv-phenology-gxe",
                         "large-files", "SVD_upland_lowland_continuous_k20.rds"))

snp <- snp_attach(file.path("~", "pvdiv", "VCF_and_Numeric", "Pvirgatum_4x_784g_imp_phased_up_low_continuous_subset_maf0.05.rds"))

  G <- snp$genotypes
  CHR <- snp$map$chromosome
  POS <- snp$map$physical.pos
  plants <- snp$fam$sample.ID
  bonferroni <- -log10(0.05/length(POS))
  NCORES <- nb_cores()

  CHRN <- enframe(CHR, name = NULL) %>%
    dplyr::rename(CHR = .data$value) %>%
    mutate(CHRN = case_when(.data$CHR == "Chr01K" ~ 1,
                            .data$CHR == "Chr01N" ~ 2,
                            .data$CHR == "Chr02K" ~ 3,
                            .data$CHR == "Chr02N" ~ 4,
                            .data$CHR == "Chr03K" ~ 5,
                            .data$CHR == "Chr03N" ~ 6,
                            .data$CHR == "Chr04K" ~ 7,
                            .data$CHR == "Chr04N" ~ 8,
                            .data$CHR == "Chr05K" ~ 9,
                            .data$CHR == "Chr05N" ~ 10,
                            .data$CHR == "Chr06K" ~ 11,
                            .data$CHR == "Chr06N" ~ 12,
                            .data$CHR == "Chr07K" ~ 13,
                            .data$CHR == "Chr07N" ~ 14,
                            .data$CHR == "Chr08K" ~ 15,
                            .data$CHR == "Chr08N" ~ 16,
                            .data$CHR == "Chr09K" ~ 17,
                            .data$CHR == "Chr09N" ~ 18,
                            TRUE ~ 19
    ))
  markers <- tibble(CHR = CHR, POS = POS)

  plants <- plants %>% enframe(value="PLANT_ID") %>% dplyr::select(-name)
  plants %>%
    left_join(phe_gwas_4big) %>% arrange(PLANT_ID)

  df_phe <- plants %>%
    left_join(phe_gwas_4big)

  gwas <- pvdiv_gwas(df = df_phe, type = "linear", snp = snp, ncores = NCORES, 
                     covar = svd, npcs = 9)

```

# Mash

STIL, KBSM, PKLE, CLMB

GR50
EM50
FL50
TC_EOS
HT_PAN_EOS
T_GR_EM
T_GR_FL
T_EM_FL

```{r}
markers2sub <- attr(svd, "subset")
markers2 <- tibble(CHR = CHR[markers2sub], POS = POS[markers2sub])

gwas_rds <- pvdiv_results_in_folder(path = ".", pattern = "GWAS_object_") # 8 phenotypes from 4 sites - 32 phenotypes total with > 200 individuals phenotyped and sequenced.

phenotype_names <- str_sub(gwas_rds, start = 13, end = -5)

numSNPs <- ceiling(1500000/length(phenotype_names)^2)
numSNPs <- 1500 # the above gave 1465, so just rounded this to 1500.

mash_input <- pvdiv_bigsnp2mashr(path = ".", gwas_rds = gwas_rds, 
                           phenotypes = phenotype_names, numSNPs = numSNPs, 
                           markers = markers, markers2 = markers2, G = G, 
                           model = "linear", saveoutput = TRUE)

mash_output <- mash_standard_run(path = ".", list_input = mash_input, 
                                  numSNPs = numSNPs, saveoutput = TRUE)
```

## Visualize bigsnpr GWAS

```{r}
for(i in 2:ncol(phe_gwas_4big)){

  df1 <- phe_gwas_4big %>%
    dplyr::select(PLANT_ID, i)
  gwas <- pvdiv_gwas(df = df1, type = "linear", snp = snp, ncores = NCORES, 
                     covar = svd, npcs = 9)

  ggsave(snp_manhattan(gwas, infos.chr = CHRN$CHRN,
                       infos.pos = snp$map$physical.pos, npoints = 2.1E06) +
           geom_hline(yintercept = bonferroni, lty = 2),
         filename = paste0("Manhattan_logistic_winter_survival", 
                           "_subset_9_PCs.png"), width = 10, height = 6,
         path = file.path("~", "pvdiv", "climateGWAS", 
                          "results-upland-vs-lowland", 
                          "Logistic_winter_survival"))
  anno_tables <- pvdiv_table_topsnps(df = gwas, type = "bigsnp", 
                                     n = c(10,100,1000), FDRalpha = 0.1, 
                                     rangevector = c(0, 20000, 100000),
                                     markers = markers, 
                                     anno_info = gene_anno_info,
                                     txdb = txdb)
  saveRDS(anno_tables, file.path("~", "pvdiv", "climateGWAS", 
                                 "results-upland-vs-lowland", 
                                 "Logistic_winter_survival",
                                 paste0("Annotation_tables_", 
                                        "logistic_winter_survival", 
                                        "_subset_9_PCs.rds")))
  ggsave(snp_qq(gwas[which(!is.na(gwas$score)),]),
         filename = paste0("QQplot_logistic_winter_survival_", 
                           "subset_9_PCs.png"), width = 4,
         height = 4, path = file.path("~", "pvdiv", "climateGWAS", 
                                      "results-upland-vs-lowland", 
                                      "Logistic_winter_survival"))
}

```

# Local score post processing of bigsnpr GWAS

```{r}
log10p <- predict(gwas)
pval <- 10^log10p

data1 <- data.table(chr = CHR, pos = POS, estim = gwas$estim, 
                    std.err = gwas$std.err, bigsnpscore = gwas$score, 
                    pval = pval)

setkey(data1, chr)
data1$pval[data1$pval==0]=1e-16#min(data1$pval[-which(data1$pval ==0)])
Nchr=length(data1[,unique(chr)])

### Computation of absolute position in the genome. 
#This is useful for doing genomewide plots. 
chrInfo=data1[,.(L=.N,cor=autocor(pval)),chr]
setkey(chrInfo,chr)
data.table(chr=data1[,unique(chr),], S=cumsum(c(0,chrInfo$L[-Nchr])))

# The score mean must be negative, ksi must be chosen between mean and max of -log10(p-value)

xi=2
data1[,score:= -log10(pval)-xi]
data1[,lindley:=lindley(score),chr]

### Choice of $\xi$ (1,2,3 or 4)

mean(data1$score)

# verify mean is negative and max is positive
mean(-log10(data1$pval));max(-log10(data1$pval))

#hist(data1$score)
max(data1$lindley)

# Compute significance threshold for each chromosome
## Uniform distribution of p-values
chrInfo[,th:=thresUnif(L, cor, xi),chr]
(data1=data1[chrInfo])

(sigZones=data1[chrInfo,sig_sl(lindley, pos, unique(th)),chr])

saveRDS(data1, "Logistic_winter_tolerance_gwas_lindley_scores.rds")
saveRDS(sigZones, "Logistic_winter_tolerance_Lindley_sigZones.rds")
```

# Visualize local score GWAS

```{r}
phenos <- "winter_tolerance"
i = 1

  ggmanobject <- ggplot(data = data1, aes(x = pos, y = lindley)) +
    geom_point(aes(color = chr, fill = chr)) +
    geom_hline(yintercept = data1$th[1], linetype = 2) + 
    facet_wrap(~ chr, nrow = 1, scales = "free_x", 
               strip.position = "bottom") +
    scale_color_viridis(option = "B", end = 0.8, discrete = TRUE) + 
    scale_fill_viridis(option = "B", end = 0.8, discrete = TRUE) +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          panel.background = element_rect(fill=NA)) +
    labs(x = "Chromosome", y = "Lindley") +
    scale_x_continuous(expand = c(0.22, 0.22)) #+
    #ylim(0, 40)
    #scale_shape_manual(values = rep(c(21,22),9), guide = FALSE)

    save_plot(paste0("Manhattan_", phenos[i], "_ksi_2_ylim_", 
                     str_replace_all(Sys.time(), ":", "."), ".png"), 
              plot = ggmanobject, base_aspect_ratio = 2.5, base_height = 5)
```

### 2020-01-07 analysis/1_bigsnpr_gwas_prep.Rmd
Mash analysis of GR50 and FL50 with plots

@@
 
+# T_GR_FL
+
+```{r}
+colremove <- names(which(colSums(!is.na(phenotypes_wide_2019)) < 200))
+
+phe_gwas_tgrfl <- phenotypes_wide_2019 %>%
+  dplyr::select(-colremove) %>%
+  dplyr::select(PLANT_ID, ends_with("T_GR_FL"))
+
+txdb <- loadDb(file = file.path("/", "home", "alice", "pvdiv", "switchgrassGWAS", 
+                                "data",
+                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
+
+anno_info <-
+  read_delim(file = file.path("/", "home", "alice",  "pvdiv", "switchgrassGWAS", 
+                              "data",
+                              "Pvirgatum_516_v5.1.annotation_info.txt"),
+             col_names = TRUE, delim = "\t")
+gene_anno_info <- tbl_df(anno_info) %>%
+  distinct(locusName, .keep_all = TRUE)
+
+svd <- readRDS(file.path("/", "home", "alice", "Github", "pvdiv-phenology-gxe",
+                         "large-files", "SVD_upland_lowland_continuous_k20.rds"))
+
+snp <- snp_attach(file.path("~", "pvdiv", "VCF_and_Numeric", "Pvirgatum_4x_784g_imp_phased_up_low_continuous_subset_maf0.05.rds"))
+
+  G <- snp$genotypes
+  CHR <- snp$map$chromosome
+  POS <- snp$map$physical.pos
+  plants <- snp$fam$sample.ID
+  bonferroni <- -log10(0.05/length(POS))
+  NCORES <- nb_cores()
+
+  CHRN <- enframe(CHR, name = NULL) %>%
+    dplyr::rename(CHR = .data$value) %>%
+    mutate(CHRN = case_when(.data$CHR == "Chr01K" ~ 1,
+                            .data$CHR == "Chr01N" ~ 2,
+                            .data$CHR == "Chr02K" ~ 3,
+                            .data$CHR == "Chr02N" ~ 4,
+                            .data$CHR == "Chr03K" ~ 5,
+                            .data$CHR == "Chr03N" ~ 6,
+                            .data$CHR == "Chr04K" ~ 7,
+                            .data$CHR == "Chr04N" ~ 8,
+                            .data$CHR == "Chr05K" ~ 9,
+                            .data$CHR == "Chr05N" ~ 10,
+                            .data$CHR == "Chr06K" ~ 11,
+                            .data$CHR == "Chr06N" ~ 12,
+                            .data$CHR == "Chr07K" ~ 13,
+                            .data$CHR == "Chr07N" ~ 14,
+                            .data$CHR == "Chr08K" ~ 15,
+                            .data$CHR == "Chr08N" ~ 16,
+                            .data$CHR == "Chr09K" ~ 17,
+                            .data$CHR == "Chr09N" ~ 18,
+                            TRUE ~ 19
+    ))
+  markers <- tibble(CHR = CHR, POS = POS)
+
+  plants <- plants %>% enframe(value="PLANT_ID") %>% dplyr::select(-name)
+  plants %>%
+    left_join(phe_gwas_tgrfl) %>% arrange(PLANT_ID)
+
+  df_phe <- plants %>%
+    left_join(phe_gwas_tgrfl)
+
+  gwas <- pvdiv_gwas(df = df_phe, type = "linear", snp = snp, ncores = NCORES, 
+                     covar = svd, npcs = 9)
+
+```
+
+# Mash 8 sites T_GR_FL
+```{r}
+load(file = "../markers.rda")
+load(file = "../markers_unlinked.rda")
+gwas_rds <- pvdiv_results_in_folder(path = ".", pattern = "GWAS_object_") # 8 phenotypes from 4 sites - 32 phenotypes total with > 200 individuals phenotyped and sequenced.
+
+phenotype_names <- str_sub(gwas_rds, start = 13, end = -5)
+
+numSNPs <- ceiling(1000000/length(phenotype_names)^2)
+numSNPs <- 20000 # the above gave 20409, so just rounded this to 20000.
+
+mash_input <- pvdiv_bigsnp2mashr(path = ".", gwas_rds = gwas_rds, 
+                           phenotypes = phenotype_names, numSNPs = numSNPs, 
+                           markers = markers, markers2 = markers2, G = G, 
+                           model = "linear", saveoutput = TRUE)
+
+mash_output <- mash_standard_run(path = ".", list_input = mash_input, 
+                                  numSNPs = numSNPs, saveoutput = TRUE)
+
+anno_tables <- pvdiv_table_topsnps(df = mash_output, type = "mash", 
+                                   n = c(10,100), FDRalpha = NA, 
+                                   rangevector = c(0, 20000, 100000),
+                                   markers = markers, 
+                                   anno_info = gene_anno_info,
+                                   txdb = txdb)
+saveRDS(anno_tables, file.path("~", "Github", "pvdiv-phenology-gxe", 
+                               "large-files", "T_GR_FL", 
+                               paste0("Annotation_tables_", 
+                                      "Time_greenup_to_flowering_date_mash_", 
+                                      "7_sites_9_PCs.rds")))
+nbycond <- mash_plot_sig_by_condition(m = mash_output, saveoutput = TRUE)
+mashhattan <- mash_plot_manhattan_by_condition(m = mash_output, saveoutput = TRUE)
+
+mhtn <- mashhattan$ggmanobject + scale_color_viridis(option = "B",  end = 0.8) + scale_fill_viridis(option = "B", end = 0.8)
+
+save_plot("Manhattan_T_GR_FL.png", plot = last_plot(), base_height = 4, base_asp = 2.5)
+
+pairwise_plot <- mash_plot_pairwise_sharing(effectRDS = "Pairwise_sharing_Strong_Effects_26989SNPs.rds", reorder = TRUE, saveoutput = TRUE)
+save_plot("Pairwise_test.png", plot = pairwise_plot$gg_corr, base_height = 2.5, base_asp = 1.618)
+```
+
+# FL50
+
+```{r}
+colremove <- names(which(colSums(!is.na(phenotypes_wide_2019)) < 200))
+
+phe_gwas_fl50 <- phenotypes_wide_2019 %>%
+  dplyr::select(-colremove) %>%
+  dplyr::select(PLANT_ID, ends_with("FL50"))
+
+txdb <- loadDb(file = file.path("/", "home", "alice", "pvdiv", "switchgrassGWAS", 
+                                "data",
+                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
+
+anno_info <-
+  read_delim(file = file.path("/", "home", "alice",  "pvdiv", "switchgrassGWAS", 
+                              "data",
+                              "Pvirgatum_516_v5.1.annotation_info.txt"),
+             col_names = TRUE, delim = "\t")
+gene_anno_info <- tbl_df(anno_info) %>%
+  distinct(locusName, .keep_all = TRUE)
+
+svd <- readRDS(file.path("/", "home", "alice", "Github", "pvdiv-phenology-gxe",
+                         "large-files", "SVD_upland_lowland_continuous_k20.rds"))
+
+snp <- snp_attach(file.path("~", "pvdiv", "VCF_and_Numeric", "Pvirgatum_4x_784g_imp_phased_up_low_continuous_subset_maf0.05.rds"))
+
+  G <- snp$genotypes
+  CHR <- snp$map$chromosome
+  POS <- snp$map$physical.pos
+  plants <- snp$fam$sample.ID
+  bonferroni <- -log10(0.05/length(POS))
+  NCORES <- nb_cores()
+
+  CHRN <- enframe(CHR, name = NULL) %>%
+    dplyr::rename(CHR = .data$value) %>%
+    mutate(CHRN = case_when(.data$CHR == "Chr01K" ~ 1,
+                            .data$CHR == "Chr01N" ~ 2,
+                            .data$CHR == "Chr02K" ~ 3,
+                            .data$CHR == "Chr02N" ~ 4,
+                            .data$CHR == "Chr03K" ~ 5,
+                            .data$CHR == "Chr03N" ~ 6,
+                            .data$CHR == "Chr04K" ~ 7,
+                            .data$CHR == "Chr04N" ~ 8,
+                            .data$CHR == "Chr05K" ~ 9,
+                            .data$CHR == "Chr05N" ~ 10,
+                            .data$CHR == "Chr06K" ~ 11,
+                            .data$CHR == "Chr06N" ~ 12,
+                            .data$CHR == "Chr07K" ~ 13,
+                            .data$CHR == "Chr07N" ~ 14,
+                            .data$CHR == "Chr08K" ~ 15,
+                            .data$CHR == "Chr08N" ~ 16,
+                            .data$CHR == "Chr09K" ~ 17,
+                            .data$CHR == "Chr09N" ~ 18,
+                            TRUE ~ 19
+    ))
+  markers <- tibble(CHR = CHR, POS = POS)
+
+  plants <- plants %>% enframe(value="PLANT_ID") %>% dplyr::select(-name)
+  plants %>%
+    left_join(phe_gwas_fl50) %>% arrange(PLANT_ID)
+
+  df_phe <- plants %>%
+    left_join(phe_gwas_fl50)
+
+  gwas <- pvdiv_gwas(df = df_phe, type = "linear", snp = snp, ncores = NCORES, 
+                     covar = svd, npcs = 9)
+
+```
+
+# Mash FL50
+```{r}
+load(file = "../markers.rda")
+load(file = "../markers_unlinked.rda")
+gwas_rds <- pvdiv_results_in_folder(path = ".", pattern = "GWAS_object_") # 8 phenotypes from 4 sites - 32 phenotypes total with > 200 individuals phenotyped and sequenced.
+
+phenotype_names <- str_sub(gwas_rds, start = 13, end = -5)
+
+numSNPs <- ceiling(1000000/length(phenotype_names)^2)
+numSNPs <- 20000 # the above gave 20409, so just rounded this to 20000.
+
+mash_input <- pvdiv_bigsnp2mashr(path = ".", gwas_rds = gwas_rds, 
+                           phenotypes = phenotype_names, numSNPs = numSNPs, 
+                           markers = markers, markers2 = markers2, G = G, 
+                           model = "linear", saveoutput = TRUE)
+
+mash_output <- mash_standard_run(path = ".", list_input = mash_input, 
+                                  numSNPs = numSNPs, saveoutput = TRUE)
+
+anno_tables <- pvdiv_table_topsnps(df = mash_output, type = "mash", 
+                                   n = c(10,100), FDRalpha = NA, 
+                                   rangevector = c(0, 20000, 100000),
+                                   markers = markers, 
+                                   anno_info = gene_anno_info,
+                                   txdb = txdb)
+saveRDS(anno_tables, file.path("~", "Github", "pvdiv-phenology-gxe", 
+                               "large-files", "T_GR_FL", 
+                               paste0("Annotation_tables_", 
+                                      "Flowering_date_mash_", 
+                                      "7_sites_9_PCs.rds")))
+```
+
+```{r}
+anno_tables <- readRDS("../PAG/FL50/Annotation_tables_Flowering_date_mash_7_sites_9_PCs.rds")
+mash_output <- readRDS(file = "../PAG/FL50/Strong_Effects29475SNPs.rds")
+anno_tables$top100SNPs_within0bp %>%
+  arrange(desc(log10BayesFactor))
+
+nbycond <- mash_plot_sig_by_condition(m = mash_output, saveoutput = TRUE)
+mashhattan <- mash_plot_manhattan_by_condition(m = mash_output)
+
+mhtn <- mashhattan$ggmanobject + scale_color_viridis(option = "B",  end = 0.8) + scale_fill_viridis(option = "B", end = 0.8)
+
+save_plot("Manhattan_FL50.png", plot = last_plot(), base_height = 4, base_asp = 2.5)
+
+pairwise_plot <- mash_plot_pairwise_sharing(effectRDS = "../PAG/FL50/Pairwise_sharing_Strong_Effects_29475SNPs.rds", reorder = FALSE)
+save_plot("Pairwise_FL50_unordered.svg", plot = pairwise_plot$gg_corr, base_height = 2.5, base_asp = 1.618)
+
+pairwise_plot2 <- mash_plot_pairwise_sharing(effectRDS = "../PAG/T_GR_FL/Pairwise_sharing_Strong_Effects_26989SNPs.rds", reorder = FALSE)
+save_plot("Pairwise_T_GR_FL_unordered.svg", plot = pairwise_plot2$gg_corr, base_height = 2.5, base_asp = 1.618)
+
+pairwise_plot3 <- mash_plot_pairwise_sharing(effectRDS = "../PAG/GR50/Pairwise_sharing_Strong_Effects_31355SNPs.rds", reorder = FALSE)
+save_plot("Pairwise_GR50_unordered.svg", plot = pairwise_plot3$gg_corr, base_height = 2.5, base_asp = 1.618)
+
+library(ggcorrplot)
+ggcorrplot(pairwise_plot$corr_matrix[c(4,6,8,7,2,5,3,1),c(4,6,8,7,2,5,3,1)])
+save_plot("Pairwise_FL50_latitude.svg", plot = last_plot(), base_height = 2.5, base_asp = 1.618)
+
+ggcorrplot(pairwise_plot2$corr_matrix[c(5,7,6,2,4,3,1),c(5,7,6,2,4,3,1)])
+save_plot("Pairwise_T_GR_FL_latitude.svg", plot = last_plot(), base_height = 2.5, base_asp = 1.618)
+
+ggcorrplot(pairwise_plot3$corr_matrix[c(5,8,10,7,9,2,6,3,4,1),c(5,8,10,7,9,2,6,3,4,1)])
+save_plot("Pairwise_GR50_latitude.svg", plot = last_plot(), base_height = 2.5, base_asp = 1.618)
+
+```
+
 # Mash
@@
 ```{r}
+svd <- readRDS(file.path("/", "home", "alice", "Github", "pvdiv-phenology-gxe",
+                         "large-files", "SVD_upland_lowland_continuous_k20.rds"))
+
+snp <- snp_attach(file.path("~", "pvdiv", "VCF_and_Numeric", "Pvirgatum_4x_784g_imp_phased_up_low_continuous_subset_maf0.05.rds"))
+
+  G <- snp$genotypes
+  CHR <- snp$map$chromosome
+  POS <- snp$map$physical.pos
+  plants <- snp$fam$sample.ID
+
+markers <- tibble(CHR = CHR, POS = POS)
+
 markers2sub <- attr(svd, "subset")
@@
 
+save(markers, file = "markers.rda")
+save(markers2, file = "markers_unlinked.rda")
+
+```
+
+```{r}
+load(file = "markers.rda")
+load(file = "markers_unlinked.rda")
 gwas_rds <- pvdiv_results_in_folder(path = ".", pattern = "GWAS_object_") # 8 phenotypes from 4 sites - 32 phenotypes total with > 200 individuals phenotyped and sequenced.

### 2020-01-08 analysis/1_bigsnpr_gwas_prep.Rmd
Slide 18 edits for pairwise mash results

@@
 save_plot("Manhattan_FL50.png", plot = last_plot(), base_height = 4, base_asp = 2.5)
+```
+
+# Pairwise plots GR FL
+```{r}
 
@@
 
-library(ggcorrplot)
-ggcorrplot(pairwise_plot$corr_matrix[c(4,6,8,7,2,5,3,1),c(4,6,8,7,2,5,3,1)])
-save_plot("Pairwise_FL50_latitude.svg", plot = last_plot(), base_height = 2.5, base_asp = 1.618)
+library(GGally)
+pairplotFL <- mash_plot_pairwise_sharing(corrmatrix = pairwise_plot$corr_matrix[c(4,6,8,7,2,5,3,1),c(4,6,8,7,2,5,3,1)], reorder = FALSE)
 
-ggcorrplot(pairwise_plot2$corr_matrix[c(5,7,6,2,4,3,1),c(5,7,6,2,4,3,1)])
-save_plot("Pairwise_T_GR_FL_latitude.svg", plot = last_plot(), base_height = 2.5, base_asp = 1.618)
+corrplot1 <- ggcorr(data = NULL, cor_matrix = pairplotFL$corr_matrix, 
+                       label = FALSE, label_alpha = TRUE,
+                       label_size = 3, hjust = 0.95, vjust = 0.3,
+                       layout.exp = 9, min_size = 0,
+                       max_size = 3.5, limits = c(0,1), low = "#000077", mid = "#EEEEEE", high = "#FF5500")
+save_plot("Pairwise_FL50_latitude_orange.svg", plot = corrplot1, base_height = 2.5, base_asp = 1.618)
 
-ggcorrplot(pairwise_plot3$corr_matrix[c(5,8,10,7,9,2,6,3,4,1),c(5,8,10,7,9,2,6,3,4,1)])
-save_plot("Pairwise_GR50_latitude.svg", plot = last_plot(), base_height = 2.5, base_asp = 1.618)
+ggcorr(data = NULL, cor_matrix = pairwise_plot2$corr_matrix[c(5,7,6,2,4,3,1),
+                                                            c(5,7,6,2,4,3,1)], 
+       label = FALSE, label_alpha = TRUE, label_size = 3, hjust = 0.95, 
+       vjust = 0.3, layout.exp = 9, min_size = 0, max_size = 3.5, 
+       limits = c(0,1), low = "#000077", mid = "#EEEEEE", high = "#FF5500")
+save_plot("Pairwise_T_GR_FL_latitude_orange.svg", plot = last_plot(), base_height = 2.5, base_asp = 1.618)
+
+ggcorr(data = NULL, cor_matrix = pairwise_plot3$corr_matrix[c(5,8,10,9,2,6,4,1),
+                                                            c(5,8,10,9,2,6,4,1)], 
+       label = FALSE, label_alpha = TRUE, label_size = 3, hjust = 0.95, 
+       vjust = 0.3, layout.exp = 9, min_size = 0, max_size = 3.5, 
+       limits = c(0,1), low = "#000077", mid = "#EEEEEE", high = "#FF5500")
+save_plot("Pairwise_GR50_latitude_orange.svg", plot = last_plot(), base_height = 2.5, base_asp = 1.618)
 

### 2020-01-21 analysis/FL50/Overlap_with_4way_QTL.Rmd
Start analysis of diversity panel and 4way overlaps

---
title: "FL50_4way_overlap"
author: "Alice MacQueen"
date: "1/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(switchgrassGWAS)
library(cowplot)
library(viridis)
```

##

```{r}
mash <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/FL50/Strong_Effects29475SNPs.rds")
```

```{r}
df <- get_significant_results(m=mash) %>%
  enframe(name = "SNP")
df1 <- switchgrassGWAS:::get_marker_df(mash)
df1 <- df1 %>%
  mutate(log10bf = switchgrassGWAS:::get_log10bf(m=mash),
         sigcond = switchgrassGWAS:::get_n_significant_conditions(mash)
         )
switchgrassGWAS:::get_log10bf(m=mash)
switchgrassGWAS::mash_plot_effects(m=mash, i = 26711)

df1 %>%
  filter(log10bf > 5)

panela <- df1 %>%
  filter(grepl("Chr09K_5", Marker) & value > 26500) %>%
  separate(Marker, into = c("Chr", "_", "Position"), sep = c(6,7), remove = FALSE) %>%
  mutate(Position = as.numeric(Position)) %>%
  ggplot(aes(x=Position, y = log10bf)) +
  geom_point(aes(color = sigcond)) +
  scale_color_viridis(option = "B") # 26711
save_plot(filename = "Chr09K_FL50_mash_result_zoom.png", plot = last_plot(), base_asp = 1.5)

p3 <- plot_grid(panela, panelb$ggobject, rel_widths = c(2,1))
save_plot(filename = "Chr09K_FL50_mash_result_zoom_2panel.png", plot = p3, base_asp = 3)
```
#Chr05N
```{r}
panela <- df1 %>%
  filter(grepl("Chr05N_1", Marker) & value > 15717) %>%
  separate(Marker, into = c("Chr", "_", "Position"), sep = c(6,7), remove = FALSE) %>%
  mutate(Position = as.numeric(Position)) %>%
  ggplot(aes(x=Position, y = log10bf)) +
  geom_point(aes(color = sigcond)) +
  scale_color_viridis(option = "B", end = 0.9) 
panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 15864)
panelb$ggobject 
p3 <- plot_grid(panela, panelb$ggobject, rel_widths = c(2,1))
save_plot(filename = "Chr05N_FL50_mash_result_zoom.png", plot = p3, base_asp = 3)

```
Chr05N_15045423	15847	4.688166	8	  affects flowering at the 3 major sites
Chr05N_15048576	15849	4.101891	8	
Chr05N_15951722	15863	4.048425	7	  affects flowering at KBSM & CLMB
Chr05N_16084715	15864	4.226715	8	  affects flowering at the 3 major sites

```{r}
panela <- df1 %>%
  filter(grepl("Chr04K_1", Marker) & value > 11158) %>%
  separate(Marker, into = c("Chr", "_", "Position"), sep = c(6,7), remove = FALSE) %>%
  mutate(Position = as.numeric(Position)) %>%
  ggplot(aes(x=Position, y = log10bf)) +
  geom_point(aes(color = sigcond)) +
  scale_color_viridis(option = "B", end = 0.9) 
panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11493)
panelb$ggobject 
p3 <- plot_grid(panela, panelb$ggobject, rel_widths = c(2,1))
save_plot(filename = "Chr04K_FL50_mash_result_zoom.png", plot = p3, base_asp = 3)
df1 %>%
  filter(grepl("Chr04K_1", Marker) & value > 11158 & log10bf > 4)
```
Chr04K_14442056	11332	4.954623	7	
Chr04K_14919129	11382	4.263826	8	
Chr04K_14920267	11384	4.655388	8	
Chr04K_14955235	11388	4.693233	7	
Chr04K_16627327	11460	4.240948	7	
Chr04K_16687695	11493	4.405048	7	
Chr04K_18870690	11610	5.308748	6

### 2020-01-22 analysis/common-garden-allele-balancing/1_VCF_file_filtering_by_common_garden_site.Rmd
Site by year subsets for common garden allele balancing

---
title: "VCF file filtering by common garden site"
author: "Alice MacQueen"
date: "1/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(switchgrassGWAS)
load("~/Github/pvdiv-phenology-gxe/data/phenotypes_2019.rda")
load("~/Github/pvdiv-phenology-gxe/data/sites.rda")
```

## Goals & Outline

    1. Use the 33.9 million SNP file, located at:
    `/home/SharedUser/PV/NatPop/All_chr_combined_636g_80pct_maf0.005.vcf.bz2`
      Convert this file to vcf.gz first.
    2. Use or make a data frame of PLANT_ID's that were present at each site in 2019.
    3. Subset the 33.9 million SNP file to 5% MAF for each site.
    4. Use vcftools to find allele frequencies for each SNP at each site.

Below are the steps necessary to make a subset of a VCF file for use for GWAS.

## Subset function

```{r}
pvdiv_pop_subset_bigsnp <- function(subset, selector){
  subset_name <- paste0("Pvirgatum_4x_784g_imp_phased_", subset, "_subset")
  snpfile_subset <- subset(snp, ind.row = selector)
  snp_subset <- snp_attach(snpfile_subset)
  snp_writeBed(snp_subset, bedfile = paste0(subset_name, ".bed"))
  mac <- round(25/length(selector), digits = 2) # what should the MAF be?
  if(mac > 0.05){
    mac <- 0.05
  }
  plink <- download_plink()
  bedfile <- snp_plinkQC(plink, prefix.in = subset_name,
                         maf = mac, geno = 0.1, hwe = 1e-50, 
                         prefix.out = paste0(subset_name, "_maf", mac),
                         extra.options = "--allow-no-sex --allow-extra-chr --chr Chr01K Chr01N Chr02K Chr02N Chr03K Chr03N Chr04K Chr04N Chr05K Chr05N Chr06K Chr06N Chr07K Chr07N Chr08K Chr08N Chr09K Chr09N")
  rdsfile <- snp_readBed(bedfile)
}

# Load data - if you get errors, make sure to run R from the GWAS/ folder in pubjuenger
snp <- snp_attach("/home/pubjuenger/GWAS/Pvirgatum_4x_784g_imp_phased_maf0.005.rds")
switchgrassGWAS::phenotypes

# Use phenotypes to set up selection criteria for subsets. These need to be vectors of the row numbers in the Taxa dataframe that you have phenotypes for.
gwasct_sel <- which(!is.na(phenotypes$GWAS_CT)) # gwas_ct GWAS

## Run function to generate subset SNP files for the subset
pvdiv_pop_subset_bigsnp(subset = "gwasct", selector = gwasct_sel)

```

## 2. Use or make a data frame of PLANT_ID's that were present at each site in 2019.

Plants alive EOS 2019 (higher of HT_PAN_EOS and TC_EOS):
BRKG: 254
CLMB: 496
FRMI: 185
KBSM: 516
KING: 286
LINC: 260
OVTN: 331
PKLE: 614
STIL: 451
TMPL: 373

For each site: # SNPs above a MAF of 5%
% SNPs above a MAF of 5% (at all 10 sites), # SNPs > 5% MAF at all 10 sites, distribution of these SNPs across the chromosomes
Number of plants alive from each subpopulation (sort of did this already)

Average, minimum, maximum, standard deviation, coefficient of variation of MAF across all 10 sites
histograms of these values across all SNPs

Compare these values to their values in 2018 to see how much frequencies shifted in the last year?

```{r}
colSums(!is.na(phenotypes_wide_2019))

for(i in 1:10){
site <- sites$Site[i]

site_2019 <- phenotypes_wide_2019 %>%
  pivot_longer(cols = BRKG_EM1:TMPL_TC_EOS, names_to = "SITE_PHE", values_to = "Measurement") %>%
  separate(col = SITE_PHE, into = c("SITE", "_", "PHE"), sep = c(4,5)) %>%
  filter(SITE == site & !is.na(Measurement)) %>%
  group_by(PLANT_ID) %>%
  summarise(count = n())

writeLines(site_2019$PLANT_ID, con = paste0(site, "_2019"))
}

for(i in 1:10){
site <- sites$Site[i]
site_2018 <- phenotypes_2018 %>%
  pivot_longer(cols = BRKG_DEAD_2018:TMPL_TC_EOS_2018, names_to = "SITE_PHE", values_to = "Measurement") %>%
  separate(col = SITE_PHE, into = c("SITE", "_", "PHE"), sep = c(4,5)) %>%
  filter(SITE == site & !is.na(Measurement)) %>%
  group_by(PLANT_ID) %>%
  summarise(count = n()) 
writeLines(site_2018$PLANT_ID, con = paste0(site, "_2018"))
}

```

Plants alive EOS 2018, then EOS 2019:
BRKG: 412 down to 254
CLMB: 591 down to 496
FRMI: 273 down to 185
KBSM: 544 down to 516
KING: 305 down to 286
LINC: 437 down to 260
OVTN: 346 down to 331
PKLE: 623 down to 614
STIL: 469 down to 451
TMPL: 388 down to 373

# VCFtools lines
```{}
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/PKLE_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/PKLE_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/BRKG_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/BRKG_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/CLMB_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/CLMB_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/FRMI_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/FRMI_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KBSM_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KBSM_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KING_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KING_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/LINC_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/LINC_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/OVTN_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/OVTN_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/STIL_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/STIL_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/TMPL_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/TMPL_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/PKLE_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/PKLE_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/BRKG_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/BRKG_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/CLMB_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/CLMB_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/FRMI_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/FRMI_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KBSM_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KBSM_2018j &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KING_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KING_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/LINC_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/LINC_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/OVTN_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/OVTN_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/STIL_2018 --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/STIL_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/TMPL_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/TMPL_2018 &
```

### 2020-01-22 analysis/common-garden-allele-balancing/1_SNP_file_filtering_by_common_garden_site.R

@@ -1,0 +1,32 @@
+library(tidyverse)
+library(bigsnpr)
+library(switchgrassGWAS)
+
+
+pvdiv_pop_subset_bigsnp <- function(subset, selector){
+  subset_name <- paste0("Pvirgatum_4x_784g_imp_phased_", subset, "_subset")
+  snpfile_subset <- subset(snp, ind.row = selector)
+  snp_subset <- snp_attach(snpfile_subset)
+  snp_writeBed(snp_subset, bedfile = paste0(subset_name, ".bed"))
+  mac <- round(25/length(selector), digits = 2) # what should the MAF be?
+  if(mac > 0.05){
+    mac <- 0.05
+  }
+  plink <- download_plink()
+  bedfile <- snp_plinkQC(plink, prefix.in = subset_name,
+                         maf = mac, geno = 0.1, hwe = 1e-50, 
+                         prefix.out = paste0(subset_name, "_maf", mac),
+                         extra.options = "--allow-no-sex --allow-extra-chr --chr Chr01K Chr01N Chr02K Chr02N Chr03K Chr03N Chr04K Chr04N Chr05K Chr05N Chr06K Chr06N Chr07K Chr07N Chr08K Chr08N Chr09K Chr09N")
+  rdsfile <- snp_readBed(bedfile)
+}
+
+
+# Load data - if you get errors, make sure to run R from the GWAS/ folder in pubjuenger
+snp <- snp_attach("/home/pubjuenger/GWAS/Pvirgatum_4x_784g_imp_phased_maf0.005.rds")
+switchgrassGWAS::phenotypes
+
+# Use phenotypes to set up selection criteria for subsets. These need to be vectors of the row numbers in the Taxa dataframe that you have phenotypes for.
+gwasct_sel <- which(!is.na(phenotypes$GWAS_CT)) # gwas_ct GWAS
+
+## Run function to generate subset SNP files for the subset
+pvdiv_pop_subset_bigsnp(subset = "gwasct", selector = gwasct_sel)

### 2020-01-23 analysis/common-garden-allele-balancing/1_VCF_file_filtering_by_common_garden_site.Rmd
Common garden allele balancing with 33M SNP file

@@
 colSums(!is.na(phenotypes_wide_2019))
-
+i=1
 for(i in 1:10){
@@
 Plants alive EOS 2018, then EOS 2019:
-BRKG: 412 down to 254
-CLMB: 591 down to 496
-FRMI: 273 down to 185
-KBSM: 544 down to 516
-KING: 305 down to 286
-LINC: 437 down to 260
-OVTN: 346 down to 331
-PKLE: 623 down to 614
-STIL: 469 down to 451
-TMPL: 388 down to 373
+BRKG: 412 down to 254 (377 in 2018; 234 in 2019)
+CLMB: 591 down to 496 (547 in 2018; 462 in 2019)
+FRMI: 273 down to 185 (245 in 2018; 163 in 2019)
+KBSM: 544 down to 516 (501 in 2018; 480 in 2019)
+KING: 305 down to 286 (275 in 2018; 267 in 2019)
+LINC: 437 down to 260 (400 in 2018; 247 in 2019)
+OVTN: 346 down to 331 (312 in 2018; 304 in 2019)
+PKLE: 623 down to 614 (579 in 2018; 577 in 2019 out of 732 individuals in 2020 genome file)
+STIL: 469 down to 451 (433 in 2018; 419 in 2019)
+TMPL: 388 down to 373 (351 in 2018; 341 in 2019)
 
@@
 ```{}
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/PKLE_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/PKLE_2019 &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/BRKG_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/BRKG_2019 &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/CLMB_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/CLMB_2019 &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/FRMI_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/FRMI_2019 &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KBSM_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KBSM_2019 &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KING_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KING_2019 &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/LINC_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/LINC_2019 &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/OVTN_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/OVTN_2019 &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/STIL_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/STIL_2019 &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/TMPL_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/TMPL_2019 &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/PKLE_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/PKLE_2018 &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/BRKG_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/BRKG_2018 &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/CLMB_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/CLMB_2018 &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/FRMI_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/FRMI_2018 &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KBSM_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KBSM_2018j &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KING_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KING_2018 &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/LINC_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/LINC_2018 &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/OVTN_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/OVTN_2018 &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/STIL_2018 --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/STIL_2018 &
-vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/TMPL_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/TMPL_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/PKLE_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/PKLE_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/BRKG_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/BRKG_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/CLMB_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/CLMB_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/FRMI_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/FRMI_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KBSM_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KBSM_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KING_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KING_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/LINC_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/LINC_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/OVTN_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/OVTN_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/STIL_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/STIL_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/TMPL_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/TMPL_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/PKLE_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/PKLE_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/BRKG_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/BRKG_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/CLMB_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/CLMB_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/FRMI_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/FRMI_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KBSM_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KBSM_2018j &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KING_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KING_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/LINC_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/LINC_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/OVTN_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/OVTN_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/STIL_2018 --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/STIL_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/TMPL_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/TMPL_2018 &
 ```
+
+```{r}
+snpfile <- read_csv(file = "33M_SNP_file_PLANT_ID")
+pkle2019 <- read_csv(file = "PKLE_2019", col_names = "PLANT_ID")
+
+snpfile %>%
+  left_join(phenotypes_wide_2019)
+
+metadata %>%
+  filter(LIB_CLIMATE == "Y") %>%
+  anti_join(snpfile)
+
+snpfile %>%
+  anti_join(metadata) # NL94SL93 is the only one not in the metadata. So what happened?
+
+metadata2 %>%
+  anti_join(snpfile) # NL94SL93 is the only one not in the 732 metadata. Ehhhh? 732 metadata has NL94 & SL93. 
+# So the error isn't because I gave Taslima the wrong metadata for the 732 file.
+
+snpfile %>%
+  anti_join(pkle2019)
+pkle2019 %>%
+  anti_join(snpfile)
+site_2019 %>%
+  anti_join(snpfile) %>%
+  left_join(metadata)
+
+metadata %>%
+  filter(LIB_CLIMATE == "Y")
+metadata %>%
+  filter(LIB_PHENO == "Y")
+```
+NL94SL93 is the only one not in the 732 metadata. Ehhhh? 732 metadata has NL94 & SL93.
+
+So the error isn't because I gave Taslima the wrong metadata for the 732 file. But there are still 44 missing individuals that were planted in the field but aren't in the 732 genotype file.
+
+I think maybe this file has the
+
+```{r}
+
+phenotypes_wide_2019 %>%
+  anti_join(snpfile) %>%
+  dplyr::select(PLANT_ID) %>%
+  left_join(metadata) %>%
+  arrange(LIB_PHENO) %>%
+  filter(!(PLANT_ID %in% c("J661.C", "J491.A", "J385.C", "J471.A", "J243.A")))
+
+phenotypes_wide_2019 %>%
+  filter(!(PLANT_ID %in% c("J661.C", "J491.A", "J385.C", "J471.A", "J243.A", "J428.A", "J399.B", "J344.B", "J003.C", "J624.B", "J682.C", "NL94×SL93_370", "NL94×SL93_305"))) # None of these have any 2019 data.
+
+phe_2019 <- phenotypes_wide_2019[which(rowSums(!is.na(phenotypes_wide_2019))>1),] 
+# Only 630 individuals have phenotyping data in 2019. But 38 aren't in 732 file. That's 10% of data.
+phe_2018 <- phenotypes_2018[which(rowSums(!is.na(phenotypes_2018))>1),] # 636 have 2018 data.
+phe_csv <- phe_2018 %>%
+  dplyr::select(PLANT_ID) %>%
+  #anti_join(snpfile) %>%
+  left_join(metadata) %>% 
+  filter(LIB_PHENO == "Y") %>%
+  dplyr::select(PLANT_ID:LIB_PHENO) #%>% anti_join(snpfile)
+
+write_csv(phe_csv, path = "Individuals_needed_for_33M_SNPs.csv")
+```
+
+Individuals that are in the phenotyping panel and should be resequenced:
+
+J428.A	IGRB	N	N	N	 Highest heterozygoisty and low-ish coverage. Fishy, but also, probaably OK. Exclude
+J399.B	IHZG	N	N	N	low coverage, high het. Exclude
+J344.B	IHZN	N	N	N	low coverage, high het. Exclude
+J003.C	IIAH	N	N	N	J003 is a mess. Exclude
+J243.A	IRUH	N	N	N	This has really bad coverage and should be excluded. Exclude
+J661.C	IRXF	N	N	N		Probable Error, Plant does not match others. Exclude
+J682.C	ITCI	N	N N Very high heterozygosity, coverage on the low side. Exclude

### 2020-01-24 analysis/common-garden-allele-balancing/2_allele_frequencies_common_gardens.Rmd
MAF Analysis and Plots for Common Garden Allele Balancing

---
title: "Allele Frequencies Common Gardens"
author: "Alice MacQueen"
date: "1/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(switchgrassGWAS)
library(viridis)
library(cowplot)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

##

```{r}
load("~/Github/pvdiv-phenology-gxe/data/sites.rda")
```

```{r}
year <- "2019"
i=1
k= i+1
nmax = 33890000

site_year <- read_delim(paste0(sites$Site[i],"_", year, ".frq"), delim = "\t", 
                        n_max = nmax, col_types = "cdddcc", skip = 1, 
                        col_names=c("CHR", "POS", "N_ALLELES", 
                                    paste0("N_CHR_", sites$Site[i],"_", year),
                                    paste0("ALLELEFREQ1_", sites$Site[i],"_", year),
                                    paste0("ALLELEFREQ2_", sites$Site[i],"_", year))) %>%
  separate(col = paste0("ALLELEFREQ1_", sites$Site[i],"_", year), 
           into = c("MAJOR",
                    paste0("MAJF_", sites$Site[i],"_", year)), sep = ":", 
           convert = TRUE) %>%
  separate(col = paste0("ALLELEFREQ2_", sites$Site[i],"_", year), 
           into = c("MINOR",
                    paste0("MINF_", sites$Site[i],"_", year)), sep = ":", 
           convert = TRUE)
site_year2 <- read_delim(paste0(sites$Site[k],"_", year, ".frq"), delim = "\t", 
                        n_max = nmax, col_types = "cdddcc", skip = 1, 
                        col_names=c("CHR", "POS", "N_ALLELES", 
                                    paste0("N_CHR_", sites$Site[k],"_", year),
                                    paste0("ALLELEFREQ1_", sites$Site[k],"_", year),
                                    paste0("ALLELEFREQ2_", sites$Site[k],"_", year))) %>%
  separate(col = paste0("ALLELEFREQ1_", sites$Site[k],"_", year), 
           into = c("MAJOR",
                    paste0("MAJF_", sites$Site[k],"_", year)), sep = ":", 
           convert = TRUE) %>%
  separate(col = paste0("ALLELEFREQ2_", sites$Site[k],"_", year), 
           into = c("MINOR",
                    paste0("MINF_", sites$Site[k],"_", year)), sep = ":", 
           convert = TRUE)
freq_sitebyyear <- site_year %>%
  full_join(site_year2, by = c("CHR", "POS", "N_ALLELES", "MAJOR", "MINOR"))

for(k in 3:10){
  site_year2 <- read_delim(paste0(sites$Site[k],"_", year, ".frq"), delim = "\t", 
                        n_max = nmax, col_types = "cdddcc", skip = 1, 
                        col_names=c("CHR", "POS", "N_ALLELES", 
                                    paste0("N_CHR_", sites$Site[k],"_", year),
                                    paste0("ALLELEFREQ1_", sites$Site[k],"_", year),
                                    paste0("ALLELEFREQ2_", sites$Site[k],"_", year))) %>%
  separate(col = paste0("ALLELEFREQ1_", sites$Site[k],"_", year), 
           into = c("MAJOR",
                    paste0("MAJF_", sites$Site[k],"_", year)), sep = ":", 
           convert = TRUE) %>%
  separate(col = paste0("ALLELEFREQ2_", sites$Site[k],"_", year), 
           into = c("MINOR",
                    paste0("MINF_", sites$Site[k],"_", year)), sep = ":", 
           convert = TRUE)
  freq_sitebyyear <- freq_sitebyyear %>%
    full_join(site_year2, by = c("CHR", "POS", "N_ALLELES", "MAJOR", "MINOR"))
}
year <- "2018"
for(k in 1:10){
  site_year2 <- read_delim(paste0(sites$Site[k],"_", year, ".frq"), delim = "\t", 
                        n_max = nmax, col_types = "cdddcc", skip = 1, 
                        col_names=c("CHR", "POS", "N_ALLELES", 
                                    paste0("N_CHR_", sites$Site[k],"_", year),
                                    paste0("ALLELEFREQ1_", sites$Site[k],"_", year),
                                    paste0("ALLELEFREQ2_", sites$Site[k],"_", year))) %>%
  separate(col = paste0("ALLELEFREQ1_", sites$Site[k],"_", year), 
           into = c("MAJOR",
                    paste0("MAJF_", sites$Site[k],"_", year)), sep = ":", 
           convert = TRUE) %>%
  separate(col = paste0("ALLELEFREQ2_", sites$Site[k],"_", year), 
           into = c("MINOR",
                    paste0("MINF_", sites$Site[k],"_", year)), sep = ":", 
           convert = TRUE)
  freq_sitebyyear <- freq_sitebyyear %>%
    full_join(site_year2, by = c("CHR", "POS", "N_ALLELES", "MAJOR", "MINOR"))
}
saveRDS(freq_sitebyyear, file = paste0("Allele_frequencies_site_by_year_", nmax, "SNPs.rds"))
```

```{r}
freq_sitebyyear <- readRDS("Allele_frequencies_site_by_year_100000SNPs.rds")
freq_sitebyyear
```

```{r}
sites
freq_sitebyyear %>%
  group_by(between(MINF_PKLE_2018, 0.05, 0.95)) %>% tally()
freq_sitebyyear %>%
  filter(between(MINF_BRKG_2018, 0.05, 0.95) & 
           between(MINF_PKLE_2018, 0.05, 0.95) &
           between(MINF_FRMI_2018, 0.05, 0.95) & 
           between(MINF_KING_2018, 0.05, 0.95) & 
           between(MINF_OVTN_2018, 0.05, 0.95) & 
           between(MINF_KBSM_2018, 0.05, 0.95) & 
           between(MINF_LINC_2018, 0.05, 0.95) & 
           between(MINF_STIL_2018, 0.05, 0.95) & 
           between(MINF_TMPL_2018, 0.05, 0.95) & 
           between(MINF_CLMB_2018, 0.05, 0.95)) 
freq_sitebyyear %>%
  group_by(between(MINF_BRKG_2018, 0.05, 0.95), 
           between(MINF_PKLE_2018, 0.05, 0.95),
           between(MINF_FRMI_2018, 0.05, 0.95), 
           between(MINF_KING_2018, 0.05, 0.95), 
           between(MINF_OVTN_2018, 0.05, 0.95),  
           between(MINF_KBSM_2018, 0.05, 0.95),  
           between(MINF_LINC_2018, 0.05, 0.95),  
           between(MINF_STIL_2018, 0.05, 0.95),  
           between(MINF_TMPL_2018, 0.05, 0.95),  
           between(MINF_CLMB_2018, 0.05, 0.95),
           between(MINF_BRKG_2019, 0.05, 0.95), 
           between(MINF_PKLE_2019, 0.05, 0.95),
           between(MINF_FRMI_2019, 0.05, 0.95), 
           between(MINF_KING_2019, 0.05, 0.95), 
           between(MINF_OVTN_2019, 0.05, 0.95),  
           between(MINF_KBSM_2019, 0.05, 0.95),  
           between(MINF_LINC_2019, 0.05, 0.95),  
           between(MINF_STIL_2019, 0.05, 0.95),  
           between(MINF_TMPL_2019, 0.05, 0.95),  
           between(MINF_CLMB_2019, 0.05, 0.95)) %>% 
  tally() %>% arrange(desc(n)) %>%
  mutate(frac = n/100000)# %>%
  #write_csv(., path = "MAF_allows_gwas_10_sites_2_years.csv")
freq_sitebyyear %>%
  group_by(between(MINF_BRKG_2018, 0.05, 0.95), 
           between(MINF_PKLE_2018, 0.05, 0.95),
           between(MINF_FRMI_2018, 0.05, 0.95), 
           between(MINF_KING_2018, 0.05, 0.95), 
           between(MINF_OVTN_2018, 0.05, 0.95),  
           between(MINF_KBSM_2018, 0.05, 0.95),  
           between(MINF_LINC_2018, 0.05, 0.95),  
           between(MINF_STIL_2018, 0.05, 0.95),  
           between(MINF_TMPL_2018, 0.05, 0.95),  
           between(MINF_CLMB_2018, 0.05, 0.95)) %>% 
  tally() %>% arrange(desc(n)) %>%
  mutate(approx = n/100000*33880000) %>%
  write_csv(., path = "MAF_allows_gwas_10_sites_2018.csv")

freq_sitebyyear %>%
  filter(between(MINF_BRKG_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_BRKG_2018, y = MINF_BRKG_2019)) + 
  geom_point(alpha = 0.05) + geom_abline(slope = 1, intercept = 0, color = "red")

36007-28544
28544/36007
36007/100000*33880000
28544/100000*33880000
```

# 36007 out of 100k sites had MAF > 0.05 at Pickle in 2018, our best site.
12.2M SNPs total.
# 28544 out of 100k sites have MAF > 0.05 at all 10 locations.
That's 7463 SNPs less, or 79.27% of the SNPs we would have had.
9.67M SNPs total. Lose 20% of the SNPs by constraining to 5% MAF at all sites.

BRKG 35452

```{r}
freq_sitebyyear %>%
  filter(between(MINF_TMPL_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_TMPL_2018, y = MINF_TMPL_2019)) + 
  geom_point(alpha = 0.05) + geom_abline(slope = 1, intercept = 0, color = "red")
freq_sitebyyear %>%
  filter(between(MINF_LINC_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_LINC_2018, y = MINF_LINC_2019)) + 
  geom_point(alpha = 0.05) + geom_abline(slope = 1, intercept = 0, color = "red")
freq_sitebyyear %>%
  filter(between(MINF_CLMB_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_CLMB_2018, y = MINF_CLMB_2019)) + 
  geom_point(alpha = 0.05) + geom_abline(slope = 1, intercept = 0, color = "red")
freq_sitebyyear %>%
  filter(between(MINF_BRKG_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_BRKG_2018, y = MINF_BRKG_2019)) + 
  geom_point(alpha = 0.05) + geom_abline(slope = 1, intercept = 0, color = "red")
freq_sitebyyear %>%
  ggplot(aes(x = MINF_BRKG_2019, y = MINF_KING_2019)) + 
  geom_point(alpha = 0.05) + geom_abline(slope = 1, intercept = 0, color = "red")
```

```{r}
freq_sitebyyear %>%
  filter(between(MINF_BRKG_2018, 0.05, 0.95) & between(MINF_KING_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_BRKG_2018, y = MINF_KING_2018)) + 
  geom_hex() + geom_abline(slope = 1, intercept = 0, color = "red") + 
  scale_fill_viridis(option = "B", trans = "log2") 
freq_sitebyyear %>%
  filter(between(MINF_BRKG_2019, 0.05, 0.95) & between(MINF_KING_2019, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_BRKG_2019, y = MINF_KING_2019)) + 
  geom_hex() + geom_abline(slope = 1, intercept = 0, color = "red") + 
  scale_fill_viridis(option = "B", trans = "log2") 
freq_sitebyyear %>%
  filter(between(MINF_LINC_2019, 0.05, 0.95) & between(MINF_PKLE_2019, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_LINC_2019, y = MINF_PKLE_2019)) + 
  geom_hex() + geom_abline(slope = 1, intercept = 0, color = "red") + 
  scale_fill_viridis(option = "B", trans = "log2") 
```

```{r}
freq_sitebyyear %>%
  filter(between(MINF_BRKG_2018, 0.05, 0.95) & between(MINF_BRKG_2019, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_BRKG_2018, y = MINF_BRKG_2019)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.43)) + 
  xlab("MAF in Brookings, 2018") +
  ylab("MAF in Brookings, 2019")

freq_sitebyyear %>%
  filter(between(MINF_KING_2019, 0.05, 0.95) & between(MINF_BRKG_2019, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_KING_2019, y = MINF_BRKG_2019)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.43)) + 
  xlab("MAF in Kingsville, 2019") +
  ylab("MAF in Brookings, 2019")

freq_sitebyyear %>%
  filter(between(MINF_OVTN_2019, 0.05, 0.95) & between(MINF_BRKG_2019, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_OVTN_2019, y = MINF_BRKG_2019)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.43)) + 
  xlab("MAF in Overton, 2019") +
  ylab("MAF in Brookings, 2019")

freq_sitebyyear %>%
  filter(between(MINF_OVTN_2018, 0.05, 0.95) & between(MINF_BRKG_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_OVTN_2018, y = MINF_BRKG_2018)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.43)) + 
  xlab("MAF in Overton, 2018") +
  ylab("MAF in Brookings, 2018")

freq_sitebyyear %>%
  filter(between(MINF_OVTN_2018, 0.05, 0.95) & between(MINF_LINC_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_OVTN_2018, y = MINF_LINC_2018)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.43)) + 
  xlab("MAF in Overton, 2018") +
  ylab("MAF in Lincoln, 2018")
```

# Plots

```{r}
maf_long <- freq_sitebyyear %>%
  dplyr::select(-starts_with("MAJF"), -starts_with("N_CHR_")) %>%
  pivot_longer(cols = starts_with("MINF"), names_to = "SITE_YEAR", values_to = "MINF") %>%
  separate(col = "SITE_YEAR", into = c("MINF_", "SITE", "YEAR"), sep = "_")

```

### 2020-01-24 analysis/common-garden-allele-balancing/2_allele_frequencies_common_gardens.Rmd
Clarify analysis in Rmd file

@@
 
-##
+# Join data on server
 
@@
 load("~/Github/pvdiv-phenology-gxe/data/sites.rda")
-```
-
-```{r}
 year <- "2019"
@@
 
+## Load data
+
 ```{r}
@@
 ```
+
+## Tally SNPs by site
 
@@
 
-# 36007 out of 100k sites had MAF > 0.05 at Pickle in 2018, our best site.
-12.2M SNPs total.
-# 28544 out of 100k sites have MAF > 0.05 at all 10 locations.
+    *36007 out of 100k sites had MAF > 0.05 at Pickle in 2018, our best site.
+    *12.2M SNPs total.
+    *28544 out of 100k sites have MAF > 0.05 at all 10 locations.
+
 That's 7463 SNPs less, or 79.27% of the SNPs we would have had.
@@
 BRKG 35452
+
+# ------------------
+# Start of Plots
 
@@
 ```
+
+## Labeled hex plots
 
@@
 
-# Plots
+## Grid of plots?
 

### 2020-01-25 analysis/common-garden-allele-balancing/1_VCF_file_filtering_by_common_garden_site.Rmd
Plots and Writeup for Common Garden Allele Balancing

@@
 load("~/Github/pvdiv-phenology-gxe/data/phenotypes_2019.rda")
+load("~/Github/pvdiv-phenology-gxe/data/phenotypes_2018.rda")
 load("~/Github/pvdiv-phenology-gxe/data/sites.rda")
@@
 colSums(!is.na(phenotypes_wide_2019))
-i=1
+i=7
 for(i in 1:10){
@@
 ```{}
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/PKLE_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/PKLE_2019 &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/BRKG_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/BRKG_2019 &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/CLMB_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/CLMB_2019 &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/FRMI_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/FRMI_2019 &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KBSM_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KBSM_2019 &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KING_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KING_2019 &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/LINC_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/LINC_2019 &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/OVTN_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/OVTN_2019 &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/STIL_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/STIL_2019 &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/TMPL_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/TMPL_2019 &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/PKLE_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/PKLE_2018 &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/BRKG_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/BRKG_2018 &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/CLMB_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/CLMB_2018 &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/FRMI_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/FRMI_2018 &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KBSM_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KBSM_2018j &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KING_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KING_2018 &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/LINC_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/LINC_2018 &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/OVTN_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/OVTN_2018 &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/STIL_2018 --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/STIL_2018 &
-vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/TMPL_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/TMPL_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/PKLE_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/PKLE_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/BRKG_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/BRKG_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/CLMB_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/CLMB_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/FRMI_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/FRMI_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KBSM_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KBSM_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KING_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KING_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/LINC_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/LINC_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/OVTN_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/OVTN_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/STIL_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/STIL_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/TMPL_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/TMPL_2019 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/PKLE_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/PKLE_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/BRKG_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/BRKG_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/CLMB_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/CLMB_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/FRMI_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/FRMI_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KBSM_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KBSM_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KING_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KING_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/LINC_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/LINC_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/OVTN_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/OVTN_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/TMPL_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/TMPL_2018 &
+vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_GWAS_630g_33dot9M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/STIL_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/STIL_2018 &
 ```

### 2020-01-25 analysis/common-garden-allele-balancing/2_allele_frequencies_common_gardens.Rmd
Plots and Writeup for Common Garden Allele Balancing

@@
 ```
+
+# ---------------------------
+## Full SNP file analyses
+
+```{r}
+
+freq_sitebyyear %>%
+  group_by(between(MINF_BRKG_2018, 0.05, 0.95), 
+           between(MINF_PKLE_2018, 0.05, 0.95),
+           between(MINF_FRMI_2018, 0.05, 0.95), 
+           between(MINF_KING_2018, 0.05, 0.95), 
+           between(MINF_OVTN_2018, 0.05, 0.95),  
+           between(MINF_KBSM_2018, 0.05, 0.95),  
+           between(MINF_LINC_2018, 0.05, 0.95),  
+           between(MINF_STIL_2018, 0.05, 0.95),  
+           between(MINF_TMPL_2018, 0.05, 0.95),  
+           between(MINF_CLMB_2018, 0.05, 0.95),
+           between(MINF_BRKG_2019, 0.05, 0.95), 
+           between(MINF_PKLE_2019, 0.05, 0.95),
+           between(MINF_FRMI_2019, 0.05, 0.95), 
+           between(MINF_KING_2019, 0.05, 0.95), 
+           between(MINF_OVTN_2019, 0.05, 0.95),  
+           between(MINF_KBSM_2019, 0.05, 0.95),  
+           between(MINF_LINC_2019, 0.05, 0.95),  
+           between(MINF_STIL_2019, 0.05, 0.95),  
+           between(MINF_TMPL_2019, 0.05, 0.95),  
+           between(MINF_CLMB_2019, 0.05, 0.95)) %>% 
+  tally() %>% arrange(desc(n)) %>%
+  write_csv(., path = "SNPs_where_MAF_allows_gwas_10_sites_2_years.csv")
+freq_sitebyyear %>%
+  group_by(between(MINF_BRKG_2018, 0.05, 0.95), 
+           between(MINF_PKLE_2018, 0.05, 0.95),
+           between(MINF_FRMI_2018, 0.05, 0.95), 
+           between(MINF_KING_2018, 0.05, 0.95), 
+           between(MINF_OVTN_2018, 0.05, 0.95),  
+           between(MINF_KBSM_2018, 0.05, 0.95),  
+           between(MINF_LINC_2018, 0.05, 0.95),  
+           between(MINF_STIL_2018, 0.05, 0.95),  
+           between(MINF_TMPL_2018, 0.05, 0.95),  
+           between(MINF_CLMB_2018, 0.05, 0.95)) %>% 
+  tally() %>% arrange(desc(n)) %>%
+  write_csv(., path = "SNPs_where_MAF_allows_gwas_10_sites_2018.csv")
+```
+
+```{r}
+freq_maf_2018 <- freq_sitebyyear %>%
+  filter(between(MINF_BRKG_2018, 0.05, 0.95) & 
+           between(MINF_PKLE_2018, 0.05, 0.95) &
+           between(MINF_FRMI_2018, 0.05, 0.95) & 
+           between(MINF_KING_2018, 0.05, 0.95) & 
+           between(MINF_OVTN_2018, 0.05, 0.95) & 
+           between(MINF_KBSM_2018, 0.05, 0.95) & 
+           between(MINF_LINC_2018, 0.05, 0.95) & 
+           between(MINF_STIL_2018, 0.05, 0.95) & 
+           between(MINF_TMPL_2018, 0.05, 0.95) & 
+           between(MINF_CLMB_2018, 0.05, 0.95)) 
+saveRDS(freq_maf_2018, "~/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/Allele_frequencies_with_maf_above_5percent.rds")
+```
+
+## Labeled hex plots
+
+```{r}
+brkg2yr <- freq_maf_2018 %>%
+  ggplot(aes(x = MINF_BRKG_2018, y = MINF_BRKG_2019)) + 
+  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
+  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.95, 0.475)) + 
+  xlab("MAF in Brookings, 2018") +
+  ylab("MAF in Brookings, 2019")
+save_plot(filename = "BRKG_2_years_MAF_change.png", plot = brkg2yr)
+
+brkg_king9 <- freq_maf_2018 %>%
+  ggplot(aes(x = MINF_KING_2019, y = MINF_BRKG_2019)) + 
+  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
+  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.95, 0.475)) + 
+  xlab("MAF in Kingsville, 2019") +
+  ylab("MAF in Brookings, 2019")
+save_plot(filename = "BRKG_2019_vs_KING_2019_MAF.png", plot = brkg_king9)
+
+brkg_ovtn9 <- freq_maf_2018 %>%
+  ggplot(aes(x = MINF_OVTN_2019, y = MINF_BRKG_2019)) + 
+  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
+  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.95, 0.475)) + 
+  xlab("MAF in Overton, 2019") +
+  ylab("MAF in Brookings, 2019")
+save_plot(filename = "BRKG_2019_vs_OVTN_2019_MAF.png", plot = brkg_ovtn9)
+
+brkg_ovtn8 <- freq_maf_2018 %>%
+  ggplot(aes(x = MINF_OVTN_2018, y = MINF_BRKG_2018)) + 
+  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
+  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.95, 0.475)) + 
+  xlab("MAF in Overton, 2018") +
+  ylab("MAF in Brookings, 2018")
+save_plot(filename = "BRKG_2018_vs_OVTN_2018_MAF.png", plot = brkg_ovtn8)
+
+linc_ovtn8 <- freq_maf_2018 %>%
+  ggplot(aes(x = MINF_OVTN_2018, y = MINF_LINC_2018)) + 
+  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
+  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.95, 0.475)) + 
+  xlab("MAF in Overton, 2018") +
+  ylab("MAF in Lincoln, 2018")
+save_plot(filename = "LINC_2018_vs_OVTN_2018_MAF.png", plot = linc_ovtn8)
+###
+linc_ovtn9 <- freq_maf_2018 %>%
+  ggplot(aes(x = MINF_OVTN_2019, y = MINF_LINC_2019)) + 
+  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
+  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.95, 0.475)) + 
+  xlab("MAF in Overton, 2019") +
+  ylab("MAF in Lincoln, 2019")
+save_plot(filename = "LINC_2019_vs_OVTN_2019_MAF.png", plot = linc_ovtn9)
+
+frmi_king9 <- freq_maf_2018 %>%
+  ggplot(aes(x = MINF_FRMI_2019, y = MINF_KING_2019)) + 
+  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
+  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.95, 0.475)) + 
+  xlab("MAF in Fermilab, IL, 2019") +
+  ylab("MAF in Kingsville, TX, 2019")
+save_plot(filename = "FRMI_2019_vs_KING_2019_MAF.png", plot = frmi_king9)
+##
+
+frmi_2yr <- freq_maf_2018 %>%
+  ggplot(aes(x = MINF_FRMI_2018, y = MINF_FRMI_2019)) + 
+  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
+  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.95, 0.475)) + 
+  xlab("MAF in Fermilab, IL, 2018") +
+  ylab("MAF in Fermilab, IL, 2019")
+save_plot(filename = "FRMI_2018_vs_2019_MAF.png", plot = frmi_2yr)
+##
+
+```
+
+# Minor allele count instead of MAF
+
+Minor allele needs to be present at least 15 times (for a MAF of 5% in 2019 at FRMI). 18 for mean
+
+But for BRKG/LINC/KING 21 - 24 present is better. Let's do 25, every location has to have 25 or more of an allele to keep that SNP.
+
+Or (if FRMI leads to dropping a ton of SNPs) possibly keep some SNPs that are above 15 for FRMI, and above 25 everywhere else.
+
+```{r}
+freq
+mean(freq$N_CHR_FRMI_2019)*.05 # 286 # 416 BRKG 428 LINC 478 KING
+(286)*.05
+mean(freq$N_CHR_BRKG_2019)*.05
+mean(freq$N_CHR_LINC_2019)*.05
+mean(freq$N_CHR_KING_2019)*.05
+```
+STIL 762
+
+```{r}
+mac_sitebyyear <- freq_sitebyyear %>%
+  mutate(MAC_PKLE_2019 = ifelse(round(N_CHR_PKLE_2019*MINF_PKLE_2019)>=N_CHR_PKLE_2019/2, round(N_CHR_PKLE_2019*MAJF_PKLE_2019), round(N_CHR_PKLE_2019*MINF_PKLE_2019)),
+         MAC_FRMI_2019 = ifelse(round(N_CHR_FRMI_2019*MINF_FRMI_2019)>=N_CHR_FRMI_2019/2, round(N_CHR_FRMI_2019*MAJF_FRMI_2019), round(N_CHR_FRMI_2019*MINF_FRMI_2019)),
+         MAC_LINC_2019 = ifelse(round(N_CHR_LINC_2019*MINF_LINC_2019)>=N_CHR_LINC_2019/2, round(N_CHR_LINC_2019*MAJF_LINC_2019), round(N_CHR_LINC_2019*MINF_LINC_2019)),
+         MAC_BRKG_2019 = ifelse(round(N_CHR_BRKG_2019*MINF_BRKG_2019)>=N_CHR_BRKG_2019/2, round(N_CHR_BRKG_2019*MAJF_BRKG_2019), round(N_CHR_BRKG_2019*MINF_BRKG_2019)),
+         MAC_CLMB_2019 = ifelse(round(N_CHR_CLMB_2019*MINF_CLMB_2019)>=N_CHR_CLMB_2019/2, round(N_CHR_CLMB_2019*MAJF_CLMB_2019), round(N_CHR_CLMB_2019*MINF_CLMB_2019)),
+         MAC_OVTN_2019 = ifelse(round(N_CHR_OVTN_2019*MINF_OVTN_2019)>=N_CHR_OVTN_2019/2, round(N_CHR_OVTN_2019*MAJF_OVTN_2019), round(N_CHR_OVTN_2019*MINF_OVTN_2019)),
+         MAC_STIL_2019 = ifelse(round(N_CHR_STIL_2019*MINF_STIL_2019)>=N_CHR_STIL_2019/2, round(N_CHR_STIL_2019*MAJF_STIL_2019), round(N_CHR_STIL_2019*MINF_STIL_2019)),
+         MAC_TMPL_2019 = ifelse(round(N_CHR_TMPL_2019*MINF_TMPL_2019)>=N_CHR_TMPL_2019/2, round(N_CHR_TMPL_2019*MAJF_TMPL_2019), round(N_CHR_TMPL_2019*MINF_TMPL_2019)),
+         MAC_KBSM_2019 = ifelse(round(N_CHR_KBSM_2019*MINF_KBSM_2019)>=N_CHR_KBSM_2019/2, round(N_CHR_KBSM_2019*MAJF_KBSM_2019), round(N_CHR_KBSM_2019*MINF_KBSM_2019)),
+         MAC_KING_2019 = ifelse(round(N_CHR_KING_2019*MINF_KING_2019)>=N_CHR_KING_2019/2, round(N_CHR_KING_2019*MAJF_KING_2019), round(N_CHR_KING_2019*MINF_KING_2019)),
+         MAC_KING_2018 = ifelse(round(N_CHR_KING_2018*MINF_KING_2018)>=N_CHR_KING_2018/2, round(N_CHR_KING_2018*MAJF_KING_2018), round(N_CHR_KING_2018*MINF_KING_2018)),
+         MAC_KBSM_2018 = ifelse(round(N_CHR_KBSM_2018*MINF_KBSM_2018)>=N_CHR_KBSM_2018/2, round(N_CHR_KBSM_2018*MAJF_KBSM_2018), round(N_CHR_KBSM_2018*MINF_KBSM_2018)),
+         MAC_PKLE_2018 = ifelse(round(N_CHR_PKLE_2018*MINF_PKLE_2018)>=N_CHR_PKLE_2018/2, round(N_CHR_PKLE_2018*MAJF_PKLE_2018), round(N_CHR_PKLE_2018*MINF_PKLE_2018)),
+         MAC_OVTN_2018 = ifelse(round(N_CHR_OVTN_2018*MINF_OVTN_2018)>=N_CHR_OVTN_2018/2, round(N_CHR_OVTN_2018*MAJF_OVTN_2018), round(N_CHR_OVTN_2018*MINF_OVTN_2018)),
+         MAC_BRKG_2018 = ifelse(round(N_CHR_BRKG_2018*MINF_BRKG_2018)>=N_CHR_BRKG_2018/2, round(N_CHR_BRKG_2018*MAJF_BRKG_2018), round(N_CHR_BRKG_2018*MINF_BRKG_2018)),
+         MAC_CLMB_2018 = ifelse(round(N_CHR_CLMB_2018*MINF_CLMB_2018)>=N_CHR_CLMB_2018/2, round(N_CHR_CLMB_2018*MAJF_CLMB_2018), round(N_CHR_CLMB_2018*MINF_CLMB_2018)),
+         MAC_TMPL_2018 = ifelse(round(N_CHR_TMPL_2018*MINF_TMPL_2018)>=N_CHR_TMPL_2018/2, round(N_CHR_TMPL_2018*MAJF_TMPL_2018), round(N_CHR_TMPL_2018*MINF_TMPL_2018)),
+         MAC_LINC_2018 = ifelse(round(N_CHR_LINC_2018*MINF_LINC_2018)>=N_CHR_LINC_2018/2, round(N_CHR_LINC_2018*MAJF_LINC_2018), round(N_CHR_LINC_2018*MINF_LINC_2018)),
+         MAC_STIL_2018 = ifelse(round(N_CHR_STIL_2018*MINF_STIL_2018)>=N_CHR_STIL_2018/2, round(N_CHR_STIL_2018*MAJF_STIL_2018), round(N_CHR_STIL_2018*MINF_STIL_2018)),
+         MAC_FRMI_2018 = ifelse(round(N_CHR_FRMI_2018*MINF_FRMI_2018)>=N_CHR_FRMI_2018/2, round(N_CHR_FRMI_2018*MAJF_FRMI_2018), round(N_CHR_FRMI_2018*MINF_FRMI_2018))) %>%
+  dplyr::select(CHR, POS, MAC_PKLE_2019:MAC_FRMI_2018)
+
+saveRDS(mac_sitebyyear, file =  "SNPs_where_minor_allele_count_allows_gwas_10_sites_2_years.rds")
+
+mac_sitebyyear %>%
+    group_by(MAC_PKLE_2019 > 25, MAC_PKLE_2018 > 25,
+           MAC_FRMI_2019 > 15, MAC_FRMI_2018 > 15,
+           MAC_OVTN_2019 > 25, MAC_OVTN_2018 > 25,
+           MAC_LINC_2019 > 25, MAC_LINC_2018 > 25,
+           MAC_STIL_2019 > 25, MAC_STIL_2018 > 25,
+           MAC_TMPL_2019 > 25, MAC_TMPL_2018 > 25,
+           MAC_CLMB_2019 > 25, MAC_CLMB_2018 > 25,
+           MAC_BRKG_2019 > 25, MAC_BRKG_2018 > 25,
+           MAC_KBSM_2019 > 25, MAC_KBSM_2018 > 25,
+           MAC_KING_2019 > 25, MAC_KING_2018 > 25) %>%
+  tally() %>%
+  write_csv(., path = "Minor_allele_count_that_allows_gwas_10_sites_2_years_relaxed_FRMI.csv")
+
+mac_sitebyyear %>%
+    group_by(MAC_PKLE_2019 > 25, MAC_PKLE_2018 > 25,
+           MAC_FRMI_2019 > 25, MAC_FRMI_2018 > 25,
+           MAC_OVTN_2019 > 25, MAC_OVTN_2018 > 25,
+           MAC_LINC_2019 > 25, MAC_LINC_2018 > 25,
+           MAC_STIL_2019 > 25, MAC_STIL_2018 > 25,
+           MAC_TMPL_2019 > 25, MAC_TMPL_2018 > 25,
+           MAC_CLMB_2019 > 25, MAC_CLMB_2018 > 25,
+           MAC_BRKG_2019 > 25, MAC_BRKG_2018 > 25,
+           MAC_KBSM_2019 > 25, MAC_KBSM_2018 > 25,
+           MAC_KING_2019 > 25, MAC_KING_2018 > 25) %>%
+  tally() %>%
+  write_csv(., path = "Minor_allele_count_that_allows_gwas_10_sites_2_years.csv")
+```
+
+```{r}
+mac %>%
+  group_by(MAC_PKLE_2019 > 25, MAC_PKLE_2018 > 25,
+           MAC_FRMI_2019 > 15, MAC_FRMI_2018 > 15,
+           MAC_OVTN_2019 > 25, MAC_OVTN_2018 > 25,
+           MAC_LINC_2019 > 25, MAC_LINC_2018 > 25,
+           MAC_STIL_2019 > 25, MAC_STIL_2018 > 25,
+           MAC_TMPL_2019 > 25, MAC_TMPL_2018 > 25,
+           MAC_CLMB_2019 > 25, MAC_CLMB_2018 > 25,
+           MAC_BRKG_2019 > 25, MAC_BRKG_2018 > 25,
+           MAC_KBSM_2019 > 25, MAC_KBSM_2018 > 25,
+           MAC_KING_2019 > 25, MAC_KING_2018 > 25) %>%
+  tally()
+```
+
+```{r}
+mac_sitebyyear <- readRDS(file =  "SNPs_where_minor_allele_count_allows_gwas_10_sites_2_years.rds")
+
+mac_2018_gt25 <- mac_sitebyyear %>%
+  filter(MAC_PKLE_2018 >= 25,
+         MAC_FRMI_2018 >= 15,
+         MAC_OVTN_2018 >= 25,
+         MAC_LINC_2018 >= 25,
+         MAC_STIL_2018 >= 25,
+         MAC_TMPL_2018 >= 25,
+         MAC_CLMB_2018 >= 25,
+         MAC_BRKG_2018 >= 25,
+         MAC_KBSM_2018 >= 25,
+         MAC_KING_2018 >= 25)
+
+brkg_2yr <- mac_2018_gt25 %>%
+  ggplot(aes(x = MAC_BRKG_2018, y = MAC_BRKG_2019)) + 
+  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.2, 0.7)) + 
+  xlab("Minor Allele Count in Brookings, 2018") +
+  ylab("Minor Allele Count in Brookings, 2019")
+save_plot(filename = "BRKG_2018_v_2019_minor_allele_count_change.png", plot = brkg_2yr)
+
+brkg_king9 <- mac_2018_gt25 %>%
+  ggplot(aes(x = MAC_KING_2019, y = MAC_BRKG_2019)) + 
+  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.2, 0.7)) + 
+  xlab("Minor Allele Count in Kingsville, 2019") +
+  ylab("Minor Allele Count in Brookings, 2019")
+save_plot(filename = "BRKG_2019_vs_KING_2019_minor_allele_count.png", plot = brkg_king9)
+
+brkg_ovtn9 <- mac_2018_gt25 %>%
+  ggplot(aes(x = MAC_OVTN_2019, y = MAC_BRKG_2019)) + 
+  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.2, 0.7)) + + 
+  xlab("Minor Allele Count in Overton, 2019") +
+  ylab("Minor Allele Count in Brookings, 2019")
+save_plot(filename = "BRKG_2019_vs_OVTN_2019_minor_allele_count.png", plot = brkg_ovtn9)
+
+brkg_ovtn8 <- mac_2018_gt25 %>%
+  ggplot(aes(x = MAC_OVTN_2018, y = MAC_BRKG_2018)) + 
+  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.2, 0.7)) + 
+  xlab("Minor Allele Count in Overton, TX, 2018") +
+  ylab("Minor Allele Count in Brookings, SD, 2018")
+save_plot(filename = "BRKG_2018_vs_OVTN_2018_minor_allele_count.png", plot = brkg_ovtn8)
+
+linc_ovtn8 <- mac_2018_gt25 %>%
+  ggplot(aes(x = MAC_OVTN_2018, y = MAC_LINC_2018)) + 
+  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.2, 0.7)) +  
+  xlab("Minor Allele Count in Overton, 2018") +
+  ylab("Minor Allele Count in Lincoln, 2018")
+save_plot(filename = "LINC_2018_vs_OVTN_2018_minor_allele_count.png", plot = linc_ovtn8)
+###
+
+linc_ovtn9 <- mac_2018_gt25 %>%
+  ggplot(aes(x = MAC_OVTN_2019, y = MAC_LINC_2019)) + 
+  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.2, 0.7)) + 
+  xlab("Minor Allele Count in Overton, TX, 2019") +
+  ylab("Minor Allele Count in Lincoln, NE, 2019")
+save_plot(filename = "LINC_2019_vs_OVTN_2019_minor_allele_count.png", plot = linc_ovtn9)
+
+frmi_king9 <- mac_2018_gt25 %>%
+  ggplot(aes(x = MAC_FRMI_2019, y = MAC_KING_2019)) + 
+  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.2, 0.7)) + 
+  xlab("Minor Allele Count in Fermilab, IL, 2019") +
+  ylab("Minor Allele Count in Kingsville, TX, 2019")
+save_plot(filename = "FRMI_2019_vs_KING_2019_minor_allele_count.png", plot = frmi_king9)
+##
+
+frmi_2yr <- mac_2018_gt25 %>%
+  ggplot(aes(x = MAC_FRMI_2018, y = MAC_FRMI_2019)) + 
+  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.2, 0.7)) + 
+  xlab("Minor Allele Count in Fermilab, IL, 2018") +
+  ylab("Minor Allele Count in Fermilab, IL, 2019")
+save_plot(filename = "FRMI_2018_vs_2019_minor_allele_count.png", plot = frmi_2yr)
+##
+linc_2yr <- mac_2018_gt25 %>%
+  ggplot(aes(x = MAC_LINC_2018, y = MAC_LINC_2019)) + 
+  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.2, 0.7)) + 
+  xlab("Minor Allele Count in Lincoln, NE, 2018") +
+  ylab("Minor Allele Count in Lincoln, NE, 2019")
+save_plot(filename = "LINC_2018_vs_2019_minor_allele_count.png", plot = linc_2yr)
+
+clmb_2yr <- mac_2018_gt25 %>%
+  ggplot(aes(x = MAC_CLMB_2018, y = MAC_CLMB_2019)) + 
+  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.2, 0.7)) + 
+  xlab("Minor Allele Count in Columbia, MO, 2018") +
+  ylab("Minor Allele Count in Columbia, MO, 2019")
+save_plot(filename = "CLMB_2018_vs_2019_minor_allele_count.png", plot = clmb_2yr)
+
+kbsm_2yr <- mac_2018_gt25 %>%
+  ggplot(aes(x = MAC_KBSM_2018, y = MAC_KBSM_2019)) + 
+  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.2, 0.7)) + 
+  xlab("Minor Allele Count in KBS, MI, 2018") +
+  ylab("Minor Allele Count in KBS, MI, 2019")
+save_plot(filename = "KBSM_2018_vs_2019_minor_allele_count.png", plot = kbsm_2yr)
+###
+frmi_king8 <- mac_2018_gt25 %>%
+  ggplot(aes(x = MAC_FRMI_2018, y = MAC_KING_2018)) + 
+  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.2, 0.7)) + 
+  xlab("Minor Allele Count in Fermilab, IL, 2018") +
+  ylab("Minor Allele Count in Kingsville, TX, 2018")
+save_plot(filename = "FRMI_2018_vs_KING_2018_minor_allele_count.png", plot = frmi_king9)
+brkg_king8 <- mac_2018_gt25 %>%
+  ggplot(aes(x = MAC_KING_2018, y = MAC_BRKG_2018)) + 
+  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.2, 0.7)) + 
+  xlab("Minor Allele Count in Kingsville, 2018") +
+  ylab("Minor Allele Count in Brookings, 2018")
+save_plot(filename = "BRKG_2018_vs_KING_2018_minor_allele_count.png", plot = brkg_king9)
+#
+clmb_pkle9 <- mac_2018_gt25 %>%
+  ggplot(aes(x = MAC_CLMB_2019, y = MAC_PKLE_2019)) + 
+  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.2, 0.7)) + 
+  xlab("Minor Allele Count in Columbia, MO, 2019") +
+  ylab("Minor Allele Count in Pickle, TX, 2019")
+save_plot(filename = "CLMB_2019_vs_PKLE_2019_minor_allele_count.png", plot = clmb_pkle9)
+
+kbsm_pkle9 <- mac_2018_gt25 %>%
+  ggplot(aes(x = MAC_KBSM_2019, y = MAC_PKLE_2019)) + 
+  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.2, 0.7)) + 
+  xlab("Minor Allele Count in KBS, MI, 2019") +
+  ylab("Minor Allele Count in Pickle, TX, 2019")
+save_plot(filename = "KBSM_2019_vs_PKLE_2019_minor_allele_count.png", plot = kbsm_pkle9)
+
+brkg_pkle9 <- mac_2018_gt25 %>%
+  ggplot(aes(x = MAC_BRKG_2019, y = MAC_PKLE_2019)) + 
+  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
+  geom_hex() + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
+  scale_fill_viridis(option = "B", trans = "log2") + 
+  theme(legend.position = c(0.2, 0.7)) + 
+  xlab("Minor Allele Count in Brookings, SD, 2019") +
+  ylab("Minor Allele Count in Pickle, TX, 2019")
+save_plot(filename = "BRKG_2019_vs_PKLE_2019_minor_allele_count.png", plot = brkg_pkle9)
+```
+
+5 plots of 2018 vs 2019: any place where more than 20 plants died. That's basically just the north.
+
+Some 2019 north vs south plots.

### 2020-02-06 analysis/common-garden-allele-balancing/1_VCF_file_filtering_by_common_garden_site.Rmd
Common garden allele balancing four 33M SNP data subsets

@@
 load("~/Github/pvdiv-phenology-gxe/data/sites.rda")
+load("~/Github/pvdiv-phenology-gxe/data/metadata.rda")
 ```
@@
 phenotypes_wide_2019 %>%
-  filter(!(PLANT_ID %in% c("J661.C", "J491.A", "J385.C", "J471.A", "J243.A", "J428.A", "J399.B", "J344.B", "J003.C", "J624.B", "J682.C", "NL94×SL93_370", "NL94×SL93_305"))) # None of these have any 2019 data.
+  filter((PLANT_ID %in% c("J661.C", "J491.A", "J385.C", "J471.A", "J243.A", "J428.A", "J399.B", "J344.B", "J003.C", "J624.B", "J682.C", "NL94×SL93_370", "NL94×SL93_305"))) # None of these have any 2019 data.
 
@@
 write_csv(phe_csv, path = "Individuals_needed_for_33M_SNPs.csv")
+
+phe_2018 %>%
+  dplyr::select(PLANT_ID) %>%
+  anti_join(phe_csv)
 ```

### 2020-02-06 analysis/common-garden-allele-balancing/2_allele_frequencies_common_gardens.Rmd
Common garden allele balancing four 33M SNP data subsets

@@
 Some 2019 north vs south plots.
+
+# -----------------
+
+# Impute on 11M SNPs
+
+```{r}
+snps <- readRDS("SNPs_where_minor_allele_count_allows_gwas_10_sites_2_years_.rds")
+```
+
+Here is a breakdown of the number of individuals we have for GWAS at each site, for phenotypes from 2018 and from 2019.
+The first two data columns have individuals with high quality SNP data and phenotypes.
+The next two have tetraploid plants with phenotypes – there is a lot of overlap, and only 5-6 plants that could have their libraries resequenced to help the GWAS data.
+
+	Plants with SNP data & phenotypes	Plants with phenotypes	Plants Dead
+	2018	2019	2018	2019	2018 to 2019
+BRKG	408	254	412	254	154
+KBSM	539	517	544	517	22
+FRMI	272	185	273	185	87
+LINC	433	262	437	262	171
+CLMB	585	496	591	496	89
+STIL	465	451	469	451	14
+OVTN	343	332	346	332	11
+TMPL	384	373	388	373	11
+PKLE	617	615	623	615	2
+KING	303	294	305	294	9
+
+Given that information, I then subdivided the SNPs from John’s new 33.9M SNP dataset.
+These are SNPs missing in 20% or less of individuals from each subpopulation.
+Instead of a minor allele frequency, I used a minor allele count of 25 or more (except for FRMI where I used 15).
+This was the same as a MAF of 5% at BRKG, FRMI, and LINC, and a slightly lower MAF for the other sites with more individuals.
+Here’s the breakdown of the site*year combinations where SNPs can and can’t be used for GWAS:
+
+Number of SNPs	Category
+12536071
+	Can't be used for GWAS – these SNPs have less than 25 copies of the minor allele in the 631 individuals planted at the common gardens.
+10452584	Can be used for GWAS at all 10 sites and both 2018 and 2019 data.
+1160326	Can’t be used in LINC & BRKG in 2019
+1155906	Only at high enough frequency to be used at the big 3 sites
+546381	Can only be used at PKLE & CLMB
+439918	Can only be used at PKLE
+346189	Can only be used at big 4 sites – KBSM CLMB PKLE STIL
+341773	Can’t be used at KING OVTN FRMI
+331082	Can’t be used at LINC in 2019 
+327364	Can’t be used at KING
+315522	Can only be used at big 4 and LINC 2018
+305573	Missing 2019 at 4 freezing sites – LINC FRMI CLMB BRKG
+285838	Can only be used at PKLE x2yr & CLMB 2018
+246340	Missing 2019 at 3 freezing sites, not big site (BRKG)
+231650	Missing KING x2yr and FRMI 2019
+192819	Missing KING x2yr and FRMI x2yr
+……	4.67M SNPs have some other combination of Site x Year that is useable.
+
+10_site_2_yr 10452584 all MAC > 25 except FRMI 2019 > 15.
+big_three   10452584 + 1155906 PKLE KBSM CLMB > 25
+fourway_seven   PKLE TMPL BRKG KBSM LINC CLMB STIL > 25
+ 10452584 + 341773 + 327364 + 231650 + 192819 + ...
+miss_freeze 305573 + 246340 + 1160326 + 331082 4 sites 2018 > 25 and 1-4 sites 2019 < 25
+
+## 10site_2yr
+all MAC >= 25 except FRMI 2019 > 15.
+
+10,183,856 SNPs
+```{r}
+tensite_twoyear <- snps %>%
+  filter(MAC_PKLE_2019 >= 25, MAC_PKLE_2018 >= 25,
+           MAC_FRMI_2019 >= 15, MAC_FRMI_2018 >= 25,
+           MAC_OVTN_2019 >= 25, MAC_OVTN_2018 >= 25,
+           MAC_LINC_2019 >= 25, MAC_LINC_2018 >= 25,
+           MAC_STIL_2019 >= 25, MAC_STIL_2018 >= 25,
+           MAC_TMPL_2019 >= 25, MAC_TMPL_2018 >= 25,
+           MAC_CLMB_2019 >= 25, MAC_CLMB_2018 >= 25,
+           MAC_BRKG_2019 >= 25, MAC_BRKG_2018 >= 25,
+           MAC_KBSM_2019 >= 25, MAC_KBSM_2018 >= 25,
+           MAC_KING_2019 >= 25, MAC_KING_2018 >= 25)
+saveRDS(tensite_twoyear, file = "SNPs_above_MAC_25_in_10sites_2years.rds")
+```
+
+## big_three
+
+18,333,684
+
+PKLE KBSM CLMB >= 25
+```{r}
+big_three <- snps %>%
+  filter(MAC_PKLE_2019 >= 25, MAC_PKLE_2018 >= 25,
+           MAC_CLMB_2019 >= 25, MAC_CLMB_2018 >= 25,
+           MAC_KBSM_2019 >= 25, MAC_KBSM_2018 >= 25)
+saveRDS(big_three, "SNPs_above_MAC_25_in_big_three_sites.rds")
+```
+
+## fourway_seven
+
+12,358,701 SNPs
+
+PKLE TMPL BRKG KBSM LINC CLMB STIL >= 25
+```{r}
+fourway_seven <- snps %>%
+  filter(MAC_PKLE_2019 >= 25, MAC_PKLE_2018 >= 25,
+           MAC_LINC_2019 >= 25, MAC_LINC_2018 >= 25,
+           MAC_STIL_2019 >= 25, MAC_STIL_2018 >= 25,
+           MAC_TMPL_2019 >= 25, MAC_TMPL_2018 >= 25,
+           MAC_CLMB_2019 >= 25, MAC_CLMB_2018 >= 25,
+           MAC_BRKG_2019 >= 25, MAC_BRKG_2018 >= 25,
+           MAC_KBSM_2019 >= 25, MAC_KBSM_2018 >= 25)
+saveRDS(fourway_seven, file = "SNPs_above_MAC_25_in_fourway_7_sites.rds")
+```
+
+## missing_frz
+
+4 sites 2018 >= 25 and 1-4 sites 2019 < 25
+
+1,963,771 SNPs
+```{r}
+missing_frz <- snps %>%
+  filter(MAC_FRMI_2018 >= 25 & MAC_LINC_2018 >= 25 & MAC_CLMB_2018 >= 25 &
+           MAC_BRKG_2018 >= 25, MAC_FRMI_2019 < 15 | MAC_LINC_2018 < 25 | 
+           MAC_CLMB_2018 < 25 | MAC_BRKG_2019 < 25) %>%
+  mutate(avediff = ((MAC_FRMI_2018 - MAC_FRMI_2019) + (MAC_LINC_2018 - MAC_LINC_2019) + (MAC_CLMB_2018 - MAC_CLMB_2019) + (MAC_BRKG_2018 - MAC_BRKG_2019))/4)
+
+saveRDS(missing_frz, file = "Missing_SNPs_2019_4_freezing_sites.rds")
+```
+
+```{r}
+missing_frz <- readRDS(file = "Missing_SNPs_2019_4_freezing_sites.rds")
+```
+
+```{r}
+missing_frz %>%
+  ggplot(aes(x = POS, y = avediff)) + 
+  geom_point() + 
+  facet_wrap(~CHR)
+save_plot(filename = "Distribution_of_SNPs_lost_in_freezing_four_sites.png", plot = last_plot())
+
+missing_frz %>%
+  ggplot(aes(x = POS, y = avediff)) + 
+  geom_hex() + 
+  facet_wrap(~CHR) + 
+  scale_color_viridis() + 
+  scale_fill_viridis()
+save_plot(filename = "Distribution_of_SNPs_lost_in_freezing_four_sites_hex.png", plot = last_plot())
+```

### 2020-02-06 analysis/common-garden-allele-balancing/2_allele_frequencies_common_gardens.Rmd
Analyze genomic distribution of winter survival selected SNPs

@@
 
-18,333,684
+18,704,908 SNPs
 
@@
 saveRDS(big_three, "SNPs_above_MAC_25_in_big_three_sites.rds")
+10183856+6027776+2493276
 ```
@@
 
-12,358,701 SNPs
+12,677,132 SNPs
 
@@
 saveRDS(fourway_seven, file = "SNPs_above_MAC_25_in_fourway_7_sites.rds")
+10183856+2493276
+```
+
+## microbiome
+
+Just got the results from PKLE thus far.
+
+21,525,180 SNPs
+
+```{r}
+microbiome <- snps %>%
+  filter(MAC_PKLE_2019 >= 25)
+10183856+6027776+2493276+2820272
 ```
@@
 
-1,963,771 SNPs
+2,288,528 SNPs
 ```{r}
@@
   filter(MAC_FRMI_2018 >= 25 & MAC_LINC_2018 >= 25 & MAC_CLMB_2018 >= 25 &
-           MAC_BRKG_2018 >= 25, MAC_FRMI_2019 < 15 | MAC_LINC_2018 < 25 | 
-           MAC_CLMB_2018 < 25 | MAC_BRKG_2019 < 25) %>%
+           MAC_BRKG_2018 >= 25, MAC_FRMI_2019 < 15 | MAC_LINC_2019 < 25 | 
+           MAC_CLMB_2019 < 25 | MAC_BRKG_2019 < 25) %>%
   mutate(avediff = ((MAC_FRMI_2018 - MAC_FRMI_2019) + (MAC_LINC_2018 - MAC_LINC_2019) + (MAC_CLMB_2018 - MAC_CLMB_2019) + (MAC_BRKG_2018 - MAC_BRKG_2019))/4)
@@
 saveRDS(missing_frz, file = "Missing_SNPs_2019_4_freezing_sites.rds")
+
+25/(615*2)
+25/(496*2)
+```
+
+```{r}
+snp_subsets <- snps %>%
+  mutate(SNP_tensite_twoyear = ifelse(MAC_PKLE_2019 >= 25 & 
+                                        MAC_PKLE_2018 >= 25 & 
+                                        MAC_FRMI_2019 >= 15 & 
+                                        MAC_FRMI_2018 >= 25 & 
+                                        MAC_OVTN_2019 >= 25 & 
+                                        MAC_OVTN_2018 >= 25 & 
+                                        MAC_LINC_2019 >= 25 & 
+                                        MAC_LINC_2018 >= 25 & 
+                                        MAC_STIL_2019 >= 25 & 
+                                        MAC_STIL_2018 >= 25 & 
+                                        MAC_TMPL_2019 >= 25 & 
+                                        MAC_TMPL_2018 >= 25 & 
+                                        MAC_CLMB_2019 >= 25 & 
+                                        MAC_CLMB_2018 >= 25 & 
+                                        MAC_BRKG_2019 >= 25 & 
+                                        MAC_BRKG_2018 >= 25 & 
+                                        MAC_KBSM_2019 >= 25 & 
+                                        MAC_KBSM_2018 >= 25 & 
+                                        MAC_KING_2019 >= 25 & 
+                                        MAC_KING_2018 >= 25, "Y", "N"),
+         SNP_big_three = ifelse(MAC_PKLE_2019 >= 25 & 
+                                        MAC_PKLE_2018 >= 25 & 
+                                        MAC_CLMB_2019 >= 25 & 
+                                        MAC_CLMB_2018 >= 25 & 
+                                        MAC_KBSM_2019 >= 25 & 
+                                        MAC_KBSM_2018 >= 25, "Y", "N"),
+         SNP_fourway_seven = ifelse(MAC_PKLE_2019 >= 25 & 
+                                        MAC_PKLE_2018 >= 25 & 
+                                        MAC_LINC_2019 >= 25 & 
+                                        MAC_LINC_2018 >= 25 & 
+                                        MAC_STIL_2019 >= 25 & 
+                                        MAC_STIL_2018 >= 25 & 
+                                        MAC_TMPL_2019 >= 25 & 
+                                        MAC_TMPL_2018 >= 25 & 
+                                        MAC_CLMB_2019 >= 25 & 
+                                        MAC_CLMB_2018 >= 25 & 
+                                        MAC_BRKG_2019 >= 25 & 
+                                        MAC_BRKG_2018 >= 25 & 
+                                        MAC_KBSM_2019 >= 25 & 
+                                        MAC_KBSM_2018 >= 25, "Y", "N"),
+         SNP_microbiome = ifelse(MAC_PKLE_2019 >= 25 & 
+                                        MAC_PKLE_2018 >= 25, "Y", "N"),
+         SNP_missing_frz = ifelse((MAC_FRMI_2018 >= 25 & MAC_BRKG_2018 >= 25 &
+                                     MAC_LINC_2018 >= 25 & 
+                                     MAC_CLMB_2018 >= 25) &
+                                     (MAC_LINC_2019 < 25 | 
+                                        MAC_CLMB_2019 < 25 | 
+                                        MAC_BRKG_2019 < 25 |
+                                    MAC_FRMI_2019 < 15), "Y", "N")
+         )
+
+snp_subsets %>%
+  dplyr::select(CHR, POS, starts_with("SNP_")) %>%
+  saveRDS(., file = "SNP_subsets_33M_markers.rds")
+
+missing_frz <- snp_subsets %>%
+  filter(SNP_missing_frz == "Y")
+
+snp_subsets %>%
+  filter(SNP_missing_frz == "Y") %>%
+  saveRDS(., file = "Missing_SNPs_2019_4_freezing_sites.rds")
+
+snp_subsets %>%
+  group_by(SNP_missing_frz) %>% tally()
+snp_subsets %>%
+  group_by(SNP_tensite_twoyear, SNP_big_three, SNP_fourway_seven, SNP_microbiome) %>% tally()
 ```
@@
 ```{r}
+theme_set(theme_poster)
 missing_frz %>%
@@
   geom_hex() + 
-  facet_wrap(~CHR) + 
-  scale_color_viridis() + 
-  scale_fill_viridis()
-save_plot(filename = "Distribution_of_SNPs_lost_in_freezing_four_sites_hex.png", plot = last_plot())
-```
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position") + ylab("Minor Allele Loss, 4 sites") +
+  theme(legend.position = "right")
+save_plot(filename = "Distribution_of_SNPs_lost_in_freezing_four_sites_log10.png", plot = last_plot(), base_asp = 4, base_height = 4)
+
+missing_frz %>%
+  filter(MAC_FRMI_2018 >= 25 & MAC_LINC_2018 >= 25 & MAC_CLMB_2018 >= 25 &
+           MAC_BRKG_2018 >= 25 & MAC_FRMI_2019 < 15 & MAC_LINC_2019 < 25 & 
+           MAC_CLMB_2019 < 25 & MAC_BRKG_2019 < 25) %>%
+  ggplot(aes(x = POS, y = avediff)) + 
+  geom_hex() + 
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position") + ylab("Minor Allele Loss, 4 sites") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "Distribution_of_SNPs_lost_in_all_four_freezing_sites_log10.png", plot = last_plot(), base_asp = 4, base_height = 4)
+
+missing_frz %>%
+  filter(MAC_BRKG_2018 >= 25 & MAC_BRKG_2019 < 15) %>%
+  mutate(Allele_loss = MAC_BRKG_2018 - MAC_BRKG_2019) %>%
+  ggplot(aes(x = POS, y = Allele_loss)) + 
+  geom_hex() + 
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position") + ylab("Minor Allele Loss") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "Distribution_of_SNPs_lost_in_BRKG_log10.png", plot = last_plot(), base_asp = 4, base_height = 4)
+
+missing_frz %>%
+  filter(MAC_LINC_2018 >= 25 & MAC_LINC_2019 < 15) %>%
+  mutate(Allele_loss = MAC_LINC_2018 - MAC_LINC_2019) %>%
+  ggplot(aes(x = POS, y = Allele_loss)) + 
+  geom_hex() + 
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position") + ylab("Minor Allele Loss") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "Distribution_of_SNPs_lost_in_LINC_log10.png", plot = last_plot(), base_asp = 4, base_height = 4)
+
+missing_frz %>%
+  filter(MAC_FRMI_2018 >= 25 & MAC_FRMI_2019 < 15) %>%
+  mutate(Allele_loss = MAC_FRMI_2018 - MAC_FRMI_2019) %>%
+  ggplot(aes(x = POS, y = Allele_loss)) + 
+  geom_hex() + 
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position") + ylab("Minor Allele Loss") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "Distribution_of_SNPs_lost_in_FRMI_log10.png", plot = last_plot(), base_asp = 4, base_height = 4)
+
+missing_frz %>%
+  filter(MAC_CLMB_2018 >= 25 & MAC_CLMB_2019 < 15) %>%
+  mutate(Allele_loss = MAC_CLMB_2018 - MAC_CLMB_2019) %>%
+  ggplot(aes(x = POS, y = Allele_loss)) + 
+  geom_hex() + 
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position") + ylab("Minor Allele Loss") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "Distribution_of_SNPs_lost_in_CLMB_log10.png", plot = last_plot(), base_asp = 4, base_height = 4)
+
+missing_frz <- missing_frz %>%
+  mutate(Loss_LINC = MAC_LINC_2018 - MAC_LINC_2019,
+         Loss_BRKG = MAC_BRKG_2018 - MAC_BRKG_2019,
+         Loss_FRMI = MAC_FRMI_2018 - MAC_FRMI_2019,
+         Loss_CLMB = MAC_CLMB_2018 - MAC_CLMB_2019)
+
+missing_frz %>%
+  group_by(Loss_BRKG > 75, Loss_LINC > 78, Loss_FRMI > 54, Loss_CLMB > 51) %>%
+  tally() %>%
+  write_csv(., path = "SNPs_loss_overlaps_1percent_tail_2019_four_freezing_sites.csv")
+
+plot1 <- missing_frz %>%
+  ggplot(aes(x = Loss_CLMB)) + geom_histogram(binwidth = 2)
+save_plot(filename = "Histogram_SNPs_lost_CLMB.png", plot = plot1)
+plot1 <- missing_frz %>%
+  ggplot(aes(x = Loss_FRMI)) + geom_histogram(binwidth = 2)
+save_plot(filename = "Histogram_SNPs_lost_FRMI.png", plot = plot1)
+
+plot1 <- missing_frz %>%
+  ggplot(aes(x = Loss_BRKG)) + geom_histogram(binwidth = 2)
+save_plot(filename = "Histogram_SNPs_lost_BRKG.png", plot = plot1)
+
+plot1 <- missing_frz %>%
+  ggplot(aes(x = Loss_LINC)) + geom_histogram(binwidth = 2)
+save_plot(filename = "Histogram_SNPs_lost_LINC.png", plot = plot1)
+
+missing_frz %>%
+  summarise(top_LINC = quantile(Loss_LINC, 0.99),
+            top_BRKG = quantile(Loss_BRKG, 0.99),
+            top_FRMI = quantile(Loss_FRMI, 0.99),
+            top_CLMB = quantile(Loss_CLMB, 0.99),
+            )
+
+# LINC & BRKG are the most similar
+
+all_frz <- snp_subsets %>%
+  mutate(Loss_LINC = (MAC_LINC_2018 - MAC_LINC_2019)/MAC_LINC_2018,
+         Loss_BRKG = (MAC_BRKG_2018 - MAC_BRKG_2019)/MAC_BRKG_2018,
+         Loss_FRMI = (MAC_FRMI_2018 - MAC_FRMI_2019)/MAC_FRMI_2018,
+         Loss_CLMB = (MAC_CLMB_2018 - MAC_CLMB_2019)/MAC_CLMB_2018)
+all_frz %>%
+  summarise(top_LINC = quantile(Loss_LINC, 0.99, na.rm = TRUE),
+            top_BRKG = quantile(Loss_BRKG, 0.99, na.rm = TRUE),
+            top_FRMI = quantile(Loss_FRMI, 0.99, na.rm = TRUE),
+            top_CLMB = quantile(Loss_CLMB, 0.99, na.rm = TRUE)
+            )
+all_frz %>%
+  group_by(Loss_BRKG > 159, Loss_LINC > 179, Loss_FRMI > 84, Loss_CLMB > 94) %>%
+  tally() %>%
+  write_csv(., path = "All_SNPs_lost_alleles_1percent_tail_2019_four_freezing_sites.csv")
+
+missing_frz %>%
+  filter(Loss_BRKG > 75 | Loss_LINC > 78 | Loss_FRMI > 54 | Loss_CLMB > 51) %>% saveRDS(., file = "Lost_SNPs_with_1percent_tail_frequency_change_freezing_sites.rds")
+all_frz %>%
+  filter(Loss_BRKG > 159 | Loss_LINC > 179 | Loss_FRMI > 84 | Loss_CLMB > 94) %>% saveRDS(., file = "SNPs_with_1percent_tail_frequency_change_freezing_sites.rds")
+```
+
+  top_LINC top_BRKG top_FRMI top_CLMB
+     <dbl>    <dbl>    <dbl>    <dbl>
+1       78       75       54       51
+
+# Frequencies lost 
+```{r}
+frz_freq <- readRDS("SNPs_with_1percent_tail_frequency_change_freezing_sites.rds") 
+loss_freq <- readRDS("Lost_SNPs_with_1percent_tail_frequency_change_freezing_sites.rds")
+```
+
+```{r}
+loss_freq %>%
+ggplot(aes(x = POS/1000000, y = Loss_LINC/MAC_LINC_2018)) + 
+  geom_hex() + 
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "Distribution_of_SNPs_1per_delta_lost_in_LINC.png", plot = last_plot(), base_asp = 4, base_height = 4)
+frz_freq %>%
+ggplot(aes(x = POS/1000000, y = Loss_LINC/MAC_LINC_2018)) + 
+  geom_hex() + 
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "Distribution_of_SNPs_1per_delta_LINC.png", plot = last_plot(), base_asp = 4, base_height = 4)
+```
+
+```{r}
+loss_freq %>%
+ggplot(aes(x = POS/1000000, y = Loss_CLMB/MAC_CLMB_2018)) + 
+  geom_hex() + 
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "Distribution_of_SNPs_1per_delta_lost_in_CLMB.png", plot = last_plot(), base_asp = 4, base_height = 4)
+frz_freq %>%
+ggplot(aes(x = POS/1000000, y = Loss_CLMB/MAC_CLMB_2018)) + 
+  geom_hex() + 
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "Distribution_of_SNPs_1per_delta_CLMB.png", plot = last_plot(), base_asp = 4, base_height = 4)
+
+loss_freq %>%
+ggplot(aes(x = POS/1000000, y = Loss_FRMI/MAC_FRMI_2018)) + 
+  geom_hex() + 
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "Distribution_of_SNPs_1per_delta_lost_in_FRMI.png", plot = last_plot(), base_asp = 4, base_height = 4)
+frz_freq %>%
+ggplot(aes(x = POS/1000000, y = Loss_FRMI/MAC_FRMI_2018)) + 
+  geom_hex() + 
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "Distribution_of_SNPs_1per_delta_FRMI.png", plot = last_plot(), base_asp = 4, base_height = 4)
+
+loss_freq %>%
+ggplot(aes(x = POS/1000000, y = Loss_BRKG/MAC_BRKG_2018)) + 
+  geom_hex() + 
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "Distribution_of_SNPs_1per_delta_lost_in_BRKG.png", plot = last_plot(), base_asp = 4, base_height = 4)
+frz_freq %>%
+ggplot(aes(x = POS/1000000, y = Loss_BRKG/MAC_BRKG_2018)) + 
+  geom_hex() + 
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "Distribution_of_SNPs_1per_delta_BRKG.png", plot = last_plot(), base_asp = 4, base_height = 4)
+```
+
+```{r}
+loss_freq %>%
+    filter(Loss_BRKG > 75 | Loss_LINC > 78 | Loss_FRMI > 54 | 
+             Loss_CLMB > 51) %>% 
+  mutate(Loss_freq = (Loss_BRKG/MAC_BRKG_2018 + Loss_FRMI/MAC_FRMI_2018 +
+                            Loss_LINC/MAC_LINC_2018 +
+                            Loss_CLMB/MAC_CLMB_2018)/4) %>% 
+  ggplot(aes(x = POS/1000000, y = Loss_freq)) + 
+  geom_hex() + 
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "Distribution_of_SNPs_1per_delta_one_lost_all_four_sites.png", plot = last_plot(), base_asp = 4, base_height = 4)
+frz_freq %>%
+  filter(Loss_BRKG > 159 | Loss_LINC > 179 | Loss_FRMI > 84 | 
+           Loss_CLMB > 94) %>%
+  mutate(Loss_freq = (Loss_BRKG/MAC_BRKG_2018 + Loss_FRMI/MAC_FRMI_2018 +
+                            Loss_LINC/MAC_LINC_2018 +
+                            Loss_CLMB/MAC_CLMB_2018)/4) %>% 
+  ggplot(aes(x = POS/1000000, y = Loss_freq)) +
+  geom_hex() + 
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "Distribution_of_SNPs_1per_delta_all_four_sites.png", plot = last_plot(), base_asp = 4, base_height = 4)
+```

### 2020-02-11 analysis/common-garden-allele-balancing/2_allele_frequencies_common_gardens.Rmd
Subset SNP files into site-specific sets
GWAS outputs FL50 for new 33M set

@@
 ```
+
+```{r}
+loss_freq %>% 
+  group_by(CHR) %>%
+  tally() %>% arrange(n)
+```
+
+## Allele frequencies k=5
+
+For the SNPs in the 1% tail for the most shift in allele frequencies at the freezing sites, these are the minor allele frequencies in each of the old k=5 subpopulations.
+If this is a useful way to look at the data, I will have to redo it for whatever new groupings we decide on.
+
+```{r}
+frz_1per_freq <- readRDS("Five_subpop_allele_frequencies_for_SNPs_with_1percent_tail_frequency_change_freezing_sites.rds")
+
+frz_1per_freq
+```
+
+```{r}
+frz_1per_freq %>%
+  filter(CHR == "Chr02N" & between(POS, 20000000, 40000000) & Loss_CLMB/MAC_CLMB_2018 >= 0.4) %>%
+  ggplot(aes(x = POS/1000000, y = mafmw)) + 
+  geom_hex() + 
+  #facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "Distribution_of_SNPs_1per_delta_LINC.png", plot = last_plot(), base_asp = 4, base_height = 4)
+
+frz_1per_freq %>%
+  filter(Loss_CLMB/MAC_CLMB_2018 >= 0.42) %>%
+  ggplot(aes(x = POS/1000000, y = mafmw)) + 
+  geom_hex() + 
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position (Mb)") + ylab("MAF Midwest") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "Distribution_of_SNPs_1per_delta_LINC.png", plot = last_plot(), base_asp = 4, base_height = 4)
+
+frz_1per_freq <- frz_1per_freq %>%
+  separate(col = "mafgc", into = c("magc", "mafgc"), sep = ":", convert = TRUE)
+frz_1per_freq %>%
+  filter(Loss_CLMB/MAC_CLMB_2018 >= 0.4) %>%
+  ggplot(aes(x = POS/1000000)) + 
+  geom_point(aes(y = maftx), color = "red", alpha = 0.2) +
+  geom_point(aes(y = mafgc), color = "orange", alpha = 0.2) +
+  geom_point(aes(y = mafmw), color = "darkblue", alpha = 0.2) + 
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position (Mb)") + ylab("MAF Midwest") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "MAF_of_5000_SNPs_where_1per_delta_winter_at_CLMB_TX_GC_MW.png", plot = last_plot(), base_asp = 4, base_height = 4)
+
+frz_1per_freq %>%
+  filter(Loss_CLMB/MAC_CLMB_2018 >= 0.4) %>%
+  ggplot(aes(x = POS/1000000)) + 
+  geom_point(aes(y = mafse), color = "#FFCC33", alpha = 0.4) +
+    geom_point(aes(y = mafne), color = "#33FFFF", alpha = 0.2) +
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position (Mb)") + ylab("MAF Midwest") +
+  ylim(c(0,1)) + 
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "MAF_of_5000_SNPs_where_1per_delta_winter_at_CLMB_NEC_SEC.png", plot = last_plot(), base_asp = 4, base_height = 4)
+
+randomfreq <- readRDS("Random_5000SNP_subset_allele_Freq.rds")
+
+randomfreq %>%
+  filter(!grepl("scaffold", CHR)) %>%
+  separate(col = "mafgc", into = c("magc", "mafgc"), sep = ":", convert = TRUE) %>%
+  ggplot(aes(x = POS/1000000)) + 
+  geom_point(aes(y = maftx), color = "red", alpha = 0.2) +
+  geom_point(aes(y = mafse), color = "yellow", alpha = 0.2) +
+  geom_point(aes(y = mafne), color = "lightblue", alpha = 0.2) +
+  geom_point(aes(y = mafgc), color = "orange", alpha = 0.2) +
+  geom_point(aes(y = mafmw), color = "darkblue", alpha = 0.2) + 
+  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
+  scale_fill_viridis(trans = "log10") + 
+  xlab("Position (Mb)") + ylab("MAF Midwest") +
+  theme(legend.position = "right", 
+        axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "MAF_of_5000_random_SNPs_where_maf_20per_one_subpop.png", plot = last_plot(), base_asp = 4, base_height = 4)  
+```

### 2020-02-11 analysis/FL50/gwas_33M_subset_fourway_seven.R

@@ -1,0 +1,123 @@
+library(tidyverse)
+library(bigsnpr)
+library(switchgrassGWAS)
+
+NCORES <- nb_cores()
+
+snp <- snp_attach(file.path("/", "home", "alice", "Github", "pvdiv-genome",
+                            "fourway_seven",
+                            "Pvirgatum_V5_GWAS_630g_33M_fourway_seven.rds"))
+
+G <- snp$genotypes
+CHR <- snp$map$chromosome
+POS <- snp$map$physical.pos
+plants <- snp$fam$sample.ID
+bonferroni <- -log10(0.05/length(POS))
+
+CHRN <- enframe(CHR, name = NULL) %>%
+  dplyr::rename(CHR = .data$value) %>%
+  mutate(CHRN = case_when(.data$CHR == "Chr01K" ~ 1,
+                          .data$CHR == "Chr01N" ~ 2,
+                          .data$CHR == "Chr02K" ~ 3,
+                          .data$CHR == "Chr02N" ~ 4,
+                          .data$CHR == "Chr03K" ~ 5,
+                          .data$CHR == "Chr03N" ~ 6,
+                          .data$CHR == "Chr04K" ~ 7,
+                          .data$CHR == "Chr04N" ~ 8,
+                          .data$CHR == "Chr05K" ~ 9,
+                          .data$CHR == "Chr05N" ~ 10,
+                          .data$CHR == "Chr06K" ~ 11,
+                          .data$CHR == "Chr06N" ~ 12,
+                          .data$CHR == "Chr07K" ~ 13,
+                          .data$CHR == "Chr07N" ~ 14,
+                          .data$CHR == "Chr08K" ~ 15,
+                          .data$CHR == "Chr08N" ~ 16,
+                          .data$CHR == "Chr09K" ~ 17,
+                          .data$CHR == "Chr09N" ~ 18,
+                          TRUE ~ 19
+  ))
+markers <- tibble(CHR = CHR, POS = POS)
+
+svd <- readRDS(file.path("/", "home", "alice", "Github", "pvdiv-genome",
+                         "fourway_seven", "SVD_20PCs_630g_33M_fourway_seven.rds"))
+# load("~/Github/pvdiv-phenology-gxe/data/phenotypes_2018.rda")
+load("~/Github/pvdiv-phenology-gxe/data/phenotypes_2019.rda")
+
[…]

### 2020-02-11 analysis/phenotypes-for-candidate-search/gwas_33M_subset_tensite_twoyear.R

@@ -1,0 +1,113 @@
+library(tidyverse)
+library(bigsnpr)
+library(switchgrassGWAS)
+
+NCORES <- nb_cores()
+
+snp <- snp_attach(file.path("/", "home", "alice", "Github", "pvdiv-genome",
+                            "tensite_twoyear",
+                            "Pvirgatum_V5_GWAS_630g_33M_tensite_twoyear.rds"))
+
+G <- snp$genotypes
+CHR <- snp$map$chromosome
+POS <- snp$map$physical.pos
+plants <- snp$fam$sample.ID
+bonferroni <- -log10(0.05/length(POS))
+
+CHRN <- enframe(CHR, name = NULL) %>%
+  dplyr::rename(CHR = .data$value) %>%
+  mutate(CHRN = case_when(.data$CHR == "Chr01K" ~ 1,
+                          .data$CHR == "Chr01N" ~ 2,
+                          .data$CHR == "Chr02K" ~ 3,
+                          .data$CHR == "Chr02N" ~ 4,
+                          .data$CHR == "Chr03K" ~ 5,
+                          .data$CHR == "Chr03N" ~ 6,
+                          .data$CHR == "Chr04K" ~ 7,
+                          .data$CHR == "Chr04N" ~ 8,
+                          .data$CHR == "Chr05K" ~ 9,
+                          .data$CHR == "Chr05N" ~ 10,
+                          .data$CHR == "Chr06K" ~ 11,
+                          .data$CHR == "Chr06N" ~ 12,
+                          .data$CHR == "Chr07K" ~ 13,
+                          .data$CHR == "Chr07N" ~ 14,
+                          .data$CHR == "Chr08K" ~ 15,
+                          .data$CHR == "Chr08N" ~ 16,
+                          .data$CHR == "Chr09K" ~ 17,
+                          .data$CHR == "Chr09N" ~ 18,
+                          TRUE ~ 19
+  ))
+markers <- tibble(CHR = CHR, POS = POS)
+
+svd <- readRDS(file.path("/", "home", "alice", "Github", "pvdiv-genome",
+                         "tensite_twoyear", "SVD_20PCs_630g_33M_tensite_twoyear.rds"))
+load("~/Github/pvdiv-phenology-gxe/data/phenotypes_2018.rda")
+load("~/Github/pvdiv-phenology-gxe/data/phenotypes_2019.rda")
+
[…]

### 2020-02-11 analysis/2_phenotypes_best_lambda_GC.Rmd
Pick best PCs for 150 phenotypes 10site_2yr; start GWAS

---
title: "Phenotypes_best_lambda_GC"
author: "Alice MacQueen"
date: "2/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

##

```{r}
lambdadf <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/Lambda_GC_BRKG_to_OVTN_FL1.csv")
```

```{r}
df <- lambdadf %>% t(.)

df %>% as_tibble(rownames = "Phenotype") %>%
  filter(!(Phenotype %in% c("NumPCs", "LATITUDE", "LONGITUDE", "ELEVATION"))) %>%
  separate(Phenotype, into = c("SITE", "_", "PHE"), sep = c(4,5)) %>%
  dplyr::select(-"_") %>%
  arrange(PHE, SITE) %>%
  filter(PHE != "DEAD_2018") %>%
  rename(V0 = V1, V1 = V2, V2 = V3, V3 = V4, V4 = V5, V5 = V6, V6 = V7)
```
One/two site: 
PKLE leaf k = 4
KING CHLR k = 3
KING SPAD k = 3
RUST k = 3 or k = 2 for PKLE?
KING/TMPL DAM k = 3

Tensite k = 4:
EM1 k = 4
EM50 k = 4
FL1 k = 4 (OVTN issue but it should be dropped anyway)
FL50 k = 4 (KING k=5...)
HT_PAN_EOS k = 4
T_EM_FL k = 4 (esp for LINC KBSM)
T_GR_EM k = 4
T_GR_FL k = 4

TC_EOS_2018 k = 2 (or possibly k = 4, PKLE)

TC_EOS k = 3 (or possibly k = 4, KBSM, LINC)
GR1.x/y k = 3
GR100.x k = 3 (or k = 5 for PKLE, TMPL, OVTN)
GR50.x  k = 3 (or k = 5 for PKLE, TMPL, OVTN)

### 2020-02-11 analysis/phenotypes-for-candidate-search/gwas_33M_subset_tensite_twoyear.R

@@ -1,6 +1,11 @@
 library(tidyverse)
 library(bigsnpr)
 library(switchgrassGWAS)
+library(AnnotationDbi)
+library(multtest)
+library(viridis)
+library(cowplot)
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
 
 NCORES <- nb_cores()
 
@@ -45,6 +50,11 @@ load("~/Github/pvdiv-phenology-gxe/data/phenotypes_2019.rda")
 
 setwd("~/Github/pvdiv-phenology-gxe/analysis/all-phenotypes-two-years/")
 
+
+# -------------------
+
+
+
 # Prepare phenotype data frame.
 df <- plants %>%
   enframe(value = "PLANT_ID") %>%
@@ -53,14 +63,44 @@ df <- plants %>%
   dplyr::select(-name)
 
 
-
 # Look for the number of PCs to correct for population structure
 # that makes lambda_GC closest to 1.05.
+
 lambdagc <- pvdiv_lambda_GC(df = df, type = "linear", snp = snp,
                             covar = svd, ncores = NCORES,
                             npcs = c(0:8), saveoutput = TRUE)
-## for CHLR, k = 3; for DAM, k = 2 and k = 5
+# From 2_phenotypes_best_lambda_GC.Rmd
+## One/two site:
+# k = 2
+## TC_EOS_2018 k = 2 (or possibly k = 4, PKLE)
+
+# k = 3
+## KING CHLR k = 3
+## KING SPAD k = 3
+## RUST k = 3 or k = 2 for PKLE?
+##  KING/TMPL DAM k = 3
+## TC_EOS k = 3 (or possibly k = 4, KBSM, LINC)
+## GR1.x/y k = 3
+## GR100.x k = 3 (or k = 5 for PKLE, TMPL, OVTN)
+## GR50.x  k = 3 (or k = 5 for PKLE, TMPL, OVTN)
+
+# k = 4
+## Tensite k = 4:
+##   EM1 k = 4
+## EM50 k = 4
+## FL1 k = 4 (OVTN issue but it should be dropped anyway)
+## FL50 k = 4 (KING k=5...)
+## HT_PAN_EOS k = 4
+## T_EM_FL k = 4 (esp for LINC KBSM)
+## T_GR_EM k = 4
+## T_GR_FL k = 4
+## PKLE leaf k = 4
+
+
+
+
 # ------------------------------------
+# Prepare phenotype data frame.
 
[…]

### 2020-02-13 analysis/common-garden-allele-balancing/1_VCF_file_filtering_by_common_garden_site.Rmd
Allow John to repeat freezing site selection analysis

@@
 colSums(!is.na(phenotypes_wide_2019))
+
+libswitch <- metadata %>%
+  dplyr::select(PLANT_ID, LIBRARY)
 i=7
@@
   group_by(PLANT_ID) %>%
-  summarise(count = n())
+  summarise(count = n()) %>%
+  left_join(libswitch, by = "PLANT_ID")
 
-writeLines(site_2019$PLANT_ID, con = paste0(site, "_2019"))
+writeLines(site_2019$LIBRARY, con = paste0("LIBRARY/", site, "_2019"))
 }
@@
   group_by(PLANT_ID) %>%
-  summarise(count = n()) 
-writeLines(site_2018$PLANT_ID, con = paste0(site, "_2018"))
+  summarise(count = n())  %>%
+  left_join(libswitch, by = "PLANT_ID")
+writeLines(site_2018$LIBRARY, con = paste0("LIBRARY/", site, "_2018"))
 }

### 2020-02-13 data-raw/phenotypes_2019.Rmd
Edit 2019 phenotype data to include BIOMASS REGR LD_EOS

@@
          HT_PAN_EOS = as.numeric(HT_PAN_EOS),
+         REGR = as.numeric(REGR),
+         LD_EOS = as.numeric(LD_EOS),
+         MOISTURE = as.numeric(MOISTURE),
+         BIOMASS = as.numeric(BIOMASS),
          DEAD_2018 = ifelse(is.na(DEAD_2018), 0, DEAD_2018),
@@
   dplyr::select(SITE:PLANT_ID, starts_with("NOTE"), everything()) %>%
-  pivot_longer(cols = FRZ_DMG:SPADCOR_158, names_to = "PHE", values_to = "Value") %>%
+  pivot_longer(cols = FRZ_DMG:T_EM_FL, names_to = "PHE", values_to = "Value") %>%
   filter(!is.na(Value)) %>%
@@
 phenotypes_long_2019$SITE <- factor(phenotypes_long_2019$SITE, levels = c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING"))
-phenotypes_long_2019$PHE <- factor(phenotypes_long_2019$PHE, levels = c("GR1", "GR50", "GR100", "EM1", "EM50", "FL1", "FL50", "TC_EOS", "HT_PAN_EOS", "FRZ_DMG", "CHLR_101", "CHLR_121", "CHLR_136", "CHLR_JASON_138", "CHLR_150.", "SPAD_158", "SPADCOR_158"))
+
+unique(phenotypes_long_2019$PHE) # check what phenotypes there are
+phenotypes_long_2019$PHE <- factor(phenotypes_long_2019$PHE, levels = c("GR1", "GR50", "GR100", "EM1", "EM50", "FL1", "FL50", "T_GR_EM", "T_GR_FL", "T_EM_FL", "TC_EOS", "HT_PAN_EOS", "REGR", "LD_EOS", "BIOMASS", "MOISTURE", "FRZ_DMG", "CHLR_101", "CHLR_121", "CHLR_136", "CHLR_JASON_138", "CHLR_150.", "SPAD_158", "SPADCOR_158"))
 
@@
 phenotypes_long_2019 %>%
-  filter(grepl("TC_EOS", PHE) | grepl("HT_PAN_EOS", PHE) | grepl("FRZ", PHE)) %>%
+  filter(grepl("TC_EOS", PHE) | grepl("HT_PAN_EOS", PHE) | grepl("BIOMASS", PHE) | grepl("MOIST", PHE)) %>%
   ggplot(aes(x = Value)) +
@@
 save_plot("2019_Phenotypic_Data_by_Site_2019-12-30.png", plot = last_plot(), base_height = 10, base_width = 6)
+
+phenotypes_long_2019 %>%
+  filter(grepl("MOIST", PHE)) %>% arrange(desc(Value))
+phenotypes_long_2019 %>%
+  filter(grepl("BIOMASS", PHE)) %>% arrange(desc(Value))
 ```
+
+SITE  PLOT_GL   PLOT_LC  ACC  PLANT_ID  PHE  Value
+KBSM	M5710	5710	J476	J476.C	MOISTURE	-1154.137931	NA	NA	
+KBSM	M4412	4412	J024	J024.A	MOISTURE	-195.065147
+KBSM	M6607	6607	J646	J646.A	MOISTURE	973.34815
+
+KBSM	M5710	5710	J476	J476.C	BIOMASS	19439.138
 

### 2020-02-13 data-raw/phenotypes_2019.R

@@ -16,15 +16,15 @@ theme_set(theme_poster)
 # Load all data
 
 load("data/sites.rda")
-load("data/metadata.rda")
-load("data/Taxa.rda")
+metadata <- read_csv(file = "~/Github/pvdiv-climate-gwas/data/Metadata_2020-02.csv")
+#load("data/Taxa.rda")
 
 usa <- map_data("state")
 wb_pheno <- loadWorkbook(file.path("C:", "Users", "ahm543", "Dropbox",
                                    paste0("GWAS DOE Switchgrass, ",
                                           "Consolidated Data"),
                                    paste0("DOE_GWAS_2019_Consolidated ",
-                                          "Pheno Data_12-25-19.xlsx")))
+                                          "Pheno Data_2-12-20.xlsx")))
 lst_pheno <- readWorksheet(wb_pheno, sheet = getSheets(wb_pheno))
 phenos_planting1 <- lst_pheno$`GWAS 2019 Greenup Data`
 phenos_planting2 <- lst_pheno$`GWAS 2019 Core Phenotype Data`
@@ -70,14 +70,15 @@ phenos_2019 <- phenos_planting1 %>%
   mutate(PLANT_ID = ifelse(ACC == "AP13",
                            "AP13",
                            PLANT_ID)) %>%
-  dplyr::select(-TC_EOS_2018, -`PLANT_.ID_GL`, -INDV, -(LD_EOS:BIOMASS))
+  dplyr::select(-TC_EOS_2018, -`PLANT_.ID_GL`, -INDV, -(VERIFY:PLOT_GL_VER))
+  # %>%  filter(PLANT_ID == "AP13")
 
 # Drop 31 plants with confusion on whether plant died for now.
 
 phenotypes_2019 <- phenos_2019 %>%
-  filter(!(DEAD_2018 %in% c(1,2)) | is.na(DEAD_2018)) %>% # 1 = dead in 2018, and 2 = not a virgatum plant. Drop 1's and 2's. Keep 0's and NA's.
-  filter(SRV != "N") %>% # Drop individuals that did not survive
-  dplyr::select(-VERIFY, -PLOT_GL_CHK)
+  filter(!(DEAD_2018 %in% c(1,2)) | is.na(DEAD_2018)) %>% # 1 = dead in 2018,
[…]

### 2020-02-13 data-raw/phenotypes_2019.R
Second change to MOISTURE for new phenotypes for 2019

@@ -112,7 +112,7 @@ phenotypes_2019 <- phenotypes_2019 %>%
          T_EM_FL = FL50 - EM50,
          EM50 = ifelse(EM50 - EM1 < 0, NA, EM50),
          FL50 = ifelse(FL50 - FL1 < 0, NA, FL50),
-         MOISTURE = ifelse(between(BIOMASS, 0, 100), MOISTURE, NA),
+         MOISTURE = ifelse(between(MOISTURE, 0, 100), MOISTURE, NA),
          BIOMASS = ifelse(BIOMASS < 10000, BIOMASS, NA)
   )
 

### 2020-02-20 presentations/Plots_for_DOE_Poster.Rmd
Figures for DOE Poster Final

---
title: "Plots_for_DOE_poster"
author: "Alice MacQueen"
date: "2/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(PopGenome)
library(geosphere)
library(cowplot)
library(viridis)
library(maps)
library(ggmap)
library(XLConnect)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
theme_set(theme_poster)
usa <- map_data("usa")
load("~/Github/pvdiv-phenology-gxe/data/sites.rda")
```

```{r}
metadata <- read_csv("~/Github/pvdiv-phenology-gxe/data/metadata_hierarchical_kof3.csv")

```

##

# Map of Diversity Panel

```{r}
metaplot <- metadata %>%
  dplyr::select(PLANT_ID:CO_ecotype, LATITUDE, LONGITUDE) %>%
  filter(LIB_GROWN == "Y" | LIB_BIOCLIM == "Y") %>%
  mutate(GENETIC_SUBPOP = case_when(is.na(Genetic_subpop_50per) ~ "Unknown",
                                    Genetic_subpop_50per == "Eastcoast" ~ "Atlantic",
                                    Genetic_subpop_50per == "Texas" ~ "Gulf",
                                    TRUE ~ Genetic_subpop_50per),
         ECOTYPE = case_when(is.na(Ecotype) ~ "Unknown", 
                             Ecotype == "TX" ~ "Lowland",
                             Ecotype == "Texas" ~ "Lowland",
                             TRUE ~ Ecotype)) %>%
  dplyr::select(PLANT_ID, GENETIC_SUBPOP, ECOTYPE, LATITUDE, LONGITUDE, everything())

metaplot %>% group_by(GENETIC_SUBPOP) %>% tally()

metaplot$GENETIC_SUBPOP <- factor(metaplot$GENETIC_SUBPOP, levels = c("Midwest", "Atlantic", "Gulf", "Admixed", "Unknown"))
metaplot$ECOTYPE <- factor(metaplot$ECOTYPE, levels = c("Upland", "Coastal", "Lowland", "Unknown"))

theme_set(theme_poster)
ggplot() + 
  geom_polygon(data = usa, aes(x=long, y = lat, group = group), fill = "white", 
               color = "black") +
  coord_quickmap(xlim = c(-105, -66), ylim = c(25, 50)) +
  geom_jitter(data = metaplot, aes(x = LONGITUDE, y = LATITUDE, 
                                  shape = GENETIC_SUBPOP, fill = ECOTYPE, 
                                  alpha = ECOTYPE,
                                  color = GENETIC_SUBPOP:ECOTYPE), 
             size = 5, width = 0.5, height = 0.5) + 
  labs(x = "Longitude", y = "Latitude") + 
  theme(legend.position = "right") +
  scale_shape_manual(values = c(21,22,24,4,3)) + 
  scale_fill_manual(values = c("#280B54FF","#FAC127FF", "#D44842FF", "white")) +
  scale_color_manual(values = c(rep("black", 10), "#280B54FF","#FAC127FF", "#D44842FF", "white", "#280B54FF","#FAC127FF", "#D44842FF", "white"), guide = FALSE) +
  scale_alpha_manual(values = c(0.7, 0.7, 0.7, 0.7), guide = FALSE)
save_plot(filename = "Map_subpops_and_Ecotypes.svg", plot = last_plot(), base_height = 5)

library(scales)
show_col(viridis(8, option = "B"))
```

## Plot of Sites and planting numbers

```{r}

plants_2018 <- phenotypes_2018 %>%
  pivot_longer(cols = BRKG_DEAD_2018:TMPL_TC_EOS_2018, names_to = "SITE_PHE", values_to = "Measurement") %>%
  separate(col = SITE_PHE, into = c("SITE", "_", "PHE"), sep = c(4,5)) %>%
  dplyr::select(-(LATITUDE:ELEVATION), -`_`) %>%
  dplyr::filter(!is.na(Measurement)) %>%
  group_by(SITE, PLANT_ID) %>% tally() %>%
  summarise(Phenotyped_2018 = n())

died_winter <- phenotypes_2018 %>%
  pivot_longer(cols = ends_with("DEAD_2018"), names_to = "SITE_PHE", values_to = "Measurement") %>%
  filter(Measurement == 0) %>%
  separate(col = SITE_PHE, into = c("SITE", "_", "PHE"), sep = c(4,5)) %>%
  dplyr::select(-`_`) %>%
  dplyr::filter(!is.na(Measurement)) %>%
  group_by(SITE, PLANT_ID) %>% tally() %>%
  summarise(Alive_2019 = n())

plants_2019 <- phenotypes_wide_2019 %>%
  pivot_longer(cols = BRKG_BIOMASS:TMPL_TC_EOS, names_to = "SITE_PHE", values_to = "Measurement") %>%
  separate(col = SITE_PHE, into = c("SITE", "_", "PHE"), sep = c(4,5)) %>%
  dplyr::select(-`_`) %>%
  dplyr::filter(!is.na(Measurement)) %>%
  group_by(SITE, PLANT_ID) %>% tally() %>%
  summarise(Phenotyped_2019 = n())

siteplot <- sites %>%
  left_join(plants_2018, by = c("Site" = "SITE")) %>%
  left_join(died_winter, by = c("Site" = "SITE")) %>%
  left_join(plants_2019, by = c("Site" = "SITE")) %>%
  mutate(Winter_killed_2018 = Phenotyped_2018 - Alive_2019,
         Not_phenotyped_2019 = Alive_2019 - Phenotyped_2019)

theme_set(theme_poster)
map1 <- ggplot() + 
  geom_polygon(data = usa, aes(x=long, y = lat, group = group), fill = "white", 
               color = "black") +
  coord_quickmap(xlim = c(-105, -66), ylim = c(25, 50)) +
  geom_point(data = siteplot, aes(x = Longitude, y = Latitude, 
                                  fill = Phenotyped_2018), 
             size = 4, shape = 21) + 
  labs(x = "Longitude", y = "Latitude") + 
  scale_fill_viridis(option = "B", limits = c(250, 650)) + 
  theme(legend.position = c(0.95, 0.3), plot.title = element_text(size = 16)) + labs(title = "Plants Grown in 2018")

map2 <- ggplot() + 
  geom_polygon(data = usa, aes(x=long, y = lat, group = group), fill = "white", 
               color = "black") +
  coord_quickmap(xlim = c(-105, -66), ylim = c(25, 50)) +
  geom_point(data = siteplot, aes(x = Longitude, y = Latitude, 
                                  fill = Winter_killed_2018), 
             size = 4, shape = 21) + 
  labs(x = "Longitude", y = "Latitude") + 
  scale_fill_viridis(option = "B", limits = c(0, 180)) + 
  theme(legend.position = c(0.95, 0.3), plot.title = element_text(size = 16)) + labs(title = "Plants Killed Winter 2018")

plot_grid(map1, map2)
save_plot(filename = "Map_plants_winter_killed_2018.svg", plot = last_plot(), base_height = 5)

library(scales)
show_col(viridis(8, option = "B"))

```

```{r}
winterlost <- read_csv("twoHit.Chi2PeakIntervals.csv")

winterlost %>%
  ggplot(aes(x = peak)) + geom_histogram() + facet_wrap(~chr)
```

# Frequencies lost 
```{r}
frz_freq <- readRDS("../analysis/common-garden-allele-balancing/SNPs_with_1percent_tail_frequency_change_freezing_sites.rds")

theme_set(theme_poster)
frz_freq %>%
  filter(Loss_BRKG > 159 | Loss_LINC > 179 | Loss_FRMI > 84 | 
           Loss_CLMB > 94) %>%
  mutate(Loss_freq = (Loss_BRKG/MAC_BRKG_2018 + Loss_FRMI/MAC_FRMI_2018 +
                            Loss_LINC/MAC_LINC_2018 +
                            Loss_CLMB/MAC_CLMB_2018)/4) %>% 
  filter(Loss_freq > 0.5) %>%
  ggplot(aes(x = POS/1000000, y = Loss_freq)) +
  geom_point(aes(color = CHR), alpha = 0.35) + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10", option = "B") + 
  xlab("Position (Mb)") + ylab("Fraction of Minor Allele Lost") +
    scale_x_continuous(expand = c(0.12, 0.12)) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = c(rep(c("#1B0C42FF", "grey"), 9)))

save_plot(filename = "Distribution_of_SNPs_1per_delta_all_four_sites_lost_50per.png", plot = last_plot(), base_asp = 4, base_height = 4)

```

# Phenology histograms

```{r}
phenotypes_long_2019 <- phenotypes_wide_2019 %>%
  pivot_longer(cols = BRKG_BIOMASS:TMPL_TC_EOS, names_to = "SITE_PHE", values_to = "Measurement") %>%
  separate(col = SITE_PHE, into = c("SITE", "_", "PHE"), sep = c(4,5)) %>%
  dplyr::select(-`_`)

phenotypes_long_2019$SITE <- factor(phenotypes_long_2019$SITE, levels = c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING"))

unique(phenotypes_long_2019$PHE) # check what phenotypes there are
phenotypes_long_2019$PHE <- factor(phenotypes_long_2019$PHE, levels = c("GR1", "GR50", "GR100", "EM1", "EM50", "FL1", "FL50", "T_GR_EM", "T_GR_FL", "T_EM_FL", "TC_EOS", "HT_PAN_EOS", "REGR", "LD_EOS", "BIOMASS", "MOISTURE", "FRZ_DMG", "CHLR_101", "CHLR_121", "CHLR_136", "CHLR_JASON_138", "CHLR_150.", "SPAD_158", "SPADCOR_158"))

phenology_long_2019 <- phenotypes_long_2019 %>%
  filter(!(grepl("CHLR", PHE)) & !(grepl("SPAD", PHE)) & !grepl("FRZ", PHE) & !is.na(PHE) & !grepl("TC_EOS", PHE) & !grepl("HT_PAN_EOS", PHE) & !grepl("BIOMASS", PHE) & !grepl("REGR", PHE) & !grepl("MOISTURE", PHE) & !grepl("LD_EOS", PHE) & !grepl("T_", PHE))

pheplot <- phenotypes_long_2019 %>%
  filter(PHE %in% c("TC_EOS", "HT_PAN_EOS", "BIOMASS")) %>%
  mutate(PHE = case_when(PHE == "TC_EOS" ~ "Tiller Count",
                         PHE == "HT_PAN_EOS" ~ "Plant Height (cm)",
                         PHE == "BIOMASS" ~ "Biomass (kg)"))

```

```{r}
library(lubridate)
phenplot <- phenology_long_2019 %>%
  filter(PHE %in% c("GR50")) %>%
  mutate(PHE = case_when(PHE == "GR50" ~ "Greenup",
                         PHE == "EM50" ~ "Panicle Emergence",
                         PHE == "FL50" ~ "Flowering"))
phenplot$PHE <- factor(phenplot$PHE, levels = c("Greenup", "Panicle Emergence", "Flowering"))

phenplot %>%
  left_join(metaplot, by = "PLANT_ID") %>%
  mutate(Days = as_date(Measurement)) %>%
  ggplot(aes(x = Days)) + 
  geom_histogram(aes(fill = ECOTYPE)) + 
  facet_grid(SITE~PHE, scales = "free") +
  scale_fill_manual(values = c("#280B54FF","#FAC127FF", "#D44842FF", "lightgrey", "grey")) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) + xlab("Value") + ylab("Count of Individuals") +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 45, hjust = 1)) +#,
        #panel.spacing = unit(2, "lines")) +
  xlab("Date") + 
  ylab("Number of Individuals") +
  geom_vline(xintercept = c(120), linetype = 2) +
  geom_vline(xintercept = c(181), linetype = 3, color = "grey20") +
  geom_vline(xintercept = c(243), linetype = 4, color = "grey80") +
  scale_y_continuous(breaks = c(10, 50, 150))

save_plot("Greenup_Data_by_Site_and_Ecotype_2020-02-19.svg", plot = last_plot(), base_height = 6, base_asp = 1.3/3)
```

```{r}
pheplot$PHE <- factor(pheplot$PHE, levels = c("Biomass (kg)", "Tiller Count", "Plant Height (cm)"))

pheplot %>%
  left_join(metaplot, by = "PLANT_ID") %>%
  mutate(Measurement = ifelse(PHE == "Biomass (kg)", Measurement/1000, Measurement)) %>%
  ggplot(aes(x = Measurement)) + 
  geom_histogram(aes(fill = ECOTYPE)) + 
  facet_grid(SITE~PHE, scales = "free") +
  scale_fill_manual(values = c("#280B54FF","#FAC127FF", "#D44842FF", "lightgrey", "grey")) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) + xlab("Value") + ylab("Count of Individuals") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1)) +#,
        #panel.spacing = unit(2, "lines")) +
  xlab("") + 
  ylab("Number of Individuals") +
  scale_y_continuous(breaks = c(10, 50, 150))

save_plot("EOS_Phenotype_Data_by_Site_and_Ecotype_2020-02-19.svg", plot = last_plot(), base_height = 6, base_asp = 1.3)

pheplot2 <- phenotypes_long_2019 %>%
  filter(PHE %in% c("BIOMASS")) %>%
  mutate(PHE = case_when(PHE == "TC_EOS" ~ "Tiller Count",
                         PHE == "HT_PAN_EOS" ~ "Plant Height (cm)",
                         PHE == "BIOMASS" ~ "Biomass (kg)"))
pheplot2 %>%
  left_join(metaplot, by = "PLANT_ID") %>%
  mutate(Measurement = ifelse(PHE == "Biomass (kg)", Measurement/1000, Measurement)) %>%
  ggplot(aes(x = Measurement)) + 
  geom_histogram(aes(fill = ECOTYPE)) + 
  facet_grid(SITE~PHE, scales = "free") +
  scale_fill_manual(values = c("#280B54FF","#FAC127FF", "#D44842FF", "lightgrey", "grey")) +
  theme(legend.position = "none") +#,
        #panel.spacing = unit(2, "lines")) +
  xlab("") + 
  geom_vline(xintercept = c(2.0), linetype = 2) +
  ylab("Number of Individuals") +
  scale_y_continuous(breaks = c(10, 50, 150))

save_plot("Biomass_Phenotype_Data_by_Site_and_Ecotype_2020-02-19.svg", plot = last_plot(), base_height = 6, base_asp = 1.3/3)
```

# Manhattan plots for GR50 in KBSM and biomass in TMPL
```{r}
gwas_data <- readRDS("GWAS_datatable_TMPL_BIOMASS_5_PCs.rds")
gwas_data[,FDR_adj := p.adjust(pvalue, method = "BH")]
FDRthreshhi <- gwas_data %>%
    as_tibble() %>%
    filter(between(FDR_adj, 0.000501, 1)) %>%  # 5% FDR threshold currently.
    summarise(thresh = max(log10p))
  FDRthreshlo <- gwas_data %>%
    as_tibble() %>%
    filter(between(FDR_adj, 0, 0.000499)) %>%  # 5% FDR threshold currently.
    summarise(thresh = min(log10p))
  if(FDRthreshhi$thresh[1] > 0 & FDRthreshlo$thresh[1] > 0){
    FDRthreshold = (FDRthreshhi$thresh[1] + FDRthreshlo$thresh[1])/2
  } else if(FDRthreshhi$thresh[1] > 0){
    FDRthreshold = FDRthreshhi$thresh[1]
  } else if(FDRthreshlo$thresh[1] > 0){
    FDRthreshold = FDRthreshlo$thresh[1]
  } else {
    FDRthreshold = NA
  }

  ggmanobject1 <- gwas_data %>%
    filter(log10p > 1) %>%
    ggplot(aes(x = pos, y = log10p)) +
    geom_hline(yintercept = c(10,15,20), color = "lightgrey") +
    geom_point(aes(color = chr, fill = chr)) +
    geom_hline(yintercept = FDRthreshold, color = "black", linetype = 2,
               size = 1) +
    facet_wrap(~ chr, nrow = 1, scales = "free_x",
               strip.position = "bottom") +
    #scale_color_viridis(option = "B", end = 0.8, discrete = TRUE) +
    #scale_fill_viridis(option = "B", end = 0.8, discrete = TRUE) +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          panel.background = element_rect(fill=NA), legend.position = "none") +
    labs(x = "Chromosome", y = "-log10(p value)") +
    scale_x_continuous(expand = c(0.22, 0.22)) +
    scale_color_manual(values = rep(c("#1B0C42FF", "grey"),9), guide = FALSE)
   #scale_shape_manual(values = rep(c(21,22),9), guide = FALSE)
  theme_set(theme_poster)
  save_plot(paste0("Manhattan_TMPL_BIOMASS_5_PCs_FDR_", str_replace_all(Sys.time(), ":", "."),
                   ".png"),
            plot = ggmanobject1, base_aspect_ratio = 4, base_height = 3)
```

```{r}
gwas_data <- readRDS("GWAS_datatable_KBSM_GR50_3_PCs_.rds")
gwas_data[,FDR_adj := p.adjust(pvalue, method = "BH")]
FDRthreshhi <- gwas_data %>%
    as_tibble() %>%
    filter(between(FDR_adj, 0.0501, 1)) %>%  # 5% FDR threshold currently.
    summarise(thresh = max(log10p))
  FDRthreshlo <- gwas_data %>%
    as_tibble() %>%
    filter(between(FDR_adj, 0, 0.0499)) %>%  # 5% FDR threshold currently.
    summarise(thresh = min(log10p))
  if(FDRthreshhi$thresh[1] > 0 & FDRthreshlo$thresh[1] > 0){
    FDRthreshold = (FDRthreshhi$thresh[1] + FDRthreshlo$thresh[1])/2
  } else if(FDRthreshhi$thresh[1] > 0){
    FDRthreshold = FDRthreshhi$thresh[1]
  } else if(FDRthreshlo$thresh[1] > 0){
    FDRthreshold = FDRthreshlo$thresh[1]
  } else {
    FDRthreshold = NA
  }

  ggmanobject1 <- gwas_data %>%
    filter(log10p > 1) %>%
    ggplot(aes(x = pos, y = log10p)) +
    geom_hline(yintercept = c(10,15), color = "lightgrey") +
    geom_point(aes(color = chr, fill = chr)) +
    geom_hline(yintercept = FDRthreshold, color = "black", linetype = 2,
               size = 1) +
    facet_wrap(~ chr, nrow = 1, scales = "free_x",
               strip.position = "bottom") +
    #scale_color_viridis(option = "B", end = 0.8, discrete = TRUE) +
    #scale_fill_viridis(option = "B", end = 0.8, discrete = TRUE) +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          panel.background = element_rect(fill=NA), legend.position = "none") +
    labs(x = "Chromosome", y = "-log10(p value)") +
    scale_x_continuous(expand = c(0.22, 0.22)) +
    scale_color_manual(values = rep(c("#1B0C42FF", "grey"),9), guide = FALSE)
   #scale_shape_manual(values = rep(c(21,22),9), guide = FALSE)
  theme_set(theme_poster)
  save_plot(paste0("Manhattan_KBSM_GR50_3_PCs_FDR_", str_replace_all(Sys.time(), ":", "."),
                   ".png"),
            plot = ggmanobject1, base_aspect_ratio = 4, base_height = 3)
```

### 2020-02-20 data-raw/metadata_hierarchical_kof3.R
New hierarchical k=3 metadata and nice Manhattan plots

@@ -1,0 +1,25 @@
+library(tidyverse)
+
+metadata732 <- read_csv("~/Github/pvdiv-climate-gwas/data/Metadata_2020-02.csv")
+load("~/Github/pvdiv-phenology-gxe/data/metadata.rda")
+ind_needed <- read_csv("~/Github/pvdiv-climate-gwas/results-33M-SNPs/Individuals_needed_for_climate_and_phenotype_GWAS.csv")
+
+metadata732 <- metadata732 %>%
+  add_row(PLANT_ID = "AP13", LIBRARY = "REF", Genetic_subpop_50per = "Texas",
+          Genetic_subpop_95per = "Texas", Texas_Q = 1, Eastcoast_Q = 0,
+          Midwest_DA = 0, LIB_GROWN = "Y", include_envGWAS = "N") %>%
+  dplyr::select(PLANT_ID:LIB_GROWN)
+
+metadata <- metadata %>%
+  left_join(ind_needed) %>%
+  dplyr::select(PLANT_ID, LIBRARY, LIB_GROWN, everything()) %>%
+  mutate(LIB_GROWN = ifelse(is.na(LIB_GROWN), "N", LIB_GROWN),
+         LIB_BIOCLIM = ifelse(is.na(include_clim), "N", include_clim)) %>%
+  dplyr::select(PLANT_ID, LIBRARY, LIB_GROWN, LIB_BIOCLIM, everything())
+
+metadata %>%
+  left_join(metadata732) %>%
+  unique() %>%
+  dplyr::select(PLANT_ID, LIBRARY, LIB_GROWN, LIB_BIOCLIM, Genetic_subpop_50per:Midwest_DA, Ecotype, Ecotype_PC1, UP_ecotype:CO_ecotype, everything()) %>%
+  dplyr::select(-(JL_genetic_subpop:include_clim)) %>%
+  write_csv(., path = "data/metadata_hierarchical_kof3.csv")

### 2020-03-01 analysis/DEAD_2018/DEAD_tensite_twoyear.R
Add GWAS for winter killed plants 2018-2019

@@ -1,0 +1,159 @@
+library(tidyverse)
+library(bigsnpr)
+library(switchgrassGWAS)
+library(AnnotationDbi)
+library(multtest)
+library(viridis)
+library(cowplot)
+library(data.table)
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+
+NCORES <- nb_cores()
+
+snp <- snp_attach(file.path("/", "home", "alice", "Github", "pvdiv-genome",
+                            "tensite_twoyear",
+                            "Pvirgatum_V5_GWAS_630g_33M_tensite_twoyear.rds"))
+
+G <- snp$genotypes
+CHR <- snp$map$chromosome
+POS <- snp$map$physical.pos
+plants <- snp$fam$sample.ID
+bonferroni <- -log10(0.05/length(POS))
+
+CHRN <- enframe(CHR, name = NULL) %>%
+  dplyr::rename(CHR = .data$value) %>%
+  mutate(CHRN = case_when(.data$CHR == "Chr01K" ~ 1,
+                          .data$CHR == "Chr01N" ~ 2,
+                          .data$CHR == "Chr02K" ~ 3,
+                          .data$CHR == "Chr02N" ~ 4,
+                          .data$CHR == "Chr03K" ~ 5,
+                          .data$CHR == "Chr03N" ~ 6,
+                          .data$CHR == "Chr04K" ~ 7,
+                          .data$CHR == "Chr04N" ~ 8,
+                          .data$CHR == "Chr05K" ~ 9,
+                          .data$CHR == "Chr05N" ~ 10,
+                          .data$CHR == "Chr06K" ~ 11,
+                          .data$CHR == "Chr06N" ~ 12,
+                          .data$CHR == "Chr07K" ~ 13,
+                          .data$CHR == "Chr07N" ~ 14,
+                          .data$CHR == "Chr08K" ~ 15,
+                          .data$CHR == "Chr08N" ~ 16,
+                          .data$CHR == "Chr09K" ~ 17,
+                          .data$CHR == "Chr09N" ~ 18,
+                          TRUE ~ 19
+  ))
+markers <- tibble(CHR = CHR, POS = POS)
+
+svd <- readRDS(file.path("/", "home", "alice", "Github", "pvdiv-genome",
+                         "tensite_twoyear", "SVD_20PCs_630g_33M_tensite_twoyear.rds"))
+load("~/Github/pvdiv-phenology-gxe/data/phenotypes_2018.rda")
+
[…]

### 2020-03-01 analysis/DEAD_2018/DEAD_tensite_twoyear_logistic.R

@@ -1,0 +1,169 @@
+library(tidyverse)
+library(bigsnpr)
+library(switchgrassGWAS)
+library(AnnotationDbi)
+library(multtest)
+library(viridis)
+library(cowplot)
+library(data.table)
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+
+NCORES <- nb_cores()
+
+snp <- snp_attach(file.path("/", "home", "alice", "Github", "pvdiv-genome",
+                            "tensite_twoyear",
+                            "Pvirgatum_V5_GWAS_630g_33M_tensite_twoyear.rds"))
+
+G <- snp$genotypes
+CHR <- snp$map$chromosome
+POS <- snp$map$physical.pos
+plants <- snp$fam$sample.ID
+bonferroni <- -log10(0.05/length(POS))
+
+CHRN <- enframe(CHR, name = NULL) %>%
+  dplyr::rename(CHR = .data$value) %>%
+  mutate(CHRN = case_when(.data$CHR == "Chr01K" ~ 1,
+                          .data$CHR == "Chr01N" ~ 2,
+                          .data$CHR == "Chr02K" ~ 3,
+                          .data$CHR == "Chr02N" ~ 4,
+                          .data$CHR == "Chr03K" ~ 5,
+                          .data$CHR == "Chr03N" ~ 6,
+                          .data$CHR == "Chr04K" ~ 7,
+                          .data$CHR == "Chr04N" ~ 8,
+                          .data$CHR == "Chr05K" ~ 9,
+                          .data$CHR == "Chr05N" ~ 10,
+                          .data$CHR == "Chr06K" ~ 11,
+                          .data$CHR == "Chr06N" ~ 12,
+                          .data$CHR == "Chr07K" ~ 13,
+                          .data$CHR == "Chr07N" ~ 14,
+                          .data$CHR == "Chr08K" ~ 15,
+                          .data$CHR == "Chr08N" ~ 16,
+                          .data$CHR == "Chr09K" ~ 17,
+                          .data$CHR == "Chr09N" ~ 18,
+                          TRUE ~ 19
+  ))
+markers <- tibble(CHR = CHR, POS = POS)
+
+svd <- readRDS(file.path("/", "home", "alice", "Github", "pvdiv-genome",
+                         "tensite_twoyear", "SVD_20PCs_630g_33M_tensite_twoyear.rds"))
[…]

### 2020-03-01 analysis/common-garden-allele-balancing/2_allele_frequencies_common_gardens.Rmd

@@
   facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
-  scale_fill_viridis(trans = "log10") + 
+  scale_fill_viridis(trans = "log10", option = "B") + 
   xlab("Position (Mb)") + ylab("Minor Allele Loss") +

### 2020-03-01 data-raw/phenotypes_2019.Rmd

@@
 load("../data/metadata.rda")
+meta732 <- read_csv("~/Github/pvdiv-climate-gwas/data/Metadata_2020-02.csv")
+meta732 <- meta732 %>%
+  add_row(PLANT_ID = "AP13", LIBRARY = "REF", Genetic_subpop_50per = "Texas", Genetic_subpop_95per = "Texas", Ecotype = "Texas", Texas_Q = 1, Eastcoast_Q = 0, Midwest_DA = 0, LIB_GROWN = "Y")
+
 load("../data/Taxa.rda")
@@
 The notes might be relevant for outlier analysis.
+```{r}
+load("~/Github/pvdiv-phenology-gxe/data/phenotypes_2019.rda")
+
+phenotypes_long_2019 <- phenotypes_wide_2019 %>%
+  pivot_longer(cols = BRKG_BIOMASS:TMPL_TC_EOS, names_to = "SITE_PHE", values_to = "Value") %>%
+  separate(col = "SITE_PHE", into = c("SITE", "_", "PHE"), sep = c(4,5), remove = FALSE) %>%
+  dplyr::select(-"_")
+```
 
@@
 
+### BIOMASS
+
+```{r}
+meta732$Genetic_subpop_50per <- factor(meta732$Genetic_subpop_50per, levels = c("Texas", "Eastcoast", "Midwest"))
+
+phenotypes_long_2019 %>%
+  filter(grepl("BIOMASS", PHE)) %>%
+  left_join(meta732, by = "PLANT_ID") %>%
+  filter(!(Genetic_subpop_50per %in% c("Admixed", "NA")), !is.na(Genetic_subpop_50per)) %>%
+  ggplot(aes(x = Value)) +
+  geom_histogram(aes(fill = Genetic_subpop_50per), binwidth = 500) + 
+  facet_grid(SITE~PHE, scales = "free_y") +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = -1) +
+  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) + xlab("Value") + ylab("Count of Individuals") +
+  geom_vline(xintercept = 2800, linetype = 3) +
+  scale_y_continuous(breaks = c(10, 50, 100))
+
+save_plot("2019_Phenotypic_Data_by_Site_and_Genetic_Subpop_scale_2020-02-14.png", plot = last_plot(), base_height = 7, base_asp = 0.4)
+
+```
+
 # Do dates make sense for each SITE*PLANT_ID?

### 2020-03-06 R/3_subset_SNP_files.Rmd
Move R and Rmd files to R directory and rda files to data

---
title: "3_subset_SNP_files"
author: "Alice MacQueen"
date: "2/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(switchgrassGWAS)
library(viridis)
library(cowplot)
library(bigsnpr)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

##

```{r}
snp_subsets <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/SNP_subsets_33M_markers.rds")

rds <- snp_readBed("Pvirgatum_V5_GWAS_630g_33M.bed")
snp <- snp_attach(rds)
```

```{r}
snp_subsets
```

# Impute 33M SNP file

```{r}

G2 <- snp_fastImputeSimple(snp$genotypes, method = "random", ncores = NCORES)
snp$genotypes$code256 <- CODE_IMPUTE_PRED
big_counts(snp$genotypes, ind.col = 1:10) # Make sure imputation succeeded
snp <- snp_save(snp)

```

# Use previously determined SNP sets as selection criteria for subsets.

These need to be vectors of the row numbers in the Taxa dataframe that you have phenotypes for.

```{r}  
tensite_twoyear_sel <- which(snp_subsets$SNP_tensite_twoyear == "Y")
fourway_seven_sel <- which(snp_subsets$SNP_fourway_seven == "Y")
microbiome_sel <- which(snp_subsets$SNP_microbiome == "Y")
missing_frz_sel <- which(snp_subsets$SNP_missing_frz == "Y")

tensite_twoyear_subset <- subset(snp, ind.col = tensite_twoyear_sel)
fourway_seven_subset <- subset(snp, ind.col = fourway_seven_sel)
microbiome_subset <- subset(snp, ind.col = microbiome_sel)
missing_frz_subset <- subset(snp, ind.colstr = missing_frz_sel)

```

```{r}
freq <- readRDS("~/Github/pvdiv-climate-gwas/results-upland-vs-lowland/2DSFS/Alleles_five_populations_subpops0.95.rds")
missing_frz_freq <- snp_subsets %>% filter(SNP_missing_frz == "Y") %>% left_join(freq, by = c("CHR", "POS"))

 frz_freq_subpop <- frz_freq %>% left_join(freq)
saveRDS(frz_freq_subpop, "Five_subpop_allele_frequencies_for_SNPs_with_1percent_tail_frequency_change_freezing_sites.rds")

```

```{r}
ind_gwas <- read_csv("~/Github/pvdiv-climate-gwas/results-33M-SNPs/Individuals_needed_for_climate_and_phenotype_GWAS.csv")
```

# Adding AP13 to vcf files

For each chromosome, sequentially (to cut down on the size of the intermediate files):

gunzip Pvirgatum_V5_GWAS_758g_Chr01K_33dot9M.vcf.gz
head -n 9 Pvirgatum_V5_GWAS_758g_Chr01K_33dot9M.vcf > Pvirgatum_header.txt
# Edit this file so it has AP13 as the last column; make sure there is a \n after AP13
# Save Pvirgatum_header.txt as Pvirgatum_V5_GWAS_758g_33M.vcf

tail -n +9 Pvirgatum_V5_GWAS_758g_Chr01K_33dot9M.vcf | sed 's/$/\t0,0:50,0/' >> Pvirgatum_V5_GWAS_758g_33M.vcf
rm Pvirgatum_V5_GWAS_758g_Chr01K_33dot9M.vcf

gunzip Pvirgatum_V5_GWAS_758g_Chr01N_33dot9M.vcf.gz
tail -n +9 Pvirgatum_V5_GWAS_758g_Chr01N_33dot9M.vcf | sed 's/$/\t0,0:50,0/' >> Pvirgatum_V5_GWAS_758g_33M.vcf
rm Pvirgatum_V5_GWAS_758g_Chr01N_33dot9M.vcf

gunzip Pvirgatum_V5_GWAS_758g_Chr02K_33dot9M.vcf.gz
tail -n +9 Pvirgatum_V5_GWAS_758g_Chr02K_33dot9M.vcf | sed 's/$/\t0,0:50,0/' >> Pvirgatum_V5_GWAS_758g_33M.vcf
rm Pvirgatum_V5_GWAS_758g_Chr02K_33dot9M.vcf

gunzip Pvirgatum_V5_GWAS_758g_Chr02N_33dot9M.vcf.gz
tail -n +9 Pvirgatum_V5_GWAS_758g_Chr02N_33dot9M.vcf | sed 's/$/\t0,0:50,0/' >> Pvirgatum_V5_GWAS_758g_33M.vcf
rm Pvirgatum_V5_GWAS_758g_Chr02N_33dot9M.vcf

gunzip Pvirgatum_V5_GWAS_758g_Chr03K_33dot9M.vcf.gz
tail -n +9 Pvirgatum_V5_GWAS_758g_Chr03K_33dot9M.vcf | sed 's/$/\t0,0:50,0/' >> Pvirgatum_V5_GWAS_758g_33M.vcf
rm Pvirgatum_V5_GWAS_758g_Chr03K_33dot9M.vcf

gunzip Pvirgatum_V5_GWAS_758g_Chr03N_33dot9M.vcf.gz
tail -n +9 Pvirgatum_V5_GWAS_758g_Chr03N_33dot9M.vcf | sed 's/$/\t0,0:50,0/' >> Pvirgatum_V5_GWAS_758g_33M.vcf
rm Pvirgatum_V5_GWAS_758g_Chr03N_33dot9M.vcf

gunzip Pvirgatum_V5_GWAS_758g_Chr04K_33dot9M.vcf.gz
tail -n +9 Pvirgatum_V5_GWAS_758g_Chr04K_33dot9M.vcf | sed 's/$/\t0,0:50,0/' >> Pvirgatum_V5_GWAS_758g_33M.vcf
rm Pvirgatum_V5_GWAS_758g_Chr04K_33dot9M.vcf

gunzip Pvirgatum_V5_GWAS_758g_Chr04N_33dot9M.vcf.gz
tail -n +9 Pvirgatum_V5_GWAS_758g_Chr04N_33dot9M.vcf | sed 's/$/\t0,0:50,0/' >> Pvirgatum_V5_GWAS_758g_33M.vcf
rm Pvirgatum_V5_GWAS_758g_Chr04N_33dot9M.vcf

gunzip Pvirgatum_V5_GWAS_758g_Chr05K_33dot9M.vcf.gz
tail -n +9 Pvirgatum_V5_GWAS_758g_Chr05K_33dot9M.vcf | sed 's/$/\t0,0:50,0/' >> Pvirgatum_V5_GWAS_758g_33M.vcf
rm Pvirgatum_V5_GWAS_758g_Chr05K_33dot9M.vcf

gunzip Pvirgatum_V5_GWAS_758g_Chr05N_33dot9M.vcf.gz
tail -n +9 Pvirgatum_V5_GWAS_758g_Chr05N_33dot9M.vcf | sed 's/$/\t0,0:50,0/' >> Pvirgatum_V5_GWAS_758g_33M.vcf
rm Pvirgatum_V5_GWAS_758g_Chr05N_33dot9M.vcf

gunzip Pvirgatum_V5_GWAS_758g_Chr06K_33dot9M.vcf.gz
tail -n +9 Pvirgatum_V5_GWAS_758g_Chr06K_33dot9M.vcf | sed 's/$/\t0,0:50,0/' >> Pvirgatum_V5_GWAS_758g_33M.vcf
rm Pvirgatum_V5_GWAS_758g_Chr06K_33dot9M.vcf

gunzip Pvirgatum_V5_GWAS_758g_Chr06N_33dot9M.vcf.gz
tail -n +9 Pvirgatum_V5_GWAS_758g_Chr06N_33dot9M.vcf | sed 's/$/\t0,0:50,0/' >> Pvirgatum_V5_GWAS_758g_33M.vcf
rm Pvirgatum_V5_GWAS_758g_Chr06N_33dot9M.vcf

gunzip Pvirgatum_V5_GWAS_758g_Chr07K_33dot9M.vcf.gz
tail -n +9 Pvirgatum_V5_GWAS_758g_Chr07K_33dot9M.vcf | sed 's/$/\t0,0:50,0/' >> Pvirgatum_V5_GWAS_758g_33M.vcf
rm Pvirgatum_V5_GWAS_758g_Chr07K_33dot9M.vcf

gunzip Pvirgatum_V5_GWAS_758g_Chr07N_33dot9M.vcf.gz
tail -n +9 Pvirgatum_V5_GWAS_758g_Chr07N_33dot9M.vcf | sed 's/$/\t0,0:50,0/' >> Pvirgatum_V5_GWAS_758g_33M.vcf
rm Pvirgatum_V5_GWAS_758g_Chr07N_33dot9M.vcf

gunzip Pvirgatum_V5_GWAS_758g_Chr08K_33dot9M.vcf.gz
tail -n +9 Pvirgatum_V5_GWAS_758g_Chr08K_33dot9M.vcf | sed 's/$/\t0,0:50,0/' >> Pvirgatum_V5_GWAS_758g_33M.vcf
rm Pvirgatum_V5_GWAS_758g_Chr08K_33dot9M.vcf

gunzip Pvirgatum_V5_GWAS_758g_Chr08N_33dot9M.vcf.gz
tail -n +9 Pvirgatum_V5_GWAS_758g_Chr08N_33dot9M.vcf | sed 's/$/\t0,0:50,0/' >> Pvirgatum_V5_GWAS_758g_33M.vcf
rm Pvirgatum_V5_GWAS_758g_Chr08N_33dot9M.vcf

gunzip Pvirgatum_V5_GWAS_758g_Chr09K_33dot9M.vcf.gz
tail -n +9 Pvirgatum_V5_GWAS_758g_Chr09K_33dot9M.vcf | sed 's/$/\t0,0:50,0/' >> Pvirgatum_V5_GWAS_758g_33M.vcf
rm Pvirgatum_V5_GWAS_758g_Chr09K_33dot9M.vcf

gunzip Pvirgatum_V5_GWAS_758g_Chr09N_33dot9M.vcf.gz
tail -n +9 Pvirgatum_V5_GWAS_758g_Chr09N_33dot9M.vcf | sed 's/$/\t0,0:50,0/' >> Pvirgatum_V5_GWAS_758g_33M.vcf
rm Pvirgatum_V5_GWAS_758g_Chr09N_33dot9M.vcf

gzip Pvirgatum_V5_GWAS_758g_33M.vcf

###

sed 's/$/\t0,0:50,0/' Pvirgatum_header.txt > Pvirgatum_header2.txt

Then join all chromosomes

Combine individual chromosome files:

  	ls * > filelist.txt
    bcftools concat -n -f filelist.txt -o Pvirgatum_V5_GWAS_757g_33M.vcf.gz -O z

### 2020-03-06 R/Fast_QQ_plot.Rmd
Move R and Rmd files to R directory and rda files to data

---
title: "Fast QQ plots"
author: "Alice MacQueen"
date: "3/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(switchgrassGWAS)
library(AnnotationDbi)
library(multtest)
library(viridis)
library(cowplot)
library(data.table)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")

NCORES <- nb_cores()
```

##

```{r}
gwas_data2 <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/all-phenotypes-two-years/GWAS_datatable_KBSM_GR50_3_PCs_.rds")

```

# Faster QQ plotting

```{r}
round2 <- function(x, at) ceiling(x / at) * at
round_xy <- function(x, y, cl = NA, cu = NA, roundby = 0.001){
  expected <- round2(x, at = roundby)
  observed <- round2(y, at = roundby)
  if(!is.na(cl[1]) & !is.na(cu[1])){
    clower <- round2(cl, at = roundby)
    cupper <- round2(cu, at = roundby)
    tp <- cbind(expected, observed, clower, cupper)
    return(tp[!duplicated(tp),])
  } else {
    tp <- cbind(expected, observed)
    return(tp[!duplicated(tp),])
  }
}

  # Save a QQ plot
  ggsave(snp_qq(gwas[which(!is.na(gwas$score)),]),
         filename = paste0("QQplot_", names(df1)[2], "_", PCdf1$NumPCs,
                           "_PCs.png"), width = 4,
         height = 4, path = file.path("/", "home", "alice", "Github",
                                      "pvdiv-phenology-gxe", "analysis",
                                      "all-phenotypes-two-years"))
```

```{r}
#' Create a quantile-quantile plot with ggplot2.
#'
#' Assumptions:
#'   - Expected P values are uniformly distributed.
#'   - Confidence intervals assume independence between tests.
#'     We expect deviations past the confidence intervals if the tests are
#'     not independent.
#'     For example, in a genome-wide association study, the genotype at any
#'     position is correlated to nearby positions. Tests of nearby genotypes
#'     will result in similar test statistics.
#'
#' @param ps Vector of p-values.
#' @param ci Size of the confidence interval, 95% by default.
#' @param lambdaGC Logical. Add the Genomic Control coefficient as subtitle to the plot?
#' @return A ggplot2 plot.
#' @examples
#' library(ggplot2)
#' gg_qqplot(runif(1e2)) + theme_grey(base_size = 24)
gg_qqplot <- function(ps, ci = 0.95, lambdaGC = FALSE) {
  n  <- length(ps)
  df <- data.frame(
    observed = -log10(sort(ps)),
    expected = -log10(ppoints(n)),
    clower   = -log10(qbeta(p = (1 - ci) / 2, shape1 = 1:n, shape2 = n:1)),
    cupper   = -log10(qbeta(p = (1 + ci) / 2, shape1 = 1:n, shape2 = n:1))
  )
  df_round <- round_xy(df$expected, df$observed, cl = df$clower, cu = df$cupper)
  log10Pe <- expression(paste("Expected -log"[10], plain("("), italic(p-value),
                              plain(")")))
  log10Po <- expression(paste("Observed -log"[10], plain("("), italic(p-value),
                              plain(")")))
  p1 <- ggplot(as_tibble(df_round)) +
    geom_point(aes(expected, observed), shape = 1, size = 1) +
    geom_abline(intercept = 0, slope = 1, size = 1.5, color = "red") +
    geom_line(aes(expected, cupper), linetype = 2) +
    geom_line(aes(expected, clower), linetype = 2) +
    xlab(log10Pe) +
    ylab(log10Po)

  if (lambdaGC) {
      xtr <- log10(ps)
      MEDIAN <- log10(0.5)
      f.opt <- function(x) (x - MEDIAN)
      xtr_p <- stats::median(xtr) / stats::uniroot(f.opt, interval = range(xtr),
                                                   check.conv = TRUE, 
                                                   tol = tol)$root
      lamGC <- signif(xtr_p)
      expr <- substitute(expression(lambda[GC] == l), list(l = lamGC))
      p1 + labs(subtitle = eval(expr))
    } else {
      p1
    }
}

  xtr <- -stats::na.omit(gwas$log10p)
  MEDIAN <- log10(0.5)
  f.opt <- function(x) (x - MEDIAN)
  stats::median(xtr) / stats::uniroot(f.opt, 
                                              interval = range(xtr),
                                      check.conv = TRUE, tol = tol)$root

  # Save a QQ plot
  ggsave(gg1,
         filename = paste0("QQplot_", "test2", ".png"), width = 4,
         height = 4, path = file.path("/", "home", "alice", "Github",
                                      "pvdiv-phenology-gxe", "analysis",
                                      "all-phenotypes-two-years"))
```

### 2020-03-06 R/Plots_for_DOE_Poster.Rmd
Move R and Rmd files to R directory and rda files to data

--- 2020-02-20 presentations/Plots_for_DOE_Poster.Rmd
@@
 library(scales)
-show_col(viridis(8, option = "B"))
+show_col(viridis(2, option = "D", end = 0.8))
+
+```
+
+# 4X vs 8X for Jason
+
+```{r}
+
+wb_pheno <- loadWorkbook(file.path("C:", "Users", "ahm543", "Dropbox",
+                                   paste0("GWAS DOE Switchgrass, ",
+                                          "Consolidated Data"),
+                                   paste0("DOE_GWAS_Master Plant List.xlsx")))
+lst_pheno <- readWorksheet(wb_pheno, sheet = getSheets(wb_pheno))
+planting_2018 <- lst_pheno$`GWAS 2018 Master Plant List`
+
+ploidyplot <- planting_2018 %>%
+  left_join(metadata) %>%
+   mutate(PLOIDY = ifelse(ACC == "AP13", "4X", PLOIDY)) %>%
+   filter(!is.na(PLOIDY)) %>%
+    mutate(GENETIC_SUBPOP = case_when(is.na(Genetic_subpop_50per) ~ "Unknown",
+                                    Genetic_subpop_50per == "Eastcoast" ~ "Atlantic",
+                                    Genetic_subpop_50per == "Texas" ~ "Gulf",
+                                    TRUE ~ Genetic_subpop_50per),
+         ECOTYPE = case_when(is.na(Ecotype) ~ "Unknown", 
+                             Ecotype == "TX" ~ "Lowland",
+                             Ecotype == "Texas" ~ "Lowland",
+                             TRUE ~ Ecotype))
+
+ploidyplot$PLOIDY <- factor(ploidyplot$PLOIDY, levels = c("8X", "4X"))
+ploidyplot$SITE <- factor(ploidyplot$SITE, levels = c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING"))
+ploidyplot$GENETIC_SUBPOP <- factor(ploidyplot$GENETIC_SUBPOP, levels = c("Midwest", "Atlantic", "Gulf", "Admixed", "Unknown"))
+ploidyplot$ECOTYPE <- factor(ploidyplot$ECOTYPE, levels = c("Upland", "Coastal", "Lowland", "Unknown"))
+
+ploidyplot %>%
+  #group_by(SITE, PLOIDY) %>% tally() %>%
+   ggplot(aes(x = SITE)) + 
+   geom_bar(aes(fill = PLOIDY)) +
+   scale_fill_manual(values = c("black", "darkgreen")) + 
+  xlab("Sites (North to South)") + ylab("Number of Plants") + 
+  theme(legend.position = "right",
+        axis.text.x = element_text(angle = 45, hjust = 1))
+
+save_plot(filename = "Ploidy of Plants at Common Garden Sites.png", plot = last_plot())
+
+ploidyplot %>%
+  #group_by(SITE, PLOIDY) %>% tally() %>%
+   ggplot(aes(x = SITE)) + 
+   geom_bar(aes(fill = GENETIC_SUBPOP)) +
+   scale_fill_manual(values = c("darkblue", "orange", "red", "grey", "lightgrey")) + 
+  xlab("Sites (North to South)") + ylab("Number of Plants") + 
+  theme(legend.position = "right",
+        axis.text.x = element_text(angle = 45, hjust = 1))
+
+save_plot(filename = "Ploidy of Plants at Common Garden Sites.png", plot = last_plot())
 
@@
 theme_set(theme_poster)
-frz_freq %>%
+frzlostplot <- frz_freq %>%
   filter(Loss_BRKG > 159 | Loss_LINC > 179 | Loss_FRMI > 84 | 
@@
                             Loss_CLMB/MAC_CLMB_2018)/4) %>% 
-  filter(Loss_freq > 0.5) %>%
+  filter(Loss_freq > 0.42) %>%
   ggplot(aes(x = POS/1000000, y = Loss_freq)) +
   geom_point(aes(color = CHR), alpha = 0.35) + 
-  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
-  scale_fill_viridis(trans = "log10", option = "B") + 
-  xlab("Position (Mb)") + ylab("Fraction of Minor Allele Lost") +
-    scale_x_continuous(expand = c(0.12, 0.12)) +
+  facet_wrap(~CHR, scales = "free_x", nrow = 1,
+            strip.position = "bottom") + 
+  #scale_fill_viridis(trans = "log10", option = "B") + 
+  xlab("Chromosome") + ylab("Fraction of Minor Allele Lost") +
+    scale_x_continuous(expand = c(0.03, 0.03)) +
   theme(legend.position = "none", 
-        axis.text.x = element_text(angle = 45, hjust = 1)) +
+        axis.text.x = element_blank(),
+          axis.ticks.x = element_blank()) +
+        #axis.text.x = element_text(angle = 45, hjust = 1)) +
   scale_color_manual(values = c(rep(c("#1B0C42FF", "grey"), 9)))
 
-save_plot(filename = "Distribution_of_SNPs_1per_delta_all_four_sites_lost_50per.png", plot = last_plot(), base_asp = 4, base_height = 4)
+save_plot(filename = paste0("Distribution_of_SNPs_1per_delta_all_four_sites_lost_50per", str_replace_all(Sys.time(), ":", "."),
+                   ".png"), plot = frzlostplot, base_asp = 4, base_height = 4)
 
@@
     as_tibble() %>%
-    filter(between(FDR_adj, 0.000501, 1)) %>%  # 5% FDR threshold currently.
+    filter(between(FDR_adj, 0.1001, 1)) %>%  # 5% FDR threshold currently.
     summarise(thresh = max(log10p))
@@
     as_tibble() %>%
-    filter(between(FDR_adj, 0, 0.000499)) %>%  # 5% FDR threshold currently.
+    filter(between(FDR_adj, 0, 0.0999)) %>%  # 5% FDR threshold currently.
+    summarise(thresh = min(log10p))
+  if(FDRthreshhi$thresh[1] > 0 & FDRthreshlo$thresh[1] > 0){
+    FDRthreshold = (FDRthreshhi$thresh[1] + FDRthreshlo$thresh[1])/2
+  } else if(FDRthreshhi$thresh[1] > 0){
+    FDRthreshold = FDRthreshhi$thresh[1]
+  } else if(FDRthreshlo$thresh[1] > 0){g
+    FDRthreshold = FDRthreshlo$thresh[1]
+  } else {
+    FDRthreshold = NA
+  }
+
+  ggmanobject1 <- gwas_data %>%
+    filter(log10p > 1) %>%
+    ggplot(aes(x = pos, y = log10p)) +
+    geom_hline(yintercept = c(10,15,20), color = "lightgrey") +
+    geom_point(aes(color = chr, fill = chr)) +
+    geom_point(aes(color = chr, fill = chr)) +
+    #geom_hline(yintercept = FDRthreshold, color = "black", linetype = 2,
+    #           size = 0.7) +
+    geom_hline(yintercept = bonferroni, color = "#FAC127FF", linetype = 2,
+               size = 1) +
+    facet_wrap(~ chr, nrow = 1, scales = "free_x",
+               strip.position = "bottom") +
+    #scale_color_viridis(option = "B", end = 0.8, discrete = TRUE) +
+    #scale_fill_viridis(option = "B", end = 0.8, discrete = TRUE) +
+    theme(axis.text.x = element_blank(),
+          axis.ticks.x = element_blank(),
+          panel.background = element_rect(fill=NA), legend.position = "none") +
+    labs(x = "Chromosome", y = "-log10(p value)") +
+    scale_x_continuous(expand = c(0.18, 0.18)) +
+    scale_color_manual(values = rep(c("#1B0C42FF", "grey"),9), guide = FALSE)
+   #scale_shape_manual(values = rep(c(21,22),9), guide = FALSE)
+  theme_set(theme_poster)
+  save_plot(paste0("Manhattan_TMPL_BIOMASS_5_PCs_FDR_", str_replace_all(Sys.time(), ":", "."),
+                   ".png"),
+            plot = ggmanobject1, base_aspect_ratio = 4, base_height = 4)
+```
+
+```{r}
+gwas_data2 <- readRDS("GWAS_datatable_KBSM_GR50_3_PCs_.rds")
+gwas_data2[,FDR_adj := p.adjust(pvalue, method = "BH")]
+FDRthreshhi <- gwas_data2 %>%
+    as_tibble() %>%
+    filter(between(FDR_adj, 0.1001, 1)) %>%  # 5% FDR threshold currently.
+    summarise(thresh = max(log10p))
+  FDRthreshlo <- gwas_data2 %>%
+    as_tibble() %>%
+    filter(between(FDR_adj, 0, 0.0999)) %>%  # 5% FDR threshold currently.
     summarise(thresh = min(log10p))
@@
 
-  ggmanobject1 <- gwas_data %>%
+  ggmanobject1 <- gwas_data2 %>%
     filter(log10p > 1) %>%
     ggplot(aes(x = pos, y = log10p)) +
-    geom_hline(yintercept = c(10,15,20), color = "lightgrey") +
+    geom_hline(yintercept = c(10,15), color = "lightgrey") +
     geom_point(aes(color = chr, fill = chr)) +
-    geom_hline(yintercept = FDRthreshold, color = "black", linetype = 2,
+    #geom_hline(yintercept = FDRthreshold, color = "black", linetype = 2,
+    #           size = 0.7) +
+    geom_hline(yintercept = bonferroni, color = "#FAC127FF", linetype = 2,
                size = 1) +
@@
     labs(x = "Chromosome", y = "-log10(p value)") +
-    scale_x_continuous(expand = c(0.22, 0.22)) +
+    scale_x_continuous(expand = c(0.18, 0.18)) +
     scale_color_manual(values = rep(c("#1B0C42FF", "grey"),9), guide = FALSE)
@@
   theme_set(theme_poster)
-  save_plot(paste0("Manhattan_TMPL_BIOMASS_5_PCs_FDR_", str_replace_all(Sys.time(), ":", "."),
+  save_plot(paste0("Manhattan_KBSM_GR50_3_PCs_FDR_", str_replace_all(Sys.time(), ":", "."),
                    ".png"),
-            plot = ggmanobject1, base_aspect_ratio = 4, base_height = 3)
+            plot = ggmanobject1, base_aspect_ratio = 4, base_height = 4)
 ```
 
+# Manhattan plots for ASV's with 10% FDR & Bonferroni
 ```{r}
-gwas_data <- readRDS("GWAS_datatable_KBSM_GR50_3_PCs_.rds")
-gwas_data[,FDR_adj := p.adjust(pvalue, method = "BH")]
+asvs <- c("ASV89_2_PCs", "ASV113_3_PCs", "ASV1197_3_PCs")
+for(i in seq_along(asvs)){
+gwas_data <- readRDS(paste0("../GWAS_datatable_", asvs[i], "_.rds"))
+#gwas_data[,FDR_adj := p.adjust(pvalue, method = "BH")]
 FDRthreshhi <- gwas_data %>%
     as_tibble() %>%
-    filter(between(FDR_adj, 0.0501, 1)) %>%  # 5% FDR threshold currently.
+    filter(between(FDR_adj, 0.1001, 1)) %>%  # 5% FDR threshold currently.
     summarise(thresh = max(log10p))
@@
     as_tibble() %>%
-    filter(between(FDR_adj, 0, 0.0499)) %>%  # 5% FDR threshold currently.
+    filter(between(FDR_adj, 0, 0.0999)) %>%  # 5% FDR threshold currently.
     summarise(thresh = min(log10p))
@@
     FDRthreshold = FDRthreshhi$thresh[1]
-  } else if(FDRthreshlo$thresh[1] > 0){
+  } else if(FDRthreshlo$thresh[1] > 0){g
     FDRthreshold = FDRthreshlo$thresh[1]
@@
     geom_point(aes(color = chr, fill = chr)) +
+    geom_point(aes(color = chr, fill = chr)) +
     geom_hline(yintercept = FDRthreshold, color = "black", linetype = 2,
+               size = 0.7) +
+    geom_hline(yintercept = bonferroni, color = "#FAC127FF", linetype = 2,
                size = 1) +
@@
   theme_set(theme_poster)
-  save_plot(paste0("Manhattan_KBSM_GR50_3_PCs_FDR_", str_replace_all(Sys.time(), ":", "."),
+  save_plot(paste0("../../presentations/Nice_Manhattans/Manhattan_", asvs[i], "_Bonferroni_10percent_FDR_", str_replace_all(Sys.time(), ":", "."),
                    ".png"),
-            plot = ggmanobject1, base_aspect_ratio = 4, base_height = 3)
+            plot = ggmanobject1, base_aspect_ratio = 4, base_height = 4)
+}
 ```

### 2020-03-06 R/DEAD_tensite_twoyear_linear.R

@@ -1,0 +1,162 @@
+library(tidyverse)
+library(bigsnpr)
+library(switchgrassGWAS)
+library(AnnotationDbi)
+library(multtest)
+library(viridis)
+library(cowplot)
+library(data.table)
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+
+NCORES <- nb_cores()
+
+snp <- snp_attach(file.path("/", "home", "alice", "Github", "pvdiv-genome",
+                            "tensite_twoyear",
+                            "Pvirgatum_V5_GWAS_630g_33M_tensite_twoyear.rds"))
+
+G <- snp$genotypes
+CHR <- snp$map$chromosome
+POS <- snp$map$physical.pos
+plants <- snp$fam$sample.ID
+bonferroni <- -log10(0.05/length(POS))
+
+CHRN <- enframe(CHR, name = NULL) %>%
+  dplyr::rename(CHR = .data$value) %>%
+  mutate(CHRN = case_when(.data$CHR == "Chr01K" ~ 1,
+                          .data$CHR == "Chr01N" ~ 2,
+                          .data$CHR == "Chr02K" ~ 3,
+                          .data$CHR == "Chr02N" ~ 4,
+                          .data$CHR == "Chr03K" ~ 5,
+                          .data$CHR == "Chr03N" ~ 6,
+                          .data$CHR == "Chr04K" ~ 7,
+                          .data$CHR == "Chr04N" ~ 8,
+                          .data$CHR == "Chr05K" ~ 9,
+                          .data$CHR == "Chr05N" ~ 10,
+                          .data$CHR == "Chr06K" ~ 11,
+                          .data$CHR == "Chr06N" ~ 12,
+                          .data$CHR == "Chr07K" ~ 13,
+                          .data$CHR == "Chr07N" ~ 14,
+                          .data$CHR == "Chr08K" ~ 15,
+                          .data$CHR == "Chr08N" ~ 16,
+                          .data$CHR == "Chr09K" ~ 17,
+                          .data$CHR == "Chr09N" ~ 18,
+                          TRUE ~ 19
+  ))
+markers <- tibble(CHR = CHR, POS = POS)
+
+svd <- readRDS(file.path("/", "home", "alice", "Github", "pvdiv-genome",
+                         "tensite_twoyear", "SVD_20PCs_630g_33M_tensite_twoyear.rds"))
+load("~/Github/pvdiv-phenology-gxe/data/phenotypes_2018.rda")
+
[…]

### 2020-03-06 analysis/DEAD_2018/DEAD_tensite_twoyear_logistic.R

@@ -1,169 +1,0 @@
-library(tidyverse)
-library(bigsnpr)
-library(switchgrassGWAS)
-library(AnnotationDbi)
-library(multtest)
-library(viridis)
-library(cowplot)
-library(data.table)
-source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
-
-NCORES <- nb_cores()
-
-snp <- snp_attach(file.path("/", "home", "alice", "Github", "pvdiv-genome",
-                            "tensite_twoyear",
-                            "Pvirgatum_V5_GWAS_630g_33M_tensite_twoyear.rds"))
-
-G <- snp$genotypes
-CHR <- snp$map$chromosome
-POS <- snp$map$physical.pos
-plants <- snp$fam$sample.ID
-bonferroni <- -log10(0.05/length(POS))
-
-CHRN <- enframe(CHR, name = NULL) %>%
-  dplyr::rename(CHR = .data$value) %>%
-  mutate(CHRN = case_when(.data$CHR == "Chr01K" ~ 1,
-                          .data$CHR == "Chr01N" ~ 2,
-                          .data$CHR == "Chr02K" ~ 3,
-                          .data$CHR == "Chr02N" ~ 4,
-                          .data$CHR == "Chr03K" ~ 5,
-                          .data$CHR == "Chr03N" ~ 6,
-                          .data$CHR == "Chr04K" ~ 7,
-                          .data$CHR == "Chr04N" ~ 8,
-                          .data$CHR == "Chr05K" ~ 9,
-                          .data$CHR == "Chr05N" ~ 10,
-                          .data$CHR == "Chr06K" ~ 11,
-                          .data$CHR == "Chr06N" ~ 12,
-                          .data$CHR == "Chr07K" ~ 13,
-                          .data$CHR == "Chr07N" ~ 14,
-                          .data$CHR == "Chr08K" ~ 15,
-                          .data$CHR == "Chr08N" ~ 16,
-                          .data$CHR == "Chr09K" ~ 17,
-                          .data$CHR == "Chr09N" ~ 18,
-                          TRUE ~ 19
-  ))
-markers <- tibble(CHR = CHR, POS = POS)
-
-svd <- readRDS(file.path("/", "home", "alice", "Github", "pvdiv-genome",
-                         "tensite_twoyear", "SVD_20PCs_630g_33M_tensite_twoyear.rds"))
[…]

### 2020-03-13 R/Flowering_time_environmental_index_analysis.Rmd
Start flowering time environmental index analysis

---
title: "Environmental Index"
author: "Alice MacQueen"
date: "3/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(corrplot)
library(zoo)
```

##

```{r}
weather <- read_rds("~/Github/pvdiv-phenotypes/data/weather_10_sites_2019.rds") %>%
  dplyr::select(DOY:DYLN) %>%
  group_by(SITE) %>%
  mutate(TMIN = ifelse(TMIN < -50, NA, TMIN),
         TMAX_SM = ifelse(is.na(TMAX),
                          (na.locf(TMAX) + na.locf(TMAX, fromLast = TRUE))/2,
                          TMAX),
         TMIN_SM = ifelse(is.na(TMIN),
                          (na.locf(TMIN) + na.locf(TMIN, fromLast = TRUE))/2,
                          TMIN),
         RAIN_SM = ifelse(is.na(RAIN), 0, RAIN))

metadata <- read_rds("~/Github/pvdiv-phenotypes/data/metadata.rds")
phe_all <- read_rds("~/Github/pvdiv-phenotypes/data/Phenotypes_long_non_Blackwell_replants_pre_and_post_2019_replant.rds")
sites <- read_rds("~/Github/pvdiv-phenotypes/data/sites.rds")
fourway_sites <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL")
```

# My analysis based on Li's code (below)

From my pvdiv-phenotypes repository, I already had generated a data frame of weather data for all 10 GWAS sites with daylength included.

Actually, get and load smoothed weather data from pvdiv-phenotypes.

Make D2F variable which is FL50 - GR50.
```{r}
gwas_d2f <- phe_all %>%
  filter(SITE %in% fourway_sites) %>%
  filter(PHE %in% c("GR50", "FL50")) %>%
  pivot_wider(names_from = PHE, values_from = MEAS) %>%
  mutate(D2F = FL50 - GR50) %>%
  filter(!is.na(D2F))

weather <- weather %>%
  filter(SITE %in% fourway_sites)
```
fourway_seven set should include these 7 site_year combinations:
BRKG_2019
CLMB_2019
KBSM_2019
KING_2019
LINC_2019
PKLE_2019
STIL_2019

Make GDD & PTT using a range of base temperatures
```{r}

Tbase <- seq(0, 6)
Topt <- seq(8,30)

GR_start <- gwas_d2f %>% 
  group_by(SITE) %>%
  summarise(GR_start = round(mean(GR50)))
FL_end <- gwas_d2f %>% 
  group_by(SITE) %>%
  summarise(FL_end = round(mean(FL50)))

GDD_PT <- NULL

for (j in Tbase){
  for (k in Topt){

    tmp <- weather %>% 
      left_join(GR_start, by = "SITE") %>%
      left_join(FL_end, by = "SITE") %>%
      group_by(SITE, Day_of_year) %>%
      filter(between(Day_of_year, GR_start, FL_end)) %>% 
      ungroup %>%
      mutate(TMEAN = (TMAX_SM + TMIN_SM) / 2,
             GDD = ifelse(TMEAN >= j, TMEAN - j, 0),
             GDD = ifelse(TMEAN >= k, k - j, GDD),
             PT = GDD * DYLN) %>%
      group_by(SITE) %>%
      summarize_at(vars(GDD, PT), funs(sum))
    colnames(tmp)[2:3] = c(paste0('GDD_', j, "_", k), paste0('PT_', j, "_", k))
    GDD_PT = c(GDD_PT, tmp)
    }
}

  GDD_PT <- GDD_PT %>% 
    as.data.frame()

  GDD_PT <- GDD_PT %>%
    select(SITE, starts_with("GDD"), starts_with("PT"))

  gwas_avegdd <- gwas_d2f %>%
    left_join(GDD_PT, by = "SITE") %>%
    ungroup()

###correlation

M = cor(gwas_avegdd %>% select(D2F, starts_with('GDD')) %>% filter(complete.cases(.)))

# corrplot::corrplot.mixed(M)

(best1 <- which(M[1,] == max(M[1,2:ncol(M)])))

M[1,best1]

M = cor(gwas_avegdd %>% select(D2F, starts_with('PT')) %>% filter(complete.cases(.)))

# corrplot::corrplot.mixed(M)

(best1 <- which(M[1,] == max(M[1,2:ncol(M)])))

M[1,best1]
```

# Li's code

```{r}
rm(list=ls())

dir = "~/Desktop/Phenology_4way/"

setwd(paste0(dir,"0-data/"))

library(tidyverse)

###1. combine weather data together

##function to calculate daylength

day.length.hrs <- function(day, latitude) {

  daylength.coeff <- asin(0.39795*cos(0.2163108 + 2*atan(0.9671396*tan(0.00860*(day - 186)))))

  daylength.hrs <- 24 - (24/pi)*acos((sin(0.8333*pi/180) + sin(latitude*pi/180)*sin(daylength.coeff))/(cos(latitude*pi/180)*cos(daylength.coeff)))

  return(daylength.hrs)

}

##read into the latitude of the 7 sites to calculate the day length

Lat = read.csv('4WCR_GWAS_PlantingLocations.csv')

wthfiles = list.files(pattern='[A-Z]_2019.csv')

wth = lapply(wthfiles, function(x){

  tmp = read.csv(x)

  SITE = strsplit(x,'_',fixed = T)[[1]][1]

  lat = Lat %>% filter(Site.Code==SITE)

  tmp = tmp %>% dplyr::select(MONTH, DAY,DOY, TMAX, TMIN, RAIN)%>%

    mutate(DL = round(day.length.hrs(DOY, lat$Latitude),2)) %>% mutate(SITE =SITE, YEAR='2019')

})

wth_2019 = do.call(rbind, wth)

write.csv(wth_2019,'WeatherFile_2019_7_Sites.csv',row.names = F)

###2. read in GR50, FL50 phenology data for reation norm plot

FL50 = read.csv('FL50_7Sites.csv')

FL50 = FL50 %>% separate(ENVI, c('SITE','YEAR'))%>% mutate(D2F = FL50-GR50)

Tbase = seq(8, 13)

Topt = seq(30, 35)

FL50_Tbase = NULL

for (i in 1:nrow(FL50)){

  s = FL50$SITE[i]

  gr = floor(FL50$GR50[i])

  fl = floor(FL50$FL50[i])

  GDD_PT = NULL

  for (j in Tbase){

    for (k in Topt){

      tmp = wth_2019 %>% filter(SITE==s)%>% filter(DOY>= gr & DOY <= fl) %>%

            mutate(TMEAN=(TMAX+TMIN)/2)%>%  mutate(GDD =ifelse(TMEAN>=j, TMEAN-j, 0))%>%

            mutate(GDD =ifelse(TMEAN>=k, k-j, GDD) )%>% mutate(PT = GDD*DL)%>%

            summarize_at(vars(GDD,PT),funs(sum))

            colnames(tmp) = c(paste0('GDD_',j, k), paste0('PT_',j, k))

            GDD_PT = c(GDD_PT,tmp)

            }

  }

    GDD_PT = GDD_PT %>% as.data.frame()

    FL50_Tbase = rbind(FL50_Tbase, GDD_PT)

}

FL50 = FL50 %>% bind_cols(FL50_Tbase)

###correlation

M = cor(FL50 %>% select(D2F, matches('PT')) %>% filter(complete.cases(.)))

corrplot::corrplot.mixed(M)
```

### 2020-03-13 R/Visualize_BIOMASS_mash_GxE.Rmd
Start flowering time environmental index analysis

---
title: "biomass mash"
author: "Alice MacQueen"
date: "3/12/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(switchgrassGWAS)
library(ashr)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

##

```{r, cache = TRUE, include = FALSE}
m <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/all-phenotypes-two-years/BIOMASS/Strong_Effects25684SNPs.rds")
gxe <- get_GxE(m)
covariances <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/all-phenotypes-two-years/BIOMASS/Mash_data_driven_covariances_25684.rds")
```

```{r, include=FALSE, cache = TRUE}
#' Return the estimated mixture proportions
#' @param m the mash result
#' @param dimension indicates whether you want the mixture proportions for the covariances, grid, or all
#' @return a named vector containing the estimated mixture proportions.
#' @details If the fit was done with `usepointmass=TRUE` then the first element of the returned vector will correspond to the null, and the remaining elements
#' to the non-null covariance matrices.
#' Suppose the fit was done with $K$ covariances and a grid of length $L$.
#' If `dimension=cov` then the returned vector will be of length $K$ (or $K+1$ if
#' `usepointmass=TRUE`).  If `dimension=grid` then the returned vector will be of length $L$ (or $L+1$).
#' If `dimension=all` then the returned vector will be of length $LK$ (or $LK+1$).
#' The names of the vector will be informative for which combination each element corresponds to.
#' @importFrom ashr get_fitted_g
#' @export
get_estimated_pi = function(m, dimension = c("cov","grid","all")){
  dimension = match.arg(dimension)
  if(dimension=="all"){
    get_estimated_pi_no_collapse(m)
  } else {
    g = get_fitted_g(m)
    pihat = g$pi
    pihat_names = NULL
    pi_null = NULL

    if(g$usepointmass){
      pihat_names=c("null",pihat_names)
      pi_null = pihat[1]
      pihat = pihat[-1]
    }

    pihat = matrix(pihat,nrow=length(g$Ulist))
    if(dimension=="cov"){
      pihat = rowSums(pihat)
      pihat_names = c(pihat_names,names(g$Ulist))
    } else if(dimension=="grid"){
      pihat = colSums(pihat)
      pihat_names = c(pihat_names,1:length(g$grid))
    }

    pihat = c(pi_null,pihat)
    names(pihat) = pihat_names
    return(pihat)
  }
}

get_estimated_pi_no_collapse = function(m){
  g = get_fitted_g(m)
  pihat = g$pi
  names(pihat) = names(expand_cov(g$Ulist, g$grid, g$usepointmass))
  pihat
}

```

```{r, cache = TRUE}
covar <- enframe(get_estimated_pi(m), name = "Covariances")

covar[21,2]*100
enframe(get_estimated_pi(m), name = "Covariances") %>%
  ggplot(aes(x = Covariances, y = value)) + geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(hjust = 1, angle = 45))
```
`r covar[21,2]*100`% of all effects load on to the "Equal Effects" covariance matrix.

Here is the pattern of effects for ED_PCA_1. The SITES are in alphabetical order - BRKG is X1,  CLMB X2, etc. 
```{r, cache = TRUE}
mash_plot_pairwise_sharing(corrmatrix = covariances$ED_PCA_1)
```

# Differential Sensitivity
```{r, cache = TRUE}
length(get_significant_results(m))
mash_plot_pairwise_sharing(corrmatrix = gxe$S_DS)
max(gxe$S_DS)
min(gxe$S_DS)
```
`r length(get_significant_results(m))` SNPs have significant effects in the mash analysis. Of those, between `r min(gxe$S_DS)` and `r max(gxe$S_DS)` have differentially significant effects between pairs of conditions.

# Antagonistic Pleiotropy
```{r, cache = TRUE}
mash_plot_pairwise_sharing(corrmatrix = gxe$S_AP)
```
`r length(get_significant_results(m))` SNPs have significant effects in the mash analysis. Of those, between `r min(gxe$S_AP)` and `r max(gxe$S_AP)` have antagonistically pleiotropic effects between pairs of conditions.

Between `r min(gxe$S_2_no)` and `r max(gxe$S_2_no)` have effects of the same magnitude and same sign between pairs of conditions.

### 2020-03-24 R/Make_kinship_van_Raden.Rmd
Make van Raden Kinship matrices for switchgrass

---
title: "Make Kinship Van Raden"
author: "Alice MacQueen"
date: "3/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
```

##

In the GAPIT numerical object, columns are SNPs and rows are individuals. It's the same for bigsnpr objects, luckily.

# Making a kinship matrix using the van Raden method.

If you don't need to subset by individuals or SNPs, you can use this function as-is. Otherwise see the code in the next section for options for subsetting.

```{r}
snp <- snp_attach("Pvirgatum_V5_GWAS_630g_33M_tensite_twoyear.rds")

pvdiv_kinship_van_raden <- function(snp){
  G <- snp$genotypes
  nInd <- snp$genotypes$nrow
  # Centered (mean of rows subtracted) transposed crossproduct of snp file.
  K <- big_tcrossprodSelf(G, fun.scaling = big_scale(center = TRUE, 
                                                                 scale = FALSE))

  #Extract diagonals
  i = 1:nInd
  j=(i-1)*nInd
  index=i+j
  d=K[index]
  DL=min(d)
  DU=max(d)
  floor=min(K[i,i])

  K=(K[i,i]-floor)/(DL-floor)
  MD=(DU-floor)/(DL-floor)

  rownames(K) <- snp$fam$sample.ID
  colnames(K) <- snp$fam$sample.ID

  if(MD>2){
    K[index]=K[index]/(MD-1)+1
  }
  return(K)
}
```

# Subset to just single ecotypes, and to just single genetic subpopulations.

This method could be used to remove any individuals that don't meet specific criteria.

```{r}
metadata <- read_rds("~/Github/pvdiv-phenotypes/data/metadata.rds")

upeco <- metadata[which(metadata$ECOTYPE_NNET == "Upland"),]$PLANT_ID
loeco <- metadata[which(metadata$ECOTYPE_NNET == "Lowland"),]$PLANT_ID
coeco <- metadata[which(metadata$ECOTYPE_NNET == "Coastal"),]$PLANT_ID

mgeno <- metadata[which(metadata$GENETIC_SUBPOP_50PER == "Midwest"),]$PLANT_ID
ageno <- metadata[which(metadata$GENETIC_SUBPOP_50PER == "Atlantic"),]$PLANT_ID
ggeno <- metadata[which(metadata$GENETIC_SUBPOP_50PER == "Gulf"),]$PLANT_ID

upland_eco <- which(snp$fam$sample.ID %in% upeco)
lowland_eco <- which(snp$fam$sample.ID %in% loeco)
coastal_eco <- which(snp$fam$sample.ID %in% coeco)
midwest_geno <- which(snp$fam$sample.ID %in% mgeno)
atlantic_geno <- which(snp$fam$sample.ID %in% ageno)
gulf_geno <- which(snp$fam$sample.ID %in% ggeno)

subsets <- list(upland_eco = upland_eco, lowland_eco = lowland_eco, coastal_eco = coastal_eco, midwest_geno = midwest_geno, atlantic_geno = atlantic_geno, gulf_geno = gulf_geno)

for(q in seq_along(subsets)){
    ind.row <- subsets[[q]]
    G <- snp$genotypes
    nInd <- length(ind.row)
    # Centered (mean of rows subtracted) transposed crossproduct of snp file.
    K <- big_tcrossprodSelf(G, ind.row = ind.row, 
                            fun.scaling = big_scale(center = TRUE, 
                                                    scale = FALSE))

    #Extract diagonals
    i = 1:nInd
    j=(i-1)*nInd
    index=i+j
    d=K[index]
    DL=min(d)
    DU=max(d)
    floor=min(K[i,i])

    K2=(K[i,i]-floor)/(DL-floor)
    MD=(DU-floor)/(DL-floor)
    # Name rows and columns
    rownames(K2) <- snp$fam$sample.ID[ind.row]
    colnames(K2) <- snp$fam$sample.ID[ind.row]

    if(MD>2){
      K2[index]=K2[index]/(MD-1)+1
    }

  write_rds(K2, paste0("Kinship_van_Raden_method_", nInd, "_individuals_", names(subsets)[q], "type.rds")) 
}
```

Other bigstatsr commands that I didn't end up using.
`snpSum <- big_colstats(G)# GAPIT: apply(snps,2,mean) Take the mean over columns.`
`snpMean <- snpSum$sum / nInd # bigsnpr has a function that can center like this.`

# GAPIT's method, for comparison purposes
```{r}
`GAPIT.kinship.VanRaden` <-
function(snps,hasInbred=TRUE) {
# Object: To calculate the kinship matrix using the method of VanRaden (2009, J. Dairy Sci. 91:4414???C4423)
# Authors: Zhwiu Zhang
# Last update: August 15, 2011 
##############################################################################################

print("Calculating kinship with VanRaden method...")

#snps=hm$GD 
nSNP=ncol(snps)
nInd=nrow(snps)
n=nInd

snpMean= apply(snps,2,mean)
snps=snps-snpMean
print("Getting X'X...")
K=tcrossprod((snps), (snps))

print("Adjusting...")
#Extract diagonals
i =1:n
j=(i-1)*n
index=i+j
d=K[index]
DL=min(d)
DU=max(d)
floor=min(K)

K=(K-floor)/(DL-floor)
MD=(DU-floor)/(DL-floor)

#Handler of diagonals over 2
#print("MD")
#print(MD)
#print(K[1:5,1:5])

if(is.na(K[1,1])) stop ("GAPIT says: Missing data is not allowed for numerical genotype data")
if(MD>2)K[index]=K[index]/(MD-1)+1

#Handler of inbred
if(MD<2 & hasInbred) K=2*K/((DU-floor)/(DL-floor))

print("Calculating kinship with VanRaden method: done")

return(K)
}
```

### 2020-04-04 R/GRM_kinship_for_fourway.Rmd
Find GRM (kinship) for fourway cross individuals for Li

--- 2020-03-24 R/Make_kinship_van_Raden.Rmd
@@
 ---
-title: "Make Kinship Van Raden"
+title: "GRM_fourway"
 author: "Alice MacQueen"
-date: "3/24/2020"
+date: "4/2/2020"
 output: html_document
@@
 library(tidyverse)
-library(bigsnpr)
+library(rlang)
 ```
@@
 
-In the GAPIT numerical object, columns are SNPs and rows are individuals. It's the same for bigsnpr objects, luckily.
+```{r}
+# paste0(rep(c(1:9),each = 2), rep(c("K", "N"), 9))
 
-# Making a kinship matrix using the van Raden method.
+# str(cross$geno$`2N`$data)
 
-If you don't need to subset by individuals or SNPs, you can use this function as-is. Otherwise see the code in the next section for options for subsetting.
+geno_df <- cross$geno[[1]]$data
 
-```{r}
-snp <- snp_attach("Pvirgatum_V5_GWAS_630g_33M_tensite_twoyear.rds")
-
-pvdiv_kinship_van_raden <- function(snp){
-  G <- snp$genotypes
-  nInd <- snp$genotypes$nrow
-  # Centered (mean of rows subtracted) transposed crossproduct of snp file.
-  K <- big_tcrossprodSelf(G, fun.scaling = big_scale(center = TRUE, 
-                                                                 scale = FALSE))
-
-  #Extract diagonals
-  i = 1:nInd
-  j=(i-1)*nInd
-  index=i+j
-  d=K[index]
-  DL=min(d)
-  DU=max(d)
-  floor=min(K[i,i])
-
-  K=(K[i,i]-floor)/(DL-floor)
-  MD=(DU-floor)/(DL-floor)
-
-  rownames(K) <- snp$fam$sample.ID
-  colnames(K) <- snp$fam$sample.ID
-
-  if(MD>2){
-    K[index]=K[index]/(MD-1)+1
-  }
-  return(K)
-}
-```
-
-# Subset to just single ecotypes, and to just single genetic subpopulations.
-
-This method could be used to remove any individuals that don't meet specific criteria.
-
-```{r}
-metadata <- read_rds("~/Github/pvdiv-phenotypes/data/metadata.rds")
-
-upeco <- metadata[which(metadata$ECOTYPE_NNET == "Upland"),]$PLANT_ID
-loeco <- metadata[which(metadata$ECOTYPE_NNET == "Lowland"),]$PLANT_ID
-coeco <- metadata[which(metadata$ECOTYPE_NNET == "Coastal"),]$PLANT_ID
-
-mgeno <- metadata[which(metadata$GENETIC_SUBPOP_50PER == "Midwest"),]$PLANT_ID
-ageno <- metadata[which(metadata$GENETIC_SUBPOP_50PER == "Atlantic"),]$PLANT_ID
-ggeno <- metadata[which(metadata$GENETIC_SUBPOP_50PER == "Gulf"),]$PLANT_ID
-
-upland_eco <- which(snp$fam$sample.ID %in% upeco)
-lowland_eco <- which(snp$fam$sample.ID %in% loeco)
-coastal_eco <- which(snp$fam$sample.ID %in% coeco)
-midwest_geno <- which(snp$fam$sample.ID %in% mgeno)
-atlantic_geno <- which(snp$fam$sample.ID %in% ageno)
-gulf_geno <- which(snp$fam$sample.ID %in% ggeno)
-
-subsets <- list(upland_eco = upland_eco, lowland_eco = lowland_eco, coastal_eco = coastal_eco, midwest_geno = midwest_geno, atlantic_geno = atlantic_geno, gulf_geno = gulf_geno)
-
-for(q in seq_along(subsets)){
-    ind.row <- subsets[[q]]
-    G <- snp$genotypes
-    nInd <- length(ind.row)
-    # Centered (mean of rows subtracted) transposed crossproduct of snp file.
-    K <- big_tcrossprodSelf(G, ind.row = ind.row, 
-                            fun.scaling = big_scale(center = TRUE, 
-                                                    scale = FALSE))
-
-    #Extract diagonals
-    i = 1:nInd
-    j=(i-1)*nInd
-    index=i+j
-    d=K[index]
-    DL=min(d)
-    DU=max(d)
-    floor=min(K[i,i])
-
-    K2=(K[i,i]-floor)/(DL-floor)
-    MD=(DU-floor)/(DL-floor)
-    # Name rows and columns
-    rownames(K2) <- snp$fam$sample.ID[ind.row]
-    colnames(K2) <- snp$fam$sample.ID[ind.row]
-
-    if(MD>2){
-      K2[index]=K2[index]/(MD-1)+1
+for(i in 2:18){
+geno_df <- geno_df %>%
+  cbind(cross$geno[[i]]$data)
 }
 
-  write_rds(K2, paste0("Kinship_van_Raden_method_", nInd, "_individuals_", names(subsets)[q], "type.rds")) 
-}
+# geno_df[1:20,1000:1005]
+# length(which(!is.na(geno_df)))
+# 50666/6170584 # less than 1% missing data, but it majorly fucks up the kinship matrix.
+
+geno_df2 <- geno_df %>%
+  as_tibble() %>%
+  mutate_if(is.numeric, list(~ifelse(is.na(.), sample(1:4, 1), .))) # replace NA's with a random value from 1 to 4.
+
+snps <- as.matrix(geno_df2)
+# snps[1:20,1000:1005]
 ```
-
-Other bigstatsr commands that I didn't end up using.
-`snpSum <- big_colstats(G)# GAPIT: apply(snps,2,mean) Take the mean over columns.`
-`snpMean <- snpSum$sum / nInd # bigsnpr has a function that can center like this.`
 
@@
 ```
+
+```{r}
+K[1:10,1:5]
+library(switchgrassGWAS)
+
+mash_plot_pairwise_sharing(corrmatrix = K[1:30,1:30])
+
+write_rds(K, path = "~/Github/pvdiv-phenology-gxe/data/Kinship_van_Raden_5250g_fourway_cross.rds")
+```

### 2020-04-08 R/D2F_annotations.Rmd
Start D2F and FL50 analysis for fourway overlaps on Gulf and Midwest pops

---
title: "Annos_D2F"
author: "Alice MacQueen"
date: "4/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

##

```{r}
anno_d2f <- read_delim("../analysis/gwas/D2Fannos.txt", delim = " ", col_names = "filename")
inputdir <- "~/Github/pvdiv-phenology-gxe/analysis/gwas/"
i=1
  annos <- read_rds(paste0(inputdir, anno_d2f$filename[i]))
  anno_overlap <- annos$top500SNPs_within0bp %>%
  filter((CHR == c("Chr02K") & between(POS, 63687000, 64420000)) | (CHR == "Chr02N" & between(POS, 60250000, 61280000)) | (CHR == "Chr04K" & between(POS, 1950000, 5450000)) | (CHR == "Chr05N" & between(POS, 20000, 4330000)) | (CHR == "Chr05N" & between(POS, 58910000, 65260000)))
  anno_overlap$SITE_PHE <- str_sub(anno_d2f$filename[i], start = 19, end = 26)

for(i in 2:nrow(anno_d2f)){
  annos <- read_rds(paste0(inputdir, anno_d2f$filename[i]))
  anno_overlap1 <- annos$top500SNPs_within0bp %>%
  filter((CHR == c("Chr02K") & between(POS, 63687000, 64420000)) | (CHR == "Chr02N" & between(POS, 60250000, 61280000)) | (CHR == "Chr04K" & between(POS, 1950000, 5450000)) | (CHR == "Chr05N" & between(POS, 20000, 4330000)) | (CHR == "Chr05N" & between(POS, 58910000, 65260000)))
    anno_overlap1$SITE_PHE <- str_sub(anno_d2f$filename[i], start = 19, end = 26)
  anno_overlap <- anno_overlap %>%
    full_join(anno_overlap1)

}
  anno_overlap %>% 
    mutate(group = ceiling(POS/1000000)) %>%
    group_by(CHR, group, SITE_PHE) %>%
    tally() %>% arrange(CHR, group)
anno_overlap %>% 
  filter(CHR == "Chr05N" & between(POS, 63800000, 65200000)) %>% arrange(POS)
```
Pavir.5NG107400 is a gene found in the major QTL region in the fourway cross on Chr05N. It is also found in GWAS of Gulf and Midwestern individuals in the diversity panel, and has SNPs that fall within the gene at PKLE (28d effect) & CLMB (11d effect). Pavir.5NG107400 is located on Chr05N at 64.39Mb, and is homologous to ATCPL2, a carboxyl-terminal domain (CTD) phosphatase that regulates plant growth, stress, and auxin responses. In A. thaliana, mutants in CPL2 have early flowering, low fertility, and increased salt sensitivity. In this 100kb region of Chr05N, we also see associations in the top 500 SNPs by p-value for TMPL (13d effect size, 90kb away from this gene), LINC (-13d, 3kb), KING (34d, 61kb).
Overall there are many overlaps on Chr05N between the QTL and GWAS analyses for days to flowering (FL50 - GR50), and overlaps at single sites for Chr02K (TMPL) and Chr04K (KBSM).

2K tmpl
4k KBSM
5N: KBSM, STIL, PKLE, CLMB, KING (early) tmpl LINC (500)
5N (LATE) PKLE CLMB KBSM KING STIL (TOP 500: CLMB KING STIL KBSM TMPL PKLE)
```{r}
fourways
```
Chr02K_63.687871	Chr02K_64.419301    no
Chr02N_60.251689	Chr02N_61.274395    Chr02N	60985754	60985754	
Chr04K_1.959391	Chr04K_5.499253       no
Chr05N_2.1318	Chr05N_4.328965         Chr05N	4568275	4568275
Chr05N_58.916323	Chr05N_65.255869    Chr05N	58930463	58930463 and 11 SNPs to 
Chr05N	64393503	64393503

# D2F
```{r}
annos$FDR0.1_within0bp %>%
  filter(CHR %in% c("Chr02K", "Chr02N", "Chr04K", "Chr05N"))

annos$FDR0.1_within0bp %>%
  filter((CHR == c("Chr02K") & between(POS, 63687000, 64420000)) | (CHR == "Chr02N" & between(POS, 60250000, 61280000)) | (CHR == "Chr04K" & between(POS, 1950000, 5450000)) | (CHR == "Chr05N" & between(POS, 20000, 4330000)) | (CHR == "Chr05N" & between(POS, 58910000, 65260000)))

annos$FDR0.1_within50000bp %>%
  filter(CHR %in% c("Chr05N") & between(POS, 58830460, 64493505))
```
Chr02N: No gene within 50kb
```{r}
-log10(2.03e-06)
```

```{r}
phe <- Phenotypes_D2F_fourway_project

phe %>%
  dplyr::select(PLANT_ID, ends_with("D2F"), ends_with("FL50"),
                ends_with("GR50")) %>%
  write_rds("../data/Phenotypes_D2F_fourway_project.rds")
```

### 2020-04-08 R/GWAS_standard_kinship.R

@@ -1,0 +1,20 @@
+library(bigsnpr)
+library(tidyverse)
+library(switchgrassGWAS)
+library(AnnotationDbi)
+txdb <- loadDb(file = file.path("~", "Github", "pvdiv-genome",
+                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
+
+outputdir <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", "gwas")
+
+inputfiles <- read_delim("~/Github/pvdiv-phenology-gxe/analysis/gwas/inputkinship.txt", delim = " ", col_names = "SNPfiles")
+
+phe_df <- read_rds(file.path("~", "Github",
+                             "pvdiv-phenology-gxe", "data",
+                             "Phenotypes_D2F_fourway_project.rds"))
+for(i in 2:nrow(inputfiles)){
+  snp <- snp_attach(inputfiles$SNPfiles[i])
+  pvdiv_standard_gwas(snp, df = phe_df, type = "linear", outputdir = outputdir,
+                      savegwas = TRUE, saveannos = TRUE, txdb = txdb,
+                      minphe = 120)
+}

### 2020-04-11 R/D2F_annotations.Rmd
Analysis for three most compelling flowering time candidate genes

@@
 library(tidyverse)
+library(switchgrassGWAS)
 ```
@@
 ```{r}
+anno_all <- pvdiv_results_in_folder(path = "../analysis/gwas/", pattern = "Annotation")
 anno_d2f <- read_delim("../analysis/gwas/D2Fannos.txt", delim = " ", col_names = "filename")
+
+annos_fl <- anno_all %>% 
+  enframe(name = NULL, value = "filename") %>%
+  filter(!grepl("GR50", filename) & !grepl("SEN50", filename))
+```
+# Top 100 or so overlaps, not filtering to fourway hits
+```{r}
+inputdir <- "~/Github/pvdiv-phenology-gxe/analysis/gwas/"
+i=1
+  annos <- read_rds(paste0(inputdir, annos_fl$filename[i]))
+  anno_overlap100 <- annos$top500SNPs_within50000bp
+
+  anno_overlap100$Anno_descr <- paste0(str_sub(annos_fl$filename[i], start = 19, 
+                                          end = -12))
+
+for(i in 2:nrow(annos_fl)){
+  annos <- read_rds(paste0(inputdir, annos_fl$filename[i]))
+  anno_overlap1001 <- annos$top500SNPs_within50000bp
+
+    anno_overlap1001$Anno_descr <- str_sub(annos_fl$filename[i], start = 19, end = -12)
+  anno_overlap100 <- anno_overlap100 %>%
+    dplyr::select(-TXID) %>%
+    full_join(anno_overlap1001)
+
+}
+anno_overlap100 <- anno_overlap100 %>%
+    separate(Anno_descr, into = c("SITE", "PHE", "type", "genos", "nSNP", 
+                                  "SNP"), sep = "_") %>%
+    mutate(genos = str_sub(genos, start = 6, end = -2),
+           subpop = case_when(nSNP == "11.5M" ~ "Atlantic_Gulf",
+                              nSNP == "8.8M" ~ "Midwest",
+                              nSNP == "12.3M" ~ "Atlantic_Midwest",
+                              nSNP == "12.1M" ~ "Gulf_Midwest",
+                              nSNP == "10.3M" ~ "Gulf",
+                              nSNP == "10.6M" ~ "Atlantic",
+                              TRUE ~ "Other"),
+           Atlantic = case_when(grepl("Atlantic", subpop) ~ 1,
+                                TRUE ~ 0),
+           Midwest = case_when(grepl("Midwest", subpop) ~ 1,
+                                TRUE ~ 0),
+           Gulf = case_when(grepl("Gulf", subpop) ~ 1,
+                                TRUE ~ 0)
+           )
+
+anno_overlap100 %>% 
+  dplyr::select(SITE, PHE, subpop:Gulf, everything()) %>%
+  mutate(group = ceiling(POS/1000000)) %>%
+  group_by(CHR, group, SITE) %>%
+  tally() %>% tally() %>% arrange(desc(n), CHR, group)
+
+anno_overlap100 %>% 
+  dplyr::select(SITE, PHE, subpop:Gulf, everything()) %>%
+  filter(Atlantic == 0 & Midwest == 1) %>%
+  mutate(group = round(POS/1000000, digits = 1)) %>%
+  group_by(CHR, group, SITE) %>%
+  tally() %>% tally() %>% filter(n >= 5) %>% arrange(CHR, group)
+
+anno_overlap100 %>%
+  dplyr::select(PHE, SITE, subpop, POS, everything()) %>%
+  mutate(group = round(POS/1000000, digits = 1),
+         log10p = -log10(`p value`)) %>%
+  filter((CHR %in% c("Chr07K") & group %in% c(10.4)) | (CHR %in% c("Chr07N") & group %in% c(45.7))) %>%
+  dplyr::select(CHR, group, log10p, `SNP Effect`, PHE, SITE, `Gene ID`, everything()) %>%
+  arrange(CHR, group)
+```
+Chr07K	10.4	5	 	Pavir.7KG042400	 PKLE LINC CLMB FRMI KBSM KING   	Encodes a cold-activated plasma membrane protein cold-responsive protein kinase that phosphorylates 14-3-3 proteins.
+The phosphorylated 14-3-3 proteins shuttle from the cytosol to the nucleus, where they interact with and destabilize the key cold-responsive C-repeat-binding factor (CBF) proteins, modulate CBF stability and the response to cold stress.
+~9-11d
+Chr07N	45.7	5	 LINC STIL TMPL KBSM CLMB ~6-8d 
+Pavir.7NG394500 (expressed protein in rice)
+Pavir.7NG399825
+
+##----------------------------
+KBSM & CLMB Chr07N 45.7 ~15d KBSM ~8d CLMB Gulf_Midwest	Pavir.7NG394600	AT1G30620
+HIGH SUGAR RESPONSE8, HSR8, MUR4, MURUS 4, UDP-D-XYLOSE 4-EPIMERASE 1, UXE1
+Description (?) 	  	encodes a type-II membrane protein that catalyzes 4-epimerization of UDP-D-Xylose to UDP-L-Arabinose in vitro, the nucleotide sugar used by glycosyltransferases in the arabinosylation of cell wall polysaccharides and wall-resident proteoglycans.
+Highly expressed in axis of inflorescence, flower, carpel, sensecent leaf
+
+# FL50 & D2F annos 2020-04-10
+
+As FL50 and D2F QTL regions are basically the same, just a few more FL50 QTL regions, and as the FL50 & D2F GWAS look quite similar in most cases, I will just compare annotations for all the flowering results here.
+```{r}
+inputdir <- "~/Github/pvdiv-phenology-gxe/analysis/gwas/"
+i=1
+  annos <- read_rds(paste0(inputdir, annos_fl$filename[i]))
+  anno_overlap <- annos$top500SNPs_within50000bp %>%
+  filter((CHR == "Chr02K" & between(POS, 63687000, 65490000)) | 
+           (CHR == "Chr02N" & between(POS, 58030000, 61280000)) | 
+           (CHR == "Chr03K" & between(POS, 17360000, 26180000)) | 
+           (CHR == "Chr04K" & between(POS, 1950000, 5500000)) | 
+           (CHR == "Chr05N" & between(POS, 20000, 4330000)) | 
+           (CHR == "Chr05N" & between(POS, 58240000, 66000000)) | 
+           (CHR == "Chr08N" & between(POS, 550000, 34420000)) | 
+           (CHR == "Chr09K" & between(POS, 13650000, 18180000)))
+
+  anno_overlap$Anno_descr <- paste0(str_sub(annos_fl$filename[i], start = 19, 
+                                          end = -12))
+
+for(i in 2:nrow(annos_fl)){
+  annos <- read_rds(paste0(inputdir, annos_fl$filename[i]))
+  anno_overlap1 <- annos$top500SNPs_within50000bp %>%
+    dplyr::select(-TXID) %>%
+  filter((CHR == "Chr02K" & between(POS, 63687000, 65490000)) | 
+           (CHR == "Chr02N" & between(POS, 58030000, 61280000)) | 
+           (CHR == "Chr03K" & between(POS, 17360000, 26180000)) | 
+           (CHR == "Chr04K" & between(POS, 1950000, 5500000)) | 
+           (CHR == "Chr05N" & between(POS, 20000, 4330000)) | 
+           (CHR == "Chr05N" & between(POS, 58240000, 66000000)) | 
+           (CHR == "Chr08N" & between(POS, 550000, 34420000)) | 
+           (CHR == "Chr09K" & between(POS, 13650000, 18180000)))
+
+    anno_overlap1$Anno_descr <- str_sub(annos_fl$filename[i], start = 19, end = -12)
+  anno_overlap <- anno_overlap %>%
+    full_join(anno_overlap1)
+
+}
+anno_overlap <- anno_overlap %>%
+    separate(Anno_descr, into = c("SITE", "PHE", "type", "genos", "nSNP", 
+                                  "SNP"), sep = "_") %>%
+    mutate(genos = str_sub(genos, start = 6, end = -2),
+           subpop = case_when(nSNP == "11.5M" ~ "Atlantic_Gulf",
+                              nSNP == "8.8M" ~ "Midwest",
+                              nSNP == "12.3M" ~ "Atlantic_Midwest",
+                              nSNP == "12.1M" ~ "Gulf_Midwest",
+                              nSNP == "10.3M" ~ "Gulf",
+                              nSNP == "10.6M" ~ "Atlantic",
+                              TRUE ~ "Other"),
+           Atlantic = case_when(grepl("Atlantic", subpop) ~ 1,
+                                TRUE ~ 0),
+           Midwest = case_when(grepl("Midwest", subpop) ~ 1,
+                                TRUE ~ 0),
+           Gulf = case_when(grepl("Gulf", subpop) ~ 1,
+                                TRUE ~ 0)
+           )
+
+anno_overlap %>% 
+  dplyr::select(SITE, PHE, subpop:Gulf, everything()) %>%
+  mutate(group = ceiling(POS/1000000)) %>%
+  group_by(CHR, group, SITE) %>%
+  tally() %>% tally() %>% arrange(desc(n), CHR, group)
+
+anno_overlap %>% 
+  dplyr::select(SITE, PHE, subpop:Gulf, everything()) %>%
+  filter(Atlantic == 0 & Midwest == 1) %>%
+  mutate(group = round(POS/1000000, digits = 1)) %>%
+  group_by(CHR, group, SITE) %>%
+  tally() %>% tally() %>% filter(n >= 3) %>% arrange(CHR, group)
+
+anno_overlap %>% 
+  mutate(group = round(POS/1000000, digits = 1)) %>%
+  filter(CHR == "Chr08N" & group %in% c(0.8, 15.6, 27.7)) %>%
+  dplyr::select(CHR, POS, subpop, `SNP Effect`, PHE, SITE, `Gene ID`, everything()) %>%
+  arrange(CHR, POS)
+```
+Chr08N 0.8Mb in Gulf Midwest LINC TMPL STIL BRKG (colder) ~13 - 14d
+Pavir.8NG355401 AT3G05550 Hypoxia responsive family protein only sort of highly expressed in early flower
+Pavir.8NG355404 AT2G30330 Protein - protein interaction with BLOS2 and also with SNX1.Located in endomembrane system and hypothesized to be involved in endomembrane transport.
+Pavir.8NG355405	AT5G05340 Pavir.8NG355404 Homolog to a peroxidase superfamily protein highly expressed only in senescent leaf petioles.
+Pavir.8NG002300	AT5G03610 highly expressed in axis of inflorescence  	GDSL-motif esterase/acyltransferase/lipase.
+Enzyme group with broad substrate specificity that may catalyze acyltransfer or hydrolase reactions with lipid and non-lipid substrates.
+
+I gotta say, all of these hits I've been looking at, I've been shocked by the number of hits that have very high gene expression in the axis of inflorescence (in Arabidopsis).
+This is almost certainly a good sign, given that I've been looking at flowering hits!
+
+Chr02N	59	8	CLMB BRKG TMPL PKLE STIL ... CLMB ... KBSM ... LINC FRMI KBSM
+Pavir.2NG496800 58.19  AT4G25230 RIN2 RPM1 interacting protein.
+Positive regulator of RPM1 and PRS2 mediated hypersensitive response.Functions as ubiquitin ligase and binds to RPM1.
+Could certainly hasten flowering if it increases a basal defense response (MacQueen et al 2016).
+
+KBSM STIL?
+Pavir.2NG491000 58.61Mb AT2G35660 CTF2A: Encodes a member of a novel gene family with homology to known proteins involved in hydroxylation and oxidation of an aromatic ring.
+Highly expressed in sepals, pedicel, leaves, and siliques.
+
+Chr02N	59	8	
+Chr03K	24	8	
+Chr03K	26	9	
+Chr04K	3	8		
+Chr04K	4	8	
+Chr05N	2	9	
+Chr05N	3	8	
+Chr05N	4	9	
+Chr05N	60	8	
+Chr05N	64	9		
+Chr05N	65	9		
+Chr05N	66	9		
+Chr08N	10	9		
+Chr08N	13	9
+
+Without Atlantic group:
+Chr05N	65	9		
+Chr02N	59	7		
+Chr05N	4	7		
+Chr05N	66	7
+
+Chr08N	0.8	4		
+Chr08N	15.6	4		
+Chr08N	27.7	4
+
+Without Midwest group:
+Chr03K	26	8		
+Chr05N	2	8		
+Chr05N	4	8		
+Chr08N	10	8		
+Chr04K	3	7		
+Chr05N	3	7		
+Chr05N	65	7
+
+Without Gulf group:
+Chr08N	3	6
+
+With Gulf group:
+Chr03K	26	8		
+Chr04K	4	8		
+Chr05N	2	8		
+Chr05N	3	7		
+Chr05N	4	9		
+Chr05N	62	7		
+Chr05N	64	7		
+Chr05N	65	9		
+Chr05N	66	7		
+Chr08N	10	8	
+Chr08N	13	7
+
+With Midwest group:
+Chr02N	59	7		
+Chr03K	24	7		
+Chr03K	26	7		
+Chr05N	4	7		
+Chr05N	5	7		
+Chr05N	64	7		
+Chr05N	65	8		
+Chr05N	66	9		
+Chr08N	13	7
+
+Anything by 100Kb interval
+Chr03K	24.9	5		
+Chr04K	3.7	5		
+Chr05N	61.4	5		
+Chr05N	65.9	5		
+Chr08N	15.6	5		
+Chr08N	17.7	5		
+Chr08N	33.1	5
+
+Midwest & Gulf
+Chr05N	4.3	3		
+Chr08N	0.8	4		
+Chr08N	15.6	3		
+Chr08N	27.7	3
+
+Gulf, no Atlantic
+Chr08N	0.8	4		
+Chr08N	15.6	4		
+Chr08N	27.7	4
+
+Midwest, no Atlantic
+Chr05N	4.3	3		
+Chr05N	59.1	3		
+Chr08N	0.8	4		
+Chr08N	15.6	3		
+Chr08N	27.7	3
+
+Other intervals to look in: 
+Ch05N 4.3
+Chr05N 59.1
+Chr05N 61.4
+Chr08N 0.8
+Chr08N 15.6
+Chr08N 27.7
+
+```{r}
+anno_overlap %>% 
+  dplyr::select(PHE, SITE, subpop, POS, everything()) %>%
+  mutate(group = round(POS/1000000, digits = 1),
+         log10p = -log10(`p value`)) %>%
+  filter((CHR %in% c("Chr05N") & group %in% c(4.3, 59.1, 61.4)) | (CHR %in% c("Chr08N") & group %in% c(0.8, 15.6, 27.7))) %>%
+  dplyr::select(CHR, group, log10p, `SNP Effect`, PHE, SITE, `Gene ID`, everything()) %>%
+  arrange(CHR, group)
+
+anno_overlap %>% 
+  dplyr::select(PHE, SITE, subpop, POS, everything()) %>%
+  mutate(group = round(POS/1000000, digits = 1),
+         log10p = -log10(`p value`)) %>%
+  filter((CHR %in% c("Chr03K") & group %in% c(24.9)) | (CHR %in% c("Chr04K") & group %in% c(3.7))) %>%
+  dplyr::select(CHR, group, log10p, `SNP Effect`, PHE, SITE, `Gene ID`, everything()) %>%
+  arrange(CHR, group)
+
+```
+Chr03K 24.9 CLMB FRMI KBSM PKLE STIL probably Atlantic
+	AT1G34340 mature & senescent leaf; senescent flower
+Chr04K 3.7 BRKG CLMB PKLE STIL TMPL Atlantic or Gulf ~5d effect Pavir.4KG031100 homolog to AT1G71696 which is highly expressed in root, root apical meristem, axis of inflorescence, carpel, and senescing flowers.
+sol1 partially suppresses the short root phenotype caused by CLE19 overexpression.
+
+5N 4.3 Mb: TMPL, STIL, KING, CLMB. In populations with Midwest plants.
+Pavir.5N036700 (oxidation related zinc finger), to Pavir.5N037500
+Best candidate:
+Pavir.5NG037100. in A. thaliana, highly expressed in the axis of inflorescence and in dividing cells.
+Plant Aurora kinases play a role in the maintenance of primary meristems and control of endoreduplication (Petrovska et al 2011 New Phytologist)
+
+Chr05N 61.4 Mb KING & BRKG (LINC & PKLE & TMPL close) Pavir.5NG557850 is a
+homolog to AT3G17450 which is a hAT dimerization domain-containing protein that is highly expressed in the axis of inflorescence, the carpel of developing and mature flowers, and early seeds.
+AT4G10360 (less likely though is expressed in pollen germination and tube growth).
+
+Chr08N 0.8Mb LINC STIL TMPL BRKG Gulf & Midwest. ~13d effect at all sites but BRKG. Pavir.8NG355404 Homolog to a peroxidase superfamily protein highly expressed only in senescent leaf petioles.
+
+Chr08N 15.6Mb PKLE LINC KING KBSM CLMB 14-18 d effect. Mostly in Atlantic subpop.
+Pavir.8NG252400 (Gulf effect probably)
+
+Chr08N 27.7Mb CLMB LINC STIL TMPL. Gulf effect probably.
+
+# D2F annos 2020-04-08
+```{r}
 inputdir <- "~/Github/pvdiv-phenology-gxe/analysis/gwas/"
@@
 
+# Most permissive set of FL50 & D2F hits - seeing as they all basically overlap #
 annos$FDR0.1_within0bp %>%
-  filter((CHR == c("Chr02K") & between(POS, 63687000, 64420000)) | (CHR == "Chr02N" & between(POS, 60250000, 61280000)) | (CHR == "Chr04K" & between(POS, 1950000, 5450000)) | (CHR == "Chr05N" & between(POS, 20000, 4330000)) | (CHR == "Chr05N" & between(POS, 58910000, 65260000)))
+  filter((CHR == c("Chr02K") & between(POS, 63687000, 65490000)) | 
+           (CHR == "Chr02N" & between(POS, 58030000, 61280000)) | 
+           (CHR == "Chr03K" & between(POS, 17360000, 26180000)) | 
+           (CHR == "Chr04K" & between(POS, 1950000, 5500000)) | 
+           (CHR == "Chr05N" & between(POS, 20000, 4330000)) | 
+           (CHR == "Chr05N" & between(POS, 58240000, 66000000)) | 
+           (CHR == "Chr08N" & between(POS, 550000, 34420000)) | 
+           (CHR == "Chr09K" & between(POS, 13650000, 18180000)))
 

### 2020-04-24 R/D2F_breakdown_by_site.Rmd
D2F environmental index analysis for two indices and CV of phenotypes

---
title: "Untitled"
author: "Alice MacQueen"
date: "4/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
```

#

```{r}
phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Phenotypes_D2F_fourway_project.rds")
metadata <- metadata %>%
  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
                            !is.na(PLOIDY) ~ as.character(PLOIDY),
                            TRUE ~ NA_character_))
```

```{r}
lodToPval <-  function(x){
  pchisq(x * (2 * log(10)), df = 1, lower.tail = FALSE) / 2
}
```

```{r}
phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(!(SITE %in% c("FRMI_D2F", "TMPL_D2F", "BRKG_D2F"))) %>%
  ggplot(aes(x = D2F)) + 
  geom_histogram() + 
  facet_wrap(~SITE)

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("D2F")) %>%
  pivot_longer(cols = ends_with("FL50"), names_to = "SITE", values_to = "FL50") %>%
  #filter(!(SITE %in% c("FRMI_FL50", "TMPL_FL50", "BRKG_FL50"))) %>%
  ggplot(aes(x = FL50)) + 
  geom_histogram() + 
  facet_wrap(~SITE)

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F < 30)
```

```{r}
phe[1:15,] %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  ggplot(aes(x = SITE, y = D2F)) + 
  geom_line(aes(group = PLANT_ID, color = PLANT_ID))

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(GENETIC_SUBPOP_KINSHIP %in% c("Midwest", "Gulf")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10) %>%
  group_by(SITE, GENETIC_SUBPOP_KINSHIP) %>%
  summarise(mean_D2F = mean(D2F, na.rm = TRUE), 
            median_D2F = median(D2F, na.rm = TRUE),
            sd_D2F = sd(D2F, na.rm = TRUE)) %>%
  ggplot(aes(x = SITE, y = mean_D2F)) + 
  geom_point(aes(color = GENETIC_SUBPOP_KINSHIP)) +
  geom_line(aes(group = GENETIC_SUBPOP_KINSHIP, color = GENETIC_SUBPOP_KINSHIP))

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  #filter(GENETIC_SUBPOP_KINSHIP %in% c("Midwest", "Gulf")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10) %>%
  group_by(SITE, GENETIC_SUBPOP_KINSHIP) %>%
  summarise(mean_D2F = mean(D2F, na.rm = TRUE), 
            median_D2F = median(D2F, na.rm = TRUE),
            sd_D2F = sd(D2F, na.rm = TRUE),
            cv_D2F = sd_D2F/mean_D2F) %>%
  arrange(GENETIC_SUBPOP_KINSHIP, cv_D2F) %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  left_join(sites) %>%
  ggplot(aes(x = Latitude, y = cv_D2F)) + 
  geom_point(aes(color = GENETIC_SUBPOP_KINSHIP)) +
  geom_line(aes(group = GENETIC_SUBPOP_KINSHIP, color = GENETIC_SUBPOP_KINSHIP)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(x = Latitude, label = SITE), y = 0.02, angle = 45) + ylim(c(0,0.6)) +
  scale_color_viridis(discrete = TRUE, end = 0.8)

?sd()
?var() # - variance (square of sd)

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(GENETIC_SUBPOP_KINSHIP %in% c("Midwest", "Gulf")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10) %>%
  ggplot(aes(x = D2F)) + 
  geom_histogram(aes(fill = GENETIC_SUBPOP_KINSHIP)) + facet_grid(~SITE)
```
Very similar sites: BRKG, CLMB, LINC, (FRMI)
STIL
KBSM, PKLE
KING
...
TMPL

Let's consider the idea that the Midwest and Gulf populations are receiving extremely different environmental signals at each of the planting locations. What can I use to support this idea? Well, the days to flowering

```{r}
sites
```
From Houle 1992:

Crow 1958: "opportunity for selection": I = V_P / X_bar^2
X_bar represents the population mean before selection.
V_P phenotypic trait variance. `var()`

To calculate CV (coefficient of variation) you take the standard deviation of the data and divide it by the mean of the data.  sqrt(V_P) / X_bar

I is related to the phenotypic coefficient of variation, CV_P = 100*sqrt(V_P) / X_bar
(this is just CV * 100 to be a percentage).

Relative evolvability of fitness:

I_A = V_A / X_bar^2 = (CV_A / 100)^2
I_A is also an appropriate indicator of the variability of fitness as large values must be the result of powerful forces maintaining genetic variance.

A variance to mean ratio is the most appropriate measure of the evolvability of a trait.

Compared three quantities:

Residual coefficient of variation CV_R = 100 * sqrt(V_P - V_A) / X_bar

h^2 narrow sense heritability = V_A / V_P

CV_A = 100*sqrt(V_A) / X_bar

```{r}
phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  #filter(GENETIC_SUBPOP_KINSHIP %in% c("Midwest", "Gulf")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10) %>%
  group_by(SITE, GENETIC_SUBPOP_KINSHIP) %>%
  summarise(mean_D2F = mean(D2F, na.rm = TRUE), 
            median_D2F = median(D2F, na.rm = TRUE),
            sd_D2F = sd(D2F, na.rm = TRUE),
            cv_D2F = sd_D2F/mean_D2F) %>%
  arrange(GENETIC_SUBPOP_KINSHIP, cv_D2F) %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  left_join(sites) %>%
  ggplot(aes(x = Latitude, y = cv_D2F)) + 
  geom_point(aes(color = GENETIC_SUBPOP_KINSHIP)) +
  geom_line(aes(group = GENETIC_SUBPOP_KINSHIP, color = GENETIC_SUBPOP_KINSHIP)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(x = Latitude, label = SITE), y = 0.02, angle = 45) + ylim(c(0,0.6)) +
  scale_color_viridis(discrete = TRUE, end = 0.8)

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  #filter(GENETIC_SUBPOP_KINSHIP %in% c("Midwest", "Gulf")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10) %>%
  group_by(SITE, GENETIC_SUBPOP_KINSHIP) %>%
  summarise(mean_D2F = mean(D2F, na.rm = TRUE), 
            median_D2F = median(D2F, na.rm = TRUE),
            sd_D2F = sd(D2F, na.rm = TRUE),
            cv_D2F = sd_D2F/mean_D2F) %>%
  arrange(GENETIC_SUBPOP_KINSHIP, cv_D2F) %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  left_join(sites) %>%
  ggplot(aes(x = Latitude, y = mean_D2F)) + 
  geom_point(aes(color = GENETIC_SUBPOP_KINSHIP)) +
  geom_line(aes(group = GENETIC_SUBPOP_KINSHIP, color = GENETIC_SUBPOP_KINSHIP)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(x = Latitude, label = SITE), y = 50, angle = 45) +
  scale_color_viridis(discrete = TRUE, end = 0.8) +
  geom_errorbar(aes(x = Latitude, ymin = mean_D2F - sd_D2F, ymax = mean_D2F + sd_D2F, group = GENETIC_SUBPOP_KINSHIP, color = GENETIC_SUBPOP_KINSHIP))

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  #filter(GENETIC_SUBPOP_KINSHIP %in% c("Midwest", "Gulf")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10) %>%
  group_by(SITE, GENETIC_SUBPOP_KINSHIP) %>%
  summarise(mean_D2F = mean(D2F, na.rm = TRUE), 
            median_D2F = median(D2F, na.rm = TRUE),
            sd_D2F = sd(D2F, na.rm = TRUE),
            cv_D2F = sd_D2F/mean_D2F) %>%
  arrange(GENETIC_SUBPOP_KINSHIP, cv_D2F) %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  left_join(sites) %>%
  ggplot(aes(x = Latitude, y = sd_D2F^2)) + 
  geom_point(aes(color = GENETIC_SUBPOP_KINSHIP)) +
  geom_line(aes(group = GENETIC_SUBPOP_KINSHIP, color = GENETIC_SUBPOP_KINSHIP)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(x = Latitude, label = SITE), y = 0.02, angle = 45) +
  scale_color_viridis(discrete = TRUE, end = 0.8)
```

# Fourway site plots
```{r}
phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10) %>%
  group_by(SITE, SUBPOP) %>%
  summarise(mean_D2F = mean(D2F, na.rm = TRUE), 
            median_D2F = median(D2F, na.rm = TRUE),
            sd_D2F = sd(D2F, na.rm = TRUE),
            cv_D2F = sd_D2F/mean_D2F) %>%
  arrange(SUBPOP, cv_D2F) %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  filter(!(SITE %in% c("FRMI", "TMPL"))) %>%
  left_join(sites) %>%
  ggplot(aes(x = Latitude, y = cv_D2F)) + 
  geom_point(aes(color = SUBPOP)) +
  geom_line(aes(group = SUBPOP, color = SUBPOP)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(x = Latitude, label = SITE), y = 0.02, angle = 45) + ylim(c(0,0.6)) +
  scale_color_viridis(discrete = TRUE)

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  #filter(GENETIC_SUBPOP_KINSHIP %in% c("Midwest", "Gulf")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10) %>%
  group_by(SITE, SUBPOP) %>%
  summarise(mean_D2F = mean(D2F, na.rm = TRUE), 
            median_D2F = median(D2F, na.rm = TRUE),
            sd_D2F = sd(D2F, na.rm = TRUE),
            cv_D2F = sd_D2F/mean_D2F) %>%
  arrange(SUBPOP, cv_D2F) %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  filter(!(SITE %in% c("FRMI", "TMPL"))) %>%
  left_join(sites) %>%
  ggplot(aes(x = Latitude, y = mean_D2F)) + 
  geom_line(aes(group = SUBPOP, color = SUBPOP)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(x = Latitude, label = SITE), y = 50, angle = 45) +
  scale_color_viridis(discrete = TRUE) +
  geom_errorbar(aes(x = Latitude, ymin = mean_D2F - sd_D2F, ymax = mean_D2F + sd_D2F, group = SUBPOP, color = SUBPOP)) +
  geom_point(aes(color = SUBPOP))

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  #filter(GENETIC_SUBPOP_KINSHIP %in% c("Midwest", "Gulf")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10) %>%
  group_by(SITE, SUBPOP) %>%
  summarise(mean_D2F = mean(D2F, na.rm = TRUE), 
            median_D2F = median(D2F, na.rm = TRUE),
            sd_D2F = sd(D2F, na.rm = TRUE),
            cv_D2F = sd_D2F/mean_D2F) %>%
  arrange(SUBPOP, cv_D2F) %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  filter(!(SITE %in% c("FRMI", "TMPL"))) %>%
  left_join(sites) %>%
  ggplot(aes(x = Latitude, y = sd_D2F^2)) + 
  geom_point(aes(color = SUBPOP)) +
  geom_line(aes(group = SUBPOP, color = SUBPOP)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(x = Latitude, label = SITE), y = 2000, angle = 45) + ylim(c(0, 4200)) +
  scale_color_viridis(discrete = TRUE, end = 1)

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10 & SUBPOP == "Gulf") %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  filter(SITE %in% c("KING", "STIL")) %>%
  ggplot(aes(x = SITE, y = D2F)) + 
  geom_line(aes(group = PLANT_ID, color = PLANT_ID)) + 
  theme(legend.position = "none")
phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10 & SUBPOP == "Midwest") %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  filter(SITE %in% c("KING", "STIL")) %>%
  ggplot(aes(x = SITE, y = D2F)) + 
  geom_line(aes(group = PLANT_ID, color = PLANT_ID)) + 
  theme(legend.position = "none")

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10 & SUBPOP == "Gulf") %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  filter(SITE %in% c("PKLE", "STIL")) %>%
  ggplot(aes(x = SITE, y = D2F)) + 
  geom_line(aes(group = PLANT_ID, color = PLANT_ID)) + 
  theme(legend.position = "none")
phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10 & SUBPOP == "Midwest") %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  filter(SITE %in% c("PKLE", "STIL")) %>%
  ggplot(aes(x = SITE, y = D2F)) + 
  geom_line(aes(group = PLANT_ID, color = PLANT_ID)) + 
  theme(legend.position = "none")

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10 & SUBPOP == "Gulf") %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  filter(SITE %in% c("STIL", "CLMB", "KING", "LINC", "KBSM")) %>%   # "PKLE", "BRKG"
  left_join(sites) %>%
  ggplot(aes(x = Latitude, y = D2F)) + 
  geom_line(aes(group = PLANT_ID, color = PLANT_ID)) + 
  theme(legend.position = "none") +
  geom_text(aes(x = Latitude, label = SITE), y = 130, angle = 45)
phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10 & SUBPOP == "Midwest") %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  filter(SITE %in% c("STIL","KING","KBSM")) %>%   # "BRKG", "CLMB", "LINC", 
  left_join(sites) %>%
  ggplot(aes(x = Latitude, y = D2F)) + 
  geom_line(aes(group = PLANT_ID, color = PLANT_ID)) + 
  theme(legend.position = "none") +
  geom_text(aes(x = Latitude, label = SITE), y = 130, angle = 45)
```

Instead of D2F, PTT2F? photothermal time to flowering, from greenup.

```{r}
ptt_df <- readRDS(file = "~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")

gdd_df <- ptt_df %>%
  mutate(GDD_0C_SM = ifelse((TMAX_SM + TMIN_SM)/2 >= 0,
                      (TMAX_SM + TMIN_SM)/2 - 0,
                      0)) %>%
  dplyr::select(DOY, SITE, Day_of_year, GDD_13C_SM:GDD_8C_SM, GDD_0C_SM)

GR_start <- phe %>% 
  dplyr::select(-ends_with("D2F"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("GR50"), names_to = "SITE", values_to = "GR50") %>%
  #filter(D2F > 10 & SUBPOP == "Midwest") %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  group_by(SITE, PLANT_ID) %>%
  summarise(GR_start = floor(min(GR50)))
FL_end <- phe %>% 
  dplyr::select(-ends_with("D2F"), -ends_with("GR50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("FL50"), names_to = "SITE", values_to = "FL50") %>%
  #filter(D2F > 10 & SUBPOP == "Midwest") %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  group_by(SITE, PLANT_ID) %>%
  summarise(FL_end = ceiling(max(FL50)))

GDD_PT <- NULL

tmp3 <- gdd_df %>%
  dplyr::select(DOY, SITE, Day_of_year, GDD_13C_SM:GDD_0C_SM) %>%
  left_join(GR_start, by = "SITE") %>%
  left_join(FL_end, by = c("SITE", "PLANT_ID")) %>%
  group_by(SITE, Day_of_year, PLANT_ID) %>%
  filter(between(Day_of_year, GR_start, FL_end)) %>% 
  ungroup()

tmp1 <- ptt_df %>%
  dplyr::select(DOY, SITE, Day_of_year, PTT_8C:PTT_13C) %>%
  left_join(GR_start, by = "SITE") %>%
  left_join(FL_end, by = c("SITE", "PLANT_ID")) %>%
  group_by(SITE, Day_of_year, PLANT_ID) %>%
  filter(between(Day_of_year, GR_start, FL_end)) %>% 
  ungroup()

tmp1 %>%
  group_by(SITE, PLANT_ID, GR_start, FL_end) %>%
  summarise(CPTT_8C = sum(PTT_8C),
            CPTT_9C = sum(PTT_9C),
            CPTT_10C = sum(PTT_10C),
            CPTT_11C = sum(PTT_11C),
            CPTT_12C = sum(PTT_12C),
            CPTT_13C = sum(PTT_13C)
            ) %>%
  mutate(D2F = FL_end - GR_start) %>%
  filter(D2F > 10) %>%
  left_join(metadata) %>%
  left_join(sites) %>%
  filter(!(SITE %in% c("TMPL", "BRKG", "FRMI"))) %>%
  filter(SUBPOP %in% c("Gulf", "Atlantic", "Midwest")) %>%
  ggplot(aes(x = Latitude, y = CPTT_8C)) + 
  geom_line(aes(group = PLANT_ID, color = SUBPOP)) + 
  #theme(legend.position = "none") +
  geom_text(aes(x = Latitude, label = SITE), y = 58000, angle = 45)
```

```{r}
pid_all_sites <- phe %>%
  dplyr::select(PLANT_ID, ends_with("D2F")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  group_by(PLANT_ID) %>%
  filter(!is.na(D2F) & !(SITE %in% c("FRMI_D2F", "BRKG_D2F", "TMPL_D2F"))) %>%
  count() %>% 
  arrange(desc(n)) %>%
  filter(n >= 4)

tmp3 %>%
  group_by(SITE, PLANT_ID, GR_start, FL_end) %>%
  summarise(CGDD_0C = sum(GDD_0C_SM),
            CGDD_8C = sum(GDD_8C_SM),
            CGDD_13C = sum(GDD_13C_SM)
            ) %>%
  mutate(D2F = FL_end - GR_start) %>%
  filter(D2F > 10) %>%
  left_join(metadata) %>%
  left_join(sites) %>%
  #filter(!(SITE %in% c("TMPL"))) %>%
  filter(PLANT_ID %in% pid_all_sites$PLANT_ID & SUBPOP %in% c("Midwest", "Gulf")) %>%
  ggplot(aes(x = Latitude, y = D2F)) + 
  geom_line(aes(group = PLANT_ID, color = SUBPOP)) + 
  #theme(legend.position = "none") +
  geom_text(aes(x = Latitude, label = SITE), y = 10, angle = 45) +
  geom_smooth(aes(group = SUBPOP), method = "lm")  +
  #geom_smooth(aes(group = SUBPOP), color = "green") +
  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1) +
  ylim(c(0, 275))
save_plot(filename = "../analysis/environmental-index/Reaction_norm_days_to_flowering_Gulf_and_Midwest_grown_4_or_more_4way_sites.png", plot = last_plot(), base_height = 6)

tmp3 %>%
  group_by(SITE, PLANT_ID, GR_start, FL_end) %>%
  summarise(CGDD_0C = sum(GDD_0C_SM),
            CGDD_8C = sum(GDD_8C_SM),
            CGDD_13C = sum(GDD_13C_SM)
            ) %>%
  mutate(D2F = FL_end - GR_start) %>%
  filter(D2F > 10) %>%
  left_join(metadata) %>%
  left_join(sites) %>%
  #filter(!(SITE %in% c("TMPL"))) %>%
  filter(PLANT_ID %in% pid_all_sites$PLANT_ID & SUBPOP %in% c("Midwest", "Gulf")) %>%
  ggplot(aes(x = Latitude, y = CGDD_13C)) + 
  geom_line(aes(group = PLANT_ID, color = SUBPOP)) + 
  #theme(legend.position = "none") +
  geom_text(aes(x = Latitude, label = SITE), y = 100, angle = 45) +
  geom_smooth(aes(group = SUBPOP), method = "lm")  +
  #geom_smooth(aes(group = SUBPOP), color = "green") +
  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1)  + 
  ylim(c(0,3200))
save_plot(filename = "../analysis/environmental-index/Reaction_norm_cumulative_GDD_to_flowering_Gulf_and_Midwest_grown_4_or_more_4way_sites.png", plot = last_plot(), base_height = 6)
```
 61 PLANT ID have D2F at all 6 fourway sites.  209 at 5 fourway sites.
PTTGR2FL

### 2020-04-24 R/Overlap_with_4way_QTL_D2F.Rmd
D2F environmental index analysis for two indices and CV of phenotypes

---
title: "D2F_4way_overlap"
author: "Alice MacQueen"
date: 2020-04-22
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(switchgrassGWAS)
library(cowplot)
library(viridis)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

##

```{r}
mash <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_Midwest/Strong_Effects49260SNPs.rds")
qtl <- read_csv(file = "~/Github/pvdiv-phenology-gxe/data/QTLs_with_flank_markers_FL50_7_sites_full_model.csv")
sites <- readRDS("~/Github/pvdiv-phenotypes/data/sites.rds")
```

```{r}
df <- get_significant_results(m=mash) %>%
  enframe(name = "SNP")
df1 <- switchgrassGWAS:::get_marker_df(mash)
df1 <- df1 %>%
  mutate(log10bf = switchgrassGWAS:::get_log10bf(m=mash),
         sigcond = switchgrassGWAS:::get_n_significant_conditions(mash)
         )
effects <- ashr::get_pm(mash)
sd <- ashr::get_psd(mash)
colnames(sd) <- str_replace(colnames(sd), "Bhat", "Shat")

df1 <- cbind(df1, effects, sd)
```

```{r}
qtl %>%
  arrange(desc(lod)) %>%
  dplyr::select(lodcolumn, lod, chr, flank_lo, flank_hi)
```
From most to least significant, for QTL:
	Chr02N_58.039498 61.274395
	Chr05N_1.502713 4.328965
	Chr05N_58.916323 65.990782
	Chr04K_1.959391 5.499253
	Chr02K_63.687871 65.484856
	  Chr09K_13.655791 18.177373
		Chr03K_17.368264	Chr03K_26.170048
		Chr08N_5.54356	Chr08N_34.415808

# Chr02N  58.039498 61.274395
```{r}
switchgrassGWAS::mash_plot_effects(m=mash, i = 11087)
switchgrassGWAS::mash_plot_effects(m=mash, i = 11468)

df1 %>%
  filter(log10bf > 5) # 25 values
df1 %>%
  filter(log10bf > 4) # 93 values (1800 above 2)
```
Close
	Chr02N_58.039498 61.274395
Chr02N_15281005	9405	4.057999	1	
Chr02N_50051440	10742	4.288475	5	5.8214656195	
Chr02N_56422155	11087	5.158628	6	5.9439138054	
Chr02N_57947932	11178	4.390894	6	4.7648643926	
Chr02N_63420472	11467	4.374355	5	5.5078867975	
Chr02N_63420484	11468	5.022484	6	-5.0720890743

Close, first is better
	Chr05N_1.502713 4.328965
	Chr05N_58.916323 65.990782
Chr05N_4512043	25413	5.929161	6	8.5445301065	
Chr05N_20476771	26421	4.303708	5	6.4982493387	
Chr05N_24505333	26560	4.225575	5	7.4869178552	
Chr05N_25731508	26609	4.081614	1	-0.0036192576	
Chr05N_28066146	26673	4.727816	6	8.9999136556	
Chr05N_36999363	26879	5.038580	5	-5.0653499084	
Chr05N_37110671	26885	4.151390	5	6.1070001182	
Chr05N_38945328	26961	4.980246	6
Chr05N_46107016	27330	4.099486	6	3.6428083879	
Chr05N_65588952	28354	4.003577	6

Close
	Chr04K_1.959391 5.499253
Chr04K_1935684	17557	4.235319	5	6.1526138902	
Chr04K_6537440	17814	4.403893	5	4.7144521481	
Chr04K_6567024	17818	4.962056	6	5.1767913940	
Chr04K_8657535	17940	5.205059  6
Chr04K_12651532	18152	6.096971  6

Not close
	Chr02K_63.687871 65.484856
Chr02K_23509869	6711	5.595545	6	-6.7548782752	
Chr02K_23682427	6713	5.430460	5	6.9669557271	
Chr02K_35881338	6882	4.095730	6

```{r}
panela <- df1 %>%
  filter(grepl("Chr02N_(5|6)", Marker) & between(value, 11000, 11500)) %>%
  separate(Marker, into = c("Chr", "_", "Position"), sep = c(6,7), remove = FALSE) %>%
  mutate(Position = as.numeric(Position)) %>%
  ggplot(aes(x=Position, y = log10bf)) +
  geom_point(aes(color = sigcond)) +
  scale_color_viridis(option = "B", end = 0.8) # 26711
save_plot(filename = "Chr02N_58_to_61Mb_D2F_mash_result_zoom.png", plot = panela, base_asp = 1.2)

panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11087)
#panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11467)
##panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11468)
#panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11178)

p3 <- plot_grid(panela, panelb$ggobject, rel_widths = c(2,1))
save_plot(filename = "Chr02N_58_to_61Mb_mash_result_zoom_2panel.png", plot = p3, base_asp = 1.5)
```
# Chr05N 
```{r}

panela <- df1 %>%
  filter(grepl("Chr05N_", Marker) & between(value, 24000, 25600)) %>%
  separate(Marker, into = c("Chr", "_", "Position"), sep = c(6,7), remove = FALSE) %>%
  mutate(Position = as.numeric(Position)) %>%
  ggplot(aes(x=Position, y = log10bf)) +
  geom_point(aes(color = sigcond)) +
  scale_color_viridis(option = "B", end = 0.8) # 26711

panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 25413)
#panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11467)
##panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11468)
#panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11178)

(p3 <- plot_grid(panela, panelb$ggobject, rel_widths = c(2,1)))
save_plot(filename = "Chr05N_0_to_8Mb_D2F_mash_result_zoom.png", plot = panela, base_asp = 1.2)
save_plot(filename = "Chr05N_0_to_8Mb_mash_result_zoom_2panel.png", plot = p3, base_asp = 1.5)
```

# Chr05N 60Mb
```{r}

panela <- df1 %>%
  filter(grepl("Chr05N_(5|6)", Marker) & between(value, 28000, 28500)) %>%
  separate(Marker, into = c("Chr", "_", "Position"), sep = c(6,7), remove = FALSE) %>%
  mutate(Position = as.numeric(Position)) %>%
  ggplot(aes(x=Position, y = log10bf)) +
  geom_point(aes(color = sigcond)) +
  scale_color_viridis(option = "B", end = 0.8) # 26711

panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 28354)
#panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11467)
##panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11468)
#panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11178)

(p3 <- plot_grid(panela, panelb$ggobject, rel_widths = c(2,1)))
save_plot(filename = "Chr05N_58_to_68Mb_D2F_mash_result_zoom.png", plot = panela, base_asp = 1.2)
save_plot(filename = "Chr05N_58_to_68Mb_mash_result_zoom_2panel.png", plot = p3, base_asp = 1.5)
```

# Chr04K

Close
	Chr04K_1.959391 5.499253
Chr04K_1935684	17557	4.235319	5	6.1526138902	
Chr04K_6537440	17814	4.403893	5	4.7144521481	
Chr04K_6567024	17818	4.962056	6	5.1767913940	
Chr04K_8657535	17940	5.205059  6
Chr04K_12651532	18152	6.096971  6
```{r}

panela <- df1 %>%
  filter(grepl("Chr04K_", Marker) & between(value, 17500, 18200)) %>%
  separate(Marker, into = c("Chr", "_", "Position"), sep = c(6,7), remove = FALSE) %>%
  mutate(Position = as.numeric(Position)) %>%
  ggplot(aes(x=Position, y = log10bf)) +
  geom_point(aes(color = sigcond)) +
  scale_color_viridis(option = "B", end = 0.8) # 26711

panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 18152)
#panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 17557)
##panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 17814)
#panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 17818)
 switchgrassGWAS::mash_plot_effects(m=mash, i = 17940)

(p3 <- plot_grid(panela, panelb$ggobject, rel_widths = c(2,1)))
save_plot(filename = "Chr04K_1_to_12Mb_D2F_mash_result_zoom.png", plot = panela, base_asp = 1.2)
save_plot(filename = "Chr04K_1_to_12Mb_mash_result_zoom_2panel.png", plot = p3, base_asp = 1.5)
```

# Chr08N_5.54356	Chr08N_34.415808
```{r}

panela <- df1 %>%
  filter(grepl("Chr08N_", Marker) & between(value, 40000, 41100)) %>%
  separate(Marker, into = c("Chr", "_", "Position"), sep = c(6,7), remove = FALSE) %>%
  mutate(Position = as.numeric(Position)) %>%
  ggplot(aes(x=Position, y = log10bf)) +
  geom_point(aes(color = sigcond)) +
  scale_color_viridis(option = "B", end = 0.8) # 26711

panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 40305)
#panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 40514)

(p3 <- plot_grid(panela, panelb$ggobject, rel_widths = c(2,1)))
save_plot(filename = "Chr08N_0_to_1Mb_D2F_mash_result_zoom.png", plot = panela, base_asp = 1.2)
save_plot(filename = "Chr08N_0_to_1Mb_mash_result_zoom_2panel.png", plot = p3, base_asp = 1.5)
```

# Li's QTL CSV

```{r}
qtl <- qtl %>%
  separate(flank_hi, into = c("Chr_hi", "_hi", "Position_hi"), sep = c(6,7), remove = FALSE) %>%
  separate(flank_lo, into = c("Chr_lo", "_lo", "Position_lo"), sep = c(6,7), remove = FALSE) %>%
  mutate(Position_hi = as.numeric(Position_hi)*1000000,
         Position_lo = as.numeric(Position_lo)*1000000)
qtl
```

```{r}

# load("~/Github/pvdiv-phenology-gxe/data/sites.rda")

levelsSN = rev(c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING"))
levelsNS = c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING")
```

```{r}
i=1
width = 1000000

for(i in 1:nrow(qtl)){
df_a <- df1 %>%
  separate(Marker, into = c("Chr", "_", "Position"), sep = c(6,7), remove = FALSE) %>%
  mutate(Position = as.numeric(Position)) %>%
  filter(grepl(qtl$chr[i], Chr) & between(Position, qtl$Position_lo[i]-width, qtl$Position_hi[i]+width)) 
panel_a <- df_a %>%
  ggplot(aes(x=Position, y = log10bf)) +
  geom_point(aes(color = sigcond)) +
  scale_color_viridis(option = "B", end = 0.9)  + 
  labs(x = "Position", y = "log10(Bayes Factor)") +
  geom_label(data = df_a[df_a$log10bf==max(df_a$log10bf),], mapping = aes(x = Position, y = log10bf, label = Marker), nudge_y = -0.4)

panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = df_a[df_a$log10bf==max(df_a$log10bf),]$value)
df_b <- panelb$effect_df %>%
  mutate(Site = str_sub(value, start =11, end = 14)) %>%
  left_join(sites) 
sites$Site <- factor(sites$Site, levels = levelsSN)
df_b$Site <- factor(df_b$Site, levels = levelsSN)
panel_b <- df_b %>%
  ggplot(aes(x = Site)) +
    geom_point(mapping = aes(y = mn)) +
    geom_errorbar(mapping = aes(ymin = mn - se,
                                ymax = mn + se, width = 0.3)) +
    geom_hline(yintercept = 0, lty = 2) +
    labs(x = "Conditions", y = "Effect Size") + coord_flip()

p3 <- plot_grid(panel_a, panel_b, rel_widths = c(2,1))
save_plot(filename = paste(qtl$Chr_lo[i], qtl$Position_lo[i], "to", qtl$Position_hi[i], "FL50_mash_result_zoom.png", sep = "_"), plot = p3, base_asp = 3)
}
```

### 2020-04-24 R/rrBLUP_analysis_D2F.Rmd
D2F environmental index analysis for two indices and CV of phenotypes

---
title: "BLUP anaysis"
author: "Alice MacQueen"
date: "8/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rlang)
library(rrBLUP)
library(switchgrassGWAS)
```

## Present BLUPs/breeding values

```{r load data}
V_g <- readRDS(file.path("~", "Github", "pvdiv-phenology-gxe", "data",
                            paste0("Kinship_van_Raden_method_363g_",
                                   "Midwest_and_Gulf_subset_maf0.05.rds")))
phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Phenotypes_D2F_fourway_project.rds")
metadata <- read_rds(file.path("~", "Github", "pvdiv-phenotypes", "data", "metadata.rds")) %>%
  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
                            !is.na(PLOIDY) ~ as.character(PLOIDY),
                            TRUE ~ NA_character_))

phe_site <- phe %>%
  pivot_longer(cols = -PLANT_ID, names_to = "SITE_PHE", values_to = "MEAS") %>%
  separate(SITE_PHE, into = c("SITE", "PHE")) %>%
  pivot_wider(names_from = "PHE", values_from = "MEAS")

sites <- read_rds("~/Github/pvdiv-phenotypes/data/sites.rds")
fourway_sites <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL")
```

```{r}
phe_post <- readRDS("~/Github/pvdiv-phenotypes/data/post_replant/Phenotypes_long_non_Blackwell_replants_post_2019_replant.rds")
phe_pre <- readRDS("~/Github/pvdiv-phenotypes/data/pre_replant/Phenotypes_all_pre_2019_replant.rds")

phe_pre2 <- phe_pre %>% 
  mutate(PLANT_ID = case_when(grepl("AP13", PLANT_ID) ~ "AP13",
                              TRUE ~ PLANT_ID)) %>%
  dplyr::select(SITE, PLOT_GL, PLANT_ID, TC_EOS_2018, GR1:GR50)

phe_post2 <- phe_post %>%
  mutate(PLANT_ID = case_when(grepl("AP13", PLANT_ID) ~ "AP13",
                              TRUE ~ PLANT_ID)) %>%
  separate(SITE_PHE, into = c("site", "PHE"), sep = c(5)) %>%
  dplyr::select(-site) %>%
  filter(PHE %in% c("GR50","FL1", "FL50", "BIOMASS", "TC_EOS", "HT_PAN_EOS", "EM1", "EM50")) %>%
   pivot_wider(names_from = "PHE", values_from = "MEAS") %>%
  group_by(PLANT_ID, SITE) %>% add_tally(name = "Per_site_count") %>%
  dplyr::select(SITE:PLANT_ID, Per_site_count, everything())

phe_all <- phe_post2 %>%
  left_join(phe_pre2, by = c("SITE", "PLOT_GL", "PLANT_ID"))
```

# Instead of D2F, PTT2F? photothermal time to flowering, from greenup.

```{r}
ptt_df <- readRDS(file = "~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")

gdd_df <- ptt_df %>%
  mutate(GDD_0C_SM = ifelse((TMAX_SM + TMIN_SM)/2 >= 0,
                      (TMAX_SM + TMIN_SM)/2 - 0,
                      0)) %>%
  dplyr::select(DOY, SITE, Day_of_year, DYLN, GDD_13C_SM:GDD_8C_SM, GDD_0C_SM)

GR_start <- phe_all %>% 
  left_join(metadata, by = "PLANT_ID") %>%
  #filter(D2F > 10 & SUBPOP == "Midwest") %>%
  ungroup() %>%
  group_by(SITE, PLOT_GL, PLANT_ID) %>%
  summarise(GR_start = floor(min(GR50)))
FL_end <- phe_all %>% 
  ungroup() %>%
  group_by(SITE, PLOT_GL, PLANT_ID) %>%
  summarise(FL_end = ceiling(max(FL50)))

tmp2 <- gdd_df %>%
  dplyr::select(DOY, SITE, Day_of_year, DYLN, GDD_13C_SM:GDD_0C_SM) %>%
  left_join(GR_start, by = "SITE") %>%
  left_join(FL_end, by = c("SITE", "PLOT_GL", "PLANT_ID")) %>%
  group_by(SITE, Day_of_year, PLOT_GL, PLANT_ID) %>%
  filter(between(Day_of_year, GR_start, FL_end)) %>% 
  ungroup()

tmp2 %>%
  group_by(SITE, PLOT_GL, PLANT_ID, GR_start, FL_end) %>%
  summarise(CGDD_0C = sum(GDD_0C_SM),
            CGDD_8C = sum(GDD_8C_SM),
            CGDD_13C = sum(GDD_13C_SM),
            CDYLN = sum(DYLN),
            min_dyln = min(DYLN),
            mean_dyln = mean(DYLN),
            max_dyln = max(DYLN),
            sd_dyln = sd(DYLN),
            cv_dyln = mean_dyln / sd_dyln
            ) %>%
  left_join(metadata) %>%
  dplyr::select(-(LIBRARY:PLANTED_2018)) %>%
  left_join(phe_all)
  mutate(D2F = FL_end - GR_start) %>%
  filter(D2F > 10) %>%
  ggplot(aes(x = CGDD_13C, y = CDYLN)) + 
  geom_point(aes(color = SUBPOP)) + facet_wrap(~SITE)

phe_meta <- phe_all %>%
  left_join(metadata) %>%
  dplyr::select(-(LIBRARY:PLANTED_2018))

library(lubridate)
library(viridis)

tmp2$SITE <- factor(tmp2$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
phe_meta$SITE <- factor(phe_meta$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
tmp2 %>%
  dplyr::select(DOY, DYLN, SITE) %>%
  unique() %>%
  filter(SITE != "OVTN") %>%
  ggplot(aes(x = DOY, y = DYLN)) + 
    geom_point() + 
    facet_wrap(~SITE) + 
    geom_jitter(data = filter(phe_meta, SITE != "OVTN" & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(FL1, origin = "2018-12-31"), color = SUBPOP, y = 6.5), width = 0, height = 1, alpha = 0.15) + ylim(c(3,16)) + 
    geom_jitter(data = filter(phe_meta, SITE != "OVTN" & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(FL50, origin = "2018-12-31"), color = SUBPOP, y = 4), width = 0, height = 1, alpha = 0.2) +
  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1) + xlab("Day of year") + ylab("Daylength (hrs)") +
  geom_vline(xintercept = as_date(365/2-8, origin = "2018-12-31"), linetype = 2) +
  geom_text(aes(y = 4, x = as_date(40, origin = "2018-12-31"), label = "FL50")) +
  geom_text(aes(y = 6.5, x = as_date(32, origin = "2018-12-31"), label = "FL1")) +
  geom_text(aes(y = 9, x = as_date(40, origin = "2018-12-31"), label = "GR50")) + 
    geom_jitter(data = filter(phe_meta, SITE != "OVTN" & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(GR50, origin = "2018-12-31"), color = SUBPOP, y = 9), width = 0, height = 1, alpha = 0.1)

save_plot(filename = "../analysis/environmental-index/Daylength_vs_GR_FL1_FL50_Gulf_and_Midwest_10_sites_2019_longest_day_line.png", plot = last_plot(), base_height = 6)

```

# Analysis for pvdiv like I did for the CDBN
```{r}
# PHE <- list(quote(BRKG_D2F), quote(CLMB_D2F))
PHE <- list(quote(FL50), quote(GR50), quote(BIOMASS), quote(TC_EOS), 
            quote(HT_PAN_EOS), quote(FL1), quote(EM1), quote(EM50),
            quote(TC_EOS_2018), quote(GR1))
rrkinblupout <- list()
kinseqblup <- list()
bluptable <- list()

for(i in seq_along(PHE)){

  testdf <- phe_all %>%
    left_join(metadata) %>%
    filter(!is.na(eval(PHE[[i]])) & GENETIC_SUBPOP_KINSHIP %in% c("Midwest", 
                                                                    "Gulf")) %>%
    as.data.frame()

  bluprrk <- kin.blup(data = testdf, geno = "PLANT_ID", pheno = deparse(PHE[[i]]),
                        K = V_g, fixed = "SITE") # , fixed = c("Location_code", "LbY")

  rrkinblupout[[i]] <- enframe(bluprrk$g, name = "PLANT_ID", 
                               value = deparse(PHE[[i]]))

    # Which Taxa were phenotyped for this phenotype? Just keep those for GWAS.
  taxa_phenotyped <- testdf %>%
    dplyr::select(PLANT_ID, !! PHE[[i]]) %>%
    as_tibble() %>%
    group_by(PLANT_ID) %>%
    summarise(count = n())
  kinseqblup[[i]] <- taxa_phenotyped %>%
    left_join(rrkinblupout[[i]], by = "PLANT_ID")

  bluptable[[i]] <- tibble(Phenotype = deparse(PHE[[i]]), V_a = bluprrk$Vg, 
                      V_e = bluprrk$Ve) %>%
    mutate(h2 = V_a / (V_a + V_e))

  saveRDS(bluprrk, paste0("~/Github/pvdiv-phenology-gxe/data-raw/ignore/", 
                          PHE[[i]],
                          "_363g_Gulf_and_Midwest_10Site_BLUPs_kinship.rds"))
}
```
From Houle 1992:

Crow 1958: "opportunity for selection": I = V_P / X_bar^2
X_bar represents the population mean before selection.
V_P phenotypic trait variance. `var()`

To calculate CV (coefficient of variation) you take the standard deviation of the data and divide it by the mean of the data.  sqrt(V_P) / X_bar

I is related to the phenotypic coefficient of variation, CV_P = 100*sqrt(V_P) / X_bar
(this is just CV * 100 to be a percentage).

Relative evolvability of fitness:

I_A = V_A / X_bar^2 = (CV_A / 100)^2
I_A is also an appropriate indicator of the variability of fitness as large values must be the result of powerful forces maintaining genetic variance.

A variance to mean ratio is the most appropriate measure of the evolvability of a trait.

Compared three quantities:

Residual coefficient of variation CV_R = 100 * sqrt(V_P - V_A) / X_bar

h^2 narrow sense heritability = V_A / V_P

CV_A = 100*sqrt(V_A) / X_bar

```{r}
phe_all %>% left_join(metadata) %>%
  pivot_longer(values_to = "MEAS", cols = TC_EOS:GR50, names_to = "PHE") %>%
  dplyr::select(-LIBRARY, -(GENETIC_SUBPOP_KINSHIP:PLANTED_2018)) %>%
  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
  group_by(SUBPOP, SITE, PHE) %>%
  summarise(X_bar = mean(MEAS, na.rm = TRUE),
            V_P = var(MEAS, na.rm = TRUE),
            I = V_P / X_bar^2, 
            CV_P = 100*sqrt(V_P) / X_bar) %>%
  arrange(desc(I))

BLUPs <- select(metadata, PLANT_ID, SUBPOP) %>%
  filter(SUBPOP %in% c("Midwest", "Gulf")) %>%
  left_join(dplyr::select(kinseqblup[[1]], PLANT_ID, 3))

for(i in seq_along(kinseqblup)[-1]){
  BLUPs <- BLUPs %>%
    left_join(dplyr::select(kinseqblup[[i]], PLANT_ID, 3))
}

BLUPtable <- bluptable[[1]]

for(i in seq_along(bluptable)[-1]){
  BLUPtable <- BLUPtable %>%
    full_join(bluptable[[i]])
}

BLUPtable
phe_all %>% left_join(metadata) %>%
  pivot_longer(values_to = "MEAS", cols = TC_EOS:GR50, names_to = "PHE") %>%
  dplyr::select(-LIBRARY, -(GENETIC_SUBPOP_KINSHIP:PLANTED_2018)) %>%
  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
  group_by(PHE) %>%
  summarise(X_bar = mean(MEAS, na.rm = TRUE),
            V_P = var(MEAS, na.rm = TRUE),
            I = V_P / X_bar^2, 
            CV_P = 100*sqrt(V_P) / X_bar) %>%
  arrange(desc(I)) %>%
  full_join(BLUPtable, by = c("PHE" = "Phenotype")) %>%
  mutate(I_A = V_a / X_bar^2,
         CV_R = 100 * sqrt(V_P - V_a) / X_bar, 
         CV_A = 100 * sqrt(V_a) / X_bar,
         CV_E = 100 * sqrt(V_e) / X_bar) %>%
  write_csv(., "~/Github/pvdiv-phenology-gxe/data-raw/ignore/Phenotypic_CVs_10_site_fixed_effects_363g_Gulf_and_Midwest.csv")
```
# Start here
- want to break down heritabilities etc by site and by subpop
- want to do this for D2F and GDD_GR2FL also. Find these quantities for phe_all.

# h^2 etc for Midwest
```{r}
# PHE <- list(quote(BRKG_D2F), quote(CLMB_D2F))
PHE <- list(quote(FL50), quote(GR50), quote(BIOMASS), quote(TC_EOS), 
            quote(HT_PAN_EOS), quote(FL1), quote(EM1), quote(EM50),
            quote(TC_EOS_2018), quote(GR1))
rrkinblupout <- list()
kinseqblup <- list()
bluptable <- list()

for(i in seq_along(PHE)){

  testdf <- phe_all %>%
    left_join(metadata) %>%
    filter(!is.na(eval(PHE[[i]])) & GENETIC_SUBPOP_KINSHIP %in% c("Midwest", 
                                                                    "Gulf")) %>%
    as.data.frame()

  bluprrk <- kin.blup(data = testdf, geno = "PLANT_ID", pheno = deparse(PHE[[i]]),
                        K = V_g, fixed = "SITE") # , fixed = c("Location_code", "LbY")

  rrkinblupout[[i]] <- enframe(bluprrk$g, name = "PLANT_ID", 
                               value = deparse(PHE[[i]]))

    # Which Taxa were phenotyped for this phenotype? Just keep those for GWAS.
  taxa_phenotyped <- testdf %>%
    dplyr::select(PLANT_ID, !! PHE[[i]]) %>%
    as_tibble() %>%
    group_by(PLANT_ID) %>%
    summarise(count = n())
  kinseqblup[[i]] <- taxa_phenotyped %>%
    left_join(rrkinblupout[[i]], by = "PLANT_ID")

  bluptable[[i]] <- tibble(Phenotype = deparse(PHE[[i]]), V_a = bluprrk$Vg, 
                      V_e = bluprrk$Ve) %>%
    mutate(h2 = V_a / (V_a + V_e))

  saveRDS(bluprrk, paste0("~/Github/pvdiv-phenology-gxe/data-raw/ignore/", 
                          PHE[[i]],
                          "_363g_Gulf_and_Midwest_10Site_BLUPs_kinship.rds"))
}
```
From Houle 1992:

Crow 1958: "opportunity for selection": I = V_P / X_bar^2
X_bar represents the population mean before selection.
V_P phenotypic trait variance. `var()`

To calculate CV (coefficient of variation) you take the standard deviation of the data and divide it by the mean of the data.  sqrt(V_P) / X_bar

I is related to the phenotypic coefficient of variation, CV_P = 100*sqrt(V_P) / X_bar
(this is just CV * 100 to be a percentage).

Relative evolvability of fitness:

I_A = V_A / X_bar^2 = (CV_A / 100)^2
I_A is also an appropriate indicator of the variability of fitness as large values must be the result of powerful forces maintaining genetic variance.

A variance to mean ratio is the most appropriate measure of the evolvability of a trait.

Compared three quantities:

Residual coefficient of variation CV_R = 100 * sqrt(V_P - V_A) / X_bar

h^2 narrow sense heritability = V_A / V_P

CV_A = 100*sqrt(V_A) / X_bar

```{r}
phe_all %>% left_join(metadata) %>%
  pivot_longer(values_to = "MEAS", cols = TC_EOS:GR50, names_to = "PHE") %>%
  dplyr::select(-LIBRARY, -(GENETIC_SUBPOP_KINSHIP:PLANTED_2018)) %>%
  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
  group_by(SUBPOP, SITE, PHE) %>%
  summarise(X_bar = mean(MEAS, na.rm = TRUE),
            V_P = var(MEAS, na.rm = TRUE),
            I = V_P / X_bar^2, 
            CV_P = 100*sqrt(V_P) / X_bar) %>%
  arrange(desc(I))

BLUPs <- select(metadata, PLANT_ID, SUBPOP) %>%
  filter(SUBPOP %in% c("Midwest", "Gulf")) %>%
  left_join(dplyr::select(kinseqblup[[1]], PLANT_ID, 3))

for(i in seq_along(kinseqblup)[-1]){
  BLUPs <- BLUPs %>%
    left_join(dplyr::select(kinseqblup[[i]], PLANT_ID, 3))
}

BLUPtable <- bluptable[[1]]

for(i in seq_along(bluptable)[-1]){
  BLUPtable <- BLUPtable %>%
    full_join(bluptable[[i]])
}

BLUPtable
phe_all %>% left_join(metadata) %>%
  pivot_longer(values_to = "MEAS", cols = TC_EOS:GR50, names_to = "PHE") %>%
  dplyr::select(-LIBRARY, -(GENETIC_SUBPOP_KINSHIP:PLANTED_2018)) %>%
  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
  group_by(PHE) %>%
  summarise(X_bar = mean(MEAS, na.rm = TRUE),
            V_P = var(MEAS, na.rm = TRUE),
            I = V_P / X_bar^2, 
            CV_P = 100*sqrt(V_P) / X_bar) %>%
  arrange(desc(I)) %>%
  full_join(BLUPtable, by = c("PHE" = "Phenotype")) %>%
  mutate(I_A = V_a / X_bar^2,
         CV_R = 100 * sqrt(V_P - V_a) / X_bar, 
         CV_A = 100 * sqrt(V_a) / X_bar,
         CV_E = 100 * sqrt(V_e) / X_bar) %>%
  write_csv(., "~/Github/pvdiv-phenology-gxe/data-raw/ignore/Phenotypic_CVs_10_site_fixed_effects_363g_Gulf_and_Midwest.csv")
```

Table:
Phenotype: Used to get the BLUP solution for the genetic values, or breeding values
Vg: REML estimate of the genetic variance
Ve: REML estimate of the error variance
h2: Narrow sense heritability, Vg / (Vg + Ve)

```{r}
i = 1

rrblupout <- readRDS(paste0("../data-raw/", PHE[[i]], "_BLUPs_kinship.rds"))

bluprange <- max(rrblupout$g) - min(rrblupout$g)
ggplot(METG_Seq327) +
  geom_histogram(aes(x = !! sym(PHE[[i]])))

CIrow <- METG_Seq327 %>%
  summarise(lowCI = quantile(!! sym(PHE[[i]]), 0.025, na.rm = TRUE),
            upCI = quantile(!! sym(PHE[[i]]), 0.975, na.rm = TRUE),
            count = sum(!is.na(!! sym(PHE[[i]]))))
CItable <- tibble(Phenotype = deparse(PHE[[i]]), `lower bound` = CIrow$lowCI,
                  `upper bound` = CIrow$upCI, range = bluprange, count = CIrow$count)

for(i in seq_along(PHE)[-1]){
  rrblupout <- readRDS(paste0("../data-raw/", PHE[[i]], "_BLUPs_kinship.rds"))
  bluprange <- max(rrblupout$g) - min(rrblupout$g)
  CIrow <- METG_Seq327 %>%
  summarise(lowCI = quantile(!! sym(PHE[[i]]), 0.025, na.rm = TRUE),
            upCI = quantile(!! sym(PHE[[i]]), 0.975, na.rm = TRUE),
            count = sum(!is.na(!! sym(PHE[[i]]))))
  CItable <- add_row(CItable, Phenotype = deparse(PHE[[i]]), `lower bound` = CIrow$lowCI,
                  `upper bound` = CIrow$upCI, range = bluprange, count = CIrow$count)
}

blupstable <- bluptable %>% 
  left_join(CItable)

write_csv(blupstable, path = "BLUP_Descriptive_Statistics.csv")

blupstable <- blupstable %>%
  add_column(sigFDR = c(0, 0, 0, 0, 1, 1, 0, 0, 1, 0, 1, 1, 0, 0, 1, 0, 0, 1, 1, 0, 0))

wilcox.test(x = blupstable$count[blupstable$sigFDR == 0], y = blupstable$count[blupstable$sigFDR == 1], paired = FALSE, alternative = "l")
wilcox.test(x = blupstable$h2[blupstable$sigFDR == 0], y = blupstable$h2[blupstable$sigFDR == 1], paired = FALSE, alternative = "l")
```

### 2020-04-24 R/get_results_from_mash.R

@@ -1,0 +1,54 @@
+#' @title Get column names from a mash object
+#'
+#' @description This function extracts the column names from the local false
+#' sign rate table of a mash object's results. This can tell you the condition
+#' names or phenotype names used in the mash object. That can be useful for
+#' looking at a subset of these columns, say.
+#'
+#' @param m An object of type mash
+#'
+#' @return A vector of phenotype names
+#'
+#' @examples
+#'     \dontrun{get_colnames(m = mash_obj)}
+#'
+get_colnames <- function(m){
+  column_names <- colnames(m$result$lfsr)
+  return(column_names)
+}
+
+#' From a mash result, get effects that are significant in at least
+#'     one condition.
+#'
+#' @param m the mash result (from joint or 1by1 analysis)
+#' @param thresh indicates the threshold below which to call signals significant
+#' @param conditions which conditions to include in check (default to all)
+#' @param sig_fn the significance function used to extract significance from
+#'     mash object; eg could be ashr::get_lfsr or ashr::get_lfdr. (Small values
+#'     must indicate significant.)
+#'
+#' @return a vector containing the indices of the significant effects, by
+#'     order of most significant to least
+#'
+#' @importFrom ashr get_lfsr
+#'
+#' @export
+get_significant_results = function(m, thresh = 0.05, conditions = NULL,
+                                   sig_fn = ashr::get_lfsr) {
+  if (is.null(conditions)) {
+    conditions = 1:get_ncond(m)
+  }
+  top = apply(sig_fn(m)[,conditions,drop=FALSE],1,min) # find top effect
+  # in each condition
+  sig = which(top < thresh)
+  ord = order(top[sig],decreasing=FALSE)
+  sig[ord]
+}
+
+require(ashr)
+require(tidyverse)
+effectplot <- get_colnames(m) %>%
+  enframe(name = NULL, value = "Conditions") %>%
+  mutate(Conditions = str_sub(.data$Conditions, start = 6),
+         mn = get_pm(m)[i,],
+         se = get_psd(m)[i,])

### 2020-04-24 R/mash_d2f_gulf_midwest_set.R

@@ -1,0 +1,152 @@
+library(tidyverse)
+library(bigsnpr)
+library(switchgrassGWAS)
+
+#  devtools::install_github("alice-macqueen/switchgrassgwas")
+inputfiles <- read_delim("~/Github/pvdiv-fitness-2019/analysis/inputkinship.txt", 
+                         delim = " ", col_names = "SNPfiles")
+outputdir <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", "gwas", 
+                       "Gulf_Midwest")
+
+snp <- snp_attach(inputfiles$SNPfiles[4])
+gwas_rds <- pvdiv_results_in_folder(path = outputdir, 
+                                    pattern = "GWAS_datatable_(PKLE|STIL|KING|KBSM|CLMB|BRKG|LINC)_D2F") 
+# Not enough Gulf & Midwest individuals at BRKG to run the univariate GWAS (<120).
+phenotype_names <- str_sub(gwas_rds, start = 16, end = 23)
+numSNPs <- ceiling(ceiling(1000000/length(phenotype_names)^2)/500)*500 # round 
+# up number of SNPs to nearest 500.
+
+mashin <- pvdiv_bigsnp2mashr(path = outputdir, 
+                             snp = snp, gwas_rds = gwas_rds, numSNPs = numSNPs, 
+                             phenotypes = phenotype_names, model = "linear", 
+                             saveoutput = TRUE, scale = FALSE)
+
+m <- mash_standard_run(path = outputdir, list_input = mashin,
+                       saveoutput = TRUE, ref = NA)
+
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+mash_plot_manhattan_by_condition(m = m, saveoutput = TRUE)
+mash_plot_sig_by_condition(m = m, saveoutput = TRUE)
+
+
+library(AnnotationDbi)
+txdb <- loadDb(file = file.path("~", "Github", "pvdiv-genome",
+                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
+annos <- pvdiv_table_topsnps(df = m, type = "mash", txdb = txdb, 
+                             snp = snp)
+saveRDS(annos, file = file.path(outputdir, paste0("Annotation_tables_mash_", 
+                                                  "fourway_sites_", 
+                                                  "Midwest_Gulf_D2F.rds")))
+
+library(ashr)
+for(n in 1:10){
[…]

### 2020-04-29 R/h2_var_p_and_cov_p.Rmd
rrBLUP and h^2 analysis for D2F

---
title: "h2 and cov_P"
author: "Alice MacQueen"
date: "4/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
```

#

```{r}
phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Phenotypes_D2F_fourway_project.rds")
metadata <- readRDS("~/Github/pvdiv-phenotypes/data/metadata.rds")
metadata <- metadata %>%
  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
                            !is.na(PLOIDY) ~ as.character(PLOIDY),
                            TRUE ~ NA_character_))
```

### 2020-04-29 R/rrBLUP_analysis_D2F.Rmd
rrBLUP and h^2 analysis for D2F

@@
 
+cov
+
 BLUPs <- select(metadata, PLANT_ID, SUBPOP) %>%
@@
   dplyr::select(-LIBRARY, -(GENETIC_SUBPOP_KINSHIP:PLANTED_2018)) %>%
-  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
+  filter(SUBPOP %in% c("Gulf", "Midwest", "Atlantic")) %>%
   group_by(SUBPOP, SITE, PHE) %>%
@@
             CV_P = 100*sqrt(V_P) / X_bar) %>%
-  arrange(desc(I))
+  arrange(PHE, SITE, SUBPOP)
 
@@
 ```
+
+# Per subpop analysis
+
+# Analysis for pvdiv like I did for the CDBN
+```{r}
+# PHE <- list(quote(BRKG_D2F), quote(CLMB_D2F))
+PHE <- list(quote(FL50), quote(GR50), quote(BIOMASS), quote(TC_EOS), 
+            quote(HT_PAN_EOS), quote(FL1), quote(EM1), quote(EM50),
+            quote(TC_EOS_2018), quote(GR1))
+rrkinblupout <- list()
+kinseqblup <- list()
+bluptable <- list()
+
+for(i in seq_along(PHE)){
+
+  testdf <- phe_all %>%
+    left_join(metadata) %>%
+    filter(!is.na(eval(PHE[[i]])) & GENETIC_SUBPOP_KINSHIP %in% c("Midwest")) %>%
+    as.data.frame()
+
+  bluprrk <- kin.blup(data = testdf, geno = "PLANT_ID", pheno = deparse(PHE[[i]]),
+                        K = V_g, fixed = "SITE") # , fixed = c("Location_code", "LbY")
+
+  rrkinblupout[[i]] <- enframe(bluprrk$g, name = "PLANT_ID", 
+                               value = deparse(PHE[[i]]))
+
+    # Which Taxa were phenotyped for this phenotype? Just keep those for GWAS.
+  taxa_phenotyped <- testdf %>%
+    dplyr::select(PLANT_ID, !! PHE[[i]]) %>%
+    as_tibble() %>%
+    group_by(PLANT_ID) %>%
+    summarise(count = n())
+  kinseqblup[[i]] <- taxa_phenotyped %>%
+    left_join(rrkinblupout[[i]], by = "PLANT_ID")
+
+  bluptable[[i]] <- tibble(Phenotype = deparse(PHE[[i]]), V_a = bluprrk$Vg, 
+                      V_e = bluprrk$Ve) %>%
+    mutate(h2 = V_a / (V_a + V_e))
+
+  saveRDS(bluprrk, paste0("~/Github/pvdiv-phenology-gxe/data-raw/ignore/", 
+                          PHE[[i]],
+                          "_134g_Midwest_10Site_BLUPs_kinship.rds"))
+}
+```
+
+```{r}
+BLUPs <- select(metadata, PLANT_ID, SUBPOP) %>%
+  filter(SUBPOP %in% c("Midwest")) %>%
+  left_join(dplyr::select(kinseqblup[[1]], PLANT_ID, 3))
+
+for(i in seq_along(kinseqblup)[-1]){
+  BLUPs <- BLUPs %>%
+    left_join(dplyr::select(kinseqblup[[i]], PLANT_ID, 3))
+}
+
+BLUPtable <- bluptable[[1]]
+
+for(i in seq_along(bluptable)[-1]){
+  BLUPtable <- BLUPtable %>%
+    full_join(bluptable[[i]])
+}
+
+BLUPtable
+phe_all %>% left_join(metadata) %>%
+  pivot_longer(values_to = "MEAS", cols = TC_EOS:GR50, names_to = "PHE") %>%
+  dplyr::select(-LIBRARY, -(GENETIC_SUBPOP_KINSHIP:PLANTED_2018)) %>%
+  filter(SUBPOP %in% c("Midwest")) %>%
+  group_by(PHE) %>%
+  summarise(X_bar = mean(MEAS, na.rm = TRUE),
+            V_P = var(MEAS, na.rm = TRUE),
+            I = V_P / X_bar^2, 
+            CV_P = 100*sqrt(V_P) / X_bar) %>%
+  arrange(desc(I)) %>%
+  full_join(BLUPtable, by = c("PHE" = "Phenotype")) %>%
+  mutate(I_A = V_a / X_bar^2,
+         CV_R = 100 * sqrt(V_P - V_a) / X_bar, 
+         CV_A = 100 * sqrt(V_a) / X_bar,
+         CV_E = 100 * sqrt(V_e) / X_bar) %>%
+  write_csv(., "~/Github/pvdiv-phenology-gxe/data-raw/ignore/Phenotypic_CVs_10_site_fixed_effects_134g_Midwest.csv")
+```

### 2020-04-30 R/rrBLUP_analysis_D2F.Rmd
Analysis for Var_P and heritability within sites and subpops for flowering

@@
 library(switchgrassGWAS)
+library(viridis)
+library(cowplot)
 ```
@@
 ```
+
+# Per site analysis
+
+GR50, D2F, FL50, and CGDD_13C
+
+```{r}
+gdd_all <- tmp2 %>%
+  group_by(SITE, PLOT_GL, PLANT_ID, FL_end, GR_start) %>%
+  summarise(CGDD_0C = sum(GDD_0C_SM),
+            CGDD_8C = sum(GDD_8C_SM),
+            CGDD_13C = sum(GDD_13C_SM),
+            CDYLN = sum(DYLN)
+            ) %>%
+  mutate(D2F = FL_end - GR_start) %>%
+  left_join(metadata) %>%
+  dplyr::select(-(LIBRARY:PLANTED_2018)) %>%
+  left_join(phe_all)
+```
+
+```{r}
+i=1; j = 7; k=3
+(i-1)*length(fourway_site_list)*length(subpop_list) + (k-1)*length(subpop_list) + j
+```
+15 phe (i), 7 sites (k), 3 subpops (j)
+
+Would like to compare subpops at each site and for each phe.
+so same phe number is closest together
+
+i*j*k + j*k + j
+
+```{r}
+PHE <- list(quote(CGDD_13C), quote(CGDD_8C), quote(CGDD_0C), quote(D2F), quote(FL50), quote(GR50), quote(CDYLN), quote(BIOMASS), quote(TC_EOS), 
+            quote(HT_PAN_EOS), quote(FL1), quote(EM1), quote(EM50),
+             quote(GR1))  # quote(TC_EOS_2018),
+fourway_site_list <- list(quote(BRKG), quote(CLMB), quote(KBSM), quote(KING), quote(LINC), quote(PKLE), quote(STIL))
+subpop_list <- list(Midwest_and_Gulf = c("Midwest", "Gulf"), Gulf = c("Gulf"),
+                    Midwest = c("Midwest")) #, Full = c("Atlantic", "Gulf",
+                                             #        "Midwest"))
+
+rrkinblupout <- list()
+kinseqblup <- list()
+bluptable <- list()
+
+for(i in seq_along(PHE)){
+  for(k in seq_along(fourway_site_list)){
+    for(j in seq_along(subpop_list)){
+
+      list_counter <- (i-1)*length(fourway_site_list)*length(subpop_list) + (k-1)*length(subpop_list) + j
+  testdf <- gdd_all %>%
+    left_join(metadata, by = c("PLANT_ID", "SUBPOP")) %>%
+    filter(!is.na(eval(PHE[[i]])) & 
+             GENETIC_SUBPOP_KINSHIP %in% subpop_list[[j]] & 
+             SITE %in% deparse(fourway_site_list[[k]])) %>%
+    as.data.frame()
+
+  bluprrk <- kin.blup(data = testdf, geno = "PLANT_ID", 
+                      pheno = deparse(PHE[[i]]), K = V_g) #, fixed = c("SITE")
+
+  rrkinblupout[[list_counter]] <- enframe(bluprrk$g, name = "PLANT_ID", 
+                               value = deparse(PHE[[i]]))
+
+    # Which Taxa were phenotyped for this phenotype? Just keep those for GWAS.
+  taxa_phenotyped <- testdf %>%
+    dplyr::select(PLANT_ID, !! PHE[[i]]) %>%
+    as_tibble() %>%
+    group_by(PLANT_ID) %>%
+    summarise(count = n())
+  kinseqblup[[list_counter]] <- taxa_phenotyped %>%
+    left_join(rrkinblupout[[list_counter]], by = "PLANT_ID")
+
+  bluptable[[list_counter]] <- tibble(Phenotype = deparse(PHE[[i]]), 
+                                      V_a = bluprrk$Vg, V_e = bluprrk$Ve,
+                                      SITE = deparse(fourway_site_list[[k]]),
+                                      Subpop = names(subpop_list)[j]) %>%
+    mutate(h2 = V_a / (V_a + V_e))
+
+  #saveRDS(bluprrk, paste0("~/Github/pvdiv-phenology-gxe/data-raw/ignore/", 
+  #                        "Kinship_BLUPs_", fourway_site_list[[k]], "_",
+  #                        PHE[[i]], "_", nrow(taxa_phenotyped), "g_",
+  #                        names(subpop_list)[j], ".rds"))
+    }
+  }
+}
+
+```
+
+```{r}
+BLUPs <- select(metadata, PLANT_ID, SUBPOP) %>%
+  filter(SUBPOP %in% c("Midwest", "Gulf")) %>%
+  left_join(dplyr::select(kinseqblup[[1]], PLANT_ID, 3))
+
+for(i in seq_along(kinseqblup)[-1]){
+  BLUPs <- BLUPs %>%
+    left_join(dplyr::select(kinseqblup[[i]], PLANT_ID, 3))
+}
+
+BLUPtable <- bluptable[[1]]
+
+for(i in seq_along(bluptable)[-1]){
+  if(!is.null(bluptable[[i]])){
+  BLUPtable <- BLUPtable %>%
+    full_join(bluptable[[i]])
+  }
+}
+
+BLUPtable
+gdd_all %>% left_join(metadata) %>%
+  pivot_longer(values_to = "MEAS", cols = c(CGDD_0C:D2F, TC_EOS:GR50), names_to = "PHE") %>%
+  dplyr::select(-LIBRARY, -(GENETIC_SUBPOP_KINSHIP:PLANTED_2018)) %>%
+  filter(SUBPOP %in% c("Midwest", "Gulf")) %>%
+  group_by(PHE, SITE, SUBPOP) %>%
+  summarise(X_bar = mean(MEAS, na.rm = TRUE),
+            V_P = var(MEAS, na.rm = TRUE),
+            I = V_P / X_bar^2, 
+            CV_P = 100*sqrt(V_P) / X_bar) %>%
+  #arrange(desc(I)) %>%
+  full_join(BLUPtable, by = c("PHE" = "Phenotype", "SITE", "SUBPOP" = "Subpop")) %>%
+  mutate(I_A = V_a / X_bar^2,
+         CV_R = 100 * sqrt(V_P - V_a) / X_bar, 
+         CV_A = 100 * sqrt(V_a) / X_bar,
+         CV_E = 100 * sqrt(V_e) / X_bar) %>%
+  filter(SITE %in% fourway_sites) %>%
+  arrange(SUBPOP, PHE, SITE) %>%
+  write_csv(., "~/Github/pvdiv-phenology-gxe/data-raw/ignore/Phenotypic_CVs_Individual_sites.csv")
+
+gdd_all %>% left_join(metadata) %>%
+  pivot_longer(values_to = "MEAS", cols = c(CGDD_0C:D2F, TC_EOS:GR50), names_to = "PHE") %>%
+  dplyr::select(-LIBRARY, -(GENETIC_SUBPOP_KINSHIP:PLANTED_2018)) %>%
+  filter(SUBPOP %in% c("Midwest", "Gulf")) %>%
+  group_by(PHE, SITE) %>%
+  summarise(X_bar = mean(MEAS, na.rm = TRUE),
+            V_P = var(MEAS, na.rm = TRUE),
+            I = V_P / X_bar^2, 
+            CV_P = 100*sqrt(V_P) / X_bar) %>%
+  #arrange(desc(I)) %>%
+  full_join(BLUPtable, by = c("PHE" = "Phenotype", "SITE")) %>%
+  mutate(I_A = V_a / X_bar^2,
+         CV_R = 100 * sqrt(V_P - V_a) / X_bar, 
+         CV_A = 100 * sqrt(V_a) / X_bar,
+         CV_E = 100 * sqrt(V_e) / X_bar) %>%
+  filter(SITE %in% fourway_sites) %>%
+  arrange(Subpop, PHE, SITE) %>%
+  write_csv(., "~/Github/pvdiv-phenology-gxe/data-raw/ignore/Phenotypic_CVs_Individual_sites_2.csv")
+```
+
+```{r}
+h2table <- read_csv("~/Github/pvdiv-phenology-gxe/data-raw/ignore/Phenotypic_CVs_Individual_sites_Gulf_Midwest.csv")
+
+h2table %>%
+  mutate(I_A = V_a / X_bar^2,
+         CV_R = 100 * sqrt(V_P - V_a) / X_bar, 
+         CV_A = 100 * sqrt(V_a) / X_bar,
+         CV_E = 100 * sqrt(V_e) / X_bar) %>%
+  filter(PHE == "D2F") %>%
+  ggplot(aes(x = SUBPOP, y = h2)) + 
+  geom_bar(stat = "identity") + facet_wrap(~SITE) + ylim(c(0,1)) + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "../analysis/Heritability_D2F_by_Site_and_Subpop.png", 
+          plot = last_plot())
+
+h2table %>%
+  mutate(I_A = V_a / X_bar^2,
+         CV_R = 100 * sqrt(V_P - V_a) / X_bar, 
+         CV_A = 100 * sqrt(V_a) / X_bar,
+         CV_E = 100 * sqrt(V_e) / X_bar) %>%
+  filter(PHE == "FL50") %>%
+  ggplot(aes(x = SUBPOP, y = h2)) + 
+  geom_bar(stat = "identity") + facet_wrap(~SITE) + ylim(c(0,1)) + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "../analysis/Heritability_FL50_by_Site_and_Subpop.png", 
+          plot = last_plot())
+
+h2table %>%
+  mutate(I_A = V_a / X_bar^2,
+         CV_R = 100 * sqrt(V_P - V_a) / X_bar, 
+         CV_A = 100 * sqrt(V_a) / X_bar,
+         CV_E = 100 * sqrt(V_e) / X_bar) %>%
+  filter(PHE == "GR50") %>%
+  ggplot(aes(x = SUBPOP, y = h2)) + 
+  geom_bar(stat = "identity") + facet_wrap(~SITE) + ylim(c(0,1)) + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+save_plot(filename = "../analysis/Heritability_GR50_by_Site_and_Subpop.png", 
+          plot = last_plot())
+
+h2table %>%
+  mutate(I_A = V_a / X_bar^2,
+         CV_R = 100 * sqrt(V_P - V_a) / X_bar, 
+         CV_A = 100 * sqrt(V_a) / X_bar,
+         CV_E = 100 * sqrt(V_e) / X_bar) %>%
+  filter(PHE %in% c("CGDD_0C", "CGDD_8C", "CGDD_13C", "D2F", "FL50", "CDYLN", "FL1", "GR50")) %>%
+  ggplot(aes(x = PHE, y = h2)) + 
+  geom_point(aes(color = SUBPOP)) + facet_wrap(~SITE) + ylim(c(0,1)) + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1)
+save_plot(filename = "../analysis/Heritability_Various_flowering_time_measures_by_Site_and_Subpop.png", 
+          plot = last_plot())
+```

### 2020-05-05 R/rrBLUP_analysis_D2F.Rmd
Abstract and Intro work

@@
                               TRUE ~ PLANT_ID)) %>%
-  dplyr::select(SITE, PLOT_GL, PLANT_ID, TC_EOS_2018, GR1:GR50)
+  dplyr::select(SITE, PLOT_GL, PLANT_ID, TC_EOS_2018, GR1:GR50, SRV)
 
@@
   dplyr::select(-site) %>%
-  filter(PHE %in% c("GR50","FL1", "FL50", "BIOMASS", "TC_EOS", "HT_PAN_EOS", "EM1", "EM50")) %>%
+  filter(PHE %in% c("GR50","FL1", "FL50", "BIOMASS", "TC_EOS", "HT_PAN_EOS", "EM1", "EM50", "SRV")) %>%
    pivot_wider(names_from = "PHE", values_from = "MEAS") %>%
@@
   left_join(phe_pre2, by = c("SITE", "PLOT_GL", "PLANT_ID"))
+
+phe_all %>%
+  dplyr::select(SITE:PLANT_ID, BIOMASS, SRV) %>%
+  write_rds(., path = "~/Github/pvdiv-fitness-2019/data/All_phenotypes_fitness.rds")
 ```

### 2020-05-21 R/D2F_exploration_fourway_and_gwas.Rmd
Writeup of fourway flowering QTL and overall phenotypic distributions

---
title: "D2F_exploration_fourway_and_GWAS"
author: "Alice MacQueen"
date: "5/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(lubridate)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

# Load data

```{r}
gwas <- readRDS("~/Github/pvdiv-phenology-gxe/data/All_phenotypes_flowering.rds")
metadata <- readRDS("~/Github/pvdiv-phenotypes/data/metadata.rds")
metadata <- metadata %>%
  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
                            !is.na(PLOIDY) ~ as.character(PLOIDY),
                            TRUE ~ NA_character_))
fourway <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/0-data/Phenology_2019_Multiple_Environment.csv")
fwy_prev_raw <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/0-data/Phenology_2016_2018.csv")
```

# Set up phenotype data in long format for analysis and plots
```{r}
fwy_prev <- fwy_prev_raw %>%
  separate(ENVI, into = c("SITE", "YEAR"), sep = "_") %>%
  mutate(D2F = FL50 - GR50,
         YEAR = as.numeric(YEAR),
         WHERE = "FWCR") %>%
  pivot_longer(cols = BIOMASS:D2F, names_to = "PHE", values_to = "MEAS") %>%
  rename(PLANT_ID = LINE)
fwy_2019 <- fourway %>%
  separate(ENVI, into = c("SITE", "YEAR"), sep = "_") %>%
  mutate(D2F = FL50 - GR50,
         YEAR = as.numeric(YEAR),
         WHERE = "FWCR") %>%
  pivot_longer(cols = FL50:D2F, names_to = "PHE", values_to = "MEAS") %>%
  rename(PLANT_ID = LINE)
gwas_2019 <- gwas %>%
  left_join(metadata, by = "PLANT_ID") %>%
  dplyr::select(PLANT_ID, SITE, PLOT_GL, GENETIC_SUBPOP_KINSHIP, SUBPOP,
                LATITUDE, LONGITUDE, TC_EOS:SRV) %>%
  pivot_longer(cols = TC_EOS:SRV, names_to = "PHE", values_to = "MEAS") %>%
  mutate(YEAR = 2019,
         WHERE = "GWAS") %>%
  filter(!is.na(MEAS))

phe_all <- gwas_2019 %>%
  full_join(fwy_2019, by = c("SITE", "YEAR", "PLANT_ID", "PHE", "MEAS",
                             "WHERE")) %>%
  full_join(fwy_prev, by = c("SITE", "YEAR", "PLANT_ID", "PHE", "MEAS",
                             "WHERE")) %>%
  filter(!is.na(MEAS))

phe_all %>%
  group_by(WHERE, PLANT_ID) %>%
  tally()

```
FWCR	AP13	2921		parent - gulf
FWCR	AxD	2124		F1
FWCR	DAC	493		  parent - midwest
FWCR	VS16	1135		parent - midwest
FWCR	VxW	2242		F1
FWCR	WBC	1945		parent - gulf
FWCR	NA	4		
GWAS	AP13	4410	in FWCR also
What J #'s (if any) are DAC, VS16, WBC in the GWAS panel?
DAC J440.A, J440.B Source_ID_2 Dacotah
J230.A Source_ID_1  WBC
J292.A SOURCE_ID_1 VS16 (counted as 8X in metadata, may be an issue)

J045.A SOURCE_ID_2 NSL 29896, Summer
```{r}
phe_parents <- phe_all %>%
  filter(PLANT_ID %in% c("DAC", "VS16", "J440.A", "J036.A", "J037.A")) 
  filter(PLANT_ID %in% c("AP13", "WBC", "J230.A")) 
  filter(PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "J440.A", "J440.B", "J230.A", "J292.A", "J045.A", "J036.A", "J037.A", "J022.B", "J210.A")) 
phe_F1 <- phe_all %>%
  filter(PLANT_ID %in% c("AxD", "VxW")) 
phe_F2 <- phe_all %>%
  filter(!(WHERE == "FWCR" & PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "AxD", "VxW"))) 
```
Have J230.A, J440.A (2 sites: CLMB & PKLE)
For 440.A: J036.A at CLMB, KBSM, STIL; J037.A at 5 sites.
For 230.A: J022.B at 8 sites. J210.A at 4 sites. J660.A at two sites. Ok, ditch 660.A.

```{r}
metadata %>% 
  filter(KINSHIP_ORDER %in% 560:574) %>% # 278:290
  arrange(KINSHIP_ORDER)
  #filter(PLANT_ID %in% c("J230.A", "J440.A")) # 567, 283
```
Closest relatives to J440.A are J036.A and J037.A. One of these seems to have its location metadata off, or something odd has happened. 
Closest relatives to J230.A are J022.B, J210.A, J660.A. Those are all in Texas.

Looked up in PVDIV_Master Metadata File_5-20-20.xlsx which is in Dropbox/Panicum Populations/Panicum Virgatum/

Are any of these J #'s phenotyped and genotyped in the diversity panel? 
Compare phenotypic values for these guys for the FWCR and GWAS panel for 2019.

# Set up 2019 weather data for plots
```{r}
ptt_df <- readRDS(file = "~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")

wea_df <- ptt_df %>%
  dplyr::select(DOY:GDD_13C_SM, PTT_13C)

GR_start <- phe_all %>% 
  ungroup() %>%
  group_by(SITE, YEAR) %>%
  filter(PHE == "GR50") %>%
  summarise(GR_start = floor(min(MEAS)))
FL_end <- phe_all %>% 
  ungroup() %>%
  group_by(SITE, YEAR) %>%
  filter(PHE == "FL50") %>%
  summarise(FL_end = ceiling(max(MEAS)))

wea_gr_fl <- wea_df %>%
  left_join(GR_start, by = "SITE", "YEAR") %>%
  left_join(FL_end, by = c("SITE", "YEAR")) %>%
  group_by(Day_of_year, SITE, YEAR) %>%
  filter(between(Day_of_year, GR_start, FL_end)) %>% 
  ungroup() # Just the weather data between the minimum GR50 and max FL50.
# wea_df has all the weather, wea_gr_fl has the weather for the subset between
# greenup and flowering.
```

# Remake daylength plot

```{r}
wea_gr_fl$SITE <- factor(wea_gr_fl$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
phe_all$SITE <- factor(phe_all$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
phe_parents$SITE <- factor(phe_parents$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
phe_F2$SITE <- factor(phe_F2$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))

wea_gr_fl %>%
  dplyr::select(DOY, DYLN, SITE, RAIN) %>%
  unique() %>%
  filter(SITE != "OVTN") %>%
  ggplot(aes(x = DOY, y = DYLN)) + 
  geom_point() +
  facet_wrap(~SITE) + 
  geom_jitter(data = filter(phe_all, SITE != "OVTN" & PHE == "GR50" & YEAR == 2019 & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = SUBPOP, y = 9), width = 0, height = 1, alpha = 0.1) +
  geom_jitter(data = filter(phe_all, SITE != "OVTN" & PHE == "GR50" & YEAR == 2019 & WHERE == "FWCR"), aes(x = as_date(MEAS, origin = "2018-12-31"), color = as_factor(YEAR), y = 6.5), width = 0, height = 1, alpha = 0.1) + 
  geom_jitter(data = filter(phe_all, SITE != "OVTN" & PHE == "FL50" & YEAR %in% c(2018,2019) & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = SUBPOP, y = 4), width = 0, height = 1, alpha = 0.1) +
  geom_jitter(data = filter(phe_all, SITE != "OVTN" & PHE == "FL50" & (YEAR %in% 2019 | YEAR == 2018 & SITE == "TMPL") & WHERE %in% c("FWCR")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = as_factor(YEAR), y = 1.5), width = 0, height = 1, alpha = 0.1) +
  ylim(c(0,16)) + 
  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1) + xlab("Day of year") + ylab("Daylength (hrs)") +
  geom_vline(xintercept = as_date(365/2-8, origin = "2018-12-31"), linetype = 2) +
  geom_text(aes(y = 1.5, x = as_date(52, origin = "2018-12-31"), label = "FWCR FL50")) +
  geom_text(aes(y = 4, x = as_date(52, origin = "2018-12-31"), label = "GWAS FL50")) +
  geom_text(aes(y = 6.5, x = as_date(242, origin = "2018-12-31"), label = "FWCR GR50")) +
  geom_text(aes(y = 9, x = as_date(242, origin = "2018-12-31"), label = "GWAS GR50")) + 
  geom_point(data = filter(wea_gr_fl, RAIN > 0 & SITE != "OVTN"), aes(y = RAIN/10), color = "blue", pch = 20)

save_plot(filename = "../analysis/environmental-index/Daylength_vs_GR50_FL50_GWAS_and_FWCR_longest_day_line_with_rainfall.png", plot = last_plot(), base_height = 7)
```

Would also like to plot rainfall here somehow to see if that's a factor - that some sites wait through drought until there is a big rainfall to flower, which is what Tom thinks.
# F2 plots
```{r}

wea_gr_fl %>%
  dplyr::select(DOY, DYLN, SITE, RAIN) %>%
  unique() %>%
  filter(SITE != "OVTN") %>%
  ggplot(aes(x = DOY, y = DYLN)) + 
  geom_point() +
  facet_wrap(~SITE) + 
  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "GR50" & YEAR == 2019 & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = SUBPOP, y = 9), width = 0, height = 1, alpha = 0.1) +
  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "GR50" & YEAR == 2019 & WHERE == "FWCR"), aes(x = as_date(MEAS, origin = "2018-12-31"), color = as_factor(YEAR), y = 6.5), width = 0, height = 1, alpha = 0.1) + 
  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "FL50" & YEAR %in% c(2018,2019) & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = SUBPOP, y = 4), width = 0, height = 1, alpha = 0.1) +
  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "FL50" & (YEAR %in% 2019 | YEAR == 2018 & SITE == "TMPL") & WHERE %in% c("FWCR")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = as_factor(YEAR), y = 1.5), width = 0, height = 1, alpha = 0.1) +
  ylim(c(0,16)) +
  theme(legend.position = "right") +
  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1) + xlab("Day of year") + ylab("Daylength (hrs)") +
  geom_vline(xintercept = as_date(365/2-8, origin = "2018-12-31"), linetype = 2) +
  geom_text(aes(y = 1.5, x = as_date(52, origin = "2018-12-31"), label = "FWCR FL50")) +
  geom_text(aes(y = 4, x = as_date(52, origin = "2018-12-31"), label = "GWAS FL50")) +
  geom_text(aes(y = 6.5, x = as_date(242, origin = "2018-12-31"), label = "FWCR GR50")) +
  geom_text(aes(y = 9, x = as_date(242, origin = "2018-12-31"), label = "GWAS GR50")) + 
  geom_point(data = filter(wea_gr_fl, RAIN > 0 & SITE != "OVTN"), aes(y = RAIN/10), color = "blue", pch = 20)

save_plot(filename = "../analysis/environmental-index/Daylength_vs_GR50_FL50_GWAS_and_FWCR_F2_only_longest_day_line_with_rainfall.png", plot = last_plot(), base_height = 7)
```

Maybe at some sites they flower when they get > 2.5 cm of rain. Hard to say. It'd be nice to have the 2020 data to check this for the GWAS, if we really want to say something about this.

# Parent to GWAS Panel Plots
```{r}

wea_gr_fl %>%
  dplyr::select(DOY, DYLN, SITE, RAIN) %>%
  unique() %>%
  filter(SITE != "OVTN") %>%
  ggplot(aes(x = DOY, y = DYLN)) + 
  geom_point() +
  facet_wrap(~SITE) + 
  geom_jitter(data = filter(phe_all, SITE != "OVTN" & PHE == "GR50" & YEAR == 2019 & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = as_factor(YEAR), y = 9, shape = SUBPOP), width = 0, height = 1, alpha = 0.25) +
  geom_jitter(data = filter(phe_parents, SITE != "OVTN" & PHE == "GR50" & YEAR == 2019 & WHERE == "FWCR"), aes(x = as_date(MEAS, origin = "2018-12-31"), color = as_factor(YEAR), y = 6.5, shape = PLANT_ID == "DAC"), width = 0, height = 1, alpha = 0.55) + 
  geom_jitter(data = filter(phe_all, SITE != "OVTN" & PHE == "FL50" & YEAR %in% c(2016:2019) & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = as_factor(YEAR), y = 4, shape = SUBPOP), width = 0, height = 1, alpha = 0.25) +
  geom_jitter(data = filter(phe_parents, SITE != "OVTN" & PHE == "FL50" & (YEAR %in% c(2016:2019) | YEAR == 2018 & SITE == "TMPL") & WHERE %in% c("FWCR")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = as_factor(YEAR), y = 1.5, shape = PLANT_ID == "DAC"), width = 0, height = 1, alpha = 0.55) +
  ylim(c(0,16)) + 
  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1) + xlab("Day of year") + ylab("Daylength (hrs)") +
  geom_vline(xintercept = as_date(365/2-8, origin = "2018-12-31"), linetype = 2) +
  geom_text(aes(y = 1.5, x = as_date(52, origin = "2018-12-31"), label = "FWCR FL50")) +
  geom_text(aes(y = 4, x = as_date(52, origin = "2018-12-31"), label = "GWAS FL50")) +
  geom_text(aes(y = 6.5, x = as_date(242, origin = "2018-12-31"), label = "FWCR GR50")) +
  geom_text(aes(y = 9, x = as_date(242, origin = "2018-12-31"), label = "GWAS GR50")) + 
  geom_point(data = filter(wea_gr_fl, RAIN > 0 & SITE != "OVTN"), aes(y = RAIN/10), color = "blue", pch = 20)
```

# F1 plots
```{r}

wea_gr_fl %>%
  dplyr::select(DOY, DYLN, SITE, RAIN) %>%
  unique() %>%
  filter(SITE != "OVTN") %>%
  ggplot(aes(x = DOY, y = DYLN)) + 
  geom_point() +
  facet_wrap(~SITE) + 
  geom_jitter(data = filter(phe_all, SITE != "OVTN" & PHE == "GR50" & YEAR == 2019 & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = SUBPOP, y = 9), width = 0, height = 1, alpha = 0.45) +
  geom_jitter(data = filter(phe_F1, SITE != "OVTN" & PHE == "GR50" & YEAR %in% c(2019) & WHERE == "FWCR"), aes(x = as_date(MEAS, origin = "2018-12-31"), color = as_factor(YEAR), y = 6.5, shape = PLANT_ID == "VxW"), width = 0, height = 1, alpha = 0.45) + 
  geom_jitter(data = filter(phe_all, SITE != "OVTN" & PHE == "FL50" & YEAR %in% c(2019) & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = SUBPOP, y = 4), width = 0, height = 1, alpha = 0.45) +
  geom_jitter(data = filter(phe_F1, SITE != "OVTN" & PHE == "FL50" & (YEAR %in% c(2019) | YEAR == 2018 & SITE == "TMPL") & WHERE %in% c("FWCR")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = as_factor(YEAR), y = 1.5, shape = PLANT_ID == "VxW"), width = 0, height = 1, alpha = 0.45) +
  ylim(c(0,16)) + 
  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1) + xlab("Day of year") + ylab("Daylength (hrs)") +
  geom_vline(xintercept = as_date(365/2-8, origin = "2018-12-31"), linetype = 2) +
  geom_text(aes(y = 1.5, x = as_date(52, origin = "2018-12-31"), label = "FWCR FL50")) +
  geom_text(aes(y = 4, x = as_date(52, origin = "2018-12-31"), label = "GWAS FL50")) +
  geom_text(aes(y = 6.5, x = as_date(242, origin = "2018-12-31"), label = "FWCR GR50")) +
  geom_text(aes(y = 9, x = as_date(242, origin = "2018-12-31"), label = "GWAS GR50")) + 
  geom_point(data = filter(wea_gr_fl, RAIN > 0 & SITE != "OVTN"), aes(y = RAIN/10), color = "blue", pch = 20)

```

# 1.	First, introduce the diversity panel at the common gardens, and introduce what we know or hypothesize about environmental indices triggering flowering in the different subpopulations – maybe even discuss Gulf, Midwest, Atlantic, and 8X subpopulations. Point out that the Gulf and Midwest subpopulations have the most distinctive flowering responses out of all of the populations.

```{r}
phe_all %>%
  filter(WHERE == "GWAS") %>%
  group_by(PLANT_ID, SUBPOP) %>%
  tally()
```

## We divided these switchgrass genotypes into five categories: tetraploid individuals in the Atlantic, Midwest, and Gulf genetic subpopulations (Lovell et al 202X), admixed tetraploid individuals, and octoploid individuals. We then explored the overall patterns of greenup and flowering within these five categories. We first explored these patterns as a function of cumulative GDD (cite some switchgrass flowering time study, and Li’s study?).

# Setup for CGDD plots
```{r}
phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Phenotypes_D2F_fourway_project.rds")
sites <- readRDS("~/Github/pvdiv-phenotypes/data/sites.rds")

pid_all_sites <- phe %>%
  dplyr::select(PLANT_ID, ends_with("D2F")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  group_by(PLANT_ID) %>%
  filter(!is.na(D2F) & !(SITE %in% c("FRMI_D2F", "BRKG_D2F", "TMPL_D2F"))) %>%
  count() %>% 
  arrange(desc(n)) %>%
  filter(n >= 4)
GR_start <- phe %>% 
  dplyr::select(-ends_with("D2F"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("GR50"), names_to = "SITE", values_to = "GR50") %>% 
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  group_by(SITE, PLANT_ID) %>%
  summarise(GR_start = floor(min(GR50)))
FL_end <- phe %>% 
  dplyr::select(-ends_with("D2F"), -ends_with("GR50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("FL50"), names_to = "SITE", values_to = "FL50") %>%
  #filter(D2F > 10 & SUBPOP == "Midwest") %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  group_by(SITE, PLANT_ID) %>%
  summarise(FL_end = ceiling(max(FL50)))
tmp3 <- wea_df %>%
  left_join(GR_start, by = "SITE") %>%
  left_join(FL_end, by = c("SITE", "PLANT_ID")) %>%
  group_by(SITE, Day_of_year, PLANT_ID) %>%
  filter(between(Day_of_year, GR_start, FL_end)) %>% 
  ungroup()
```
CGDD plots
```{r}
tmp3 %>%
  group_by(SITE, PLANT_ID, GR_start, FL_end) %>%
  summarise(CGDD_13C = sum(GDD_13C_SM)
            ) %>%
  mutate(D2F = FL_end - GR_start) %>%
  filter(D2F > 10) %>%
  left_join(metadata) %>%
  left_join(sites) %>%
  filter(!(SITE %in% c("FRMI"))) %>%
  filter(PLANT_ID %in% pid_all_sites$PLANT_ID & SUBPOP %in% c("Midwest", "Gulf")) %>%
  ggplot(aes(x = Latitude, y = CGDD_13C)) + 
  geom_line(aes(group = PLANT_ID, color = SUBPOP)) + 
  #theme(legend.position = "none") +
  geom_text(aes(x = Latitude, label = SITE), y = 100, angle = 45) +
  geom_smooth(aes(group = SUBPOP), method = "lm")  +
  #geom_smooth(aes(group = SUBPOP), color = "green") +
  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1)  + 
  ylim(c(0,3200))

tmp3 %>%
  group_by(SITE, PLANT_ID, GR_start, FL_end) %>%
  summarise(CGDD_13C = sum(GDD_13C_SM)
            ) %>%
  mutate(D2F = FL_end - GR_start) %>%
  filter(D2F > 10) %>%
  left_join(metadata) %>%
  left_join(sites) %>%
  filter(!(SITE %in% c("FRMI"))) %>%
  filter(PLANT_ID %in% pid_all_sites$PLANT_ID) %>%
  ggplot(aes(x = Latitude, y = CGDD_13C)) + 
  geom_line(aes(group = PLANT_ID, color = PLANT_ID)) + 
  facet_wrap(~SUBPOP) +
  theme(legend.position = "none") +
  geom_text(aes(x = Latitude, label = SITE), y = 100, angle = 45) +
  geom_smooth(aes(group = SUBPOP), method = "lm", color = "darkorange")  +
  #geom_smooth(aes(group = SUBPOP), color = "green") +
  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1)  + 
  ylim(c(0,3200)) + ylab(label = "Cumulative GDD (13C)")
  save_plot(filename = "../manuscript/plots/Cumulative_GDD_13C_by_site_and_subpop.png", plot = last_plot(), base_height = 8)

tmp3 %>%
  group_by(SITE, PLANT_ID, GR_start, FL_end) %>%
  summarise(CGDD_13C = sum(GDD_13C_SM)
            ) %>%
  mutate(D2F = FL_end - GR_start) %>%
  filter(D2F > 10) %>%
  left_join(metadata) %>%
  left_join(sites) %>%
  filter(!(SITE %in% c("FRMI"))) %>%
  filter(PLANT_ID %in% pid_all_sites$PLANT_ID) %>%
  ggplot(aes(x = Latitude, y = CGDD_13C)) + 
  #geom_line(aes(group = PLANT_ID, color = PLANT_ID)) + 
  #facet_wrap(~SUBPOP) +
  #theme(legend.position = "none") +
  geom_text(aes(x = Latitude, label = SITE), y = 100, angle = 45) +
  geom_smooth(aes(group = SUBPOP, color = SUBPOP), method = "lm")  +
  #geom_smooth(aes(group = SUBPOP), color = "green") +
  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1)  + 
  ylim(c(0,3200)) + ylab(label = "Cumulative GDD (13C)")
  save_plot(filename = "../manuscript/plots/Cumulative_GDD_13C_smoothed_subpops.png", plot = last_plot(), base_height = 4.1, base_asp = 1.1)
```

```{r}
lmdata <- tmp3 %>%
  group_by(SITE, PLANT_ID, GR_start, FL_end) %>%
  summarise(CGDD_13C = sum(GDD_13C_SM)
            ) %>%
  mutate(D2F = FL_end - GR_start) %>%
  filter(D2F > 10) %>%
  left_join(metadata) %>%
  left_join(sites) %>%
  filter(!(SITE %in% c("FRMI"))) %>%
  filter(PLANT_ID %in% pid_all_sites$PLANT_ID)
summary(lm(CGDD_13C ~ SUBPOP + Latitude:SUBPOP, data = lmdata))
lmdata$prediction <- predict(lm(CGDD_13C ~ SUBPOP + Latitude:SUBPOP, data = lmdata))

lmdata$resid <- residuals(lm(CGDD_13C ~ SUBPOP + Latitude:SUBPOP, data = lmdata))

ggplot(lmdata) + 
  geom_point(aes(x = prediction, y = resid, color = SUBPOP)) + 
  facet_wrap(~SITE)

lmdata$SITE <- factor(lmdata$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
lmdata %>%
  group_by(SUBPOP, SITE) %>%
  summarise(prediction = mean(prediction),
            resid = mean(abs(resid))) %>%
  ggplot(aes(x = SITE, y = prediction)) + 
  geom_point() +
  geom_errorbar(mapping = aes(x = SITE, ymin = prediction - resid, ymax = prediction + resid)) +
  facet_wrap(~SUBPOP)
lmdata %>%
  group_by(SUBPOP, SITE) %>%
  summarise(prediction = mean(prediction),
            resid = mean(abs(resid))) %>%
 # write_csv("Subpop_CGDD_predictions.csv")
  summarise(prediction = mean(prediction),
            stderr = mean(resid)) # BAD! But sd(prediction) was giving NA's

```

### 2020-05-27 R/D2F_exploration_fourway_and_gwas.Rmd
Manuscript additions looking for switchgrass floewring when exposed to shortening days of a specific length

@@
 ```{r}
-phe_parents <- phe_all %>%
+phe_F0_up <- phe_all %>%
   filter(PLANT_ID %in% c("DAC", "VS16", "J440.A", "J036.A", "J037.A")) 
+phe_F0_low <- phe_all %>%
   filter(PLANT_ID %in% c("AP13", "WBC", "J230.A")) 
+phe_F0 <- phe_all %>%
   filter(PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "J440.A", "J440.B", "J230.A", "J292.A", "J045.A", "J036.A", "J037.A", "J022.B", "J210.A")) 
@@
 phe_all$SITE <- factor(phe_all$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
-phe_parents$SITE <- factor(phe_parents$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
+phe_F0$SITE <- factor(phe_F0$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
 phe_F2$SITE <- factor(phe_F2$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
@@
   scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1) + xlab("Day of year") + ylab("Daylength (hrs)") +
-  geom_vline(xintercept = as_date(365/2-8, origin = "2018-12-31"), linetype = 2) +
+  geom_vline(xintercept = as_date(173, origin = "2018-12-31"), linetype = 2) +
   geom_text(aes(y = 1.5, x = as_date(52, origin = "2018-12-31"), label = "FWCR FL50")) +
@@
 Maybe at some sites they flower when they get > 2.5 cm of rain. Hard to say. It'd be nice to have the 2020 data to check this for the GWAS, if we really want to say something about this.
+
+# Tie to previous research
+## DYLN at FL plots
+While the rate and extent of vegetative development is closely related to GDD, reproductive development is more tied to day-of-the-year, or photoperiod.
+Switchgrass is a short-day plant; it flowers when exposed to shortening days of a specific length (Benedict, 1940).
+[so what is daylength of flowering for each plant_id * site – and is it the same across sites for individual plant_ids?
+Or different?
+If the same, this is the driver; if different, something else is the driver.]
+```{r}
+fl50_dyln <- phe_all %>%
+  filter(PHE == "FL50") %>%
+  left_join(wea_gr_fl, by = c("SITE", "YEAR", "MEAS" = "Day_of_year"))
+fl50_dyln$SITE <- factor(fl50_dyln$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
+
+fl1_dyln <- phe_all %>%
+  filter(PHE == "FL1") %>%
+  left_join(wea_gr_fl, by = c("SITE", "YEAR", "MEAS" = "Day_of_year"))
+fl1_dyln$SITE <- factor(fl1_dyln$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
+
+fl50_dyln %>%
+  group_by(SUBPOP) %>%
+  summarise(min_dyln = min(DYLN),
+            max_dyln = max(DYLN))
+max_dyln <- fl50_dyln %>%
+  group_by(SITE) %>%
+  summarise(min_dyln = min(DYLN, na.rm = TRUE),
+            max_dyln = max(DYLN, na.rm = TRUE))
+max_dyln <- wea_gr_fl %>%
+  group_by(SITE) %>%
+  summarise(min_dyln = min(DYLN, na.rm = TRUE),
+            max_dyln = max(DYLN, na.rm = TRUE))
+
+fl1_dyln %>%
+  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "DAYS_SHORTENING",
+                               MEAS <= 172 ~ "DAYS_LENGTHENING",
+                               TRUE ~ NA_character_)) %>%
+  filter(YEAR == 2019 & WHERE == "GWAS" & SUBPOP %in% c("8X", "Atlantic")) %>% # , "Gulf", "Midwest"
+  group_by(PLANT_ID) %>%
+  add_tally() %>%
+  filter(n == 8 | n > 13) %>%
+  ggplot(aes(x = PLANT_ID, y = DYLN)) + 
+  geom_point(aes(color = SITE, shape = SUBPOP)) + 
+  facet_wrap(~FL_B4_SOL) +
+  theme(axis.text.x = element_text(hjust = 1, angle = 45), legend.position = "right") + 
+  scale_color_viridis(discrete = TRUE, direction = -1) +
+  geom_hline(data = max_dyln, aes(yintercept = max_dyln, color = SITE), linetype = 2) + 
+  geom_label(data = max_dyln, aes(x = 4, y = max_dyln, label = SITE))
+
+fl1_dyln %>%
+  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "DAYS_SHORTENING",
+                               MEAS <= 172 ~ "DAYS_LENGTHENING",
+                               TRUE ~ NA_character_)) %>%
+  filter(YEAR == 2019 & WHERE == "GWAS" & SUBPOP %in% c("Gulf", "Midwest")) %>% # , 
+  group_by(PLANT_ID) %>%
+  add_tally() %>%
+  filter(n == 8 | n > 13) %>%
+  ggplot(aes(x = PLANT_ID, y = DYLN)) + 
+  geom_point(aes(color = SITE, shape = SUBPOP)) + 
+  facet_wrap(~FL_B4_SOL) +
+  theme(axis.text.x = element_text(hjust = 1, angle = 45), legend.position = "right", panel.spacing.x = unit(x = 5, units = "points")) + 
+  scale_color_viridis(discrete = TRUE, direction = -1) +
+  geom_hline(data = max_dyln, aes(yintercept = max_dyln, color = SITE), linetype = 2) + 
+  geom_label(data = max_dyln, aes(x = 4, y = max_dyln, label = SITE))
+
+fl1_dyln %>%
+  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "DAYS_SHORTENING",
+                               MEAS <= 172 ~ "DAYS_LENGTHENING",
+                               TRUE ~ NA_character_)) %>%
+  filter(YEAR == 2019 & WHERE == "GWAS" & SUBPOP %in% c("Gulf", "Midwest")) %>% # , 
+  group_by(PLANT_ID) %>%
+  add_tally() %>%
+  filter(n == 9 | n > 13) %>%
+  ggplot(aes(x = PLANT_ID, y = DYLN)) +
+  geom_violin(color = "grey", scale = "area") +
+  geom_jitter(aes(color = SITE, shape = SUBPOP, fill = FL_B4_SOL), stroke = 1.2, size = 1.5, height = 0, width = 0.1) + 
+  theme(axis.text.x = element_text(hjust = 1, angle = 45), legend.position = "right", panel.spacing.x = unit(x = 5, units = "points")) + 
+  scale_color_viridis(discrete = TRUE, direction = -1) +
+  geom_hline(data = max_dyln, aes(yintercept = max_dyln, color = SITE), linetype = 2) + 
+  #facet_wrap(~SUBPOP) +
+  scale_shape_manual(values = c(21, 24)) +
+  scale_fill_viridis(option = "B", discrete = TRUE, direction = -1)
+  geom_label(data = max_dyln, aes(x = 4, y = max_dyln, label = SITE))
+
+fl1_dyln %>%
+  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "DAYS_SHORTENING",
+                               MEAS <= 172 ~ "DAYS_LENGTHENING",
+                               TRUE ~ NA_character_)) %>%
+  filter(YEAR == 2019 & WHERE == "GWAS" & SUBPOP %in% c("8X", "Atlantic")) %>% # , 
+  group_by(PLANT_ID) %>%
+  add_tally() %>%
+  filter(n == 9 | n > 13) %>%
+  ggplot(aes(x = PLANT_ID, y = DYLN)) +
+  geom_violin(color = "grey", scale = "area") +
+  geom_jitter(aes(color = SITE, shape = SUBPOP, fill = FL_B4_SOL), stroke = 1.2, size = 1.5, height = 0, width = 0.1) + 
+  theme(axis.text.x = element_text(hjust = 1, angle = 45), legend.position = "right", panel.spacing.x = unit(x = 5, units = "points")) + 
+  scale_color_viridis(discrete = TRUE, direction = -1) +
+  geom_hline(data = max_dyln, aes(yintercept = max_dyln, color = SITE), linetype = 2) + 
+  #facet_wrap(~SUBPOP) +
+  scale_shape_manual(values = c(21, 24)) +
+  scale_fill_viridis(option = "B", discrete = TRUE, direction = -1)
+  geom_label(data = max_dyln, aes(x = 4, y = max_dyln, label = SITE))
+
+```
+Maybe also plot the max daylength at each site? 
+Atlantic may be interesting to differentiate by ecotype or latitude of origin.
+
+"The photoperiod response displayed within a given strain naturally differs with its latitude of origin and can be a point of exploitation in forage and biomass production systems."
+```{r}
+
+wea_gr_fl %>%
+  dplyr::select(DOY, DYLN, SITE, RAIN) %>%
+  unique() %>%
+  filter(SITE != "OVTN") %>%
+  ggplot(aes(x = DOY, y = DYLN)) + 
+  geom_point() +
+  facet_wrap(~SITE) + 
+  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "GR50" & YEAR == 2019 & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = LATITUDE, y = 9), width = 0, height = 1, alpha = 0.25) +
+  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "GR50" & YEAR == 2019 & WHERE == "FWCR"), aes(x = as_date(MEAS, origin = "2018-12-31"), y = 6.5), width = 0, height = 1, alpha = 0.1) + 
+  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "FL50" & YEAR %in% c(2018,2019) & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = LATITUDE, y = 4), width = 0, height = 1, alpha = 0.25) +
+  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "FL50" & (YEAR %in% 2019 | YEAR == 2018 & SITE == "TMPL") & WHERE %in% c("FWCR")), aes(x = as_date(MEAS, origin = "2018-12-31"), y = 1.5), width = 0, height = 1, alpha = 0.1) +
+  ylim(c(0,16)) +
+  theme(legend.position = "right") +
+  scale_color_viridis(end = 0.8, direction = -1) + xlab("Day of year") + ylab("Daylength (hrs)") +
+  geom_vline(xintercept = as_date(365/2-8, origin = "2018-12-31"), linetype = 2) +
+  geom_text(aes(y = 1.5, x = as_date(52, origin = "2018-12-31"), label = "FWCR FL50")) +
+  geom_text(aes(y = 4, x = as_date(52, origin = "2018-12-31"), label = "GWAS FL50")) +
+  geom_text(aes(y = 6.5, x = as_date(242, origin = "2018-12-31"), label = "FWCR GR50")) +
+  geom_text(aes(y = 9, x = as_date(242, origin = "2018-12-31"), label = "GWAS GR50")) + 
+  geom_point(data = filter(wea_gr_fl, RAIN > 0 & SITE != "OVTN"), aes(y = RAIN/10), color = "blue", pch = 20)
+
+```
 

### 2020-06-02 R/D2F_exploration_fourway_and_gwas.Rmd
Writeup Results photoperiod cue section

@@
 
+# saveRDS(phe_all, file = "../data/all_phenotype_meas_4wcr_gwas.rds")
 ```
 FWCR	AP13	2921		parent - gulf
-FWCR	AxD	2124		F1
+FWCR	AxD	2124		F1 (AxB in results)
 FWCR	DAC	493		  parent - midwest
 FWCR	VS16	1135		parent - midwest
-FWCR	VxW	2242		F1
+FWCR	VxW	2242		F1 (CxD in results)
 FWCR	WBC	1945		parent - gulf

### 2020-06-02 R/Draft_figures_2020-06-02.Rmd
Writeup Results photoperiod cue section

---
title: "Draft_figures_2020-06-02"
author: "Alice MacQueen"
date: "6/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(lubridate)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

#

```{r}
metadata <- readRDS("~/Github/pvdiv-phenotypes/data/metadata.rds")
metadata <- metadata %>%
  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
                            !is.na(PLOIDY) ~ as.character(PLOIDY),
                            TRUE ~ NA_character_))
phe_all <- read_rds("../data/all_phenotype_meas_4wcr_gwas.rds")
sites <- readRDS("~/Github/pvdiv-phenotypes/data/sites.rds")
```

FWCR	AP13	2921		parent - gulf
FWCR	AxD	2124		F1 (AxB in results)
FWCR	DAC	493		  parent - midwest
FWCR	VS16	1135		parent - midwest
FWCR	VxW	2242		F1 (CxD in results)
FWCR	WBC	1945		parent - gulf
FWCR	NA	4		
GWAS	AP13	4410	in FWCR also
What J #'s (if any) are DAC, VS16, WBC in the GWAS panel?
DAC J440.A, J440.B Source_ID_2 Dacotah
J230.A Source_ID_1  WBC
J292.A SOURCE_ID_1 VS16 (counted as 8X in metadata, may be an issue)

J045.A SOURCE_ID_2 NSL 29896, Summer
```{r}
phe_F0_up <- phe_all %>%
  filter(PLANT_ID %in% c("DAC", "VS16", "J440.A", "J036.A", "J037.A")) 
phe_F0_low <- phe_all %>%
  filter(PLANT_ID %in% c("AP13", "WBC", "J230.A")) 
phe_F0 <- phe_all %>%
  filter(PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "J440.A", "J440.B", "J230.A", "J292.A", "J045.A", "J036.A", "J037.A", "J022.B", "J210.A")) 
phe_F1 <- phe_all %>%
  filter(PLANT_ID %in% c("AxD", "VxW")) 
phe_F2 <- phe_all %>%
  filter(!(WHERE == "FWCR" & PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "AxD", "VxW"))) 
```

# Set up 2019 weather data for plots
```{r}
ptt_df <- readRDS(file = "~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")

wea_df <- ptt_df %>%
  dplyr::select(DOY:GDD_13C_SM, PTT_13C)

GR_start <- phe_all %>% 
  ungroup() %>%
  group_by(SITE, YEAR) %>%
  filter(PHE == "GR50") %>%
  summarise(GR_start = floor(min(MEAS)))
FL_end <- phe_all %>% 
  ungroup() %>%
  group_by(SITE, YEAR) %>%
  filter(PHE == "FL50") %>%
  summarise(FL_end = ceiling(max(MEAS)))

wea_gr_fl <- wea_df %>%
  left_join(GR_start, by = "SITE", "YEAR") %>%
  left_join(FL_end, by = c("SITE", "YEAR")) %>%
  group_by(Day_of_year, SITE, YEAR) %>%
  filter(between(Day_of_year, GR_start, FL_end)) %>% 
  ungroup() # Just the weather data between the minimum GR50 and max FL50.
# wea_df has all the weather, wea_gr_fl has the weather for the subset between
# greenup and flowering.
```

# --
# Run all chunks above as setup

# -- Plots

We first evaluated the evidence that switchgrass photoperiodicity was genotype dependent – specifically, that switchgrass flowers when exposed to shortening days of a specific length. 
```{r}
fl50_dyln <- phe_all %>%
  filter(PHE == "FL50") %>%
  left_join(wea_gr_fl, by = c("SITE", "YEAR", "MEAS" = "Day_of_year"))
fl50_dyln$SITE <- factor(fl50_dyln$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))

fl1_dyln <- phe_all %>%
  filter(PHE == "FL1") %>%
  left_join(wea_gr_fl, by = c("SITE", "YEAR", "MEAS" = "Day_of_year"))
fl1_dyln$SITE <- factor(fl1_dyln$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))

fl50_dyln %>%
  group_by(SUBPOP) %>%
  summarise(min_dyln = min(DYLN),
            max_dyln = max(DYLN))
max_dyln <- wea_gr_fl %>%
  group_by(SITE) %>%
  summarise(min_dyln = min(DYLN, na.rm = TRUE),
            max_dyln = max(DYLN, na.rm = TRUE))
```

Many genotypes flowered while days were lengthening at the three southernmost sites (Table X). 
Specifically, at Kingsville, 107 genotypes (38.9%) had 50% of tillers flowering before the summer solstice, after which, days start shortening. At Pickle, 518 genotypes (62.3%) were flowering while days were lengthening. At Temple, 107 genotype (26.9%) were flowering while days were lengthening, while at Stillwater and Columbia, 5 and 9 genotypes (2.0%, 1.2%) flowered while days were lengthening. These observations indicated that photoperiodicity could not be the only environmental triggering flowering in this species. 
```{r}
fl50_dyln %>%
  filter(WHERE == "GWAS") %>%
  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "DAYS_SHORTENING",
                               MEAS <= 172 ~ "DAYS_LENGTHENING",
                               TRUE ~ NA_character_)) %>%
  group_by(SITE, FL_B4_SOL, PLANT_ID) %>%
  summarise(count = n()) %>%
  group_by(SITE, FL_B4_SOL) %>%
  tally() %>%
  group_by(SITE) %>%
  mutate(frac = n/sum(n))

fl50_dyln %>%
  filter(WHERE == "GWAS") %>%
  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "DAYS_SHORTENING",
                               MEAS <= 172 ~ "DAYS_LENGTHENING",
                               TRUE ~ NA_character_)) %>%
  group_by(SITE, FL_B4_SOL, SUBPOP, PLANT_ID) %>%
  summarise(count = n()) %>%
  group_by(SITE, FL_B4_SOL, SUBPOP) %>%
  tally() %>%
  group_by(SITE, SUBPOP) %>%
  #mutate(frac = n/sum(n)) %>%
  pivot_wider(names_from = SUBPOP, values_from = n) %>%
  left_join(sites) %>%
  write_csv("../manuscript/plots/Plant_flowering_b4_or_after_solstice.csv")

fl50_dyln$SITE <- factor(fl50_dyln$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))

daylength_change <- fl50_dyln %>%
  filter(WHERE == "GWAS" & !(SITE %in% c("OVTN", "FRMI"))) %>%
  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "Shortening",
                               MEAS <= 172 ~ "Lengthening",
                               TRUE ~ NA_character_)) %>%
  group_by(SITE, FL_B4_SOL, SUBPOP, PLANT_ID) %>%
  summarise(count = n()) %>%
  group_by(SITE, FL_B4_SOL, SUBPOP) %>%
  tally() #%>%
  #group_by(SITE, SUBPOP) %>%
  #mutate(frac = n/sum(n)) %>%
  #pivot_wider(names_from = SUBPOP, values_from = n) %>%
  #left_join(sites) %>%
dyln_change <- daylength_change %>%
  ggplot(aes(x = FL_B4_SOL, y = n, width = 0.7)) + 
  geom_bar(aes(fill = SUBPOP), stat = "identity", position = position_dodge()) +
  facet_wrap(~SITE, nrow = 2) +
  scale_fill_viridis(discrete = TRUE) +
  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
        legend.position = "right", panel.spacing.x = unit(x = 0.01, unit = "npc")) +
  ylab("Count of Genotypes") +
  xlab("Does the genotype flower when days are 
       lengthening or shortening?") 
save_plot(filename = "../manuscript/plots/daylength_change_subpops.png", plot = dyln_change)

?unit
?geom_bar
```
At Kingsville, all five subpopulations had individuals that flowered while days were lengthening. Only the admixed 4X and the Gulf subpops had a majority of individuals that flowered while days were shortening (55% and 85.3%).
At Pickle, all five subpopulations had individuals that flowered while days were lengthening. Only the Gulf subpop had a majority of individuals that flowered while days were shortening (94.0%).
At Temple,  all five subpopulations had individuals that flowered while days were lengthening. The admixed 4X, Atlantic, and Gulf had a majority of individuals that flowered while days were shortening (65.2%, 57.9%, and 99.5%).
At Stillwater, 4X, Atlantic, and Midwest individuals flowered while days were lengthening, but these individuals were in the minority (5.3%, 2.3%, 3.7%).
At Columbia, 8X and Midwest individuals flowered while days were lengthening (1.4% and 5%).
Further north than Columbia, virtually all individuals flowered while days were shortening.

```{r}
fl50_dyln %>%
  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "DAYS_SHORTENING",
                               MEAS <= 172 ~ "DAYS_LENGTHENING",
                               TRUE ~ NA_character_)) %>%
  filter(YEAR == 2019 & WHERE == "GWAS" & SUBPOP %in% c("Gulf", "Midwest")) %>% # , 
  group_by(PLANT_ID) %>%
  add_tally() %>%
  filter(n == 9 | n > 13) %>%
  ggplot(aes(x = PLANT_ID, y = DYLN)) +
  geom_violin(color = "grey", scale = "area") +
  geom_jitter(aes(color = SITE, shape = SUBPOP, fill = FL_B4_SOL), stroke = 1.2, size = 1.5, height = 0, width = 0.1) + 
  theme(axis.text.x = element_text(hjust = 1, angle = 45), legend.position = "right", panel.spacing.x = unit(x = 5, units = "points")) + 
  scale_color_viridis(discrete = TRUE, direction = -1) +
  geom_hline(data = max_dyln, aes(yintercept = max_dyln, color = SITE), linetype = 2) + 
  #facet_wrap(~SUBPOP) +
  scale_shape_manual(values = c(21, 24)) +
  scale_fill_viridis(option = "B", discrete = TRUE, direction = -1)
  geom_label(data = max_dyln, aes(x = 4, y = max_dyln, label = SITE))

fl1_dyln %>%
  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "DAYS_SHORTENING",
                               MEAS <= 172 ~ "DAYS_LENGTHENING",
                               TRUE ~ NA_character_)) %>%
  filter(YEAR == 2019 & WHERE == "GWAS" & SUBPOP %in% c("8X", "Atlantic")) %>% # , 
  group_by(PLANT_ID) %>%
  add_tally() %>%
  filter(n == 9 | n > 13) %>%
  ggplot(aes(x = PLANT_ID, y = DYLN)) +
  geom_violin(color = "grey", scale = "area") +
  geom_jitter(aes(color = SITE, shape = SUBPOP, fill = FL_B4_SOL), stroke = 1.2, size = 1.5, height = 0, width = 0.1) + 
  theme(axis.text.x = element_text(hjust = 1, angle = 45), legend.position = "right", panel.spacing.x = unit(x = 5, units = "points")) + 
  scale_color_viridis(discrete = TRUE, direction = -1) +
  geom_hline(data = max_dyln, aes(yintercept = max_dyln, color = SITE), linetype = 2) + 
  #facet_wrap(~SUBPOP) +
  scale_shape_manual(values = c(21, 24)) +
  scale_fill_viridis(option = "B", discrete = TRUE, direction = -1)
  geom_label(data = max_dyln, aes(x = 4, y = max_dyln, label = SITE))

```

### 2020-06-04 R/Draft_figures_2020-06-02.Rmd
New back to back plots of flowering in shortening vs lengthening days

@@
                             !is.na(PLOIDY) ~ as.character(PLOIDY),
-                            TRUE ~ NA_character_))
+                            TRUE ~ NA_character_),
+         LATITUDE = abs(LATITUDE),
+         LONGITUDE = abs(LONGITUDE)) 
 phe_all <- read_rds("../data/all_phenotype_meas_4wcr_gwas.rds")
 sites <- readRDS("~/Github/pvdiv-phenotypes/data/sites.rds")
+phe_all$SUBPOP <- factor(phe_all$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
+sites <- sites %>%
+  mutate(manu_site = case_when(State != "TX" ~ State,
+                               SITE == "KING" ~ "TX1",
+                               SITE == "PKLE" ~ "TX2",
+                               SITE == "TMPL" ~ "TX3",
+                               SITE == "OVTN" ~ "TX4"))
+sites$manu_site <- factor(sites$manu_site, levels = c("TX1", "TX2", "TX3", "TX4", "OK", "MO", "NE", "IL", "MI", "SD"))
+
+subpops <- enframe(unique(metadata$SUBPOP), name = NULL, value = "SUBPOP") %>%
+  mutate(color_code = case_when(SUBPOP == "Atlantic" ~ "#6E91CB",
+                                SUBPOP == "Gulf" ~ "#F47F72",    
+                                SUBPOP == "Midwest" ~ "#442C83",
+                                SUBPOP == "4X" ~ "grey", 
+                                SUBPOP == "8X" ~ "#090906"))
+subpops$SUBPOP <- factor(subpops$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
+
+# gulf: #F47F72, atlantic: #6E91CB, midwest: #442C83
 ```
@@
   #left_join(sites) %>%
+daylength_change$SUBPOP <- factor(daylength_change$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
+color_code <- c("grey", "#6E91CB","#F47F72", "#442C83", "#090906")
 dyln_change <- daylength_change %>%
+  left_join(sites) %>%
+  left_join(subpops) %>%
+  dplyr::select(SITE, FL_B4_SOL, SUBPOP, n) %>%
+  ungroup() %>%
+  #add_row(SITE = "BRKG", FL_B4_SOL = "Shortening") %>%
+  #add_row(SITE = "LINC", FL_B4_SOL = "Shortening") %>%
+  complete(SITE, FL_B4_SOL, SUBPOP, fill = list(n = 0))
+delta_dyln %>%
   ggplot(aes(x = FL_B4_SOL, y = n, width = 0.7)) + 
   geom_bar(aes(fill = SUBPOP), stat = "identity", position = position_dodge()) +
-  facet_wrap(~SITE, nrow = 2) +
-  scale_fill_viridis(discrete = TRUE) +
+  facet_wrap(~manu_site, nrow = 2) +
+  scale_fill_manual(values = color_code) +
   theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
@@
        lengthening or shortening?") 
-save_plot(filename = "../manuscript/plots/daylength_change_subpops.png", plot = dyln_change)
+save_plot(filename = "../manuscript/plots/daylength_change_subpops.png", plot = delta_dyln)
 
-?unit
-?geom_bar
 ```
+```{r}
+dyln_change$SUBPOP <- factor(dyln_change$SUBPOP, levels = rev(c("4X", "Atlantic", "Gulf", "Midwest", "8X")))
+color_code <- rev(c("grey", "#6E91CB","#F47F72", "#442C83", "#090906"))
+
+site_subpop_list <- list()
+eightsites <- c("KING", "PKLE", "TMPL", "STIL", "CLMB", "LINC", "KBSM", "BRKG")
+
+for(i in seq_along(eightsites)){
+p1 <- dyln_change %>%
+  filter((SITE %in% eightsites[i]) & FL_B4_SOL == "Lengthening") %>%
+  ggplot(aes(x = SUBPOP, y = n, width = 1)) + 
+  geom_col(aes(fill = SUBPOP), position = "dodge") +
+  scale_y_reverse(expand = c(0.002, 0.002), limits = c(240, 0)) + coord_flip() +
+  scale_fill_manual(values = color_code) +
+  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
+        legend.position = "none", panel.spacing.x = unit(x = 0, unit = "mm"),
+        plot.margin = unit(c(5.5, 0, 5.5, 5.5), "pt")) +
+  ylab("Genotypes") +
+  xlab("Subpopulation") + labs(subtitle = "Lengthening", title = sites$manu_site[sites$SITE == eightsites[i]])
+p2 <- dyln_change %>%
+  filter((SITE %in% eightsites[i]) & FL_B4_SOL == "Shortening") %>%
+  ggplot(aes(x = SUBPOP, y = n, width = 1)) + 
+  geom_col(aes(fill = SUBPOP), position = "dodge") +
+  scale_y_continuous(expand = c(0.002, 0.002), limits = c(0,240)) +
+  coord_flip() +
+  scale_fill_manual(values = color_code) +
+  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
+        axis.text.y = element_blank(), axis.title.y = element_blank(),
+        legend.position = "none", panel.spacing.x = unit(x = 0, unit = "mm"),        axis.line.y = element_blank(), axis.ticks.y = element_blank(),
+        plot.margin = unit(c(5.5, 5.5, 5.5, -3.5), "pt")) +
+  geom_hline(yintercept = 0, linetype = 2) + labs(subtitle = "Shortening", title = " ") + ylab(" ")
+
+site_subpop_list[[i]] <- plot_grid(p1, p2, rel_widths = c(1.66,1))
+}
+
+subpop_dyln_plot <- plot_grid(site_subpop_list[[1]], site_subpop_list[[2]], site_subpop_list[[3]],
+          site_subpop_list[[4]], site_subpop_list[[5]], site_subpop_list[[6]],
+          site_subpop_list[[7]], site_subpop_list[[8]], nrow = 2)
+
+save_plot(filename = "../manuscript/plots/daylength_change_subpops_back_to_back.png", plot = subpop_dyln_plot, base_asp = 2.2, base_height = 5)
+```
+
 At Kingsville, all five subpopulations had individuals that flowered while days were lengthening.
@@
 
+Latitude of origin bins for flowering while day lengthening/shortening plot
+
 ```{r}
 fl50_dyln %>%
-  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "DAYS_SHORTENING",
-                               MEAS <= 172 ~ "DAYS_LENGTHENING",
+  filter(!is.na(LATITUDE)) %>%
+  mutate(latbin = floor(abs(LATITUDE)/5)) %>%
+  filter(WHERE == "GWAS") %>%
+  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "Shortening",
+                               MEAS <= 172 ~ "Lengthening",
                                TRUE ~ NA_character_)) %>%
-  filter(YEAR == 2019 & WHERE == "GWAS" & SUBPOP %in% c("Gulf", "Midwest")) %>% # , 
-  group_by(PLANT_ID) %>%
-  add_tally() %>%
-  filter(n == 9 | n > 13) %>%
-  ggplot(aes(x = PLANT_ID, y = DYLN)) +
-  geom_violin(color = "grey", scale = "area") +
-  geom_jitter(aes(color = SITE, shape = SUBPOP, fill = FL_B4_SOL), stroke = 1.2, size = 1.5, height = 0, width = 0.1) + 
-  theme(axis.text.x = element_text(hjust = 1, angle = 45), legend.position = "right", panel.spacing.x = unit(x = 5, units = "points")) + 
-  scale_color_viridis(discrete = TRUE, direction = -1) +
-  geom_hline(data = max_dyln, aes(yintercept = max_dyln, color = SITE), linetype = 2) + 
-  #facet_wrap(~SUBPOP) +
-  scale_shape_manual(values = c(21, 24)) +
-  scale_fill_viridis(option = "B", discrete = TRUE, direction = -1)
-  geom_label(data = max_dyln, aes(x = 4, y = max_dyln, label = SITE))
+  group_by(SITE, latbin) %>% tally()
 
-fl1_dyln %>%
-  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "DAYS_SHORTENING",
-                               MEAS <= 172 ~ "DAYS_LENGTHENING",
+dyln_latbin <- fl50_dyln %>%
+  filter(!is.na(LATITUDE)) %>%
+  mutate(latbin = floor(abs(LATITUDE))) %>%
+  filter(WHERE == "GWAS" & !(SITE %in% c("OVTN", "FRMI"))) %>%
+  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "Shortening",
+                               MEAS <= 172 ~ "Lengthening",
                                TRUE ~ NA_character_)) %>%
-  filter(YEAR == 2019 & WHERE == "GWAS" & SUBPOP %in% c("8X", "Atlantic")) %>% # , 
-  group_by(PLANT_ID) %>%
-  add_tally() %>%
-  filter(n == 9 | n > 13) %>%
-  ggplot(aes(x = PLANT_ID, y = DYLN)) +
-  geom_violin(color = "grey", scale = "area") +
-  geom_jitter(aes(color = SITE, shape = SUBPOP, fill = FL_B4_SOL), stroke = 1.2, size = 1.5, height = 0, width = 0.1) + 
-  theme(axis.text.x = element_text(hjust = 1, angle = 45), legend.position = "right", panel.spacing.x = unit(x = 5, units = "points")) + 
-  scale_color_viridis(discrete = TRUE, direction = -1) +
-  geom_hline(data = max_dyln, aes(yintercept = max_dyln, color = SITE), linetype = 2) + 
-  #facet_wrap(~SUBPOP) +
-  scale_shape_manual(values = c(21, 24)) +
-  scale_fill_viridis(option = "B", discrete = TRUE, direction = -1)
-  geom_label(data = max_dyln, aes(x = 4, y = max_dyln, label = SITE))
+  group_by(SITE, FL_B4_SOL, latbin, PLANT_ID, SUBPOP) %>%
+  summarise(count = n()) %>%
+  dplyr::select(-PLANT_ID) %>%
+  complete(SITE, FL_B4_SOL, latbin, SUBPOP) %>%
+  group_by(SITE, FL_B4_SOL, latbin, SUBPOP) %>%
+  summarise(count = sum(!is.na(count))) %>%
+  complete(SITE, FL_B4_SOL, latbin, SUBPOP, fill = list(count = 0)) %>%
+  arrange(latbin)
 
+#dyln_change <- 
+dyln_latbin %>%
+  left_join(sites, by = "SITE") %>%
+  filter(!(SITE %in% c("FRMI", "OVTN"))) %>%
+  ggplot(aes(x = FL_B4_SOL, y = count, width = 0.8)) + 
+  geom_bar(aes(fill = as_factor(latbin)), stat = "identity", position = position_dodge()) +
+  facet_wrap(~manu_site, nrow = 2) +
+  scale_fill_viridis(option = "B", discrete = TRUE, direction = -1) +
+  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
+        legend.position = "right", panel.spacing.x = unit(x = 0.01, unit = "npc")) +
+  ylab("Count of Genotypes") +
+  xlab("Do genotypes from that latitude of origin flower when days are 
+       lengthening or shortening?") 
+save_plot(filename = "../manuscript/plots/daylength_change_subpops.png", plot = dyln_change)
+dyln_latbin2 <- dyln_latbin %>% ungroup() %>% add_row(latbin = 24, count = 0) %>%
+  complete(SITE, FL_B4_SOL, latbin, SUBPOP, fill = list(count = 0)) %>%
+  filter(!is.na(FL_B4_SOL) & !is.na(SUBPOP)) %>%
+  mutate(count = case_when(latbin == 24 ~ 48,
+                           TRUE ~ count)) %>%
+  unique()
 ```
+
+```{r}
+latranges <- metadata %>%
+  group_by(SUBPOP) %>%
+  summarise(min = min(LATITUDE, na.rm = TRUE),
+            max = max(LATITUDE, na.rm = TRUE))
+
+metadata$SUBPOP <- factor(metadata$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
+paneld <- metadata %>%
+  ggplot(aes(x = SUBPOP, y = LATITUDE)) +
+  geom_violin(aes(color = SUBPOP)) + geom_jitter(aes(color = SUBPOP), alpha = 0.1, height = 0, width = 0.1) + scale_color_manual(values = rev(color_code)) + theme(legend.position = "none", axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45)) + ylab("Latitude of origin") + xlab("Subpopulation") + labs(title = "", subtitle = "")
+```
+
+```{r}
+
+panela <- dyln_latbin2 %>%
+  left_join(sites, by = "SITE") %>%
+  filter((SITE %in% c("KING"))) %>%
+  mutate(n = case_when(FL_B4_SOL == "Lengthening" ~ -count,
+                       TRUE ~ count)) %>%
+  ggplot(aes(x = latbin, y = n, width = 1)) + 
+  geom_col(aes(fill = as_factor(latbin)), position = "dodge") +
+  coord_flip() +
+  facet_wrap(~FL_B4_SOL, ncol = 2, scales = "free_x") +
+  scale_fill_viridis(option = "B", discrete = TRUE, direction = -1) +
+  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
+        legend.position = "none", panel.spacing.x = unit(x = 0, unit = "mm")) +
+  scale_y_continuous(expand = c(0.002, 0.002), 
+    labels = function(x) signif(abs(x), 3)) +
+  ylab("Number of genotypes") +
+  xlab("Latitude of origin") + labs(title = "TX1", subtitle = "Day length change on date of flowering") +
+  xlim(c(25,45))
+panelb <- dyln_latbin2 %>%
+  left_join(sites, by = "SITE") %>%
+  filter((SITE %in% c("PKLE"))) %>%
+  mutate(n = case_when(FL_B4_SOL == "Lengthening" ~ -count,
+                       TRUE ~ count)) %>%
+  ggplot(aes(x = latbin, y = n, width = 1)) + 
+  geom_col(aes(fill = as_factor(latbin)), position = "dodge") +
+  coord_flip() +
+  facet_wrap(~FL_B4_SOL, ncol = 2, scales = "free_x") +
+  scale_fill_viridis(option = "B", discrete = TRUE, direction = -1) +
+  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
+        legend.position = "none", panel.spacing.x = unit(x = 0, unit = "mm")) +
+  scale_y_continuous(expand = c(0.002, 0.002), 
+    labels = function(x) signif(abs(x), 3)) +
+  ylab("Number of genotypes") +
+  xlab("Latitude of origin") + labs(title = "TX2", subtitle = "Day length change on date of flowering") +
+  xlim(c(25,45))
+panelc <- dyln_latbin2 %>%
+  left_join(sites, by = "SITE") %>%
+  filter((SITE %in% c("TMPL"))) %>%
+  mutate(n = case_when(FL_B4_SOL == "Lengthening" ~ -count,
+                       TRUE ~ count)) %>%
+  ggplot(aes(x = latbin, y = n, width = 1)) + 
+  geom_col(aes(fill = as_factor(latbin)), position = "dodge") +
+  coord_flip() +
+  facet_wrap(~FL_B4_SOL, ncol = 2, scales = "free_x") +
+  scale_fill_viridis(option = "B", discrete = TRUE, direction = -1) +
+  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
+        legend.position = "none", panel.spacing.x = unit(x = 0, unit = "mm")) +
+  scale_y_continuous(expand = c(0.002, 0.002), 
+    labels = function(x) signif(abs(x), 3)) +
+  ylab("Number of genotypes") +
+  xlab("Latitude of origin") + labs(title = "TX3", subtitle = "Day length change on date of flowering") +
+  xlim(c(25,45))
+
+dyln_lat_plot <- plot_grid(paneld, panela, panelb, panelc, nrow = 1)
+save_plot(filename = "../manuscript/plots/Daylength_changes_flowering_date_by_latitude_of_origin_subpops.svg", plot = dyln_lat_plot, base_asp = 2.6, base_height = 4)
+```
+
+```{r}
+filter(unique(dyln_latbin2), count > 200)
+panela <- dyln_latbin2 %>%
+  left_join(sites, by = "SITE") %>%
+  filter((SITE %in% c("KING"))) %>%
+  mutate(n = case_when(FL_B4_SOL == "Lengthening" ~ -count,
+                       TRUE ~ count)) %>%
+  ggplot(aes(x = latbin, y = n, width = 1)) + 
+  geom_col(aes(fill = SUBPOP)) +
+  coord_flip() +
+  facet_wrap(~FL_B4_SOL, ncol = 2, scales = "free_x") +
+  scale_fill_manual(values = rev(color_code)) +
+  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
+        legend.position = "none", panel.spacing.x = unit(x = 0, unit = "mm")) +
+  scale_y_continuous(expand = c(0.002, 0.002), 
+    labels = function(x) signif(abs(x), 3)) +
+  ylab("Number of genotypes") +
+  xlab("Latitude of origin") + labs(title = "TX1", subtitle = "Day length change on date of flowering") +
+  xlim(c(24,45))
+panelb <- dyln_latbin2 %>%
+  left_join(sites, by = "SITE") %>%
+  filter((SITE %in% c("PKLE"))) %>%
+  mutate(n = case_when(FL_B4_SOL == "Lengthening" ~ -count,
+                       TRUE ~ count)) %>%
+  ggplot(aes(x = latbin, y = n, width = 1)) + 
+  geom_col(aes(fill = SUBPOP)) +
+  coord_flip() +
+  facet_wrap(~FL_B4_SOL, ncol = 2, scales = "free_x") +
+  scale_fill_manual(values = rev(color_code)) +
+  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
+        legend.position = "none", panel.spacing.x = unit(x = 0, unit = "mm")) +
+  scale_y_continuous(expand = c(0.002, 0.002), 
+    labels = function(x) signif(abs(x), 3)) +
+  ylab("Number of genotypes") +
+  xlab("Latitude of origin") + labs(title = "TX2", subtitle = "Day length change on date of flowering") +
+  xlim(c(24,45))
+panelc <- dyln_latbin2 %>%
+  left_join(sites, by = "SITE") %>%
+  filter((SITE %in% c("TMPL"))) %>%
+  mutate(n = case_when(FL_B4_SOL == "Lengthening" ~ -count,
+                       TRUE ~ count)) %>%
+  ggplot(aes(x = latbin, y = n, width = 1)) + 
+  geom_col(aes(fill = SUBPOP)) +
+  coord_flip() +
+  facet_wrap(~FL_B4_SOL, ncol = 2, scales = "free_x") +
+  scale_fill_manual(values = rev(color_code)) +
+  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
+        legend.position = "right", panel.spacing.x = unit(x = 0, unit = "mm")) +
+  scale_y_continuous(expand = c(0.002, 0.002), 
+    labels = function(x) signif(abs(x), 3)) +
+  ylab("Number of genotypes") +
+  xlab("Latitude of origin") + labs(title = "TX3", subtitle = "Day length change on date of flowering") +
+  xlim(c(24,45))
+
+dyln_lat_plot <- plot_grid(panela, panelb, panelc, nrow = 1, rel_widths = c(1, 1, 1.318))
+save_plot(filename = "../manuscript/plots/Daylength_changes_flowering_date_by_latitude_of_origin_and_subpop.png", plot = dyln_lat_plot, base_asp = 2.4, base_height = 4)
+```

### 2020-06-05 R/D2F_exploration_fourway_and_gwas.Rmd
Manuscript changes completing draft of two environmental cues section

@@
   filter(!is.na(D2F) & !(SITE %in% c("FRMI_D2F", "BRKG_D2F", "TMPL_D2F"))) %>%
-  count() %>% 
+  count() %>% a
   arrange(desc(n)) %>%

### 2020-06-05 R/Draft_figures_2020-06-02.Rmd
4WCR F0 F1 and F2 figures and comparisons

@@
 phe_F0_up <- phe_all %>%
-  filter(PLANT_ID %in% c("DAC", "VS16", "J440.A", "J036.A", "J037.A")) 
+  filter(WHERE == "FWCR" & PLANT_ID %in% c("DAC", "VS16", "J440.A", "J036.A", "J037.A")) 
 phe_F0_low <- phe_all %>%
-  filter(PLANT_ID %in% c("AP13", "WBC", "J230.A")) 
+  filter(WHERE == "FWCR" & PLANT_ID %in% c("AP13", "WBC", "J230.A")) 
 phe_F0 <- phe_all %>%
-  filter(PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "J440.A", "J440.B", "J230.A", "J292.A", "J045.A", "J036.A", "J037.A", "J022.B", "J210.A")) 
+  filter(WHERE == "FWCR" & PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "J440.A", "J440.B", "J230.A", "J292.A", "J045.A", "J036.A", "J037.A", "J022.B", "J210.A")) 
 phe_F1 <- phe_all %>%
-  filter(PLANT_ID %in% c("AxD", "VxW")) 
+  filter(WHERE == "FWCR" & PLANT_ID %in% c("AxD", "VxW")) 
 phe_F2 <- phe_all %>%
-  filter(!(WHERE == "FWCR" & PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "AxD", "VxW"))) 
+  filter(WHERE == "FWCR" & !(PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "AxD", "VxW"))) 
 ```
@@
 ```
+
+# --------------
+# 4WCR Plots
+
+We then planted the four parents, the two F1 individuals (AP13xDAC, and VS16xWBC), and XXX F2 individuals at eight field sites, and recorded greenup and flowering date for the 2016-2019 seasons.
+Though there was a X day gap in flowering date between the two F1 crosses, F1 individuals flowered at similar dates as the Midwest parents and Midwest subpopulation individuals in general, indicating that the flowering time genetic response to the Midwestern GDD levels was dominant to the photoperiod response and the higher GDD levels required by the Gulf subpopulation.
+```{r}
+phe_F1 %>%
+  filter(PHE == "FL50" & YEAR == 2019) %>%
+  group_by(SITE, PLANT_ID) %>%
+  summarise(fl50 = mean(MEAS, na.rm = TRUE)) %>%
+  pivot_wider(names_from = PLANT_ID, values_from = fl50) %>%
+  mutate(diff = VxW - AxD) %>%
+  group_by(SITE) %>%
+  summarise(meandiff = mean(diff)) %>%
+  arrange(meandiff) %>%
+  filter(!(SITE %in% c("OVTN", "MNHT"))) %>%
+  summarise(meandiff = mean(meandiff))
+
+F1plot <- phe_F1 %>%
+  filter(PHE == "FL50" & (YEAR %in% c(2019) | (SITE == "TMPL" & YEAR == 2018)) & !(SITE %in% c("OVTN", "MNHT"))) %>%
+  left_join(sites, by = "SITE") %>%
+  group_by(manu_site, YEAR, PLANT_ID) %>%
+  summarise(MEAS = mean(MEAS, na.rm = TRUE))
+phe_all$SUBPOP <- factor(phe_all$SUBPOP, levels = rev(c("4X", "Atlantic", "Gulf", "Midwest", "8X")))
+color_code <- c("grey", "#6E91CB","#F47F72", "#442C83", "#090906")
+
+F1vsGWAS <- phe_all %>%
+  left_join(sites, by = "SITE") %>%
+  filter(PHE == "FL50" & WHERE == "GWAS" & !(SITE %in% c("FRMI", "OVTN")) & SUBPOP %in% c("Midwest", "Gulf")) %>%
+  ggplot(aes(x = as_date(MEAS, origin = "2019-01-01"))) +
+  geom_vline(xintercept = as_date(173, origin = "2019-01-01"), color = "#FDE725FF", size = 1) +
+  geom_density(aes(color = SUBPOP)) + 
+  facet_wrap(~manu_site, nrow = 2) +
+  geom_vline(data = F1plot, aes(xintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 2) +
+  scale_color_manual(values = c("grey", "#F47F72", "#442C83",  "#090906")) +
+  theme(legend.position = "right", axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45)) +
+  ylab("Subpop Fraction") + xlab("Flowering Date") +
+  geom_hline(yintercept = 0, color = "darkgrey")
+save_plot(filename = "../manuscript/plots/F1_vs_GWAS_distribution.png", plot = F1vsGWAS, base_asp = 2.5, base_height = 2.5)
+
+F1vsGWAS_violin <- phe_all %>%
+  left_join(sites, by = "SITE") %>%
+  filter(PHE == "FL50" & WHERE == "GWAS" & !(SITE %in% c("FRMI", "OVTN")) & SUBPOP %in% c("Midwest", "Gulf")) %>%
+  ggplot(aes(x = SUBPOP, y = as_date(MEAS, origin = "2019-01-01"))) +
+  geom_hline(yintercept = as_date(173, origin = "2019-01-01"), color = "#FDE725FF", size = 1) +
+  geom_violin(aes(fill = SUBPOP)) + 
+  facet_wrap(~manu_site, nrow = 2) +
+  geom_hline(data = F1plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 2) +
+  scale_color_manual(values = c("grey",  "#090906", "#F47F72", "#442C83")) +
+  scale_fill_manual(values = c("#442C83", "#F47F72")) +
+  theme(legend.position = "right", axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45)) +
+  xlab("Subpopulation") + ylab("Flowering Date") + coord_flip()
+
+save_plot(filename = "../manuscript/plots/F1_vs_GWAS_distribution_violin.png", plot = F1vsGWAS_violin, base_asp = 2.5, base_height = 2.5)
+
+library(scales)
+show_col(viridis(8, option = "B"))
+```
+## F0 Plot
+
+"#D44842FF", "#280B54FF", "#9F2A63FF", "#FAC127FF"
+```{r}
+F0plot <- phe_F0 %>%
+  filter(PHE == "FL50" & (YEAR %in% c(2016:2019)) & !(SITE %in% c("OVTN", "MNHT"))) %>%
+  left_join(sites, by = "SITE") %>%
+  group_by(manu_site, PLANT_ID) %>%
+  summarise(MEAS = mean(MEAS, na.rm = TRUE))
+
+F0vsGWAS <- phe_all %>%
+  left_join(sites, by = "SITE") %>%
+  filter(PHE == "FL50" & WHERE == "GWAS" & !(SITE %in% c("FRMI", "OVTN")) & SUBPOP %in% c("Midwest", "Gulf")) %>%
+  ggplot(aes(x = SUBPOP, y = as_date(MEAS, origin = "2019-01-01"))) +
+  geom_hline(yintercept = as_date(173, origin = "2019-01-01"), color = "#FDE725FF", size = 1) +
+  geom_violin(aes(fill = SUBPOP)) + 
+  facet_wrap(~manu_site, nrow = 2) +
+  geom_hline(data = F0plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 2) +
+  scale_fill_manual(values = c("#442C83", "#F47F72")) +
+  scale_color_manual(values = c("#D44842FF", "#280B54FF", "#9F2A63FF", "#FAC127FF")) +
+  coord_flip() + theme(legend.position = "right", axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45)) +
+  xlab("Subpopulation") + ylab("Flowering Date")
+F0vsGWAS
+save_plot(filename = "../manuscript/plots/F0_vs_GWAS_distribution.png", plot = F0vsGWAS)
+```
+
+```{r}
+phe_F2 %>%
+  group_by(PLANT_ID) %>%
+  summarise(count = n()) %>% arrange(count)
+
+FWCRvsGWAS <- phe_all %>%
+  filter((WHERE == "FWCR") | (WHERE == "GWAS" & SUBPOP %in% c("Midwest", "Gulf"))) %>%
+  filter(PHE == "FL50" & !(SITE %in% c("OVTN", "MNHT", "FRMI"))) %>%
+  left_join(sites, by = "SITE") %>%
+  dplyr::select(-(Longitude:Note))  %>%
+  mutate(POP = case_when(WHERE == "FWCR" & !(PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "AxD", "VxW")) ~ "F2",
+                         WHERE == "FWCR" & (PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC")) ~ "F0",
+                         WHERE == "FWCR" & (PLANT_ID %in% c("AxD", "VxW")) ~ "F1",
+                            WHERE == "GWAS" ~ as.character(SUBPOP)))
+
+summary(as_factor(FWCRvsGWAS$POP))
+F0plot <- phe_F0 %>%
+  filter(PHE == "FL50" & (YEAR %in% c(2016:2019)) & !(SITE %in% c("OVTN", "MNHT"))) %>%
+  left_join(sites, by = "SITE") %>%
+  group_by(manu_site, PLANT_ID) %>%
+  summarise(MEAS = mean(MEAS, na.rm = TRUE))
+
+FWCRvsGWAS %>%
+  filter(!(POP %in% c("F0", "F1")) & (YEAR == 2019 | (YEAR == 2018 & SITE == "TMPL"))) %>%
+  ggplot(aes(x = POP, y = as_date(MEAS, origin = "2019-01-01"))) +
+  geom_hline(yintercept = as_date(173, origin = "2019-01-01"), color = "#FDE725FF", size = 1) +
+  geom_violin(aes(fill = POP)) + 
+  facet_wrap(~manu_site, nrow = 2) +
+  geom_hline(data = F0plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 2) +
+  geom_hline(data = F1plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 3) +
+  scale_fill_manual(values = c("lightgrey", "#F47F72", "#442C83")) +
+  scale_color_manual(values = c("#D44842FF", "grey", "#280B54FF", "#9F2A63FF", "black", "#FAC127FF")) +
+  coord_flip() + theme(legend.position = "right", axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45)) +
+  xlab("Subpopulation") + ylab("Flowering Date")
+
+FWCRvsGWAS %>%
+  filter(!(POP %in% c("F0", "F1")) & (YEAR == 2018 | (WHERE == "GWAS" & YEAR == 2019))) %>%
+  ggplot(aes(x = POP, y = as_date(MEAS, origin = "2019-01-01"))) +
+  geom_hline(yintercept = as_date(173, origin = "2019-01-01"), color = "#FDE725FF", size = 1) +
+  geom_violin(aes(fill = POP)) + 
+  facet_wrap(~manu_site, nrow = 2) +
+  geom_hline(data = F0plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 2) +
+  geom_hline(data = F1plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 3) +
+  scale_fill_manual(values = c("lightgrey", "#F47F72", "#442C83")) +
+  scale_color_manual(values = c("#D44842FF", "grey", "#280B54FF", "#9F2A63FF", "black", "#FAC127FF")) +
+  coord_flip() + theme(legend.position = "right", axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45)) +
+  xlab("Subpopulation") + ylab("Flowering Date")
+
+FWCRvsGWAS %>%
+  filter(!(POP %in% c("F0", "F1")) & WHERE == "FWCR") %>%
+  ggplot(aes(x = as_factor(YEAR), y = as_date(MEAS, origin = "2019-01-01"))) +
+  geom_hline(yintercept = as_date(173, origin = "2019-01-01"), color = "#FDE725FF", size = 1) +
+  geom_violin(aes(fill = as_factor(YEAR))) + 
+  facet_wrap(~manu_site, nrow = 2) +
+  geom_jitter(data = filter(FWCRvsGWAS, POP %in% c("F0", "F1")), aes(color = PLANT_ID), shape = 4, height = 0) +
+  #geom_hline(data = F0plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 2) +
+  #geom_hline(data = F1plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 3) +
+  scale_fill_viridis(discrete = TRUE, end = 0.8) + 
+  scale_color_manual(values = c("#D44842FF", "grey", "#280B54FF", "#9F2A63FF", "black", "#FAC127FF")) +
+  coord_flip() + theme(legend.position = "right", axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45)) +
+  xlab("Year") + ylab("Flowering Date") + ylim(c(as_date(50, origin = "2019-01-01"), as_date(300, origin = "2019-01-01")))
+
+save_plot(filename = "../manuscript/plots/FWCR_distribution_by_year.png", plot = last_plot())
+
+FWCRvsGWAS %>%
+  filter(!(POP %in% c("F0", "F1")) & WHERE == "FWCR") %>%
+  ggplot(aes(x = as_factor(YEAR), y = as_date(MEAS, origin = "2019-01-01"))) +
+  geom_hline(yintercept = as_date(173, origin = "2019-01-01"), color = "#FDE725FF", size = 1) +
+  geom_jitter(data = filter(FWCRvsGWAS, POP %in% c("F0")), aes(color = PLANT_ID), shape = 4, height = 0) +
+  geom_violin(aes(fill = as_factor(YEAR))) + 
+  facet_wrap(~manu_site, nrow = 2) +
+
+  #geom_hline(data = F0plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 2) +
+  #geom_hline(data = F1plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 3) +
+  scale_fill_viridis(discrete = TRUE, end = 0.8) + 
+  scale_color_manual(values = c("#D44842FF", "#280B54FF", "#9F2A63FF",  "#FAC127FF")) +
+  coord_flip() + theme(legend.position = "right", axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45)) +
+  xlab("Year") + ylab("Flowering Date") + ylim(c(as_date(50, origin = "2019-01-01"), as_date(300, origin = "2019-01-01")))
+
+save_plot(filename = "../manuscript/plots/FWCR_F0_distribution_by_year.png", plot = last_plot())
+
+```
+
+(How many days behind the parents was each F1 cross, on average, across all 4 years of data?)  
+```{r}
+F1_F0_diff <- FWCRvsGWAS %>%
+  filter((POP %in% c("F0", "F1")) & WHERE == "FWCR") %>%
+  group_by(SITE, YEAR, PLANT_ID) %>%
+  summarise(meanFL = mean(MEAS, na.rm = TRUE)) %>%
+  pivot_wider(names_from = PLANT_ID, values_from = meanFL) %>%
+  mutate(earlyFL_F1_F0_diff = AxD - DAC,
+         lateFL_F1_F0_diff = VxW - VS16,
+         earlyFL_F1_F0_diff_g = AP13 - AxD,
+         lateFL_F1_F0_diff_g = WBC - VxW
+         )  %>%
+  ungroup()
+F1_F0_diff %>%
+  summarise(mean_early_F1_F0_diff = mean(earlyFL_F1_F0_diff, na.rm = TRUE),
+            mean_late_F1_F0_diff = mean(lateFL_F1_F0_diff, na.rm = TRUE),
+            mean_early_F1_F0_diff_g = mean(earlyFL_F1_F0_diff_g, na.rm = TRUE),
+            mean_late_F1_F0_diff_g = mean(lateFL_F1_F0_diff_g, na.rm = TRUE))
+
+sd(F1_F0_diff$earlyFL_F1_F0_diff, na.rm = TRUE)
+sd(F1_F0_diff$lateFL_F1_F0_diff, na.rm = TRUE)
+sd(F1_F0_diff$earlyFL_F1_F0_diff_g, na.rm = TRUE)
+sd(F1_F0_diff$lateFL_F1_F0_diff_g, na.rm = TRUE)
+```
+
+```{r}
+Gulf_minFL <- FWCRvsGWAS %>%
+  filter((POP %in% c("F0"))) %>%
+  group_by(SITE, YEAR, PLANT_ID) %>%
+  summarise(minFL = min(MEAS, na.rm = TRUE)) %>%
+  filter(PLANT_ID %in% c("AP13", "WBC")) %>%
+  pivot_wider(names_from = PLANT_ID, values_from = minFL, names_prefix = "minFL_")
+Gulf_meanFL <- FWCRvsGWAS %>%
+  filter((POP %in% c("F0"))) %>%
+  group_by(SITE, YEAR, PLANT_ID) %>%
+  summarise(meanFL = mean(MEAS, na.rm = TRUE)) %>%
+  filter(PLANT_ID %in% c("AP13", "WBC")) %>%
+  pivot_wider(names_from = PLANT_ID, values_from = meanFL, names_prefix = "meanFL_")
+
+F2_meanFL <- FWCRvsGWAS %>%
+  filter((POP %in% c("F2"))) %>%
+  group_by(SITE, YEAR, PLANT_ID) %>%
+  summarise(meanFL = mean(MEAS, na.rm = TRUE)) %>% 
+  left_join(Gulf_meanFL) %>%
+  filter(meanFL > meanFL_AP13 | meanFL > meanFL_WBC)
+F2_minFL <- FWCRvsGWAS %>%
+  filter((POP %in% c("F2"))) %>%
+  group_by(SITE, YEAR, PLANT_ID) %>%
+  summarise(meanFL = mean(MEAS, na.rm = TRUE)) %>% 
+  left_join(Gulf_minFL) %>%
+  filter(meanFL >= minFL_AP13 | meanFL >= minFL_WBC)
+
+F2_meanFL %>%
+  group_by(YEAR, PLANT_ID) %>%
+  tally() %>% arrange(desc(n))
+F2_minFL %>%
+  group_by( PLANT_ID) %>%
+  tally() %>% arrange(desc(n))
+
+478/14672
+2308/14672
+```
+Only 3.26% of F2 flowering dates occurred after the average flowering date for one or both Gulf parents; 13.3% of F2 flowering dates occurred after the minimum flowering date for one or both Gulf parents.
+
+Only 358 of 801 F2 individuals ever flowered after the average flowering date Gulf parent, and only 102 did so in more than one site or year. (13 in three + site/years, 2 in 5-6).
+
+minimum flowering date for Gulf parents
+PLANT_ID  n
+91	15			
+402	13			
+105	12			
+194	11			
+210	11			
+248	11			
+196	10			
+214	10

### 2020-06-11 data-raw/Data_Setup.Rmd
New data setup for manuscript - metadata, sites, GDD and Julian date phenotypes for greenup and flowering

---
title: "Data_Setup"
author: "Alice MacQueen"
date: "6/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# metadata (germplasm)
```{r}
metadata <- readRDS("~/Github/pvdiv-phenotypes/data/metadata.rds")
metadata <- metadata %>%
  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
                            !is.na(PLOIDY) ~ as.character(PLOIDY),
                            TRUE ~ NA_character_))
write_rds(metadata, "../data/metadata.rds")
fourway <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/0-data/Phenology_2019_Multiple_Environment.csv")
```

# common gardens
```{r}
sites <- readRDS("~/Github/pvdiv-phenotypes/data/sites.rds")

sites <- sites %>%
  mutate(manu_site = case_when(State != "TX" ~ State,
                               SITE == "KING" ~ "TX1",
                               SITE == "PKLE" ~ "TX2",
                               SITE == "TMPL" ~ "TX3",
                               SITE == "OVTN" ~ "TX4"))
sites$manu_site <- factor(sites$manu_site, levels = c("TX1", "TX2", "TX3", "TX4", "OK", "MO", "NE", "IL", "MI", "SD"))
write_rds(sites, "../data/sites.rds")
```

# Subpop color coding
```{r}
subpops <- enframe(unique(metadata$SUBPOP), name = NULL, value = "SUBPOP") %>%
  mutate(color_code = case_when(SUBPOP == "Atlantic" ~ "#6E91CB",
                                SUBPOP == "Gulf" ~ "#F47F72",    
                                SUBPOP == "Midwest" ~ "#442C83",
                                SUBPOP == "4X" ~ "grey", 
                                SUBPOP == "8X" ~ "#090906"))
subpops$SUBPOP <- factor(subpops$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
write_rds(subpops, path = "../data/subpop_color_coding.rds")
```

# Make phenotypes for flowering time paper from all the cleaned phenotypes.

Specifically using GR50 & FL50 Julian dates from 2019. And using them to make cGDD between greenup and flowering for another GWAS phenotype.
```{r}
phenotypes <- read_rds("~/Github/pvdiv-phenotypes/data/Phenotypes_cleaned.rds")
ptt_df <- read_rds("~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")
fourway <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/0-data/Phenology_2019_Multiple_Environment.csv")
fwy_prev_raw <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/0-data/Phenology_2016_2018.csv")
```

Also made a year-specific "PLOT_GL" for the 4WCR. Can use this to calculate cGDD by specific plant, when a GR50 and FL50 measurement are linked for the same plant for a F0 or F1, for example.
```{r}
phe_gwas <- phenotypes %>%
  filter(PHE %in% c("EM1", "EM50", "FL1", "FL50", "GR1", "GR50", "GR100")) %>%
  mutate(YEAR = 2019,
         WHERE = "GWAS") %>%
  select(-manu_site)

fwy_prev <- fwy_prev_raw %>%
  separate(ENVI, into = c("SITE", "YEAR"), sep = "_") %>%
  group_by(LINE, SITE, YEAR) %>%
  mutate(PLOT_GL = paste0(str_sub(SITE, end = 1), str_sub(YEAR, start =3, end = 4), "_", row_number())) %>%
  mutate(YEAR = as.numeric(YEAR),
         WHERE = "FWCR") %>%
  pivot_longer(cols = BIOMASS:HT, names_to = "PHE", values_to = "MEAS") %>%
  rename(PLANT_ID = LINE) %>%
  filter(PHE %in% c("FL50", "GR50"))
fwy_2019 <- fourway %>%
  separate(ENVI, into = c("SITE", "YEAR"), sep = "_") %>%
  group_by(LINE, SITE, YEAR) %>%
  mutate(PLOT_GL = paste0(str_sub(SITE, end = 1), str_sub(YEAR, start =3, end = 4), "_", row_number())) %>%
  mutate(YEAR = as.numeric(YEAR),
         WHERE = "FWCR") %>%
  pivot_longer(cols = FL50:HT, names_to = "PHE", values_to = "MEAS") %>%
  rename(PLANT_ID = LINE) %>%
  filter(PHE %in% c("FL50", "GR50"))

phe_all <- phe_gwas %>%
  full_join(fwy_2019, by = c("SITE", "PLOT_GL", "PLANT_ID", "YEAR", "PHE", "MEAS", "WHERE")) %>%
  full_join(fwy_prev, by = c("SITE", "PLOT_GL", "PLANT_ID", "YEAR", "PHE", "MEAS", "WHERE")) %>%
  filter(!is.na(MEAS))

phe_all %>%
  group_by(PLOT_GL, SITE, PLANT_ID, YEAR) %>%
  tally() %>% arrange((n))

write_rds(phe_all, path = "../data/Phenology_phenotypes_4wcr_and_gwas.rds")
```

cGDD between GR50 and FL50 for each YEAR, each PLANT_ID (and PLOT_GL)
```{r}
wea_df <- ptt_df %>%
  dplyr::select(DOY:GDD_12C_SM, PTT_12C)

fl_timing <- phe_all %>% 
  filter(YEAR == 2019 & PHE %in% c("GR50", "FL50")) %>%
  pivot_wider(names_from = PHE, values_from = MEAS) %>%
  select(-SUBPOP) %>%
  drop_na()

wea_gr_fl <- wea_df %>%
  left_join(fl_timing)

wea_gr_fl2 <- wea_gr_fl %>%
  group_by(Day_of_year, PLOT_GL, PLANT_ID) %>%
  filter(between(Day_of_year, GR50, FL50)) %>% 
  ungroup()

cgdd_df <- wea_gr_fl2 %>%
  group_by(SITE, PLANT_ID, PLOT_GL, GR50, FL50) %>%
  summarise(CGDD_12C = sum(GDD_12C_SM),
            PTT_12C = sum(PTT_12C),
            CRAIN = sum(RAIN_SM)
            )

write_rds(cgdd_df, path = "../data/Julian_date_and_GDD_phenotypes_4wcr_and_gwas.rds")

wea_gr_fl3 <- wea_gr_fl %>%
  group_by(Day_of_year, PLOT_GL, PLANT_ID) %>%
  filter(between(Day_of_year, FL50-5, FL50)) %>% 
  ungroup() 
crain_df <- wea_gr_fl3 %>%
  group_by(SITE, PLANT_ID, PLOT_GL, GR50, FL50) %>%
  summarise(CGDD_12C_5d = sum(GDD_12C_SM),
            PTT_12C_5d = sum(PTT_12C),
            CRAIN_5d = sum(RAIN_SM)
            )

cumweadf <- cgdd_df %>%
  left_join(crain_df)

write_rds(cumweadf, path = "../data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")

```

### 2020-06-11 data-raw/Flowering_GWAS_Phenotypes.Rmd
New data setup for manuscript - metadata, sites, GDD and Julian date phenotypes for greenup and flowering

--- 2020-06-11 data-raw/Data_Setup.Rmd
@@
 ---
-title: "Data_Setup"
+title: "Flowering_GWAS_Phenotypes"
 author: "Alice MacQueen"
@@
 library(tidyverse)
-```
-
-# metadata (germplasm)
-```{r}
-metadata <- readRDS("~/Github/pvdiv-phenotypes/data/metadata.rds")
-metadata <- metadata %>%
-  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
-                            !is.na(PLOIDY) ~ as.character(PLOIDY),
-                            TRUE ~ NA_character_))
-write_rds(metadata, "../data/metadata.rds")
-fourway <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/0-data/Phenology_2019_Multiple_Environment.csv")
-```
-
-# common gardens
-```{r}
-sites <- readRDS("~/Github/pvdiv-phenotypes/data/sites.rds")
-
-sites <- sites %>%
-  mutate(manu_site = case_when(State != "TX" ~ State,
-                               SITE == "KING" ~ "TX1",
-                               SITE == "PKLE" ~ "TX2",
-                               SITE == "TMPL" ~ "TX3",
-                               SITE == "OVTN" ~ "TX4"))
-sites$manu_site <- factor(sites$manu_site, levels = c("TX1", "TX2", "TX3", "TX4", "OK", "MO", "NE", "IL", "MI", "SD"))
-write_rds(sites, "../data/sites.rds")
-```
-
-# Subpop color coding
-```{r}
-subpops <- enframe(unique(metadata$SUBPOP), name = NULL, value = "SUBPOP") %>%
-  mutate(color_code = case_when(SUBPOP == "Atlantic" ~ "#6E91CB",
-                                SUBPOP == "Gulf" ~ "#F47F72",    
-                                SUBPOP == "Midwest" ~ "#442C83",
-                                SUBPOP == "4X" ~ "grey", 
-                                SUBPOP == "8X" ~ "#090906"))
-subpops$SUBPOP <- factor(subpops$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
-write_rds(subpops, path = "../data/subpop_color_coding.rds")
+phenotypes <- read_rds("~/Github/pvdiv-phenotypes/data/Phenotypes_cleaned.rds")
 ```
@@
 ```{r}
-phenotypes <- read_rds("~/Github/pvdiv-phenotypes/data/Phenotypes_cleaned.rds")
-ptt_df <- read_rds("~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")
-fourway <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/0-data/Phenology_2019_Multiple_Environment.csv")
-fwy_prev_raw <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/0-data/Phenology_2016_2018.csv")
-```
-
-Also made a year-specific "PLOT_GL" for the 4WCR. Can use this to calculate cGDD by specific plant, when a GR50 and FL50 measurement are linked for the same plant for a F0 or F1, for example.
-```{r}
-phe_gwas <- phenotypes %>%
-  filter(PHE %in% c("EM1", "EM50", "FL1", "FL50", "GR1", "GR50", "GR100")) %>%
-  mutate(YEAR = 2019,
-         WHERE = "GWAS") %>%
-  select(-manu_site)
-
-fwy_prev <- fwy_prev_raw %>%
-  separate(ENVI, into = c("SITE", "YEAR"), sep = "_") %>%
-  group_by(LINE, SITE, YEAR) %>%
-  mutate(PLOT_GL = paste0(str_sub(SITE, end = 1), str_sub(YEAR, start =3, end = 4), "_", row_number())) %>%
-  mutate(YEAR = as.numeric(YEAR),
-         WHERE = "FWCR") %>%
-  pivot_longer(cols = BIOMASS:HT, names_to = "PHE", values_to = "MEAS") %>%
-  rename(PLANT_ID = LINE) %>%
-  filter(PHE %in% c("FL50", "GR50"))
-fwy_2019 <- fourway %>%
-  separate(ENVI, into = c("SITE", "YEAR"), sep = "_") %>%
-  group_by(LINE, SITE, YEAR) %>%
-  mutate(PLOT_GL = paste0(str_sub(SITE, end = 1), str_sub(YEAR, start =3, end = 4), "_", row_number())) %>%
-  mutate(YEAR = as.numeric(YEAR),
-         WHERE = "FWCR") %>%
-  pivot_longer(cols = FL50:HT, names_to = "PHE", values_to = "MEAS") %>%
-  rename(PLANT_ID = LINE) %>%
-  filter(PHE %in% c("FL50", "GR50"))
-
-phe_all <- phe_gwas %>%
-  full_join(fwy_2019, by = c("SITE", "PLOT_GL", "PLANT_ID", "YEAR", "PHE", "MEAS", "WHERE")) %>%
-  full_join(fwy_prev, by = c("SITE", "PLOT_GL", "PLANT_ID", "YEAR", "PHE", "MEAS", "WHERE")) %>%
-  filter(!is.na(MEAS))
-
-phe_all %>%
-  group_by(PLOT_GL, SITE, PLANT_ID, YEAR) %>%
-  tally() %>% arrange((n))
-
-write_rds(phe_all, path = "../data/Phenology_phenotypes_4wcr_and_gwas.rds")
-```
-
-cGDD between GR50 and FL50 for each YEAR, each PLANT_ID (and PLOT_GL)
-```{r}
-wea_df <- ptt_df %>%
-  dplyr::select(DOY:GDD_12C_SM, PTT_12C)
-
-fl_timing <- phe_all %>% 
-  filter(YEAR == 2019 & PHE %in% c("GR50", "FL50")) %>%
-  pivot_wider(names_from = PHE, values_from = MEAS) %>%
-  select(-SUBPOP) %>%
-  drop_na()
-
-wea_gr_fl <- wea_df %>%
-  left_join(fl_timing)
-
-wea_gr_fl2 <- wea_gr_fl %>%
-  group_by(Day_of_year, PLOT_GL, PLANT_ID) %>%
-  filter(between(Day_of_year, GR50, FL50)) %>% 
-  ungroup()
-
-cgdd_df <- wea_gr_fl2 %>%
-  group_by(SITE, PLANT_ID, PLOT_GL, GR50, FL50) %>%
-  summarise(CGDD_12C = sum(GDD_12C_SM),
-            PTT_12C = sum(PTT_12C),
-            CRAIN = sum(RAIN_SM)
-            )
-
-write_rds(cgdd_df, path = "../data/Julian_date_and_GDD_phenotypes_4wcr_and_gwas.rds")
-
-wea_gr_fl3 <- wea_gr_fl %>%
-  group_by(Day_of_year, PLOT_GL, PLANT_ID) %>%
-  filter(between(Day_of_year, FL50-5, FL50)) %>% 
-  ungroup() 
-crain_df <- wea_gr_fl3 %>%
-  group_by(SITE, PLANT_ID, PLOT_GL, GR50, FL50) %>%
-  summarise(CGDD_12C_5d = sum(GDD_12C_SM),
-            PTT_12C_5d = sum(PTT_12C),
-            CRAIN_5d = sum(RAIN_SM)
-            )
-
-cumweadf <- cgdd_df %>%
-  left_join(crain_df)
-
-write_rds(cumweadf, path = "../data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
 

### 2020-06-12 R/GWAS_GDD.R
Modify GDD GWAS script and remove old GWAS run

@@ -5,7 +5,7 @@ library(AnnotationDbi)
 txdb <- loadDb(file = file.path("~", "Github", "pvdiv-genome",
                                 "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
 
-outputdir <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", "gwas")
+outputdir <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", "gwas", "GDD")
 
 inputfiles <- read_delim("~/Github/pvdiv-phenology-gxe/analysis/gwas/inputkinship.txt", delim = " ", col_names = "SNPfiles") # Midwest, Gulf, Atlantic, Midwest & Gulf
 

### 2020-06-12 R/Draft_figures_2020-06-02.Rmd
Make phenotype data for GWAS and add to Data_setup.Rmd

@@
 #
+```{r}
+metadata <- read_rds("../data/metadata.rds")
+subpops <- read_rds("../data/subpop_color_coding.rds")
+phenotypes <- 
+  read_rds("../data/Phenology_CGDD_PTT_and_rainfall_phenotypes.rds")
+sites <- read_rds("../data/sites.rds")
+```
+```{r}
+phe_df <- phenotypes %>%
+  filter(WHERE == "GWAS" & !(SITE %in% c("OVTN", "FRMI"))) %>%
+  unite(col = "SITE_PHE", SITE, PHE) %>%
+  filter(SUBPOP %in% c("Atlantic", "Midwest", "Gulf")) %>%
+  select(-PLOT_GL, -WHERE, -YEAR) %>%
+  pivot_wider(names_from = SITE_PHE, values_from = MEAS, values_fn = mean)
+
+fwcr_df <- phenotypes %>%
+  filter(WHERE == "FWCR" & !(SITE %in% c("OVTN", "FRMI", "MNHT")) & YEAR == 2019) %>%
+  unite(col = "SITE_PHE", SITE, PHE) %>%
+  select( -WHERE, -SUBPOP, -PLOT_GL) %>%
+  pivot_wider(names_from = SITE_PHE, values_from = MEAS, values_fn = mean)
+
+phe_df %>%
+  ggplot(aes(x = PKLE_CRAIN, TMPL_CRAIN)) + geom_point(aes(color = SUBPOP)) + geom_abline(slope = 1, linetype = 2) +
+  scale_color_manual(values = subpops$color_code)
+
+fwcr_df %>%
+  ggplot(aes(x = PKLE_CRAIN, TMPL_CRAIN)) + geom_point() + geom_abline(slope = 1, linetype = 2)
+
+fwcr_df %>%
+  filter(PKLE_CGDD_12C > 1200)
+fwcr_df %>%
+  filter(KBSM_CGDD_12C < 200)
+
+gddcor <- phe_df %>%
+  select(matches("CRAIN$")) %>%
+  cor(., use = "pairwise")
+
+gddcor %>%
+  as_tibble(rownames = "SITE") %>%
+  pivot_longer(cols = -SITE, names_to = "SITE2", values_to = "COR") %>%
+  ggplot(aes(x = SITE, y = SITE2)) + 
+  geom_tile(aes(fill = COR)) +
+  scale_fill_viridis(option = "B") +
+  theme(legend.position = "right", axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
+
+phe_df <- phe_df %>% select(PLANT_ID, SUBPOP, matches("GDD"), matches("GR50"), matches("FL50"), matches("CRAIN"), matches("PTT"), everything())
+
+atl_df <- phe_df %>% filter(SUBPOP == "Atlantic") 
+atl_df <- atl_df %>%
+  select(names(which(colSums(!is.na(atl_df)) > 100)))
+midw_df <- phe_df %>% filter(SUBPOP == "Midwest") 
+midw_df <- midw_df %>%
+  select(names(which(colSums(!is.na(midw_df)) > 100)))
+gulf_df <- phe_df %>% filter(SUBPOP == "Gulf") 
+gulf_df <- gulf_df %>%
+  select(names(which(colSums(!is.na(gulf_df)) > 100)))
+midwgulf_df <- phe_df %>% filter(SUBPOP %in% c("Gulf", "Midwest")) 
+midwgulf_df <- midwgulf_df %>%
+  select(names(which(colSums(!is.na(midwgulf_df)) > 100)))
+
+```
+I've messed up making CGDD for the fourway cross somehow. especially for KBSM. Ugh.
+
+KBSM & LINC most similar. CLMB, BRKG, LINC, and KBSM (and less so STIL) very similar. All corr > 0.6 though.
 
 ```{r}
-metadata <- readRDS("~/Github/pvdiv-phenotypes/data/metadata.rds")
-metadata <- metadata %>%
-  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
-                            !is.na(PLOIDY) ~ as.character(PLOIDY),
-                            TRUE ~ NA_character_),
-         LATITUDE = abs(LATITUDE),
-         LONGITUDE = abs(LONGITUDE)) 
-phe_all <- read_rds("../data/all_phenotype_meas_4wcr_gwas.rds")
-sites <- readRDS("~/Github/pvdiv-phenotypes/data/sites.rds")
+phe_all <- read_rds("../data/older/all_phenotype_meas_4wcr_gwas.rds")
 phe_all$SUBPOP <- factor(phe_all$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
-sites <- sites %>%
-  mutate(manu_site = case_when(State != "TX" ~ State,
-                               SITE == "KING" ~ "TX1",
-                               SITE == "PKLE" ~ "TX2",
-                               SITE == "TMPL" ~ "TX3",
-                               SITE == "OVTN" ~ "TX4"))
-sites$manu_site <- factor(sites$manu_site, levels = c("TX1", "TX2", "TX3", "TX4", "OK", "MO", "NE", "IL", "MI", "SD"))
-
-subpops <- enframe(unique(metadata$SUBPOP), name = NULL, value = "SUBPOP") %>%
-  mutate(color_code = case_when(SUBPOP == "Atlantic" ~ "#6E91CB",
-                                SUBPOP == "Gulf" ~ "#F47F72",    
-                                SUBPOP == "Midwest" ~ "#442C83",
-                                SUBPOP == "4X" ~ "grey", 
-                                SUBPOP == "8X" ~ "#090906"))
-subpops$SUBPOP <- factor(subpops$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
 

### 2020-06-12 data-raw/Data_Setup.Rmd
Make phenotype data for GWAS and add to Data_setup.Rmd

@@
                             !is.na(PLOIDY) ~ as.character(PLOIDY),
-                            TRUE ~ NA_character_))
+                            TRUE ~ NA_character_),
+         LATITUDE = abs(LATITUDE),
+         LONGITUDE = abs(LONGITUDE))
 write_rds(metadata, "../data/metadata.rds")
-fourway <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/0-data/Phenology_2019_Multiple_Environment.csv")
 ```
@@
 ```
+# All phenotypes
+```{r}
+phe_all <- read_rds("../data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
+phenology <- read_rds("../data/Phenology_phenotypes_4wcr_and_gwas.rds")
+```
+
+```{r}
+phenology %>%
+  pivot_wider(names_from = PHE, values_from = MEAS) %>%
+  left_join(phe_all) %>%
+  dplyr::select(-CGDD_12C_5d, -PTT_12C_5d) %>%
+  pivot_longer(cols = EM1:CRAIN_5d, names_to = "PHE", values_to = "MEAS") %>%
+  filter(!is.na(MEAS)) %>%
+  group_by(PHE, WHERE, YEAR) %>%
+  tally() %>% arrange(desc(YEAR))
+
+phenology %>%
+  pivot_wider(names_from = PHE, values_from = MEAS) %>%
+  left_join(phe_all) %>%
+  dplyr::select(-CGDD_12C_5d, -PTT_12C_5d) %>%
+  pivot_longer(cols = EM1:CRAIN_5d, names_to = "PHE", values_to = "MEAS") %>%
+  filter(!is.na(MEAS)) %>%
+  write_rds("../data/Phenology_CGDD_PTT_and_rainfall_phenotypes.rds")
+```
+
+# Phenotypes for GWAS
+```{r}
+phenotypes <- 
+  read_rds("../data/Phenology_CGDD_PTT_and_rainfall_phenotypes.rds")
+
+phe_df <- phenotypes %>%
+  filter(WHERE == "GWAS" & !(SITE %in% c("OVTN", "FRMI"))) %>%
+  unite(col = "SITE_PHE", SITE, PHE) %>%
+  filter(SUBPOP %in% c("Atlantic", "Midwest", "Gulf")) %>%
+  select(-PLOT_GL, -WHERE, -YEAR) %>%
+  pivot_wider(names_from = SITE_PHE, values_from = MEAS, values_fn = mean) %>%
+  select(PLANT_ID, SUBPOP, matches("GDD"), matches("GR50"), matches("FL50"),
+         matches("CRAIN"), matches("PTT"), everything())
+
+midwgulf_df <- phe_df %>% filter(SUBPOP %in% c("Gulf", "Midwest")) 
+midwgulf_df <- midwgulf_df %>%
+  select(names(which(colSums(!is.na(midwgulf_df)) > 100))) %>%
+  select(-SUBPOP)
+midw_df <- phe_df %>% filter(SUBPOP == "Midwest") 
+midw_df <- midw_df %>%
+  select(names(which(colSums(!is.na(midw_df)) > 100))) %>%
+  select(-SUBPOP)
+gulf_df <- phe_df %>% filter(SUBPOP == "Gulf") 
+gulf_df <- gulf_df %>%
+  select(names(which(colSums(!is.na(gulf_df)) > 100))) %>%
+  select(-SUBPOP)
+atl_df <- phe_df %>% filter(SUBPOP == "Atlantic") 
+atl_df <- atl_df %>%
+  select(names(which(colSums(!is.na(atl_df)) > 100))) %>%
+  select(-SUBPOP)
+midwatl_df <- phe_df %>% filter(SUBPOP %in% c("Atlantic", "Midwest")) 
+midwatl_df <- midwatl_df %>%
+  select(names(which(colSums(!is.na(midwatl_df)) > 100))) %>%
+  select(-SUBPOP)
+gulfatl_df <- phe_df %>% filter(SUBPOP %in% c("Gulf", "Atlantic")) 
+gulfatl_df <- gulfatl_df %>%
+  select(names(which(colSums(!is.na(gulfatl_df)) > 100))) %>%
+  select(-SUBPOP)
+
+phe_subsets <- list(Midwest_and_Gulf = midwgulf_df, Midwest = midw_df, Gulf = gulf_df, Atlantic = atl_df, Midwest_and_Atlantic = midwatl_df, Gulf_and_Atlantic = gulfatl_df)
+write_rds(phe_subsets, "../data/Phenotype_Subsets_for_Subpop_GWAS.rds")
+names(phe_subsets)
+```

### 2020-06-12 R/GWAS_GDD.R

@@ -1,0 +1,20 @@
+library(bigsnpr)
+library(tidyverse)
+library(switchgrassGWAS)
+library(AnnotationDbi)
+txdb <- loadDb(file = file.path("~", "Github", "pvdiv-genome",
+                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
+
+outputdir <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", "gwas")
+
+inputfiles <- read_delim("~/Github/pvdiv-phenology-gxe/analysis/gwas/inputkinship.txt", delim = " ", col_names = "SNPfiles") # Midwest, Gulf, Atlantic, Midwest & Gulf
+
+phe_subsets <- read_rds("../data/Phenotype_Subsets_for_Subpop_GWAS.rds")
+
+for(i in 1:nrow(inputfiles)){
+  phe_df <- phe_subsets[i]
+  snp <- snp_attach(inputfiles$SNPfiles[i])
+  pvdiv_standard_gwas(snp, df = phe_df, type = "linear", outputdir = outputdir,
+                      savegwas = TRUE, saveannos = TRUE, txdb = txdb,
+                      minphe = 100)
+}

### 2020-06-12 R/GWAS_GDD.R
GWAS script for GDD phenotypes

@@ -15,6 +15,5 @@ for(i in 1:nrow(inputfiles)){
   phe_df <- phe_subsets[i]
   snp <- snp_attach(inputfiles$SNPfiles[i])
   pvdiv_standard_gwas(snp, df = phe_df, type = "linear", outputdir = outputdir,
-                      savegwas = TRUE, saveannos = TRUE, txdb = txdb,
-                      minphe = 100)
+                      savegwas = TRUE, saveannos = FALSE, minphe = 100)
 }

### 2020-06-17 R/Analyse_univarate_gwas.Rmd
GWAS with GDD results overviews for writeup

---
title: "Analyze Univariate GWAS"
author: "Alice MacQueen"
date: "6/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

#
```{r}
summary <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/gwas/GDD/Bonferroni/GWAS_summary.csv",
                    col_names = "filename")

summary %>%
  separate(col = "filename", into = c("path", "site", "phe", NA, NA, "individuals", "SNPs", NA, "PCs", NA, NA, "date_run", "time_run"), sep = "_") %>%
  separate(individuals, into = c("individuals", NA), sep = -1, convert = TRUE) %>%
  mutate(subpop = case_when(SNPs == "12.1M" ~ "Midwest_Gulf",
                            SNPs == "10.6M" ~ "Atlantic",
                            SNPs == "8.8M" ~ "Midwest",
                            SNPs == "10.3M" ~ "Gulf",
                            SNPs == "12.3M" ~ "Midwest_Atlantic",
                            SNPs == "11.5M" ~ "Atlantic_Gulf",
                       TRUE ~ NA_character_)) %>%
  write_csv("~/Github/pvdiv-phenology-gxe/analysis/gwas/GDD/Bonferroni/GWAS_summary.csv", 
            col_names = TRUE)

```

```{r}
summary1 <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/gwas/GDD/GWAS_summary_Bonferroni.csv", col_names = TRUE)
summary2 <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/gwas/GDD/Bonferroni/GWAS_summary.csv", col_names = TRUE)

summary2 %>%
  rename(Manh_time = time_run) %>%
  select(-path) %>%
  left_join(summary1) %>%
  mutate(time_run = coalesce(time_run, Manh_time),
         time_run = str_sub(time_run, end = -5)) %>%
  select(-Qqplot_order, -path, -Manh_time) %>%
  write_csv("~/Github/pvdiv-phenology-gxe/analysis/gwas/GDD/Bonferroni/GWAS_summary.csv", 
            col_names = TRUE, na = "")
```

```{r}
summary <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/gwas/GDD/Bonferroni/GWAS_summary_Bonferroni_mash_decisions.csv", col_names = TRUE)

summary %>%
  filter(grepl("GR", phe)) %>%
  group_by(subpop, site, qqplot) %>%
  add_tally(sort = FALSE) %>%
  select(indv:notes, n, everything()) %>%
  arrange(subpop, site, phe)

summary %>%
  filter(grepl("GR", phe), qqplot != "drop", qqplot != "iffy") %>%
  mutate(bonferroni = as.numeric(bonferroni)) %>%
  arrange(subpop, site, phe) %>%
  filter(qqplot == "signal" & bonferroni > 0) %>%
  group_by(subpop) %>% 
  tally()

```

```{r}
summary %>%
  filter(grepl("FL50$", phe), qqplot != "drop", qqplot != "iffy") %>%
  mutate(bonferroni = as.numeric(bonferroni)) %>%
  arrange(subpop, site, phe) %>%
  filter(qqplot == "signal" & bonferroni > 0) %>%
  group_by(subpop) %>% 
  tally()
summary %>%
  filter(grepl("FL50$", phe))
summary %>%
  filter(grepl("FL50$", phe), qqplot != "drop", qqplot != "iffy") %>%
  mutate(bonferroni = as.numeric(bonferroni)) %>%
  arrange(subpop, site, phe) %>%
  filter(qqplot == "signal" & bonferroni > 0)

summary %>%
  filter(qqplot != "drop", qqplot != "iffy", notes != "structure") %>%
  mutate(bonferroni = as.numeric(bonferroni)) %>%
  group_by(subpop, phe) %>%
  summarise(sum = sum(bonferroni, na.rm = TRUE))
```
CGDD: 37 total. 14 have signal in QQplot & hits above Bonferroni
Atlantic	1			
Atlantic_Gulf	4			
Gulf	1			
Midwest	2			
Midwest_Atlantic	2			
Midwest_Gulf	5	
PTT: 37 total. 13 have signal & hits above BOnferroni.
Atlantic	1			
Atlantic_Gulf	3			
Midwest	1			
Midwest_Atlantic	3			
Midwest_Gulf	5	
CRAIN: 17 of 37.
Atlantic	2			
Atlantic_Gulf	4			
Gulf	3			
Midwest	1			
Midwest_Atlantic	3			
Midwest_Gulf	4	
CRAIN.5d: 16 of 37. 
Atlantic	3			
Atlantic_Gulf	3			
Gulf	2			
Midwest	1			
Midwest_Atlantic	4			
Midwest_Gulf	3

Very few GR have signal in QQplot & hits above Bonferroni. 
GR1 13 of 40. GR50 12 of 40. GR100 5 of 38.
Atlantic	3			
Atlantic_Gulf	5			
Gulf	7			
Midwest 0 
Midwest_Atlantic	7			
Midwest_Gulf	8

EM has more: 33 out of 70. EM1 16 of 35.  EM50 17 of 35.
Atlantic	6			
Atlantic_Gulf	6			
Gulf	5			
Midwest	1			
Midwest_Atlantic	7			
Midwest_Gulf	8

FL also has more: 33 out of 76. FL1 17 of 39. FL50 16 of 37.
Atlantic	4			
Atlantic_Gulf	10			
Gulf	4			
Midwest	2			
Midwest_Atlantic	7			
Midwest_Gulf	6

### 2020-06-19 R/Draft_models_and_stats_2020-06-19.Rmd
Start writing fourway GDD section

---
title: "Draft_models_and_stats_2020-06-19"
author: "Alice MacQueen"
date: "6/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(switchgrassGWAS)
library(cowplot)
library(corpcor)
library(viridis)
```

# ASReml-R analyses on server

```{r}
library(tidyverse)
library(asreml)
library(switchgrassGWAS)
library(cowplot)
library(viridis)
library(corpcor)
```

# Load data

```{r}
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "manuscript", "plots")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe <- read_rds(file.path(workingdir, "data",
                          "Phenology_CGDD_PTT_and_rainfall_phenotypes.rds"))
```

Unfortunately, no parents flowered exclusively during shortening days at the three Texas sites; WBC flowered during shortening days in the majority years in the Texas common gardens.
```{r}
phe %>%
  filter(WHERE == "FWCR" & PLANT_ID %in% c("AP13", "WBC", "DAC", "VS16", "VxW", "AxD")) %>%
  filter(PHE == "FL50" & SITE %in% c("PKLE", "TMPL", "KING", "OVTN", "STIL")) %>%
  group_by(PLANT_ID, SITE, MEAS >= 173) %>%
  tally() 
```
AxD KING 0 / 41
PKLE 0 / 39
TMPL 0/39

VxW 
PKLE 2 / 43
TMPL: 21 / 41
KING: 0 / 43

AP13:
TMPL: 56 / 66 >= 173
PKLE: 52 / 67 
KING: 0 / 40

WBC: 
TMPL: 40/40
PKLE: 39 / 46
KING: 10 / 18

DAC: 0% of 15
VS16: 3.6% of 56

```{r}
56/66
52/67
0
(56+52)/(66+67+40)

100
39/46
10/18
(40+39+10)/(40+46+18)
2/56
23/(43+41+43)
41+39+39
```

### 2020-06-30 R/make_mash_objects_FL50_and_GDD.Rmd
Mash analysis on FL50 and GDD and Li's methods

---
title: "Mash analysis FL50 and GDD"
author: "Alice MacQueen"
date: "6/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(switchgrassGWAS)
# need these three for the last code chunk only: visualizing mash results
library(GGally)
library(cowplot)
library(viridis)
```

# Run bigsnp2mashr

Need to do this on the server, that's where I have saved these GWAS_datatables.

Midwest subpop:
CGDD_12C, FL50 x3

```{r}
trim_name <- function(x){
    for(i in seq_along(x)){
  if((str_sub(x[i], start = -1) %in% c("_")))
  { x[i] <- str_sub(x[i], end = -2) }
    }
  return(x)
}

path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                       "gwas", "GDD")

inputfiles <- read_delim(file.path("~", "Github", "pvdiv-phenology-gxe", 
                                   "analysis", "gwas", "inputkinship.txt"), 
                         delim = " ", col_names = "SNPfiles") 
# Midwest & Gulf, Midwest, Gulf, Atlantic, Midwest & Atlantic, Gulf & Atlantic

# Midwest
snp <- snp_attach(inputfiles$SNPfiles[2])
gwas_rds <- pvdiv_results_in_folder(path = path, 
                                    pattern = paste0("GWAS_datatable_",
                                                     "(KBSM|PKLE|CLMB)", ".+",
                                                     "(CGDD_12C|FL50)", 
                                                     ".+",
                                                     "\\d+g_8.8M"))

site_phe <- trim_name(str_sub(gwas_rds, start = 16, end = -41))
numSNPs <- ceiling((1000000 / length(site_phe)^2)/1000)*1000

midwest_mashin <- 
  pvdiv_bigsnp2mashr(path = path, snp = snp,
                    gwas_rds = gwas_rds, phenotypes = site_phe, 
                    numSNPs = numSNPs, scale = TRUE, suffix = "midwest", 
                    saveoutput = TRUE)
# Gulf [3]
snp <- snp_attach(inputfiles$SNPfiles[3])
gwas_rds <- pvdiv_results_in_folder(path = path, 
                                    pattern = paste0("GWAS_datatable_", ".+",
                                                     "(CGDD_12C|FL50)", 
                                                     ".+",
                                                     "\\d+g_10.3M"))

site_phe <- trim_name(str_sub(gwas_rds, start = 16, end = -42))
numSNPs <- ceiling((1000000 / length(site_phe)^2)/1000)*1000

midwest_mashin <- 
  pvdiv_bigsnp2mashr(path = path, snp = snp,
                    gwas_rds = gwas_rds, phenotypes = site_phe, 
                    numSNPs = numSNPs, scale = TRUE, suffix = "gulf", 
                    saveoutput = TRUE)

# Midwest_Gulf [1]
snp <- snp_attach(inputfiles$SNPfiles[1])
gwas_rds <- pvdiv_results_in_folder(path = path, 
                                    pattern = paste0("GWAS_datatable_", ".+",
                                                     "(CGDD_12C|FL50)", 
                                                     ".+",
                                                     "\\d+g_12.1M"))

site_phe <- trim_name(str_sub(gwas_rds, start = 16, end = -42))
numSNPs <- ceiling((1000000 / length(site_phe)^2)/1000)*1000

midwest_mashin <- 
  pvdiv_bigsnp2mashr(path = path, snp = snp,
                    gwas_rds = gwas_rds, phenotypes = site_phe, 
                    numSNPs = numSNPs, scale = TRUE, suffix = "midwest_gulf", 
                    saveoutput = TRUE)

# Atlantic [4]
snp <- snp_attach(inputfiles$SNPfiles[4])
gwas_rds <- pvdiv_results_in_folder(path = path, 
                                    pattern = paste0("GWAS_datatable_",
                                                     "(KBSM|PKLE|CLMB|LINC|TMPL)", 
                                                     ".+", "(CGDD_12C|FL50)", 
                                                     ".+", "\\d+g_10.6M"))
# drop BRKG for the Atlantic subpopulation because it had obvious population structure.

site_phe <- trim_name(str_sub(gwas_rds, start = 16, end = -42))
numSNPs <- ceiling((1000000 / length(site_phe)^2)/1000)*1000

atlantic_mashin <- 
  pvdiv_bigsnp2mashr(path = path, snp = snp,
                    gwas_rds = gwas_rds, phenotypes = site_phe, 
                    numSNPs = numSNPs, scale = TRUE, suffix = "atlantic", 
                    saveoutput = TRUE)

# other things to try... CRAIN... etc.

```

# mash runs
```{r}
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                       "gwas", "GDD")
phe_suffix <- c("midwest", "gulf", "midwest_gulf", "atlantic")
numSNPs <- c(28000, 10000, 4000, 10000)

m_out <- list()
for(i in 2){
  m_out[[i]] <- mash_standard_run(path = path, numSNPs = numSNPs[i], 
                                  suffix = phe_suffix[i], saveoutput = TRUE)
}
```

```{r}

path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                       "gwas", "GDD")
phe_suffix <- c("midwest", "gulf", "midwest_gulf", "atlantic")
numSNPs <- c(28000, 10000, 4000, 10000)

i = 4
m <- readRDS(file.path(path, paste0("Strong_Effects", numSNPs[i], "SNPs_", 
                                    phe_suffix[i], ".rds")))
U_ed <- readRDS(file.path(path, paste0("Mash_data_driven_covariances_",
                                       numSNPs[i], "_", phe_suffix[i], ".rds")))

mash_plot_manhattan_by_condition(m)
mash_plot_sig_by_condition(m)
mash_plot_pairwise_sharing(m)
mash_plot_effects(m, n = 1)
mash_plot_effects(m, n = 2)

get_significant_results(m)[1:10]
barplot(switchgrassGWAS:::get_estimated_pi(m))

ggcorr(data = NULL, cor_matrix = m$fitted_g$Ulist$`Stand_BhatCLMB_CGDD_12C-mean`)
#ggcorr(data = NULL, cor_matrix = m$fitted_g$Ulist$ED_PCA_2)
thing <- U_ed$ED_PCA_1
ggcorr(data = NULL, cor_matrix = thing, limits = c(min(thing)-.1, max(thing)+.1))

```

Midwest mostly follows covariance matrix ED_PCA_1 and 3.
Gulf ED_PCA_2, 4, and 9.
Midwest_Gulf ED_PCA_1 through 4.

Atlantic hardly any significant effects for GDD/FL50. 10. 2N, 5K, 5N, 7K, 9K, 9N.
ED_PCA_1.

Try CRAIN and other stuff? FL50 & GDD alone? Hmm.

### 2020-07-08 R/Make_mash_U_hyp.Rmd
Make U_hyp object for gulf subpop and use in mash

---
title: "Make_mash_U_hyp"
author: "Alice MacQueen"
date: 2020-07-07
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(geosphere)
library(spdep)
library(rlist)
sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
sitedf <- as.data.frame(sites)
coordinates(sitedf) <- c("Longitude", "Latitude")
```

# Covariance matrices for mash

Want to make these based on the inverse of the distance between sites, and also 
on the covariances

```{r}
sitenb <- tri2nb(sitedf, row.names = sitedf$manu_site)
summary(sitenb)
wm <- nb2mat(sitenb, style = 'W')
ww <- nb2listw(sitenb, style = 'W')
moran(sitedf$Elevation, ww, n=length(ww$neighbours), S0=Szero(ww), NAOK = TRUE)
moran.test(sitedf$Elevation, ww, randomisation=FALSE, na.action = na.omit)
moran.mc(sitedf$Elevation, ww, nsim=499)

wmw <- nb2mat(sitenb, style = 'W')
colnames(wmw) <- rownames(wmw)

U1 <- as_tibble(wmw, .name_repair = "unique") %>%
    mutate(rowU = sitedf$manu_site) %>%
    pivot_longer(cols = -.data$rowU, names_to = "colU",
                 values_to = "covar") %>%
    filter(!is.na(.data$covar))

  U1 %>%
    ggplot(aes(x = .data$rowU, y = .data$colU)) +
    CDBNgenomics::theme_oeco +
    geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
    scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                    "#27AD81FF", "#5DC863FF", "#FDE725FF")) +
    geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
    theme(legend.position = "right",
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    xlab("") + ylab("")
```

## Inverse distances between sites, in 100km
```{r}
distsites <- distm(sitedf)
colnames(distsites) <- sites$SITE
rownames(distsites) <- sites$SITE
W <- 1 / (distsites/100000) # 100km distance measure
W[is.infinite(W)] <- NA
diag(W) <- max(W, na.rm = TRUE) + min(W, na.rm = TRUE)
U_dist <- W
U1 <- as_tibble(W, .name_repair = "unique") %>%
    mutate(rowU = rownames(W)) %>%
    pivot_longer(cols = -.data$rowU, names_to = "colU",
                 values_to = "covar") %>%
    filter(!is.na(.data$covar))

  U1 %>%
    ggplot(aes(x = .data$rowU, y = .data$colU)) +
    CDBNgenomics::theme_oeco +
    geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
    scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                    "#27AD81FF", "#5DC863FF", "#FDE725FF")) +
    geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
    theme(legend.position = "right",
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    xlab("") + ylab("")
```

## Weather covariances in the time period between greenup and flowering, plus in the 5d before flowering
```{r}
phe_wea <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")

weadf_cov <- phe_wea %>%
  pivot_longer(cols = GR50:CRAIN_5d, names_to = "phe", values_to = "MEAS") %>%
  unite(col = "site_phe", SITE, phe, sep = ".") %>%
  pivot_wider(id_cols = PLANT_ID, names_from = site_phe, values_from = MEAS, values_fn = mean)

weadf_cov2 <- cov(weadf_cov[-1], use = "pairwise")
colnames(weadf_cov2)

pattern <- c("GR50", "FL50", "CGDD_12C$", "CGDD_12C_5d", "PTT_12C$", 
             "PTT_12C_5d", "CRAIN$", "CRAIN_5d" )
Ureturn <- list()

for(i in seq_along(pattern)){
  U_wea <- weadf_cov2 %>%
    as_tibble(rownames = "rowU") %>%
    dplyr::select(rowU, matches(pattern[i])) %>%
    filter(grepl(pattern[i], rowU)) %>%
    separate(rowU, into = c("site", "phe"), sep = 4) %>%
    dplyr::select(-phe) %>%
    column_to_rownames(var = "site") %>%
    as.data.frame()
  colnames(U_wea) <- str_sub(colnames(U_wea), end = 4)

  Ureturn <- list.append(Ureturn, U_wea = U_wea)
  names(Ureturn)[i] <- paste0("U_", str_extract(pattern[i], "[^$]+")) # matches one or more characters other than $
}
names(Ureturn)
Ureturn <- list.append(Ureturn, U_dist = U_dist)
```

# Use these U_hyp in mash_standard_run....
# mash runs

Midwest: CLMB, KBSM, PKLE
Gulf: CLMB KBSM KING PKLE TMPL
Midwest_Gulf: BRKG CLMB KBSM KING LINC PKLE STIL TMPL
Atlantic: CLMB KBSM LINC PKLE TMPL
always in alphabetical order, so that's good.
Also, each location is repeated twice. Once for CGDD and once for FL50. So covar matrices should be duplicated also? Ehhhh might also consider running these separately first! Honestly. And join them in a mash analysis later if I want to compare them. But just duplicate them now to test??

```{r}
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                  "gwas", "GDD")
phe_suffix <- c("midwest", "gulf", "midwest_gulf", "atlantic")
numSNPs <- c(28000, 10000, 4000, 10000)

gulf_loc <- c("CLMB", "KBSM", "KING", "PKLE", "TMPL")
midw_loc <- c("CLMB", "KBSM", "PKLE")
atl_loc <- c("CLMB", "KBSM", "LINC", "PKLE", "TMPL")
midwgulf_loc <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL",
                  "TMPL")

Ugulf <- list()
i = 1 
for(i in seq_along(names(Ureturn))){
  Utest <- Ureturn[[i]][which(rownames(Ureturn[[i]]) %in% gulf_loc),
                 which(colnames(Ureturn[[i]]) %in% gulf_loc)]

  Urow <- rep(Utest, each = 2)
  Utr <- matrix(unlist(Urow), nrow = nrow(Utest), ncol = ncol(Utest)*2)
  Ut <- rep(Utr, each = 2)
  Udup <- matrix(unlist(Ut), nrow = nrow(Utr)*2, ncol = ncol(Utr))
  colnames(Udup) <- rep(gulf_loc, each = 2)
  rownames(Udup) <- rep(gulf_loc, each = 2)
  Ugulf <- list.append(Ugulf, Udup = Udup)
  names(Ugulf)[i] <- names(Ureturn)[i]
}

write_rds(Ugulf, path = file.path(path, "U_hyp_gulf.rds"))
```

```{r}
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                  "gwas", "GDD")
phe_suffix <- c("midwest", "gulf", "midwest_gulf", "atlantic")
numSNPs <- c(28000, 10000, 4000, 10000)

gulf_loc <- c("CLMB", "KBSM", "KING", "PKLE", "TMPL")
midw_loc <- c("CLMB", "KBSM", "PKLE")
atl_loc <- c("CLMB", "KBSM", "LINC", "PKLE", "TMPL")
midwgulf_loc <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL",
                  "TMPL")
U_gulf <- read_rds(file.path(path, "U_hyp_gulf.rds"))
i = 2
  m_out <- mash_standard_run(path = path, numSNPs = numSNPs[i], U_hyp = U_gulf,
                             suffix = paste0(phe_suffix[i], "_U_hyp_v1"),
                             saveoutput = TRUE)
```

## Look at old mash_standard_runs

Need to rerun these without using ref = 'mean'. Then I'll be able to look at the covariance matrices.
```{r}
library(switchgrassGWAS)

path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                  "gwas", "GDD")
phe_suffix <- c( "atlantic", "gulf", "midwest", "midwest_gulf")
numSNPs <- c(10000, 10000, 28000, 4000)

gulf_loc <- c("CLMB", "KBSM", "KING", "PKLE", "TMPL")
midw_loc <- c("CLMB", "KBSM", "PKLE")
atl_loc <- c("CLMB", "KBSM", "LINC", "PKLE", "TMPL")
midwgulf_loc <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL",
                  "TMPL")
mash_rds <- pvdiv_results_in_folder(path = path, pattern = "^Strong_Effects")
mash_suffix <- str_sub(mash_rds, start = 24, end = -5) %>%
  str_replace(., "^_", "")

i = 2
for(i in seq_along(mash_rds)){
  m <- readRDS(file.path(path, mash_rds[i]))
  suffix <- mash_suffix[i]
  p1 <- mash_plot_manhattan_by_condition(m, saveoutput = TRUE, suffix = suffix)
  p2 <- mash_plot_sig_by_condition(m, saveoutput = TRUE, suffix = suffix)
  p3 <- mash_plot_pairwise_sharing(m, saveoutput = TRUE, suffix = suffix)
  p4 <- mash_plot_effects(m, n = 1, saveoutput = TRUE, suffix = suffix)
  p5 <- mash_plot_effects(m, n = 2, saveoutput = TRUE, suffix = suffix)
  p6 <- mash_plot_covar(m, saveoutput = TRUE, suffix = suffix)
  p7 <- mash_plot_Ulist(m, range = get_U_by_mass(m), saveoutput = TRUE, 
                  suffix = suffix)
}
```

### 2020-07-09 R/Make_mash_U_hyp.Rmd
Make symmetric U_hyp matrices

@@
 write_rds(Ugulf, path = file.path(path, "U_hyp_gulf.rds"))
+
+Ugulf$U_GR50
+U_gulf_sym <- list()
+i = 2
+for(i in seq_along(names(Ugulf))){
+  Usym <- Ugulf[[i]]
+  Usym1 <- Usym
+  Usym2 <- Usym
+  for(j in nrow(Usym1)){
+    for(k in ncol(Usym1)){
+      if(j < k){
+        Usym1[j, k] <- Usym[j, k]
+        Usym2[j, k] <- Usym[k, j]
+      } else {
+        Usym1[j, k] <- Usym[k, j]
+        Usym2[j, k] <- Usym[j, k]
+      }
+    }
+  }
+  U_gulf_sym <- list.append(U_gulf_sym, U1 = Usym1, U2 = Usym2)
+  length(names(U_gulf_sym))
+  names(U_gulf_sym)[length(names(U_gulf_sym))-1] <- paste0(names(Ugulf)[i], "_1")
+  names(U_gulf_sym)[length(names(U_gulf_sym))] <- paste0(names(Ugulf)[i], "_2")
+}
+
+write_rds(U_gulf_sym, path = file.path(path, "U_hyp_gulf_symmetric.rds"))
+
 ```
@@
   m_out <- mash_standard_run(path = path, numSNPs = numSNPs[i], U_hyp = U_gulf,
-                             suffix = paste0(phe_suffix[i], "_U_hyp_v1"),
+                             suffix = phe_suffix[i],
                              saveoutput = TRUE)

### 2020-07-09 R/Make_mash_U_hyp.Rmd
Fix symmetry of U_dist U_hyp

--- 2020-07-08 R/Make_mash_U_hyp.Rmd
@@
 names(Ureturn)
+
+Ureturn$U_dist <- Ureturn$U_dist %>% as.data.frame()
 Ureturn <- list.append(Ureturn, U_dist = U_dist)
+
+str(Ureturn)
 ```
@@
 }
+str(Ugulf)
+write_rds(Ugulf, path = file.path(path, "U_hyp_gulf.rds"))
 
-write_rds(Ugulf, path = file.path(path, "U_hyp_gulf.rds"))
+Ugulf$U_GR50
+U_gulf_sym <- list()
+isSymmetric(Ugulf$U_dist)
+i = 2
+for(i in seq_along(names(Ugulf))){
+  Usym <- Ugulf[[i]]
+  Usym1 <- Usym
+  Usym2 <- Usym
+  for(j in nrow(Usym1)){
+    for(k in ncol(Usym1)){
+      if(j < k){
+        Usym1[j, k] <- Usym[j, k]
+        Usym2[j, k] <- Usym[k, j]
+      } else {
+        Usym1[j, k] <- Usym[k, j]
+        Usym2[j, k] <- Usym[j, k]
+      }
+    }
+  }
+  U_gulf_sym <- list.append(U_gulf_sym, U1 = Usym1, U2 = Usym2)
+  length(names(U_gulf_sym))
+  names(U_gulf_sym)[length(names(U_gulf_sym))-1] <- paste0(names(Ugulf)[i], "_1")
+  names(U_gulf_sym)[length(names(U_gulf_sym))] <- paste0(names(Ugulf)[i], "_2")
+}
+
+write_rds(U_gulf_sym, path = file.path(path, "U_hyp_gulf_symmetric.rds"))
+
 ```
@@
                   "TMPL")
-U_gulf <- read_rds(file.path(path, "U_hyp_gulf.rds"))
+U_gulf <- read_rds(file.path(path, "U_hyp_gulf_symmetric.rds"))
 i = 2
   m_out <- mash_standard_run(path = path, numSNPs = numSNPs[i], U_hyp = U_gulf,
-                             suffix = paste0(phe_suffix[i], "_U_hyp_v1"),
+                             suffix = phe_suffix[i],
                              saveoutput = TRUE)

### 2020-07-09 R/Make_mash_U_hyp.Rmd
Test smaller Uhyp matrices for midwest and gulf subpops

@@
                   "TMPL")
-U_gulf <- read_rds(file.path(path, "U_hyp_gulf_symmetric.rds"))
+U_gulf <- read_rds(file.path(path, "U_hyp_gulf.rds"))
 i = 2
-  m_out <- mash_standard_run(path = path, numSNPs = numSNPs[i], U_hyp = U_gulf,
-                             suffix = phe_suffix[i],
+U_ed <- file.path(path, "Mash_data_driven_covariances_10000_gulf.rds")
+  m_out <- mash_standard_run(path = path, numSNPs = numSNPs[i], U_ed = U_ed,
+                             U_hyp = U_gulf, suffix = phe_suffix[i],
                              saveoutput = TRUE)
@@
 }
+load_all("~/Github/switchgrassGWAS")
 ```
+
+# Test mash with U_hyp
+
+I got lots of errors the first time I ran mash to completion with U_hyp, and I am not yet certain why.
+I'd like to run mash on just FL50 in the gulf, and GDD in the midwest maybe, to see if I can diagnose what's going on or what the covariance matrices should look like.
+
+## Rerun bigsnp2mashr
+
+```{r}
+trim_name <- function(x){
+    for(i in seq_along(x)){
+  if((str_sub(x[i], start = -1) %in% c("_")))
+  { x[i] <- str_sub(x[i], end = -2) }
+    }
+  return(x)
+}
+
+pathin <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
+                       "gwas", "GDD")
+path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
+                       "gwas", "GDD", "Uhyp")
+
+inputfiles <- read_delim(file.path("~", "Github", "pvdiv-phenology-gxe", 
+                                   "analysis", "gwas", "inputkinship.txt"), 
+                         delim = " ", col_names = "SNPfiles") 
+# Midwest & Gulf, Midwest, Gulf, Atlantic, Midwest & Atlantic, Gulf & Atlantic
+
+# Midwest
+snp <- snp_attach(inputfiles$SNPfiles[2])
+gwas_rds <- pvdiv_results_in_folder(path = pathin, 
+                                    pattern = paste0("GWAS_datatable_",
+                                                     "(KBSM|PKLE|CLMB)", ".+",
+                                                     "(CGDD_12C)", 
+                                                     ".+",
+                                                     "\\d+g_8.8M"))
+
+site_phe <- trim_name(str_sub(gwas_rds, start = 16, end = -41))
+numSNPs <- 20000 #ceiling((1000000 / length(site_phe)^2)/1000)*1000
+
+midwest_mashin <- 
+  pvdiv_bigsnp2mashr(path = pathin, snp = snp,
+                    gwas_rds = gwas_rds, phenotypes = site_phe, 
+                    numSNPs = numSNPs, scale = FALSE, 
+                    suffix = "midwest_uhyp1", saveoutput = TRUE)
+# Gulf [3]
+snp <- snp_attach(inputfiles$SNPfiles[3])
+gwas_rds <- pvdiv_results_in_folder(path = pathin, 
+                                    pattern = paste0("GWAS_datatable_", ".+",
+                                                     "(FL50)", 
+                                                     ".+",
+                                                     "\\d+g_10.3M"))
+
+site_phe <- trim_name(str_sub(gwas_rds, start = 16, end = -42))
+numSNPs <- 30000 #ceiling((1000000 / length(site_phe)^2)/1000)*1000
+
+gulf_mashin <- 
+  pvdiv_bigsnp2mashr(path = pathin, snp = snp,
+                    gwas_rds = gwas_rds, phenotypes = site_phe, 
+                    numSNPs = numSNPs, scale = FALSE, suffix = "gulf_uhyp1", 
+                    saveoutput = TRUE)
+```
+
+# Uhyp for midwest and gulf
+
+```{r}
+gulf_loc <- c("CLMB", "KBSM", "KING", "PKLE", "TMPL")
+midw_loc <- c("CLMB", "KBSM", "PKLE")
+
+Ugulfhyp <- list()
+Umidwhyp <- list()
+
+for(i in seq_along(names(Ureturn))){
+  Ugulf1 <- Ureturn[[i]][which(rownames(Ureturn[[i]]) %in% gulf_loc),
+                 which(colnames(Ureturn[[i]]) %in% gulf_loc)]
+  Umidw1 <- Ureturn[[i]][which(rownames(Ureturn[[i]]) %in% midw_loc),
+                 which(colnames(Ureturn[[i]]) %in% midw_loc)]
+
+  Ugulfhyp <- list.append(Ugulfhyp, Ugulf1 = Ugulf1)
+  Umidwhyp <- list.append(Umidwhyp, Umidw1 = Umidw1)
+  names(Ugulfhyp)[i] <- names(Ureturn)[i]
+  names(Umidwhyp)[i] <- names(Ureturn)[i]
+}
+str(Umidwhyp)
+write_rds(Ugulfhyp, path = file.path(path, "U_hyp_gulf_1phe.rds"))
+write_rds(Umidwhyp, path = file.path(path, "U_hyp_midw_1phe.rds"))
+```
+
+```{r}
+path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
+                       "gwas", "GDD", "Uhyp")
+numSNPs <- 20000
+suffix = "midwest_uhyp1"
+Umidwhyp <- read_rds(path = file.path(path, "U_hyp_midw_1phe.rds"))
+
+m <- mash_standard_run(path = path, numSNPs = numSNPs, U_hyp = Umidwhyp, 
+                       suffix = suffix, saveoutput = TRUE)
+```

### 2020-07-09 R/Make_mash_U_hyp.Rmd
specify Uhyp as matrices

@@
 
-  Ugulfhyp <- list.append(Ugulfhyp, Ugulf1 = Ugulf1)
-  Umidwhyp <- list.append(Umidwhyp, Umidw1 = Umidw1)
+  Ugulfhyp <- list.append(Ugulfhyp, Ugulf1 = as.matrix(Ugulf1))
+  Umidwhyp <- list.append(Umidwhyp, Umidw1 = as.matrix(Umidw1))
   names(Ugulfhyp)[i] <- names(Ureturn)[i]
@@
 str(Umidwhyp)
+is.matrix(Umidwhyp$U_GR50)
+isSymmetric(Umidwhyp$U_GR50)
 write_rds(Ugulfhyp, path = file.path(path, "U_hyp_gulf_1phe.rds"))
@@
 Umidwhyp <- read_rds(path = file.path(path, "U_hyp_midw_1phe.rds"))
+Udist <- Umidwhyp$U_dist
+
+m <- mash_standard_run(path = path, numSNPs = numSNPs, U_hyp = Udist, 
+                       suffix = suffix, saveoutput = TRUE)
 
@@
 ```
+Error in mashr::mash_set_data(Bhat_random, Shat_random) :
+  Shat cannot contain NaN/Inf values
+  problem for midwest...
+
+U_hyp must be matrices. They must be symmetric, in the same order as Bhat etc, and with the same # rows and cols as Bhat etc.
+
+```{r}
+path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
+                       "gwas", "GDD", "Uhyp")
+numSNPs <- 30000
+suffix = "gulf_uhyp1"
+Ugulfhyp <- read_rds(path = file.path(path, "U_hyp_gulf_1phe.rds"))
+Udistg <- Ugulfhyp$U_dist
+U_h1 <- list(Udist = as.matrix(Udistg))
+isSymmetric(U_h1$Udist)
+typeof(Udistg)
+
+U_ed <- file.path(path, "Mash_data_driven_covariances_30000_gulf_uhyp1.rds")
+
+m <- mash_standard_run(path = path, numSNPs = numSNPs, U_ed = U_ed,
+                       U_hyp = U_h1, suffix = suffix, saveoutput = TRUE)
+```
+
+```{r}
+path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
+                       "gwas", "GDD", "Uhyp")
+m <- readRDS(file.path(path, "Strong_Effects30000SNPs_gulf_uhyp1.rds"))
+suffix = "gulf_uhyp1"
+mash_plot_covar(m, saveoutput = TRUE, suffix = suffix)
+mash_plot_Ulist(m, range = get_U_by_mass(m), saveoutput = TRUE, 
+                  suffix = suffix)
+mash_plot_sig_by_condition(m, saveoutput = TRUE, suffix = suffix)
+mash_plot_pairwise_sharing(m, saveoutput = TRUE, suffix = suffix)
+mash_plot_effects(m, n = 1, saveoutput = TRUE, suffix = suffix)
+mash_plot_effects(m, n = 2, saveoutput = TRUE, suffix = suffix)
+mash_plot_effects(m, n = 8)
+
+mash_plot_manhattan_by_condition(m, saveoutput = TRUE, suffix = suffix)
+```

### 2020-07-09 R/Make_mash_U_hyp.Rmd
Try rescaling all FL50 gulf effects to fall on same after-solstice scale

@@
 
-U_hyp must be matrices. They must be symmetric, in the same order as Bhat etc, and with the same # rows and cols as Bhat etc.
+U_hyp must be numeric matrices. They must be symmetric, square, in the same order as Bhat etc, and with the same # rows and cols as Bhat etc.
+
+They can't contain NaN, NA, Inf values. And it might help if they're positive definite.
+
+That's all the info in the mash package that I can find.
+
+I suppose that I want them to be on the same scale as the mash input data, not the same scale as whatever weather or distance data they came from.
+This is a sense in which correlation matrices would be better.
+Take a look at the U_ed.
+Rescale to a similar scale as ED_PCA_1.
+But then...
 
@@
 
+U_h2 <- list(Udist = Ugulfhyp$U_dist, UGDD = Ugulfhyp$U_CGDD_12C)
+
 U_ed <- file.path(path, "Mash_data_driven_covariances_30000_gulf_uhyp1.rds")
@@
                        U_hyp = U_h1, suffix = suffix, saveoutput = TRUE)
+m <- mash_standard_run(path = path, numSNPs = numSNPs, U_ed = U_ed,
+                       U_hyp = U_h2, suffix = suffix, saveoutput = TRUE)
 ```
@@
 mash_plot_covar(m, saveoutput = TRUE, suffix = suffix)
-mash_plot_Ulist(m, range = get_U_by_mass(m), saveoutput = TRUE, 
-                  suffix = suffix)
+mash_plot_Ulist(m, range = get_U_by_mass(m))
 mash_plot_sig_by_condition(m, saveoutput = TRUE, suffix = suffix)
@@
 ```
+
+```{r}
+load_all("~/Github/switchgrassGWAS")
+fitted_g <- list(Ulist = u_ed)
+m_mock <- list(fitted_g = fitted_g)
+mash_plot_Ulist(m_mock, limits = FALSE)
+
+bhat %>% as_tibble() %>%
+  colMeans()
+summary(bhat)
+
+# CLMB_FL50 KBSM_FL50 KING_FL50 PKLE_FL50 TMPL_FL50 
+240.7987 - 173  
+261.9135 - 173   
+209.8265 - 173  
+210.2195 - 173  
+240.0100 - 173
+
+B_hat_strong_df_30000topSNPs_gulf_solstice <- bhat %>%
+  mutate(Bhat_CLMB_FL50 = Bhat_CLMB_FL50 + (240.7987 - 173),
+         Bhat_KING_FL50 = Bhat_KING_FL50 + (209.8265 - 173 ),
+         Bhat_KBSM_FL50 = Bhat_KBSM_FL50 + (261.9135 - 173),
+         Bhat_PKLE_FL50 = Bhat_PKLE_FL50 + (210.2195 - 173),
+         Bhat_TMPL_FL50 = Bhat_TMPL_FL50 + (240.0100 - 173),
+         ) #%>%
+  #summary()
+write_rds(B_hat_strong_df_30000topSNPs_gulf_solstice, file.path(path, "B_hat_strong_df_30000topSNPs_gulf_solstice.rds"))
+
+B_hat_random_df_30000topSNPs_gulf_solstice <- B_hat_random_df_30000topSNPs_gulf_uhyp1 %>%
+  mutate(Bhat_CLMB_FL50 = Bhat_CLMB_FL50 + (240.7987 - 173),
+         Bhat_KING_FL50 = Bhat_KING_FL50 + (209.8265 - 173 ),
+         Bhat_KBSM_FL50 = Bhat_KBSM_FL50 + (261.9135 - 173),
+         Bhat_PKLE_FL50 = Bhat_PKLE_FL50 + (210.2195 - 173),
+         Bhat_TMPL_FL50 = Bhat_TMPL_FL50 + (240.0100 - 173),
+         ) 
+write_rds(B_hat_random_df_30000topSNPs_gulf_solstice, file.path(path, "B_hat_random_df_30000topSNPs_gulf_solstice.rds"))
+```
+what is each site's mean FL50? Just for Gulf plants.
+
+For interpretability of the mash results... may want to scale all these effects so that 0 is, say, the summer solstice? Instead of the mean flowering time at that site for that pop.
+```{r}
+phe_gulf <- phenotypes$Gulf %>%
+  select(PLANT_ID, ends_with("FL50"))
+
+sd(phe_gulf$CLMB_FL50, na.rm = TRUE)
+sd(phe_gulf$KBSM_FL50, na.rm = TRUE)
+sd(phe_gulf$KING_FL50, na.rm = TRUE)
+sd(phe_gulf$PKLE_FL50, na.rm = TRUE)
+sd(phe_gulf$TMPL_FL50, na.rm = TRUE)
+
+summary(phe_gulf$CLMB_FL50, na.rm = TRUE)
+summary(phe_gulf$KBSM_FL50, na.rm = TRUE)
+summary(phe_gulf$KING_FL50, na.rm = TRUE)
+summary(phe_gulf$PKLE_FL50, na.rm = TRUE)
+summary(phe_gulf$TMPL_FL50, na.rm = TRUE)
+summary(phe_gulf)
+
+```
+CLMB_FL50 KBSM_FL50 KING_FL50 PKLE_FL50 TMPL_FL50 
+240.7987   261.9135  209.8265  210.2195 240.0100
+
+241 +/- 18.6
+262 +/- 10.5
+210 +/- 34
+210 +/- 35
+240 +/- 28

### 2020-07-14 R/dyln_lag_plots.Rmd
Plotting daylength lag by site

---
title: "DYLN lag plots"
author: "Alice MacQueen"
date: "7/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(lubridate)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

#

```{r}
wea_df <- readRDS("~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")
metadata <- readRDS("~/Github/pvdiv-phenology-gxe/data/metadata.rds")
sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
phenotypes <- readRDS("~/Github/pvdiv-phenology-gxe/data/Phenology_CGDD_PTT_and_rainfall_phenotypes.rds")
jday_phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
```

```{r}
jday_phe <- wea_df %>%
  group_by(SITE) %>%
  mutate(dyln_change_sec = (DYLN - lag(DYLN))*60*60,
         FL50 = yday(DOY)) %>%
  select(DOY, DYLN, dyln_change_sec, everything(), -PTT_12C) %>%
  right_join(jday_phe, by = c("SITE", "FL50"))

jday_phe <- jday_phe %>%
  select(-(TMAX:CPTT_11C))

sites <- wea_df %>%
  group_by(SITE) %>%
  filter(DOY == "2019-06-22") %>%
  select(SITE, DYLN) %>%
  rename(Longest_dyln = DYLN) %>%
  right_join(sites)

sites <- wea_df %>%
  group_by(SITE) %>%
  filter(DOY <= "2019-06-22") %>%
  summarise(CGDD_solstice = sum(GDD_12C_SM)) %>%
  right_join(sites)
```

```{r}
jday_phe <- metadata %>%
  rename(Lat_origin = LATITUDE, Long_origin = LONGITUDE) %>%
  select(PLANT_ID, SUBPOP, Lat_origin, Long_origin) %>%
  right_join(jday_phe)
jday_phe <- jday_phe %>%
  left_join(sites)

jday_phe$SUBPOP <- factor(jday_phe$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
jday_phe$SITE <- factor(jday_phe$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
jday_phe$manu_site <- factor(jday_phe$manu_site, levels = rev(c("TX1", "TX2", "TX3", "TX4", "OK", "MO", "NE", "IL", "MI", "SD")))
```

```{r}
color_code <- c("grey", "#6E91CB","#F47F72", "#442C83", "#090906", "grey")
col1 <- jday_phe %>%
  filter(!(SITE %in% c("OVTN", "FRMI")) & !is.na(SUBPOP)) %>%
  ggplot(aes(x = dyln_change_sec)) + 
  geom_density(aes(group = SUBPOP, color = SUBPOP), size = 0.75) + 
  facet_wrap(~manu_site, scales = "free", nrow = 8) + 
  geom_vline(xintercept = 0, linetype = 2) +
  theme(legend.position = "none", axis.text = element_text(size = 8)) +
  scale_color_manual(values = color_code) + 
  xlim(c(-200, 150)) +
  xlab("Daylength change (s) at 
       flowering") + 
  ylab("Density")
save_plot("../manuscript/plots/Dyln_change_by_site.png", plot = col1, base_height = 8, base_asp = 0.3)

col2 <- jday_phe %>%
  filter(!(SITE %in% c("OVTN", "FRMI")) & !is.na(SUBPOP)) %>%
  ggplot(aes(x = DYLN)) + 
  geom_density(aes(group = SUBPOP, color = SUBPOP), size = 0.75) + 
  facet_wrap(~manu_site, scales = "free_y", nrow = 8) + 
  geom_vline(data = filter(sites, !(SITE %in% c("OVTN", "FRMI"))),
             aes(xintercept = Longest_dyln), linetype = 2) +
  theme(legend.position = "none", axis.text = element_text(size = 8)) +
  scale_color_manual(values = color_code) + 
  xlim(c(10, 16)) +
  xlab("Daylength (hrs) at 
       flowering") + 
  ylab("Density")
save_plot("../manuscript/plots/Dyln_by_site.png", plot = col2, base_height = 8, base_asp = 0.3)

col3 <- jday_phe %>%
  filter(!(SITE %in% c("OVTN", "FRMI")) & !is.na(SUBPOP)) %>%
  ggplot(aes(x = CRAIN_5d)) + 
  geom_density(aes(group = SUBPOP, color = SUBPOP), size = 0.75) + 
  facet_wrap(~manu_site, scales = "free_y", nrow = 8) + 
  theme(legend.position = "none", axis.text = element_text(size = 8)) +
  scale_color_manual(values = color_code) + 
  xlab("Rainfall (mm) five days
       before flowering") + 
  ylab("Density")
save_plot("../manuscript/plots/CRAIN_5d_by_site.png", plot = col3, base_height = 8, base_asp = 0.3)

col4 <- jday_phe %>%
  filter(!(SITE %in% c("OVTN", "FRMI")) & !is.na(SUBPOP)) %>%
  ggplot(aes(x = CGDD_12C)) + 
  geom_density(aes(group = SUBPOP, color = SUBPOP), size = 0.75) + 
  facet_wrap(~manu_site, scales = "free_y", nrow = 8) + 
  theme(legend.position = "none", axis.text = element_text(size = 8)) +
  scale_color_manual(values = color_code) + 
  xlab("Cumulative GDD 
       (12C) at flowering") + 
  ylab("Density")
save_plot("../manuscript/plots/CGDD_by_site.png", plot = col4, base_height = 8, base_asp = 0.3)

col5 <- jday_phe %>%
  filter(!(SITE %in% c("OVTN", "FRMI")) & !is.na(SUBPOP)) %>%
  ggplot(aes(x = as_date(FL50, origin = "2019-01-01"))) + 
  geom_density(aes(group = SUBPOP, color = SUBPOP), size = 0.75) + 
  geom_vline(xintercept = as_date(173, origin = "2019-01-01"), linetype = 2) +
  facet_wrap(~manu_site, scales = "free_y", nrow = 8) + 
  theme(legend.position = "none", axis.text = element_text(size = 8)) +
  scale_color_manual(values = color_code) + 
  xlab("Flowering 
       Date") + 
  ylab("Density")
save_plot("../manuscript/plots/FL50_by_site.png", plot = col5, base_height = 8, base_asp = 0.3)

FL50_1 <- plot_grid(col5, col2, col1, col4, col3, nrow = 1)
save_plot("../manuscript/plots/Five_FL50_weather_cues_by_site.png", plot = FL50_1, base_height = 8, base_width = 9.85)
```

Same plots but histograms for just Gulf and Midwest
consign all the other groups to the supplement
```{r}
hist_phe <- jday_phe %>%
  filter(!(SITE %in% c("OVTN", "FRMI"))  & SUBPOP %in% c("Gulf", "Midwest"))
color_code <- c("#F47F72", "#442C83")
col1 <- hist_phe %>%
  ggplot(aes(x = dyln_change_sec)) + 
  geom_histogram(aes(group = SUBPOP, fill = SUBPOP), binwidth = 10) + 
  facet_wrap(~manu_site, scales = "free_y", nrow = 8) + 
  geom_vline(xintercept = 0, linetype = 2) +
  theme(legend.position = "none", axis.text = element_text(size = 8)) +
  scale_fill_manual(values = color_code) + 
  xlim(c(-200, 105)) +
  xlab("Daylength change (s) at 
       flowering") + 
  ylab("")
save_plot("../manuscript/plots/Dyln_change_by_site_histogram.png", plot = col1, base_height = 7.5, base_asp = 0.27)

col2 <- hist_phe %>%
  ggplot(aes(x = DYLN)) + 
  geom_histogram(aes(group = SUBPOP, fill = SUBPOP), binwidth = 0.15) + 
  facet_wrap(~manu_site, scales = "free_y", nrow = 8) + 
  geom_vline(data = filter(sites, !(SITE %in% c("OVTN", "FRMI"))),
             aes(xintercept = Longest_dyln), linetype = 2) +
  theme(legend.position = "none", axis.text = element_text(size = 8)) +
  scale_fill_manual(values = color_code) + 
  xlim(c(11, 15.55)) +
  xlab("Daylength (hrs) at 
       flowering") + 
  ylab("")
save_plot("../manuscript/plots/Dyln_by_site_histogram.png", plot = col2, base_height = 7.5, base_asp = 0.27)

col3 <- hist_phe %>%
  ggplot(aes(x = CRAIN_5d)) + 
  geom_histogram(aes(group = SUBPOP, fill = SUBPOP), binwidth = 5) + 
  facet_wrap(~manu_site, scales = "free_y", nrow = 8) + 
  theme(legend.position = "none", axis.text = element_text(size = 8)) +
  scale_fill_manual(values = color_code) + 
  xlab("Rainfall (mm) five days
       before flowering") + 
  ylab("")
save_plot("../manuscript/plots/CRAIN_5d_by_site_histogram.png", plot = col3, base_height = 7.5, base_asp = 0.27)

col4 <- hist_phe %>%
  ggplot(aes(x = CGDD_12C)) + 
  geom_histogram(aes(group = SUBPOP, fill = SUBPOP), binwidth = 100) + 
  geom_vline(data = filter(sites, !(SITE %in% c("OVTN", "FRMI"))),
             aes(xintercept = CGDD_solstice), linetype = 2) +
  facet_wrap(~manu_site, scales = "free_y", nrow = 8) + 
  theme(legend.position = "none", axis.text = element_text(size = 8)) +
  scale_fill_manual(values = color_code) + 
  xlab("Cumulative GDD 
       (12C) at flowering") + 
  ylab("")
save_plot("../manuscript/plots/CGDD_by_site_histogram.png", plot = col4, base_height = 7.5, base_asp = 0.27)

col5 <- hist_phe %>%
  ggplot(aes(x = as_date(FL50, origin = "2019-01-01"))) + 
  geom_histogram(aes(group = SUBPOP, fill = SUBPOP), binwidth = 8) + 
  geom_vline(xintercept = as_date(173, origin = "2019-01-01"), linetype = 2) +
  facet_wrap(~manu_site, scales = "free_y", nrow = 8) + 
  theme(legend.position = "none", axis.text = element_text(size = 8)) +
  scale_fill_manual(values = color_code) + 
  xlab("Flowering 
       Date") + 
  ylab("Number of Individuals")
save_plot("../manuscript/plots/FL50_by_site_histogram.png", plot = col5, base_height = 7.5, base_asp = 0.27)

FL50_1 <- plot_grid(col5, col2, col1, col4, col3, nrow = 1)
save_plot("../manuscript/plots/Five_FL50_weather_cues_by_site_histogram.png", plot = FL50_1, base_height = 7.5, base_asp = 1.3)
```

also dyln at flowering
also dyln change as a function of latitude of origin
also dyln as a function of lat of origin?

```{r}
jday_phe <- jday_phe %>%
  mutate(Lat_origin_1 = round(Lat_origin))

jday_phe %>%
  filter(!(SITE %in% c("OVTN", "FRMI")) & !is.na(SUBPOP)) %>%
  ggplot(aes(x = dyln_change_sec, y = Lat_origin)) + 
  geom_point(aes(group = SUBPOP, color = SUBPOP), alpha = 0.75) + 
  facet_grid(manu_site~SUBPOP, scales = "free") + geom_vline(xintercept = 0, linetype = 2) +
  theme(legend.position = "right") +
  scale_color_manual(values = color_code) + xlim(c(-200, 150)) +
  xlab("Daylength change on day of flowering (s)")

jday_phe %>%
  filter(!(SITE %in% c("OVTN", "FRMI")) & !is.na(SUBPOP)) %>%
  ggplot(aes(x = DYLN, y = Lat_origin)) + 
  geom_point(aes(group = manu_site, color = manu_site), alpha = 0.4) + 
  facet_wrap(~SUBPOP, scales = "free_y", nrow = 2) + geom_vline(xintercept = 0, linetype = 2) +
  theme(legend.position = "right") +
  scale_color_viridis(discrete = TRUE, end = 0.95) + xlim(c(10.5, 16.5))
  xlab("Daylength on day of flowering")

  jday_phe %>%
  filter((manu_site %in% c("TX1", "TX2", "TX3")) & SUBPOP %in% c("Gulf", "Midwest")) %>%
  ggplot(aes(x = dyln_change_sec, y = Lat_origin_1)) + 
  geom_boxplot(aes(group = Lat_origin_1), orientation = "y") + 
    geom_point(aes(y = Lat_origin, color = SUBPOP), alpha = 0.3) +
  facet_wrap(~manu_site, scales = "free") + geom_vline(xintercept = 0, linetype = 2) +
  theme(legend.position = "right") +
  scale_color_manual(values = color_code[3:4]) + 
  xlab("Daylength change on day of flowering (s)")

  ?geom_boxplot
```

### 2020-07-17 R/Plots_manuscript_v0.2.Rmd
Add GR50 and FL50 specific functions of weather for ASReml h2 analysis

---
title: "Analyses v0.2"
author: "Alice MacQueen"
date: "7/17/2020"
output: html_document
---

Add any libraries/scripts that need to be loaded for these analyses here.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(lubridate)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

# Load data
Add any data that needs to be loaded for these analyses here.
```{r}
wea_df <- readRDS("~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")
metadata <- readRDS("~/Github/pvdiv-phenology-gxe/data/metadata.rds")
sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
phenotypes <- readRDS("~/Github/pvdiv-phenology-gxe/data/Phenology_CGDD_PTT_and_rainfall_phenotypes.rds")
jday_phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
```
```{r}
jday_phe <- wea_df %>%
  group_by(SITE) %>%
  mutate(dyln_change_sec = (DYLN - lag(DYLN))*60*60,
         FL50 = yday(DOY)) %>%
  select(DOY, DYLN, dyln_change_sec, everything(), -PTT_12C) %>%
  right_join(jday_phe, by = c("SITE", "FL50")) %>%
  select(-(TMAX:CPTT_11C))
jday_phe <- metadata %>%
  rename(Lat_origin = LATITUDE, Long_origin = LONGITUDE) %>%
  select(PLANT_ID, SUBPOP, Lat_origin, Long_origin) %>%
  right_join(jday_phe) %>%
  left_join(sites)

jday_phe$SUBPOP <- factor(jday_phe$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
jday_phe$SITE <- factor(jday_phe$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
jday_phe$manu_site <- factor(jday_phe$manu_site, levels = rev(c("TX1", "TX2", "TX3", "TX4", "OK", "MO", "NE", "IL", "MI", "SD")))
```

# Overview
Here I'm putting the code for any new analyses that I'm running to do new writeup for the manuscript v0.2.

# Greenup h2
Greenup as a function of Julian date had extremely low heritability within subpopulations (0.005 +/- 0.003 Gulf; 0.021 +/0 0.011 for Midwest) across all eight common gardens, but high heritability across both subpopulations (h2 = 0.987 +/- 0.006).

<did any environmental cue beat GR50 to estimate greenup?>.

## define environmental cues for greenup
Probably all CGDD related.
```{r}

asreml_fl <- jday_phe %>%
  select(PLANT_ID:Long_origin, PLOT_GL, SITE, FL50, GR50, CGDD_12C, CRAIN, CRAIN_5d, DYLN, dyln_change_sec) %>%
  rename(dyln_fl50 = DYLN, cgdd_12c_gr2fl = CGDD_12C, crain_gr2fl = CRAIN) %>%
  left_join(wea_df) %>%
  group_by(SITE, PLOT_GL) %>%
  filter(between(DOY, as_date(FL50, origin = "2019-01-01") - 6, as_date(FL50, origin = "2019-01-01"))) %>%
  mutate(crain_2d = RAIN_SM + lag(RAIN_SM),
         crain_3d = RAIN_SM + lag(RAIN_SM) + lag(RAIN_SM, n = 2),
         crain_7d = sum(RAIN_SM)) %>%
  filter(DOY == as_date(FL50, origin = "2019-01-01")) %>%
  filter(SUBPOP %in% c("Midwest", "Gulf", "Atlantic") & !(SITE %in% c("OVTN"))) %>%
  select(-(TMAX:TMIN_SM), -(GDD_13C_SM:CPTT_13C)) %>%
  rename(crain_1d = RAIN_SM, crain_5d = CRAIN_5d)
saveRDS(asreml_fl, file = "../data/Weather_related_phe_for_asreml_h2_FL50.rds")
asreml_gr <- jday_phe %>%
  select(PLANT_ID:Long_origin, PLOT_GL, SITE, GR50) %>%
  left_join(wea_df) %>%
  group_by(SITE, PLOT_GL) %>%
  filter(DOY <= as_date(GR50, origin = "2019-01-01")) %>%
  mutate(cgdd_12c_jan2gr = sum(GDD_12C_SM),
         cgdd_8c_jan2gr = sum(GDD_8C_SM)
         ) %>%
  select(-(GDD_13C_SM:CPTT_13C), -(TMAX:Longitude), GDD_12C_SM) %>%
  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 17, as_date(GR50, origin = "2019-01-01"))) %>%
  mutate(tmax_18d = mean(TMAX_SM),
         tmin_18d = mean(TMIN_SM),
         tave_18d = (tmax_18d + tmin_18d)/2,
         cgdd_12c_18d = mean(GDD_12C_SM)) %>%
  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 9, as_date(GR50, origin = "2019-01-01"))) %>%
  mutate(tmax_10d = mean(TMAX_SM),
         tmin_10d = mean(TMIN_SM),
         tave_10d = (tmax_10d + tmin_10d)/2,
         cgdd_12c_10d = mean(GDD_12C_SM)) %>%
  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 4, as_date(GR50, origin = "2019-01-01"))) %>%
  mutate(tmax_5d = mean(TMAX_SM),
         tmin_5d = mean(TMIN_SM),
         tave_5d = (tmax_5d + tmin_5d)/2,
         cgdd_12c_5d = mean(GDD_12C_SM))%>%
  filter(DOY == as_date(GR50, origin = "2019-01-01")) %>%
  filter(SUBPOP %in% c("Midwest", "Gulf", "Atlantic") & !(SITE %in% c("OVTN"))) %>%
  select(-(TMAX_SM:RAIN_SM), -Day_of_year)
saveRDS(asreml_gr, file = "../data/Weather_related_phe_for_asreml_h2_GR50.rds")
```

## ASReml h2 greenup analysis
```{r}
library(tidyverse)
library(asreml)
library(switchgrassGWAS)
library(cowplot)
```

```{r}
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "heritability")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")

```

### loop to make heritability table using ASReml mixed models and vpredict.
```{r}
site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
               NE = "NE", IL = "IL", MI = "MI", SD = "SD", eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), nine_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD"))
subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))
phe_fl <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl",
           "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d",
           "crain_7d")
phe_gr <- c("GR50", "DYLN", "cgdd_12c_jan2gr", "cgdd_8c_jan2gr", "tmax_18d", 
            "tmin_18d", "tave_18d", "cgdd_12c_18d", "tmax_10d", "tmin_10d",
            "tave_10d", "cgdd_12c_10d", "tmax_5d", "tmin_5d", "tave_5d",
            "cgdd_12c_5d")
phe_dfs <- list(phe_gr_df, phe_fl_df)
phe_lists <- list(phe_gr, phe_fl)
h2_table <- tibble()

for(p in seq_along(phe_dfs)){
  phe_v <- phe_lists[[p]]
  phe <- phe_dfs[[p]]

  for(i in seq_along(site_v)){
    for(j in seq_along(subpop_v)){
      for(k in seq_along(phe_v)){
        phe_single <- phe %>%
    left_join(sites) %>%
    filter(manu_site %in% site_v[[i]] & SUBPOP %in% subpop_v[[j]]) %>%
    rename(phenotype = eval(phe_v[k])) %>%
    mutate(PLANT_ID = as_factor(PLANT_ID),
           SITE = as_factor(SITE),
           PLOT_GL = as_factor(PLOT_GL)) %>%
    as.data.frame()

  if(length(site_v[[i]]) > 1){  # need to add site as a random factor if 
    # there is more than one site included in the dataset.
     asr_out <- asreml(phenotype ~ 1,
                       random = ~vm(PLANT_ID, k_full) +
                         ~idv(SITE),
                       residual = ~idv(units),
                       data = phe_single, 
                       workspace = "3gb")

    h2_est <- vpredict(asr_out, h2 ~ V2/(V1+V2+V3)) %>% as_tibble() %>%
      mutate(site = site_v[i],
             subpop = subpop_v[j],
             phe = phe_v[k],
             bic = summary(asr_out)$bic[1],
             loglik = summary(asr_out)$loglik)
  } else {
    asr_out <- asreml(phenotype ~ 1,
                       random = ~vm(PLANT_ID, k_full),
                       residual = ~idv(units),
                       data = phe_single, 
                       workspace = "3gb")

    h2_est <- vpredict(asr_out, h2 ~ V1/(V1+V2)) %>% as_tibble() %>%
      mutate(site = names(site_v)[i],
             subpop = names(subpop_v)[j],
             phe = phe_v[k],
             bic = summary(asr_out)$bic[1],
             loglik = summary(asr_out)$loglik)
  }

  if(i == 1 & j == 1 & k == 1){
    h2_table <- h2_est
  } else {
    h2_table <- add_row(h2_table, h2_est)
  }
  tryCatch({
      save_plot(filename=file.path(outputdir, paste0("ASReml_performance_",
                                                 names(site_v)[i], "_",
                                                 names(subpop_v)[j],
                                                 "_", phe_v[k], ".png")), 
            plot = plot(asr_out))
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})

      }
    }
  }
  write_csv(h2_table, file.path(outputdir, paste0("h2_df_", phe_v[1],
                                                  "_related_", 
                                                  "functions_of_weather.csv")))
}

```

# Make new environmental variables for Li to use?
Can I estimate heritability for the fourway cross?

### 2020-07-17 R/dyln_lag_plots.Rmd
Add GR50 and FL50 specific functions of weather for ASReml h2 analysis

@@
   right_join(sites)
+
+asreml_h2 <- jday_phe %>%
+  select(PLANT_ID:SITE, FL50:GR50, CGDD_12C, CRAIN, CRAIN_5d, -DOY) %>%
+  rename(dyln_fl50 = DYLN, cgdd_12c_gr2fl = CGDD_12C, crain_gr2fl = CRAIN) %>%
+  left_join(wea_df) %>%
+  group_by(SITE, PLOT_GL) %>%
+  filter(between(DOY, as_date(FL50, origin = "2019-01-01") - 6, as_date(FL50, origin = "2019-01-01"))) %>%
+  mutate(crain_2d = RAIN_SM + lag(RAIN_SM),
+         crain_3d = RAIN_SM + lag(RAIN_SM) + lag(RAIN_SM, n = 2),
+         crain_7d = sum(RAIN_SM)) %>%
+  filter(DOY == as_date(FL50, origin = "2019-01-01")) %>%
+  filter(SUBPOP %in% c("Midwest", "Gulf") & !(SITE %in% c("OVTN", "FRMI")))
+asreml_h2 <- asreml_h2 %>%
+  select(-(TMAX:CPTT_13C))
+saveRDS(asreml_h2, file = "../data/Weather_related_phe_for_asreml_h2.rds")
 ```
@@
   ggplot(aes(x = CRAIN_5d)) + 
-  geom_histogram(aes(group = SUBPOP, fill = SUBPOP), binwidth = 5) + 
+  geom_histogram(aes(group = SUBPOP, fill = SUBPOP), binwidth = 10) + 
   facet_wrap(~manu_site, scales = "free_y", nrow = 8) + 
@@
 
+col7 <- hist_phe %>%
+  ggplot(aes(x = as_date(GR50, origin = "2019-01-01"))) + 
+  geom_histogram(aes(group = SUBPOP, fill = SUBPOP), binwidth = 8) + 
+  geom_vline(xintercept = as_date(173, origin = "2019-01-01"), linetype = 2) +
+  facet_wrap(~manu_site, scales = "free_y", nrow = 8) + 
+  theme(legend.position = "none", axis.text = element_text(size = 8)) +
+  scale_fill_manual(values = color_code) + 
+  xlab("Greenup 
+       Date") + 
+  ylab("Number of Individuals")
+save_plot("../manuscript/plots/GR50_by_site_histogram.png", plot = col7, base_height = 7.5, base_asp = 0.27)
+
+col6 <- hist_phe %>%
+  ggplot(aes(x = CRAIN)) + 
+  geom_histogram(aes(group = SUBPOP, fill = SUBPOP), binwidth = 50) + 
+  facet_wrap(~manu_site, scales = "free_y", nrow = 8) + 
+  theme(legend.position = "none", axis.text = element_text(size = 8)) +
+  scale_fill_manual(values = color_code) + 
+  xlab("Rainfall (mm) greenup
+       to flowering") + 
+  ylab("")
+save_plot("../manuscript/plots/CRAIN_by_site_histogram.png", plot = col6, base_height = 7.5, base_asp = 0.27)
+
 FL50_1 <- plot_grid(col5, col2, col1, col4, col3, nrow = 1)
 save_plot("../manuscript/plots/Five_FL50_weather_cues_by_site_histogram.png", plot = FL50_1, base_height = 7.5, base_asp = 1.3)
+
+FL50_2 <- plot_grid(col5, col2, col1, col4, nrow = 1)
+save_plot("../manuscript/plots/Four_FL50_weather_cues_by_site_histogram.png", plot = FL50_2, base_height = 7.5, base_asp = 1)
+
+FL50_3 <- plot_grid(col7, col5, col6, col3, nrow = 1)
+save_plot("../manuscript/plots/GR50_FL50_CRAIN_by_site_histogram.png", plot = FL50_3, base_height = 7.5, base_asp = 1)
+```
+## ASReml: Heritabilities for these phenotypes, defined by particular environmental cues.
+The reasoning being that if a phenotype defined by a cue has higher heritability, it's more likely that that cue is playing a role in flowering.
+I could just introduce all of these traits this way and talk about them this way, instead of introducing and defeating hypotheses.
+That might also be a softer start.
+
+# ASReml-R analyses on server
+
+```{r}
+library(tidyverse)
+library(asreml)
+library(switchgrassGWAS)
+library(cowplot)
+library(viridis)
+library(corpcor)
+```
+
+```{r}
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "manuscript", "plots")
+metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
+sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
+subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
+phe <- read_rds(file.path(workingdir, "data",
+                          "Weather_related_phe_for_asreml_h2.rds"))
+```
+
+```{r}
+k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
+#k_mid <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_method_138_individuals_midwest_genotype_and_admixed.rds")
+#k_gulf <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_method_214_individuals_gulf_genotype_and_admixed.rds")
+```
+
+```{r}
+site_v <- c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD")
+subpop_v <- c("Gulf", "Midwest")
+phe_v <- c("dyln_fl50", "dyln_change_sec", "FL50", "GR50", "cgdd_12c_gr2fl",
+           "crain_gr2fl", "CRAIN_5d", "crain_2d", "crain_3d", "crain_7d")
+i = 1
+j = 1
+k = 1
+
+phe_single$SUBPOP <- phe_single(dyln_cue$SUBPOP, levels = c("Midwest", "Gulf",
+                                                            "Atlantic", "4X",
+                                                            "8X"))
+h2_table <- tibble()
+
+for(i in 4:8){
+  for(j in seq_along(subpop_v)){
+    for(k in seq_along(phe_v)){
+      phe_single <- phe %>%
+  left_join(sites) %>%
+  filter(manu_site == site_v[i] & SUBPOP %in% subpop_v[j]) %>%
+  rename(phenotype = eval(phe_v[k])) %>%
+  mutate(PLANT_ID = as_factor(PLANT_ID),
+         SITE = as_factor(SITE),
+         PLOT_GL = as_factor(PLOT_GL)) %>%
+  as.data.frame()
+
+asr_out <- asreml(phenotype ~ 1,
+                   random = ~vm(PLANT_ID, k_full),
+                   residual = ~idv(units),
+                   data = phe_single, 
+                   workspace = "3gb")
+#phe_v[k]
+#summary(asr_out)$varcomp
+#summary(asr_out)$bic
+h2_est <- vpredict(asr_out, h2 ~ V1/(V1+V2)) %>% as_tibble() %>%
+  mutate(site = site_v[i],
+         subpop = subpop_v[j],
+         phe = phe_v[k],
+         bic = summary(asr_out)$bic[1],
+         loglik = summary(asr_out)$loglik)
+
+if(i == 1 & j == 1 & k == 1){
+  h2_table <- add_row(h2_est)
+} else {
+  h2_table <- add_row(h2_table, h2_est)
+}
+tryCatch({
+    save_plot(filename=file.path(outputdir, paste0("ASReml_performance_",
+                                               site_v[i], "_", subpop_v[j],
+                                               "_", phe_v[k], ".png")), 
+          plot = plot(asr_out))
+  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+
+    }
+  }
+}
+
+```
+factors:
+PLANT_ID, SUBPOP (if both Gulf & Midwest are included), SITE/manu_site (if multiple sites are included)
+covariates:
+Lat_origin
+
+```{r}
+site_v <- c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD")
+subpop_v <- c("Gulf", "Midwest")
+phe_v <- c("dyln_fl50", "dyln_change_sec", "FL50", "GR50", "cgdd_12c_gr2fl",
+           "crain_gr2fl", "CRAIN_5d", "crain_2d", "crain_3d", "crain_7d")
+i = 1
+k = 1
+
+phe_single$SUBPOP <- phe_single(dyln_cue$SUBPOP, levels = c("Midwest", "Gulf",
+                                                            "Atlantic", "4X",
+                                                            "8X"))
+h2_table <- tibble()
+
+for(i in seq_along(site_v)){
+    for(k in seq_along(phe_v)){
+phe_single <- phe %>%
+  left_join(sites) %>%
+  filter(manu_site == site_v[i] & SUBPOP %in% subpop_v) %>%
+  rename(phenotype = eval(phe_v[k])) %>%
+  mutate(PLANT_ID = as_factor(PLANT_ID),
+         SITE = as_factor(SITE),
+         PLOT_GL = as_factor(PLOT_GL)) %>%
+  as.data.frame()
+
+asr_out <- asreml(phenotype ~ 1,
+                   random = ~vm(PLANT_ID, k_full),
+                   residual = ~idv(units),
+                   data = phe_single, 
+                   workspace = "3gb")
+#phe_v[k]
+#summary(asr_out)$varcomp
+#summary(asr_out)$bic
+h2_est <- vpredict(asr_out, h2 ~ V1/(V1+V2)) %>% as_tibble() %>%
+  mutate(site = site_v[i],
+         subpop = "Gulf_and_Midwest",
+         phe = phe_v[k],
+         bic = summary(asr_out)$bic[1],
+         loglik = summary(asr_out)$loglik)
+
+#if(i == 1 & j == 1 & k == 1){
+#  h2_table <- add_row(h2_est)
+#} else {
+  h2_table <- add_row(h2_table, h2_est)
+#}
+tryCatch({
+    save_plot(filename=file.path(outputdir, paste0("ASReml_performance_",
+                                               site_v[i], "_Gulf_and_Midwest_",
+                                               phe_v[k], ".png")), 
+          plot = plot(asr_out))
+  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+
+    }
+  }
+
+# all sites for subpops separately
+for(j in seq_along(subpop_v)){
+  for(k in seq_along(phe_v)){
+  phe_single <- phe %>%
+    left_join(sites) %>%
+    filter(manu_site %in% site_v & SUBPOP %in% subpop_v[j]) %>%
+    rename(phenotype = eval(phe_v[k])) %>%
+    mutate(PLANT_ID = as_factor(PLANT_ID),
+           SITE = as_factor(SITE),
+           PLOT_GL = as_factor(PLOT_GL)) %>%
+    as.data.frame()
+
+  asr_out <- asreml(phenotype ~ 1,
+                     random = ~vm(PLANT_ID, k_full) +
+                       ~idv(SITE),
+                     residual = ~idv(units),
+                     data = phe_single, 
+                     workspace = "3gb")
+  #phe_v[k]
+  #summary(asr_out)$varcomp
+  #summary(asr_out)$bic
+  h2_est <- vpredict(asr_out, h2 ~ V2/(V1+V2+V3)) %>% as_tibble() %>%
+    mutate(site = site_v[i],
+           subpop = subpop_v[j],
+           phe = phe_v[k],
+           bic = summary(asr_out)$bic[1],
+           loglik = summary(asr_out)$loglik)
+
+#if(i == 1 & j == 1 & k == 1){
+#  h2_table <- add_row(h2_est)
+#} else {
+  h2_table <- add_row(h2_table, h2_est)
+#}
+  tryCatch({
+      save_plot(filename=file.path(outputdir, paste0("ASReml_performance_",
+                                                 "eight_sites_", subpop_v[j],
+                                                 "_", phe_v[k], ".png")), 
+            plot = plot(asr_out))
+    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+
+  }
+}
+
+# all sites for both midwest & gulf
+for(k in seq_along(phe_v)){
+      phe_single <- phe %>%
+  left_join(sites) %>%
+  filter(manu_site %in% site_v & SUBPOP %in% subpop_v) %>%
+  rename(phenotype = eval(phe_v[k])) %>%
+  mutate(PLANT_ID = as_factor(PLANT_ID),
+         SITE = as_factor(SITE),
+         PLOT_GL = as_factor(PLOT_GL)) %>%
+  as.data.frame()
+
+asr_out <- asreml(phenotype ~ 1,
+                   random = ~vm(PLANT_ID, k_full),
+                   residual = ~idv(units),
+                   data = phe_single, 
+                   workspace = "3gb")
+#phe_v[k]
+#summary(asr_out)$varcomp
+#summary(asr_out)$bic
+h2_est <- vpredict(asr_out, h2 ~ V2/(V1+V2+V3)) %>% as_tibble() %>%
+  mutate(site = site_v[i],
+         subpop = subpop_v[j],
+         phe = phe_v[k],
+         bic = summary(asr_out)$bic[1],
+         loglik = summary(asr_out)$loglik)
+
+if(i == 1 & j == 1 & k == 1){
+  h2_table <- add_row(h2_est)
+} else {
+  h2_table <- add_row(h2_table, h2_est)
+}
+tryCatch({
+    save_plot(filename=file.path(outputdir, paste0("ASReml_performance_",
+                                               "eight_sites_Gulf_and_Midwest",
+                                               "_", phe_v[k], ".png")), 
+          plot = plot(asr_out))
+  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+
+}
 ```
@@
 ```
+
+# Plot ASReml h2 on map
+load data from previous section also (line 268; ASReml-R analyses on server)
+```{r}
+h2_table <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/heritability/h2_table_single_site_and_eight_site_midwest_and_gulf_subpops_FL50.csv")
+library(maps)
+library(ggmap)
+usa <- map_data("usa")
+```
+
+```{r}
+h2plot <- h2_table %>%
+  left_join(sites, by = c("site" = "manu_site"))
+```
+
+```{r}
+h2plot %>%
+  filter(subpop == "Gulf_and_Midwest" & phe %in% c("FL50", "crain_gr2fl", "dyln_change_sec", "cgdd_12c_gr2fl", "dyln_fl50", "crain_2d", "CRAIN_5d")) %>%
+  ggplot() + 
+  geom_polygon(data = usa, aes(x=long, y = lat, group = group), fill = "white", 
+               color = "black") +
+  coord_quickmap(xlim = c(-105, -66), ylim = c(25, 50)) +
+  geom_jitter(aes(x = Longitude, y = Latitude, 
+                                  color = Estimate, shape = phe),
+              width = 1, height = 1) + 
+  labs(x = "Longitude", y = "Latitude") + 
+  theme(legend.position = "right") +
+  scale_color_viridis(option = "B", end = 0.95) +
+  scale_shape_manual(values = c(21,22,24,20,3,4,5)) 
+  #scale_fill_manual(values = c("#280B54FF","#FAC127FF", "#D44842FF", "white")) +
+  #scale_color_manual(values = c(rep("black", 10), "#280B54FF","#FAC127FF", "#D44842FF", "white", "#280B54FF","#FAC127FF", "#D44842FF", "white"), guide = FALSE) +
+  #scale_alpha_manual(values = c(0.7, 0.7, 0.7, 0.7), guide = FALSE)
+save_plot(filename = "Map_subpops_and_Ecotypes.svg", plot = last_plot(), base_height = 5)
+
+```

### 2020-07-18 R/Plots_manuscript_v0.2.Rmd
phenotype df for gwas for G_BLUP GR & FL

@@
 library(lubridate)
+library(switchgrassGWAS)
 source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
@@
 
-  for(i in seq_along(site_v)){
+  for(i in 6:9){
     for(j in seq_along(subpop_v)){
@@
         phe_single <- phe %>%
-    left_join(sites) %>%
+    left_join(sites, by = "SITE") %>%
     filter(manu_site %in% site_v[[i]] & SUBPOP %in% subpop_v[[j]]) %>%
@@
     # there is more than one site included in the dataset.
+    tryCatch({
      asr_out <- asreml(phenotype ~ 1,
@@
     h2_est <- vpredict(asr_out, h2 ~ V2/(V1+V2+V3)) %>% as_tibble() %>%
-      mutate(site = site_v[i],
-             subpop = subpop_v[j],
+      mutate(site = names(site_v)[i],
+             subpop = names(subpop_v)[j],
              phe = phe_v[k],
@@
              loglik = summary(asr_out)$loglik)
+    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
   } else {
+    tryCatch({
     asr_out <- asreml(phenotype ~ 1,
@@
              loglik = summary(asr_out)$loglik)
+  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
   }
@@
                                                   "_related_", 
-                                                  "functions_of_weather.csv")))
+                                                  "functions_of_weather_v2.csv")))
 }
@@
 
+## BLUPs for G
+
+We next evaluated genetic effects of greenup as functions of GDD and minimum temperature 10 days prior to greenup, and genetic effects of flowering as functions of daylength at flowering and GDD between greenup and flowering.
+To do this, we calculated BLUPs for each individual using mixed models with common garden and a kinship matrix as random effects.
+We then conducted genome-wide association on these BLUPs both within and between genetic subpopulations.
+```{r}
+site_v <- list(eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
+subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"))
+phe_fl <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl")
+phe_gr <- c("GR50", "tmin_10d", "cgdd_12c_10d")
+phe_dfs <- list(phe_gr_df, phe_fl_df)
+phe_lists <- list(phe_gr, phe_fl)
+h2_table <- tibble()
+
+i = 1
+j = 1
+k = 1
+p = 2
+
+for(p in seq_along(phe_dfs)){
+  phe_v <- phe_lists[[p]]
+  phe <- phe_dfs[[p]]
+
+  for(i in seq_along(site_v)){
+    for(j in seq_along(subpop_v)){
+      for(k in seq_along(phe_v)){
+        phe_single <- phe %>%
+    left_join(sites, by = "SITE") %>%
+    filter(manu_site %in% site_v[[i]] & SUBPOP %in% subpop_v[[j]]) %>%
+    rename(phenotype = eval(phe_v[k])) %>%
+    mutate(PLANT_ID = as_factor(PLANT_ID),
+           SITE = as_factor(SITE),
+           PLOT_GL = as_factor(PLOT_GL)) %>%
+    as.data.frame()
+
+  if(length(site_v[[i]]) > 1){  # need to add site as a random factor if 
+    # there is more than one site included in the dataset.
+    tryCatch({
+     asr_out <- asreml(phenotype ~ 1,
+                       random = ~vm(PLANT_ID, k_full) +
+                         ~idv(manu_site),
+                       residual = ~idv(units),
+                       data = phe_single, 
+                       workspace = "3gb")
+
+    h2_est <- vpredict(asr_out, h2 ~ V2/(V1+V2+V3)) %>% as_tibble() %>%
+      mutate(site = names(site_v)[i],
+             subpop = names(subpop_v)[j],
+             phe = phe_v[k],
+             bic = summary(asr_out)$bic[1],
+             loglik = summary(asr_out)$loglik)
+    blup <- summary(asr_out, coef=TRUE)$coef.random %>% as_tibble(rownames = "Effect")
+    write_csv(blup, file.path(outputdir, paste0("blups_", names(site_v)[i], "_",
+                                                 names(subpop_v)[j],
+                                                 "_", phe_v[k], ".csv")))
+    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+  } else {
+    tryCatch({
+    asr_out <- asreml(phenotype ~ 1,
+                       random = ~vm(PLANT_ID, k_full),
+                       residual = ~idv(units),
+                       data = phe_single, 
+                       workspace = "3gb")
+
+    h2_est <- vpredict(asr_out, h2 ~ V1/(V1+V2)) %>% as_tibble() %>%
+      mutate(site = names(site_v)[i],
+             subpop = names(subpop_v)[j],
+             phe = phe_v[k],
+             bic = summary(asr_out)$bic[1],
+             loglik = summary(asr_out)$loglik)
+  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+  }
+
+  if(i == 1 & j == 1 & k == 1){
+    h2_table <- h2_est
+  } else {
+    h2_table <- add_row(h2_table, h2_est)
+  }
+  tryCatch({
+      save_plot(filename=file.path(outputdir, paste0("ASReml_performance_",
+                                                 names(site_v)[i], "_",
+                                                 names(subpop_v)[j],
+                                                 "_", phe_v[k], ".png")), 
+            plot = plot(asr_out))
+    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+
+      }
+    }
+  }
+
+}
+
+```
+
+After this I'll need to select from each csv just the individuals in the subpop that I included phenotypes for, to do GWAS on. Then do GWAS with bigsnpr, and go ahead and look at those results.
+
+# gwas
+
+set up phenotype df for gwas
+```{r}
+site_v <- list(eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
+subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"))
+phe_fl <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl")
+phe_gr <- c("GR50", "tmin_10d", "cgdd_12c_10d")
+phe_dfs <- list(phe_gr_df, phe_fl_df)
+phe_lists <- list(phe_gr, phe_fl)
+
+i = 1
+j = 1
+k = 1
+p = 2
+phe_gwas <- tibble()
+
+for(j in seq_along(subpop_v)){
+  for(p in seq_along(phe_dfs)){
+    phe_v <- phe_lists[[p]]
+    phe <- phe_dfs[[p]]
+    for(i in seq_along(site_v)){
+      for(k in seq_along(phe_v)){
+        phe_raw <- read_csv(file.path(outputdir, paste0("blups_",
+                                                        names(site_v)[i], "_",
+                                                        names(subpop_v)[j],
+                                                        "_", phe_v[k],
+                                                        ".csv")))
+        names(phe_raw)[2] <- paste0(phe_v[k], "_G_BLUP")
+        phe_proc <- phe_raw %>%
+          separate(Effect, into = c("factor", "PLANT_ID"), sep = "\\)_") %>%
+          filter(!is.na(PLANT_ID)) %>%
+          left_join(metadata, by = "PLANT_ID") %>%
+          filter(SUBPOP %in% subpop_v[[j]]) %>%
+          select(PLANT_ID, 3)
+        if(i == 1 & k == 1 & p == 1){
+          phe_gwas <- phe_proc
+        } else {
+          phe_gwas <- phe_gwas %>%
+            left_join(phe_proc, by = "PLANT_ID")
+        }
+      }
+    }
+  }
+write_csv(phe_gwas, path = file.path(outputdir, paste0("G_BLUP_phenotypes_",
+                                                       "for_gwas_GR_FL_",
+                                                       names(subpop_v)[j],
+                                                       ".csv")))
+}
+```
+
+```{r}
+library(bigsnpr)
+library(tidyverse)
+library(switchgrassGWAS)
+library(AnnotationDbi)
+txdb <- loadDb(file = file.path("~", "Github", "pvdiv-genome",
+                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+inputdir <- file.path(workingdir, "analysis", "heritability")
+outputdir <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", "gwas", "G_BLUP")
+subpop_v <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", Gulf = "Gulf")
+
+inputfiles <- read_delim("~/Github/pvdiv-phenology-gxe/analysis/gwas/inputkinship.txt", delim = " ", col_names = "SNPfiles")
+
+for(i in seq_along(subpop_v)){
+  snp <- snp_attach(inputfiles$SNPfiles[i])
+  phe_df <- read_rds(file.path(inputdir, paste0("G_BLUP_phenotypes_",
+                                                "for_gwas_GR_FL_",
+                                                names(subpop_v)[i], ".csv")))
+  pvdiv_standard_gwas(snp, df = phe_df, type = "linear", outputdir = outputdir,
+                      savegwas = TRUE, saveannos = TRUE, txdb = txdb,
+                      minphe = 130)
+}
+```
+
 # Make new environmental variables for Li to use?
+
+We next evaluated genetic effects of greenup as functions of GDD (12C) and minimum temperature 10 days prior to greenup, and genetic effects of flowering as functions of daylength at flowering and GDD between greenup and flowering.
+
 Can I estimate heritability for the fourway cross?

### 2020-07-20 R/Plots_manuscript_v0.2.Rmd
Add GDD QTL file to data

@@
   snp <- snp_attach(inputfiles$SNPfiles[i])
-  phe_df <- read_rds(file.path(inputdir, paste0("G_BLUP_phenotypes_",
+  phe_df <- read_csv(file.path(inputdir, paste0("G_BLUP_phenotypes_",
                                                 "for_gwas_GR_FL_",
@@
 }
+snp_v <- c("12\\.1M", "8\\.8M", "10\\.3M")
+# Make output of top 1000 SNPs
+for(i in seq_along(subpop_v)){
+  snp <- snp_attach(inputfiles$SNPfiles[i])
+  gwas_rds <- pvdiv_results_in_folder(path = outputdir, 
+                                      pattern = paste0("GWAS_datatable*.*",
+                                                       snp_v[i]))
+  phenotypes <- str_sub(gwas_rds, start = 16, end = -43)
+  out1 <- pvdiv_bigsnp2mashr(path = outputdir, snp = snp, gwas_rds = gwas_rds, 
+                             phenotypes = phenotypes, numSNPs = 1000, 
+                             clump = FALSE, scaled = FALSE, saveoutput = TRUE,
+                             suffix = paste0("GR_FL_", names(subpop_v)[i]))
+}
+
 ```
+# examine top 1000 SNPs
+```{r}
+gulf_top1000 <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/gwas/G_BLUP/Part-One-Output_Top-Effects-1000-SNPs_GR_FL_Gulf.rds")
+gandm_top1000 <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/gwas/G_BLUP/Part-One-Output_Top-Effects-1000-SNPs_GR_FL_Gulf_and_Midwest.rds")
+midw_top1000 <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/gwas/G_BLUP/Part-One-Output_Top-Effects-1000-SNPs_GR_FL_Midwest.rds")
+```
+
+```{r}
+gulf_top1000 %>%
+  arrange(desc(max_score_log10p)) %>%
+  ungroup() %>%
+  #top_n(100, max_score_log10p) %>%
+  left_join(log10p_Gulf_top1000) %>%
+  #select(-starts_with("GR50"), -starts_with("tmin"), -starts_with("cgdd_12c_10d"), -marker, -CHRN) %>%
+  ggplot(aes(x = FL50_G_BLUP_log10p, y = cgdd_12c_10d_G_BLUP_log10p)) +
+  geom_point(aes(shape = as_factor(n_cond_in_top_set), color = max_score_log10p),
+             size = 2) +
+  geom_abline(slope = 1, intercept = 0) +
+  scale_color_viridis(end = 0.9, option = "B") +
+  scale_shape_manual(values = c(0, 1, 2, 5, 3)) +
+  theme(legend.position = "right")
+names(log10p_Gulf_top1000)
+
+midw_top1000 %>%
+  arrange(desc(max_score_log10p)) %>%
+  ungroup() %>%
+  #top_n(100, max_score_log10p) %>%
+  left_join(log10p_midwest) %>%
+  #select(-starts_with("GR50"), -starts_with("tmin"), -starts_with("cgdd_12c_10d"), -marker, -CHRN) %>%
+  ggplot(aes(x = FL50_G_BLU_log10p, y = cgdd_12c_gr2fl_G_BLU_log10p)) +
+  geom_point(aes(shape = as_factor(n_cond_in_top_set), color = max_score_log10p),
+             size = 2) +
+  geom_abline(slope = 1, intercept = 0) +
+  scale_color_viridis(end = 0.9, option = "B") +
+  scale_shape_manual(values = c(0, 1, 2, 5, 3)) +
+  theme(legend.position = "right")
+names(log10p_midwest)
+midw_top1000 %>%
+  arrange(desc(max_score_log10p)) %>%
+  ungroup() %>%
+  top_n(100, max_score_log10p)
+```
+Midwest has a lot more top hits for these G BLUPs that are at least on the same chromosomes as Li's QTLs.
+FWIW.
+GxE may be better, and need to get Li's intervals again and do a formal comparison, but just at first glance.
+top hits for Midwest are mostly for dyln_fl50...
+```{r}
+qtl <- read_csv("../data/QTLsWithFlankMarkers_FL50_GDD_7_SITES_Full_model.csv")
+qtldf <- qtl %>%
+  separate(flank_lo, into = c("CHR", "POS_lo"), sep = "_", convert = TRUE) %>%
+  separate(flank_hi, into = c("CHR1", "POS_hi"), sep = "_", convert = TRUE) %>%
+  select(-chr, -CHR1, -X1, -lodindex) %>%
+  mutate(POS_lo = POS_lo*1000000,
+         POS_hi = POS_hi*1000000) %>%
+  filter(lodcolumn != "D2F")
+
+i=1
+midw_qtl_overlap <- tibble()
+gulf_qtl_overlap <- tibble()
+gandm_qtl_overlap <- tibble()
+for(i in 1:nrow(qtldf)){
+  gam_qtl1 <- gandm_top1000 %>%
+    filter(CHR == qtldf$CHR[i] & between(POS, qtldf$POS_lo[i], qtldf$POS_hi[i]))
+  gulf_qtl1 <- gulf_top1000 %>%
+    filter(CHR == qtldf$CHR[i] & between(POS, qtldf$POS_lo[i], qtldf$POS_hi[i]))
+  mid_qtl1 <- midw_top1000 %>%
+    filter(CHR == qtldf$CHR[i] & between(POS, qtldf$POS_lo[i], qtldf$POS_hi[i]))
+  if(i == 1){
+    midw_qtl_overlap <- mid_qtl1
+    gulf_qtl_overlap <- gulf_qtl1
+    gandm_qtl_overlap <- gam_qtl1
+  } else {
+    midw_qtl_overlap <- midw_qtl_overlap %>%
+      full_join(mid_qtl1, by = c("CHR", "POS", "Condition", "max_score_log10p",
+                                 "n_cond_in_top_set", "marker", "CHRN"))
+    gandm_qtl_overlap <- gandm_qtl_overlap %>%
+      full_join(gam_qtl1, by = c("CHR", "POS", "Condition", "max_score_log10p",
+                                 "n_cond_in_top_set", "marker", "CHRN"))
+    gulf_qtl_overlap <- gulf_qtl_overlap %>%
+      full_join(gulf_qtl1, by = c("CHR", "POS", "Condition", "max_score_log10p",
+                                 "n_cond_in_top_set", "marker", "CHRN"))
+
+  }
+}
+
+nrow(midw_top1000)
+midw_qtl_overlap %>%
+  arrange(desc(max_score_log10p))
+nrow(gandm_top1000)
+gandm_qtl_overlap %>%
+  arrange(desc(max_score_log10p))
+nrow(gulf_top1000)
+gulf_qtl_overlap %>%
+  arrange(desc(max_score_log10p))
+```
+How many SNPs are in the QTL regions? And how many SNPs are in the full genome? Therefore how many SNPs would we expect at random?
 

### 2020-07-22 data-raw/Data_Setup.Rmd
Add new trait documentation for Li for QTL mapping

@@
 
+### ------
 # Make phenotypes for flowering time paper from all the cleaned phenotypes.
@@
 wea_df <- ptt_df %>%
-  dplyr::select(DOY:GDD_12C_SM, PTT_12C)
+  dplyr::select(DOY:GDD_12C_SM, PTT_12C) %>%
+  group_by(SITE) %>%
+  mutate(dyln_change_sec = (DYLN - lag(DYLN))*60*60)
 
@@
             )
+wea_gr_fl4 <- wea_gr_fl %>%
+  ungroup() %>%
+  group_by(Day_of_year, PLOT_GL, PLANT_ID) %>%
+  filter(Day_of_year == FL50) %>% 
+  rename(dyln_fl50 = DYLN) %>%
+  dplyr::select(-(TMAX:Day_of_year), -(TMAX_SM:PTT_12C))
 
 cumweadf <- cgdd_df %>%
-  left_join(crain_df)
+  left_join(crain_df) %>%
+  left_join(wea_gr_fl4)
 
 write_rds(cumweadf, path = "../data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
+```
+### -------
 
-```
 # All phenotypes

### 2020-07-22 R/Plots_manuscript_v0.2.Rmd
Wrap up work on manuscript v0.2 and start outline and analysis for v0.3

@@
 midw_top1000 <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/gwas/G_BLUP/Part-One-Output_Top-Effects-1000-SNPs_GR_FL_Midwest.rds")
+log10p_gulf <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/gwas/G_BLUP/log10p_strong_df_1000topSNPs_GR_FL_Gulf.rds")
+log10p_gandm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/gwas/G_BLUP/log10p_strong_df_1000topSNPs_GR_FL_Gulf_and_Midwest.rds")
+log10p_midw <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/gwas/G_BLUP/log10p_strong_df_1000topSNPs_GR_FL_Midwest.rds")
 ```
@@
 ```{r}
-qtl <- read_csv("../data/QTLsWithFlankMarkers_FL50_GDD_7_SITES_Full_model.csv")
+qtl <- read_csv("~/Github/pvdiv-phenology-gxe/data/QTLsWithFlankMarkers_FL50_GDD_7_SITES_Full_model.csv")
 qtldf <- qtl %>%
@@
   separate(flank_hi, into = c("CHR1", "POS_hi"), sep = "_", convert = TRUE) %>%
-  select(-chr, -CHR1, -X1, -lodindex) %>%
+  dplyr::select(-chr, -CHR1, -X1, -lodindex) %>%
   mutate(POS_lo = POS_lo*1000000,
@@
 
+i=3
+snp <- snp_attach(inputfiles$SNPfiles[i])
+markers <- tibble(CHR = snp$map$chromosome, POS = snp$map$physical.pos)
+marker_qtl_overlap <- tibble()
+for(i in 1:nrow(qtldf)){
+  qtl1 <- markers %>%
+    filter(CHR == qtldf$CHR[i] & between(POS, qtldf$POS_lo[i], qtldf$POS_hi[i]))
+  if(i == 1){
+    marker_qtl_overlap <- qtl1
+  } else {
+    marker_qtl_overlap <- marker_qtl_overlap %>%
+      full_join(qtl1, by = c("CHR", "POS"))
+  }
+} 
+nrow(markers)
+nrow(marker_qtl_overlap)
+
 nrow(midw_top1000)
@@
 gulf_qtl_overlap %>%
-  arrange(desc(max_score_log10p))
+  arrange(desc(max_score_log10p)) %>%
+  group_by(Condition, n_cond_in_top_set) %>% tally()
+
+names(log10p_Gulf_top1000)
+log10p_Gulf_top1000 %>%
+  ungroup() %>%
+  top_n(1000, cgdd_12c_gr2fl_G_BLUP_log10p) %>%
+  inner_join(gulf_qtl_overlap)
+snpmatrix <- matrix(c(49, 1000, 565807, 10320201), nrow = 2, dimnames = list(Observed = c("SNPs in QTL", "Total SNPs"), Expected = c("Sig SNPs", "Total SNPs")))
+fisher.test(snpmatrix, alternative = "t")
 ```
-How many SNPs are in the QTL regions? And how many SNPs are in the full genome? Therefore how many SNPs would we expect at random?
+FL50; dyln_fl50 in Gulf: 6.67 (52 / 1000)
+dyln_change_sec in Gulf: 8.21 (64 / 1000)
+cgdd_12c_10d_G_BLUP_log10p 37 / 1000 4.75
 
-# Make new environmental variables for Li to use?
+How many SNPs are in the QTL regions? And how many SNPs are in the full genome?
+565807 in QTL region out of 10320201 SNPs in Gulf set. 241 / 5043   5.48%
+650092 in QTL region out of 12083955 SNPs in Gulf and Midwest set. 276 / 5320 5.38%
+475951 in QTL region out of 8824339 SNPs in Midwest set. Whoa. That's already a striking difference. ~5.39 %ofthegenome.
+Therefore how many SNPs would we expect at random?
+
+Fisher's exact test for SNPs above Bonferroni in QTL intervals, vs elsewhere.
+```{r}
+-log10(0.05/10320201)  # 8.31 Gulf
+-log10(0.05/12083955)  # 8.38 GandM
+-log10(0.05/8824339)   # 8.25 Midwest
+subpop_v
+bonferroni_v <- c(-log10(0.05/12083955), -log10(0.05/8824339), 
+                  -log10(0.05/10320201))
+log10p_list <- list(gandm = log10p_gandm, midwest = log10p_midw, gulf = log10p_gulf)
+
+```
+
+```{r}
+sig_df <- tibble()
+
+for(j in seq_along(bonferroni_v)){
+  log10p_df <- log10p_list[[j]] %>%
+    mutate_at(.vars = vars(ends_with("log10p")),
+              .funs = ~ if_else(. >= bonferroni_v[j], ., NA_real_)) %>%
+    ungroup %>%
+    select(-(1)) 
+  if(j == 1){
+    sig_df <- enframe(colSums(!is.na(log10p_df[,-1])), name = "Column", 
+                      value = paste0("Num_sig_", names(subpop_v)[j])) %>%
+      mutate(Column = str_sub(Column, end = 10))
+  } else {
+    sig1 <- enframe(colSums(!is.na(log10p_df[,-1])), name = "Column", 
+                    value = paste0("Num_sig_", names(subpop_v)[j])) %>%
+      mutate(Column = str_sub(Column, end = 10))
+    sig_df <- sig_df %>%
+      full_join(sig1, by = "Column")
+  }
+}
+sig_df
+write_csv(sig_df, "G_BLUP_Num_SNPs_above_Bonferroni_GR_FL.csv")
+```
+
+```{r}
+j = 1
+i=4
+names(log10p_gandm)[4:ncol(log10p_gandm)]
+log10p_df <- log10p_gandm
+
+marker_qtl_overlap <- tibble()
+
+for(i in 4:ncol(log10p_df)){
+  phe_use <- names(log10p_df)[i]
+  sig_markers <- log10p_df %>%
+    select(SNP:POS, eval(phe_use)) %>%
+    ungroup() %>%
+    filter(!! sym(phe_use) >= bonferroni_v[j])
+
+  for(k in 1:nrow(qtldf)){
+  qtl1 <- sig_markers %>%
+    filter(CHR == qtldf$CHR[k] & between(POS, qtldf$POS_lo[k], qtldf$POS_hi[k]))
+  if(k == 1){
+    marker_qtl_overlap <- qtl1
+  } else {
+    marker_qtl_overlap <- marker_qtl_overlap %>%
+      full_join(qtl1, by = c("CHR", "POS"))
+  }
+} 
+nrow(sig_markers)
+nrow(marker_qtl_overlap)
+
+}
+log10p_df <- log10p_df %>%
+  mutate_at(.vars = vars(ends_with("log10p")), .funs = ~ if_else(. >= bonferroni_v[j], ., NA_real_)) %>%
+  ungroup %>%
+  select(-(1))
+
+colSums(!is.na(log10p_df[,-1]))
+marker_qtl_overlap <- tibble()
+
+for(k in 1:nrow(qtldf)){
+  phe_use <- names(log10p_df)[3]
+  qtl1 <- log10p_df %>%
+    filter(!is.na(!! sym(phe_use)) & CHR == qtldf$CHR[k] & between(POS, qtldf$POS_lo[k], qtldf$POS_hi[k]))
+  if(k == 1){
+    marker_qtl_overlap <- qtl1
+  } else {
+    marker_qtl_overlap <- marker_qtl_overlap %>%
+      full_join(qtl1, by = c("CHR", "POS"))
+  }
+}
+k=1
+
+```
+
+Um, so by Fisher's exact test, there are fewer Midwest SNPs in the sig set than we would expect at random. Rad. What the heck.
+And no significant enrichment of Gulf SNPs in the sig set.
+And no significant enrichment in the Gulf & Midwest set. So what does this mean?
+```{r}
+snpmatrix <- matrix(c(115, 4876, 475951, 8824339), nrow = 2, dimnames = list(Observed = c("SNPs in QTL", "Total SNPs"), Expected = c("Sig SNPs", "Total SNPs")))
+matrix(c(3, 40, 500, 80278), nrow = 2, dimnames = list(Observed = c("GO in Sample", "Genes in Sample"), Expected = c("GO in Genome", "Genes in Genome")))
+fisher.test(snpmatrix, alternative = "t")
+(115/4876)/(475951/8824339)
+snpmatrix <- matrix(c(241, 5043, 565807, 10320201), nrow = 2, dimnames = list(Observed = c("SNPs in QTL", "Total SNPs"), Expected = c("Sig SNPs", "Total SNPs")))
+fisher.test(snpmatrix, alternative = "t")
+
+snpmatrix <- matrix(c(276, 5320, 650092, 12083955), nrow = 2, dimnames = list(Observed = c("SNPs in QTL", "Total SNPs"), Expected = c("Sig SNPs", "Total SNPs")))
+fisher.test(snpmatrix, alternative = "t")
+```
+genematrix <- matrix(c(3, 40, 500, 80278), nrow = 2, dimnames = list(Observed = c("GO in Sample", "Genes in Sample"), Expected = c("GO in Genome", "Genes in Genome")))
+fisher.test(GOmatrix, alternative = "t")
+
+# Make new environmental variables for Li to use? Did this in Data_Setup.
 

### 2020-07-22 R/Plots_manuscript_v0.3.Rmd
Wrap up work on manuscript v0.2 and start outline and analysis for v0.3

---
title: "Plots_manuscript_v0.3"
author: "Alice MacQueen"
date: "7/22/2020"
output: html_document
---

Add any libraries/scripts that need to be loaded for these analyses here.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(lubridate)
library(switchgrassGWAS)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

# Load data
Add any data that needs to be loaded for these analyses here.
```{r}
metadata <- readRDS("~/Github/pvdiv-phenology-gxe/data/metadata.rds")
sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
phenotypes <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
```

# Overview
Here I'm putting the code for any new analyses that I'm running to do new writeup for the manuscript v0.3.

v0.2 has the ASReml h2 analysis, has BLUPs for G (which as of v0.3 probably needs a GxE term...)

### 2020-07-23 R/Plots_manuscript_v0.3.Rmd
Add manuscript work and GR50 and FL50 distribution correlation and h2 figure

@@
 library(switchgrassGWAS)
+library(GGally)
 source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
@@
 ```{r}
-metadata <- readRDS("~/Github/pvdiv-phenology-gxe/data/metadata.rds")
-sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
+sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
+subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
+phe_gr_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
+phe_fl_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
 phenotypes <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
+h2fl <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/heritability/h2_df_FL50_related_functions_of_weather_v3.csv")
+h2gr <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/heritability/h2_df_GR50_related_functions_of_weather_v2.csv")
 ```
@@
 
-v0.2 has the ASReml h2 analysis, has BLUPs for G (which as of v0.3 probably needs a GxE term...)
+v0.2 has the ASReml h2 analysis, has BLUPs for G (which as of v0.3 probably needs a GxE term... define some possibilities below.)
+
+Results: h2 at single sites. Then h2 at all eight sites. And so there must be GxE, and there must be a lot of rank-changing GxE, in that we can’t predict ranks across sites.
+
+To show this, make barplots of the different variables and facet them by site.
+Then, show heritability across all eight sites (in a model with a random effect of site, interacted with some number of eigenvectors from a SVD of the kinship matrix.)
+
+# Heritability Plots
+
+## h2 as functions of GR50 and FL50
+```{r}
+h2all <- h2fl %>%
+  full_join(h2gr)
+
+f1h2 <- h2all %>%
+  filter(phe %in% c("FL50", "GR50")) %>%
+  mutate(site = case_when(is.na(site) ~ "all",
+                          site == "eight_sites" ~ "all",
+                          TRUE ~ site),
+         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
+                                 subpop == "three_subpops" ~ "all",
+                                 TRUE ~ subpop),
+         phe_name = case_when(phe == "GR50" ~ "50% greenup",
+                              phe == "FL50" ~ "50% flowering",
+                              TRUE ~ phe))
+f1h2$site <- factor(f1h2$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "all", "nine_sites"))
+f1h2$phe_name <- factor(f1h2$phe_name, levels = c("50% greenup", "50% flowering"))
+f1h2$subpop_name <- factor(f1h2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "all"))
+ggh21 <- f1h2 %>%
+  filter(!(site %in% c("nine_sites", "IL")) & !(subpop %in% c("three_subpops", "Atlantic"))) %>%
+  ggplot(aes(x = site, y = Estimate)) +
+  geom_bar(aes(fill = site == "all"), stat = "identity") +
+  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
+  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
+  facet_grid(phe_name ~ subpop_name) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), 
+        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
+        legend.position = "none") + 
+  ylim(c(0, 1)) +
+  xlab("Common Gardens") + 
+  ylab(bquote('h'^2)) +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
+  #
+save_plot("../manuscript/plots/h2_GR_FL_site_by_subpop.png", plot = ggh21, base_height = 3.5, base_asp = 1.618)
+
+f1h2 %>%
+  filter(!(site %in% c("all", "nine_sites", "IL")) & !(subpop %in% c("three_subpops", "Atlantic"))) %>%
+  ungroup() %>%
+  group_by(phe) %>%
+  summarise(mean = mean(Estimate))
+f1h2 %>%
+  filter((site %in% c("all")) & !(subpop %in% c("three_subpops", "Atlantic"))) %>%
+  ungroup() %>%
+  group_by(phe) %>%
+  summarise(mean = mean(Estimate))
+```
+
+## h2 flowering weather
+```{r}
+flh2 <- h2fl %>%
+  filter(phe %in% c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d")) %>%
+  mutate(site = case_when(is.na(site) ~ "all",
+                          site == "eight_sites" ~ "all",
+                          TRUE ~ site),
+         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
+                                 subpop == "three_subpops" ~ "all",
+                                 TRUE ~ subpop),
+         phe_name = case_when(phe == "GR50" ~ "50% greenup",
+                              phe == "FL50" ~ "50% flowering",
+                              TRUE ~ phe))
+flh2$site <- factor(flh2$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "all", "nine_sites"))
+flh2$phe_name <- factor(flh2$phe_name, levels = c("50% greenup", "50% flowering"))
+flh2$subpop_name <- factor(flh2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "all"))
+
+flh2_p1 <- flh2 %>%
+  filter(subpop == "Gulf" & !(site %in% c("nine_sites")) & !is.na(site)) %>%
+  ggplot(aes(x = phe, y = Estimate)) +
+  geom_bar(stat = "identity") +
+  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
+  facet_wrap(~ site) +
+  geom_hline(data = filter(flh2, subpop == "Gulf" & !(site %in% c("NA", "nine_sites")) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + ylim(c(0, 1))
+save_plot("../manuscript/plots/h2_FL_site_weather_functions.png", plot = flh2_p1, base_height = 3.5, base_asp = 1.618)
+flh2 %>%
+  filter(subpop == "Gulf_and_Midwest" & !(site %in% c("eight_sites", "nine_sites")) & !is.na(site)) %>%
+  ggplot(aes(x = phe, y = Estimate)) +
+  geom_bar(stat = "identity") +
+  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
+  facet_wrap(~ site) +
+  geom_hline(data = filter(flh2, subpop == "Gulf_and_Midwest" & !(site %in% c("NA", "nine_sites")) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + ylim(c(0,1))
+
+```
+## h2 greenup weather
+```{r}
+
+h2gr$site <- factor(h2gr$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
+flh2 %>%
+  filter(subpop == "Gulf" & !(site %in% c("eight_sites", "nine_sites")))
+
+grh2_p1 <- h2gr %>%
+  filter(subpop == "Gulf" & !(site %in% c("eight_sites", "nine_sites")) & !is.na(site)) %>%
+  ggplot(aes(x = phe, y = Estimate)) +
+  geom_bar(stat = "identity") +
+  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
+  facet_wrap(~ site) +
+  geom_hline(data = filter(flh2, subpop == "Gulf" & !(site %in% c("NA", "nine_sites")) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + ylim(c(0, 1))
+save_plot("../manuscript/plots/h2_GR_site_weather_functions.png", plot = grh2_p1, base_height = 3.5, base_asp = 1.618)
+```
+## Phenotype correlations
+can you easily look at the correlation between greenup ranks at OK versus NE?
+```{r}
+phe_gr_df %>%
+  filter(SITE %in% c("LINC", "STIL")) %>%
+  ungroup() %>%
+  select(PLANT_ID:GR50, -PLOT_GL) %>%
+  pivot_wider(names_from = "SITE", values_from = "GR50", values_fn = mean) %>%
+  ggplot(aes(x = STIL, y = LINC)) +
+  geom_hex() + #aes(color = SUBPOP), height = 1, width = 1) +
+  scale_fill_viridis(option = "B",
+                     end = 0.9, direction = -1)
+
+phe_gr_df %>%
+  filter(SITE %in% c("STIL", "PKLE")) %>%
+  ungroup() %>%
+  select(PLANT_ID:GR50, -PLOT_GL) %>%
+  pivot_wider(names_from = "SITE", values_from = "GR50", values_fn = mean) %>%
+  ggplot(aes(x = STIL, y = Lat_origin)) +
+  geom_hex() + #aes(color = SUBPOP), height = 1, width = 1) +
+  scale_fill_viridis(option = "B",
+                     end = 0.9, direction = -1)
+
+cor_gr <- phe_gr_df %>%
+  left_join(select(sites, SITE, manu_site)) %>%
+  ungroup() %>%
+  select(PLANT_ID:GR50, manu_site, -PLOT_GL, -SITE) %>%
+  pivot_wider(names_from = "manu_site", values_from = "GR50", values_fn = mean)
+cor_fl <- phe_fl_df %>%
+  left_join(select(sites, SITE, manu_site)) %>%
+  ungroup() %>%
+  select(PLANT_ID:FL50, manu_site, -PLOT_GL, -SITE) %>%
+  pivot_wider(names_from = "manu_site", values_from = "FL50", values_fn = mean)
+
+cor_north <- phe_gr_df %>%
+  left_join(select(sites, SITE, manu_site)) %>%
+  ungroup() %>%
+  filter(Lat_origin > 40) %>%
+  select(PLANT_ID:GR50, manu_site, -PLOT_GL, -SITE) %>%
+  pivot_wider(names_from = "manu_site", values_from = "GR50", values_fn = mean)
+cor_south <- phe_gr_df %>%
+  left_join(select(sites, SITE, manu_site)) %>%
+  ungroup() %>%
+  filter(Lat_origin < 32) %>%
+  select(PLANT_ID:GR50, manu_site, -PLOT_GL, -SITE) %>%
+  pivot_wider(names_from = "manu_site", values_from = "GR50", values_fn = mean)
+cor(cor_north[,5:13], use = "pairwise") %>%
+  switchgrassGWAS:::reorder_cormat(.) %>%
+  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
+cor(cor_south[,5:13], use = "pairwise") %>%
+  switchgrassGWAS:::reorder_cormat(.) %>%
+  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
+
+cor(cor_gr$STIL, cor_gr$LINC, use = "pairwise")
+cor(cor_gr[,5:13], use = "pairwise") %>%
+  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
+cor(cor_fl[,5:13], use = "pairwise") %>%
+  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
+
+# two different ways. Think I like the second more. Force latitudinal order.
+cor_gr %>%
+  filter(SUBPOP %in% c("Midwest", "Gulf")) %>%
+  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
+  cor(use = "pairwise") %>%
+  switchgrassGWAS:::reorder_cormat(.) %>%
+  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
+```
+
+```{r}
+pand <- cor_fl %>%
+  filter(SUBPOP %in% c("Gulf")) %>%
+  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
+  ggcorr(method = "pairwise", label = TRUE) +
+  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                      "white", "#27AD81FF", "#5DC863FF",
+                                      "#FDE725FF"), limits = c(-1,1)) +
+  theme(legend.position = "none")
+pane <- cor_fl %>%
+  filter(SUBPOP %in% c("Midwest")) %>%
+  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
+  ggcorr(method = "pairwise", label = TRUE) +
+  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                      "white", "#27AD81FF", "#5DC863FF",
+                                      "#FDE725FF"), limits = c(-1,1)) +
+  theme(legend.position = "none")
+panf <- cor_fl %>%
+  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
+  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
+  ggcorr(method = "pairwise", label = TRUE) +
+  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                      "white", "#27AD81FF", "#5DC863FF",
+                                      "#FDE725FF"), limits = c(-1,1)) +
+  theme(legend.position = "none")
+pana <- cor_gr %>%
+  filter(SUBPOP %in% c("Gulf")) %>%
+  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
+  ggcorr(method = "pairwise", label = TRUE) +
+  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                      "white", "#27AD81FF", "#5DC863FF",
+                                      "#FDE725FF"), limits = c(-1,1)) +
+  theme(legend.position = c(0.4, 0.85))
+panb <- cor_gr %>%
+  filter(SUBPOP %in% c("Midwest")) %>%
+  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
+  ggcorr(method = "pairwise", label = TRUE) +
+  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                      "white", "#27AD81FF", "#5DC863FF",
+                                      "#FDE725FF"), limits = c(-1,1)) +
+  theme(legend.position = "none")
+panc <- cor_gr %>%
+  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
+  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
+  ggcorr(method = "pairwise", label = TRUE) +
+  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                      "white", "#27AD81FF", "#5DC863FF",
+                                      "#FDE725FF"), limits = c(-1,1)) +
+  theme(legend.position = "none")
+
+pheno_corr <- plot_grid(pana, panb, panc, pand, pane, panf, nrow = 2)
+save_plot(filename = "../manuscript/plots/Phenotypic_correlations_GR50_FL50_2.png", plot = pheno_corr, base_height = 5)
+```
+
+# --------------
+# SVD of K
+
+## Plot eigenvectors to capture predominant patterns in K matrix.
+```{r}
+k_full <- read_rds("../data/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
+```
+
+```{r}
+b1 <- svd(k_full)
+b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 11.5
+b1$d[2] / sum(b1$d) # 7.1
+b1$d[3] / sum(b1$d) # 6.1
+b1$d[4] / sum(b1$d) # 1.7
+b1$d[5] / sum(b1$d) # 1.3
+sum(b1$d[1:5]) / sum(b1$d) # First 5 eigenvectors explain 1% or more variance, and 27.7% of total variance.
+for(g in 1:5){
+barplot(b1$v[,g]/k_full[which.max(abs(k_full[,g])),g])
+}
+order <- enframe(colnames(k_full), name = NULL, value = "PLANT_ID")
+ggb1 <- tibble(`First eigenvector (PVE = 11.5%)` = b1$v[,1]/k_full[which.max(abs(k_full[,1])),1],
+       `Second eigenvector (PVE = 7.1%)` = b1$v[,2]/k_full[which.max(abs(k_full[,2])),2],
+       `Third eigenvector (PVE = 6.1%)` = b1$v[,3]/k_full[which.max(abs(k_full[,3])),3],
+       `Fourth eigenvector (PVE = 1.7%)` = b1$v[,4]/k_full[which.max(abs(k_full[,4])),4],
+       `Fifth eigenvector (PVE = 1.3%)` = b1$v[,5]/k_full[which.max(abs(k_full[,5])),5],
+       `Sixth eigenvector (PVE = 0.9%)` = b1$v[,6]/k_full[which.max(abs(k_full[,6])),6],
+       PLANT_ID = order$PLANT_ID) %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  dplyr::select(PLANT_ID, SUBPOP, ECOTYPE_NNET, LATITUDE, ends_with("%)"))
+
+ggb1 <- ggb1 %>%
+  arrange(LATITUDE) %>%
+  mutate(rank = row_number()) %>%
+  pivot_longer(cols = ends_with("%)"), names_to = "Eigenvector", values_to = "value")
+ggb1$Eigenvector <- factor(ggb1$Eigenvector, 
+                           levels = c("First eigenvector (PVE = 11.5%)", 
+                                      "Second eigenvector (PVE = 7.1%)", 
+                                      "Third eigenvector (PVE = 6.1%)", 
+                                      "Fourth eigenvector (PVE = 1.7%)",
+                                      "Fifth eigenvector (PVE = 1.3%)",
+                                      "Sixth eigenvector (PVE = 0.9%)"
+                                      ))
+ggb1_plot <- ggb1 %>%
+  ggplot(aes(x = rank, y = value)) +
+  geom_bar(aes(color = SUBPOP, fill = SUBPOP), stat = "identity") +
+  facet_wrap(~ Eigenvector) +
+  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
+  scale_color_viridis(discrete = TRUE, end = 0.9, direction = -1) +
+  scale_fill_viridis(discrete = TRUE, end = 0.9, direction = -1) +
+  xlab("Latitude of Origin (Rank Order)") + ylab("") +
+  geom_vline(xintercept = 607, linetype = 2)
+save_plot("../manuscript/plots/SVD_of_Kinship_matrix_first_six_eigenvectors.png", base_height = 5, base_asp = 2, plot = ggb1_plot)
+
+ggb1_plot2 <- ggb1 %>%
+  ggplot(aes(x = rank, y = value)) +
+  geom_bar(aes(color = ECOTYPE_NNET, fill = ECOTYPE_NNET), stat = "identity") +
+  facet_wrap(~ Eigenvector) +
+  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
+  scale_color_viridis(discrete = TRUE, option = "B", end = 0.9, 
+                      direction = -1) +
+  scale_fill_viridis(discrete = TRUE, option = "B", end = 0.9, direction = -1) +
+  xlab("Latitude of Origin (Rank Order)") + ylab("") +
+  geom_vline(xintercept = 607, linetype = 2)
+save_plot("../manuscript/plots/SVD_of_Kinship_matrix_first_six_eigenvectors_ecotype.png", base_height = 5, base_asp = 2, plot = ggb1_plot2)
+```
+
+The eigenvectors capture the predominant patterns in the Uk3 covariance matrix.
+```
+v <- m$fitted_g$Ulist$ED_PCA_1
+b1=svd(v)
+for(g in 1){
+barplot(b1$v[,g]/v[which.max(abs(v[,g])),g])
+}
+
+b1$d # the top eigenvector captures the predominant pattern in the ED_PCA_1 covariance matrix. 
+```
+
+WAIT. I could also plot this for the svd from the whole SNP set. That might make more sense than doing a svd on the kinship matrix...
+
+# -------------
+# ASReml models for eight sites including GxE
+```{r}
+
+```
+
+## ASReml h2 greenup analysis
+```{r}
+library(tidyverse)
+library(asreml)
+library(switchgrassGWAS)
+library(cowplot)
+```
+
+```{r}
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "analysis", "heritability")
+metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
+sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
+subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
+phe_gr_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
+phe_fl_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
+k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
+
+```
+
+### loop to make heritability table using ASReml mixed models and vpredict.
+
+## Include just texas sites for greenup and just northern four sites.
+```{r}
+site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
+               NE = "NE", IL = "IL", MI = "MI", SD = "SD", eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), nine_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
+subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))
+phe_fl <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl",
+           "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d",
+           "crain_7d")
+phe_gr <- c("GR50", "DYLN", "cgdd_12c_jan2gr", "cgdd_8c_jan2gr", "tmax_18d", 
+            "tmin_18d", "tave_18d", "cgdd_12c_18d", "tmax_10d", "tmin_10d",
+            "tave_10d", "cgdd_12c_10d", "tmax_5d", "tmin_5d", "tave_5d",
+            "cgdd_12c_5d")
+phe_dfs <- list(phe_gr_df, phe_fl_df)
+phe_lists <- list(phe_gr, phe_fl)
+h2_table <- tibble()
+
+for(p in 1){#seq_along(phe_dfs)){
+  phe_v <- phe_lists[[p]]
+  phe <- phe_dfs[[p]]
+
+  for(i in 12:13){#seq_along(site_v)){
+    for(j in 1:3){#seq_along(subpop_v)){
+      for(k in seq_along(phe_v)){
+        phe_single <- phe %>%
+    left_join(sites, by = "SITE") %>%
+    filter(manu_site %in% site_v[[i]] & SUBPOP %in% subpop_v[[j]]) %>%
+    rename(phenotype = eval(phe_v[k])) %>%
+    mutate(PLANT_ID = as_factor(PLANT_ID),
+           SITE = as_factor(SITE),
+           PLOT_GL = as_factor(PLOT_GL)) %>%
+    as.data.frame()
+
+  if(length(site_v[[i]]) > 1){  # need to add site as a random factor if 
+    # there is more than one site included in the dataset.
+    tryCatch({
+     asr_out <- asreml(phenotype ~ 1,
+                       random = ~vm(PLANT_ID, k_full) +
+                         ~idv(SITE), #:vm(PLANT_ID, k_full),
+                       residual = ~idv(units),
+                       data = phe_single, 
+                       workspace = "3gb")
+
+    h2_est <- vpredict(asr_out, h2 ~ V2/(V1+V2+V3)) %>% as_tibble() %>%
+      mutate(site = names(site_v)[i],
+             subpop = names(subpop_v)[j],
+             phe = phe_v[k],
+             bic = summary(asr_out)$bic[1],
+             loglik = summary(asr_out)$loglik)
+    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+  } else {
+    tryCatch({
+    asr_out <- asreml(phenotype ~ 1,
+                       random = ~vm(PLANT_ID, k_full),
+                       residual = ~idv(units),
+                       data = phe_single, 
+                       workspace = "3gb")
+
+    h2_est <- vpredict(asr_out, h2 ~ V1/(V1+V2)) %>% as_tibble() %>%
+      mutate(site = names(site_v)[i],
+             subpop = names(subpop_v)[j],
+             phe = phe_v[k],
+             bic = summary(asr_out)$bic[1],
+             loglik = summary(asr_out)$loglik)
+  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+  }
+
+  if(i == 1 & j == 1 & k == 1){
+    h2_table <- h2_est
+  } else {
+    h2_table <- add_row(h2_table, h2_est)
+  }
+  tryCatch({
+      save_plot(filename=file.path(outputdir, paste0("ASReml_performance_",
+                                                 names(site_v)[i], "_",
+                                                 names(subpop_v)[j],
+                                                 "_", phe_v[k], ".png")), 
+            plot = plot(asr_out))
+    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+
+      }
+    }
+  }
+  write_csv(h2_table, file.path(outputdir, paste0("h2_df_", phe_v[1],
+                                                  "_related_functions_", 
+                                                  "of_weather_v4.csv")))
+}

### 2020-07-25 R/Plots_manuscript_v0.3.Rmd
Fix phenotypes for 4WCR and plot FL and GR as functions of weather

@@
 h2gr <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/heritability/h2_df_GR50_related_functions_of_weather_v2.csv")
+h2gr4 <- read_csv(file.path(workingdir, "analysis", "heritability", "h2_df_GR50_related_functions_of_weather_v4.csv"))
+h2gr <- h2gr %>%
+  full_join(h2gr4)
 ```
@@
   mutate(site = case_when(is.na(site) ~ "all",
-                          site == "eight_sites" ~ "all",
+                          site == "eight_sites" ~ "All",
                           TRUE ~ site),
@@
                                  TRUE ~ subpop),
-         phe_name = case_when(phe == "GR50" ~ "50% greenup",
-                              phe == "FL50" ~ "50% flowering",
+         phe_name = case_when(phe == "FL50" ~ "50% flowering",
+                              phe == "dyln_fl50" ~ "daylength",
+                              phe == "dyln_change_sec" ~ "daylength change (s)",
+                              phe == "cgdd_12c_gr2fl" ~ "GDD, GR to FL",
+                              phe == "crain_gr2fl" ~ "Rainfall, GR to FL",
+                              phe == "crain_1d" ~ "Rainfall, 1 day",
+                              phe == "crain_2d" ~ "Rainfall, 2d sum",
+                              phe == "crain_3d" ~ "Rainfall, 3d sum",
+                              phe == "crain_5d" ~ "Rainfall, 5d sum",
+                              phe == "crain_7d" ~ "Rainfall, 7d sum",
                               TRUE ~ phe))
-flh2$site <- factor(flh2$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "all", "nine_sites"))
-flh2$phe_name <- factor(flh2$phe_name, levels = c("50% greenup", "50% flowering"))
+flh2$site <- factor(flh2$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "nine_sites"))
+flh2$phe_name <- factor(flh2$phe_name, levels = c("50% flowering", "daylength",
+                                                  "daylength change (s)",
+                                                  "GDD, GR to FL",
+                                                  "Rainfall, GR to FL",
+                                                  "Rainfall, 1 day",
+                                                  "Rainfall, 2d sum",
+                                                  "Rainfall, 3d sum",
+                                                  "Rainfall, 5d sum",
+                                                  "Rainfall, 7d sum"
+                                                  ))
 flh2$subpop_name <- factor(flh2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "all"))
+site_plot <- c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD", "All", "Texas", "North")
 
 flh2_p1 <- flh2 %>%
-  filter(subpop == "Gulf" & !(site %in% c("nine_sites")) & !is.na(site)) %>%
-  ggplot(aes(x = phe, y = Estimate)) +
-  geom_bar(stat = "identity") +
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site)) %>% # subpop == subpop_v[[i]] & 
+  ggplot(aes(x = phe_name, y = Estimate)) +
+  geom_bar(stat = "identity", aes(fill = phe_name == "50% flowering")) +
   geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
-  facet_wrap(~ site) +
-  geom_hline(data = filter(flh2, subpop == "Gulf" & !(site %in% c("NA", "nine_sites")) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + ylim(c(0, 1))
-save_plot("../manuscript/plots/h2_FL_site_weather_functions.png", plot = flh2_p1, base_height = 3.5, base_asp = 1.618)
-flh2 %>%
-  filter(subpop == "Gulf_and_Midwest" & !(site %in% c("eight_sites", "nine_sites")) & !is.na(site)) %>%
-  ggplot(aes(x = phe, y = Estimate)) +
-  geom_bar(stat = "identity") +
+  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
+  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
+  facet_grid(site ~ subpop_name) +
+  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 1.1)) +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
+
+save_plot("../manuscript/plots/h2_FL_site_weather_functions_All_Sites_v2.png", plot = flh2_p1, base_height = 7, base_asp = 1.3)
+
+site_plot <- c("All", "Texas", "North")
+
+flh2_p2 <- flh2 %>%
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site)) %>% # subpop == subpop_v[[i]] & 
+  ggplot(aes(x = phe_name, y = Estimate)) +
+  geom_bar(stat = "identity", aes(fill = phe_name == "50% flowering")) +
   geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
-  facet_wrap(~ site) +
-  geom_hline(data = filter(flh2, subpop == "Gulf_and_Midwest" & !(site %in% c("NA", "nine_sites")) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + ylim(c(0,1))
+  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
+  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
+  facet_grid(site ~ subpop_name) +
+  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.5)) +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")
 
+save_plot("../manuscript/plots/h2_FL_site_weather_functions_All_Sites_v2.png", plot = flh2_p2, base_height = 4, base_asp = 1.5)
 ```
@@
 ```{r}
+h2gr
+```
 
-h2gr$site <- factor(h2gr$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
-flh2 %>%
-  filter(subpop == "Gulf" & !(site %in% c("eight_sites", "nine_sites")))
+```{r}
+unique(h2gr$site)
+grh2 <- h2gr %>%
+  filter(!(phe %in% c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d", "DYLN"))) %>%
+  mutate(site = case_when(is.na(site) ~ "all",
+                          site == "eight_sites" ~ "All",
+                          site == "tx_sites" ~ "Texas",
+                          site == "north_sites" ~ "North",
+                          TRUE ~ site),
+         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
+                                 subpop == "three_subpops" ~ "All",
+                                 TRUE ~ subpop),
+         phe_name = case_when(phe == "GR50" ~ "50% greenup",
+                              phe == "FL50" ~ "50% flowering",
+                              phe == "cgdd_12c_jan2gr" ~ "GDD 01-01 to greenup",
+                              phe == "cgdd_8c_jan2gr" ~ "GDD 8C 01-01 to greenup",
+                              phe == "tmax_18d" ~ "Max T, 18d mean",
+                              phe == "tmin_18d" ~ "Min T, 18d mean",
+                              phe == "tave_18d" ~ "Mean T, 18d mean",
+                              phe == "tmax_10d" ~ "Max T, 10d mean",
+                              phe == "tmin_10d" ~ "Min T, 10d mean",
+                              phe == "tave_10d" ~ "Mean T, 10d mean",
+                              phe == "tmax_5d" ~ "Max T, 5d mean",
+                              phe == "tmin_5d" ~ "Min T, 5d mean",
+                              phe == "tave_5d" ~ "Mean T, 5d mean",
+                              phe == "cgdd_12c_10d" ~ "GDD, 10d",
+                              phe == "cgdd_12c_18d" ~ "GDD, 18d",
+                              phe == "cgdd_12c_5d" ~ "GDD, 5d",
+                              TRUE ~ phe))
+grh2$site <- factor(grh2$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North", "nine_sites"))
+grh2$phe_name <- factor(grh2$phe_name, levels = c("50% greenup", "GDD, 5d",
+                                                  "Min T, 5d mean",
+                                                  "Mean T, 5d mean",
+                                                  "Max T, 5d mean",
+                                                  "GDD, 10d",
+                                                  "Min T, 10d mean",
+                                                  "Mean T, 10d mean",
+                                                  "Max T, 10d mean",
+                                                  "GDD, 18d",
+                                                  "Min T, 18d mean",
+                                                  "Mean T, 18d mean",
+                                                  "Max T, 18d mean",
+                                                  "GDD 8C 01-01 to greenup",
+                                                  "GDD 01-01 to greenup"))
+grh2$subpop_name <- factor(grh2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "All"))
 
-grh2_p1 <- h2gr %>%
-  filter(subpop == "Gulf" & !(site %in% c("eight_sites", "nine_sites")) & !is.na(site)) %>%
-  ggplot(aes(x = phe, y = Estimate)) +
-  geom_bar(stat = "identity") +
+subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))
+i=3
+subpop_v[[i]]
+site_plot <- c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North")
+
+grh2_p1 <- grh2 %>%
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site)) %>% # subpop == subpop_v[[i]] & 
+  ggplot(aes(x = phe_name, y = Estimate)) +
+  geom_bar(stat = "identity", aes(fill = phe_name == "50% greenup")) +
   geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
-  facet_wrap(~ site) +
-  geom_hline(data = filter(flh2, subpop == "Gulf" & !(site %in% c("NA", "nine_sites")) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + ylim(c(0, 1))
-save_plot("../manuscript/plots/h2_GR_site_weather_functions.png", plot = grh2_p1, base_height = 3.5, base_asp = 1.618)
+  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
+  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
+  facet_grid(site ~ subpop_name) +
+  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 1.1)) +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
+save_plot("../manuscript/plots/h2_GR_site_weather_functions_All_Sites_v2.png", plot = grh2_p1, base_height = 7, base_asp = 1.3)
+
+site_plot <- c( "All", "Texas", "North")
+
+grh2_p2 <- grh2 %>%
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site)) %>% # subpop == subpop_v[[i]] & 
+  ggplot(aes(x = phe_name, y = Estimate)) +
+  geom_bar(stat = "identity", aes(fill = phe_name == "50% greenup")) +
+  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
+  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
+  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
+  facet_grid(site ~ subpop_name) +
+  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.5)) +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
+save_plot("../manuscript/plots/h2_GR_site_weather_functions_All_Sites_v3.png", plot = grh2_p2, base_height = 5, base_asp = 1.5)
+
+site_plot <- c( "All")
+
+grh2_p3 <- grh2 %>%
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site)) %>% # subpop == subpop_v[[i]] & 
+  ggplot(aes(x = phe_name, y = Estimate)) +
+  geom_bar(stat = "identity", aes(fill = phe_name == "50% greenup")) +
+  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
+  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
+  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
+  facet_grid(site ~ subpop_name) +
+  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.15)) +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")
+save_plot("../manuscript/plots/h2_GR_site_weather_functions_All_Sites_v4.png", plot = grh2_p3, base_height = 4, base_asp = 1.5)
+
+gr_fl_p1 <- plot_grid(grh2_p3, flh2_p2, nrow = 2)
 ```
@@
 library(cowplot)
-```
 
-```{r}
 source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
@@
 ## Include just texas sites for greenup and just northern four sites.
+e.g.:
+for(p in 1){#seq_along(phe_dfs)){
+  phe_v <- phe_lists[[p]]
+  phe <- phe_dfs[[p]]
+
+  for(i in 12:13){#seq_along(site_v)){ # may need to adjust below for h2_table
+    for(j in 1:3){#seq_along(subpop_v)){
+      for(k in seq_along(phe_v)){
+
+## Model eight, texas, and north sites with a kinship : site interaction.
+Then, find GxE and G proportions of variance.
 ```{r}
@@
 
-for(p in 1){#seq_along(phe_dfs)){
+for(p in seq_along(phe_dfs)){
   phe_v <- phe_lists[[p]]
@@
 
-  for(i in 12:13){#seq_along(site_v)){
+  for(i in c(10,12:13)){#seq_along(site_v)){ # may need to adjust below for h2_table
     for(j in 1:3){#seq_along(subpop_v)){
@@
                        random = ~vm(PLANT_ID, k_full) +
-                         ~idv(SITE), #:vm(PLANT_ID, k_full),
+                         ~idv(SITE):vm(PLANT_ID, k_full),
                        residual = ~idv(units),
@@
 
-  if(i == 1 & j == 1 & k == 1){
+  if(i == 12 & j == 1 & k == 1){
     h2_table <- h2_est

### 2020-07-29 R/Plots_manuscript_v0.3.Rmd

@@
                               TRUE ~ phe))
-flh2$site <- factor(flh2$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "nine_sites"))
+flh2$site <- factor(flh2$site, levels = rev(c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "nine_sites")))
 flh2$phe_name <- factor(flh2$phe_name, levels = c("50% flowering", "daylength",
@@
                                                   "Rainfall, 1 day",
-                                                  "Rainfall, 2d sum",
+                                                  #"Rainfall, 2d sum",
                                                   "Rainfall, 3d sum",
-                                                  "Rainfall, 5d sum",
-                                                  "Rainfall, 7d sum"
+                                                  "Rainfall, 5d sum"#,
+                                                  #"Rainfall, 7d sum"
                                                   ))
 flh2$subpop_name <- factor(flh2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "all"))
-site_plot <- c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD", "All", "Texas", "North")
+site_plot <- rev(c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD", "All", "Texas", "North"))
+subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))
 
 flh2_p1 <- flh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site)) %>% # subpop == subpop_v[[i]] & 
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
   ggplot(aes(x = phe_name, y = Estimate)) +
@@
 
-save_plot("../manuscript/plots/h2_FL_site_weather_functions_All_Sites_v2.png", plot = flh2_p1, base_height = 7, base_asp = 1.3)
+save_plot("../manuscript/plots/h2_FL_site_weather_functions_All_Sites_v3.png", plot = flh2_p1, base_height = 7, base_asp = 1.1)
 
@@
 flh2_p2 <- flh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site)) %>% # subpop == subpop_v[[i]] & 
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
   ggplot(aes(x = phe_name, y = Estimate)) +
@@
   geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.5)) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.45)) +
   scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")
@@
                               phe == "FL50" ~ "50% flowering",
-                              phe == "cgdd_12c_jan2gr" ~ "GDD 01-01 to greenup",
+                              phe == "cgdd_12c_jan2gr" ~ "GDD 01-01 to GR",
                               phe == "cgdd_8c_jan2gr" ~ "GDD 8C 01-01 to greenup",
@@
                               TRUE ~ phe))
-grh2$site <- factor(grh2$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North", "nine_sites"))
+grh2$site <- factor(grh2$site, levels = rev(c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North", "nine_sites")))
 grh2$phe_name <- factor(grh2$phe_name, levels = c("50% greenup", "GDD, 5d",
                                                   "Min T, 5d mean",
-                                                  "Mean T, 5d mean",
-                                                  "Max T, 5d mean",
+                                                  #"Mean T, 5d mean",
+                                                  #"Max T, 5d mean",
                                                   "GDD, 10d",
                                                   "Min T, 10d mean",
-                                                  "Mean T, 10d mean",
-                                                  "Max T, 10d mean",
+                                                  #"Mean T, 10d mean",
+                                                  #"Max T, 10d mean",
                                                   "GDD, 18d",
                                                   "Min T, 18d mean",
-                                                  "Mean T, 18d mean",
-                                                  "Max T, 18d mean",
-                                                  "GDD 8C 01-01 to greenup",
-                                                  "GDD 01-01 to greenup"))
+                                                  #"Mean T, 18d mean",
+                                                  #"Max T, 18d mean",
+                                                  #"GDD 8C 01-01 to greenup",
+                                                  "GDD 01-01 to GR"))
 grh2$subpop_name <- factor(grh2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "All"))
@@
 grh2_p1 <- grh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site)) %>% # subpop == subpop_v[[i]] & 
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
   ggplot(aes(x = phe_name, y = Estimate)) +
@@
   scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
-save_plot("../manuscript/plots/h2_GR_site_weather_functions_All_Sites_v2.png", plot = grh2_p1, base_height = 7, base_asp = 1.3)
+save_plot("../manuscript/plots/h2_GR_site_weather_functions_All_Sites_v2.png", plot = grh2_p1, base_height = 7.5, base_asp = 1.1)
 
@@
 grh2_p2 <- grh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site)) %>% # subpop == subpop_v[[i]] & 
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
   ggplot(aes(x = phe_name, y = Estimate)) +
@@
   scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
-save_plot("../manuscript/plots/h2_GR_site_weather_functions_All_Sites_v3.png", plot = grh2_p2, base_height = 5, base_asp = 1.5)
+save_plot("../manuscript/plots/h2_GR_site_weather_functions_All_Sites_v3.png", plot = grh2_p2, base_height = 5, base_asp = 1.3)
 
@@
 grh2_p3 <- grh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site)) %>% # subpop == subpop_v[[i]] & 
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
   ggplot(aes(x = phe_name, y = Estimate)) +
@@
   geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.15)) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.11)) +
   scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")
@@
 gr_fl_p1 <- plot_grid(grh2_p3, flh2_p2, nrow = 2)
+
+save_plot("../manuscript/plots/h2_GR_and_FL_site_weather_functions_All_Sites.png", plot = gr_fl_p1, base_height = 4.6, base_asp = 1.35)
 ```
@@
 }
+```
+
+# D2F
+
+If flowering date frequently varies as a function of GDD in switchgrass, this explains observations that moving southern populations northwards delays flowering, and moving northern populations south hastens flowering (Sanderson et al 1996).
+< Indeed, we observed D2F for Midwest plants was shorter at TX than BRKG, and vice versa?>
+
+```{r}
+d2f_lat <- phenotypes %>%
+  mutate(D2F = FL50 - GR50) %>%
+  left_join(select(metadata, PLANT_ID, SUBPOP)) %>%
+  filter(!(SITE %in% c("OVTN", "FRMI")) & WHERE == "GWAS" & SUBPOP %in% c("Gulf", "Midwest")) %>%
+  group_by(SUBPOP, SITE) %>%
+  summarise(mean_D2F = mean(D2F, na.rm = TRUE)) %>%
+  left_join(select(sites, SITE, Latitude, manu_site)) %>%
+  arrange(SUBPOP, Latitude) 
+d2f_lat %>%
+  ggplot(aes(x = Latitude, y = mean_D2F)) +
+  geom_point(aes(color = SUBPOP))
+
+summary(lm(mean_D2F ~ Latitude*SUBPOP, data = d2f_lat))
+82-55
+```
+
+## G_BLUP GWAS
+
+```{r}
+gwasdir <- file.path(workingdir, "analysis", "gwas", "G_BLUP")
+log10p_gulf <- readRDS(file.path(gwasdir,                                 "log10p_strong_df_1000topSNPs_GR_FL_Gulf.rds"))
+log10p_midw <- readRDS(file.path(gwasdir,                                 "log10p_strong_df_1000topSNPs_GR_FL_Midwest.rds"))
+log10p_gandm <- readRDS(file.path(gwasdir,                                 "log10p_strong_df_1000topSNPs_GR_FL_Gulf_and_Midwest.rds"))
+```
+
+```{r}
+log10p_gulf %>%
+  pivot_longer(cols = -(SNP:POS), names_to = "PHE", values_to = "log10p") %>%
+  ungroup() %>%
+  filter(log10p > -log10(0.05/10300000)) %>%
+  group_by(PHE) %>%
+  tally() %>%
+  write_csv(., "../manuscript/tables/Gulf_SNPs_above_Bonferroni_by_trait.csv")
+```
+
+```{r}
+log10p_midw %>%
+  pivot_longer(cols = -(SNP:POS), names_to = "PHE", values_to = "log10p") %>%
+  ungroup() %>%
+  filter(log10p > -log10(0.05/8800000)) %>%
+  group_by(PHE) %>%
+  tally() %>%
+  write_csv(., "../manuscript/tables/Midwest_SNPs_above_Bonferroni_by_trait.csv")
+
+log10p_gandm %>%
+  pivot_longer(cols = -(SNP:POS), names_to = "PHE", values_to = "log10p") %>%
+  ungroup() %>%
+  filter(log10p > -log10(0.05/12100000)) %>%
+  group_by(PHE) %>%
+  tally() %>%
+  write_csv(., "../manuscript/tables/Gulf_and_Midwest_diversity_panel_SNPs_above_Bonferroni_by_trait.csv")
+```
+
+```{r}
+gulf_top <- readRDS(file.path(gwasdir, "Part-One-Output_Top-Effects-1000-SNPs_GR_FL_Gulf.rds"))
+
+gulf_top %>%
+  group_by(n_cond_in_top_set) %>%
+  tally()
+```

### 2020-07-29 data-raw/Data_Setup.Rmd

@@
   group_by(LINE, SITE, YEAR) %>%
-  mutate(PLOT_GL = paste0(str_sub(SITE, end = 1), str_sub(YEAR, start =3, end = 4), "_", row_number())) %>%
+  mutate(PLOT_GL = paste0(str_sub(SITE, end = 1), str_sub(YEAR, start = 3, 
+                                                          end = 4), "_", row_number())) %>%
   mutate(YEAR = as.numeric(YEAR),
@@
 phe_all %>%
-  group_by(PLOT_GL, SITE, PLANT_ID, YEAR) %>%
+  group_by(WHERE) %>%
   tally() %>% arrange((n))
@@
 cgdd_df <- wea_gr_fl2 %>%
-  group_by(SITE, PLANT_ID, PLOT_GL, GR50, FL50) %>%
+  group_by(WHERE, SITE, PLANT_ID, PLOT_GL, GR50, FL50) %>%
   summarise(CGDD_12C = sum(GDD_12C_SM),
@@
   ungroup()
+
+cumweadf %>%
+  group_by(WHERE) %>%
+  tally() %>% arrange((n))
+
 crain_df <- wea_gr_fl3 %>%
-  group_by(SITE, PLANT_ID, PLOT_GL, GR50, FL50) %>%
+  group_by(WHERE, SITE, PLANT_ID, PLOT_GL, GR50, FL50) %>%
   summarise(CGDD_12C_5d = sum(GDD_12C_SM),
@@
 ```
+
+```{r}
+fourway = cumweadf[cumweadf$WHERE == "FWCR",]
+df
+fourway
+```
+
 ### -------

### 2020-07-30 R/Plots_manuscript_v0.3.Rmd
Finish drafting first Results section v0.3 comparing heritabilities and GxE

@@
 library(GGally)
+library(maps)
+library(ggmap)
+usa <- map_data("usa")
 source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
@@
 phenotypes <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
+h2gr <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "h2_df_GR50_related_functions_of_weather_v1.csv"))
+h2fl <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "h2_df_FL50_related_functions_of_weather_v1.csv"))
+h2gr_gxe <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "gxe_and_h2_df_GR50_related_functions_of_weather_v1.csv"))
+h2fl_gxe <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "gxe_and_h2_df_FL50_related_functions_of_weather_v1.csv"))
+
+h2gr <- h2gr %>%
+  full_join(h2gr_gxe)
+h2fl <- h2fl %>%
+  full_join(h2fl_gxe)
+```
+### older files to load
 h2fl <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/heritability/h2_df_FL50_related_functions_of_weather_v3.csv")
@@
   full_join(h2gr4)
-```
 
@@
 
-# Heritability Plots
+# Map Fig 1A
 
-## h2 as functions of GR50 and FL50
 ```{r}
-h2all <- h2fl %>%
-  full_join(h2gr)
+subpops
+library(ggrepel)
 
-f1h2 <- h2all %>%
-  filter(phe %in% c("FL50", "GR50")) %>%
-  mutate(site = case_when(is.na(site) ~ "all",
-                          site == "eight_sites" ~ "all",
-                          TRUE ~ site),
-         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
-                                 subpop == "three_subpops" ~ "all",
-                                 TRUE ~ subpop),
-         phe_name = case_when(phe == "GR50" ~ "50% greenup",
-                              phe == "FL50" ~ "50% flowering",
-                              TRUE ~ phe))
-f1h2$site <- factor(f1h2$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "all", "nine_sites"))
-f1h2$phe_name <- factor(f1h2$phe_name, levels = c("50% greenup", "50% flowering"))
-f1h2$subpop_name <- factor(f1h2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "all"))
-ggh21 <- f1h2 %>%
-  filter(!(site %in% c("nine_sites", "IL")) & !(subpop %in% c("three_subpops", "Atlantic"))) %>%
-  ggplot(aes(x = site, y = Estimate)) +
-  geom_bar(aes(fill = site == "all"), stat = "identity") +
-  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
-  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
-  facet_grid(phe_name ~ subpop_name) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), 
-        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
-        legend.position = "none") + 
-  ylim(c(0, 1)) +
-  xlab("Common Gardens") + 
-  ylab(bquote('h'^2)) +
-  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
-  #
-save_plot("../manuscript/plots/h2_GR_FL_site_by_subpop.png", plot = ggh21, base_height = 3.5, base_asp = 1.618)
+metaplot <- metadata %>%
+  filter(SUBPOP %in% c("Gulf", "Midwest") & !is.na(LATITUDE)) %>%
+  mutate(LONGITUDE = -LONGITUDE)
+metaplot$ECOTYPE_NNET <- factor(metaplot$ECOTYPE_NNET, levels = c("Coastal", "Lowland", "Upland", "Unknown"))
+siteplot <- sites %>%
+  filter(!(SITE %in% c("FRMI", "OVTN")))
 
-f1h2 %>%
-  filter(!(site %in% c("all", "nine_sites", "IL")) & !(subpop %in% c("three_subpops", "Atlantic"))) %>%
-  ungroup() %>%
-  group_by(phe) %>%
-  summarise(mean = mean(Estimate))
-f1h2 %>%
-  filter((site %in% c("all")) & !(subpop %in% c("three_subpops", "Atlantic"))) %>%
-  ungroup() %>%
-  group_by(phe) %>%
-  summarise(mean = mean(Estimate))
+ggplot() + 
+  geom_polygon(data = usa, aes(x=long, y = lat, group = group), 
+               fill = "white", color = "black") +
+  coord_quickmap(xlim = c(-105, -66), ylim = c(25, 50)) +
+  geom_jitter(data = metaplot, aes(x = LONGITUDE, y = LATITUDE, 
+                                   color = SUBPOP, shape = ECOTYPE_NNET),
+              width = 0.45, height = 0.45) + 
+  geom_point(data = siteplot, aes(x = Longitude, y = Latitude), shape = 21, size = 4, stroke = 1.5, fill = "white") +
+  geom_point(data = siteplot, aes(x = Longitude, y = Latitude), shape = 4, size = 3, stroke = 1.5) +
+  geom_label_repel(data = siteplot, aes(x = Longitude, y = Latitude, label = manu_site), alpha = 0.7) +
+  labs(x = "Longitude", y = "Latitude") + 
+  theme(legend.position = "right") +
+  scale_color_manual(values = c("#F47F72", "#442C83")) +
+  scale_shape_manual(values = c(0, 6, 1, 4)) 
+  #scale_fill_manual(values = c("#280B54FF","#FAC127FF", "#D44842FF", "white")) +
+  #scale_color_manual(values = c(rep("black", 10), "#280B54FF","#FAC127FF", "#D44842FF", "white", "#280B54FF","#FAC127FF", "#D44842FF", "white"), guide = FALSE) +
+  #scale_alpha_manual(values = c(0.7, 0.7, 0.7, 0.7), guide = FALSE)
+save_plot(filename = "../manuscript/plots/Map_subpops_and_Ecotypes.png", plot = last_plot(), base_height = 3)
+?geom_point
 ```
 
-## h2 flowering weather
-```{r}
-flh2 <- h2fl %>%
-  filter(phe %in% c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d")) %>%
-  mutate(site = case_when(is.na(site) ~ "all",
-                          site == "eight_sites" ~ "All",
-                          TRUE ~ site),
-         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
-                                 subpop == "three_subpops" ~ "all",
-                                 TRUE ~ subpop),
-         phe_name = case_when(phe == "FL50" ~ "50% flowering",
-                              phe == "dyln_fl50" ~ "daylength",
-                              phe == "dyln_change_sec" ~ "daylength change (s)",
-                              phe == "cgdd_12c_gr2fl" ~ "GDD, GR to FL",
-                              phe == "crain_gr2fl" ~ "Rainfall, GR to FL",
-                              phe == "crain_1d" ~ "Rainfall, 1 day",
-                              phe == "crain_2d" ~ "Rainfall, 2d sum",
-                              phe == "crain_3d" ~ "Rainfall, 3d sum",
-                              phe == "crain_5d" ~ "Rainfall, 5d sum",
-                              phe == "crain_7d" ~ "Rainfall, 7d sum",
-                              TRUE ~ phe))
-flh2$site <- factor(flh2$site, levels = rev(c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "nine_sites")))
-flh2$phe_name <- factor(flh2$phe_name, levels = c("50% flowering", "daylength",
-                                                  "daylength change (s)",
-                                                  "GDD, GR to FL",
-                                                  "Rainfall, GR to FL",
-                                                  "Rainfall, 1 day",
-                                                  #"Rainfall, 2d sum",
-                                                  "Rainfall, 3d sum",
-                                                  "Rainfall, 5d sum"#,
-                                                  #"Rainfall, 7d sum"
-                                                  ))
-flh2$subpop_name <- factor(flh2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "all"))
-site_plot <- rev(c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD", "All", "Texas", "North"))
-subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))
-
-flh2_p1 <- flh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
-  ggplot(aes(x = phe_name, y = Estimate)) +
-  geom_bar(stat = "identity", aes(fill = phe_name == "50% flowering")) +
-  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
-  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
-  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
-  facet_grid(site ~ subpop_name) +
-  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 1.1)) +
-  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
-
-save_plot("../manuscript/plots/h2_FL_site_weather_functions_All_Sites_v3.png", plot = flh2_p1, base_height = 7, base_asp = 1.1)
-
-site_plot <- c("All", "Texas", "North")
-
-flh2_p2 <- flh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
-  ggplot(aes(x = phe_name, y = Estimate)) +
-  geom_bar(stat = "identity", aes(fill = phe_name == "50% flowering")) +
-  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
-  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
-  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
-  facet_grid(site ~ subpop_name) +
-  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.45)) +
-  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")
-
-save_plot("../manuscript/plots/h2_FL_site_weather_functions_All_Sites_v2.png", plot = flh2_p2, base_height = 4, base_asp = 1.5)
-```
-## h2 greenup weather
-```{r}
-h2gr
-```
-
-```{r}
-unique(h2gr$site)
-grh2 <- h2gr %>%
-  filter(!(phe %in% c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d", "DYLN"))) %>%
-  mutate(site = case_when(is.na(site) ~ "all",
-                          site == "eight_sites" ~ "All",
-                          site == "tx_sites" ~ "Texas",
-                          site == "north_sites" ~ "North",
-                          TRUE ~ site),
-         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
-                                 subpop == "three_subpops" ~ "All",
-                                 TRUE ~ subpop),
-         phe_name = case_when(phe == "GR50" ~ "50% greenup",
-                              phe == "FL50" ~ "50% flowering",
-                              phe == "cgdd_12c_jan2gr" ~ "GDD 01-01 to GR",
-                              phe == "cgdd_8c_jan2gr" ~ "GDD 8C 01-01 to greenup",
-                              phe == "tmax_18d" ~ "Max T, 18d mean",
-                              phe == "tmin_18d" ~ "Min T, 18d mean",
-                              phe == "tave_18d" ~ "Mean T, 18d mean",
-                              phe == "tmax_10d" ~ "Max T, 10d mean",
-                              phe == "tmin_10d" ~ "Min T, 10d mean",
-                              phe == "tave_10d" ~ "Mean T, 10d mean",
-                              phe == "tmax_5d" ~ "Max T, 5d mean",
-                              phe == "tmin_5d" ~ "Min T, 5d mean",
-                              phe == "tave_5d" ~ "Mean T, 5d mean",
-                              phe == "cgdd_12c_10d" ~ "GDD, 10d",
-                              phe == "cgdd_12c_18d" ~ "GDD, 18d",
-                              phe == "cgdd_12c_5d" ~ "GDD, 5d",
-                              TRUE ~ phe))
-grh2$site <- factor(grh2$site, levels = rev(c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North", "nine_sites")))
-grh2$phe_name <- factor(grh2$phe_name, levels = c("50% greenup", "GDD, 5d",
-                                                  "Min T, 5d mean",
-                                                  #"Mean T, 5d mean",
-                                                  #"Max T, 5d mean",
-                                                  "GDD, 10d",
-                                                  "Min T, 10d mean",
-                                                  #"Mean T, 10d mean",
-                                                  #"Max T, 10d mean",
-                                                  "GDD, 18d",
-                                                  "Min T, 18d mean",
-                                                  #"Mean T, 18d mean",
-                                                  #"Max T, 18d mean",
-                                                  #"GDD 8C 01-01 to greenup",
-                                                  "GDD 01-01 to GR"))
-grh2$subpop_name <- factor(grh2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "All"))
-
-subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))
-i=3
-subpop_v[[i]]
-site_plot <- c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North")
-
-grh2_p1 <- grh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
-  ggplot(aes(x = phe_name, y = Estimate)) +
-  geom_bar(stat = "identity", aes(fill = phe_name == "50% greenup")) +
-  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
-  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
-  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
-  facet_grid(site ~ subpop_name) +
-  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 1.1)) +
-  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
-save_plot("../manuscript/plots/h2_GR_site_weather_functions_All_Sites_v2.png", plot = grh2_p1, base_height = 7.5, base_asp = 1.1)
-
-site_plot <- c( "All", "Texas", "North")
-
-grh2_p2 <- grh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
-  ggplot(aes(x = phe_name, y = Estimate)) +
-  geom_bar(stat = "identity", aes(fill = phe_name == "50% greenup")) +
-  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
-  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
-  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
-  facet_grid(site ~ subpop_name) +
-  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.5)) +
-  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
-save_plot("../manuscript/plots/h2_GR_site_weather_functions_All_Sites_v3.png", plot = grh2_p2, base_height = 5, base_asp = 1.3)
-
-site_plot <- c( "All")
-
-grh2_p3 <- grh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
-  ggplot(aes(x = phe_name, y = Estimate)) +
-  geom_bar(stat = "identity", aes(fill = phe_name == "50% greenup")) +
-  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
-  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
-  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
-  facet_grid(site ~ subpop_name) +
-  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.11)) +
-  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")
-save_plot("../manuscript/plots/h2_GR_site_weather_functions_All_Sites_v4.png", plot = grh2_p3, base_height = 4, base_asp = 1.5)
-
-gr_fl_p1 <- plot_grid(grh2_p3, flh2_p2, nrow = 2)
-
-save_plot("../manuscript/plots/h2_GR_and_FL_site_weather_functions_All_Sites.png", plot = gr_fl_p1, base_height = 4.6, base_asp = 1.35)
-```
-## Phenotype correlations
+# -- Figure 1 B
+# Phenotype correlations
 can you easily look at the correlation between greenup ranks at OK versus NE?
@@
   select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
-  ggcorr(method = "pairwise", label = TRUE) +
+  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
   scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
@@
 pheno_corr <- plot_grid(pana, panb, panc, pand, pane, panf, nrow = 2)
-save_plot(filename = "../manuscript/plots/Phenotypic_correlations_GR50_FL50_2.png", plot = pheno_corr, base_height = 5)
+save_plot(filename = "../manuscript/plots/Phenotypic_correlations_GR50_FL50_2.svg", plot = pheno_corr, base_width = 6.5)
+```
+
+# --- Fig 1 C & D
+# Heritability Plots
+
+Within common gardens, h2 were typically quite high: 62% on average for greenup date, and 87% for flowering date.
+However, h2 were not consistently high, notably for greenup at our OK and NE common gardens.
+```{r}
+h2all <- h2fl %>%
+  full_join(h2gr)
+
+h2all %>%
+  filter(phe %in% c("FL50", "GR50") & site %in% c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD")) %>%
+  group_by(phe) %>%
+  summarise(meanh2 = mean(Estimate))
+
+h2all %>%
+  filter(phe %in% c("FL50", "GR50") & site %in% c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD")) %>%
+  filter(Estimate < 0.2)
+
+h2all %>%
+  filter(phe %in% c("FL50", "GR50") & site %in% c("eight_sites")) %>%
+  group_by(phe) %>%
+  summarise(meanh2 = mean(Estimate))
+
+h2all %>%
+  filter(phe %in% c("GR50") & site %in% c("tx_sites", "north_sites")) %>%
+  mutate(GandGxE = Estimate + GxE_Estimate) %>%
+  select(GandGxE, Estimate:phe, GxE_Estimate) %>%
+  group_by(site) %>%
+  summarise(meanh2 = mean(Estimate + GxE_Estimate))
+
+h2all %>%
+  filter(site %in% c("eight_sites")) %>%
+  mutate(GandGxE = Estimate + GxE_Estimate)  %>%
+  select(GandGxE, site:phe, Estimate, GxE_Estimate)
+```
+
+## h2 as functions of GR50 and FL50
+```{r}
+h2all <- h2fl %>%
+  full_join(h2gr)
+unique(h2all$site)
+f1h2 <- h2all %>%
+  filter(phe %in% c("FL50", "GR50")) %>%
+  mutate(site = case_when(is.na(site) ~ "all",
+                          site == "eight_sites" ~ "All",
+                          site == "tx_sites" ~ "Texas",
+                          site == "north_sites" ~ "North",
+                          TRUE ~ site),
+         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
+                                 subpop == "three_subpops" ~ "all",
+                                 TRUE ~ subpop),
+         phe_name = case_when(phe == "GR50" ~ "50% greenup",
+                              phe == "FL50" ~ "50% flowering",
+                              TRUE ~ phe))
+f1h2$site <- factor(f1h2$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All"))#, "Texas", "North"))
+f1h2$phe_name <- factor(f1h2$phe_name, levels = c("50% greenup", "50% flowering"))
+f1h2$subpop_name <- factor(f1h2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "all"))
+ggh21 <- f1h2 %>%
+  filter(!(site %in% c("nine_sites", "IL")) & !is.na(subpop_name) & !is.na(phe_name) & !is.na(site)) %>%
+  ggplot(aes(x = site, y = Estimate)) +
+  geom_bar(aes(fill = site %in% c("All", "Texas", "North")), stat = "identity") +
+  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
+  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
+  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
+  facet_grid(phe_name ~ subpop_name) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), 
+        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
+        legend.position = "none") + 
+  ylim(c(0, 1)) +
+  xlab("Common Gardens") + 
+  ylab(bquote('h'^2)) +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
+  #
+save_plot("../manuscript/plots/h2_GR_FL_site_by_subpop_k_specific.png", plot = ggh21, base_width = 5.5, base_height = 3)
+```
+
+## h2 flowering weather
+```{r}
+flh2 <- h2fl %>%
+  filter(phe %in% c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d")) %>%
+  mutate(site = case_when(is.na(site) ~ "all",
+                          site == "eight_sites" ~ "All",
+                          site == "tx_sites" ~ "Texas",
+                          site == "north_sites" ~ "North",
+                          TRUE ~ site),
+         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
+                                 subpop == "three_subpops" ~ "all",
+                                 TRUE ~ subpop),
+         phe_name = case_when(phe == "FL50" ~ "50% flowering",
+                              phe == "dyln_fl50" ~ "daylength",
+                              phe == "dyln_change_sec" ~ "daylength change (s)",
+                              phe == "cgdd_12c_gr2fl" ~ "GDD, GR to FL",
+                              phe == "crain_gr2fl" ~ "Rainfall, GR to FL",
+                              phe == "crain_1d" ~ "Rainfall, 1 day",
+                              phe == "crain_2d" ~ "Rainfall, 2d sum",
+                              phe == "crain_3d" ~ "Rainfall, 3d sum",
+                              phe == "crain_5d" ~ "Rainfall, 5d sum",
+                              phe == "crain_7d" ~ "Rainfall, 7d sum",
+                              TRUE ~ phe))
+flh2$site <- factor(flh2$site, levels = rev(c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North")))
+flh2$phe_name <- factor(flh2$phe_name, levels = c("50% flowering", "daylength",
+                                                  "daylength change (s)",
+                                                  "GDD, GR to FL",
+                                                  "Rainfall, GR to FL",
+                                                  "Rainfall, 1 day",
+                                                  #"Rainfall, 2d sum",
+                                                  "Rainfall, 3d sum",
+                                                  "Rainfall, 5d sum"#,
+                                                  #"Rainfall, 7d sum"
+                                                  ))
+flh2$subpop_name <- factor(flh2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "all"))
+site_plot <- rev(c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD", "All", "Texas", "North"))
+subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))
+
+flh2_p1 <- flh2 %>%
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
+  ggplot(aes(x = phe_name, y = Estimate)) +
+  geom_bar(stat = "identity", aes(fill = phe_name == "50% flowering")) +
+  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
+  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
+  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
+  facet_grid(site ~ subpop_name) +
+  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 1.1)) +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
+
+save_plot("../manuscript/plots/h2_FL_site_weather_functions_single_sites_k_specific.png", plot = flh2_p1, base_height = 7, base_asp = 1.1)
+
+site_plot <- c("All", "Texas", "North")
+
+flh2_p2 <- flh2 %>%
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
+  ggplot(aes(x = phe_name, y = Estimate)) +
+  geom_bar(stat = "identity", aes(fill = phe_name == "50% flowering")) +
+  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
+  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
+  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
+  facet_grid(site ~ subpop_name) +
+  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.8)) +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")
+
+save_plot("../manuscript/plots/h2_FL_site_weather_functions_site_groups_k_specific.png", plot = flh2_p2, base_height = 4, base_asp = 1.5)
+
+site_plot <- c("All")
+
+flh2_p3 <- flh2 %>%
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
+  ggplot(aes(x = phe_name, y = Estimate)) +
+  geom_bar(stat = "identity", aes(fill = phe_name == "50% flowering")) +
+  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
+  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
+  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
+  facet_grid(site ~ subpop_name) +
+  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.5)) +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")
+save_plot("../manuscript/plots/h2_FL_site_weather_functions_All_Sites_k_specific.png", plot = flh2_p3, base_height = 3.5)
+```
+## gxe flowering all, north, texas
+```{r}
+site_plot <- c("All", "Texas", "North")
+
+fl_barplot2 <- flh2 %>%
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>%
+  select(site:phe, subpop_name, phe_name, ends_with("SE")) %>%
+  pivot_longer(cols = ends_with("SE"), names_to = "Type", values_to = "SE") %>%
+  mutate(type_name = case_when(Type == "SE" ~ "G",
+                               Type == "GxE_SE" ~ "GxE",
+                               Type == "error_SE" ~ "error",
+                               Type == "Env_SE" ~ "E")) %>%
+  select(-Type)
+fl_barplot <- flh2 %>%
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>%
+  select(site:phe, subpop_name, phe_name, ends_with("Estimate")) %>%
+  pivot_longer(cols = ends_with("Estimate"), names_to = "Type", values_to = "Estimate") %>%
+  mutate(type_name = case_when(Type == "Estimate" ~ "G",
+                               Type == "GxE_Estimate" ~ "GxE",
+                               Type == "error_Estimate" ~ "error",
+                               Type == "Env_Estimate" ~ "E"))
+fl_barplot <- fl_barplot %>%
+  left_join(fl_barplot2)
+fl_barplot$type_name <- factor(fl_barplot$type_name, levels = rev(c("G", "GxE", "E", "error")))
+
+fl_bar_gxe <- fl_barplot %>%
+  ggplot(aes(x = phe_name, y = Estimate)) +
+  geom_bar(stat = "identity", aes(fill = type_name)) + 
+  geom_errorbar(data = filter(fl_barplot, type_name == "G"), aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
+  facet_grid(site ~ subpop_name)+
+  scale_fill_viridis(discrete = TRUE, direction = -1) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") +
+  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = GxE_Estimate+Estimate), linetype = 2) + xlab("") +
+  ylab("Variance Explained")
+
+save_plot("../manuscript/plots/Var_Comp_flowering_site_groups.png", plot = fl_bar_gxe, base_width = 5.5, base_height = 4)
+
+flh2_p4 <- flh2 %>%
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
+  ggplot(aes(x = phe_name, y = GxE_Estimate)) +
+  geom_bar(stat = "identity", aes(fill = phe_name == "50% flowering")) +
+  geom_errorbar(aes(ymin = GxE_Estimate - GxE_SE, ymax = GxE_Estimate + GxE_SE), width = 0.3) +
+  geom_linerange(aes(ymin = GxE_Estimate, ymax = GxE_Estimate + GxE_SE)) +
+  geom_linerange(aes(ymin = GxE_Estimate - GxE_SE, ymax = GxE_Estimate)) +
+  facet_grid(site ~ subpop_name) +
+  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = GxE_Estimate), linetype = 2) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 1)) +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")
+
+save_plot("../manuscript/plots/h2_FL_site_weather_functions_site_groups_k_specific.png", plot = flh2_p2, base_height = 4, base_asp = 1.5)
+
+library(scales)
+show_col(viridis(4)) # #440154FF   #31688EFF
+```
+
+## h2 greenup weather
+
+```{r}
+unique(h2gr$site)
+grh2 <- h2gr %>%
+  filter(!(phe %in% c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d", "DYLN"))) %>%
+  mutate(site = case_when(is.na(site) ~ "all",
+                          site == "eight_sites" ~ "All",
+                          site == "tx_sites" ~ "Texas",
+                          site == "north_sites" ~ "North",
+                          TRUE ~ site),
+         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
+                                 subpop == "three_subpops" ~ "All",
+                                 TRUE ~ subpop),
+         phe_name = case_when(phe == "GR50" ~ "50% greenup",
+                              phe == "FL50" ~ "50% flowering",
+                              phe == "cgdd_12c_jan2gr" ~ "GDD 01-01 to GR",
+                              phe == "cgdd_8c_jan2gr" ~ "GDD 8C 01-01 to greenup",
+                              phe == "tmax_18d" ~ "Max T, 18d mean",
+                              phe == "tmin_18d" ~ "Min T, 18d mean",
+                              phe == "tave_18d" ~ "Mean T, 18d mean",
+                              phe == "tmax_10d" ~ "Max T, 10d mean",
+                              phe == "tmin_10d" ~ "Min T, 10d mean",
+                              phe == "tave_10d" ~ "Mean T, 10d mean",
+                              phe == "tmax_5d" ~ "Max T, 5d mean",
+                              phe == "tmin_5d" ~ "Min T, 5d mean",
+                              phe == "tave_5d" ~ "Mean T, 5d mean",
+                              phe == "cgdd_12c_10d" ~ "GDD, 10d",
+                              phe == "cgdd_12c_18d" ~ "GDD, 18d",
+                              phe == "cgdd_12c_5d" ~ "GDD, 5d",
+                              TRUE ~ phe))
+grh2$site <- factor(grh2$site, levels = rev(c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North", "nine_sites")))
+grh2$phe_name <- factor(grh2$phe_name, levels = c("50% greenup", "GDD, 5d",
+                                                  "Min T, 5d mean",
+                                                  #"Mean T, 5d mean",
+                                                  #"Max T, 5d mean",
+                                                  "GDD, 10d",
+                                                  "Min T, 10d mean",
+                                                  #"Mean T, 10d mean",
+                                                  #"Max T, 10d mean",
+                                                  "GDD, 18d",
+                                                  "Min T, 18d mean",
+                                                  #"Mean T, 18d mean",
+                                                  #"Max T, 18d mean",
+                                                  #"GDD 8C 01-01 to greenup",
+                                                  "GDD 01-01 to GR"))
+grh2$subpop_name <- factor(grh2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "All"))
+
+subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))
+
+site_plot <- c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North")
+
+grh2_p1 <- grh2 %>%
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
+  ggplot(aes(x = phe_name, y = Estimate)) +
+  geom_bar(stat = "identity", aes(fill = phe_name == "50% greenup")) +
+  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
+  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
+  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
+  facet_grid(site ~ subpop_name) +
+  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 1.1)) +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
+save_plot("../manuscript/plots/h2_GR_site_weather_functions_single_sites_k_specific.png", plot = grh2_p1, base_height = 7.5, base_asp = 1.1)
+
+site_plot <- c( "All", "Texas", "North")
+
+grh2_p2 <- grh2 %>%
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
+  ggplot(aes(x = phe_name, y = Estimate)) +
+  geom_bar(stat = "identity", aes(fill = phe_name == "50% greenup")) +
+  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
+  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
+  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
+  facet_grid(site ~ subpop_name) +
+  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.6)) +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
+save_plot("../manuscript/plots/h2_GR_site_weather_functions_site_groups_k_specific.png", plot = grh2_p2, base_height = 5, base_asp = 1.3)
+
+site_plot <- c( "All")
+
+grh2_p3 <- grh2 %>%
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
+  ggplot(aes(x = phe_name, y = Estimate)) +
+  geom_bar(stat = "identity", aes(fill = phe_name == "50% greenup")) +
+  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
+  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
+  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
+  facet_grid(site ~ subpop_name) +
+  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.11)) +
+  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")
+save_plot("../manuscript/plots/h2_GR_site_weather_functions_All_Sites_k_specific.png", plot = grh2_p3, base_height = 3.5)
+
+gr_fl_p1 <- plot_grid(grh2_p3, flh2_p3, nrow = 2)
+
+save_plot("../manuscript/plots/h2_GR_and_FL_site_weather_functions_All_Sites_k_specific.png", plot = gr_fl_p1, base_height = 4.6)
+```
+## gxe greenup
+
+```{r}
+site_plot <- c("All", "Texas", "North")
+
+gr_barplot2 <- grh2 %>%
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>%
+  select(site:phe, subpop_name, phe_name, ends_with("SE")) %>%
+  pivot_longer(cols = ends_with("SE"), names_to = "Type", values_to = "SE") %>%
+  mutate(type_name = case_when(Type == "SE" ~ "G",
+                               Type == "GxE_SE" ~ "GxE",
+                               Type == "error_SE" ~ "error",
+                               Type == "Env_SE" ~ "E")) %>%
+  select(-Type)
+gr_barplot <- grh2 %>%
+  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>%
+  select(site:phe, subpop_name, phe_name, ends_with("Estimate")) %>%
+  pivot_longer(cols = ends_with("Estimate"), names_to = "Type", values_to = "Estimate") %>%
+  mutate(type_name = case_when(Type == "Estimate" ~ "G",
+                               Type == "GxE_Estimate" ~ "GxE",
+                               Type == "error_Estimate" ~ "error",
+                               Type == "Env_Estimate" ~ "E"))
+gr_barplot <- gr_barplot %>%
+  left_join(gr_barplot2)
+gr_barplot$type_name <- factor(gr_barplot$type_name, levels = rev(c("G", "GxE", "E", "error")))
+
+gr_bar_gxe <- gr_barplot %>%
+  ggplot(aes(x = phe_name, y = Estimate)) +
+  geom_bar(stat = "identity", aes(fill = type_name)) + 
+  geom_errorbar(data = filter(gr_barplot, type_name == "G"), aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
+  facet_grid(site ~ subpop_name)+
+  scale_fill_viridis(discrete = TRUE, direction = -1) +
+  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "right") +
+  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = GxE_Estimate+Estimate), linetype = 2)
+
+save_plot("../manuscript/plots/Var_Comp_greenup_site_groups.png", plot = gr_bar_gxe, base_height = 4, base_asp = 1.5)
 ```
@@
 k_full <- read_rds("../data/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
+k_gandm <- read_rds("../data/Kinship_van_Raden_method_363g_Gulf_and_Midwest_subset_maf0.05.rds")
+k_midw <- read_rds("../data/Kinship_van_Raden_method_134g_Midwest_subset_maf0.05.rds")
+k_gulf <- read_rds("../data/Kinship_van_Raden_method_229g_Gulf_subset_maf0.05.rds")
 ```
@@
 
+## k_gandm
+
+```{r}
+b1 <- svd(k_gandm)
+b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 29.2
+b1$d[2] / sum(b1$d) # 18.2
+b1$d[3] / sum(b1$d) # 3.7
+b1$d[4] / sum(b1$d) # 1.4
+b1$d[5] / sum(b1$d) # 0.008
+sum(b1$d[1:4]) / sum(b1$d) # First 4 eigenvectors explain 1% or more variance, and 52.5% of total variance.
+
+for(g in 1:5){
+barplot(b1$v[,g]/k_gandm[which.max(abs(k_gandm[,g])),g])
+}
+order <- enframe(colnames(k_gandm), name = NULL, value = "PLANT_ID")
+ggb1 <- tibble(`First eigenvector (PVE = 29.2%)` = b1$v[,1]/k_gandm[which.max(abs(k_gandm[,1])),1],
+       `Second eigenvector (PVE = 18.2%)` = b1$v[,2]/k_gandm[which.max(abs(k_gandm[,2])),2],
+       `Third eigenvector (PVE = 3.7%)` = b1$v[,3]/k_gandm[which.max(abs(k_gandm[,3])),3],
+       `Fourth eigenvector (PVE = 1.4%)` = b1$v[,4]/k_gandm[which.max(abs(k_gandm[,4])),4],
+       #`Fifth eigenvector (PVE = 0.8%)` = b1$v[,5]/k_gandm[which.max(abs(k_gandm[,5])),5],
+       PLANT_ID = order$PLANT_ID) %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  dplyr::select(PLANT_ID, SUBPOP, ECOTYPE_NNET, LATITUDE, ends_with("%)"))
+
+ggb1 <- ggb1 %>%
+  arrange(LATITUDE) %>%
+  mutate(rank = row_number()) %>%
+  pivot_longer(cols = ends_with("%)"), names_to = "Eigenvector", values_to = "value")
+ggb1$Eigenvector <- factor(ggb1$Eigenvector, 
+                           levels = c("First eigenvector (PVE = 29.2%)", 
+                                      "Second eigenvector (PVE = 18.2%)", 
+                                      "Third eigenvector (PVE = 3.7%)", 
+                                      "Fourth eigenvector (PVE = 1.4%)",
+                                      "Fifth eigenvector (PVE = 0.8%)"
+                                      ))
+ggb1$ECOTYPE_NNET <- factor(ggb1$ECOTYPE_NNET, levels = c("Coastal", "Lowland", "Upland", "Unknown"))
+ggb1_plot <- ggb1 %>%
+  ggplot(aes(x = rank, y = value)) +
+  geom_bar(aes(color = SUBPOP, fill = SUBPOP), stat = "identity") +
+  facet_wrap(~ Eigenvector) +
+  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
+  scale_color_manual(values = c("#F47F72", "#442C83")) +
+  scale_fill_manual(values = c("#F47F72", "#442C83")) +
+  xlab("Latitude of Origin (Rank Order)") + ylab("") +
+  geom_vline(xintercept = 342, linetype = 2)
+save_plot("../manuscript/plots/SVD_of_Kinship_matrix_Gulf_and_Midwest_first_four_eigenvectors.png", base_height = 5, base_asp = 2, plot = ggb1_plot)
+
+ggb1_plot2 <- ggb1 %>%
+  ggplot(aes(x = rank, y = value)) +
+  geom_bar(aes(color = ECOTYPE_NNET, fill = ECOTYPE_NNET), stat = "identity") +
+  facet_wrap(~ Eigenvector) +
+  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
+  scale_color_viridis(discrete = TRUE, option = "B", end = 0.9, 
+                      direction = -1) +
+  scale_fill_viridis(discrete = TRUE, option = "B", end = 0.9, direction = -1) +
+  xlab("Latitude of Origin (Rank Order)") + ylab("") +
+  geom_vline(xintercept = 342, linetype = 2)
+save_plot("../manuscript/plots/SVD_of_Kinship_matrix_Gulf_and_Midwest_first_four_eigenvectors_ecotype.png", base_height = 5, base_asp = 2, plot = ggb1_plot2)
+```
+
+## k_midw
+
+```{r}
+b1 <- svd(k_midw)
+b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 10.2
+b1$d[2] / sum(b1$d) # 4.8
+b1$d[3] / sum(b1$d) # 2.4
+b1$d[4] / sum(b1$d) # 2.0
+b1$d[5] / sum(b1$d) # 1.7
+b1$d[6] / sum(b1$d) # 1.3
+b1$d[7] / sum(b1$d) # 1.3
+b1$d[8] / sum(b1$d) # 1.1
+b1$d[9] / sum(b1$d) # 1.1
+b1$d[10] / sum(b1$d) # 1.1
+b1$d[11] / sum(b1$d) # 1.1
+b1$d[12] / sum(b1$d) # 1.0
+b1$d[13] / sum(b1$d) # 1.0
+sum(b1$d[1:13]) / sum(b1$d) # First 13 eigenvectors explain 1% or more variance, and 30% of total variance. Phew! Exhausting.
+```
+
+```{r}
+k_use <- k_midw
+for(g in 1:5){
+barplot(b1$v[,g]/k_use[which.max(abs(k_use[,g])),g])
+}
+order <- enframe(colnames(k_use), name = NULL, value = "PLANT_ID")
+ggb1 <- tibble(`First eigenvector (PVE = 10.2%)` = b1$v[,1]/k_use[which.max(abs(k_use[,1])),1],
+       `Second eigenvector (PVE = 4.8%)` = b1$v[,2]/k_use[which.max(abs(k_use[,2])),2],
+       `Third eigenvector (PVE = 2.4%)` = b1$v[,3]/k_use[which.max(abs(k_use[,3])),3],
+       `Fourth eigenvector (PVE = 2.0%)` = b1$v[,4]/k_use[which.max(abs(k_use[,4])),4],
+       `Fifth eigenvector (PVE = 1.7%)` = b1$v[,5]/k_use[which.max(abs(k_use[,5])),5],
+       `Sixth eigenvector (PVE = 1.3%)` = b1$v[,6]/k_use[which.max(abs(k_use[,6])),6],
+       PLANT_ID = order$PLANT_ID) %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  dplyr::select(PLANT_ID, SUBPOP, ECOTYPE_NNET, LATITUDE, ends_with("%)"))
+
+ggb1 <- ggb1 %>%
+  arrange(LATITUDE) %>%
+  mutate(rank = row_number()) %>%
+  pivot_longer(cols = ends_with("%)"), names_to = "Eigenvector", values_to = "value")
+ggb1$Eigenvector <- factor(ggb1$Eigenvector, 
+                           levels = c("First eigenvector (PVE = 10.2%)", 
+                                      "Second eigenvector (PVE = 4.8%)", 
+                                      "Third eigenvector (PVE = 2.4%)", 
+                                      "Fourth eigenvector (PVE = 2.0%)",
+                                      "Fifth eigenvector (PVE = 1.7%)",
+                                      "Sixth eigenvector (PVE = 1.3%)"
+                                      ))
+ggb1$ECOTYPE_NNET <- factor(ggb1$ECOTYPE_NNET, levels = c("Coastal", "Lowland", "Upland", "Unknown"))
+
+ggb1_plot2 <- ggb1 %>%
+  ggplot(aes(x = rank, y = value)) +
+  geom_bar(aes(color = ECOTYPE_NNET, fill = ECOTYPE_NNET), stat = "identity") +
+  facet_wrap(~ Eigenvector) +
+  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
+  scale_color_viridis(discrete = TRUE, option = "B", end = 0.9, 
+                      direction = -1) +
+  scale_fill_viridis(discrete = TRUE, option = "B", end = 0.9, direction = -1) +
+  xlab("Latitude of Origin (Rank Order)") + ylab("") +
+  geom_vline(xintercept = 133, linetype = 2)
+save_plot("../manuscript/plots/SVD_of_Kinship_matrix_Midwest_first_six_eigenvectors_ecotype.png", base_height = 5, base_asp = 2, plot = ggb1_plot2)
+```
+## k_gulf
+
+```{r}
+b1 <- svd(k_gulf)
+b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 22.5
+b1$d[2] / sum(b1$d) # 9.4
+b1$d[3] / sum(b1$d) # 3.1
+b1$d[4] / sum(b1$d) # 1.8
+b1$d[5] / sum(b1$d) # 1.6
+b1$d[6] / sum(b1$d) # 1.3
+b1$d[7] / sum(b1$d) # 1.1
+b1$d[8] / sum(b1$d) # 1.0
+sum(b1$d[1:8]) / sum(b1$d) # First 8 eigenvectors explain 1% or more variance, and 41.8% of total variance. 
+```
+
+```{r}
+k_use <- k_gulf
+for(g in 1:8){
+barplot(b1$v[,g]/k_use[which.max(abs(k_use[,g])),g])
+}
+order <- enframe(colnames(k_use), name = NULL, value = "PLANT_ID")
+ggb1 <- tibble(`First eigenvector (PVE = 22.5%)` = b1$v[,1]/k_use[which.max(abs(k_use[,1])),1],
+       `Second eigenvector (PVE = 9.4%)` = b1$v[,2]/k_use[which.max(abs(k_use[,2])),2],
+       `Third eigenvector (PVE = 3.1%)` = b1$v[,3]/k_use[which.max(abs(k_use[,3])),3],
+       `Fourth eigenvector (PVE = 1.8%)` = b1$v[,4]/k_use[which.max(abs(k_use[,4])),4],
+       `Fifth eigenvector (PVE = 1.6%)` = b1$v[,5]/k_use[which.max(abs(k_use[,5])),5],
+       `Sixth eigenvector (PVE = 1.3%)` = b1$v[,6]/k_use[which.max(abs(k_use[,6])),6],
+       PLANT_ID = order$PLANT_ID) %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  dplyr::select(PLANT_ID, SUBPOP, ECOTYPE_NNET, LATITUDE, ends_with("%)"))
+
+ggb1 <- ggb1 %>%
+  arrange(LATITUDE) %>%
+  mutate(rank = row_number()) %>%
+  pivot_longer(cols = ends_with("%)"), names_to = "Eigenvector", values_to = "value")
+ggb1$Eigenvector <- factor(ggb1$Eigenvector, 
+                           levels = c("First eigenvector (PVE = 22.5%)", 
+                                      "Second eigenvector (PVE = 9.4%)", 
+                                      "Third eigenvector (PVE = 3.1%)", 
+                                      "Fourth eigenvector (PVE = 1.8%)",
+                                      "Fifth eigenvector (PVE = 1.6%)",
+                                      "Sixth eigenvector (PVE = 1.3%)"
+                                      ))
+ggb1$ECOTYPE_NNET <- factor(ggb1$ECOTYPE_NNET, levels = c("Coastal", "Lowland", "Upland", "Unknown"))
+
+ggb1_plot2 <- ggb1 %>%
+  ggplot(aes(x = rank, y = value)) +
+  geom_bar(aes(color = ECOTYPE_NNET, fill = ECOTYPE_NNET), stat = "identity") +
+  facet_wrap(~ Eigenvector) +
+  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
+  scale_color_viridis(discrete = TRUE, option = "B", end = 0.9, 
+                      direction = -1) +
+  scale_fill_viridis(discrete = TRUE, option = "B", end = 0.9, direction = -1) +
+  xlab("Latitude of Origin (Rank Order)") + ylab("") +
+  geom_vline(xintercept = 210, linetype = 2)
+save_plot("../manuscript/plots/SVD_of_Kinship_matrix_Gulf_first_six_eigenvectors_ecotype.png", base_height = 5, base_asp = 2, plot = ggb1_plot2)
+```
+
 The eigenvectors capture the predominant patterns in the Uk3 covariance matrix.
@@
 WAIT. I could also plot this for the svd from the whole SNP set. That might make more sense than doing a svd on the kinship matrix...
+
+```{r}
+snp <- snp_attach("~/Github/pvdiv-genome/genetic_subpops/kinship/Pvirgatum_V5_GWAS_134g_Midwest_K_subset_maf0.05.rds")
+
+```
 
@@
 
+# ASReml 07-29
+
+## Model eight, texas, and north sites with a kinship : site interaction.
+Then, find GxE and G proportions of variance.
+## Use Midwest-specific kinship matrix for Midwest-only, and Gulf-specific kinship matrix for Gulf-only comparisons.
+And use (and visualize eigenvectors of?
+Midwest and Gulf kinship matrix for the comparisons with both.)
+```{r}
+library(tidyverse)
+library(asreml)
+library(switchgrassGWAS)
+library(cowplot)
+
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "analysis", "heritability", "kinship_subpop")
+metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
+sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
+subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
+phe_gr_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
+phe_fl_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
+
+kinshipdir <- file.path(workingdir, "data")
+k_gandm <- read_rds(file.path(kinshipdir, "Kinship_van_Raden_method_363g_Gulf_and_Midwest_subset_maf0.05.rds"))
+k_midw <- read_rds(file.path(kinshipdir,
+                "Kinship_van_Raden_method_134g_Midwest_subset_maf0.05.rds"))
+k_gulf <- read_rds(file.path(kinshipdir, "Kinship_van_Raden_method_229g_Gulf_subset_maf0.05.rds"))
+
+```
+
+```{r}
+site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
+               NE = "NE", MI = "MI", SD = "SD", eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
+subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"))
+phe_fl <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl",
+           "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")
+phe_gr <- c("GR50", "DYLN", "cgdd_12c_jan2gr", "tmin_18d", "cgdd_12c_18d",
+            "tmin_10d", "cgdd_12c_10d", "tmin_5d", "cgdd_12c_5d")
+phe_dfs <- list(phe_gr_df, phe_fl_df)
+phe_lists <- list(phe_gr, phe_fl)
+k_list <- list(Gulf = k_gulf, Midwest = k_midw, Gulf_and_Midwest = k_gandm)
+h2_table <- tibble()
+gxe_table <- tibble()
+
+for(p in seq_along(phe_dfs)){
+  phe_v <- phe_lists[[p]]
+  phe <- phe_dfs[[p]]
+
+  for(i in seq_along(site_v)){ # may need to adjust below for h2_table
+    dir.create(file.path(outputdir, names(site_v)[i]), showWarnings = FALSE)
+    for(j in seq_along(subpop_v)){
+      k_use <- k_list[[j]]
+      for(k in seq_along(phe_v)){
+        phe_single <- phe %>%
+          left_join(sites, by = "SITE") %>%
+          filter(manu_site %in% site_v[[i]] & SUBPOP %in% subpop_v[[j]]) %>%
+          rename(phenotype = eval(phe_v[k])) %>%
+          mutate(PLANT_ID = as_factor(PLANT_ID),
+                 SITE = as_factor(SITE),
+                 PLOT_GL = as_factor(PLOT_GL)) %>%
+          as.data.frame()
+
+  if(length(site_v[[i]]) > 1){  # need to add site as a random factor if 
+    # there is more than one site included in the dataset.
+    tryCatch({
+     asr_out <- asreml(phenotype ~ 1,
+                       random = ~idv(SITE)*vm(PLANT_ID, k_use),
+                       residual = ~idv(units),
+                       data = phe_single, 
+                       workspace = "3gb")
+
+    Env_est <- vpredict(asr_out, E ~ V1/(V1+V2+V3+V4))
+    GxE_est <- vpredict(asr_out, GxE_h2 ~ V3/(V1+V2+V3+V4))
+    err_est <- vpredict(asr_out, err ~ V4/(V1+V2+V3+V4))
+    h2_est <- vpredict(asr_out, G_h2 ~ V2/(V1+V2+V3+V4)) %>% as_tibble() %>%
+      mutate(site = names(site_v)[i],
+             subpop = names(subpop_v)[j],
+             phe = phe_v[k],
+             bic = summary(asr_out)$bic[1],
+             loglik = summary(asr_out)$loglik,
+             Env_Estimate = Env_est$Estimate[1],
+             Env_SE = Env_est$SE[1],
+             GxE_Estimate = GxE_est$Estimate[1],
+             GxE_SE = GxE_est$SE[1],
+             error_Estimate = err_est$Estimate[1],
+             error_SE = err_est$SE[1])
+
+    blup <- summary(asr_out, coef=TRUE)$coef.random %>% 
+    as_tibble(rownames = "Effect")
+    write_csv(blup, file.path(outputdir, names(site_v)[i], 
+                              paste0("blups_", names(site_v)[i], "_",
+                                     names(subpop_v)[j], "_", phe_v[k],
+                                     ".csv")))
+    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+  } else {
+    tryCatch({
+    asr_out <- asreml(phenotype ~ 1,
+                       random = ~vm(PLANT_ID, k_use),
+                       residual = ~idv(units),
+                       data = phe_single, 
+                       workspace = "3gb")
+
+    h2_est <- vpredict(asr_out, h2 ~ V1/(V1+V2)) %>% as_tibble() %>%
+      mutate(site = names(site_v)[i],
+             subpop = names(subpop_v)[j],
+             phe = phe_v[k],
+             bic = summary(asr_out)$bic[1],
+             loglik = summary(asr_out)$loglik)
+    blup <- summary(asr_out, coef=TRUE)$coef.random %>% 
+    as_tibble(rownames = "Effect")
+    write_csv(blup, file.path(outputdir, names(site_v)[i],
+                              paste0("blups_", names(site_v)[i], "_",
+                                     names(subpop_v)[j], "_", phe_v[k],
+                                     ".csv")))
+  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+  }
+
+  if(i == 1 & j == 1 & k == 1){
+    h2_table <- h2_est
+  } else if(i == 9 & j == 1 & k == 1){
+    gxe_table <- h2_est
+  } else if(length(site_v[[i]]) > 1){
+    gxe_table <- add_row(gxe_table, h2_est)
+  } else {
+    h2_table <- add_row(h2_table, h2_est)
+  }
+  tryCatch({
+      save_plot(filename=file.path(outputdir, names(site_v)[i],
+                                   paste0("ASReml_performance_",
+                                                 names(site_v)[i], "_",
+                                                 names(subpop_v)[j],
+                                                 "_", phe_v[k], ".png")), 
+            plot = plot(asr_out))
+    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+
+      }
+    }
+  }
+  write_csv(h2_table, file.path(outputdir, paste0("h2_df_", phe_v[1],
+                                                  "_related_functions_", 
+                                                  "of_weather_v1.csv")))
+  write_csv(gxe_table, file.path(outputdir, paste0("gxe_and_h2_df_", phe_v[1],
+                                                  "_related_functions_", 
+                                                  "of_weather_v1.csv")))
+}
+```
+
 # D2F
@@
   tally()
+
+gulf_top_manh2 <- gulf_top %>%
+  filter(max_score_log10p > 5.897) %>%
+  ggplot(aes(x = POS, y = max_score_log10p)) + 
+  geom_point(aes(color = Condition, shape = as_factor(n_cond_in_top_set))) + 
+  #geom_line(aes(group = Condition)) +
+  facet_wrap( ~ CHR, nrow = 1, scales = "free_x", strip.position = "bottom") +
+  theme(legend.position = "top",
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        panel.background = element_rect(fill=NA)) +
+        labs(x = "Chromosome", y = "-log10(p value)") +
+        scale_x_continuous(expand = c(0.18, 0.18)) +
+  scale_color_viridis(discrete = TRUE, end = 0.97) +
+  scale_shape_manual(values = c(4, 16, 17, 18, 14, 3, 4))
+save_plot("../manuscript/plots/Gulf_G_BLUP_top_500_Manhattan_2.png", base_height = 5, base_asp = 2.5, plot = gulf_top_manh2)
+
+gulf_top_manh <- gulf_top %>%
+  filter(max_score_log10p > 5.897) %>%
+  ggplot(aes(x = POS, y = max_score_log10p)) + 
+  geom_point(aes(color = n_cond_in_top_set, shape = Condition)) + 
+  #geom_line(aes(group = Condition)) +
+  facet_wrap( ~ CHR, nrow = 1, scales = "free_x", strip.position = "bottom") +
+  theme(legend.position = "top",
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        panel.background = element_rect(fill=NA)) +
+        labs(x = "Chromosome", y = "-log10(p value)") +
+        scale_x_continuous(expand = c(0.18, 0.18)) +
+  scale_color_viridis(option = "B", end = 0.9) +
+  scale_shape_manual(values = c(15, 16, 17, 18, 14, 3, 4))
+
+save_plot("../manuscript/plots/Gulf_G_BLUP_top_500_Manhattan.png", base_height = 5, base_asp = 2.5, plot = gulf_top_manh)
 ```
+
+```{r}
+midw_top <- readRDS(file.path(gwasdir, "Part-One-Output_Top-Effects-1000-SNPs_GR_FL_Midwest.rds"))
+
+midw_top %>%
+  group_by(n_cond_in_top_set) %>%
+  tally()
+
+midw_top_manh2 <- midw_top %>%
+  filter(max_score_log10p > 6) %>%
+  ggplot(aes(x = POS, y = max_score_log10p)) + 
+  geom_point(aes(color = Condition, shape = as_factor(n_cond_in_top_set))) + 
+  #geom_line(aes(group = Condition)) +
+  facet_wrap( ~ CHR, nrow = 1, scales = "free_x", strip.position = "bottom") +
+  theme(legend.position = "top",
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        panel.background = element_rect(fill=NA)) +
+        labs(x = "Chromosome", y = "-log10(p value)") +
+        scale_x_continuous(expand = c(0.18, 0.18)) +
+  scale_color_viridis(discrete = TRUE, end = 0.97) +
+  scale_shape_manual(values = c(4, 16, 17, 18, 14, 3, 4))
+save_plot("../manuscript/plots/Midwest_G_BLUP_top_500_Manhattan_2.png", base_height = 5, base_asp = 2.5, plot = midw_top_manh2)
+
+midw_top_manh <- midw_top %>%
+  filter(max_score_log10p > 5.897) %>%
+  ggplot(aes(x = POS, y = max_score_log10p)) + 
+  geom_point(aes(color = n_cond_in_top_set, shape = Condition)) + 
+  #geom_line(aes(group = Condition)) +
+  facet_wrap( ~ CHR, nrow = 1, scales = "free_x", strip.position = "bottom") +
+  theme(legend.position = "top",
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        panel.background = element_rect(fill=NA)) +
+        labs(x = "Chromosome", y = "-log10(p value)") +
+        scale_x_continuous(expand = c(0.18, 0.18)) +
+  scale_color_viridis(option = "B", end = 0.9) +
+  scale_shape_manual(values = c(15, 16, 17, 18, 14, 3, 4))
+
+save_plot("../manuscript/plots/Midwest_G_BLUP_top_500_Manhattan.png", base_height = 5, base_asp = 2.5, plot = midw_top_manh)
+```
+
+```{r}
+gandm_top <- readRDS(file.path(gwasdir, "Part-One-Output_Top-Effects-1000-SNPs_GR_FL_Gulf_and_Midwest.rds"))
+
+gandm_top %>%
+  group_by(n_cond_in_top_set) %>%
+  tally()
+
+gandm_top_manh2 <- gandm_top %>%
+  filter(max_score_log10p > 6) %>%
+  ggplot(aes(x = POS, y = max_score_log10p)) + 
+  geom_point(aes(color = Condition, shape = as_factor(n_cond_in_top_set))) + 
+  #geom_line(aes(group = Condition)) +
+  facet_wrap( ~ CHR, nrow = 1, scales = "free_x", strip.position = "bottom") +
+  theme(legend.position = "top",
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        panel.background = element_rect(fill=NA)) +
+        labs(x = "Chromosome", y = "-log10(p value)") +
+        scale_x_continuous(expand = c(0.18, 0.18)) +
+  scale_color_viridis(discrete = TRUE, end = 0.97) +
+  scale_shape_manual(values = c(4, 16, 17, 18, 14, 3, 4))
+save_plot("../manuscript/plots/Gulf_and_Midwest_G_BLUP_top_500_Manhattan_2.png", base_height = 5, base_asp = 2.5, plot = gandm_top_manh2)
+
+gandm_top_manh <- gandm_top %>%
+  filter(max_score_log10p > 5.897) %>%
+  ggplot(aes(x = POS, y = max_score_log10p)) + 
+  geom_point(aes(color = n_cond_in_top_set, shape = Condition)) + 
+  #geom_line(aes(group = Condition)) +
+  facet_wrap( ~ CHR, nrow = 1, scales = "free_x", strip.position = "bottom") +
+  theme(legend.position = "top",
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        panel.background = element_rect(fill=NA)) +
+        labs(x = "Chromosome", y = "-log10(p value)") +
+        scale_x_continuous(expand = c(0.18, 0.18)) +
+  scale_color_viridis(option = "B", end = 0.9) +
+  scale_shape_manual(values = c(15, 16, 17, 18, 14, 3, 4))
+
+save_plot("../manuscript/plots/Gulf_and_Midwest_G_BLUP_top_500_Manhattan.png", base_height = 5, base_asp = 2.5, plot = gandm_top_manh)
+```
+
+## Overlap with QTL and between subpops
+```{r}
+
+```

### 2020-07-30 R/Plots_manuscript_v0.3.Rmd
Add BLUPs for subpop specific kinship GWAS

@@
   mutate(GandGxE = Estimate + GxE_Estimate)  %>%
-  select(GandGxE, site:phe, Estimate, GxE_Estimate)
+  select(GandGxE, site:phe, Estimate, SE, GxE_Estimate, GxE_SE)
 ```
@@
 
+# ----
 # D2F
@@
 
-## G_BLUP GWAS
+# ----------------------------
+# G_BLUP GWAS
+
+The plan is to do GWAS on G_BLUP from an eight site model with GxE modeled as site:kinship for each subpopulation and both combined. The traits I will do GWAS on for G_BLUPs are as follows:
+
+eight_sites
+FL50 (to show other assoc are stronger)
+dyln_fl50
+cgdd_12c_gr2fl
+
+Maybe also do GWAS on G_BLUP from the North and Texas models as above, for the traits:
+tx_sites & north_sites
+FL50 (to show other assoc are stronger)
+dyln_fl50
+cgdd_12c_gr2fl
+crain_gr2fl
+crain_1d
+GR50
+cgdd_12c_10d
+
+All I really want to show are that there are stronger G associations for the weather-derived variables than for FL50 and GR50.
+And possibly exploring G associations at subpopulation home and away sites, for the second bit.
+That might be too complicated. Eh.
+
+It's possible I might want to confirm some of these G candidates with the fourway - how we see similar patterns in this cross for FL50 vs dyln/GDD, and there are some similar associations.
+
+THEN, for the next section, do GWAS on the BLUPs from the eight single site models. For these traits:
+all single sites...
+FL50 (actually, maybe not)
+dyln_fl50
+cgdd_12c_gr2fl
+crain_gr2fl
+crain_1d
+--
+GR50 (or maybe not)
+cgdd_12c_10d
+
+Then do mash runs.
+First for single traits across all eight sites, to make U_hyp for the next step - take the U_ed with large percentage variance explained.
+Make U_hyp with those covariances but no covariance with anything else...
+Then run all the flowering-related traits together.
+Talk about how then SNPs can have effects that are strongest in response to one type of environmental cue.
+And how cues can be present or absent at different sites.
+And how this kind of analysis can help reveal all the patterns at play, because it doesn't make assumptions about how effects are structured.
+
+Maybe bring back the QTL maps after this and talk about how patterns of effects differ across sites in these models, too. Though the Genstat model is less sensitive...
+
+```{r}
+h2all %>%
+  filter(site %in% c("eight_sites"))
+```
+
+## Make phenotype df for GWAS
+```{r}
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "analysis", "heritability", "kinship_subpop")
+site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
+               NE = "NE", MI = "MI", SD = "SD", eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
+subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"))
+phe_eight <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl")
+phe_txano <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
+               "crain_1d", "GR50", "cgdd_12c_10d")
+phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
+               "crain_1d", "GR50", "cgdd_12c_10d")
+phe_lists <- list(TX1 = phe_single, TX2 = phe_single, TX3 = phe_single, OK = phe_single, MO = phe_single,
+               NE = phe_single, MI = phe_single, SD = phe_single, eight_sites = phe_eight, tx_sites = phe_txano, north_sites = phe_txano)
+
+i = 9
+j = 1
+k = 1
+phe_gwas <- tibble()
+
+for(j in seq_along(subpop_v)){
+    for(i in seq_along(site_v)){
+      phe_v <- phe_lists[[i]]
+      for(k in seq_along(phe_v)){
+        phe_raw <- read_csv(file.path(outputdir, names(site_v)[i],
+                                      paste0("blups_", names(site_v)[i], "_",
+                                             names(subpop_v)[j], "_", phe_v[k],
+                                             ".csv")))
+        names(phe_raw)[2] <- paste0(phe_v[k], "_", names(site_v)[i])
+        phe_proc <- phe_raw %>%
+          separate(Effect, into = c("factor", "PLANT_ID"), sep = "\\)_") %>%
+          filter(!is.na(PLANT_ID) & !grepl("^SITE", factor)) %>%
+          left_join(metadata, by = "PLANT_ID") %>%
+          filter(SUBPOP %in% subpop_v[[j]]) %>%
+          select(PLANT_ID, 3)
+        if(i == 1 & k == 1){
+          phe_gwas <- phe_proc
+        } else {
+          phe_gwas <- phe_gwas %>%
+            left_join(phe_proc, by = "PLANT_ID")
+        }
+      }
+    }
+  phe_gwas <- phe_gwas %>%
+  select(PLANT_ID, ends_with("eight_sites"), ends_with("tx_sites"), ends_with("north_sites"), everything())
+  write_csv(phe_gwas, path = file.path(workingdir, "analysis", "gwas", "BLUPs",
+                                       paste0("BLUP_phenotypes_",
+                                              "for_gwas_GR_FL_",
+                                              names(subpop_v)[j],
+                                              ".csv")))
+}
+
+ggplot(data = phe_gwas, aes(x = phe_gwas$cgdd_12c_gr2fl_eight_sites)) + geom_histogram()
+
+```
+
+## Run GWAS 
+```{r}
+library(bigsnpr)
+library(tidyverse)
+library(switchgrassGWAS)
+library(AnnotationDbi)
+txdb <- loadDb(file = file.path("~", "Github", "pvdiv-genome",
+                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+inputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs")
+outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs")
+subpop_v <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", Gulf = "Gulf")
+snp_v <- c("12\\.1M", "8\\.8M", "10\\.3M")
+
+inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
+                                   "inputkinship.txt"),  delim = " ", 
+                                   col_names = "SNPfiles")
+set_v <- list(eight_sites = "eight_sites", tx_sites = "tx_sites", north_sites = "north_sites", single_sites = "!sites")
+
+for(i in seq_along(subpop_v)){
+  for(j in seq_along(set_v)){
+    dir.create(file.path(outputdir, names(set_v)[j]), showWarnings = FALSE)
+    snp <- snp_attach(inputfiles$SNPfiles[i])
+    phe_df <- read_csv(file.path(inputdir, paste0("BLUP_phenotypes_",
+                                                  "for_gwas_GR_FL_",
+                                                  names(subpop_v)[i],
+                                                  ".csv"))) 
+  if(names(set_v)[j] != "single_sites"){
+    phe_df <- phe_df %>%
+    select(PLANT_ID, ends_with(names(set_v)[j]))
+  } else {
+    phe_df <- phe_df %>%
+    select(PLANT_ID, !ends_with("sites"))
+  }
+  pvdiv_standard_gwas(snp, df = phe_df, type = "linear", 
+                      outputdir = file.path(outputdir, names(set_v)[j]),
+                      savegwas = TRUE, saveannos = TRUE, txdb = txdb,
+                      minphe = 130)
+
+  gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir,
+                                                       names(set_v)[j]), 
+                                      pattern = paste0("GWAS_datatable*.*",
+                                                       snp_v[i]))
+  phenotypes <- str_sub(gwas_rds, start = 16, end = -42) %>%
+    str_replace(., "_$", "")
+  out1 <- pvdiv_bigsnp2mashr(path = file.path(outputdir, names(set_v)[j]), 
+                             snp = snp, gwas_rds = gwas_rds, 
+                             phenotypes = phenotypes, numSNPs = 1000, 
+                             clump = FALSE, scaled = FALSE, saveoutput = TRUE,
+                             suffix = paste0("FL_", names(set_v)[j], "_",
+                                             names(subpop_v)[i]))
+  }
+}
+
+```
 

### 2020-07-31 R/Plots_manuscript_v0.3.Rmd
Start GWAS G BLUP eight site Results section

@@
   full_join(h2fl_gxe)
-```
-### older files to load
-h2fl <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/heritability/h2_df_FL50_related_functions_of_weather_v3.csv")
-h2gr <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/heritability/h2_df_GR50_related_functions_of_weather_v2.csv")
-h2gr4 <- read_csv(file.path(workingdir, "analysis", "heritability", "h2_df_GR50_related_functions_of_weather_v4.csv"))
-h2gr <- h2gr %>%
-  full_join(h2gr4)
-
-# Overview
-Here I'm putting the code for any new analyses that I'm running to do new writeup for the manuscript v0.3.
+```a the code for any new analyses that I'm running to do new writeup for the manuscript v0.3.
 
@@
 
+sites %>% arrange(manu_site)
 ```
@@
 
+for(j in 4){#seq_along(set_v)){
   for(i in seq_along(subpop_v)){
-  for(j in seq_along(set_v)){
     dir.create(file.path(outputdir, names(set_v)[j]), showWarnings = FALSE)
@@
     phe_df <- phe_df %>%
-    select(PLANT_ID, ends_with(names(set_v)[j]))
+    dplyr::select(PLANT_ID, ends_with(names(set_v)[j]))
   } else {
     phe_df <- phe_df %>%
-    select(PLANT_ID, !ends_with("sites"))
+    dplyr::select(PLANT_ID, !ends_with("sites"))
+    if(names(subpop_v)[i] != "Gulf"){
+      phe_df <- phe_df %>%
+    dplyr::select(PLANT_ID, ends_with("NE"), ends_with("MI"), ends_with("SD"))
+    }
   }
@@
 
-```{r}
 gwasdir <- file.path(workingdir, "analysis", "gwas", "G_BLUP")
@@
 log10p_gandm <- readRDS(file.path(gwasdir,                                 "log10p_strong_df_1000topSNPs_GR_FL_Gulf_and_Midwest.rds"))
+
+```{r}
+j=1
+i=3
+
+names(set_v)[j]
+names(subpop_v)[i]
+bonferroni_v <- c(-log10(0.05/12100000), -log10(0.05/8800000), -log10(0.05/10300000))
+
+gwasdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", names(set_v)[j])
+log10p_gulf <- readRDS(file.path(gwasdir, 
+                                 paste0("log10p_strong_df_1000topSNPs_",
+                                        "FL_", names(set_v)[j], "_Gulf",
+                                        ".rds")))
+log10p_midw <- readRDS(file.path(gwasdir,
+                                 paste0("log10p_strong_df_1000topSNPs_", "FL_",
+                                 names(set_v)[j], "_", "Midwest.rds")))
+log10p_gandm <- readRDS(file.path(gwasdir,
+                                  paste0("log10p_strong_df_1000topSNPs_",
+                                         "FL_", names(set_v)[j], "_",
+                                         "Gulf_and_Midwest.rds")))
+log10p_df <- readRDS(file.path(gwasdir,
+                                  paste0("log10p_strong_df_1000topSNPs_",
+                                         "FL_", names(set_v)[j], "_",
+                                         names(subpop_v)[i], ".rds")))
 ```
@@
   group_by(PHE) %>%
-  tally() %>%
-  write_csv(., "../manuscript/tables/Gulf_SNPs_above_Bonferroni_by_trait.csv")
-```
-
-```{r}
+  tally() #%>%
+  #write_csv(., paste0("../manuscript/tables/Gulf_", names(set_v)[j], "_SNPs_above_Bonferroni_by_trait.csv"))
+
 log10p_midw %>%
@@
   group_by(PHE) %>%
-  tally() %>%
-  write_csv(., "../manuscript/tables/Midwest_SNPs_above_Bonferroni_by_trait.csv")
+  tally() #%>%
+  #write_csv(., paste0("../manuscript/tables/Midwest_", names(set_v)[j], "_SNPs_above_Bonferroni_by_trait.csv"))
 
@@
   group_by(PHE) %>%
-  tally() %>%
-  write_csv(., "../manuscript/tables/Gulf_and_Midwest_diversity_panel_SNPs_above_Bonferroni_by_trait.csv")
-```
-
-```{r}
-gulf_top <- readRDS(file.path(gwasdir, "Part-One-Output_Top-Effects-1000-SNPs_GR_FL_Gulf.rds"))
-
-gulf_top %>%
+  tally() #%>%
+  #write_csv(., paste0("../manuscript/tables/Gulf_and_Midwest_", names(set_v)[j], "__SNPs_above_Bonferroni_by_trait.csv"))
+  #
+log10p_gandm %>%
+  pivot_longer(cols = -(SNP:POS), names_to = "PHE", values_to = "log10p") %>%
+  ungroup() %>%
+  filter(log10p > -log10(0.05/12100000)) %>%
+  mutate(Pos_Mb = floor(POS/100000)/10) %>%
+  group_by(CHR, Pos_Mb) %>% tally()
+bonferroni <- -log10(0.05/12100000)
+log10p_gandm %>%
+  filter(cgdd_12c_gr2fl_eight_sites_log10p > -log10(0.05/12100000) |
+         FL50_eight_sites_log10p > -log10(0.05/12100000)) %>%
+  group_by(cgdd_12c_gr2fl_eight_sites_log10p > FL50_eight_sites_log10p) %>%
+  tally()
+log10p_gandm %>%
+  filter(dyln_fl50_eight_sites_log10p > -log10(0.05/12100000)  |
+         FL50_eight_sites_log10p > -log10(0.05/12100000)) %>%
+  group_by(dyln_fl50_eight_sites_log10p > FL50_eight_sites_log10p) %>%
+  tally()
+
+log10p_gandm %>%
+  ggplot(aes(x = dyln_fl50_eight_sites_log10p, y = FL50_eight_sites_log10p)) + 
+  geom_point(alpha = 0.5) + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) +
+  geom_vline(xintercept = bonferroni, color = "red", linetype = 2) + 
+  geom_hline(yintercept = bonferroni, color = "red", linetype = 2)
+```
+```{r}
+i=1 # subpop
+j=3 # site set
+
+names(set_v)[j]
+names(subpop_v)[i]
+bonferroni_v <- c(-log10(0.05/12100000), -log10(0.05/8800000), -log10(0.05/10300000))
+
+gwasdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", names(set_v)[j])
+log10p_df <- readRDS(file.path(gwasdir,
+                                  paste0("log10p_strong_df_1000topSNPs_",
+                                         "FL_", names(set_v)[j], "_",
+                                         names(subpop_v)[i], ".rds")))
+top_set <- readRDS(file.path(gwasdir, 
+                                paste0("Part-One-Output_Top-Effects-1000-SNPs_",
+                                        "FL_", names(set_v)[j], "_",
+                                        names(subpop_v)[i], ".rds")))
+
+log10p_df %>%
+  pivot_longer(cols = -(SNP:POS), names_to = "PHE", values_to = "log10p") %>%
+  ungroup() %>%
+  filter(log10p > bonferroni_v[i]) %>%
+  group_by(CHR, POS) %>% tally()
+log10p_df %>%
+  pivot_longer(cols = -(SNP:POS), names_to = "PHE", values_to = "log10p") %>%
+  ungroup() %>%
+  filter(log10p > bonferroni_v[i]) %>%
+  mutate(Pos_Mb = floor(POS/100000)/10) %>%
+  group_by(CHR, Pos_Mb) %>% tally()
+
+top50 <- top_set %>% ungroup() %>%
+  top_n(100, max_score_log10p) %>%
+  mutate(SNP = paste(CHR, POS, sep = "_"))
+dyln_binom <- log10p_df %>%
+  filter(SNP %in% top50$SNP) %>%
+  group_by(dyln_fl50_north_sites_log10p > FL50_north_sites_log10p) %>%
+  tally()
+binom.test(x = dyln_binom$n[2], n = 100, alternative = "greater")
+
+cgdd_binom <- log10p_df %>%
+  filter(SNP %in% top50$SNP) %>%
+  group_by(cgdd_12c_gr2fl_north_sites_log10p > FL50_north_sites_log10p) %>%
+  tally()
+binom.test(x = cgdd_binom$n[2], n = 100, alternative = "greater")
+
+log10p_df %>%
+  ggplot(aes(x = dyln_fl50_north_sites_log10p, y = FL50_north_sites_log10p)) + 
+  geom_point(alpha = 0.5) + 
+  geom_abline(slope = 1, intercept = 0, linetype = 2) +
+  geom_vline(xintercept = bonferroni_v[i], color = "red", linetype = 2) + 
+  geom_hline(yintercept = bonferroni_v[i], color = "red", linetype = 2)
+```
+
+## Plot top GWAS hits
+```{r}
+for(i in seq_along(subpop_v)){
+  for(j in 1:3){
+    gwasdir <- file.path(workingdir, "analysis", "gwas", "BLUPs",
+                         names(set_v)[j])
+    top_set <- readRDS(file.path(gwasdir, 
+                                paste0("Part-One-Output_Top-Effects-1000-SNPs_",
+                                        "FL_", names(set_v)[j], "_",
+                                        names(subpop_v)[i], ".rds")))
+    top_set %>%
       group_by(n_cond_in_top_set) %>%
@@
 
-gulf_top_manh2 <- gulf_top %>%
+    top_manh2 <- top_set %>%
       filter(max_score_log10p > 5.897) %>%
@@
       #geom_line(aes(group = Condition)) +
-  facet_wrap( ~ CHR, nrow = 1, scales = "free_x", strip.position = "bottom") +
+      facet_wrap( ~ CHR, nrow = 1, scales = "free_x", 
+                  strip.position = "bottom") +
       theme(legend.position = "top",
@@
       scale_shape_manual(values = c(4, 16, 17, 18, 14, 3, 4))
-save_plot("../manuscript/plots/Gulf_G_BLUP_top_500_Manhattan_2.png", base_height = 5, base_asp = 2.5, plot = gulf_top_manh2)
-
-gulf_top_manh <- gulf_top %>%
-  filter(max_score_log10p > 5.897) %>%
-  ggplot(aes(x = POS, y = max_score_log10p)) + 
-  geom_point(aes(color = n_cond_in_top_set, shape = Condition)) + 
-  #geom_line(aes(group = Condition)) +
-  facet_wrap( ~ CHR, nrow = 1, scales = "free_x", strip.position = "bottom") +
-  theme(legend.position = "top",
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        panel.background = element_rect(fill=NA)) +
-        labs(x = "Chromosome", y = "-log10(p value)") +
-        scale_x_continuous(expand = c(0.18, 0.18)) +
-  scale_color_viridis(option = "B", end = 0.9) +
-  scale_shape_manual(values = c(15, 16, 17, 18, 14, 3, 4))
-
-save_plot("../manuscript/plots/Gulf_G_BLUP_top_500_Manhattan.png", base_height = 5, base_asp = 2.5, plot = gulf_top_manh)
-```
-
-```{r}
-midw_top <- readRDS(file.path(gwasdir, "Part-One-Output_Top-Effects-1000-SNPs_GR_FL_Midwest.rds"))
-
-midw_top %>%
-  group_by(n_cond_in_top_set) %>%
+    save_plot(paste0("../manuscript/plots/", "FL_", names(set_v)[j], "_",
+                     names(subpop_v)[i], "_top_500_Manhattan.png"), 
+              base_height = 5, base_asp = 2.5, plot = top_manh2)
+
+  }
+}
+```
+
+## Overlap with QTL and between subpops
+
+All I really want to show are that there are stronger G associations for the weather-derived variables than for FL50 and GR50.
+And possibly exploring G associations at subpopulation home and away sites, for the second bit.
+That might be too complicated. Eh.
+
+It's possible I might want to confirm some of these G candidates with the fourway - how we see similar patterns in this cross for FL50 vs dyln/GDD, and there are some similar associations.
+
+Comparing the top 100 hits within and between subpopulations, X of Y were found in the same 100kb regions.
+```{r}
+i=3 # subpop
+j=1 # site set
+
+names(set_v)[j]
+names(subpop_v)[i]
+bonferroni_v <- c(-log10(0.05/12100000), -log10(0.05/8800000), -log10(0.05/10300000))
+
+top_pos <- list()
+
+gwasdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", names(set_v)[j])
+log10p_df <- readRDS(file.path(gwasdir,
+                                  paste0("log10p_strong_df_1000topSNPs_",
+                                         "FL_", names(set_v)[j], "_",
+                                         names(subpop_v)[i], ".rds")))
+for(i in 1:3){
+top_set <- readRDS(file.path(gwasdir, 
+                                paste0("Part-One-Output_Top-Effects-1000-SNPs_",
+                                        "FL_", names(set_v)[j], "_",
+                                        names(subpop_v)[i], ".rds")))
+
+top_pos[[i]] <- top_set %>%
+  ungroup() %>%
+  top_n(n = 100, wt = max_score_log10p) %>%
+  mutate(Pos_Mb = floor(POS/100000)/10) %>%
+  group_by(CHR, Pos_Mb) %>%
   tally()
-
-midw_top_manh2 <- midw_top %>%
-  filter(max_score_log10p > 6) %>%
-  ggplot(aes(x = POS, y = max_score_log10p)) + 
-  geom_point(aes(color = Condition, shape = as_factor(n_cond_in_top_set))) + 
-  #geom_line(aes(group = Condition)) +
-  facet_wrap( ~ CHR, nrow = 1, scales = "free_x", strip.position = "bottom") +
-  theme(legend.position = "top",
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        panel.background = element_rect(fill=NA)) +
-        labs(x = "Chromosome", y = "-log10(p value)") +
-        scale_x_continuous(expand = c(0.18, 0.18)) +
-  scale_color_viridis(discrete = TRUE, end = 0.97) +
-  scale_shape_manual(values = c(4, 16, 17, 18, 14, 3, 4))
-save_plot("../manuscript/plots/Midwest_G_BLUP_top_500_Manhattan_2.png", base_height = 5, base_asp = 2.5, plot = midw_top_manh2)
-
-midw_top_manh <- midw_top %>%
-  filter(max_score_log10p > 5.897) %>%
-  ggplot(aes(x = POS, y = max_score_log10p)) + 
-  geom_point(aes(color = n_cond_in_top_set, shape = Condition)) + 
-  #geom_line(aes(group = Condition)) +
-  facet_wrap( ~ CHR, nrow = 1, scales = "free_x", strip.position = "bottom") +
-  theme(legend.position = "top",
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        panel.background = element_rect(fill=NA)) +
-        labs(x = "Chromosome", y = "-log10(p value)") +
-        scale_x_continuous(expand = c(0.18, 0.18)) +
-  scale_color_viridis(option = "B", end = 0.9) +
-  scale_shape_manual(values = c(15, 16, 17, 18, 14, 3, 4))
-
-save_plot("../manuscript/plots/Midwest_G_BLUP_top_500_Manhattan.png", base_height = 5, base_asp = 2.5, plot = midw_top_manh)
-```
-
-```{r}
-gandm_top <- readRDS(file.path(gwasdir, "Part-One-Output_Top-Effects-1000-SNPs_GR_FL_Gulf_and_Midwest.rds"))
-
-gandm_top %>%
-  group_by(n_cond_in_top_set) %>%
+}
+
+gandm_m <- top_pos[[1]] %>%
+  inner_join(top_pos[[2]])
+gandm_g <- top_pos[[1]] %>%
+  inner_join(top_pos[[3]])
+m_g <- top_pos[[2]] %>%
+  inner_join(top_pos[[3]])
+c(nrow(gandm_m)/nrow(top_pos[[1]]), nrow(gandm_m)/nrow(top_pos[[2]]),
+  nrow(gandm_g)/nrow(top_pos[[1]]), nrow(gandm_g)/nrow(top_pos[[3]]),
+  nrow(m_g)/nrow(top_pos[[2]]), nrow(m_g)/nrow(top_pos[[3]])
+  )
+
+top50 <- top_set %>% ungroup() %>%
+  top_n(100, max_score_log10p) %>%
+  mutate(SNP = paste(CHR, POS, sep = "_"))
+dyln_binom <- log10p_df %>%
+  filter(SNP %in% top50$SNP) 
+```
+
+```{r}
+i=3 # subpop
+j=1 # site set
+
+names(set_v)[j]
+names(subpop_v)[i]
+bonferroni_v <- c(-log10(0.05/12100000), -log10(0.05/8800000), -log10(0.05/10300000))
+
+top_pos <- list()
+
+log10p_df <- readRDS(file.path(gwasdir,
+                                  paste0("log10p_strong_df_1000topSNPs_",
+                                         "FL_", names(set_v)[j], "_",
+                                         names(subpop_v)[i], ".rds")))
+for(j in 1:3){
+  gwasdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", names(set_v)[j])
+  top_set <- readRDS(file.path(gwasdir, 
+                                paste0("Part-One-Output_Top-Effects-1000-SNPs_",
+                                        "FL_", names(set_v)[j], "_",
+                                        names(subpop_v)[i], ".rds")))
+
+top_pos[[j]] <- top_set %>%
+  ungroup() %>%
+  top_n(n = 500, wt = max_score_log10p) %>%
+  mutate(Pos_Mb = floor(POS/100000)/10) %>%
+  group_by(CHR, Pos_Mb) %>%
   tally()
-
-gandm_top_manh2 <- gandm_top %>%
-  filter(max_score_log10p > 6) %>%
-  ggplot(aes(x = POS, y = max_score_log10p)) + 
-  geom_point(aes(color = Condition, shape = as_factor(n_cond_in_top_set))) + 
-  #geom_line(aes(group = Condition)) +
-  facet_wrap( ~ CHR, nrow = 1, scales = "free_x", strip.position = "bottom") +
-  theme(legend.position = "top",
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        panel.background = element_rect(fill=NA)) +
-        labs(x = "Chromosome", y = "-log10(p value)") +
-        scale_x_continuous(expand = c(0.18, 0.18)) +
-  scale_color_viridis(discrete = TRUE, end = 0.97) +
-  scale_shape_manual(values = c(4, 16, 17, 18, 14, 3, 4))
-save_plot("../manuscript/plots/Gulf_and_Midwest_G_BLUP_top_500_Manhattan_2.png", base_height = 5, base_asp = 2.5, plot = gandm_top_manh2)
-
-gandm_top_manh <- gandm_top %>%
-  filter(max_score_log10p > 5.897) %>%
-  ggplot(aes(x = POS, y = max_score_log10p)) + 
-  geom_point(aes(color = n_cond_in_top_set, shape = Condition)) + 
-  #geom_line(aes(group = Condition)) +
-  facet_wrap( ~ CHR, nrow = 1, scales = "free_x", strip.position = "bottom") +
-  theme(legend.position = "top",
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        panel.background = element_rect(fill=NA)) +
-        labs(x = "Chromosome", y = "-log10(p value)") +
-        scale_x_continuous(expand = c(0.18, 0.18)) +
-  scale_color_viridis(option = "B", end = 0.9) +
-  scale_shape_manual(values = c(15, 16, 17, 18, 14, 3, 4))
-
-save_plot("../manuscript/plots/Gulf_and_Midwest_G_BLUP_top_500_Manhattan.png", base_height = 5, base_asp = 2.5, plot = gandm_top_manh)
-```
-
-## Overlap with QTL and between subpops
-```{r}
-
-```
+}
+
+gandm_m <- top_pos[[1]] %>%
+  inner_join(top_pos[[2]])
+gandm_g <- top_pos[[1]] %>%
+  inner_join(top_pos[[3]])
+m_g <- top_pos[[2]] %>%
+  inner_join(top_pos[[3]])
+c(nrow(gandm_m)/nrow(top_pos[[1]]), nrow(gandm_m)/nrow(top_pos[[2]]),
+  nrow(gandm_g)/nrow(top_pos[[1]]), nrow(gandm_g)/nrow(top_pos[[3]]),
+  nrow(m_g)/nrow(top_pos[[2]]), nrow(m_g)/nrow(top_pos[[3]])
+  )
+
+top50 <- top_set %>% ungroup() %>%
+  top_n(100, max_score_log10p) %>%
+  mutate(SNP = paste(CHR, POS, sep = "_"))
+dyln_binom <- log10p_df %>%
+  filter(SNP %in% top50$SNP) 
+```
+for gulf and midwest:
+4.3-6.5% overlap in top 100 between eight and tx; no overlap in top 100 between eight and north or north and texas.
+500: 8% eight and texas; 1.6% eight and north. 0.9% north and texas.
+
+for midwest:
+5% overlap in top 500 eight and texas; 18% overlap eight and north; 1.7% north and texas.
+
+for gulf:
+6.2% overlap eight and texas in top 500; 5.2% overlap eight and north; 2.7% overlap north and texas.
+
+I think the top 500 SNPs are good here - give you large enough numbers.

### 2020-08-02 R/Plots_manuscript_v0.3.Rmd
Ignore plots in gwas/BLUPs/single_sites

@@
 ```
+## Run mash
+Then do mash runs.
+First for single traits across all eight sites, to make U_hyp for the next step - take the U_ed with large percentage variance explained.
+Make U_hyp with those covariances but no covariance with anything else...
+Then run all the flowering-related traits together.
+Talk about how then SNPs can have effects that are strongest in response to one type of environmental cue.
+And how cues can be present or absent at different sites.
+And how this kind of analysis can help reveal all the patterns at play, because it doesn't make assumptions about how effects are structured.
+```{r}
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
+site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
+               NE = "NE", MI = "MI", SD = "SD")#, eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
+subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"))
+phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
+               "crain_1d", "GR50", "cgdd_12c_10d")
+```
+
+single traits across all eight sites...
+```{r}
+i=1 # site
+j=1 # phe
+k=1 # subpop
+phe_suffix <- paste(phe_single[j], names(subpop_v)[k], 
+                    sep = "_")
+gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir,
+                                                       names(set_v)[j]), 
+                                      pattern = paste0("GWAS_datatable*.*",
+                                                       snp_v[i]))
+phenotypes <- str_sub(gwas_rds, start = 16, end = -42) %>%
+    str_replace(., "_$", "")
+out1 <- pvdiv_bigsnp2mashr(path = file.path(outputdir, names(set_v)[j]), 
+                           snp = snp, gwas_rds = gwas_rds, 
+                           phenotypes = phenotypes, numSNPs = 1000, 
+                           clump = FALSE, scaled = FALSE, saveoutput = TRUE,
+                           suffix = paste0("FL_", names(set_v)[j], "_",
+                                           names(subpop_v)[i]))
+numSNPs <- c(28000, 10000, 4000, 10000)
+
+gulf_loc <- c("CLMB", "KBSM", "KING", "PKLE", "TMPL")
+midw_loc <- c("CLMB", "KBSM", "PKLE")
+atl_loc <- c("CLMB", "KBSM", "LINC", "PKLE", "TMPL")
+midwgulf_loc <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL",
+                  "TMPL")
+U_gulf <- read_rds(file.path(path, "U_hyp_gulf.rds"))
+i = 2
+U_ed <- file.path(path, "Mash_data_driven_covariances_10000_gulf.rds")
+  m_out <- mash_standard_run(path = path, numSNPs = numSNPs[i], U_ed = U_ed,
+                             U_hyp = U_gulf, suffix = phe_suffix[i],
+                             saveoutput = TRUE)
+```
+
+## Analyze GWAS Results
+
+May want to rerun pvdiv_bigsnp2mashr function so that I'm looking at GR & FL top 1000 SNPs separately.
 

### 2020-08-06 R/Plots_manuscript_v0.3.Rmd
Work on combined Manhattan plot for BLUPs for G for subpop specific K

@@
   full_join(h2fl_gxe)
-```a the code for any new analyses that I'm running to do new writeup for the manuscript v0.3.
+```
+as well as the code for any new analyses that I'm running to do new writeup for the manuscript v0.3.
 
@@
 ```
+
 ```{r}
@@
 I think the top 500 SNPs are good here - give you large enough numbers.
+
+## View Annotations
+```{r}
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+snp_v <- c("12\\.1M", "8\\.8M", "10\\.3M")
+subpop_v <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", Gulf = "Gulf")
+set_v <- list(eight_sites = "eight_sites", tx_sites = "tx_sites", north_sites = "north_sites", single_sites = "!sites")
+
+for(j in seq_along(set_v)){
+  anno_list <- list()
+  annodir <- file.path(workingdir, "analysis", "gwas", "BLUPs",
+                       names(set_v)[j])
+  anno_v <- pvdiv_results_in_folder(path = annodir, pattern = "Annotation")
+
+  for(i in seq_along(anno_v)){
+  annos <- readRDS(file.path(annodir, anno_v[i]))
+  if(names(set_v)[j] != "single_sites"){
+  anno_info <- anno_v[i] %>% enframe(name = NULL, value = "File") %>%
+    mutate(phe = case_when(grepl("dyln_fl50", File) ~ "dyln_fl50",
+                           grepl("cgdd_12c_gr2fl", File) ~ "cgdd_12c_gr2fl",
+                           grepl("FL50", File) ~ "FL50",
+                           grepl("GR50", File) ~ "GR50",
+                           grepl("crain_gr2fl", File) ~ "crain_gr2fl",
+                           grepl("crain_1d", File) ~ "crain_1d",
+                           grepl("cgdd_12c_10d", File) ~ "cgdd_12c_10d"),
+           sites = case_when(grepl("eight_sites", File) ~ "eight_sites",
+                             grepl("north_sites", File) ~ "north_sites",
+                             grepl("tx_sites", File) ~ "tx_sites"),
+           subpop = case_when(grepl("8\\.8M", File) ~ "Midwest",
+                              grepl("10\\.3M", File) ~ "Gulf",
+                              grepl("12\\.1M", File) ~ "Gulf_and_Midwest"))
+  } else {
+    anno_info <- anno_v[i] %>% enframe(name = NULL, value = "File") %>%
+    mutate(phe = case_when(grepl("dyln_fl50", File) ~ "dyln_fl50",
+                           grepl("cgdd_12c_gr2fl", File) ~ "cgdd_12c_gr2fl",
+                           grepl("FL50", File) ~ "FL50",
+                           grepl("GR50", File) ~ "GR50",
+                           grepl("crain_gr2fl", File) ~ "crain_gr2fl",
+                           grepl("crain_1d", File) ~ "crain_1d",
+                           grepl("cgdd_12c_10d", File) ~ "cgdd_12c_10d"),
+           sites = case_when(grepl("TX1", File) ~ "TX1",
+                             grepl("TX2", File) ~ "TX2",
+                             grepl("TX3", File) ~ "TX3",
+                             grepl("OK", File) ~ "OK",
+                             grepl("MO", File) ~ "MO",
+                             grepl("MI", File) ~ "MI",
+                             grepl("NE", File) ~ "NE",
+                             grepl("SD", File) ~ "SD"),
+           subpop = case_when(grepl("8\\.8M", File) ~ "Midwest",
+                              grepl("10\\.3M", File) ~ "Gulf",
+                              grepl("12\\.1M", File) ~ "Gulf_and_Midwest"))
+  }
+  anno_list[[i]] <- annos$top500SNPs_within50000bp %>%
+    arrange(QUERYID, `distance from gene`) %>%
+    group_by(QUERYID) %>%
+    slice(1) %>%
+    ungroup() %>%
+    mutate(POS_Mb = floor(POS/10000)/100,
+           phe = anno_info$phe,
+           sites = anno_info$sites,
+           subpop = anno_info$subpop) %>%
+    select(sites, subpop, phe, CHR, POS_Mb, `Gene ID`, `distance from gene`, `FDR Adjusted p value`, `Arabidopsis thaliana homolog`:`A.
+thaliana gene annotation`, `Rice gene annotation`, `Rice homolog`)
+  }
+
+  anno_df <- anno_list[[1]]
+  for(i in seq_along(anno_list)[-1]){
+    anno_df <- anno_df %>%
+      full_join(anno_list[[i]])
+  }
+  anno_df %>%
+    arrange(CHR, POS_Mb) %>%
+    filter(`FDR Adjusted p value` <= 0.1) %>%
+write_csv(., path = file.path(workingdir, "analysis", "gwas", "BLUPs",
+                                    paste0(names(set_v)[j],     
+                                           "_top_500_SNPs_above_FDR",
+                                           "_closest_annotations.csv")))
+}
+```
+Taylor et al 2019, Crop Science: switchgrass GWAS
+Grabowski et al 2017, New Phytologist - GWAS on ~509 individuals on heading date in New York
+Switchgrass GWAS at two locations: Frontiers in Plant Science
+Pavir.8NG050900: Pseudo Response Regulator 5 (PRR5) - important in flowering date
+Flowering locus T (on Chromosome 5N). FKF1, PRR5 COP1 GASA8 P450 - differentially expressed flowering genes between early and late flowering individuals.
+
+Pavir.2NG448600: PRR5
+Pavir.5KG044000 (FT)
+Pavir.2NG126000 (PRR7)
+Pavir.
+
+```{r}
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+set_v <- list(eight_sites = "eight_sites", tx_sites = "tx_sites", north_sites = "north_sites", single_sites = "!sites")
+j = 1
+annos_df <- annos1
+for(j in seq_along(set_v)[-1]){
+annos1 <- read_csv(file = file.path(workingdir, "analysis", "gwas", "BLUPs",
+                                    paste0(names(set_v)[j],     
+                                           "_top_500_SNPs_above_FDR",
+                                           "_closest_annotations.csv")))
+annos_df <- annos_df %>%
+  full_join(annos1)
+}
+
+annos_plot <- annos_df %>%
+  filter(sites %in% c("eight_sites", "tx_sites", "north_sites") & !(phe %in% c("GR50", "cgdd_12c_10d"))) %>%
+  group_by(CHR, POS_Mb, sites, subpop, phe) %>% add_tally(name = "n_SNPs") %>%
+  arrange(CHR, POS_Mb) %>%
+  #filter(n_SNPs > 1) %>%
+  slice(1) %>%
+  group_by(CHR, POS_Mb) %>% add_tally(name = "n_Mb") %>%
+  filter(n_Mb > 7) %>% arrange(CHR, POS_Mb) %>%
+  select(n_SNPs, n_Mb, everything())
+```
+
+dyln, fl50, cgdd eight sites gulf & midwest, north sites gulf & midwest cgdd & dyln, tx crain gr2fl, dyln, cgdd
+```{r}
+qtldf
+```
+
+3 regions with the most 'independent' hits: Chr03K 50.84, Chr03N 31.73, Chr05N 64.86
+3N isn't potentially confirmed by any QTL. 3K doesn't overlap. 
+5N does!All four traits for QTL (dyln much higher lod).
+Then: Chr03K 38.23, Chr04K 8.07, Chr05N 2.79, Chr09N 69.33
+3K ~ 17 - 26. Not really.  4K 2.28-5.5 & 9.92 - 13.93, so definitely close/overlaps. 5N overlaps. 9N does not.
+
+So 5N (late), 5N (early), and 4K have overlaps thus far. Very strong result.
+
+Then: 
+Chr01K 56.75 (no), Chr02K 4.25 (yes), 3K 38.23, 38.27, 38.35, 40.87 (no), Chr04N 42.92 (no), Chr05N 49.05 (very close ~ 44.5 - 46.6), Chr07K 33.33 (very close 34.82), Chr08N 28.5 (yes), Chr08N 43.77 (no), Chr09N 21.45 26.06 30.04 (no but 21 is close: 6.88 - 17.31)
+
+So for 7 regions sig above FDR in >=8 subsets, 3 overlap, and overlap the 3 of the 4 biggest QTL.
+For the next 14 regions sig above FDR in 7 or more subsets, 2 overlap and 3 are very close. 
+So of the 12-13 QTL, we find overlap for 3 of the 4 biggest, and 2-5 of the remainder. We don't find big hits in GWAS on 2N.
+
+Biggest QTL: 2N, 4K, 5N (x2-3) (4-5)
+Other QTL: 2K (x2), 3K, 7K, 7N, 8N, 9K, 9N (8)
+
+nothing on 1K, 1N, 3N, 4N, 5K, 6K, 6N, 8K. Nothing!!
+
+I think this is good. And can write up this section focusing on this.
+```{r}
+qtldf <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv")
+
+qtldf <- qtldf %>%
+  separate(flank_lo, into = c("CHR", "Pos_Mb_lo"), sep = "_", convert = TRUE) %>%
+  separate(flank_hi, into = c("flank", "Pos_Mb_hi"), sep = "_", convert = TRUE) %>%
+  mutate(Pos_Mb_lo = round(Pos_Mb_lo, digits = 2),
+         Pos_Mb_hi = round(Pos_Mb_hi, digits = 2)) %>%
+  select(CHR, Pos_Mb_lo, Pos_Mb_hi, lodcolumn, lod, everything()) %>%
+  arrange(CHR, Pos_Mb_lo)
+```
+
+## Combined Manhattan
+
+facet grid - rows are three sets of sites, colors are phenotypes, shapes are subpopulations. Then I can plot -log10 p values on the y axis and Chr/Pos on the x axis. 
+```{r}
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+subpop_v <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", Gulf = "Gulf")
+set_v <- list(eight_sites = "eight_sites", tx_sites = "tx_sites", north_sites = "north_sites", single_sites = "!sites")
+j = 1
+i = 1
+
+for(i in seq_along(subpop_v)){
+  for(j in 1:3){
+    if(i == 1 & j ==1){
+      logpdf <- readRDS(file.path(workingdir, "analysis", "gwas", "BLUPs",
+                                  names(set_v)[j],
+                                  paste0("log10p_strong_df_1000topSNPs_FL_",
+                                         names(set_v)[j], "_",
+                                         names(subpop_v)[i], ".rds"))) %>%
+        mutate(subpop = names(subpop_v)[i]) %>% ungroup()
+      } else {
+        logpdf1 <- readRDS(file.path(workingdir, "analysis", "gwas", "BLUPs",
+                                  names(set_v)[j],
+                                  paste0("log10p_strong_df_1000topSNPs_FL_",
+                                         names(set_v)[j], "_", 
+                                         names(subpop_v)[i], ".rds"))) %>%
+        mutate(subpop = names(subpop_v)[i]) %>% ungroup()
+        logpdf <- logpdf %>%
+          full_join(logpdf1)
+      }
+  }
+}
+
+log10p_long <- logpdf %>%
+  select(-starts_with("GR50"), -starts_with("cgdd_12c_10d")) %>%
+  pivot_longer(cols = ends_with("log10p"), names_to = "File", values_to = "log10p") %>%
+  mutate(phe = case_when(grepl("dyln_fl50", File) ~ "dyln_fl50",
+                           grepl("cgdd_12c_gr2fl", File) ~ "cgdd_12c_gr2fl",
+                           grepl("FL50", File) ~ "FL50",
+                           grepl("GR50", File) ~ "GR50",
+                           grepl("crain_gr2fl", File) ~ "crain_gr2fl",
+                           grepl("crain_1d", File) ~ "crain_1d",
+                           grepl("cgdd_12c_10d", File) ~ "cgdd_12c_10d"),
+           sites = case_when(grepl("eight_site", File) ~ "eight_sites",
+                             grepl("north_site", File) ~ "north_sites",
+                             grepl("tx_site", File) ~ "tx_sites")) %>%
+  select(SNP:POS, sites, subpop, phe, log10p) %>%
+  filter(!is.na(log10p))
+
+```
+
+```{r}
+
+annos_plot1 <- annos_plot %>% group_by(CHR, POS_Mb) %>% tally() %>%
+  mutate(POS = POS_Mb*1000000+5000)
+annos_plot2 <- annos_plot %>% group_by(CHR, POS_Mb) %>% tally() %>%
+  mutate(POS = POS_Mb*1000000) %>% filter(POS_Mb > 50)
+log10p_long$phe <- factor(log10p_long$phe, levels = c("crain_gr2fl", "cgdd_12c_gr2fl", "dyln_fl50", "crain_1d", "FL50"))
+
+log10p_plot <- log10p_long %>%
+  mutate(site = case_when(is.na(sites) ~ "all",
+                          sites == "eight_sites" ~ "All",
+                          sites == "tx_sites" ~ "Texas",
+                          sites == "north_sites" ~ "North",
+                          TRUE ~ sites),
+         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
+                                 subpop == "three_subpops" ~ "all",
+                                 TRUE ~ subpop),
+         phe = as.character(phe),
+         phe_name = case_when(phe == "FL50" ~ "50% flowering",
+                              phe == "dyln_fl50" ~ "Daylength",
+                              phe == "dyln_change_sec" ~ "daylength change (s)",
+                              phe == "cgdd_12c_gr2fl" ~ "GDD, GR to FL",
+                              phe == "crain_gr2fl" ~ "Rainfall, GR to FL",
+                              phe == "crain_1d" ~ "Rainfall, 1 day",
+                              phe == "crain_2d" ~ "Rainfall, 2d sum",
+                              phe == "crain_3d" ~ "Rainfall, 3d sum",
+                              phe == "crain_5d" ~ "Rainfall, 5d sum",
+                              phe == "crain_7d" ~ "Rainfall, 7d sum",
+                              TRUE ~ phe)) %>%
+  filter(!(phe %in% c("crain_1d", "FL50")) & !(subpop == "Midwest" & between(log10p, 9.242223, 9.242225)) & !(subpop == "Midwest" & between(log10p, 12.466318, 12.46632)))
+
+log10p_plot$site <- factor(log10p_plot$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North"))
+log10p_plot$phe_name <- factor(log10p_plot$phe_name, 
+                               levels = c("50% flowering", "Rainfall, GR to FL",
+                                          "GDD, GR to FL", "Daylength",
+                                          "daylength change (s)",
+                                          "Rainfall, 1 day",
+                                                  #"Rainfall, 2d sum",
+                                                  "Rainfall, 3d sum",
+                                                  "Rainfall, 5d sum"#,
+                                                  #"Rainfall, 7d sum"
+                                                  ))
+
+log10p_plot %>%
+  ggplot(aes(x = POS, y = log10p)) +
+  geom_vline(data = annos_plot1, aes(xintercept = POS), color = "grey70", 
+             size = 1.1) +
+  geom_point(aes(color = phe_name, shape = subpop_name)) +
+  facet_grid(site ~ CHR, scales = "free_x", space = "free_x", switch = "x") +
+      theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
+              legend.position = "top",
+            panel.background = element_rect(fill = "grey85")) +
+            labs(x = "Chromosome", y = "-log10(p value)") +
+            scale_x_continuous(breaks = c(20000000, 40000000, 60000000, 80000000, 100000000), labels = c("20 Mb", "40 Mb", "60 Mb", "80 Mb", ""), expand = c(0.4, 0.4)) +
+      scale_color_viridis(discrete = TRUE, end = 0.93,
+                          option = "B") +
+      scale_shape_manual(values = c(6, 4, 1, 4, 16, 4, 17, 18, 14, 3, 4)) +
+  ylim(c(6, 15)) + geom_hline(yintercept = c(8.3), linetype = 2)
+save_plot(paste0("../manuscript/plots/", "G_BLUP_Combined_Manhattans_fl_phe_subpops_sites.png"), 
+              base_height = 4.5, base_asp = 3.5, plot = last_plot())
+
+log10p_long %>%
+  filter(subpop == "Midwest" & log10p > 7) %>% group_by(log10p) %>% tally() %>% arrange(desc(n)) # 9.242224 12.466319
+
+-log10(0.05/1200000)
+
+log10p_plot %>%
+  ungroup() %>%
+  group_by(CHR, POS) %>%
+  arrange(CHR, desc(POS)) %>%
+  slice(1)
+  summarise(minP = min(POS),
+            maxP = max(POS), .groups = "keep")
+
+log10p_plot %>%
+  filter(grepl("K", CHR)) %>%
+  ggplot(aes(x = POS, y = log10p)) +
+  geom_vline(data = filter(annos_plot1, grepl("K", CHR)),
+             aes(xintercept = POS), color = "grey70", size = 1.1) +
+  geom_point(aes(color = phe_name, shape = subpop_name)) +
+  facet_grid(site ~ CHR, scales = "free_x", space = "free_x", switch = "x", ) +
+      theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), 
+              legend.position = "top",
+            panel.background = element_rect(fill = "grey85")) +
+            labs(x = "Chromosome", y = "-log10(p value)") +
+            scale_x_continuous(breaks = c(20000000, 40000000, 60000000, 80000000, 100000000), labels = c("20 Mb", "40 Mb", "60 Mb", "80 Mb", ""), expand = c(0.15, 0.15)) +
+      scale_color_viridis(discrete = TRUE, end = 0.93,
+                          option = "B") +
+      scale_shape_manual(values = c(6, 4, 1, 4, 16, 4, 17, 18, 14, 3, 4)) +
+  ylim(c(6, 15)) + geom_hline(yintercept = c(8.3), linetype = 2) + labs(subtitle = "K subgenome")
+save_plot(paste0("../manuscript/plots/", "G_BLUP_Combined_Manhattans_fl_phe_subpops_sites_K_subgenome.png"), 
+              base_height = 4, plot = last_plot())
+```

### 2020-08-08 R/Plots_manuscript_v0.3.Rmd
Add GWAS G BLUP results and fourway confirmation sections

@@
 j = 1
-annos_df <- annos1
+annos_df <- annos1 <- read_csv(file = file.path(workingdir, "analysis", "gwas", "BLUPs",
+                                    paste0(names(set_v)[j],     
+                                           "_top_500_SNPs_above_FDR",
+                                           "_closest_annotations.csv")))
 for(j in seq_along(set_v)[-1]){
@@
 annos_plot <- annos_df %>%
-  filter(sites %in% c("eight_sites", "tx_sites", "north_sites") & !(phe %in% c("GR50", "cgdd_12c_10d"))) %>%
+  filter(sites %in% c("eight_sites", "tx_sites", "north_sites") & !(phe %in% c("GR50", "cgdd_12c_10d", "FL50", "crain_1d"))) %>%
   group_by(CHR, POS_Mb, sites, subpop, phe) %>% add_tally(name = "n_SNPs") %>%
@@
   group_by(CHR, POS_Mb) %>% add_tally(name = "n_Mb") %>%
-  filter(n_Mb > 7) %>% arrange(CHR, POS_Mb) %>%
+  filter(n_Mb > 5) %>% arrange(CHR, POS_Mb) %>%
   select(n_SNPs, n_Mb, everything())
+annos_plot1 <- annos_plot %>% group_by(CHR, POS_Mb) %>% tally() %>%
+  mutate(POS_lo = POS_Mb*1000000,
+         POS_hi = POS_Mb*1000000+10000)
+
+annos_plot <- annos_df %>% 
+  mutate(POS_bin_lo = floor(POS_Mb*100/2),
+         POS_bin_hi = ceiling(POS_Mb*100/2)) %>%
+  filter(sites %in% c("eight_sites", "tx_sites", "north_sites") & !(phe %in% c("GR50", "cgdd_12c_10d", "FL50", "crain_1d", "crain_gr2fl"))) %>%
+  group_by(CHR, POS_bin_hi, sites, subpop, phe) %>% add_tally(name = "n_SNPs") %>%
+  arrange(CHR, POS_bin_hi) %>%
+  #filter(n_SNPs > 1) %>%
+  slice(1) %>%
+  group_by(CHR, POS_bin_hi) %>% add_tally(name = "n_Mb") %>%
+  filter(n_Mb > 4) %>% arrange(CHR, POS_bin_hi) %>%
+  select(n_SNPs, n_Mb, everything())
+annos_plot1 <- annos_plot %>% group_by(CHR, POS_bin_hi, POS_Mb) %>% tally() %>%
+  mutate(POS_lo = POS_Mb*1000000,
+         POS_hi = POS_Mb*1000000+10000)
 ```
@@
 ```{r}
-qtldf
+qtldf %>% arrange(CHR, Pos_Mb_lo) %>%
+  #filter(lodcolumn != "dyln_change_sec") %>%
+  ggplot(aes(x = Pos_Mb_lo, y = lod)) +
+  geom_jitter(aes(color = lodcolumn), height = 2) +
+  geom_errorbarh(aes(y = lod, xmin = Pos_Mb_lo, xmax = Pos_Mb_hi, color = lodcolumn), size = 1.5) +
+  facet_wrap(~CHR) + scale_color_manual(values = c("magenta", "yellow", "darkgreen", "blue")) + theme(legend.position = c(0.9, 0.15), panel.background = element_rect(fill = "grey90")) + xlim(c(0, 70)) + geom_vline(xintercept = 0)
 ```
@@
 
-annos_plot1 <- annos_plot %>% group_by(CHR, POS_Mb) %>% tally() %>%
-  mutate(POS = POS_Mb*1000000+5000)
-annos_plot2 <- annos_plot %>% group_by(CHR, POS_Mb) %>% tally() %>%
-  mutate(POS = POS_Mb*1000000) %>% filter(POS_Mb > 50)
 log10p_long$phe <- factor(log10p_long$phe, levels = c("crain_gr2fl", "cgdd_12c_gr2fl", "dyln_fl50", "crain_1d", "FL50"))
@@
 
-log10p_long %>%
-  filter(subpop == "Midwest" & log10p > 7) %>% group_by(log10p) %>% tally() %>% arrange(desc(n)) # 9.242224 12.466319
-
--log10(0.05/1200000)
-
-log10p_plot %>%
-  ungroup() %>%
-  group_by(CHR, POS) %>%
-  arrange(CHR, desc(POS)) %>%
-  slice(1)
-  summarise(minP = min(POS),
-            maxP = max(POS), .groups = "keep")
-
-log10p_plot %>%
+log10p_lim <- log10p_plot %>%
+  dplyr::group_by(CHR) %>%
+  dplyr::summarise(maxP = max(POS)) %>%
+  mutate(minP = 0, maxP = maxP+2000000, log10p = 2) %>%
+  pivot_longer(names_to = "type", values_to = "POS", cols = minP:maxP)
+
+k_subgenome <- log10p_plot %>%
   filter(grepl("K", CHR)) %>%
@@
              aes(xintercept = POS), color = "grey70", size = 1.1) +
-  geom_point(aes(color = phe_name, shape = subpop_name)) +
-  facet_grid(site ~ CHR, scales = "free_x", space = "free_x", switch = "x", ) +
+  geom_point(aes(color = phe_name, shape = subpop_name), size = 0.7, stroke = 0.8) +
+  geom_blank(data = filter(log10p_lim, grepl("K", CHR))) +
+  facet_grid(site ~ CHR, scales = "free_x", space = "free_x", switch = "x") +
       theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), 
               legend.position = "top",
-            panel.background = element_rect(fill = "grey85")) +
+            panel.background = element_rect(fill = "grey85"), panel.spacing.x = unit(0, 'cm')) +
             labs(x = "Chromosome", y = "-log10(p value)") +
-            scale_x_continuous(breaks = c(20000000, 40000000, 60000000, 80000000, 100000000), labels = c("20 Mb", "40 Mb", "60 Mb", "80 Mb", ""), expand = c(0.15, 0.15)) +
-      scale_color_viridis(discrete = TRUE, end = 0.93,
+            scale_x_continuous(breaks = c(20000000, 40000000, 60000000, 80000000), labels = c("20 Mb", "40 Mb", "60 Mb", "80 Mb"), expand = c(0.1, 0.1)) +
+      scale_color_viridis(discrete = TRUE, end = 0.88,
                           option = "B") +
-      scale_shape_manual(values = c(6, 4, 1, 4, 16, 4, 17, 18, 14, 3, 4)) +
-  ylim(c(6, 15)) + geom_hline(yintercept = c(8.3), linetype = 2) + labs(subtitle = "K subgenome")
-save_plot(paste0("../manuscript/plots/", "G_BLUP_Combined_Manhattans_fl_phe_subpops_sites_K_subgenome.png"), 
-              base_height = 4, plot = last_plot())
-```
+      scale_shape_manual(values = c(1, 4, 6, 4, 16, 4, 17, 18, 14, 3, 4)) +
+  ylim(c(6, 15)) + geom_hline(yintercept = c(8.3), linetype = 2) + labs(subtitle = "K subgenome", x = "")
+
+n_subgenome <- log10p_plot %>%
+  filter(grepl("N", CHR)) %>%
+  ggplot(aes(x = POS, y = log10p)) +
+  geom_vline(data = filter(annos_plot1, grepl("N", CHR)),
+             aes(xintercept = POS), color = "grey70", size = 1.1) +
+  geom_point(aes(color = phe_name, shape = subpop_name), size = 0.7, stroke = 0.8) +
+  geom_blank(data = filter(log10p_lim, grepl("N", CHR))) +
+  facet_grid(site ~ CHR, scales = "free_x", space = "free_x", switch = "x") +
+      theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), 
+              legend.position = "none",
+            panel.background = element_rect(fill = "grey85"), panel.spacing.x = unit(0, 'cm')) +
+            scale_x_continuous(breaks = c(20000000, 40000000, 60000000, 80000000), labels = c("20 Mb", "40 Mb", "60 Mb", "80 Mb"), expand = c(0.1, 0.1)) +
+      scale_color_viridis(discrete = TRUE, end = 0.88,
+                          option = "B") +
+      scale_shape_manual(values = c(1, 4, 6, 4, 16, 4, 17, 18, 14, 3, 4)) +
+  ylim(c(6, 15)) + geom_hline(yintercept = c(8.3), linetype = 2) + labs(subtitle = "N subgenome", x = "", y = "-log10(p value)")
+
+both_subgenomes <- plot_grid(k_subgenome, n_subgenome, nrow = 2, rel_heights = c(0.52, 0.47))
+save_plot(paste0("../manuscript/plots/", "G_BLUP_Combined_Manhattans_fl_phe_subpops_sites_by_subgenome.svg"), 
+              base_height = 7.5, base_width = 6.5, plot = both_subgenomes)
+```
+
+## Manhattan QTL windows
+
+```{r}
+qtldf <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv")
+
+qtldf <- qtldf %>%
+  separate(flank_lo, into = c("CHR", "Pos_Mb_lo"), sep = "_", convert = TRUE) %>%
+  separate(flank_hi, into = c("flank", "Pos_Mb_hi"), sep = "_", convert = TRUE) %>%
+  mutate(Pos_Mb_lo = round(Pos_Mb_lo, digits = 2),
+         Pos_Mb_hi = round(Pos_Mb_hi, digits = 2)) %>%
+  select(CHR, Pos_Mb_lo, Pos_Mb_hi, lodcolumn, lod, everything()) %>%
+  arrange(CHR, Pos_Mb_lo)
+
+qtlplot <- qtldf %>%
+  mutate(phe = case_when(lodcolumn == "FL50" ~ "50% flowering",
+                              lodcolumn == "dyln_fl50" ~ "Daylength",
+                              lodcolumn == "dyln_change_sec" ~ "daylength change (s)",
+                              lodcolumn == "CGDD_12C" ~ "GDD, GR to FL",
+                              lodcolumn == "crain_gr2fl" ~ "Rainfall, GR to FL")) %>%
+  filter(phe %in% c("Daylength", "GDD, GR to FL")) %>%
+  mutate(POS_lo = Pos_Mb_lo * 1000000,
+         POS_hi = Pos_Mb_hi * 1000000)
+
+qtlplot$phe <- factor(qtlplot$phe, 
+                               levels = c("50% flowering", "Rainfall, GR to FL",
+                                          "GDD, GR to FL", "Daylength",
+                                          "daylength change (s)",
+                                          "Rainfall, 1 day",
+                                                  #"Rainfall, 2d sum",
+                                                  "Rainfall, 3d sum",
+                                                  "Rainfall, 5d sum"#,
+                                                  #"Rainfall, 7d sum"
+                                                  ))
+```
+
+```{r}
+k_subgenome <- log10p_plot %>%
+  filter(grepl("K", CHR) & phe != "crain_gr2fl") %>%
+  ggplot() +
+  geom_rect(data = filter(qtlplot, grepl("K", CHR)), mapping = aes(ymin = 6, ymax = 15, xmin = POS_lo, xmax = POS_hi, fill = phe), alpha = 0.2, color = "grey70", size = 0.4) +
+  geom_rect(data = filter(annos_plot1, grepl("K", CHR)), mapping = aes(ymin = 6, ymax = 15, xmin = POS_lo, xmax = POS_hi), fill = "grey70", color = "grey10", linetype = 2) +
+  geom_point(aes(x = POS, y = log10p, color = phe_name, shape = subpop_name),
+             size = 0.65, stroke = 0.75) +
+  geom_blank(data = filter(log10p_lim, grepl("K", CHR)), aes(x = POS, y = log10p)) +
+  facet_grid(site ~ CHR, scales = "free_x", space = "free_x", switch = "x") +
+      theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), 
+              legend.position = "top",
+            panel.background = element_rect(fill = "grey90"), panel.spacing.x = unit(0, 'cm')) +
+            labs(x = "Chromosome", y = "-log10(p value)") +
+            scale_x_continuous(breaks = c(20000000, 40000000, 60000000,
+                                          80000000), 
+                               labels = c("20 Mb", "40 Mb", "60 Mb", "80 Mb"),
+                               expand = c(0.1, 0.1)) +
+      scale_color_viridis(discrete = TRUE, begin = 0.35, end = 0.75, option = "B") +
+      scale_fill_viridis(discrete = TRUE, begin = 0.35, end = 0.75, option = "B") +
+      scale_shape_manual(values = c(1, 4, 6, 4, 16, 4, 17, 18, 14, 3, 4), guide = FALSE) +
+  ylim(c(6, 15)) + geom_hline(yintercept = c(8.3), linetype = 2) + labs(subtitle = "K subgenome", x = "")
+
+n_subgenome <- log10p_plot %>%
+  filter(grepl("N", CHR) & phe != "crain_gr2fl") %>%
+  ggplot() +
+  geom_rect(data = filter(qtlplot, grepl("N", CHR)), mapping = aes(ymin = 6, ymax = 15, xmin = POS_lo, xmax = POS_hi, fill = phe), alpha = 0.2, color = "grey70", size = 0.4) +
+  geom_rect(data = filter(annos_plot1, grepl("N", CHR)), mapping = aes(ymin = 6, ymax = 15, xmin = POS_lo, xmax = POS_hi), fill = "grey70", color = "grey10", linetype = 2) +
+  geom_point(aes(x = POS, y = log10p, color = phe_name, shape = subpop_name),
+             size = 0.65, stroke = 0.75) +
+  geom_blank(data = filter(log10p_lim, grepl("N", CHR)), aes(x = POS, y = log10p)) +
+  facet_grid(site ~ CHR, scales = "free_x", space = "free_x", switch = "x") +
+      theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), 
+              legend.position = "top",
+            panel.background = element_rect(fill = "grey90"), panel.spacing.x = unit(0, 'cm')) +
+            labs(x = "Chromosome", y = "-log10(p value)") +
+            scale_x_continuous(breaks = c(20000000, 40000000, 60000000,
+                                          80000000), 
+                               labels = c("20 Mb", "40 Mb", "60 Mb", "80 Mb"),
+                               expand = c(0.1, 0.1)) +
+      scale_color_viridis(discrete = TRUE, begin = 0.35, end = 0.75, option = "B", guide = FALSE) +
+      scale_fill_viridis(discrete = TRUE, begin = 0.35, end = 0.75, option = "B", guide = FALSE) +
+      scale_shape_manual(values = c(1, 4, 6, 4, 16, 4, 17, 18, 14, 3, 4)) +
+  ylim(c(6, 15)) + geom_hline(yintercept = c(8.3), linetype = 2) + labs(subtitle = "N subgenome", x = "")
+
+both_subgenomes <- plot_grid(k_subgenome, n_subgenome, nrow = 2)
+save_plot(paste0("../manuscript/plots/", "G_BLUP_Combined_Manhattans_GDD_dyln_",
+                 "subpops_sites_by_subgenome_with_QTL_and_20kb_",
+                 "common_associations.png"), base_height = 7, 
+          base_width = 7, plot = both_subgenomes)
+```
+
+## Manhattan QTL windows
+
+### GDD & DYLN only

### 2020-08-13 R/Plots_manuscript_v0.3.Rmd
Add example annotation script for lab

@@
 j = 1
-annos_df <- annos1 <- read_csv(file = file.path(workingdir, "analysis", "gwas", "BLUPs",
+annos_df <- read_csv(file = file.path(workingdir, "analysis", "gwas", "BLUPs",
                                     paste0(names(set_v)[j],     

### 2020-08-13 R/Example_annotation_script.R

@@ -1,0 +1,10 @@
+library(tibble)
+library(switchgrassGWAS)
+library(AnnotationDbi)
+txdb <- loadDb(file = file.path("/", "home", "SharedUser", "PV",
+                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
+
+df <- tibble(CHR = "Chr03K", start = 50800000, end = 50900000)
+
+annos <- pvdiv_table_topsnps(df = df, type = "table", txdb = txdb)
+# annos should be a dataframe containing one gene, Pavir.3KG393200

### 2020-08-31 R/Plots_manuscript_v0.3.Rmd
Add more information about DOF candidate gene

@@
 Pavir.
+```{r}
+annos_dyln$top500SNPs_within50000bp %>%
+  filter(CHR == "Chr05N" & region_start > 64700000)
+
+annos_df %>%
+  filter(CHR == "Chr05N" & POS_Mb > 64.6)
+
+annos_dyln$top500SNPs_within50000bp %>%
+  filter(CHR == "Chr05N" & region_start < 3000000 & region_start > 2000000) %>%
+  select(`Gene ID`, `distance from gene`, `Arabidopsis thaliana homolog`, `Rice homolog`)
+```
+	Pavir.5NG109746 (18.7kb), 	AT1G49200 RING/U-box superfamily protein
+		LOC_Os01g64620.1 highly expressed hypocotyl, silique
+	Pavir.5NG109743 (22.5kb), 	AT2G22590 UDP-Glycosyltransferase superfamily pt.
+	LOC_Os01g64540.1  highly expressed in mature flower petals
+	Pavir.5NG109744 (17.9kb), 		AT2G35930 (PUB23) plant U-box 23
+		LOC_Os01g64570.1 highly expressed in senescent leaf petiole; negative regulation of drought tolerance
+	Pavir.5NG109745 (4.1kb)     AT2G28510 Dof-type zinc finger DNA-binding family protein  highly expressed in the carpel of the young flower & in seeds during germination day 1; promotes leaf senescence in arabidopsis
+		LOC_Os01g64590.1
 

### 2020-09-30 R/Plots_manuscript_G_BLUP.Rmd
Start comparisons of GWAS and QTL intervals with FT annotations

--- 2020-08-13 R/Plots_manuscript_v0.3.Rmd
@@
 ---
-title: "Plots_manuscript_v0.3"
+title: "Plots_manuscript_G_BLUP"
 author: "Alice MacQueen"
-date: "7/22/2020"
+date: "9/30/2020"
 output: html_document
@@
 
-Add any libraries/scripts that need to be loaded for these analyses here.
 ```{r setup, include=FALSE}
@@
 library(tidyverse)
-library(cowplot)
-library(viridis)
-library(lubridate)
-library(switchgrassGWAS)
-library(GGally)
-library(maps)
-library(ggmap)
-usa <- map_data("usa")
-source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
 ```
@@
 # Load data
-Add any data that needs to be loaded for these analyses here.
+
+Xiaoyu's annotation file, QTL 
 ```{r}
@@
 sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
-subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
-phe_gr_df <- read_rds(file.path(workingdir, "data",
-                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
-phe_fl_df <- read_rds(file.path(workingdir, "data",
-                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
-phenotypes <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
-h2gr <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "h2_df_GR50_related_functions_of_weather_v1.csv"))
-h2fl <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "h2_df_FL50_related_functions_of_weather_v1.csv"))
-h2gr_gxe <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "gxe_and_h2_df_GR50_related_functions_of_weather_v1.csv"))
-h2fl_gxe <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "gxe_and_h2_df_FL50_related_functions_of_weather_v1.csv"))
-
-h2gr <- h2gr %>%
-  full_join(h2gr_gxe)
-h2fl <- h2fl %>%
-  full_join(h2fl_gxe)
-```
-as well as the code for any new analyses that I'm running to do new writeup for the manuscript v0.3.
-
-v0.2 has the ASReml h2 analysis, has BLUPs for G (which as of v0.3 probably needs a GxE term... define some possibilities below.)
-
-Results: h2 at single sites. Then h2 at all eight sites. And so there must be GxE, and there must be a lot of rank-changing GxE, in that we can’t predict ranks across sites.
-
-To show this, make barplots of the different variables and facet them by site.
-Then, show heritability across all eight sites (in a model with a random effect of site, interacted with some number of eigenvectors from a SVD of the kinship matrix.)
-
-# Map Fig 1A
-
-```{r}
-subpops
-library(ggrepel)
-
-metaplot <- metadata %>%
-  filter(SUBPOP %in% c("Gulf", "Midwest") & !is.na(LATITUDE)) %>%
-  mutate(LONGITUDE = -LONGITUDE)
-metaplot$ECOTYPE_NNET <- factor(metaplot$ECOTYPE_NNET, levels = c("Coastal", "Lowland", "Upland", "Unknown"))
-siteplot <- sites %>%
-  filter(!(SITE %in% c("FRMI", "OVTN")))
-
-ggplot() + 
-  geom_polygon(data = usa, aes(x=long, y = lat, group = group), 
-               fill = "white", color = "black") +
-  coord_quickmap(xlim = c(-105, -66), ylim = c(25, 50)) +
-  geom_jitter(data = metaplot, aes(x = LONGITUDE, y = LATITUDE, 
-                                   color = SUBPOP, shape = ECOTYPE_NNET),
-              width = 0.45, height = 0.45) + 
-  geom_point(data = siteplot, aes(x = Longitude, y = Latitude), shape = 21, size = 4, stroke = 1.5, fill = "white") +
-  geom_point(data = siteplot, aes(x = Longitude, y = Latitude), shape = 4, size = 3, stroke = 1.5) +
-  geom_label_repel(data = siteplot, aes(x = Longitude, y = Latitude, label = manu_site), alpha = 0.7) +
-  labs(x = "Longitude", y = "Latitude") + 
-  theme(legend.position = "right") +
-  scale_color_manual(values = c("#F47F72", "#442C83")) +
-  scale_shape_manual(values = c(0, 6, 1, 4)) 
-  #scale_fill_manual(values = c("#280B54FF","#FAC127FF", "#D44842FF", "white")) +
-  #scale_color_manual(values = c(rep("black", 10), "#280B54FF","#FAC127FF", "#D44842FF", "white", "#280B54FF","#FAC127FF", "#D44842FF", "white"), guide = FALSE) +
-  #scale_alpha_manual(values = c(0.7, 0.7, 0.7, 0.7), guide = FALSE)
-save_plot(filename = "../manuscript/plots/Map_subpops_and_Ecotypes.png", plot = last_plot(), base_height = 3)
-?geom_point
+ftanno <- read_csv(file.path(workingdir, "data/Pvirgatum_v5.1.annotation_FTGeneHomologs.csv"))
 ```
 
-# -- Figure 1 B
-# Phenotype correlations
-can you easily look at the correlation between greenup ranks at OK versus NE?
-```{r}
-phe_gr_df %>%
-  filter(SITE %in% c("LINC", "STIL")) %>%
-  ungroup() %>%
-  select(PLANT_ID:GR50, -PLOT_GL) %>%
-  pivot_wider(names_from = "SITE", values_from = "GR50", values_fn = mean) %>%
-  ggplot(aes(x = STIL, y = LINC)) +
-  geom_hex() + #aes(color = SUBPOP), height = 1, width = 1) +
-  scale_fill_viridis(option = "B",
-                     end = 0.9, direction = -1)
+# Create dataframes to plot (usually end with "_plot")
 
-phe_gr_df %>%
-  filter(SITE %in% c("STIL", "PKLE")) %>%
-  ungroup() %>%
-  select(PLANT_ID:GR50, -PLOT_GL) %>%
-  pivot_wider(names_from = "SITE", values_from = "GR50", values_fn = mean) %>%
-  ggplot(aes(x = STIL, y = Lat_origin)) +
-  geom_hex() + #aes(color = SUBPOP), height = 1, width = 1) +
-  scale_fill_viridis(option = "B",
-                     end = 0.9, direction = -1)
-
-cor_gr <- phe_gr_df %>%
-  left_join(select(sites, SITE, manu_site)) %>%
-  ungroup() %>%
-  select(PLANT_ID:GR50, manu_site, -PLOT_GL, -SITE) %>%
-  pivot_wider(names_from = "manu_site", values_from = "GR50", values_fn = mean)
-cor_fl <- phe_fl_df %>%
-  left_join(select(sites, SITE, manu_site)) %>%
-  ungroup() %>%
-  select(PLANT_ID:FL50, manu_site, -PLOT_GL, -SITE) %>%
-  pivot_wider(names_from = "manu_site", values_from = "FL50", values_fn = mean)
-
-cor_north <- phe_gr_df %>%
-  left_join(select(sites, SITE, manu_site)) %>%
-  ungroup() %>%
-  filter(Lat_origin > 40) %>%
-  select(PLANT_ID:GR50, manu_site, -PLOT_GL, -SITE) %>%
-  pivot_wider(names_from = "manu_site", values_from = "GR50", values_fn = mean)
-cor_south <- phe_gr_df %>%
-  left_join(select(sites, SITE, manu_site)) %>%
-  ungroup() %>%
-  filter(Lat_origin < 32) %>%
-  select(PLANT_ID:GR50, manu_site, -PLOT_GL, -SITE) %>%
-  pivot_wider(names_from = "manu_site", values_from = "GR50", values_fn = mean)
-cor(cor_north[,5:13], use = "pairwise") %>%
-  switchgrassGWAS:::reorder_cormat(.) %>%
-  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
-cor(cor_south[,5:13], use = "pairwise") %>%
-  switchgrassGWAS:::reorder_cormat(.) %>%
-  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
-
-cor(cor_gr$STIL, cor_gr$LINC, use = "pairwise")
-cor(cor_gr[,5:13], use = "pairwise") %>%
-  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
-cor(cor_fl[,5:13], use = "pairwise") %>%
-  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
-
-# two different ways. Think I like the second more. Force latitudinal order.
-cor_gr %>%
-  filter(SUBPOP %in% c("Midwest", "Gulf")) %>%
-  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
-  cor(use = "pairwise") %>%
-  switchgrassGWAS:::reorder_cormat(.) %>%
-  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
-```
-
-```{r}
-pand <- cor_fl %>%
-  filter(SUBPOP %in% c("Gulf")) %>%
-  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
-  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
-  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
-                                      "white", "#27AD81FF", "#5DC863FF",
-                                      "#FDE725FF"), limits = c(-1,1)) +
-  theme(legend.position = "none")
-pane <- cor_fl %>%
-  filter(SUBPOP %in% c("Midwest")) %>%
-  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
-  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
-  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
-                                      "white", "#27AD81FF", "#5DC863FF",
-                                      "#FDE725FF"), limits = c(-1,1)) +
-  theme(legend.position = "none")
-panf <- cor_fl %>%
-  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
-  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
-  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
-  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
-                                      "white", "#27AD81FF", "#5DC863FF",
-                                      "#FDE725FF"), limits = c(-1,1)) +
-  theme(legend.position = "none")
-pana <- cor_gr %>%
-  filter(SUBPOP %in% c("Gulf")) %>%
-  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
-  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
-  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
-                                      "white", "#27AD81FF", "#5DC863FF",
-                                      "#FDE725FF"), limits = c(-1,1)) +
-  theme(legend.position = c(0.4, 0.85))
-panb <- cor_gr %>%
-  filter(SUBPOP %in% c("Midwest")) %>%
-  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
-  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
-  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
-                                      "white", "#27AD81FF", "#5DC863FF",
-                                      "#FDE725FF"), limits = c(-1,1)) +
-  theme(legend.position = "none")
-panc <- cor_gr %>%
-  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
-  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
-  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
-  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
-                                      "white", "#27AD81FF", "#5DC863FF",
-                                      "#FDE725FF"), limits = c(-1,1)) +
-  theme(legend.position = "none")
-
-pheno_corr <- plot_grid(pana, panb, panc, pand, pane, panf, nrow = 2)
-save_plot(filename = "../manuscript/plots/Phenotypic_correlations_GR50_FL50_2.svg", plot = pheno_corr, base_width = 6.5)
-```
-
-# --- Fig 1 C & D
-# Heritability Plots
-
-Within common gardens, h2 were typically quite high: 62% on average for greenup date, and 87% for flowering date.
-However, h2 were not consistently high, notably for greenup at our OK and NE common gardens.
-```{r}
-h2all <- h2fl %>%
-  full_join(h2gr)
-
-h2all %>%
-  filter(phe %in% c("FL50", "GR50") & site %in% c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD")) %>%
-  group_by(phe) %>%
-  summarise(meanh2 = mean(Estimate))
-
-h2all %>%
-  filter(phe %in% c("FL50", "GR50") & site %in% c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD")) %>%
-  filter(Estimate < 0.2)
-
-h2all %>%
-  filter(phe %in% c("FL50", "GR50") & site %in% c("eight_sites")) %>%
-  group_by(phe) %>%
-  summarise(meanh2 = mean(Estimate))
-
-h2all %>%
-  filter(phe %in% c("GR50") & site %in% c("tx_sites", "north_sites")) %>%
-  mutate(GandGxE = Estimate + GxE_Estimate) %>%
-  select(GandGxE, Estimate:phe, GxE_Estimate) %>%
-  group_by(site) %>%
-  summarise(meanh2 = mean(Estimate + GxE_Estimate))
-
-h2all %>%
-  filter(site %in% c("eight_sites")) %>%
-  mutate(GandGxE = Estimate + GxE_Estimate)  %>%
-  select(GandGxE, site:phe, Estimate, SE, GxE_Estimate, GxE_SE)
-```
-
-## h2 as functions of GR50 and FL50
-```{r}
-h2all <- h2fl %>%
-  full_join(h2gr)
-unique(h2all$site)
-f1h2 <- h2all %>%
-  filter(phe %in% c("FL50", "GR50")) %>%
-  mutate(site = case_when(is.na(site) ~ "all",
-                          site == "eight_sites" ~ "All",
-                          site == "tx_sites" ~ "Texas",
-                          site == "north_sites" ~ "North",
-                          TRUE ~ site),
-         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
-                                 subpop == "three_subpops" ~ "all",
-                                 TRUE ~ subpop),
-         phe_name = case_when(phe == "GR50" ~ "50% greenup",
-                              phe == "FL50" ~ "50% flowering",
-                              TRUE ~ phe))
-f1h2$site <- factor(f1h2$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All"))#, "Texas", "North"))
-f1h2$phe_name <- factor(f1h2$phe_name, levels = c("50% greenup", "50% flowering"))
-f1h2$subpop_name <- factor(f1h2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "all"))
-ggh21 <- f1h2 %>%
-  filter(!(site %in% c("nine_sites", "IL")) & !is.na(subpop_name) & !is.na(phe_name) & !is.na(site)) %>%
-  ggplot(aes(x = site, y = Estimate)) +
-  geom_bar(aes(fill = site %in% c("All", "Texas", "North")), stat = "identity") +
-  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
-  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
-  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
-  facet_grid(phe_name ~ subpop_name) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), 
-        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
-        legend.position = "none") + 
-  ylim(c(0, 1)) +
-  xlab("Common Gardens") + 
-  ylab(bquote('h'^2)) +
-  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
-  #
-save_plot("../manuscript/plots/h2_GR_FL_site_by_subpop_k_specific.png", plot = ggh21, base_width = 5.5, base_height = 3)
-```
-
-## h2 flowering weather
-```{r}
-flh2 <- h2fl %>%
-  filter(phe %in% c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d")) %>%
-  mutate(site = case_when(is.na(site) ~ "all",
-                          site == "eight_sites" ~ "All",
-                          site == "tx_sites" ~ "Texas",
-                          site == "north_sites" ~ "North",
-                          TRUE ~ site),
-         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
-                                 subpop == "three_subpops" ~ "all",
-                                 TRUE ~ subpop),
-         phe_name = case_when(phe == "FL50" ~ "50% flowering",
-                              phe == "dyln_fl50" ~ "daylength",
-                              phe == "dyln_change_sec" ~ "daylength change (s)",
-                              phe == "cgdd_12c_gr2fl" ~ "GDD, GR to FL",
-                              phe == "crain_gr2fl" ~ "Rainfall, GR to FL",
-                              phe == "crain_1d" ~ "Rainfall, 1 day",
-                              phe == "crain_2d" ~ "Rainfall, 2d sum",
-                              phe == "crain_3d" ~ "Rainfall, 3d sum",
-                              phe == "crain_5d" ~ "Rainfall, 5d sum",
-                              phe == "crain_7d" ~ "Rainfall, 7d sum",
-                              TRUE ~ phe))
-flh2$site <- factor(flh2$site, levels = rev(c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North")))
-flh2$phe_name <- factor(flh2$phe_name, levels = c("50% flowering", "daylength",
-                                                  "daylength change (s)",
-                                                  "GDD, GR to FL",
-                                                  "Rainfall, GR to FL",
-                                                  "Rainfall, 1 day",
-                                                  #"Rainfall, 2d sum",
-                                                  "Rainfall, 3d sum",
-                                                  "Rainfall, 5d sum"#,
-                                                  #"Rainfall, 7d sum"
-                                                  ))
-flh2$subpop_name <- factor(flh2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "all"))
-site_plot <- rev(c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD", "All", "Texas", "North"))
-subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))
-
-flh2_p1 <- flh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
-  ggplot(aes(x = phe_name, y = Estimate)) +
-  geom_bar(stat = "identity", aes(fill = phe_name == "50% flowering")) +
-  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
-  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
-  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
-  facet_grid(site ~ subpop_name) +
-  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 1.1)) +
-  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
-
-save_plot("../manuscript/plots/h2_FL_site_weather_functions_single_sites_k_specific.png", plot = flh2_p1, base_height = 7, base_asp = 1.1)
-
-site_plot <- c("All", "Texas", "North")
-
-flh2_p2 <- flh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
-  ggplot(aes(x = phe_name, y = Estimate)) +
-  geom_bar(stat = "identity", aes(fill = phe_name == "50% flowering")) +
-  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
-  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
-  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
-  facet_grid(site ~ subpop_name) +
-  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.8)) +
-  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")
-
-save_plot("../manuscript/plots/h2_FL_site_weather_functions_site_groups_k_specific.png", plot = flh2_p2, base_height = 4, base_asp = 1.5)
-
-site_plot <- c("All")
-
-flh2_p3 <- flh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
-  ggplot(aes(x = phe_name, y = Estimate)) +
-  geom_bar(stat = "identity", aes(fill = phe_name == "50% flowering")) +
-  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
-  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
-  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
-  facet_grid(site ~ subpop_name) +
-  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.5)) +
-  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")
-save_plot("../manuscript/plots/h2_FL_site_weather_functions_All_Sites_k_specific.png", plot = flh2_p3, base_height = 3.5)
-```
-## gxe flowering all, north, texas
-```{r}
-site_plot <- c("All", "Texas", "North")
-
-fl_barplot2 <- flh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>%
-  select(site:phe, subpop_name, phe_name, ends_with("SE")) %>%
-  pivot_longer(cols = ends_with("SE"), names_to = "Type", values_to = "SE") %>%
-  mutate(type_name = case_when(Type == "SE" ~ "G",
-                               Type == "GxE_SE" ~ "GxE",
-                               Type == "error_SE" ~ "error",
-                               Type == "Env_SE" ~ "E")) %>%
-  select(-Type)
-fl_barplot <- flh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>%
-  select(site:phe, subpop_name, phe_name, ends_with("Estimate")) %>%
-  pivot_longer(cols = ends_with("Estimate"), names_to = "Type", values_to = "Estimate") %>%
-  mutate(type_name = case_when(Type == "Estimate" ~ "G",
-                               Type == "GxE_Estimate" ~ "GxE",
-                               Type == "error_Estimate" ~ "error",
-                               Type == "Env_Estimate" ~ "E"))
-fl_barplot <- fl_barplot %>%
-  left_join(fl_barplot2)
-fl_barplot$type_name <- factor(fl_barplot$type_name, levels = rev(c("G", "GxE", "E", "error")))
-
-fl_bar_gxe <- fl_barplot %>%
-  ggplot(aes(x = phe_name, y = Estimate)) +
-  geom_bar(stat = "identity", aes(fill = type_name)) + 
-  geom_errorbar(data = filter(fl_barplot, type_name == "G"), aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
-  facet_grid(site ~ subpop_name)+
-  scale_fill_viridis(discrete = TRUE, direction = -1) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") +
-  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = GxE_Estimate+Estimate), linetype = 2) + xlab("") +
-  ylab("Variance Explained")
-
-save_plot("../manuscript/plots/Var_Comp_flowering_site_groups.png", plot = fl_bar_gxe, base_width = 5.5, base_height = 4)
-
-flh2_p4 <- flh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
-  ggplot(aes(x = phe_name, y = GxE_Estimate)) +
-  geom_bar(stat = "identity", aes(fill = phe_name == "50% flowering")) +
-  geom_errorbar(aes(ymin = GxE_Estimate - GxE_SE, ymax = GxE_Estimate + GxE_SE), width = 0.3) +
-  geom_linerange(aes(ymin = GxE_Estimate, ymax = GxE_Estimate + GxE_SE)) +
-  geom_linerange(aes(ymin = GxE_Estimate - GxE_SE, ymax = GxE_Estimate)) +
-  facet_grid(site ~ subpop_name) +
-  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = GxE_Estimate), linetype = 2) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 1)) +
-  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")
-
-save_plot("../manuscript/plots/h2_FL_site_weather_functions_site_groups_k_specific.png", plot = flh2_p2, base_height = 4, base_asp = 1.5)
-
-library(scales)
-show_col(viridis(4)) # #440154FF   #31688EFF
-```
-
-## h2 greenup weather
-
-```{r}
-unique(h2gr$site)
-grh2 <- h2gr %>%
-  filter(!(phe %in% c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d", "DYLN"))) %>%
-  mutate(site = case_when(is.na(site) ~ "all",
-                          site == "eight_sites" ~ "All",
-                          site == "tx_sites" ~ "Texas",
-                          site == "north_sites" ~ "North",
-                          TRUE ~ site),
-         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
-                                 subpop == "three_subpops" ~ "All",
-                                 TRUE ~ subpop),
-         phe_name = case_when(phe == "GR50" ~ "50% greenup",
-                              phe == "FL50" ~ "50% flowering",
-                              phe == "cgdd_12c_jan2gr" ~ "GDD 01-01 to GR",
-                              phe == "cgdd_8c_jan2gr" ~ "GDD 8C 01-01 to greenup",
-                              phe == "tmax_18d" ~ "Max T, 18d mean",
-                              phe == "tmin_18d" ~ "Min T, 18d mean",
-                              phe == "tave_18d" ~ "Mean T, 18d mean",
-                              phe == "tmax_10d" ~ "Max T, 10d mean",
-                              phe == "tmin_10d" ~ "Min T, 10d mean",
-                              phe == "tave_10d" ~ "Mean T, 10d mean",
-                              phe == "tmax_5d" ~ "Max T, 5d mean",
-                              phe == "tmin_5d" ~ "Min T, 5d mean",
-                              phe == "tave_5d" ~ "Mean T, 5d mean",
-                              phe == "cgdd_12c_10d" ~ "GDD, 10d",
-                              phe == "cgdd_12c_18d" ~ "GDD, 18d",
-                              phe == "cgdd_12c_5d" ~ "GDD, 5d",
-                              TRUE ~ phe))
-grh2$site <- factor(grh2$site, levels = rev(c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North", "nine_sites")))
-grh2$phe_name <- factor(grh2$phe_name, levels = c("50% greenup", "GDD, 5d",
-                                                  "Min T, 5d mean",
-                                                  #"Mean T, 5d mean",
-                                                  #"Max T, 5d mean",
-                                                  "GDD, 10d",
-                                                  "Min T, 10d mean",
-                                                  #"Mean T, 10d mean",
-                                                  #"Max T, 10d mean",
-                                                  "GDD, 18d",
-                                                  "Min T, 18d mean",
-                                                  #"Mean T, 18d mean",
-                                                  #"Max T, 18d mean",
-                                                  #"GDD 8C 01-01 to greenup",
-                                                  "GDD 01-01 to GR"))
-grh2$subpop_name <- factor(grh2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "All"))
-
-subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))
-
-site_plot <- c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North")
-
-grh2_p1 <- grh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
-  ggplot(aes(x = phe_name, y = Estimate)) +
-  geom_bar(stat = "identity", aes(fill = phe_name == "50% greenup")) +
-  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
-  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
-  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
-  facet_grid(site ~ subpop_name) +
-  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 1.1)) +
-  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
-save_plot("../manuscript/plots/h2_GR_site_weather_functions_single_sites_k_specific.png", plot = grh2_p1, base_height = 7.5, base_asp = 1.1)
-
-site_plot <- c( "All", "Texas", "North")
-
-grh2_p2 <- grh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
-  ggplot(aes(x = phe_name, y = Estimate)) +
-  geom_bar(stat = "identity", aes(fill = phe_name == "50% greenup")) +
-  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
-  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
-  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
-  facet_grid(site ~ subpop_name) +
-  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.6)) +
-  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
-save_plot("../manuscript/plots/h2_GR_site_weather_functions_site_groups_k_specific.png", plot = grh2_p2, base_height = 5, base_asp = 1.3)
-
-site_plot <- c( "All")
-
-grh2_p3 <- grh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
-  ggplot(aes(x = phe_name, y = Estimate)) +
-  geom_bar(stat = "identity", aes(fill = phe_name == "50% greenup")) +
-  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
-  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
-  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
-  facet_grid(site ~ subpop_name) +
-  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.11)) +
-  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")
-save_plot("../manuscript/plots/h2_GR_site_weather_functions_All_Sites_k_specific.png", plot = grh2_p3, base_height = 3.5)
-
-gr_fl_p1 <- plot_grid(grh2_p3, flh2_p3, nrow = 2)
-
-save_plot("../manuscript/plots/h2_GR_and_FL_site_weather_functions_All_Sites_k_specific.png", plot = gr_fl_p1, base_height = 4.6)
-```
-## gxe greenup
-
-```{r}
-site_plot <- c("All", "Texas", "North")
-
-gr_barplot2 <- grh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>%
-  select(site:phe, subpop_name, phe_name, ends_with("SE")) %>%
-  pivot_longer(cols = ends_with("SE"), names_to = "Type", values_to = "SE") %>%
-  mutate(type_name = case_when(Type == "SE" ~ "G",
-                               Type == "GxE_SE" ~ "GxE",
-                               Type == "error_SE" ~ "error",
-                               Type == "Env_SE" ~ "E")) %>%
-  select(-Type)
-gr_barplot <- grh2 %>%
-  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>%
-  select(site:phe, subpop_name, phe_name, ends_with("Estimate")) %>%
-  pivot_longer(cols = ends_with("Estimate"), names_to = "Type", values_to = "Estimate") %>%
-  mutate(type_name = case_when(Type == "Estimate" ~ "G",
-                               Type == "GxE_Estimate" ~ "GxE",
-                               Type == "error_Estimate" ~ "error",
-                               Type == "Env_Estimate" ~ "E"))
-gr_barplot <- gr_barplot %>%
-  left_join(gr_barplot2)
-gr_barplot$type_name <- factor(gr_barplot$type_name, levels = rev(c("G", "GxE", "E", "error")))
-
-gr_bar_gxe <- gr_barplot %>%
-  ggplot(aes(x = phe_name, y = Estimate)) +
-  geom_bar(stat = "identity", aes(fill = type_name)) + 
-  geom_errorbar(data = filter(gr_barplot, type_name == "G"), aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
-  facet_grid(site ~ subpop_name)+
-  scale_fill_viridis(discrete = TRUE, direction = -1) +
-  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "right") +
-  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = GxE_Estimate+Estimate), linetype = 2)
-
-save_plot("../manuscript/plots/Var_Comp_greenup_site_groups.png", plot = gr_bar_gxe, base_height = 4, base_asp = 1.5)
-```
-
-# --------------
-# SVD of K
-
-## Plot eigenvectors to capture predominant patterns in K matrix.
-```{r}
-k_full <- read_rds("../data/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
-k_gandm <- read_rds("../data/Kinship_van_Raden_method_363g_Gulf_and_Midwest_subset_maf0.05.rds")
-k_midw <- read_rds("../data/Kinship_van_Raden_method_134g_Midwest_subset_maf0.05.rds")
-k_gulf <- read_rds("../data/Kinship_van_Raden_method_229g_Gulf_subset_maf0.05.rds")
-```
-
-```{r}
-b1 <- svd(k_full)
-b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 11.5
-b1$d[2] / sum(b1$d) # 7.1
-b1$d[3] / sum(b1$d) # 6.1
-b1$d[4] / sum(b1$d) # 1.7
-b1$d[5] / sum(b1$d) # 1.3
-sum(b1$d[1:5]) / sum(b1$d) # First 5 eigenvectors explain 1% or more variance, and 27.7% of total variance.
-for(g in 1:5){
-barplot(b1$v[,g]/k_full[which.max(abs(k_full[,g])),g])
-}
-order <- enframe(colnames(k_full), name = NULL, value = "PLANT_ID")
-ggb1 <- tibble(`First eigenvector (PVE = 11.5%)` = b1$v[,1]/k_full[which.max(abs(k_full[,1])),1],
-       `Second eigenvector (PVE = 7.1%)` = b1$v[,2]/k_full[which.max(abs(k_full[,2])),2],
-       `Third eigenvector (PVE = 6.1%)` = b1$v[,3]/k_full[which.max(abs(k_full[,3])),3],
-       `Fourth eigenvector (PVE = 1.7%)` = b1$v[,4]/k_full[which.max(abs(k_full[,4])),4],
-       `Fifth eigenvector (PVE = 1.3%)` = b1$v[,5]/k_full[which.max(abs(k_full[,5])),5],
-       `Sixth eigenvector (PVE = 0.9%)` = b1$v[,6]/k_full[which.max(abs(k_full[,6])),6],
-       PLANT_ID = order$PLANT_ID) %>%
-  left_join(metadata, by = "PLANT_ID") %>%
-  dplyr::select(PLANT_ID, SUBPOP, ECOTYPE_NNET, LATITUDE, ends_with("%)"))
-
-ggb1 <- ggb1 %>%
-  arrange(LATITUDE) %>%
-  mutate(rank = row_number()) %>%
-  pivot_longer(cols = ends_with("%)"), names_to = "Eigenvector", values_to = "value")
-ggb1$Eigenvector <- factor(ggb1$Eigenvector, 
-                           levels = c("First eigenvector (PVE = 11.5%)", 
-                                      "Second eigenvector (PVE = 7.1%)", 
-                                      "Third eigenvector (PVE = 6.1%)", 
-                                      "Fourth eigenvector (PVE = 1.7%)",
-                                      "Fifth eigenvector (PVE = 1.3%)",
-                                      "Sixth eigenvector (PVE = 0.9%)"
-                                      ))
-ggb1_plot <- ggb1 %>%
-  ggplot(aes(x = rank, y = value)) +
-  geom_bar(aes(color = SUBPOP, fill = SUBPOP), stat = "identity") +
-  facet_wrap(~ Eigenvector) +
-  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
-  scale_color_viridis(discrete = TRUE, end = 0.9, direction = -1) +
-  scale_fill_viridis(discrete = TRUE, end = 0.9, direction = -1) +
-  xlab("Latitude of Origin (Rank Order)") + ylab("") +
-  geom_vline(xintercept = 607, linetype = 2)
-save_plot("../manuscript/plots/SVD_of_Kinship_matrix_first_six_eigenvectors.png", base_height = 5, base_asp = 2, plot = ggb1_plot)
-
-ggb1_plot2 <- ggb1 %>%
-  ggplot(aes(x = rank, y = value)) +
-  geom_bar(aes(color = ECOTYPE_NNET, fill = ECOTYPE_NNET), stat = "identity") +
-  facet_wrap(~ Eigenvector) +
-  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
-  scale_color_viridis(discrete = TRUE, option = "B", end = 0.9, 
-                      direction = -1) +
-  scale_fill_viridis(discrete = TRUE, option = "B", end = 0.9, direction = -1) +
-  xlab("Latitude of Origin (Rank Order)") + ylab("") +
-  geom_vline(xintercept = 607, linetype = 2)
-save_plot("../manuscript/plots/SVD_of_Kinship_matrix_first_six_eigenvectors_ecotype.png", base_height = 5, base_asp = 2, plot = ggb1_plot2)
-```
-
-## k_gandm
-
-```{r}
-b1 <- svd(k_gandm)
-b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 29.2
-b1$d[2] / sum(b1$d) # 18.2
-b1$d[3] / sum(b1$d) # 3.7
-b1$d[4] / sum(b1$d) # 1.4
-b1$d[5] / sum(b1$d) # 0.008
-sum(b1$d[1:4]) / sum(b1$d) # First 4 eigenvectors explain 1% or more variance, and 52.5% of total variance.
-
-for(g in 1:5){
-barplot(b1$v[,g]/k_gandm[which.max(abs(k_gandm[,g])),g])
-}
-order <- enframe(colnames(k_gandm), name = NULL, value = "PLANT_ID")
-ggb1 <- tibble(`First eigenvector (PVE = 29.2%)` = b1$v[,1]/k_gandm[which.max(abs(k_gandm[,1])),1],
-       `Second eigenvector (PVE = 18.2%)` = b1$v[,2]/k_gandm[which.max(abs(k_gandm[,2])),2],
-       `Third eigenvector (PVE = 3.7%)` = b1$v[,3]/k_gandm[which.max(abs(k_gandm[,3])),3],
-       `Fourth eigenvector (PVE = 1.4%)` = b1$v[,4]/k_gandm[which.max(abs(k_gandm[,4])),4],
-       #`Fifth eigenvector (PVE = 0.8%)` = b1$v[,5]/k_gandm[which.max(abs(k_gandm[,5])),5],
-       PLANT_ID = order$PLANT_ID) %>%
-  left_join(metadata, by = "PLANT_ID") %>%
-  dplyr::select(PLANT_ID, SUBPOP, ECOTYPE_NNET, LATITUDE, ends_with("%)"))
-
-ggb1 <- ggb1 %>%
-  arrange(LATITUDE) %>%
-  mutate(rank = row_number()) %>%
-  pivot_longer(cols = ends_with("%)"), names_to = "Eigenvector", values_to = "value")
-ggb1$Eigenvector <- factor(ggb1$Eigenvector, 
-                           levels = c("First eigenvector (PVE = 29.2%)", 
-                                      "Second eigenvector (PVE = 18.2%)", 
-                                      "Third eigenvector (PVE = 3.7%)", 
-                                      "Fourth eigenvector (PVE = 1.4%)",
-                                      "Fifth eigenvector (PVE = 0.8%)"
-                                      ))
-ggb1$ECOTYPE_NNET <- factor(ggb1$ECOTYPE_NNET, levels = c("Coastal", "Lowland", "Upland", "Unknown"))
-ggb1_plot <- ggb1 %>%
-  ggplot(aes(x = rank, y = value)) +
-  geom_bar(aes(color = SUBPOP, fill = SUBPOP), stat = "identity") +
-  facet_wrap(~ Eigenvector) +
-  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
-  scale_color_manual(values = c("#F47F72", "#442C83")) +
-  scale_fill_manual(values = c("#F47F72", "#442C83")) +
-  xlab("Latitude of Origin (Rank Order)") + ylab("") +
-  geom_vline(xintercept = 342, linetype = 2)
-save_plot("../manuscript/plots/SVD_of_Kinship_matrix_Gulf_and_Midwest_first_four_eigenvectors.png", base_height = 5, base_asp = 2, plot = ggb1_plot)
-
-ggb1_plot2 <- ggb1 %>%
-  ggplot(aes(x = rank, y = value)) +
-  geom_bar(aes(color = ECOTYPE_NNET, fill = ECOTYPE_NNET), stat = "identity") +
-  facet_wrap(~ Eigenvector) +
-  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
-  scale_color_viridis(discrete = TRUE, option = "B", end = 0.9, 
-                      direction = -1) +
-  scale_fill_viridis(discrete = TRUE, option = "B", end = 0.9, direction = -1) +
-  xlab("Latitude of Origin (Rank Order)") + ylab("") +
-  geom_vline(xintercept = 342, linetype = 2)
-save_plot("../manuscript/plots/SVD_of_Kinship_matrix_Gulf_and_Midwest_first_four_eigenvectors_ecotype.png", base_height = 5, base_asp = 2, plot = ggb1_plot2)
-```
-
-## k_midw
-
-```{r}
-b1 <- svd(k_midw)
-b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 10.2
-b1$d[2] / sum(b1$d) # 4.8
-b1$d[3] / sum(b1$d) # 2.4
-b1$d[4] / sum(b1$d) # 2.0
-b1$d[5] / sum(b1$d) # 1.7
-b1$d[6] / sum(b1$d) # 1.3
-b1$d[7] / sum(b1$d) # 1.3
-b1$d[8] / sum(b1$d) # 1.1
-b1$d[9] / sum(b1$d) # 1.1
-b1$d[10] / sum(b1$d) # 1.1
-b1$d[11] / sum(b1$d) # 1.1
-b1$d[12] / sum(b1$d) # 1.0
-b1$d[13] / sum(b1$d) # 1.0
-sum(b1$d[1:13]) / sum(b1$d) # First 13 eigenvectors explain 1% or more variance, and 30% of total variance. Phew! Exhausting.
-```
-
-```{r}
-k_use <- k_midw
-for(g in 1:5){
-barplot(b1$v[,g]/k_use[which.max(abs(k_use[,g])),g])
-}
-order <- enframe(colnames(k_use), name = NULL, value = "PLANT_ID")
-ggb1 <- tibble(`First eigenvector (PVE = 10.2%)` = b1$v[,1]/k_use[which.max(abs(k_use[,1])),1],
-       `Second eigenvector (PVE = 4.8%)` = b1$v[,2]/k_use[which.max(abs(k_use[,2])),2],
-       `Third eigenvector (PVE = 2.4%)` = b1$v[,3]/k_use[which.max(abs(k_use[,3])),3],
-       `Fourth eigenvector (PVE = 2.0%)` = b1$v[,4]/k_use[which.max(abs(k_use[,4])),4],
-       `Fifth eigenvector (PVE = 1.7%)` = b1$v[,5]/k_use[which.max(abs(k_use[,5])),5],
-       `Sixth eigenvector (PVE = 1.3%)` = b1$v[,6]/k_use[which.max(abs(k_use[,6])),6],
-       PLANT_ID = order$PLANT_ID) %>%
-  left_join(metadata, by = "PLANT_ID") %>%
-  dplyr::select(PLANT_ID, SUBPOP, ECOTYPE_NNET, LATITUDE, ends_with("%)"))
-
-ggb1 <- ggb1 %>%
-  arrange(LATITUDE) %>%
-  mutate(rank = row_number()) %>%
-  pivot_longer(cols = ends_with("%)"), names_to = "Eigenvector", values_to = "value")
-ggb1$Eigenvector <- factor(ggb1$Eigenvector, 
-                           levels = c("First eigenvector (PVE = 10.2%)", 
-                                      "Second eigenvector (PVE = 4.8%)", 
-                                      "Third eigenvector (PVE = 2.4%)", 
-                                      "Fourth eigenvector (PVE = 2.0%)",
-                                      "Fifth eigenvector (PVE = 1.7%)",
-                                      "Sixth eigenvector (PVE = 1.3%)"
-                                      ))
-ggb1$ECOTYPE_NNET <- factor(ggb1$ECOTYPE_NNET, levels = c("Coastal", "Lowland", "Upland", "Unknown"))
-
-ggb1_plot2 <- ggb1 %>%
-  ggplot(aes(x = rank, y = value)) +
-  geom_bar(aes(color = ECOTYPE_NNET, fill = ECOTYPE_NNET), stat = "identity") +
-  facet_wrap(~ Eigenvector) +
-  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
-  scale_color_viridis(discrete = TRUE, option = "B", end = 0.9, 
-                      direction = -1) +
-  scale_fill_viridis(discrete = TRUE, option = "B", end = 0.9, direction = -1) +
-  xlab("Latitude of Origin (Rank Order)") + ylab("") +
-  geom_vline(xintercept = 133, linetype = 2)
-save_plot("../manuscript/plots/SVD_of_Kinship_matrix_Midwest_first_six_eigenvectors_ecotype.png", base_height = 5, base_asp = 2, plot = ggb1_plot2)
-```
-## k_gulf
-
-```{r}
-b1 <- svd(k_gulf)
-b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 22.5
-b1$d[2] / sum(b1$d) # 9.4
-b1$d[3] / sum(b1$d) # 3.1
-b1$d[4] / sum(b1$d) # 1.8
-b1$d[5] / sum(b1$d) # 1.6
-b1$d[6] / sum(b1$d) # 1.3
-b1$d[7] / sum(b1$d) # 1.1
-b1$d[8] / sum(b1$d) # 1.0
-sum(b1$d[1:8]) / sum(b1$d) # First 8 eigenvectors explain 1% or more variance, and 41.8% of total variance. 
-```
-
-```{r}
-k_use <- k_gulf
-for(g in 1:8){
-barplot(b1$v[,g]/k_use[which.max(abs(k_use[,g])),g])
-}
-order <- enframe(colnames(k_use), name = NULL, value = "PLANT_ID")
-ggb1 <- tibble(`First eigenvector (PVE = 22.5%)` = b1$v[,1]/k_use[which.max(abs(k_use[,1])),1],
-       `Second eigenvector (PVE = 9.4%)` = b1$v[,2]/k_use[which.max(abs(k_use[,2])),2],
-       `Third eigenvector (PVE = 3.1%)` = b1$v[,3]/k_use[which.max(abs(k_use[,3])),3],
-       `Fourth eigenvector (PVE = 1.8%)` = b1$v[,4]/k_use[which.max(abs(k_use[,4])),4],
-       `Fifth eigenvector (PVE = 1.6%)` = b1$v[,5]/k_use[which.max(abs(k_use[,5])),5],
-       `Sixth eigenvector (PVE = 1.3%)` = b1$v[,6]/k_use[which.max(abs(k_use[,6])),6],
-       PLANT_ID = order$PLANT_ID) %>%
-  left_join(metadata, by = "PLANT_ID") %>%
-  dplyr::select(PLANT_ID, SUBPOP, ECOTYPE_NNET, LATITUDE, ends_with("%)"))
-
-ggb1 <- ggb1 %>%
-  arrange(LATITUDE) %>%
-  mutate(rank = row_number()) %>%
-  pivot_longer(cols = ends_with("%)"), names_to = "Eigenvector", values_to = "value")
-ggb1$Eigenvector <- factor(ggb1$Eigenvector, 
-                           levels = c("First eigenvector (PVE = 22.5%)", 
-                                      "Second eigenvector (PVE = 9.4%)", 
-                                      "Third eigenvector (PVE = 3.1%)", 
-                                      "Fourth eigenvector (PVE = 1.8%)",
-                                      "Fifth eigenvector (PVE = 1.6%)",
-                                      "Sixth eigenvector (PVE = 1.3%)"
-                                      ))
-ggb1$ECOTYPE_NNET <- factor(ggb1$ECOTYPE_NNET, levels = c("Coastal", "Lowland", "Upland", "Unknown"))
-
-ggb1_plot2 <- ggb1 %>%
-  ggplot(aes(x = rank, y = value)) +
-  geom_bar(aes(color = ECOTYPE_NNET, fill = ECOTYPE_NNET), stat = "identity") +
-  facet_wrap(~ Eigenvector) +
-  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
-  scale_color_viridis(discrete = TRUE, option = "B", end = 0.9, 
-                      direction = -1) +
-  scale_fill_viridis(discrete = TRUE, option = "B", end = 0.9, direction = -1) +
-  xlab("Latitude of Origin (Rank Order)") + ylab("") +
-  geom_vline(xintercept = 210, linetype = 2)
-save_plot("../manuscript/plots/SVD_of_Kinship_matrix_Gulf_first_six_eigenvectors_ecotype.png", base_height = 5, base_asp = 2, plot = ggb1_plot2)
-```
-
-The eigenvectors capture the predominant patterns in the Uk3 covariance matrix.
-```
-v <- m$fitted_g$Ulist$ED_PCA_1
-b1=svd(v)
-for(g in 1){
-barplot(b1$v[,g]/v[which.max(abs(v[,g])),g])
-}
-
-b1$d # the top eigenvector captures the predominant pattern in the ED_PCA_1 covariance matrix. 
-```
-
-WAIT. I could also plot this for the svd from the whole SNP set. That might make more sense than doing a svd on the kinship matrix...
-
-```{r}
-snp <- snp_attach("~/Github/pvdiv-genome/genetic_subpops/kinship/Pvirgatum_V5_GWAS_134g_Midwest_K_subset_maf0.05.rds")
-
-```
-
-# -------------
-# ASReml models for eight sites including GxE
-```{r}
-
-```
-
-## ASReml h2 greenup analysis
-```{r}
-library(tidyverse)
-library(asreml)
-library(switchgrassGWAS)
-library(cowplot)
-
-source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
-workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
-outputdir <- file.path(workingdir, "analysis", "heritability")
-metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
-sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
-subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
-phe_gr_df <- read_rds(file.path(workingdir, "data",
-                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
-phe_fl_df <- read_rds(file.path(workingdir, "data",
-                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
-k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
-
-```
-
-### loop to make heritability table using ASReml mixed models and vpredict.
-
-## Include just texas sites for greenup and just northern four sites.
-e.g.:
-for(p in 1){#seq_along(phe_dfs)){
-  phe_v <- phe_lists[[p]]
-  phe <- phe_dfs[[p]]
-
-  for(i in 12:13){#seq_along(site_v)){ # may need to adjust below for h2_table
-    for(j in 1:3){#seq_along(subpop_v)){
-      for(k in seq_along(phe_v)){
-
-## Model eight, texas, and north sites with a kinship : site interaction.
-Then, find GxE and G proportions of variance.
-```{r}
-site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
-               NE = "NE", IL = "IL", MI = "MI", SD = "SD", eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), nine_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
-subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))
-phe_fl <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl",
-           "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d",
-           "crain_7d")
-phe_gr <- c("GR50", "DYLN", "cgdd_12c_jan2gr", "cgdd_8c_jan2gr", "tmax_18d", 
-            "tmin_18d", "tave_18d", "cgdd_12c_18d", "tmax_10d", "tmin_10d",
-            "tave_10d", "cgdd_12c_10d", "tmax_5d", "tmin_5d", "tave_5d",
-            "cgdd_12c_5d")
-phe_dfs <- list(phe_gr_df, phe_fl_df)
-phe_lists <- list(phe_gr, phe_fl)
-h2_table <- tibble()
-
-for(p in seq_along(phe_dfs)){
-  phe_v <- phe_lists[[p]]
-  phe <- phe_dfs[[p]]
-
-  for(i in c(10,12:13)){#seq_along(site_v)){ # may need to adjust below for h2_table
-    for(j in 1:3){#seq_along(subpop_v)){
-      for(k in seq_along(phe_v)){
-        phe_single <- phe %>%
-    left_join(sites, by = "SITE") %>%
-    filter(manu_site %in% site_v[[i]] & SUBPOP %in% subpop_v[[j]]) %>%
-    rename(phenotype = eval(phe_v[k])) %>%
-    mutate(PLANT_ID = as_factor(PLANT_ID),
-           SITE = as_factor(SITE),
-           PLOT_GL = as_factor(PLOT_GL)) %>%
-    as.data.frame()
-
-  if(length(site_v[[i]]) > 1){  # need to add site as a random factor if 
-    # there is more than one site included in the dataset.
-    tryCatch({
-     asr_out <- asreml(phenotype ~ 1,
-                       random = ~vm(PLANT_ID, k_full) +
-                         ~idv(SITE):vm(PLANT_ID, k_full),
-                       residual = ~idv(units),
-                       data = phe_single, 
-                       workspace = "3gb")
-
-    h2_est <- vpredict(asr_out, h2 ~ V2/(V1+V2+V3)) %>% as_tibble() %>%
-      mutate(site = names(site_v)[i],
-             subpop = names(subpop_v)[j],
-             phe = phe_v[k],
-             bic = summary(asr_out)$bic[1],
-             loglik = summary(asr_out)$loglik)
-    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
-  } else {
-    tryCatch({
-    asr_out <- asreml(phenotype ~ 1,
-                       random = ~vm(PLANT_ID, k_full),
-                       residual = ~idv(units),
-                       data = phe_single, 
-                       workspace = "3gb")
-
-    h2_est <- vpredict(asr_out, h2 ~ V1/(V1+V2)) %>% as_tibble() %>%
-      mutate(site = names(site_v)[i],
-             subpop = names(subpop_v)[j],
-             phe = phe_v[k],
-             bic = summary(asr_out)$bic[1],
-             loglik = summary(asr_out)$loglik)
-  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
-  }
-
-  if(i == 12 & j == 1 & k == 1){
-    h2_table <- h2_est
-  } else {
-    h2_table <- add_row(h2_table, h2_est)
-  }
-  tryCatch({
-      save_plot(filename=file.path(outputdir, paste0("ASReml_performance_",
-                                                 names(site_v)[i], "_",
-                                                 names(subpop_v)[j],
-                                                 "_", phe_v[k], ".png")), 
-            plot = plot(asr_out))
-    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
-
-      }
-    }
-  }
-  write_csv(h2_table, file.path(outputdir, paste0("h2_df_", phe_v[1],
-                                                  "_related_functions_", 
-                                                  "of_weather_v4.csv")))
-}
-```
-
-# ASReml 07-29
-
-## Model eight, texas, and north sites with a kinship : site interaction.
-Then, find GxE and G proportions of variance.
-## Use Midwest-specific kinship matrix for Midwest-only, and Gulf-specific kinship matrix for Gulf-only comparisons.
-And use (and visualize eigenvectors of?
-Midwest and Gulf kinship matrix for the comparisons with both.)
-```{r}
-library(tidyverse)
-library(asreml)
-library(switchgrassGWAS)
-library(cowplot)
-
-source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
-workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
-outputdir <- file.path(workingdir, "analysis", "heritability", "kinship_subpop")
-metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
-sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
-subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
-phe_gr_df <- read_rds(file.path(workingdir, "data",
-                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
-phe_fl_df <- read_rds(file.path(workingdir, "data",
-                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
-
-kinshipdir <- file.path(workingdir, "data")
-k_gandm <- read_rds(file.path(kinshipdir, "Kinship_van_Raden_method_363g_Gulf_and_Midwest_subset_maf0.05.rds"))
-k_midw <- read_rds(file.path(kinshipdir,
-                "Kinship_van_Raden_method_134g_Midwest_subset_maf0.05.rds"))
-k_gulf <- read_rds(file.path(kinshipdir, "Kinship_van_Raden_method_229g_Gulf_subset_maf0.05.rds"))
-
-```
-
-```{r}
-site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
-               NE = "NE", MI = "MI", SD = "SD", eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
-subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"))
-phe_fl <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl",
-           "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")
-phe_gr <- c("GR50", "DYLN", "cgdd_12c_jan2gr", "tmin_18d", "cgdd_12c_18d",
-            "tmin_10d", "cgdd_12c_10d", "tmin_5d", "cgdd_12c_5d")
-phe_dfs <- list(phe_gr_df, phe_fl_df)
-phe_lists <- list(phe_gr, phe_fl)
-k_list <- list(Gulf = k_gulf, Midwest = k_midw, Gulf_and_Midwest = k_gandm)
-h2_table <- tibble()
-gxe_table <- tibble()
-
-for(p in seq_along(phe_dfs)){
-  phe_v <- phe_lists[[p]]
-  phe <- phe_dfs[[p]]
-
-  for(i in seq_along(site_v)){ # may need to adjust below for h2_table
-    dir.create(file.path(outputdir, names(site_v)[i]), showWarnings = FALSE)
-    for(j in seq_along(subpop_v)){
-      k_use <- k_list[[j]]
-      for(k in seq_along(phe_v)){
-        phe_single <- phe %>%
-          left_join(sites, by = "SITE") %>%
-          filter(manu_site %in% site_v[[i]] & SUBPOP %in% subpop_v[[j]]) %>%
-          rename(phenotype = eval(phe_v[k])) %>%
-          mutate(PLANT_ID = as_factor(PLANT_ID),
-                 SITE = as_factor(SITE),
-                 PLOT_GL = as_factor(PLOT_GL)) %>%
-          as.data.frame()
-
-  if(length(site_v[[i]]) > 1){  # need to add site as a random factor if 
-    # there is more than one site included in the dataset.
-    tryCatch({
-     asr_out <- asreml(phenotype ~ 1,
-                       random = ~idv(SITE)*vm(PLANT_ID, k_use),
-                       residual = ~idv(units),
-                       data = phe_single, 
-                       workspace = "3gb")
-
-    Env_est <- vpredict(asr_out, E ~ V1/(V1+V2+V3+V4))
-    GxE_est <- vpredict(asr_out, GxE_h2 ~ V3/(V1+V2+V3+V4))
-    err_est <- vpredict(asr_out, err ~ V4/(V1+V2+V3+V4))
-    h2_est <- vpredict(asr_out, G_h2 ~ V2/(V1+V2+V3+V4)) %>% as_tibble() %>%
-      mutate(site = names(site_v)[i],
-             subpop = names(subpop_v)[j],
-             phe = phe_v[k],
-             bic = summary(asr_out)$bic[1],
-             loglik = summary(asr_out)$loglik,
-             Env_Estimate = Env_est$Estimate[1],
-             Env_SE = Env_est$SE[1],
-             GxE_Estimate = GxE_est$Estimate[1],
-             GxE_SE = GxE_est$SE[1],
-             error_Estimate = err_est$Estimate[1],
-             error_SE = err_est$SE[1])
-
-    blup <- summary(asr_out, coef=TRUE)$coef.random %>% 
-    as_tibble(rownames = "Effect")
-    write_csv(blup, file.path(outputdir, names(site_v)[i], 
-                              paste0("blups_", names(site_v)[i], "_",
-                                     names(subpop_v)[j], "_", phe_v[k],
-                                     ".csv")))
-    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
-  } else {
-    tryCatch({
-    asr_out <- asreml(phenotype ~ 1,
-                       random = ~vm(PLANT_ID, k_use),
-                       residual = ~idv(units),
-                       data = phe_single, 
-                       workspace = "3gb")
-
-    h2_est <- vpredict(asr_out, h2 ~ V1/(V1+V2)) %>% as_tibble() %>%
-      mutate(site = names(site_v)[i],
-             subpop = names(subpop_v)[j],
-             phe = phe_v[k],
-             bic = summary(asr_out)$bic[1],
-             loglik = summary(asr_out)$loglik)
-    blup <- summary(asr_out, coef=TRUE)$coef.random %>% 
-    as_tibble(rownames = "Effect")
-    write_csv(blup, file.path(outputdir, names(site_v)[i],
-                              paste0("blups_", names(site_v)[i], "_",
-                                     names(subpop_v)[j], "_", phe_v[k],
-                                     ".csv")))
-  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
-  }
-
-  if(i == 1 & j == 1 & k == 1){
-    h2_table <- h2_est
-  } else if(i == 9 & j == 1 & k == 1){
-    gxe_table <- h2_est
-  } else if(length(site_v[[i]]) > 1){
-    gxe_table <- add_row(gxe_table, h2_est)
-  } else {
-    h2_table <- add_row(h2_table, h2_est)
-  }
-  tryCatch({
-      save_plot(filename=file.path(outputdir, names(site_v)[i],
-                                   paste0("ASReml_performance_",
-                                                 names(site_v)[i], "_",
-                                                 names(subpop_v)[j],
-                                                 "_", phe_v[k], ".png")), 
-            plot = plot(asr_out))
-    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
-
-      }
-    }
-  }
-  write_csv(h2_table, file.path(outputdir, paste0("h2_df_", phe_v[1],
-                                                  "_related_functions_", 
-                                                  "of_weather_v1.csv")))
-  write_csv(gxe_table, file.path(outputdir, paste0("gxe_and_h2_df_", phe_v[1],
-                                                  "_related_functions_", 
-                                                  "of_weather_v1.csv")))
-}
-```
-
-# ----
-# D2F
-
-If flowering date frequently varies as a function of GDD in switchgrass, this explains observations that moving southern populations northwards delays flowering, and moving northern populations south hastens flowering (Sanderson et al 1996).
-< Indeed, we observed D2F for Midwest plants was shorter at TX than BRKG, and vice versa?>
-
-```{r}
-d2f_lat <- phenotypes %>%
-  mutate(D2F = FL50 - GR50) %>%
-  left_join(select(metadata, PLANT_ID, SUBPOP)) %>%
-  filter(!(SITE %in% c("OVTN", "FRMI")) & WHERE == "GWAS" & SUBPOP %in% c("Gulf", "Midwest")) %>%
-  group_by(SUBPOP, SITE) %>%
-  summarise(mean_D2F = mean(D2F, na.rm = TRUE)) %>%
-  left_join(select(sites, SITE, Latitude, manu_site)) %>%
-  arrange(SUBPOP, Latitude) 
-d2f_lat %>%
-  ggplot(aes(x = Latitude, y = mean_D2F)) +
-  geom_point(aes(color = SUBPOP))
-
-summary(lm(mean_D2F ~ Latitude*SUBPOP, data = d2f_lat))
-82-55
-```
-
-# ----------------------------
-# G_BLUP GWAS
-
-The plan is to do GWAS on G_BLUP from an eight site model with GxE modeled as site:kinship for each subpopulation and both combined. The traits I will do GWAS on for G_BLUPs are as follows:
-
-eight_sites
-FL50 (to show other assoc are stronger)
-dyln_fl50
-cgdd_12c_gr2fl
-
-Maybe also do GWAS on G_BLUP from the North and Texas models as above, for the traits:
-tx_sites & north_sites
-FL50 (to show other assoc are stronger)
-dyln_fl50
-cgdd_12c_gr2fl
-crain_gr2fl
-crain_1d
-GR50
-cgdd_12c_10d
-
-All I really want to show are that there are stronger G associations for the weather-derived variables than for FL50 and GR50.
-And possibly exploring G associations at subpopulation home and away sites, for the second bit.
-That might be too complicated. Eh.
-
-It's possible I might want to confirm some of these G candidates with the fourway - how we see similar patterns in this cross for FL50 vs dyln/GDD, and there are some similar associations.
-
-THEN, for the next section, do GWAS on the BLUPs from the eight single site models. For these traits:
-all single sites...
-FL50 (actually, maybe not)
-dyln_fl50
-cgdd_12c_gr2fl
-crain_gr2fl
-crain_1d
---
-GR50 (or maybe not)
-cgdd_12c_10d
-
-Then do mash runs.
-First for single traits across all eight sites, to make U_hyp for the next step - take the U_ed with large percentage variance explained.
-Make U_hyp with those covariances but no covariance with anything else...
-Then run all the flowering-related traits together.
-Talk about how then SNPs can have effects that are strongest in response to one type of environmental cue.
-And how cues can be present or absent at different sites.
-And how this kind of analysis can help reveal all the patterns at play, because it doesn't make assumptions about how effects are structured.
-
-Maybe bring back the QTL maps after this and talk about how patterns of effects differ across sites in these models, too. Though the Genstat model is less sensitive...
-
-```{r}
-h2all %>%
-  filter(site %in% c("eight_sites"))
-```
-
-## Make phenotype df for GWAS
-```{r}
-workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
-outputdir <- file.path(workingdir, "analysis", "heritability", "kinship_subpop")
-site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
-               NE = "NE", MI = "MI", SD = "SD", eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
-subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"))
-phe_eight <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl")
-phe_txano <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
-               "crain_1d", "GR50", "cgdd_12c_10d")
-phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
-               "crain_1d", "GR50", "cgdd_12c_10d")
-phe_lists <- list(TX1 = phe_single, TX2 = phe_single, TX3 = phe_single, OK = phe_single, MO = phe_single,
-               NE = phe_single, MI = phe_single, SD = phe_single, eight_sites = phe_eight, tx_sites = phe_txano, north_sites = phe_txano)
-
-i = 9
-j = 1
-k = 1
-phe_gwas <- tibble()
-
-for(j in seq_along(subpop_v)){
-    for(i in seq_along(site_v)){
-      phe_v <- phe_lists[[i]]
-      for(k in seq_along(phe_v)){
-        phe_raw <- read_csv(file.path(outputdir, names(site_v)[i],
-                                      paste0("blups_", names(site_v)[i], "_",
-                                             names(subpop_v)[j], "_", phe_v[k],
-                                             ".csv")))
-        names(phe_raw)[2] <- paste0(phe_v[k], "_", names(site_v)[i])
-        phe_proc <- phe_raw %>%
-          separate(Effect, into = c("factor", "PLANT_ID"), sep = "\\)_") %>%
-          filter(!is.na(PLANT_ID) & !grepl("^SITE", factor)) %>%
-          left_join(metadata, by = "PLANT_ID") %>%
-          filter(SUBPOP %in% subpop_v[[j]]) %>%
-          select(PLANT_ID, 3)
-        if(i == 1 & k == 1){
-          phe_gwas <- phe_proc
-        } else {
-          phe_gwas <- phe_gwas %>%
-            left_join(phe_proc, by = "PLANT_ID")
-        }
-      }
-    }
-  phe_gwas <- phe_gwas %>%
-  select(PLANT_ID, ends_with("eight_sites"), ends_with("tx_sites"), ends_with("north_sites"), everything())
-  write_csv(phe_gwas, path = file.path(workingdir, "analysis", "gwas", "BLUPs",
-                                       paste0("BLUP_phenotypes_",
-                                              "for_gwas_GR_FL_",
-                                              names(subpop_v)[j],
-                                              ".csv")))
-}
-
-ggplot(data = phe_gwas, aes(x = phe_gwas$cgdd_12c_gr2fl_eight_sites)) + geom_histogram()
-
-sites %>% arrange(manu_site)
-```
-
-## Run GWAS 
-```{r}
-library(bigsnpr)
-library(tidyverse)
-library(switchgrassGWAS)
-library(AnnotationDbi)
-txdb <- loadDb(file = file.path("~", "Github", "pvdiv-genome",
-                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
-workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
-inputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs")
-outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs")
-subpop_v <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", Gulf = "Gulf")
-snp_v <- c("12\\.1M", "8\\.8M", "10\\.3M")
-
-inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
-                                   "inputkinship.txt"),  delim = " ", 
-                                   col_names = "SNPfiles")
-set_v <- list(eight_sites = "eight_sites", tx_sites = "tx_sites", north_sites = "north_sites", single_sites = "!sites")
-
-for(j in 4){#seq_along(set_v)){
-  for(i in seq_along(subpop_v)){
-    dir.create(file.path(outputdir, names(set_v)[j]), showWarnings = FALSE)
-    snp <- snp_attach(inputfiles$SNPfiles[i])
-    phe_df <- read_csv(file.path(inputdir, paste0("BLUP_phenotypes_",
-                                                  "for_gwas_GR_FL_",
-                                                  names(subpop_v)[i],
-                                                  ".csv"))) 
-  if(names(set_v)[j] != "single_sites"){
-    phe_df <- phe_df %>%
-    dplyr::select(PLANT_ID, ends_with(names(set_v)[j]))
-  } else {
-    phe_df <- phe_df %>%
-    dplyr::select(PLANT_ID, !ends_with("sites"))
-    if(names(subpop_v)[i] != "Gulf"){
-      phe_df <- phe_df %>%
-    dplyr::select(PLANT_ID, ends_with("NE"), ends_with("MI"), ends_with("SD"))
-    }
-  }
-  pvdiv_standard_gwas(snp, df = phe_df, type = "linear", 
-                      outputdir = file.path(outputdir, names(set_v)[j]),
-                      savegwas = TRUE, saveannos = TRUE, txdb = txdb,
-                      minphe = 130)
-
-  gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir,
-                                                       names(set_v)[j]), 
-                                      pattern = paste0("GWAS_datatable*.*",
-                                                       snp_v[i]))
-  phenotypes <- str_sub(gwas_rds, start = 16, end = -42) %>%
-    str_replace(., "_$", "")
-  out1 <- pvdiv_bigsnp2mashr(path = file.path(outputdir, names(set_v)[j]), 
-                             snp = snp, gwas_rds = gwas_rds, 
-                             phenotypes = phenotypes, numSNPs = 1000, 
-                             clump = FALSE, scaled = FALSE, saveoutput = TRUE,
-                             suffix = paste0("FL_", names(set_v)[j], "_",
-                                             names(subpop_v)[i]))
-  }
-}
-
-```
-## Run mash
-Then do mash runs.
-First for single traits across all eight sites, to make U_hyp for the next step - take the U_ed with large percentage variance explained.
-Make U_hyp with those covariances but no covariance with anything else...
-Then run all the flowering-related traits together.
-Talk about how then SNPs can have effects that are strongest in response to one type of environmental cue.
-And how cues can be present or absent at different sites.
-And how this kind of analysis can help reveal all the patterns at play, because it doesn't make assumptions about how effects are structured.
-```{r}
-workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
-outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
-site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
-               NE = "NE", MI = "MI", SD = "SD")#, eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
-subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"))
-phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
-               "crain_1d", "GR50", "cgdd_12c_10d")
-```
-
-single traits across all eight sites...
-```{r}
-i=1 # site
-j=1 # phe
-k=1 # subpop
-phe_suffix <- paste(phe_single[j], names(subpop_v)[k], 
-                    sep = "_")
-gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir,
-                                                       names(set_v)[j]), 
-                                      pattern = paste0("GWAS_datatable*.*",
-                                                       snp_v[i]))
-phenotypes <- str_sub(gwas_rds, start = 16, end = -42) %>%
-    str_replace(., "_$", "")
-out1 <- pvdiv_bigsnp2mashr(path = file.path(outputdir, names(set_v)[j]), 
-                           snp = snp, gwas_rds = gwas_rds, 
-                           phenotypes = phenotypes, numSNPs = 1000, 
-                           clump = FALSE, scaled = FALSE, saveoutput = TRUE,
-                           suffix = paste0("FL_", names(set_v)[j], "_",
-                                           names(subpop_v)[i]))
-numSNPs <- c(28000, 10000, 4000, 10000)
-
-gulf_loc <- c("CLMB", "KBSM", "KING", "PKLE", "TMPL")
-midw_loc <- c("CLMB", "KBSM", "PKLE")
-atl_loc <- c("CLMB", "KBSM", "LINC", "PKLE", "TMPL")
-midwgulf_loc <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL",
-                  "TMPL")
-U_gulf <- read_rds(file.path(path, "U_hyp_gulf.rds"))
-i = 2
-U_ed <- file.path(path, "Mash_data_driven_covariances_10000_gulf.rds")
-  m_out <- mash_standard_run(path = path, numSNPs = numSNPs[i], U_ed = U_ed,
-                             U_hyp = U_gulf, suffix = phe_suffix[i],
-                             saveoutput = TRUE)
-```
-
-## Analyze GWAS Results
-
-May want to rerun pvdiv_bigsnp2mashr function so that I'm looking at GR & FL top 1000 SNPs separately.
-
-gwasdir <- file.path(workingdir, "analysis", "gwas", "G_BLUP")
-log10p_gulf <- readRDS(file.path(gwasdir,                                 "log10p_strong_df_1000topSNPs_GR_FL_Gulf.rds"))
-log10p_midw <- readRDS(file.path(gwasdir,                                 "log10p_strong_df_1000topSNPs_GR_FL_Midwest.rds"))
-log10p_gandm <- readRDS(file.path(gwasdir,                                 "log10p_strong_df_1000topSNPs_GR_FL_Gulf_and_Midwest.rds"))
-
-```{r}
-j=1
-i=3
-
-names(set_v)[j]
-names(subpop_v)[i]
-bonferroni_v <- c(-log10(0.05/12100000), -log10(0.05/8800000), -log10(0.05/10300000))
-
-gwasdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", names(set_v)[j])
-log10p_gulf <- readRDS(file.path(gwasdir, 
-                                 paste0("log10p_strong_df_1000topSNPs_",
-                                        "FL_", names(set_v)[j], "_Gulf",
-                                        ".rds")))
-log10p_midw <- readRDS(file.path(gwasdir,
-                                 paste0("log10p_strong_df_1000topSNPs_", "FL_",
-                                 names(set_v)[j], "_", "Midwest.rds")))
-log10p_gandm <- readRDS(file.path(gwasdir,
-                                  paste0("log10p_strong_df_1000topSNPs_",
-                                         "FL_", names(set_v)[j], "_",
-                                         "Gulf_and_Midwest.rds")))
-log10p_df <- readRDS(file.path(gwasdir,
-                                  paste0("log10p_strong_df_1000topSNPs_",
-                                         "FL_", names(set_v)[j], "_",
-                                         names(subpop_v)[i], ".rds")))
-```
-
-```{r}
-log10p_gulf %>%
-  pivot_longer(cols = -(SNP:POS), names_to = "PHE", values_to = "log10p") %>%
-  ungroup() %>%
-  filter(log10p > -log10(0.05/10300000)) %>%
-  group_by(PHE) %>%
-  tally() #%>%
-  #write_csv(., paste0("../manuscript/tables/Gulf_", names(set_v)[j], "_SNPs_above_Bonferroni_by_trait.csv"))
-
-log10p_midw %>%
-  pivot_longer(cols = -(SNP:POS), names_to = "PHE", values_to = "log10p") %>%
-  ungroup() %>%
-  filter(log10p > -log10(0.05/8800000)) %>%
-  group_by(PHE) %>%
-  tally() #%>%
-  #write_csv(., paste0("../manuscript/tables/Midwest_", names(set_v)[j], "_SNPs_above_Bonferroni_by_trait.csv"))
-
-log10p_gandm %>%
-  pivot_longer(cols = -(SNP:POS), names_to = "PHE", values_to = "log10p") %>%
-  ungroup() %>%
-  filter(log10p > -log10(0.05/12100000)) %>%
-  group_by(PHE) %>%
-  tally() #%>%
-  #write_csv(., paste0("../manuscript/tables/Gulf_and_Midwest_", names(set_v)[j], "__SNPs_above_Bonferroni_by_trait.csv"))
-  #
-log10p_gandm %>%
-  pivot_longer(cols = -(SNP:POS), names_to = "PHE", values_to = "log10p") %>%
-  ungroup() %>%
-  filter(log10p > -log10(0.05/12100000)) %>%
-  mutate(Pos_Mb = floor(POS/100000)/10) %>%
-  group_by(CHR, Pos_Mb) %>% tally()
-bonferroni <- -log10(0.05/12100000)
-log10p_gandm %>%
-  filter(cgdd_12c_gr2fl_eight_sites_log10p > -log10(0.05/12100000) |
-         FL50_eight_sites_log10p > -log10(0.05/12100000)) %>%
-  group_by(cgdd_12c_gr2fl_eight_sites_log10p > FL50_eight_sites_log10p) %>%
-  tally()
-log10p_gandm %>%
-  filter(dyln_fl50_eight_sites_log10p > -log10(0.05/12100000)  |
-         FL50_eight_sites_log10p > -log10(0.05/12100000)) %>%
-  group_by(dyln_fl50_eight_sites_log10p > FL50_eight_sites_log10p) %>%
-  tally()
-
-log10p_gandm %>%
-  ggplot(aes(x = dyln_fl50_eight_sites_log10p, y = FL50_eight_sites_log10p)) + 
-  geom_point(alpha = 0.5) + 
-  geom_abline(slope = 1, intercept = 0, linetype = 2) +
-  geom_vline(xintercept = bonferroni, color = "red", linetype = 2) + 
-  geom_hline(yintercept = bonferroni, color = "red", linetype = 2)
-```
-
-```{r}
-i=1 # subpop
-j=3 # site set
-
-names(set_v)[j]
-names(subpop_v)[i]
-bonferroni_v <- c(-log10(0.05/12100000), -log10(0.05/8800000), -log10(0.05/10300000))
-
-gwasdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", names(set_v)[j])
-log10p_df <- readRDS(file.path(gwasdir,
-                                  paste0("log10p_strong_df_1000topSNPs_",
-                                         "FL_", names(set_v)[j], "_",
-                                         names(subpop_v)[i], ".rds")))
-top_set <- readRDS(file.path(gwasdir, 
-                                paste0("Part-One-Output_Top-Effects-1000-SNPs_",
-                                        "FL_", names(set_v)[j], "_",
-                                        names(subpop_v)[i], ".rds")))
-
-log10p_df %>%
-  pivot_longer(cols = -(SNP:POS), names_to = "PHE", values_to = "log10p") %>%
-  ungroup() %>%
-  filter(log10p > bonferroni_v[i]) %>%
-  group_by(CHR, POS) %>% tally()
-log10p_df %>%
-  pivot_longer(cols = -(SNP:POS), names_to = "PHE", values_to = "log10p") %>%
-  ungroup() %>%
-  filter(log10p > bonferroni_v[i]) %>%
-  mutate(Pos_Mb = floor(POS/100000)/10) %>%
-  group_by(CHR, Pos_Mb) %>% tally()
-
-top50 <- top_set %>% ungroup() %>%
-  top_n(100, max_score_log10p) %>%
-  mutate(SNP = paste(CHR, POS, sep = "_"))
-dyln_binom <- log10p_df %>%
-  filter(SNP %in% top50$SNP) %>%
-  group_by(dyln_fl50_north_sites_log10p > FL50_north_sites_log10p) %>%
-  tally()
-binom.test(x = dyln_binom$n[2], n = 100, alternative = "greater")
-
-cgdd_binom <- log10p_df %>%
-  filter(SNP %in% top50$SNP) %>%
-  group_by(cgdd_12c_gr2fl_north_sites_log10p > FL50_north_sites_log10p) %>%
-  tally()
-binom.test(x = cgdd_binom$n[2], n = 100, alternative = "greater")
-
-log10p_df %>%
-  ggplot(aes(x = dyln_fl50_north_sites_log10p, y = FL50_north_sites_log10p)) + 
-  geom_point(alpha = 0.5) + 
-  geom_abline(slope = 1, intercept = 0, linetype = 2) +
-  geom_vline(xintercept = bonferroni_v[i], color = "red", linetype = 2) + 
-  geom_hline(yintercept = bonferroni_v[i], color = "red", linetype = 2)
-```
-
-## Plot top GWAS hits
-```{r}
-for(i in seq_along(subpop_v)){
-  for(j in 1:3){
-    gwasdir <- file.path(workingdir, "analysis", "gwas", "BLUPs",
-                         names(set_v)[j])
-    top_set <- readRDS(file.path(gwasdir, 
-                                paste0("Part-One-Output_Top-Effects-1000-SNPs_",
-                                        "FL_", names(set_v)[j], "_",
-                                        names(subpop_v)[i], ".rds")))
-    top_set %>%
-      group_by(n_cond_in_top_set) %>%
-      tally()
-
-    top_manh2 <- top_set %>%
-      filter(max_score_log10p > 5.897) %>%
-      ggplot(aes(x = POS, y = max_score_log10p)) + 
-      geom_point(aes(color = Condition, shape = as_factor(n_cond_in_top_set))) + 
-      #geom_line(aes(group = Condition)) +
-      facet_wrap( ~ CHR, nrow = 1, scales = "free_x", 
-                  strip.position = "bottom") +
-      theme(legend.position = "top",
-            axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-            panel.background = element_rect(fill=NA)) +
-            labs(x = "Chromosome", y = "-log10(p value)") +
-            scale_x_continuous(expand = c(0.18, 0.18)) +
-      scale_color_viridis(discrete = TRUE, end = 0.97) +
-      scale_shape_manual(values = c(4, 16, 17, 18, 14, 3, 4))
-    save_plot(paste0("../manuscript/plots/", "FL_", names(set_v)[j], "_",
-                     names(subpop_v)[i], "_top_500_Manhattan.png"), 
-              base_height = 5, base_asp = 2.5, plot = top_manh2)
-
-  }
-}
-```
-
-## Overlap with QTL and between subpops
-
-All I really want to show are that there are stronger G associations for the weather-derived variables than for FL50 and GR50.
-And possibly exploring G associations at subpopulation home and away sites, for the second bit.
-That might be too complicated. Eh.
-
-It's possible I might want to confirm some of these G candidates with the fourway - how we see similar patterns in this cross for FL50 vs dyln/GDD, and there are some similar associations.
-
-Comparing the top 100 hits within and between subpopulations, X of Y were found in the same 100kb regions.
-```{r}
-i=3 # subpop
-j=1 # site set
-
-names(set_v)[j]
-names(subpop_v)[i]
-bonferroni_v <- c(-log10(0.05/12100000), -log10(0.05/8800000), -log10(0.05/10300000))
-
-top_pos <- list()
-
-gwasdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", names(set_v)[j])
-log10p_df <- readRDS(file.path(gwasdir,
-                                  paste0("log10p_strong_df_1000topSNPs_",
-                                         "FL_", names(set_v)[j], "_",
-                                         names(subpop_v)[i], ".rds")))
-for(i in 1:3){
-top_set <- readRDS(file.path(gwasdir, 
-                                paste0("Part-One-Output_Top-Effects-1000-SNPs_",
-                                        "FL_", names(set_v)[j], "_",
-                                        names(subpop_v)[i], ".rds")))
-
-top_pos[[i]] <- top_set %>%
-  ungroup() %>%
-  top_n(n = 100, wt = max_score_log10p) %>%
-  mutate(Pos_Mb = floor(POS/100000)/10) %>%
-  group_by(CHR, Pos_Mb) %>%
-  tally()
-}
-
-gandm_m <- top_pos[[1]] %>%
-  inner_join(top_pos[[2]])
-gandm_g <- top_pos[[1]] %>%
-  inner_join(top_pos[[3]])
-m_g <- top_pos[[2]] %>%
-  inner_join(top_pos[[3]])
-c(nrow(gandm_m)/nrow(top_pos[[1]]), nrow(gandm_m)/nrow(top_pos[[2]]),
-  nrow(gandm_g)/nrow(top_pos[[1]]), nrow(gandm_g)/nrow(top_pos[[3]]),
-  nrow(m_g)/nrow(top_pos[[2]]), nrow(m_g)/nrow(top_pos[[3]])
-  )
-
-top50 <- top_set %>% ungroup() %>%
-  top_n(100, max_score_log10p) %>%
-  mutate(SNP = paste(CHR, POS, sep = "_"))
-dyln_binom <- log10p_df %>%
-  filter(SNP %in% top50$SNP) 
-```
-
-```{r}
-i=3 # subpop
-j=1 # site set
-
-names(set_v)[j]
-names(subpop_v)[i]
-bonferroni_v <- c(-log10(0.05/12100000), -log10(0.05/8800000), -log10(0.05/10300000))
-
-top_pos <- list()
-
-log10p_df <- readRDS(file.path(gwasdir,
-                                  paste0("log10p_strong_df_1000topSNPs_",
-                                         "FL_", names(set_v)[j], "_",
-                                         names(subpop_v)[i], ".rds")))
-for(j in 1:3){
-  gwasdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", names(set_v)[j])
-  top_set <- readRDS(file.path(gwasdir, 
-                                paste0("Part-One-Output_Top-Effects-1000-SNPs_",
-                                        "FL_", names(set_v)[j], "_",
-                                        names(subpop_v)[i], ".rds")))
-
-top_pos[[j]] <- top_set %>%
-  ungroup() %>%
-  top_n(n = 500, wt = max_score_log10p) %>%
-  mutate(Pos_Mb = floor(POS/100000)/10) %>%
-  group_by(CHR, Pos_Mb) %>%
-  tally()
-}
-
-gandm_m <- top_pos[[1]] %>%
-  inner_join(top_pos[[2]])
-gandm_g <- top_pos[[1]] %>%
-  inner_join(top_pos[[3]])
-m_g <- top_pos[[2]] %>%
-  inner_join(top_pos[[3]])
-c(nrow(gandm_m)/nrow(top_pos[[1]]), nrow(gandm_m)/nrow(top_pos[[2]]),
-  nrow(gandm_g)/nrow(top_pos[[1]]), nrow(gandm_g)/nrow(top_pos[[3]]),
-  nrow(m_g)/nrow(top_pos[[2]]), nrow(m_g)/nrow(top_pos[[3]])
-  )
-
-top50 <- top_set %>% ungroup() %>%
-  top_n(100, max_score_log10p) %>%
-  mutate(SNP = paste(CHR, POS, sep = "_"))
-dyln_binom <- log10p_df %>%
-  filter(SNP %in% top50$SNP) 
-```
-for gulf and midwest:
-4.3-6.5% overlap in top 100 between eight and tx; no overlap in top 100 between eight and north or north and texas.
-500: 8% eight and texas; 1.6% eight and north. 0.9% north and texas.
-
-for midwest:
-5% overlap in top 500 eight and texas; 18% overlap eight and north; 1.7% north and texas.
-
-for gulf:
-6.2% overlap eight and texas in top 500; 5.2% overlap eight and north; 2.7% overlap north and texas.
-
-I think the top 500 SNPs are good here - give you large enough numbers.
-
-## View Annotations
-```{r}
-workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
-snp_v <- c("12\\.1M", "8\\.8M", "10\\.3M")
-subpop_v <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", Gulf = "Gulf")
-set_v <- list(eight_sites = "eight_sites", tx_sites = "tx_sites", north_sites = "north_sites", single_sites = "!sites")
-
-for(j in seq_along(set_v)){
-  anno_list <- list()
-  annodir <- file.path(workingdir, "analysis", "gwas", "BLUPs",
-                       names(set_v)[j])
-  anno_v <- pvdiv_results_in_folder(path = annodir, pattern = "Annotation")
-
-  for(i in seq_along(anno_v)){
-  annos <- readRDS(file.path(annodir, anno_v[i]))
-  if(names(set_v)[j] != "single_sites"){
-  anno_info <- anno_v[i] %>% enframe(name = NULL, value = "File") %>%
-    mutate(phe = case_when(grepl("dyln_fl50", File) ~ "dyln_fl50",
-                           grepl("cgdd_12c_gr2fl", File) ~ "cgdd_12c_gr2fl",
-                           grepl("FL50", File) ~ "FL50",
-                           grepl("GR50", File) ~ "GR50",
-                           grepl("crain_gr2fl", File) ~ "crain_gr2fl",
-                           grepl("crain_1d", File) ~ "crain_1d",
-                           grepl("cgdd_12c_10d", File) ~ "cgdd_12c_10d"),
-           sites = case_when(grepl("eight_sites", File) ~ "eight_sites",
-                             grepl("north_sites", File) ~ "north_sites",
-                             grepl("tx_sites", File) ~ "tx_sites"),
-           subpop = case_when(grepl("8\\.8M", File) ~ "Midwest",
-                              grepl("10\\.3M", File) ~ "Gulf",
-                              grepl("12\\.1M", File) ~ "Gulf_and_Midwest"))
-  } else {
-    anno_info <- anno_v[i] %>% enframe(name = NULL, value = "File") %>%
-    mutate(phe = case_when(grepl("dyln_fl50", File) ~ "dyln_fl50",
-                           grepl("cgdd_12c_gr2fl", File) ~ "cgdd_12c_gr2fl",
-                           grepl("FL50", File) ~ "FL50",
-                           grepl("GR50", File) ~ "GR50",
-                           grepl("crain_gr2fl", File) ~ "crain_gr2fl",
-                           grepl("crain_1d", File) ~ "crain_1d",
-                           grepl("cgdd_12c_10d", File) ~ "cgdd_12c_10d"),
-           sites = case_when(grepl("TX1", File) ~ "TX1",
-                             grepl("TX2", File) ~ "TX2",
-                             grepl("TX3", File) ~ "TX3",
-                             grepl("OK", File) ~ "OK",
-                             grepl("MO", File) ~ "MO",
-                             grepl("MI", File) ~ "MI",
-                             grepl("NE", File) ~ "NE",
-                             grepl("SD", File) ~ "SD"),
-           subpop = case_when(grepl("8\\.8M", File) ~ "Midwest",
-                              grepl("10\\.3M", File) ~ "Gulf",
-                              grepl("12\\.1M", File) ~ "Gulf_and_Midwest"))
-  }
-  anno_list[[i]] <- annos$top500SNPs_within50000bp %>%
-    arrange(QUERYID, `distance from gene`) %>%
-    group_by(QUERYID) %>%
-    slice(1) %>%
-    ungroup() %>%
-    mutate(POS_Mb = floor(POS/10000)/100,
-           phe = anno_info$phe,
-           sites = anno_info$sites,
-           subpop = anno_info$subpop) %>%
-    select(sites, subpop, phe, CHR, POS_Mb, `Gene ID`, `distance from gene`, `FDR Adjusted p value`, `Arabidopsis thaliana homolog`:`A.
-thaliana gene annotation`, `Rice gene annotation`, `Rice homolog`)
-  }
-
-  anno_df <- anno_list[[1]]
-  for(i in seq_along(anno_list)[-1]){
-    anno_df <- anno_df %>%
-      full_join(anno_list[[i]])
-  }
-  anno_df %>%
-    arrange(CHR, POS_Mb) %>%
-    filter(`FDR Adjusted p value` <= 0.1) %>%
-write_csv(., path = file.path(workingdir, "analysis", "gwas", "BLUPs",
-                                    paste0(names(set_v)[j],     
-                                           "_top_500_SNPs_above_FDR",
-                                           "_closest_annotations.csv")))
-}
-```
-Taylor et al 2019, Crop Science: switchgrass GWAS
-Grabowski et al 2017, New Phytologist - GWAS on ~509 individuals on heading date in New York
-Switchgrass GWAS at two locations: Frontiers in Plant Science
-Pavir.8NG050900: Pseudo Response Regulator 5 (PRR5) - important in flowering date
-Flowering locus T (on Chromosome 5N). FKF1, PRR5 COP1 GASA8 P450 - differentially expressed flowering genes between early and late flowering individuals.
-
-Pavir.2NG448600: PRR5
-Pavir.5KG044000 (FT)
-Pavir.2NG126000 (PRR7)
-Pavir.
-
-```{r}
-workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
-set_v <- list(eight_sites = "eight_sites", tx_sites = "tx_sites", north_sites = "north_sites", single_sites = "!sites")
-j = 1
-annos_df <- read_csv(file = file.path(workingdir, "analysis", "gwas", "BLUPs",
-                                    paste0(names(set_v)[j],     
-                                           "_top_500_SNPs_above_FDR",
-                                           "_closest_annotations.csv")))
-for(j in seq_along(set_v)[-1]){
-annos1 <- read_csv(file = file.path(workingdir, "analysis", "gwas", "BLUPs",
-                                    paste0(names(set_v)[j],     
-                                           "_top_500_SNPs_above_FDR",
-                                           "_closest_annotations.csv")))
-annos_df <- annos_df %>%
-  full_join(annos1)
-}
-
-annos_plot <- annos_df %>%
-  filter(sites %in% c("eight_sites", "tx_sites", "north_sites") & !(phe %in% c("GR50", "cgdd_12c_10d", "FL50", "crain_1d"))) %>%
-  group_by(CHR, POS_Mb, sites, subpop, phe) %>% add_tally(name = "n_SNPs") %>%
-  arrange(CHR, POS_Mb) %>%
-  #filter(n_SNPs > 1) %>%
-  slice(1) %>%
-  group_by(CHR, POS_Mb) %>% add_tally(name = "n_Mb") %>%
-  filter(n_Mb > 5) %>% arrange(CHR, POS_Mb) %>%
-  select(n_SNPs, n_Mb, everything())
-annos_plot1 <- annos_plot %>% group_by(CHR, POS_Mb) %>% tally() %>%
-  mutate(POS_lo = POS_Mb*1000000,
-         POS_hi = POS_Mb*1000000+10000)
-
-annos_plot <- annos_df %>% 
-  mutate(POS_bin_lo = floor(POS_Mb*100/2),
-         POS_bin_hi = ceiling(POS_Mb*100/2)) %>%
-  filter(sites %in% c("eight_sites", "tx_sites", "north_sites") & !(phe %in% c("GR50", "cgdd_12c_10d", "FL50", "crain_1d", "crain_gr2fl"))) %>%
-  group_by(CHR, POS_bin_hi, sites, subpop, phe) %>% add_tally(name = "n_SNPs") %>%
-  arrange(CHR, POS_bin_hi) %>%
-  #filter(n_SNPs > 1) %>%
-  slice(1) %>%
-  group_by(CHR, POS_bin_hi) %>% add_tally(name = "n_Mb") %>%
-  filter(n_Mb > 4) %>% arrange(CHR, POS_bin_hi) %>%
-  select(n_SNPs, n_Mb, everything())
-annos_plot1 <- annos_plot %>% group_by(CHR, POS_bin_hi, POS_Mb) %>% tally() %>%
-  mutate(POS_lo = POS_Mb*1000000,
-         POS_hi = POS_Mb*1000000+10000)
-```
-
-dyln, fl50, cgdd eight sites gulf & midwest, north sites gulf & midwest cgdd & dyln, tx crain gr2fl, dyln, cgdd
-```{r}
-qtldf %>% arrange(CHR, Pos_Mb_lo) %>%
-  #filter(lodcolumn != "dyln_change_sec") %>%
-  ggplot(aes(x = Pos_Mb_lo, y = lod)) +
-  geom_jitter(aes(color = lodcolumn), height = 2) +
-  geom_errorbarh(aes(y = lod, xmin = Pos_Mb_lo, xmax = Pos_Mb_hi, color = lodcolumn), size = 1.5) +
-  facet_wrap(~CHR) + scale_color_manual(values = c("magenta", "yellow", "darkgreen", "blue")) + theme(legend.position = c(0.9, 0.15), panel.background = element_rect(fill = "grey90")) + xlim(c(0, 70)) + geom_vline(xintercept = 0)
-```
-
-3 regions with the most 'independent' hits: Chr03K 50.84, Chr03N 31.73, Chr05N 64.86
-3N isn't potentially confirmed by any QTL. 3K doesn't overlap. 
-5N does!All four traits for QTL (dyln much higher lod).
-Then: Chr03K 38.23, Chr04K 8.07, Chr05N 2.79, Chr09N 69.33
-3K ~ 17 - 26. Not really.  4K 2.28-5.5 & 9.92 - 13.93, so definitely close/overlaps. 5N overlaps. 9N does not.
-
-So 5N (late), 5N (early), and 4K have overlaps thus far. Very strong result.
-
-Then: 
-Chr01K 56.75 (no), Chr02K 4.25 (yes), 3K 38.23, 38.27, 38.35, 40.87 (no), Chr04N 42.92 (no), Chr05N 49.05 (very close ~ 44.5 - 46.6), Chr07K 33.33 (very close 34.82), Chr08N 28.5 (yes), Chr08N 43.77 (no), Chr09N 21.45 26.06 30.04 (no but 21 is close: 6.88 - 17.31)
-
-So for 7 regions sig above FDR in >=8 subsets, 3 overlap, and overlap the 3 of the 4 biggest QTL.
-For the next 14 regions sig above FDR in 7 or more subsets, 2 overlap and 3 are very close. 
-So of the 12-13 QTL, we find overlap for 3 of the 4 biggest, and 2-5 of the remainder. We don't find big hits in GWAS on 2N.
-
-Biggest QTL: 2N, 4K, 5N (x2-3) (4-5)
-Other QTL: 2K (x2), 3K, 7K, 7N, 8N, 9K, 9N (8)
-
-nothing on 1K, 1N, 3N, 4N, 5K, 6K, 6N, 8K. Nothing!!
-
-I think this is good. And can write up this section focusing on this.
+## QTL
 ```{r}
@@
   arrange(CHR, Pos_Mb_lo)
+
+qtlplot <- qtldf %>%
+  mutate(phe = case_when(lodcolumn == "FL50" ~ "50% flowering",
+                              lodcolumn == "dyln_fl50" ~ "Daylength",
+                              lodcolumn == "dyln_change_sec" ~ "daylength change (s)",
+                              lodcolumn == "CGDD_12C" ~ "GDD, GR to FL",
+                              lodcolumn == "crain_gr2fl" ~ "Rainfall, GR to FL")) %>%
+  filter(phe %in% c("Daylength", "GDD, GR to FL")) %>%
+  mutate(POS_lo = Pos_Mb_lo * 1000000,
+         POS_hi = Pos_Mb_hi * 1000000)
+
+qtlplot$phe <- factor(qtlplot$phe, 
+                      levels = c("50% flowering", "Rainfall, GR to FL",
+                                 "GDD, GR to FL", "Daylength",
+                                          "daylength change (s)",
+                                          "Rainfall, 1 day",
+                                                  #"Rainfall, 2d sum",
+                                                  "Rainfall, 3d sum",
+                                                  "Rainfall, 5d sum"#,
+                                                  #"Rainfall, 7d sum"
+                                                  ))
 ```
 
-## Combined Manhattan
-
-facet grid - rows are three sets of sites, colors are phenotypes, shapes are subpopulations. Then I can plot -log10 p values on the y axis and Chr/Pos on the x axis. 
+## G BLUP GWAS P Values
 ```{r}
-workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
 subpop_v <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", Gulf = "Gulf")
@@
 
+logpdf <- tibble()
 for(i in seq_along(subpop_v)){
   for(j in 1:3){
-    if(i == 1 & j ==1){
-      logpdf <- readRDS(file.path(workingdir, "analysis", "gwas", "BLUPs",
-                                  names(set_v)[j],
-                                  paste0("log10p_strong_df_1000topSNPs_FL_",
-                                         names(set_v)[j], "_",
-                                         names(subpop_v)[i], ".rds"))) %>%
-        mutate(subpop = names(subpop_v)[i]) %>% ungroup()
-      } else {
         logpdf1 <- readRDS(file.path(workingdir, "analysis", "gwas", "BLUPs",
@@
         mutate(subpop = names(subpop_v)[i]) %>% ungroup()
-        logpdf <- logpdf %>%
-          full_join(logpdf1)
-      }
+        logpdf <- bind_rows(logpdf, logpdf1)
       }
@@
   filter(!is.na(log10p))
-
-```
-
-```{r}
-
 log10p_long$phe <- factor(log10p_long$phe, levels = c("crain_gr2fl", "cgdd_12c_gr2fl", "dyln_fl50", "crain_1d", "FL50"))
@@
 
-log10p_plot %>%
-  ggplot(aes(x = POS, y = log10p)) +
-  geom_vline(data = annos_plot1, aes(xintercept = POS), color = "grey70", 
-             size = 1.1) +
-  geom_point(aes(color = phe_name, shape = subpop_name)) +
-  facet_grid(site ~ CHR, scales = "free_x", space = "free_x", switch = "x") +
-      theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
-              legend.position = "top",
-            panel.background = element_rect(fill = "grey85")) +
-            labs(x = "Chromosome", y = "-log10(p value)") +
-            scale_x_continuous(breaks = c(20000000, 40000000, 60000000, 80000000, 100000000), labels = c("20 Mb", "40 Mb", "60 Mb", "80 Mb", ""), expand = c(0.4, 0.4)) +
-      scale_color_viridis(discrete = TRUE, end = 0.93,
-                          option = "B") +
-      scale_shape_manual(values = c(6, 4, 1, 4, 16, 4, 17, 18, 14, 3, 4)) +
-  ylim(c(6, 15)) + geom_hline(yintercept = c(8.3), linetype = 2)
-save_plot(paste0("../manuscript/plots/", "G_BLUP_Combined_Manhattans_fl_phe_subpops_sites.png"), 
-              base_height = 4.5, base_asp = 3.5, plot = last_plot())
-
 log10p_lim <- log10p_plot %>%
@@
 
-k_subgenome <- log10p_plot %>%
-  filter(grepl("K", CHR)) %>%
-  ggplot(aes(x = POS, y = log10p)) +
-  geom_vline(data = filter(annos_plot1, grepl("K", CHR)),
-             aes(xintercept = POS), color = "grey70", size = 1.1) +
-  geom_point(aes(color = phe_name, shape = subpop_name), size = 0.7, stroke = 0.8) +
-  geom_blank(data = filter(log10p_lim, grepl("K", CHR))) +
-  facet_grid(site ~ CHR, scales = "free_x", space = "free_x", switch = "x") +
-      theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), 
-              legend.position = "top",
-            panel.background = element_rect(fill = "grey85"), panel.spacing.x = unit(0, 'cm')) +
-            labs(x = "Chromosome", y = "-log10(p value)") +
-            scale_x_continuous(breaks = c(20000000, 40000000, 60000000, 80000000), labels = c("20 Mb", "40 Mb", "60 Mb", "80 Mb"), expand = c(0.1, 0.1)) +
-      scale_color_viridis(discrete = TRUE, end = 0.88,
-                          option = "B") +
-      scale_shape_manual(values = c(1, 4, 6, 4, 16, 4, 17, 18, 14, 3, 4)) +
-  ylim(c(6, 15)) + geom_hline(yintercept = c(8.3), linetype = 2) + labs(subtitle = "K subgenome", x = "")
-
-n_subgenome <- log10p_plot %>%
-  filter(grepl("N", CHR)) %>%
-  ggplot(aes(x = POS, y = log10p)) +
-  geom_vline(data = filter(annos_plot1, grepl("N", CHR)),
-             aes(xintercept = POS), color = "grey70", size = 1.1) +
-  geom_point(aes(color = phe_name, shape = subpop_name), size = 0.7, stroke = 0.8) +
-  geom_blank(data = filter(log10p_lim, grepl("N", CHR))) +
-  facet_grid(site ~ CHR, scales = "free_x", space = "free_x", switch = "x") +
-      theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), 
-              legend.position = "none",
-            panel.background = element_rect(fill = "grey85"), panel.spacing.x = unit(0, 'cm')) +
-            scale_x_continuous(breaks = c(20000000, 40000000, 60000000, 80000000), labels = c("20 Mb", "40 Mb", "60 Mb", "80 Mb"), expand = c(0.1, 0.1)) +
-      scale_color_viridis(discrete = TRUE, end = 0.88,
-                          option = "B") +
-      scale_shape_manual(values = c(1, 4, 6, 4, 16, 4, 17, 18, 14, 3, 4)) +
-  ylim(c(6, 15)) + geom_hline(yintercept = c(8.3), linetype = 2) + labs(subtitle = "N subgenome", x = "", y = "-log10(p value)")
-
-both_subgenomes <- plot_grid(k_subgenome, n_subgenome, nrow = 2, rel_heights = c(0.52, 0.47))
-save_plot(paste0("../manuscript/plots/", "G_BLUP_Combined_Manhattans_fl_phe_subpops_sites_by_subgenome.svg"), 
-              base_height = 7.5, base_width = 6.5, plot = both_subgenomes)
 ```
 
-## Manhattan QTL windows
+## Top annotations from G_BLUP GWAS
 
 ```{r}
-qtldf <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv")
+set_v <- list(eight_sites = "eight_sites", tx_sites = "tx_sites", north_sites = "north_sites", single_sites = "!sites")
+j = 1
+annos_df <- tibble()
+for(j in seq_along(set_v)){
+annos1 <- read_csv(file = file.path(workingdir, "analysis", "gwas", "BLUPs",
+                                    paste0(names(set_v)[j],     
+                                           "_top_500_SNPs_above_FDR",
+                                           "_closest_annotations.csv")))
+annos_df <- bind_rows(annos_df, annos1)
+}
 
-qtldf <- qtldf %>%
-  separate(flank_lo, into = c("CHR", "Pos_Mb_lo"), sep = "_", convert = TRUE) %>%
-  separate(flank_hi, into = c("flank", "Pos_Mb_hi"), sep = "_", convert = TRUE) %>%
-  mutate(Pos_Mb_lo = round(Pos_Mb_lo, digits = 2),
-         Pos_Mb_hi = round(Pos_Mb_hi, digits = 2)) %>%
-  select(CHR, Pos_Mb_lo, Pos_Mb_hi, lodcolumn, lod, everything()) %>%
-  arrange(CHR, Pos_Mb_lo)
+annos_plot <- annos_df %>%
+  filter(sites %in% c("eight_sites", "tx_sites", "north_sites") & !(phe %in% c("GR50", "cgdd_12c_10d", "FL50", "crain_1d"))) %>%
+  group_by(CHR, POS_Mb, sites, subpop, phe) %>% add_tally(name = "n_SNPs") %>%
+  arrange(CHR, POS_Mb) %>%
+  #filter(n_SNPs > 1) %>%
+  slice(1) %>%
+  group_by(CHR, POS_Mb) %>% add_tally(name = "n_Mb") %>%
+  filter(n_Mb > 5) %>% arrange(CHR, POS_Mb) %>%
+  select(n_SNPs, n_Mb, everything())
+annos_plot1 <- annos_plot %>% group_by(CHR, POS_Mb) %>% tally() %>%
+  mutate(POS_lo = POS_Mb*1000000,
+         POS_hi = POS_Mb*1000000+10000)
 
-qtlplot <- qtldf %>%
-  mutate(phe = case_when(lodcolumn == "FL50" ~ "50% flowering",
-                              lodcolumn == "dyln_fl50" ~ "Daylength",
-                              lodcolumn == "dyln_change_sec" ~ "daylength change (s)",
-                              lodcolumn == "CGDD_12C" ~ "GDD, GR to FL",
-                              lodcolumn == "crain_gr2fl" ~ "Rainfall, GR to FL")) %>%
-  filter(phe %in% c("Daylength", "GDD, GR to FL")) %>%
-  mutate(POS_lo = Pos_Mb_lo * 1000000,
-         POS_hi = Pos_Mb_hi * 1000000)
+annos_plot <- annos_df %>% 
+  mutate(POS_bin_lo = floor(POS_Mb*100/2),
+         POS_bin_hi = ceiling(POS_Mb*100/2)) %>%
+  filter(sites %in% c("eight_sites", "tx_sites", "north_sites") & !(phe %in% c("GR50", "cgdd_12c_10d", "FL50", "crain_1d", "crain_gr2fl"))) %>%
+  group_by(CHR, POS_bin_hi, sites, subpop, phe) %>% add_tally(name = "n_SNPs") %>%
+  arrange(CHR, POS_bin_hi) %>%
+  #filter(n_SNPs > 1) %>%
+  slice(1) %>%
+  group_by(CHR, POS_bin_hi) %>% add_tally(name = "n_Mb") %>%
+  filter(n_Mb > 4) %>% arrange(CHR, POS_bin_hi) %>%
+  select(n_SNPs, n_Mb, everything())
+annos_plot %>% group_by(CHR, POS_bin_hi, POS_Mb) %>% tally() %>%
+  mutate(POS_lo = POS_Mb*1000000,
+         POS_hi = POS_Mb*1000000+10000)
+```
 
-qtlplot$phe <- factor(qtlplot$phe, 
-                               levels = c("50% flowering", "Rainfall, GR to FL",
-                                          "GDD, GR to FL", "Daylength",
-                                          "daylength change (s)",
-                                          "Rainfall, 1 day",
-                                                  #"Rainfall, 2d sum",
-                                                  "Rainfall, 3d sum",
-                                                  "Rainfall, 5d sum"#,
-                                                  #"Rainfall, 7d sum"
-                                                  ))
-```
+# G_BLUP Manhattan with overlapping top hits
 
@@
 
-## Manhattan QTL windows
+# G_BLUP 20kb windows Upset plot
 
-### GDD & DYLN only
+anno_df: All are above a 10% FDR.
+If there are more than 500 SNPs above the FDR, only returns the top 500 by p value.
+Only gets the closest gene to each SNP.
+Should be three site sets and three subpop sets, several (four?) flowering phenotypes.
+
+ftanno: OsGene and AtGene columns of this df indicate if the gene is a FT homolog in rice (8544 rows) or A. thaliana  (865 rows). Key is the locusName column.
+
+log10p_plot: Top 1000 SNPs for each GWAS.
+
+qtlplot: Significant QTL for flowering in fourway cross.
+```{r}
+annos_df
+ft_annos <- ftanno %>%
+  filter(!is.na(OsGene) | !is.na(AtGene)) %>%
+  mutate(Pos_Mb = round(start/10000)/100,
+         #Pos_Mb_e = floor(start/10000)/100,
+         POS_bin = round(Pos_Mb*100/2))# ,
+         #POS_bin_hi = round(Pos_Mb_e*100/2))
+
+  # checked that gene start & end never fall into different 20kb bins.   filter(POS_bin_lo != POS_bin_hi)  filter(Pos_Mb_s != Pos_Mb_e)
+log10p_plot
+qtlplot
+```
+
+Are anno of the annos in annos_df in ftanno?
+```{r}
+annos_df %>%
+  inner_join(ft_annos, by = c("Gene ID" = "locusName")) %>%
+  group_by(`Gene ID`) %>% tally()
+annos_df %>%
+  group_by(`Gene ID`) %>% tally() %>%
+  inner_join(ft_annos, by = c("Gene ID" = "locusName"))
+```
+2682 of 19963...
+
+Are any of the 20kb regions in log10p_plot in the 20kb regions in ftanno?
+```{r}
+log10p_plot %>%
+  mutate(Pos_Mb = round(POS/10000)/100) %>%
+  inner_join(ft_annos, by = c("CHR" = "seqid", "Pos_Mb"))
+```
+11324 of 110822 rows...
+
+Do the QTL overlap any of the genes in ftanno?
+```{r}
+ft_annos %>%
+  filter(seqid %in% qtldf$CHR[1] & between(Pos_Mb, qtldf$Pos_Mb_lo[1], qtldf$Pos_Mb_hi[1]))
+
+qtldf
+i=27
+ft_annos %>%
+  filter(seqid %in% qtldf$CHR[i] & between(Pos_Mb, qtldf$Pos_Mb_lo[i], qtldf$Pos_Mb_hi[i])) %>%
+  select(OsGene, AtGene, everything())
+```
+
+Yes. Lots. Very encouraging.

### 2020-09-30 R/Plots_manuscript_v0.3.Rmd
Start comparisons of GWAS and QTL intervals with FT annotations

@@
 		LOC_Os01g64590.1
-
+## ---- #### ---- ####
 ```{r}

### 2020-10-01 R/Plots_manuscript_G_BLUP.Rmd
Add functionally tested OsGene keywords to FT analysis

@@
 library(tidyverse)
+library(ComplexUpset)
+library(cowplot)
+
 ```
@@
 
-Xiaoyu's annotation file, QTL 
+Xiaoyu's annotation file, and FT keywords for rice from functionally annotated genes
 ```{r}
@@
 ftanno <- read_csv(file.path(workingdir, "data/Pvirgatum_v5.1.annotation_FTGeneHomologs.csv"))
+ftkey <- read_delim(file.path(workingdir, "data", "OsGene_keyword_table.txt"), delim = "\t")
+ftanno %>%
+  filter(!is.na(OsGene)) %>%
+  left_join(ftkey, by = c("OsGene" = "Symbol")) %>%
+  filter(Keyword %in% c("flowering time", "heading date")) %>% unique()
+
+osgene_keyword_df <- ftkey %>%
+  select(-RAPdb, -MSU) %>%
+  group_by(Symbol, Keyword) %>%
+  slice(1) %>%
+  pivot_wider(names_from = Keyword, values_from = Title)
+
+osgene_ft_key <- osgene_keyword_df %>%
+  select(Symbol, `flowering time`, `heading date`, anther, flower, `floral meristem`, `floral meristem determinacy`, flowering, `flower development`, `carpel differentiation`, stamen, `stamen number`) %>%
+  pivot_longer(cols = -Symbol, names_to = "Keyword", values_to = "Title") %>%
+  filter(!is.na(Title)) %>%
+  pivot_wider(names_from = Keyword, values_from = Title)
+
+ft_annos <- ftanno %>%
+  filter(OsGene %in% osgene_ft_key$Symbol | !is.na(AtGene)) %>%
+  mutate(Pos_Mb = round(start/10000)/100,
+         #Pos_Mb_e = floor(start/10000)/100,
+         POS_bin = round(Pos_Mb*100/2)) %>%# ,
+         #POS_bin_hi = round(Pos_Mb_e*100/2))
+  select(OsGene, AtGene, everything())   
+1599 / 83260
 ```
+~1600 virgatum genes have rice or thaliana homologs with roles in flowering.
+1599 / 83260 total genes. 1.92% of genes.
 
@@
 
+log10p_long %>% arrange(log10p) %>%
+  group_by(sites, subpop, phe) %>%
+  slice(1) %>%
+  arrange(sites, subpop, phe) %>%
+  mutate(pval = 10^-log10p)
 ```
@@
          POS_hi = POS_Mb*1000000+10000)
+
+annos_df %>% 
+  arrange(desc(`FDR Adjusted p value`)) %>%
+  group_by(sites, subpop, phe) %>%
+  slice(1) %>%
+  arrange(sites, subpop, phe) %>%
+  filter(sites %in% c("eight_sites", "tx_sites", "north_sites") & !(phe %in% c("GR50", "cgdd_12c_10d", "FL50", "crain_1d", "crain_gr2fl"))) 
+  mutate(pval = 10^-log10p)
 ```
 
-# G_BLUP Manhattan with overlapping top hits
+# Manhattan with overlapping top hits (G_BLUP)
 
@@
                                expand = c(0.1, 0.1)) +
-      scale_color_viridis(discrete = TRUE, begin = 0.35, end = 0.75, option = "B") +
-      scale_fill_viridis(discrete = TRUE, begin = 0.35, end = 0.75, option = "B") +
+      scale_color_viridis_d(begin = 0.35, end = 0.75, option = "B") +
+      scale_fill_viridis_d(begin = 0.35, end = 0.75, option = "B") +
       scale_shape_manual(values = c(1, 4, 6, 4, 16, 4, 17, 18, 14, 3, 4), guide = FALSE) +
@@
 
-# G_BLUP 20kb windows Upset plot
+# Definitions
+# Rice/thaliana FT
 
@@
 annos_df
-ft_annos <- ftanno %>%
-  filter(!is.na(OsGene) | !is.na(AtGene)) %>%
-  mutate(Pos_Mb = round(start/10000)/100,
-         #Pos_Mb_e = floor(start/10000)/100,
-         POS_bin = round(Pos_Mb*100/2))# ,
-         #POS_bin_hi = round(Pos_Mb_e*100/2))
+ft_annos
+osgene_ft_key
 
@@
   inner_join(ft_annos, by = c("Gene ID" = "locusName"))
+
+ft_annos <- ft_annos %>%
+  mutate(Pos_Mb1 = round(Pos_Mb, digits = 1))
+annos_plot %>%
+  mutate(Pos_Mb1 = round(POS_Mb, digits = 1)) %>%
+  inner_join(ft_annos, by = c("CHR" = "seqid", "Pos_Mb1"))
 ```
-2682 of 19963...
+528 of 19963...
+528 of 1599
 
@@
 ```
-11324 of 110822 rows...
+1780 of 110822 rows...
 
@@
 qtldf
-i=27
+i=24
 ft_annos %>%
@@
 Yes. Lots. Very encouraging.
+
+# ----------------------------
+# Upset plots
+## 20kb windows Upset plot (G_BLUP)
+
+```{r}
+#annos_upset <- 
+annos_df %>% 
+  mutate(POS_bin_lo = floor(POS_Mb*100/2),
+         POS_bin_hi = ceiling(POS_Mb*100/2)) %>%
+  filter(sites %in% c("eight_sites", "tx_sites", "north_sites") & !(phe %in% c("GR50", "cgdd_12c_10d", "FL50", "crain_1d", "crain_gr2fl"))) %>%
+  select(sites, subpop, phe, CHR, POS_Mb) %>%
+  mutate(All = case_when(sites == "eight_sites" ~ TRUE,
+                         TRUE ~ FALSE),
+         North = case_when(sites == "north_sites" ~ TRUE,
+                         TRUE ~ FALSE),
+         Texas = case_when(sites == "tx_sites" ~ TRUE,
+                         TRUE ~ FALSE),
+         Midwest = case_when(subpop == "Midwest" ~ TRUE,
+                         TRUE ~ FALSE),
+         Gulf = case_when(subpop == "Gulf" ~ TRUE,
+                         TRUE ~ FALSE),
+         Gulf_and_Midwest = case_when(subpop == "Gulf_and_Midwest" ~ TRUE,
+                         TRUE ~ FALSE),
+         )
+
+upsetdf1 <- annos_df %>% 
+  mutate(POS_bin_lo = floor(POS_Mb*100/2),
+         POS_bin_hi = ceiling(POS_Mb*100/2)) %>%
+  filter(sites %in% c("eight_sites", "tx_sites", "north_sites") & !(phe %in% c("GR50", "cgdd_12c_10d", "FL50", "crain_1d", "crain_gr2fl"))) %>%
+  select(sites, subpop, phe, CHR, POS_Mb) %>%
+  unite(col = "Group", sites, subpop, phe) %>%
+  unite(col = "CHR_POS", CHR, POS_Mb) %>%
+  mutate(Sig = TRUE) %>%
+  complete(Group, CHR_POS, fill = list(Sig = FALSE)) %>%
+  group_by(Group, CHR_POS, Sig) %>%
+  slice(1) %>%
+  pivot_wider(names_from = Group, values_from = Sig)
+upset_v1 <- names(upsetdf1)[2:16]
+```
+
+```{r}
+freqplot_u <- upset(data = upsetdf1, intersect = upset_v1, 
+      name="20kb region frequency changes by sites/subpops/phenotype",
+      base_annotations=list(
+        'Intersection size'=intersection_size(text_aes=aes(label=paste0(round(intersection_size/nrow(upsetdf1) * 100), '%')))),
+      min_size = 10,
+      width_ratio = 0.125,
+      themes=upset_modify_themes(
+        list(
+            'overall_sizes'=theme(axis.text.x=element_text(angle=45, hjust =1))
+        ))) +
+  labs(caption = "Number of 20kb windows with a significant GWAS hit.")
+save_plot(filename = "../writeup/figures/Upset_plot_20kb_sig_allele_freq.png", plot = freqplot_u, base_height = 3)
+```

### 2020-10-02 R/Plots_manuscript_G_BLUP.Rmd
Start manuscript v0.4 with FT genes from Xiaoyu incorporated

@@
   arrange(sites, subpop, phe) %>%
-  mutate(pval = 10^-log10p)
+  mutate(pval = 10^-log10p) %>%
+  arrange(pval)
 ```
@@
   arrange(sites, subpop, phe) %>%
-  filter(sites %in% c("eight_sites", "tx_sites", "north_sites") & !(phe %in% c("GR50", "cgdd_12c_10d", "FL50", "crain_1d", "crain_gr2fl"))) 
-  mutate(pval = 10^-log10p)
+  filter(sites %in% c("eight_sites", "tx_sites", "north_sites") & !(phe %in% c("GR50", "cgdd_12c_10d", "FL50", "crain_1d", "crain_gr2fl")) & !subpop %in% c("Gulf_and_Midwest")) %>%
+  arrange(`FDR Adjusted p value`)
+
+```
+# Get all SNPs above FDR from server.
+phe %in% c("dyln_fl50", "cgdd_12c_gr2fl", "FL50")
+sites %in% c("eight_sites", "tx_sites", "north_sites")
+```{r}
+phe_v <- c("dyln_fl50", "cgdd_12c_gr2fl", "FL50")
+site_v <- c("eight_sites", "tx_sites", "north_sites")
+i=1
+j=1
+gwas_FDR <- tibble()
+for(i in seq_along(phe_v)){
+  for(j in seq_along(site_v)){
+  gwas_files <- list.files(path = paste0("./", site_v[j]), pattern = paste0("GWAS_datatable_", phe_v[i]))
+  for(k in seq_along(gwas_files)){
+    gwas1 <- readRDS(file.path(site_v[j], gwas_files[k])) %>%
+      filter(FDR_adj < 0.1) %>%
+      mutate(sites = site_v[j],
+             phe = phe_v[i],
+             gwas = gwas_files[k])
+    gwas_FDR <- bind_rows(gwas_FDR, gwas1)
+  }
+}
+}
+saveRDS(gwas_FDR, file = "Univariate_G_BLUP_27_flowering_GWAS_SNPs_above_FDR10per.rds")
+```
+
+# FDR GWAS hits
+```{r}
+gwas_fdr <- read_rds("../analysis/gwas/BLUPs/Univariate_G_BLUP_27_flowering_GWAS_SNPs_above_FDR10per.rds")
+
+gwas_fdr <- gwas_fdr %>%
+  mutate(subpop = case_when(grepl("8\\.8M", gwas) ~ "Midwest",
+                              grepl("10\\.3M", gwas) ~ "Gulf",
+                              grepl("12\\.1M", gwas) ~ "Gulf_and_Midwest"),
+         Pos_Mb = round(POS/10000)/100,
+         POS_bin = round(Pos_Mb*100/2),
+         Pos_Mb1 = round(Pos_Mb, digits = 1)) %>%
+  filter(!(subpop == "Midwest" & between(log10p, 9.242223, 9.242225)) & !(subpop == "Midwest" & between(log10p, 12.466318, 12.46632)))
+
+gwas_fdr %>%
+  filter(subpop %in% c("Midwest", "Gulf") & phe %in% c("dyln_fl50", "cgdd_12c_gr2fl")) %>%
+  group_by(sites, phe, subpop) %>%
+  tally()
+
+gwas_fdr %>%
+  filter(subpop %in% c("Midwest", "Gulf") & phe %in% c("dyln_fl50", "cgdd_12c_gr2fl")) %>%
+  group_by(CHR, POS) %>%
+  tally() %>% arrange(desc(n))
+
+gwas_fdr %>%
+  filter(subpop %in% c("Midwest", "Gulf") & phe %in% c("dyln_fl50", "cgdd_12c_gr2fl")) %>%
+  group_by(CHR, Pos_Mb1, POS_bin) %>%
+  tally() %>% arrange(desc(n))
+annos_plot1
+ft_annos
+
+gwas_fdr %>%
+  filter(subpop %in% c("Midwest", "Gulf") & phe %in% c("dyln_fl50", "cgdd_12c_gr2fl")) %>%
+  group_by(CHR, POS_bin) %>%
+  tally() %>% arrange(desc(n))
+
+gwas_fdr %>%
+  group_by(log10p) %>%
+  tally() %>% arrange(desc(n))
+```
+## Upset plot FDR
+```{r}
+fdr_upset <- gwas_fdr %>%
+  filter(subpop %in% c("Midwest", "Gulf") & phe %in% c("dyln_fl50", "cgdd_12c_gr2fl")) %>%
+  select(sites, subpop, phe, CHR, POS_bin) %>%
+  unite(col = "GWAS", sites, subpop, phe) %>%
+  unite(col = "CHR_POS", CHR, POS_bin) %>%
+  mutate(Sig = TRUE) %>%
+  complete(GWAS, CHR_POS, fill = list(Sig = FALSE)) %>%
+  group_by(GWAS, CHR_POS, Sig) %>%
+  slice(1) %>%
+  pivot_wider(names_from = GWAS, values_from = Sig)
+upset_v0 <- names(fdr_upset)[2:ncol(fdr_upset)]
+freqplot_u <- upset(data = fdr_upset, intersect = upset_v0, 
+      name="20kb region associations by univariate GWAS",
+      base_annotations=list(
+        'Intersection size'=intersection_size(text_aes=aes(label=paste0(round(intersection_size/nrow(fdr_upset) * 100), '%')))),
+      min_size = 10,
+      width_ratio = 0.125,
+      themes=upset_modify_themes(
+        list(
+            'overall_sizes'=theme(axis.text.x=element_text(angle=45, hjust =1))
+        ))) +
+  labs(caption = "Number of 20kb windows with a significant GWAS hit.")
+save_plot(filename = "../writeup/figures/Upset_plot_20kb_sig_allele_freq.png", plot = freqplot_u, base_height = 3)
+```
+
+Check out various smaller plots
+```{r}
+fdr_upset <- gwas_fdr %>%
+  filter(subpop %in% c("Midwest", "Gulf") & !phe %in% c("FL50")) %>%
+  select(sites, subpop, phe, CHR, POS_bin) %>%
+  unite(col = "GWAS", sites, subpop, phe) %>%
+  unite(col = "CHR_POS", CHR, POS_bin) %>%
+  mutate(Sig = TRUE) %>%
+  complete(GWAS, CHR_POS, fill = list(Sig = FALSE)) %>%
+  group_by(GWAS, CHR_POS, Sig) %>%
+  slice(1) %>%
+  pivot_wider(names_from = GWAS, values_from = Sig)
+upset_v0 <- names(fdr_upset)[2:ncol(fdr_upset)]
+upset(data = fdr_upset, intersect = upset_v0, 
+      name="20kb region associations by univariate GWAS",
+      base_annotations=list(
+        'Intersection size'=intersection_size()),
+      min_size = 0,
+      width_ratio = 0.125,
+      themes=upset_modify_themes(
+        list(
+            'overall_sizes'=theme(axis.text.x=element_text(angle=45, hjust =1))
+        ))) +
+  labs(caption = "Number of 20kb windows with a significant GWAS hit.")
+
+three_uni_gwas <- gwas_fdr %>%
+  filter(subpop %in% c("Midwest", "Gulf") & !phe %in% c("FL50")) %>%
+  group_by(sites, subpop, phe, CHR, POS_bin) %>%
+  tally() %>%
+  group_by(CHR, POS_bin) %>%
+  summarise(count = n()) %>% arrange(desc(count)) %>%
+  filter(count > 2)
+
+three_uni_gwas %>%
+  arrange(CHR, POS_bin) %>%
+  left_join(ft_annos, by = c("CHR" = "seqid", "POS_bin")) %>%
+  select(CHR, Pos_Mb, everything())
+
+gwas_fdr %>%
+  filter(log10p > 7) %>%
+  filter(subpop %in% c("Gulf", "Midwest") & !phe %in% c("FL50")) %>%
+  group_by(sites, subpop, phe, CHR, POS_bin) %>%
+  tally() %>%
+  group_by(CHR, POS_bin) %>%
+  summarise(count = n()) %>%
+  left_join(ft_annos, by = c("CHR" = "seqid", "POS_bin")) %>%
+  select(CHR, Pos_Mb, everything()) %>%
+  filter(!is.na(OsGene) | !is.na(AtGene)) %>%
+  left_join(osgene_ft_key, by = c("OsGene" = "Symbol")) # 93 out of 3148 rows contain candidates
+(9/223)/(1599/83260) # all regions log10p > 7   OR 2.1
+(26/793)/(1599/83260) # all regions log10p > 6   OR 1.7
+(65/2240)/(1599/83260) # all regions log10p > 5   OR 1.51
+(10/280)/(1599/83260) # all regions FDR 1%
+(43/1393)/(1599/83260) # all regions FDR 5%
+(95/3137)/(1599/83260) # all regions FDR 10%
+
+(10/268)/(1599/83260)
+(81/2551)/(1599/83260) # OR north sites 1.65
+(8/420)/(1599/83260) # OR tx sites 0.99! NS!
+(16/463)/(1599/83260) # OR all sites 1.80
+(95/3137)/(1599/83260)
+fisher.test(matrix(c(95, 1599, 3137, 83260), nrow = 2))
+fisher.test(matrix(c(16, 1599, 463, 83260), nrow = 2))
+fisher.test(matrix(c(81, 1599, 2551, 83260), nrow = 2))
+fisher.test(matrix(c(8, 1599, 420, 83260), nrow = 2))
+```
+For the Midwest subpop, there aren't sig assoc for dyln in TX. Just CGDD.
+For the Gulf, there aren't sig assoc for CGDD in TX. Just dyln.
+For G&M in Texas, 61% signals are just for dyln. 33% both dyln & CGDD. 6% CGDD.
+For G&M in North, 76% signals are just for CGDD. 20% both. 2% dyln.
+
+2K 52: FT
+3K 19 FTL10
+
+```{r}
+
 ```
@@
 ```{r}
-annos_df
+annos_df %>%
+  arrange(desc(`FDR Adjusted p value`))
 ft_annos

### 2020-10-10 R/Analysis_manuscript_mash_GxE.Rmd
View single phenotype eight site mash results

---
title: "Analysis_manuscript_mash_GxE"
author: "Alice MacQueen"
date: "10/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(bigsnpr)
#library(mashr)
#install_github("alice-macqueen/switchgrassGWAS")
library(switchgrassGWAS)
```

THEN, for the next section, do GWAS on the BLUPs from the eight single site models. For these traits:
all single sites...
FL50 (actually, maybe not)
dyln_fl50
cgdd_12c_gr2fl
crain_gr2fl
crain_1d
--
GR50 (or maybe not)
cgdd_12c_10d

Then do mash runs. First for single traits across all eight sites, to make U_hyp for the next step - take the U_ed with large percentage variance explained. Make U_hyp with those covariances but no covariance with anything else... 
Then run all the flowering-related traits together. Talk about how then SNPs can have effects that are strongest in response to one type of environmental cue. And how cues can be present or absent at different sites. And how this kind of analysis can help reveal all the patterns at play, because it doesn't make assumptions about how effects are structured.

Maybe bring back the QTL maps after this and talk about how patterns of effects differ across sites in these models, too. Though the Genstat model is less sensitive...

# Genotype-by-environment effects of flowering as functions of environmental cues

While the presence of different associations at the Texas and North gardens demonstrated GxE in flowering at a continental scale, our analysis of G and GxE within each of our eight common gardens also suggested the presence of GxE for rainfall, GDD, and photoperiod cues for flowering, for which variation was more visible outside of each subpopulations’ native range. We evaluated GxE in flowering as a function of five environmental cues: as a function of daylength at flowering (‘flowering daylength’), as a function of daylength change from the previous day on the day of flowering (‘flowering daylength change’) as a function of cumulative GDD between greenup and flowering (‘flowering GDD’), as a function of rainfall on the day of flowering (‘flowering rainfall’), and as a function of rainfall between greenup and flowering (‘cumulative rainfall’). We conducted univariate GWAS at each common garden for these flowering functions, then analyzed the allelic effects of unlinked SNPs across common garden sites for the top XK SNPs using mash. When the same SNP set is used in multiple univariate GWAS, a subsequent mash analysis shares information on patterns of effect size and direction for SNPs across these GWAS, improving the power to detect significant, shared results.

## Make phenotype df for GWAS
```{r}
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "heritability", "kinship_subpop")
site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
               NE = "NE", MI = "MI", SD = "SD", eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"))
phe_eight <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl")
phe_txano <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
               "crain_1d", "GR50", "cgdd_12c_10d")
phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
               "crain_1d", "GR50", "cgdd_12c_10d")
phe_lists <- list(TX1 = phe_single, TX2 = phe_single, TX3 = phe_single, OK = phe_single, MO = phe_single,
               NE = phe_single, MI = phe_single, SD = phe_single, eight_sites = phe_eight, tx_sites = phe_txano, north_sites = phe_txano)

i = 9
j = 1
k = 1
phe_gwas <- tibble()

for(j in seq_along(subpop_v)){
    for(i in seq_along(site_v)){
      phe_v <- phe_lists[[i]]
      for(k in seq_along(phe_v)){
        phe_raw <- read_csv(file.path(outputdir, names(site_v)[i],
                                      paste0("blups_", names(site_v)[i], "_",
                                             names(subpop_v)[j], "_", phe_v[k],
                                             ".csv")))
        names(phe_raw)[2] <- paste0(phe_v[k], "_", names(site_v)[i])
        phe_proc <- phe_raw %>%
          separate(Effect, into = c("factor", "PLANT_ID"), sep = "\\)_") %>%
          filter(!is.na(PLANT_ID) & !grepl("^SITE", factor)) %>%
          left_join(metadata, by = "PLANT_ID") %>%
          filter(SUBPOP %in% subpop_v[[j]]) %>%
          select(PLANT_ID, 3)
        if(i == 1 & k == 1){
          phe_gwas <- phe_proc
        } else {
          phe_gwas <- phe_gwas %>%
            left_join(phe_proc, by = "PLANT_ID")
        }
      }
    }
  phe_gwas <- phe_gwas %>%
  select(PLANT_ID, ends_with("eight_sites"), ends_with("tx_sites"), ends_with("north_sites"), everything())
  write_csv(phe_gwas, path = file.path(workingdir, "analysis", "gwas", "BLUPs",
                                       paste0("BLUP_phenotypes_",
                                              "for_gwas_GR_FL_",
                                              names(subpop_v)[j],
                                              ".csv")))
}
```

## Run GWAS 
In Plots_manuscript_v0.3.Rmd

## Run mash
Then do mash runs. First for single traits across all eight sites, to make U_hyp for the next step - take the U_ed with large percentage variance explained. Make U_hyp with those covariances but no covariance with anything else... 
Then run all the flowering-related traits together. Talk about how then SNPs can have effects that are strongest in response to one type of environmental cue. And how cues can be present or absent at different sites. And how this kind of analysis can help reveal all the patterns at play, because it doesn't make assumptions about how effects are structured.
```{r}
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
               NE = "NE", MI = "MI", SD = "SD")#, eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", Gulf = "229g_10.3M")
phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
               "crain_1d", "GR50", "cgdd_12c_10d")
inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
                                   "inputkinship.txt"),  delim = " ", 
                                   col_names = "SNPfiles")
```

single traits across all eight sites...
```{r}
i=1 # site
j=1 # phe
k=3 # subpop: G&M, M, Gulf
for(j in seq_along(phe_single)){
  for(k in c(1,3)){
phe_suffix <- paste(phe_single[j], subpop_v[k], 
                    sep = "*.*")
snp <- snp_attach(inputfiles$SNPfiles[k])
gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
                                      pattern = paste0("GWAS_datatable*.*",
                                                       phe_suffix))
phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
    str_replace(., "_$", "")
phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
out1phe <- pvdiv_bigsnp2mashr(path = file.path(outputdir), 
                           snp = snp, gwas_rds = gwas_rds, 
                           phenotypes = phenotypes, numSNPs = 1000, 
                           clump = TRUE, scaled = TRUE, saveoutput = TRUE,
                           suffix = paste0("single_phe_", names(subpop_v)[k], 
                                           "_", phe_single[j]))

m_out <- mash_standard_run(path = outputdir, numSNPs = 1000, 
                           suffix = paste0("single_phe_", names(subpop_v)[k], 
                                           "_", phe_single[j]),
                           saveoutput = TRUE)
  }
}
```
Some of the Midwest traits don't work well. Haven't run that subpop yet.

Looking at the results for the other subsets first before I go back and fix the Midwest.

# View single trait mash outputs
```{r}
i=1 # site
j=1 # phe
k=1 # subpop: G&M, M, Gulf
suffix <- paste0("single_phe_", names(subpop_v)[k], "_", phe_single[j])

m_out <- readRDS(file.path(outputdir, paste0("Strong_Effects1000SNPs_", 
                                                     suffix, ".rds")))
switchgrassGWAS::get_U_by_mass(m_out)
switchgrassGWAS::mash_plot_covar(m_out)
switchgrassGWAS::mash_plot_Ulist(m_out, range = get_U_by_mass(m_out))
mash_plot_sig_by_condition(m_out)
mash_plot_manhattan_by_condition(m_out)
mash_plot_pairwise_sharing(m_out, reorder = FALSE)

```

Keep covariance matrices with point mass > 0.05 and add them to U_hyp

### 2020-10-29 R/Analysis_manuscript_mash_GxE.Rmd
Start manuscript 0.5 incorporating TEJ changes and genetic covariances

@@
 i=1 # site
-j=1 # phe
+j=7 # phe
 k=1 # subpop: G&M, M, Gulf
@@
 
-Keep covariance matrices with point mass > 0.05 and add them to U_hyp
+Keep covariance matrices with point mass > 0.05 and add them to U_hyp.
+Add them as single effects (the remaining phe are 0)
+And add them as combinations if possible.
+
+Need to figure out what order the traits are in, after I set up Shat and Bhat matrices for mash. Maybe do this for GR50 first. Only two phenotypes instead of five.
+
+1. Make mash inputs for five & two phenotypes for eight sites. Standardize phenotypes and clump SNPs.
+2. Find trait order.  The trait order is the same order as the 'phenotypes' vector.
+3. Make U_hyp in trait order.
+4. Run mash with U_hyp.
+
+```{r}
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
+site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
+               NE = "NE", MI = "MI", SD = "SD")#, eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
+subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", Gulf = "229g_10.3M")
+phe_mult <- list(flowering_5 = c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
+               "crain_1d"), greenup_2 = c("GR50", "cgdd_12c_10d"))
+inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
+                                   "inputkinship.txt"),  delim = " ", 
+                                   col_names = "SNPfiles")
+```
+
+```{r}
+j=2
+k=3
+for(j in seq_along(phe_mult)){
+  for(k in c(1,3)){
+phe_suffix <- paste(phe_mult[[j]], subpop_v[k], 
+                    sep = "*.*")
+snp <- snp_attach(inputfiles$SNPfiles[k])
+gwas_rds <- c()
+for(i in seq_along(phe_suffix)){
+gwas_rds1 <- pvdiv_results_in_folder(path = file.path(outputdir), 
+                                      pattern = paste0("GWAS_datatable*.*",
+                                                       phe_suffix[i]))
+gwas_rds <- c(gwas_rds, gwas_rds1)
+}
+phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
+    str_replace(., "_$", "")
+
+phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
+numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
+out1phe <- pvdiv_bigsnp2mashr(path = file.path(outputdir), 
+                           snp = snp, gwas_rds = gwas_rds, 
+                           phenotypes = phenotypes, numSNPs = numSNPs, 
+                           clump = TRUE, scaled = TRUE, saveoutput = TRUE,
+                           suffix = paste0("single_phe_", names(subpop_v)[k], 
+                                           "_", names(phe_mult)[j]))
+  }
+}
+```
+
+# Now make U_hyp in trait order. Try programmatically.
+```{r}
+library(rlist)
+v <- matrix(data = 0, nrow=length(phenotypes), ncol=length(phenotypes))
+v <- matrix(data = 0, nrow=16, ncol=16)
+
+i=2 # site
+j=2 # phe
+k=1 # subpop: G&M, M, Gulf
+suffix <- paste0("single_phe_", names(subpop_v)[k], "_", phe_mult[[j]][i])
+
+m_single <- readRDS(file.path(outputdir, paste0("Strong_Effects1000SNPs_", 
+                                                     suffix, ".rds")))
+
+U_single <- m_single$fitted_g$Ulist[get_U_by_mass(m_single)]
+names(m_single$fitted_g$Ulist)
+U_s <- U_single[!names(U_single) %in% c("equal_effects", "identity", "simple_het_1", "simple_het_2", "simple_het_3")]
+
+U_hyp <- list()
+
+  for(q in seq_along(names(U_s))){
+    s_length <- dim(U_s[[1]])[1]
+    hyp1 <- v
+    hyp1[c((i*s_length-s_length+1):(i*s_length)),c((i*s_length-s_length+1):(i*s_length))] <- U_s[[q]]
+    U_hyp <- list.append(U_hyp, hyp1)
+  }
+q=1
+j=1
+U_single_list <- list()
+for(j in seq_along(phe_mult)){
+  suffix <- paste0("single_phe_", names(subpop_v)[k], "_", phe_mult[[j]][i])
+  m_single <- readRDS(file.path(outputdir, paste0("Strong_Effects1000SNPs_", 
+                                                  suffix, ".rds")))
+  U_single1 <- m_single$fitted_g$Ulist[get_U_by_mass(m_single)]
+  U_single_list <- list.append(U_single_list, U_single1)
+}
+
+for(n in )
+
+gwas_rds1 <- pvdiv_results_in_folder(path = file.path(outputdir), 
+                                      pattern = paste0("GWAS_datatable*.*",
+                                                       phe_suffix[i]))
+phenotypes_1 <- str_sub(gwas_rds1, start = 16, end = -42) %>%
+    str_replace(., "_$", "")
+```
+
+```{r}
+m_out <- mash_standard_run(path = outputdir, numSNPs = numSNPs, 
+                           suffix = paste0(names(subpop_v)[k], 
+                                           "_", names(phe_mult)[j]),
+                           saveoutput = TRUE)
+```
+
+The trait order is the same order as the 'phenotypes' vector.

### 2020-10-29 R/Manuscript_v0.5_genetic_covariances.Rmd
Start manuscript 0.5 incorporating TEJ changes and genetic covariances

---
title: "Manuscript_v0.5_genetic_covariances"
author: "Alice MacQueen"
date: "10/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(asreml)
library(cowplot)
library(switchgrassGWAS)
```

# Load data & prep multivariate model data frames.
Add any data that needs to be loaded for these analyses here.
```{r}
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "heritability")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")

fl50_cov_df <- phe_fl_df %>%
  ungroup() %>%
  left_join(select(sites, SITE, manu_site)) %>%
  select(PLANT_ID, SUBPOP, manu_site, FL50) %>%
  group_by(PLANT_ID, SUBPOP, manu_site) %>%
  mutate(IND = row_number()) %>%
  pivot_wider(names_from = manu_site, values_from = "FL50") %>%
  mutate(PLANT_ID = as.factor(PLANT_ID),
         IND = as.factor(IND)) %>%
  select(-IL) %>%
  select(PLANT_ID:TX2, TX3, OK, MO, NE, MI, SD)
gr50_cov_df <- phe_fl_df %>%
  ungroup() %>%
  left_join(select(sites, SITE, manu_site)) %>%
  select(PLANT_ID, SUBPOP, manu_site, GR50) %>%
  group_by(PLANT_ID, SUBPOP, manu_site) %>%
  mutate(IND = row_number()) %>%
  pivot_wider(names_from = manu_site, values_from = "GR50") %>%
  mutate(PLANT_ID = as.factor(PLANT_ID),
         IND = as.factor(IND)) %>%
  select(-IL) %>%
  select(PLANT_ID:TX2, TX3, OK, MO, NE, MI, SD)
```

# In this document:
For both GR50 and FL50, in Gulf, Midwest, and G&M subpops:
fit 7 (or more) bivariate models; compare h^2 estimates between them for A-H.
Fit trivariate models?
Try to fit 8-way model? Run these models on the server so my laptop isn't overloaded.

## bivariate analyses
```{r}
asr_1 <- asreml(cbind(PKLE, KBSM) ~ trait,
                random = ~us(trait, init = c(1,0.1,1)):vm(PLANT_ID, k_full), #+
                  #us(trait, init = c(1, 0.1, 1)):SUBPOP,
                residual = ~units:us(trait, init = c(0.1, 0.1, 0.1)), 
                #na.method(y = "omit", x = "omit"),
                data = fl50_cor_df, 
                workspace = "3gb")
asr_1 <- update.asreml(asr_1)
summary(asr_1)$varcomp

vpredict(asr_1, h2_PKLE ~ V1/(V1+V5))
vpredict(asr_1, h2_KBSM ~ V3/(V3+V7))
vpredict(asr_1, r_G ~ V2/((V1*V3)^0.5))
vpredict(asr_1, r_R ~ V6/((V5*V7)^0.5))

# now try adding a factor to split trait by subpopulation. 
# Nope. I don't think this is a good idea with the kinship matrix there too.
# Maybe try this later if I ever try and partition variance by individuals. Could do for subpops and for individuals/genotypes, also.
asr_2 <- asreml(cbind(PKLE, KBSM) ~ trait,
                random = ~us(trait, init = c(1,0.1,1)):vm(PLANT_ID, k_full) +
                  us(trait, init = c(1, 0.1, 1)):SUBPOP,
                residual = ~units:us(trait, init = c(0.1, 0.1, 0.1)), 
                #na.method(y = "omit", x = "omit"),
                data = fl50_cor_df, 
                workspace = "3gb")
asr_2 <- update.asreml(asr_2)
summary(asr_2)$varcomp
vpredict(asr_2, h2_PKLE ~ V4/(V1+V4+V8))
vpredict(asr_2, h2_KBSM ~ V3/(V3+V6+V10))
vpredict(asr_2, r_S ~ V2/((V1*V3)^0.5))
vpredict(asr_2, r_G ~ V5/((V4*V6)^0.5))

# below works way better (with fewer warnings) when initial values are given. Order for init argument is specified below this code box.
asr_3 <- asreml(cbind(PKLE, KBSM, CLMB) ~ trait,
                random = ~us(trait, init = c(1,0.1, 1, 0.1, 0.1, 1)):vm(PLANT_ID, k_full), 
                residual = ~units:us(trait, init = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1)), 
                na.action = na.method(x = "omit"),
                data = fl50_cor_df, 
                workspace = "3gb")
asr_3 <- update.asreml(asr_3)
summary(asr_3)$varcomp

vpredict(asr_3, h2_PKLE ~ V1/(V1+V8)) # 93.80
vpredict(asr_3, h2_KBSM ~ V3/(V3+V10)) # 90.19
vpredict(asr_3, h2_CLMB ~ V6/(V6+V13)) # 70.58
vpredict(asr_3, r_PK ~ V2/((V1*V3)^0.5)) # 41.23
vpredict(asr_3, r_KC ~ V5/((V3*V6)^0.5)) # 76.57
vpredict(asr_3, r_CP ~ V4/((V1*V6)^0.5)) # 52.84
?na.method()
```
order for init: AA, BA, BB, CA, CB, CC
AA BA BB CA CB CC DA DB DC DD EA EB EC ED EE FA FB FC FD FE FF (6int) GG (7int) HH

2 garden model: h2_PKLE	0.9328744	0.009197664	
h2_KBSM	0.9151788	0.01180919	
r_G	0.4239493	0.04072152	
r_R	0.9709351	0.006573728

### 2020-11-04 R/Analysis_manuscript_mash_GxE.Rmd

--- 2020-10-10 R/Analysis_manuscript_mash_GxE.Rmd
@@
 library(bigsnpr)
+library(rlist)
 #library(mashr)
@@
 
-Keep covariance matrices with point mass > 0.05 and add them to U_hyp
+Keep covariance matrices with point mass > 0.05 and add them to U_hyp.
+Add them as single effects (the remaining phe are 0)
+And add them as combinations if possible.
+
+1. Make mash inputs for five & two phenotypes for eight sites. Standardize phenotypes and clump SNPs.
+2. Find trait order.  The trait order is the same order as the 'phenotypes' vector.
+3. Make U_hyp in trait order.
+4. Run mash with U_hyp.
+# ------- load for m_single to full mash part --
+```{r}
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
+site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
+               NE = "NE", MI = "MI", SD = "SD")#, eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
+subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", Gulf = "229g_10.3M")
+phe_mult <- list(flowering_5 = c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
+               "crain_1d"), greenup_2 = c("GR50", "cgdd_12c_10d"))
+inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
+                                   "inputkinship.txt"),  delim = " ", 
+                                   col_names = "SNPfiles")
+```
+# make mashr inputs for 2 greenup and 5 flowering phenotypes via bigsnp2mashr
+```{r}
+j=2
+k=3
+for(j in seq_along(phe_mult)){
+  for(k in c(1,3)){
+phe_suffix <- paste(phe_mult[[j]], subpop_v[k], 
+                    sep = "*.*")
+snp <- snp_attach(inputfiles$SNPfiles[k])
+gwas_rds <- c()
+for(i in seq_along(phe_suffix)){
+gwas_rds1 <- pvdiv_results_in_folder(path = file.path(outputdir), 
+                                      pattern = paste0("GWAS_datatable*.*",
+                                                       phe_suffix[i]))
+gwas_rds <- c(gwas_rds, gwas_rds1)
+}
+phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
+    str_replace(., "_$", "")
+
+phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
+numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
+out1phe <- pvdiv_bigsnp2mashr(path = file.path(outputdir), 
+                           snp = snp, gwas_rds = gwas_rds, 
+                           phenotypes = phenotypes, numSNPs = numSNPs, 
+                           clump = TRUE, scaled = TRUE, saveoutput = TRUE,
+                           suffix = paste0(names(phe_mult)[j], "_phe_",
+                                           names(subpop_v)[k], "_subpop"))
+  }
+}
+```
+
+# Now make U_hyp in trait order. Try programmatically. Tests here
+```{r}
+v <- matrix(data = 0, nrow=16, ncol=16)
+
+i=2 # site
+j=1 # phe
+k=1 # subpop: G&M, M, Gulf
+# test generate U_ed with point mass > 0.05 from one m_single run.
+suffix <- paste0("single_phe_", names(subpop_v)[k], "_", phe_mult[[j]][i])
+
+m_single <- readRDS(file.path(outputdir, paste0("Strong_Effects1000SNPs_", 
+                                                     suffix, ".rds")))
+
+U_single <- m_single$fitted_g$Ulist[get_U_by_mass(m_single)]
+names(m_single$fitted_g$Ulist)
+U_s <- U_single[!names(U_single) %in% c("equal_effects", "identity", "simple_het_1", "simple_het_2", "simple_het_3")]
+
+q=1
+U_hyp <- list()
+
+# test generate U_hyp for multiple phenotypes from one m_single run.
+for(q in seq_along(names(U_s))){
+    s_length <- dim(U_s[[1]])[1]
+    hyp1 <- v
+    hyp1[c((i*s_length-s_length+1):(i*s_length)),
+         c((i*s_length-s_length+1):(i*s_length))] <- U_s[[q]]
+    U_hyp <- list.append(U_hyp, hyp1)
+}
+```
+
+# -------------------------
+# Create U_hyp programmatically
+```{r}
+library(tidyverse)
+library(cowplot)
+library(bigsnpr)
+library(rlist)
+library(switchgrassGWAS)
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
+site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
+               NE = "NE", MI = "MI", SD = "SD")#, eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
+subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", Gulf = "229g_10.3M")
+phe_mult <- list(flowering_5 = c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
+               "crain_1d"), greenup_2 = c("GR50", "cgdd_12c_10d"))
+```
+
+```{r}
+j=1 # phe_mult: 1 is fl
+i=1
+k=1 # subpop 1 = Gulf   3 = G&M
+
+U_single_list <- list()
+# Make a list (from each phenotype) of lists of U_ed with point mass > 0.05.
+for(i in seq_along(phe_mult[[j]])){
+  suffix <- paste0("single_phe_", names(subpop_v)[k], "_", phe_mult[[j]][i])
+  m_single <- readRDS(file.path(outputdir, paste0("Strong_Effects1000SNPs_", 
+                                                  suffix, ".rds")))
+  U_single1 <- m_single$fitted_g$Ulist[get_U_by_mass(m_single)]
+  U_s1 <- U_single1[!names(U_single1) %in% c("equal_effects", "identity", "simple_het_1", "simple_het_2", "simple_het_3")]
+  U_single_list <- list.append(U_single_list, U_s1)
+}
+
+# Using U_single_list, generate U_hyp for multiple phenotypes.
+
+## part 1: determine size of matrix that U_ed will need to be
+s_total <- 0
+
+for(i in seq_along(U_single_list)){
+    s_length <- dim(as.data.frame(U_single_list[[i]][1]))[1]
+    s_total <- s_total+s_length 
+  }
+
+## part 2: Define null U_ed matrix
+v <- matrix(data = 0, nrow=s_total, ncol=s_total) 
+U_hyp <- list()
+for(i in seq_along(U_single_list)){
+  for(q in seq_along(names(U_single_list[[i]]))){
+    s_length <- dim(as.data.frame(U_single_list[[i]][q]))[1]
+    U_s1 <- matrix(unlist(U_single_list[[i]][q]), nrow = s_length, ncol = s_length)
+    hyp1 <- v
+    hyp1[c((i*s_length-s_length+1):(i*s_length)),c((i*s_length-s_length+1):(i*s_length))] <- U_s1
+    U_hyp <- list.append(U_hyp, hyp1)
+  }
+}
+
+# part 3: determine U_ed with the same names
+all_names <- c()
+for(i in seq_along(U_single_list)){
+  all_names <- c(all_names, names(U_single_list[[i]]))
+}
+U_hyp_comb_df <- enframe(all_names) %>% group_by(value) %>% tally() %>% filter(n > 1)  # these U_ed names have two or more phenotypes with point mass > 0.05.
+
+n=1
+U_hyp_comb <- list()
+for(n in seq_along(U_hyp_comb_df$value)){
+  name_U <- U_hyp_comb_df$value[n]
+  for(i in seq_along(U_single_list)){
+    for(q in seq_along(names(U_single_list[[i]]))){
+      if(i == 1 & q == 1){
+        s_length <- dim(as.data.frame(U_single_list[[i]][q]))[1]
+        hyp1 <- v
+      }
+      if(name_U == names(U_single_list[[i]])[q]){
+        U_s1 <- matrix(unlist(U_single_list[[i]][q]), nrow = s_length, 
+                       ncol = s_length)
+        hyp1[c((i*s_length-s_length+1):(i*s_length)),c((i*s_length-s_length+1):(i*s_length))] <- U_s1
+      }
+    }
+  }
+
+  U_hyp_comb <- list.append(U_hyp_comb, hyp1)
+}
+# part 4: Combine U_single into a combined U_hyp if the names of the U_ed match. 
+# Probably there are better ways to do this but this is the best I can think of, apart from doing every last possible combination.
+That will also suck because I don't have any way to generate covariances between phenotypes.
+So, eh.
+Most likely none of these will get used.
+
+View(U_hyp_comb[[2]])
+U_hyp <- c(U_hyp, U_hyp_comb)
+```
+So between the 'single phenotype' U_hyp and the combined U_hyp, that will give me 16 U_hyp for the five flowering phenotypes, and 8 U_hyp for the greenup phenotypes.
+
+Got it! Now need to either save these U_hyp, or just use them in mash after I generate the relevant list.
+
+# Run mash on multiple phenotypes using these U_hyp
+
+j = 2; k = 1: greenup, Gulf
+
+```{r}
+j=2 # phe_mult 1 = flowering 2 = greenup
+k=3 # subpop 1 = Gulf   3 = G&M
+
+phe_suffix <- paste(phe_mult[[j]], subpop_v[k], 
+                    sep = "*.*")
+gwas_rds <- c()
+for(i in seq_along(phe_suffix)){
+gwas_rds1 <- pvdiv_results_in_folder(path = file.path(outputdir), 
+                                      pattern = paste0("GWAS_datatable*.*",
+                                                       phe_suffix[i]))
+gwas_rds <- c(gwas_rds, gwas_rds1)
+}
+phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
+    str_replace(., "_$", "")
+
+phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
+numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
+
+m_out <- mash_standard_run(path = outputdir, numSNPs = numSNPs, 
+                           U_hyp = U_hyp,
+                           suffix = paste0(names(phe_mult)[j], "_phe_",
+                                           names(subpop_v)[k], "_subpop"),
+                           saveoutput = TRUE)
+```
+
+Another option, I suppose, is just running mash on multiple phenotypes for a very small subset of strong SNPs...
+just a thought.
+Would give covar between phenotypes, too, when now I just have covar within phenotypes.
+Really this is just using a much smaller subset of strong SNPs to define U_ed than I have been doing.
+
+For whatever reason (unknown as of 2020-11-02), finding the data-driven covariance matrices has taken more than a whole weekend for flowering phenotypes (5) in the Gulf subpop.
+I fear errors/the worst but will let it run a while longer to see.
+In case it's the number of conditions, I tried to run a separate mash on greenup phenotypes (2) in the Gulf subpop.
+We'll see if that completes faster.
+
+```{r}
+(18+17+16+15+14+13+12+11+10+9+8+7+6+5+4+3+2+1)*3
+10/513
+```

### 2020-11-04 R/Manuscript_v0.5_genetic_covariances.Rmd

@@
 library(tidyverse)
-library(asreml)
+#library(asreml)
 library(cowplot)
@@
 ## bivariate analyses
+
+asr_1: test bivariate analysis.
 ```{r}
-asr_1 <- asreml(cbind(PKLE, KBSM) ~ trait,
+asr_1 <- asreml(cbind(TX2, MI) ~ trait,
                 random = ~us(trait, init = c(1,0.1,1)):vm(PLANT_ID, k_full), #+
@@
                 #na.method(y = "omit", x = "omit"),
-                data = fl50_cor_df, 
+                data = fl50_cov_df, 
                 workspace = "3gb")
@@
 
-vpredict(asr_1, h2_PKLE ~ V1/(V1+V5))
-vpredict(asr_1, h2_KBSM ~ V3/(V3+V7))
+vpredict(asr_1, h2_TX2 ~ V1/(V1+V5))
+vpredict(asr_1, h2_MO ~ V3/(V3+V7))
 vpredict(asr_1, r_G ~ V2/((V1*V3)^0.5))
@@
 summary(asr_2)$varcomp
-vpredict(asr_2, h2_PKLE ~ V4/(V1+V4+V8))
-vpredict(asr_2, h2_KBSM ~ V3/(V3+V6+V10))
+vpredict(asr_2, h2_TX2 ~ V4/(V1+V4+V8))
+vpredict(asr_2, h2_MI ~ V3/(V3+V6+V10))
 vpredict(asr_2, r_S ~ V2/((V1*V3)^0.5))
 vpredict(asr_2, r_G ~ V5/((V4*V6)^0.5))
+```
 
+asr_3: trivariate analysis.  specifying initial values for variances and covariances.
+```{r}
 # below works way better (with fewer warnings) when initial values are given. Order for init argument is specified below this code box.
-asr_3 <- asreml(cbind(PKLE, KBSM, CLMB) ~ trait,
+asr_3 <- asreml(cbind(TX2, MI, MO) ~ trait,
                 random = ~us(trait, init = c(1,0.1, 1, 0.1, 0.1, 1)):vm(PLANT_ID, k_full), 
@@
                 na.action = na.method(x = "omit"),
-                data = fl50_cor_df, 
-                workspace = "3gb")
+                data = fl50_cov_df, 
+                workspace = "3gb", maxiter = 30)
 asr_3 <- update.asreml(asr_3)
@@
 
-vpredict(asr_3, h2_PKLE ~ V1/(V1+V8)) # 93.80
-vpredict(asr_3, h2_KBSM ~ V3/(V3+V10)) # 90.19
-vpredict(asr_3, h2_CLMB ~ V6/(V6+V13)) # 70.58
-vpredict(asr_3, r_PK ~ V2/((V1*V3)^0.5)) # 41.23
-vpredict(asr_3, r_KC ~ V5/((V3*V6)^0.5)) # 76.57
-vpredict(asr_3, r_CP ~ V4/((V1*V6)^0.5)) # 52.84
+vpredict(asr_3, h2_TX2 ~ V1/(V1+V8)) # 93.80
+vpredict(asr_3, h2_MI ~ V3/(V3+V10)) # 90.19
+vpredict(asr_3, h2_MO ~ V6/(V6+V13)) # 70.58
+vpredict(asr_3, r_TX2MI ~ V2/((V1*V3)^0.5)) # 41.23
+vpredict(asr_3, r_MIMO ~ V5/((V3*V6)^0.5)) # 76.57
+vpredict(asr_3, r_TX2MO ~ V4/((V1*V6)^0.5)) # 52.84
 ?na.method()
+
+asr_4 <- asreml(cbind(TX1, TX2, TX3, OK) ~ trait,
+                random = ~us(trait, init = c(1, 0.1, 1, 0.1, 0.1, 1, 0.1, 0.1, 
+                                             0.1, 1)):vm(PLANT_ID, k_full), 
+                residual = ~units:us(trait, init = c(0.1, 0.1, 0.1, 0.1, 0.1, 
+                                                     0.1, 0.1, 0.1, 0.1, 0.1)), 
+                na.action = na.method(x = "omit"),
+                data = fl50_cov_df, 
+                workspace = "3gb", maxiter = 20)
+asr_4 <- update.asreml(asr_4)
+summary(asr_4)$varcomp
+
+vpredict(asr_4, h2_TX1 ~ V1/(V1+V12)) # 93.80
+vpredict(asr_4, h2_TX2 ~ V3/(V3+V14)) # 90.19
+vpredict(asr_4, h2_TX3 ~ V6/(V6+V17)) # 70.58
+vpredict(asr_4, h2_OK ~ V10/(V10+V21))
+vpredict(asr_4, r_TX2MI ~ V2/((V1*V3)^0.5)) # 41.23
+vpredict(asr_4, r_MIMO ~ V5/((V3*V6)^0.5)) # 76.57
+vpredict(asr_4, r_TX2MO ~ V4/((V1*V6)^0.5)) # 52.84
 ```
@@
 r_R	0.9709351	0.006573728
+
+# On server:
+For both GR50 and FL50, in Gulf, Midwest, and G&M subpops.
+-> subset fl50_cov_df and gr50_cov_df to these three individual subsets. Six subsets.
+```{r}
+subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", 
+                 Gulf_and_Midwest = c("Gulf", "Midwest"))
+i=3
+phe_df <- list(FL50 = fl50_cov_df, GR50 = gr50_cov_df)
+j=2
+
+cov_df_use <- phe_df[[j]] %>%
+  filter(SUBPOP %in% subpop_v[[i]])
+```
+
+for Midwest, try just gardens with n>100 first.
+```{r}
+i=2
+j=2
+asr_3 <- asreml(cbind(TX2, MI, MO) ~ trait,
+                random = ~us(trait, init = c(1,0.1, 1, 0.1, 0.1, 1)):vm(PLANT_ID, k_full), 
+                residual = ~units:us(trait, init = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1)), 
+                na.action = na.method(x = "omit"),
+                data = cov_df_use, 
+                workspace = "6gb", maxiter = 50)
+asr_3 <- update.asreml(asr_3, maxiter = 100)
+summary(asr_3)$varcomp
+
+vpredict(asr_3, h2_TX2 ~ V1/(V1+V8)) # 93.80
+vpredict(asr_3, h2_MI ~ V3/(V3+V10)) # 90.19
+vpredict(asr_3, h2_MO ~ V6/(V6+V13)) # 70.58
+vpredict(asr_3, r_TX2MI ~ V2/((V1*V3)^0.5)) # 41.23
+vpredict(asr_3, r_MIMO ~ V5/((V3*V6)^0.5)) # 76.57
+vpredict(asr_3, r_TX2MO ~ V4/((V1*V6)^0.5)) # 52.84
+saveRDS(asr_3, "analysis/genetic-covariances/GR50_asreml_cov_model_3_gardens_Midwest_subpop.rds") 
+# for the model running 2020-11-02 that will probably have a loglik around -778, which was run with i=2 and j=2.
+```
+
+```{r}
+
+# try all eight gardens first.
+asr_8 <- asreml(cbind(TX1, TX2, TX3, OK, MO, NE, MI, SD) ~ trait,
+                random = ~us(trait, init = c(1, 0.1, 1, 0.1, 0.1, 1, 
+                                             0.1, 0.1, 0.1, 1,
+                                             0.1, 0.1, 0.1, 0.1, 1,
+                                             0.1, 0.1, 0.1, 0.1, 0.1, 1,
+                                             0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 1,
+                                             0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 1)):vm(PLANT_ID, k_full), 
+                residual = ~units:us(trait, init = rep(0.1, 36)), 
+                na.action = na.method(x = "omit"),
+                data = cov_df_use, 
+                workspace = "6gb", maxiter = 240)
+saveRDS(asr_8, "analysis/genetic-covariances/GR50_asreml_cov_model_8_gardens_Gulf_and_Midwest_subpop.rds") 
+summary(asr_8)$varcomp
+asr_8 <- update.asreml(asr_8)
+
+summary(asr_8)$varcomp
+
+saveRDS(asr_8, "analysis/genetic-covariances/GR50_asreml_cov_model_8_gardens_Gulf_and_Midwest_subpop.rds")
+
+# for the model running 2020-11-03 that will probably have a loglik around ____, which was run with i=3 and j=2.
+
+#saveRDS(asr_8, "analysis/genetic-covariances/FL50_asreml_cov_model_8_gardens_Gulf_subpop.rds") 
+# for the model running 2020-11-02 that will probably have a loglik around -3978, which was run with i=1 and j=1.
+
+8+7+6+5+4+3+2+1 # 36 var and covar in eight garden model.
+# programmatically, each pair of two gardens, k and k+1. (special case of 8 & 1 also).
+```
+check model ran and saved ok on the server
+```{r}
+asr_8 <- readRDS("../analysis/genetic-covariances/GR50_asreml_cov_model_8_gardens_Gulf_subpop.rds")
+summary(asr_8)$varcomp
+vpredict(asr_8, h2_tx1 ~ V1/(V1+V38))
+vpredict(asr_8, h2_tx2 ~ V3/(V3+V40))
+vpredict(asr_8, h2_tx3 ~ V6/(V6+V43))
+vpredict(asr_8, h2_ok ~ V10/(V10+V47))  # all sig > 0 minus 2SE, ok is the closest to zero... only 3.8%
+vpredict(asr_8, h2_mo ~ V15/(V15+V52))
+vpredict(asr_8, h2_ne ~ V21/(V21+V58))
+vpredict(asr_8, h2_mi ~ V28/(V28+V65))
+vpredict(asr_8, h2_sd ~ V36/(V36+V73))
+
+vpredict(asr_8, cov_tx2tx1 ~ V2/((V1*V3)^0.5))  # sig pos
+
+vpredict(asr_8, cov_tx3tx1 ~ V4/((V1*V6)^0.5))   # sig pos
+vpredict(asr_8, cov_tx3tx2 ~ V5/((V6*V3)^0.5))   # sig pos
+
+vpredict(asr_8, cov_oktx1 ~ V7/((V1*V10)^0.5)) 
+vpredict(asr_8, cov_oktx2 ~ V8/((V3*V10)^0.5)) 
+vpredict(asr_8, cov_oktx3 ~ V9/((V6*V10)^0.5))
+
+vpredict(asr_8, cov_motx1 ~ V11/((V1*V15)^0.5)) 
+vpredict(asr_8, cov_motx2 ~ V12/((V3*V15)^0.5))  # ** sig neg even plus 2SE
+vpredict(asr_8, cov_motx3 ~ V13/((V6*V15)^0.5)) 
+vpredict(asr_8, cov_mook ~ V14/((V10*V15)^0.5))
+
+vpredict(asr_8, cov_netx1 ~ V16/((V1*V21)^0.5)) 
+vpredict(asr_8, cov_netx2 ~ V17/((V3*V21)^0.5)) 
+vpredict(asr_8, cov_netx3 ~ V18/((V6*V21)^0.5)) 
+vpredict(asr_8, cov_neok ~ V19/((V10*V21)^0.5)) 
+vpredict(asr_8, cov_nemo ~ V20/((V15*V21)^0.5))   # sig pos
+
+vpredict(asr_8, cov_mitx1 ~ V22/((V1*V28)^0.5)) 
+vpredict(asr_8, cov_mitx2 ~ V23/((V3*V28)^0.5)) # ** sig neg even plus 2SE
+vpredict(asr_8, cov_mitx3 ~ V24/((V6*V28)^0.5)) 
+vpredict(asr_8, cov_miok ~ V25/((V10*V28)^0.5)) 
+vpredict(asr_8, cov_mimo ~ V26/((V15*V28)^0.5)) # sig pos
+vpredict(asr_8, cov_mine ~ V27/((V21*V28)^0.5))
+
+vpredict(asr_8, cov_sdtx1 ~ V29/((V1*V36)^0.5)) # ** sig neg even plus 2SE
+vpredict(asr_8, cov_sdtx2 ~ V30/((V3*V36)^0.5)) # ** sig neg even plus 2SE
+vpredict(asr_8, cov_sdtx3 ~ V31/((V6*V36)^0.5)) # ** sig neg even plus 2SE
+vpredict(asr_8, cov_sdok ~ V32/((V10*V36)^0.5)) 
+vpredict(asr_8, cov_sdmo ~ V33/((V15*V36)^0.5)) # sig pos
+vpredict(asr_8, cov_sdne ~ V34/((V21*V36)^0.5)) 
+vpredict(asr_8, cov_sdmi ~ V35/((V28*V36)^0.5))
+
+#36.80566-(16.49249*2)
+```
+36.80566-(16.49249*2)
+
+fit 7 (or more) bivariate models; compare h^2 estimates between them for A-H.
+-> collect h^2 estimates and SE in tibbles, and compare/plot them.
+
+----------------------
+Fit trivariate models?
+Try to fit 8-way model? Run these models on the server so my laptop isn't overloaded.

### 2020-11-13 R/Manuscript_v0.5_genetic_covariances.Rmd
Bivariate genetic covariances in Gulf subpop

@@
 library(tidyverse)
-#library(asreml)
+library(asreml)
 library(cowplot)
@@
 outputdir <- file.path(workingdir, "analysis", "heritability")
+inputdir <- file.path(workingdir, "analysis", "genetic-covariances")
 metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
@@
 
+# Will Midwest models converge for bivariate analyses?
+
+Can't get trivariate analysis to converge ever for the Midwest subpop, though the Gulf & G&M models do converge. 
+Unfortunately I will need to fit ~28 models per phenotype and subpop for bivariate models to get all of the genetic covariances. Blehhhhh.
+```{r}
+garden_v <- c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD")
+subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", 
+                 Gulf_and_Midwest = c("Gulf", "Midwest"))
+phe_df <- list(FL50 = fl50_cov_df, GR50 = gr50_cov_df)
+k=3; j = 1; i = 1; q = 1
+var_table <- tibble(Value = "Value", Estimate = 0, SE = 0, Phenotype = "Phenotype", Subpop = "Subpop", Garden_1 = "garden1", Garden_2 = "garden2", loglik = 0)
+
+append_vpredict_bivar <- function(vpredict_df, var_df, 
+                                  phenotype = names(phe_df)[j], 
+                                  subpop = names(subpop_v)[i], 
+                                  garden1 = garden_v[k],
+                                  garden2 = garden_v2[q], loglik = mloglik){
+  var_df <- vpredict_df %>%
+    as_tibble(rownames = "Value") %>%
+    mutate(Phenotype = phenotype,
+           Subpop = subpop,
+           Garden_1 = garden1,
+           Garden_2 = garden2,
+           loglik = mloglik) %>%
+       add_row(var_df, .)
+  return(var_df)
+}
+
+for(i in seq_along(subpop_v)){
+  for(j in seq_along(phe_df)){
+    for(k in seq_along(garden_v)){
+      garden_v2 <- garden_v[-(1:k)]
+      for(q in seq_along(garden_v2)){
+    cov_df_use <- phe_df[[j]] %>%
+  filter(SUBPOP %in% subpop_v[[i]]) %>%
+      rename(garden1 = garden_v[k], garden2 = garden_v2[q]) %>%
+      filter(!(is.na(garden1) & is.na(garden2)))
+    tryCatch({
+      asr_2 <- asreml(cbind(garden1, garden2) ~ trait,
+                      random = ~us(trait, init = c(1, 0.1, 1)):vm(PLANT_ID,
+                                                                    k_full), 
+                      residual = ~units:us(trait, init = c(0.1, 0.1, 0.1)), 
+                      na.action = na.method(x = "omit"),
+                      data = cov_df_use, 
+                      workspace = "8gb", maxiter = 300)
+      asr_2 <- update.asreml(asr_2, maxiter = 20)
+      saveRDS(asr_2, file.path(inputdir, paste0(names(phe_df)[j], "_asreml_cov_model_", garden_v[k], "_", garden_v2[q], "_gardens_", names(subpop_v)[i], "_subpop_200_iterations.rds")))
+      mloglik = asr_2$loglik[1]
+      var_table <- vpredict(asr_2, h2_garden1 ~ V1/(V1+V5)) %>% 
+        append_vpredict_bivar(., var_table)
+      var_table <- vpredict(asr_2, h2_garden2 ~ V3/(V3+V7)) %>% 
+        append_vpredict_bivar(.,var_table)
+      var_table <- vpredict(asr_2, r_G ~ V2/((V1*V3)^0.5)) %>% 
+        append_vpredict_bivar(., var_table)
+      }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+    saveRDS(var_table, file.path(inputdir,     
+                                 "Two_site_var_table_up_to_300_iterations.rds"))
+      }
+    }
+  }
+}
+
+```
+
+```{r}
+var_table <- read_rds(file.path(inputdir,     
+                                "Two_site_var_table_up_to_300_iterations.rds"))
+
+var_table %>%
+  write_csv(., file.path(inputdir,     
+                                "Two_site_var_table_up_to_300_iterations.csv"))
+```
+
 asr_3: trivariate analysis.  specifying initial values for variances and covariances.
@@
 
+#------------
+# Three site covariances loop
+```{r}
+subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", 
+                 Gulf_and_Midwest = c("Gulf", "Midwest"))
+phe_df <- list(FL50 = fl50_cov_df, GR50 = gr50_cov_df)
+
+var_table <- tibble(Value = "Value", Estimate = 0, SE = 0, Phenotype = "Phenotype", Subpop = "Subpop")
+
+for(i in seq_along(subpop_v)){
+  for(j in seq_along(phe_df)){
+    cov_df_use <- phe_df[[j]] %>%
+  filter(SUBPOP %in% subpop_v[[i]])
+      asr_3 <- asreml(cbind(TX2, MI, MO) ~ trait,
+                      random = ~us(trait, init = c(1, 0.1, 1, 
+                                                   0.1, 0.1, 1)):vm(PLANT_ID,
+                                                                    k_full), 
+                      residual = ~units:us(trait, init = c(0.1, 0.1, 0.1, 
+                                                           0.1, 0.1, 0.1)), 
+                      na.action = na.method(x = "omit"),
+                      data = cov_df_use, 
+                      workspace = "6gb", maxiter = 2600)
+      saveRDS(asr_3, paste0("analysis/genetic-covariances/", names(phe_df)[j], "_asreml_cov_model_3_gardens_", names(subpop_v)[i], "_subpop_2600_iterations.rds"))
+
+     var_table <- vpredict(asr_3, h2_TX2 ~ V1/(V1+V8)) %>% 
+       as_tibble(rownames = "Value") %>%
+       mutate(Phenotype = names(phe_df)[j],
+              Subpop = names(subpop_v)[i]) %>%
+       add_row(var_table, .)
+     var_table <- vpredict(asr_3, h2_MI ~ V3/(V3+V10)) %>% 
+       as_tibble(rownames = "Value") %>%
+       mutate(Phenotype = names(phe_df)[j],
+              Subpop = names(subpop_v)[i]) %>%
+       add_row(var_table, .)
+     var_table <- vpredict(asr_3, h2_MO ~ V6/(V6+V13)) %>% 
+       as_tibble(rownames = "Value") %>%
+       mutate(Phenotype = names(phe_df)[j],
+              Subpop = names(subpop_v)[i]) %>%
+       add_row(var_table, .) 
+     var_table <- vpredict(asr_3, r_TX2MI ~ V2/((V1*V3)^0.5)) %>% 
+       as_tibble(rownames = "Value") %>%
+       mutate(Phenotype = names(phe_df)[j],
+              Subpop = names(subpop_v)[i]) %>%
+       add_row(var_table, .) 
+     var_table <- vpredict(asr_3, r_MIMO ~ V5/((V3*V6)^0.5))  %>% 
+       as_tibble(rownames = "Value") %>%
+       mutate(Phenotype = names(phe_df)[j],
+              Subpop = names(subpop_v)[i]) %>%
+       add_row(var_table, .)
+     var_table <- vpredict(asr_3, r_TX2MO ~ V4/((V1*V6)^0.5)) %>% 
+       as_tibble(rownames = "Value") %>%
+       mutate(Phenotype = names(phe_df)[j],
+              Subpop = names(subpop_v)[i]) %>%
+       add_row(var_table, .) 
+      saveRDS(var_table, "Three_site_var_table_50_iterations_usually.rds")
+
+  }
+}
+asr_3 <- update.asreml(asr_3, maxiter = 100)
+```
+
+## view ASReml output from three site covariance models
+```{r}
+var_table <- read_rds(file.path(inputdir,
+                          "Three_site_var_table_50_iterations_usually_v2.rds"))
+var_table %>% write_csv(., file.path(inputdir, "Three_site_var_table_50_iterations_or_more.csv"))
+```
+
 for Midwest, try just gardens with n>100 first.
@@
 ```
-check model ran and saved ok on the server
+check models ran and saved ok on the server
+
+## View 8 output and save to a var_table
 ```{r}
 asr_8 <- readRDS("../analysis/genetic-covariances/GR50_asreml_cov_model_8_gardens_Gulf_subpop.rds")
+asr_8 <- readRDS("../analysis/genetic-covariances/FL50_asreml_cov_model_8_gardens_Gulf_subpop.rds")
 summary(asr_8)$varcomp
-vpredict(asr_8, h2_tx1 ~ V1/(V1+V38))
-vpredict(asr_8, h2_tx2 ~ V3/(V3+V40))
-vpredict(asr_8, h2_tx3 ~ V6/(V6+V43))
-vpredict(asr_8, h2_ok ~ V10/(V10+V47))  # all sig > 0 minus 2SE, ok is the closest to zero... only 3.8%
-vpredict(asr_8, h2_mo ~ V15/(V15+V52))
-vpredict(asr_8, h2_ne ~ V21/(V21+V58))
-vpredict(asr_8, h2_mi ~ V28/(V28+V65))
-vpredict(asr_8, h2_sd ~ V36/(V36+V73))
 
-vpredict(asr_8, cov_tx2tx1 ~ V2/((V1*V3)^0.5))  # sig pos
+subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", 
+                 Gulf_and_Midwest = c("Gulf", "Midwest"))
+phe_df <- list(FL50 = fl50_cov_df, GR50 = gr50_cov_df)
+j=1
+i=1
+summary(asr_8)$call
+var_table <- tibble(Value = "Value", Estimate = 0, SE = 0, Phenotype = "Phenotype", Subpop = "Subpop")
+append_vpredict <- function(vpredict_df, var_df, phenotype = names(phe_df)[j], 
+                            subpop = names(subpop_v)[i]){
+  var_df <- vpredict_df %>%
+    as_tibble(rownames = "Value") %>%
+    mutate(Phenotype = phenotype,
+           Subpop = subpop) %>%
+       add_row(var_df, .)
+  return(var_df)
+}
 
-vpredict(asr_8, cov_tx3tx1 ~ V4/((V1*V6)^0.5))   # sig pos
-vpredict(asr_8, cov_tx3tx2 ~ V5/((V6*V3)^0.5))   # sig pos
+var_table <- vpredict(asr_8, h2_tx1 ~ V1/(V1+V38)) %>% 
+  append_vpredict(., var_table)
+var_table <- vpredict(asr_8, h2_tx2 ~ V3/(V3+V40)) %>% 
+  append_vpredict(., var_table)
+var_table <- vpredict(asr_8, h2_tx3 ~ V6/(V6+V43)) %>% 
+  append_vpredict(., var_table)
+var_table <- vpredict(asr_8, h2_ok ~ V10/(V10+V47)) %>% 
+  append_vpredict(., var_table)  # all sig > 0 minus 2SE, ok is the closest to zero... only 3.8%
+var_table <- vpredict(asr_8, h2_mo ~ V15/(V15+V52)) %>% 
+  append_vpredict(., var_table)
+var_table <- vpredict(asr_8, h2_ne ~ V21/(V21+V58)) %>% 
+  append_vpredict(., var_table)
+var_table <- vpredict(asr_8, h2_mi ~ V28/(V28+V65)) %>% 
+  append_vpredict(., var_table)
+var_table <- vpredict(asr_8, h2_sd ~ V36/(V36+V73)) %>% 
+  append_vpredict(., var_table)
 
-vpredict(asr_8, cov_oktx1 ~ V7/((V1*V10)^0.5)) 
-vpredict(asr_8, cov_oktx2 ~ V8/((V3*V10)^0.5)) 
-vpredict(asr_8, cov_oktx3 ~ V9/((V6*V10)^0.5))
+var_table <- vpredict(asr_8, cov_tx2tx1 ~ V2/((V1*V3)^0.5)) %>% 
+  append_vpredict(., var_table)  # sig pos
 
-vpredict(asr_8, cov_motx1 ~ V11/((V1*V15)^0.5)) 
-vpredict(asr_8, cov_motx2 ~ V12/((V3*V15)^0.5))  # ** sig neg even plus 2SE
-vpredict(asr_8, cov_motx3 ~ V13/((V6*V15)^0.5)) 
-vpredict(asr_8, cov_mook ~ V14/((V10*V15)^0.5))
+var_table <- vpredict(asr_8, cov_tx3tx1 ~ V4/((V1*V6)^0.5)) %>% 
+  append_vpredict(., var_table)   # sig pos
+var_table <- vpredict(asr_8, cov_tx3tx2 ~ V5/((V6*V3)^0.5)) %>% 
+  append_vpredict(., var_table)   # sig pos
 
-vpredict(asr_8, cov_netx1 ~ V16/((V1*V21)^0.5)) 
-vpredict(asr_8, cov_netx2 ~ V17/((V3*V21)^0.5)) 
-vpredict(asr_8, cov_netx3 ~ V18/((V6*V21)^0.5)) 
-vpredict(asr_8, cov_neok ~ V19/((V10*V21)^0.5)) 
-vpredict(asr_8, cov_nemo ~ V20/((V15*V21)^0.5))   # sig pos
+var_table <- vpredict(asr_8, cov_oktx1 ~ V7/((V1*V10)^0.5)) %>% 
+  append_vpredict(., var_table) 
+var_table <- vpredict(asr_8, cov_oktx2 ~ V8/((V3*V10)^0.5))  %>% 
+  append_vpredict(., var_table)
+var_table <- vpredict(asr_8, cov_oktx3 ~ V9/((V6*V10)^0.5))  %>% 
+  append_vpredict(., var_table)
 
-vpredict(asr_8, cov_mitx1 ~ V22/((V1*V28)^0.5)) 
-vpredict(asr_8, cov_mitx2 ~ V23/((V3*V28)^0.5)) # ** sig neg even plus 2SE
-vpredict(asr_8, cov_mitx3 ~ V24/((V6*V28)^0.5)) 
-vpredict(asr_8, cov_miok ~ V25/((V10*V28)^0.5)) 
-vpredict(asr_8, cov_mimo ~ V26/((V15*V28)^0.5)) # sig pos
-vpredict(asr_8, cov_mine ~ V27/((V21*V28)^0.5))
+var_table <- vpredict(asr_8, cov_motx1 ~ V11/((V1*V15)^0.5)) %>% 
+  append_vpredict(., var_table) 
+var_table <- vpredict(asr_8, cov_motx2 ~ V12/((V3*V15)^0.5)) %>% 
+  append_vpredict(., var_table)  # ** sig neg even plus 2SE
+var_table <- vpredict(asr_8, cov_motx3 ~ V13/((V6*V15)^0.5)) %>% 
+  append_vpredict(., var_table) 
+var_table <- vpredict(asr_8, cov_mook ~ V14/((V10*V15)^0.5)) %>% 
+  append_vpredict(., var_table)
 
-vpredict(asr_8, cov_sdtx1 ~ V29/((V1*V36)^0.5)) # ** sig neg even plus 2SE
-vpredict(asr_8, cov_sdtx2 ~ V30/((V3*V36)^0.5)) # ** sig neg even plus 2SE
-vpredict(asr_8, cov_sdtx3 ~ V31/((V6*V36)^0.5)) # ** sig neg even plus 2SE
-vpredict(asr_8, cov_sdok ~ V32/((V10*V36)^0.5)) 
-vpredict(asr_8, cov_sdmo ~ V33/((V15*V36)^0.5)) # sig pos
-vpredict(asr_8, cov_sdne ~ V34/((V21*V36)^0.5)) 
-vpredict(asr_8, cov_sdmi ~ V35/((V28*V36)^0.5))
+var_table <- vpredict(asr_8, cov_netx1 ~ V16/((V1*V21)^0.5)) %>% 
+  append_vpredict(., var_table) 
+var_table <- vpredict(asr_8, cov_netx2 ~ V17/((V3*V21)^0.5)) %>% 
+  append_vpredict(., var_table) 
+var_table <- vpredict(asr_8, cov_netx3 ~ V18/((V6*V21)^0.5)) %>% 
+  append_vpredict(., var_table) 
+var_table <- vpredict(asr_8, cov_neok ~ V19/((V10*V21)^0.5)) %>% 
+  append_vpredict(., var_table) 
+var_table <- vpredict(asr_8, cov_nemo ~ V20/((V15*V21)^0.5)) %>% 
+  append_vpredict(., var_table)   # sig pos
 
+var_table <- vpredict(asr_8, cov_mitx1 ~ V22/((V1*V28)^0.5)) %>% 
+  append_vpredict(., var_table) 
+var_table <- vpredict(asr_8, cov_mitx2 ~ V23/((V3*V28)^0.5)) %>% 
+  append_vpredict(., var_table) # ** sig neg even plus 2SE
+var_table <- vpredict(asr_8, cov_mitx3 ~ V24/((V6*V28)^0.5)) %>% 
+  append_vpredict(., var_table) 
+var_table <- vpredict(asr_8, cov_miok ~ V25/((V10*V28)^0.5)) %>% 
+  append_vpredict(., var_table) 
+var_table <- vpredict(asr_8, cov_mimo ~ V26/((V15*V28)^0.5)) %>% 
+  append_vpredict(., var_table) # sig pos
+var_table <- vpredict(asr_8, cov_mine ~ V27/((V21*V28)^0.5)) %>% 
+  append_vpredict(., var_table)
+
+var_table <- vpredict(asr_8, cov_sdtx1 ~ V29/((V1*V36)^0.5)) %>% 
+  append_vpredict(., var_table) # ** sig neg even plus 2SE
+var_table <- vpredict(asr_8, cov_sdtx2 ~ V30/((V3*V36)^0.5)) %>% 
+  append_vpredict(., var_table) # ** sig neg even plus 2SE
+var_table <- vpredict(asr_8, cov_sdtx3 ~ V31/((V6*V36)^0.5)) %>% 
+  append_vpredict(., var_table) # ** sig neg even plus 2SE
+var_table <- vpredict(asr_8, cov_sdok ~ V32/((V10*V36)^0.5)) %>% 
+  append_vpredict(., var_table) 
+var_table <- vpredict(asr_8, cov_sdmo ~ V33/((V15*V36)^0.5)) %>% 
+  append_vpredict(., var_table) # sig pos
+var_table <- vpredict(asr_8, cov_sdne ~ V34/((V21*V36)^0.5)) %>% 
+  append_vpredict(., var_table) 
+var_table <- vpredict(asr_8, cov_sdmi ~ V35/((V28*V36)^0.5)) %>% 
+  append_vpredict(., var_table)
+
+write_csv(var_table, file.path(inputdir, "Var_Table_8_garden_Gulf_models_20_iter.csv"))
 #36.80566-(16.49249*2)

### 2020-11-13 R/Analysis_manuscript_mash_GxE.Rmd
Add greenup mash analysis section to manuscript

@@
 library(switchgrassGWAS)
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
 ```
@@
 ```{r}
-j=1 # phe_mult: 1 is fl
+j=2 # phe_mult: 1 is fl
 i=1
-k=1 # subpop 1 = Gulf   3 = G&M
+k=3 # subpop 1 = Gulf_and_Midwest   3 = Gulf
 
@@
 
-View(U_hyp_comb[[2]])
+#View(U_hyp_comb[[2]])
 U_hyp <- c(U_hyp, U_hyp_comb)
@@
 j=2 # phe_mult 1 = flowering 2 = greenup
-k=3 # subpop 1 = Gulf   3 = G&M
+k=3 # subpop 3 = Gulf   1 = G&M
 
@@
 
-m_out <- mash_standard_run(path = outputdir, numSNPs = numSNPs, 
-                           U_hyp = U_hyp,
+U_ed <- readRDS(paste0("analysis/gwas/BLUPs/single_sites/Mash_data_driven_covariances_", numSNPs, "_", names(phe_mult)[j], "_phe_", names(subpop_v)[k],
+                       "_subpop", ".rds"))
+U_hyp <- readRDS(paste0("analysis/gwas/BLUPs/single_sites/Mash_data_driven_covariances_", numSNPs, "_", names(phe_mult)[j], "_phe_", names(subpop_v)[3],
+                       "_subpop", ".rds"))
+
+m_out2 <- mash_standard_run(path = outputdir, numSNPs = numSNPs, 
+                           U_hyp = U_hyp, U_ed = U_ed,
                            suffix = paste0(names(phe_mult)[j], "_phe_",
@@
                            saveoutput = TRUE)
+
 ```
@@
 ```
+
+```{r}
+subpop_v  # 1 is G&M  3 is Gulf
+numSNPs <- 5000
+m1 <- readRDS(paste0("../analysis/gwas/BLUPs/single_sites/Strong_Effects", numSNPs, "SNPs_", names(phe_mult)[j], "_phe_", names(subpop_v)[1],
+                       "_subpop", ".rds"))
+m3 <- readRDS(paste0("../analysis/gwas/BLUPs/single_sites/Strong_Effects", numSNPs, "SNPs_", names(phe_mult)[j], "_phe_", names(subpop_v)[3],
+                       "_subpop", ".rds"))
+
+mash_plot_Ulist(m1,1)
+mash_plot_Ulist(m3,1)
+mash_plot_covar(m1)
+mash_plot_covar(m3)
+m1$loglik
+m3$loglik
+length(get_significant_results(m3))
+```
+
+```{r}
+m <- readRDS(paste0("../analysis/gwas/BLUPs/single_sites/Strong_Effects5000SNPs_greenup_2_phe_Gulf_and_Midwest_subpop_Gulf_Uhyp.rds"))
+
+mash_plot_covar(m)
+mash_plot_sig_by_condition(m)
+mash_plot_manhattan_by_condition(m)
+mash_plot_pairwise_sharing(m)
+mash_plot_effects(m, n = 6)
+
+switchgrassGWAS::get_U_by_mass(m)
+```
+
+```{r}
+library(mashr)
+mash_plot_covar <- function(m, saveoutput = FALSE, suffix = ""){
+  df <- get_estimated_pi(m)
+  df <- enframe(df, name = "Covariance Matrix", value = "Mass") %>%
+    mutate(`Covariance Matrix` = str_replace(.data$`Covariance Matrix`,
+                                             "(Stand_)??Bhat_?",
+                                             "single_effect_"),
+           `Covariance Matrix` = str_replace(.data$`Covariance Matrix`,
+                                             "^null$", "no_effects")) %>%
+    arrange(desc(.data$`Covariance Matrix`))
+  df$`Covariance Matrix` <- factor(df$`Covariance Matrix`,
+                                   levels = (df$`Covariance Matrix`))
+  ggobject <- ggplot(df) +
+    geom_bar(aes(x = .data$Mass, y = .data$`Covariance Matrix`),
+             stat = "identity") +
+    switchgrassGWAS::theme_oeco
+
+  if(saveoutput == TRUE){
+    if(!(str_sub(suffix, end = 1) %in% c("", "_"))){
+      suffix <- paste0("_", suffix)
+    }
+    if(str_sub(suffix, end = 1) %in% c("")){
+      suffix <- paste0("_", get_date_filename())
+    }
+    save_plot(paste0("Covariance_matrix_mass_plot", suffix,
+                     ".png"), plot = ggobject, base_aspect_ratio = 1,
+              base_height = 4)
+  }
+  return(list(covar_df = df, ggobject = ggobject))
+}
+```
+
+# ---------------------
+
+# Greenup mash results
+
+#' @param posterior_weights Vector containing the posterior
+#'     probability of each mixture component in Ulist for the data.
+```{r}
+length(get_significant_results(m))
+get_significant_results(m)[1:10]
+get_significant_results(m)[1:10]
+
+str(m$posterior_weights)
+
+n=7500
+i = get_significant_results(m)[n]
+
+# largest probability that this SNP loads on this covariance matrix
+max(m$posterior_weights[i,])
+which(m$posterior_weights[i,] == max(m$posterior_weights[i,]))
+
+# second largest probability that this SNP loads on this covariance matrix
+max( m$posterior_weights[i,][m$posterior_weights[i,]!=max(m$posterior_weights[i,])])
+which(m$posterior_weights[i,] == max(m$posterior_weights[i,][m$posterior_weights[i,]!=max(m$posterior_weights[i,])]))
+
+sum(m$posterior_weights[,1])
+
+mash_plot_covar(m, saveoutput = TRUE)
+
+colnames(m$posterior_weights)
+Ulist <- switchgrassGWAS:::get_Ulist(m)
+names(Ulist)
+
+df <- mashr:::get_estimated_pi(m, dimension = "all")
+```
+
+Gulf
+```{r}
+inputdir <- file.path("../analysis/gwas/BLUPs/single_sites", "Gulf")
+mash_out <- pvdiv_results_in_folder(inputdir, "Strong")
+mash_out
+m <- read_rds(file.path(inputdir, mash_out[11]))
+mash_plot_covar(m, saveoutput = TRUE)
+get_U_by_mass(m)
+mash_plot_Ulist(m, range = get_U_by_mass(m))
+mash_plot_effects(m, n = 1, saveoutput = TRUE)
+mash_plot_sig_by_condition(m, saveoutput = TRUE)
+mash_plot_manhattan_by_condition(m, saveoutput = TRUE)
+mash_plot_pairwise_sharing(effectRDS = file.path(inputdir, mash_out[1])) # sign
+mash_plot_pairwise_sharing(effectRDS = file.path(inputdir, mash_out[2])) # sign & size
+Gulf_greenup_gxe <- get_GxE(m)
+
+mash_plot_pairwise_sharing(corrmatrix = Gulf_greenup_gxe$S_AP/Gulf_greenup_gxe$S_all_pairwise, saveoutput = TRUE)
+mash_plot_pairwise_sharing(corrmatrix = Gulf_greenup_gxe$S_DS/Gulf_greenup_gxe$S_all_pairwise, reorder = TRUE, saveoutput = TRUE)
+mash_plot_pairwise_sharing(corrmatrix = Gulf_greenup_gxe$S_2_no, reorder = FALSE)
+mash_plot_pairwise_sharing(corrmatrix = Gulf_greenup_gxe$S_all_pairwise, reorder = FALSE)
+
+```
+
+```{r}
+821+2309+789+1401+1090+1277+658+1075
+9420/10581
+```
+
+# SVD of a covariance matrix
+
+Work through an two-condition example here to help me understand how to explain this.
+
+Sample points from a bivariate Gaussian distribution with a standard deviation of 3 in roughly the lower left-upper right direction and of 1 in the orthogonal direction.
+Because the x and y components co-vary, the variances of x and y do not fully describe the distribution.
+A 2 × 2 covariance matrix is needed; the directions of the arrows correspond to the eigenvectors of this covariance matrix and their lengths to the square roots of the eigenvalues.
+```{r}
+cov1 <- matrix(c(1, -0.4, 0.3, -0.4, 1, 0.7, 0.3, 0.7, 1), ncol = 3 )
+
+b1 <- svd(cov1)
+
+b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 80%
+b1$d[2] / sum(b1$d) # 20%
+b1$d[3] / sum(b1$d) # 20%
+# b1$u symmetric with v if first matrix is symmetric
+b1$v
+b1$d
+
+#sum(b1$d[1:2]) / sum(b1$d) # First 2 eigenvectors explain 100% of total variance.
+for(g in 1:3){
+barplot(b1$v[,g]/cov1[which.max(abs(cov1[,g])),g])
+}
+
+L = chol(cov1)
+nvars = dim(L)[1]
+nobs=200
+# Random variables that follow an M correlation matrix
+r = t(L) %*% matrix(rnorm(nvars*nobs), nrow=nvars, ncol=nobs)
+r = t(r)
+
+rdata = as.data.frame(r)
+names(rdata) = c('cond1', 'cond2', 'cond3')
+library(lattice)
+splom(rdata)
+```
+The closer the covariance between x and y is to zero, the more equal the eigenvalues will be, and the more equal the proportion of variance explained by each eigenvector will be.
+If the covariance is negative the first eigenvector will show a sign change.
+
+# ------------------
+# Gulf & Midwest
+```{r}
+inputdir <- file.path("../analysis/gwas/BLUPs/single_sites", "Gulf_and_Midwest")
+mash_out <- pvdiv_results_in_folder(inputdir, "Strong")
+mash_out
+m <- read_rds(file.path(inputdir, mash_out[13]))
+
+mash_plot_covar(m)
+get_U_by_mass(m)
+mash_plot_Ulist(m, range = get_U_by_mass(m))
+mash_plot_effects(m, n = 1)
+mash_plot_sig_by_condition(m)
+mash_plot_manhattan_by_condition(m)
+mash_plot_pairwise_sharing(effectRDS = file.path(inputdir, mash_out[1])) # sign
+mash_plot_pairwise_sharing(effectRDS = file.path(inputdir, mash_out[2])) # sign & size
+Gulf_greenup_gxe <- get_GxE(m)
+```
+
+```{r}
+get_U_by_mass(m)
+i=2
+names(Ulist)[i]
+Ulist <- switchgrassGWAS:::get_Ulist(m)
+b1 <- svd(Ulist[[i]])
+b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 11.5
+b1$d[2] / sum(b1$d) # get to 60%
+b1$d[3] / sum(b1$d) # 
+b1$d[4] / sum(b1$d) # 
+b1$d[5] / sum(b1$d) #
+
+sum(b1$d[1:2]) / sum(b1$d) # First 5 eigenvectors explain 1% or more variance, and 27.7% of total variance.
+for(g in 1:5){
+barplot(b1$v[,g]/Ulist[[1]][which.max(abs(Ulist[[1]][,g])),g])
+}
+
+ggb1 <- tibble(`First eigenvector (PVE = 100%)` = b1$v[,1]/Ulist[[1]][which.max(abs(Ulist[[1]][,1])),1],
+       Phenotype = switchgrassGWAS:::get_colnames(m)) %>%
+  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_")) %>%
+  separate(Phenotype, into = c("Function", "Garden"), sep = -3) %>%
+  mutate(Function = str_remove(Function, "_$"),
+         Garden = str_remove(Garden, "^_"))
+ggb1$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
+ggb1_plot <- ggb1 %>%
+  ggplot(aes(x = Garden, y = `First eigenvector (PVE = 100%)`)) +
+  geom_bar(aes(group = Function, color = Function, fill = Function), stat = "identity", position = "dodge") +
+  theme(legend.position = "top", panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(hjust = 1, angle = 45)) +
+  geom_hline(yintercept = 0, linetype = 2) +
+  scale_color_viridis_d(option = "B", end = 0.7, 
+                      direction = -1) +
+  scale_fill_viridis_d(option = "B", begin = 0.1, end = 0.7, direction = -1) 
+save_plot(paste0("../manuscript/plots/SVD_of_greenup_Gulf_", names(Ulist)[i], ".png"), base_height = 4, plot = ggb1_plot)
+
+ggb2 <- tibble(`First eigenvector (PVE = 21.7%)` = b1$v[,2]/Ulist[[1]][which.max(abs(Ulist[[1]][,2])),2],
+       Phenotype = switchgrassGWAS:::get_colnames(m)) %>%
+  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_")) %>%
+  separate(Phenotype, into = c("Function", "Garden"), sep = -3) %>%
+  mutate(Function = str_remove(Function, "_$"),
+         Garden = str_remove(Garden, "^_"))
+ggb2$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
+ggb2_plot <- ggb2 %>%
+  ggplot(aes(x = Garden, y = `First eigenvector (PVE = 21.7%)`)) +
+  geom_bar(aes(group = Function, color = Function, fill = Function), stat = "identity", position = "dodge") +
+  theme(legend.position = "top", panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(hjust = 1, angle = 45)) +
+  geom_hline(yintercept = 0, linetype = 2) +
+  scale_color_viridis_d(option = "B", end = 0.7, 
+                      direction = -1) +
+  scale_fill_viridis_d(option = "B", begin = 0.1, end = 0.7, direction = -1) 
+save_plot(paste0("../manuscript/plots/SVD_of_greenup_Gulf_", names(Ulist)[i], "Eigenvector_2.png"), base_height = 4, plot = ggb2_plot)
+
+```
+GO enrichment amongst things that load onto different covariance matrices?

### 2020-11-19 R/Allele_frequencies_and_positions.Rmd
Add mash analysis with Mashattan by covariance matrix and other allele features

---
title: "Allele frequencies and positions"
author: "Alice MacQueen"
date: "11/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(cowplot)
library(maps)
library(ggmap)
library(ggrepel)
library(lubridate)
usa <- map_data("usa")
metadata <- read_rds("../data/metadata.rds")
```

#
```{r}
metadata %>%
  group_by(GENETIC_SUBPOP_KINSHIP) %>%
  tally()
Atlantic <- metadata %>%
  filter(GENETIC_SUBPOP_KINSHIP == "Atlantic")
Gulf <- metadata %>%
  filter(GENETIC_SUBPOP_KINSHIP == "Gulf")
Midwest <- metadata %>%
  filter(GENETIC_SUBPOP_KINSHIP == "Midwest")

metadata %>%
  group_by(ECOTYPE_NNET) %>%
  tally()
Coastal <- metadata %>%
  filter(ECOTYPE_NNET == "Coastal")
Lowland <- metadata %>%
  filter(ECOTYPE_NNET == "Lowland")
Upland <- metadata %>%
  filter(ECOTYPE_NNET == "Upland")
metadata %>%
  filter(GENETIC_SUBPOP_KINSHIP %in% c("Atlantic", "Gulf", "Midwest")) %>%
  group_by(COLLECTION_TYPE) %>%
  tally()

# vcf.gz in SharedUser has PLANT_ID as the library names now.
writeLines(Atlantic$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Atlantic", "_subpop"))
writeLines(Gulf$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Gulf", "_subpop"))
writeLines(Midwest$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Midwest", "_subpop"))
writeLines(Coastal$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Coastal", "_ecotype"))
writeLines(Lowland$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Lowland", "_ecotype"))
writeLines(Upland$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Upland", "_ecotype"))
```

What are things I could color Manhattans by? For mash results?

Ideally, always for alternate alleles (aka not AP13's allele, in reality):

The covariance matrix that the SNP has the highest posterior probability for. Or second highest. Or whatever.

Their allele frequency in each subpopulation.

The average latitudinal and/or longitudinal position in our diversity panel. Or, as a factor, which state or region is this position in?

Which ecotype the allele is predominantly found in? Or is it common everywhere?

```{r}
allele_freq <- readRDS("~/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/Allele_frequencies_site_by_year_100000SNPs.rds")
allele_freq
```

Maybe big_count in bigsnpr could take care of this for me.
```{r}
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")

inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
                                   "inputkinship.txt"),  delim = " ", 
                                   col_names = "SNPfiles")

metadata <- read_rds("~/Github/pvdiv-phenology-gxe/data/metadata.rds")
```

```{r}
#snp <- snp_attach("~/Github/pvdiv-genome/genetic_subpops/Pvirgatum_V5_GWAS_229g_Gulf_K_subset_maf0.05.rds")
snp <- snp_attach(inputfiles$SNPfiles[1])

metadata1 <- metadata %>%
  filter(PLANT_ID %in% snp$fam$sample.ID) %>%
  mutate(Atlantic = ifelse(GENETIC_SUBPOP_KINSHIP == "Atlantic", 1, 0),
         Gulf = ifelse(GENETIC_SUBPOP_KINSHIP == "Gulf", 1, 0),
         Midwest = ifelse(GENETIC_SUBPOP_KINSHIP == "Midwest", 1, 0),
         Coastal = ifelse(ECOTYPE_NNET == "Coastal", 1, 0),
         Upland = ifelse(ECOTYPE_NNET == "Upland", 1, 0),
         Lowland = ifelse(ECOTYPE_NNET == "Lowland", 1, 0),
         COLL_YEAR = year(COLL_DATE))
i=1
m1 <- metadata1 %>%
  mutate(SNP = snp$genotypes[,i]) %>%
  summarise(Alt_freq = mean(SNP/2, na.rm = TRUE),
            Alt_Latitude = weighted.mean(LATITUDE, SNP, na.rm = TRUE),
            Alt_Longitude = weighted.mean(-LONGITUDE, SNP, na.rm = TRUE),
            Alt_Elevation = weighted.mean(ELEVATION, SNP, na.rm = TRUE),
            Alt_Year = weighted.mean(COLL_YEAR, SNP, na.rm = TRUE),
            Alt_Coastal_d = weighted.mean(Coastal, SNP, na.rm = TRUE),
            Alt_Coastal_c = weighted.mean(COASTAL_ASSIGNMENT, SNP, 
                                          na.rm = TRUE),
            Alt_Lowland_d = weighted.mean(Lowland, SNP, na.rm = TRUE),
            Alt_Lowland_c = weighted.mean(LOWLAND_ASSIGNMENT, SNP, 
                                          na.rm = TRUE),
            Alt_Upland_d = weighted.mean(Upland, SNP, na.rm = TRUE),
            Alt_Upland_c = weighted.mean(UPLAND_ASSIGNMENT, SNP, na.rm = TRUE),
            Alt_Atlantic_d = weighted.mean(Atlantic, SNP, na.rm = TRUE),
            Alt_Atlantic_c = weighted.mean(ATLANTIC_Q, SNP, na.rm = TRUE),
            Alt_Gulf_d = weighted.mean(Gulf, SNP, na.rm = TRUE),
            Alt_Gulf_c = weighted.mean(GULF_Q, SNP, na.rm = TRUE),
            Alt_Midwest_d = weighted.mean(Midwest, SNP, na.rm = TRUE),
            Alt_Midwest_c = weighted.mean(MIDWEST_DA, SNP, na.rm = TRUE)
            ) %>%
  mutate(Marker = snp$map$marker.ID[i]) %>%
  select(Marker, everything())

for(i in 2:snp$genotypes$ncol){
  m_i <- metadata1 %>%
  mutate(SNP = snp$genotypes[,i]) %>%
  summarise(Alt_freq = mean(SNP/2, na.rm = TRUE),
            Alt_Latitude = weighted.mean(LATITUDE, SNP, na.rm = TRUE),
            Alt_Longitude = weighted.mean(-LONGITUDE, SNP, na.rm = TRUE),
            Alt_Elevation = weighted.mean(ELEVATION, SNP, na.rm = TRUE),
            Alt_Year = weighted.mean(COLL_YEAR, SNP, na.rm = TRUE),
            Alt_Coastal_d = weighted.mean(Coastal, SNP, na.rm = TRUE),
            Alt_Coastal_c = weighted.mean(COASTAL_ASSIGNMENT, SNP, 
                                          na.rm = TRUE),
            Alt_Lowland_d = weighted.mean(Lowland, SNP, na.rm = TRUE),
            Alt_Lowland_c = weighted.mean(LOWLAND_ASSIGNMENT, SNP, 
                                          na.rm = TRUE),
            Alt_Upland_d = weighted.mean(Upland, SNP, na.rm = TRUE),
            Alt_Upland_c = weighted.mean(UPLAND_ASSIGNMENT, SNP, na.rm = TRUE),
            Alt_Atlantic_d = weighted.mean(Atlantic, SNP, na.rm = TRUE),
            Alt_Atlantic_c = weighted.mean(ATLANTIC_Q, SNP, na.rm = TRUE),
            Alt_Gulf_d = weighted.mean(Gulf, SNP, na.rm = TRUE),
            Alt_Gulf_c = weighted.mean(GULF_Q, SNP, na.rm = TRUE),
            Alt_Midwest_d = weighted.mean(Midwest, SNP, na.rm = TRUE),
            Alt_Midwest_c = weighted.mean(MIDWEST_DA, SNP, na.rm = TRUE)
            ) %>%
  mutate(Marker = snp$map$marker.ID[i]) %>%
  select(Marker, everything())
  m1 <- m1 %>% add_row(m_i)
}

write_rds(m1, "~/Github/pvdiv-phenology-gxe/large-files/Metadata_associated_with_Alternate_alleles_363g_Midwest_and_Gulf.rds", compress = "gz")
```
Of the 363 Gulf & Midwest plants:
83.7 have ecotype assignment without the neural network.
67.7% have collection date
93.9% have lat & long; 89% have elevation

# Map this test Lat/Long
```{r}
ggplot() + 
  geom_polygon(data = usa, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  coord_quickmap(xlim = c(-105, -66), ylim = c(25, 50)) + 
  geom_point(data = m1, aes(x = SNP_Long, y = SNP_Lat), shape = 13, 
             size = 4, stroke = 1.5) + 
  labs(x = "Longitude", y = "Latitude")
```

### 2020-11-19 R/Analysis_manuscript_mash_GxE.Rmd
Add mash analysis with Mashattan by covariance matrix and other allele features

@@
 
-#' @param posterior_weights Vector containing the posterior
-#'     probability of each mixture component in Ulist for the data.
+@param posterior_weights Vector containing the posterior
+     probability of each mixture component in Ulist for the data.
 ```{r}
@@
 
-n=7500
+n=1
 i = get_significant_results(m)[n]
@@
 
-Gulf
+## Load m for Gulf
 ```{r}
@@
 m <- read_rds(file.path(inputdir, mash_out[11]))
+```
+
+Mashhattan where points are colored by which covariance matrix they load most strongly on to. Do different regions of the genome have distinct patterns, or are they all jumbled up?
+```{r}
+
+  log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
+    as.data.frame() %>%
+    rownames_to_column(var = "value") %>%
+    mutate(value = as.integer(.data$value)) %>%
+    as_tibble() %>%
+    left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
+    dplyr::rename(log10BayesFactor = .data$V1) %>%
+    dplyr::select(-.data$value)
+
+#' Return the Bayes Factor for each effect
+#'
+#' @param m the mash result (from joint or 1by1 analysis); must have been
+#'     computed using usepointmass=TRUE
+#'
+#' @return if m was fitted using usepointmass=TRUE then returns a vector of
+#'     the log10(bf) values for each effect. That is, the jth element
+#'     lbf_j is log10(Pr(Bj | g=ghat-nonnull)/Pr(Bj | g = 0)) where gha
+#'     t-nonnull is the non-null part of ghat.  Otherwise returns NULL.
+#'
+get_log10bf = function(m) {
+  if(is.null(m$null_loglik)){
+    return(NULL)
+  } else {
+    return((m$alt_loglik - m$null_loglik)/log(10))
+  }
+}
+
+#' Return the name of the covariance matrix with the highest posterior weight
+#'     for each effect
+#'
+#' @param m the mash result (from joint or 1by1 analysis)
+#'
+#' @return returns a df with Marker, Covariance_Matrix, and Grid_Value. I'm assuming that the grid value is an indication of like a svd decomp
+#'
+get_topcov = function(m) {
+  df <- switchgrassGWAS:::get_marker_df(m = m)
+  covdf <- purrr::map_chr(df$value, ~ names(which(m$posterior_weights[.x,] ==
+                                           max(m$posterior_weights[.x,])))) %>%
+    enframe(name = "value", value = "Covariance_Grid") %>%
+    separate("Covariance_Grid", into = c("Covariance_Matrix", "Grid_Value"),
+             sep = "\\.", remove = FALSE) 
+  df <- df %>% full_join(covdf, by = "value")
+  return(df)
+  }
+
+names(which(m$posterior_weights[i,] == max(m$posterior_weights[i,])))
+
+  ggman_df <- df %>%
+    separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
+             extra = "merge", convert = TRUE) %>%
+    left_join(log10bf_df, by = "Marker") %>%
+    arrange(.data$Chr, .data$Pos) %>%
+    group_by(Covariance_Grid) %>% add_tally() %>%
+    arrange(desc(n))
+
+  log10BF <- expression(paste("log"[10], plain("(Bayes Factor)")))
+
+  ggmanobject <- ggplot(data = ggman_df, aes(x = .data$Pos, y = .data$log10BayesFactor)) +
+    switchgrassGWAS::theme_oeco +
+    geom_point(aes(color = .data$Covariance_Grid, fill = .data$Covariance_Grid,
+                   shape = as.factor(.data$Chr))) +
+    facet_wrap(~ .data$Chr, nrow = 1, scales = "free_x", strip.position = "bottom") +
+    scale_color_viridis_d(option = "D", end = 0.95) + scale_fill_viridis_d(option = "D", end = 0.95) +
+    theme(strip.text = element_text(size = 8),
+          axis.text.y = element_text(size = 8),
+          axis.text.x = element_blank(),
+          legend.text = element_text(size = 8),
+          axis.ticks.x = element_blank(),
+          panel.background = element_rect(fill=NA)) +
+    labs(x = "Chromosome", y = log10BF) +
+    scale_x_continuous(expand = c(0.32, 0.32)) +
+    scale_shape_manual(values = rep(c(21,22),9), guide = FALSE) +
+    geom_hline(yintercept = 2, linetype = 2)
+```
+
+```{r}
+
 mash_plot_covar(m, saveoutput = TRUE)

### 2020-11-21 R/Analysis_manuscript_mash_GxE.Rmd
Mashattan R scripts with allele frequency and subpop information

--- 2020-11-13 R/Analysis_manuscript_mash_GxE.Rmd
@@
 ```{r}
-m <- readRDS(paste0("../analysis/gwas/BLUPs/single_sites/Strong_Effects5000SNPs_greenup_2_phe_Gulf_and_Midwest_subpop_Gulf_Uhyp.rds"))
+m <- readRDS(paste0("~/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/Strong_Effects5000SNPs_greenup_2_phe_Gulf_and_Midwest_subpop_Gulf_Uhyp.rds"))
 
@@
 
-Gulf
+## Load m for Gulf
 ```{r}
-inputdir <- file.path("../analysis/gwas/BLUPs/single_sites", "Gulf")
+inputdir <- file.path("../analysis/gwas/BLUPs/single_sites", "Gulf_and_Midwest")
 mash_out <- pvdiv_results_in_folder(inputdir, "Strong")
 mash_out
-m <- read_rds(file.path(inputdir, mash_out[11]))
+m <- read_rds(file.path(inputdir, mash_out[13]))
+m_meta <- read_rds("../analysis/Metadata_associated_with_alternate_alleles_snp_630g_PKLE_m_greenup_2_phe_Gulf_subpop_full.rds")
+m_meta <- read_rds("../analysis/Metadata_associated_with_alternate_alleles_snp_363g_m_greenup_2_phe_Gulf_and_Midwest_subpop.rds")
+```
+
+Mashhattan where points are colored by which covariance matrix they load most strongly on to. Do different regions of the genome have distinct patterns, or are they all jumbled up?
+```{r}
+log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
+    as.data.frame() %>%
+    rownames_to_column(var = "value") %>%
+    mutate(value = as.integer(.data$value)) %>%
+    as_tibble() %>%
+    left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
+    dplyr::rename(log10BayesFactor = .data$V1) %>%
+    dplyr::select(-.data$value)
+
+#' Return the name of the covariance matrix with the highest posterior weight
+#'     for each effect
+#'
+#' @param m the mash result (from joint or 1by1 analysis)
+#'
+#' @return returns a df with Marker, Covariance_Matrix, and Grid_Value. I'm assuming that the grid value is an indication of like a svd decomp
+#'
+get_topcov = function(m){
+  df <- switchgrassGWAS:::get_marker_df(m = m)
+  covdf <- purrr::map_chr(df$value, ~ names(which(m$posterior_weights[.x,] ==
+                                           max(m$posterior_weights[.x,])))) %>%
+    enframe(name = "value", value = "Covariance_Grid") %>%
+    separate("Covariance_Grid", into = c("Covariance_Matrix", "Grid_Value"),
+             sep = "\\.", remove = FALSE) 
+  df <- df %>% full_join(covdf, by = "value")
+  return(df)
+  }
+
+df <- get_topcov(m)
+
+  ggman_df <- df %>%
+    left_join(m_meta) %>%
+    separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
+             extra = "merge", convert = TRUE) %>%
+    left_join(log10bf_df, by = "Marker") %>%
+    arrange(.data$Chr, .data$Pos)
+
+  log10BF <- expression(paste("log"[10], plain("(Bayes Factor)")))
+
+  ggmanobject <- ggman_df %>%
+    group_by(Covariance_Grid) %>% add_tally() %>%
+    arrange(desc(n)) %>%
+    ggplot(aes(x = .data$Pos, y = .data$log10BayesFactor)) +
+    switchgrassGWAS::theme_oeco +
+    geom_point(aes(color = .data$Covariance_Grid, fill = .data$Covariance_Grid,
+                   shape = as.factor(.data$Chr))) +
+    facet_wrap(~ .data$Chr, nrow = 1, scales = "free_x", strip.position = "bottom") +
+    scale_color_viridis_d(option = "D", end = 0.95) + scale_fill_viridis_d(option = "D", end = 0.95) +
+    theme(strip.text = element_text(size = 8),
+          axis.text.y = element_text(size = 8),
+          axis.text.x = element_blank(),
+          legend.text = element_text(size = 8),
+          axis.ticks.x = element_blank(),
+          panel.background = element_rect(fill=NA)) +
+    labs(x = "Chromosome", y = log10BF) +
+    scale_x_continuous(expand = c(0.32, 0.32)) +
+    scale_shape_manual(values = rep(c(21,22),9), guide = FALSE) +
+    geom_hline(yintercept = 2, linetype = 2)
+
+ggman_df %>%
+    group_by(Covariance_Grid) %>% add_tally() %>%
+    arrange(desc(n)) %>%
+    ggplot(aes(x = .data$Pos, y = .data$log10BayesFactor)) +
+    switchgrassGWAS::theme_oeco +
+    geom_point(aes(color = .data$Covariance_Grid, fill = .data$Covariance_Grid,
+                   alpha = Alt_freq,
+                   shape = as.factor(.data$Chr))) +
+    facet_wrap(~ .data$Chr, nrow = 1, scales = "free_x", strip.position = "bottom") +
+    scale_color_viridis_d(option = "D", end = 0.95) + scale_fill_viridis_d(option = "D", end = 0.95) +
+    theme(strip.text = element_text(size = 8),
+          axis.text.y = element_text(size = 8),
+          axis.text.x = element_blank(),
+          legend.text = element_text(size = 8),
+          axis.ticks.x = element_blank(),
+          panel.background = element_rect(fill=NA)) +
+    labs(x = "Chromosome", y = log10BF) +
+    scale_x_continuous(expand = c(0.32, 0.32)) +
+    scale_shape_manual(values = rep(c(21,22),9), guide = FALSE) +
+    geom_hline(yintercept = 2, linetype = 2)
+
+ggman_df %>%
+  #arrange((Alt_Latitude)) %>%
+    ggplot(aes(x = .data$Pos, y = .data$log10BayesFactor)) +
+    switchgrassGWAS::theme_oeco +
+    geom_point(aes(color = .data$Alt_Midwest_c,
+                   #alpha = Alt_freq,
+                   shape = as.factor(.data$Chr))) +
+    facet_wrap(~ .data$Chr, nrow = 1, scales = "free_x", 
+               strip.position = "bottom") +
+    scale_color_viridis_c(option = "B", end = 0.95) +
+    #scale_fill_viridis_d(option = "D", end = 0.95) +
+    theme(strip.text = element_text(size = 8),
+          axis.text.y = element_text(size = 8),
+          axis.text.x = element_blank(),
+          legend.text = element_text(size = 8),
+          axis.ticks.x = element_blank(),
+          panel.background = element_rect(fill=NA)) +
+    labs(x = "Chromosome", y = log10BF) +
+    scale_x_continuous(expand = c(0.32, 0.32)) +
+    scale_shape_manual(values = rep(c(15, 16),9), guide = FALSE) +
+    geom_hline(yintercept = 2, linetype = 2)
+
+ggman_df
+```
+
+enrichment of sig SNPs for these different Alt_features.
+Enrichment of covariance matrices for these different Alt_features
+```{r}
+library(lubridate)
+metadata1 <- pvdiv_metadata %>%
+  filter(SUBPOP %in% c("Atlantic", "Midwest", "Gulf")) %>%
+    mutate(Atlantic = ifelse(GENETIC_SUBPOP_KINSHIP == "Atlantic", 1, 0),
+           Gulf = ifelse(GENETIC_SUBPOP_KINSHIP == "Gulf", 1, 0),
+           Midwest = ifelse(GENETIC_SUBPOP_KINSHIP == "Midwest", 1, 0),
+           Coastal = ifelse(ECOTYPE_NNET == "Coastal", 1, 0),
+           Upland = ifelse(ECOTYPE_NNET == "Upland", 1, 0),
+           Lowland = ifelse(ECOTYPE_NNET == "Lowland", 1, 0),
+           COLL_YEAR = year(COLL_DATE))
+
+ave630 <- metadata1 %>%
+  summarise(All_Latitude = mean(LATITUDE, na.rm = TRUE),
+            All_Longitude = mean(-LONGITUDE, na.rm = TRUE),
+            All_Elevation = mean(ELEVATION, na.rm = TRUE),
+            All_Year = mean(COLL_YEAR, na.rm = TRUE),
+            All_Coastal_d = mean(Coastal, na.rm = TRUE),
+            All_Coastal_c = mean(COASTAL_ASSIGNMENT, na.rm = TRUE),
+            All_Lowland_d = mean(Lowland, na.rm = TRUE),
+            All_Lowland_c = mean(LOWLAND_ASSIGNMENT, na.rm = TRUE),
+            All_Upland_d = mean(Upland, na.rm = TRUE),
+            All_Upland_c = mean(UPLAND_ASSIGNMENT, na.rm = TRUE),
+            All_Atlantic_d = mean(Atlantic),
+            All_Atlantic_c = mean(ATLANTIC_Q, na.rm = TRUE),
+            All_Gulf_d = mean(Gulf, na.rm = TRUE),
+            All_Gulf_c = mean(GULF_Q, na.rm = TRUE),
+            All_Midwest_d = mean(Midwest, na.rm = TRUE),
+            All_Midwest_c = mean(MIDWEST_DA, na.rm = TRUE))
+
+ave363 <- metadata1 %>%
+  filter(SUBPOP %in% c("Midwest", "Gulf")) %>%
+  summarise(All_Latitude = mean(LATITUDE, na.rm = TRUE),
+            All_Longitude = mean(-LONGITUDE, na.rm = TRUE),
+            All_Elevation = mean(ELEVATION, na.rm = TRUE),
+            All_Year = mean(COLL_YEAR, na.rm = TRUE),
+            All_Coastal_d = mean(Coastal, na.rm = TRUE),
+            All_Coastal_c = mean(COASTAL_ASSIGNMENT, na.rm = TRUE),
+            All_Lowland_d = mean(Lowland, na.rm = TRUE),
+            All_Lowland_c = mean(LOWLAND_ASSIGNMENT, na.rm = TRUE),
+            All_Upland_d = mean(Upland, na.rm = TRUE),
+            All_Upland_c = mean(UPLAND_ASSIGNMENT, na.rm = TRUE),
+            All_Atlantic_d = mean(Atlantic),
+            All_Atlantic_c = mean(ATLANTIC_Q, na.rm = TRUE),
+            All_Gulf_d = mean(Gulf, na.rm = TRUE),
+            All_Gulf_c = mean(GULF_Q, na.rm = TRUE),
+            All_Midwest_d = mean(Midwest, na.rm = TRUE),
+            All_Midwest_c = mean(MIDWEST_DA, na.rm = TRUE))
+```
+
+```{r}
+# --------
+
+P1 <- ggman_df %>%
+  mutate(Dev_Latitude = Alt_Latitude - ave363$All_Latitude,
+         Dev_Longitude = Alt_Longitude - ave363$All_Longitude,
+         Dev_Elevation = Alt_Elevation - ave363$All_Elevation,
+         Dev_Year = Alt_Year - ave363$All_Year,
+         Dev_Coastal = Alt_Coastal_d - ave363$All_Coastal_d,
+         Dev_Lowland = Alt_Lowland_d - ave363$All_Lowland_d,
+         Dev_Upland = Alt_Upland_d - ave363$All_Upland_d,
+         Dev_Atlantic = Alt_Atlantic_d - ave363$All_Atlantic_d,
+         Dev_Gulf = Alt_Gulf_d - ave363$All_Gulf_d,
+         Dev_Midwest = Alt_Midwest_d - ave363$All_Midwest_d,) 
+testvar <- "Dev_Midwest"
+P1 <- P1 %>% rename(testvar = testvar)
+
+lim_lower <- min(P1$testvar)
+lim_upper <- max(P1$testvar) + max(P1$testvar) * 0.01
+
+P1 %>%
+  filter(log10BayesFactor>2) %>%
+  ggplot(aes(x = testvar)) + 
+  geom_histogram() + 
+  geom_vline(xintercept = 0) +
+  facet_wrap(~Covariance_Grid)
+
+P1 %>%
+  ggplot(aes(x = .data$Pos, y = .data$log10BayesFactor)) +
+  switchgrassGWAS::theme_oeco +
+  geom_point(aes(color = .data$testvar,
+                 #alpha = Alt_freq,
+                 shape = as.factor(.data$Chr))) +
+  facet_grid(.data$Covariance_Grid ~ .data$Chr, switch = "x", 
+             scales = "free_x") +
+  scale_color_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                    "white", "#27AD81FF", "#5DC863FF",
+                                    "#FDE725FF"),
+                      limits = c(lim_lower, lim_upper)) +
+  theme(strip.text = element_text(size = 8),
+        axis.text.y = element_text(size = 8),
+        axis.text.x = element_blank(),
+        legend.text = element_text(size = 8),
+        axis.ticks.x = element_blank(),
+        panel.background = element_rect(fill=NA)) +
+  labs(x = "Chromosome", y = log10BF) +
+  scale_x_continuous(expand = c(0.32, 0.32)) +
+  scale_shape_manual(values = rep(c(15, 16),10), guide = FALSE) +
+  geom_hline(yintercept = 2, linetype = 2)
+```
+
+```{r}
+
 mash_plot_covar(m, saveoutput = TRUE)

### 2020-12-03 R/Analysis_v0.5_mash_GxE.Rmd
Move R scripts not used in current manuscript and rename those that are

--- 2020-11-21 R/Analysis_manuscript_mash_GxE.Rmd
@@
 ```
+
+# TRIAL script
+# -------------------------
+NB: Trial of generating covariance matrices for combinations of environmental phenotypes from the U_ed from single environmental phenotypes.
 
@@
                NE = "NE", MI = "MI", SD = "SD")#, eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
-subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", Gulf = "229g_10.3M")
-phe_mult <- list(flowering_5 = c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
-               "crain_1d"), greenup_2 = c("GR50", "cgdd_12c_10d"))
+subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", 
+                 Gulf = "229g_10.3M")
+phe_mult <- list(flowering_5 = c("FL50", "dyln_fl50", "cgdd_12c_gr2fl",
+                                 "crain_gr2fl", "crain_1d"), greenup_2 = c("GR50", "cgdd_12c_10d"))
 ```
@@
 
+# --- New Mashhattan --
 Mashhattan where points are colored by which covariance matrix they load most strongly on to. Do different regions of the genome have distinct patterns, or are they all jumbled up?

### 2020-12-03 R/Analysis_v0.5_snp_allele_frequencies_and_other_info.Rmd
Move R scripts not used in current manuscript and rename those that are

---
title: "Allele frequencies and positions"
author: "Alice MacQueen"
date: "11/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(cowplot)
#library(maps)
#library(ggmap)
#library(ggrepel)
#usa <- map_data("usa")
library(switchgrassGWAS)

workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")

inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
                                   "inputkinship.txt"),  delim = " ", 
                                   col_names = "SNPfiles")

metadata <- read_rds("~/Github/pvdiv-phenology-gxe/data/metadata.rds")
```

#
```{r}

metadata %>%
  group_by(GENETIC_SUBPOP_KINSHIP) %>%
  tally()
Atlantic <- metadata %>%
  filter(GENETIC_SUBPOP_KINSHIP == "Atlantic")
Gulf <- metadata %>%
  filter(GENETIC_SUBPOP_KINSHIP == "Gulf")
Midwest <- metadata %>%
  filter(GENETIC_SUBPOP_KINSHIP == "Midwest")

metadata %>%
  group_by(ECOTYPE_NNET) %>%
  tally()
Coastal <- metadata %>%
  filter(ECOTYPE_NNET == "Coastal")
Lowland <- metadata %>%
  filter(ECOTYPE_NNET == "Lowland")
Upland <- metadata %>%
  filter(ECOTYPE_NNET == "Upland")
metadata %>%
  filter(GENETIC_SUBPOP_KINSHIP %in% c("Atlantic", "Gulf", "Midwest")) %>%
  group_by(COLLECTION_TYPE) %>%
  tally()

# vcf.gz in SharedUser has PLANT_ID as the library names now.
writeLines(Atlantic$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Atlantic", "_subpop"))
writeLines(Gulf$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Gulf", "_subpop"))
writeLines(Midwest$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Midwest", "_subpop"))
writeLines(Coastal$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Coastal", "_ecotype"))
writeLines(Lowland$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Lowland", "_ecotype"))
writeLines(Upland$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Upland", "_ecotype"))
```

What are things I could color Manhattans by? For mash results?

Ideally, always for alternate alleles (aka not AP13's allele, in reality):

The covariance matrix that the SNP has the highest posterior probability for. Or second highest. Or whatever.

Their allele frequency in each subpopulation.

The average latitudinal and/or longitudinal position in our diversity panel. Or, as a factor, which state or region is this position in?

Which ecotype the allele is predominantly found in? Or is it common everywhere?

If I use vcftools to get allele frequencies, what is the kind of information that I get?
```{r}
allele_freq <- readRDS("~/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/Allele_frequencies_site_by_year_100000SNPs.rds")
allele_freq
```

Maybe big_count in bigsnpr could take care of this for me.

# get_mash_metadata
```{r}
inputdir <- file.path(workingdir, "analysis/gwas/BLUPs/single_sites", "Gulf_and_Midwest")
mash_out <- pvdiv_results_in_folder(inputdir, "Strong")
mash_out
m <- read_rds(file.path(inputdir, mash_out[12]))
markers <- switchgrassGWAS:::get_marker_df(m)
```

```{r}
snp <- snp_attach("~/Github/pvdiv-genome/genetic_subpops/Pvirgatum_V5_GWAS_229g_Gulf_K_subset_maf0.05.rds")
snp <- snp_attach("~/Github/pvdiv-genome/envGWAS/Pvirgatum_V5_GWAS_678g_33M_envGWAS_maf_0.03.rds")
#snp <- snp_attach(inputfiles$SNPfiles[1])
#
metadata_env <- get_mash_metadata(m, snp, suffix = "678g_envGWAS_m_greenup_2_phe_Gulf_subpop", saveoutput = TRUE)

snp <- snp_attach("~/Github/pvdiv-genome/microbiome/Pvirgatum_V5_GWAS_630g_33M_microbiome.rds")

metadata_630 <- get_mash_metadata(m, snp, suffix = "snp_630g_PKLE_m_greenup_2_phe_Gulf_subpop", saveoutput = TRUE)
metadata_363 <- get_mash_metadata(m, snp, suffix = "snp_363g_m_greenup_2_phe_Gulf_and_Midwest_subpop", saveoutput = TRUE)

```

```{r function to get metadata}

#' @title Get switchgrass metadata for SNPs in a mash object.
#'
#' @description Takes a mash object and a bigsnp object and returns
#'     a dataframe containing summary metadata for each SNP in the mash object.
#'     This information includes details for the alternate allele of each SNP
#'     such as its frequency in the bigsnp object, the average latitude, 
#'     longitude, elevation, and year that the SNP was sampled, and how 
#'     frequently individuals with the SNP have specific subpop and ecotype 
#'     identities in the bigsnp object.
#'     NB: This is VERY slow and best for up to tens of thousands of
#'     SNPs, NOT millions of SNPs.
#'
#' @param m A mash object (outputted by mash).
#' @param snp A bigsnp object 
#' @param metadata 
#' @param saveoutput Logical. Should the output be saved to the path?
#' @param suffix Character. Optional. A unique suffix used to save the files,
#'     instead of the current date & time.
get_mash_metadata <- function(m, snp, metadata, suffix = "", 
                              saveoutput = FALSE){
  markers <- switchgrassGWAS:::get_marker_df(m)
  filtercol <- which(snp$map$marker.ID %in% markers$Marker)
  if(str_sub(suffix, end = 1) %in% c("")){
      suffix <- paste0("_", switchgrassGWAS:::get_date_filename())
    }

  metadata1 <- metadata %>%
    filter(PLANT_ID %in% snp$fam$sample.ID) %>%
    mutate(Atlantic = ifelse(GENETIC_SUBPOP_KINSHIP == "Atlantic", 1, 0),
           Gulf = ifelse(GENETIC_SUBPOP_KINSHIP == "Gulf", 1, 0),
           Midwest = ifelse(GENETIC_SUBPOP_KINSHIP == "Midwest", 1, 0),
           Coastal = ifelse(ECOTYPE_NNET == "Coastal", 1, 0),
           Upland = ifelse(ECOTYPE_NNET == "Upland", 1, 0),
           Lowland = ifelse(ECOTYPE_NNET == "Lowland", 1, 0),
           COLL_YEAR = year(COLL_DATE))
  m1 <- metadata1 %>%
    mutate(SNP = snp$genotypes[,filtercol[1]]) %>%
    summarise(Alt_freq = mean(SNP/2, na.rm = TRUE),
              Alt_Latitude = weighted.mean(LATITUDE, SNP, na.rm = TRUE),
              Alt_Longitude = weighted.mean(-LONGITUDE, SNP, na.rm = TRUE),
              Alt_Elevation = weighted.mean(ELEVATION, SNP, na.rm = TRUE),
              Alt_Year = weighted.mean(COLL_YEAR, SNP, na.rm = TRUE),
              Alt_Coastal_d = weighted.mean(Coastal, SNP, na.rm = TRUE),
              Alt_Coastal_c = weighted.mean(COASTAL_ASSIGNMENT, SNP, 
                                            na.rm = TRUE),
              Alt_Lowland_d = weighted.mean(Lowland, SNP, na.rm = TRUE),
              Alt_Lowland_c = weighted.mean(LOWLAND_ASSIGNMENT, SNP, 
                                            na.rm = TRUE),
              Alt_Upland_d = weighted.mean(Upland, SNP, na.rm = TRUE),
              Alt_Upland_c = weighted.mean(UPLAND_ASSIGNMENT, SNP, na.rm = TRUE),
              Alt_Atlantic_d = weighted.mean(Atlantic, SNP, na.rm = TRUE),
              Alt_Atlantic_c = weighted.mean(ATLANTIC_Q, SNP, na.rm = TRUE),
              Alt_Gulf_d = weighted.mean(Gulf, SNP, na.rm = TRUE),
              Alt_Gulf_c = weighted.mean(GULF_Q, SNP, na.rm = TRUE),
              Alt_Midwest_d = weighted.mean(Midwest, SNP, na.rm = TRUE),
              Alt_Midwest_c = weighted.mean(MIDWEST_DA, SNP, na.rm = TRUE)
              ) %>%
    mutate(Marker = snp$map$marker.ID[1]) %>%
    select(Marker, everything())

  for(i in 2:length(filtercol)){
    m_i <- metadata1 %>%
    mutate(SNP = snp$genotypes[,filtercol[i]]) %>%
    summarise(Alt_freq = mean(SNP/2, na.rm = TRUE),
              Alt_Latitude = weighted.mean(LATITUDE, SNP, na.rm = TRUE),
              Alt_Longitude = weighted.mean(-LONGITUDE, SNP, na.rm = TRUE),
              Alt_Elevation = weighted.mean(ELEVATION, SNP, na.rm = TRUE),
              Alt_Year = weighted.mean(COLL_YEAR, SNP, na.rm = TRUE),
              Alt_Coastal_d = weighted.mean(Coastal, SNP, na.rm = TRUE),
              Alt_Coastal_c = weighted.mean(COASTAL_ASSIGNMENT, SNP, 
                                            na.rm = TRUE),
              Alt_Lowland_d = weighted.mean(Lowland, SNP, na.rm = TRUE),
              Alt_Lowland_c = weighted.mean(LOWLAND_ASSIGNMENT, SNP, 
                                            na.rm = TRUE),
              Alt_Upland_d = weighted.mean(Upland, SNP, na.rm = TRUE),
              Alt_Upland_c = weighted.mean(UPLAND_ASSIGNMENT, SNP, na.rm = TRUE),
              Alt_Atlantic_d = weighted.mean(Atlantic, SNP, na.rm = TRUE),
              Alt_Atlantic_c = weighted.mean(ATLANTIC_Q, SNP, na.rm = TRUE),
              Alt_Gulf_d = weighted.mean(Gulf, SNP, na.rm = TRUE),
              Alt_Gulf_c = weighted.mean(GULF_Q, SNP, na.rm = TRUE),
              Alt_Midwest_d = weighted.mean(Midwest, SNP, na.rm = TRUE),
              Alt_Midwest_c = weighted.mean(MIDWEST_DA, SNP, na.rm = TRUE)
              ) %>%
    mutate(Marker = snp$map$marker.ID[filtercol[i]]) %>%
    select(Marker, everything())
    m1 <- m1 %>% add_row(m_i)
    if(saveoutput == TRUE){
    if(i%%1000 == 0){
    write_rds(m1, paste0("~/Github/pvdiv-phenology-gxe/large-files/Metadata_associated_with_alternate_alleles_", suffix, ".rds"), compress = "gz")
    }
    }
  }
  if(saveoutput == TRUE){
    write_rds(m1, paste0("~/Github/pvdiv-phenology-gxe/large-files/Metadata_associated_with_alternate_alleles_", suffix, "_full.rds"), compress = "gz")
  }
}

((1000/(100))*25)/60
((12100000/(100))*25)/60/60/24 # 35 days to run for all 12M SNPs. Yikes.
((21697/(100))*25)/60/60  # 1.5 hours to run just for the Gulf mash SNP subset.
write_rds(m1, "~/Github/pvdiv-phenology-gxe/large-files/Metadata_associated_with_alternate_alleles_229g_Gulf_subpop_m_greenup_2_phe_full.rds", compress = "gz")

m1
```
Of the 363 Gulf & Midwest plants:
83.7 have ecotype assignment without the neural network.
67.7% have collection date
93.9% have lat & long; 89% have elevation

25 seconds to run 100 SNPs.
# Map this test Lat/Long
```{r}
ggplot() + 
  geom_polygon(data = usa, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  coord_quickmap(xlim = c(-105, -66), ylim = c(25, 50)) + 
  geom_point(data = m1, aes(x = SNP_Long, y = SNP_Lat), shape = 13, 
             size = 4, stroke = 1.5) + 
  labs(x = "Longitude", y = "Latitude")
```

### 2020-12-03 R/Analysis_v0.5_weather_covariance_Uhyp_flowering.Rmd
Move R scripts not used in current manuscript and rename those that are

---
title: "Weather_covariance_matrices_flowering"
author: "Alice MacQueen"
date: "11/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(mashr)
library(switchgrassGWAS)
library(rlist)
```
# TRIAL script
# -----------------------------
NB: Trial using covariance matrices derived from weather-related environmental cues as U_hyp in mash.

# Load data

```{r}
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")

site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
               NE = "NE", MI = "MI", SD = "SD")#, eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", 
                 Gulf = "229g_10.3M")
subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", 
                 Gulf = "Gulf")
phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
               "crain_1d", "GR50", "cgdd_12c_10d")
phe_mult <- list(flowering_5 = c("FL50", "dyln_fl50", "cgdd_12c_gr2fl",
                                 "crain_gr2fl", "crain_1d"), 
                 greenup_2 = c("GR50", "cgdd_12c_10d"))
inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
                                   "inputkinship.txt"),  delim = " ", 
                                   col_names = "SNPfiles")
```
# 1. Weather U_hyp

Make weather corr/covar matrices

1. Make dataframe of each env variable across all 8 sites
2. Compute correlations with pairwise complete observations
  2a. Compare correlation matrices - how dissimilar are they?
  2b. Compute correlations for Gulf, Midwest, & G&M subsets & compare
3. Experiment with changing within-site covariance to coefficient of variation for each site. Wouldn't expect these to be 1, really.

Four phenotypes to do this for, and three pop subsets.

### plot covariance matrices to look at their differences
```{r}
pvdiv_plot_cov <- function(covdf){
  U1 <- cor_phe
  if(isSymmetric(U1)){
        for(i in 1:nrow(U1)){
          for(j in 1:ncol(U1)){
            if(i < j){
              U1[i, j] <- NA
            }
          }
        }
  }
  U1 <- as_tibble(U1, rownames = "rowU", .name_repair = "unique") %>%
          pivot_longer(cols = -.data$rowU, names_to = "colU",
                       values_to = "covar") %>%
          filter(!is.na(.data$covar))
  U1$colU <- factor(U1$colU, levels = colnames(cor_phe))
  U1$rowU <- factor(U1$rowU, levels = colnames(cor_phe))

  U1_covar <- U1 %>%
    ggplot(aes(x = .data$rowU, y = .data$colU)) +
    switchgrassGWAS::theme_oeco +
    geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
    scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                    "white", "#27AD81FF", "#5DC863FF",
                                    "#FDE725FF"), limits = c(-1,1)) +
    geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
    theme(legend.position = c(0.2, 0.9),
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    xlab("") + ylab("") 
  return(U1_covar)
}
```

```{r}
U_hyp <- list()
ggcov <- list()
phe_hyp <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d")

for(i in seq_along(phe_hyp)){
  for(k in seq_along(subpop_v2)){

  cov_df <- phe_fl_df %>%
    ungroup() %>%
    left_join(select(sites, SITE, manu_site), by = "SITE") %>%
    filter(SUBPOP %in% subpop_v2[[k]]) %>%
    select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
    group_by(PLANT_ID, SUBPOP, manu_site) %>%
    mutate(IND = row_number()) %>%
    pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
    mutate(PLANT_ID = as.factor(PLANT_ID),
           IND = as.factor(IND)) %>%
    select(-IL) %>%
    select(PLANT_ID, SUBPOP, IND, MI, MO, NE, OK, SD, TX1, TX2, TX3)

  cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
  diag(cor_phe) <- cov_df %>%
    ungroup() %>%
    summarise(MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),
              MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
              NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
              OK = sqrt(var(OK, na.rm = TRUE))/mean(OK, na.rm = TRUE),
              SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
              TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
              TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
              TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE)) %>%
    as_vector()

  ggcov1 <- pvdiv_plot_cov(cor_phe) + ggtitle(label = paste0(phe_hyp[i], "_", names(subpop_v2)[k]))
  ggcov <- list.prepend(ggcov, ggcov1)
  U_hyp <- list.prepend(U_hyp, cor_phe)
  names(U_hyp)[1] <- paste0(phe_hyp[i], "_", names(subpop_v2)[k])
  }
}

ggcov
```

# 2. make mashr inputs for flowering 50 phenotype

FL50 for GWAS with 90 or more individuals. 8 gardens for Gulf & G&M. 3 gardens (TX2, MO, MI) for Midwest.

## Midwest NB
Some univariate GWAS for the Midwest had population structure issues. Mostly for greenup - even GR50 at MO was quite bad. Drop MO for GR50. Otherwise they look ok for FL50 and GR50.

## Set up mash inputs
```{r}
i=1 # site
j=1 # phe
k=3 # subpop: G&M, M, Gulf
for(k in c(1:3)){
phe_suffix <- paste("FL50", subpop_v[k], 
                    sep = "*.*")
snp <- snp_attach(inputfiles$SNPfiles[k])
gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
                                      pattern = paste0("GWAS_datatable*.*",
                                                       phe_suffix))
phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
    str_replace(., "_$", "")
phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
out1phe <- pvdiv_bigsnp2mashr(path = file.path(outputdir), 
                           snp = snp, gwas_rds = gwas_rds, 
                           phenotypes = phenotypes, numSNPs = numSNPs, 
                           clump = TRUE, scaled = TRUE, saveoutput = TRUE,
                           suffix = paste0("FL50_", names(subpop_v)[k]))
}
```

## Run mash
```{r}
for(k in c(1:3)){
m_out <- mash_standard_run(path = outputdir, numSNPs = numSNPs, 
                           U_hyp = U_hyp,
                           suffix = paste0("FL50_", names(subpop_v)[k]),
                           saveoutput = TRUE)
  }
```

# 3. Analyse mash output

### 2020-12-03 R/Plots_v0.2_manuscript_figures.Rmd
Move R scripts not used in current manuscript and rename those that are

--- 2020-07-22 R/Plots_manuscript_v0.2.Rmd
@@
 
-for(p in seq_along(phe_dfs)){
+for(p in 2){#seq_along(phe_dfs)){
   phe_v <- phe_lists[[p]]
@@
 
-  for(i in 6:9){
+  for(i in seq_along(site_v)){
     for(j in seq_along(subpop_v)){
@@
                        random = ~vm(PLANT_ID, k_full) +
-                         ~idv(SITE),
+                         ~idv(SITE):vm(PLANT_ID, k_full),
                        residual = ~idv(units),
@@
                                                   "_related_", 
-                                                  "functions_of_weather_v2.csv")))
+                                                  "functions_of_weather_v3.csv")))
 }

### 2020-12-03 R/Plots_v0.3_G_BLUP.Rmd
Move R scripts not used in current manuscript and rename those that are

--- 2020-10-02 R/Plots_manuscript_G_BLUP.Rmd
@@
 
+# Total # 20kb regions
+Midwest: 8824339 SNPs       52797 20kb regions
+Gulf: 10320201 SNPs         53025 20kb regions
+Gulf_and_Midwest: 12083955  53086 20kb regions
+53240 20kb regions total
 ```{r}
+phe_v <- c("dyln_fl50", "cgdd_12c_gr2fl", "FL50")
+site_v <- c("eight_sites", "tx_sites", "north_sites")
+i=1
+j=1
+k=1
 
+  gwas_files <- list.files(path = paste0("./", site_v[j]), pattern = paste0("GWAS_datatable_", phe_v[i]))
+
+  gwas_files
+
+  gwas_20kb <- tibble()
+subpop_v <- c("Midwest", "Gulf", "Gulf_and_Midwest")
+
+for(k in seq_along(gwas_files)){
+  gwas1 <- readRDS(file.path(site_v[j], gwas_files[k]))
+gwas1 <- gwas1 %>%
+  mutate(Pos_Mb = round(POS/10000)/100,
+         POS_bin = round(Pos_Mb*100/2)) %>%
+  group_by(CHR, POS_bin) %>% 
+  tally() %>%
+  mutate(subpop = subpop_v[k])
+gwas_20kb <- bind_rows(gwas_20kb, gwas1)
+}
 ```
+
+# Overlaps phe across subpops
+
+Associations did or did not overlap significantly in the same phenotypes among subpopulations? What about in the same subpop between sets of sites? This is GxE at the continental scale, point out.
+In both Gulf & Midwest:
+All  dyln_fl50  Chr01N	229	2	   1 region 1/463     NS 0.08496942
+All cgdd_12c_gr2fl  0 regions   0/2
+All FL50  0 regions   0/18
+North  dyln_fl50 39 regions  39/2530      sig 0.00688
+North cgdd_12c_gr2fl  0 regions  0/59
+North  FL50    6 regions 5Nx2 6Kx2 8N 9K   6/1097     NS 0.141
+Texas  dyln_fl50         0 regions  0/232
+Texas  cgdd_12c_gr2fl    0 regions  0/190
+Texas  FL50             0 regions   0/64
+```{r}
+sitev <- "eight_sites"
+phev <- "dyln_fl50"
+popv <- list(Gulf = 53025, Midwest = 52797, Gulf_and_Midwest = 53086)
+i=2
+
+AandB <- gwas_fdr %>%
+  filter(subpop %in% names(popv)[c(1,2)] & phe %in% phev & sites %in% sitev) %>%
+    group_by(sites, subpop, phe, CHR, POS_bin) %>%
+    tally() %>%
+    group_by(CHR, POS_bin) %>%
+    summarise(count = n()) %>%
+    filter(count > 1)
+inB <- gwas_fdr %>%
+  filter(subpop %in% names(popv)[1] & phe %in% phev & sites %in% sitev) %>%
+    group_by(sites, subpop, phe, CHR, POS_bin) %>%
+    tally() %>%
+    group_by(CHR, POS_bin) %>%
+    summarise(count = n()) 
+notinB <- popv[[i]] - nrow(inB)
+inA <- gwas_fdr %>%
+  filter(subpop %in% names(popv)[2] & phe %in% phev & sites %in% sitev) %>%
+    group_by(sites, subpop, phe, CHR, POS_bin) %>%
+    tally() %>%
+    group_by(CHR, POS_bin) %>%
+    summarise(count = n())
+
+phyper(nrow(AandB), nrow(inB), notinB, nrow(inA), lower.tail = FALSE)
+```
+
+Eight & Texas:
+In Gulf
+dyln_fl50  22/274
+FL50
+cgdd_12c_gr2fl 0/2
+
+## Hypergeometric test for significance of overlaps
+phyper(10, 50, 200 - 50, 25, lower.tail = FALSE)
+phyper(A&B, inB, notinB, inA, lower.tail = FALSE)
+```{r}
+A <- "north_sites"
+B <- "tx_sites"
+phev <- "cgdd_12c_gr2fl"
+popv <- list(Gulf = 53025, Midwest = 52797, Gulf_and_Midwest = 53086)
+i=2
+
+AandB <- gwas_fdr %>%
+  filter(subpop %in% names(popv)[i] & phe %in% phev & sites %in% c(A,B)) %>%
+    group_by(sites, subpop, phe, CHR, POS_bin) %>%
+    tally() %>%
+    group_by(CHR, POS_bin) %>%
+    summarise(count = n()) %>%
+    filter(count > 1)
+inB <- gwas_fdr %>%
+  filter(subpop %in% names(popv)[i] & phe %in% phev & sites %in% c(B)) %>%
+    group_by(sites, subpop, phe, CHR, POS_bin) %>%
+    tally() %>%
+    group_by(CHR, POS_bin) %>%
+    summarise(count = n()) 
+notinB <- popv[[i]] - nrow(inB)
+inA <- gwas_fdr %>%
+  filter(subpop %in% names(popv)[i] & phe %in% phev & sites %in% c(A)) %>%
+    group_by(sites, subpop, phe, CHR, POS_bin) %>%
+    tally() %>%
+    group_by(CHR, POS_bin) %>%
+    summarise(count = n())
+
+phyper(nrow(AandB), nrow(inB), notinB, nrow(inA), lower.tail = FALSE)
+8/232
+```
+tx & north dyln Gulf 0.0078 (8 overlaps)
+eight & north cgdd Gulf 1.13e-06
+eight & north dyln Gulf 5.36e-31 (25 overlaps)
+eight & tx dyln Gulf 2.2e-38 (22 overlaps)
+
+eight & north dyln Midwest 4.42e-242
+Very few overlaps between site sets for the Midwest subpopulation.
+Lots of overlaps & sig for Gulf_and_Midwest, if we trust this.
 
@@
 ft_annos %>%
-  filter(seqid %in% qtldf$CHR[1] & between(Pos_Mb, qtldf$Pos_Mb_lo[1], qtldf$Pos_Mb_hi[1]))
+  filter(seqid %in% gtldf3$CHR[1] & between(Pos_Mb, gtldf3$Pos_Mb_lo[1], gtldf3$Pos_Mb_hi[1]))
 
-qtldf
-i=24
+gtldf3 <- qtldf %>% filter(lod > 30 & lodcolumn == "dyln_fl50")
+i=4
 ft_annos %>%
-  filter(seqid %in% qtldf$CHR[i] & between(Pos_Mb, qtldf$Pos_Mb_lo[i], qtldf$Pos_Mb_hi[i])) %>%
-  select(OsGene, AtGene, everything())
+  filter(seqid %in% gtldf3$CHR[i] & between(Pos_Mb, gtldf3$Pos_Mb_lo[i], gtldf3$Pos_Mb_hi[i])) %>%
+  select(OsGene, AtGene, everything()) %>%
+  left_join(osgene_ft_key, by = c("OsGene" = "Symbol"))
 ```
+## start here QTL candidate genes
+Chr02N	57.23	58.81	   SIP1|OsSUF4	SUF4
+SIP1 participates in regulation of flowering time in rice by recruiting OsTrx1 to Ehd1.
+
+Chr04K	4.23	5.50     6 genes including Hd3a	FT
+	Hd3a, a rice ortholog of the Arabidopsis FT gene, promotes transition to flowering downstream of Hd1 under short-day conditions
+	Phytochrome dependent quantitative control of Hd3a transcription is the basis of the night break effect in rice flowering
+	Variations in Hd1 proteins, Hd3a promoters, and Ehd1 expression levels contribute to diversity of flowering time in cultivated rice
+
+Chr04K	9.92	13.93    3 genes including Hd3a	FT ( a different copy)
+Hd3a, a rice ortholog of the Arabidopsis FT gene, promotes transition to flowering downstream of Hd1 under short-day conditions
+	Phytochrome dependent quantitative control of Hd3a transcription is the basis of the night break effect in rice flowering
+	OsMFT1 increases spikelets per panicle and delays heading date in rice by suppressing Ehd1, FZP and SEPALLATA-like genes.
+
+Chr05N	2.04	3.44     5 genes including FT and OsFTL1	FT &&&&& OSMADS3:
+Genetic Enhancer Analysis Reveals that FLORAL ORGAN NUMBER2 and OsMADS3 Cooperatively Regulate Maintenance and Determinacy of the Flower Meristem in Rice.
+
+Chr05N	64.27	65.98	  3 genes including OsFTL10 and OsABF1|OsABI5|OREB1|OsbZIP10: 	A Drought-inducible bZIP Transcription Factor OsABF1 Delays Reproductive Timing in Rice.
 

### 2020-12-03 R/Pre_manuscript/1_bigsnpr_gwas_prep.Rmd
Move R scripts not used in current manuscript and rename those that are

--- 2020-01-08 analysis/1_bigsnpr_gwas_prep.Rmd
@@
 
+## Visualize mash effects
+
+```{r}
+mash_output <- readRDS("FL50/Strong_Effects29475SNPs.rds")
+```
+
+```{r}
+i=10
+get_significant_results(mash_output)[i]
+plot1 <- switchgrassGWAS::mash_plot_effects(m = mash_output, n=i)
+
+xlabs <- str_sub(plot1$effect_df$value, start = 11, end = -6)
+plot1$effect_df$SITE <- xlabs
+
+#plot1$effect_df$SITE <- factor(plot1$effect_df$SITE, levels = c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING"))
+plot1$effect_df$SITE  <- factor(plot1$effect_df$SITE , levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
+
+plot1$ggobject + 
+  theme(axis.text.x = element_text(angle = 45, vjust = 1)) + 
+  scale_x_discrete(labels = xlabs)
+
+ggplot(data = plot1$effect_df) +
+    geom_point(mapping = aes(x = .data$SITE, y = .data$mn)) +
+    geom_errorbar(mapping = aes(ymin = .data$mn - .data$se,
+                                ymax = .data$mn + .data$se,
+                                x = .data$SITE), width = 0.3) +
+    geom_hline(yintercept = 0, lty = 2) +
+    labs(x = "Conditions", y = "Effect Size") +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  coord_flip()
+
+save_plot(paste0(names(get_significant_results(mash_output)[i]), "_Effect_Plot.png"), plot = last_plot(), base_height = 4, base_asp = 0.8)
+
+```
+
 ## Visualize bigsnpr GWAS

### 2020-12-03 R/Pre_manuscript/4_kmer_and_PAV_vcf_conversion.Rmd
Move R scripts not used in current manuscript and rename those that are

---
title: "4_kmer_and_PAV_vcf_conversion"
author: "Alice MacQueen"
date: "3/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##

wget --user=switchgrass --ask-password http://hagsc.org/restricted_access/switchgrass/newKmerPAVMatrix.Nov25.19.0.005MAF.wHeader.vcf.gz

plink --vcf newKmerPAVMatrix.Nov25.19.0.005MAF.wHeader.vcf.gz --const-fid Pvirgatum --allow-extra-chr --make-bed --out newKmerPAVMatrix.Nov25.19.0.005MAF

# Alter fam and nosex files so they have PLANT_ID not LIBRARY
```{r}
metadata <- read_csv("~/Github/pvdiv-phenology-gxe/data/metadata_hierarchical_kof3.csv")

kmerfam <- read_delim("~/Github/pvdiv-kmer-gwas/data-raw/Pvirgatum_gene_PAV_2019_08_16.fam", delim = " ", col_names = c("FID", "LIBRARY", "X3", "X4", "X5", "X6")) %>%
  left_join(metadata) %>%
  dplyr::select(FID, PLANT_ID, X3:X6, LIBRARY) %>%
  #filter(!is.na(PLANT_ID))
  mutate(PLANT_ID = coalesce(PLANT_ID, LIBRARY)) %>%
  dplyr::select(FID, PLANT_ID, X3:X6)
write_delim(kmerfam, path = "~/Github/pvdiv-kmer-gwas/data-raw/Pvirgatum_gene_PAV_2019_08_16.fam", delim = " ", col_names = FALSE)

kmernosex <- read_delim("~/Github/pvdiv-kmer-gwas/data-raw/Pvirgatum_gene_PAV_2019_08_16.nosex", delim = "\t", col_names = c("FID", "LIBRARY")) %>%
  left_join(metadata) %>%
  dplyr::select(FID, PLANT_ID, LIBRARY) %>%
  #filter(!is.na(PLANT_ID))
  mutate(PLANT_ID = coalesce(PLANT_ID, LIBRARY)) %>%
  dplyr::select(FID, PLANT_ID)
write_delim(kmernosex, path = "~/Github/pvdiv-kmer-gwas/data-raw/Pvirgatum_gene_PAV_2019_08_16.nosex", delim = "\t", col_names = FALSE)
```

plink --vcf Pvirgatum_gene_PAV_2019_08_16.vcf --const-fid Pvirgatum --allow-extra-chr --make-bed --out Pvirgatum_gene_PAV_2019_08_16

```{r}
pavfam <- read_delim("~/Github/pvdiv-kmer-gwas/data-raw/newKmerPAVMatrix.Nov25.19.0.005MAF.fam", delim = " ", col_names = c("FID", "LIBRARY", "X3", "X4", "X5", "X6")) %>%
  left_join(metadata) %>%
  dplyr::select(FID, PLANT_ID, X3:X6, LIBRARY) %>%
  #filter(!is.na(PLANT_ID))
  mutate(PLANT_ID = coalesce(PLANT_ID, LIBRARY)) %>%
  dplyr::select(FID, PLANT_ID, X3:X6)
write_delim(pavfam, path = "~/Github/pvdiv-kmer-gwas/data-raw/newKmerPAVMatrix.Nov25.19.0.005MAF.fam", delim = " ", col_names = FALSE)

pavnosex <- read_delim("~/Github/pvdiv-kmer-gwas/data-raw/newKmerPAVMatrix.Nov25.19.0.005MAF.nosex", delim = "\t", col_names = c("FID", "LIBRARY")) %>%
  left_join(metadata) %>%
  dplyr::select(FID, PLANT_ID, LIBRARY) %>%
  #filter(!is.na(PLANT_ID))
  mutate(PLANT_ID = coalesce(PLANT_ID, LIBRARY)) %>%
  dplyr::select(FID, PLANT_ID)
write_delim(pavnosex, path = "~/Github/pvdiv-kmer-gwas/data-raw/newKmerPAVMatrix.Nov25.19.0.005MAF.nosex", delim = "\t", col_names = FALSE)
```

# Make bigsnp file and subset it to plants in ind_gwas

Have to do this from the working directory of the bed file. Can't point to it, even with an absolute path. OR you'll get an error like: `Error in readbina(bedfile, bigGeno, getCode()) :`
  `Wrong magic number. Aborting..`

```{r}
rds <- snp_readBed("Pvirgatum_gene_PAV_2019_08_16.bed")

ind_gwas <- metadata %>%
  filter(LIB_GROWN == "Y" | LIB_BIOCLIM == "Y")

kmersubset <- which(kmerfam$PLANT_ID %in% ind_gwas$PLANT_ID)
pavsubset <- which(pavfam$PLANT_ID %in% ind_gwas$PLANT_ID)
saveRDS(kmersubset, file = "../data/individuals_subset.rds")

which(kmersubset %in% pavsubset)
```

```{r}
pvdiv_pop_subset_bigsnp <- function(subset, selector){
  subset_name <- paste0("Pvirgatum_4x_784g_imp_phased_", subset, "_subset")
  snpfile_subset <- subset(snp, ind.row = selector)
  snp_subset <- snp_attach(snpfile_subset)
  snp_writeBed(snp_subset, bedfile = paste0(subset_name, ".bed"))
  mac <- round(25/length(selector), digits = 2) # what should the MAF be?
  if(mac > 0.05){
    mac <- 0.05
  }
  plink <- download_plink()
  bedfile <- snp_plinkQC(plink, prefix.in = subset_name,
                         maf = mac, geno = 0.1, hwe = 1e-50, 
                         prefix.out = paste0(subset_name, "_maf", mac),
                         extra.options = "--allow-no-sex --allow-extra-chr --chr Chr01K Chr01N Chr02K Chr02N Chr03K Chr03N Chr04K Chr04N Chr05K Chr05N Chr06K Chr06N Chr07K Chr07N Chr08K Chr08N Chr09K Chr09N")
  rdsfile <- snp_readBed(bedfile)
}

```

### 2020-12-03 R/Pre_manuscript/D2F_breakdown_by_site.Rmd
Move R scripts not used in current manuscript and rename those that are

--- 2020-04-24 R/D2F_breakdown_by_site.Rmd
@@
 phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Phenotypes_D2F_fourway_project.rds")
+metadata <- readRDS("~/Github/pvdiv-phenotypes/data/metadata.rds")
 metadata <- metadata %>%
@@
                             TRUE ~ NA_character_))
+sites <- readRDS("~/Github/pvdiv-phenotypes/data/sites.rds")
 ```

### 2020-12-03 R/Pre_manuscript/Draft_models_and_stats_2020-06-04.Rmd
Move R scripts not used in current manuscript and rename those that are

---
title: "Draft models and statistics"
author: "Alice MacQueen"
date: "6/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(switchgrassGWAS)
library(corpcor)
library(viridis)
```

# ASReml-R analyses on server

```{r}
library(tidyverse)
library(asreml)
library(switchgrassGWAS)
library(corpcor)
```

```{r}
metadata <- readRDS("~/Github/pvdiv-phenotypes/data/metadata.rds")
metadata <- metadata %>%
  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
                            !is.na(PLOIDY) ~ as.character(PLOIDY),
                            TRUE ~ NA_character_),
         LATITUDE = abs(LATITUDE),
         LONGITUDE = abs(LONGITUDE)) 
phe_all <- read_rds("~/Github/pvdiv-phenology-gxe/data/all_phenotype_meas_4wcr_gwas.rds")
sites <- readRDS("~/Github/pvdiv-phenotypes/data/sites.rds")
phe_all$SUBPOP <- factor(phe_all$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
sites <- sites %>%
  mutate(manu_site = case_when(State != "TX" ~ State,
                               SITE == "KING" ~ "TX1",
                               SITE == "PKLE" ~ "TX2",
                               SITE == "TMPL" ~ "TX3",
                               SITE == "OVTN" ~ "TX4"))
sites$manu_site <- factor(sites$manu_site, levels = c("TX1", "TX2", "TX3", "TX4", "OK", "MO", "NE", "IL", "MI", "SD"))

subpops <- enframe(unique(metadata$SUBPOP), name = NULL, value = "SUBPOP") %>%
  mutate(color_code = case_when(SUBPOP == "Atlantic" ~ "#6E91CB",
                                SUBPOP == "Gulf" ~ "#F47F72",    
                                SUBPOP == "Midwest" ~ "#442C83",
                                SUBPOP == "4X" ~ "grey", 
                                SUBPOP == "8X" ~ "#090906"))
subpops$SUBPOP <- factor(subpops$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))

# gulf: #F47F72, atlantic: #6E91CB, midwest: #442C83
```

```{r}
sites %>% arrange(Latitude)
```

We observed a strong signal of latitude of origin on whether plants grown at Texas common gardens flowered in lengthening or shortening days (generalized linear model on binomial output, p-value, Supp. Figure 1). 
```{r}
lat_photo_cue <- phe_all %>%
  filter(!is.na(LATITUDE) & LATITUDE > 0 & WHERE == "GWAS" & PHE == "FL50" & SITE %in% c("PKLE", "KING", "TMPL")) %>%
  mutate(Photo_cue = case_when(MEAS > 173 ~ 1,
                               MEAS <= 172 ~ 0,
                               TRUE ~ NA_real_))
lat_photo_cue$SUBPOP <- factor(lat_photo_cue$SUBPOP, levels = c("Midwest", "Gulf", "Atlantic", "4X", "8X"))

lat1 <- glm(Photo_cue ~ LATITUDE:SUBPOP + SUBPOP, data = lat_photo_cue, family = "binomial")

summary(lat1)

#lat2 <- glm(Photo_cue ~ LATITUDE*SUBPOP, data = lat_photo_cue, family = "binomial")
#summary(lat2)  # Think the first model is better. AIC 4610 vs 4604.5. 
#
#

# Create some new data
n2 <- 100
new.df <- data.frame(LATITUDE = 26:45, SUBPOP = rep(c("Midwest", "Gulf", "Atlantic", "4X", "8X"), length(25:45)/5))
new.df <- new.df %>% complete(LATITUDE, SUBPOP)

# Predict the probabilities
# predict(model, new.df) will just give you the logit (z) values
# We want the probabilities so use the "response" type
new.df$probs <- predict(lat1, new.df, "response")

# Draw some random values between 0 and 1
draws <- runif(n2)

# Check if draw value is less than threshold probability
new.df$results <- (draws < probs)

new.df %>%
  filter((SUBPOP == "Midwest" & between(LATITUDE, 30, 46)) | 
           !(SUBPOP %in% c("Midwest", "Gulf")) | 
          (SUBPOP == "Gulf" & between(LATITUDE, 25, 43)) ) %>%
  ggplot(aes(x= SUBPOP, y = LATITUDE)) + 
  geom_tile(aes(fill = probs)) +
  scale_fill_viridis(option = "B")

lat_photo_cue <- lat_photo_cue %>%
  ungroup() %>%
  mutate(predphoto = predict(lat1, ., "response"),
         residphoto = resid(lat1))

color_code <- c("#442C83", "#F47F72", "#6E91CB", "grey", "#090906")

lat_photo_cue %>%
  ggplot(aes(x = LATITUDE, y = predphoto)) + 
  geom_point(aes(color = SUBPOP), alpha = 0.4) + 
  scale_color_manual(values = color_code) + 
  geom_hline(yintercept = 0, linetype = 2) +
  geom_smooth(aes(color = SUBPOP)) 
lat_photo_cue %>%
  ggplot(aes(x = LATITUDE, y = residphoto)) + 
  geom_point(aes(color = SUBPOP), alpha = 0.4) + 
  scale_color_viridis(discrete = TRUE, option = "B") + 
  geom_hline(yintercept = 0, linetype = 2) +
  geom_smooth(aes(color = SUBPOP))  + 
  facet_wrap(~SUBPOP)

photo_cue_prob <- lat_photo_cue %>%
  group_by(PLANT_ID) %>%
  summarise(photo_prob = mean(predphoto))
```

# Set up weather data for 2019
```{r}
ptt_df <- readRDS(file = "~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")

wea_df <- ptt_df %>%
  dplyr::select(DOY:GDD_12C_SM, PTT_12C)

GR_start <- phe_all %>% 
  filter(WHERE == "GWAS") %>%
  ungroup() %>%
  group_by(SITE, YEAR, PLANT_ID, PLOT_GL) %>%
  filter(PHE == "GR50") %>%
  summarise(GR_start = floor(min(MEAS)))
FL_end <- phe_all %>% 
  filter(WHERE == "GWAS") %>%
  ungroup() %>%
  group_by(SITE, YEAR, PLANT_ID, PLOT_GL) %>%
  filter(PHE == "FL50") %>%
  summarise(FL_end = ceiling(max(MEAS)))

wea_gr_fl <- wea_df %>%
  left_join(GR_start) %>%
  left_join(FL_end) %>%
  group_by(Day_of_year, SITE, YEAR, PLANT_ID, PLOT_GL) %>%
  filter(between(Day_of_year, GR_start, FL_end)) %>% 
  ungroup() # Just the weather data between the minimum GR50 and max FL50.
# wea_df has all the weather, wea_gr_fl has the weather for the subset between
# greenup and flowering.
# 
gdd_12c_cue <- wea_gr_fl %>%
  group_by(SITE, YEAR, PLANT_ID, PLOT_GL, FL_end) %>%
  summarise(CGDD_13C = sum(GDD_13C_SM),
            CGDD_12C = sum(GDD_12C_SM),
            CPTT_12C = sum(PTT_12C)) %>%
  rename(FL50 = FL_end) %>%
  left_join(metadata)

```

# daylength cue

Additionally, we observed no evidence for any specific day length that triggered flowering (model: DYLN at FL50 ~ SUBPOP + PLANT_ID + SUBPOP:LATITUDE does not explain much variance). 
```{r}
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
dyln_cue <- wea_gr_fl %>%
  filter(Day_of_year == FL_end) %>%
  left_join(select(metadata, PLANT_ID, SUBPOP, LATITUDE)) %>%
  mutate(Photo_cue = case_when(FL_end > 173 ~ 1,
                               FL_end <= 172 ~ 0,
                               TRUE ~ NA_real_)) %>%
  filter(Photo_cue == 1) %>%
  filter(!is.na(SUBPOP) & !is.na(LATITUDE)) %>%
  filter(PLANT_ID %in% row.names(k_full)) %>%
  mutate(PLANT_ID = as_factor(PLANT_ID),
         SITE = as_factor(SITE)) %>%
  as.data.frame()
dyln_cue$SUBPOP <- factor(dyln_cue$SUBPOP, levels = c("Midwest", "Gulf", "Atlantic", "4X", "8X"))

asr_dyln <- asreml(DYLN ~ 1, random = 
                     ~idv(SITE) + 
                    ~vm(PLANT_ID, k_full),
                  residual = ~idv(units), data = dyln_cue, 
                  workspace = "3gb")

summary(asr_dyln)$varcomp %>%
  as_tibble(rownames = "Effect") %>%
  mutate(`%Var` = (component*100)/sum(component),
         `VarSE` = (std.error*100)/sum(component),
         lowVar = `%Var` - VarSE,
         hiVar = `%Var` + VarSE,
         loglik = asr_dyln$loglik) %>%
  write_csv(., "manuscript/plots/Variance_components_dyln_fl50_by_site_and_k_full.csv",
            col_names = TRUE)

dyln_cue$predicted <- asr_dyln$linear.predictors
dyln_cue %>% 
  as_tibble() %>%
  group_by(SUBPOP) %>%
  summarise(meanpred = mean(predicted),
            fivepred = quantile(predicted, 0.025),
            ninefivepred = quantile(predicted, 0.975),
            fiveobs = quantile(DYLN, 0.025),
            ninefiveobs = quantile(DYLN, 0.975))

blup <- summary(asr_dyln, coef=TRUE)$coef.random %>% as_tibble(rownames = "Effect")
int <- summary(asr_dyln, coef=TRUE)$coef.fixed %>% as_tibble(rownames = "Effect") %>% rename(std.error = `std error`)

write_csv(int, "manuscript/plots/Effects_dyln_fl50_by_site_and_k_full.csv", col_names = TRUE)
write_csv(blup, "manuscript/plots/Effects_dyln_fl50_by_site_and_k_full.csv", col_names = TRUE, append = TRUE)
```

To identify an additional environmental cue for flowering, we evaluated flowering date as a function of cumulative GDD between greenup and flowering at the northern seven sites (Kiniry et al 2005, Behrman 2013). A model with effects of subpopulation and an interaction between subpopulation and latitude of origin (linear model table) explained 55% of the variation in flowering GDD. The Midwest required the smallest GDD before flowering (585 +/- 12 GDD), and the Gulf required the largest before flowering (1238 +/- 378 GDD), with the other three subpopulations falling in between (708 – 750 +/- 106-123 GDD). 
```{r}
gdd_12c_7site <- filter(gdd_12c_cue, !(SITE %in% c("PKLE", "TMPL", "KING", "OVTN"))) %>%
  dplyr::select(-(LIBRARY:PLANTED_2018), LATITUDE) %>%
  filter(!is.na(SUBPOP) & !is.na(LATITUDE)) %>%
  left_join(photo_cue_prob)
```

```{r}

# gdd3 is the best model
gdd_12c_cue$SUBPOP <- factor(gdd_12c_cue$SUBPOP, levels = c("Midwest", "Gulf", "Atlantic", "4X", "8X"))
gdd1 <- lm(CGDD_12C ~ SUBPOP+LATITUDE:SUBPOP, data = filter(gdd_12c_cue, !(SITE %in% c("OVTN"))))
summary(gdd1)
gdd9site <- filter(gdd_12c_cue, !(SITE %in% c("OVTN")) & !is.na(LATITUDE) &
                     !is.na(SUBPOP)) %>%
  ungroup() %>%
  mutate(pred9 = predict(gdd1, .),
         resid9 = resid(gdd1))
gdd9site %>% 
  dplyr::select(SUBPOP:resid9, everything()) %>%
  ggplot(aes(x = LATITUDE, y = resid9)) + 
  geom_point(aes(color = SUBPOP), alpha = 0.4) + 
  scale_color_viridis(discrete = TRUE, option = "B") + 
  geom_hline(yintercept = 0, linetype = 2) +
  geom_smooth(aes(color = SUBPOP)) + 
  facet_wrap(~SUBPOP)
gdd9site %>% 
  dplyr::select(SUBPOP:resid9, everything()) %>%
  ggplot(aes(x = pred9, y = resid9)) + 
  geom_jitter(aes(color = SUBPOP), alpha = 0.4) + 
  scale_color_viridis(discrete = TRUE, option = "B") + 
  geom_hline(yintercept = 0, linetype = 2) +
  geom_smooth(aes(color = SUBPOP)) + 
  facet_wrap(~SUBPOP)
```
# ASReml
```{r}
gdd_12c_7site <- filter(gdd_12c_cue, !(SITE %in% c("OVTN"))) %>%
  left_join(select(sites, SITE:Latitude)) %>%
  left_join(photo_cue_prob) %>%
  dplyr::select(-(LIBRARY:PLANTED_2018), LATITUDE) %>%
  filter(!is.na(SUBPOP) & !is.na(LATITUDE)) %>%
  filter(!(photo_prob > 0.5 & SITE %in% c("KING", "TMPL", "PKLE", "STIL"))) %>%
    as.data.frame()

gdd_12c_7site$SUBPOP <- factor(gdd_12c_7site$SUBPOP, levels = c("Midwest", "Gulf", "Atlantic", "4X", "8X"))

asr_gdd <- asreml(CGDD_12C ~ 1, random = ~idv(SUBPOP) +
                    ~id(LATITUDE):idv(SUBPOP),
                  residual = ~idv(units), data = gdd_12c_7site, 
                  workspace = "3gb")

summary(asr_gdd)$varcomp %>%
  as_tibble(rownames = "Effect") %>%
  mutate(`%Var` = (component*100)/sum(component),
         `VarSE` = (std.error*100)/sum(component),
         lowVar = `%Var` - VarSE,
         hiVar = `%Var` + VarSE,
         loglik = asr_gdd$loglik) %>%
  write_csv(., "manuscript/plots/Variance_components_CGDD_12C_Subpop_LatofOrigin_filteredphoto_cue.csv",
            col_names = TRUE)

gdd_12c_7site$predicted <- asr_gdd$linear.predictors
gdd_12c_7site %>% 
  as_tibble() %>%
  group_by(SUBPOP) %>%
  summarise(meanpred = mean(predicted),
            fivepred = quantile(predicted, 0.025),
            ninefivepred = quantile(predicted, 0.975),
            fiveobs = quantile(CGDD_12C, 0.025),
            ninefiveobs = quantile(CGDD_12C, 0.975))

 blup <- summary(asr_gdd, coef=TRUE)$coef.random %>% as_tibble(rownames = "Effect")
int <- summary(asr_gdd, coef=TRUE)$coef.fixed %>% as_tibble(rownames = "Effect") %>% rename(std.error = `std error`)

write_csv(int, "manuscript/plots/Effects_CGDD_12C_Subpop_LatofOrigin_filteredphoto_cue.csv", col_names = TRUE)
write_csv(blup, "manuscript/plots/Effects_CGDD_12C_Subpop_LatofOrigin_filteredphoto_cue.csv", col_names = TRUE, append = TRUE)

gdd_12c_7site %>%
   as_tibble() %>%
   filter(SUBPOP == "Midwest") %>% 
  filter(LATITUDE > quantile(LATITUDE, 0.8)) %>%
   summarise(meanpred = mean(predicted))
gdd_12c_7site %>%
   as_tibble() %>%
   filter(SUBPOP == "Midwest") %>%
  filter(LATITUDE < quantile(LATITUDE, 0.2) & CGDD_12C < 1095) %>%
   summarise(meanpred = mean(predicted))

gdd_12c_7site %>%
   as_tibble() %>%
   filter(SUBPOP == "Gulf") %>% 
  filter(LATITUDE > quantile(LATITUDE, 0.8)) %>%
   summarise(meanpred = mean(predicted))
gdd_12c_7site %>%
   as_tibble() %>%
   filter(SUBPOP == "Gulf") %>%
  filter(LATITUDE < quantile(LATITUDE, 0.2)) %>%
   summarise(meanpred = mean(predicted))

```

# K GDD ASReml
```{r}
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
gdd_12c_h2 <- filter(gdd_12c_cue, !(SITE %in% c("OVTN"))) %>%
  left_join(select(sites, SITE:Latitude)) %>%
  left_join(photo_cue_prob) %>%
  dplyr::select(-(LIBRARY:PLANTED_2018), LATITUDE) %>%
  filter(!is.na(SUBPOP) & !is.na(LATITUDE)) %>%
  filter(PLANT_ID %in% row.names(k_full)) %>%
  filter(!(photo_prob > 0.5 & SITE %in% c("KING", "TMPL", "PKLE", "STIL"))) %>%
  ungroup() %>%
  mutate(PLANT_ID = as_factor(PLANT_ID),
         SITE = as_factor(SITE)) %>%
    as.data.frame()

gdd_12c_h2$SUBPOP <- factor(gdd_12c_h2$SUBPOP, levels = c("Midwest", "Gulf", "Atlantic", "4X", "8X"))

asr_gddh2 <- asreml(CGDD_12C ~ 1, random = ~idv(SITE) + ~vm(PLANT_ID, k_full),
                  residual = ~idv(units), data = gdd_12c_h2, 
                  workspace = "3gb")

summary(asr_gddh2)$varcomp %>%
  as_tibble(rownames = "Effect") %>%
  mutate(`%Var` = (component*100)/sum(component),
         `VarSE` = (std.error*100)/sum(component),
         lowVar = `%Var` - VarSE,
         hiVar = `%Var` + VarSE,
         loglik = asr_gddh2$loglik) %>%
  write_csv(., "manuscript/plots/Variance_components_gdd12c_fl50_by_site_and_k_full.csv",
            col_names = TRUE)
```

# K flowering date ASReml
```{r}
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
fl50_h2 <- filter(gdd_12c_cue, !(SITE %in% c("OVTN"))) %>%
  left_join(select(sites, SITE:Latitude)) %>%
  left_join(photo_cue_prob) %>%
  dplyr::select(-(LIBRARY:PLANTED_2018), LATITUDE) %>%
  filter(!is.na(SUBPOP) & !is.na(LATITUDE)) %>%
  filter(PLANT_ID %in% row.names(k_full)) %>%
  filter(!(photo_prob > 0.5 & SITE %in% c("KING", "TMPL", "PKLE", "STIL"))) %>%
  ungroup() %>%
  mutate(PLANT_ID = as_factor(PLANT_ID),
         SITE = as_factor(SITE)) %>%
    as.data.frame()

fl50_h2$SUBPOP <- factor(fl50_h2$SUBPOP, levels = c("Midwest", "Gulf", "Atlantic", "4X", "8X"))

asr_fl50 <- asreml(FL50 ~ 1, random = ~idv(SITE) + ~vm(PLANT_ID, k_full),
                  residual = ~idv(units), data = fl50_h2, 
                  workspace = "3gb")

summary(asr_fl50)$varcomp %>%
  as_tibble(rownames = "Effect") %>%
  mutate(`%Var` = (component*100)/sum(component),
         `VarSE` = (std.error*100)/sum(component),
         lowVar = `%Var` - VarSE,
         hiVar = `%Var` + VarSE,
         loglik = asr_fl50$loglik) %>%
  write_csv(., "manuscript/plots/Variance_components_fl50_photo_insensitive_only_by_site_and_k_full.csv",
            col_names = TRUE)
```

```{r}
gdd2 <- lm(CGDD_12C ~ SUBPOP:LATITUDE + LATITUDE, data = gdd_12c_7site)
summary(gdd2)
gdd3 <- lm(CGDD_12C ~ LATITUDE:SUBPOP + SUBPOP, data = gdd_12c_7site)
summary(gdd3)

gdd4 <- lm(CGDD_12C ~ LATITUDE*photo_prob*SUBPOP, data = gdd_12c_7site)
summary(gdd4)

gdd7site <- gdd_12c_7site %>%
  filter(!is.na(LATITUDE) & !is.na(SUBPOP)) %>%
  ungroup() %>%
  mutate(pred7lat = predict(gdd2, .),
         resid7lat = resid(gdd2),
         pred7int = predict(gdd3,.),
         resid7int = resid(gdd3))

gdd7photo <- gdd_12c_7site %>%
  ungroup() %>%
  filter(!is.na(LATITUDE) & !is.na(SUBPOP) & !is.na(photo_prob)) %>%
  mutate(pred7photo = predict(gdd4, .),
         resid7photo = resid(gdd4))

gdd7site %>%
  filter((SUBPOP == "Midwest" & between(LATITUDE, 30, 46)) | 
           !(SUBPOP %in% c("Midwest", "Gulf")) | 
          (SUBPOP == "Gulf" & between(LATITUDE, 25, 43)) ) %>%
  ggplot(aes(x= SUBPOP, y = LATITUDE)) + 
  geom_jitter(aes(color = pred7int)) +
  scale_color_viridis(option = "B")
gdd7site %>% 
  ggplot(aes(x = LATITUDE, y = resid7int)) + 
  geom_point(aes(color = SUBPOP), alpha = 0.4) + 
  scale_color_viridis(discrete = TRUE, option = "B") + 
  geom_hline(yintercept = 0, linetype = 2) +
  geom_smooth(aes(color = SUBPOP)) + 
  facet_wrap(~SUBPOP) # int better than lat in terms of residuals

gdd7site %>% 
  ggplot(aes(x = pred7int , y = resid7int)) + 
  geom_point(aes(color = SUBPOP), alpha = 0.4) + 
  scale_color_viridis(discrete = TRUE, option = "B") + 
  geom_hline(yintercept = 0, linetype = 2) +
  geom_smooth(aes(color = SUBPOP)) + 
  facet_wrap(~SUBPOP) # int better than lat in terms of residuals
```

```{r}

gdd7site %>%
  group_by(SUBPOP) %>%
  summarise(meanpred = mean(pred7int),
            fivepred = quantile(pred7int, 0.05),
            ninefivepred = quantile(pred7int, 0.95),
            fiveobs = quantile(CGDD_12C, 0.05),
            ninefiveobs = quantile(CGDD_12C, 0.95))
```

```{r}

gdddf <- data.frame(LATITUDE = 26:45, SUBPOP = rep(c("Midwest", "Gulf", "Atlantic", "4X", "8X"), length(25:45)/5))
gdddf <- gdddf %>% complete(LATITUDE, SUBPOP)

# Predict the probabilities
gdddf$pred9 <- predict(gdd1, gdddf)
gdddf$pred7int <- predict(gdd3, gdddf)
gdddf$pred7lat <- predict(gdd2, gdddf)
resid(gdd2)

gdddf %>%
  filter((SUBPOP == "Midwest" & between(LATITUDE, 30, 46)) | 
           !(SUBPOP %in% c("Midwest", "Gulf")) | 
          (SUBPOP == "Gulf" & between(LATITUDE, 25, 43)) ) %>%
  ggplot(aes(x= SUBPOP, y = LATITUDE)) + 
  geom_tile(aes(fill = pred7int)) +
  scale_fill_viridis(option = "B")

gdddf %>%
  mutate(diff = pred7int-pred7lat) %>%
  ggplot(aes(x= SUBPOP, y = LATITUDE)) + 
  geom_tile(aes(fill = diff)) +
  scale_fill_viridis(option = "B")
```

### 2020-12-03 R/Pre_manuscript/Draft_models_and_stats_2020-06-19.Rmd
Move R scripts not used in current manuscript and rename those that are

---
title: "Draft_models_and_stats_2020-06-19"
author: "Alice MacQueen"
date: "6/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(switchgrassGWAS)
library(cowplot)
library(corpcor)
library(viridis)
```

# ASReml-R analyses on server

```{r}
library(tidyverse)
library(asreml)
library(switchgrassGWAS)
library(cowplot)
library(viridis)
library(corpcor)
```

# Load data

```{r}
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "manuscript", "plots")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe <- read_rds(file.path(workingdir, "data",
                          "Phenology_CGDD_PTT_and_rainfall_phenotypes.rds"))
```

Unfortunately, no parents flowered exclusively during shortening days at the three Texas sites; WBC flowered during shortening days in the majority years in the Texas common gardens.
```{r}
phe %>%
  filter(WHERE == "FWCR" & PLANT_ID %in% c("AP13", "WBC", "DAC", "VS16", "VxW", "AxD")) %>%
  filter(PHE == "FL50" & SITE %in% c("PKLE", "TMPL", "KING", "OVTN", "STIL")) %>%
  group_by(PLANT_ID, SITE, MEAS >= 173) %>%
  tally() 
```
AxD KING 0 / 41
PKLE 0 / 39
TMPL 0/39

VxW 
PKLE 2 / 43
TMPL: 21 / 41
KING: 0 / 43

AP13:
TMPL: 56 / 66 >= 173
PKLE: 52 / 67 
KING: 0 / 40

WBC: 
TMPL: 40/40
PKLE: 39 / 46
KING: 10 / 18

DAC: 0% of 15
VS16: 3.6% of 56

```{r}
56/66
52/67
0
(56+52)/(66+67+40)

100
39/46
10/18
(40+39+10)/(40+46+18)
2/56
23/(43+41+43)
41+39+39
```

# daylength cue

Additionally, we observed no evidence for any specific day length that triggered flowering (model: DYLN at FL50 ~ SUBPOP + PLANT_ID + SUBPOP:LATITUDE does not explain much variance). 
```{r}
ptt_df <- readRDS(file = "~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")

wea_df <- ptt_df %>%
  dplyr::select(DOY:GDD_12C_SM, PTT_12C)

GR_start <- phe %>% 
  filter(WHERE == "GWAS") %>%
  ungroup() %>%
  group_by(SITE, YEAR, PLANT_ID, PLOT_GL) %>%
  filter(PHE == "GR50") %>%
  summarise(GR_start = floor(min(MEAS)))
FL_end <- phe %>% 
  filter(WHERE == "GWAS") %>%
  ungroup() %>%
  group_by(SITE, YEAR, PLANT_ID, PLOT_GL) %>%
  filter(PHE == "FL50") %>%
  summarise(FL_end = ceiling(max(MEAS)))

wea_gr_fl <- wea_df %>%
  left_join(GR_start) %>%
  left_join(FL_end) %>%
  group_by(Day_of_year, SITE, YEAR, PLANT_ID, PLOT_GL) %>%
  filter(between(Day_of_year, GR_start, FL_end)) %>% 
  ungroup() #
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")

wea_gr_fl %>%
  filter(Day_of_year == FL_end & FL_end == 175)
```

```{r}
dyln_cue <- wea_gr_fl %>%
  filter(Day_of_year == FL_end) %>%
  left_join(select(metadata, PLANT_ID, SUBPOP, LATITUDE)) %>%
  mutate(Photo_cue = case_when(FL_end >= 174 ~ 1,
                               FL_end <= 173 ~ 0,
                               TRUE ~ NA_real_)) %>%
  filter(Photo_cue == 1) %>%
  filter(!is.na(SUBPOP) & !is.na(LATITUDE)) %>%
  filter(PLANT_ID %in% row.names(k_full)) %>%
  arrange(SITE) %>%
  mutate(PLANT_ID = as_factor(PLANT_ID),
         SITE = as_factor(SITE),
         PLOT_GL = as_factor(PLOT_GL)) %>%
  as.data.frame()
dyln_cue$SUBPOP <- factor(dyln_cue$SUBPOP, levels = c("Midwest", "Gulf",
                                                      "Atlantic", "4X", "8X"))

asr_dyln <- asreml(DYLN ~ 1, random = 
                    ~vm(PLANT_ID, k_full) + ~idv(SITE),
                  residual = ~dsum(~units|SITE), data = dyln_cue, 
                  workspace = "3gb")
summary(asr_dyln)$varcomp
summary(asr_dyln)$bic
vpredict(asr_dyln, h2 ~ V2/(V1+V2+V3+V4+V5+V6+V7+V8+V9+V10+V11))
save_plot(filename=file.path(outputdir, "dyln_asreml_1.png"), 
          plot = plot(asr_dyln))

asr_dsum <- asreml(DYLN ~ 1,
                   random = ~vm(PLANT_ID, k_full) +
                     ~idv(SITE),
                   residual = ~idv(units),
                   data = dyln_cue, 
                   workspace = "3gb")
summary(asr_dsum)$varcomp
summary(asr_dsum)$bic
vpredict(asr_dsum, h2 ~ V2/(V1+V2+V3))
save_plot(filename=file.path(outputdir, "dyln_asreml_dsum.png"), 
          plot = plot(asr_dsum))

summary(asr_dyln)$bic    # first model is better
summary(asr_dsum)$bic

asr_corv <- asreml(DYLN ~ SITE,
                   random = ~vm(PLANT_ID, k_full),
                   residual = ~id(units):corv(SITE),
                   data = dyln_cue,
                   workspace = "3gb"
                   )

model.corv <- asreml(fixed = RESP ~ TIME,
                     residual = ~id(RAT):corv(TIME),
                     data = LUNG)

summary(asr_dyln)$varcomp %>%
  as_tibble(rownames = "Effect") %>%
  mutate(`%Var` = (component*100)/sum(component),
         `VarSE` = (std.error*100)/sum(component),
         lowVar = `%Var` - VarSE,
         hiVar = `%Var` + VarSE,
         loglik = asr_dyln$loglik) %>%
  write_csv(., "manuscript/plots/Variance_components_dyln_fl50_by_site_and_k_full.csv",
            col_names = TRUE)

dyln_cue$predicted <- asr_dyln$linear.predictors
dyln_cue %>% 
  as_tibble() %>%
  group_by(SUBPOP) %>%
  summarise(meanpred = mean(predicted),
            fivepred = quantile(predicted, 0.025),
            ninefivepred = quantile(predicted, 0.975),
            fiveobs = quantile(DYLN, 0.025),
            ninefiveobs = quantile(DYLN, 0.975))

blup <- summary(asr_dyln, coef=TRUE)$coef.random %>% as_tibble(rownames = "Effect")
int <- summary(asr_dyln, coef=TRUE)$coef.fixed %>% as_tibble(rownames = "Effect") %>% rename(std.error = `std error`)

write_csv(int, "manuscript/plots/Effects_dyln_fl50_by_site_and_k_full.csv", col_names = TRUE)
write_csv(blup, "manuscript/plots/Effects_dyln_fl50_by_site_and_k_full.csv", col_names = TRUE, append = TRUE)
```

### 2020-12-03 R/Pre_manuscript/Flowering_time_environmental_index_analysis.Rmd
Move R scripts not used in current manuscript and rename those that are

--- 2020-03-13 R/Flowering_time_environmental_index_analysis.Rmd
@@
 library(zoo)
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
 ```
@@
 fourway_sites <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL")
+metadata$ECOTYPE <- factor(metadata$ECOTYPE, levels = c("Upland", "Coastal", "Lowland", "Unknown"))
+metadata$GENETIC_SUBPOP_50PER <- factor(metadata$GENETIC_SUBPOP_50PER, levels = c("Midwest", "Atlantic", "Gulf", "Admixed", "Unknown"))
 ```
@@
 
-Make GDD & PTT using a range of base temperatures
+# Make GDD & PTT using a range of base temperatures
 ```{r}
 
-Tbase <- seq(0, 6)
-Topt <- seq(8,30)
+Tbase <- seq(5, 13)
+Topt <- seq(24,33)
 
@@
 M[1,best1]
+```
+
+# Make GDD & PTT using a range of base temperatures
+
+And starting every individual's GDD and PTT from its own GR50 time
+```{r}
+
+Tbase <- seq(5, 13)
+Topt <- seq(24,33)
+
+GR_start <- gwas_d2f %>% 
+  group_by(SITE) %>%
+  summarise(GR_start = floor(min(GR50)))
+FL_end <- gwas_d2f %>% 
+  group_by(SITE) %>%
+  summarise(FL_end = ceiling(max(FL50)))
+
+GDD_PT <- NULL
+
+tmp1 <- weather %>% 
+      left_join(GR_start, by = "SITE") %>%
+      left_join(FL_end, by = "SITE") %>%
+      group_by(SITE, Day_of_year) %>%
+      filter(between(Day_of_year, GR_start, FL_end)) %>% 
+      ungroup %>%
+  left_join(gwas_d2f)
+
+for (j in Tbase){
+  for (k in Topt){
+
+    tmp <- tmp1 %>%
+      group_by(SITE, PLANT_ID, Day_of_year) %>%
+      filter(between(Day_of_year, floor(GR50), ceiling(FL50))) %>%
+      mutate(TMEAN = (TMAX_SM + TMIN_SM) / 2,
+             GDD = ifelse(TMEAN >= j, TMEAN - j, 0),
+             GDD = ifelse(TMEAN >= k, k - j, GDD),
+             PT = GDD * DYLN) %>%
+      ungroup %>%
+      group_by(SITE, PLANT_ID) %>%
+      summarize_at(vars(GDD, PT), funs(sum))
+    colnames(tmp)[3:4] = c(paste0('GDD_', j, "_", k), paste0('PT_', j, "_", k))
+    GDD_PT = c(GDD_PT, tmp)
+    }
+}
+
+  GDD_PT <- GDD_PT %>% 
+    as.data.frame()
+
+  GDD_PT <- GDD_PT %>%
+    select(SITE, PLANT_ID, starts_with("GDD"), starts_with("PT"))
+
+  gwas_avegdd <- gwas_d2f %>%
+    left_join(GDD_PT, by = c("PLANT_ID", "SITE")) %>%
+    ungroup()
+
+###correlation
+
+M = cor(gwas_avegdd %>% select(D2F, starts_with('GDD')) %>% filter(complete.cases(.)))
+
+# corrplot::corrplot.mixed(M)
+
+(best1 <- which(M[1,] == max(M[1,2:ncol(M)])))
+
+M[1,best1]
+
+M = cor(gwas_avegdd %>% select(D2F, starts_with('PT')) %>% filter(complete.cases(.)))
+
+# corrplot::corrplot.mixed(M[1:15,1:15])
+
+(best1 <- which(M[1,] == max(M[1,2:ncol(M)])))
+
+M[1,best1]
+
+ggplot(data = M, aes(x = D2F, y = PT_8_33)) + 
+  geom_point()
+M2 <- enframe(M[1,], name = "PT", value = "corr") %>%
+  filter(PT != "D2F") %>%
+  separate(col = PT, into = c("Var", "Tbase", "Topt")) %>%
+  dplyr::select(-Var) %>%
+  pivot_wider(names_from = Topt, values_from = corr, names_prefix = "Topt_") %>%
+  mutate(Tbase = paste0("Tbase_", Tbase))
+M3 <- M2 %>%
+  column_to_rownames(var = "Tbase") %>%
+  as.matrix(.)
+corrplot::corrplot.mixed(M3)
+```
+```{r}
+write_rds(gwas_avegdd, path = "../analysis/environmental-index/GDD_and_PTT_GWAS.rds")
+```
+
+# Do analysis just using daylength, not temperature
+
+And starting every individual's daylength received from its own GR50 time
+```{r}
+
+Tbase <- seq(5, 13)
+Topt <- seq(24,33)
+
+GR_start <- gwas_d2f %>% 
+  group_by(SITE) %>%
+  summarise(GR_start = floor(min(GR50)))
+FL_end <- gwas_d2f %>% 
+  group_by(SITE) %>%
+  summarise(FL_end = ceiling(max(FL50)))
+
+GDD_PT <- NULL
+
+tmp1 <- weather %>% 
+      left_join(GR_start, by = "SITE") %>%
+      left_join(FL_end, by = "SITE") %>%
+      group_by(SITE, Day_of_year) %>%
+      filter(between(Day_of_year, GR_start, FL_end)) %>% 
+      ungroup %>%
+  left_join(gwas_d2f)
+
+    tmp <- tmp1 %>%
+      group_by(SITE, PLANT_ID, Day_of_year) %>%
+      filter(between(Day_of_year, floor(GR50), ceiling(FL50))) %>%
+      ungroup %>%
+      group_by(SITE, PLANT_ID) %>%
+      summarize_at(vars(DYLN), funs(sum))
+    GDD_PT = c(GDD_PT, tmp)
+
+  GDD_PT <- GDD_PT %>% 
+    as.data.frame()
+
+  GDD_PT <- GDD_PT %>%
+    select(SITE, PLANT_ID, starts_with("DYLN"), starts_with("PT"))
+
+  gwas_avegdd <- gwas_d2f %>%
+    left_join(GDD_PT, by = c("PLANT_ID", "SITE")) %>%
+    ungroup()
+
+###correlation
+
+M = cor(gwas_avegdd %>% select(D2F, starts_with('DYLN')) %>% filter(complete.cases(.)))
+
+ corrplot::corrplot.mixed(M)
+
+(best1 <- which(M[1,] == max(M[1,2:ncol(M)])))
+
+M[1,best1]
+
+gwas_avegdd %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  filter(PLANT_ID %in% c("AP13", "J001.B", "J005.B", "J006.B", "J016.B", "J024.A", "J026.B", "J350.C", "J499.B", "J517.B")) %>%
+  ggplot(aes(x = D2F, y = DYLN)) + 
+  geom_point(aes(color = PLANT_ID)) + 
+  geom_smooth(aes(color = PLANT_ID), se = FALSE, method = lm) +
+  scale_color_viridis(discrete = TRUE)
+
+gwas_avegdd %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  filter(!(GENETIC_SUBPOP_50PER %in% c("Admixed", "Unknown", "Midwest"))) %>%
+  filter(!(ECOTYPE %in% c("Upland", "Unknown"))) %>%
+  ggplot(aes(x = D2F, y = DYLN.8)) + 
+  geom_point(aes(color = GENETIC_SUBPOP_50PER)) + 
+  geom_smooth(aes(color = GENETIC_SUBPOP_50PER), se = FALSE, method = lm) +
+  scale_color_viridis(discrete = TRUE) + ylab("One site-year of cumulative daylength") + xlab("Days to flowering (FL50 - GR50)")
+
+gwas_avegdd %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  filter(!(GENETIC_SUBPOP_50PER %in% c("Admixed", "Unknown", "Midwest"))) %>%
+  filter(!(ECOTYPE %in% c("Upland", "Unknown"))) %>%
+  ggplot(aes(x = D2F, y = DYLN.8)) + 
+  geom_point(aes(color = ECOTYPE)) + 
+  geom_smooth(aes(color = ECOTYPE), se = FALSE, method = lm) +
+  scale_color_viridis(discrete = TRUE) + ylab("One site-year of cumulative daylength") + xlab("Days to flowering (FL50 - GR50)")
+```
+
+```{r}
+gwas_avegdd %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  filter(!(GENETIC_SUBPOP_50PER %in% c("Admixed", "Unknown"))) %>%
+  filter(!(ECOTYPE %in% c("Unknown"))) %>%
+  ggplot(aes(x = D2F, y = DYLN)) + 
+  geom_point(aes(color = GENETIC_SUBPOP_50PER)) + 
+  geom_smooth(aes(color = GENETIC_SUBPOP_50PER), se = TRUE, method = "gam",
+              formula = y ~ poly(x,2)) +
+  scale_color_viridis(discrete = TRUE, end = 0.8) + ylab("One site-year of cumulative daylength") + xlab("Days to flowering (FL50 - GR50)")
+
+save_plot(filename = "../analysis/environmental-index/Days_to_flowering_by_Genetic_subpop.png", plot = last_plot())
+
+gwas_avegdd %>%
+  left_join(metadata, by = "PLANT_ID") %>%
+  filter(!(GENETIC_SUBPOP_50PER %in% c("Admixed", "Unknown"))) %>%
+  filter(!(ECOTYPE %in% c("Unknown"))) %>%
+  ggplot(aes(x = D2F, y = DYLN)) + 
+  geom_point(aes(color = ECOTYPE)) + 
+  geom_smooth(aes(color = ECOTYPE), se = TRUE, method = "gam", 
+              formula = y ~ poly(x,2)) +
+  scale_color_viridis(discrete = TRUE, end = 0.8) + 
+  ylab("One site-year of cumulative daylength") + 
+  xlab("Days to flowering (FL50 - GR50)") + 
+  theme(legend.position = "right")
+
+save_plot(filename = "../analysis/environmental-index/Days_to_flowering_by_Ecotype.png", plot = last_plot())
+
 ```

### 2020-12-03 R/Pre_manuscript/GRM_kinship_for_fourway.Rmd
Move R scripts not used in current manuscript and rename those that are

--- 2020-04-04 R/GRM_kinship_for_fourway.Rmd
@@
 ## 
+```{r}
+write_rds(cross, "../data/fourway_cross_data.rds", compress = "gz")
+```
 
@@
 geno_df <- cross$geno[[1]]$data
-
+as_tibble(cross$pheno)
 for(i in 2:18){
@@
 # snps[1:20,1000:1005]
+snps1 <- snps[1:750,]
+snps <- snps1
 ```
@@
 
-mash_plot_pairwise_sharing(corrmatrix = K[1:30,1:30])
+mash_plot_pairwise_sharing(corrmatrix = K[1:20,1:20])
 
-write_rds(K, path = "~/Github/pvdiv-phenology-gxe/data/Kinship_van_Raden_5250g_fourway_cross.rds")
+write_rds(K, path = "~/Github/pvdiv-phenology-gxe/data/Kinship_van_Raden_750g_fourway_cross_withInbred.rds")
 ```
+
+```{r}
+mash_plot_pairwise_sharing(corrmatrix = K[1:20,1:20])
+
+```

### 2020-12-03 R/Pre_manuscript/Make_mash_U_hyp.Rmd
Move R scripts not used in current manuscript and rename those that are

--- 2020-07-09 R/Make_mash_U_hyp.Rmd
@@
 mash_plot_effects(m, n = 2, saveoutput = TRUE, suffix = suffix)
-mash_plot_effects(m, n = 8)
+mash_plot_effects(m, n = 8, saveoutput = TRUE)
 
@@
 For interpretability of the mash results... may want to scale all these effects so that 0 is, say, the summer solstice? Instead of the mean flowering time at that site for that pop.
+
+No, this doesn't work, because then there are no "0" effects in the random effects. Humm.
 ```{r}
@@
 240 +/- 28
+
+```{r}
+path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
+                       "gwas", "GDD", "Uhyp")
+numSNPs <- 30000
+suffix = "gulf_solstice"
+Ugulfhyp <- read_rds(path = file.path(path, "U_hyp_gulf_1phe.rds"))
+U_h2 <- list(Udist = Ugulfhyp$U_dist, UGDD = Ugulfhyp$U_CGDD_12C)
+
+#U_ed <- file.path(path, "Mash_data_driven_covariances_30000_gulf_uhyp1.rds")
+
+m <- mash_standard_run(path = path, numSNPs = numSNPs, suffix = suffix,
+                       saveoutput = TRUE)
+```
+
+Maybe...
+I could scale everything by the maximum value across the whole dataset.
+This would be a different type of scaling that could not handle multiple different phenotypes.
+But, for the same phenotype, it would 1) allow covariance matrices to be scaled from -1 to 1; 2) allow us to still visualize the fact that effect sizes are different at different sites - that there is variance-changing GxE.
+I think.
+Hopefully.
+
+I should think a little more on how mash is dealing with this... and look into the mash code...
+
+## Plot eigenvectors to capture predominant patterns in covariance matrices.
+```{r}
+sum(m$fitted_g$pi)
+m$fitted_g$pi[which(m$fitted_g$pi > 0.005)]
+
+v <- m$fitted_g$Ulist$ED_PCA_1
+b1=svd(v)
+for(g in 1){
+barplot(b1$v[,g]/v[which.max(abs(v[,g])),g])
+}
+
+b1$d # the top eigenvector captures the predominant pattern in the ED_PCA_1 covariance matrix.
+
+m$fitted_g$grid
+```
+Plot eigenvectors capturing predominant patterns
+
+The eigenvectors capture the predominant patterns in the Uk3 covariance matrix.
+```
+col = as.character(color.gtex[,2])
+for(g in 1:3){
+  v=svd(covmat[[k]])$v[h,];d=svd(covmat[[k]])$d
+  par(mar=c(8,4.1,4.1,2.1))
+  barplot(v[,g]/v[which.max(abs(v[,g])),g],las=2,
+          main=paste("Eigenvector",g,"of Uk",k),
+          cex.names = 0.5,col=col[h],names=names[h])
+}
+```
+# Matrix math with V_j
+
+Each covariance matrix Vj is specified as Vj = SjCSj, where C is a correlation matrix that accounts for correlations among the measurements in the R conditions, and Sj is the R x R diagonal matrix with diagonal elements s^j1, … , s^jR.
+If the measurements in the R conditions are independent, one would set C = IR, the R x R identity matrix, and therefore Vj = Sj2.
+
+Interestingly, they assume row-wise normal distribution of effect sizes, so a SNP has effects taken from a normal distribution.
+They assume b^j is normally distributed about bj with variance-covariance matrix Vj.
+The more highly correlated two conditions are, the
+```{r}
+Sj <- matrix(data = c(3, 0 , 0, 2), nrow = 2, ncol = 2)
+Clo <- matrix(data = c(1, 0.1, 0.1, 1), nrow = 2, ncol = 2)
+isSymmetric(Clo)
+Cid <- matrix(data = c(1, 0, 0, 1), nrow = 2, ncol = 2)
+Chi <- matrix(data = c(1, 1, 1, 1), nrow = 2, ncol = 2)
+
+Sj %*% Cid %*% Sj
+Sj %*% Clo %*% Sj
+Sj %*% Chi %*% Sj
+?cov2cor
+
+```
+
+```{r}
+dataf = list(Bhat = B_hat_random_df_30000topSNPs_gulf_uhyp1, Shat = S_hat_random_df_30000topSNPs_gulf_uhyp1)
+estimate_null_correlation_simple = function(dataf, z_thresh=2, est_cor = TRUE){
+  z = dataf$Bhat/dataf$Shat
+  max_absz = apply(abs(z),1, max)
+  nullish = which(max_absz < z_thresh)
+  if(length(nullish)<n_conditions(dataf)){
+    stop("not enough null data to estimate null correlation")
+  }
+  nullish_z = z[nullish,]
+  Vhat = cov(nullish_z)
+  if(est_cor){
+    Vhat = cov2cor(Vhat)
+  }
+  return(Vhat)
+}
+```

### 2020-12-03 R/Pre_manuscript/Map_all_metadata.Rmd
Move R scripts not used in current manuscript and rename those that are

---
title: "map all metadata"
author: "Alice MacQueen"
date: "7/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(maps)
library(ggmap)
usa <- map_data("usa")
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
```
Metadata sent to David Lowry 2020-07-31
```{r}
metadata %>%
  dplyr::select(PLANT_ID:GENETIC_SUBPOP_KINSHIP, SUBPOP, ECOTYPE_NNET, PLOIDY:PLANTED_2018) %>%
  mutate(LONGITUDE = -LONGITUDE) %>%
  write_csv(., "pvdiv_gwas_metadata.csv")
```

#

```{r cars}
library(ggrepel)

metaplot <- metadata %>%
  filter(!is.na(LATITUDE)) %>%
  mutate(LONGITUDE = -LONGITUDE)
metaplot$ECOTYPE_NNET <- factor(metaplot$ECOTYPE_NNET, levels = c("Coastal", "Lowland", "Upland", "Unknown"))
siteplot <- sites %>%
  filter(!(SITE %in% c("FRMI", "OVTN")))

ggplot() + 
  geom_polygon(data = usa, aes(x=long, y = lat, group = group), 
               fill = "white", color = "black") +
  coord_quickmap(xlim = c(-105, -66), ylim = c(25, 50)) +
  geom_jitter(data = metaplot, aes(x = LONGITUDE, y = LATITUDE, 
                                   color = SUBPOP, shape = ECOTYPE_NNET),
              width = 0.45, height = 0.45) + 
  geom_point(data = siteplot, aes(x = Longitude, y = Latitude), shape = 21, size = 4, stroke = 1.5, fill = "white") +
  geom_point(data = siteplot, aes(x = Longitude, y = Latitude), shape = 4, size = 3, stroke = 1.5) +
  geom_label_repel(data = siteplot, aes(x = Longitude, y = Latitude, label = manu_site), alpha = 0.7) +
  labs(x = "Longitude", y = "Latitude") + 
  theme(legend.position = "right") +
  #scale_color_viridis(discrete = TRUE) +
  scale_shape_manual(values = c(0, 6, 1, 4, 5)) +
  #scale_fill_manual(values = c("#280B54FF","#FAC127FF", "#D44842FF", "white")) +
  scale_color_manual(values = c(rep("grey", 1), "#090906","#6E91CB", "#F47F72", "#442C83", "#280B54FF","#FAC127FF", "#D44842FF", "white")) 
  #scale_alpha_manual(values = c(0.7, 0.7, 0.7, 0.7), guide = FALSE)
save_plot(filename = "../manuscript/plots/Map_all_subpops_and_Ecotypes.png", plot = last_plot(), base_height = 4)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2020-12-03 R/Pre_manuscript/Overlap_with_4way_QTL.Rmd
Move R scripts not used in current manuscript and rename those that are

--- 2020-04-24 R/Overlap_with_4way_QTL_D2F.Rmd
@@
 ---
-title: "D2F_4way_overlap"
+title: "FL50_4way_overlap"
 author: "Alice MacQueen"
-date: 2020-04-22
+date: "1/21/2020"
 output: html_document
@@
 ```{r}
-mash <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_Midwest/Strong_Effects49260SNPs.rds")
-qtl <- read_csv(file = "~/Github/pvdiv-phenology-gxe/data/QTLs_with_flank_markers_FL50_7_sites_full_model.csv")
-sites <- readRDS("~/Github/pvdiv-phenotypes/data/sites.rds")
+mash <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/FL50/Strong_Effects29475SNPs.rds")
+qtl <- read_csv("QTLsWithFlankMarkers_full_multiple_sites.csv")
+load("~/Github/pvdiv-phenology-gxe/data/sites.rda")
 ```
@@
          )
-effects <- ashr::get_pm(mash)
-sd <- ashr::get_psd(mash)
-colnames(sd) <- str_replace(colnames(sd), "Bhat", "Shat")
-
-df1 <- cbind(df1, effects, sd)
 ```
@@
 ```{r}
-qtl %>%
-  arrange(desc(lod)) %>%
-  dplyr::select(lodcolumn, lod, chr, flank_lo, flank_hi)
-```
-From most to least significant, for QTL:
-	Chr02N_58.039498 61.274395
-	Chr05N_1.502713 4.328965
-	Chr05N_58.916323 65.990782
-	Chr04K_1.959391 5.499253
-	Chr02K_63.687871 65.484856
-	  Chr09K_13.655791 18.177373
-		Chr03K_17.368264	Chr03K_26.170048
-		Chr08N_5.54356	Chr08N_34.415808
-
-# Chr02N  58.039498 61.274395
-```{r}
-switchgrassGWAS::mash_plot_effects(m=mash, i = 11087)
-switchgrassGWAS::mash_plot_effects(m=mash, i = 11468)
+switchgrassGWAS:::get_log10bf(m=mash)
+switchgrassGWAS::mash_plot_effects(m=mash, i = 26711)
 
 df1 %>%
-  filter(log10bf > 5) # 25 values
-df1 %>%
-  filter(log10bf > 4) # 93 values (1800 above 2)
-```
-Close
-	Chr02N_58.039498 61.274395
-Chr02N_15281005	9405	4.057999	1	
-Chr02N_50051440	10742	4.288475	5	5.8214656195	
-Chr02N_56422155	11087	5.158628	6	5.9439138054	
-Chr02N_57947932	11178	4.390894	6	4.7648643926	
-Chr02N_63420472	11467	4.374355	5	5.5078867975	
-Chr02N_63420484	11468	5.022484	6	-5.0720890743
+  filter(log10bf > 5)
 
-Close, first is better
-	Chr05N_1.502713 4.328965
-	Chr05N_58.916323 65.990782
-Chr05N_4512043	25413	5.929161	6	8.5445301065	
-Chr05N_20476771	26421	4.303708	5	6.4982493387	
-Chr05N_24505333	26560	4.225575	5	7.4869178552	
-Chr05N_25731508	26609	4.081614	1	-0.0036192576	
-Chr05N_28066146	26673	4.727816	6	8.9999136556	
-Chr05N_36999363	26879	5.038580	5	-5.0653499084	
-Chr05N_37110671	26885	4.151390	5	6.1070001182	
-Chr05N_38945328	26961	4.980246	6
-Chr05N_46107016	27330	4.099486	6	3.6428083879	
-Chr05N_65588952	28354	4.003577	6
-
-Close
-	Chr04K_1.959391 5.499253
-Chr04K_1935684	17557	4.235319	5	6.1526138902	
-Chr04K_6537440	17814	4.403893	5	4.7144521481	
-Chr04K_6567024	17818	4.962056	6	5.1767913940	
-Chr04K_8657535	17940	5.205059  6
-Chr04K_12651532	18152	6.096971  6
-
-Not close
-	Chr02K_63.687871 65.484856
-Chr02K_23509869	6711	5.595545	6	-6.7548782752	
-Chr02K_23682427	6713	5.430460	5	6.9669557271	
-Chr02K_35881338	6882	4.095730	6
-
-```{r}
 panela <- df1 %>%
-  filter(grepl("Chr02N_(5|6)", Marker) & between(value, 11000, 11500)) %>%
+  filter(grepl("Chr09K_5", Marker) & value > 26500) %>%
   separate(Marker, into = c("Chr", "_", "Position"), sep = c(6,7), remove = FALSE) %>%
@@
   geom_point(aes(color = sigcond)) +
-  scale_color_viridis(option = "B", end = 0.8) # 26711
-save_plot(filename = "Chr02N_58_to_61Mb_D2F_mash_result_zoom.png", plot = panela, base_asp = 1.2)
-
-panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11087)
-#panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11467)
-##panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11468)
-#panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11178)
+  scale_color_viridis(option = "B") # 26711
+save_plot(filename = "Chr09K_FL50_mash_result_zoom.png", plot = last_plot(), base_asp = 1.5)
 
 p3 <- plot_grid(panela, panelb$ggobject, rel_widths = c(2,1))
-save_plot(filename = "Chr02N_58_to_61Mb_mash_result_zoom_2panel.png", plot = p3, base_asp = 1.5)
+save_plot(filename = "Chr09K_FL50_mash_result_zoom_2panel.png", plot = p3, base_asp = 3)
 ```
-# Chr05N 
+#Chr05N
 ```{r}
-
 panela <- df1 %>%
-  filter(grepl("Chr05N_", Marker) & between(value, 24000, 25600)) %>%
+  filter(grepl("Chr05N_1", Marker) & value > 15717) %>%
   separate(Marker, into = c("Chr", "_", "Position"), sep = c(6,7), remove = FALSE) %>%
@@
   geom_point(aes(color = sigcond)) +
-  scale_color_viridis(option = "B", end = 0.8) # 26711
+  scale_color_viridis(option = "B", end = 0.9) 
+panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 15864)
+panelb$ggobject 
+p3 <- plot_grid(panela, panelb$ggobject, rel_widths = c(2,1))
+save_plot(filename = "Chr05N_FL50_mash_result_zoom.png", plot = p3, base_asp = 3)
 
-panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 25413)
-#panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11467)
-##panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11468)
-#panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11178)
+```
+Chr05N_15045423	15847	4.688166	8	  affects flowering at the 3 major sites
+Chr05N_15048576	15849	4.101891	8	
+Chr05N_15951722	15863	4.048425	7	  affects flowering at KBSM & CLMB
+Chr05N_16084715	15864	4.226715	8	  affects flowering at the 3 major sites
 
-(p3 <- plot_grid(panela, panelb$ggobject, rel_widths = c(2,1)))
-save_plot(filename = "Chr05N_0_to_8Mb_D2F_mash_result_zoom.png", plot = panela, base_asp = 1.2)
-save_plot(filename = "Chr05N_0_to_8Mb_mash_result_zoom_2panel.png", plot = p3, base_asp = 1.5)
-```
-
-# Chr05N 60Mb
 ```{r}
-
 panela <- df1 %>%
-  filter(grepl("Chr05N_(5|6)", Marker) & between(value, 28000, 28500)) %>%
+  filter(grepl("Chr04K_1", Marker) & value > 11158) %>%
   separate(Marker, into = c("Chr", "_", "Position"), sep = c(6,7), remove = FALSE) %>%
@@
   geom_point(aes(color = sigcond)) +
-  scale_color_viridis(option = "B", end = 0.8) # 26711
-
-panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 28354)
-#panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11467)
-##panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11468)
-#panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11178)
-
-(p3 <- plot_grid(panela, panelb$ggobject, rel_widths = c(2,1)))
-save_plot(filename = "Chr05N_58_to_68Mb_D2F_mash_result_zoom.png", plot = panela, base_asp = 1.2)
-save_plot(filename = "Chr05N_58_to_68Mb_mash_result_zoom_2panel.png", plot = p3, base_asp = 1.5)
+  scale_color_viridis(option = "B", end = 0.9) 
+panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11493)
+panelb$ggobject 
+p3 <- plot_grid(panela, panelb$ggobject, rel_widths = c(2,1))
+save_plot(filename = "Chr04K_FL50_mash_result_zoom.png", plot = p3, base_asp = 3)
+df1 %>%
+  filter(grepl("Chr04K_1", Marker) & value > 11158 & log10bf > 4)
 ```
-
-# Chr04K
-
-Close
-	Chr04K_1.959391 5.499253
-Chr04K_1935684	17557	4.235319	5	6.1526138902	
-Chr04K_6537440	17814	4.403893	5	4.7144521481	
-Chr04K_6567024	17818	4.962056	6	5.1767913940	
-Chr04K_8657535	17940	5.205059  6
-Chr04K_12651532	18152	6.096971  6
-```{r}
-
-panela <- df1 %>%
-  filter(grepl("Chr04K_", Marker) & between(value, 17500, 18200)) %>%
-  separate(Marker, into = c("Chr", "_", "Position"), sep = c(6,7), remove = FALSE) %>%
-  mutate(Position = as.numeric(Position)) %>%
-  ggplot(aes(x=Position, y = log10bf)) +
-  geom_point(aes(color = sigcond)) +
-  scale_color_viridis(option = "B", end = 0.8) # 26711
-
-panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 18152)
-#panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 17557)
-##panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 17814)
-#panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 17818)
- switchgrassGWAS::mash_plot_effects(m=mash, i = 17940)
-
-(p3 <- plot_grid(panela, panelb$ggobject, rel_widths = c(2,1)))
-save_plot(filename = "Chr04K_1_to_12Mb_D2F_mash_result_zoom.png", plot = panela, base_asp = 1.2)
-save_plot(filename = "Chr04K_1_to_12Mb_mash_result_zoom_2panel.png", plot = p3, base_asp = 1.5)
-```
-
-# Chr08N_5.54356	Chr08N_34.415808
-```{r}
-
-panela <- df1 %>%
-  filter(grepl("Chr08N_", Marker) & between(value, 40000, 41100)) %>%
-  separate(Marker, into = c("Chr", "_", "Position"), sep = c(6,7), remove = FALSE) %>%
-  mutate(Position = as.numeric(Position)) %>%
-  ggplot(aes(x=Position, y = log10bf)) +
-  geom_point(aes(color = sigcond)) +
-  scale_color_viridis(option = "B", end = 0.8) # 26711
-
-panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 40305)
-#panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 40514)
-
-(p3 <- plot_grid(panela, panelb$ggobject, rel_widths = c(2,1)))
-save_plot(filename = "Chr08N_0_to_1Mb_D2F_mash_result_zoom.png", plot = panela, base_asp = 1.2)
-save_plot(filename = "Chr08N_0_to_1Mb_mash_result_zoom_2panel.png", plot = p3, base_asp = 1.5)
-```
+Chr04K_14442056	11332	4.954623	7	
+Chr04K_14919129	11382	4.263826	8	qtl
+Chr04K_14920267	11384	4.655388	8	
+Chr04K_14955235	11388	4.693233	7	
+Chr04K_16627327	11460	4.240948	7	
+Chr04K_16687695	11493	4.405048	7	
+Chr04K_18870690	11610	5.308748	6
 
@@
 i=1
-width = 1000000
+width = 200000
 

### 2020-12-03 R/Pre_manuscript/Plots_PAG_2020.Rmd
Move R scripts not used in current manuscript and rename those that are

---
title: "Plots_PAG_2020"
author: "Alice MacQueen"
date: "1/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(XLConnect)
library(rlang)
library(switchgrassGWAS)
library(cowplot)
library(viridis)
library(bigsnpr)
library(scales)
library(lubridate)

source("../../../Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
theme_set(theme_poster)

load("../data/sites.rda")
load("../data/metadata.rda")
load("../data/Taxa.rda")
load("../data/phenotypes_2019.rda")

usa <- map_data("state")
```

## Set colors

John's scheme
dark blue = MW
light blue = NEC
yellow = SEC
orange = GC
red = TX

blue = UP
orange = SEC, GC
red = TX

Plot: GR50, EM50, FL50, T_GR_FL as facets on x
10 sites as facets on y, in latitudinal order
colored by ecotype

```{r}
phenotypes_dates <- phenotypes_wide_2019 %>%
  dplyr::select(PLANT_ID, ends_with("GR50"), ends_with("EM50"), ends_with("FL50")) %>%
  pivot_longer(cols = BRKG_GR50:TMPL_FL50, names_to = "SITE_PHE", values_to = "Value") %>%
  filter(!is.na(Value)) %>%
  separate(SITE_PHE, into = c("SITE", "_", "PHE"), sep = c(4,5)) %>%
  dplyr::select(-`_`)

phenotypes_dates$SITE <- factor(phenotypes_dates$SITE, levels = c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING"))
phenotypes_dates$PHE <- factor(phenotypes_dates$PHE, levels = c("GR50", "EM50", "FL50"))

phenotypes_values <- phenotypes_wide_2019 %>%
  dplyr::select(PLANT_ID, ends_with("TC_EOS"), ends_with("HT_PAN_EOS"), ends_with("T_GR_FL")) %>%
  pivot_longer(cols = BRKG_TC_EOS:TMPL_T_GR_FL, names_to = "SITE_PHE", values_to = "Value") %>%
  filter(!is.na(Value)) %>%
  separate(SITE_PHE, into = c("SITE", "_", "PHE"), sep = c(4,5)) %>%
  dplyr::select(-`_`)

phenotypes_values$SITE <- factor(phenotypes_values$SITE, levels = c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING"))
phenotypes_values$PHE <- factor(phenotypes_values$PHE, levels = c("T_GR_FL", "TC_EOS", "HT_PAN_EOS"))

```

```{r}
 metadata <- metadata %>%
   filter(PLANT_ID != "AP13")

metadata <- add_row(metadata, PLANT_ID = "AP13", LIBRARY = "REF", LIB_CLIMATE = "Y", LIB_PHENO = "Y", LIB_ACC = "Y", JL_genetic_subpop = "TX", JL_detail_genetic = "TX", PLOIDY = "4X", Ecotype = "Texas", Genetic_subpop = "TX")

metadata$Genetic_subpop <- factor(metadata$Genetic_subpop, levels = c("TX", "GC,TX", "GC", "SEC", "NEC,SEC", "NEC", "MW", "Admixed", "Unknown"))
metadata$Ecotype <- factor(metadata$Ecotype, levels = c("Texas", "Coastal", "Upland"))
```

```{r}

phenotypes_dates %>%
  mutate(Days = as_date(Value)) %>%
  filter(PHE != "EM50") %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(Genetic_subpop != "Unknown" & PLANT_ID != "AP13") %>%
  ggplot(aes(x = Days)) + 
  geom_histogram(aes(fill = Ecotype)) + 
  facet_grid(SITE~PHE, scales = "free") +
  scale_fill_manual(values = c("#CC0000", "#FF9900", "#000077")) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) + 
  scale_y_continuous(breaks = c(20,60,100)) + 
  xlab("Date") + 
  ylab("Number of Individuals") +
  geom_vline(xintercept = c(120), linetype = 2) +
  geom_vline(xintercept = c(181), linetype = 3, color = "grey20") +
  geom_vline(xintercept = c(243), linetype = 4, color = "grey80")

save_plot("Phenology_Data_by_Site_Ecotype_noAP13_2020-01-07.png", plot = last_plot(), base_height = 8, base_asp = 1.618/2)

phenotypes_values %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(PHE != "T_GR_FL") %>%
  filter(Genetic_subpop != "Unknown" & PLANT_ID != "AP13") %>%
  ggplot(aes(x = Value)) + 
  geom_histogram(aes(fill = Ecotype)) + 
  facet_grid(SITE~PHE, scales = "free") +
  scale_fill_manual(values = c("#CC0000", "#FF9900", "#000077")) +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) + 
  scale_y_continuous(breaks = c(20,60,100)) +
  xlab("Value") + 
  ylab("Number of Individuals") +
  geom_vline(xintercept = c(100), linetype = 2) +
  geom_vline(xintercept = c(300), linetype = 3, color = "grey20")

save_plot("Phenotype_Data_by_Site_Ecotype_noAP13_2020-01-07.png", plot = last_plot(), base_height = 8, base_asp = 1.8/2)

phenotypes_values %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(PHE == "T_GR_FL") %>%
  filter(Genetic_subpop != "Unknown" & PLANT_ID != "AP13") %>%
  ggplot(aes(x = Value)) + 
  geom_histogram(aes(fill = Ecotype)) + 
  facet_grid(SITE~PHE, scales = "free") +
  scale_fill_manual(values = c("#CC0000", "#FF9900", "#000077")) +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) + 
  scale_y_continuous(breaks = c(20,60,100)) +
  xlab("Value") + 
  ylab("Number of Individuals") +
  geom_vline(xintercept = c(100), linetype = 2) +
  geom_vline(xintercept = c(300), linetype = 3, color = "grey20")

save_plot("T_GR_FL_Data_by_Site_Ecotype_noAP13_2020-01-07.png", plot = last_plot(), base_height = 8, base_asp = 1.8/3)
```

```{r}
mashhattan <- mash_plot_manhattan_by_condition(m = mash_output)
mhtn <- mashhattan$ggmanobject + scale_color_viridis(option = "B",  end = 0.8) + scale_fill_viridis(option = "B", end = 0.8)

save_plot("Manhattan_test.png", plot = last_plot(), base_height = 4, base_asp = 2.5)

pairwise_plot <- mash_plot_pairwise_sharing(effectRDS = "Pairwise_sharing_Strong_Effects_21349SNPs.rds", reorder = FALSE)
save_plot("Pairwise_test_unordered.svg", plot = pairwise_plot$gg_corr, base_height = 2.5, base_asp = 1.618)

```

```{r}
pairwise_plot <- mash_plot_pairwise_sharing(effectRDS = "T_GR_FL/Pairwise_sharing_sign_Strong_Effects_26989SNPs.rds", reorder = TRUE)
save_plot("T_GR_FL/Pairwise_test_T_GR_FL.svg", plot = pairwise_plot$gg_corr, base_height = 2.5, base_asp = 1.618)

anno_dfs <- readRDS("T_GR_FL/Annotation_tables_Time_greenup_to_flowering_date_mash_7_sites_9_PCs.rds")
anno_dfs$top10SNPs_within20000bp
```

# Logistic winter survival

```{r}
annoswk <- readRDS("~/Github/pvdiv-climate-gwas/results-upland-vs-lowland/Logistic_winter_survival/Annotation_tables_logistic_winter_survival_subset_9_PCs.rds")
annosldly_wk <- readRDS("~/Github/pvdiv-climate-gwas/results-upland-vs-lowland/Logistic_winter_survival/Logistic_winter_survival_Lindley_sigZones_with_allele_frequencies.rds")
annoswk$top1000SNPs_within0bp %>%
  mutate(Posbin = ceiling(region_end/10000))
annosldly <- annosldly_wk %>% arrange(chr, end) %>%
  mutate(Posbin = ceiling(end/10000))
annosuncor <- annoswk$top1000SNPs_within100000bp %>%
  mutate(Posbin = ceiling(region_end/10000))
annosldly %>%
  left_join(annosuncor, by = c("chr" = "CHR", "Posbin")) %>%
  dplyr::select(chr, Posbin, everything()) %>%
  group_by(chr, Posbin) %>% tally()

pvdiv_table_topsnps(df = annosldly_wk, type = "table")

annosuncor %>%
  filter(CHR == "Chr03K")
```

```{r}
annos_winterkill <- annosldly %>%
  left_join(annosuncor, by = c("chr" = "CHR", "Posbin")) %>%
  dplyr::select(chr, Posbin, everything()) %>%
  arrange(desc(peak))

write_csv(annos_winterkill, path = "Annotations_Lindley_Winter_Killed.csv")
```

```{r}
winterkill_df <- annos_winterkill %>%
  dplyr::select(chr, beg, end) %>%
  dplyr::rename(CHR = chr, start = beg)  %>%
      mutate(start = as.integer(.data$start),
             end = as.integer(.data$end),
             POS = as.integer(.data$end))

library(VariantAnnotation)
save(winterkill_df, file = "winterkill_df.rda")

anno_info <-
  read_delim(file = file.path("~",  "Github", "pvdiv-phenology-gxe", 
                              "large-files",
                              "Pvirgatum_516_v5.1.annotation_info.txt"),
             col_names = TRUE, delim = "\t")
gene_anno_info <- tbl_df(anno_info) %>%
  distinct(locusName, .keep_all = TRUE)
txdb <- loadDb(file = file.path("~",  "Github", "pvdiv-phenology-gxe", 
                              "large-files",
                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))

 target <- with(winterkill_df, GRanges(seqnames = Rle(CHR),
                                      ranges = IRanges(start = start,
                                                       end = end,
                                                       names = NULL),
                                      strand = Rle(strand("*"))))

        ### Find genes that overlap input SNPs with VariantAnnotation
        loc <-
          VariantAnnotation::locateVariants(target, txdb,
                                            VariantAnnotation::AllVariants(promoter =
                                                                             VariantAnnotation::PromoterVariants(upstream = 2000L,
                                                        downstream = 2000L),
                                     intergenic =
                                       VariantAnnotation::IntergenicVariants(upstream = 20000L,
                                                          downstream = 20000L,
                                                          idType = "gene")))

test_df <- pvdiv_table_topsnps(df = winterkill_df, type = "table", anno_info = gene_anno_info, txdb = txdb)

load("Winterkill_lindley_annotation.rda")
library(tidyverse)
winterkill_anno_lindley <- test_df[[1]] %>% # CHR POS
  left_join(annosldly, by = c("CHR" = "chr", "POS" = "end"))
write_csv(winterkill_anno_lindley, "Winterkill_lindley_annotation_df.csv")
```

### 2020-12-03 R/Pre_manuscript/Visualize_BIOMASS_mash_GxE.Rmd
Move R scripts not used in current manuscript and rename those that are

--- 2020-03-13 R/Visualize_BIOMASS_mash_GxE.Rmd
@@
 length(get_significant_results(m))
-mash_plot_pairwise_sharing(corrmatrix = gxe$S_DS)
+mash_plot_pairwise_sharing(corrmatrix = gxe$S_all)
 max(gxe$S_DS)
@@
 mash_plot_pairwise_sharing(corrmatrix = gxe$S_AP)
+sum(gxe$S_AP)/78411 # 510
+sum(gxe$S_DS) # 3188
+sum(gxe$S_2_no) # 118913
+sum(gxe$S_all) # 130939
+sum(gxe$S_CN) # 11516
+
+(sum(gxe$S_AP)/sum(gxe$S_all))*length(get_significant_results(m))
+(sum(gxe$S_CN)/sum(gxe$S_all))*length(get_significant_results(m))
+(sum(gxe$S_2_no)/sum(gxe$S_all))*length(get_significant_results(m))
+(sum(gxe$S_DS)/sum(gxe$S_all))*length(get_significant_results(m))
+19780+85+1915+530
+length(gxe$S_all)
+130939/81
+# 21781
+# 
+mash_plot_pairwise_sharing(corrmatrix = (gxe$S_AP/gxe$S_all))
+max((gxe$S_AP/gxe$S_all))
+max((gxe$S_CN/gxe$S_all))
+max((gxe$S_2_no/gxe$S_all))
+max((gxe$S_DS/gxe$S_all))
+mean((gxe$S_AP/gxe$S_all))
+mean((gxe$S_CN/gxe$S_all))
+mean((gxe$S_2_no/gxe$S_all)) + mean((gxe$S_DS/gxe$S_all))
+
+AP <- (gxe$S_AP)
+CN <- (gxe$S_CN)
+DS <- (gxe$S_DS)
+twono <- (gxe$S_2_no)
+all <- gxe$S_all
+diag(twono) <- NA
+diag(AP) <- NA
+diag(CN) <- NA
+diag(DS) <- NA
+diag(all) <- NA
+
+median(AP, na.rm = TRUE)
+median(CN, na.rm = TRUE)
+median(DS, na.rm = TRUE)
+median(twono, na.rm = TRUE)
+median(all, na.rm = TRUE)
+
+sum(AP, na.rm = TRUE)/sum(diag(all))
+sum(CN, na.rm = TRUE)/sum(diag(all))
+sum(DS, na.rm = TRUE)/sum(diag(all)) + sum(twono, na.rm = TRUE)/sum(diag(all))
+sum(all, na.rm = TRUE)/sum(diag(all))
+
+sum(AP, na.rm = TRUE)/(sum(all, na.rm = TRUE))
+#sum(CN, na.rm = TRUE)/(2*sum(all, na.rm = TRUE))
+sum(DS, na.rm = TRUE)/(sum(all, na.rm = TRUE))
+sum(twono, na.rm = TRUE)/(sum(all, na.rm = TRUE))
+sum(all, na.rm = TRUE)/sum(all, na.rm = TRUE)
+
+AP/diag(all)
+CN/diag(all)
+DS/diag(all)
+twono/diag(all)
+mean(diag(all))
+sum(diag(all))
+
+mean(AP, na.rm = TRUE)
+mean(CN, na.rm = TRUE)
+mean(DS + twono, na.rm = TRUE) 
+mean(twono, na.rm = TRUE)
+mean(all, na.rm = TRUE)
+7+160+44+76
+max(all, na.rm = TRUE)
+120+159+7
+
+1.5+33+9+16
 ```
+
 `r length(get_significant_results(m))` SNPs have significant effects in the mash analysis.
@@
 Between `r min(gxe$S_2_no)` and `r max(gxe$S_2_no)` have effects of the same magnitude and same sign between pairs of conditions.
+
+Instead, the vast majority of SNPs have common effects across sites (
+`r mean((gxe$S_2_no/gxe$S_all)) + mean((gxe$S_DS/gxe$S_all))`) or are conditionally neutral (`r mean((gxe$S_CN/gxe$S_all))`) ( e.g., having effects at some sites but not others) with only `r mean((gxe$S_AP/gxe$S_all))` showing sign change.

### 2020-12-03 R/Pre_manuscript/Visualize_FL50_33M_Results.Rmd
Move R scripts not used in current manuscript and rename those that are

---
title: "FL50_33M_Results"
author: "Alice MacQueen"
date: "2/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(switchgrassGWAS)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

##

```{r}
annos$top10SNPs_within0bp
```

```{r}
m <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/FL50/gwas_33M/Strong_Effects28384SNPs.rds")
switchgrassGWAS::mash_plot_pairwise_sharing(effectRDS = "Pairwise_sharing_Strong_Effects_28384SNPs.rds")

mash_plot_manhattan_by_condition(m = m)
mash_plot_sig_by_condition(m=m)
mash_plot_effects(m=m, n=1)
```

# Overlaps with fourway set!

### 2020-12-03 R/Analysis_v0.5_weather_covariance_Uhyp_flowering.Rmd
Add analysis with weather phenotypes as covariance U_hyp

@@
 
-  ggcov1 <- pvdiv_plot_cov(cor_phe) + ggtitle(label = paste0(phe_hyp[i], "_", names(subpop_v2)[k]))
-  ggcov <- list.prepend(ggcov, ggcov1)
+  #ggcov1 <- pvdiv_plot_cov(cor_phe) + ggtitle(label = paste0(phe_hyp[i], "_", names(subpop_v2)[k]))
+  #ggcov <- list.prepend(ggcov, ggcov1)
   U_hyp <- list.prepend(U_hyp, cor_phe)
@@
 
-## Midwest NB
-Some univariate GWAS for the Midwest had population structure issues. Mostly for greenup - even GR50 at MO was quite bad. Drop MO for GR50. Otherwise they look ok for FL50 and GR50.
+## Set up mash inputs
 
-## Set up mash inputs
+### mash G&M
 ```{r}
@@
 k=3 # subpop: G&M, M, Gulf
-for(k in c(1:3)){
+for(k in c(3)){
 phe_suffix <- paste("FL50", subpop_v[k], 
@@
 ```
+Getting a lot of this: "error: chol(): decomposition failed"
 
-## Run mash
+## Run mash Gulf
 ```{r}
-for(k in c(1:3)){
+for(k in c(1)){
+  phe_suffix <- paste("FL50", subpop_v[k], 
+                      sep = "*.*")
+  gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
+                                        pattern = paste0("GWAS_datatable*.*",
+                                                         phe_suffix))
+  phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
+      str_replace(., "_$", "")
+  phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
+  numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
   m_out <- mash_standard_run(path = outputdir, numSNPs = numSNPs, 
@@
 
+## 2b. Midwest NB
+Some univariate GWAS for the Midwest had population structure issues. Mostly for greenup - even GR50 at MO was quite bad. Drop MO for GR50. Otherwise they look ok for FL50 and GR50.
+
+This is the error I get if I try and run mash on all eight sites with just the Midwest subpopulation as input:
+Error in mashr::mash_set_data(Bhat_random, Shat_random) :
+  Both Bhat and Shat are zero (or near zero) for some input data.
+Please check your input.
+If it is expected please set Shat to a positive number to avoid numerical issues; or lower the threshold to call zeros using zero_check_tol (currently set to 2.22044604925031e-16).
+To replace zero elements you can use `zero_Bhat_Shat_reset`.
+
+What's happening with Midwest input?  NE & OK inputs are all zeros.
+```{r}
+mwdf <- readRDS(file.path(outputdir,
+                          "B_hat_strong_df_19000topSNPs_FL50_Midwest.rds"))
+
+summary(mwdf) # NE, OK random Bhats are all 0's. Shat random are all vv small.
+
+k=2
+phe_suffix <- paste("FL50", subpop_v[k], 
+                    sep = "*.*")
+snp <- snp_attach(inputfiles$SNPfiles[k])
+gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
+                                      pattern = paste0("GWAS_datatable*.*",
+                                                       phe_suffix))
+gwas_rds <- gwas_rds[c(1:2,5:8)]
+
+phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
+    str_replace(., "_$", "")
+phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
+numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
+out1phe <- pvdiv_bigsnp2mashr(path = file.path(outputdir), 
+                           snp = snp, gwas_rds = gwas_rds, 
+                           phenotypes = phenotypes, numSNPs = numSNPs, 
+                           clump = TRUE, scaled = TRUE, saveoutput = TRUE,
+                           suffix = paste0("FL50_", names(subpop_v)[k]))
+
+U_hyp <- list()
+ggcov <- list()
+phe_hyp <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d")
+
+for(i in seq_along(phe_hyp)){
+  for(k in seq_along(subpop_v2)){
+
+  cov_df <- phe_fl_df %>%
+    ungroup() %>%
+    left_join(select(sites, SITE, manu_site), by = "SITE") %>%
+    filter(SUBPOP %in% subpop_v2[[k]]) %>%
+    select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
+    group_by(PLANT_ID, SUBPOP, manu_site) %>%
+    mutate(IND = row_number()) %>%
+    pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
+    mutate(PLANT_ID = as.factor(PLANT_ID),
+           IND = as.factor(IND)) %>%
+    select(-IL) %>%
+    select(PLANT_ID, SUBPOP, IND, MI, MO, SD, TX1, TX2, TX3)
+
+  cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
+  diag(cor_phe) <- cov_df %>%
+    ungroup() %>%
+    summarise(MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),
+              MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
+              SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
+              TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
+              TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
+              TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE)) %>%
+    as_vector()
+
+  #ggcov1 <- pvdiv_plot_cov(cor_phe) + ggtitle(label = paste0(phe_hyp[i], "_", names(subpop_v2)[k]))
+  #ggcov <- list.prepend(ggcov, ggcov1)
+  U_hyp <- list.prepend(U_hyp, cor_phe)
+  names(U_hyp)[1] <- paste0(phe_hyp[i], "_", names(subpop_v2)[k])
+  }
+}
+
+for(k in c(2)){
+  phe_suffix <- paste("FL50", subpop_v[k], 
+                      sep = "*.*")
+  gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
+                                        pattern = paste0("GWAS_datatable*.*",
+                                                         phe_suffix))
+  phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
+      str_replace(., "_$", "")
+  phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
+  numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
+  m_out <- mash_standard_run(path = outputdir, numSNPs = numSNPs, 
+                             U_hyp = U_hyp,
+                             suffix = paste0("FL50_", names(subpop_v)[k]),
+                             saveoutput = TRUE)
+  }
+```
+
+# -----------------------------
+
 # 3. Analyse mash output

### 2020-12-03 R/Analysis_v0.5_Gulf_mash_FL50_weather_Uhyp.R

@@ -1,0 +1,97 @@
+# Gulf mash on FL50 non-interactive
+
+## Load data
+library(tidyverse)
+library(bigsnpr)
+library(mashr)
+library(switchgrassGWAS)
+library(rlist)
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
+metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
+sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
+subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
+phe_gr_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
+phe_fl_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
+k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
+
+site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
+               NE = "NE", MI = "MI", SD = "SD")#, eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
+subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M",
+                 Gulf = "229g_10.3M")
+subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest",
+                  Gulf = "Gulf")
+phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
+                "crain_1d", "GR50", "cgdd_12c_10d")
+phe_mult <- list(flowering_5 = c("FL50", "dyln_fl50", "cgdd_12c_gr2fl",
+                                 "crain_gr2fl", "crain_1d"),
+                 greenup_2 = c("GR50", "cgdd_12c_10d"))
[…]

### 2020-12-03 R/Analysis_v0.5_Gulf_and_Midwest_mash_FL50_weather_Uhyp.R
Add noninteractive mash FL50 scripts for Midwest & G&M with weather Uhyp

@@ -1,0 +1,97 @@
+# Gulf mash on FL50 non-interactive
+
+## Load data
+library(tidyverse)
+library(bigsnpr)
+library(mashr)
+library(switchgrassGWAS)
+library(rlist)
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
+metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
+sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
+subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
+phe_gr_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
+phe_fl_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
+k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
+
+site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
+               NE = "NE", MI = "MI", SD = "SD")#, eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
+subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M",
+                 Gulf = "229g_10.3M")
+subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest",
+                  Gulf = "Gulf")
+phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
+                "crain_1d", "GR50", "cgdd_12c_10d")
+phe_mult <- list(flowering_5 = c("FL50", "dyln_fl50", "cgdd_12c_gr2fl",
+                                 "crain_gr2fl", "crain_1d"),
+                 greenup_2 = c("GR50", "cgdd_12c_10d"))
[…]

### 2020-12-03 R/Analysis_v0.5_Midwest_mash_FL50_weather_Uhyp.R

@@ -1,0 +1,102 @@
+# Midwest mash on FL50 non-interactive
+
+## Load data
+library(tidyverse)
+library(bigsnpr)
+library(mashr)
+library(switchgrassGWAS)
+library(rlist)
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
+metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
+sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
+subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
+phe_gr_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
+phe_fl_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
+k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
+
+site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
+               NE = "NE", MI = "MI", SD = "SD")#, eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
+subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M",
+                 Gulf = "229g_10.3M")
+subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest",
+                  Gulf = "Gulf")
+phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
+                "crain_1d", "GR50", "cgdd_12c_10d")
+phe_mult <- list(flowering_5 = c("FL50", "dyln_fl50", "cgdd_12c_gr2fl",
+                                 "crain_gr2fl", "crain_1d"),
[…]

### 2020-12-03 R/Analysis_v0.5_Midwest_mash_FL50_weather_Uhyp.R
Edits to Midwest mash run with six gardens

@@ -39,6 +39,12 @@ U_hyp <- list()
 ggcov <- list()
 phe_hyp <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d")
 
+k=2
+phe_suffix <- paste("FL50", subpop_v[k],
+                    sep = "*.*")
+gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir),
+                                    pattern = paste0("GWAS_datatable*.*",
+                                                     phe_suffix))
 gwas_rds <- gwas_rds[c(1:2,5:8)]
 
 phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
@@ -85,18 +91,8 @@ for(i in seq_along(phe_hyp)){
   }
 }
 
-for(k in c(2)){
-  phe_suffix <- paste("FL50", subpop_v[k],
-                      sep = "*.*")
-  gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir),
-                                      pattern = paste0("GWAS_datatable*.*",
-                                                       phe_suffix))
-  phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
-    str_replace(., "_$", "")
-  phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
-  numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
+k=2
   m_out <- mash_standard_run(path = outputdir, numSNPs = numSNPs,
                              U_hyp = U_hyp,
                              suffix = paste0("FL50_", names(subpop_v)[k]),
                              saveoutput = TRUE)
-}

### 2020-12-05 R/Analysis_v0.5_weather_covariance_Uhyp_flowering.Rmd
Add Mashhattan plotted by covariance matrix posterior weights exploration for FL50

@@
 library(rlist)
+library(rlang)
 ```
@@
   2b. Compute correlations for Gulf, Midwest, & G&M subsets & compare
-3. Experiment with changing within-site covariance to coefficient of variation for each site. Wouldn't expect these to be 1, really.
+3.
+Experiment with changing within-site covariance to coefficient of variation for each site.
+Wouldn't expect these to be 1, really.
+Although mash does set the maximum of the diagonal to 1...
+hmm.
+I wonder why.
 
@@
 
-### plot covariance matrices to look at their differences
+## 1a. Covar plot function
+Function to plot covariance matrices to look at their differences
 ```{r}
@@
 
+## 1b. Make U_hyp
 ```{r}
@@
 
-## Set up mash inputs
+## 2a. Set up mash inputs
 
-### mash G&M
+### i. mash G&M
 ```{r}
@@
 
-## Run mash Gulf
+### ii. Run mash Gulf
 ```{r}
@@
 # 3. Analyse mash output
+## -- analysis outline --
+•	How many SNP effects load onto each covariance matrix?
+•	Mashhattans colored by covariance matrices.
+•	What patterns do the data-driven matrices capture that our weather matrices don’t capture?
+•	We characterized patterns of differential sensitivity and antagonistic pleiotropy between all SNPs and all pairs of gardens.
+•	We looked for enrichments of genes that could potentially affect flowering in the 20kb genomic windows surrounding the SNPs with significant effects on flowering.
+•	Rerun greenup mash and rewrite greenup mash section.
+
+## 3a. Covar Loadings
+How many SNP effects load onto each covariance matrix?
+```{r}
+gulfm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/Strong_Effects19000SNPs_FL50_Gulf.rds")
+midwm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/Strong_Effects33000SNPs_FL50_Midwest.rds")
+gandm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/Strong_Effects19000SNPs_FL50_Gulf_and_Midwest.rds")
+```
+
+### i. Midwest
+```{r}
+m <- midwm
+length(get_significant_results(m))  # 9907/63367
+get_significant_results(m)[1:10]
+
+covarm <- mash_plot_covar(m, saveoutput = TRUE)
+get_U_by_mass(m)
+mash_plot_Ulist(m, range = get_U_by_mass(m))
+mash_plot_effects(m, n = 1, saveoutput = TRUE)
+mash_plot_sig_by_condition(m, saveoutput = TRUE)
+mash_plot_manhattan_by_condition(m, saveoutput = TRUE)
+
+mash_out <- pvdiv_results_in_folder(outputdir, "Strong")
+mash_plot_pairwise_sharing(effectRDS = file.path(outputdir, mash_out[3])) # sign
+mash_plot_pairwise_sharing(effectRDS = file.path(outputdir, mash_out[6])) # sign & size
+Midw_greenup_gxe <- get_GxE(m)
+
+mash_plot_pairwise_sharing(corrmatrix = Midw_greenup_gxe$S_AP/Midw_greenup_gxe$S_all_pairwise, saveoutput = TRUE)
+mash_plot_pairwise_sharing(corrmatrix = Midw_greenup_gxe$S_DS/Midw_greenup_gxe$S_all_pairwise, reorder = TRUE, saveoutput = TRUE)
+mash_plot_pairwise_sharing(corrmatrix = Midw_greenup_gxe$S_2_no, reorder = FALSE)
+mash_plot_pairwise_sharing(corrmatrix = Midw_greenup_gxe$S_all_pairwise, reorder = FALSE)
+
+```
+
+### ii. G&M
+```{r}
+length(get_significant_results(gandm))  # 24518/35735
+get_significant_results(gandm)[1:10]
+
+covargam <- mash_plot_covar(gandm, saveoutput = TRUE)
+get_U_by_mass(gandm)
+mash_plot_Ulist(gandm, range = get_U_by_mass(gandm), limits = FALSE)
+```
+
+### iii. Gulf
+```{r}
+length(get_significant_results(gulfm))  # 24518/35735
+get_significant_results(gulfm)[1:10]
+
+covarG <- mash_plot_covar(gulfm, saveoutput = TRUE)
+get_U_by_mass(gulfm)
+mash_plot_Ulist(gulfm, range = get_U_by_mass(gulfm))
+```
+```{r}
+covargam$covar_df %>% 
+  mutate(Type = case_when(grepl("ED_", `Covariance Matrix`) ~ "data-driven",
+                          grepl("single_effect", `Covariance Matrix`) ~ "canonical",
+                          grepl("simple_het", `Covariance Matrix`) ~ "canonical",
+                          grepl("no_effects", `Covariance Matrix`) ~ "canonical",
+                          grepl("identity", `Covariance Matrix`) ~ "canonical",
+                          grepl("equal_effects", `Covariance Matrix`) ~ "canonical",
+                          grepl("_and_Midwest$", `Covariance Matrix`) ~ "GAM_hyp",
+                          grepl("Midwest$", `Covariance Matrix`) ~ "Midw_hyp",
+                          grepl("Gulf$", `Covariance Matrix`) ~ "Gulf_hyp",
+                          TRUE ~ "other"
+                          )) %>%
+  group_by(Type) %>%
+  summarise(Mass = sum(Mass))
+
+21.65+10.9795
+covarG$ggobject
+covarG$covar_df
+8.38457+18.782188+18.38669
+```
+
+## 3b.	Mashhattans 
+Mashhattans colored by covariance matrices.
+
+### i. Function
+
+I'm not sure that this function is really the way to go... I don't understand how SNPs really load onto these different covariance matrices.
+
+Seems that the posterior weights for each SNP sum to one across this whole covariance grid.
+Sometimes there is a pw near 1, but more often they're a mix.
+So I think this is still useful for exploratory purposes - which covar have the max posterior weights?
+But I probably want to plot the posterior weights for each cov matrix one by one, which is much simpler to accomplish anyway.
+```{r}
+#' Return the name of the covariance matrix with the highest posterior weight
+#'     for each effect
+#'
+#' @param m the mash result (from joint or 1by1 analysis)
+#'
+#' @return returns a df with Marker, Covariance_Matrix, and Grid_Value. I'm assuming that the grid value is an indication of like a svd decomp
+#'
+get_topcov = function(m){
+  df <- switchgrassGWAS:::get_marker_df(m = m)
+  covdf <- purrr::map_chr(df$value, ~ names(which(m$posterior_weights[.x,] ==
+                                           max(m$posterior_weights[.x,])))) %>%
+    enframe(name = "value", value = "Covariance_Grid") %>%
+    separate("Covariance_Grid", into = c("Covariance_Matrix", "Grid_Value"),
+             sep = "\\.", remove = FALSE) 
+  df <- df %>% full_join(covdf, by = "value") %>% 
+  mutate(Type = case_when(grepl("ED_", Covariance_Matrix) ~ "data-driven",
+                          grepl("single_effect", Covariance_Matrix) ~ "canonical",
+                          grepl("simple_het", Covariance_Matrix) ~ "canonical",
+                          grepl("no_effects", Covariance_Matrix) ~ "none",
+                          grepl("identity", Covariance_Matrix) ~ "canonical",
+                          grepl("equal_effects", Covariance_Matrix) ~ "equal",
+                          grepl("_and_Midwest$", Covariance_Matrix) ~ "GAM_hyp",
+                          grepl("Midwest$", Covariance_Matrix) ~ "Midw_hyp",
+                          grepl("Gulf$", Covariance_Matrix) ~ "Gulf_hyp",
+                          grepl("_TX2$", Covariance_Matrix) ~ "canonical",
+                          grepl("_TX1$", Covariance_Matrix) ~ "canonical",
+                          grepl("_TX3$", Covariance_Matrix) ~ "canonical",
+                          grepl("_M$", Covariance_Matrix) ~ "canonical",
+                          grepl("_MO$", Covariance_Matrix) ~ "canonical",
+                          grepl("_SD$", Covariance_Matrix) ~ "canonical",
+                          grepl("_OK$", Covariance_Matrix) ~ "canonical",
+                          grepl("_NE$", Covariance_Matrix) ~ "canonical",
+                          grepl("_MI$", Covariance_Matrix) ~ "canonical",
+                          TRUE ~ "other"
+                          ))
+  return(df)
+}
+```
+
+### ii. Plot
+```{r}
+m <- midwm
+log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
+    as.data.frame() %>%
+    rownames_to_column(var = "value") %>%
+    mutate(value = as.integer(.data$value)) %>%
+    as_tibble() %>%
+    left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
+    dplyr::rename(log10BayesFactor = .data$V1) %>%
+    dplyr::select(-.data$value)
+
+#df <- get_topcov(m)
+
+pw_df <- as_tibble(m$posterior_weights)
+pw_df <- switchgrassGWAS:::get_marker_df(m = m) %>%
+  bind_cols(pw_df)
+
+ggman_df <- pw_df %>%
+  #left_join(m_meta) %>%
+  separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
+           extra = "merge", convert = TRUE) %>%
+  left_join(log10bf_df, by = "Marker") %>%
+  arrange(.data$Chr, .data$Pos)
+
+log10BF <- expression(paste("log"[10], plain("(Bayes Factor)")))
+
+mash_plot_covar(m)
+```
+
+```{r}
+gglist <- list()
+for(i in seq_along(names(pw_df)[-(1:2)])){
+var <- names(pw_df)[i+2]   # 20 or "Stand_BhatMidwest_FL50_M.21" is GIGO
+gglist[[i]] <- ggman_df %>%
+  #filter(Type != "canonical") %>%
+  #group_by(crain_gr2fl_Gulf.8 > 0.01) %>% add_tally() %>%
+  arrange(!! sym(var)) %>%
+  ggplot(aes(x = .data$Pos, y = .data$log10BayesFactor)) +
+  switchgrassGWAS::theme_oeco +
+  geom_point(aes(color = !! sym(var), fill = !! sym(var),
+                 shape = as.factor(.data$Chr))) +
+  facet_wrap(~ .data$Chr, nrow = 1, scales = "free_x", strip.position = "bottom") +
+  scale_color_viridis_c(option = "D", end = 0.95) + 
+  scale_fill_viridis_c(option = "D", end = 0.95) +
+  theme(strip.text = element_text(size = 8),
+        axis.text.y = element_text(size = 8),
+        axis.text.x = element_blank(),
+        legend.text = element_text(size = 8),
+        axis.ticks.x = element_blank(),
+        panel.background = element_rect(fill=NA)) +
+  labs(x = "Chromosome", y = log10BF) +
+  scale_x_continuous(expand = c(0.32, 0.32)) +
+  scale_shape_manual(values = rep(c(21,22),9), guide = FALSE) +
+  geom_hline(yintercept = 2, linetype = 2) +
+  ggtitle(var)
+}
+gglist
+```
+Midwest 4 cgdd gr2fl G&M background, everywhere, v low pw
+
+G&M Mashhattans plotting SNP cov matrix posterior weights vs log10BF
+4 "dyln_fl50_Gulf.3" lots of background
+5 same lots of background
+6 same background. Some above 2: Chr07N, Chr05K. 
+7 same background. Some above 2 but all v low posterior weights.
+8 background. Some above 2 Chr04N Chr09K.
+9 PCA1 everywhere 
+13 FL50 TX1 some very high log10BF and high posterior weights on many Chr.
+
+Gulf Mashhattans by posterior weights for single covar mat/grids:
+Gulf subpop covar matrix 11 (equal effects) has really interesting results - high posterior weights and high log10BF for several regions of the genome, esp end of Chr05N, end of Chr04K, Chr06N. Cool!
+ED tPCA and ED PCA1 are of course everywhere. 19 & 17 especially.
+NE one specific effect on Chr02N. (14, 15), maybe on Chr01N also.
+SD (18) some specific high log10BF hits also: Chr03K, Chr09N, Chr09K, others.
+TX1 (20) some specific high log10BF hits also: Chr07K, Chr05N, Chr09N, others.
+TX2 (21) similar: Chr09N, Chr09N, Chr03N.
+
+## 3c. Data-driven mat
+
+What patterns do the data-driven matrices capture that our weather matrices don’t capture?
+
+SVD and plots of effects for the ED_PCA matrices.
+
+## 3d. DS & AP
+
+We characterized patterns of differential sensitivity and antagonistic pleiotropy between all SNPs and all pairs of gardens.
+
+## 3e. Enrichments
+We looked for enrichments of genes that could potentially affect flowering in the 20kb genomic windows surrounding the SNPs with significant effects on flowering.
+
+## 3f. Loglikelihood
+
+log likelihood of models with and without U_hyp.
+
+# -----------------------------
+## In another script:
+•	Rerun greenup mash and rewrite greenup mash section.
+
+•	Rerun FL50 mash for Midwest without MI (or with reduced #'s for MI, OK, NE?). MI showing pop structure effects in mash results.

### 2021-01-05 R/Analysis_v0.5_snp_allele_frequencies_and_other_info.Rmd
Start greenup mash analysis using weather covariances matrices as U_hyp

@@
 ---
-
+# TRIAL script
+# -----------------------------
+NB: Trial to get other metadata for mash SNPs and to plot that data as colors in the mashhattan to add information to these plots.
 ```{r setup, include=FALSE}

### 2021-01-05 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Start greenup mash analysis using weather covariances matrices as U_hyp

---
title: "Weather covariance matrices as U_hyp for GR50"
author: "Alice MacQueen"
date: 2021-01-05
output: html_document
---

# TRIAL script
# -------------------------
NB: Trial script on mash-env-cov branch to use weather covariance matrices as 
U_hyp in a mash run of GR50 phenotypes from 8 gardens.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(switchgrassGWAS)
library(mashr)
```

# Load project data

```{r}
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
mashdir <- file.path(workingdir, "analysis", "mash")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
# pvdiv PLANT_ID metadata
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
# pvdiv common garden metadata
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
# all greenup related phenotypes, including greenup as functions of weather variables.
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
# all flowering related phenotypes, including functions of weather variables.
k_full <- read_rds(file.path(workingdir, "data",
                    "Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds"))
# kinship matrix for 630 tetraploid individuals.

## vectors with subsets of gardens, phenotypes, and genetic subpopulations.
site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
               NE = "NE", MI = "MI", SD = "SD")#, eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", 
                 Gulf = "229g_10.3M")
subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", 
                 Gulf = "Gulf")
phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
               "crain_1d", "GR50", "cgdd_12c_10d")
phe_mult <- list(flowering_5 = c("FL50", "dyln_fl50", "cgdd_12c_gr2fl",
                                 "crain_gr2fl", "crain_1d"), 
                 greenup_2 = c("GR50", "cgdd_12c_10d"))
inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
                                   "inputkinship.txt"),  delim = " ", 
                                   col_names = "SNPfiles") 
# paths to large SNP files
```

# 1. Weather U_hyp

Make weather corr/covar matrices

1. Make dataframe of each env variable across all 8 sites
2. Compute correlations with pairwise complete observations
3. Change within-site covariance to coefficient of variation for each site. NB: mash does set the maximum of the diagonal to 1... hmm. I wonder why.

Do this for three pop subsets for 'cgdd_12c_10d' and maybe for 'cgdd_12c_18d' and 'cgdd_12c_5d'.

Function to plot covariance matrices to look at their differences
```{r}
pvdiv_plot_cov <- function(covdf){
  U1 <- cor_phe
  if(isSymmetric(U1)){
        for(i in 1:nrow(U1)){
          for(j in 1:ncol(U1)){
            if(i < j){
              U1[i, j] <- NA
            }
          }
        }
  }
  U1 <- as_tibble(U1, rownames = "rowU", .name_repair = "unique") %>%
          pivot_longer(cols = -.data$rowU, names_to = "colU",
                       values_to = "covar") %>%
          filter(!is.na(.data$covar))
  U1$colU <- factor(U1$colU, levels = colnames(cor_phe))
  U1$rowU <- factor(U1$rowU, levels = colnames(cor_phe))

  U1_covar <- U1 %>%
    ggplot(aes(x = .data$rowU, y = .data$colU)) +
    switchgrassGWAS::theme_oeco +
    geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
    scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                    "white", "#27AD81FF", "#5DC863FF",
                                    "#FDE725FF"), limits = c(-1,1)) +
    geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
    theme(legend.position = c(0.2, 0.9),
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    xlab("") + ylab("") 
  return(U1_covar)
}

```

## 1a. Make U_hyp

Specify in the phe_hyp vector which functions of weather I want to make U_hyp for mash.
```{r}
U_hyp <- list()
ggcov <- list()
phe_hyp <- c("GR50", "cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", "tave_10d")

for(i in seq_along(phe_hyp)){
  for(k in seq_along(subpop_v2)){

  cov_df <- phe_gr_df %>%
    ungroup() %>%
    left_join(select(sites, SITE, manu_site), by = "SITE") %>%
    filter(SUBPOP %in% subpop_v2[[k]]) %>%
    select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
    group_by(PLANT_ID, SUBPOP, manu_site) %>%
    mutate(IND = row_number()) %>%
    pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
    mutate(PLANT_ID = as.factor(PLANT_ID),
           IND = as.factor(IND)) %>%
    select(-IL) %>%
    select(PLANT_ID, SUBPOP, IND, MI, MO, NE, OK, SD, TX1, TX2, TX3)

  cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
  diag(cor_phe) <- cov_df %>%
    ungroup() %>%
    summarise(MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),
              MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
              NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
              OK = sqrt(var(OK, na.rm = TRUE))/mean(OK, na.rm = TRUE),
              SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
              TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
              TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
              TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE)) %>%
    as_vector()

  ggcov1 <- pvdiv_plot_cov(cor_phe) + ggtitle(label = paste0(phe_hyp[i], "_", names(subpop_v2)[k]))
  ggcov <- list.prepend(ggcov, ggcov1)
  U_hyp <- list.prepend(U_hyp, cor_phe)
  names(U_hyp)[1] <- paste0(phe_hyp[i], "_", names(subpop_v2)[k])
  }
}

ggcov
```

# 2. Make mashr inputs for 8 GR50 phenotypes & run mash.

Some univariate GWAS for the Midwest had population structure issues. Mostly for greenup - even GR50 at MO was quite bad. Drop MO for GR50. Otherwise they look ok for FL50 and GR50.

## 2a. Gulf & Gulf&Midwest.

### mash inputs 
NB: Do this on server, where the large GWAS_datatable objects were created.
```{r}
library
k = 3 # subpop: G&M, M, Gulf
for(k in c(1, 3)){
phe_suffix <- paste("GR50", subpop_v[k], 
                    sep = "*.*")
snp <- snp_attach(inputfiles$SNPfiles[k])
gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
                                      pattern = paste0("GWAS_datatable*.*",
                                                       phe_suffix))
phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
    str_replace(., "_$", "")
phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
out1phe <- pvdiv_bigsnp2mashr(path = file.path(outputdir), 
                           snp = snp, gwas_rds = gwas_rds, 
                           phenotypes = phenotypes, numSNPs = numSNPs, 
                           clump = TRUE, scaled = TRUE, saveoutput = TRUE,
                           suffix = paste0("GR50_", names(subpop_v)[k]))
}
```

Use scp to move these output files (which all have the same suffix) to genegenie.

e.g. 
`scp -i Github/jls_alice alice@popgen.biosci.utexas.edu:~/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/*19000topSNPs_GR50_Midwest*  /home/alice/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/Midwest/`

`scp -i ~/Github/jls_alice alice@popgen.biosci.utexas.edu:~/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/Part-One-Output*  /home/alice/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/
`

### run mash 
```{r}
for(k in c(1, 3)){

  numSNPs <- 19000
  m_out <- mash_standard_run(path = file.path(outputdir, names(subpop_v)[k]),
                             numSNPs = numSNPs, 
                             U_hyp = U_hyp,
                             suffix = paste0("GR50_", names(subpop_v)[k]),
                             saveoutput = TRUE)
}

inputfiles
```

## 2b. Midwest, dropping some gardens.

# -----------------------------

# 3. Analyse mash output
## -- analysis outline --
•	How many SNP effects load onto each covariance matrix?
•	Mashhattans colored by covariance matrices.
•	What patterns do the data-driven matrices capture that our weather matrices don’t capture?
•	We characterized patterns of differential sensitivity and antagonistic pleiotropy between all SNPs and all pairs of gardens.
•	We looked for enrichments of genes that could potentially affect flowering in the 20kb genomic windows surrounding the SNPs with significant effects on flowering.
•	Rerun greenup mash and rewrite greenup mash section.

### 2021-01-12 R/Analysis_v0.5_weather_covariance_Uhyp_flowering.Rmd
Move mash files to analysis/mash directory

@@
 ```{r}
-gulfm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/Strong_Effects19000SNPs_FL50_Gulf.rds")
-midwm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/Strong_Effects33000SNPs_FL50_Midwest.rds")
-gandm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/Strong_Effects19000SNPs_FL50_Gulf_and_Midwest.rds")
+gulfm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects19000SNPs_FL50_Gulf.rds")
+midwm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects33000SNPs_FL50_Midwest.rds")
+gandm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects19000SNPs_FL50_Gulf_and_Midwest.rds")
 ```

### 2021-01-12 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Move mash files to analysis/mash directory

@@
 library(mashr)
+library(rlist)
 ```
@@
 ```{r}
-for(k in c(1, 3)){
+for(k in c(1,3,2)){
 
   numSNPs <- 19000
-  m_out <- mash_standard_run(path = file.path(outputdir, names(subpop_v)[k]),
+  m_out <- mash_standard_run(path = file.path(outputdir), #, names(subpop_v)[k]), # use to run on genegenie within mash dirs
                              numSNPs = numSNPs, 
@@
 •	Rerun greenup mash and rewrite greenup mash section.
+
+## 3a. Covar Loadings
+How many SNP effects load onto each covariance matrix?
+```{r}
+gulfm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects19000SNPs_GR50_Gulf.rds")
+
+gandm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects19000SNPs_GR50_Gulf_and_Midwest.rds")
+```
+44765 SNPs tested for G&M, 45931 SNPs tested for Gulf.
+
+Midwest I haven't run successfully yet as of 2021-01-12 - I will need to remove some locations with all-zero effects, probably.
+
+### ii. Gulf
+```{r}
+length(get_significant_results(gulfm))  # 9142/45931  # 19.9%
+get_significant_results(gulfm)[1:10]
+
+covargulf <- mash_plot_covar(gulfm, suffix = "19000SNPs_GR50_Gulf", saveoutput = TRUE)
+covargulf
+get_U_by_mass(gulfm)
+mash_plot_Ulist(gulfm, range = get_U_by_mass(gulfm), limits = FALSE)
+mash_plot_manhattan_by_condition(gulfm, saveoutput = FALSE)
+mash_plot_sig_by_condition(gulfm, saveoutput = FALSE)
+```
+
+### ii. G&M
+```{r}
+length(get_significant_results(gandm))  # 15104/44765  # 33.7%
+get_significant_results(gandm)[1:10]
+
+covargam <- mash_plot_covar(gandm, suffix = "19000SNPs_GR50_Gulf_and_Midwest", saveoutput = TRUE)
+covargam
+get_U_by_mass(gandm)
+mash_plot_Ulist(gandm, range = get_U_by_mass(gandm), limits = FALSE)
+mash_plot_manhattan_by_condition(gandm, saveoutput = FALSE)
+mash_plot_sig_by_condition(gandm, saveoutput = FALSE)
+```
+
+```{r}
+covargam$covar_df %>% 
+  mutate(Type = case_when(grepl("ED_", `Covariance Matrix`) ~ "data-driven",
+                          grepl("single_effect", `Covariance Matrix`) ~ "canonical",
+                          grepl("simple_het", `Covariance Matrix`) ~ "canonical",
+                          grepl("no_effects", `Covariance Matrix`) ~ "canonical",
+                          grepl("identity", `Covariance Matrix`) ~ "canonical",
+                          grepl("equal_effects", `Covariance Matrix`) ~ "canonical",
+                          grepl("_and_Midwest$", `Covariance Matrix`) ~ "GAM_hyp",
+                          grepl("Midwest$", `Covariance Matrix`) ~ "Midw_hyp",
+                          grepl("Gulf$", `Covariance Matrix`) ~ "Gulf_hyp",
+                          TRUE ~ "other"
+                          )) %>%
+  group_by(Type) %>%
+  summarise(Mass = sum(Mass))
+```
+Gulf: 20% canonical, 68.5% data-driven, 9.5% Gulf hyp, 1.6% Midwest hyp
+Gulf & Midwest: 28% canonical, 62% data-driven, 5.5% Gulf, 1.2% Midwest, 2.8% G&M
+
+Not stunning, especially compared to how the flowering hyp do.
+
+## 3b.	Mashhattans 
+Mashhattans colored by covariance matrices.
+
+### ii. Plot setup
+
+Get posterior weight dataframe for the covariance matrix grid. Per SNP/row, entries in this grid sum to 1.
+
+Get log10BayesFactor for mash SNPs also and combine these into a dataframe for ggplot.
+```{r}
+m <- gandm
+log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
+    as.data.frame() %>%
+    rownames_to_column(var = "value") %>%
+    mutate(value = as.integer(.data$value)) %>%
+    as_tibble() %>%
+    left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
+    dplyr::rename(log10BayesFactor = .data$V1) %>%
+    dplyr::select(-.data$value)
+
+#df <- get_topcov(m)
+
+pw_df <- as_tibble(m$posterior_weights)
+pw_df <- switchgrassGWAS:::get_marker_df(m = m) %>%
+  bind_cols(pw_df)
+
+ggman_df <- pw_df %>%
+  #left_join(m_meta) %>%
+  separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
+           extra = "merge", convert = TRUE) %>%
+  left_join(log10bf_df, by = "Marker") %>%
+  arrange(.data$Chr, .data$Pos)
+
+log10BF <- expression(paste("log"[10], plain("(Bayes Factor)")))
+
+mash_plot_covar(m)
+```
+
+Make a list containing all of the ggplots for all entries in the covariance grid represented by `pw_df`.
+Then view all the ggplots in gglist.
+```{r}
+gglist <- list()
+for(i in seq_along(names(pw_df)[-(1:2)])){
+var <- names(pw_df)[i+2]   
+gglist[[i]] <- ggman_df %>%
+  arrange(!! sym(var)) %>%
+  ggplot(aes(x = .data$Pos, y = .data$log10BayesFactor)) +
+  switchgrassGWAS::theme_oeco +
+  geom_point(aes(color = !! sym(var), fill = !! sym(var),
+                 shape = as.factor(.data$Chr))) +
+  facet_wrap(~ .data$Chr, nrow = 1, scales = "free_x", strip.position = "bottom") +
+  scale_color_viridis_c(option = "D", end = 0.95) + 
+  scale_fill_viridis_c(option = "D", end = 0.95) +
+  theme(strip.text = element_text(size = 8),
+        axis.text.y = element_text(size = 8),
+        axis.text.x = element_blank(),
+        legend.text = element_text(size = 8),
+        axis.ticks.x = element_blank(),
+        panel.background = element_rect(fill=NA)) +
+  labs(x = "Chromosome", y = log10BF) +
+  scale_x_continuous(expand = c(0.32, 0.32)) +
+  scale_shape_manual(values = rep(c(21,22),9), guide = FALSE) +
+  geom_hline(yintercept = 2, linetype = 2) +
+  ggtitle(var)
+}
+gglist
+```
+
+Which covariance grids have SNPs with log10BF > 2 that load strongly onto them?
+For Gulf mash:
+GR50_Gulf.6: many, almost all < 2
+GR50_Midwest.6: many, almost all < 2
+equal_effects.7: many, almost all < 2
+GR50_Gulf.6.7: many, almost all < 2
+equal_effects.8: some, all <2
+ED_PCA_1.9 many, many > 2
+ED_PCA_2.9 many, few > 2
+ED_PCA_4.9 some, few > 2
+ED_tPCA.9: many, many > 2
+ED_PCA_3.10: many, few > 2
+ED_PCA_5.10: some, few > 2
+OK.10 many, some > 2, many >0.8 loading & > 2
+SD.10 many, some > 2
+cgdd_12c_G&M.10: few, few > 2, some with 80%+ loading here
+
+ED_tPCA.12: many, many > 2, many >0.8 loading esp at high log10BF
+SD.12: many, many > 2, many > 0.8 loading esp at high log10BF
+
+TX1.15: few bg, few >2
+
+```{r}
+var <- names(ggman_df)[ncol(ggman_df)-1]  
+ggman_df %>%
+  pivot_longer(cols = null:var, names_to = "covar_grid", values_to = "loading") %>%
+  filter(loading > 0.5) %>%
+  group_by(log10BayesFactor > 2, covar_grid, loading > 0.8) %>%
+  tally()
+```
+## 3c. Data-driven mat
+
+What patterns do the data-driven matrices capture that our weather matrices don’t capture?
+
+SVD and plots of effects for the ED_PCA matrices.
+
+## 3d. DS & AP
+
+We characterized patterns of differential sensitivity and antagonistic pleiotropy between all SNPs and all pairs of gardens.
+
+## 3e. Enrichments
+We looked for enrichments of genes that could potentially affect flowering in the 20kb genomic windows surrounding the SNPs with significant effects on flowering.
+
+## 3f. Loglikelihood
+
+log likelihood of models with and without U_hyp.

### 2021-01-13 R/Analysis_v0.5_weather_covariance_Uhyp_flowering.Rmd
Add nohup scripts and mash metadata

@@
 ```{r}
-gulfm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects19000SNPs_FL50_Gulf.rds")
-midwm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects33000SNPs_FL50_Midwest.rds")
+gulfm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects19000SNPs_FL50_Gulf.rds")
+midwm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects33000SNPs_FL50_Midwest.rds")
 gandm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects19000SNPs_FL50_Gulf_and_Midwest.rds")

### 2021-01-13 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Add nohup scripts and mash metadata

@@
 
-```{r}
+```{r load data}
 source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
@@
 Function to plot covariance matrices to look at their differences
-```{r}
+```{r pvdiv_plot_cov function}
 pvdiv_plot_cov <- function(covdf){
@@
 Specify in the phe_hyp vector which functions of weather I want to make U_hyp for mash.
-```{r}
+```{r make u_hyp}
 U_hyp <- list()
@@
 NB: Do this on server, where the large GWAS_datatable objects were created.
-```{r}
+```{r mash inputs with gulf}
 library
@@
 
+Files with the six suffixes I have/want mash results for:
+B_hat_strong_df_19000topSNPs_FL50_Gulf_and_Midwest.rds
+B_hat_strong_df_19000topSNPs_FL50_Gulf.rds
+B_hat_strong_df_19000topSNPs_GR50_Gulf_and_Midwest.rds
+B_hat_strong_df_19000topSNPs_GR50_Gulf.rds
+B_hat_strong_df_24000topSNPs_GR50_Midwest.rds
+B_hat_strong_df_33000topSNPs_FL50_Midwest.rds
+
 ### run mash 
-```{r}
+```{r run mash with gulf}
 for(k in c(1,3,2)){
@@
 
-## 2b. Midwest, dropping some gardens.
+## 2b. Midwest, dropping gardens with zero values.
+Midwest_GR50_SD have values of zero for Bhat and Shat. Drop this from Midwest analyses.
+```{r check midwest}
+mwdf <- readRDS(file.path(outputdir,
+                          "B_hat_strong_df_19000topSNPs_GR50_Midwest.rds"))
+
+summary(mwdf)
+```
+
+```{r mash midwest}
+k=2
+phe_suffix <- paste("GR50", subpop_v[k], 
+                    sep = "*.*")
+snp <- snp_attach(inputfiles$SNPfiles[k])
+gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
+                                      pattern = paste0("GWAS_datatable*.*",
+                                                       phe_suffix))
+gwas_rds <- gwas_rds[c(1:4,6:8)]
+
+phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -41) %>%
+    str_replace(., "_$", "")
+phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
+numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
+out1phe <- pvdiv_bigsnp2mashr(path = file.path(outputdir), 
+                           snp = snp, gwas_rds = gwas_rds, 
+                           phenotypes = phenotypes, numSNPs = numSNPs, 
+                           clump = TRUE, scaled = TRUE, saveoutput = TRUE,
+                           suffix = paste0("GR50_", names(subpop_v)[k]))
+m_out <- mash_standard_run(path = file.path(outputdir), #, names(subpop_v)[k]), # use to run on genegenie within mash dirs
+                             numSNPs = numSNPs, 
+                             suffix = paste0("GR50_", names(subpop_v)[k]),
+                             saveoutput = TRUE)
+saveRDS(m_out, file = file.path(mashdir, paste0("Strong_Effects_", numSNPs,
+                                                "GR50_", names(subpop_v)[k],
+                                                "_no_Uhyp.rds")))
+m_out1 <- mash_standard_run(path = file.path(outputdir), numSNPs = numSNPs, 
+                           U_ed = file.path(outputdir, paste0("Model_of_",
+                           "random_tests_24000SNPs_GR50_Midwest.rds")),
+                           U_hyp = U_hyp,
+                           suffix = paste0("GR50_", names(subpop_v)[k]),
+                           saveoutput = FALSE)
+saveRDS(m_out1, file = file.path(mashdir, paste0("Strong_Effects_", numSNPs,
+                                                "GR50_", names(subpop_v)[k],
+                                                "_no_Uhyp.rds")))
+```
 
@@
 # 3. Analyse mash output
-## -- analysis outline --
-•	How many SNP effects load onto each covariance matrix?
-•	Mashhattans colored by covariance matrices.
-•	What patterns do the data-driven matrices capture that our weather matrices don’t capture?
-•	We characterized patterns of differential sensitivity and antagonistic pleiotropy between all SNPs and all pairs of gardens.
-•	We looked for enrichments of genes that could potentially affect flowering in the 20kb genomic windows surrounding the SNPs with significant effects on flowering.
-•	Rerun greenup mash and rewrite greenup mash section.
+## -- analysis outline 2020-01-12 --
+•	3a What is the log likelihood of the mash models (x3) with and without U_hyp?
+•	3b. How many SNP effects load onto each covariance matrix? What about SNP effects with log10BF > 2?
+•	3c. Mashhattans colored by covariance matrices. Report (and plot/table) SNPs with log10BF > 2 and high loading (>0.5? 0.8?) onto one covar matrix.
+  • Stress any Uhyp where the above is true and talk about candidate genes here. We have both an environmental driver and a small genomic interval here, these are of particular interest.
+•	3d. We looked for enrichments of genes that could potentially affect greenup/flowering in the 20kb genomic windows surrounding the SNPs with significant effects on flowering. 
+  •	Also, we looked at other metadata enrichments for these top SNPs.
+•	3e. What patterns do the data-driven matrices capture that our weather matrices don’t capture?
+•	3f. Finally, talk about overall SNP patterns seen in the phenotype: patterns of DS & AP.
 
-## 3a. Covar Loadings
+## 3b. Covar Loadings
 How many SNP effects load onto each covariance matrix?
-```{r}
-gulfm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects19000SNPs_GR50_Gulf.rds")
+```{r load mash objects}
+gulfm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects19000SNPs_GR50_Gulf_Uhyp.rds")
 
-gandm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects19000SNPs_GR50_Gulf_and_Midwest.rds")
+gandm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects19000SNPs_GR50_Gulf_and_Midwest_Uhyp.rds")
 ```
@@
 ### ii. Gulf
-```{r}
+```{r gulf mash results}
 length(get_significant_results(gulfm))  # 9142/45931  # 19.9%
@@
 ### ii. G&M
-```{r}
+```{r gulf and midwest mash results}
 length(get_significant_results(gandm))  # 15104/44765  # 33.7%
@@
 
-## 3b.	Mashhattans 
+## 3c.	Mashhattans 
 Mashhattans colored by covariance matrices.
@@
 Get log10BayesFactor for mash SNPs also and combine these into a dataframe for ggplot.
-```{r}
+```{r mashhattan plot setup}
 m <- gandm
@@
 Then view all the ggplots in gglist.
-```{r}
+```{r mashhattan plots loop}
 gglist <- list()
@@
 
-```{r}
+```{r snps with high sig and high covar loadings}
 var <- names(ggman_df)[ncol(ggman_df)-1]  
@@
 ```
-## 3c. Data-driven mat
+# -----------
+## 3d. Enrichments
+•	3d. We looked for enrichments of genes that could potentially affect greenup/flowering in the 20kb genomic windows surrounding the SNPs with significant effects on flowering. 
+  •	Also, we looked at other metadata enrichments for these top SNPs.
+
+mashdir is "~/Github/pvdiv-phenology-gxe/analysis/mash/", then there are three subpop-specific directories within that.
+
+```{r}
+mash_obj <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects33000SNPs_FL50_Midwest_Uhyp.rds")
+snp <- snp_attach("~/Github/pvdiv-genome/genetic_subpops/kinship/Pvirgatum_V5_GWAS_363g_Midwest_and_Gulf_subset_maf0.05.rds")
+
+meta_mash_obj <- get_mash_metadata(m = mash_obj, snp = snp, saveoutput = FALSE)
+saveRDS(meta_mash_obj, file = file.path(mashdir, names(subpop_v)[1], 
+                                        paste0("Metadata_associated_with_",
+                                               "alternate_alleles_",
+                                              "33000SNPs_FL50_Midwest",
+                                               "_m_", "363g_snp", ".rds")))
+rm(meta_mash_obj)
+```
+
+## 3e. Data-driven mat
 
@@
 
-## 3d. DS & AP
+## 3f. DS & AP
 
 We characterized patterns of differential sensitivity and antagonistic pleiotropy between all SNPs and all pairs of gardens.
-
-## 3e. Enrichments
-We looked for enrichments of genes that could potentially affect flowering in the 20kb genomic windows surrounding the SNPs with significant effects on flowering.
-
-## 3f. Loglikelihood
-
-log likelihood of models with and without U_hyp.

### 2021-01-13 R/Analysis_v0.6_mash_Midwest_greenup_ weather_covariance_Uhyp.R

@@ -1,0 +1,99 @@
+library(tidyverse)
+library(bigsnpr)
+library(switchgrassGWAS)
+library(mashr)
+library(rlist)
+
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
+mashdir <- file.path(workingdir, "analysis", "mash")
+
+metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
+# pvdiv PLANT_ID metadata
+sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
+# pvdiv common garden metadata
+subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
+phe_gr_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
+# all greenup related phenotypes, including greenup as functions of weather variables.
+phe_fl_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
+# all flowering related phenotypes, including functions of weather variables.
+k_full <- read_rds(file.path(workingdir, "data",
+                             "Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds"))
+# kinship matrix for 630 tetraploid individuals.
+
+## vectors with subsets of gardens, phenotypes, and genetic subpopulations.
+site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
+               NE = "NE", MI = "MI", SD = "SD")#, eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
+subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", 
+                 Gulf = "229g_10.3M")
+subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", 
+                  Gulf = "Gulf")
[…]

### 2021-01-13 R/nohup_v0.6_mash_FL50_Gulf_ weather_covariance_Uhyp.R

@@ -1,0 +1,123 @@
+library(tidyverse)
+library(bigsnpr)
+library(switchgrassGWAS)
+library(mashr)
+library(rlist)
+
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
+mashdir <- file.path(workingdir, "analysis", "mash")
+
+metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
+# pvdiv PLANT_ID metadata
+sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
+# pvdiv common garden metadata
+subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
+phe_gr_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
+# all greenup related phenotypes, including greenup as functions of weather variables.
+phe_fl_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
+# all flowering related phenotypes, including functions of weather variables.
+k_full <- read_rds(file.path(workingdir, "data",
+                             "Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds"))
+# kinship matrix for 630 tetraploid individuals.
+
+## vectors with subsets of gardens, phenotypes, and genetic subpopulations.
+site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
+               NE = "NE", MI = "MI", SD = "SD")
+subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", 
+                 Gulf = "229g_10.3M")
+subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", 
+                  Gulf = "Gulf")
+phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
+                "crain_1d", "GR50", "cgdd_12c_10d")
[…]

### 2021-01-13 R/nohup_v0.6_mash_FL50_Midwest_weather_covariance_Uhyp.R

@@ -1,0 +1,120 @@
+library(tidyverse)
+library(bigsnpr)
+library(switchgrassGWAS)
+library(mashr)
+library(rlist)
+
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
+mashdir <- file.path(workingdir, "analysis", "mash")
+
+metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
+# pvdiv PLANT_ID metadata
+sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
+# pvdiv common garden metadata
+subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
+phe_gr_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
+# all greenup related phenotypes, including greenup as functions of weather variables.
+phe_fl_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
+# all flowering related phenotypes, including functions of weather variables.
+k_full <- read_rds(file.path(workingdir, "data",
+                             "Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds"))
+# kinship matrix for 630 tetraploid individuals.
+
+## vectors with subsets of gardens, phenotypes, and genetic subpopulations.
+site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
+               NE = "NE", MI = "MI", SD = "SD")
+subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", 
+                 Gulf = "229g_10.3M")
+subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", 
+                  Gulf = "Gulf")
+phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
+                "crain_1d", "GR50", "cgdd_12c_10d")
[…]

### 2021-01-13 R/nohup_v0.6_mash_GR50_Gulf_ and_Midwest_weather_covariance_Uhyp.R

@@ -1,0 +1,123 @@
+library(tidyverse)
+library(bigsnpr)
+library(switchgrassGWAS)
+library(mashr)
+library(rlist)
+
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
+mashdir <- file.path(workingdir, "analysis", "mash")
+
+metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
+# pvdiv PLANT_ID metadata
+sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
+# pvdiv common garden metadata
+subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
+phe_gr_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
+# all greenup related phenotypes, including greenup as functions of weather variables.
+phe_fl_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
+# all flowering related phenotypes, including functions of weather variables.
+k_full <- read_rds(file.path(workingdir, "data",
+                             "Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds"))
+# kinship matrix for 630 tetraploid individuals.
+
+## vectors with subsets of gardens, phenotypes, and genetic subpopulations.
+site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
+               NE = "NE", MI = "MI", SD = "SD")
+subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", 
+                 Gulf = "229g_10.3M")
+subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", 
+                  Gulf = "Gulf")
+phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
+                "crain_1d", "GR50", "cgdd_12c_10d")
[…]

### 2021-01-13 R/nohup_v0.6_mash_GR50_Midwest_weather_covariance_Uhyp.R

@@ -1,0 +1,122 @@
+library(tidyverse)
+library(bigsnpr)
+library(switchgrassGWAS)
+library(mashr)
+library(rlist)
+
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
+mashdir <- file.path(workingdir, "analysis", "mash")
+
+metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
+# pvdiv PLANT_ID metadata
+sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
+# pvdiv common garden metadata
+subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
+phe_gr_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
+# all greenup related phenotypes, including greenup as functions of weather variables.
+phe_fl_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
+# all flowering related phenotypes, including functions of weather variables.
+k_full <- read_rds(file.path(workingdir, "data",
+                             "Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds"))
+# kinship matrix for 630 tetraploid individuals.
+
+## vectors with subsets of gardens, phenotypes, and genetic subpopulations.
+site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
+               NE = "NE", MI = "MI", SD = "SD")
+subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", 
+                 Gulf = "229g_10.3M")
+subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", 
+                  Gulf = "Gulf")
+phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
+                "crain_1d", "GR50", "cgdd_12c_10d")
[…]

### 2021-01-13 R/nohup_v0.6_mash_Gulf_GR50_ weather_covariance_Uhyp.R

@@ -1,0 +1,115 @@
+library(tidyverse)
+library(bigsnpr)
+library(switchgrassGWAS)
+library(mashr)
+library(rlist)
+
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
+mashdir <- file.path(workingdir, "analysis", "mash")
+
+metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
+# pvdiv PLANT_ID metadata
+sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
+# pvdiv common garden metadata
+subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
+phe_gr_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
+# all greenup related phenotypes, including greenup as functions of weather variables.
+phe_fl_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
+# all flowering related phenotypes, including functions of weather variables.
+k_full <- read_rds(file.path(workingdir, "data",
+                             "Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds"))
+# kinship matrix for 630 tetraploid individuals.
+
+## vectors with subsets of gardens, phenotypes, and genetic subpopulations.
+site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
+               NE = "NE", MI = "MI", SD = "SD")#, eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
+subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", 
+                 Gulf = "229g_10.3M")
+subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", 
[…]

### 2021-01-13 R/nohup_v0.6_mash_Gulf_and_Midwest_FL50_ weather_covariance_Uhyp.R

@@ -1,0 +1,122 @@
+library(tidyverse)
+library(bigsnpr)
+library(switchgrassGWAS)
+library(mashr)
+library(rlist)
+
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
+mashdir <- file.path(workingdir, "analysis", "mash")
+
+metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
+# pvdiv PLANT_ID metadata
+sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
+# pvdiv common garden metadata
+subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
+phe_gr_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
+# all greenup related phenotypes, including greenup as functions of weather variables.
+phe_fl_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
+# all flowering related phenotypes, including functions of weather variables.
+k_full <- read_rds(file.path(workingdir, "data",
+                             "Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds"))
+# kinship matrix for 630 tetraploid individuals.
+
+## vectors with subsets of gardens, phenotypes, and genetic subpopulations.
+site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
+               NE = "NE", MI = "MI", SD = "SD")
+subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", 
+                 Gulf = "229g_10.3M")
+subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", 
+                  Gulf = "Gulf")
+phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
+                "crain_1d", "GR50", "cgdd_12c_10d")
[…]

### 2021-01-13 R/nohup_v0.6_mash_FL50_Gulf_and_Midwest_weather_covariance_Uhyp.R
Edit nohup scripts for mash

@@ -43,8 +43,7 @@ inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
 
 
 
-# Define U_hyp
-# needs to change in two (Untitled) places for Midwest to drop gardens. --------
+## Define U_hyp: needs to change for Midwest, which drops gardens. -------------
 
 U_hyp <- list()
 # change based on phenotype ---------------
@@ -57,7 +56,7 @@ phe_hyp <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl",
 for(i in seq_along(phe_hyp)){
   for(k in seq_along(subpop_v2)){
     
-    cov_df <- phe_gr_df %>%
+    cov_df <- phe_fl_df %>%
       ungroup() %>%
       left_join(select(sites, SITE, manu_site), by = "SITE") %>%
       filter(SUBPOP %in% subpop_v2[[k]]) %>%
@@ -93,14 +92,14 @@ for(i in seq_along(phe_hyp)){
 numSNPs_v <- c(19000, 19000, 19000, 19000, 24000, 33000)
 suffix_v <- c("FL50_Gulf_and_Midwest", "FL50_Gulf", "GR50_Gulf_and_Midwest",
               "GR50_Gulf", "GR50_Midwest", "FL50_Midwest")
-i = 2  # change this for different scripts -------------------------------------
+i = 1  # change this for different scripts -------------------------------------
 # B_hat_strong_df_19000topSNPs_FL50_Gulf_and_Midwest.rds
 # B_hat_strong_df_19000topSNPs_FL50_Gulf.rds
 # B_hat_strong_df_19000topSNPs_GR50_Gulf_and_Midwest.rds
 # B_hat_strong_df_19000topSNPs_GR50_Gulf.rds
 # B_hat_strong_df_24000topSNPs_GR50_Midwest.rds
 # B_hat_strong_df_33000topSNPs_FL50_Midwest.rds
-k = 3  # 1 G&M 2 M 3 G change based on subpop -------
+k = 1  # 1 G&M 2 M 3 G change based on subpop -------
 
 
 numSNPs <- numSNPs_v[i]

### 2021-01-13 R/nohup_v0.6_mash_FL50_Gulf_weather_covariance_Uhyp.R

@@ -43,7 +43,8 @@ inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
 
 
 
-## Define U_hyp: needs to change for Midwest, which drops gardens. -------------
+# Define U_hyp
+# needs to change in two (Untitled) places for Midwest to drop gardens. --------
 
 U_hyp <- list()
 # change based on phenotype ---------------
@@ -56,7 +57,7 @@ phe_hyp <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl",
 for(i in seq_along(phe_hyp)){
   for(k in seq_along(subpop_v2)){
     
-    cov_df <- phe_gr_df %>%
+    cov_df <- phe_fl_df %>%
       ungroup() %>%
       left_join(select(sites, SITE, manu_site), by = "SITE") %>%
       filter(SUBPOP %in% subpop_v2[[k]]) %>%
@@ -92,14 +93,14 @@ for(i in seq_along(phe_hyp)){
 numSNPs_v <- c(19000, 19000, 19000, 19000, 24000, 33000)
 suffix_v <- c("FL50_Gulf_and_Midwest", "FL50_Gulf", "GR50_Gulf_and_Midwest",
               "GR50_Gulf", "GR50_Midwest", "FL50_Midwest")
-i = 1  # change this for different scripts -------------------------------------
+i = 2  # change this for different scripts -------------------------------------
 # B_hat_strong_df_19000topSNPs_FL50_Gulf_and_Midwest.rds
 # B_hat_strong_df_19000topSNPs_FL50_Gulf.rds
 # B_hat_strong_df_19000topSNPs_GR50_Gulf_and_Midwest.rds
 # B_hat_strong_df_19000topSNPs_GR50_Gulf.rds
 # B_hat_strong_df_24000topSNPs_GR50_Midwest.rds
 # B_hat_strong_df_33000topSNPs_FL50_Midwest.rds
-k = 1  # 1 G&M 2 M 3 G change based on subpop -------
+k = 3  # 1 G&M 2 M 3 G change based on subpop -------
 
 
 numSNPs <- numSNPs_v[i]

### 2021-01-13 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Modify nohup mash scripts to remove self-correlation of phenotypes and add greenup corr for flowering

@@
 ### run mash
-```{r run mash with gulf}
-for(k in c(1,3,2)){
 
-  numSNPs <- 19000
-  m_out <- mash_standard_run(path = file.path(outputdir), #, names(subpop_v)[k]), # use to run on genegenie within mash dirs
-                             numSNPs = numSNPs, 
-                             U_hyp = U_hyp,
-                             suffix = paste0("GR50_", names(subpop_v)[k]),
-                             saveoutput = TRUE)
-}
-
-inputfiles
-```
+Ran mash using six 'nohup_v0.6' R scripts on the server 2021-01-12.
 
@@
                            suffix = paste0("GR50_", names(subpop_v)[k]))
-m_out <- mash_standard_run(path = file.path(outputdir), #, names(subpop_v)[k]), # use to run on genegenie within mash dirs
-                             numSNPs = numSNPs, 
-                             suffix = paste0("GR50_", names(subpop_v)[k]),
-                             saveoutput = TRUE)
-saveRDS(m_out, file = file.path(mashdir, paste0("Strong_Effects_", numSNPs,
-                                                "GR50_", names(subpop_v)[k],
-                                                "_no_Uhyp.rds")))
-m_out1 <- mash_standard_run(path = file.path(outputdir), numSNPs = numSNPs, 
-                           U_ed = file.path(outputdir, paste0("Model_of_",
-                           "random_tests_24000SNPs_GR50_Midwest.rds")),
-                           U_hyp = U_hyp,
-                           suffix = paste0("GR50_", names(subpop_v)[k]),
-                           saveoutput = FALSE)
-saveRDS(m_out1, file = file.path(mashdir, paste0("Strong_Effects_", numSNPs,
-                                                "GR50_", names(subpop_v)[k],
-                                                "_no_Uhyp.rds")))
+# Ran mash using a 'nohup_v0.6' R script.
 ```
@@
 
+## 3a. Log-likelihoods
+```{r}
+m_gulf_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_GR50_Gulf_Uhyp.rds")
+m_gulf_gr_no <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_GR50_Gulf_no_Uhyp.rds")
+m_gulf_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_FL50_Gulf_Uhyp.rds")
+m_gulf_fl_no <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_FL50_Gulf_no_Uhyp.rds")
+
+m_midw_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_24000_SNPs_GR50_Midwest_Uhyp.rds")
+m_midw_gr_no <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_24000_SNPs_GR50_Midwest_no_Uhyp.rds")  
+m_midw_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_33000_SNPs_FL50_Midwest_Uhyp.rds")
+m_midw_fl_no <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_33000_SNPs_FL50_Midwest_no_Uhyp.rds")
+
+m_gam_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_GR50_Gulf_and_Midwest_Uhyp.rds")
+m_gam_gr_no <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_GR50_Gulf_and_Midwest_no_Uhyp.rds")
+m_gam_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_FL50_Gulf_and_Midwest_Uhyp.rds")
+m_gam_fl_no <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_FL50_Gulf_and_Midwest_no_Uhyp.rds")
+```
+
+```{r}
+m_gulf_gr_hyp$loglik
+m_gulf_gr_no$loglik
+m_gulf_fl_hyp$loglik
+m_gulf_fl_no$loglik
+
+-2*(m_gulf_gr_no$loglik - m_gulf_gr_hyp$loglik) # negative
+-2*(m_gulf_fl_no$loglik - m_gulf_fl_hyp$loglik) # positive
+# test statistic equal to twice the absolute difference in these log likelihoods is assumed to be distributed as Chi square with one degree of freedom. Larger positive statistics are more significant.
+1 - pchisq(-2*(m_gulf_fl_no$loglik - m_gulf_fl_hyp$loglik), 30) #sig
+1 - pchisq(-2*(m_gulf_gr_no$loglik - m_gulf_gr_hyp$loglik), 15)
+# I don't know how many df each covariance matrix (each Uhyp) would take. At least one per added Uhyp? Up to (ncond*ncond)/2 per added Uhyp? What do they do in the mash paper?
+
+-2*(m_midw_gr_no$loglik - m_midw_gr_hyp$loglik) # positive
+-2*(m_midw_fl_no$loglik - m_midw_fl_hyp$loglik) # v. negative 
+1 - pchisq(-2*(m_midw_fl_no$loglik - m_midw_fl_hyp$loglik), 30*32)
+1 - pchisq(-2*(m_midw_gr_no$loglik - m_midw_gr_hyp$loglik), 15*32) # sig
+
+-2*(m_gam_gr_no$loglik - m_gam_gr_hyp$loglik)  # positive
+-2*(m_gam_fl_no$loglik - m_gam_fl_hyp$loglik) # v. positive
+1 - pchisq(-2*(m_gam_fl_no$loglik - m_gam_fl_hyp$loglik), 30*32) # sig
+1 - pchisq(-2*(m_gam_gr_no$loglik - m_gam_gr_hyp$loglik), 15*32) # sig
+
+1-pchisq(0.0279,1) # LR -2(L0 - L1) gives LR of 0.0279 which gives p of 0.87
+1-pchisq((6.89), 1) # LR of 6.89 gives p of ~0.01
+```
+
+A test using the log-likelihood
+
+For the MLR estimator there is an additional test for nested models.
+This test compares the log-likelihoods for the null and alternative models rather than the chi-square values.
+Below is a list of the information needed, along with the symbol (i.e., letter and number) used to represent each value.
+
+L0 = log-likelihood for the null model.
+L1 = log-likelihood for the alternative model.
+
+c0 = scaling correction factor for the null model.
+c1 = scaling correction factor for the alternative model.
+
+p0 = number of parameters estimated in the null model.
+p1 = number of parameters estimated in the alternative model.
+
+From this information the tests statistic TRd can be calculated (note that cd is calculated first then used to calculate TRd), along with the degrees of freedom:
+
+cd = (p0*c0-p1*c1)/(p0-p1)
+TRd = -2*(L0-L1)/cd
+df = p1-p0
+
 ## 3b. Covar Loadings
 How many SNP effects load onto each covariance matrix?
-```{r load mash objects}
-gulfm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects19000SNPs_GR50_Gulf_Uhyp.rds")
 
-gandm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects19000SNPs_GR50_Gulf_and_Midwest_Uhyp.rds")
+```{r}
+mash_plot_covar(m_gulf_gr_no)
+mash_plot_covar(m_gulf_gr_hyp)
+mash_plot_covar(m_gulf_fl_no)
+mash_plot_covar(m_gulf_fl_hyp)
+
+mash_plot_covar(m_midw_gr_no)
+mash_plot_covar(m_midw_gr_hyp)
+mash_plot_covar(m_midw_fl_no)
+mash_plot_covar(m_midw_fl_hyp)
+
+mash_plot_covar(m_gam_gr_no)
+mash_plot_covar(m_gam_gr_hyp)
+mash_plot_covar(m_gam_fl_no)
+mash_plot_covar(m_gam_fl_hyp)
 ```
+
+```{r}
+cov_gg <- mash_plot_covar(m_gulf_gr_hyp)
+cov_gf <- mash_plot_covar(m_gulf_fl_hyp)
+
+cov_mg <- mash_plot_covar(m_midw_gr_hyp)
+cov_mf <- mash_plot_covar(m_midw_fl_hyp)
+
+cov_gamg <- mash_plot_covar(m_gam_gr_hyp)
+cov_gamf <- mash_plot_covar(m_gam_fl_hyp)
+
+cov_gg$ggobject + ggtitle("Gulf greenup 11.4%")
+cov_gf$ggobject + ggtitle("Gulf flowering 32.6%")
+cov_mg$ggobject + ggtitle("Midwest greenup 35.6%")
+cov_mf$ggobject + ggtitle("Midwest flowering 45.7%")
+cov_gamg$ggobject + ggtitle("Gulf and Midwest greenup 9.6%")
+cov_gamf$ggobject + ggtitle("Gulf and Midwest flowering 46%")
+```
+
+46% of effect mass for both subpop flowering was on hyp matrices, while only 9.6% of greenup was.
+Gulf greenup: 11.4%  flowering: 32.6%
+Midwest greenup: 35.6% flowering 45.7%
+
+```{r}
+cov_gamf$covar_df %>% 
+  mutate(Type = case_when(grepl("ED_", `Covariance Matrix`) ~ "data-driven",
+                          grepl("single_effect", `Covariance Matrix`) ~ "canonical",
+                          grepl("simple_het", `Covariance Matrix`) ~ "canonical",
+                          grepl("no_effects", `Covariance Matrix`) ~ "canonical",
+                          grepl("identity", `Covariance Matrix`) ~ "canonical",
+                          grepl("equal_effects", `Covariance Matrix`) ~ "canonical",
+                          grepl("_and_Midwest$", `Covariance Matrix`) ~ "GAM_hyp",
+                          grepl("Midwest$", `Covariance Matrix`) ~ "Midw_hyp",
+                          grepl("Gulf$", `Covariance Matrix`) ~ "Gulf_hyp",
+                          TRUE ~ "other"
+                          )) %>%
+  group_by(Type) %>%
+  summarise(Mass = sum(Mass))
+
+cov_gamg$covar_df %>% 
+  mutate(Type = case_when(grepl("ED_", `Covariance Matrix`) ~ "data-driven",
+                          grepl("single_effect", `Covariance Matrix`) ~ "canonical",
+                          grepl("simple_het", `Covariance Matrix`) ~ "canonical",
+                          grepl("no_effects", `Covariance Matrix`) ~ "canonical",
+                          grepl("identity", `Covariance Matrix`) ~ "canonical",
+                          grepl("equal_effects", `Covariance Matrix`) ~ "canonical",
+                          grepl("_and_Midwest$", `Covariance Matrix`) ~ "hyp",
+                          grepl("Midwest$", `Covariance Matrix`) ~ "hyp",
+                          grepl("Gulf$", `Covariance Matrix`) ~ "hyp",
+                          TRUE ~ "other"
+                          )) %>%
+  group_by(Type) %>%
+  summarise(Mass = sum(Mass))
+```
+Gulf: 20% canonical, 68.5% data-driven, 9.5% Gulf hyp, 1.6% Midwest hyp
+Gulf & Midwest: 28% canonical, 62% data-driven, 5.5% Gulf, 1.2% Midwest, 2.8% G&M
+
+46% of effect mass for both subpop flowering was on hyp matrices, while only 9.6% of greenup was.
+Gulf greenup: 11.4%  flowering: 32.6%
+Midwest greenup: 35.6% flowering 45.7%
+
 44765 SNPs tested for G&M, 45931 SNPs tested for Gulf.
-
-Midwest I haven't run successfully yet as of 2021-01-12 - I will need to remove some locations with all-zero effects, probably.
 
@@
 
-covargulf <- mash_plot_covar(gulfm, suffix = "19000SNPs_GR50_Gulf", saveoutput = TRUE)
+covargulf <- mash_plot_covar(m_gulf_fl_hyp, suffix = "19000SNPs_GR50_Gulf", saveoutput = FALSE)
 covargulf
@@
 ```
-
-```{r}
-covargam$covar_df %>% 
-  mutate(Type = case_when(grepl("ED_", `Covariance Matrix`) ~ "data-driven",
-                          grepl("single_effect", `Covariance Matrix`) ~ "canonical",
-                          grepl("simple_het", `Covariance Matrix`) ~ "canonical",
-                          grepl("no_effects", `Covariance Matrix`) ~ "canonical",
-                          grepl("identity", `Covariance Matrix`) ~ "canonical",
-                          grepl("equal_effects", `Covariance Matrix`) ~ "canonical",
-                          grepl("_and_Midwest$", `Covariance Matrix`) ~ "GAM_hyp",
-                          grepl("Midwest$", `Covariance Matrix`) ~ "Midw_hyp",
-                          grepl("Gulf$", `Covariance Matrix`) ~ "Gulf_hyp",
-                          TRUE ~ "other"
-                          )) %>%
-  group_by(Type) %>%
-  summarise(Mass = sum(Mass))
-```
-Gulf: 20% canonical, 68.5% data-driven, 9.5% Gulf hyp, 1.6% Midwest hyp
-Gulf & Midwest: 28% canonical, 62% data-driven, 5.5% Gulf, 1.2% Midwest, 2.8% G&M
 

### 2021-01-13 R/Analysis_v0.6_mash_Midwest_greenup_ weather_covariance_Uhyp.R

@@ -1,99 +1,0 @@
-library(tidyverse)
-library(bigsnpr)
-library(switchgrassGWAS)
-library(mashr)
-library(rlist)
-
-source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
-workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
-outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
-mashdir <- file.path(workingdir, "analysis", "mash")
-
-metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
-# pvdiv PLANT_ID metadata
-sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
-# pvdiv common garden metadata
-subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
-phe_gr_df <- read_rds(file.path(workingdir, "data",
-                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
-# all greenup related phenotypes, including greenup as functions of weather variables.
-phe_fl_df <- read_rds(file.path(workingdir, "data",
-                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
-# all flowering related phenotypes, including functions of weather variables.
-k_full <- read_rds(file.path(workingdir, "data",
-                             "Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds"))
-# kinship matrix for 630 tetraploid individuals.
-
-## vectors with subsets of gardens, phenotypes, and genetic subpopulations.
-site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
-               NE = "NE", MI = "MI", SD = "SD")#, eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
-subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", 
-                 Gulf = "229g_10.3M")
-subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", 
-                  Gulf = "Gulf")
[…]

### 2021-01-13 R/nohup_v0.6_mash_FL50_Gulf_and_Midwest_weather_covariance_Uhyp.R

@@ -47,10 +47,10 @@ inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
 
 U_hyp <- list()
 # change based on phenotype ---------------
-#phe_hyp <- c("GR50", "cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", "tave_10d")
-phe_hyp <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
-             "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", 
-             "crain_7d")
+#phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", 
+#             "tave_5d", "tave_10d", "tave_18d")
+phe_hyp <- c("GR50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
+             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")
 
 
 for(i in seq_along(phe_hyp)){
@@ -104,8 +104,11 @@ k = 1  # 1 G&M 2 M 3 G change based on subpop -------
 
 numSNPs <- numSNPs_v[i]
 suffix <- suffix_v[i]
+U_ed <- file.path(mashdir, paste0("Model_of_random_tests_", numSNPs, "SNPs_", 
+                                  suffix, ".rds"))
 
 m_out <- mash_standard_run(path = file.path(outputdir), 
+                           U_ed = U_ed,
                            numSNPs = numSNPs, 
                            suffix = suffix,
                            saveoutput = FALSE)
@@ -113,10 +116,11 @@ saveRDS(m_out, file = file.path(mashdir, names(subpop_v)[k],
                                 paste0("Strong_Effects_", numSNPs, "_SNPs_",
                                        suffix, "_no_Uhyp.rds")))
 m_out1 <- mash_standard_run(path = file.path(outputdir), 
-                           numSNPs = numSNPs, 
-                           U_hyp = U_hyp,
-                           suffix = suffix,
-                           saveoutput = FALSE)
+                            U_ed = U_ed,
+                            numSNPs = numSNPs, 
+                            U_hyp = U_hyp,
+                            suffix = suffix,
+                            saveoutput = FALSE)
 saveRDS(m_out1, file = file.path(mashdir, names(subpop_v)[k], 
                                 paste0("Strong_Effects_", numSNPs, "_SNPs_",
                                        suffix, "_Uhyp.rds")))

### 2021-01-13 R/nohup_v0.6_mash_FL50_Gulf_weather_covariance_Uhyp.R

@@ -48,10 +48,10 @@ inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
 
 U_hyp <- list()
 # change based on phenotype ---------------
-#phe_hyp <- c("GR50", "cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", "tave_10d")
-phe_hyp <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
-             "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", 
-             "crain_7d")
+#phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", 
+#             "tave_5d", "tave_10d", "tave_18d")
+phe_hyp <- c("GR50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
+             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")
 
 
 for(i in seq_along(phe_hyp)){
@@ -105,8 +105,11 @@ k = 3  # 1 G&M 2 M 3 G change based on subpop -------
 
 numSNPs <- numSNPs_v[i]
 suffix <- suffix_v[i]
+U_ed <- file.path(mashdir, paste0("Model_of_random_tests_", numSNPs, "SNPs_", 
+                                  suffix, ".rds"))
 
 m_out <- mash_standard_run(path = file.path(outputdir), 
+                           U_ed = U_ed,
                            numSNPs = numSNPs, 
                            suffix = suffix,
                            saveoutput = FALSE)
@@ -114,10 +117,11 @@ saveRDS(m_out, file = file.path(mashdir, names(subpop_v)[k],
                                 paste0("Strong_Effects_", numSNPs, "_SNPs_",
                                        suffix, "_no_Uhyp.rds")))
 m_out1 <- mash_standard_run(path = file.path(outputdir), 
-                           numSNPs = numSNPs, 
-                           U_hyp = U_hyp,
-                           suffix = suffix,
-                           saveoutput = FALSE)
+                            U_ed = U_ed,
+                            numSNPs = numSNPs, 
+                            U_hyp = U_hyp,
+                            suffix = suffix,
+                            saveoutput = FALSE)
 saveRDS(m_out1, file = file.path(mashdir, names(subpop_v)[k], 
-                                paste0("Strong_Effects_", numSNPs, "_SNPs_",
[…]

### 2021-01-13 R/nohup_v0.6_mash_FL50_Midwest_weather_covariance_Uhyp.R

@@ -48,10 +48,10 @@ inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
 
 U_hyp <- list()
 # change based on phenotype ---------------
-#phe_hyp <- c("GR50", "cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", "tave_10d")
-phe_hyp <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
-             "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", 
-             "crain_7d")
+#phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", 
+#             "tave_5d", "tave_10d", "tave_18d")
+phe_hyp <- c("GR50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
+             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")
 
 
 for(i in seq_along(phe_hyp)){
@@ -102,8 +102,11 @@ k = 2  # 1 G&M 2 M 3 G change based on subpop -------
 
 numSNPs <- numSNPs_v[i]
 suffix <- suffix_v[i]
+U_ed <- file.path(mashdir, paste0("Model_of_random_tests_", numSNPs, "SNPs_", 
+                                  suffix, ".rds"))
 
 m_out <- mash_standard_run(path = file.path(outputdir), 
+                           U_ed = U_ed,
                            numSNPs = numSNPs, 
                            suffix = suffix,
                            saveoutput = FALSE)
@@ -111,10 +114,11 @@ saveRDS(m_out, file = file.path(mashdir, names(subpop_v)[k],
                                 paste0("Strong_Effects_", numSNPs, "_SNPs_",
                                        suffix, "_no_Uhyp.rds")))
 m_out1 <- mash_standard_run(path = file.path(outputdir), 
-                           numSNPs = numSNPs, 
-                           U_hyp = U_hyp,
-                           suffix = suffix,
-                           saveoutput = FALSE)
+                            U_ed = U_ed,
+                            numSNPs = numSNPs, 
+                            U_hyp = U_hyp,
+                            suffix = suffix,
+                            saveoutput = FALSE)
 saveRDS(m_out1, file = file.path(mashdir, names(subpop_v)[k], 
-                                paste0("Strong_Effects_", numSNPs, "_SNPs_",
[…]

### 2021-01-13 R/nohup_v0.6_mash_GR50_Gulf_ and_Midwest_weather_covariance_Uhyp.R

@@ -48,10 +48,10 @@ inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
 
 U_hyp <- list()
 # change based on phenotype ---------------
-phe_hyp <- c("GR50", "cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", "tave_10d")
-#phe_hyp <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
-#             "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", 
-#             "crain_7d")
+phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", 
+             "tave_5d", "tave_10d", "tave_18d")
+#phe_hyp <- c("dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
+#             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")
 
 
 for(i in seq_along(phe_hyp)){
@@ -105,8 +105,11 @@ k = 1  # 1 G&M 2 M 3 G change based on subpop -------
 
 numSNPs <- numSNPs_v[i]
 suffix <- suffix_v[i]
+U_ed <- file.path(mashdir, paste0("Model_of_random_tests_", numSNPs, "SNPs_", 
+                                  suffix, ".rds"))
 
 m_out <- mash_standard_run(path = file.path(outputdir), 
+                           U_ed = U_ed,
                            numSNPs = numSNPs, 
                            suffix = suffix,
                            saveoutput = FALSE)
@@ -114,10 +117,11 @@ saveRDS(m_out, file = file.path(mashdir, names(subpop_v)[k],
                                 paste0("Strong_Effects_", numSNPs, "_SNPs_",
                                        suffix, "_no_Uhyp.rds")))
 m_out1 <- mash_standard_run(path = file.path(outputdir), 
-                           numSNPs = numSNPs, 
-                           U_hyp = U_hyp,
-                           suffix = suffix,
-                           saveoutput = FALSE)
+                            U_ed = U_ed,
+                            numSNPs = numSNPs, 
+                            U_hyp = U_hyp,
+                            suffix = suffix,
+                            saveoutput = FALSE)
 saveRDS(m_out1, file = file.path(mashdir, names(subpop_v)[k], 
-                                paste0("Strong_Effects_", numSNPs, "_SNPs_",
[…]

### 2021-01-13 R/nohup_v0.6_mash_GR50_Gulf_weather_covariance_Uhyp.R

@@ -45,7 +45,10 @@ inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
 ## Define Uhyp
 
 U_hyp <- list()
-phe_hyp <- c("GR50", "cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", "tave_10d")
+phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", 
+             "tave_5d", "tave_10d", "tave_18d")
+#phe_hyp <- c("dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
+#             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")
 
 for(i in seq_along(phe_hyp)){
   for(k in seq_along(subpop_v2)){
@@ -97,19 +100,23 @@ i = 4  # change this for different scripts
 k = 3
 numSNPs <- numSNPs_v[i]
 suffix <- suffix_v[i]
+U_ed <- file.path(mashdir, paste0("Model_of_random_tests_", numSNPs, "SNPs_", 
+                                  suffix, ".rds"))
 
-m_out <- mash_standard_run(path = file.path(outputdir), #, names(subpop_v)[k]), # use to run on genegenie within mash dirs
+m_out <- mash_standard_run(path = file.path(outputdir), 
+                           U_ed = U_ed,
                            numSNPs = numSNPs, 
                            suffix = suffix,
                            saveoutput = FALSE)
 saveRDS(m_out, file = file.path(mashdir, names(subpop_v)[k], 
                                 paste0("Strong_Effects_", numSNPs, "_SNPs_",
                                        suffix, "_no_Uhyp.rds")))
-m_out1 <- mash_standard_run(path = file.path(outputdir), #, names(subpop_v)[k]), # use to run on genegenie within mash dirs
-                           numSNPs = numSNPs, 
-                           U_hyp = U_hyp,
-                           suffix = suffix,
-                           saveoutput = FALSE)
+m_out1 <- mash_standard_run(path = file.path(outputdir), 
+                            U_ed = U_ed,
+                            numSNPs = numSNPs, 
+                            U_hyp = U_hyp,
+                            suffix = suffix,
+                            saveoutput = FALSE)
 saveRDS(m_out1, file = file.path(mashdir, names(subpop_v)[k], 
-                                paste0("Strong_Effects_", numSNPs, "_SNPs_",
[…]

### 2021-01-13 R/nohup_v0.6_mash_GR50_Midwest_weather_covariance_Uhyp.R

@@ -48,10 +48,10 @@ inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
 
 U_hyp <- list()
 # change based on phenotype ---------------
-phe_hyp <- c("GR50", "cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", "tave_10d")
-#phe_hyp <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
-#             "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", 
-#             "crain_7d")
+phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", 
+             "tave_5d", "tave_10d", "tave_18d")
+#phe_hyp <- c("dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
+#             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")
 
 
 for(i in seq_along(phe_hyp)){
@@ -104,8 +104,11 @@ k = 2  # 1 G&M 2 M 3 G change based on subpop -------
 
 numSNPs <- numSNPs_v[i]
 suffix <- suffix_v[i]
+U_ed <- file.path(mashdir, paste0("Model_of_random_tests_", numSNPs, "SNPs_", 
+                                  suffix, ".rds"))
 
 m_out <- mash_standard_run(path = file.path(outputdir), 
+                           U_ed = U_ed,
                            numSNPs = numSNPs, 
                            suffix = suffix,
                            saveoutput = FALSE)
@@ -113,10 +116,11 @@ saveRDS(m_out, file = file.path(mashdir, names(subpop_v)[k],
                                 paste0("Strong_Effects_", numSNPs, "_SNPs_",
                                        suffix, "_no_Uhyp.rds")))
 m_out1 <- mash_standard_run(path = file.path(outputdir), 
-                           numSNPs = numSNPs, 
-                           U_hyp = U_hyp,
-                           suffix = suffix,
-                           saveoutput = FALSE)
+                            U_ed = U_ed,
+                            numSNPs = numSNPs, 
+                            U_hyp = U_hyp,
+                            suffix = suffix,
+                            saveoutput = FALSE)
 saveRDS(m_out1, file = file.path(mashdir, names(subpop_v)[k], 
-                                paste0("Strong_Effects_", numSNPs, "_SNPs_",
[…]

### 2021-01-13 R/nohup_v0.6_mash_FL50_Gulf_and_Midwest_weather_covariance_Uhyp.R
Fix specification of U_ed in nohup mash scripts

@@ -104,7 +104,7 @@ k = 1  # 1 G&M 2 M 3 G change based on subpop -------
 
 numSNPs <- numSNPs_v[i]
 suffix <- suffix_v[i]
-U_ed <- file.path(mashdir, paste0("Model_of_random_tests_", numSNPs, "SNPs_", 
+U_ed <- file.path(mashdir, paste0("Mash_data_driven_covariances_", numSNPs, "_", 
                                   suffix, ".rds"))
 
 m_out <- mash_standard_run(path = file.path(outputdir), 

### 2021-01-13 R/nohup_v0.6_mash_FL50_Gulf_weather_covariance_Uhyp.R

@@ -105,7 +105,7 @@ k = 3  # 1 G&M 2 M 3 G change based on subpop -------
 
 numSNPs <- numSNPs_v[i]
 suffix <- suffix_v[i]
-U_ed <- file.path(mashdir, paste0("Model_of_random_tests_", numSNPs, "SNPs_", 
+U_ed <- file.path(mashdir, paste0("Mash_data_driven_covariances_", numSNPs, "_", 
                                   suffix, ".rds"))
 
 m_out <- mash_standard_run(path = file.path(outputdir), 

### 2021-01-13 R/nohup_v0.6_mash_FL50_Midwest_weather_covariance_Uhyp.R

@@ -102,7 +102,7 @@ k = 2  # 1 G&M 2 M 3 G change based on subpop -------
 
 numSNPs <- numSNPs_v[i]
 suffix <- suffix_v[i]
-U_ed <- file.path(mashdir, paste0("Model_of_random_tests_", numSNPs, "SNPs_", 
+U_ed <- file.path(mashdir, paste0("Mash_data_driven_covariances_", numSNPs, "_", 
                                   suffix, ".rds"))
 
 m_out <- mash_standard_run(path = file.path(outputdir), 

### 2021-01-13 R/nohup_v0.6_mash_GR50_Gulf_and_Midwest_weather_covariance_Uhyp.R

@@ -105,7 +105,7 @@ k = 1  # 1 G&M 2 M 3 G change based on subpop -------
 
 numSNPs <- numSNPs_v[i]
 suffix <- suffix_v[i]
-U_ed <- file.path(mashdir, paste0("Model_of_random_tests_", numSNPs, "SNPs_", 
+U_ed <- file.path(mashdir, paste0("Mash_data_driven_covariances_", numSNPs, "_", 
                                   suffix, ".rds"))
 
 m_out <- mash_standard_run(path = file.path(outputdir), 

### 2021-01-13 R/nohup_v0.6_mash_GR50_Gulf_weather_covariance_Uhyp.R

@@ -100,7 +100,7 @@ i = 4  # change this for different scripts
 k = 3
 numSNPs <- numSNPs_v[i]
 suffix <- suffix_v[i]
-U_ed <- file.path(mashdir, paste0("Model_of_random_tests_", numSNPs, "SNPs_", 
+U_ed <- file.path(mashdir, paste0("Mash_data_driven_covariances_", numSNPs, "_", 
                                   suffix, ".rds"))
 
 m_out <- mash_standard_run(path = file.path(outputdir), 

### 2021-01-13 R/nohup_v0.6_mash_GR50_Midwest_weather_covariance_Uhyp.R

@@ -104,7 +104,7 @@ k = 2  # 1 G&M 2 M 3 G change based on subpop -------
 
 numSNPs <- numSNPs_v[i]
 suffix <- suffix_v[i]
-U_ed <- file.path(mashdir, paste0("Model_of_random_tests_", numSNPs, "SNPs_", 
+U_ed <- file.path(mashdir, paste0("Mash_data_driven_covariances_", numSNPs, "_", 
                                   suffix, ".rds"))
 
 m_out <- mash_standard_run(path = file.path(outputdir), 

### 2021-01-14 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Start manuscript writeup with mash env cov matrices

@@
 library(rlist)
+library(cowplot)
 ```
@@
 ```{r}
-mash_plot_covar(m_gulf_gr_no)
-mash_plot_covar(m_gulf_gr_hyp)
-mash_plot_covar(m_gulf_fl_no)
-mash_plot_covar(m_gulf_fl_hyp)
-
-mash_plot_covar(m_midw_gr_no)
-mash_plot_covar(m_midw_gr_hyp)
-mash_plot_covar(m_midw_fl_no)
-mash_plot_covar(m_midw_fl_hyp)
-
-mash_plot_covar(m_gam_gr_no)
-mash_plot_covar(m_gam_gr_hyp)
-mash_plot_covar(m_gam_fl_no)
-mash_plot_covar(m_gam_fl_hyp)
-```
-
-```{r}
 cov_gg <- mash_plot_covar(m_gulf_gr_hyp)
@@
 
-cov_gg$ggobject + ggtitle("Gulf greenup 11.4%")
-cov_gf$ggobject + ggtitle("Gulf flowering 32.6%")
-cov_mg$ggobject + ggtitle("Midwest greenup 35.6%")
-cov_mf$ggobject + ggtitle("Midwest flowering 45.7%")
-cov_gamg$ggobject + ggtitle("Gulf and Midwest greenup 9.6%")
-cov_gamf$ggobject + ggtitle("Gulf and Midwest flowering 46%")
+cov_gg$ggobject + 
+  ggtitle("Gulf greenup", subtitle = "8.4% SNP mass")
+cov_gf$ggobject + 
+  ggtitle("Gulf flowering", subtitle = "31.8% SNP mass")
+cov_mg$ggobject + 
+  ggtitle("Midwest greenup", subtitle = "28.6% SNP mass")
+cov_mf$ggobject + 
+  ggtitle("Midwest flowering", subtitle = "45.1% SNP mass")
+cov_gamg$ggobject + 
+  ggtitle("Gulf and Midwest greenup", subtitle = "8.5% SNP mass")
+cov_gamf$ggobject + 
+  ggtitle("Gulf and Midwest flowering", subtitle = "47.3% SNP mass")
+```
+47.3% of effect mass for both subpop flowering was on hyp matrices, while only 8.5% of greenup was.
+Gulf greenup: 8.4%  flowering: 31.8%
+Midwest greenup: 28.6% flowering 45.1%
+
+```{r}
+get_cov_type <- function(m, summarise = FALSE, subpop = FALSE){
+  cov_df <- mash_plot_covar(m)
+  if (subpop == FALSE) {
+  cov_type <- cov_df$covar_df %>% 
+    mutate(Type = case_when(grepl("ED_", `Covariance Matrix`) ~ "data-driven",
+                            grepl("single_effect", `Covariance Matrix`) ~
+                              "canonical",
+                            grepl("simple_het", `Covariance Matrix`) ~
+                              "canonical",
+                            grepl("no_effects", `Covariance Matrix`) ~
+                              "canonical",
+                            grepl("identity", `Covariance Matrix`) ~
+                              "canonical",
+                            grepl("equal_effects", `Covariance Matrix`) ~
+                              "canonical",
+                            grepl("_and_Midwest$", `Covariance Matrix`) ~
+                              "hypothesized",
+                            grepl("Midwest$", `Covariance Matrix`) ~ "hypothesized",
+                            grepl("Gulf$", `Covariance Matrix`) ~ "hypothesized",
+                            TRUE ~ "other"
+                            )) 
+  } else {
+    cov_type <- cov_df$covar_df %>% 
+      mutate(Type = case_when(grepl("ED_", `Covariance Matrix`) ~
+                                "data-driven",
+                              grepl("single_effect", `Covariance Matrix`) ~
+                                "canonical",
+                              grepl("simple_het", `Covariance Matrix`) ~
+                                "canonical",
+                              grepl("no_effects", `Covariance Matrix`) ~
+                                "canonical",
+                              grepl("identity", `Covariance Matrix`) ~
+                                "canonical",
+                              grepl("equal_effects", `Covariance Matrix`) ~
+                                "canonical",
+                              grepl("_and_Midwest$", `Covariance Matrix`) ~
+                                "GAM_hyp",
+                              grepl("Midwest$", `Covariance Matrix`) ~
+                                "Midw_hyp",
+                              grepl("Gulf$", `Covariance Matrix`) ~ "Gulf_hyp",
+                              TRUE ~ "other"
+                              )) 
+  }
+  if (summarise == TRUE) {
+    cov_type <- cov_type %>%
+      group_by(Type) %>%
+      summarise(Mass = sum(Mass))
+  }
+  return(cov_type)
+}
+
+get_cov_type(m_gulf_gr_hyp, summarise = TRUE, subpop = TRUE)
+get_cov_type(m_gulf_fl_hyp, summarise = TRUE, subpop = TRUE)
+get_cov_type(m_midw_gr_hyp, summarise = TRUE, subpop = TRUE)
+get_cov_type(m_midw_fl_hyp, summarise = TRUE, subpop = TRUE)
+get_cov_type(m_gam_gr_hyp, summarise = TRUE, subpop = TRUE)
+get_cov_type(m_gam_fl_hyp, summarise = TRUE, subpop = TRUE)
 ```
 
-46% of effect mass for both subpop flowering was on hyp matrices, while only 9.6% of greenup was.
-Gulf greenup: 11.4%  flowering: 32.6%
-Midwest greenup: 35.6% flowering 45.7%
+```{r make rough draft figure 2}
+df_2a <- get_cov_type(m_gulf_gr_hyp, summarise = TRUE) %>%
+  mutate(subpop = "Gulf",
+         phenotype = "greenup")
+df_2b <- get_cov_type(m_midw_gr_hyp, summarise = TRUE) %>%
+  mutate(subpop = "Midwest",
+         phenotype = "greenup")
+df_2c <- get_cov_type(m_gam_gr_hyp, summarise = TRUE) %>%
+  mutate(subpop = "Both",
+         phenotype = "greenup")
+df_2d <- get_cov_type(m_gulf_fl_hyp, summarise = TRUE) %>%
+  mutate(subpop = "Gulf",
+         phenotype = "flowering")
+df_2e <- get_cov_type(m_midw_fl_hyp, summarise = TRUE) %>%
+  mutate(subpop = "Midwest",
+         phenotype = "flowering")
+df_2f <- get_cov_type(m_gam_fl_hyp, summarise = TRUE) %>%
+  mutate(subpop = "Both",
+         phenotype = "flowering")
+df_2gr_b <- df_2a %>% full_join(df_2b) %>% full_join(df_2c)
+df_2fl_b <- df_2d %>% full_join(df_2e) %>% full_join(df_2f)
 
-```{r}
-cov_gamf$covar_df %>% 
-  mutate(Type = case_when(grepl("ED_", `Covariance Matrix`) ~ "data-driven",
-                          grepl("single_effect", `Covariance Matrix`) ~ "canonical",
-                          grepl("simple_het", `Covariance Matrix`) ~ "canonical",
-                          grepl("no_effects", `Covariance Matrix`) ~ "canonical",
-                          grepl("identity", `Covariance Matrix`) ~ "canonical",
-                          grepl("equal_effects", `Covariance Matrix`) ~ "canonical",
-                          grepl("_and_Midwest$", `Covariance Matrix`) ~ "GAM_hyp",
-                          grepl("Midwest$", `Covariance Matrix`) ~ "Midw_hyp",
-                          grepl("Gulf$", `Covariance Matrix`) ~ "Gulf_hyp",
-                          TRUE ~ "other"
-                          )) %>%
-  group_by(Type) %>%
-  summarise(Mass = sum(Mass))
+df_2a <- get_cov_type(m_gulf_gr_hyp) %>%
+  mutate(subpop = "Gulf",
+         phenotype = "greenup")
+df_2b <- get_cov_type(m_midw_gr_hyp) %>%
+  mutate(subpop = "Midwest",
+         phenotype = "greenup")
+df_2c <- get_cov_type(m_gam_gr_hyp) %>%
+  mutate(subpop = "Both",
+         phenotype = "greenup")
+df_2d <- get_cov_type(m_gulf_fl_hyp) %>%
+  mutate(subpop = "Gulf",
+         phenotype = "flowering")
+df_2e <- get_cov_type(m_midw_fl_hyp) %>%
+  mutate(subpop = "Midwest",
+         phenotype = "flowering")
+df_2f <- get_cov_type(m_gam_fl_hyp) %>%
+  mutate(subpop = "Both",
+         phenotype = "flowering")
+df_2gr_a <- df_2a %>% full_join(df_2b) %>% full_join(df_2c) %>%
+  mutate(`Covariance Matrix` = str_replace(`Covariance Matrix`, "single_effect_Gulf_and_Midwest_", "single_effect_"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "single_effect_Midwest_", "single_effect_"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "single_effect_Gulf_", "single_effect_")
+         )
+df_2fl_a <- df_2d %>% full_join(df_2e) %>% full_join(df_2f) %>%
+  mutate(`Covariance Matrix` = str_replace(`Covariance Matrix`, "single_effect_Gulf_and_Midwest_", "single_effect_"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "single_effect_Midwest_", "single_effect_"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "single_effect_Gulf_", "single_effect_")
+         )
 
-cov_gamg$covar_df %>% 
-  mutate(Type = case_when(grepl("ED_", `Covariance Matrix`) ~ "data-driven",
-                          grepl("single_effect", `Covariance Matrix`) ~ "canonical",
-                          grepl("simple_het", `Covariance Matrix`) ~ "canonical",
-                          grepl("no_effects", `Covariance Matrix`) ~ "canonical",
-                          grepl("identity", `Covariance Matrix`) ~ "canonical",
-                          grepl("equal_effects", `Covariance Matrix`) ~ "canonical",
-                          grepl("_and_Midwest$", `Covariance Matrix`) ~ "hyp",
-                          grepl("Midwest$", `Covariance Matrix`) ~ "hyp",
-                          grepl("Gulf$", `Covariance Matrix`) ~ "hyp",
-                          TRUE ~ "other"
-                          )) %>%
-  group_by(Type) %>%
-  summarise(Mass = sum(Mass))
+f2b <- df_2gr_b %>%
+  filter(Mass > 1e-05) %>%
+  ggplot(aes(x = subpop, y = Mass)) + 
+  geom_bar(aes(fill = Type), stat = "identity") +
+  scale_fill_viridis_d(end = 0.9) +
+  theme(axis.title.x=element_blank(),
+        legend.position = "none",
+        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45))
+f2d <- df_2fl_b %>%
+  filter(Mass > 1e-05) %>%
+  ggplot(aes(x = subpop, y = Mass)) + 
+  geom_bar(aes(fill = Type), stat = "identity") +
+  scale_fill_viridis_d(end = 0.9) +
+  theme(axis.title.x=element_blank(),
+        legend.position = "none",
+        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45))
+f2a <- df_2gr_a %>%
+  filter(Mass > 1e-06) %>%
+  ggplot(aes(x = `Covariance Matrix`, y = Mass)) + 
+  geom_bar(aes(fill = Type), stat = "identity") +
+  facet_wrap(~subpop) +
+  coord_flip() + scale_fill_viridis_d(end = 0.9) +
+  theme(legend.position = c(-0.30, 0.3),
+        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45), 
+        panel.spacing.x = unit(-0, 'cm'))
+f2c <- df_2fl_a %>%
+  filter(Mass > 1e-06) %>%
+  ggplot(aes(x = `Covariance Matrix`, y = Mass)) + 
+  geom_bar(aes(fill = Type), stat = "identity") +
+  facet_wrap(~subpop) +
+  coord_flip() + scale_fill_viridis_d(end = 0.9) +
+  theme(legend.position = "none",
+        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45), 
+        panel.spacing.x = unit(-0, 'cm'))
+
+fig2 <- plot_grid(f2a, f2b, f2c, f2d, nrow = 2, rel_widths = c(0.75, 0.25), rel_heights = c(0.53, 0.47))
+save_plot("../manuscript/plots/Figure_2_mash_covariance_masses.png", base_height = 8, base_width = 7, plot = fig2)
 ```
-Gulf: 20% canonical, 68.5% data-driven, 9.5% Gulf hyp, 1.6% Midwest hyp
-Gulf & Midwest: 28% canonical, 62% data-driven, 5.5% Gulf, 1.2% Midwest, 2.8% G&M
-
-46% of effect mass for both subpop flowering was on hyp matrices, while only 9.6% of greenup was.
-Gulf greenup: 11.4%  flowering: 32.6%
-Midwest greenup: 35.6% flowering 45.7%
+47.3% of effect mass for both subpop flowering was on hyp matrices, while only 8.5% of greenup was.
+Gulf greenup: 8.4%  flowering: 31.8%
+Midwest greenup: 28.6% flowering 45.1%
 
@@
 ```{r gulf mash results}
-length(get_significant_results(gulfm))  # 9142/45931  # 19.9%
-get_significant_results(gulfm)[1:10]
+length(get_significant_results(m_gulf_gr_hyp))  # 9645/45931  # 20.999%
+length(get_significant_results(m_gulf_gr_no))   # 10427/45931 # 22.7%
+get_significant_results(m_gulf_gr_hyp)[1:10]
 
-covargulf <- mash_plot_covar(m_gulf_fl_hyp, suffix = "19000SNPs_GR50_Gulf", saveoutput = FALSE)
-covargulf
-get_U_by_mass(gulfm)
-mash_plot_Ulist(gulfm, range = get_U_by_mass(gulfm), limits = FALSE)
-mash_plot_manhattan_by_condition(gulfm, saveoutput = FALSE)
-mash_plot_sig_by_condition(gulfm, saveoutput = FALSE)
+get_U_by_mass(m_gulf_gr_hyp)
+mash_plot_Ulist(m_gulf_gr_hyp, range = get_U_by_mass(m_gulf_gr_hyp), limits = FALSE)
+mash_plot_manhattan_by_condition(m_gulf_gr_hyp, saveoutput = FALSE)
+mash_plot_sig_by_condition(m_gulf_gr_hyp, saveoutput = FALSE)
 ```
@@
 ```{r gulf and midwest mash results}
-length(get_significant_results(gandm))  # 15104/44765  # 33.7%
-get_significant_results(gandm)[1:10]
+length(get_significant_results(m_gam_gr_hyp))  # 15786/44765  # 35.26
+length(get_significant_results(m_gam_gr_no))  # 15873/44765  # 35.46%
+get_significant_results(m_gam_gr_hyp)[1:10]
 
-covargam <- mash_plot_covar(gandm, suffix = "19000SNPs_GR50_Gulf_and_Midwest", saveoutput = TRUE)
-covargam
-get_U_by_mass(gandm)
-mash_plot_Ulist(gandm, range = get_U_by_mass(gandm), limits = FALSE)
-mash_plot_manhattan_by_condition(gandm, saveoutput = FALSE)
-mash_plot_sig_by_condition(gandm, saveoutput = FALSE)
+get_U_by_mass(m_gam_gr_hyp)
+mash_plot_Ulist(m_gam_gr_hyp, range = get_U_by_mass(m_gam_gr_hyp), limits = FALSE)
+mash_plot_manhattan_by_condition(m_gam_fl_hyp, saveoutput = FALSE)
+mash_plot_sig_by_condition(m_gam_fl_hyp, saveoutput = FALSE)
 ```
@@
 
-### ii. Plot setup
+### ii. Functions for posterior weights
 
@@
 ```{r mashhattan plot setup}
-m <- gandm
+get_post_weights <- function(m){
   log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
@@
       dplyr::select(-.data$value)
-
-#df <- get_topcov(m)
 
@@
     arrange(.data$Chr, .data$Pos) 
+  return(ggman_df)
+}
 
-log10BF <- expression(paste("log"[10], plain("(Bayes Factor)")))
+pw_gg <- get_post_weights(m_gulf_gr_hyp)
+pw_gf <- get_post_weights(m_gulf_fl_hyp)
 
-mash_plot_covar(m)
+pw_mg <- get_post_weights(m_midw_gr_hyp)
+pw_mf <- get_post_weights(m_midw_fl_hyp)
+
+pw_gamg <- get_post_weights(m_gam_gr_hyp)
+pw_gamf <- get_post_weights(m_gam_fl_hyp)
+
+get_pw_filtered <- function(m, pw_thresh = 0.5, BF_thresh = 2){
+  ggman_df <- get_post_weights(m)
+  var <- names(ggman_df)[ncol(ggman_df)-1]  
+  pw_summary <- ggman_df %>%
+    pivot_longer(cols = null:var, names_to = "covar_grid", 
+                 values_to = "loading") %>%
+    filter(log10BayesFactor > BF_thresh) %>%
+    group_by(covar_grid, loading > pw_thresh) %>%
+    tally()
+return(pw_summary)
+}
+
+get_pw_filtered(m_gulf_gr_hyp, pw_thresh = 0.1, BF_thresh = 2) %>% 
+  mutate(Type = case_when(grepl("ED_", covar_grid) ~ "data-driven",
+                          grepl("Stand", covar_grid) ~ "canonical",
+                          grepl("simple_het", covar_grid) ~ "canonical",
+                          grepl("null", covar_grid) ~ "canonical",
+                          grepl("identity", covar_grid) ~ "canonical",
+                          grepl("equal_effects", covar_grid) ~ "canonical",
+                          grepl("_and_Midwest", covar_grid) ~ "hyp",
+                          grepl("Midwest", covar_grid) ~ "hyp",
+                          grepl("Gulf", covar_grid) ~ "hyp",
+                          TRUE ~ "other"
+                          ))
+
+get_significant_results(m_gulf_gr_hyp)[1:10]
+
 ```
+Not many SNPs with U_hyp loading > 10% and SNP with log10BF > 2:
+Both GR: 10 & 9 SNPs   cgdd_12c_18d_Gulf_and_Midwest x2 (0.9 and 0.10)
+Both FL: 2 SNPs        dyln_change_sec_Midwest
 
+Gulf FL: 0 SNPs
+Gulf GR: 11 & 2 SNPs   cgdd_12c_18d_Gulf_and_Midwest.10 & tave_18d_Gulf_and_Midwest.6
+
+Midw GR: 0 SNPs
+Midw FL: 11 SNPs       cgdd_12c_gr2fl_Gulf_and_Midwest
+
+### iii. Plot setup
 Make a list containing all of the ggplots for all entries in the covariance grid represented by `pw_df`.
@@
 ```{r mashhattan plots loop}
+pw_df <- pw_gg
+
 gglist <- list()
-for(i in seq_along(names(pw_df)[-(1:2)])){
-var <- names(pw_df)[i+2]   
-gglist[[i]] <- ggman_df %>%
+log10BF <- expression(paste("log"[10], plain("(Bayes Factor)")))
+for(i in seq_along(names(pw_df)[-(1:5)])){
+  var <- names(pw_df)[i+4]  
+  gglist[[i]] <- pw_df %>%
     arrange(!! sym(var)) %>%
@@
                    shape = as.factor(.data$Chr))) +
-  facet_wrap(~ .data$Chr, nrow = 1, scales = "free_x", strip.position = "bottom") +
+    facet_wrap(~ .data$Chr, nrow = 1, scales = "free_x", 
+               strip.position = "bottom") +
     scale_color_viridis_c(option = "D", end = 0.95) + 
@@
 
-Which covariance grids have SNPs with log10BF > 2 that load strongly onto them?
-For Gulf mash:
-GR50_Gulf.6: many, almost all < 2
-GR50_Midwest.6: many, almost all < 2
-equal_effects.7: many, almost all < 2
-GR50_Gulf.6.7: many, almost all < 2
-equal_effects.8: some, all <2
-ED_PCA_1.9 many, many > 2
-ED_PCA_2.9 many, few > 2
-ED_PCA_4.9 some, few > 2
-ED_tPCA.9: many, many > 2
-ED_PCA_3.10: many, few > 2
-ED_PCA_5.10: some, few > 2
-OK.10 many, some > 2, many >0.8 loading & > 2
-SD.10 many, some > 2
-cgdd_12c_G&M.10: few, few > 2, some with 80%+ loading here
-
-ED_tPCA.12: many, many > 2, many >0.8 loading esp at high log10BF
-SD.12: many, many > 2, many > 0.8 loading esp at high log10BF
-
-TX1.15: few bg, few >2
-
-```{r snps with high sig and high covar loadings}
-var <- names(ggman_df)[ncol(ggman_df)-1]  
-ggman_df %>%
-  pivot_longer(cols = null:var, names_to = "covar_grid", values_to = "loading") %>%
-  filter(loading > 0.5) %>%
-  group_by(log10BayesFactor > 2, covar_grid, loading > 0.8) %>%
-  tally()
-```
 # -----------
@@
 
+### Rice/thaliana FT
+
+anno_df: All are above a 10% FDR.
+If there are more than 500 SNPs above the FDR, only returns the top 500 by p value.
+Only gets the closest gene to each SNP.
+Should be three site sets and three subpop sets, several (four?) flowering phenotypes.
+
+os_annos: OsGene and AtGene columns of this df indicate if the gene has a homolog in rice or A. thaliana with a functional study. Key is the locusName column.
+```{r}
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+ftanno <- read_csv(file.path(workingdir, "data/Pvirgatum_v5.1.annotation_FTGeneHomologs.csv"))
+ftkey <- read_delim(file.path(workingdir, "data", "OsGene_keyword_table.txt"),
+                    delim = "\t")
+ftanno %>%
+  filter(!is.na(OsGene)) %>%
+  left_join(ftkey, by = c("OsGene" = "Symbol")) %>%
+  filter(Keyword %in% c("flowering time", "heading date")) %>% unique()
+
+osgene_keyword_df <- ftkey %>%
+  dplyr::select(-RAPdb, -MSU) %>%
+  group_by(Symbol, Keyword) %>%
+  dplyr::slice(1) %>%
+  pivot_wider(names_from = Keyword, values_from = Title)
+
+osgene_key <- osgene_keyword_df %>%
+  dplyr::select(Symbol, `flowering time`, `heading date`, anther, flower,
+                `floral meristem`, `floral meristem determinacy`, flowering,
+                `flower development`, `carpel differentiation`, stamen, 
+                `stamen number`, everything()) %>%
+  pivot_longer(cols = -Symbol, names_to = "Keyword", values_to = "Title") %>%
+  filter(!is.na(Title)) %>%
+  pivot_wider(names_from = Keyword, values_from = Title)
+
+os_annos <- ftanno %>%
+  filter(OsGene %in% osgene_key$Symbol | !is.na(AtGene)) %>%
+  mutate(Pos_Mb = round(start/10000)/100,
+         #Pos_Mb_e = floor(start/10000)/100,
+         POS_bin = round(Pos_Mb*100/2)) %>%# ,
+         #POS_bin_hi = round(Pos_Mb_e*100/2))
+  dplyr::select(OsGene, AtGene, locusName, everything()) %>%
+  dplyr::rename(`Gene ID` = locusName)
+1599 / 83260
+```
+
+# Annotations with rice/thaliana functional annotation papers for specific SNPs
+```{r}
+library(AnnotationDbi)
+txdb <- loadDb("~/Github/pvdiv-genome/Pvirgatum_516_v5.1.gene.txdb.sqlite")
+not_all_na <- function(x) any(!is.na(x))
+not_any_na <- function(x) all(!is.na(x))
+region_size <- 20000
+
+gg_annos <- pw_gg %>% 
+  filter((cgdd_12c_18d_Gulf_and_Midwest.10 > .1 | tave_18d_Gulf_and_Midwest.6 > 0.1) & log10BayesFactor > 2) %>%
+  mutate(CHR = Chr,
+         start = Pos - (region_size/2),
+         end = Pos + (region_size/2)) %>%
+  pvdiv_table_topsnps(., type = "table", txdb = txdb)
+
+gg_annos %>%
+  inner_join(os_annos, by = c("Gene ID")) %>%
+  left_join(osgene_key, by = c("OsGene" = "Symbol")) %>%
+  dplyr::select(where(not_all_na))
+
+top_annos <- pw_gg %>%
+  filter(log10BayesFactor > 2) %>%
+  mutate(CHR = Chr,
+         start = Pos - (region_size/2),
+         end = Pos + (region_size/2)) %>%
+  pvdiv_table_topsnps(., type = "table", txdb = txdb) %>%
+  inner_join(os_annos, by = c("Gene ID")) %>%
+  left_join(osgene_key, by = c("OsGene" = "Symbol")) %>%
+  dplyr::select(where(not_all_na)) %>%
+  group_by(Chr, POS, `Gene ID`) %>%
+  dplyr::slice(1)
+```
+
+```{r function}
+library(AnnotationDbi)
+not_all_na <- function(x) any(!is.na(x))
+txdb <- loadDb("~/Github/pvdiv-genome/Pvirgatum_516_v5.1.gene.txdb.sqlite")
+
+get_post_weights <- function(m){
+  log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
+      as.data.frame() %>%
+      rownames_to_column(var = "value") %>%
+      mutate(value = as.integer(.data$value)) %>%
+      as_tibble() %>%
+      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
+      dplyr::rename(log10BayesFactor = .data$V1) %>%
+      dplyr::select(-.data$value)
+
+  pw_df <- as_tibble(m$posterior_weights)
+  pw_df <- switchgrassGWAS:::get_marker_df(m = m) %>%
+    bind_cols(pw_df)
+
+  ggman_df <- pw_df %>%
+    #left_join(m_meta) %>%
+    separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
+             extra = "merge", convert = TRUE) %>%
+    left_join(log10bf_df, by = "Marker") %>%
+    arrange(.data$Chr, .data$Pos) 
+  return(ggman_df)
+}
+
+get_post_annos <- function(m, region_size = 20000, BFthresh = 2){
+  pw_df <- get_post_weights(m)
+  pw_annos <- pw_df %>%
+    filter(log10BayesFactor > BFthresh) %>%
+    mutate(CHR = Chr,
+           start = Pos - (region_size/2),
+           end = Pos + (region_size/2)) %>%
+    pvdiv_table_topsnps(., type = "table", txdb = txdb) %>%
+    inner_join(os_annos, by = c("Gene ID")) %>%
+    left_join(osgene_key, by = c("OsGene" = "Symbol")) %>%
+    dplyr::select(where(not_all_na)) %>%
+    group_by(Chr, POS, `Gene ID`) %>%
+    dplyr::slice(1) %>% 
+    arrange(desc(log10BayesFactor))
+  return(pw_annos)
+}
+```
+
+```{r}
+annos_gg <- get_post_annos(m_gulf_gr_hyp)
+annos_gf <- get_post_annos(m_gulf_fl_hyp)
+annos_mg <- get_post_annos(m_midw_gr_hyp)
+annos_mf <- get_post_annos(m_midw_fl_hyp)
+annos_gamg <- get_post_annos(m_gam_gr_hyp)
+annos_gamf <- get_post_annos(m_gam_fl_hyp)
+```
+
+```{r}
+annos_gg
+```
+
+### Get mash SNP metadata
 ```{r}
@@
 
+Example to modify from a previous script:
+```{r}
+get_U_by_mass(m)
+i=2
+names(Ulist)[i]
+Ulist <- switchgrassGWAS:::get_Ulist(m)
+b1 <- svd(Ulist[[i]])
+b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 11.5
+b1$d[2] / sum(b1$d) # get to 60%
+b1$d[3] / sum(b1$d) # 
+b1$d[4] / sum(b1$d) # 
+b1$d[5] / sum(b1$d) #
+
+sum(b1$d[1:2]) / sum(b1$d) # First 5 eigenvectors explain 1% or more variance, and 27.7% of total variance.
+for(g in 1:5){
+barplot(b1$v[,g]/Ulist[[1]][which.max(abs(Ulist[[1]][,g])),g])
+}
+
+ggb1 <- tibble(`First eigenvector (PVE = 100%)` = b1$v[,1]/Ulist[[1]][which.max(abs(Ulist[[1]][,1])),1],
+       Phenotype = switchgrassGWAS:::get_colnames(m)) %>%
+  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_")) %>%
+  separate(Phenotype, into = c("Function", "Garden"), sep = -3) %>%
+  mutate(Function = str_remove(Function, "_$"),
+         Garden = str_remove(Garden, "^_"))
+ggb1$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
+ggb1_plot <- ggb1 %>%
+  ggplot(aes(x = Garden, y = `First eigenvector (PVE = 100%)`)) +
+  geom_bar(aes(group = Function, color = Function, fill = Function), stat = "identity", position = "dodge") +
+  theme(legend.position = "top", panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(hjust = 1, angle = 45)) +
+  geom_hline(yintercept = 0, linetype = 2) +
+  scale_color_viridis_d(option = "B", end = 0.7, 
+                      direction = -1) +
+  scale_fill_viridis_d(option = "B", begin = 0.1, end = 0.7, direction = -1) 
+save_plot(paste0("../manuscript/plots/SVD_of_greenup_Gulf_", names(Ulist)[i], ".png"), base_height = 4, plot = ggb1_plot)
+
+ggb2 <- tibble(`First eigenvector (PVE = 21.7%)` = b1$v[,2]/Ulist[[1]][which.max(abs(Ulist[[1]][,2])),2],
+       Phenotype = switchgrassGWAS:::get_colnames(m)) %>%
+  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_")) %>%
+  separate(Phenotype, into = c("Function", "Garden"), sep = -3) %>%
+  mutate(Function = str_remove(Function, "_$"),
+         Garden = str_remove(Garden, "^_"))
+ggb2$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
+ggb2_plot <- ggb2 %>%
+  ggplot(aes(x = Garden, y = `First eigenvector (PVE = 21.7%)`)) +
+  geom_bar(aes(group = Function, color = Function, fill = Function), stat = "identity", position = "dodge") +
+  theme(legend.position = "top", panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(hjust = 1, angle = 45)) +
+  geom_hline(yintercept = 0, linetype = 2) +
+  scale_color_viridis_d(option = "B", end = 0.7, 
+                      direction = -1) +
+  scale_fill_viridis_d(option = "B", begin = 0.1, end = 0.7, direction = -1) 
+save_plot(paste0("../manuscript/plots/SVD_of_greenup_Gulf_", names(Ulist)[i], "Eigenvector_2.png"), base_height = 4, plot = ggb2_plot)
+
+```
+
 ## 3f. DS & AP
@@
 We characterized patterns of differential sensitivity and antagonistic pleiotropy between all SNPs and all pairs of gardens.
+
+`get_GxE` and `mash_plot_pairwise_sharing()` should do the trick here.

### 2021-01-15 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Writeup for mash figure 2 covariance matrices

@@
 
-get_pw_filtered(m_gulf_gr_hyp, pw_thresh = 0.1, BF_thresh = 2) %>% 
+get_pw_filtered(m_midw_fl_hyp, pw_thresh = 0.1, BF_thresh = 2) %>% 
   mutate(Type = case_when(grepl("ED_", covar_grid) ~ "data-driven",
@@
 Both GR: 10 & 9 SNPs   cgdd_12c_18d_Gulf_and_Midwest x2 (0.9 and 0.10)
-Both FL: 2 SNPs        dyln_change_sec_Midwest
+Both FL: 2 SNPs        dyln_change_sec_Midwest.7
 
@@
 Midw GR: 0 SNPs
-Midw FL: 11 SNPs       cgdd_12c_gr2fl_Gulf_and_Midwest
+Midw FL: 11 SNPs       cgdd_12c_gr2fl_Gulf_and_Midwest.8
 
@@
 # Annotations with rice/thaliana functional annotation papers for specific SNPs
-```{r}
-library(AnnotationDbi)
-txdb <- loadDb("~/Github/pvdiv-genome/Pvirgatum_516_v5.1.gene.txdb.sqlite")
-not_all_na <- function(x) any(!is.na(x))
-not_any_na <- function(x) all(!is.na(x))
-region_size <- 20000
-
-gg_annos <- pw_gg %>% 
-  filter((cgdd_12c_18d_Gulf_and_Midwest.10 > .1 | tave_18d_Gulf_and_Midwest.6 > 0.1) & log10BayesFactor > 2) %>%
-  mutate(CHR = Chr,
-         start = Pos - (region_size/2),
-         end = Pos + (region_size/2)) %>%
-  pvdiv_table_topsnps(., type = "table", txdb = txdb)
-
-gg_annos %>%
-  inner_join(os_annos, by = c("Gene ID")) %>%
-  left_join(osgene_key, by = c("OsGene" = "Symbol")) %>%
-  dplyr::select(where(not_all_na))
-
-top_annos <- pw_gg %>%
-  filter(log10BayesFactor > 2) %>%
-  mutate(CHR = Chr,
-         start = Pos - (region_size/2),
-         end = Pos + (region_size/2)) %>%
-  pvdiv_table_topsnps(., type = "table", txdb = txdb) %>%
-  inner_join(os_annos, by = c("Gene ID")) %>%
-  left_join(osgene_key, by = c("OsGene" = "Symbol")) %>%
-  dplyr::select(where(not_all_na)) %>%
-  group_by(Chr, POS, `Gene ID`) %>%
-  dplyr::slice(1)
-```
 
@@
 ```{r}
-annos_gg
+annos_gg %>%
+  filter(cgdd_12c_18d_Gulf_and_Midwest.10 > 0.05 |
+           tave_18d_Gulf_and_Midwest.6 > 0.05) %>%
+  dplyr::select(where(not_all_na))
+
+annos_gamg %>%
+  filter(cgdd_12c_18d_Gulf_and_Midwest.10 > 0.05 |
+           cgdd_12c_18d_Gulf_and_Midwest.9 > 0.05) %>%
+  dplyr::select(where(not_all_na))
+
+annos_gamf %>%
+  # dplyr::select(starts_with("dyln_change_sec"))
+  filter(dyln_change_sec_Midwest.7 > 0.05) %>%
+  dplyr::select(where(not_all_na))
+annos_mf %>%
+  # dplyr::select(starts_with("cgdd_12c_gr2fl_Gulf_and"))
+  filter(cgdd_12c_gr2fl_Gulf_and_Midwest.8 > 0.05) %>%
+  dplyr::select(where(not_all_na))
 ```
+Not many SNPs with U_hyp loading > 10% and SNP with log10BF > 2:
+Both GR: 10 & 9 SNPs   cgdd_12c_18d_Gulf_and_Midwest x2 (0.9 and 0.10)
+Both FL: 2 SNPs        dyln_change_sec_Midwest.7
+
+Gulf FL: 0 SNPs
+Gulf GR: 11 & 2 SNPs   cgdd_12c_18d_Gulf_and_Midwest.10 & tave_18d_Gulf_and_Midwest.6
+
+Midw GR: 0 SNPs
+Midw FL: 11 SNPs       cgdd_12c_gr2fl_Gulf_and_Midwest.8
 

### 2021-01-19 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Edit manuscript to add cov matrix examples to Figure 2

@@
 
+# ggcov
 ggcov
@@
 -2*(m_midw_fl_no$loglik - m_midw_fl_hyp$loglik) # v. negative 
-1 - pchisq(-2*(m_midw_fl_no$loglik - m_midw_fl_hyp$loglik), 30*32)
-1 - pchisq(-2*(m_midw_gr_no$loglik - m_midw_gr_hyp$loglik), 15*32) # sig
+1 - pchisq(-2*(m_midw_fl_no$loglik - m_midw_fl_hyp$loglik), 30*36)
+1 - pchisq(-2*(m_midw_gr_no$loglik - m_midw_gr_hyp$loglik), 15*36) # sig
 
@@
 -2*(m_gam_fl_no$loglik - m_gam_fl_hyp$loglik) # v. positive
-1 - pchisq(-2*(m_gam_fl_no$loglik - m_gam_fl_hyp$loglik), 30*32) # sig
-1 - pchisq(-2*(m_gam_gr_no$loglik - m_gam_gr_hyp$loglik), 15*32) # sig
+1 - pchisq(-2*(m_gam_fl_no$loglik - m_gam_fl_hyp$loglik), 30*36) # sig
+1 - pchisq(-2*(m_gam_gr_no$loglik - m_gam_gr_hyp$loglik), 15*36) # sig
 
@@
 1-pchisq((6.89), 1) # LR of 6.89 gives p of ~0.01
+8+7+6+5+4+3+2+1
 ```
@@
 
+```{r}
+get_U_by_mass(m_gulf_gr_hyp,thresh = 0.01)
+cov_gr1 <- mash_plot_Ulist(m_gulf_gr_hyp, range = 18, limits = TRUE)
+cov_gr2 <- mash_plot_Ulist(m_gam_gr_hyp, 13, limits = TRUE)
+
+fig_eqeff <- cov_gr1$equal_effects_df %>%
+  ggplot(aes(x = .data$rowU, y = .data$colU)) +
+  switchgrassGWAS::theme_oeco +
+  geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
+  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                "white", "#27AD81FF", "#5DC863FF",
+                                "#FDE725FF"),
+                       limits = c(-1, 1)) +
+  theme(legend.position = "none",
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        axis.line.x = element_blank(),
+        axis.line.y = element_blank(),
+        axis.title.y.right = element_text()) +
+  labs(x = "", y = "", subtitle = "equal_effects")
+
+fig_ok_eff <- cov_gr2$Stand_BhatGulf_and_Midwest_GR50_OK_df %>%
+  ggplot(aes(x = .data$rowU, y = .data$colU)) +
+  switchgrassGWAS::theme_oeco +
+  geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
+  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                "white", "#27AD81FF", "#5DC863FF",
+                                "#FDE725FF"),
+                       limits = c(-1, 1)) +
+  theme(legend.position = "none",
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        axis.line.x = element_blank(),
+        axis.line.y = element_blank()) +
+  xlab("") + ylab("") + ggtitle("single_effect_OK")
+```
+
+```{r second try making covar examples}
+phe_hyp <- c("tave_10d", "tave_18d", "dyln_fl50", "cgdd_12c_gr2fl")
+
+for(i in seq_along(phe_hyp)){
+  for(k in seq_along(subpop_v2)){
+
+    if (i %in% c(1,2)) {
+      cov_df <- phe_gr_df %>%
+        ungroup() %>%
+        left_join(select(sites, SITE, manu_site), by = "SITE") %>%
+        filter(SUBPOP %in% subpop_v2[[k]]) %>%
+        select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
+        group_by(PLANT_ID, SUBPOP, manu_site) %>%
+        mutate(IND = row_number()) %>%
+        pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
+        mutate(PLANT_ID = as.factor(PLANT_ID),
+               IND = as.factor(IND)) %>%
+        select(-IL) %>%
+        select(PLANT_ID, SUBPOP, IND, TX1, TX2, TX3, OK, MO, NE, MI, SD)
+    } else {
+      cov_df <- phe_fl_df %>%
+        ungroup() %>%
+        left_join(select(sites, SITE, manu_site), by = "SITE") %>%
+        filter(SUBPOP %in% subpop_v2[[k]]) %>%
+        select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
+        group_by(PLANT_ID, SUBPOP, manu_site) %>%
+        mutate(IND = row_number()) %>%
+        pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
+        mutate(PLANT_ID = as.factor(PLANT_ID),
+               IND = as.factor(IND)) %>%
+        select(-IL) %>%
+        select(PLANT_ID, SUBPOP, IND, TX1, TX2, TX3, OK, MO, NE, MI, SD)
+    }
+
+  cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
+  diag(cor_phe) <- cov_df %>%
+    ungroup() %>%
+    summarise(TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
+              TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
+              TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE),
+              OK = sqrt(var(OK, na.rm = TRUE))/mean(OK, na.rm = TRUE),
+              MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
+              NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
+              MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),
+              SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE)) %>%
+    as_vector()
+
+  if(isSymmetric(cor_phe)){
+        for(m in 1:nrow(cor_phe)){
+          for(j in 1:ncol(cor_phe)){
+            if(m < j){
+              cor_phe[m, j] <- NA
+            }
+          }
+        }
+    }
+  U1 <- as_tibble(cor_phe, rownames = "rowU", .name_repair = "unique") %>%
+          pivot_longer(cols = -.data$rowU, names_to = "colU",
+                       values_to = "covar") %>%
+          filter(!is.na(.data$covar))
+  U1$colU <- factor(U1$colU, levels = colnames(cor_phe))
+  U1$rowU <- factor(U1$rowU, levels = colnames(cor_phe))
+  U1 <- U1 %>%
+    mutate(phe = phe_hyp[i],
+           subpop = names(subpop_v2)[k],
+           subpop = case_when(subpop == "Gulf_and_Midwest" ~ "Both",
+                              TRUE ~ subpop))
+
+  if(i == 1 & k == 1){
+    cov_flex <- U1
+  } else {
+    cov_flex <- cov_flex %>%
+      full_join(U1)
+  }
+    }
+}
+
+cov_flex$phe <- factor(cov_flex$phe, levels = c("tave_18d", "tave_10d", "dyln_fl50", "cgdd_12c_gr2fl"))
+
+figure2_a <- cov_flex %>%
+  filter(phe %in% c("tave_10d", "tave_18d")) %>%
+  ggplot(aes(x = .data$rowU, y = .data$colU)) +
+  switchgrassGWAS::theme_oeco +
+  geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
+  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                "white", "#27AD81FF", "#5DC863FF",
+                                "#FDE725FF"),
+                       limits = c(-1, 1)) +
+  #geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
+  facet_grid(rows = vars(phe), cols = vars(subpop)) + 
+  theme(legend.position = "right",
+        #axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        axis.line.x = element_blank(),
+        axis.line.y = element_blank(),
+        panel.spacing.x = unit(-0, 'cm')) +
+  xlab("") + ylab("")
+
+legend <- get_legend(figure2_a)  
+figure2_a <- cov_flex %>%
+  filter(phe %in% c("tave_10d", "tave_18d")) %>%
+  ggplot(aes(x = .data$rowU, y = .data$colU)) +
+  switchgrassGWAS::theme_oeco +
+  geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
+  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                "white", "#27AD81FF", "#5DC863FF",
+                                "#FDE725FF"),
+                       limits = c(-1, 1)) +
+  #geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
+  facet_grid(rows = vars(phe), cols = vars(subpop)) + 
+  theme(legend.position = "none",
+        #axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        axis.line.x = element_blank(),
+        axis.line.y = element_blank(),
+        panel.spacing.x = unit(-0, 'cm')) +
+  xlab("") + ylab("")
+
+figure2_d <- cov_flex %>%
+  filter(phe %in% c("cgdd_12c_gr2fl", "dyln_fl50")) %>%
+  ggplot(aes(x = .data$rowU, y = .data$colU)) +
+  switchgrassGWAS::theme_oeco +
+  geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
+  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                "white", "#27AD81FF", "#5DC863FF",
+                                "#FDE725FF"),
+                       limits = c(-1, 1)) +
+  #geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
+  facet_grid(rows = vars(phe), cols = vars(subpop)) + 
+  theme(legend.position = "none",
+        #axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        axis.line.x = element_blank(),
+        axis.line.y = element_blank(),
+        panel.spacing.x = unit(-0, 'cm')) +
+  xlab("") + ylab("")
+```
+
 ```{r make rough draft figure 2}
@@
   coord_flip() + scale_fill_viridis_d(end = 0.9) +
-  theme(legend.position = c(-0.30, 0.3),
+  theme(legend.position = c(-0.4, 0.3),
         axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45), 
-        panel.spacing.x = unit(-0, 'cm'))
+        panel.spacing.x = unit(-0, 'cm')) #+ labs(x = "")
 f2c <- df_2fl_a %>%
@@
         axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45), 
-        panel.spacing.x = unit(-0, 'cm'))
+        panel.spacing.x = unit(-0, 'cm')) #+ labs(x = "")
 
@@
 Midwest greenup: 28.6% flowering 45.1%
+
+```{r}
+ggdraw() +
+  draw_plot(figure2_a, x = -.02, y = 0.51, width = 0.32, height = 0.35) +
+  draw_plot(fig_eqeff, x = 0.02, y = 0.35, width = 0.125, height = 0.2) +
+  draw_plot(figure2_d, x = -.02, y = 0.03, width = 0.32, height = 0.35) +
+  draw_plot(f2a, x = 0.3, y = 0.47, width = 0.55, height = 0.53) +
+  draw_plot(legend, x = 0.15, y = 0.34, width = 0.1, height = 0.2) +
+  draw_plot(f2b, x = 0.85, y = 0.47, width = 0.15, height = 0.53) +
+  draw_plot(f2c, x = 0.3, y = 0, width = 0.55, height = 0.47) +
+  draw_plot(f2d, x = 0.85, y = 0, width = 0.15, height = 0.47)
+
+save_plot(filename = "Figure_2_Covariance_matrices_and_posterior_weights.png", plot = last_plot(), base_height = 8, base_width = 9)
+```
 

### 2021-01-23 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Finish mash hypothesis-driven covariance matrix Results section

@@
 annos_gg %>%
-  filter(cgdd_12c_18d_Gulf_and_Midwest.10 > 0.05 |
-           tave_18d_Gulf_and_Midwest.6 > 0.05) %>%
+  filter(cgdd_12c_18d_Gulf_and_Midwest.10 > 0.1 |
+           tave_18d_Gulf_and_Midwest.6 > 0.1) %>%
   dplyr::select(where(not_all_na))
@@
 annos_gamg %>%
-  filter(cgdd_12c_18d_Gulf_and_Midwest.10 > 0.05 |
-           cgdd_12c_18d_Gulf_and_Midwest.9 > 0.05) %>%
+  filter(cgdd_12c_18d_Gulf_and_Midwest.10 > 0.1 |
+           cgdd_12c_18d_Gulf_and_Midwest.9 > 0.1) %>%
   dplyr::select(where(not_all_na))
 
+annos_mf %>%
+  # dplyr::select(starts_with("cgdd_12c_gr2fl_Gulf_and"))
+  filter(cgdd_12c_gr2fl_Gulf_and_Midwest.8 > 0.1) %>%
+  dplyr::select(where(not_all_na))
 annos_gamf %>%
   # dplyr::select(starts_with("dyln_change_sec"))
-  filter(dyln_change_sec_Midwest.7 > 0.05) %>%
+  filter(dyln_change_sec_Midwest.7 > 0.1) %>%
   dplyr::select(where(not_all_na))
-annos_mf %>%
-  # dplyr::select(starts_with("cgdd_12c_gr2fl_Gulf_and"))
-  filter(cgdd_12c_gr2fl_Gulf_and_Midwest.8 > 0.05) %>%
-  dplyr::select(where(not_all_na))
+
 ```

### 2021-01-26 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Add data-driven results draft figure and start Results for this figure

@@
 pvdiv_plot_cov <- function(covdf){
-  U1 <- cor_phe
+  U1 <- covdf
   if(isSymmetric(U1)){
@@
           filter(!is.na(.data$covar))
-  U1$colU <- factor(U1$colU, levels = colnames(cor_phe))
-  U1$rowU <- factor(U1$rowU, levels = colnames(cor_phe))
+  U1$colU <- factor(U1$colU, levels = colnames(covdf))
+  U1$rowU <- factor(U1$rowU, levels = colnames(covdf))
 
@@
 
-## 3e. Data-driven mat
+# -------------------------
+
+## 3e. Data-driven matrices
+
+Run functions on line 743 first - to get posterior weights.
+```{r}
+pw_gamg %>%
+  filter(log10BayesFactor >= 2)
+pw_gg %>%
+  filter(log10BayesFactor >= 2) %>%
+  select(-(Marker:value), -log10BayesFactor) %>%
+  colSums()/902 
+pw_mg %>%
+  filter(log10BayesFactor >= 2)
+pw_gamf %>%
+  filter(log10BayesFactor >= 2)
+pw_gf %>%
+  filter(log10BayesFactor >= 2)
+pw_mf %>%
+  filter(log10BayesFactor >= 2)
+```
+Gulf greenup sig 902 SNPs: 15.1% ED_PCA_1.9 
+24.713% ED_tPCA.9
+24.315% ED_tPCA.12
+5.722 ED_tPCA.13
 
@@
 
-Example to modify from a previous script:
-```{r}
-get_U_by_mass(m)
-i=2
-names(Ulist)[i]
+ED_tPCA is made up of the 5 ED_PCA_#, each of which explains a different PVE of the tPCA.
+SNPs can have high pw's for just one of these PCA's, and/or for the tPCA, which is a weighted combination of all 5.
+While the PCA's typically have simple, garden-based effects, the tPCA is a more complicated weighted combination of these effects.
+
+### i. Function
+
+```{r functions to plot svd of data driven covar matrices}
+get_svd_df <- function(m, i = NULL, eigen_v = 1){
   Ulist <- switchgrassGWAS:::get_Ulist(m)
+  stopifnot(eigen_v %in% c(1,2,3,4,5))
+  if(is.null(i)){
+    i_v <- which(names(Ulist) %in% names(Ulist)[get_U_by_mass(m, thresh = 0.1)])
+    i = i_v[1]
+  } else if(is.character(i)){
+    i <- which(names(Ulist) %in% i)
+  }
+  stopifnot(i %in% 1:length(Ulist))
+
   b1 <- svd(Ulist[[i]])
-b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 11.5
-b1$d[2] / sum(b1$d) # get to 60%
-b1$d[3] / sum(b1$d) # 
-b1$d[4] / sum(b1$d) # 
-b1$d[5] / sum(b1$d) #
-
-sum(b1$d[1:2]) / sum(b1$d) # First 5 eigenvectors explain 1% or more variance, and 27.7% of total variance.
-for(g in 1:5){
-barplot(b1$v[,g]/Ulist[[1]][which.max(abs(Ulist[[1]][,g])),g])
-}
-
-ggb1 <- tibble(`First eigenvector (PVE = 100%)` = b1$v[,1]/Ulist[[1]][which.max(abs(Ulist[[1]][,1])),1],
+  PVE_EV <- b1$d[eigen_v] / sum(b1$d) # Proportion of variance explained by this eigenvector
+  ggb1 <- tibble(`Eigenvector` = b1$v[,eigen_v]/Ulist[[i]][which.max(abs(Ulist[[i]][,eigen_v])),eigen_v],
        Phenotype = switchgrassGWAS:::get_colnames(m)) %>%
+  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_and_Midwest_")) %>%
   mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_")) %>%
-  separate(Phenotype, into = c("Function", "Garden"), sep = -3) %>%
-  mutate(Function = str_remove(Function, "_$"),
-         Garden = str_remove(Garden, "^_"))
+  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatMidwest_")) %>%
+  separate(Phenotype, into = c("Phenotype", "Garden"), sep = -3) %>%
+  mutate(Phenotype = str_remove(Phenotype, "_$"),
+         Garden = str_remove(Garden, "^_"),
+         PVE_EV = PVE_EV)
   ggb1$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
-ggb1_plot <- ggb1 %>%
-  ggplot(aes(x = Garden, y = `First eigenvector (PVE = 100%)`)) +
-  geom_bar(aes(group = Function, color = Function, fill = Function), stat = "identity", position = "dodge") +
-  theme(legend.position = "top", panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(hjust = 1, angle = 45)) +
+  return(ggb1)
+}
+
+mash_plot_svd <- function(m, i = NULL, eigen_v = 1){
+  b1_df <- get_svd_df(m = m, i = i, eigen_v = eigen_v)
+  if(eigen_v == 1){
+    ylab <- paste0("First eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
+  } else if(eigen_v == 2){
+    ylab <- paste0("Second eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
+  } else if(eigen_v == 3){
+    ylab <- paste0("Third eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
+  } else if(eigen_v == 4){
+    ylab <- paste0("Fourth eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
+  } else if(eigen_v == 5){
+    ylab <- paste0("Fifth eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
+  }
+  ggb1 <- b1_df %>%
+    ggplot(aes(x = Garden, y = `Eigenvector`)) +
+    switchgrassGWAS::theme_oeco + 
+    geom_bar(stat = "identity", position = "dodge") +
+    theme(legend.position = "top", panel.spacing.x = unit(0.1, 'cm'),
+          axis.text.x = element_text(hjust = 1, angle = 45)) +
     geom_hline(yintercept = 0, linetype = 2) +
@@
                         direction = -1) +
-  scale_fill_viridis_d(option = "B", begin = 0.1, end = 0.7, direction = -1) 
-save_plot(paste0("../manuscript/plots/SVD_of_greenup_Gulf_", names(Ulist)[i], ".png"), base_height = 4, plot = ggb1_plot)
-
-ggb2 <- tibble(`First eigenvector (PVE = 21.7%)` = b1$v[,2]/Ulist[[1]][which.max(abs(Ulist[[1]][,2])),2],
-       Phenotype = switchgrassGWAS:::get_colnames(m)) %>%
-  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_")) %>%
-  separate(Phenotype, into = c("Function", "Garden"), sep = -3) %>%
-  mutate(Function = str_remove(Function, "_$"),
-         Garden = str_remove(Garden, "^_"))
-ggb2$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
-ggb2_plot <- ggb2 %>%
-  ggplot(aes(x = Garden, y = `First eigenvector (PVE = 21.7%)`)) +
-  geom_bar(aes(group = Function, color = Function, fill = Function), stat = "identity", position = "dodge") +
-  theme(legend.position = "top", panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(hjust = 1, angle = 45)) +
-  geom_hline(yintercept = 0, linetype = 2) +
-  scale_color_viridis_d(option = "B", end = 0.7, 
+    scale_fill_viridis_d(option = "B", begin = 0.1, end = 0.7, 
                          direction = -1)  +
-  scale_fill_viridis_d(option = "B", begin = 0.1, end = 0.7, direction = -1) 
-save_plot(paste0("../manuscript/plots/SVD_of_greenup_Gulf_", names(Ulist)[i], "Eigenvector_2.png"), base_height = 4, plot = ggb2_plot)
-
+    labs(y = ylab)
+  return(ggb1)
+}
+```
+
+```{r}
+gxe_gg <- get_GxE(m_gulf_gr_hyp)
+gxe_mg <- get_GxE(m_midw_gr_hyp)
+gxe_gamg <- get_GxE(m_gam_gr_hyp)
+gxe_gf <- get_GxE(m_gulf_fl_hyp)
+gxe_mf <- get_GxE(m_midw_fl_hyp)
+gxe_gamf <- get_GxE(m_gam_fl_hyp)
+```
+
+```{r}
+f3a <- mash_plot_svd(m_gulf_gr_hyp, i = "ED_tPCA") 
+f3b <- mash_plot_svd(m_gulf_gr_hyp, i = "ED_tPCA", eigen_v = 2) 
+f3e <- mash_plot_svd(m_midw_gr_hyp, i = "ED_tPCA") 
+f3f <- mash_plot_svd(m_midw_gr_hyp, i = "ED_tPCA", eigen_v = 2) 
+f3i <- mash_plot_svd(m_gam_gr_hyp, i = "ED_tPCA") 
+f3j <- mash_plot_svd(m_gam_gr_hyp, i = "ED_tPCA", eigen_v = 2) 
+f3m <- mash_plot_svd(m_gulf_fl_hyp, i = "ED_tPCA") 
+f3n <- mash_plot_svd(m_gulf_fl_hyp, i = "ED_tPCA", eigen_v = 2) 
+f3q <- mash_plot_svd(m_midw_fl_hyp, i = "ED_tPCA") 
+f3r <- mash_plot_svd(m_midw_fl_hyp, i = "ED_tPCA", eigen_v = 2) 
+f3u <- mash_plot_svd(m_gam_fl_hyp, i = "ED_tPCA") 
+f3v <- mash_plot_svd(m_gam_fl_hyp, i = "ED_tPCA", eigen_v = 2)
+
+f3c <- mash_plot_pairwise_sharing(corrmatrix = gxe_gg$S_AP, reorder = FALSE)
+f3d <- mash_plot_pairwise_sharing(corrmatrix = gxe_gg$S_DS, reorder = FALSE)
+f3g <- mash_plot_pairwise_sharing(corrmatrix = gxe_mg$S_AP, reorder = FALSE)
+f3h <- mash_plot_pairwise_sharing(corrmatrix = gxe_mg$S_DS, reorder = FALSE)
+f3k <- mash_plot_pairwise_sharing(corrmatrix = gxe_gamg$S_AP, reorder = FALSE)
+f3l <- mash_plot_pairwise_sharing(corrmatrix = gxe_gamg$S_DS, reorder = FALSE)
+f3o <- mash_plot_pairwise_sharing(corrmatrix = gxe_gf$S_AP, reorder = FALSE)
+f3p <- mash_plot_pairwise_sharing(corrmatrix = gxe_gf$S_DS, reorder = FALSE)
+f3s <- mash_plot_pairwise_sharing(corrmatrix = gxe_mf$S_AP, reorder = FALSE)
+f3t <- mash_plot_pairwise_sharing(corrmatrix = gxe_mf$S_DS, reorder = FALSE)
+f3w <- mash_plot_pairwise_sharing(corrmatrix = gxe_gamf$S_AP, reorder = FALSE)
+f3x <- mash_plot_pairwise_sharing(corrmatrix = gxe_gamf$S_DS, reorder = FALSE)
+```
+put labels in latitudinal order
+get rid of all the STand_Bhat_pop_GR50_ garbage
+
+```{r}
+ggdraw() +
+  draw_plot(f3a, x = 0, y = 0.825, width = 0.33, height = 0.175) +
+  draw_plot(f3b, x = 0, y = 0.65, width = 0.33, height = 0.175) +
+  draw_plot(f3d$gg_corr, x = 0.165, y = 0.5, width = 0.165, height = 0.15, scale = 2) +
+  draw_plot(f3c$gg_corr, x = 0, y = 0.5, width = 0.165, height = 0.15, scale = 2) +
+  draw_plot(f3e, x = 0.33, y = 0.825, width = 0.33, height = 0.175) +
+  draw_plot(f3f, x = 0.33, y = 0.65, width = 0.33, height = 0.175) +
+  draw_plot(f3h$gg_corr, x = 0.165+.33, y = 0.5, width = 0.165, height = 0.15, scale = 2) +
+  draw_plot(f3g$gg_corr, x = 0+.33, y = 0.5, width = 0.165, height = 0.15, scale = 2)  +
+  draw_plot(f3i, x = 0.66, y = 0.825, width = 0.33, height = 0.175) +
+  draw_plot(f3j, x = 0.66, y = 0.65, width = 0.33, height = 0.175) +
+  draw_plot(f3l$gg_corr, x = 0.165+.66, y = 0.5, width = 0.165, height = 0.15, scale = 2) +
+  draw_plot(f3k$gg_corr, x = 0+.66, y = 0.5, width = 0.165, height = 0.15, scale = 2) +  ## greenup above here, flowering below
+  draw_plot(f3m, x = 0, y = 0.825-0.5, width = 0.33, height = 0.175) +
+  draw_plot(f3n, x = 0, y = 0.65-0.5, width = 0.33, height = 0.175) +
+  draw_plot(f3p$gg_corr, x = 0.165, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2) +
+  draw_plot(f3o$gg_corr, x = 0, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2) +
+  draw_plot(f3q, x = 0.33, y = 0.825-0.5, width = 0.33, height = 0.175) +
+  draw_plot(f3r, x = 0.33, y = 0.65-0.5, width = 0.33, height = 0.175) +
+  draw_plot(f3t$gg_corr, x = 0.165+.33, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2) +
+  draw_plot(f3s$gg_corr, x = 0+.33, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2)  +
+  draw_plot(f3u, x = 0.66, y = 0.825-0.5, width = 0.33, height = 0.175) +
+  draw_plot(f3v, x = 0.66, y = 0.65-0.5, width = 0.33, height = 0.175) +
+  draw_plot(f3x$gg_corr, x = 0.165+.66, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2) +
+  draw_plot(f3w$gg_corr, x = 0+.66, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2)
+
+save_plot(filename = "Figure_3_Data_driven_Covariance_matrices.png", plot = last_plot(), base_height = 8, base_width = 9)
 ```

### 2021-01-26 R/Analysis_v0.6_QTL_GxE_and_mash_comparison.Rmd
Finish mash Results section writeup

---
title: "Analysis_v0.7_QTL_GxE"
author: "Alice MacQueen"
date: "1/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TRIAL script
# -------------------------
NB: Trial script to compare the QTL to the mash results using env covar matrices.

## 3a. Log-likelihoods
```{r}
m_gulf_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_GR50_Gulf_Uhyp.rds")
m_gulf_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_FL50_Gulf_Uhyp.rds")

m_midw_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_24000_SNPs_GR50_Midwest_Uhyp.rds")
m_midw_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_33000_SNPs_FL50_Midwest_Uhyp.rds")

m_gam_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_GR50_Gulf_and_Midwest_Uhyp.rds")
m_gam_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_FL50_Gulf_and_Midwest_Uhyp.rds")

```

## QTL
```{r}
qtldf <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv")

qtldf <- qtldf %>%
  separate(flank_lo, into = c("CHR", "Pos_Mb_lo"), sep = "_", convert = TRUE) %>%
  separate(flank_hi, into = c("flank", "Pos_Mb_hi"), sep = "_", convert = TRUE) %>%
  mutate(Pos_Mb_lo = round(Pos_Mb_lo, digits = 2),
         Pos_Mb_hi = round(Pos_Mb_hi, digits = 2)) %>%
  select(CHR, Pos_Mb_lo, Pos_Mb_hi, lodcolumn, lod, everything()) %>%
  arrange(CHR, Pos_Mb_lo)

qtlplot <- qtldf %>%
  mutate(phe = case_when(lodcolumn == "FL50" ~ "50% flowering",
                              lodcolumn == "dyln_fl50" ~ "Daylength",
                              lodcolumn == "dyln_change_sec" ~ "daylength change (s)",
                              lodcolumn == "CGDD_12C" ~ "GDD, GR to FL",
                              lodcolumn == "crain_gr2fl" ~ "Rainfall, GR to FL")) %>%
  filter(phe %in% c("Daylength", "GDD, GR to FL")) %>%
  mutate(POS_lo = Pos_Mb_lo * 1000000,
         POS_hi = Pos_Mb_hi * 1000000)

qtlplot$phe <- factor(qtlplot$phe, 
                      levels = c("50% flowering", "Rainfall, GR to FL",
                                 "GDD, GR to FL", "Daylength",
                                          "daylength change (s)",
                                          "Rainfall, 1 day",
                                                  #"Rainfall, 2d sum",
                                                  "Rainfall, 3d sum",
                                                  "Rainfall, 5d sum"#,
                                                  #"Rainfall, 7d sum"
                                                  ))
```

```{r}
get_post_weights <- function(m){
  log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
      as.data.frame() %>%
      rownames_to_column(var = "value") %>%
      mutate(value = as.integer(.data$value)) %>%
      as_tibble() %>%
      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
      dplyr::rename(log10BayesFactor = .data$V1) %>%
      dplyr::select(-.data$value)

  pw_df <- as_tibble(m$posterior_weights)
  pw_df <- switchgrassGWAS:::get_marker_df(m = m) %>%
    bind_cols(pw_df)

  ggman_df <- pw_df %>%
    #left_join(m_meta) %>%
    separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
             extra = "merge", convert = TRUE) %>%
    left_join(log10bf_df, by = "Marker") %>%
    arrange(.data$Chr, .data$Pos) 
  return(ggman_df)
}

pw_gg <- get_post_weights(m_gulf_gr_hyp) %>%
  mutate(subpop = "Gulf",
         phenotype = "greenup")
pw_gf <- get_post_weights(m_gulf_fl_hyp) %>%
  mutate(subpop = "Gulf",
         phenotype = "flowering")

pw_mg <- get_post_weights(m_midw_gr_hyp) %>%
  mutate(subpop = "Midwest",
         phenotype = "greenup")
pw_mf <- get_post_weights(m_midw_fl_hyp) %>%
  mutate(subpop = "Midwest",
         phenotype = "flowering")

pw_gamg <- get_post_weights(m_gam_gr_hyp) %>%
  mutate(subpop = "Both",
         phenotype = "greenup")
pw_gamf <- get_post_weights(m_gam_fl_hyp) %>%
  mutate(subpop = "Both",
         phenotype = "flowering")
```

```{r}
pw_3f <- pw_gf %>%
  full_join(pw_mf) %>%
  full_join(pw_gamf) %>%
  arrange(Chr, Pos) %>%
  filter(log10BayesFactor > 2) %>%
  select(Chr:Pos, log10BayesFactor, subpop, phenotype, everything())

qtlplot

pw_3f %>%
  filter(Chr %in% qtlplot$CHR[1] & between(Pos, qtlplot$POS_lo[1], qtlplot$POS_hi[1])) %>%
  group_by(subpop) %>% tally()
```

### 2021-01-26 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Finish mash Results section writeup

@@
 ```{r functions to plot svd of data driven covar matrices}
+m <- m_midw_fl_hyp
 get_svd_df <- function(m, i = NULL, eigen_v = 1){
@@
          Garden = str_remove(Garden, "^_"),
-         PVE_EV = PVE_EV)
+         PVE_EV = PVE_EV,
+         Garden = case_when(grepl("^0_M$", Garden) ~ "MI",
+                            TRUE ~ Garden))
   ggb1$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
@@
   if(eigen_v == 1){
-    ylab <- paste0("First eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
+    ylab <- paste0("First eigenvector\n(PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
   } else if(eigen_v == 2){
-    ylab <- paste0("Second eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
+    ylab <- paste0("Second eigenvector\n(PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
   } else if(eigen_v == 3){
-    ylab <- paste0("Third eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
+    ylab <- paste0("Third eigenvector\n(PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
   } else if(eigen_v == 4){
-    ylab <- paste0("Fourth eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
+    ylab <- paste0("Fourth eigenvector\n(PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
   } else if(eigen_v == 5){
-    ylab <- paste0("Fifth eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
+    ylab <- paste0("Fifth eigenvector\n(PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
   }
@@
 ```{r}
+corrmatrix <-  gxe_mf$S_AP
+pvdiv_plot_cor <- function(corrmatrix, upper = NULL, legend = FALSE){
+  U1 <- corrmatrix
+
+  U1 <- as_tibble(U1, rownames = "rowU", .name_repair = "unique") %>%
+          pivot_longer(cols = -.data$rowU, names_to = "colU",
+                       values_to = "covar") %>%
+          filter(!is.na(.data$covar)) %>%
+    mutate(rowU = str_remove(rowU, 
+                             "Stand_BhatGulf_and_Midwest_")) %>%
+    mutate(rowU = str_remove(rowU, "Stand_BhatGulf_")) %>%
+    mutate(rowU = str_remove(rowU, "Stand_BhatMidwest_")) %>%
+    mutate(rowU = str_remove(rowU, "GR50_")) %>%
+    mutate(rowU = str_remove(rowU, "FL50_")) %>%
+    mutate(colU = str_remove(colU, 
+                             "Stand_BhatGulf_and_Midwest_")) %>%
+    mutate(colU = str_remove(colU, "Stand_BhatGulf_")) %>%
+    mutate(colU = str_remove(colU, "Stand_BhatMidwest_")) %>%
+    mutate(colU = str_remove(colU, "GR50_")) %>%
+    mutate(colU = str_remove(colU, "FL50_")) %>%
+    mutate(rowU = case_when(grepl("^M$", rowU) ~ "MI",
+                            TRUE ~ rowU),
+           colU = case_when(grepl("^M$", colU) ~ "MI",
+                            TRUE ~ colU),)
+  U1$colU <- factor(U1$colU, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
+  U1$rowU <- factor(U1$rowU, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
+  U1 <- arrange(U1, rowU, colU) %>%
+    pivot_wider(names_from = colU, values_from = covar) 
+  for(i in 1:nrow(U1)){
+    for(j in 2:ncol(U1)){
+      if(i <= j-1){
+        U1[i, j] <- NA
+      }
+    }
+  }
+  U1 <- U1 %>%
+    pivot_longer(cols = -.data$rowU, names_to = "colU",
+                       values_to = "covar")
+  U1$colU <- factor(U1$colU, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD")) 
+  if(is.null(upper)){
+  upper <- max(U1$covar, na.rm = TRUE)+5
+  }
+  U1_covar <- U1 %>%
+    filter(!is.na(.data$covar)) %>%
+    ggplot(aes(x = .data$rowU, y = .data$colU)) +
+    switchgrassGWAS::theme_oeco +
+    geom_tile(aes(fill = .data$covar)) +
+    scale_fill_viridis_c(option = "B", limits = c(0, upper)) +
+    #geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
+    theme(legend.position = "none",
+          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
+    xlab("") + ylab("") 
+  if(legend == TRUE){
+    U1_covar <- U1_covar +
+      theme(legend.position = c(0.2, 0.9)) +
+      geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey")
+  }
+  return(U1_covar)
+}
+
+```
+
+```{r}
 f3a <- mash_plot_svd(m_gulf_gr_hyp, i = "ED_tPCA") 
@@
 
-f3c <- mash_plot_pairwise_sharing(corrmatrix = gxe_gg$S_AP, reorder = FALSE)
-f3d <- mash_plot_pairwise_sharing(corrmatrix = gxe_gg$S_DS, reorder = FALSE)
-f3g <- mash_plot_pairwise_sharing(corrmatrix = gxe_mg$S_AP, reorder = FALSE)
-f3h <- mash_plot_pairwise_sharing(corrmatrix = gxe_mg$S_DS, reorder = FALSE)
-f3k <- mash_plot_pairwise_sharing(corrmatrix = gxe_gamg$S_AP, reorder = FALSE)
-f3l <- mash_plot_pairwise_sharing(corrmatrix = gxe_gamg$S_DS, reorder = FALSE)
-f3o <- mash_plot_pairwise_sharing(corrmatrix = gxe_gf$S_AP, reorder = FALSE)
-f3p <- mash_plot_pairwise_sharing(corrmatrix = gxe_gf$S_DS, reorder = FALSE)
-f3s <- mash_plot_pairwise_sharing(corrmatrix = gxe_mf$S_AP, reorder = FALSE)
-f3t <- mash_plot_pairwise_sharing(corrmatrix = gxe_mf$S_DS, reorder = FALSE)
-f3w <- mash_plot_pairwise_sharing(corrmatrix = gxe_gamf$S_AP, reorder = FALSE)
-f3x <- mash_plot_pairwise_sharing(corrmatrix = gxe_gamf$S_DS, reorder = FALSE)
+f3c <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_AP, upper = 5500, legend = TRUE)
+f3c_leg <- get_legend(f3c)
+f3c <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_AP, upper = 5500, legend = FALSE)
+f3d <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_DS, upper = 5500, legend = FALSE)
+
+f3g <- pvdiv_plot_cor(corrmatrix = gxe_mg$S_AP, upper = 7777, legend = TRUE)
+f3g_leg <- get_legend(f3g)
+f3g <- pvdiv_plot_cor(corrmatrix = gxe_mg$S_AP, upper = 7777, legend = FALSE)
+f3h <- pvdiv_plot_cor(corrmatrix = gxe_mg$S_DS, upper = 7777, legend = FALSE)
+
+f3k <- pvdiv_plot_cor(corrmatrix = gxe_gamg$S_AP, upper = 7580, legend = TRUE)
+f3k_leg <- get_legend(f3k)
+f3k <- pvdiv_plot_cor(corrmatrix = gxe_gamg$S_AP, upper = 7580, legend = FALSE)
+f3l <- pvdiv_plot_cor(corrmatrix = gxe_gamg$S_DS, upper = 7580,  legend = FALSE)
+
+f3o <- pvdiv_plot_cor(corrmatrix = gxe_gf$S_AP, upper = 1850, legend = TRUE)
+f3o_leg <- get_legend(f3o)
+f3o <- pvdiv_plot_cor(corrmatrix = gxe_gf$S_AP, upper = 1850, legend = FALSE)
+f3p <- pvdiv_plot_cor(corrmatrix = gxe_gf$S_DS, upper = 1850, legend = FALSE)
+
+f3s <- pvdiv_plot_cor(corrmatrix = gxe_mf$S_AP, upper = 3636, legend = TRUE)
+f3s_leg <- get_legend(f3s)
+f3s <- pvdiv_plot_cor(corrmatrix = gxe_mf$S_AP, upper = 3636, legend = FALSE)
+f3t <- pvdiv_plot_cor(corrmatrix = gxe_mf$S_DS, upper = 3636, legend = FALSE)
+
+f3w <- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_AP, upper = 13438, legend = TRUE)
+f3w_leg <- get_legend(f3w)
+f3w <- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_AP, upper = 13438, legend = FALSE)
+f3x <- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_DS, upper = 13438, legend = FALSE)
 ```
@@
 ggdraw() +
-  draw_plot(f3a, x = 0, y = 0.825, width = 0.33, height = 0.175) +
-  draw_plot(f3b, x = 0, y = 0.65, width = 0.33, height = 0.175) +
-  draw_plot(f3d$gg_corr, x = 0.165, y = 0.5, width = 0.165, height = 0.15, scale = 2) +
-  draw_plot(f3c$gg_corr, x = 0, y = 0.5, width = 0.165, height = 0.15, scale = 2) +
-  draw_plot(f3e, x = 0.33, y = 0.825, width = 0.33, height = 0.175) +
-  draw_plot(f3f, x = 0.33, y = 0.65, width = 0.33, height = 0.175) +
-  draw_plot(f3h$gg_corr, x = 0.165+.33, y = 0.5, width = 0.165, height = 0.15, scale = 2) +
-  draw_plot(f3g$gg_corr, x = 0+.33, y = 0.5, width = 0.165, height = 0.15, scale = 2)  +
-  draw_plot(f3i, x = 0.66, y = 0.825, width = 0.33, height = 0.175) +
-  draw_plot(f3j, x = 0.66, y = 0.65, width = 0.33, height = 0.175) +
-  draw_plot(f3l$gg_corr, x = 0.165+.66, y = 0.5, width = 0.165, height = 0.15, scale = 2) +
-  draw_plot(f3k$gg_corr, x = 0+.66, y = 0.5, width = 0.165, height = 0.15, scale = 2) +  ## greenup above here, flowering below
-  draw_plot(f3m, x = 0, y = 0.825-0.5, width = 0.33, height = 0.175) +
-  draw_plot(f3n, x = 0, y = 0.65-0.5, width = 0.33, height = 0.175) +
-  draw_plot(f3p$gg_corr, x = 0.165, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2) +
-  draw_plot(f3o$gg_corr, x = 0, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2) +
-  draw_plot(f3q, x = 0.33, y = 0.825-0.5, width = 0.33, height = 0.175) +
-  draw_plot(f3r, x = 0.33, y = 0.65-0.5, width = 0.33, height = 0.175) +
-  draw_plot(f3t$gg_corr, x = 0.165+.33, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2) +
-  draw_plot(f3s$gg_corr, x = 0+.33, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2)  +
-  draw_plot(f3u, x = 0.66, y = 0.825-0.5, width = 0.33, height = 0.175) +
-  draw_plot(f3v, x = 0.66, y = 0.65-0.5, width = 0.33, height = 0.175) +
-  draw_plot(f3x$gg_corr, x = 0.165+.66, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2) +
-  draw_plot(f3w$gg_corr, x = 0+.66, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2)
-
-save_plot(filename = "Figure_3_Data_driven_Covariance_matrices.png", plot = last_plot(), base_height = 8, base_width = 9)
+  draw_plot(f3a, x = 0, y = 0.84, width = 0.33, height = 0.16) +
+  draw_plot(f3b, x = 0, y = 0.68, width = 0.33, height = 0.16) +
+  draw_plot(f3d, x = 0.165, y = 0.5, width = 0.165, height = 0.18) +
+  draw_plot(f3c, x = 0, y = 0.5, width = 0.165, height = 0.18) +
+  draw_plot(f3c_leg, x = 0.095, y = 0.595, width = 0.1, height = 0.1) +
+  draw_plot(f3e, x = 0.33, y = 0.84, width = 0.33, height = 0.16) +
+  draw_plot(f3f, x = 0.33, y = 0.68, width = 0.33, height = 0.16) +
+  draw_plot(f3h, x = 0.165+.33, y = 0.5, width = 0.165, height = 0.18) +
+  draw_plot(f3g, x = 0+.33, y = 0.5, width = 0.165, height = 0.18)  +
+  draw_plot(f3g_leg, x = 0.095+.33, y = 0.595, width = 0.1, height = 0.1) +
+  draw_plot(f3i, x = 0.66, y = 0.84, width = 0.33, height = 0.16) +
+  draw_plot(f3j, x = 0.66, y = 0.68, width = 0.33, height = 0.16) +
+  draw_plot(f3l, x = 0.165+.66, y = 0.5, width = 0.165, height = 0.18) +
+  draw_plot(f3k, x = 0+.66, y = 0.5, width = 0.165, height = 0.18) +  
+  draw_plot(f3k_leg, x = 0.095+.66, y = 0.595, width = 0.1, height = 0.1) +## greenup above here, flowering below
+  draw_plot(f3m, x = 0, y = 0.84-0.5, width = 0.33, height = 0.16) +
+  draw_plot(f3n, x = 0, y = 0.68-0.5, width = 0.33, height = 0.16) +
+  draw_plot(f3p, x = 0.165, y = 0.5-0.5, width = 0.165, height = 0.18) +
+  draw_plot(f3o, x = 0, y = 0.5-0.5, width = 0.165, height = 0.18) +  
+  draw_plot(f3o_leg, x = 0.095, y = 0.595-0.5, width = 0.1, height = 0.1) +
+  draw_plot(f3q, x = 0.33, y = 0.84-0.5, width = 0.33, height = 0.16) +
+  draw_plot(f3r, x = 0.33, y = 0.68-0.5, width = 0.33, height = 0.16) +
+  draw_plot(f3t, x = 0.165+.33, y = 0.5-0.5, width = 0.165, height = 0.18) +
+  draw_plot(f3s, x = 0+.33, y = 0.5-0.5, width = 0.165, height = 0.18)  +
+  draw_plot(f3s_leg, x = 0.095+.33, y = 0.595-0.5, width = 0.1, height = 0.1) +
+  draw_plot(f3u, x = 0.66, y = 0.84-0.5, width = 0.33, height = 0.16) +
+  draw_plot(f3v, x = 0.66, y = 0.68-0.5, width = 0.33, height = 0.16) +
+  draw_plot(f3x, x = 0.165+.66, y = 0.5-0.5, width = 0.165, height = 0.18) +
+  draw_plot(f3w, x = 0+.66, y = 0.5-0.5, width = 0.165, height = 0.18) +
+  draw_plot(f3w_leg, x = 0.095+.66, y = 0.595-0.5, width = 0.1, height = 0.1)
+
+save_plot(filename = "../manuscript/plots/Figure_3_Data_driven_Covariance_matrices.png", plot = last_plot(), base_height = 10, base_width = 11)
 ```

### 2021-01-26 phenology-fourway/5-QTLmappingForAlice/4-QTLinterval&Plotting.R

@@ -1,0 +1,76 @@
+setwd("~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/")
+library(qtl)
+library(qtl2)
+library(plyr)
+library(tidyverse)
+library(qtl2convert)
+
+############# WORK ON GENO
+rm(list=ls())
+load('cross_reduced_marker_FL50_phenotypes_Alice_7_SITES.RData')
+load('scan1Results_FL50_phenotypes_Alice_7_SITES.RData')
+
+map  =  lapply(pull.map(cross), function(x) x[1,])
+###get the threshold value for each model and output the QTL intervals
+threshold_full = summary(s1permfull,alpha = 0.05)
+threshold_reduced = summary(s1permreduced,alpha = 0.05)
+
+###  g x e
+gxe = s1full - s1reduced
+gxe_perm = threshold_full - threshold_reduced ###or gxe_perm = summary(gxe, alpha=0.05)
+
+###get the peak for full model and gxe model
+peaks_full = find_peaks(s1full, map, threshold = threshold_full, drop = 1.5, peakdrop = 5)
+peaks_gxe = find_peaks(gxe, map, threshold = gxe_perm, drop = 1.5)
+
+flank_marker_full = peaks_full %>% mutate(flank_lo = find_marker(map, chr, ci_lo), flank_hi=find_marker(map, chr, ci_hi))
+write.csv(flank_marker_full,'QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv') ###change the name you want
+
+flank_marker_gxe = peaks_gxe %>% mutate(flank_lo = find_marker(map, chr, ci_lo), flank_hi=find_marker(map, chr, ci_hi))
+write.csv(flank_marker_gxe,'QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_gxe_model.csv')
+
+###########
+png('../2-figures/FL50_7_SITES_LOD_FL50_phenotypes_Alice_7_SITES.png', width = 800)
+plot(s1full, map, lodcolumn=2, col="slateblue", ylim=c(0,56))
+plot(s1full,map, lodcolumn=3, col="violetred", add=T)
+plot(s1full,map, lodcolumn=4, col="darkgreen", add=T)
+plot(s1full,map, lodcolumn=5, col="yellow", add=T)
[…]

### 2021-01-26 phenology-fourway/5-QTLmappingForAlice/5-QTLEffectAtEachSite.R

@@ -1,0 +1,312 @@
+###QTL Effect plot
+rm(list=ls())
+library(qtl)
+library(qtl2)
+library(qtl2convert)
+library(tidyverse)
+setwd("~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/")
+###load the original cross and reduce the markers from 4700 to 1185
+load('sexAveraged.cross.rda', verbose = T)
+map  =  lapply(pull.map(cross), function(x) x[1,])
+# reduced_markers  =  unlist(
+#   lapply(
+#     reduce_markers(
+#       lapply(pull.map(cross), function(x)
+#         x[1,]), min_distance = 1), names))
+#
+# mars2drop  =  markernames(cross)[!markernames(cross) %in% reduced_markers]
+# cross  =  drop.markers(cross, mars2drop)
+cross = calc.genoprob(cross, step=1, error.prob = 1e-3,map.function = "kosambi",stepwidth = "max")
+######### Converting probalibity from qtl1 to qtl2
+probs  =  probs_qtl_to_qtl2(cross)$probs
+##calculate the kinship matrix for single site
+kinship_loco_Single = calc_kinship(probs,use_allele_probs = F,type = "loco")
+
+###extract the genotype ID from the original cross file, so we can align the phenology file with genotype ID
+ID = data.frame(ID=cross$pheno$id)
+ID = ID %>% mutate_at(vars('ID'), list(as.character)) %>% mutate_at(vars('ID'), list(as.numeric))
+
+######get the phenotype at each site
+phenology = read.csv('FL50_phenotypes_Alice.csv')
+colnames(phenology)[1:2] = c('ID','Covariate')
+str(phenology) ###make sure your phenology traits are numeric, if not, change them into numeric
+phenos = phenology %>% mutate_at(vars('ID'), list(as.character)) %>%
+  mutate_at(vars('ID'), list(as.numeric))%>% group_by(Covariate,ID ) %>%
+  summarize_all(mean, na.rm=T) %>% ungroup()%>% as.data.frame()
+
+###QTL effect for each trait---FL50
+EffectMatrix = vector('list', length=0)
+t=0
[…]

### 2021-01-27 R/Analysis_v0.6_QTL_GxE_and_mash_comparison.Rmd
Add Results QTL interval and mash 1% tail comparisons

@@
 knitr::opts_chunk$set(echo = TRUE)
+library(tidyverse)
+library(cowplot)
 ```
@@
                               lodcolumn == "crain_gr2fl" ~ "Rainfall, GR to FL")) %>%
-  filter(phe %in% c("Daylength", "GDD, GR to FL")) %>%
+  #filter(phe %in% c("Daylength", "GDD, GR to FL")) %>%
   mutate(POS_lo = Pos_Mb_lo * 1000000,
@@
 
+# Get any mash overlaps for QTL intervals
 ```{r}
@@
   arrange(Chr, Pos) %>%
-  filter(log10BayesFactor > 2) %>%
+  filter(log10BayesFactor > 3) %>%
+  select(Chr:Pos, log10BayesFactor, subpop, phenotype, everything())
+
+pw_3f_all <- pw_gf %>%
+  full_join(pw_mf) %>%
+  full_join(pw_gamf) %>%
+  arrange(Chr, Pos) %>%
   select(Chr:Pos, log10BayesFactor, subpop, phenotype, everything())
@@
 qtlplot
+summary(pw_3f$log10BayesFactor)
+quantile(pw_gf$log10BayesFactor, 0.96196895)
+quantile(pw_gf$log10BayesFactor, 0.96196895)
+quantile(pw_gamf$log10BayesFactor, 0.8006157)
+1-0.96196895
+1-0.8006157
+2/464  # GT 0.5%? No 0.0043%
+22/418 # GT 3.8031? Yes 5.26%
+42/255 # GT 19.93843%? No 16.47%
+```
 
-pw_3f %>%
-  filter(Chr %in% qtlplot$CHR[1] & between(Pos, qtlplot$POS_lo[1], qtlplot$POS_hi[1])) %>%
+```{r}
+i=16  # from 1 to 16
+## In block & > 2
+QTLandBF <- pw_3f %>%
+  filter(Chr %in% qtlplot$CHR[i] & between(Pos, qtlplot$POS_lo[i], qtlplot$POS_hi[i])) %>%
   group_by(subpop) %>% tally()
+## BF>2
+inBF <- pw_3f %>%
+  group_by(subpop) %>% tally()
+## BF not > 2 (all minus BF > 2, all is below)
+allwrtBF <- pw_3f_all %>%
+  group_by(subpop) %>% tally()
+## All in block
+inQTL <- pw_3f_all %>%
+  filter(Chr %in% qtlplot$CHR[i] & between(Pos, qtlplot$POS_lo[i], qtlplot$POS_hi[i])) %>%
+  group_by(subpop) %>% tally()
+
+QTLandBF
+inQTL
+
+phyper(QTLandBF$n[1], inBF$n[1], allwrtBF$n[1]-inBF$n[1], inQTL$n[1], 
+       lower.tail = FALSE) # Both
+phyper(QTLandBF$n[2], inBF$n[2], allwrtBF$n[2]-inBF$n[2], inQTL$n[2], 
+       lower.tail = FALSE) # Gulf
+phyper(QTLandBF$n[3], inBF$n[3], allwrtBF$n[3]-inBF$n[3], inQTL$n[3], 
+       lower.tail = FALSE) # Midwest
 ```
+BF > 2 results: 5 QTL intervals
+1 ~ G
+2: M
+6: G, M
+10: G&M, M
+11: G, M
+12: G&M, G
+14 ~ G&M
+16 ~ G&M
+
+BF > 3 results: 7 QTL intervals. Different intervals, too. So I think having the same BF cutoff for these different mash results will give misleading conclusions. Do a % cutoff instead.
+4 ~ G&M
+6: G
+8: G, M
+11: M
+12: G&M, G
+13: M
+14: G&M
+15: M
+
+1137 SNPs in 2K region. 66 sig > 2 in mash. 2/464 Midwest, 22/418 Gulf, 42/255 Both.
+
+Summary statistics for log 10 Bayes Factor for pw_3f df
+   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+  2.000   2.234   2.568   2.848   3.140  10.385 
+90% 3.936367
+95% 4.622242
+99% 6.257548
+
+Gulf
+90% 1.410253
+95% 1.828653
+99% 2.911984
+Top 3.8% of Gulf is above 2.
+
+Midwest 
+90% 0.72685
+95% 0.949753
+99% 1.616983
+Top 0.5% of Midwest is above 2.
+
+G&M
+90% 2.595963
+95% 3.196745
+99% 4.65753
+top 20% of G&M is above 2.
+
+## Hypergeometric test for significance of overlaps
+phyper(10, 50, 200 - 50, 25, lower.tail = FALSE)
+phyper(A&B, inB, notinB, inA, lower.tail = FALSE)
+```{r}
+phyper(22, 336, 50490-336, 418, lower.tail = FALSE)
+phyper(AandB$n[2], inB$n[2], allwrtB$n[2]-inB$n[2], inA$n[2], 
+       lower.tail = FALSE)
+```
+
+In block & > 2, >2, not > 2, in block
+22, 336, 50490-336, 418
+
+BF > 2
+Both	7125			
+Gulf	1921			
+Midwest	336
+
+All SNPs
+Both	35735			
+Gulf	50490			
+Midwest	63367
+
+Is there an enrichment of SNPs in the top 1% (5%? 10%?) of the distribution for each mash result in the QTL regions?
+1% is probably best out of these three arbitrary cutoffs.
+```{r}
+m <- m_gulf_fl_hyp
+qtl_df <- qtlplot
+
+get_QTL_enrichment <- function(m, qtl_df, quantile = 0.95){
+  pw_df_all <- get_post_weights(m)
+
+  BFthresh <- quantile(pw_df_all$log10BayesFactor, quantile)
+  pw_df <- pw_df_all %>%
+  filter(log10BayesFactor > BFthresh) %>%
+  select(Chr:Pos, log10BayesFactor, everything())
+
+  for(i in 1:nrow(qtl_df)){  # from 1 to 16
+  ## In block & > 2
+  QTLandBF <- pw_df %>%
+    filter(Chr %in% qtl_df$CHR[i] & between(Pos, qtl_df$POS_lo[i],
+                                             qtl_df$POS_hi[i])) %>% 
+    tally()
+  ## BF>2
+  inBF <- pw_df %>%
+    tally()
+  ## BF not > 2 (all minus BF > 2, all is below)
+  allwrtBF <- pw_df_all %>%
+    tally()
+  ## All in block
+  inQTL <- pw_df_all %>%
+    filter(Chr %in% qtl_df$CHR[i] & between(Pos, qtl_df$POS_lo[i],
+                                             qtl_df$POS_hi[i])) %>%
+    tally()
+  notinBF <- (allwrtBF$n[1]-inBF$n[1])
+  pval <- phyper(QTLandBF$n[1], inBF$n[1], allwrtBF$n[1]-inBF$n[1], 
+                 inQTL$n[1], lower.tail = FALSE)
+
+  if(i == 1){
+  outputdf <- tibble(Chr = qtl_df$CHR[i], Pos_lo = qtl_df$POS_lo[i], Pos_hi = qtl_df$POS_hi[i], inQTLandBF = QTLandBF$n[1], inBF = inBF$n[1], notinBF = notinBF, inQTL = inQTL$n[1], pvalue = pval)
+  } else {
+    outputdf <- outputdf %>%
+      add_row(Chr = qtl_df$CHR[i], Pos_lo = qtl_df$POS_lo[i], Pos_hi = qtl_df$POS_hi[i],inQTLandBF = QTLandBF$n[1], inBF = inBF$n[1], notinBF = notinBF, inQTL = inQTL$n[1], pvalue = pval)
+    }
+  }
+  return(outputdf)
+}
+
+enrich_gf <- get_QTL_enrichment(m_gulf_fl_hyp, qtl_df = qtlplot, 0.99) %>%
+  mutate(subpop = "Gulf", phenotype = "flowering")
+enrich_mf <- get_QTL_enrichment(m_midw_fl_hyp, qtl_df = qtlplot, 0.99) %>%
+  mutate(subpop = "Midwest", phenotype = "flowering")
+enrich_gamf <- get_QTL_enrichment(m_gam_fl_hyp, qtl_df = qtlplot, 0.99) %>%
+  mutate(subpop = "Both", phenotype = "flowering")
+
+enrich_df <- enrich_gf %>%
+  full_join(enrich_mf) %>%
+  full_join(enrich_gamf) %>%
+  arrange(Chr, Pos_lo)
+enrich_df <- enrich_df %>%
+  left_join(qtlplot, by = c("Chr" = "CHR", "Pos_lo" = "POS_lo", "Pos_hi" = "POS_hi")) %>%
+  select(Chr, Pos_lo, Pos_hi, inQTLandBF:subpop, lod, lodcolumn, pos) %>%
+  unique() %>%
+  arrange(Chr, Pos_lo, lodcolumn)
+
+enrich_df %>%
+  filter(pvalue < 0.05)
+
+t.test(enrich_df[enrich_df$pvalue < 0.05,]$lod, enrich_df[enrich_df$pvalue >= 0.05,]$lod) 
+qtlplot %>% group_by(CHR, POS_lo) %>% tally() # 21-24 unique QTL by Pos lo/hi. 23 by QTL maximum. 12 of 24 have enriched mash hits.
+
+```
+
+16 QTL. In all cases there was at least one SNP in the QTL interval that was in the 1% tail of SNPs in one of the three flowering mash runs.
+We also looked for enrichment of SNPs in the 1% tail of significance in the mash distribution within each QTL interval.
+(in random genomic intervals of the same size do we see a similar level of enrichment?)
+Half of these QTL had a significant enrichment of significant results for mash compared to expectations given the number of SNPs in the QTL interval.
+
+2K: 2 QTL (lod 14 & 17), 1 enriched (in Midwest) (lod 17: CGDD_12C)
+2N: 4 QTL (3 non-overlapping) (lod 19, 39, 17, 15), 2 enriched (lod 39 dylnfl50 in Midwest, in Both & 15 CGDD_12C in all three)
+4K: 2 QTL, 2 enriched (lod 40 dyln in Midwest & 33 dyln in Gulf)
+5N: 4 QTL (3 non-overlapping) (lod 55, 25, 17, 33), 3 enriched (2 non-overlapping) (lod 25 dyln in Midwest, 17 CGDD in Gulf, in Midwest, 33 dyln in Gulf, in Both)
+7K: 2 QTL, 0 en (lod 14, 13)
+8N: 1 QTL, 0 en (lod 15)
+9N: 1 QTL, 0 en (lod 18)
+
+in Midwest & for CGDD: 3
+in Midwest & for dyln:
+in Gulf & for CGDD: 
+```{r}
+qtlplot %>%
+  group_by(lodcolumn) %>%
+  tally()
+qtlplot %>%
+  group_by(CHR, Pos_Mb_lo, Pos_Mb_hi) %>%
+  tally()
+qtlplot %>%
+  filter(lodcolumn %in% c("dyln_change_sec", "FL50"))
+```
+
+# Make random QTL dataframe
+
+This random df just needs a CHR, POS_lo, and POS_hi column to work in the get_QTL_enrichment function I made. 
+```{r}
+sig_random <- c()
+
+for(j in 1:100){
+  qtl_size <- qtlplot %>%
+    select(CHR, POS_lo, POS_hi) %>%
+    unique() %>%   # don't doublecount completely overlapping QTL
+    #(does doublecount partially overlapping QTL, however.)
+    group_by(CHR, POS_lo) %>%
+    mutate(Size = POS_hi - POS_lo) %>%
+    select(CHR, POS_lo, Size)
+
+  chr_size_mash <- pw_3f_all %>%
+    group_by(Chr) %>%
+    summarise(Pos_lo = min(Pos),
+              Pos_hi = max(Pos))
+
+  rchr <- chr_size_mash$Chr[sample(1:18, nrow(qtl_size), replace = TRUE)]
+  rposlo <- c()
+  for(i in 1:length(rchr)){
+    rposlo[i] <- sample(chr_size_mash$Pos_lo[which(chr_size_mash$Chr == rchr[i])]:chr_size_mash$Pos_hi[which(chr_size_mash$Chr == rchr[i])], 1)
+  }
+
+  random_qtl <- tibble(CHR = rchr, POS_lo = rposlo, Size = qtl_size$Size) %>%
+    mutate(POS_hi = POS_lo + Size) %>%
+    arrange(CHR, POS_lo)
+
+  random_gf <- get_QTL_enrichment(m_gulf_fl_hyp, qtl_df = random_qtl, 0.99) %>%
+    mutate(subpop = "Gulf", phenotype = "flowering")
+  random_mf <- get_QTL_enrichment(m_midw_fl_hyp, qtl_df = random_qtl, 0.99) %>%
+    mutate(subpop = "Midwest", phenotype = "flowering")
+  random_gamf <- get_QTL_enrichment(m_gam_fl_hyp, qtl_df = random_qtl, 0.99) %>%
+    mutate(subpop = "Both", phenotype = "flowering")
+
+  random_df <- random_gf %>%
+    full_join(random_mf) %>%
+    full_join(random_gamf) %>%
+    arrange(Chr, Pos_lo)
+
+  sig_random[j] <- nrow(filter(random_df, pvalue < 0.05))
+}
+saveRDS(sig_random, file = "../data/Random_QTL_df_significant_enrichments_of_mash_1per_tail_SNPs_100runs.rds")
+tibble(`Significant Enrichments` = sig_random) %>%
+  ggplot(aes(x = `Significant Enrichments`)) +
+  switchgrassGWAS::theme_oeco +
+  geom_histogram(binwidth = 1) + 
+  xlim(c(1, 24)) + ylab("") +
+  geom_vline(xintercept = 23, linetype = 2, color = "red")
+save_plot(filename = "../manuscript/plots/Figure_4_QTL_overlaps_random_QTL_enrichments_with_mash.png", plot = last_plot(), base_height = 2)
+```

### 2021-01-27 R/Analysis_v0.6_QTL_interval_plotting.Rmd
Add Results QTL interval and mash 1% tail comparisons

---
title: "QTL_interval_plotting"
author: "Alice MacQueen"
date: "1/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(qtl)
library(qtl2)
library(qtl2convert)
library(tidyverse)
```

#

```{r prep QTL_Eff df}
###QTL Effect plot
###load the original cross and reduce the markers from 4700 to 1185
load('../phenology-fourway/5-QTLmappingForAlice/sexAveraged.cross.rda', verbose = T)
map  =  lapply(pull.map(cross), function(x) x[1,])
# reduced_markers  =  unlist(
#   lapply(
#     reduce_markers(
#       lapply(pull.map(cross), function(x)
#         x[1,]), min_distance = 1), names))
#
# mars2drop  =  markernames(cross)[!markernames(cross) %in% reduced_markers]
# cross  =  drop.markers(cross, mars2drop)
cross = calc.genoprob(cross, step=1, error.prob = 1e-3,map.function = "kosambi",stepwidth = "max")
######### Converting probalibity from qtl1 to qtl2
probs  =  probs_qtl_to_qtl2(cross)$probs
##calculate the kinship matrix for single site
kinship_loco_Single = calc_kinship(probs,use_allele_probs = F,type = "loco")

###extract the genotype ID from the original cross file, so we can align the phenology file with genotype ID
ID <- data.frame(ID=cross$pheno$id)
ID <- ID %>% mutate_at(vars('ID'), list(as.character)) %>% mutate_at(vars('ID'), list(as.numeric))

######get the phenotype at each site
phenology = read.csv('../phenology-fourway/5-QTLmappingForAlice/FL50_phenotypes_Alice.csv', header = TRUE)
colnames(phenology)[1:2] = c('ID','Covariate')
str(phenology) ###make sure your phenology traits are numeric, if not, change them into numeric
phenos <- phenology %>% mutate_at(vars('ID'), list(as.character)) %>%
  mutate_at(vars('ID'), list(as.numeric))%>% group_by(Covariate,ID ) %>%
  summarize_all(mean, na.rm=T) %>% ungroup()%>% as.data.frame()
```

```{r setup QTL effects for all traits}
###QTL effect for each trait
EffectMatrix = vector('list', length=0)
t=0
for (envi in unique(phenos$Covariate)){
    t=t+1
    pheno = phenos %>% filter(Covariate==envi) %>% dplyr::select(-Covariate) %>%
      select(ID, FL50) %>% right_join(ID) %>%  column_to_rownames('ID') %>% as.matrix()

    ##ran effect scan for each chromosome
    eff_out = vector('list',length = 0)
    se_out = vector('list',length = 0)
    add_contrasts=cbind(mu=c(1,1,1,1), add1=c(-1, 1, -1, 1),add2=c(-1, -1, 1, 1), dom=c(-1, 1, 1, -1))

    z = 0
    for (chr in chrnames(cross)){
      print(c(chr, envi,t))
      z= z+1
      eff = scan1coef(probs[,chr],pheno,kinship = kinship_loco_Single[chr], se=T, contrasts = add_contrasts)
      se = attr(eff, 'SE')
      eff_out[[z]] =  eff %>% as.data.frame() %>% rownames_to_column('MARKER')
      se_out[[z]] =  se %>% as.data.frame() %>% rownames_to_column('MARKER')
    }
    eff_out = bind_rows(eff_out)
    se_out = bind_rows(se_out)

    se_out2 = se_out %>% dplyr::select(-mu) %>% gather('CROSS','SE',-MARKER)
    eff_out2 = eff_out %>% dplyr::select(-mu) %>% gather('CROSS','EFF',-MARKER) %>% mutate(SE=se_out2$SE, Covariate=envi)

    EffectMatrix[[t]] = eff_out2
  }

EffectMatrix2 = bind_rows(EffectMatrix)

###unlist nmap and merge with Effect file
nmap = data.frame(pos = unlist(map))
nnmap = nmap %>% rownames_to_column('MARKER') %>% rowwise() %>%
  mutate(chr = str_sub(MARKER, 1,2), MARKER=str_sub(MARKER, 4,nchar(MARKER)))%>%
  ungroup() %>% mutate(pos=round(pos,4))

flank_marker_full = read.csv('../phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv')
flank_marker_full = flank_marker_full %>% mutate_at(vars(chr),list(as.character))

for (i in 1:nrow(flank_marker_full)){
  tmp = nnmap %>% filter(chr==flank_marker_full$chr[i] & pos==round(flank_marker_full$pos[i],4))
  flank_marker_full$MARKER[i] = tmp$MARKER
}

QTL_full_model = flank_marker_full %>% dplyr::rename(trait=lodcolumn)

```

```{r ggplots of QTL effects}
QTLdf <- QTL_full_model %>% left_join(EffectMatrix2)%>%
  mutate(CROSS = str_replace(CROSS,'add2','C x D')) %>%
  mutate(CROSS = str_replace(CROSS,'add1','A x B')) %>%
  mutate(Garden = case_when(Covariate == "BRKG" ~ "SD",
                            Covariate == "CLMB" ~ "MO",
                            Covariate == "KBSM" ~ "MI",
                            Covariate == "KING" ~ "TX1",
                            Covariate == "LINC" ~ "NE",
                            Covariate == "PKLE" ~ "TX2",
                            Covariate == "STIL" ~ "OK",
                            TRUE ~ "Other"),
         Cross = case_when(CROSS == "A x B" ~ "AP13 x DAC",
                           CROSS == "C x D" ~ "VS16 x WBC",
                           CROSS == "dom" ~ "Dominance",
                           TRUE ~ "Other")) %>%
  separate(MARKER, into = c("Chr", "Pos"), sep = "_", convert = TRUE, 
           remove = FALSE) %>%
  mutate(Pos_plot = paste0(Chr, "\n", round(Pos, 1), " Mb"))

QTLdf$Covariate = factor(QTLdf$Covariate, levels=c('KING','PKLE','STIL','CLMB',
                                                       'LINC','KBSM','BRKG'))
QTLdf$Garden = factor(QTLdf$Garden, levels = c("TX1", "TX2", "OK", "MO", 
                                                   "NE", "MI", "SD"))

QTLdf %>% 
  filter(Cross != "Dominance") %>% # trait == 'FL50' &
  ggplot(aes(x = Garden, y = EFF, group = Cross)) +
  geom_point(size = 2) + facet_grid(Cross + trait ~ Pos_plot) +
  geom_hline(yintercept = 0, linetype=2) + 
  geom_errorbar(aes(ymin=EFF-SE, ymax=EFF+SE), size=0.6, width=0.5) +
  xlab("") +
  ylab('QTL Effect') + 
  coord_flip() + 
  theme_bw()+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        text=element_text(size=12, face="bold"),
        axis.title=element_text(size=12,face="bold"),
        axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1))
# shared QTL for more than one trait differ in LOD score, but not meaningfully in effect patterns across gardens. Useful to know.

QTLdf %>% 
  filter(trait == 'CGDD_12C') 
QTLdf %>% 
  filter(trait == 'dyln_fl50')
QTLdf %>% 
  filter(trait == 'dyln_change_sec') 
```

### 2021-01-29 R/Analysis_v0.6_QTL_interval_plotting.Rmd
Prepare DOE abstract for 2021 meeting

@@
 
+QTLplot <- QTLdf %>%
+  arrange(desc(lod)) %>%
+  group_by(Chr, Pos, Garden, Cross) %>%
+  slice(1) %>% 
+  filter(Cross != "Dominance") # trait == 'FL50' &
+
+QTLplot$Pos_plot <- factor(QTLplot$Pos_plot, levels = c("Chr02K\n8.6 Mb",
+                                                        "Chr02K\n64.4 Mb",
+                                                        "Chr02N\n19.5 Mb",
+                                                        "Chr02N\n58 Mb",
+                                                        "Chr02N\n60.3 Mb",
+                                                        "Chr02N\n60.9 Mb",
+                                                        "Chr02N\n69.4 Mb",
+                                                        "Chr03K\n18.6 Mb",
+                                                        "Chr04K\n4.7 Mb",
+                                                        "Chr04K\n11.8 Mb",
+                                                        "Chr05N\n2.3 Mb",
+                                                        "Chr05N\n3.2 Mb",
+                                                        "Chr05N\n3.4 Mb",
+                                                        "Chr05N\n45.9 Mb",
+                                                        "Chr05N\n59.9 Mb",
+                                                        "Chr05N\n64.9 Mb",
+                                                        "Chr05N\n65.3 Mb",
+                                                        "Chr07K\n37 Mb",
+                                                        "Chr07N\n45.2 Mb",
+                                                        "Chr08N\n6.8 Mb",
+                                                        "Chr08N\n16.5 Mb",
+                                                        "Chr09K\n16.3 Mb",
+                                                        "Chr09N\n15.7 Mb"))
+#levels(as.factor(QTLplot$Pos_plot))
+
+QTLplot %>%
+  ggplot(aes(x = Garden, y = EFF, group = Cross)) +
+  geom_col(size = 1, aes(fill = trait)) + 
+  facet_grid(Cross ~ Pos_plot) +
+  geom_hline(yintercept = 0, linetype=2) + 
+  geom_errorbar(aes(ymin=EFF-SE, ymax=EFF+SE, color = trait), size=0.6, width=0.5) +
+  xlab("") + scale_y_continuous(limits = c(-4.25, 6.5), breaks = c(-3, 0, 3)) +
+  ylab('QTL Effect') + 
+  coord_flip() + 
+  switchgrassGWAS::theme_oeco + 
+  scale_color_manual(values = c("#CC0066", "#FFCC00", "#006600", "#6600FF")) +
+  scale_fill_manual(values = c("#CC0066", "#FFCC00", "#006600", "#6600FF")) +
+  theme(legend.position = "top",
+        panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        text = element_text(size = 10, face="bold"),
+        strip.text = element_text(size = 10, face="bold"),
+        axis.title = element_text(size = 10,face="bold"),
+        axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1),
+        )# +
+    #scale_y_continuous(expand = c(0.2, 0.2)) 
+save_plot(filename = "../manuscript/plots/Figure_4_QTL_effects_by_trait.png", base_height = 4.5, base_width = 13.5, plot = last_plot())
+
 QTLdf %>%
@@
   filter(trait == 'dyln_change_sec')
+
 ```
+
+I could combine this with some plots pulled from the mash object using `mash_plot_effects` to do a side-by-side comparison for the paper.
+That would go well maybe with the Figure 4 I started today, although of course I can't think of a good way to validate this statistically any more than I did in the other script.
+```{r}
+library(switchgrassGWAS)
+mash_plot_effects()
+```

### 2021-01-30 R/Analysis_v0.6_QTL_GxE_and_mash_comparison.Rmd
Finish manuscript v0.6 Intro-Discussion and four Figures

@@
       rownames_to_column(var = "value") %>%
-      mutate(value = as.integer(.data$value)) %>%
+      dplyr::mutate(value = as.integer(.data$value)) %>%
       as_tibble() %>%
@@
     #left_join(m_meta) %>%
-    separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
+    tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
              extra = "merge", convert = TRUE) %>%
-    left_join(log10bf_df, by = "Marker") %>%
-    arrange(.data$Chr, .data$Pos) 
+    dplyr::left_join(log10bf_df, by = "Marker") %>%
+    dplyr::arrange(.data$Chr, .data$Pos) 
   return(ggman_df)
@@
 
+qtlplot <- qtlplot %>%
+  arrange(desc(lod)) %>%
+  mutate(pos = as.integer(pos)) %>%
+  group_by(CHR, pos) %>%
+  slice(1) %>%
+  arrange(CHR, pos)
 enrich_gf <- get_QTL_enrichment(m_gulf_fl_hyp, qtl_df = qtlplot, 0.99) %>%
@@
 sig_random <- c()
+chr_size_mash <- pw_3f_all %>%
+  ungroup() %>%
+  group_by(Chr) %>%
+  dplyr::summarise(Pos_lo = min(Pos),
+                   Pos_hi = max(Pos))
 
-for(j in 1:100){
+for(j in 1:1000){
   qtl_size <- qtlplot %>%
@@
     group_by(CHR, POS_lo) %>%
-    mutate(Size = POS_hi - POS_lo) %>%
-    select(CHR, POS_lo, Size)
-
-  chr_size_mash <- pw_3f_all %>%
-    group_by(Chr) %>%
-    summarise(Pos_lo = min(Pos),
-              Pos_hi = max(Pos))
+    dplyr::mutate(Size = POS_hi - POS_lo) %>%
+    dplyr::select(CHR, POS_lo, Size)
 
@@
 }
-saveRDS(sig_random, file = "../data/Random_QTL_df_significant_enrichments_of_mash_1per_tail_SNPs_100runs.rds")
-tibble(`Significant Enrichments` = sig_random) %>%
-  ggplot(aes(x = `Significant Enrichments`)) +
+saveRDS(sig_random, file = "../data/Random_QTL_df_significant_enrichments_of_mash_1per_tail_SNPs_1000runs.rds")
+tibble(`mash 1% tail significant enrichments` = sig_random) %>%
+  ggplot(aes(x = `mash 1% tail significant enrichments` )) +
   switchgrassGWAS::theme_oeco +
@@
   xlim(c(1, 24)) + ylab("") +
-  geom_vline(xintercept = 23, linetype = 2, color = "red")
-save_plot(filename = "../manuscript/plots/Figure_4_QTL_overlaps_random_QTL_enrichments_with_mash.png", plot = last_plot(), base_height = 2)
+  geom_vline(xintercept = 16, linetype = 2, color = "red")
+save_plot(filename = "../manuscript/plots/Figure_4_QTL_overlaps_random_QTL_enrichments_with_mash_1000runs.png", plot = last_plot(), base_height = 1.8, base_asp = 1.8)
+
+tibble(sig_r = sig_random) %>%
+  arrange(desc(sig_r))
+3/1000
 ```

### 2021-01-30 R/Analysis_v0.6_QTL_interval_plotting.Rmd
Finish manuscript v0.6 Intro-Discussion and four Figures

@@
 library(switchgrassGWAS)
-mash_plot_effects()
+m_gulf_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_FL50_Gulf_Uhyp.rds")
+m_midw_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_33000_SNPs_FL50_Midwest_Uhyp.rds")
+m_gam_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_FL50_Gulf_and_Midwest_Uhyp.rds")
+
+get_post_weights <- function(m){
+  log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
+      as.data.frame() %>%
+      rownames_to_column(var = "value") %>%
+      dplyr::mutate(value = as.integer(.data$value)) %>%
+      as_tibble() %>%
+      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
+      dplyr::rename(log10BayesFactor = .data$V1) %>%
+      dplyr::select(-.data$value)
+
+  pw_df <- as_tibble(m$posterior_weights)
+  pw_df <- switchgrassGWAS:::get_marker_df(m = m) %>%
+    bind_cols(pw_df)
+
+  ggman_df <- pw_df %>%
+    #left_join(m_meta) %>%
+    tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
+             extra = "merge", convert = TRUE) %>%
+    dplyr::left_join(log10bf_df, by = "Marker") %>%
+    dplyr::arrange(.data$Chr, .data$Pos) 
+  return(ggman_df)
+}
+
+pw_gf <- get_post_weights(m = m_gulf_fl_hyp) %>%
+  mutate(subpop = "Gulf",
+         phenotype = "flowering")
+pw_mf <- get_post_weights(m_midw_fl_hyp) %>%
+  mutate(subpop = "Midwest",
+         phenotype = "flowering")
+pw_gamf <- get_post_weights(m_gam_fl_hyp) %>%
+  mutate(subpop = "Both",
+         phenotype = "flowering")
 ```
+
+```{r}
+qtldf <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv")
+
+qtldf <- qtldf %>%
+  separate(flank_lo, into = c("CHR", "Pos_Mb_lo"), sep = "_", convert = TRUE) %>%
+  separate(flank_hi, into = c("flank", "Pos_Mb_hi"), sep = "_", convert = TRUE) %>%
+  mutate(Pos_Mb_lo = round(Pos_Mb_lo, digits = 2),
+         Pos_Mb_hi = round(Pos_Mb_hi, digits = 2)) %>%
+  select(CHR, Pos_Mb_lo, Pos_Mb_hi, lodcolumn, lod, everything()) %>%
+  arrange(CHR, Pos_Mb_lo)
+
+qtlplot <- qtldf %>%
+  mutate(phe = case_when(lodcolumn == "FL50" ~ "50% flowering",
+                              lodcolumn == "dyln_fl50" ~ "Daylength",
+                              lodcolumn == "dyln_change_sec" ~ "daylength change (s)",
+                              lodcolumn == "CGDD_12C" ~ "GDD, GR to FL",
+                              lodcolumn == "crain_gr2fl" ~ "Rainfall, GR to FL")) %>%
+  #filter(phe %in% c("Daylength", "GDD, GR to FL")) %>%
+  mutate(POS_lo = Pos_Mb_lo * 1000000,
+         POS_hi = Pos_Mb_hi * 1000000)
+
+qtlplot$phe <- factor(qtlplot$phe, 
+                      levels = c("50% flowering", "Rainfall, GR to FL",
+                                 "GDD, GR to FL", "Daylength",
+                                          "daylength change (s)",
+                                          "Rainfall, 1 day",
+                                                  #"Rainfall, 2d sum",
+                                                  "Rainfall, 3d sum",
+                                                  "Rainfall, 5d sum"#,
+                                                  #"Rainfall, 7d sum"
+                                                  ))
+```
+
+```{r}
+pw_3f <- pw_gf %>%
+  full_join(pw_mf) %>%
+  full_join(pw_gamf) %>%
+  arrange(Chr, Pos) %>%
+  select(Chr:Pos, log10BayesFactor, subpop, value, phenotype, everything()) %>%
+  filter(log10BayesFactor > 2)
+pw_list <- list(Gulf = pw_gf, Midwest = pw_mf, Both = pw_gamf)
+m_list <- list(Gulf = m_gulf_fl_hyp, Midwest = m_midw_fl_hyp, Both = m_gam_fl_hyp)
+names(m_list)
+qtlplot <- qtlplot %>%
+  arrange(desc(lod)) %>%
+  mutate(pos = as.integer(pos)) %>%
+  group_by(CHR, pos) %>%
+  slice(1) %>%
+  arrange(CHR, pos)
+mash_plot_manhattan_by_condition(m_gulf_fl_hyp)
+mash_plot_effects(m_midw_fl_hyp, n = 5)
+```
+
+Make 4 plots for Figure 4: 
+i = 4, j = 1 Both BF 7
+10, 3 BF 4.37 Gulf
+
+15, 5: 5.1 BF Midwest
+16, 1 7.6 BF Both
+17, 1 7.6 BF Both (same SNP as 16)
+```{r}
+i <- c(4, 10, 15, 16)
+j <- c(1, 3, 5, 1)
+f4df <- tibble(i = i, j = j)
+
+ggdf_list <- list()
+
+k=3
+for(k in 1:nrow(f4df)){
+  pw_int <- pw_3f %>%
+    filter(.data$Chr %in% qtlplot$CHR[f4df$i[k]] & between(.data$Pos, 
+                                                   qtlplot$POS_lo[f4df$i[k]], 
+                                                   qtlplot$POS_hi[f4df$i[k]])) %>%
+    arrange(desc(log10BayesFactor))
+
+  plot_df <- 
+    mash_plot_effects(m = m_list[[which(names(m_list) == pw_int$subpop[f4df$j[k]])]], 
+                    i = pw_int$value[f4df$j[k]])
+  ggb1 <- plot_df$effect_df %>%
+    mutate(Pos_plot = plot_df$marker,
+           subpop = pw_int$subpop[f4df$j[k]]) %>%
+    mutate(Conditions = str_remove(Conditions, "Gulf_and_Midwest_")) %>%
+  mutate(Conditions = str_remove(Conditions, "Gulf_")) %>%
+  mutate(Conditions = str_remove(Conditions, "Midwest_")) %>%
+  separate(Conditions, into = c("Phenotype", "Garden"), sep = -3) %>%
+  mutate(Phenotype = str_remove(Phenotype, "_$"),
+         Garden = str_remove(Garden, "^_")) %>%
+  mutate(Garden = str_replace(Garden, "0_M", "MI"),
+         Phenotype = str_replace(Phenotype, "FL5$", "FL50"))
+  ggb1$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
+  ggdf_list[[k]] <- ggb1
+}
+
+ggdf <- ggdf_list[[1]] %>%
+  full_join(ggdf_list[[2]]) %>%
+  full_join(ggdf_list[[3]]) %>%
+  full_join(ggdf_list[[4]])
+
+ggdf %>%
+  ggplot() + 
+  geom_point(mapping = aes(x = Garden, y = .data$mn)) +
+  switchgrassGWAS::theme_oeco +
+  geom_errorbar(mapping = aes(ymin = .data$mn - .data$se,
+                              ymax = .data$mn + .data$se,
+                              x = Garden), width = 0.3) +
+  geom_hline(yintercept = 0, lty = 2) +
+  facet_wrap(~Pos_plot+subpop, nrow = 1) +
+  labs(x = "", y = "Effect Size (% total variation)") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+  coord_flip()
+
+save_plot(filename = "../manuscript/plots/Figure_4_mash_SNP_effects.png", plot = last_plot(), base_height = 2.4, base_width = 5.5)
+```

### 2021-02-01 R/Analysis_v0.7_QTL_interval_lod_plot.Rmd
Add ggplot version of QTL plot by lod score and update Figure 4

---
title: "4-QTLinterval and plotting"
author: "Alice MacQueen"
date: "2/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(qtl)
library(qtl2)
library(plyr)
library(tidyverse)
library(qtl2convert)
library(cowplot)
```

setwd("~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/")

```{r}
workingdir <- "~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/"
load(file.path(workingdir, 'cross_reduced_marker_FL50_phenotypes_Alice_7_SITES.RData'))
load(file.path(workingdir, 'scan1Results_FL50_phenotypes_Alice_7_SITES.RData.RData'))

map  =  lapply(pull.map(cross), function(x) x[1,])
###get the threshold value for each model and output the QTL intervals
threshold_full = summary(s1permfull,alpha = 0.05)
threshold_reduced = summary(s1permreduced,alpha = 0.05)
```

```{r}
###  g x e
gxe = s1full - s1reduced
gxe_perm = threshold_full - threshold_reduced ###or gxe_perm = summary(gxe, alpha=0.05)

###get the peak for full model and gxe model
peaks_full = find_peaks(s1full, map, threshold = threshold_full, drop = 1.5, peakdrop = 5)
peaks_gxe = find_peaks(gxe, map, threshold = gxe_perm, drop = 1.5)

flank_marker_full = peaks_full %>% mutate(flank_lo = find_marker(map, chr, ci_lo), flank_hi=find_marker(map, chr, ci_hi))
# write.csv(flank_marker_full,'QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv') ###change the name you want

flank_marker_gxe = peaks_gxe %>% mutate(flank_lo = find_marker(map, chr, ci_lo), flank_hi=find_marker(map, chr, ci_hi))
# write.csv(flank_marker_gxe,'QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_gxe_model.csv')

###########
```

```{r}
threshold_df <- threshold_full %>% as_tibble(rownames = "Threshold") %>%
  pivot_longer(GR50:dyln_change_sec, names_to = "Phenotype", values_to = "LOD")

s1full %>%
  as_tibble(rownames = "Marker") %>%
  pivot_longer(GR50:dyln_change_sec, names_to = "Phenotype", values_to = "LOD") %>%
  separate(Marker, into = c("Chr", "Pos"), sep = "_", remove = FALSE, convert = TRUE) %>%
  filter(!grepl("loc", Chr)) %>%
  ggplot(aes(x = Pos, y = LOD)) + 
  geom_line(aes(color = Phenotype)) +
  geom_hline(data = threshold_df, aes(yintercept = LOD, color = Phenotype, linetype = Phenotype)) +
  facet_wrap(~Chr, nrow = 1, scales = "free_x") +
  switchgrassGWAS::theme_oeco + 
  scale_color_manual(values = c("#CC0066", "#FFCC00", "#006600", "#6600FF", "grey")) + 
  scale_linetype_manual(values = c(4,5,4,4,4)) +
  theme(legend.position = "top",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        text = element_text(size = 10, face="bold"),
        strip.text = element_text(size = 10, face="bold"),
        axis.title = element_text(size = 10,face="bold"),
        axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1),
        ) + 
  xlab("Position (Mb)") + ylab("LOD score") +
  scale_x_continuous(breaks = c(0, 20, 40, 60, 80))

save_plot(filename = "../manuscript/plots/Figure_4_QTL_by_lod_score.svg", base_height = 3, base_width = 13.5, plot = last_plot())
```

Li's additional code
```{r}
###########
png('../2-figures/FL50_7_SITES_LOD_FL50_phenotypes_Alice_7_SITES.png', width = 800)
plot(s1full, map, lodcolumn=2, col="slateblue", ylim=c(0,56))
plot(s1full,map, lodcolumn=3, col="violetred", add=T)
plot(s1full,map, lodcolumn=4, col="darkgreen", add=T)
plot(s1full,map, lodcolumn=5, col="yellow", add=T)
abline(h=threshold_full[2],lty=2,lwd=3,col=c("slateblue"))
abline(h=threshold_full[3],lty=2,lwd=3,col=c("violetred"))
abline(h=threshold_full[4],lty=2,lwd=3,col=c("darkgreen"))
abline(h=threshold_full[5],lty=2,lwd=3,col=c("yellow"))

legend("topleft", lwd=2, col=c("slateblue", "violetred","darkgreen","yellow"), colnames(s1full)[c(2,3,4,5)], bg="gray90")
dev.off()

flank_marker_full = flank_marker_full %>% mutate(GXE= ifelse(pos %in% flank_marker_gxe$pos, 'Y','N'))%>%
  mutate(ChrNo = str_extract(chr, '[1-9]'))%>% mutate_at(vars(ChrNo),as.numeric) %>%
  mutate(Dist = case_when(
    str_detect(chr,'K') ~ (ChrNo*2-1),
    str_detect(chr,'N') ~ ChrNo*2
  ))
# ###6. plot QTL on genetic map
# load('sexAveraged.cross.rda')
png('../2-figures/QTLs_FL50_phenotypes_Alice_7_SITES.png',width = 900)
cols  =  c("violetred","yellow","darkgreen","slateblue") ###add your colors depending on how many traits you have
library(qtlTools)
with(flank_marker_full, segmentsOnMap(cross, phe = lodcolumn, chr = chr,
                                      l = ci_lo, h = ci_hi,
                                      peaklod = lod, peakcM = pos,
                                      showPeaks = TRUE,
                                      chrBuffer = c(0.15,0.1),
                                      tick.width=0.05, lwd="byLod",
                                      col = cols,
                                      leg.inset=.57, legendCex=1,
                                      legendPosition="topleft")
)
#mtext('QTL_7_SITES',side = 3)

for (i in 1:nrow(flank_marker_full)){
    if (flank_marker_full$GXE[i]=='Y')
    text(x=(flank_marker_full$Dist[i]+0.4), y=(flank_marker_full$ci_lo[i]),'*',col = 'red',cex = 1.5)
  }

dev.off()

```

### 2021-02-04 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Finish manuscript v0.6 Intro-Discussion and four Figures
Completed manuscript v0.8 and edited figures for consistency

@@
 Midwest greenup: 28.6% flowering 45.1%
+```{r}
+5.127744e-02/4.037520e-01
+```
 
@@
 cov_flex$phe <- factor(cov_flex$phe, levels = c("tave_18d", "tave_10d", "dyln_fl50", "cgdd_12c_gr2fl"))
+cov_flex$subpop <- factor(cov_flex$subpop, levels = c("Gulf", "Midwest", "Both)"))
 
@@
          )
+
+df_2gr_b$subpop <- factor(df_2gr_b$subpop, levels = c("Gulf", "Midwest", "Both"))
+df_2fl_b$subpop <- factor(df_2fl_b$subpop, levels = c("Gulf", "Midwest", "Both"))
+df_2gr_a$subpop <- factor(df_2gr_a$subpop, levels = c("Gulf", "Midwest", "Both"))
+df_2fl_a$subpop <- factor(df_2fl_a$subpop, levels = c("Gulf", "Midwest", "Both"))
 
@@
 
-save_plot(filename = "Figure_2_Covariance_matrices_and_posterior_weights.png", plot = last_plot(), base_height = 8, base_width = 9)
+save_plot(filename = "../manuscript/plots/Figure_2_Covariance_matrices_and_posterior_weights.svg", plot = last_plot(), base_height = 8, base_width = 9)
 ```

### 2021-02-09 R/Analysis_v0.6_QTL_interval_plotting.Rmd
Make figures for Tom's DOE talk 2021

@@
 ```
+
+```{r}
+	Chr06N_41610882
+plot_df <- mash_plot_effects(m_midw_fl_hyp, i = 41793)
+
+ggb1 <- plot_df$effect_df %>%
+    mutate(Pos_plot = plot_df$marker,
+           subpop = "Midwest") %>%
+    mutate(Conditions = str_remove(Conditions, "Gulf_and_Midwest_")) %>%
+  mutate(Conditions = str_remove(Conditions, "Gulf_")) %>%
+  mutate(Conditions = str_remove(Conditions, "Midwest_")) %>%
+  separate(Conditions, into = c("Phenotype", "Garden"), sep = -3) %>%
+  mutate(Phenotype = str_remove(Phenotype, "_$"),
+         Garden = str_remove(Garden, "^_")) %>%
+  mutate(Garden = str_replace(Garden, "0_M", "MI"),
+         Phenotype = str_replace(Phenotype, "FL5$", "FL50"))
+ggb1$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
+
+  ggb1 %>%
+  ggplot() + 
+  geom_point(mapping = aes(x = Garden, y = .data$mn)) +
+  switchgrassGWAS::theme_oeco +
+  geom_errorbar(mapping = aes(ymin = .data$mn - .data$se,
+                              ymax = .data$mn + .data$se,
+                              x = Garden), width = 0.3) +
+  geom_hline(yintercept = 0, lty = 2) +
+  facet_wrap(~Pos_plot+subpop, nrow = 1) +
+  labs(x = "", y = "Effect Size (% total variation)") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+  coord_flip()
+
+  ave_plot(filename = "../manuscript/plots/Figure_DOE_mash_SNP_effects.png", plot = last_plot(), base_height = 2.4, base_width = 2)
+```

### 2021-02-09 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Make figures for Tom's DOE talk 2021

@@
 ```{r mashhattan plots loop}
-pw_df <- pw_gg
+pw_df <- pw_mf
 
@@
     arrange(!! sym(var)) %>%
+    filter(!! sym(var) > 0.01) %>%
     ggplot(aes(x = .data$Pos, y = .data$log10BayesFactor)) +
@@
 }
-gglist
+ggdoe1 <- gglist[[6]] + ggtitle("SNPs in the Midwest subpopulation that covary strongly with cumulative GDD")
+save_plot(filename = "../manuscript/plots/Mashhattan_covar_Midwest_flowering_cgdd.png", base_width = 9, plot = ggdoe1)
+ggdoe2 <- gglist[[9]] + ggtitle("SNPs in the Midwest subpopulation with equal effects at all gardens")
+save_plot(filename = "../manuscript/plots/Mashhattan_covar_Midwest_flowering_equal_effects.png", base_width = 9, plot = ggdoe2)
 ```
@@
 
+Chr06N_41610882
+mash_plot_effects(m_midw_fl_hyp, i = 41793)
 ```

### 2021-02-10 R/Analysis_v0.6_QTL_interval_plotting.Rmd
Make DOE Poster and Figures

@@
   facet_wrap(~Pos_plot+subpop, nrow = 1) +
-  labs(x = "", y = "Effect Size (% total variation)") +
+  labs(x = "", y = "Effect Size (%)") +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
@@
 
-  ave_plot(filename = "../manuscript/plots/Figure_DOE_mash_SNP_effects.png", plot = last_plot(), base_height = 2.4, base_width = 2)
+  save_plot(filename = "../manuscript/plots/Figure_DOE_mash_SNP_effects.png", plot = last_plot(), base_height = 2.4, base_width = 2)
 ```

### 2021-02-10 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Make DOE Poster and Figures

@@
 
-Chr06N_41610882
-mash_plot_effects(m_midw_fl_hyp, i = 41793)
 ```
@@
 save_plot(filename = "../manuscript/plots/Figure_3_Data_driven_Covariance_matrices.png", plot = last_plot(), base_height = 10, base_width = 11)
+
+save_plot(filename = "../manuscript/plots/Partial/AP_Flowering_Both.png", plot = f3x, base_height = 2, base_width = 2)
+save_plot(filename = "../manuscript/plots/Partial/DS_Flowering_Both.png", plot = f3w, base_height = 2, base_width = 2)
+
+ggdraw() +
+  draw_plot(f3w, x = 0, y = 0, width = 0.5, height = 6/7) +
+  draw_plot(f3x, x = 0.5, y = 0, width = 0.5, height = 6/7) +
+  draw_plot(f3w_leg, x = 0.335, y = 0.725, width = 0.1, height = 0.1) +
+  draw_label(label = "Sign changes (AP)", x = 0.2, y = 0.92, size = 10) +
+  draw_label(label = "Magnitude changes (DS)", x = 0.73, y = 0.92, size = 10)
+save_plot(filename = "../manuscript/plots/Partial/AP_DS_Flowering_Both.png", plot = last_plot(), base_height = 2, base_width = 3.5)
 ```

### 2021-05-04 R/Analysis_v0.6_QTL_interval_plotting.Rmd
Modifications to Figures from lab meeting and creation of supplemental figures and tables

@@
          Pos_Mb_hi = round(Pos_Mb_hi, digits = 2)) %>%
-  select(CHR, Pos_Mb_lo, Pos_Mb_hi, lodcolumn, lod, everything()) %>%
+  dplyr::select(CHR, Pos_Mb_lo, Pos_Mb_hi, lodcolumn, lod, everything()) %>%
   arrange(CHR, Pos_Mb_lo)
@@
                                                   ))
+```
+# Functionally validated flowering time genes in QTL intervals (Table S2)
+```{r}
+qtl_anno_df <- qtlplot %>%
+  dplyr::rename(start = POS_lo, end = POS_hi)
+txdb <- loadDb("~/Github/pvdiv-genome/Pvirgatum_516_v5.1.gene.txdb.sqlite")
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+ftanno <- read_csv(file.path(workingdir, "data/Pvirgatum_v5.1.annotation_FTGeneHomologs.csv"))
+ftkey <- read_delim(file.path(workingdir, "data", "OsGene_keyword_table.txt"),
+                    delim = "\t")
+ftanno %>%
+  filter(!is.na(OsGene)) %>%
+  left_join(ftkey, by = c("OsGene" = "Symbol")) %>%
+  filter(Keyword %in% c("flowering time", "heading date")) %>% unique()
+
+osgene_keyword_df <- ftkey %>%
+  dplyr::select(-RAPdb, -MSU) %>%
+  group_by(Symbol, Keyword) %>%
+  dplyr::slice(1) %>%
+  pivot_wider(names_from = Keyword, values_from = Title)
+
+osgene_key <- osgene_keyword_df %>%
+  dplyr::select(Symbol, `flowering time`, `heading date`, anther, flower,
+                `floral meristem`, `floral meristem determinacy`, flowering,
+                `flower development`, `carpel differentiation`, stamen, 
+                `stamen number`) %>%
+  pivot_longer(cols = -Symbol, names_to = "Keyword", values_to = "Title") %>%
+  filter(!is.na(Title)) %>%
+  pivot_wider(names_from = Keyword, values_from = Title)
+
+os_annos <- ftanno %>%
+  filter(OsGene %in% osgene_key$Symbol | !is.na(AtGene)) %>%
+  mutate(Pos_Mb = round(start/10000)/100,
+         #Pos_Mb_e = floor(start/10000)/100,
+         POS_bin = round(Pos_Mb*100/2)) %>%# ,
+         #POS_bin_hi = round(Pos_Mb_e*100/2))
+  dplyr::select(OsGene, AtGene, locusName, everything()) %>%
+  dplyr::rename(`Gene ID` = locusName)
+
+qtl_annos <- switchgrassGWAS::pvdiv_table_topsnps(df = qtl_anno_df, type = "table", txdb = txdb) %>%
+    inner_join(os_annos, by = c("Gene ID")) %>%
+    left_join(osgene_key, by = c("OsGene" = "Symbol")) %>%
+    dplyr::select(where(not_all_na)) %>%
+    group_by(CHR, pos, `Gene ID`) %>%
+    dplyr::slice(1)
+
+names(qtl_annos)
+all_na_ft <- rowSums(is.na(qtl_annos[,58:65]))
+add_column(qtl_annos, FT_Osgene_NA = all_na_ft, .before = TRUE) %>%
+  group_by(CHR, pos, FT_Osgene_NA < 8) %>%
+  tally()
+flowering_qtl_annos <- add_column(qtl_annos, FT_Osgene_NA = all_na_ft, .before = TRUE) %>%
+  filter(FT_Osgene_NA < 8) %>%
+  dplyr::select(CHR, pos, phe, lod, Pos_Mb_lo, Pos_Mb_hi, ci_lo, ci_hi, `Gene ID`,  gene_start:gene_width, `Arabidopsis thaliana homolog`:`Rice gene annotation`, OsGene, anther:`floral meristem determinacy`)
+
+flowering_qtl_annos %>%
+  write_csv("~/Github/pvdiv-phenology-gxe/manuscript/tables/Final/Table_S2_QTL_intervals_with_functionally_validated_flowering_genes.csv")
 ```

### 2021-05-04 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Modifications to Figures from lab meeting and creation of supplemental figures and tables

@@
 ```{r}
-annos_gg %>%
+annos_s1_1 <- annos_gg %>%
   filter(cgdd_12c_18d_Gulf_and_Midwest.10 > 0.1 |
            tave_18d_Gulf_and_Midwest.6 > 0.1) %>%
-  dplyr::select(where(not_all_na))
-
-annos_gamg %>%
+  dplyr::select(where(not_all_na)) %>%
+  mutate(Phenotype = "Greenup",
+         Subpop = "Gulf")
+
+annos_s1_2 <- annos_gamg %>%
   filter(cgdd_12c_18d_Gulf_and_Midwest.10 > 0.1 |
            cgdd_12c_18d_Gulf_and_Midwest.9 > 0.1) %>%
-  dplyr::select(where(not_all_na))
-
-annos_mf %>%
+  dplyr::select(where(not_all_na)) %>%
+  mutate(Phenotype = "Greenup",
+         Subpop = "Both")
+
+annos_s1_3 <- annos_mf %>%
   # dplyr::select(starts_with("cgdd_12c_gr2fl_Gulf_and"))
   filter(cgdd_12c_gr2fl_Gulf_and_Midwest.8 > 0.1) %>%
-  dplyr::select(where(not_all_na))
-annos_gamf %>%
+  dplyr::select(where(not_all_na)) %>%
+  mutate(TXID = as.character(TXID)) %>%
+  mutate(Phenotype = "Flowering",
+         Subpop = "Midwest")
+annos_s1_4 <- annos_gamf %>%
   # dplyr::select(starts_with("dyln_change_sec"))
   filter(dyln_change_sec_Midwest.7 > 0.1) %>%
-  dplyr::select(where(not_all_na))
-
+  dplyr::select(where(not_all_na)) %>%
+  mutate(Phenotype = "Flowering",
+         Subpop = "Both")
+
+annos_s1_1 %>%
+  full_join(annos_s1_2) %>%
+  full_join(annos_s1_3) %>%
+  full_join(annos_s1_4) %>%
+  ungroup() %>%
+  dplyr::select(Marker, Phenotype, Subpop, Chr, Pos, log10BayesFactor, `distance from gene`, `Gene ID`, gene_start, gene_end, gene_width, Pfam:KOG, KO, `Arabidopsis thaliana homolog`:`Rice gene annotation`,
+         OsGene, AtGene, Best.hit.arabi.name:rice.defline, leaf:mitochondria, 	
+`transcription factor`:`blast disease`, `early leaf senescence`:dormancy, flower:`florigen transport`, everything()) %>%
+  dplyr::select(-(CHR:value), -(POS:POS_bin)) %>%
+  write_csv("~/Github/pvdiv-phenology-gxe/manuscript/tables/Table_S1_sig_hyp_matrix_SNPs.csv")
 ```

### 2021-05-04 R/Analysis_v0.7_QTL_interval_lod_plot.Rmd
Modifications to Figures from lab meeting and creation of supplemental figures and tables

@@
 
+# Rug plot of mash hits
+```{r}
+m_gulf_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_GR50_Gulf_Uhyp.rds")
+m_gulf_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_FL50_Gulf_Uhyp.rds")
+
+m_midw_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_24000_SNPs_GR50_Midwest_Uhyp.rds")
+m_midw_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_33000_SNPs_FL50_Midwest_Uhyp.rds")
+
+m_gam_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_GR50_Gulf_and_Midwest_Uhyp.rds")
+m_gam_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_FL50_Gulf_and_Midwest_Uhyp.rds")
+```
+
+```{r}
+get_post_weights <- function(m){
+  log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
+      as.data.frame() %>%
+      rownames_to_column(var = "value") %>%
+      dplyr::mutate(value = as.integer(.data$value)) %>%
+      as_tibble() %>%
+      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
+      dplyr::rename(log10BayesFactor = .data$V1) %>%
+      dplyr::select(-.data$value)
+
+  pw_df <- as_tibble(m$posterior_weights)
+  pw_df <- switchgrassGWAS:::get_marker_df(m = m) %>%
+    bind_cols(pw_df)
+
+  ggman_df <- pw_df %>%
+    #left_join(m_meta) %>%
+    tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
+             extra = "merge", convert = TRUE) %>%
+    left_join(log10bf_df, by = "Marker") %>%
+    dplyr::arrange(.data$Chr, .data$Pos) 
+  return(ggman_df)
+}
+
+pw_gg <- get_post_weights(m_gulf_gr_hyp)
+pw_gf <- get_post_weights(m_gulf_fl_hyp) %>%
+  mutate(Subpop = "Gulf")
+
+pw_mg <- get_post_weights(m_midw_gr_hyp)
+pw_mf <- get_post_weights(m_midw_fl_hyp) %>%
+  mutate(Subpop = "Midwest")
+
+pw_gamg <- get_post_weights(m_gam_gr_hyp)
+pw_gamf <- get_post_weights(m_gam_fl_hyp) %>%
+  mutate(Subpop = "Both")
+```
+
+```{r}
+test_rug1 <- pw_gf %>%
+  filter(log10BayesFactor > 2) %>% 
+  mutate(Pos = Pos/1000000, 
+         LOD = -0.5)
+test_rug2 <- pw_gamf %>%
+  filter(log10BayesFactor > 2) %>% 
+  mutate(Pos = Pos/1000000, 
+         LOD = -0.5)
+test_rug3 <- pw_mf %>%
+  filter(log10BayesFactor > 2) %>% 
+  mutate(Pos = Pos/1000000, 
+         LOD = -0.5)
+test_rug <- pw_gf %>%
+  full_join(pw_gamf) %>%
+  full_join(pw_mf) %>%
+  filter(log10BayesFactor > 2) %>% 
+  mutate(Pos = Pos/1000000, 
+         LOD = -0.5)
+```
+
+```{r}
+s1full %>%
+  as_tibble(rownames = "Marker") %>%
+  pivot_longer(GR50:dyln_change_sec, names_to = "Phenotype", values_to = "LOD") %>%
+  separate(Marker, into = c("Chr", "Pos"), sep = "_", remove = FALSE, convert = TRUE) %>%
+  filter(!grepl("loc", Chr)) %>%
+  ggplot(aes(x = Pos, y = LOD)) + 
+  geom_line(aes(color = Phenotype)) +
+  geom_rug(data = test_rug2, sides = "t", alpha = 0.1) + 
+  geom_rug(data = test_rug1, sides = "b", alpha = 0.2) + 
+  geom_rug(data = test_rug3, outside = TRUE, sides = "b", alpha = 0.2) + 
+  geom_hline(data = threshold_df, aes(yintercept = LOD, color = Phenotype, linetype = Phenotype)) +
+  facet_wrap(~Chr, nrow = 1, scales = "free_x") +
+  switchgrassGWAS::theme_oeco + 
+  scale_color_manual(values = c("#CC0066", "#FFCC00", "#006600", "#6600FF", "grey")) + 
+  scale_linetype_manual(values = c(4,5,4,4,4)) +
+  theme(legend.position = "top",
+        panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        text = element_text(size = 10, face="bold"),
+        strip.text = element_text(size = 10, face="bold"),
+        axis.title = element_text(size = 10,face="bold"),
+        axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1),
+        panel.spacing.x = unit(-0., 'cm')
+        ) + 
+  xlab("Position (Mb)") + ylab("LOD score") +
+  scale_x_continuous(breaks = c(0, 20, 40, 60, 80)) + 
+  coord_cartesian(clip = "off")
+save_plot(filename = "../manuscript/plots/Figure_4_QTL_by_lod_score_w_rugs.svg", base_height = 3, base_width = 13.5, plot = last_plot())
+```
+
 Li's additional code

### 2021-05-04 R/Plots_v0.3_manuscript_figures.Rmd
Modifications to Figures from lab meeting and creation of supplemental figures and tables

--- 2020-09-30 R/Plots_manuscript_v0.3.Rmd
@@
 h2fl <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "h2_df_FL50_related_functions_of_weather_v1.csv"))
-h2gr_gxe <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "gxe_and_h2_df_GR50_related_functions_of_weather_v1.csv"))
+h2gr_gxe <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "gxe_and_h2_df_GR50_related_functions_of_weather_v2.csv"))
 h2fl_gxe <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "gxe_and_h2_df_FL50_related_functions_of_weather_v1.csv"))
@@
   full_join(h2gr_gxe)
+h2gr <- h2gr_gxe
 h2fl <- h2fl %>%
@@
                               phe == "tmin_18d" ~ "Min T, 18d mean",
-                              phe == "tave_18d" ~ "Mean T, 18d mean",
+                              phe == "tave_18d" ~ "Temp Ave, 18d",
                               phe == "tmax_10d" ~ "Max T, 10d mean",
                               phe == "tmin_10d" ~ "Min T, 10d mean",
-                              phe == "tave_10d" ~ "Mean T, 10d mean",
+                              phe == "tave_10d" ~ "Temp Ave, 10d",
                               phe == "tmax_5d" ~ "Max T, 5d mean",
                               phe == "tmin_5d" ~ "Min T, 5d mean",
-                              phe == "tave_5d" ~ "Mean T, 5d mean",
+                              phe == "tave_5d" ~ "Temp Ave, 5d",
                               phe == "cgdd_12c_10d" ~ "GDD, 10d",
@@
 grh2$phe_name <- factor(grh2$phe_name, levels = c("50% greenup", "GDD, 5d",
-                                                  "Min T, 5d mean",
-                                                  #"Mean T, 5d mean",
+                                                  #"Min T, 5d mean",
+                                                  "Temp Ave, 5d",
                                                   #"Max T, 5d mean",
                                                   "GDD, 10d",
-                                                  "Min T, 10d mean",
-                                                  #"Mean T, 10d mean",
+                                                  #"Min T, 10d mean",
+                                                  "Temp Ave, 10d",
                                                   #"Max T, 10d mean",
                                                   "GDD, 18d",
-                                                  "Min T, 18d mean",
-                                                  #"Mean T, 18d mean",
+                                                  #"Min T, 18d mean",
+                                                  "Temp Ave, 18d",
                                                   #"Max T, 18d mean",
@@
 subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))
-
+```
+
+```{r}
 site_plot <- c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North")
@@
 
+Use code starting ~on line 966 to generate this.
 ```{r}
@@
 
-save_plot("../manuscript/plots/Var_Comp_greenup_site_groups.png", plot = gr_bar_gxe, base_height = 4, base_asp = 1.5)
+save_plot("../manuscript/plots/Var_Comp_greenup_site_groups.svg", plot = gr_bar_gxe, base_height = 4, base_asp = 1.5)
 ```
@@
            "crain_7d")
-phe_gr <- c("GR50", "DYLN", "cgdd_12c_jan2gr", "cgdd_8c_jan2gr", "tmax_18d", 
-            "tmin_18d", "tave_18d", "cgdd_12c_18d", "tmax_10d", "tmin_10d",
-            "tave_10d", "cgdd_12c_10d", "tmax_5d", "tmin_5d", "tave_5d",
-            "cgdd_12c_5d")
+phe_gr <- c("GR50", "DYLN", "cgdd_12c_jan2gr", "cgdd_8c_jan2gr", "tave_10d",
+            "tave_18d", "tave_5d", "cgdd_12c_10d",
+            "cgdd_12c_5d", "cgdd_12c_18d", "tmax_18d", 
+            "tmin_18d",  "tmax_10d", "tmin_10d",
+            "tmax_5d", "tmin_5d")
 phe_dfs <- list(phe_gr_df, phe_fl_df)
@@
 
-  if(i == 12 & j == 1 & k == 1){
+  if(i == 10 & j == 1 & k == 1){
     h2_table <- h2_est
@@
                                                   "_related_functions_", 
-                                                  "of_weather_v4.csv")))
+                                                  "of_weather_v5.csv")))
 }
@@
            "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")
-phe_gr <- c("GR50", "DYLN", "cgdd_12c_jan2gr", "tmin_18d", "cgdd_12c_18d",
-            "tmin_10d", "cgdd_12c_10d", "tmin_5d", "cgdd_12c_5d")
+phe_gr <- c("GR50", "DYLN", "cgdd_12c_jan2gr", "tave_18d", "cgdd_12c_18d",
+            "tave_10d", "cgdd_12c_10d", "tave_5d", "cgdd_12c_5d")
 phe_dfs <- list(phe_gr_df, phe_fl_df)
@@
 It's possible I might want to confirm some of these G candidates with the fourway - how we see similar patterns in this cross for FL50 vs dyln/GDD, and there are some similar associations.
-
+# ---- moved to new doc: Analysis_manuscript_mash_GxE.Rmd ----
 THEN, for the next section, do GWAS on the BLUPs from the eight single site models. For these traits:

### 2021-05-12 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Prepare manuscript and figures vsn for coauthors

@@
 ```
+1.208534e-01+1.400286e-01  photoperiod cues
+2.371512e-02  cgdd_12c_gr2fl_Gulf_and_Midwest
+
+5.718063e-02+1.461888e-01+1.398986e-01
+
 47.3% of effect mass for both subpop flowering was on hyp matrices, while only 8.5% of greenup was.

### 2021-06-24 R/SI_Appendix_Datasets.Rmd
Add SI Appendix Datasets, SI references, and author affiliations

--- 2021-01-30 R/Analysis_v0.6_QTL_GxE_and_mash_comparison.Rmd
@@
 ---
-title: "Analysis_v0.7_QTL_GxE"
+title: "SI_Appendix_Datasets"
 author: "Alice MacQueen"
-date: "1/26/2021"
+date: 2021-06-24
 output: html_document
@@
 library(cowplot)
+library(switchgrassGWAS)
+library(ashr)
 ```
 
-# TRIAL script
-# -------------------------
-NB: Trial script to compare the QTL to the mash results using env covar matrices.
+#
 
-## 3a. Log-likelihoods
+Make all these datasets (except Dataset 7 is already made):
+
+    Dataset 1. SNP-associated effects and standard errors for green-up estimated using the Gulf genetic subpopulation.
+    Dataset 2. SNP-associated effects and standard errors for green-up estimated using the Midwest genetic subpopulation.
+    Dataset 3. SNP-associated effects and standard errors for green-up estimated using Both genetic subpopulations.
+    Dataset 4. SNP-associated effects and standard errors for flowering estimated using the Gulf genetic subpopulation.
+    Dataset 5. SNP-associated effects and standard errors for flowering estimated using the Midwest genetic subpopulation.
+    Dataset 6. SNP-associated effects and standard errors for flowering estimated using Both genetic subpopulations.
+    Dataset 7. Genes within QTL that have functionally validated roles in flowering in rice.
+    Dataset 8.
+Overlap between quantitative trait loci for flowering measured using an outbred pseudo-F2 mapping population, and significant mash effect estimates (with a log10Bayes Factor > 2, or in the 1% tail of most significant mash results) found using three overlapping diversity panels.
+Genes within 20kb with a functionally validated role in flowering in rice are  indicated.
+
+# mash objects
 ```{r}
@@
 m_gam_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_FL50_Gulf_and_Midwest_Uhyp.rds")
+```
 
+```{r}
+m <- m_gam_gr_hyp
+dataset_creation_function <- function(m, filename, ncol = 27){
+
+  all_effects <- switchgrassGWAS:::get_marker_df(m) %>%
+    mutate(log10BF = switchgrassGWAS:::get_log10bf(m))
+
+  Means <- get_pm(m)
+  StdErr <- get_psd(m)
+  lfsr <- get_lfsr(m)
+  colnames(StdErr) <- colnames(StdErr) %>%
+    str_replace(., "Stand_Bhat", "Effect_SE_") %>%
+    str_replace(., "Gulf_and_Midwest", "Both")
+  colnames(Means) <- colnames(Means) %>%
+    str_replace(., "Stand_Bhat", "Effect_Mean_") %>%
+    str_replace(., "Gulf_and_Midwest", "Both")
+  colnames(lfsr) <- colnames(lfsr) %>%
+    str_replace(., "Stand_Bhat", "Effect_lfsr_") %>%
+    str_replace(., "Gulf_and_Midwest", "Both")
+
+  all_effects <- cbind(all_effects, Means, StdErr, lfsr)
+
+  effects_df <- all_effects %>%
+    pivot_longer(cols = 4:all_of(ncol), names_to = "Condition", values_to = "Value") %>%
+    separate(col = "Condition", into = c(NA, "Type", "Subpop", "Phe", "Site"), 
+             sep = "_") %>%
+    mutate(Phe = case_when(Phe == "GR50" ~ "green-up",
+                           Phe == "FL50" ~ "flowering")) %>%
+    arrange(Subpop, Phe, Site, Type) %>%
+    pivot_wider(names_from = c("Subpop", "Phe", "Site", "Type"), 
+                values_from = "Value")
+
+  effects_df %>%
+    select(-value) %>%
+    write_csv(filename)
+}
+
+dataset_creation_function(m_gulf_gr_hyp, filename = "~/Github/pvdiv-phenology-gxe/manuscript/Coauthors/PNAS/Dataset_1_gulf_green-up.csv")
+dataset_creation_function(m_gulf_fl_hyp, filename = "~/Github/pvdiv-phenology-gxe/manuscript/Coauthors/PNAS/Dataset_4_gulf_flowering.csv")
+
+dataset_creation_function(m_gam_gr_hyp, filename = "~/Github/pvdiv-phenology-gxe/manuscript/Coauthors/PNAS/Dataset_3_both_subpopulations_green-up.csv")
+dataset_creation_function(m_gam_fl_hyp, filename = "~/Github/pvdiv-phenology-gxe/manuscript/Coauthors/PNAS/Dataset_6_both_subpopulations_flowering.csv")
+
+dataset_creation_function(m_midw_fl_hyp, filename = "~/Github/pvdiv-phenology-gxe/manuscript/Coauthors/PNAS/Dataset_5_midwest_flowering.csv",
+                          ncol = 21)
+dataset_creation_function(m_midw_gr_hyp, filename = "~/Github/pvdiv-phenology-gxe/manuscript/Coauthors/PNAS/Dataset_2_midwest_green-up.csv",
+                          ncol = 21)
 ```
+
+# Dataset 8
 
@@
 ```{r}
-get_post_weights <- function(m){
-  log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
-      as.data.frame() %>%
-      rownames_to_column(var = "value") %>%
-      dplyr::mutate(value = as.integer(.data$value)) %>%
-      as_tibble() %>%
-      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
-      dplyr::rename(log10BayesFactor = .data$V1) %>%
-      dplyr::select(-.data$value)
+m <- m_gulf_fl_hyp
+qtl_df <- qtlplot
+subpop = "Gulf"
+phe = "flowering"
 
-  pw_df <- as_tibble(m$posterior_weights)
-  pw_df <- switchgrassGWAS:::get_marker_df(m = m) %>%
-    bind_cols(pw_df)
+qtl_mash_overlap_function <- function(m, subpop, phe){
 
-  ggman_df <- pw_df %>%
-    #left_join(m_meta) %>%
-    tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
-             extra = "merge", convert = TRUE) %>%
-    dplyr::left_join(log10bf_df, by = "Marker") %>%
-    dplyr::arrange(.data$Chr, .data$Pos) 
-  return(ggman_df)
+  mash_df <- switchgrassGWAS:::get_marker_df(m) %>%
+    mutate(log10BF = switchgrassGWAS:::get_log10bf(m)) %>%
+    tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, 
+                    sep = "_", extra = "merge", convert = TRUE) %>%
+    dplyr::arrange(.data$Chr, .data$Pos) %>%
+    select(-value)
+
+  for(i in 1:nrow(qtl_df)){  # from 1 to 16
+    ## In block & > 2
+    qtl_sub <- mash_df %>%
+      filter(Chr %in% qtl_df$CHR[i] & between(Pos, qtl_df$POS_lo[i],
+                                               qtl_df$POS_hi[i])) %>% 
+      filter(log10BF > 2)
+
+    if(i == 1){
+    outputdf <- tibble(Chr = qtl_df$CHR[i], QTL_Pos_lo = qtl_df$POS_lo[i], 
+                       QTL_Pos_hi = qtl_df$POS_hi[i], 
+                       QTL_phenotype = qtl_df$lodcolumn[i], 
+                       QTL_LOD = qtl_df$lod[i], mash_Pos = qtl_sub$Pos, 
+                       mash_log10BF = qtl_sub$log10BF, mash_subpop = subpop, 
+                       mash_phenotype = phe)
+    } else {
+      outputdf1 <- tibble(Chr = qtl_df$CHR[i], QTL_Pos_lo = qtl_df$POS_lo[i], 
+                          QTL_Pos_hi = qtl_df$POS_hi[i], 
+                          QTL_phenotype = qtl_df$lodcolumn[i], 
+                          QTL_LOD = qtl_df$lod[i], mash_Pos = qtl_sub$Pos, 
+                          mash_log10BF = qtl_sub$log10BF, mash_subpop = subpop, 
+                          mash_phenotype = phe)
+      outputdf <- outputdf %>%
+        full_join(outputdf1)
+    }
+  }
+  return(outputdf)
 }
 
-pw_gg <- get_post_weights(m_gulf_gr_hyp) %>%
-  mutate(subpop = "Gulf",
-         phenotype = "greenup")
-pw_gf <- get_post_weights(m_gulf_fl_hyp) %>%
-  mutate(subpop = "Gulf",
-         phenotype = "flowering")
+overlap1 <- qtl_mash_overlap_function(m_gulf_fl_hyp, subpop = "Gulf", 
+                                      phe = "flowering") 
+overlap2 <- qtl_mash_overlap_function(m_midw_fl_hyp, subpop = "Midwest", 
+                                      phe = "flowering") 
+overlap3 <- qtl_mash_overlap_function(m_gam_fl_hyp, subpop = "Both", 
+                                      phe = "flowering")
 
-pw_mg <- get_post_weights(m_midw_gr_hyp) %>%
-  mutate(subpop = "Midwest",
-         phenotype = "greenup")
-pw_mf <- get_post_weights(m_midw_fl_hyp) %>%
-  mutate(subpop = "Midwest",
-         phenotype = "flowering")
+dataset_8 <- overlap1 %>%
+  full_join(overlap2) %>%
+  full_join(overlap3) %>%
+  arrange(Chr, mash_Pos)
 
-pw_gamg <- get_post_weights(m_gam_gr_hyp) %>%
-  mutate(subpop = "Both",
-         phenotype = "greenup")
-pw_gamf <- get_post_weights(m_gam_fl_hyp) %>%
-  mutate(subpop = "Both",
-         phenotype = "flowering")
+write_csv(dataset_8, file = "~/Github/pvdiv-phenology-gxe/manuscript/Coauthors/PNAS/Dataset_8_QTL_mash_overlap.csv")
 ```
-
-# Get any mash overlaps for QTL intervals
-```{r}
-pw_3f <- pw_gf %>%
-  full_join(pw_mf) %>%
-  full_join(pw_gamf) %>%
-  arrange(Chr, Pos) %>%
-  filter(log10BayesFactor > 3) %>%
-  select(Chr:Pos, log10BayesFactor, subpop, phenotype, everything())
-
-pw_3f_all <- pw_gf %>%
-  full_join(pw_mf) %>%
-  full_join(pw_gamf) %>%
-  arrange(Chr, Pos) %>%
-  select(Chr:Pos, log10BayesFactor, subpop, phenotype, everything())
-
-qtlplot
-summary(pw_3f$log10BayesFactor)
-quantile(pw_gf$log10BayesFactor, 0.96196895)
-quantile(pw_gf$log10BayesFactor, 0.96196895)
-quantile(pw_gamf$log10BayesFactor, 0.8006157)
-1-0.96196895
-1-0.8006157
-2/464  # GT 0.5%? No 0.0043%
-22/418 # GT 3.8031? Yes 5.26%
-42/255 # GT 19.93843%? No 16.47%
-```
-
-```{r}
-i=16  # from 1 to 16
-## In block & > 2
-QTLandBF <- pw_3f %>%
-  filter(Chr %in% qtlplot$CHR[i] & between(Pos, qtlplot$POS_lo[i], qtlplot$POS_hi[i])) %>%
-  group_by(subpop) %>% tally()
-## BF>2
-inBF <- pw_3f %>%
-  group_by(subpop) %>% tally()
-## BF not > 2 (all minus BF > 2, all is below)
-allwrtBF <- pw_3f_all %>%
-  group_by(subpop) %>% tally()
-## All in block
-inQTL <- pw_3f_all %>%
-  filter(Chr %in% qtlplot$CHR[i] & between(Pos, qtlplot$POS_lo[i], qtlplot$POS_hi[i])) %>%
-  group_by(subpop) %>% tally()
-
-QTLandBF
-inQTL
-
-phyper(QTLandBF$n[1], inBF$n[1], allwrtBF$n[1]-inBF$n[1], inQTL$n[1], 
-       lower.tail = FALSE) # Both
-phyper(QTLandBF$n[2], inBF$n[2], allwrtBF$n[2]-inBF$n[2], inQTL$n[2], 
-       lower.tail = FALSE) # Gulf
-phyper(QTLandBF$n[3], inBF$n[3], allwrtBF$n[3]-inBF$n[3], inQTL$n[3], 
-       lower.tail = FALSE) # Midwest
-```
-BF > 2 results: 5 QTL intervals
-1 ~ G
-2: M
-6: G, M
-10: G&M, M
-11: G, M
-12: G&M, G
-14 ~ G&M
-16 ~ G&M
-
-BF > 3 results: 7 QTL intervals. Different intervals, too. So I think having the same BF cutoff for these different mash results will give misleading conclusions. Do a % cutoff instead.
-4 ~ G&M
-6: G
-8: G, M
-11: M
-12: G&M, G
-13: M
-14: G&M
-15: M
-
-1137 SNPs in 2K region. 66 sig > 2 in mash. 2/464 Midwest, 22/418 Gulf, 42/255 Both.
-
-Summary statistics for log 10 Bayes Factor for pw_3f df
-   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
-  2.000   2.234   2.568   2.848   3.140  10.385 
-90% 3.936367
-95% 4.622242
-99% 6.257548
-
-Gulf
-90% 1.410253
-95% 1.828653
-99% 2.911984
-Top 3.8% of Gulf is above 2.
-
-Midwest 
-90% 0.72685
-95% 0.949753
-99% 1.616983
-Top 0.5% of Midwest is above 2.
-
-G&M
-90% 2.595963
-95% 3.196745
-99% 4.65753
-top 20% of G&M is above 2.
-
-## Hypergeometric test for significance of overlaps
-phyper(10, 50, 200 - 50, 25, lower.tail = FALSE)
-phyper(A&B, inB, notinB, inA, lower.tail = FALSE)
-```{r}
-phyper(22, 336, 50490-336, 418, lower.tail = FALSE)
-phyper(AandB$n[2], inB$n[2], allwrtB$n[2]-inB$n[2], inA$n[2], 
-       lower.tail = FALSE)
-```
-
-In block & > 2, >2, not > 2, in block
-22, 336, 50490-336, 418
-
-BF > 2
-Both	7125			
-Gulf	1921			
-Midwest	336
-
-All SNPs
-Both	35735			
-Gulf	50490			
-Midwest	63367
 
@@
 
+get_post_weights <- function(m){
+  log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
+      as.data.frame() %>%
+      rownames_to_column(var = "value") %>%
+      dplyr::mutate(value = as.integer(.data$value)) %>%
+      as_tibble() %>%
+      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
+      dplyr::rename(log10BayesFactor = .data$V1) %>%
+      dplyr::select(-.data$value)
+
+  pw_df <- as_tibble(m$posterior_weights)
+  pw_df <- switchgrassGWAS:::get_marker_df(m = m) %>%
+    bind_cols(pw_df)
+
+  ggman_df <- pw_df %>%
+    #left_join(m_meta) %>%
+    tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
+             extra = "merge", convert = TRUE) %>%
+    dplyr::left_join(log10bf_df, by = "Marker") %>%
+    dplyr::arrange(.data$Chr, .data$Pos) 
+  return(ggman_df)
+}
+
 qtlplot <- qtlplot %>%
@@
 9N: 1 QTL, 0 en (lod 18)
-
-in Midwest & for CGDD: 3
-in Midwest & for dyln:
-in Gulf & for CGDD: 
-```{r}
-qtlplot %>%
-  group_by(lodcolumn) %>%
-  tally()
-qtlplot %>%
-  group_by(CHR, Pos_Mb_lo, Pos_Mb_hi) %>%
-  tally()
-qtlplot %>%
-  filter(lodcolumn %in% c("dyln_change_sec", "FL50"))
-```
-
-# Make random QTL dataframe
-
-This random df just needs a CHR, POS_lo, and POS_hi column to work in the get_QTL_enrichment function I made. 
-```{r}
-sig_random <- c()
-chr_size_mash <- pw_3f_all %>%
-  ungroup() %>%
-  group_by(Chr) %>%
-  dplyr::summarise(Pos_lo = min(Pos),
-                   Pos_hi = max(Pos))
-
-for(j in 1:1000){
-  qtl_size <- qtlplot %>%
-    select(CHR, POS_lo, POS_hi) %>%
-    unique() %>%   # don't doublecount completely overlapping QTL
-    #(does doublecount partially overlapping QTL, however.)
-    group_by(CHR, POS_lo) %>%
-    dplyr::mutate(Size = POS_hi - POS_lo) %>%
-    dplyr::select(CHR, POS_lo, Size)
-
-  rchr <- chr_size_mash$Chr[sample(1:18, nrow(qtl_size), replace = TRUE)]
-  rposlo <- c()
-  for(i in 1:length(rchr)){
-    rposlo[i] <- sample(chr_size_mash$Pos_lo[which(chr_size_mash$Chr == rchr[i])]:chr_size_mash$Pos_hi[which(chr_size_mash$Chr == rchr[i])], 1)
-  }
-
-  random_qtl <- tibble(CHR = rchr, POS_lo = rposlo, Size = qtl_size$Size) %>%
-    mutate(POS_hi = POS_lo + Size) %>%
-    arrange(CHR, POS_lo)
-
-  random_gf <- get_QTL_enrichment(m_gulf_fl_hyp, qtl_df = random_qtl, 0.99) %>%
-    mutate(subpop = "Gulf", phenotype = "flowering")
-  random_mf <- get_QTL_enrichment(m_midw_fl_hyp, qtl_df = random_qtl, 0.99) %>%
-    mutate(subpop = "Midwest", phenotype = "flowering")
-  random_gamf <- get_QTL_enrichment(m_gam_fl_hyp, qtl_df = random_qtl, 0.99) %>%
-    mutate(subpop = "Both", phenotype = "flowering")
-
-  random_df <- random_gf %>%
-    full_join(random_mf) %>%
-    full_join(random_gamf) %>%
-    arrange(Chr, Pos_lo)
-
-  sig_random[j] <- nrow(filter(random_df, pvalue < 0.05))
-}
-saveRDS(sig_random, file = "../data/Random_QTL_df_significant_enrichments_of_mash_1per_tail_SNPs_1000runs.rds")
-tibble(`mash 1% tail significant enrichments` = sig_random) %>%
-  ggplot(aes(x = `mash 1% tail significant enrichments` )) +
-  switchgrassGWAS::theme_oeco +
-  geom_histogram(binwidth = 1) + 
-  xlim(c(1, 24)) + ylab("") +
-  geom_vline(xintercept = 16, linetype = 2, color = "red")
-save_plot(filename = "../manuscript/plots/Figure_4_QTL_overlaps_random_QTL_enrichments_with_mash_1000runs.png", plot = last_plot(), base_height = 1.8, base_asp = 1.8)
-
-tibble(sig_r = sig_random) %>%
-  arrange(desc(sig_r))
-3/1000
-```

### 2021-06-29 R/UT_Dataverse_files.Rmd
Prep supplement and stable data replication links for SNP data

---
title: "UT Dataverse"
author: "Alice MacQueen"
date: 2021-06-29
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bigsnpr)
```

# Make bed files

Make bed/bim/fam SNP files for Midwest, Gulf, and Both genetic subpopulations.
I've used '.rds' and '.bk' files for the analyses in this manuscript, but these
are not common genome file formats, while bed formatted files are. So uploading
files in this format to a permanent DOI will help reproducibility for this
dataset.

I only need to do this for the 'Both' subpopulations; the Midwest and Gulf
subpopulation SNP sets were generated for the genome paper.

```{r}
snp <- snp_attach("~/Github/pvdiv-genome/genetic_subpops/kinship/Pvirgatum_V5_GWAS_363g_Midwest_and_Gulf_subset_maf0.05.rds")
```

```{r}
snp_writeBed(snp, bedfile = "~/Github/pvdiv-genome/Dataverse/Pvirgatum_V5_GWAS_363g_Midwest_and_Gulf_subset_maf0.05.bed")
```

### 2021-06-29 R/Pre_manuscript/1_SNP_file_filtering_by_common_garden_site.R
Remove from git repository files and folders that aren't pertinent to the completed manuscript

@@ -1,32 +1,0 @@
-library(tidyverse)
-library(bigsnpr)
-library(switchgrassGWAS)
-
-
-pvdiv_pop_subset_bigsnp <- function(subset, selector){
-  subset_name <- paste0("Pvirgatum_4x_784g_imp_phased_", subset, "_subset")
-  snpfile_subset <- subset(snp, ind.row = selector)
-  snp_subset <- snp_attach(snpfile_subset)
-  snp_writeBed(snp_subset, bedfile = paste0(subset_name, ".bed"))
-  mac <- round(25/length(selector), digits = 2) # what should the MAF be?
-  if(mac > 0.05){
-    mac <- 0.05
-  }
-  plink <- download_plink()
-  bedfile <- snp_plinkQC(plink, prefix.in = subset_name,
-                         maf = mac, geno = 0.1, hwe = 1e-50, 
-                         prefix.out = paste0(subset_name, "_maf", mac),
-                         extra.options = "--allow-no-sex --allow-extra-chr --chr Chr01K Chr01N Chr02K Chr02N Chr03K Chr03N Chr04K Chr04N Chr05K Chr05N Chr06K Chr06N Chr07K Chr07N Chr08K Chr08N Chr09K Chr09N")
-  rdsfile <- snp_readBed(bedfile)
-}
-
-
-# Load data - if you get errors, make sure to run R from the GWAS/ folder in pubjuenger
-snp <- snp_attach("/home/pubjuenger/GWAS/Pvirgatum_4x_784g_imp_phased_maf0.005.rds")
-switchgrassGWAS::phenotypes
-
-# Use phenotypes to set up selection criteria for subsets. These need to be vectors of the row numbers in the Taxa dataframe that you have phenotypes for.
-gwasct_sel <- which(!is.na(phenotypes$GWAS_CT)) # gwas_ct GWAS
-
-## Run function to generate subset SNP files for the subset
-pvdiv_pop_subset_bigsnp(subset = "gwasct", selector = gwasct_sel)

### 2021-06-29 R/Pre_manuscript/BIOMASS_mash.R

@@ -1,77 +1,0 @@
-library(tidyverse)
-library(bigsnpr)
-library(switchgrassGWAS)
-source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
-
-
-NCORES <- nb_cores()
-
-snp <- snp_attach(file.path("/", "home", "alice", "Github", "pvdiv-genome",
-                            "tensite_twoyear",
-                            "Pvirgatum_V5_GWAS_630g_33M_tensite_twoyear.rds"))
-
-G <- snp$genotypes
-CHR <- snp$map$chromosome
-POS <- snp$map$physical.pos
-plants <- snp$fam$sample.ID
-bonferroni <- -log10(0.05/length(POS))
-
-CHRN <- enframe(CHR, name = NULL) %>%
-  dplyr::rename(CHR = .data$value) %>%
-  mutate(CHRN = case_when(.data$CHR == "Chr01K" ~ 1,
-                          .data$CHR == "Chr01N" ~ 2,
-                          .data$CHR == "Chr02K" ~ 3,
-                          .data$CHR == "Chr02N" ~ 4,
-                          .data$CHR == "Chr03K" ~ 5,
-                          .data$CHR == "Chr03N" ~ 6,
-                          .data$CHR == "Chr04K" ~ 7,
-                          .data$CHR == "Chr04N" ~ 8,
-                          .data$CHR == "Chr05K" ~ 9,
-                          .data$CHR == "Chr05N" ~ 10,
-                          .data$CHR == "Chr06K" ~ 11,
-                          .data$CHR == "Chr06N" ~ 12,
-                          .data$CHR == "Chr07K" ~ 13,
-                          .data$CHR == "Chr07N" ~ 14,
-                          .data$CHR == "Chr08K" ~ 15,
-                          .data$CHR == "Chr08N" ~ 16,
-                          .data$CHR == "Chr09K" ~ 17,
-                          .data$CHR == "Chr09N" ~ 18,
-                          TRUE ~ 19
-  ))
-markers <- tibble(CHR = CHR, POS = POS)
-
-
-# Run bigsnp2mashr
-
-markers20per <- snp_clumping(G, infos.chr = CHRN$CHRN, ncores = NCORES)
-markers2 <- markers[markers20per,] # Get unlinked markers for mash
[…]

### 2021-06-29 R/Pre_manuscript/BIOMASS_tensite_twoyear.R

@@ -1,160 +1,0 @@
-library(tidyverse)
-library(bigsnpr)
-library(switchgrassGWAS)
-library(AnnotationDbi)
-library(multtest)
-library(viridis)
-library(cowplot)
-library(data.table)
-source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
-
-NCORES <- nb_cores()
-
-snp <- snp_attach(file.path("/", "home", "alice", "Github", "pvdiv-genome",
-                            "tensite_twoyear",
-                            "Pvirgatum_V5_GWAS_630g_33M_tensite_twoyear.rds"))
-
-G <- snp$genotypes
-CHR <- snp$map$chromosome
-POS <- snp$map$physical.pos
-plants <- snp$fam$sample.ID
-bonferroni <- -log10(0.05/length(POS))
-
-CHRN <- enframe(CHR, name = NULL) %>%
-  dplyr::rename(CHR = .data$value) %>%
-  mutate(CHRN = case_when(.data$CHR == "Chr01K" ~ 1,
-                          .data$CHR == "Chr01N" ~ 2,
-                          .data$CHR == "Chr02K" ~ 3,
-                          .data$CHR == "Chr02N" ~ 4,
-                          .data$CHR == "Chr03K" ~ 5,
-                          .data$CHR == "Chr03N" ~ 6,
-                          .data$CHR == "Chr04K" ~ 7,
-                          .data$CHR == "Chr04N" ~ 8,
-                          .data$CHR == "Chr05K" ~ 9,
-                          .data$CHR == "Chr05N" ~ 10,
-                          .data$CHR == "Chr06K" ~ 11,
-                          .data$CHR == "Chr06N" ~ 12,
-                          .data$CHR == "Chr07K" ~ 13,
-                          .data$CHR == "Chr07N" ~ 14,
-                          .data$CHR == "Chr08K" ~ 15,
-                          .data$CHR == "Chr08N" ~ 16,
-                          .data$CHR == "Chr09K" ~ 17,
-                          .data$CHR == "Chr09N" ~ 18,
-                          TRUE ~ 19
-  ))
-markers <- tibble(CHR = CHR, POS = POS)
-
-svd <- readRDS(file.path("/", "home", "alice", "Github", "pvdiv-genome",
-                         "tensite_twoyear", "SVD_20PCs_630g_33M_tensite_twoyear.rds"))
-load("~/Github/pvdiv-phenology-gxe/data/phenotypes_2018.rda")
[…]

### 2021-06-29 R/Pre_manuscript/DEAD_tensite_twoyear.R

@@ -1,159 +1,0 @@
-library(tidyverse)
-library(bigsnpr)
-library(switchgrassGWAS)
-library(AnnotationDbi)
-library(multtest)
-library(viridis)
-library(cowplot)
-library(data.table)
-source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
-
-NCORES <- nb_cores()
-
-snp <- snp_attach(file.path("/", "home", "alice", "Github", "pvdiv-genome",
-                            "tensite_twoyear",
-                            "Pvirgatum_V5_GWAS_630g_33M_tensite_twoyear.rds"))
-
-G <- snp$genotypes
-CHR <- snp$map$chromosome
-POS <- snp$map$physical.pos
-plants <- snp$fam$sample.ID
-bonferroni <- -log10(0.05/length(POS))
-
-CHRN <- enframe(CHR, name = NULL) %>%
-  dplyr::rename(CHR = .data$value) %>%
-  mutate(CHRN = case_when(.data$CHR == "Chr01K" ~ 1,
-                          .data$CHR == "Chr01N" ~ 2,
-                          .data$CHR == "Chr02K" ~ 3,
-                          .data$CHR == "Chr02N" ~ 4,
-                          .data$CHR == "Chr03K" ~ 5,
-                          .data$CHR == "Chr03N" ~ 6,
-                          .data$CHR == "Chr04K" ~ 7,
-                          .data$CHR == "Chr04N" ~ 8,
-                          .data$CHR == "Chr05K" ~ 9,
-                          .data$CHR == "Chr05N" ~ 10,
-                          .data$CHR == "Chr06K" ~ 11,
-                          .data$CHR == "Chr06N" ~ 12,
-                          .data$CHR == "Chr07K" ~ 13,
-                          .data$CHR == "Chr07N" ~ 14,
-                          .data$CHR == "Chr08K" ~ 15,
-                          .data$CHR == "Chr08N" ~ 16,
-                          .data$CHR == "Chr09K" ~ 17,
-                          .data$CHR == "Chr09N" ~ 18,
-                          TRUE ~ 19
-  ))
-markers <- tibble(CHR = CHR, POS = POS)
-
-svd <- readRDS(file.path("/", "home", "alice", "Github", "pvdiv-genome",
-                         "tensite_twoyear", "SVD_20PCs_630g_33M_tensite_twoyear.rds"))
-load("~/Github/pvdiv-phenology-gxe/data/phenotypes_2018.rda")
-
[…]

### 2021-06-29 R/Pre_manuscript/DEAD_tensite_twoyear_linear.R

@@ -1,162 +1,0 @@
-library(tidyverse)
-library(bigsnpr)
-library(switchgrassGWAS)
-library(AnnotationDbi)
-library(multtest)
-library(viridis)
-library(cowplot)
-library(data.table)
-source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
-
-NCORES <- nb_cores()
-
-snp <- snp_attach(file.path("/", "home", "alice", "Github", "pvdiv-genome",
-                            "tensite_twoyear",
-                            "Pvirgatum_V5_GWAS_630g_33M_tensite_twoyear.rds"))
-
-G <- snp$genotypes
-CHR <- snp$map$chromosome
-POS <- snp$map$physical.pos
-plants <- snp$fam$sample.ID
-bonferroni <- -log10(0.05/length(POS))
-
-CHRN <- enframe(CHR, name = NULL) %>%
-  dplyr::rename(CHR = .data$value) %>%
-  mutate(CHRN = case_when(.data$CHR == "Chr01K" ~ 1,
-                          .data$CHR == "Chr01N" ~ 2,
-                          .data$CHR == "Chr02K" ~ 3,
-                          .data$CHR == "Chr02N" ~ 4,
-                          .data$CHR == "Chr03K" ~ 5,
-                          .data$CHR == "Chr03N" ~ 6,
-                          .data$CHR == "Chr04K" ~ 7,
-                          .data$CHR == "Chr04N" ~ 8,
-                          .data$CHR == "Chr05K" ~ 9,
-                          .data$CHR == "Chr05N" ~ 10,
-                          .data$CHR == "Chr06K" ~ 11,
-                          .data$CHR == "Chr06N" ~ 12,
-                          .data$CHR == "Chr07K" ~ 13,
-                          .data$CHR == "Chr07N" ~ 14,
-                          .data$CHR == "Chr08K" ~ 15,
-                          .data$CHR == "Chr08N" ~ 16,
-                          .data$CHR == "Chr09K" ~ 17,
-                          .data$CHR == "Chr09N" ~ 18,
-                          TRUE ~ 19
-  ))
-markers <- tibble(CHR = CHR, POS = POS)
-
-svd <- readRDS(file.path("/", "home", "alice", "Github", "pvdiv-genome",
-                         "tensite_twoyear", "SVD_20PCs_630g_33M_tensite_twoyear.rds"))
-load("~/Github/pvdiv-phenology-gxe/data/phenotypes_2018.rda")
-
[…]

### 2021-06-29 R/Pre_manuscript/Example_annotation_script.R

@@ -1,10 +1,0 @@
-library(tibble)
-library(switchgrassGWAS)
-library(AnnotationDbi)
-txdb <- loadDb(file = file.path("/", "home", "SharedUser", "PV",
-                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
-
-df <- tibble(CHR = "Chr03K", start = 50800000, end = 50900000)
-
-annos <- pvdiv_table_topsnps(df = df, type = "table", txdb = txdb)
-# annos should be a dataframe containing one gene, Pavir.3KG393200

### 2021-06-29 R/Pre_manuscript/GWAS_D2F_linear_first_lambdagc.R

@@ -1,185 +1,0 @@
-library(tidyverse)
-library(bigsnpr)
-library(switchgrassGWAS)
-library(AnnotationDbi)
-library(multtest)
-library(viridis)
-library(cowplot)
-library(data.table)
-source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
-get_date_filename <- function(){
-  str_replace_all(str_replace_all(Sys.time(), ":", "."), " ", "_")
-}
-
-
-# ---------- Choose SNP file & SVD/Kinship matrix appropriate for number of sites
-snp <- snp_attach(file.path("~", "Github", "pvdiv-genome",
-                            "tensite_twoyear",
-                            "Pvirgatum_V5_GWAS_630g_33M_tensite_twoyear.rds"))
-svd <- readRDS(file.path("~", "Github", "pvdiv-genome",
-                         "tensite_twoyear",
-                         "SVD_20PCs_630g_33M_tensite_twoyear.rds"))
-
-# ---------- Load or make some useful objects ------
-
-NCORES <- nb_cores()
-G <- snp$genotypes
-CHR <- snp$map$chromosome
-POS <- snp$map$physical.pos
-plants <- snp$fam$sample.ID
-bonferroni <- -log10(0.05/length(POS))
-markers <- tibble(CHR = CHR, POS = POS)
-
-txdb <- loadDb(file = file.path("/", "home", "alice", "Github", "pvdiv-genome",
-                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
-
-anno_info <-
-  read_delim(file = file.path("/", "home", "alice", "Github", "pvdiv-genome",
-                              "Pvirgatum_516_v5.1.annotation_info.txt"),
-             col_names = TRUE, delim = "\t")
-gene_anno_info <- tbl_df(anno_info) %>%
-  distinct(locusName, .keep_all = TRUE)
-
-
-# ------- Set analysis directory and phenotypes
-analysisdir <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis",
-                         "gwas", "linear")
-setwd(analysisdir)
-phe_df <- read_rds(file.path("~", "Github",
-                             "pvdiv-phenotypes", "data",
-                             "Phenotypes_D2F_fourway_project.rds"))
-
-
-df <- plants %>%
-  enframe(name= NULL, value = "PLANT_ID") %>%
[…]

### 2021-06-29 R/Pre_manuscript/GWAS_GDD.R

@@ -1,19 +1,0 @@
-library(bigsnpr)
-library(tidyverse)
-library(switchgrassGWAS)
-library(AnnotationDbi)
-txdb <- loadDb(file = file.path("~", "Github", "pvdiv-genome",
-                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
-
-outputdir <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", "gwas", "GDD")
-
-inputfiles <- read_delim("~/Github/pvdiv-phenology-gxe/analysis/gwas/inputkinship.txt", delim = " ", col_names = "SNPfiles") # Midwest & Gulf, Midwest, Gulf, Atlantic, Midwest & Atlantic, Gulf & Atlantic
-
-phe_subsets <- read_rds("~/Github/pvdiv-phenology-gxe/data/Phenotype_Subsets_for_Subpop_GWAS.rds")
-
-for(i in 1:nrow(inputfiles)){
-  phe_df <- phe_subsets[[i]]
-  snp <- snp_attach(inputfiles$SNPfiles[i])
-  pvdiv_standard_gwas(snp, df = phe_df, type = "linear", outputdir = outputdir,
-                      savegwas = TRUE, saveannos = FALSE, minphe = 100)
-}

### 2021-06-29 R/Pre_manuscript/GWAS_standard_kinship.R

@@ -1,20 +1,0 @@
-library(bigsnpr)
-library(tidyverse)
-library(switchgrassGWAS)
-library(AnnotationDbi)
-txdb <- loadDb(file = file.path("~", "Github", "pvdiv-genome",
-                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
-
-outputdir <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", "gwas")
-
-inputfiles <- read_delim("~/Github/pvdiv-phenology-gxe/analysis/gwas/inputkinship.txt", delim = " ", col_names = "SNPfiles")
-
-phe_df <- read_rds(file.path("~", "Github",
-                             "pvdiv-phenology-gxe", "data",
-                             "Phenotypes_D2F_fourway_project.rds"))
-for(i in 2:nrow(inputfiles)){
-  snp <- snp_attach(inputfiles$SNPfiles[i])
-  pvdiv_standard_gwas(snp, df = phe_df, type = "linear", outputdir = outputdir,
-                      savegwas = TRUE, saveannos = TRUE, txdb = txdb,
-                      minphe = 120)
-}

### 2021-06-29 R/Pre_manuscript/get_results_from_mash.R

@@ -1,54 +1,0 @@
-#' @title Get column names from a mash object
-#'
-#' @description This function extracts the column names from the local false
-#' sign rate table of a mash object's results. This can tell you the condition
-#' names or phenotype names used in the mash object. That can be useful for
-#' looking at a subset of these columns, say.
-#'
-#' @param m An object of type mash
-#'
-#' @return A vector of phenotype names
-#'
-#' @examples
-#'     \dontrun{get_colnames(m = mash_obj)}
-#'
-get_colnames <- function(m){
-  column_names <- colnames(m$result$lfsr)
-  return(column_names)
-}
-
-#' From a mash result, get effects that are significant in at least
-#'     one condition.
-#'
-#' @param m the mash result (from joint or 1by1 analysis)
-#' @param thresh indicates the threshold below which to call signals significant
-#' @param conditions which conditions to include in check (default to all)
-#' @param sig_fn the significance function used to extract significance from
-#'     mash object; eg could be ashr::get_lfsr or ashr::get_lfdr. (Small values
-#'     must indicate significant.)
-#'
-#' @return a vector containing the indices of the significant effects, by
-#'     order of most significant to least
-#'
-#' @importFrom ashr get_lfsr
-#'
-#' @export
-get_significant_results = function(m, thresh = 0.05, conditions = NULL,
-                                   sig_fn = ashr::get_lfsr) {
-  if (is.null(conditions)) {
-    conditions = 1:get_ncond(m)
-  }
-  top = apply(sig_fn(m)[,conditions,drop=FALSE],1,min) # find top effect
-  # in each condition
-  sig = which(top < thresh)
-  ord = order(top[sig],decreasing=FALSE)
-  sig[ord]
-}
-
-require(ashr)
-require(tidyverse)
-effectplot <- get_colnames(m) %>%
-  enframe(name = NULL, value = "Conditions") %>%
-  mutate(Conditions = str_sub(.data$Conditions, start = 6),
-         mn = get_pm(m)[i,],
-         se = get_psd(m)[i,])

### 2021-06-29 R/Pre_manuscript/gwas_33M_subset_fourway_seven.R

@@ -1,123 +1,0 @@
-library(tidyverse)
-library(bigsnpr)
-library(switchgrassGWAS)
-
-NCORES <- nb_cores()
-
-snp <- snp_attach(file.path("/", "home", "alice", "Github", "pvdiv-genome",
-                            "fourway_seven",
-                            "Pvirgatum_V5_GWAS_630g_33M_fourway_seven.rds"))
-
-G <- snp$genotypes
-CHR <- snp$map$chromosome
-POS <- snp$map$physical.pos
-plants <- snp$fam$sample.ID
-bonferroni <- -log10(0.05/length(POS))
-
-CHRN <- enframe(CHR, name = NULL) %>%
-  dplyr::rename(CHR = .data$value) %>%
-  mutate(CHRN = case_when(.data$CHR == "Chr01K" ~ 1,
-                          .data$CHR == "Chr01N" ~ 2,
-                          .data$CHR == "Chr02K" ~ 3,
-                          .data$CHR == "Chr02N" ~ 4,
-                          .data$CHR == "Chr03K" ~ 5,
-                          .data$CHR == "Chr03N" ~ 6,
-                          .data$CHR == "Chr04K" ~ 7,
-                          .data$CHR == "Chr04N" ~ 8,
-                          .data$CHR == "Chr05K" ~ 9,
-                          .data$CHR == "Chr05N" ~ 10,
-                          .data$CHR == "Chr06K" ~ 11,
-                          .data$CHR == "Chr06N" ~ 12,
-                          .data$CHR == "Chr07K" ~ 13,
-                          .data$CHR == "Chr07N" ~ 14,
-                          .data$CHR == "Chr08K" ~ 15,
-                          .data$CHR == "Chr08N" ~ 16,
-                          .data$CHR == "Chr09K" ~ 17,
-                          .data$CHR == "Chr09N" ~ 18,
-                          TRUE ~ 19
-  ))
-markers <- tibble(CHR = CHR, POS = POS)
-
-svd <- readRDS(file.path("/", "home", "alice", "Github", "pvdiv-genome",
-                         "fourway_seven", "SVD_20PCs_630g_33M_fourway_seven.rds"))
-# load("~/Github/pvdiv-phenology-gxe/data/phenotypes_2018.rda")
-load("~/Github/pvdiv-phenology-gxe/data/phenotypes_2019.rda")
-
[…]

### 2021-06-29 R/Pre_manuscript/mash_d2f_gulf_midwest_set.R

@@ -1,152 +1,0 @@
-library(tidyverse)
-library(bigsnpr)
-library(switchgrassGWAS)
-
-#  devtools::install_github("alice-macqueen/switchgrassgwas")
-inputfiles <- read_delim("~/Github/pvdiv-fitness-2019/analysis/inputkinship.txt", 
-                         delim = " ", col_names = "SNPfiles")
-outputdir <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", "gwas", 
-                       "Gulf_Midwest")
-
-snp <- snp_attach(inputfiles$SNPfiles[4])
-gwas_rds <- pvdiv_results_in_folder(path = outputdir, 
-                                    pattern = "GWAS_datatable_(PKLE|STIL|KING|KBSM|CLMB|BRKG|LINC)_D2F") 
-# Not enough Gulf & Midwest individuals at BRKG to run the univariate GWAS (<120).
-phenotype_names <- str_sub(gwas_rds, start = 16, end = 23)
-numSNPs <- ceiling(ceiling(1000000/length(phenotype_names)^2)/500)*500 # round 
-# up number of SNPs to nearest 500.
-
-mashin <- pvdiv_bigsnp2mashr(path = outputdir, 
-                             snp = snp, gwas_rds = gwas_rds, numSNPs = numSNPs, 
-                             phenotypes = phenotype_names, model = "linear", 
-                             saveoutput = TRUE, scale = FALSE)
-
-m <- mash_standard_run(path = outputdir, list_input = mashin,
-                       saveoutput = TRUE, ref = NA)
-
-source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
-mash_plot_manhattan_by_condition(m = m, saveoutput = TRUE)
-mash_plot_sig_by_condition(m = m, saveoutput = TRUE)
-
-
-library(AnnotationDbi)
-txdb <- loadDb(file = file.path("~", "Github", "pvdiv-genome",
-                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
-annos <- pvdiv_table_topsnps(df = m, type = "mash", txdb = txdb, 
-                             snp = snp)
-saveRDS(annos, file = file.path(outputdir, paste0("Annotation_tables_mash_", 
-                                                  "fourway_sites_", 
-                                                  "Midwest_Gulf_D2F.rds")))
-
-library(ashr)
-for(n in 1:10){
[…]

### 2021-06-29 README.Rmd
Modify readme to reflect repository contents for completed manuscript

@@
 
-The goal of pvdiv-phenology-gxe is to analyze the genetics of GxE in the phenology data collected for the switchgrass diversity panel in 2018 and 2019, including tiller count in 2018, survival to 2019, and greenup to tiller count & biomass in 2019.
-This repository will be private until the associated dataset is published.
+pvdiv-phenology-gxe contains the non-genomic data and code used to analyze the genetics of GxE in the phenology data collected for the switchgrass diversity panel in 2019, including green-up dates and flowering dates in 2019.
+This repository will be private until the associated manuscript is published.
 
-As of December 2019, this analysis is on data Jason Bonette compiled 2019-12-25.
-I create downstream data objects from data collected on this date.
-At the beginning of every month, and whenever I hear from him, I will check his dataset for any changes.
+In general, if you have any questions about how to locate an analysis folder or code used for an analysis, look at this README first.
+Then, if you still have questions, feel free to open a pull request with your issue and I will be happy to help.
 
-R Files:
+## Repository Files:
 
-In data-raw/:
+### In data-raw/:
 
@@
     *Taxa.R creates the Taxa rda file used to tie sequence data to the 785 sequenced PLANT_IDs.
+
+### In data/:
+
+    * Contains datasets as ".rds" files that are used in analyses (to produce results in the analysis/ folder). These files can be read into R using the `readRDS()` function.
+
+### In analysis/:
+
+    * Data objects and draft figures created as part of the analysis of heritability (including variance components analysis, done in ASReml), genome-wide association (on phenotypic BLUPs, done using the switchgrassGWAS R package), and mash (on results from the univariate GWAS, done using the switchgrassGWAS R package).
+Generally, the code to replicate these analyses can be found in the R/ directory.
+
+### In R/:
+
+    * Contains the majority of the code used in the manuscript (some additional code used to prepare data is in data-raw/.) 
+    * Files that start with "Plots" were used to make manuscript figures
+    * Files that start with "nohup" were used to run mash
+    * Files that start with "Analysis" were used to generate files in the analysis/ folder.
+I tried to name these descriptively but there are a lot of them; please open a pull request if you want help locating specific code.
+
+### In phenology-fourway/:
+
+    * Code and data objects needed to replicate the genetic mapping of flowering-related QTL in the pseudo-F2 cross.
+
+### In manuscript/:
+
+    * Contains completed drafts of a manuscript on the genetics of GxE in phenology in the switchgrass diversity panel.

### 2021-06-29 README.Rmd
Bulleted lists in Readme

@@
 
-    *metadata.R creates the metadata rda file used for subsequent analyses.
-    *phenotypes_2018.R creates the phenotypic data rda file used for analyses of 2018 data
-    *phenotypes_2019.Rmd creates the phenotypic data rda file used for analyses of 2019 data.
+    - metadata.R creates the metadata rda file used for subsequent analyses.
+    - phenotypes_2018.R creates the phenotypic data rda file used for analyses of 2018 data
+    - phenotypes_2019.Rmd creates the phenotypic data rda file used for analyses of 2019 data.
 These plants differ from those in phenotypes_2018 - many plants that died were replaced by Blackwell cultivar clones, and any phenotypes measured on these clones were not included in downstream analyses.
-    *Taxa.R creates the Taxa rda file used to tie sequence data to the 785 sequenced PLANT_IDs.
+    - Taxa.R creates the Taxa rda file used to tie sequence data to the 785 sequenced PLANT_IDs.
 
@@
 
-    * Contains datasets as ".rds" files that are used in analyses (to produce results in the analysis/ folder). These files can be read into R using the `readRDS()` function.
+    - Contains datasets as ".rds" files that are used in analyses (to produce results in the analysis/ folder). These files can be read into R using the `readRDS()` function.
 
@@
 
-    * Data objects and draft figures created as part of the analysis of heritability (including variance components analysis, done in ASReml), genome-wide association (on phenotypic BLUPs, done using the switchgrassGWAS R package), and mash (on results from the univariate GWAS, done using the switchgrassGWAS R package).
+    - Data objects and draft figures created as part of the analysis of heritability (including variance components analysis, done in ASReml), genome-wide association (on phenotypic BLUPs, done using the switchgrassGWAS R package), and mash (on results from the univariate GWAS, done using the switchgrassGWAS R package).
 Generally, the code to replicate these analyses can be found in the R/ directory.
@@
 
-    * Contains the majority of the code used in the manuscript (some additional code used to prepare data is in data-raw/.) 
-    * Files that start with "Plots" were used to make manuscript figures
-    * Files that start with "nohup" were used to run mash
-    * Files that start with "Analysis" were used to generate files in the analysis/ folder.
+    - Contains the majority of the code used in the manuscript (some additional code used to prepare data is in data-raw/.) 
+    - Files that start with "Plots" were used to make manuscript figures
+    - Files that start with "nohup" were used to run mash
+    - Files that start with "Analysis" were used to generate files in the analysis/ folder.
 I tried to name these descriptively but there are a lot of them; please open a pull request if you want help locating specific code.
@@
 
-    * Code and data objects needed to replicate the genetic mapping of flowering-related QTL in the pseudo-F2 cross.
+    - Code and data objects needed to replicate the genetic mapping of flowering-related QTL in the pseudo-F2 cross.
 
@@
 
-    * Contains completed drafts of a manuscript on the genetics of GxE in phenology in the switchgrass diversity panel.
+    - Contains completed drafts of a manuscript on the genetics of GxE in phenology in the switchgrass diversity panel.

### 2021-08-11 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Add Table 1 with summary proportions of effects with different types of GxE

@@
 
-# TRIAL script
+# Script used to make figures
 # -------------------------
-NB: Trial script on mash-env-cov branch to use weather covariance matrices as 
+NB: This script uses weather covariance matrices as 
 U_hyp in a mash run of GR50 phenotypes from 8 gardens.
@@
 ```{r load data}
-source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
-workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+source("../../../Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+workingdir <- file.path("..", "..", "..", "Github", "pvdiv-phenology-gxe")
 outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
@@
 ```{r}
-m_gulf_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_GR50_Gulf_Uhyp.rds")
-m_gulf_gr_no <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_GR50_Gulf_no_Uhyp.rds")
-m_gulf_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_FL50_Gulf_Uhyp.rds")
-m_gulf_fl_no <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_FL50_Gulf_no_Uhyp.rds")
-
-m_midw_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_24000_SNPs_GR50_Midwest_Uhyp.rds")
-m_midw_gr_no <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_24000_SNPs_GR50_Midwest_no_Uhyp.rds")  
-m_midw_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_33000_SNPs_FL50_Midwest_Uhyp.rds")
-m_midw_fl_no <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_33000_SNPs_FL50_Midwest_no_Uhyp.rds")
-
-m_gam_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_GR50_Gulf_and_Midwest_Uhyp.rds")
-m_gam_gr_no <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_GR50_Gulf_and_Midwest_no_Uhyp.rds")
-m_gam_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_FL50_Gulf_and_Midwest_Uhyp.rds")
-m_gam_fl_no <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_FL50_Gulf_and_Midwest_no_Uhyp.rds")
+m_gulf_gr_hyp <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_GR50_Gulf_Uhyp.rds")
+m_gulf_gr_no <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_GR50_Gulf_no_Uhyp.rds")
+m_gulf_fl_hyp <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_FL50_Gulf_Uhyp.rds")
+m_gulf_fl_no <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_FL50_Gulf_no_Uhyp.rds")
+
+m_midw_gr_hyp <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_24000_SNPs_GR50_Midwest_Uhyp.rds")
+m_midw_gr_no <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_24000_SNPs_GR50_Midwest_no_Uhyp.rds")  
+m_midw_fl_hyp <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_33000_SNPs_FL50_Midwest_Uhyp.rds")
+m_midw_fl_no <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_33000_SNPs_FL50_Midwest_no_Uhyp.rds")
+
+m_gam_gr_hyp <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_GR50_Gulf_and_Midwest_Uhyp.rds")
+m_gam_gr_no <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_GR50_Gulf_and_Midwest_no_Uhyp.rds")
+m_gam_fl_hyp <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_FL50_Gulf_and_Midwest_Uhyp.rds")
+m_gam_fl_no <- readRDS("../../../Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_FL50_Gulf_and_Midwest_no_Uhyp.rds")
 ```
@@
 `get_GxE` and `mash_plot_pairwise_sharing()` should do the trick here.
+
+```{r}
+site_regions <- sites %>% 
+  mutate(Region = case_when(manu_site %in% c("SD", "MO", "IL", "MI", "NE") ~ "North",
+                            manu_site %in% c("TX1", "TX2", "TX3", "TX4") ~ "Texas",
+                                             TRUE ~ "OK")) %>%
+  select(manu_site, Region) %>%
+  rename("Site" = "manu_site")
+
+gxe_gamg$S_1_row %>%
+  as_tibble(rownames = "Condition_1") %>%
+  pivot_longer(cols = 2:9, names_to = "Condition_2", values_to = "Sig") %>%
+  separate(Condition_1, into = c(NA, NA, NA, NA, "Phenotype", "Garden_1"), sep = "_") %>%
+  separate(Condition_2, into = c(NA, NA, NA, NA, NA, "Garden_2"), sep = "_") %>%
+  left_join(site_regions, by = c("Garden_1" = "Site")) %>%
+  rename("Region_1" = "Region") %>%
+  left_join(site_regions, by = c("Garden_2" = "Site")) %>%
+  rename("Region_2" = "Region") %>%
+  filter(Region_1 != "OK" & Region_2 !="OK") %>%
+  group_by(Region_1, Region_2) %>%
+  summarise(Total_sig = sum(Sig))
+
+gxe_gf$S_AP %>%
+  as_tibble(rownames = "Condition_1") %>%
+  pivot_longer(cols = 2:9, names_to = "Condition_2", values_to = "Sig") %>%
+  separate(Condition_1, into = c(NA, NA, "Phenotype", "Garden_1"), sep = "_") %>%
+  separate(Condition_2, into = c(NA, NA, NA, "Garden_2"), sep = "_") %>%
+  left_join(site_regions, by = c("Garden_1" = "Site")) %>%
+  rename("Region_1" = "Region") %>%
+  left_join(site_regions, by = c("Garden_2" = "Site")) %>%
+  rename("Region_2" = "Region") %>%
+  filter(Region_1 != "OK" & Region_2 !="OK") %>%
+  group_by(Region_1, Region_2) %>%
+  summarise(Total_sig = sum(Sig))
+
+gxe_mg$S_2_no %>%
+  as_tibble(rownames = "Condition_1") %>%
+  pivot_longer(cols = 2:8, names_to = "Condition_2", values_to = "Sig") %>%
+  separate(Condition_1, into = c(NA, NA, "Phenotype", "Garden_1"), sep = "_") %>%
+  separate(Condition_2, into = c(NA, NA, NA, "Garden_2"), sep = "_") %>%
+  left_join(site_regions, by = c("Garden_1" = "Site")) %>%
+  rename("Region_1" = "Region") %>%
+  left_join(site_regions, by = c("Garden_2" = "Site")) %>%
+  rename("Region_2" = "Region") %>%
+  filter(Region_1 != "OK" & Region_2 !="OK") %>%
+  group_by(Region_1, Region_2) %>%
+  summarise(Total_sig = sum(Sig))
+
+gxe_mf$S_AP %>%
+  as_tibble(rownames = "Condition_1") %>%
+  pivot_longer(cols = 2:7, names_to = "Condition_2", values_to = "Sig") %>%
+  separate(Condition_1, into = c(NA, NA, "Phenotype", "Garden_1"), sep = "_") %>%
+  separate(Condition_2, into = c(NA, NA, NA, "Garden_2"), sep = "_") %>%
+  left_join(site_regions, by = c("Garden_1" = "Site")) %>%
+  rename("Region_1" = "Region") %>%
+  left_join(site_regions, by = c("Garden_2" = "Site")) %>%
+  rename("Region_2" = "Region") %>%
+  filter(Region_1 != "OK" & Region_2 !="OK") %>%
+  group_by(Region_1, Region_2) %>%
+  summarise(Total_sig = sum(Sig))
+```

### 2022-01-09 R/Analysis_v1.1_covariance_comparisons.Rmd
Add weather and phenotypes for 2020-21 & cov comparison ability

---
title: "Analysis_v1.1_covariance_comparisons"
author: "Alice MacQueen"
date: 2022-01-06
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(rlist)
library(vegan)
library(geosphere)
```

#

Compare covariance matrices - phenotypic covariance matrices used as Uhyp, some genetic covariance matrices I made with ASReml, and potentially some new genetic covariance matrices that Li will make.

Load covariance matrices
```{r}
outputdir <- file.path(here("analysis", "gwas", "BLUPs", "single_sites"))
mashdir <- file.path(here("analysis", "mash"))

metadata <- read_rds(here("data", "metadata.rds"))
# pvdiv PLANT_ID metadata
sites <- read_rds(here("data", "sites.rds"))
# pvdiv common garden metadata
subpops <- read_rds(here("data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(here("data",
                           "Weather_related_phe_for_asreml_h2_GR50.rds"))
# all greenup related phenotypes, including greenup as functions of weather variables.
phe_fl_df <- read_rds(here("data",
                           "Weather_related_phe_for_asreml_h2_FL50.rds"))
# all flowering related phenotypes, including functions of weather variables.
k_full <- read_rds(here("data",                             "Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds"))
# kinship matrix for 630 tetraploid individuals.

## vectors with subsets of gardens, phenotypes, and genetic subpopulations.
site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
               NE = "NE", MI = "MI", SD = "SD")
subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", 
                 Gulf = "229g_10.3M")
subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest",
                  Gulf = "Gulf")
phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
                "crain_1d", "GR50", "cgdd_12c_10d")
phe_mult <- list(flowering_5 = c("FL50", "dyln_fl50", "cgdd_12c_gr2fl",
                                 "crain_gr2fl", "crain_1d"), 
                 greenup_2 = c("GR50", "cgdd_12c_10d"))
# paths to large SNP files
inputfiles <- read_delim(here("analysis", "gwas", "inputkinship.txt"),  
                         delim = " ", 
                         col_names = "SNPfiles")

## Define U_hyp: needs to change for Midwest, which drops gardens. -------------

U_hyp <- list()
# change based on phenotype ---------------
#phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", 
#             "tave_5d", "tave_10d", "tave_18d")
phe_hyp <- c("FL50", "GR50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")

for(i in seq_along(phe_hyp)){
  for(k in seq_along(subpop_v2)){

    cov_df <- phe_fl_df %>%
      ungroup() %>%
      left_join(select(sites, SITE, manu_site), by = "SITE") %>%
      filter(SUBPOP %in% subpop_v2[[k]]) %>%
      select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
      group_by(PLANT_ID, SUBPOP, manu_site) %>%
      mutate(IND = row_number()) %>%
      pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
      mutate(PLANT_ID = as.factor(PLANT_ID),
             IND = as.factor(IND)) %>%
      select(-IL) %>%
      select(PLANT_ID, SUBPOP, IND, MI, MO, NE, OK, SD, TX1, TX2, TX3) #####

    cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
    # Set diagonal of this matrix = coefficient of variation within each garden.
    diag(cor_phe) <- cov_df %>%
      ungroup() %>%
      summarise(MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),  #####
                MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
                NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
                OK = sqrt(var(OK, na.rm = TRUE))/mean(OK, na.rm = TRUE),
                SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
                TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
                TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
                TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE)) %>%
      as_vector()

    U_hyp <- list.prepend(U_hyp, cor_phe)
    names(U_hyp)[1] <- paste0(phe_hyp[i], "_", names(subpop_v2)[k])
  }
}

```

How similar are the flowering Uhyp covar matrices?
```{r}
site_dists <- distm(cbind(sites_for_dist$Longitude, sites_for_dist$Latitude), fun = distHaversine)
dist_sites <- as.dist(site_dists)

U_dist <- list()
for(i in seq_along(U_hyp)) {
  U_dist[[i]] <- as.dist(U_hyp[[i]])

}
names(U_dist) <- names(U_hyp)

mantel.rtest(dist_sites, dist_cgdd_12c_gr2fl_Gulf, nrepet = 500)

i = 1; j = 10
pval_mantel <- data.frame()
for(i in seq_along(U_dist)) {
  for (j in seq_along(U_dist)) {
  pval_mantel[i, j] <- mantel.rtest(U_dist[[i]], U_dist[[j]], nrepet = 1000)[5]   
  }
}
colnames(pval_mantel) <- names(U_dist)
pval_mantel %>%
  add_column(matrix_row = names(U_dist), .before = TRUE) %>%
  pivot_longer(cols = 2:28, names_to = "matrix_col", values_to = "pvalue") %>%
  #filter(matrix_row != matrix_col) %>%
  arrange(pvalue) %>%
  filter(matrix_col %in% c("dyln_fl50_Gulf_and_Midwest", "dyln_fl50_Gulf", "dyln_change_sec_Midwest", "cgdd_12c_gr2fl_Gulf", "cgdd_12c_gr2fl_Gulf_and_Midwest",
                            "cgdd_12c_gr2fl_Midwest", "GR50_Midwest", "GR50_Gulf", "GR50_Gulf_and_Midwest", "FL50_Gulf", "FL50_Midwest",
                           "FL50_Gulf_and_Midwest")) %>%
  filter(matrix_row %in% c("dyln_fl50_Gulf_and_Midwest", "dyln_fl50_Gulf", "dyln_change_sec_Midwest", "cgdd_12c_gr2fl_Gulf", "cgdd_12c_gr2fl_Gulf_and_Midwest",
                            "cgdd_12c_gr2fl_Midwest",  "GR50_Midwest", "GR50_Gulf", "GR50_Gulf_and_Midwest", "FL50_Gulf", "FL50_Midwest",
                           "FL50_Gulf_and_Midwest")) %>%
  ggplot(aes(x = matrix_col, y = matrix_row)) + 
  geom_tile(aes(fill = -log10(pvalue) > 2.5, color = -log10(pvalue)), size = 5, 
            width = 0.75, height = 0.75) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  scale_fill_viridis_d(end = 0.95, option = "B") +
  scale_color_viridis_c(end = 0.8, option = "D")

mantel.rtest(U_dist$cgdd_12c_gr2fl_Midwest, U_dist$crain_1d_Gulf)

sites$manu_site <- factor(sites$manu_site, levels = c("MI", "MO", "NE", "OK", "SD", "TX1", "TX2", "TX3"))

sites_for_dist <- sites %>%
  filter(!is.na(manu_site)) %>%
  arrange(manu_site)

-log10(.005)
.05/50
?geom_tile
```
MI MO NE OK SD TX1 TX2 TX3

In contrast, if the tested correlation matrix does not have a significant relationship with the geographic separation of the samples (Mantel statistic R: 0.138, p value = 0.052). In that case, as samples became physically more separated their corresponding correlation matrix do not necessarily become more dissimilar.

144 of 552 pairwise comparisons (of nonidentical matrices) were significant.
26%. A reasonable chunk.
```{r}
144/552
```

```{r}
pval_mantel %>%
  add_column(matrix_row = names(U_dist), .before = TRUE) %>%
  pivot_longer(cols = 2:28, names_to = "matrix_col", values_to = "pvalue") %>%
  #filter(matrix_row != matrix_col) %>%
  arrange(pvalue) %>%
  ggplot(aes(x = matrix_col, y = matrix_row)) + 
  geom_tile(aes(fill = -log10(pvalue) > 2.5, color = -log10(pvalue)), size = 4, 
            width = 0.75, height = 0.75) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  scale_fill_viridis_d(end = 0.95, option = "B") +
  scale_color_viridis_c(end = 0.8, option = "D")
```

Correlations for greenup matrices?

Distance correlations with flowering?

Three flowering matrices correlated with physical site distance: 
crain_3d_Gulf  0.01698302
crain_3d_Gulf_and_Midwest  0.02797203
crain_1d_Gulf  0.03996004
```{r}
mantel.rtest(dist_sites, dist_cgdd_12c_gr2fl_Gulf, nrepet = 500)

pval_mantel_dist <- c()
for(i in seq_along(U_dist)) {
  pval_mantel_dist[i] <- mantel.rtest(dist_sites, U_dist[[i]], nrepet = 1000)[5] }
names(pval_mantel_dist) <- names(U_dist)
```

## Distance correlations with greenup?

## Correlations between FL G & P matrices?

Need to make FL50 P correlation matrix
```{r}
U_dist$FL50_Gulf

cov_df <- var_table %>%
  filter(Phenotype == "FL50" & Subpop == "Gulf") %>%
  select(-SE, -Phenotype, -Subpop, -loglik) %>%
  pivot_wider(names_from = Garden_2, values_from = Estimate) %>%
  filter(Value == "r_G") %>%
  select(-Value) %>%
  add_row(Garden_1 = "SD") %>%
  column_to_rownames(var = "Garden_1") %>%
  add_column(TX1 = NA, .before = TRUE) %>%
  as.matrix()
diag(cov_df)
h2_diag_mean <- var_table %>%
  filter(Phenotype == "FL50" & Subpop == "Gulf") %>%
  filter(Value %in% c("h2_garden1", "h2_garden2")) %>%
  select(-SE, -Phenotype, -Subpop, -loglik)  %>%
  pivot_longer(cols = Garden_1:Garden_2, names_to = "GardenNo", values_to = "Site") %>%
  filter(!(Value == "h2_garden1" & GardenNo == "Garden_2") & !(Value == "h2_garden2" & GardenNo == "Garden_1") ) %>%
  group_by(Site) %>%
  summarize(h2 = mean(Estimate)) %>%
  arrange(desc(Site))

diag(cov_df) <- h2_diag_mean$h2

makeSymmU <- function(m) {
   m[upper.tri(m)] <- t(m)[upper.tri(m)]
   return(m)
}
makeSymmL <- function(m) {
   m[lower.tri(m)] <- t(m)[lower.tri(m)]
   return(m)
}

cov_mat <- makeSymmL(cov_df)
cov_df <- as.data.frame(cov_mat) %>%
  rownames_to_column() %>%
  select(rowname, MI, MO, NE, OK, SD, TX1, TX2, TX3)
cov_df$rowname <- factor(cov_df$rowname, levels = c("MI", "MO", "NE", "OK", 
                                                    "SD", "TX1", "TX2", "TX3"))
cov_mat <- cov_df %>%
  arrange(rowname) %>%
  column_to_rownames() %>%
  as.matrix()

cov_dist <- as.dist(cov_mat)
U_dist$FL50_Gulf
cov_dist
mantel.rtest(cov_dist, U_dist$FL50_Gulf, nrepet = 100000)
```
G & P for FL50 have p-value of 9.9999e-05 (minimum for 100K replicates, I believe). There is a relationship between the two matrices. Well, good, I should hope so.

## Correlations between GR G & P matrices?
```{r}

var_table <- read_csv(here("analysis", "genetic-covariances",
                           "Two_site_var_table_up_to_300_iterations.csv"))
var_table$Garden_2 <- factor(var_table$Garden_2, levels = c("SD", "MI", "NE", "MO", "OK", "TX3", "TX2"))
var_table$Garden_1 <- factor(var_table$Garden_1, levels = c("SD", "MI", "NE", "MO", "OK", "TX3", "TX2", "TX1"))

var_table %>%
  ggplot(aes(x = Garden_1, y = Garden_2)) + 
  geom_raster(aes(fill = Estimate)) + 
  facet_grid(rows = vars(Phenotype), cols = vars(Subpop)) +
      scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1))

```

Basically can only meaningfully compare Gulf FL50 & GR50 at the moment. Li may have 7 more to share with me soon.

Gotta go through this process to get the matrices to be comparable.
```{r}
U_dist$GR50_Gulf

cov_df <- var_table %>%
  filter(Phenotype == "GR50" & Subpop == "Gulf") %>%
  select(-SE, -Phenotype, -Subpop, -loglik) %>%
  pivot_wider(names_from = Garden_2, values_from = Estimate) %>%
  filter(Value == "r_G") %>%
  select(-Value) %>%
  add_row(Garden_1 = "SD") %>%
  column_to_rownames(var = "Garden_1") %>%
  add_column(TX1 = NA, .before = TRUE) %>%
  as.matrix()
diag(cov_df)
h2_diag_mean <- var_table %>%
  filter(Phenotype == "GR50" & Subpop == "Gulf") %>%
  filter(Value %in% c("h2_garden1", "h2_garden2")) %>%
  select(-SE, -Phenotype, -Subpop, -loglik)  %>%
  pivot_longer(cols = Garden_1:Garden_2, names_to = "GardenNo", values_to = "Site") %>%
  filter(!(Value == "h2_garden1" & GardenNo == "Garden_2") & !(Value == "h2_garden2" & GardenNo == "Garden_1") ) %>%
  group_by(Site) %>%
  summarize(h2 = mean(Estimate)) %>%
  arrange(desc(Site))

diag(cov_df) <- h2_diag_mean$h2

makeSymmU <- function(m) {
   m[upper.tri(m)] <- t(m)[upper.tri(m)]
   return(m)
}
makeSymmL <- function(m) {
   m[lower.tri(m)] <- t(m)[lower.tri(m)]
   return(m)
}

cov_mat <- makeSymmL(cov_df)
cov_df <- as.data.frame(cov_mat) %>%
  rownames_to_column() %>%
  select(rowname, MI, MO, NE, OK, SD, TX1, TX2, TX3)
cov_df$rowname <- factor(cov_df$rowname, levels = c("MI", "MO", "NE", "OK", 
                                                    "SD", "TX1", "TX2", "TX3"))
cov_mat <- cov_df %>%
  arrange(rowname) %>%
  column_to_rownames() %>%
  as.matrix()

cov_dist <- as.dist(cov_mat)
U_dist$GR50_Gulf
cov_dist
mantel.rtest(cov_dist, U_dist$GR50_Gulf, nrepet = 100000)
```

G & P are pretty trivially not the same here. No relationship between these matrices. Too bad. p = 0.51.
Actually, belay that, I was comparing GR50 Gulf P & FL50 Gulf G! I would hope those are different.

G & P for GR50 have p-value of 0.00074 (100K replicates). There is a relationship between the two matrices. Well, good, I should hope so. Both I would count as a significant relationship by my criteria for cov matrices earlier, that is, the somewhat arbitrary/by eye cutoff of -log10p > 2.5 (p < .00316.)
```{r}
74/100000
-log10(.00073999)
-log10(9.99e-05)
-log10(.003162)
```
Mantel test, from Wikipedia:
In contrast to the ordinary use of the correlation coefficient, to assess significance of any apparent departure from a zero correlation, the rows and columns of one of the matrices are subjected to random permutations many times, with the correlation being recalculated after each permutation. The significance of the observed correlation is the proportion of such permutations that lead to a higher correlation coefficient.

The reasoning is that if the null hypothesis of there being no relation between the two matrices is true, then permuting the rows and columns of the matrix should be equally likely to produce a larger or a smaller coefficient. In addition to overcoming the problems arising from the statistical dependence of elements within each of the two matrices, use of the permutation test means that no reliance is being placed on assumptions about the statistical distributions of elements in the matrices.

### 2022-01-09 R/Analysis_v1.1_weather_2020.Rmd
Add weather and phenotypes for 2020-21 & cov comparison ability

---
title: "Analysis_v1.1_weather"
author: "Alice MacQueen"
date: "1/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(zoo)
```

#

```{r}
source("~/Github/pv-weather/R/get_functions.R")
sites <- read_rds(here("data", "sites.rds"))
metadata <- read_rds(here("data", "metadata.rds"))
mesonet_stations <- read_csv("~/Github/pv-weather/data-raw/All_Mesonet_Stations.csv")

get_station_info <- function(df, decdegrees = 0.1, loc_key = "Location_code"){
  loc_key <- sym(loc_key)
  locs <- list()
  Location_finder <- df %>%
    group_by(!! loc_key) %>%
    summarize(Lat1 = mean(Latitude, na.rm = TRUE) - decdegrees,
              Long1 = mean(Longitude, na.rm = TRUE) - decdegrees,
              Lat2 = Lat1 + decdegrees*2,
              Long2 = Long1 + decdegrees*2)
  for(i in 1:nrow(df)){
    locs[[i]] <- mesonet_stations %>%
      filter(between(lon, Location_finder$Long1[i], Location_finder$Long2[i]) &
               between(lat, Location_finder$Lat1[i], Location_finder$Lat2[i])) %>%
      arrange(begints)
  }
  station_info <- do.call(rbind.data.frame, locs)
  return(station_info)
}

Station1 <- metadata %>%
  filter(PLANT_ID == "J001.A") %>%
  rename(Latitude = LATITUDE, Longitude = LONGITUDE) %>%
  mutate(Longitude = -Longitude)
get_station_info(Station1, loc_key = "PLANT_ID")
Station1$station <- "BKV"
J001_2018 <- get_mesonet_weather(station = Station1$station, year1 = "2018", 
                                 month1 = "06", day1 = "01", year2 = "2021", 
                                 month2 = "11", day2 = "01", 
                                 tz = "America/New_York")
```

SD stid: SMCS2 (or MDYS2)
MO stid: MO027
IL stid: WSCI2 (don't need this)
MI stid: KBSM4 (or GLLM4)
NE stid: MPHN1

```{r}
meso_url <- "https://mesonet.agron.iastate.edu/cgi-bin/request/asos.py/"
library(httr)
library(data.table)
library(lubridate)
#' @import httr
#' @import data.table
#' @import dplyr
#' @import lubridate
#'
#' @export
get_mesonet_weather <- function(station, year1, month1, day1, year2, month2, day2, tz){
  MESO_WEA <- GET(url = meso_url,
                  query = list(station = station,
                               data = "all",
                               year1 = year1,
                               month1 = month1,
                               day1 = day1,
                               year2 = year2,
                               month2 = month2,
                               day2 = day2,
                               tz = tz,
                               format = "comma")) %>%
    content() %>%
    read_csv(skip = 5, na = "M") %>%
    mutate(Working_date = as_date(valid)) %>%
    group_by(Working_date) %>%
    summarise(tmin = (min(as.numeric(tmpf), na.rm = TRUE)-32)*(5/9), # convert F to C for all temperatures
              tmax = (max(as.numeric(tmpf), na.rm = TRUE)-32)*(5/9),
              RHmean = mean(as.numeric(relh), na.rm = TRUE),
              dewp = (mean(as.numeric(dwpf), na.rm = TRUE)-32)*(5/9),
              wndk = mean(as.numeric(sknt), na.rm = TRUE),
              prcp = (sum(as.numeric(p01i), na.rm = TRUE)*254), # convert inches to 0.1x mm
              Tdrnl = mean(tmax - tmin),
              Tmean = (mean(as.numeric(tmpf), na.rm = TRUE)-32)*(5/9)) %>%
    mutate_if(is.double, funs(na_if(., NaN)))
  DT <- data.table(MESO_WEA)
  invisible(lapply(names(DT),function(.name) set(DT,
                                                 which(is.infinite(DT[[.name]])),
                                                 j = .name,value =NA)))
  invisible(lapply(names(DT),function(.name) set(DT, which(is.nan(DT[[.name]])),
                                                 j = .name,value =NA)))
  return(DT)
}

get_mesonet_weather(station = "WSCI2", year = "2019", month1 = "01", day1 = "01", year2 = "2021", month2 = "10", day2 = "31", tz = "America/Chicago")
```
Only the ASOS stations seem to work. Still, nothing is farther than 0.18 decimal degrees away from the field site. Given that the techs haven't added weather data for the site proper, that's just going to have to be good enough.
```{r}
sites[1,]
get_station_info(sites[1,], decdegrees = 0.15, loc_key = "manu_site")
Station1$station <- "BKX"
test_SD <- get_mesonet_weather(station = Station1$station, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

sites[2,]
get_station_info(sites[2,], decdegrees = 0.1, loc_key = "manu_site")
Station2 <- "COU"
test_MO <- get_mesonet_weather(station = Station2, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

sites[3,]
get_station_info(sites[3,], decdegrees = 0.1, loc_key = "manu_site")
Station3 <- "DPA"
test_IL <- get_mesonet_weather(station = Station3, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

sites[4,]
get_station_info(sites[4,], decdegrees = 0.15, loc_key = "manu_site")
Station4 <- "BTL"
test_MI <- get_mesonet_weather(station = Station4, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

sites[5,]
get_station_info(sites[5,], decdegrees = 0.1, loc_key = "manu_site")
Station5 <- "NQI"
test_TX1 <- get_mesonet_weather(station = Station5, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

sites[6,]
get_station_info(sites[6,], decdegrees = 0.18, loc_key = "manu_site")
Station6 <- "AHQ"
test_NE <- get_mesonet_weather(station = Station6, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

sites[7,]
get_station_info(sites[7,], decdegrees = 0.18, loc_key = "manu_site")
Station7 <- "RFI"
test_TX4 <- get_mesonet_weather(station = Station7, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

sites[8,]
get_station_info(sites[8,], decdegrees = 0.1, loc_key = "manu_site")
Station8 <- "ATT"
test_TX2 <- get_mesonet_weather(station = Station8, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

sites[9,]
get_station_info(sites[9,], decdegrees = 0.18, loc_key = "manu_site")
Station9 <- "SWO"
test_OK <- get_mesonet_weather(station = Station9, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

sites[10,]
get_station_info(sites[10,], decdegrees = 0.15, loc_key = "manu_site")
Station10 <- "TPL"
test_TX3 <- get_mesonet_weather(station = Station10, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

```

```{r}
wea_MI <- test_MI %>%
  mutate(manu_site = "MI") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))
wea_SD <- test_SD %>%
  mutate(manu_site = "SD") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))
wea_MO <- test_MO %>%
  mutate(manu_site = "MO") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))
wea_TX1 <- test_TX1 %>%
  mutate(manu_site = "TX1") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))
wea_TX2 <- test_TX2 %>%
  mutate(manu_site = "TX2") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))
wea_TX3 <- test_TX3 %>%
  mutate(manu_site = "TX3") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))
wea_OK <- test_OK %>%
  mutate(manu_site = "OK") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))
wea_NE <- test_NE %>%
  mutate(manu_site = "NE") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))
wea_TX4 <- test_TX4 %>%
  mutate(manu_site = "TX4") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))
wea_IL <- test_IL %>%
  mutate(manu_site = "IL") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))

wea_ten_sites <- wea_MI %>%
  full_join(wea_SD) %>%
  full_join(wea_MO) %>%
  full_join(wea_TX1) %>%
  full_join(wea_TX2) %>%
  full_join(wea_TX3) %>%
  full_join(wea_OK) %>%
  full_join(wea_NE) %>%
  full_join(wea_TX4) %>%
  full_join(wea_IL)

##function to calculate daylength

day.length.hrs <- function(day, latitude) {

  daylength.coeff <- asin(0.39795*cos(0.2163108 + 2*atan(0.9671396*tan(0.00860*(day - 186)))))

  daylength.hrs <- 24 - (24/pi)*acos((sin(0.8333*pi/180) + sin(latitude*pi/180)*sin(daylength.coeff))/(cos(latitude*pi/180)*cos(daylength.coeff)))

  return(daylength.hrs)

}

```

```{r}
wea_ten_sites

phenotypes <- readRDS("~/Github/pvdiv-phenotypes/data/GWAS/Phenotypes_gwas_all_subpops_2018_to_2021.rds")

phe_ten_sites <- phenotypes %>%
  select(PLANT_ID, matches("GR50"), matches("FL50")) %>%
  pivot_longer(cols = -PLANT_ID, names_to = "SITE_YEAR_PHE", values_to = "JDAY") %>%
  separate(SITE_YEAR_PHE, into = c("SITE", "Year", "PHE"), sep = "_", convert = TRUE) %>%
  filter(!is.na(JDAY)) %>%
  left_join(select(sites, SITE, manu_site))

phe_ten_sites %>%
  write_rds(file = here("data", "Phenotypes_2018_to_2021_10_sites.rds"))
wea_ten_sites %>%
  write_rds(file = here("data", "Daily_weather_2018_to_2021_10_sites.rds"))

```

Can make most weather variables before I join the weather dataframe to the phenotype dataframe. Can make most of these with mutate and the lag function, even if it's a bit ugly looking in terms of code.

Greenup
#phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", 
#             "tave_5d", "tave_10d", "tave_18d")
```{r}
colSums(is.na(wea_ten_sites))

wea_ten_sites <- wea_ten_sites %>%
  mutate(JDAY = yday(Working_date),
         DYLN = day.length.hrs(day = JDAY, latitude = Latitude),
         tmax_sm = ifelse(is.na(tmax),
                          (na.locf(tmax) + na.locf(tmax, fromLast = TRUE))/2,
                          tmax),
         tmin_sm = ifelse(is.na(tmin),
                          (na.locf(tmin) + na.locf(tmin, fromLast = TRUE))/2,
                          tmin),
         tmean_sm = ifelse(is.na(Tmean),
                          (na.locf(Tmean) + na.locf(Tmean, fromLast = TRUE))/2,
                          Tmean),
         GDD_12Cbase = ifelse((tmax_sm + tmin_sm)/2 >= 12.0,
                              (tmax_sm + tmin_sm)/2 - 10,
                              0),
         Year = year(Working_date)) %>%
  arrange(Latitude, Year, JDAY) %>%
  group_by(manu_site) %>%
  mutate(crain_2d = prcp + lag(prcp, default = 0),
         crain_3d = prcp + lag(prcp, default = 0) + lag(prcp, default = 0, 
                                                        n = 2),
         crain_5d = prcp + lag(prcp, default = 0) + 
           lag(prcp, default = 0, n = 2) + lag(prcp, default = 0, n = 3) +
           lag(prcp, default = 0, n = 4),
         dyln_change_sec = (DYLN - lag(DYLN))*60*60,
         dyln_change_sec_7d_prior = (lag(DYLN, n = 6) - 
                                       lag(DYLN, n = 7))*60*60,
         dyln_change_sec_14d_prior = (lag(DYLN, n = 13) - 
                                       lag(DYLN, n = 14))*60*60,
         dyln_7d_prior = lag(DYLN, n = 6),
         dyln_14d_prior = lag(DYLN, n = 13),
         cgdd_12c_5d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4))/5,
         tave_5d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4))/5,
         cgdd_12c_10d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5) +
           lag(GDD_12Cbase, default = 0, n = 6) +
           lag(GDD_12Cbase, default = 0, n = 7) +
           lag(GDD_12Cbase, default = 0, n = 8) +
           lag(GDD_12Cbase, default = 0, n = 9))/10,
         tave_10d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5) +
           lag(tmean_sm, default = 0, n = 6) +
           lag(tmean_sm, default = 0, n = 7) +
           lag(tmean_sm, default = 0, n = 8) +
           lag(tmean_sm, default = 0, n = 9))/10,
         cgdd_12c_20d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5) +
           lag(GDD_12Cbase, default = 0, n = 6) +
           lag(GDD_12Cbase, default = 0, n = 7) +
           lag(GDD_12Cbase, default = 0, n = 8) +
           lag(GDD_12Cbase, default = 0, n = 9) +
           lag(GDD_12Cbase, default = 0, n = 10) +
           lag(GDD_12Cbase, default = 0, n = 11) +
           lag(GDD_12Cbase, default = 0, n = 12) +
           lag(GDD_12Cbase, default = 0, n = 13) +
           lag(GDD_12Cbase, default = 0, n = 14) +
           lag(GDD_12Cbase, default = 0, n = 15) +
           lag(GDD_12Cbase, default = 0, n = 16) +
           lag(GDD_12Cbase, default = 0, n = 17) +
           lag(GDD_12Cbase, default = 0, n = 18) +
           lag(GDD_12Cbase, default = 0, n = 19))/20,
         tave_20d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5) +
           lag(tmean_sm, default = 0, n = 6) +
           lag(tmean_sm, default = 0, n = 7) +
           lag(tmean_sm, default = 0, n = 8) +
           lag(tmean_sm, default = 0, n = 9) +
           lag(tmean_sm, default = 0, n = 10) +
           lag(tmean_sm, default = 0, n = 11) +
           lag(tmean_sm, default = 0, n = 12) +
           lag(tmean_sm, default = 0, n = 13) +
           lag(tmean_sm, default = 0, n = 14) +
           lag(tmean_sm, default = 0, n = 15) +
           lag(tmean_sm, default = 0, n = 16) +
           lag(tmean_sm, default = 0, n = 17) +
           lag(tmean_sm, default = 0, n = 18) +
           lag(tmean_sm, default = 0, n = 19))/20)

```
Flowering
phe_hyp <- c("FL50", "GR50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")

If I join the current wea_ten_sites to the greenup and flowering df, that should leave me with variables I want just for the dates & greenup & flowering.
Including some new daylength ones I made which will be 1-2 weeks before flowering.

The greenup to flowering variables will need to be made in a separate step. 
These are cgdd_12c_gr2fl and crain_gr2fl. This cumulative sum is different for each year, site, and plant id.
```{r}

Tbase <- 12

phe_gr_fl <- phe_ten_sites %>%
  mutate(JDAY_EXP = case_when(Year == 2018 ~ JDAY,
                              Year == 2019 ~ JDAY + 365,
                              Year == 2020 ~ JDAY + (365*2),
                              Year == 2021 ~ JDAY + (365*3),
                              TRUE ~ NA_real_)) %>%
  select(-JDAY) %>%
  pivot_wider(names_from = PHE, values_from = JDAY_EXP)

GR_start <- phe_gr_fl %>% 
  group_by(manu_site, Year) %>%
  filter(!is.na(GR50)) %>%
  summarise(GR_start = floor(min(GR50)))
FL_end <- phe_gr_fl %>% 
  group_by(manu_site, Year) %>%
  filter(!is.na(FL50)) %>%
  summarise(FL_end = ceiling(max(FL50)))

tmp1 <- wea_ten_sites %>%
  mutate(JDAY_EXP = case_when(Year == 2018 ~ JDAY,
                              Year == 2019 ~ JDAY + 365,
                              Year == 2020 ~ JDAY + 365 + 365,
                              Year == 2021 ~ JDAY + (365*3),
                              TRUE ~ NA_real_)) %>%
  left_join(phe_gr_fl)

tmp <- tmp1 %>%
  group_by(manu_site, PLANT_ID, Year, JDAY_EXP) %>%
  filter(between(JDAY_EXP, floor(GR50), ceiling(FL50))) %>%
  ungroup %>%
  group_by(manu_site, Year, PLANT_ID) %>%
  summarize_at(vars(GDD_12Cbase, prcp), funs(sum))

tmp %>%
  filter(manu_site == "TX1")
tmp
```

```{r}
phe_ten_sites %>%
  left_join(wea_ten_sites)
```

### 2022-01-09 R/Analysis_v1.1_weather_2020.Rmd
Add hypothesis matrices using weather data from 3 years

@@
 library(zoo)
+library(cowplot)
 ```
@@
          Year = year(Working_date)) %>%
-  arrange(Latitude, Year, JDAY) %>%
+  mutate(JDAY_EXP = case_when(Year == 2018 ~ JDAY,
+                              Year == 2019 ~ JDAY + 365,
+                              Year == 2020 ~ JDAY + (365*2),
+                              Year == 2021 ~ JDAY + (365*3),
+                              TRUE ~ NA_real_)) %>%
+  arrange(Latitude, Year, JDAY_EXP) %>%
+  select(-JDAY) %>%
   group_by(manu_site) %>%
@@
 
-GR_start <- phe_gr_fl %>% 
-  group_by(manu_site, Year) %>%
-  filter(!is.na(GR50)) %>%
-  summarise(GR_start = floor(min(GR50)))
-FL_end <- phe_gr_fl %>% 
-  group_by(manu_site, Year) %>%
-  filter(!is.na(FL50)) %>%
-  summarise(FL_end = ceiling(max(FL50)))
-
 tmp1 <- wea_ten_sites %>%
-  mutate(JDAY_EXP = case_when(Year == 2018 ~ JDAY,
-                              Year == 2019 ~ JDAY + 365,
-                              Year == 2020 ~ JDAY + 365 + 365,
-                              Year == 2021 ~ JDAY + (365*3),
-                              TRUE ~ NA_real_)) %>%
   left_join(phe_gr_fl)
@@
   filter(manu_site == "TX1")
-tmp
+tmp %>%
+  filter(!is.na(PLANT_ID))
+
 ```
@@
 ```{r}
-phe_ten_sites %>%
-  left_join(wea_ten_sites)
+cphe_gr2fl <- tmp %>%
+  filter(!is.na(PLANT_ID)) %>%
+  rename(cgdd_12c_gr2fl = GDD_12Cbase, crain_gr2fl = prcp)
+
+wea_ten_sites %>%
+  filter(manu_site == "TX1" & JDAY_EXP == 715)
+
+tmp2 <- wea_ten_sites %>%
+  select(-Year)
+
+gr_ten_sites <- phe_gr_fl %>%
+  mutate(GR50 = floor(GR50)) %>%
+  filter(!is.na(GR50)) %>%
+  left_join(tmp2, by = c("manu_site", "GR50" = "JDAY_EXP")) %>%
+  select(PLANT_ID, Year, manu_site, GR50, cgdd_12c_5d:tave_20d)
+
+fl_ten_sites <- phe_gr_fl %>%
+  filter(!is.na(FL50)) %>%
+  mutate(FL50 = floor(FL50),
+         GR50 = floor(GR50)) %>%
+  left_join(tmp2, by = c("manu_site", "FL50" = "JDAY_EXP")) %>%
+  select(PLANT_ID, Year, manu_site, GR50, FL50, prcp, DYLN, crain_2d:dyln_14d_prior) %>%
+  left_join(cphe_gr2fl) %>%
+  rename(crain_1d = prcp, dyln_fl50 = DYLN)
+# 11 September flowering dates in 2021 at Pickle have NA's for weather values. Decided to let this be.
+
+1354-(365*3)
+tmp2 %>%
+  filter(manu_site == "TX2" & JDAY_EXP > 1353)
 ```
+#phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", 
+#             "tave_5d", "tave_10d", "tave_18d")
+phe_hyp <- c("FL50", "GR50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
+             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")
+
+```{r}
+gr_ten_sites %>%
+  saveRDS(file = here("data", "Weather_related_phe_GR50_3yr.rds"))
+fl_ten_sites %>%
+  saveRDS(file = here("data", "Weather_related_phe_FL50_3yr.rds"))
+```
+
+Make phenotypic correlation matrices for these new years, and new weather variables.
+
+```{r}
+metadata <- read_rds(here("data", "metadata.rds"))
+subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest",
+                  Gulf = "Gulf")
+year_v = c(2019, 2020, 2021)
+
+fl_ten_sites <- fl_ten_sites %>%
+  left_join(select(metadata, PLANT_ID, SUBPOP))
+gr_ten_sites <- gr_ten_sites %>%
+  left_join(select(metadata, PLANT_ID, SUBPOP))
+
+# change based on phenotype ---------------
+#phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_20d", "cgdd_12c_5d", 
+#             "tave_5d", "tave_10d", "tave_20d")
+phe_hyp <- c("FL50", "GR50", "dyln_fl50", "dyln_change_sec", 
+             "dyln_change_sec_7d_prior", "dyln_change_sec_14d_prior", 
+             "dyln_7d_prior", "dyln_14d_prior", "cgdd_12c_gr2fl", 
+             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")
+i=1
+k=1
+
+U_hyp_2019 <- list()
+for(i in seq_along(phe_hyp)){
+  for(k in seq_along(subpop_v2)){
+
+    cov_df <- fl_ten_sites %>%
+      ungroup() %>%
+      filter(SUBPOP %in% subpop_v2[[k]]) %>%
+      filter(Year %in% 2019) %>%
+      select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
+      group_by(PLANT_ID, SUBPOP, manu_site) %>%
+      mutate(IND = row_number()) %>%
+      pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
+      mutate(PLANT_ID = as.factor(PLANT_ID),
+             IND = as.factor(IND)) %>%
+      select(PLANT_ID, SUBPOP, IND, IL, MI, MO, NE, SD, TX1, TX2, TX3) #####
+
+    cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
+    # Set diagonal of this matrix = coefficient of variation within each garden.
+    diag(cor_phe) <- cov_df %>%
+      ungroup() %>%
+      summarise(IL = sqrt(var(IL, na.rm = TRUE))/mean(IL, na.rm = TRUE),  #####
+                MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),  #####
+                MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
+                NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
+                SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
+                TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
+                TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
+                TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE)) %>%
+      as_vector()
+
+    U_hyp_2019 <- list.prepend(U_hyp_2019, cor_phe)
+    names(U_hyp_2019)[1] <- paste0(phe_hyp[i], "_", names(subpop_v2)[k])
+  }
+}
+
+# for 2020
+U_hyp_2020 <- list()
+for(i in seq_along(phe_hyp)){
+  for(k in seq_along(subpop_v2)){
+
+    cov_df <- fl_ten_sites %>%
+      ungroup() %>%
+      filter(SUBPOP %in% subpop_v2[[k]]) %>%
+      filter(Year %in% 2020) %>%
+      select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
+      group_by(PLANT_ID, SUBPOP, manu_site) %>%
+      mutate(IND = row_number()) %>%
+      pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
+      mutate(PLANT_ID = as.factor(PLANT_ID),
+             IND = as.factor(IND)) %>%
+      select(PLANT_ID, SUBPOP, IND, IL, MI, MO, NE, SD, TX1, TX2, TX3, TX4) #####
+
+    cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
+    # Set diagonal of this matrix = coefficient of variation within each garden.
+    diag(cor_phe) <- cov_df %>%
+      ungroup() %>%
+      summarise(IL = sqrt(var(IL, na.rm = TRUE))/mean(IL, na.rm = TRUE),  #####
+                MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),  #####
+                MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
+                NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
+                SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
+                TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
+                TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
+                TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE),
+                TX4 = sqrt(var(TX4, na.rm = TRUE))/mean(TX4, na.rm = TRUE)) %>%
+      as_vector()
+
+    U_hyp_2020 <- list.prepend(U_hyp_2020, cor_phe)
+    names(U_hyp_2020)[1] <- paste0(phe_hyp[i], "_", names(subpop_v2)[k])
+  }
+}
+
+U_hyp_2021 <- list()
+for(i in seq_along(phe_hyp)){
+  for(k in seq_along(subpop_v2)){
+
+    cov_df <- fl_ten_sites %>%
+      ungroup() %>%
+      filter(SUBPOP %in% subpop_v2[[k]]) %>%
+      filter(Year %in% 2021) %>%
+      select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
+      group_by(PLANT_ID, SUBPOP, manu_site) %>%
+      mutate(IND = row_number()) %>%
+      pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
+      mutate(PLANT_ID = as.factor(PLANT_ID),
+             IND = as.factor(IND)) %>%
+      select(PLANT_ID, SUBPOP, IND, IL, MI, MO, TX2, TX3, TX4) #####
+
+    cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
+    # Set diagonal of this matrix = coefficient of variation within each garden.
+    diag(cor_phe) <- cov_df %>%
+      ungroup() %>%
+      summarise(IL = sqrt(var(IL, na.rm = TRUE))/mean(IL, na.rm = TRUE),  #####
+                MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),  #####
+                MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
+                TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
+                TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE),
+                TX4 = sqrt(var(TX4, na.rm = TRUE))/mean(TX4, na.rm = TRUE)) %>%
+      as_vector()
+
+    U_hyp_2021 <- list.prepend(U_hyp_2021, cor_phe)
+    names(U_hyp_2021)[1] <- paste0(phe_hyp[i], "_", names(subpop_v2)[k])
+  }
+}
+
+```
+
+Visualize correlation matrices themselves
+```{r}
+library(switchgrassGWAS)
+
+i=7
+switchgrassGWAS::mash_plot_pairwise_sharing(corrmatrix = U_hyp_2019[[i]],
+                                            reorder = FALSE)
+switchgrassGWAS::mash_plot_pairwise_sharing(corrmatrix = U_hyp_2020[[i]],
+                                            reorder = FALSE)
+switchgrassGWAS::mash_plot_pairwise_sharing(corrmatrix = U_hyp_2021[[i]],
+                                            reorder = FALSE)
+```
+
+How related are correlation matrices? Mantel tests & visualization of pairwise tests.
+```{r}
+for(i in seq_along(U_hyp_2019)) {
+  U_hyp_2019[[i]] <- replace_na(U_hyp_2019[[i]], 0)
+}
+
+U_dist_2019 <- list()
+for(i in seq_along(U_hyp_2019)) {
+  U_dist_2019[[i]] <- as.dist(U_hyp_2019[[i]])
+
+}
+names(U_dist_2019) <- names(U_hyp_2019)
+
+pval_mantel_2019 <- data.frame()
+for(i in seq_along(U_dist_2019)) {
+  for (j in seq_along(U_dist_2019)) {
+      pval_mantel_2019[i, j] <- mantel.rtest(U_dist_2019[[i]], U_dist_2019[[j]], nrepet = 1000)[5] 
+  }
+}
+colnames(pval_mantel_2019) <- names(U_dist_2019)
+```
+
+```{r}
+plot_mantel_2019 <- pval_mantel_2019 %>%
+  add_column(matrix_row = names(U_dist_2019), .before = TRUE) %>%
+  pivot_longer(cols = 2:ncol(pval_mantel_2019)+1, names_to = "matrix_col", values_to = "pvalue") %>%
+  #filter(matrix_row != matrix_col) %>%
+  arrange(pvalue) %>%
+  ggplot(aes(x = matrix_col, y = matrix_row)) + 
+  geom_tile(aes(fill = -log10(pvalue) > 2.5, color = -log10(pvalue)), 
+            size = 3.5, 
+            width = 0.7, height = 0.7) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
+  scale_fill_viridis_d(end = 0.95, option = "B") +
+  scale_color_viridis_c(end = 0.8, option = "D")
+
+save_plot(plot = plot_mantel_2019, filename = here("analysis", "hypothesis_matrices", "Weather_Uhyp_Mantel_pvalue_plot_2019.png"), base_height = 13, base_width = 15)
+```
+
+```{r}
+for(i in seq_along(U_hyp_2020)) {
+  U_hyp_2020[[i]] <- replace_na(U_hyp_2020[[i]], 0)
+}
+
+U_dist_2020 <- list()
+for(i in seq_along(U_hyp_2020)) {
+  U_dist_2020[[i]] <- as.dist(U_hyp_2020[[i]])
+
+}
+names(U_dist_2020) <- names(U_hyp_2020)
+
+pval_mantel_2020 <- data.frame()
+for(i in seq_along(U_dist_2020)) {
+  for (j in seq_along(U_dist_2020)) {
+      pval_mantel_2020[i, j] <- mantel.rtest(U_dist_2020[[i]], U_dist_2020[[j]], nrepet = 1000)[5] 
+  }
+}
+colnames(pval_mantel_2020) <- names(U_dist_2020)
+```
+
+```{r}
+plot_mantel_2020 <- pval_mantel_2020 %>%
+  add_column(matrix_row = names(U_dist_2020), .before = TRUE) %>%
+  pivot_longer(cols = 2:ncol(pval_mantel_2020)+1, names_to = "matrix_col", values_to = "pvalue") %>%
+  #filter(matrix_row != matrix_col) %>%
+  arrange(pvalue) %>%
+  ggplot(aes(x = matrix_col, y = matrix_row)) + 
+  geom_tile(aes(fill = -log10(pvalue) > 2.5, color = -log10(pvalue)), 
+            size = 3.5, 
+            width = 0.7, height = 0.7) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
+  scale_fill_viridis_d(end = 0.95, option = "B") +
+  scale_color_viridis_c(end = 0.8, option = "D")
+
+save_plot(plot = plot_mantel_2020, filename = here("analysis", "hypothesis_matrices", "Weather_Uhyp_Mantel_pvalue_plot_2020.png"), base_height = 13, base_width = 15)
+```
+
+```{r}
+for(i in seq_along(U_hyp_2021)) {
+  U_hyp_2021[[i]] <- replace_na(U_hyp_2021[[i]], 0)
+}
+
+U_dist_2021 <- list()
+for(i in seq_along(U_hyp_2021)) {
+  U_dist_2021[[i]] <- as.dist(U_hyp_2021[[i]])
+
+}
+names(U_dist_2021) <- names(U_hyp_2021)
+
+pval_mantel_2021 <- data.frame()
+for(i in seq_along(U_dist_2021)) {
+  for (j in seq_along(U_dist_2021)) {
+      pval_mantel_2021[i, j] <- mantel.rtest(U_dist_2021[[i]], U_dist_2021[[j]], nrepet = 1000)[5] 
+  }
+}
+colnames(pval_mantel_2021) <- names(U_dist_2021)
+```
+
+```{r}
+plot_mantel_2021 <- pval_mantel_2021 %>%
+  add_column(matrix_row = names(U_dist_2021), .before = TRUE) %>%
+  pivot_longer(cols = 2:ncol(pval_mantel_2021)+1, names_to = "matrix_col", values_to = "pvalue") %>%
+  #filter(matrix_row != matrix_col) %>%
+  arrange(pvalue) %>%
+  ggplot(aes(x = matrix_col, y = matrix_row)) + 
+  geom_tile(aes(fill = -log10(pvalue) > 2.5, color = -log10(pvalue)), 
+            size = 3.5, 
+            width = 0.7, height = 0.7) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
+  scale_fill_viridis_d(end = 0.95, option = "B") +
+  scale_color_viridis_c(end = 0.8, option = "D")
+
+save_plot(plot = plot_mantel_2021, filename = here("analysis", "hypothesis_matrices", "Weather_Uhyp_Mantel_pvalue_plot_2021.png"), base_height = 13, base_width = 15)
+```

### 2022-01-13 R/Analysis_v1.1_weather_2020.Rmd
Add 25ish site*year combination phenotypic variance-covariance matrices and visualization

@@
 library(zoo)
+library(lubridate)
 library(cowplot)
@@
 
-##function to calculate daylength
-
-day.length.hrs <- function(day, latitude) {
-
-  daylength.coeff <- asin(0.39795*cos(0.2163108 + 2*atan(0.9671396*tan(0.00860*(day - 186)))))
-
-  daylength.hrs <- 24 - (24/pi)*acos((sin(0.8333*pi/180) + sin(latitude*pi/180)*sin(daylength.coeff))/(cos(latitude*pi/180)*cos(daylength.coeff)))
-
-  return(daylength.hrs)
-
-}
-
 ```
@@
   write_rds(file = here("data", "Daily_weather_2018_to_2021_10_sites.rds"))
+
+```
+
+# ----
+## load phe & wea data
+```{r}
+phe_ten_sites <- readRDS(here("data", "Phenotypes_2018_to_2021_10_sites.rds"))
+wea_ten_sites <- readRDS(file = here("data", "Daily_weather_2018_to_2021_10_sites.rds"))
+
+##function to calculate daylength
+
+day.length.hrs <- function(day, latitude) {
+
+  daylength.coeff <- asin(0.39795*cos(0.2163108 + 2*atan(0.9671396*tan(0.00860*(day - 186)))))
+
+  daylength.hrs <- 24 - (24/pi)*acos((sin(0.8333*pi/180) + sin(latitude*pi/180)*sin(daylength.coeff))/(cos(latitude*pi/180)*cos(daylength.coeff)))
+
+  return(daylength.hrs)
+
+}
 
@@
   saveRDS(file = here("data", "Weather_related_phe_FL50_3yr.rds"))
+```
+
+# Phe correlation matrices
+
+## load data
+```{r}
+gr_ten_sites <- readRDS(file = here("data", "Weather_related_phe_GR50_3yr.rds"))
+fl_ten_sites <- readRDS(here("data", "Weather_related_phe_FL50_3yr.rds"))
+
+```
+
+### All site*years
+Make phenotypic correlation matrices for greenup & flowering for these new years for all site*year combinations.
+
+```{r}
+metadata <- read_rds(here("data", "metadata.rds"))
+subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest",
+                  Gulf = "Gulf")
+year_v = c(2019, 2020, 2021)
+
+fl_ten_sites <- fl_ten_sites %>%
+  left_join(select(metadata, PLANT_ID, SUBPOP))
+gr_ten_sites <- gr_ten_sites %>%
+  left_join(select(metadata, PLANT_ID, SUBPOP))
+
+fl_wide <- fl_ten_sites %>%
+  select(PLANT_ID:FL50, SUBPOP) %>%
+  select(-GR50) %>%
+  pivot_wider(names_from = c("manu_site", "Year"), values_from = "FL50")
+gr_wide <- fl_ten_sites %>%
+  select(PLANT_ID:GR50, SUBPOP) %>%
+  pivot_wider(names_from = c("manu_site", "Year"), values_from = "GR50")
+
+k=1
+U_hyp_sitephe <- list()
+library(rlist)
+for(k in seq_along(subpop_v2)){
+
+    cov_df <- fl_wide %>%
+      ungroup() %>%
+      filter(SUBPOP %in% subpop_v2[[k]])
+
+    cor_phe <- cor(cov_df[,-(1:2)], use = "pairwise")
+    # Set diagonal of this matrix = coefficient of variation within each garden.
+    diag(cor_phe) <- cov_df %>%
+      ungroup() %>%
+      select(where(is.numeric)) %>%
+      summarise(across(everything(), ~ sqrt(var(.x, na.rm = TRUE))/mean(.x, na.rm = TRUE))) %>%
+      as_vector()
+
+    U_hyp_sitephe <- list.prepend(U_hyp_sitephe, cor_phe)
+    names(U_hyp_sitephe)[1] <- paste0("FL50", "_", names(subpop_v2)[k])
+}
+
+for(k in seq_along(subpop_v2)){
+
+    cov_df <- gr_wide %>%
+      ungroup() %>%
+      filter(SUBPOP %in% subpop_v2[[k]])
+
+    cor_phe <- cor(cov_df[,-(1:2)], use = "pairwise")
+    # Set diagonal of this matrix = coefficient of variation within each garden.
+    diag(cor_phe) <- cov_df %>%
+      ungroup() %>%
+      select(where(is.numeric)) %>%
+      summarise(across(everything(), ~ sqrt(var(.x, na.rm = TRUE))/mean(.x, na.rm = TRUE))) %>%
+      as_vector()
+
+    U_hyp_sitephe <- list.prepend(U_hyp_sitephe, cor_phe)
+    names(U_hyp_sitephe)[1] <- paste0("GR50", "_", names(subpop_v2)[k])
+}
+
+```
+
+```{r}
+plot_corr_fl_gulf <- U_hyp_sitephe$FL50_Gulf %>%
+  as_tibble() %>%
+  add_column(matrix_row = colnames(U_hyp_sitephe$FL50_Gulf), .before = TRUE) %>%
+  pivot_longer(cols = -matrix_row, names_to = "matrix_col", values_to = "pvalue") %>%
+  #filter(matrix_row != matrix_col) %>%
+  arrange(pvalue) %>%
+  ggplot(aes(x = matrix_col, y = matrix_row)) + 
+  geom_tile(aes(fill = pvalue)) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
+      scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                      "white", "#27AD81FF", "#5DC863FF",
+                                      "#FDE725FF"), limits = c(-1,1)) +
+  labs(title = "Gulf Flowering P Corr All Site*Years", 
+       subtitle = "Grey indicates NA values")
+
+save_plot(plot = plot_corr_fl_gulf, filename = here("analysis", "hypothesis_matrices", "Weather_Uhyp_Corr_FL_Gulf.png"), base_height = 4.5, base_width = 5)
+
+plot_corr_fl_midw <- U_hyp_sitephe$FL50_Midwest %>%
+  as_tibble() %>%
+  add_column(matrix_row = colnames(U_hyp_sitephe$FL50_Midwest), .before = TRUE) %>%
+  pivot_longer(cols = -matrix_row, names_to = "matrix_col", values_to = "pvalue") %>%
+  #filter(matrix_row != matrix_col) %>%
+  arrange(pvalue) %>%
+  ggplot(aes(x = matrix_col, y = matrix_row)) + 
+  geom_tile(aes(fill = pvalue)) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
+      scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                      "white", "#27AD81FF", "#5DC863FF",
+                                      "#FDE725FF"), limits = c(-1,1)) +
+  labs(title = "Midwest Flowering P Corr All Site*Years", 
+       subtitle = "Grey indicates NA values")
+
+save_plot(plot = plot_corr_fl_midw, filename = here("analysis", "hypothesis_matrices", "Weather_Uhyp_Corr_FL_Midwest.png"), base_height = 4.5, base_width = 5)
+
+plot_corr_fl_gam <- U_hyp_sitephe$FL50_Gulf_and_Midwest %>%
+  as_tibble() %>%
+  add_column(matrix_row = colnames(U_hyp_sitephe$FL50_Gulf_and_Midwest), .before = TRUE) %>%
+  pivot_longer(cols = -matrix_row, names_to = "matrix_col", values_to = "pvalue") %>%
+  #filter(matrix_row != matrix_col) %>%
+  arrange(pvalue) %>%
+  ggplot(aes(x = matrix_col, y = matrix_row)) + 
+  geom_tile(aes(fill = pvalue)) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
+      scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                      "white", "#27AD81FF", "#5DC863FF",
+                                      "#FDE725FF"), limits = c(-1,1)) +
+  labs(title = "Gulf & Midwest Flowering P Corr All Site*Years", 
+       subtitle = "Grey indicates NA values")
+
+save_plot(plot = plot_corr_fl_gam, filename = here("analysis", "hypothesis_matrices", "Weather_Uhyp_Corr_FL_GAM.png"), base_height = 4.5, base_width = 5)
+
+plot_corr_gr_gulf <- U_hyp_sitephe$GR50_Gulf %>%
+  as_tibble() %>%
+  add_column(matrix_row = colnames(U_hyp_sitephe$GR50_Gulf), .before = TRUE) %>%
+  pivot_longer(cols = -matrix_row, names_to = "matrix_col", values_to = "pvalue") %>%
+  #filter(matrix_row != matrix_col) %>%
+  arrange(pvalue) %>%
+  ggplot(aes(x = matrix_col, y = matrix_row)) + 
+  geom_tile(aes(fill = pvalue)) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
+      scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                      "white", "#27AD81FF", "#5DC863FF",
+                                      "#FDE725FF"), limits = c(-1,1)) +
+  labs(title = "Gulf Green Up P Corr All Site*Years", 
+       subtitle = "Grey indicates NA values")
+
+save_plot(plot = plot_corr_gr_gulf, filename = here("analysis", "hypothesis_matrices", "Weather_Uhyp_Corr_GR_Gulf.png"), base_height = 4.5, base_width = 5)
+
+plot_corr_gr_midw <- U_hyp_sitephe$GR50_Midwest %>%
+  as_tibble() %>%
+  add_column(matrix_row = colnames(U_hyp_sitephe$GR50_Midwest), .before = TRUE) %>%
+  pivot_longer(cols = -matrix_row, names_to = "matrix_col", values_to = "pvalue") %>%
+  #filter(matrix_row != matrix_col) %>%
+  arrange(pvalue) %>%
+  ggplot(aes(x = matrix_col, y = matrix_row)) + 
+  geom_tile(aes(fill = pvalue)) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
+      scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                      "white", "#27AD81FF", "#5DC863FF",
+                                      "#FDE725FF"), limits = c(-1,1)) +
+  labs(title = "Midwest Green Up P Corr All Site*Years", 
+       subtitle = "Grey indicates NA values")
+
+save_plot(plot = plot_corr_gr_midw, filename = here("analysis", "hypothesis_matrices", "Weather_Uhyp_Corr_GR_Midwest.png"), base_height = 4.5, base_width = 5)
+
+plot_corr_gr_gam <- U_hyp_sitephe$GR50_Gulf_and_Midwest %>%
+  as_tibble() %>%
+  add_column(matrix_row = colnames(U_hyp_sitephe$GR50_Gulf_and_Midwest), .before = TRUE) %>%
+  pivot_longer(cols = -matrix_row, names_to = "matrix_col", values_to = "pvalue") %>%
+  #filter(matrix_row != matrix_col) %>%
+  arrange(pvalue) %>%
+  ggplot(aes(x = matrix_col, y = matrix_row)) + 
+  geom_tile(aes(fill = pvalue)) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
+      scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                      "white", "#27AD81FF", "#5DC863FF",
+                                      "#FDE725FF"), limits = c(-1,1)) +
+  labs(title = "Gulf & Midwest Green Up P Corr All Site*Years", 
+       subtitle = "Grey indicates NA values")
+
+save_plot(plot = plot_corr_gr_gam, filename = here("analysis", "hypothesis_matrices", "Weather_Uhyp_Corr_GR_GAM.png"), base_height = 4.5, base_width = 5)
 ```

### 2022-03-15 R/Analysis_v1.1_weather_2020.Rmd
Add phenotypic covariance matrices for 8 sites for 2019 phenology data

@@
 library(cowplot)
+library(rlist)
 ```
@@
 ```
+
+# P covar for Li
+
+```{r}
+metadata <- read_rds(here("data", "metadata.rds"))
+subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest",
+                  Gulf = "Gulf")
+year_v = c(2019, 2020, 2021)
+gr_ten_sites <- readRDS(file = here("data",
+                                    "Weather_related_phe_GR50_3yr.rds"))
+fl_ten_sites <- readRDS(here("data", "Weather_related_phe_FL50_3yr.rds"))
+
+fl_ten_sites <- fl_ten_sites %>%
+  left_join(select(metadata, PLANT_ID, SUBPOP))
+gr_ten_sites <- gr_ten_sites %>%
+  left_join(select(metadata, PLANT_ID, SUBPOP))
+
+# change based on phenotype ---------------
+#phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_20d", "cgdd_12c_5d", 
+#             "tave_5d", "tave_10d", "tave_20d")
+phe_hyp <- c("FL50", "GR50", "dyln_fl50", "dyln_change_sec", 
+             "dyln_change_sec_7d_prior", "dyln_change_sec_14d_prior", 
+             "dyln_7d_prior", "dyln_14d_prior", "cgdd_12c_gr2fl", 
+             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")
+i=1
+k=1
+
+U_hyp_2019 <- list()
+for(i in seq_along(phe_hyp)){
+  for(k in seq_along(subpop_v2)){
+
+    cov_df <- fl_ten_sites %>%
+      ungroup() %>%
+      filter(SUBPOP %in% subpop_v2[[k]]) %>%
+      filter(Year %in% 2019) %>%
+      select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
+      group_by(PLANT_ID, SUBPOP, manu_site) %>%
+      mutate(IND = row_number()) %>%
+      pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
+      mutate(PLANT_ID = as.factor(PLANT_ID),
+             IND = as.factor(IND)) %>%
+      select(PLANT_ID, SUBPOP, IND, MI, MO, NE, OK, SD, TX1, TX2, TX3) #####
+
+    cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
+    # Set diagonal of this matrix = coefficient of variation within each garden.
+    diag(cor_phe) <- cov_df %>%
+      ungroup() %>%
+      summarise(MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),  #####
+                MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
+                NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
+                OK = sqrt(var(OK, na.rm = TRUE))/mean(OK, na.rm = TRUE),  #####
+                SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
+                TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
+                TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
+                TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE)) %>%
+      as_vector()
+
+    U_hyp_2019 <- list.prepend(U_hyp_2019, cor_phe)
+    names(U_hyp_2019)[1] <- paste0(phe_hyp[i], "_", names(subpop_v2)[k])
+  }
+}
+
+phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_20d", "cgdd_12c_5d", 
+             "tave_5d", "tave_10d", "tave_20d")
+
+U_hyp_2019_gr <- list()
+for(i in seq_along(phe_hyp)){
+  for(k in seq_along(subpop_v2)){
+
+    cov_df <- gr_ten_sites %>%
+      ungroup() %>%
+      filter(SUBPOP %in% subpop_v2[[k]]) %>%
+      filter(Year %in% 2019) %>%
+      select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
+      group_by(PLANT_ID, SUBPOP, manu_site) %>%
+      mutate(IND = row_number()) %>%
+      pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
+      mutate(PLANT_ID = as.factor(PLANT_ID),
+             IND = as.factor(IND)) %>%
+      select(PLANT_ID, SUBPOP, IND, MI, MO, NE, OK, SD, TX1, TX2, TX3) #####
+
+    cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
+    # Set diagonal of this matrix = coefficient of variation within each garden.
+    diag(cor_phe) <- cov_df %>%
+      ungroup() %>%
+      summarise(MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),  #####
+                MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
+                NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
+                OK = sqrt(var(OK, na.rm = TRUE))/mean(OK, na.rm = TRUE),  #####
+                SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
+                TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
+                TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
+                TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE)) %>%
+      as_vector()
+
+    U_hyp_2019_gr <- list.prepend(U_hyp_2019_gr, cor_phe)
+    names(U_hyp_2019_gr)[1] <- paste0(phe_hyp[i], "_", names(subpop_v2)[k])
+  }
+}
+
+saveRDS(U_hyp_2019, file = here("data", "P_covar_matrices_flowering_2019_8sites.rds"))
+saveRDS(U_hyp_2019_gr, file = here("data", "P_covar_matrices_greenup_2019_8sites.rds"))
+```

### 2023-08-02 R/Analysis_v1.2_mash_using_greedy_covar.Rmd
revision: More covariance matrices to add to greedy mash algorithm

---
title: "Analysis v1.2 mash"
author: "Alice MacQueen"
date: "2023-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(switchgrassGWAS)
library(mashr)
library(rlist)
library(cowplot)
library(here)
library(zoo)
```

# Installation reminders for any reinstalls
On debian: `apt install libgsl-dev` to install the GNU Scientific Library, zB.

#

Sam used a greedy mash algorithm to find the best set of covariance matrices to include, using 19K randomly associated variants.

Now use the best set of covariance matrices on 19K strongly associated variants.

Step 0. Set up.
  0a. Look at phenotype distribution & heritability; run GWAS to get univariate SNP effect estimates & standard errors. Already done.
  0b. Make hypothesis covariance matrices based on weather data from 2018 - 2020. For cumulative weather variables, use a range of days, 1 - 20, say. Come back and add more if 20 is chosen by greedy mash.
Step 1. Run greedy mash to find best set of covariance matrices that explain SNP effect variation across environments, using 19K 'randomly associated variants'. Go back to 0b. if high number of day covar matrices tend to be chosen by greedy mash.
Step 2. Use covariances from (1) to run mash on 19K 'strongly associated variants'. 
3. Look at loadings on covariance matrices and meanings of covariance matrices as the writeup (Figure 2, 3; Table 1). Look at amount of antagonistic pleiotropy vs same sign significant effects.
4. Repeat the QTL analysis... with Li.

How much of a lift is it to re-run the greedy mash algorithm? If we, or if reviewers are thinking about this in a Jiaming Yu sense, I think the next logical question to ask is 'Why these covariance matrices'? The overly honest answer for the current set is that there were computational limits to running mash with so many covariance matrices, so I chose a few arbitrary time points for each weather variable we thought might be a environmental driver. But now we have introduced a method to choose between highly correlated covariance matrices, perhaps we could do better. 
2.

# Deprecated analysis
### Code from Plots_v0.2_manuscript_figures.Rmd to make weather variables

Actually, code from Analysis_v1.1_weather_2020.Rmd might be more suitable, if I am remaking these anyway.

```{r}
wea_df <- readRDS("~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")
metadata <- readRDS("~/Github/pvdiv-phenology-gxe/data/metadata.rds")
sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
phenotypes <- readRDS("~/Github/pvdiv-phenology-gxe/data/Phenology_CGDD_PTT_and_rainfall_phenotypes.rds")
jday_phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
```

```{r}
jday_phe <- wea_df %>%
  group_by(SITE) %>%
  mutate(dyln_change_sec = (DYLN - lag(DYLN))*60*60,
         FL50 = yday(DOY)) %>%
  select(DOY, DYLN, dyln_change_sec, everything()) %>%
  select(-PTT_12C) %>%
  right_join(jday_phe, by = c("SITE", "FL50")) %>%
  select(-(TMAX:CPTT_11C))
jday_phe <- metadata %>%
  rename(Lat_origin = LATITUDE, Long_origin = LONGITUDE) %>%
  select(PLANT_ID, SUBPOP, Lat_origin, Long_origin) %>%
  right_join(jday_phe) %>%
  left_join(sites)

jday_phe$SUBPOP <- factor(jday_phe$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
jday_phe$SITE <- factor(jday_phe$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
jday_phe$manu_site <- factor(jday_phe$manu_site, levels = rev(c("TX1", "TX2", "TX3", "TX4", "OK", "MO", "NE", "IL", "MI", "SD")))
```

## define environmental cues for greenup
Probably all CGDD related.
```{r}

asreml_fl <- jday_phe %>%
  select(PLANT_ID:Long_origin, PLOT_GL, SITE, FL50, GR50, CGDD_12C, CRAIN, CRAIN_5d, DYLN, dyln_change_sec) %>%
  rename(dyln_fl50 = DYLN, cgdd_12c_gr2fl = CGDD_12C, crain_gr2fl = CRAIN) %>%
  left_join(wea_df) %>%
  group_by(SITE, PLOT_GL) %>%
  filter(between(DOY, as_date(FL50, origin = "2019-01-01") - 6, as_date(FL50, origin = "2019-01-01"))) %>%
  mutate(crain_2d = RAIN_SM + lag(RAIN_SM),
         crain_3d = RAIN_SM + lag(RAIN_SM) + lag(RAIN_SM, n = 2),
         crain_7d = sum(RAIN_SM)) %>%
  filter(DOY == as_date(FL50, origin = "2019-01-01")) %>%
  filter(SUBPOP %in% c("Midwest", "Gulf", "Atlantic") & !(SITE %in% c("OVTN"))) %>%
  select(-(TMAX:TMIN_SM), -(GDD_13C_SM:CPTT_13C)) %>%
  rename(crain_1d = RAIN_SM, crain_5d = CRAIN_5d)
saveRDS(asreml_fl, file = "../data/Weather_related_phe_for_asreml_h2_FL50.rds")
asreml_gr <- jday_phe %>%
  select(PLANT_ID:Long_origin, PLOT_GL, SITE, GR50) %>%
  left_join(wea_df) %>%
  group_by(SITE, PLOT_GL) %>%
  filter(DOY <= as_date(GR50, origin = "2019-01-01")) %>%
  mutate(cgdd_12c_jan2gr = sum(GDD_12C_SM),
         cgdd_8c_jan2gr = sum(GDD_8C_SM)
         ) %>%
  select(-(GDD_13C_SM:CPTT_13C), -(TMAX:Longitude), GDD_12C_SM) %>%
  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 17, as_date(GR50, origin = "2019-01-01"))) %>%
  mutate(tmax_18d = mean(TMAX_SM),
         tmin_18d = mean(TMIN_SM),
         tave_18d = (tmax_18d + tmin_18d)/2,
         cgdd_12c_18d = mean(GDD_12C_SM)) %>%
  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 9, as_date(GR50, origin = "2019-01-01"))) %>%
  mutate(tmax_10d = mean(TMAX_SM),
         tmin_10d = mean(TMIN_SM),
         tave_10d = (tmax_10d + tmin_10d)/2,
         cgdd_12c_10d = mean(GDD_12C_SM)) %>%
  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 4, as_date(GR50, origin = "2019-01-01"))) %>%
  mutate(tmax_5d = mean(TMAX_SM),
         tmin_5d = mean(TMIN_SM),
         tave_5d = (tmax_5d + tmin_5d)/2,
         cgdd_12c_5d = mean(GDD_12C_SM))%>%
  filter(DOY == as_date(GR50, origin = "2019-01-01")) %>%
  filter(SUBPOP %in% c("Midwest", "Gulf", "Atlantic") & !(SITE %in% c("OVTN"))) %>%
  select(-(TMAX_SM:RAIN_SM), -Day_of_year)
saveRDS(asreml_gr, file = "../data/Weather_related_phe_for_asreml_h2_GR50.rds")
```

##

# ----
## load phe & wea data
```{r}
phe_ten_sites <- readRDS(here("data", "Phenotypes_2018_to_2021_10_sites.rds"))
jday_phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
wea_ten_sites <- readRDS(file = here("data", "Daily_weather_2018_to_2021_10_sites.rds"))

##function to calculate daylength

day.length.hrs <- function(day, latitude) {

  daylength.coeff <- asin(0.39795*cos(0.2163108 + 2*atan(0.9671396*tan(0.00860*(day - 186)))))

  daylength.hrs <- 24 - (24/pi)*acos((sin(0.8333*pi/180) + sin(latitude*pi/180)*sin(daylength.coeff))/(cos(latitude*pi/180)*cos(daylength.coeff)))

  return(daylength.hrs)

}

phe_2019 <- jday_phe %>%
  select(-(CGDD_12C:dyln_change_sec)) %>%
  pivot_longer(cols = GR50:FL50, names_to = "PHE", values_to = "JDAY")
```

```{r}

wea_ten_sites <- wea_ten_sites %>%
  mutate(JDAY = yday(Working_date),
         DYLN = day.length.hrs(day = JDAY, latitude = Latitude),
         tmax_sm = ifelse(is.na(tmax),
                          (na.locf(tmax) + na.locf(tmax, fromLast = TRUE))/2,
                          tmax),
         tmin_sm = ifelse(is.na(tmin),
                          (na.locf(tmin) + na.locf(tmin, fromLast = TRUE))/2,
                          tmin),
         tmean_sm = ifelse(is.na(Tmean),
                          (na.locf(Tmean) + na.locf(Tmean, fromLast = TRUE))/2,
                          Tmean),
         GDD_12Cbase = ifelse((tmax_sm + tmin_sm)/2 >= 12.0,
                              (tmax_sm + tmin_sm)/2 - 10,
                              0),
         Year = year(Working_date)) %>%
  mutate(JDAY_EXP = case_when(Year == 2018 ~ JDAY,
                              Year == 2019 ~ JDAY + 365,
                              Year == 2020 ~ JDAY + (365*2),
                              Year == 2021 ~ JDAY + (365*3),
                              TRUE ~ NA_real_)) %>%
  arrange(Latitude, Year, JDAY_EXP) %>%
  select(-JDAY) %>%
  group_by(manu_site) %>%
  mutate(crain_2d = prcp + lag(prcp, default = 0),
         crain_3d = prcp + lag(prcp, default = 0) + 
           lag(prcp, default = 0, n = 2),
         crain_4d = prcp + lag(prcp, default = 0) + 
           lag(prcp, default = 0, n = 2) + 
           lag(prcp, default = 0, n = 3),
         crain_5d = prcp + lag(prcp, default = 0) + 
           lag(prcp, default = 0, n = 2) + lag(prcp, default = 0, n = 3) +
           lag(prcp, default = 0, n = 4),
         crain_6d = prcp + lag(prcp, default = 0) + 
           lag(prcp, default = 0, n = 2) + lag(prcp, default = 0, n = 3) +
           lag(prcp, default = 0, n = 4) +
           lag(prcp, default = 0, n = 5),
         crain_7d = prcp + lag(prcp, default = 0) + 
           lag(prcp, default = 0, n = 2) + lag(prcp, default = 0, n = 3) +
           lag(prcp, default = 0, n = 4) +
           lag(prcp, default = 0, n = 5) +
           lag(prcp, default = 0, n = 6),
         dyln_change_sec = (DYLN - lag(DYLN))*60*60,
         dyln_change_sec_2d_prior = (lag(DYLN, n = 1) - 
                                       lag(DYLN, n = 2))*60*60,
         dyln_change_sec_3d_prior = (lag(DYLN, n = 2) - 
                                       lag(DYLN, n = 3))*60*60,
         dyln_change_sec_4d_prior = (lag(DYLN, n = 3) - 
                                       lag(DYLN, n = 4))*60*60,
         dyln_change_sec_5d_prior = (lag(DYLN, n = 4) - 
                                       lag(DYLN, n = 5))*60*60,
         dyln_change_sec_6d_prior = (lag(DYLN, n = 5) - 
                                       lag(DYLN, n = 6))*60*60,
         dyln_change_sec_7d_prior = (lag(DYLN, n = 6) - 
                                       lag(DYLN, n = 7))*60*60,
         dyln_change_sec_14d_prior = (lag(DYLN, n = 13) - 
                                       lag(DYLN, n = 14))*60*60,
         dyln_2d_prior = lag(DYLN, n = 1),
         dyln_3d_prior = lag(DYLN, n = 2),
         dyln_4d_prior = lag(DYLN, n = 3),
         dyln_5d_prior = lag(DYLN, n = 4),
         dyln_6d_prior = lag(DYLN, n = 5),
         dyln_7d_prior = lag(DYLN, n = 6),
         dyln_14d_prior = lag(DYLN, n = 13),
         cgdd_12c_2d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0))/2,
         tave_2d = (tmean_sm + lag(tmean_sm, default = 0))/2,
         cgdd_12c_3d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2))/3,
         tave_3d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2))/3,
         cgdd_12c_4d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3))/4,
         tave_4d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3))/4,
         cgdd_12c_5d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4))/5,
         tave_5d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4))/5,
         cgdd_12c_6d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5))/6,
         tave_6d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5))/6,
         cgdd_12c_7d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5) +
           lag(GDD_12Cbase, default = 0, n = 6))/7,
         tave_7d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5) +
           lag(tmean_sm, default = 0, n = 6))/7,
         cgdd_12c_14d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5) +
           lag(GDD_12Cbase, default = 0, n = 6) +
           lag(GDD_12Cbase, default = 0, n = 7) +
           lag(GDD_12Cbase, default = 0, n = 8) +
           lag(GDD_12Cbase, default = 0, n = 9) +
           lag(GDD_12Cbase, default = 0, n = 10) +
           lag(GDD_12Cbase, default = 0, n = 11) +
           lag(GDD_12Cbase, default = 0, n = 12) +
           lag(GDD_12Cbase, default = 0, n = 13))/14,
         tave_14d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5) +
           lag(tmean_sm, default = 0, n = 6) +
           lag(tmean_sm, default = 0, n = 7) +
           lag(tmean_sm, default = 0, n = 8) +
           lag(tmean_sm, default = 0, n = 9) +
           lag(tmean_sm, default = 0, n = 10) +
           lag(tmean_sm, default = 0, n = 11) +
           lag(tmean_sm, default = 0, n = 12) +
           lag(tmean_sm, default = 0, n = 13))/14,
         cgdd_12c_21d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5) +
           lag(GDD_12Cbase, default = 0, n = 6) +
           lag(GDD_12Cbase, default = 0, n = 7) +
           lag(GDD_12Cbase, default = 0, n = 8) +
           lag(GDD_12Cbase, default = 0, n = 9) +
           lag(GDD_12Cbase, default = 0, n = 10) +
           lag(GDD_12Cbase, default = 0, n = 11) +
           lag(GDD_12Cbase, default = 0, n = 12) +
           lag(GDD_12Cbase, default = 0, n = 13) +
           lag(GDD_12Cbase, default = 0, n = 14) +
           lag(GDD_12Cbase, default = 0, n = 15) +
           lag(GDD_12Cbase, default = 0, n = 16) +
           lag(GDD_12Cbase, default = 0, n = 17) +
           lag(GDD_12Cbase, default = 0, n = 18) +
           lag(GDD_12Cbase, default = 0, n = 19) +
           lag(GDD_12Cbase, default = 0, n = 20))/21,
         tave_21d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5) +
           lag(tmean_sm, default = 0, n = 6) +
           lag(tmean_sm, default = 0, n = 7) +
           lag(tmean_sm, default = 0, n = 8) +
           lag(tmean_sm, default = 0, n = 9) +
           lag(tmean_sm, default = 0, n = 10) +
           lag(tmean_sm, default = 0, n = 11) +
           lag(tmean_sm, default = 0, n = 12) +
           lag(tmean_sm, default = 0, n = 13) +
           lag(tmean_sm, default = 0, n = 14) +
           lag(tmean_sm, default = 0, n = 15) +
           lag(tmean_sm, default = 0, n = 16) +
           lag(tmean_sm, default = 0, n = 17) +
           lag(tmean_sm, default = 0, n = 18) +
           lag(tmean_sm, default = 0, n = 19) +
           lag(tmean_sm, default = 0, n = 20))/21,
         cgdd_12c_28d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5) +
           lag(GDD_12Cbase, default = 0, n = 6) +
           lag(GDD_12Cbase, default = 0, n = 7) +
           lag(GDD_12Cbase, default = 0, n = 8) +
           lag(GDD_12Cbase, default = 0, n = 9) +
           lag(GDD_12Cbase, default = 0, n = 10) +
           lag(GDD_12Cbase, default = 0, n = 11) +
           lag(GDD_12Cbase, default = 0, n = 12) +
           lag(GDD_12Cbase, default = 0, n = 13) +
           lag(GDD_12Cbase, default = 0, n = 14) +
           lag(GDD_12Cbase, default = 0, n = 15) +
           lag(GDD_12Cbase, default = 0, n = 16) +
           lag(GDD_12Cbase, default = 0, n = 17) +
           lag(GDD_12Cbase, default = 0, n = 18) +
           lag(GDD_12Cbase, default = 0, n = 19) +
           lag(GDD_12Cbase, default = 0, n = 20) +
           lag(GDD_12Cbase, default = 0, n = 21) +
           lag(GDD_12Cbase, default = 0, n = 22) +
           lag(GDD_12Cbase, default = 0, n = 23) +
           lag(GDD_12Cbase, default = 0, n = 24) +
           lag(GDD_12Cbase, default = 0, n = 25) +
           lag(GDD_12Cbase, default = 0, n = 26) +
           lag(GDD_12Cbase, default = 0, n = 27))/28,
         tave_28d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5) +
           lag(tmean_sm, default = 0, n = 6) +
           lag(tmean_sm, default = 0, n = 7) +
           lag(tmean_sm, default = 0, n = 8) +
           lag(tmean_sm, default = 0, n = 9) +
           lag(tmean_sm, default = 0, n = 10) +
           lag(tmean_sm, default = 0, n = 11) +
           lag(tmean_sm, default = 0, n = 12) +
           lag(tmean_sm, default = 0, n = 13) +
           lag(tmean_sm, default = 0, n = 14) +
           lag(tmean_sm, default = 0, n = 15) +
           lag(tmean_sm, default = 0, n = 16) +
           lag(tmean_sm, default = 0, n = 17) +
           lag(tmean_sm, default = 0, n = 18) +
           lag(tmean_sm, default = 0, n = 19) +
           lag(tmean_sm, default = 0, n = 20) +
           lag(tmean_sm, default = 0, n = 21) +
           lag(tmean_sm, default = 0, n = 22) +
           lag(tmean_sm, default = 0, n = 23) +
           lag(tmean_sm, default = 0, n = 24) +
           lag(tmean_sm, default = 0, n = 25) +
           lag(tmean_sm, default = 0, n = 26) +
           lag(tmean_sm, default = 0, n = 27))/28,
         )

```

```{r}
wea_ten_sites
```

# Define phenotypes
```{r}
Tbase <- 12

phe_gr_fl <- phe_2019 %>%
  mutate(JDAY_EXP = case_when(YEAR == 2018 ~ JDAY,
                              YEAR == 2019 ~ JDAY + 365,
                              YEAR == 2020 ~ JDAY + (365*2),
                              YEAR == 2021 ~ JDAY + (365*3),
                              TRUE ~ NA_real_)) %>%
  select(-JDAY) %>%
  pivot_wider(names_from = PHE, values_from = JDAY_EXP) %>%
  left_join(select(sites, SITE, manu_site)) %>%
  mutate(Region = case_when(manu_site %in% c("SD", "MO", "IL", "MI", "NE") ~ "North",
                            manu_site %in% c("TX1", "TX2", "TX3", "TX4") ~ "Texas",
                                             TRUE ~ "OK"))

tmp2 <- wea_ten_sites %>%
  select(-Year)

gr_ten_sites <- phe_gr_fl %>%
  ungroup() %>%
  mutate(GR50 = floor(GR50)) %>%
  filter(!is.na(GR50)) %>%
  left_join(tmp2, by = c("manu_site", "GR50" = "JDAY_EXP")) %>%
  select(-SITE, -FL50, -(Working_date:wndk), -Tdrnl, -(Latitude:Elevation), -(tmax_sm:tmean_sm), -DLYN) %>%
  rename(crain_1d = prcp, dyln_gr50 = DYLN, cgdd_12c_1d = GDD_12Cbase, tave_1d = Tmean)

fl_ten_sites <- phe_gr_fl %>%
  ungroup() %>%
  filter(!is.na(FL50)) %>%
  mutate(FL50 = floor(FL50),
         GR50 = floor(GR50)) %>%
  left_join(tmp2, by = c("manu_site", "FL50" = "JDAY_EXP")) %>%
  select(-SITE, -(Working_date:wndk), -Tdrnl, -(Latitude:Elevation), -(tmax_sm:tmean_sm), -DLYN) %>%
  rename(crain_1d = prcp, dyln_gr50 = DYLN, cgdd_12c_1d = GDD_12Cbase, tave_1d = Tmean)
# 11 September flowering dates in 2021 at Pickle have NA's for weather values. Decided to let this be.

```

# Covar csv for greedy mash algorithm

Make & save hypothesis matrices in a hypothesis_matrices directory in data. 
Then hopefully it's not much of a lift for Sam to run the greedy mash algorithm again. Otherwise I will figure out how to run it.

phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", 
             "tave_5d", "tave_10d", "tave_18d")
phe_hyp <- c("FL50", "GR50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")

```{r}
gr_ten_sites %>%
  saveRDS(file = here("data", "Weather_related_phe_GR50_2019.rds"))
fl_ten_sites %>%
  saveRDS(file = here("data", "Weather_related_phe_FL50_2019.rds"))
```

### load weather phenotypes
```{r}
gr_ten_sites <- readRDS(file = here("data", 
                                    "Weather_related_phe_GR50_2019.rds"))
fl_ten_sites <- readRDS(file = here("data", 
                                    "Weather_related_phe_FL50_2019.rds"))
```

```{r}
phe_hyp <- colnames(fl_ten_sites)[c(5,9:51)]
colnames(gr_ten_sites)[8:50]
```

```{r}
metadata <- read_rds(here("data", "metadata.rds"))
subpop_v2 <- list(all = c("Gulf", "Midwest"), midwest = "Midwest",
                  gulf = "Gulf")
year_v = c(2019, 2020, 2021)

fl_ten_sites <- fl_ten_sites %>%
  left_join(select(metadata, PLANT_ID, SUBPOP)) %>%
  rename(Year = YEAR)

# change based on phenotype ---------------
#phe_hyp <- c("FL50", "GR50", "dyln_fl50", "dyln_change_sec", 
#             "dyln_change_sec_7d_prior", "dyln_change_sec_14d_prior", 
#             "dyln_7d_prior", "dyln_14d_prior", "cgdd_12c_gr2fl", 
#             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")

phe_hyp <- colnames(fl_ten_sites)[c(5,9:51)]
i=1
k=1

U_hyp_2019 <- list()
for(i in seq_along(phe_hyp)){
  for(k in seq_along(subpop_v2)){

    cov_df <- fl_ten_sites %>%
      ungroup() %>%
      filter(SUBPOP %in% subpop_v2[[k]]) %>%
      filter(Year %in% 2019) %>%
      select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
      group_by(PLANT_ID, SUBPOP, manu_site) %>%
      mutate(IND = row_number()) %>%
      pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
      mutate(PLANT_ID = as.factor(PLANT_ID),
             IND = as.factor(IND)) %>%
      select(PLANT_ID, SUBPOP, IND, MI, MO, NE, SD, TX1, TX2, TX3) #####  IL,

    cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
    # Set diagonal of this matrix = coefficient of variation within each garden.
    diag(cor_phe) <- cov_df %>%
      ungroup() %>%
      summarise(#IL = sqrt(var(IL, na.rm = TRUE))/mean(IL, na.rm = TRUE),  #####
                MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),  #####
                MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
                NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
                SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
                TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
                TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
                TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE)) %>%
      as_vector()

    U_hyp_2019 <- list.prepend(U_hyp_2019, cor_phe)
    names(U_hyp_2019)[1] <- paste0("flowering.", names(subpop_v2)[k], ".", phe_hyp[i], ".U.mat")
  }
}

U_hyp_2019
length(U_hyp_2019)
for (i in 1:length(U_hyp_2019)) {
  d_out <- as.data.frame(U_hyp_2019[[i]], optional = FALSE) %>%
    rownames_to_column(var = " ")
  write_csv(d_out, file = here("data", "greedy_more_cov", "hypothesis_matrices", paste0(names(U_hyp_2019)[i], ".csv")), )
}
```

```{r}
gr_ten_sites <- gr_ten_sites %>%
  left_join(select(metadata, PLANT_ID, SUBPOP)) %>%
  rename(Year = YEAR)

# change based on phenotype ---------------
#phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_20d", "cgdd_12c_5d", 
#             "tave_5d", "tave_10d", "tave_20d")
phe_hyp <- colnames(gr_ten_sites)[8:50]
i=1
k=1

U_hyp_2019 <- list()
for(i in seq_along(phe_hyp)){
  for(k in seq_along(subpop_v2)){

    cov_df <- gr_ten_sites %>%
      ungroup() %>%
      filter(SUBPOP %in% subpop_v2[[k]]) %>%
      filter(Year %in% 2019) %>%
      select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
      group_by(PLANT_ID, SUBPOP, manu_site) %>%
      mutate(IND = row_number()) %>%
      pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
      mutate(PLANT_ID = as.factor(PLANT_ID),
             IND = as.factor(IND)) %>%
      select(PLANT_ID, SUBPOP, IND, MI, MO, NE, SD, TX1, TX2, TX3) #####  IL,

    cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
    # Set diagonal of this matrix = coefficient of variation within each garden.
    diag(cor_phe) <- cov_df %>%
      ungroup() %>%
      summarise(#IL = sqrt(var(IL, na.rm = TRUE))/mean(IL, na.rm = TRUE),  #####
                MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),  #####
                MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
                NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
                SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
                TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
                TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
                TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE)) %>%
      as_vector()

    U_hyp_2019 <- list.prepend(U_hyp_2019, cor_phe)
    names(U_hyp_2019)[1] <- paste0("greenup.", names(subpop_v2)[k], ".", phe_hyp[i], ".U.mat")
  }
}
for (i in 1:length(U_hyp_2019)) {
  d_out <- as.data.frame(U_hyp_2019[[i]], optional = FALSE) %>%
    rownames_to_column(var = " ")
  write_csv(d_out, file = here("data", "greedy_more_cov", "hypothesis_matrices", paste0(names(U_hyp_2019)[i], ".csv")), )
}
```

Visualize correlation matrices
```{r}

i=7
switchgrassGWAS::mash_plot_pairwise_sharing(corrmatrix = U_hyp_2019[[i]],
                                            reorder = FALSE)
switchgrassGWAS::mash_plot_pairwise_sharing(corrmatrix = U_hyp_2019[[i+1]],
                                            reorder = FALSE)
switchgrassGWAS::mash_plot_pairwise_sharing(corrmatrix = U_hyp_2019[[i+2]],
                                            reorder = FALSE) 
names(U_hyp_2019)[[i+2]]
```

### 2023-08-08 R/Analysis_v1.2_mash_using_greedy_covar.Rmd
analysis: Add OK to 1-7d and 1-4w hypothesis covariance matrices

@@
              IND = as.factor(IND)) %>%
-      select(PLANT_ID, SUBPOP, IND, MI, MO, NE, SD, TX1, TX2, TX3) #####  IL,
+      select(PLANT_ID, SUBPOP, IND, MI, MO, NE, OK, SD, TX1, TX2, TX3) #####  IL,
 
@@
                 NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
+                OK = sqrt(var(OK, na.rm = TRUE))/mean(OK, na.rm = TRUE),
                 SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
@@
     rownames_to_column(var = " ")
-  write_csv(d_out, file = here("data", "greedy_more_cov", "hypothesis_matrices", paste0(names(U_hyp_2019)[i], ".csv")), )
+  write_csv(d_out, file = here("data", "greedy_more_cov", "hypothesis_matrices", paste0(names(U_hyp_2019)[i], ".csv")))
 }

### 2023-09-08 R/Analysis_v1.2_mash_using_greedy_covar.Rmd
analysis: run mash using covar mat from greedy mash algorithm

@@
 library(bigsnpr)
+library(mashr)
 library(switchgrassGWAS)
-library(mashr)
 library(rlist)
@@
 
+For mashr: Sometimes install.packages("RcppGSL") is also needed.
 #

### 2023-09-08 R/nohup_v1.2_mash_greedy_greenup_gulf_strong.R

@@ -1,0 +1,125 @@
+library(tidyverse)
+library(bigsnpr)
+library(switchgrassGWAS)
+library(mashr)
+library(rlist)
+library(here)
+
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+outputdir <- here("analysis", "mash", "Gulf", "inputs")
+mashdir <- here("analysis", "mash_greedy_algorithm")
+
+#metadata <- read_rds(here("data", "metadata.rds"))
+# pvdiv PLANT_ID metadata
+#sites <- read_rds(here("data", "sites.rds"))
+# pvdiv common garden metadata
+#subpops <- read_rds(here("data", "subpop_color_coding.rds"))
+phe_gr_df <- read_rds(here("data",
+                           "Weather_related_phe_for_asreml_h2_GR50.rds"))
+# all greenup related phenotypes, including greenup as functions of weather variables.
+phe_fl_df <- read_rds(here("data",
+                           "Weather_related_phe_for_asreml_h2_FL50.rds"))
+# all flowering related phenotypes, including functions of weather variables.
+k_full <- read_rds(here("data",
+                        "Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds"))
+# kinship matrix for 630 tetraploid individuals.
+
+## vectors with subsets of gardens, phenotypes, and genetic subpopulations.
+#site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
+#               NE = "NE", MI = "MI", SD = "SD")#, eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
+subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", 
+                 Gulf = "229g_10.3M")
+#subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", 
+#                  Gulf = "Gulf")
+#phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
+#                "crain_1d", "GR50", "cgdd_12c_10d")
+#phe_mult <- list(flowering_5 = c("FL50", "dyln_fl50", "cgdd_12c_gr2fl",
+#                                 "crain_gr2fl", "crain_1d"), 
[…]

### 2023-09-11 analysis/mash_greedy_algorithm/mash_switchgrass_models/mash/all_matrices/mash_greedy.R
analysis: greedy mash algorithm using original covariance matrices

@@ -1,0 +1,64 @@
+## Run GWAS 
+library(dplyr)
+library(bigsnpr)
+library(readr)
+library(mashr)
+library(AnnotationDbi)
+library(optparse)
+library(collections)
+library(bigparallelr)
+library('matrixcalc')
+assert_cores(1)
+
+option_list = list(
+  make_option(c("-p","--pheno"), type="character", default=NULL,help="which site to use",metavar="character"),
+  make_option(c("-c","--cohort"), type="character", default=NULL,help="gulf, midwest, or all",metavar="character"),
+  make_option(c("-m","--matrices"), type="character", default=NULL,help="gulf, midwest, or all",metavar="character"),
+  make_option(c("-s","--step"), type="character", default=NULL,help="step in the greedy algorithm",metavar="character"),
+  make_option(c("-e","--effects"), type="character", default=NULL,help="gulf, midwest, or all",metavar="character")
+)
+
+opt_parser = OptionParser(option_list=option_list)
+opt = parse_args(opt_parser)
+
+if (is.null(opt$pheno)) {
+  print_help(opt_parser)
+  stop("Missing argument for phenotype code", call.=FALSE)
+}
+
+pheno = opt$pheno
+cohort = opt$cohort
+matrices = unlist(strsplit(opt$matrices,','))
+effect = opt$effects
+step = opt$step
+
+betas = read.table(paste0('combined_mash_effects/',pheno,'.',cohort,'.betas.',effect,'.txt'))
+stderrs = read.table(paste0('combined_mash_stderrs/',pheno,'.',cohort,'.stderrs.',effect,'.txt'))
+
+mat_betas = matrix(unlist(betas),ncol=8,byrow = TRUE)
+mat_stderrs = matrix(unlist(stderrs),ncol=8,byrow = TRUE)
+
+data=mash_set_data(mat_betas,mat_stderrs) 
+
+
+covmats = list()
+labels = list()
+for (matrix in matrices) {
+
+  label=gsub('.csv','',matrix)
+  label = gsub('matrices/','',label)
+  temp = read.csv(matrix,header = TRUE, row.names = 1)
+  temp = matrix(unlist(temp),ncol=8,byrow = TRUE)
+  covmats[[label]] = temp
+  labels = append(labels,label)
+  
+}
+
+g = mash(data,covmats)
+likelihood = mash_compute_loglik(g,data)
+weights=get_estimated_pi(g)
+
+
[…]

### 2023-09-11 analysis/mash_greedy_algorithm/mash_switchgrass_models/mash/all_matrices/pod.greedy.sh

@@ -1,0 +1,12 @@
+python run.mash.greedy.search.py -p greenup -c gulf -e random
+python run.mash.greedy.search.py -p greenup -c midwest -e random
+python run.mash.greedy.search.py -p greenup -c all -e random
+python run.mash.greedy.search.py -p greenup -c gulf -e strong
+python run.mash.greedy.search.py -p greenup -c midwest -e strong
+python run.mash.greedy.search.py -p greenup -c all -e strong
+python run.mash.greedy.search.py -p flowering -c gulf -e random
+python run.mash.greedy.search.py -p flowering -c midwest -e random
+python run.mash.greedy.search.py -p flowering -c all -e random
+python run.mash.greedy.search.py -p flowering -c gulf -e strong
+python run.mash.greedy.search.py -p flowering -c midwest -e strong
+python run.mash.greedy.search.py -p flowering -c all -e strong

### 2023-09-11 analysis/mash_greedy_algorithm/mash_switchgrass_models/mash/hypothesis_matrices_only/build_covariance_matrices.py

@@ -1,0 +1,248 @@
+import pandas as pd 
+import numpy as np 
+import sys
+###################
+# The following class object is written to generate the covariance matrices hypothesized in the
+# mash framework of MacQueen et al. There are three types of matrices included in the analysis
+# the first set are the hypothesis based matrices calculated from data collected for individuals
+# at each of the eight sites that were analyzed. The second set are those that are data driven 
+# and use the PCS of the XXX matrix to generate hypothesized covariance relationships between the 
+# different sites. Finally, the canonical matrices are generated using the mash software and represent
+# site specific effects, no effects, equal site effects, and a heterogenous effect case.
+###################
+
+###################
+
+class switchgrass_covariance(object):
+
+	def __init__(self,pheno):
+
+		#read in the relevant metadata from all the individuals and sites
+		self.metadata = pd.read_csv('metadata.txt',sep = "\t")
+
+		#read in sites file that maps four letter site codes to labels used in manuscript
+		self.sites = pd.read_csv('sites.txt',sep = '\t')
+		self.site_codes = pd.unique(self.sites['manu_site'].sort_values())
+		self.SITES = pd.unique(self.sites['SITE'].sort_values())
+		self.sitemapper = {i:j for i,j in zip(self.sites['SITE'],self.sites['manu_site'])}
+		#Number of folds that were used in the analysis, here 5
+		self.folds = [i for i in range(1,6)]
+
+		#Read in all possible ids to be used for generating covariance matrices
+		self.allids = pd.read_csv('../gwas/plink_files/Pvirgatum_V5_GWAS_363g_Gulf_and_Midwest_subset_maf0.05.fam', delim_whitespace = True, header = None)
+		self.allids = self.allids[1].to_list()
+		
+		#read in the phenotypes
+		self.phenotypes = pd.read_csv(pheno + '_phenos.txt',delim_whitespace = True)
+		self.phenotypes = self.phenotypes[(self.phenotypes['SITE'].isin(self.SITES)) & (self.phenotypes['SUBPOP'].isin(['Gulf','Midwest']))].drop_duplicates(keep='first')		
+		self.phenotypes['manu_site'] = self.phenotypes['SITE'].map(self.sitemapper)
+		
+		#read in the list of cohort ids
[…]

### 2023-09-11 analysis/mash_greedy_algorithm/mash_switchgrass_models/mash/hypothesis_matrices_only/combine.bootstrap.gwas.to.mash.py

@@ -1,0 +1,35 @@
+import pandas as pd 
+import numpy as np 
+import sys
+
+
+####Create the true matrices of effects for all three populations
+bootstrap = str(sys.argv[1])
+
+populations = ['gulf','midwest','all']
+sites = ['MI','MO','NE','OK','SD','TX1','TX2','TX3']
+traits = ['flowering','greenup']
+
+for trait in traits:
+	for pop in populations:
+		for effects in ['strong','random']:
+			print(trait,pop,effects)
+			first = pd.read_csv('/stor/work/Kirkpatrick/sps2972/switchgrass/gwas/bootstrap_gwasoutputs/' + sites[0] + '/' + trait + '.' + pop + '.' + bootstrap + '.' + effects + '.results.csv')
+			betas = first[['estim']].rename(columns={'estim':sites[0]})
+			stderr = first[['std.err']].rename(columns={'std.err':sites[0]})
+			for site in sites[1:]:
+				temp = pd.read_csv('/stor/work/Kirkpatrick/sps2972/switchgrass/gwas/bootstrap_gwasoutputs/' + site + '/' + trait + '.' + pop + '.' + bootstrap + '.' + effects + '.results.csv')
+				betas[site] = temp[['estim']]
+				stderr[site] = temp[['std.err']]
+
+			betas = betas.dropna()
+			stderr = stderr.dropna()
+			if betas.shape[0] != stderr.shape[0]:
+
+				idx = betas.index.intersection(stderr.index)
+				betas = betas.loc[idx]
+				stderr = stderr.loc[idx]
+				
+			betas.to_csv('bootstrap_combined_mash_effects/' + trait + '.' + pop + '.betas.' + effects + '.' + bootstrap + '.txt',sep = '\t',header = False, index = False)
+			stderr.to_csv('bootstrap_combined_mash_stderrs/' + trait + '.' + pop + '.stderrs.' + effects + '.' + bootstrap + '.txt',sep = '\t',header = False, index = False)
+			

### 2023-09-11 analysis/mash_greedy_algorithm/mash_switchgrass_models/mash/hypothesis_matrices_only/combine.gwas.to.mash.py

@@ -1,0 +1,36 @@
+import pandas as pd 
+import numpy as np 
+import sys
+
+
+####Create the true matrices of effects for all three populations
+
+populations = ['gulf','midwest','all']
+sites = ['MI','MO','NE','OK','SD','TX1','TX2','TX3']
+traits = ['flowering','greenup']
+
+for trait in traits:
+	for pop in populations:
+		for effects in ['strong','random']:
+			print(trait,pop,effects)
+			first = pd.read_csv('/stor/work/Kirkpatrick/sps2972/switchgrass/gwas/gwasoutputs/' + sites[0] + '/' + trait + '.wc.' + pop + '.' + effects + '.results.csv')
+			betas = first[['estim']].rename(columns={'estim':sites[0]})
+			stderr = first[['std.err']].rename(columns={'std.err':sites[0]})
+			for site in sites[1:]:
+				temp = pd.read_csv('/stor/work/Kirkpatrick/sps2972/switchgrass/gwas/gwasoutputs/' + site + '/' + trait + '.wc.' + pop + '.' + effects + '.results.csv')
+				betas[site] = temp[['estim']]
+				stderr[site] = temp[['std.err']]
+
+
+			betas = betas.dropna()
+			stderr = stderr.dropna()
+			if betas.shape[0] != stderr.shape[0]:
+
+				idx = betas.index.intersection(stderr.index)
+				betas = betas.loc[idx]
+				stderr = stderr.loc[idx]
+			print(betas.shape)
+			print(stderr.shape)
+			betas.to_csv('combined_mash_effects/' + trait + '.' + pop + '.betas.' + effects + '.txt',sep = '\t',header = False, index = False)
+			stderr.to_csv('combined_mash_stderrs/' + trait + '.' + pop + '.stderrs.' + effects + '.txt',sep = '\t',header = False, index = False)
+			

### 2023-09-11 analysis/mash_greedy_algorithm/mash_switchgrass_models/mash/hypothesis_matrices_only/mash_bootstrap.R

@@ -1,0 +1,63 @@
+## Run GWAS 
+library(dplyr)
+library(bigsnpr)
+library(readr)
+library(mashr)
+library(AnnotationDbi)
+library(optparse)
+library(collections)
+library(bigparallelr)
+
+assert_cores(1)
+
+option_list = list(
+  make_option(c("-p","--pheno"), type="character", default=NULL,help="which site to use",metavar="character"),
+  make_option(c("-c","--cohort"), type="character", default=NULL,help="gulf, midwest, or all",metavar="character"),
+  make_option(c("-m","--matrices"), type="character", default=NULL,help="gulf, midwest, or all",metavar="character"),
+  make_option(c("-s","--step"), type="character", default=NULL,help="step in the greedy algorithm",metavar="character"),
+  make_option(c("-e","--effects"), type="character", default=NULL,help="gulf, midwest, or all",metavar="character"),
+  make_option(c("-b","--bootstrap"), type="character", default=NULL,help="gulf, midwest, or all",metavar="character")
+)
+
+opt_parser = OptionParser(option_list=option_list)
+opt = parse_args(opt_parser)
+
+if (is.null(opt$pheno)) {
+  print_help(opt_parser)
+  stop("Missing argument for phenotype code", call.=FALSE)
+}
+
+pheno = opt$pheno
+cohort = opt$cohort
+matrices = unlist(strsplit(opt$matrices,','))
+effect = opt$effects
+step = opt$step
+bootstrap=opt$bootstrap
+print(matrices)
+betas = read.table(paste0('bootstrap_combined_mash_effects/',pheno,'.',cohort,'.betas.',effect,'.',bootstrap,'.txt'))
+stderrs = read.table(paste0('bootstrap_combined_mash_stderrs/',pheno,'.',cohort,'.stderrs.',effect,'.',bootstrap,'.txt'))
+
+mat_betas = matrix(unlist(betas),ncol=8,byrow = TRUE)
+mat_stderrs = matrix(unlist(stderrs),ncol=8,byrow = TRUE)
+
+data=mash_set_data(mat_betas,mat_stderrs) 
+
+
+covmats = list()
+labels = list()
+for (matrix in matrices) {
+  label = unlist(strsplit(toString(matrix),'.',fixed = TRUE))
+  label=paste(label[2],label[3],sep='.')
+  temp = read.csv(matrix,header = TRUE, row.names = 1)
[…]

### 2023-09-11 analysis/mash_greedy_algorithm/mash_switchgrass_models/mash/hypothesis_matrices_only/mash_greedy.R

@@ -1,0 +1,64 @@
+## Run GWAS 
+library(dplyr)
+library(bigsnpr)
+library(readr)
+library(mashr)
+library(AnnotationDbi)
+library(optparse)
+library(collections)
+library(bigparallelr)
+library('matrixcalc')
+assert_cores(1)
+
+option_list = list(
+  make_option(c("-p","--pheno"), type="character", default=NULL,help="which site to use",metavar="character"),
+  make_option(c("-c","--cohort"), type="character", default=NULL,help="gulf, midwest, or all",metavar="character"),
+  make_option(c("-m","--matrices"), type="character", default=NULL,help="gulf, midwest, or all",metavar="character"),
+  make_option(c("-s","--step"), type="character", default=NULL,help="step in the greedy algorithm",metavar="character"),
+  make_option(c("-e","--effects"), type="character", default=NULL,help="gulf, midwest, or all",metavar="character")
+)
+
+opt_parser = OptionParser(option_list=option_list)
+opt = parse_args(opt_parser)
+
+if (is.null(opt$pheno)) {
+  print_help(opt_parser)
+  stop("Missing argument for phenotype code", call.=FALSE)
+}
+
+pheno = opt$pheno
+cohort = opt$cohort
+matrices = unlist(strsplit(opt$matrices,','))
+effect = opt$effects
+step = opt$step
+print(matrices)
+betas = read.table(paste0('combined_mash_effects/',pheno,'.',cohort,'.betas.',effect,'.txt'))
+stderrs = read.table(paste0('combined_mash_stderrs/',pheno,'.',cohort,'.stderrs.',effect,'.txt'))
+
+mat_betas = matrix(unlist(betas),ncol=8,byrow = TRUE)
+mat_stderrs = matrix(unlist(stderrs),ncol=8,byrow = TRUE)
+
+data=mash_set_data(mat_betas,mat_stderrs) 
+
+
+covmats = list()
+labels = list()
+for (matrix in matrices) {
+  label = unlist(strsplit(toString(matrix),'.',fixed = TRUE))
+  label=paste(label[2],label[3],sep='.')
+  temp = read.csv(matrix,header = TRUE, row.names = 1)
+  print(temp)
+  temp = matrix(unlist(temp),ncol=8,byrow = TRUE)
+  print(is.positive.semi.definite(temp))
+  covmats[[label]] = temp
+  labels = append(labels,label)
+}
+
[…]

### 2023-09-11 analysis/mash_greedy_algorithm/mash_switchgrass_models/mash/hypothesis_matrices_only/pod.bootstrap.sh

@@ -1,0 +1,1200 @@
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 1
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 2
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 3
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 4
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 5
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 6
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 7
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 8
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 9
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 10
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 11
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 12
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 13
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 14
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 15
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 16
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 17
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 18
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 19
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 20
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 21
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 22
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 23
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 24
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 25
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 26
+python run.mash.greedy.bootstrap.py -p greenup -c gulf -e random -b 27
[…]

### 2023-09-11 analysis/mash_greedy_algorithm/mash_switchgrass_models/mash/hypothesis_matrices_only/pod.combined.bootstrap.gwas.to.mash.sh

@@ -1,0 +1,100 @@
+python combine.bootstrap.gwas.to.mash.py 1
+python combine.bootstrap.gwas.to.mash.py 2
+python combine.bootstrap.gwas.to.mash.py 3
+python combine.bootstrap.gwas.to.mash.py 4
+python combine.bootstrap.gwas.to.mash.py 5
+python combine.bootstrap.gwas.to.mash.py 6
+python combine.bootstrap.gwas.to.mash.py 7
+python combine.bootstrap.gwas.to.mash.py 8
+python combine.bootstrap.gwas.to.mash.py 9
+python combine.bootstrap.gwas.to.mash.py 10
+python combine.bootstrap.gwas.to.mash.py 11
+python combine.bootstrap.gwas.to.mash.py 12
+python combine.bootstrap.gwas.to.mash.py 13
+python combine.bootstrap.gwas.to.mash.py 14
+python combine.bootstrap.gwas.to.mash.py 15
+python combine.bootstrap.gwas.to.mash.py 16
+python combine.bootstrap.gwas.to.mash.py 17
+python combine.bootstrap.gwas.to.mash.py 18
+python combine.bootstrap.gwas.to.mash.py 19
+python combine.bootstrap.gwas.to.mash.py 20
+python combine.bootstrap.gwas.to.mash.py 21
+python combine.bootstrap.gwas.to.mash.py 22
+python combine.bootstrap.gwas.to.mash.py 23
+python combine.bootstrap.gwas.to.mash.py 24
+python combine.bootstrap.gwas.to.mash.py 25
+python combine.bootstrap.gwas.to.mash.py 26
+python combine.bootstrap.gwas.to.mash.py 27
+python combine.bootstrap.gwas.to.mash.py 28
+python combine.bootstrap.gwas.to.mash.py 29
+python combine.bootstrap.gwas.to.mash.py 30
+python combine.bootstrap.gwas.to.mash.py 31
+python combine.bootstrap.gwas.to.mash.py 32
+python combine.bootstrap.gwas.to.mash.py 33
+python combine.bootstrap.gwas.to.mash.py 34
+python combine.bootstrap.gwas.to.mash.py 35
+python combine.bootstrap.gwas.to.mash.py 36
+python combine.bootstrap.gwas.to.mash.py 37
+python combine.bootstrap.gwas.to.mash.py 38
+python combine.bootstrap.gwas.to.mash.py 39
+python combine.bootstrap.gwas.to.mash.py 40
[…]

### 2023-09-11 analysis/mash_greedy_algorithm/mash_switchgrass_models/mash/hypothesis_matrices_only/run.mash.greedy.bootstrap.py

@@ -1,0 +1,76 @@
+import pandas as pd 
+import numpy as np 
+import glob
+import sys
+import os
+import argparse
+############
+#need to add arguments for:
+#pheno
+#strong/random
+#cohort from which effects were derived
+
+
+parser = argparse.ArgumentParser()
+parser.add_argument('-p','--phenotype',type=str,help='phenotype to analyze: greenup or flowering')
+parser.add_argument('-e','--effects',type=str,help='type of effects to analyze: random or strong')
+parser.add_argument('-c','--cohort',type=str,help='cohort to analyze: gulf, midwest, or all')
+parser.add_argument('-b','--bootstrap',type=str,help = 'bootstrap to analyze in this run')
+
+args = parser.parse_args()
+pheno = args.phenotype
+effects = args.effects
+cohort = args.cohort
+bootstrap = args.bootstrap
+
+likelihood_path = pd.read_csv('likelihood_paths/' + pheno + '.' + cohort + '.' + effects + '.csv')['matrix'].tolist()
+# #test with greenup, all, random
+all_hypothesis_matrices = glob.glob('hypothesis_matrices/' + pheno + '.*')
+# steps = len(all_hypothesis_matrices)
+# mostlikely=[]
+# testers = all_hypothesis_matrices
+
+# final_likelihoods = pd.DataFrame(np.zeros((len(all_hypothesis_matrices),2)),columns = ['matrix','likelihood'])
+counter = 1
+
+# while len(testers) > 0:
+for i,j in enumerate(likelihood_path):
+	thispath = likelihood_path[:i+1]
+	final = thispath[-1]
+	print(final)
+	matrices = ['hypothesis_matrices/' + i + '.U.mat.csv' for i in thispath]
+	
+	outputfile = final
+	
+	if os.path.isfile('bootstrap_likelihoods/step' + str(counter) + '/' +  outputfile + '.' + cohort + '.' + effects + '.' + str(bootstrap) + '.txt'):
+		print('skipped',i)
+		pass
+	else:
+		print('Rscript mash_bootstrap.R -m ' + ','.join(matrices) + ' -p ' + pheno + ' -c ' + cohort + ' -e ' + effects + ' -s ' + str(counter) + ' -b ' + str(bootstrap))
[…]

### 2023-09-11 analysis/mash_greedy_algorithm/mash_switchgrass_models/mash/hypothesis_matrices_only/run.mash.greedy.search.py

@@ -1,0 +1,87 @@
+import pandas as pd 
+import numpy as np 
+import glob
+import sys
+import os
+import argparse
+from scipy import stats
+
+############
+#need to add arguments for:
+#pheno
+#strong/random
+#cohort from which effects were derived
+
+
+parser = argparse.ArgumentParser()
+parser.add_argument('-p','--phenotype',type=str,help='phenotype to analyze: greenup or flowering')
+parser.add_argument('-e','--effects',type=str,help='type of effects to analyze: random or strong')
+parser.add_argument('-c','--cohort',type=str,help='cohort to analyze: gulf, midwest, or all')
+
+args = parser.parse_args()
+pheno = args.phenotype
+effects = args.effects
+cohort = args.cohort
+
+
+#test with greenup, all, random
+all_hypothesis_matrices = glob.glob('hypothesis_matrices/' + pheno + '.*')
+steps = len(all_hypothesis_matrices)
+mostlikely=[]
+testers = all_hypothesis_matrices
+
+final_likelihoods = pd.DataFrame(np.zeros((len(all_hypothesis_matrices),2)),columns = ['matrix','likelihood'])
+counter = 1
+
+while len(testers) > 0:
+
+	for matrix in testers:
+		outputfile = matrix.split('/')[1].split('.')
+		outputfile = '.'.join(outputfile[:3])
+
+		if os.path.isfile('likelihoods/step' + str(counter) + '/' +  outputfile + '.' + cohort + '.' + effects + '.txt'):
+			print('skipped')
+			pass
+		else:
+			print('Rscript mash_greedy.R -m ' + matrix + ',' + ','.join(mostlikely) + ' -p ' + pheno + ' -c ' + cohort + ' -e ' + effects + ' -s ' + str(counter))
+			os.system('Rscript mash_greedy.R -m ' + matrix + ',' + ','.join(mostlikely) + ' -p ' + pheno + ' -c ' + cohort + ' -e ' + effects + ' -s ' + str(counter) + '\n')
+
+	likelihood_files = glob.glob('likelihoods/step' + str(counter) + '/' + pheno + '.*.' + cohort + '.' + effects + '.txt')
+	step_likelihoods = pd.DataFrame(np.zeros((len(testers),2)),columns = ['matrix','likelihood'])
+
+	for e,f in enumerate(likelihood_files):
[…]

### 2023-09-11 analysis/mash_greedy_algorithm/mash_switchgrass_models/mash/hypothesis_matrices_only/write_pod_jobs.py

@@ -1,0 +1,16 @@
+import pandas as pd 
+import numpy as np 
+import sys
+
+newfile = open('pod.greedy.sh','w')
+for trait in ['greenup','flowering']:
+	for effect in ['random','strong']:
+		for cohort in ['gulf','midwest','all']:
+				newfile.write('python run.mash.greedy.search.py -p ' + trait + ' -c ' + cohort + ' -e ' + effect + '\n')
+
+newfile = open('pod.bootstrap.sh','w')
+for trait in ['greenup','flowering']:
+        for effect in ['random','strong']:
+                for cohort in ['gulf','midwest','all']:
+                        for bootstrap in range(1,101):
+                                newfile.write('python run.mash.greedy.bootstrap.py -p ' + trait + ' -c ' + cohort + ' -e ' + effect + ' -b ' + str(bootstrap) + '\n')

### 2023-09-11 R/Analysis_v1.2_mash_using_greedy_covar.Rmd
analysis: re-run mash using only greedy algorithm covariance matrices

@@
 ```
+
+# -----------------------------
+# Run mash
+
+```{r}
+library(tidyverse)
+library(bigsnpr)
+library(switchgrassGWAS)
+library(mashr)
+library(rlist)
+library(here)
+
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+mashdir <- here("analysis", "mash_greedy_algorithm")
+
+#phe_gr_df <- read_rds(here("data",
+#                           "Weather_related_phe_for_asreml_h2_GR50.rds"))
+# all greenup related phenotypes, including greenup as functions of weather variables.
+#phe_fl_df <- read_rds(here("data",
+#                           "Weather_related_phe_for_asreml_h2_FL50.rds"))
+# all flowering related phenotypes, including functions of weather variables.
+
+# Files with the six suffixes I have/want mash results for:
+numSNPs_v <- c(19000, 19000, 19000, 19000, 24000, 33000)
+suffix_v <- c("FL50_Gulf_and_Midwest", "FL50_Gulf", "GR50_Gulf_and_Midwest",
+              "GR50_Gulf", "GR50_Midwest", "FL50_Midwest")
+suffix_vli <- c("Gulf_and_Midwest", "Gulf", "Gulf_and_Midwest",
+              "Gulf", "Midwest", "Midwest")
+subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", 
+                 Gulf = "229g_10.3M")
+
+pop_sc <- c("all", "gulf", "all", "gulf", "midwest", "midwest")
+phe_sc <- c("flowering", "flowering", "greenup", "greenup", "greenup",
+            "flowering")
+
+Upath <- here("analysis", "mash_greedy_algorithm", "mash_switchgrass_models", 
+                "mash", "all_matrices", "matrices")
+stopcondpath <- here("analysis", "mash_greedy_algorithm",
+                     "mash_switchgrass_models", "mash", 
+                     "all_matrices", "likelihood_paths")
+
+for (i in 1:6) {  # change this for different scripts
+  # B_hat_strong_df_19000topSNPs_FL50_Gulf_and_Midwest.rds
+  # B_hat_strong_df_19000topSNPs_FL50_Gulf.rds
+  # B_hat_strong_df_19000topSNPs_GR50_Gulf_and_Midwest.rds
+  # B_hat_strong_df_19000topSNPs_GR50_Gulf.rds
+  # B_hat_strong_df_24000topSNPs_GR50_Midwest.rds
+  # B_hat_strong_df_33000topSNPs_FL50_Midwest.rds
+  ## mash Gulf greenup
+  numSNPs <- numSNPs_v[i]
+  suffix <- suffix_v[i]
+  # use this one for the likelihood paths
+  # Read in likelihood paths, then use that to read in matrices, then use those to run mash.
+
+  U_greedy <- 
+    read_csv(file = file.path(stopcondpath, 
+                              paste0(phe_sc[i], ".", pop_sc[i], 
+                                     ".random.stop.condition.csv"))) %>%
+    filter(matrix != "0.0")
+  ## Read in Uhyp
+
+  U_list <- list()
+  for (k in 1:nrow(U_greedy)) {
+    U_single <- read_csv(file = file.path(Upath, U_greedy$matrix[k]), show_col_types = FALSE)
+    # if conditions for Midwest
+    if (i == 5) {
+    # MI, MO, NE, OK, TX1, TX2, TX3 used for GR50  
+      U_single <- U_single %>%
+        filter(`...1` %in% c("MI", "MO", "NE", "OK", "TX1", "TX2", "TX3")) %>%
+        select(`...1`, .data$MI, .data$MO, .data$NE, .data$OK, .data$TX1,
+               .data$TX2, .data$TX3)
+    } else if (i == 6) {
+    # MI, MO, SD, TX1, TX2, TX3 used for FL50  
+      U_single <- U_single %>%
+        filter(`...1` %in% c("MI", "MO", "SD", "TX1", "TX2", "TX3")) %>%
+        select(`...1`, .data$MI, .data$MO, .data$SD, .data$TX1, .data$TX2,
+               .data$TX3)
+    }
+    U_single <- U_single %>%
+      select(-`...1`)
+    U_single <- as.matrix(unname(U_single))
+
+    # format cov matrices correctly for mash
+    U_list <- list.append(U_list, U_single)
+  }
+  names(U_list) <- str_remove(U_greedy$matrix, ".csv")
+
+  lipath <- here("analysis", "mash", suffix_vli[i], "inputs")
+
+  list_input <- switchgrassGWAS:::load_mash_df(path = lipath, 
+                                               numSNPs = numSNPs,
+                                               suffix = suffix)
+  # Just run mash itself as the greedy choice of cov matrices changes mash run
+  # enough that mash_standard_run will not perform correctly (eg it always adds
+  # all canonical covariance matrices).
+  Bhat_strong <- as.matrix(list_input$B_hat_strong)
+  Shat_strong <- as.matrix(list_input$S_hat_strong)
+  Bhat_random <- as.matrix(list_input$B_hat_random)
+  Shat_random <- as.matrix(list_input$S_hat_random)
+
+  data_r <- mashr::mash_set_data(Bhat_random, Shat_random)
+  Vhat <- mashr::estimate_null_correlation_simple(data = data_r)
+
+  data_strong <- mashr::mash_set_data(Bhat_strong, Shat_strong, V = Vhat)
+  data_random <- mashr::mash_set_data(Bhat_random, Shat_random, V = Vhat)
+
+  # Run mash on the random dataset using the random data w/ correlation structure
+  message(paste0("Fit mash to the random tests using both data-driven and ",
+                 "canonical covariances."))
+  # U_c <- mashr::cov_canonical(data_random)
+  m = mashr::mash(data_random, Ulist = U_list, outputlevel = 1)
+
+  # Run mash on the strong dataset (or all data) using
+  # the previous results from the random data
+  message(paste0("Compute posterior matrices for the strong effects",
+                 " using the mash fit from the
+                   random tests."))
+  m2 = mashr::mash(data_strong, g = get_fitted_g(m), fixg = TRUE)
+
+  saveRDS(m2, file = file.path(mashdir, suffix_vli[i], 
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_random_greedy_Ulist", 
+                                          numSNPs, "_SNPs_",
+                                          suffix, ".rds")))
+
+  print(paste0(suffix, " mash outputs: "))
+  print("Log likelihood with specified covariance matrices: ")
+  print(get_loglik(m2), digits = 10)
+  print("How many significant markers?")
+  print(length(get_significant_results(m2)))
+  switchgrassGWAS::mash_plot_covar(m2)
+  switchgrassGWAS::mash_plot_manhattan_by_condition(m2)
+  switchgrassGWAS::mash_plot_pairwise_sharing(m2)
+  gxe <- switchgrassGWAS::get_GxE(m2)
+  saveRDS(gxe, file = file.path(mashdir, suffix_vli[i], 
+                                   paste0("GxE_default_",
+                                          "est_w_random_greedy_Ulist",
+                                          numSNPs, "_SNPs_",
+                                          suffix, "_greedy_Ulist.rds")))
+}
+```

### 2023-09-11 R/nohup_v1.2_mash_greedy_greenup_gulf_strong.R

@@ -94,8 +94,8 @@ Shat_random <- as.matrix(list_input$S_hat_random)
 data_r <- mashr::mash_set_data(Bhat_random, Shat_random)
 Vhat <- mashr::estimate_null_correlation_simple(data = data_r)
 
-data_strong <- mashr::mash_set_data(Bhat_strong, Shat_strong, V=Vhat)
-data_random <- mashr::mash_set_data(Bhat_random, Shat_random, V=Vhat)
+data_strong <- mashr::mash_set_data(Bhat_strong, Shat_strong, V = Vhat)
+data_random <- mashr::mash_set_data(Bhat_random, Shat_random, V = Vhat)
 
 U_single <- as.matrix(U_single)
 # Run mash on the random dataset using the random data w/ correlation structure

### 2023-09-12 R/Analysis_v1.2_mash_using_greedy_covar.Rmd
analysis: Substitute h2 for CV on diagonals of hypothesis covar matrices

@@
 
-##
+# ----
 
-# ----
+# Covar 1-20d weather
+As of 2023-09-11, want to remake these per reviewer revision to have h2 on the diagonal instead of the CV. Need to estimate h2 for these new covar matrices, probably using rrBLUP.
+
 ## load phe & wea data
@@
 phe_ten_sites <- readRDS(here("data", "Phenotypes_2018_to_2021_10_sites.rds"))
+sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
 jday_phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
@@
 
+# ---
 # Covar csv for greedy mash algorithm
@@
 
+## Calculate heritability for diagonal
+I don't have access to ASREML any more, so need a different package to do this.
+```{r}
+library(rrBLUP)
+k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
+fl_ten_sites
+colnames(k_full)
+
+phe_one <- fl_ten_sites %>% filter(manu_site == "TX1") %>%
+  filter(WHERE != "FWCR") %>%
+  group_by(PLANT_ID) %>%
+  summarise(crain_1d = mean(crain_1d, na.rm = TRUE)) %>%
+  filter(!is.na(PLANT_ID))
+k_in_phe <- which(colnames(k_full) %in% phe_one$PLANT_ID)
+k_one <- k_full[k_in_phe, k_in_phe]
+
+nrow(k_one)
+nrow(phe_one)
+mod2 <- mixed.solve(y = phe_one$crain_1d, K = k_one)
+h2 <- mod2$Vu/(mod2$Vu + mod2$Ve)
+```
+
 ```{r}
@@
   rename(Year = YEAR)
+fl_ten_sites$PLANT_ID <- factor(fl_ten_sites$PLANT_ID, levels = colnames(k_full))
 
@@
 phe_hyp <- colnames(fl_ten_sites)[c(5,9:51)]
-i=1
+i=2
 k=1
@@
              IND = as.factor(IND)) %>%
-      select(PLANT_ID, SUBPOP, IND, MI, MO, NE, OK, SD, TX1, TX2, TX3) #####  IL,
+      select(PLANT_ID, SUBPOP, IND, MI, MO, NE, OK, SD, TX1, TX2, TX3) %>% #####  IL,
+      replace_na(list(MI = 0, MO = 0, NE = 0, OK = 0, SD = 0, TX1 = 0, TX2 = 0, 
+                      TX3 = 0))
 
     cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
-    # Set diagonal of this matrix = coefficient of variation within each garden.
-    diag(cor_phe) <- cov_df %>%
-      ungroup() %>%
-      summarise(#IL = sqrt(var(IL, na.rm = TRUE))/mean(IL, na.rm = TRUE),  #####
-                MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),  #####
-                MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
-                NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
-                OK = sqrt(var(OK, na.rm = TRUE))/mean(OK, na.rm = TRUE),
-                SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
-                TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
-                TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
-                TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE)) %>%
-      as_vector()
+    # Set diagonal of this matrix = heritability within each garden
+    h2_diag <- c()
+    site_names <- c("MI", "MO", "NE", "OK", "SD", "TX1", "TX2", "TX3")
+    for(j in seq_along(site_names)) {
+      phe_one <- fl_ten_sites %>% filter(manu_site == site_names[j]) %>%
+        filter(WHERE != "FWCR") %>%
+        group_by(PLANT_ID) %>%
+        summarise(phenotype = mean(!! sym(phe_hyp[i]), na.rm = TRUE)) %>%
+        filter(!is.na(PLANT_ID))
+      k_in_phe <- which(colnames(k_full) %in% phe_one$PLANT_ID)
+      k_one <- k_full[k_in_phe, k_in_phe]
+      mod2 <- mixed.solve(y = phe_one$phenotype, K = k_one)
+      h2_diag[j] <- mod2$Vu/(mod2$Vu + mod2$Ve)
+    }
 
+    diag(cor_phe) <- h2_diag
+    if (!matrixcalc::is.positive.semi.definite(cor_phe)) {
+      cor_phe <- Matrix::nearPD(cor_phe)
+    }
     U_hyp_2019 <- list.prepend(U_hyp_2019, cor_phe)
-    names(U_hyp_2019)[1] <- paste0("flowering.", names(subpop_v2)[k], ".", phe_hyp[i], ".U.mat")
+    names(U_hyp_2019)[1] <- paste0("flowering.", names(subpop_v2)[k], ".", 
+                                   phe_hyp[i], ".U.mat")
   }
@@
 U_hyp_2019
-length(U_hyp_2019)
-for (i in 1:length(U_hyp_2019)) {
-  d_out <- as.data.frame(U_hyp_2019[[i]], optional = FALSE) %>%
-    rownames_to_column(var = " ")
-  write_csv(d_out, file = here("data", "greedy_more_cov", "hypothesis_matrices", paste0(names(U_hyp_2019)[i], ".csv")))
+
+# Test that every matrix is positive semi definite. Should all be true now changed code above.
+U_psd <- c() 
+for (i in seq_along(U_hyp_2019)) {
+  if (is.matrix(U_hyp_2019[[i]])) {
+    U_psd[i] <- matrixcalc::is.positive.semi.definite(U_hyp_2019[[i]])
+  } else {
+    U_psd[i] <- matrixcalc::is.positive.semi.definite(as.matrix(U_hyp_2019[[i]]$mat))
   }
+
+}
+
+for (i in seq_along(U_hyp_2019)) {
+  if (is.matrix(U_hyp_2019[[i]])) {
+    matrixout = matrix(unlist(U_hyp_2019[[i]]), ncol = 8, byrow = TRUE)
+  } else {
+    matrixout = matrix(unlist(U_hyp_2019[[i]]$mat), ncol = 8, byrow = TRUE)
+  }
+  row.names(matrixout) = site_names
+  colnames(matrixout) = site_names
+  #print(matrixout)
+  write.csv(matrixout, file = here("analysis", "mash_greedy_algorithm", 
+                                   "mash_h2_diag_covar", "matrices", 
+                                   paste0(names(U_hyp_2019)[i], ".csv")))
+}
+
 ```
+Another error: "Error in mash(data, covmats) : 
+  All U_scaled matrices should be positive semi-definite
 
@@
     cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
-    # Set diagonal of this matrix = coefficient of variation within each garden.
-    diag(cor_phe) <- cov_df %>%
-      ungroup() %>%
-      summarise(#IL = sqrt(var(IL, na.rm = TRUE))/mean(IL, na.rm = TRUE),  #####
-                MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),  #####
-                MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
-                NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
-                OK = sqrt(var(OK, na.rm = TRUE))/mean(OK, na.rm = TRUE),
-                SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
-                TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
-                TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
-                TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE)) %>%
-      as_vector()
+    # Set diagonal of this matrix = heritability within each garden
+    h2_diag <- c()
+    site_names <- c("MI", "MO", "NE", "OK", "SD", "TX1", "TX2", "TX3")
+    for(j in seq_along(site_names)) {
+      phe_one <- fl_ten_sites %>% filter(manu_site == site_names[j]) %>%
+        filter(WHERE != "FWCR") %>%
+        group_by(PLANT_ID) %>%
+        summarise(phenotype = mean(!! sym(phe_hyp[i]), na.rm = TRUE)) %>%
+        filter(!is.na(PLANT_ID))
+      k_in_phe <- which(colnames(k_full) %in% phe_one$PLANT_ID)
+      k_one <- k_full[k_in_phe, k_in_phe]
+      mod2 <- mixed.solve(y = phe_one$phenotype, K = k_one)
+      h2_diag[j] <- mod2$Vu/(mod2$Vu + mod2$Ve)
+    }
+
+    diag(cor_phe) <- h2_diag
+    if (!matrixcalc::is.positive.semi.definite(cor_phe)) {
+      cor_phe <- Matrix::nearPD(cor_phe)
+    }
 
@@
 }
-for (i in 1:length(U_hyp_2019)) {
-  d_out <- as.data.frame(U_hyp_2019[[i]], optional = FALSE) %>%
-    rownames_to_column(var = " ")
-  write_csv(d_out, file = here("data", "greedy_more_cov", "hypothesis_matrices", paste0(names(U_hyp_2019)[i], ".csv")), )
+
+for (i in seq_along(U_hyp_2019)) {
+  if (is.matrix(U_hyp_2019[[i]])) {
+    matrixout = matrix(unlist(U_hyp_2019[[i]]), ncol = 8, byrow = TRUE)
+  } else {
+    matrixout = matrix(unlist(U_hyp_2019[[i]]$mat), ncol = 8, byrow = TRUE)
   }
+  row.names(matrixout) = site_names
+  colnames(matrixout) = site_names
+  #print(matrixout)
+  write.csv(matrixout, file = here("analysis", "mash_greedy_algorithm", 
+                                   "mash_h2_diag_covar", "matrices", 
+                                   paste0(names(U_hyp_2019)[i], ".csv")))
+}
+
 ```
@@
 ```
+
+# ---------------
+# Info to run greedy mash algo
+
+data/mash_switchgrass_models/mash/all_matrices/matrices hosts all covar mat (?)
+analysis/mash_greedy_algorithm/mash_switchgrass_models/mash/all_matrices/matrices has more? covar mat?
+
+Set up new analysis/mash_greedy_algorithm/mash_1_to_28d_covar/ or subdirectory to have all the subdir necessary to run `run.mash.greedy.search.py` python script.
+
+Make sure all matrices are present in matrices/ subdir and named appropriately for recognition
+Make sure random and strong Bhat and Shat in combined_mash_effects and combined_mash_stderrs are appropriate - perhaps replace these with my inputs in mash/Subpop/inputs; perhaps use what's here as tests first
+
+May need to adjust midwest covar & effects appropriately to reflect only 6 (floewring) and 7 (greenup) rows & columns used; other GWAS were dropped & I think including them would be misleading, likely.
 

### 2023-09-12 R/Analysis_v1.2_mash_using_greedy_covar.Rmd
analysis: rerun greedy mash algorithm with h2 diag covar and known random input df

@@
 library(zoo)
+library(rrBLUP)
 ```
@@
 
-# Deprecated analysis
-### Code from Plots_v0.2_manuscript_figures.Rmd to make weather variables
+# ----
+# Generate weather variables
 
-Actually, code from Analysis_v1.1_weather_2020.Rmd might be more suitable, if I am remaking these anyway.
-
-```{r}
-wea_df <- readRDS("~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")
-metadata <- readRDS("~/Github/pvdiv-phenology-gxe/data/metadata.rds")
-sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
-phenotypes <- readRDS("~/Github/pvdiv-phenology-gxe/data/Phenology_CGDD_PTT_and_rainfall_phenotypes.rds")
-jday_phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
-```
-
-```{r}
-jday_phe <- wea_df %>%
-  group_by(SITE) %>%
-  mutate(dyln_change_sec = (DYLN - lag(DYLN))*60*60,
-         FL50 = yday(DOY)) %>%
-  select(DOY, DYLN, dyln_change_sec, everything()) %>%
-  select(-PTT_12C) %>%
-  right_join(jday_phe, by = c("SITE", "FL50")) %>%
-  select(-(TMAX:CPTT_11C))
-jday_phe <- metadata %>%
-  rename(Lat_origin = LATITUDE, Long_origin = LONGITUDE) %>%
-  select(PLANT_ID, SUBPOP, Lat_origin, Long_origin) %>%
-  right_join(jday_phe) %>%
-  left_join(sites)
-
-jday_phe$SUBPOP <- factor(jday_phe$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
-jday_phe$SITE <- factor(jday_phe$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
-jday_phe$manu_site <- factor(jday_phe$manu_site, levels = rev(c("TX1", "TX2", "TX3", "TX4", "OK", "MO", "NE", "IL", "MI", "SD")))
-```
-
-## define environmental cues for greenup
-Probably all CGDD related.
-```{r}
-
-asreml_fl <- jday_phe %>%
-  select(PLANT_ID:Long_origin, PLOT_GL, SITE, FL50, GR50, CGDD_12C, CRAIN, CRAIN_5d, DYLN, dyln_change_sec) %>%
-  rename(dyln_fl50 = DYLN, cgdd_12c_gr2fl = CGDD_12C, crain_gr2fl = CRAIN) %>%
-  left_join(wea_df) %>%
-  group_by(SITE, PLOT_GL) %>%
-  filter(between(DOY, as_date(FL50, origin = "2019-01-01") - 6, as_date(FL50, origin = "2019-01-01"))) %>%
-  mutate(crain_2d = RAIN_SM + lag(RAIN_SM),
-         crain_3d = RAIN_SM + lag(RAIN_SM) + lag(RAIN_SM, n = 2),
-         crain_7d = sum(RAIN_SM)) %>%
-  filter(DOY == as_date(FL50, origin = "2019-01-01")) %>%
-  filter(SUBPOP %in% c("Midwest", "Gulf", "Atlantic") & !(SITE %in% c("OVTN"))) %>%
-  select(-(TMAX:TMIN_SM), -(GDD_13C_SM:CPTT_13C)) %>%
-  rename(crain_1d = RAIN_SM, crain_5d = CRAIN_5d)
-saveRDS(asreml_fl, file = "../data/Weather_related_phe_for_asreml_h2_FL50.rds")
-asreml_gr <- jday_phe %>%
-  select(PLANT_ID:Long_origin, PLOT_GL, SITE, GR50) %>%
-  left_join(wea_df) %>%
-  group_by(SITE, PLOT_GL) %>%
-  filter(DOY <= as_date(GR50, origin = "2019-01-01")) %>%
-  mutate(cgdd_12c_jan2gr = sum(GDD_12C_SM),
-         cgdd_8c_jan2gr = sum(GDD_8C_SM)
-         ) %>%
-  select(-(GDD_13C_SM:CPTT_13C), -(TMAX:Longitude), GDD_12C_SM) %>%
-  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 17, as_date(GR50, origin = "2019-01-01"))) %>%
-  mutate(tmax_18d = mean(TMAX_SM),
-         tmin_18d = mean(TMIN_SM),
-         tave_18d = (tmax_18d + tmin_18d)/2,
-         cgdd_12c_18d = mean(GDD_12C_SM)) %>%
-  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 9, as_date(GR50, origin = "2019-01-01"))) %>%
-  mutate(tmax_10d = mean(TMAX_SM),
-         tmin_10d = mean(TMIN_SM),
-         tave_10d = (tmax_10d + tmin_10d)/2,
-         cgdd_12c_10d = mean(GDD_12C_SM)) %>%
-  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 4, as_date(GR50, origin = "2019-01-01"))) %>%
-  mutate(tmax_5d = mean(TMAX_SM),
-         tmin_5d = mean(TMIN_SM),
-         tave_5d = (tmax_5d + tmin_5d)/2,
-         cgdd_12c_5d = mean(GDD_12C_SM))%>%
-  filter(DOY == as_date(GR50, origin = "2019-01-01")) %>%
-  filter(SUBPOP %in% c("Midwest", "Gulf", "Atlantic") & !(SITE %in% c("OVTN"))) %>%
-  select(-(TMAX_SM:RAIN_SM), -Day_of_year)
-saveRDS(asreml_gr, file = "../data/Weather_related_phe_for_asreml_h2_GR50.rds")
-```
-
-# ----
-
-# Covar 1-20d weather
+# Covar 1-28d weather
 As of 2023-09-11, want to remake these per reviewer revision to have h2 on the diagonal instead of the CV. Need to estimate h2 for these new covar matrices, probably using rrBLUP.
@@
 
-# ---
-# Covar csv for greedy mash algorithm
-
-Make & save hypothesis matrices in a hypothesis_matrices directory in data. 
-Then hopefully it's not much of a lift for Sam to run the greedy mash algorithm again. Otherwise I will figure out how to run it.
-
-phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", 
-             "tave_5d", "tave_10d", "tave_18d")
-phe_hyp <- c("FL50", "GR50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
-             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")
-
+## save weather-based phenotypes
 ```{r}
@@
 ```
+# --- can start here
+# Covar csv for greedy mash algorithm
+
+Make & save hypothesis matrices in a hypothesis_matrices directory in data. 
+Then run the greedy mash algorithm again.
+
+phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", 
+             "tave_5d", "tave_10d", "tave_18d")
+phe_hyp <- c("FL50", "GR50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
+             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")
 
@@
 I don't have access to ASREML any more, so need a different package to do this.
+
+### sketch out how h2 is calculated in rrBLUP
+
+Use the kinship matrix I've already calculated (using the van Raden method, using switchgrassGWAS::pvdiv_kinship()) using the SNPs as the K matrix.
 ```{r}
@@
 
+Another error: "Error in mash(data, covmats) : 
+  All U_scaled matrices should be positive semi-definite
+
+So I should adjust these to make sure all are positive semi-definite, too, in addition to replacing the diagonals with h2.
+
+### Loop to calculate covar for flowering
 ```{r}
@@
 
+### Loop to calculate covar for greenup
 ```{r}
@@
 
-Visualize correlation matrices
+### Midwest covar fewer sites
+
+#### flowering: 6 sites
+MI, MO, SD, TX1, TX2, TX3
+```{r flowering midwest}
+
+phe_hyp <- colnames(fl_ten_sites)[c(5,9:51)]
+i=2
+k=2
+
+U_hyp_2019 <- list()
+for(i in seq_along(phe_hyp)){
+
+    cov_df <- fl_ten_sites %>%
+      ungroup() %>%
+      filter(SUBPOP %in% subpop_v2[[k]]) %>%
+      filter(Year %in% 2019) %>%
+      select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
+      group_by(PLANT_ID, SUBPOP, manu_site) %>%
+      mutate(IND = row_number()) %>%
+      pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
+      mutate(PLANT_ID = as.factor(PLANT_ID),
+             IND = as.factor(IND)) %>%
+      select(PLANT_ID, SUBPOP, IND, MI, MO, SD, TX1, TX2, TX3) %>% #####  IL,
+      replace_na(list(MI = 0, MO = 0, SD = 0, TX1 = 0, TX2 = 0, 
+                      TX3 = 0))
+
+    cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
+    # Set diagonal of this matrix = heritability within each garden
+    h2_diag <- c()
+    site_names <- c("MI", "MO", "SD", "TX1", "TX2", "TX3")
+    for(j in seq_along(site_names)) {
+      phe_one <- fl_ten_sites %>% filter(manu_site == site_names[j]) %>%
+        filter(WHERE != "FWCR") %>%
+        group_by(PLANT_ID) %>%
+        summarise(phenotype = mean(!! sym(phe_hyp[i]), na.rm = TRUE)) %>%
+        filter(!is.na(PLANT_ID))
+      k_in_phe <- which(colnames(k_full) %in% phe_one$PLANT_ID)
+      k_one <- k_full[k_in_phe, k_in_phe]
+      mod2 <- mixed.solve(y = phe_one$phenotype, K = k_one)
+      h2_diag[j] <- mod2$Vu/(mod2$Vu + mod2$Ve)
+    }
+
+    diag(cor_phe) <- h2_diag
+    if (!matrixcalc::is.positive.semi.definite(cor_phe)) {
+      cor_phe <- Matrix::nearPD(cor_phe)
+    }
+    U_hyp_2019 <- list.prepend(U_hyp_2019, cor_phe)
+    names(U_hyp_2019)[1] <- paste0("flowering.", names(subpop_v2)[k], ".", 
+                                   phe_hyp[i], ".U.mat")
+}
+
+U_hyp_2019
+
+# Test that every matrix is positive semi definite. Should all be true now changed code above.
+U_psd <- c() 
+for (i in seq_along(U_hyp_2019)) {
+  if (is.matrix(U_hyp_2019[[i]])) {
+    U_psd[i] <- matrixcalc::is.positive.semi.definite(U_hyp_2019[[i]])
+  } else {
+    U_psd[i] <- matrixcalc::is.positive.semi.definite(as.matrix(U_hyp_2019[[i]]$mat))
+  }
+
+}
+
+for (i in seq_along(U_hyp_2019)) {
+  if (is.matrix(U_hyp_2019[[i]])) {
+    matrixout = matrix(unlist(U_hyp_2019[[i]]), ncol = 6, byrow = TRUE)
+  } else {
+    matrixout = matrix(unlist(U_hyp_2019[[i]]$mat), ncol = 6, byrow = TRUE)
+  }
+  row.names(matrixout) = site_names
+  colnames(matrixout) = site_names
+  #print(matrixout)
+  write.csv(matrixout, file = here("analysis", "mash_greedy_algorithm", 
+                                   "mash_h2_diag_covar", "midwest_flowering", "matrices", 
+                                   paste0(names(U_hyp_2019)[i], ".csv")))
+}
+
+```
+
+```{r}
+workingdir <- here("analysis", "mash_greedy_algorithm", 
+                                   "mash_h2_diag_covar", "midwest_flowering")
+pheno = "flowering"
+betas = read.table(here(workingdir, "combined_mash_effects", paste0( pheno,'.midwest.betas.random.txt')))
+stderrs = read.table(here(workingdir, "combined_mash_stderrs", paste0( pheno,'.midwest.stderrs.random.txt')))
+
+mat_betas = matrix(unlist(betas),ncol=6,byrow = TRUE)
+mat_stderrs = matrix(unlist(stderrs),ncol=6,byrow = TRUE)
+
+sites = c('MI','MO','SD','TX1','TX2','TX3')
+data=mash_set_data(mat_betas,mat_stderrs)
+
+canonical = cov_canonical(data = data)
+matnames = ls(canonical)
+for (matrix in matnames) {
+  if (matrix %in% c('equal_effects','identity','simple_het_1','simple_het_2','simple_het_3')) {
+  matrixout = matrix(unlist(canonical[matrix]), ncol = 6, byrow = TRUE)
+  row.names(matrixout) = sites
+  colnames(matrixout) = sites
+  write.csv(matrixout, file = here(workingdir, "matrices", paste0(pheno,'.', matrix, '.csv')))
+  }
+}
+print(canonical)
+for (i in 1:6) {
+  singleton = paste0('singletons_',i)
+  print(singleton)
+  singmat = canonical[singleton]
+  site = sites[i]
+
+  singmat = matrix(unlist(canonical[singleton]),ncol = 6, byrow = TRUE)
+  row.names(singmat) = sites
+  colnames(singmat) = sites
+  write.csv(singmat, file = here(workingdir, "matrices", paste0(pheno, '.singleton.', site, '.csv')))
+}
+
+canonical = cov_pca(data,5)
+cohort = "midwest"
+effect = "random"
+matnames = ls(canonical)
+for (mat in matnames) {
+  matrixout = matrix(unlist(canonical[mat]),ncol=6,byrow=TRUE)
+  row.names(matrixout) = sites
+  colnames(matrixout) = sites
+  print(matrixout)
+  write.csv(matrixout, file = here(workingdir, "matrices",
+                                   paste0(pheno,'.', cohort,'.', effect,'.', mat, '.csv')))
+}
+
+```
+
+#### greenup: 7 sites
+MI, MO, NE, OK, TX1, TX2, TX3
+```{r}
+phe_hyp <- colnames(gr_ten_sites)[8:50]
+i=1
+k=2
+
+U_hyp_2019 <- list()
+for(i in seq_along(phe_hyp)){
+
+    cov_df <- gr_ten_sites %>%
+      ungroup() %>%
+      filter(SUBPOP %in% subpop_v2[[k]]) %>%
+      filter(Year %in% 2019) %>%
+      select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
+      group_by(PLANT_ID, SUBPOP, manu_site) %>%
+      mutate(IND = row_number()) %>%
+      pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
+      mutate(PLANT_ID = as.factor(PLANT_ID),
+             IND = as.factor(IND)) %>%
+      select(PLANT_ID, SUBPOP, IND, MI, MO, NE, OK, TX1, TX2, TX3) #####  IL,
+
+    cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
+    # Set diagonal of this matrix = heritability within each garden
+    h2_diag <- c()
+    site_names <- c("MI", "MO", "NE", "OK", "TX1", "TX2", "TX3")
+    for(j in seq_along(site_names)) {
+      phe_one <- fl_ten_sites %>% filter(manu_site == site_names[j]) %>%
+        filter(WHERE != "FWCR") %>%
+        group_by(PLANT_ID) %>%
+        summarise(phenotype = mean(!! sym(phe_hyp[i]), na.rm = TRUE)) %>%
+        filter(!is.na(PLANT_ID))
+      k_in_phe <- which(colnames(k_full) %in% phe_one$PLANT_ID)
+      k_one <- k_full[k_in_phe, k_in_phe]
+      mod2 <- mixed.solve(y = phe_one$phenotype, K = k_one)
+      h2_diag[j] <- mod2$Vu/(mod2$Vu + mod2$Ve)
+    }
+
+    diag(cor_phe) <- h2_diag
+    if (!matrixcalc::is.positive.semi.definite(cor_phe)) {
+      cor_phe <- Matrix::nearPD(cor_phe)
+    }
+
+    U_hyp_2019 <- list.prepend(U_hyp_2019, cor_phe)
+    names(U_hyp_2019)[1] <- paste0("greenup.", names(subpop_v2)[k], ".", phe_hyp[i], ".U.mat")
+}
+
+for (i in seq_along(U_hyp_2019)) {
+  if (is.matrix(U_hyp_2019[[i]])) {
+    matrixout = matrix(unlist(U_hyp_2019[[i]]), ncol = 7, byrow = TRUE)
+  } else {
+    matrixout = matrix(unlist(U_hyp_2019[[i]]$mat), ncol = 7, byrow = TRUE)
+  }
+  row.names(matrixout) = site_names
+  colnames(matrixout) = site_names
+  #print(matrixout)
+  write.csv(matrixout, file = here("analysis", "mash_greedy_algorithm", 
+                                   "mash_h2_diag_covar", "midwest_greenup",
+                                   "matrices", 
+                                   paste0(names(U_hyp_2019)[i], ".csv")))
+}
+
+```
+
+```{r}
+workingdir <- here("analysis", "mash_greedy_algorithm", 
+                                   "mash_h2_diag_covar", "midwest_greenup")
+pheno = "greenup"
+betas = read.table(here(workingdir, "combined_mash_effects", paste0( pheno,'.midwest.betas.random.txt')))
+stderrs = read.table(here(workingdir, "combined_mash_stderrs", paste0( pheno,'.midwest.stderrs.random.txt')))
+
+mat_betas = matrix(unlist(betas),ncol=7,byrow = TRUE)
+mat_stderrs = matrix(unlist(stderrs),ncol=7,byrow = TRUE)
+
+sites = c('MI','MO','NE', 'OK','TX1','TX2','TX3')
+data=mash_set_data(mat_betas,mat_stderrs)
+
+canonical = cov_canonical(data = data)
+matnames = ls(canonical)
+for (matrix in matnames) {
+  if (matrix %in% c('equal_effects','identity','simple_het_1','simple_het_2','simple_het_3')) {
+  matrixout = matrix(unlist(canonical[matrix]), ncol = 7, byrow = TRUE)
+  row.names(matrixout) = sites
+  colnames(matrixout) = sites
+  write.csv(matrixout, file = here(workingdir, "matrices", paste0(pheno,'.', matrix, '.csv')))
+  }
+}
+print(canonical)
+for (i in 1:7) {
+  singleton = paste0('singletons_',i)
+  print(singleton)
+  singmat = canonical[singleton]
+  site = sites[i]
+
+  singmat = matrix(unlist(canonical[singleton]),ncol = 7, byrow = TRUE)
+  row.names(singmat) = sites
+  colnames(singmat) = sites
+  write.csv(singmat, file = here(workingdir, "matrices", paste0(pheno, '.singleton.', site, '.csv')))
+}
+
+canonical = cov_pca(data,5)
+cohort = "midwest"
+effect = "random"
+matnames = ls(canonical)
+for (mat in matnames) {
+  matrixout = matrix(unlist(canonical[mat]),ncol = 7,byrow=TRUE)
+  row.names(matrixout) = sites
+  colnames(matrixout) = sites
+  print(matrixout)
+  write.csv(matrixout, file = here(workingdir, "matrices",
+                                   paste0(pheno,'.', cohort,'.', effect,'.', mat, '.csv')))
+}
+
+```
+
+#### Visualize correlation matrices
 ```{r}
@@
 analysis/mash_greedy_algorithm/mash_switchgrass_models/mash/all_matrices/matrices has more? covar mat?
+Can cp from these matrices dir to mash_h2_diag_covar/matrices the following regex, which run.mash.greedy.search.py looks for:
+
+ '.singleton.*'
+'.identity.*'
+'.simple_het_*.*'
+ '.equal_effects.*'
+'.*PCA*'
 
 Set up new analysis/mash_greedy_algorithm/mash_1_to_28d_covar/ or subdirectory to have all the subdir necessary to run `run.mash.greedy.search.py` python script.
+I think these may have to be pre-generated for the script to work, and I'll make sure they're empty to keep there from being I/O issues I haven't tested for.
 
-Make sure all matrices are present in matrices/ subdir and named appropriately for recognition
-Make sure random and strong Bhat and Shat in combined_mash_effects and combined_mash_stderrs are appropriate - perhaps replace these with my inputs in mash/Subpop/inputs; perhaps use what's here as tests first
+Make sure all matrices are present in matrices/ subdir and named & output appropriately for recognition
 
-May need to adjust midwest covar & effects appropriately to reflect only 6 (floewring) and 7 (greenup) rows & columns used; other GWAS were dropped & I think including them would be misleading, likely.
+Make sure random and strong Bhat and Shat in combined_mash_effects and combined_mash_stderrs are appropriate - perhaps ultimately replace these with my inputs in mash/Subpop/inputs; perhaps use what's here as tests first
+
+When I change to my inputs for effects & stderrs, I will need to adjust midwest covar & effects appropriately to reflect only 6 (flowering) and 7 (greenup) rows & columns used; the other GWAS were dropped as too low power or due to substantial phenotype data loss, & I think including them would be misleading, likely.
+
+### Next steps
+Redo Midwest: Remake Midwest covar matrices, including hypothesis, singleton, identity, simple het, equal effects, and data-driven matrices.
+Make mash_greedy.R variants for changed input dir & for different Midwest column numbers.
+Run two new run.mash.greedy.search.py commands:
+python run.mash.greedy.search.py -p greenup -c midwest -e random
+python run.mash.greedy.search.py -p flowering -c midwest -e random
+
+Try Gulf & All with my known inputs:
+python run.mash.greedy.search.py -p greenup -c gulf -e random
+python run.mash.greedy.search.py -p greenup -c all -e random
+python run.mash.greedy.search.py -p flowering -c gulf -e random
+python run.mash.greedy.search.py -p flowering -c all -e random
+
+Write inputs to mash_h2_diag_covar/All_inputs, Gulf_inputs, and Midwest_inputs, from the original mash work. Especially important for Midwest which has some poor GWAS at dropped sites.
+The inputs Sam used have no column or rownames and look to be tab delimited.
+
+aka combined_mash_effects/greenup.all.betas.random.txt 
+and combined_mash_stderrs/flowering.gulf.betas.strong.txt
+
+I only really need the random ones here, run mash again on the strong effects below.
+```{r}
+popkey = list(all = "Gulf_and_Midwest", midwest = "Midwest", gulf = "Gulf")
+phekey = list(greenup = "GR50", flowering = "FL50")
+# i=2; j=1
+
+for (i in seq_along(popkey)) {
+  for (j in seq_along(phekey)) {
+
+    if (i == 2 & j == 1) {
+      betas <- readRDS(file = here("analysis", "mash", popkey[[i]], "inputs", 
+                                 paste0("B_hat_random_df_24000topSNPs_",
+                                        phekey[[j]], "_", popkey[[i]], ".rds")))
+      stderr <- readRDS(file = here("analysis", "mash", popkey[[i]], "inputs", 
+                                 paste0("S_hat_random_df_24000topSNPs_",
+                                        phekey[[j]], "_", popkey[[i]], ".rds")))
+    } else if (i == 2 & j == 2) {
+      betas <- readRDS(file = here("analysis", "mash", popkey[[i]], "inputs", 
+                                 paste0("B_hat_random_df_33000topSNPs_",
+                                        phekey[[j]], "_", popkey[[i]], ".rds")))
+      stderr <- readRDS(file = here("analysis", "mash", popkey[[i]], "inputs", 
+                                 paste0("S_hat_random_df_33000topSNPs_",
+                                        phekey[[j]], "_", popkey[[i]], ".rds")))
+    } else {
+      betas <- readRDS(file = here("analysis", "mash", popkey[[i]], "inputs", 
+                                 paste0("B_hat_random_df_19000topSNPs_",
+                                        phekey[[j]], "_", popkey[[i]], ".rds")))
+      stderr <- readRDS(file = here("analysis", "mash", popkey[[i]], "inputs", 
+                                 paste0("S_hat_random_df_19000topSNPs_",
+                                        phekey[[j]], "_", popkey[[i]], ".rds")))
+    }
+    write.table(betas, file = here("analysis", "mash_greedy_algorithm", 
+                                   "mash_h2_diag_covar", 
+                                   paste0(names(popkey)[i], "_inputs"), 
+                                   paste0(names(phekey)[j], ".", 
+                                          names(popkey)[i], 
+                                          ".betas.random.txt")),
+                row.names = FALSE, col.names = FALSE)
+    write.table(stderr, file = here("analysis", "mash_greedy_algorithm", 
+                                    "mash_h2_diag_covar", 
+                                    paste0(names(popkey)[i], "_inputs"), 
+                                    paste0(names(phekey)[j], ".", 
+                                           names(popkey)[i], 
+                                           ".stderrs.random.txt")),
+                row.names = FALSE, col.names = FALSE)
+  }
+}
+
+```
+
+```{r}
+for (i in 1:35) {
+  if (!dir.exists(here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar", "midwest_inputs", "likelihoods", paste0("step", i)))) {
+    dir.create(here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar", "midwest_inputs", "likelihoods", paste0("step", i)))
+  }
+}
+```
 

### 2023-09-15 R/Analysis_v0.6_QTL_GxE_and_mash_comparison.Rmd
writeup: Remake Figures 1-4 for revision, using new mash objects

@@
 library(cowplot)
+library(here)
 ```
@@
 ```{r}
-m_gulf_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_GR50_Gulf_Uhyp.rds")
-m_gulf_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_FL50_Gulf_Uhyp.rds")
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+mashdir <- here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar")
 
-m_midw_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_24000_SNPs_GR50_Midwest_Uhyp.rds")
-m_midw_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_33000_SNPs_FL50_Midwest_Uhyp.rds")
+numSNPs_v <- c(19000, 19000, 19000, 19000, 24000, 33000)
+suffix_v <- c("FL50_Gulf_and_Midwest", "FL50_Gulf", "GR50_Gulf_and_Midwest",
+              "GR50_Gulf", "GR50_Midwest", "FL50_Midwest")
+```
 
-m_gam_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_GR50_Gulf_and_Midwest_Uhyp.rds")
-m_gam_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_FL50_Gulf_and_Midwest_Uhyp.rds")
+```{r}
+m_gulf_gr_hyp <- readRDS(file.path(mashdir,  
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
+                                          numSNPs_v[4], "_SNPs_",
+                                          suffix_v[4], ".rds")))
+m_gulf_fl_hyp <- readRDS(file.path(mashdir,  
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
+                                          numSNPs_v[2], "_SNPs_",
+                                          suffix_v[2], ".rds")))
+
+m_midw_gr_hyp <- readRDS(file.path(mashdir,  
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
+                                          numSNPs_v[5], "_SNPs_",
+                                          suffix_v[5], ".rds")))
+m_midw_fl_hyp <- readRDS(file.path(mashdir,  
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
+                                          numSNPs_v[6], "_SNPs_",
+                                          suffix_v[6], ".rds")))
+
+m_gam_gr_hyp <- readRDS(file.path(mashdir,  
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
+                                          numSNPs_v[3], "_SNPs_",
+                                          suffix_v[3], ".rds")))
+m_gam_fl_hyp <- readRDS(file.path(mashdir,  
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
+                                          numSNPs_v[1], "_SNPs_",
+                                          suffix_v[1], ".rds")))
 
@@
 ```{r}
-i=16  # from 1 to 16
+i=1  # from 1 to 16
 ## In block & > 2
@@
   geom_vline(xintercept = 16, linetype = 2, color = "red")
-save_plot(filename = "../manuscript/plots/Figure_4_QTL_overlaps_random_QTL_enrichments_with_mash_1000runs.png", plot = last_plot(), base_height = 1.8, base_asp = 1.8)
+save_plot(filename = "../manuscript/Coauthors/PNAS/Revision/Figure_4_QTL_overlaps_random_QTL_enrichments_with_mash_1000runs.png", plot = last_plot(), base_height = 1.8, base_asp = 1.8)
 
@@
   arrange(desc(sig_r))
-3/1000
+11/1000
 ```

### 2023-09-15 R/Analysis_v0.6_QTL_interval_plotting.Rmd
writeup: Remake Figures 1-4 for revision, using new mash objects

@@
 library(tidyverse)
+library(AnnotationDbi)
 ```
@@
     #scale_y_continuous(expand = c(0.2, 0.2)) 
-save_plot(filename = "../manuscript/plots/Figure_4_QTL_effects_by_trait.png", base_height = 4.5, base_width = 13.5, plot = last_plot())
+save_plot(filename = "../manuscript/Coauthors/PNAS/Revision/Figure_4_QTL_effects_by_trait.png", base_height = 4.5, base_width = 13.5, plot = last_plot())
 
@@
 ```{r}
+library(AnnotationDbi)
+library(GenomicFeatures)
 qtl_anno_df <- qtlplot %>%
@@
   arrange(Chr, Pos) %>%
-  select(Chr:Pos, log10BayesFactor, subpop, value, phenotype, everything()) %>%
+  dplyr::select(Chr:Pos, log10BayesFactor, subpop, value, phenotype, everything()) %>%
   filter(log10BayesFactor > 2)
@@
 qtlplot <- qtlplot %>%
-  arrange(desc(lod)) %>%
+  arrange(dplyr::desc(lod)) %>%
   mutate(pos = as.integer(pos)) %>%
   group_by(CHR, pos) %>%
-  slice(1) %>%
+  dplyr::slice(1) %>%
   arrange(CHR, pos)
@@
                                                    qtlplot$POS_hi[f4df$i[k]])) %>%
-    arrange(desc(log10BayesFactor))
+    arrange(dplyr::desc(log10BayesFactor))
 
@@
 
-save_plot(filename = "../manuscript/plots/Figure_4_mash_SNP_effects.png", plot = last_plot(), base_height = 2.4, base_width = 5.5)
+save_plot(filename = "../manuscript/Coauthors/PNAS/Revision/Figure_4_mash_SNP_effects.png", plot = last_plot(), base_height = 2.4, base_width = 5.5)
 ```

### 2023-09-15 R/Analysis_v0.7_QTL_interval_lod_plot.Rmd
writeup: Remake Figures 1-4 for revision, using new mash objects

@@
 
-save_plot(filename = "../manuscript/plots/Figure_4_QTL_by_lod_score.svg", base_height = 3, base_width = 13.5, plot = last_plot())
+save_plot(filename = "../manuscript/Coauthors/PNAS/Revision/Figure_4_QTL_by_lod_score.svg", base_height = 3, base_width = 13.5, plot = last_plot())
 ```
@@
   coord_cartesian(clip = "off")
-save_plot(filename = "../manuscript/plots/Figure_4_QTL_by_lod_score_w_rugs.svg", base_height = 3, base_width = 13.5, plot = last_plot())
+save_plot(filename = "../manuscript/Coauthors/PNAS/Revision/Figure_4_QTL_by_lod_score_w_rugs.svg", base_height = 3, base_width = 13.5, plot = last_plot())
 ```

### 2023-09-15 R/Analysis_v1.2_mash_using_greedy_covar.Rmd
writeup: Remake Figures 1-4 for revision, using new mash objects

@@
 
+Run the pod.greedy.sh scripts from their dir, or run the python commands in those scripts.
+
+Look at the results in the likelihood_paths files for the following data:
+
+Using a set of covariance matrices with h2 on the diagonal, made positive semi-definite, and for multiple weather variable averages, 1-7d, 14,21,and28d averages before the greenup or flowering event.
+
+Using the effects and standard errors that I used for my previous mash analysis. However I originally selected them, plus the reduced number of sites for the Midwest plants.
+
+Only keep covariance matrices if they significantly improve the mash model likelihoods when added.
+This way, we know they improve the mash model and that they are distinct enough from others matrices that we added.
+
+Using the random set of SNPs. Then, look at the mash results generated using these covariance matrices using a strong set of SNPs.
+
+Given that the greedy algorithm, run on Sam's random inputs, however he generated or found those, gives many more covar matrices especially from Uhyp, I wonder if some repetition of this analysis would be valuable, before blowing these results all out of proportion in a paper.
+Either different random sets of effects, or effects from different years of data.
+Think on this.
+I definitely want to use my known inputs, not Sam's, as mine have a traceable history (or anyways more traceable).
+
 # -----------------------------
@@
 source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
-mashdir <- here("analysis", "mash_greedy_algorithm")
-
-#phe_gr_df <- read_rds(here("data",
-#                           "Weather_related_phe_for_asreml_h2_GR50.rds"))
-# all greenup related phenotypes, including greenup as functions of weather variables.
-#phe_fl_df <- read_rds(here("data",
-#                           "Weather_related_phe_for_asreml_h2_FL50.rds"))
-# all flowering related phenotypes, including functions of weather variables.
-
-# Files with the six suffixes I have/want mash results for:
+mashdir <- here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar")
+
 numSNPs_v <- c(19000, 19000, 19000, 19000, 24000, 33000)
@@
 
-Upath <- here("analysis", "mash_greedy_algorithm", "mash_switchgrass_models", 
-                "mash", "all_matrices", "matrices")
-stopcondpath <- here("analysis", "mash_greedy_algorithm",
-                     "mash_switchgrass_models", "mash", 
-                     "all_matrices", "likelihood_paths")
-
+Upaths <- c(here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar",
+                 "all_inputs"),
+            here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar",
+                 "gulf_inputs"),
+            here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar",
+                 "all_inputs"),
+            here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar",
+                 "gulf_inputs"),
+            here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar",
+                 "midwest_greenup"),
+            here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar",
+                 "midwest_flowering"))
+
+#   i = 4
 for (i in 1:6) {  # change this for different scripts
@@
   suffix <- suffix_v[i]
+  Upath <- here(Upaths[i], "matrices")
+  stopcondpath <- here(Upaths[i], "likelihood_paths")
   # use this one for the likelihood paths
@@
 
-  saveRDS(m2, file = file.path(mashdir, suffix_vli[i], 
+  saveRDS(m2, file = file.path(mashdir,  
                                    paste0("Strong_mash_posterior_",
-                                          "est_w_random_greedy_Ulist", 
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
                                           numSNPs, "_SNPs_",
@@
   gxe <- switchgrassGWAS::get_GxE(m2)
-  saveRDS(gxe, file = file.path(mashdir, suffix_vli[i], 
+  saveRDS(gxe, file = file.path(mashdir,  
                                    paste0("GxE_default_",
-                                          "est_w_random_greedy_Ulist",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist",
                                           numSNPs, "_SNPs_",
@@
 ```
+
+# ----------------
+
+Reasoning from Analysis_v0.6 script for dropping sites from the Midwest:
+
+Some univariate GWAS for the Midwest had population structure issues. Mostly for greenup - even GR50 at MO was quite bad. Drop MO for GR50.
+Midwest_GR50_SD have values of zero for Bhat and Shat. Drop this from Midwest analyses.
+
+# Remake mash figures
+
+## Load mash objects
+```{r}
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+mashdir <- here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar")
+
+numSNPs_v <- c(19000, 19000, 19000, 19000, 24000, 33000)
+suffix_v <- c("FL50_Gulf_and_Midwest", "FL50_Gulf", "GR50_Gulf_and_Midwest",
+              "GR50_Gulf", "GR50_Midwest", "FL50_Midwest")
+```
+
+```{r}
+m_gulf_gr_hyp <- readRDS(file.path(mashdir,  
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
+                                          numSNPs_v[4], "_SNPs_",
+                                          suffix_v[4], ".rds")))
+m_gulf_fl_hyp <- readRDS(file.path(mashdir,  
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
+                                          numSNPs_v[2], "_SNPs_",
+                                          suffix_v[2], ".rds")))
+
+m_midw_gr_hyp <- readRDS(file.path(mashdir,  
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
+                                          numSNPs_v[5], "_SNPs_",
+                                          suffix_v[5], ".rds")))
+m_midw_fl_hyp <- readRDS(file.path(mashdir,  
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
+                                          numSNPs_v[6], "_SNPs_",
+                                          suffix_v[6], ".rds")))
+
+m_gam_gr_hyp <- readRDS(file.path(mashdir,  
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
+                                          numSNPs_v[3], "_SNPs_",
+                                          suffix_v[3], ".rds")))
+m_gam_fl_hyp <- readRDS(file.path(mashdir,  
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
+                                          numSNPs_v[1], "_SNPs_",
+                                          suffix_v[1], ".rds")))
+
+```
+
+## Figure 2
+```{r}
+get_cov_type <- function(m, summarise = FALSE, subpop = FALSE){
+  cov_df <- mash_plot_covar(m)
+  cov_type <- cov_df$covar_df %>% 
+    mutate(Type = case_when(grepl("PCA", `Covariance Matrix`) ~ "data-driven",
+                            grepl("singleton", `Covariance Matrix`) ~
+                              "canonical",
+                            grepl("simple_het", `Covariance Matrix`) ~
+                              "canonical",
+                            grepl("no_effects", `Covariance Matrix`) ~
+                              "canonical",
+                            grepl("identity", `Covariance Matrix`) ~
+                              "canonical",
+                            grepl("equal_effects", `Covariance Matrix`) ~
+                              "canonical",
+                            grepl("U.mat$", `Covariance Matrix`) ~
+                              "hypothesized",
+                            TRUE ~ "other"
+                            )) 
+  if (summarise == TRUE) {
+    cov_type <- cov_type %>%
+      group_by(Type) %>%
+      summarise(Mass = sum(Mass))
+  }
+  return(cov_type)
+}
+
+get_cov_type(m_gulf_gr_hyp, summarise = TRUE, subpop = TRUE)
+get_cov_type(m_gulf_fl_hyp, summarise = TRUE, subpop = TRUE)
+get_cov_type(m_midw_gr_hyp, summarise = FALSE, subpop = TRUE)
+get_cov_type(m_midw_fl_hyp, summarise = TRUE, subpop = TRUE)
+get_cov_type(m_gam_gr_hyp, summarise = TRUE, subpop = TRUE)
+get_cov_type(m_gam_fl_hyp, summarise = TRUE, subpop = TRUE)
+```
+
+```{r}
+cov_gg <- mash_plot_covar(m_gulf_gr_hyp)
+cov_gf <- mash_plot_covar(m_gulf_fl_hyp)
+
+cov_mg <- mash_plot_covar(m_midw_gr_hyp)
+cov_mf <- mash_plot_covar(m_midw_fl_hyp)
+
+cov_gamg <- mash_plot_covar(m_gam_gr_hyp)
+cov_gamf <- mash_plot_covar(m_gam_fl_hyp)
+
+cov_gg$ggobject + 
+  ggtitle("Gulf greenup", subtitle = "64.5% SNP mass")
+cov_gf$ggobject + 
+  ggtitle("Gulf flowering", subtitle = "33.0% SNP mass")
+cov_mg$ggobject + 
+  ggtitle("Midwest greenup", subtitle = "0% SNP mass")
+cov_mf$ggobject + 
+  ggtitle("Midwest flowering", subtitle = "22.6% SNP mass")
+cov_gamg$ggobject + 
+  ggtitle("Gulf and Midwest greenup", subtitle = "32.7% SNP mass")
+cov_gamf$ggobject + 
+  ggtitle("Gulf and Midwest flowering", subtitle = "0% SNP mass")
+```
+### covar
+
+```{r get covar matrices picked by greedy mash}
+get_greedy_Ulist <- function(i) {
+  pop_sc <- c("all", "gulf", "all", "gulf", "midwest", "midwest")
+  phe_sc <- c("flowering", "flowering", "greenup", "greenup", "greenup",
+              "flowering")
+  Upaths <- c(here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar",
+                   "all_inputs"),
+              here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar",
+                   "gulf_inputs"),
+              here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar",
+                   "all_inputs"),
+              here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar",
+                   "gulf_inputs"),
+              here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar",
+                   "midwest_greenup"),
+              here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar",
+                   "midwest_flowering"))
+
+#   i = 4
+#for (i in 1:6) {  # change this for different scripts
+  # B_hat_strong_df_19000topSNPs_FL50_Gulf_and_Midwest.rds
+  # B_hat_strong_df_19000topSNPs_FL50_Gulf.rds
+  # B_hat_strong_df_19000topSNPs_GR50_Gulf_and_Midwest.rds
+  # B_hat_strong_df_19000topSNPs_GR50_Gulf.rds
+  # B_hat_strong_df_24000topSNPs_GR50_Midwest.rds
+  # B_hat_strong_df_33000topSNPs_FL50_Midwest.rds
+  ## mash Gulf greenup
+  Upath <- here(Upaths[i], "matrices")
+  stopcondpath <- here(Upaths[i], "likelihood_paths")
+  # use this one for the likelihood paths
+  # Read in likelihood paths, then use that to read in matrices, then use those to run mash.
+
+  U_greedy <- 
+    read_csv(file = file.path(stopcondpath, 
+                              paste0(phe_sc[i], ".", pop_sc[i], 
+                                     ".random.stop.condition.csv"))) %>%
+    filter(matrix != "0.0")
+  ## Read in Uhyp
+
+  U_list <- list()
+  for (k in 1:nrow(U_greedy)) {
+    U_single <- read_csv(file = file.path(Upath, U_greedy$matrix[k]), show_col_types = FALSE)
+    # if conditions for Midwest
+    if (i == 5) {
+    # MI, MO, NE, OK, TX1, TX2, TX3 used for GR50  
+      U_single <- U_single %>%
+        filter(`...1` %in% c("MI", "MO", "NE", "OK", "TX1", "TX2", "TX3")) %>%
+        select(`...1`, .data$MI, .data$MO, .data$NE, .data$OK, .data$TX1,
+               .data$TX2, .data$TX3)
+    } else if (i == 6) {
+    # MI, MO, SD, TX1, TX2, TX3 used for FL50  
+      U_single <- U_single %>%
+        filter(`...1` %in% c("MI", "MO", "SD", "TX1", "TX2", "TX3")) %>%
+        select(`...1`, .data$MI, .data$MO, .data$SD, .data$TX1, .data$TX2,
+               .data$TX3)
+    }
+
+    # format cov matrices correctly for mash
+    U_list <- list.append(U_list, U_single)
+  }
+  names(U_list) <- str_remove(U_greedy$matrix, ".csv") %>%
+    enframe(value = "Covariance Matrix") %>%
+  mutate(`Covariance Matrix` = str_replace(`Covariance Matrix`, "greenup.", ""),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "gulf.", ""),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "all.", ""),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "midwest.", ""),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "singleton.", "Single Effect "),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "simple_het_", "Simple Het. "),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "no_effects", "No Effects"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "tave_1d.U.mat", "Temp Ave, 1d"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "dyln_14d_prior.U.mat", "Daylength (14d prior)"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "dyln_change_sec.U.mat", "Daylength change (s)"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "flowering.", ""),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "crain_7d.U.mat", "Cumulative rainfall, 7d"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "dyln_change_sec_2d_prior.U.mat", "Daylength change (s), 2d prior")
+         ) %>% 
+    select(`Covariance Matrix`) %>%
+    as_vector()
+  return(U_list)
+}
+
+#}
+```
+
+```{r pick covar matrices to plot fig 2}
+plot_covar_Ulist_fig2 <- function(df, mf = FALSE, mg = FALSE) {
+  if (mg) {
+    df <- df %>%
+      select(`...1`, TX1, TX2, TX3, OK, MO, NE, MI) 
+  } else if (mf) {
+    df <- df %>%
+      select(`...1`, TX1, TX2, TX3, MO, MI, SD) 
+  } else {
+    df <- df %>%
+      select(`...1`, TX1, TX2, TX3, OK, MO, NE, MI, SD) 
+  }
+
+df$`...1` <- factor(df$`...1`, levels = rev(c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD")))
+
+  U1 <- df %>%
+          pivot_longer(cols = -.data$`...1`, names_to = "colU",
+                       values_to = "covar") %>%
+          filter(!is.na(.data$covar)) %>%
+    rename(rowU = `...1`)
+  U1$colU <- factor(U1$colU, levels = colnames(df)[2:ncol(df)])
+  U1$rowU <- factor(U1$rowU, levels = colnames(df)[2:ncol(df)])
+
+  U1$Site_pair <- NA
+for (i in 1:nrow(U1)) {
+  v <- sort(c(U1$rowU[i], U1$colU[i]))
+  U1$Site_pair[i] <- paste0(v[1], "_", v[2])
+}
+
+  plot_covar <- U1 %>%
+    arrange(desc(rowU), desc(colU)) %>%
+    distinct(Site_pair, .keep_all = TRUE) %>%
+    ggplot(aes(x = .data$rowU, y = .data$colU)) +
+    switchgrassGWAS::theme_oeco +
+    geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
+    scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
+                                  "white", "#27AD81FF", "#5DC863FF",
+                                  "#FDE725FF"),
+                         limits = c(-1, 1)) +
+    theme(legend.position = "none",
+          axis.text = element_blank(),
+          axis.ticks = element_blank(),
+          axis.line.x = element_blank(),
+          axis.line.y = element_blank(),
+          axis.title.y.right = element_text()) +
+    labs(x = "", y = "")
+  return(plot_covar)
+}
+```
+
+```{r}
+Ulist_gamf <- get_greedy_Ulist(1)
+Ulist_gf <- get_greedy_Ulist(2)
+Ulist_gamg <- get_greedy_Ulist(3)
+Ulist_gg <- get_greedy_Ulist(4)
+Ulist_mg <- get_greedy_Ulist(5)
+Ulist_mf <- get_greedy_Ulist(6)
+Ulist_gf$`Cumulative rainfall, 7d`
+Ulist_mg$dyln_change_sec_7d_prior.U.mat
+```
+
+```{r build panels of covar matrices for 2a}
+pa01 <- plot_covar_Ulist_fig2(Ulist_gg$`Daylength (14d prior)`) + 
+  labs(subtitle = "Daylength (14d prior)")
+pa03 <- plot_covar_Ulist_fig2(Ulist_gg$cgdd_12c_6d.U.mat) + 
+  labs(subtitle = "Cumulative GDD, 6d")
+
+pa06 <- plot_covar_Ulist_fig2(Ulist_gf$`Cumulative rainfall, 7d`) + 
+  labs(subtitle = "Cumulative rainfall, 7d")
+
+pa04 <- plot_covar_Ulist_fig2(Ulist_mg$dyln_change_sec_7d_prior.U.mat, mg = TRUE) + 
+  labs(subtitle = "Daylength change (s), 7d prior")
+pa07 <- plot_covar_Ulist_fig2(Ulist_mf$`Daylength change (s), 2d prior`, mf = TRUE) + 
+  labs(subtitle = "Daylength change (s), 2d prior")
+
+pa02 <- plot_covar_Ulist_fig2(Ulist_gamg$`Daylength change (s)`) + 
+  labs(subtitle = "Daylength change (s)")
+pa05 <- plot_covar_Ulist_fig2(Ulist_gamg$`Temp Ave, 1d`) + 
+  labs(subtitle = "Temp Ave, 1d")
+pa08 <- plot_covar_Ulist_fig2(Ulist_gamf$crain_1d.U.mat) + 
+  labs(subtitle = "Cumulative rainfall, 1d")
+
+pa13 <- plot_covar_Ulist_fig2(Ulist_gf[[1]]) + 
+  labs(subtitle = "identity (Mass <1%)") 
+pa09 <- plot_covar_Ulist_fig2(Ulist_gf[[5]]) + 
+  labs(subtitle = names(Ulist_gf)[5]) 
+pa10 <- plot_covar_Ulist_fig2(Ulist_gf[[2]]) + 
+  labs(subtitle = names(Ulist_gf)[2])
+pa12 <- plot_covar_Ulist_fig2(Ulist_gg[[2]]) + 
+  labs(subtitle = names(Ulist_gg)[2])
+pa11 <- plot_covar_Ulist_fig2(Ulist_gf[[8]]) + 
+  labs(subtitle = names(Ulist_gf)[8])
+```
+
+```{r build panels of barplots for fig 2}
+cov_gr1 <- mash_plot_Ulist(m_gulf_gr_hyp, range = 2, limits = TRUE)
+legend <- get_legend(cov_gr1) 
+fig_eqeff <- cov_gr1$greenup.simple_het_2_df %>%
+  get_covar_fig2() +
+  labs(subtitle = "simple heterozygousity")
+
+df_2a <- get_cov_type(m_gulf_gr_hyp, summarise = TRUE) %>%
+  mutate(subpop = "Gulf",
+         phenotype = "greenup")
+df_2b <- get_cov_type(m_midw_gr_hyp, summarise = TRUE) %>%
+  mutate(subpop = "Midwest",
+         phenotype = "greenup")
+df_2c <- get_cov_type(m_gam_gr_hyp, summarise = TRUE) %>%
+  mutate(subpop = "Both",
+         phenotype = "greenup")
+df_2d <- get_cov_type(m_gulf_fl_hyp, summarise = TRUE) %>%
+  mutate(subpop = "Gulf",
+         phenotype = "flowering")
+df_2e <- get_cov_type(m_midw_fl_hyp, summarise = TRUE) %>%
+  mutate(subpop = "Midwest",
+         phenotype = "flowering")
+df_2f <- get_cov_type(m_gam_fl_hyp, summarise = TRUE) %>%
+  mutate(subpop = "Both",
+         phenotype = "flowering")
+df_2gr_b <- df_2a %>% full_join(df_2b) %>% full_join(df_2c)
+df_2fl_b <- df_2d %>% full_join(df_2e) %>% full_join(df_2f)
+
+df_2a <- get_cov_type(m_gulf_gr_hyp) %>%
+  mutate(subpop = "Gulf",
+         phenotype = "greenup")
+df_2b <- get_cov_type(m_midw_gr_hyp) %>%
+  mutate(subpop = "Midwest",
+         phenotype = "greenup")
+df_2c <- get_cov_type(m_gam_gr_hyp) %>%
+  mutate(subpop = "Both",
+         phenotype = "greenup")
+df_2d <- get_cov_type(m_gulf_fl_hyp) %>%
+  mutate(subpop = "Gulf",
+         phenotype = "flowering")
+df_2e <- get_cov_type(m_midw_fl_hyp) %>%
+  mutate(subpop = "Midwest",
+         phenotype = "flowering")
+df_2f <- get_cov_type(m_gam_fl_hyp) %>%
+  mutate(subpop = "Both",
+         phenotype = "flowering")
+df_2gr_a <- df_2a %>% full_join(df_2b) %>% full_join(df_2c) %>%
+  mutate(`Covariance Matrix` = str_replace(`Covariance Matrix`, "greenup.", ""),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "gulf.", ""),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "all.", ""),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "midwest.", ""),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "singleton.", "Single Effect "),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "simple_het_", "Simple Het. "),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "no_effects", "No Effects"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "tave_1d.U.mat", " Temp Ave, 1d"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "dyln_14d_prior.U.mat", " Daylength (14d prior)"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "dyln_change_sec.U.mat", " Daylength change (s)")
+         )
+df_2fl_a <- df_2d %>% full_join(df_2e) %>% full_join(df_2f) %>%
+  mutate(`Covariance Matrix` = str_replace(`Covariance Matrix`, "flowering.", ""),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "gulf.", ""),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "all.", ""),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "midwest.", ""),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "singleton.", "Single Effect "),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "simple_het_", "Simple Het. "),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "no_effects", "No Effects"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "crain_7d.U.mat", "  Cumulative rainfall, 7d"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "dyln_change_sec_2d_prior.U.mat", " Daylength change (s), 2d prior")
+         )
+
+df_2gr_b$subpop <- factor(df_2gr_b$subpop, levels = c("Gulf", "Midwest", "Both"))
+df_2fl_b$subpop <- factor(df_2fl_b$subpop, levels = c("Gulf", "Midwest", "Both"))
+df_2gr_a$subpop <- factor(df_2gr_a$subpop, levels = c("Gulf", "Midwest", "Both"))
+df_2fl_a$subpop <- factor(df_2fl_a$subpop, levels = c("Gulf", "Midwest", "Both"))
+df_2fl_a$Type <- factor(df_2fl_a$Type, levels = c("canonical", "hypothesized"))
+df_2fl_a$Type <- factor(df_2fl_a$Type, levels = c("canonical", "hypothesized"))
+
+f2b <- df_2gr_b %>%
+  filter(Mass > 1e-05) %>%
+  ggplot(aes(x = subpop, y = Mass)) + 
+  geom_bar(aes(fill = Type), stat = "identity") +
+  scale_fill_viridis_d(end = 0.9) +
+  theme(axis.title.x=element_blank(),
+        legend.position = "none",
+        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45))
+f2d <- df_2fl_b %>%
+  filter(Mass > 1e-05) %>%
+  ggplot(aes(x = subpop, y = Mass)) + 
+  geom_bar(aes(fill = Type), stat = "identity") +
+  scale_fill_viridis_d(end = 0.9) +
+  theme(axis.title.x=element_blank(),
+        legend.position = "none",
+        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45))
+f2a <- df_2gr_a %>%
+  filter(Mass > 1e-06) %>%
+  ggplot(aes(x = `Covariance Matrix`, y = Mass)) + 
+  geom_bar(aes(fill = Type), stat = "identity") +
+  facet_wrap(~subpop) +
+  coord_flip() + scale_fill_viridis_d(end = 0.9) +
+  theme(legend.position = c(-0.4, 0.3),
+        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45), 
+        panel.spacing.x = unit(-0, 'cm')) #+ labs(x = "")
+f2c <- df_2fl_a %>%
+  filter(Mass > 1e-06) %>%
+  ggplot(aes(x = `Covariance Matrix`, y = Mass)) + 
+  geom_bar(aes(fill = Type), stat = "identity") +
+  facet_wrap(~subpop) +
+  coord_flip() + scale_fill_viridis_d(end = 0.9) +
+  theme(legend.position = "none",
+        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45), 
+        panel.spacing.x = unit(-0, 'cm')) #+ labs(x = "")
+
+```
+
+```{r build figure 2}
+ggdraw() +
+  draw_plot(pa01, x = -0.02, y = 0.4, width = 0.1, height = 0.18) +
+  draw_plot(pa02, x = -0.02, y = 0.6, width = 0.1, height = 0.18) +
+  #draw_plot(pa03, x = -0.02, y = 0.6, width = 0.1, height = 0.18) +
+  #draw_plot(pa04, x = 0.1, y = 0.6, width = 0.1, height = 0.18) +
+  draw_plot(pa05, x = -0.02, y = 0.8, width = 0.1, height = 0.18) +
+  draw_plot(pa06, x = -0.02, y = 0, width = 0.1, height = 0.18) +
+  draw_plot(pa07, x = -0.02, y = 0.2, width = 0.1, height = 0.18) +
+  #draw_plot(pa08, x = 0.2, y = 0, width = 0.1, height = 0.18) +
+  draw_plot(pa09, x = 0.13, y = 0.6, width = 0.1, height = 0.18) +
+  draw_plot(pa10, x = 0.13, y = 0.2, width = 0.1, height = 0.18) +
+  draw_plot(pa11, x = 0.13, y = 0.4, width = 0.1, height = 0.18) +
+  draw_plot(pa12, x = 0.13, y = 0, width = 0.1, height = 0.18) +
+  draw_plot(pa13, x = 0.13, y = 0.8, width = 0.1, height = 0.18) +
+  draw_plot(f2a, x = 0.3, y = 0.47, width = 0.55, height = 0.53) +
+  #draw_plot(legend, x = 0.15, y = 0.34, width = 0.1, height = 0.2) +
+  draw_plot(f2b, x = 0.85, y = 0.47, width = 0.15, height = 0.53) +
+  draw_plot(f2c, x = 0.3, y = 0, width = 0.55, height = 0.47) +
+  draw_plot(f2d, x = 0.85, y = 0, width = 0.15, height = 0.47)
+
+save_plot(filename = "../manuscript/Coauthors/PNAS/Revision/Figure_2_Covariance_matrices_and_posterior_weights.svg", plot = last_plot(), base_height = 6, base_width = 8.5, )
+```
+
+## Table 1
+
+```{r}
+gxe_gg <- get_GxE(m_gulf_gr_hyp)
+gxe_mg <- get_GxE(m_midw_gr_hyp)
+gxe_gamg <- get_GxE(m_gam_gr_hyp)
+gxe_gf <- get_GxE(m_gulf_fl_hyp)
+gxe_mf <- get_GxE(m_midw_fl_hyp)
+gxe_gamf <- get_GxE(m_gam_fl_hyp)
+```
+
+```{r}
+sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
+
+site_regions <- sites %>% 
+  mutate(Region = case_when(manu_site %in% c("SD", "MO", "IL", "MI", "NE") ~ "North",
+                            manu_site %in% c("TX1", "TX2", "TX3", "TX4") ~ "Texas",
+                                             TRUE ~ "OK")) %>%
+  select(manu_site, Region) %>%
+  rename("Site" = "manu_site") %>%
+  add_row(Site = "M", Region = "North")
+names(gxe_gg)
+```
+
+Change these to look at S_AP, S_DS, S_2_no; S_1_row and S_1_col separately
+```{r}
+
+one_gxe_summary <- function(df, n_cond = 8, n_sep = 2) {
+  nsep <- rep(NA, times = n_sep)
+  df <- df %>%
+    as_tibble(rownames = "Condition_1") %>%
+    pivot_longer(cols = 2:(n_cond+1), names_to = "Condition_2", values_to = "Sig") %>%
+    separate(Condition_1, into = c(nsep, "Phenotype", "Garden_1"), sep = "_") %>%
+    separate(Condition_2, into = c(nsep, NA, "Garden_2"), sep = "_") %>%
+    left_join(site_regions, by = c("Garden_1" = "Site")) %>%
+    rename("Region_1" = "Region") %>%
+    left_join(site_regions, by = c("Garden_2" = "Site")) %>%
+    rename("Region_2" = "Region") %>%
+    filter(Garden_1 != Garden_2) %>%
+    filter(Region_1 != "OK" & Region_2 !="OK") %>%
+    group_by(Region_1, Region_2) %>%
+    summarise(Total_sig = sum(Sig), .groups = "keep")
+  df$Region_comparison <- NA
+  for (i in 1:nrow(df)) {
+    v <- sort(c(df$Region_1[i], df$Region_2[i]))
+    df$Region_comparison[i] <- paste0(v[1], "_", v[2])
+  }
+  df <- df %>%
+    ungroup() %>%
+    select(-Region_1, -Region_2) %>%
+    group_by(Region_comparison) %>%
+    summarise(n = floor(sum(Total_sig)/2))
+  return(df)
+}
+# test output
+one_gxe_summary(df = gxe_mf$S_AP, n_cond = 6)
+
+get_gxe_summary <- function(gxe_list, n_cond = 8, n_sep = 2) {
+  AP <- one_gxe_summary(gxe_list$S_AP, n_cond = n_cond, n_sep = n_sep) %>%
+    rename(Sign = n)
+  DS <- one_gxe_summary(gxe_list$S_DS, n_cond = n_cond, n_sep = n_sep) %>%
+    rename(Magnitude = n)
+  ND <- one_gxe_summary(gxe_list$S_2_no, n_cond = n_cond, n_sep = n_sep) %>%
+    rename(None = n)
+  gxe_n <- AP %>%
+    full_join(DS, by = "Region_comparison") %>%
+    full_join(ND, by = "Region_comparison") %>%
+    pivot_longer(cols = -Region_comparison, names_to = "Type_of_GxE", 
+                 values_to = "n") %>%
+    pivot_wider(names_from = "Region_comparison", values_from = "n")
+  gxe_frac <- gxe_n %>%
+    mutate(North_North = round(North_North/sum(North_North), digits = 4),
+           North_Texas = round(North_Texas/sum(North_Texas), digits = 4),
+           Texas_Texas = round(Texas_Texas/sum(Texas_Texas), digits = 4))
+  return(list(gxe_counts = gxe_n, gxe_fractions = gxe_frac))
+}
+
+get_gxe_summary(gxe_gg) 
+```
+
+```{r}
+get_gxe_summary(gxe_gg) 
+get_gxe_summary(gxe_gf) 
+get_gxe_summary(gxe_mg, n_cond = 7) 
+get_gxe_summary(gxe_mf, n_cond = 6) 
+get_gxe_summary(gxe_gamg, n_sep = 4) 
+get_gxe_summary(gxe_gamf, n_sep = 4) 
+```
+
+## Figure 3
+
+```{r}
+pvdiv_plot_cor <- function(corrmatrix, upper = NULL, legend = FALSE){
+  U1 <- corrmatrix
+
+  U1 <- as_tibble(U1, rownames = "rowU", .name_repair = "unique") %>%
+          pivot_longer(cols = -.data$rowU, names_to = "colU",
+                       values_to = "covar") %>%
+          filter(!is.na(.data$covar)) %>%
+    mutate(rowU = str_remove(rowU, 
+                             "Stand_BhatGulf_and_Midwest_")) %>%
+    mutate(rowU = str_remove(rowU, "Stand_BhatGulf_")) %>%
+    mutate(rowU = str_remove(rowU, "Stand_BhatMidwest_")) %>%
+    mutate(rowU = str_remove(rowU, "GR50_")) %>%
+    mutate(rowU = str_remove(rowU, "FL50_")) %>%
+    mutate(colU = str_remove(colU, 
+                             "Stand_BhatGulf_and_Midwest_")) %>%
+    mutate(colU = str_remove(colU, "Stand_BhatGulf_")) %>%
+    mutate(colU = str_remove(colU, "Stand_BhatMidwest_")) %>%
+    mutate(colU = str_remove(colU, "GR50_")) %>%
+    mutate(colU = str_remove(colU, "FL50_")) %>%
+    mutate(rowU = case_when(grepl("^M$", rowU) ~ "MI",
+                            TRUE ~ rowU),
+           colU = case_when(grepl("^M$", colU) ~ "MI",
+                            TRUE ~ colU),)
+  U1$colU <- factor(U1$colU, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
+  U1$rowU <- factor(U1$rowU, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
+  U1 <- arrange(U1, rowU, colU) %>%
+    pivot_wider(names_from = colU, values_from = covar) 
+  for(i in 1:nrow(U1)){
+    for(j in 2:ncol(U1)){
+      if(i <= j-1){
+        U1[i, j] <- NA
+      }
+    }
+  }
+  U1 <- U1 %>%
+    pivot_longer(cols = -.data$rowU, names_to = "colU",
+                       values_to = "covar")
+  U1$colU <- factor(U1$colU, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD")) 
+  if(is.null(upper)){
+  upper <- max(U1$covar, na.rm = TRUE)+5
+  }
+  U1_covar <- U1 %>%
+    filter(!is.na(.data$covar)) %>%
+    ggplot(aes(x = .data$rowU, y = .data$colU)) +
+    switchgrassGWAS::theme_oeco +
+    geom_tile(aes(fill = .data$covar)) +
+    scale_fill_viridis_c(option = "B", limits = c(0, upper)) +
+    #geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
+    theme(legend.position = "none",
+          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
+    xlab("") + ylab("") 
+  if(legend == TRUE){
+    U1_covar <- U1_covar +
+      theme(legend.position = c(0.2, 0.9)) +
+      geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey")
+  }
+  return(U1_covar)
+}
+
+```
+
+```{r}
+mash_plot_Ulist(m_midw_fl_hyp)
+mash_plot_manhattan_by_condition(m_midw_fl_hyp)
+```
+
+```{r}
+p3legend <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_AP, legend = TRUE, upper = 4333) + 
+  labs(title = "Gulf", subtitle = "Sign difference")
+legend <- get_legend(p3legend)
+
+p3a1 <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_AP, legend = FALSE, upper = 4333) + 
+  labs(subtitle = " ")
+p3a2 <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_DS, legend = FALSE, upper = 4333) + 
+  labs(subtitle = " ")
+p3a3 <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_2_no, legend = FALSE, upper = 4333) + 
+  labs(subtitle = " ")
+p3b1 <- pvdiv_plot_cor(corrmatrix = gxe_mg$S_AP, legend = FALSE, upper = 4333) + 
+  labs(subtitle = "Sign difference")
+p3b2 <- pvdiv_plot_cor(corrmatrix = gxe_mg$S_DS, legend = FALSE, upper = 4333) + 
+  labs(subtitle = "Magnitude difference")
+p3b3 <- pvdiv_plot_cor(corrmatrix = gxe_mg$S_2_no, legend = FALSE, upper = 4333) + 
+  labs(subtitle = "Not distinguishable")
+p3c1 <- pvdiv_plot_cor(corrmatrix = gxe_gamg$S_AP, legend = FALSE, upper = 4333) + 
+  labs(subtitle = " ")
+p3c2 <- pvdiv_plot_cor(corrmatrix = gxe_gamg$S_DS, legend = FALSE, upper = 4333) + 
+  labs(subtitle = " ")
+p3c3 <- pvdiv_plot_cor(corrmatrix = gxe_gamg$S_2_no, legend = FALSE, upper = 4333) + 
+  labs(subtitle = " ")
+
+p3d1 <- pvdiv_plot_cor(corrmatrix = gxe_gf$S_AP, legend = FALSE, upper = 4333) + 
+  labs(subtitle = " ")
+p3d2 <- pvdiv_plot_cor(corrmatrix = gxe_gf$S_DS, legend = FALSE, upper = 4333) + 
+  labs(subtitle = " ")
+p3d3 <- pvdiv_plot_cor(corrmatrix = gxe_gf$S_2_no, legend = FALSE, upper = 4333) + 
+  labs(subtitle = " ")
+p3e1 <- pvdiv_plot_cor(corrmatrix = gxe_mf$S_AP, legend = FALSE, upper = 4333) + 
+  labs(subtitle = "Sign difference")
+p3e2 <- pvdiv_plot_cor(corrmatrix = gxe_mf$S_DS, legend = FALSE, upper = 4333) + 
+  labs(subtitle = "Magnitude difference")
+p3e3 <- pvdiv_plot_cor(corrmatrix = gxe_mf$S_2_no, legend = FALSE, upper = 4333) + 
+  labs(subtitle = "Not distinguishable")
+p3f1 <- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_AP, legend = FALSE, upper = 4333) + 
+  labs(subtitle = " ")
+p3f2 <- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_DS, legend = FALSE, upper = 4333) + 
+  labs(subtitle = " ")
+p3f3 <- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_2_no, legend = FALSE, upper = 4333) + 
+  labs(subtitle = " ")
+```
+
+```{r build figure 3}
+ggdraw() +
+  draw_plot(p3a1, x = 0, y = .82, width = 1/3, height = .18) +
+  draw_plot(p3a2, x = 0, y = 4/6, width = 1/3, height = .18) +
+  draw_plot(p3a3, x = 0, y = .51, width = 1/3, height = .18) +
+  draw_plot(p3b1, x = 0.33, y = .82, width = 1/3, height = .18) +
+  draw_plot(p3b2, x = 0.33, y = 4/6, width = 1/3, height = .18) +
+  draw_plot(p3b3, x = 0.33, y = .51, width = 1/3, height = .18) +
+  draw_plot(p3c1, x = 0.66, y = .82, width = 1/3, height = .18) +
+  draw_plot(p3c2, x = 0.66, y = 4/6, width = 1/3, height = .18) +
+  draw_plot(p3c3, x = 0.66, y = .51, width = 1/3, height = .18) +
+  draw_plot(p3d1, x = 0, y = 2/6, width = 1/3, height = .18) +
+  draw_plot(p3d2, x = 0, y = 1/6, width = 1/3, height = .18) +
+  draw_plot(p3d3, x = 0, y = 0/6, width = 1/3, height = .18) +
+  draw_plot(p3e1, x = 0.33, y = 2/6, width = 1/3, height = .18) +
+  draw_plot(p3e2, x = 0.33, y = 1/6, width = 1/3, height = .18) +
+  draw_plot(p3e3, x = 0.33, y = 0/6, width = 1/3, height = .18) +
+  draw_plot(p3f1, x = 0.66, y = 2/6, width = 1/3, height = .18) +
+  draw_plot(p3f2, x = 0.66, y = 1/6, width = 1/3, height = .18) +
+  draw_plot(p3f3, x = 0.66, y = 0/6, width = 1/3, height = .18) +
+  draw_plot(legend, x = 0.37, y = .38, width = 1/3, height = .18)
+
+save_plot(filename = "../manuscript/Coauthors/PNAS/Revision/Figure_3_Effect_Types_by_Site.svg", plot = last_plot(), base_height = 10, base_width = 5.5)
+```
+
+## Figure 4

### 2023-12-02 manuscript/revision2/revision2.qmd
Add revision Quarto document in PNAS format

---
title: Quarto Template for PNAS Submissions
format:
  pnas-pdf:
    keep-tex: true  
author:
  - name: Author One
    affiliations:
      - name: Harvard University
        id: a
        department: Department of Government
        city: Cambridge
        state: MA
        postal-code: 02138
      - name: Harvard University
        id: c
        department: Department of Statistics
        city: Cambridge
        state: MA
        postal-code: 02138
    attributes:
      equal-contributor: true
  - name: Author Two
    affiliations:
      - name: Yale University
        id: b
        department: Department of Political Science
        city: New Haven
        state: CT
        postal-code: 06511
    email: alicem@uchicago.edu
    attributes:
      equal-contributor: true
      corresponding: true
  - name: Author Three
    affiliations:
      - ref: a
abstract: |
  The timing of vegetative growth and flowering in plants ("phenological timings") multiple depend on environmental cues; thus, phenological traits can have important genotype-by-environment (GxE) interactions. Here, we map genotype-by-weather effects on phenological timings in multiple populations of two highly divergent switchgrass (Panicum virgatum) populations spanning the central United States. We distinguish GxWeather patterns covarying with interpretable weather-based cues and agnostic, empirically-inferred GxWeather patterns. Most GxWeather effects belong to the latter category and do not covary with weather-based cues. However, >26% of Gulf population effects on flowering time covary with photoperiod cues, while >__% of Midwest upland population SNPs affecting flowering had effects that covaried with __ cues. An independent pseudo-F2 cross of Gulf and Midwest individuals mapped 23 additive QTLs for flowering at the same common gardens, all with significant mash associations and ten with enrichment of highly significant mash associations. We demonstrate that we can identify QTL with GxE and assign them to specific weather-based cues and/or data-driven patterns. Breeding for particular alleles at these loci could change flowering responsiveness to photoperiod cues in switchgrass. More broadly, this approach could be used to identify genetic marker-environment interactions in any species with related populations phenotyped in multiple environments.
runningauthor: "One, Two, and Three"
significance: |
  The timing of plant seasonal development (phenology) has major impacts on fitness because of the steep price of plant-environment mismatches. Here we map the genetic basis of two phenological events, the start of above-ground growth and flowering, in two genetically and phenologically distinct populations of switchgrass. We do this at eight field locations spanning the latitudinal range of both switchgrass populations. Our approach allows us to identify regions of the genome that covary with specific weather-related environmental features at every location. For flowering, these features differed by population: the Midwest population had genetic variation that primarily covaried with cumulative growing degree days, a temperature-related measure, while the Gulf population had genetic variation that primarily covaried with photoperiod, a day-length-related measure.
author_contributions:  "T.E.J. designed research. D.B.L. contributed plant material and resources. J.B., D.B.L., and T.E.J. designed and executed field experiments. A.R.B., P.A.F., F.B.F., D.B.L., R.B.M., F.M.R., Y.W., and T.E.J. hosted field experiments. A.H.M. and L.Z. conducted statistical and computational analyses. The manuscript was written by A.H.M., L.Z., and T.E.J. with contributions from all authors."
conflict_of_interest: "Please declare any competing interests here."
equal-contributor: "A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable)."
keywords:  [allele-by-environment effect variation, antagonistic pleiotropy, photoperiod, cumulative growing degree days, genetic variation]
acknowledgements: | 
  We thank the Brackenridge Field laboratory, the Ladybird Johnson Wildflower Center, and the Juenger laboratory for support with plant care and propagation. This material is based upon work supported in part by the Great Lakes Bioenergy Research Center, U.S. Department of Energy, Office of Science, Office of Biological and Environmental Research under Award Numbers DE-SC0018409 and DE-FC02-07ER64494, the US Department of Energy Awards DESC0014156 to T.E.J., DE-SC0017883 to D.B.L, National Science Foundation PGRP Awards IOS0922457 and IOS1444533 to T.E.J, and the Long-term Ecological Research Program (DEB 1832042) at the Kellogg Biological Station.
bibliography: bibliography.bib  
fontsize: "9pt"
classoption:
 - twocolumn
 - twoside
---

\firstpage[20]{2}

The timing of plant vegetative and reproductive development are major components of plant fitness affected by multiple external environmental cues (e.g. degree of winter chilling, day length, temperature, and water availability) that signal existing or upcoming growing conditions (1--3). Genetic responses to environmental cues determine the speed, timing, and energy apportioned to vegetative and reproductive growth and shape both the individual's lifespan and its lifetime production of viable seed. Day length (or photoperiod) is one of the most predictable environmental cues, and genetic sensitivity to photoperiod protects plants from potentially fatal consequences of phenological responses to temperature cues at the "wrong" time of year. However, the usefulness of specific environmental cues depends on both features of the environment and the species' adaptive strategies (4). Species with wide natural distributions can have multiple distinct environmentally-cued phenological responses: for example, populations of sunflower (Helianthus annuus) exhibit day-neutral, facultative short day, and facultative long-day flowering responses, which vary with their environments (5, 6). Distinct genetic responses in different environments are known as genotype by environment interactions, or GxE.

Flowering time, in particular, is a common subject of GxE research (5--11), a key output of selection driving adaptation to local environments (3, 12, 13), and a selection target for crop improvement to adapt crops to local or future environments (14). Changing flowering responsiveness to photoperiod cues has allowed geographic range expansion and increased yields in a number of cereal species (15--19) and other crops (20, 21). Recent statistical advances in studying phenological GxE have involved determining critical environmental indices before the phenological event occurs, such as photothermal time within a critical growth window (10). However, most studies of flowering GxE focus on finding a single, best fitting form of genotype-environment covariance, despite the key expectation that different genetic subpopulations, and even different genomic regions, have likely evolved distinct patterns of GxE. Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or sign-changing GxE, at the level of individual loci (22--25), previous work has found limited evidence of antagonistic pleiotropy (12, 26), but has been limited by a known statistical bias that reduced detection of SNP effects that differ in sign (26--28). Thus, despite substantial interest in the frequencies of various forms of GxE, the prevalence of antagonistic pleiotropy relative to other forms of GxE remains unknown.

Switchgrass (Panicum virgatum) is considered a short-day plant with reproductive development strongly linked to day of the year (29). However, as part of its wide environmental adaptation across the eastern half of North America, its photoperiodicity has been predicted to differ by plant latitude of origin (30, 31). We previously found divergent Midwest and Gulf genetic subpopulations of switchgrass which segregate for distinct sets of environmental adaptations, based on the analysis of two measures of fitness (i.e.,biomass and survival) among 732 genotypes (32). The Midwest genetic subpopulation is primarily composed of individuals from the well-studied upland switchgrass ecotype (33, 34), while the Gulf subpopulation has individuals from the lowland ecotype and the phenotypically intermediate coastal ecotype (32).

Here, we test how the Midwest and Gulf subpopulations differ in their phenological adaptations across environments, and hence their phenological GxE. We phenotype a diversity panel of hundreds of switchgrass genotypes from the Midwest and Gulf subpopulations for the timing of vegetative development ("green-up") and reproductive development (flowering) at eight common garden locations spanning 17 degrees of latitude. These gardens cover the majority of the latitudinal and climatic range of switchgrass and capture the most comprehensive picture to date of the environmental variation this species encounters. After single-site estimation of SNP effects for these phenotypes, we use multivariate adaptive shrinkage (mash) to jointly estimate SNP effects in each subpopulation and phenotype at all eight sites (35). Mash allows us to identify and specify multiple ways genetic effects linked to SNPs may covary with the environment, and does not have a statistical bias in detecting SNP effects with the same or opposite signs (39). To confirm our genetic mapping of GxE, we compare to mapping results from an outbred pseudo-F2 cross grown at the same sites. Taken together, our results allow us to describe the environmental cues and genetic variation affecting phenology in two divergent natural populations of switchgrass.

# Results

In our diversity panel of tetraploid switchgrass (32), genotypes from the Gulf and Midwest genetic subpopulations had distinct phenological timings and distinct patterns of phenological correlations across our eight common garden sites (Fig. 1A,B). At the three Texas common gardens (hereafter 'Texas' gardens), located within the natural range of the Gulf subpopulation, Gulf green-up occurred before Midwestern green-up, and Gulf flowering occurred after Midwestern flowering (Fig. 1A). At the four northernmost common gardens (hereafter 'North' gardens), located within the natural range of the Midwest subpopulation, both Gulf green-up and flowering occurred after Midwest green-up and flowering. At the Oklahoma common garden, located near the natural range limits of both the Gulf and the Midwest subpopulations, Gulf and Midwest green-up occurred over the same time period. These patterns led to strong negative phenotypic correlations for green-up between the North and Texas gardens, particularly in the Gulf and Both subpopulations, and contributed to positive phenotypic correlations for flowering time of larger magnitude at more northern gardens (Fig. 1B).

```         
Narrow-sense heritabilities (h2) indicated that rank-changing GxE for these phenotypes was present across the common gardens (Fig 1C). h2 were typically high at individual gardens: 59% on average for green-up date, and 87% for flowering date. However, h2 were variable across gardens, and green-up dates were uncorrelated (r2 < 0.2) or negatively correlated between pairs of gardens (Fig. 1B). These negative and small correlations undoubtedly contributed to the low h2 values for green-up and flowering date when estimated jointly at all eight gardens: h2 was 0.8% for green-up and 23.2% for flowering date.
```

### Mapping major patterns of genotype-by-environment effects on green-up and flowering

```         
We next explored how genetic variation in phenology covaried with environmental cues by using mash to jointly discover trait-associated SNPs across all common gardens for the Midwest, Gulf, and ‘Both’ subpopulations (Fig 1D, SI Appendix, Datasets 1-6). We first estimated SNP effects using GWAS on site-specific BLUPs for each phenotype. We then used mash on two subsets of these SNP effects, first to determine the covariance structures that significantly improved the mash model log likelihood when included, then to re-estimate the set of SNP effects with the largest effect estimates in univariate GWAS (hereafter “strong” effects) using the set of significant covariance matrices. 
```

Three types of covariance matrices were tested for inclusion: "canonical" covariance matrices, with simple patterns of effects; "data-driven" matrices derived from common patterns of SNP effects in the data, and "hypothesis-based" covariance matrices that captured the covariance of weather-based phenological cues for cloned genotypes grown at multiple common gardens (Table 1, SI Appendix, Section S2). Mash first assigns mixture proportions for each SNP onto each provided covariance matrix using maximum likelihood. Then, in a second step, each SNP's effect estimate undergoes shrinkage such that it is more aligned with the covariance structure inferred in the first step. The hypothesis-based covariance matrices are an important advantage mash offers for studying patterns of GxE: they allow hypothesis testing of specific environmental drivers for each SNP effect, and loadings on these matrices provide information on genome-wide patterns of SNP-environment interaction. Say that a SNP in FLC controls flowering in a temperature-dependent manner. In that case, that SNP effect could have a high mixture proportion, or mass, on a covariance matrix created using a temperature-based environmental cue, such as average temperature on the day prior to flowering. In our data, we would infer that the effect of that SNP on flowering was caused by a response to that or a correlated environmental cue.

The phenotypic correlations for green up date had moderate negative correlations between the Texas and North gardens, particularly in the Gulf and in Both subpopulations (Fig 1B). If these phenotypic correlations have a genetic component, then they might be controlled by SNP effects that differed in sign across these regions, SNP effects that are large in one garden or region and non-significant in others, or a combination of these patterns. Five hypothesis-based covariance matrices were selected by the greedy mash algorithm, one to two per subpopulation. Of these five, three had mass on them in mash models of the strong effects (Fig 2A, 2B). Two of these three matrices had negative covariances between sites between Texas and North gardens (Fig 2A), while one had all positive or near-zero covariances. SNP-associated phenotypic effects covaried with different weather-based cues in the Gulf & in Both subpopulations (Fig 2B). In total, 65% of the posterior weight of strong SNP effects in the mash model of Gulf green-up fell on a covariance matrix of daylength 14 days prior to the date of green-up. The covariance matrix for this weather cue was very similar to the pattern of phenotypic correlation for green-up in the Gulf subpopulation (Fig 1B; Fig 2A). Mash models of Midwest and Both subpopulation green-up did not include this covariance matrix; the Midwest had no weight on any hypothesis-based matrix, while Both subpopulations had non-zero weights on two additional hypothesis-based matrices, average temperature one day prior to green-up, and the day length change in seconds in the day prior to green-up (Fig 2B). The average temperature covariance matrix had negative covariances between Texas and North gardens, though not as strong as the negative phenotypic correlations seen in Both subpopulations (Fig 1B; Fig 2A). Only the Gulf subpopulation and Both subpopulations had mass on any hypothesis-based covariance matrices (Fig 2C).

For flowering date, distinct weather-based cues captured SNP-associated effect patterns in the Gulf and Midwest subpopulations. 33% of SNP effects on flowering in the Gulf subpopulation covaried with cumulative rainfall in the seven days prior to flowering (Fig 2D). 22.6% of SNP effects on flowering in the Midwest subpopulation contrived with day length change in the two days prior to flowering (Fig 2D), which had negative covariances between Texas and North gardens. Neither covariance matrix was selected for in Both subpopulations, and no hypothesis-based covariance matrices had mass in Both subpopulations (Fig 2E). In five of the six mash models of strong effects, the hypothesis-based covariance matrices captured a minority of the posterior weights of the strong effects (Figure 2C,E); the majority of this mass was on various canonical covariance matrices. These matrices included simple heterozygosity, with intermediate, positive covariances between all gardens, and garden-specific effects present only at one garden.

We next characterized the pairwise patterns of effects where we were confident in the sign of the effect at both gardens. We used the local false sign rate (lfsr), an analogue of the lfdr that establishes confidence in the effect sign, not the effect's difference from zero, to determine significance. We required significance (p \< 0.05) in both gardens to include effects. This means that our tests for antagonistic pleiotropy, or a sign change between conditions, carry an equal statistical burden to those for effects with the same sign. We also summarized the pairwise patterns of effect sign and magnitude within and between the Texas and North regions (Table 2).

For greenup date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, between pairs of Texas and North gardens (Fig. 3A). 78.7% of pairwise comparisons between North and Texas gardens had a difference in sign, while only 28.6% and 0.2% of North-North or Texas-Texas comparisons had a difference in sign, respectively (Table 2). The majority of pairwise effects for greenup for the Midwest (\>55%) and Both (\>85%) subpopulations were the same sign, and effects frequently differed in magnitude between the MO and OK garden and other gardens (Table 2; Fig 3A).

For flowering date for the Gulf subpopulation, less than 2% of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, within or between regions (Table 2). More effects differed in magnitude between the Texas and North regions than within these regions (42.7% vs \<20%; Figure 3B). The Midwest population had relatively few significant effects for flowering, but a large proportion of these differed in sign between Texas and North regions (42.7%) or within the North region (65.4%). Finally, in Both subpopulations, less than 20% of pairwise effects differed in sign (Table 2). Most differences in sign were between TX1, the southernmost garden, and all other gardens (Fig 3B). Similarly, more effects that differed in magnitude included gardens in the Texas region (52.3-55.9%), and most effect pairs in the North region were not distinguishable (91.5%).

### Confirmation of genotype-by-environment effects using an independent mapping population

We sought additional experimental support for our mash intervals using an independent pseudo-F2 mapping population created from Gulf & Midwest individuals and grown at the same sites (Fig. 4A,B). We conducted quantitative trait loci (QTL) mapping of flowering as functions of four environmental cues that we also used as covariance matrices in mash, and identified eight QTL for flowering date, six QTL for flowering GDD, ten QTL for flowering day length, and eight QTL flowering day length change, all of which showed QTL by environment interactions (SI Appendix, Fig. S3). All QTL for flowering overlapped one or more homologs from rice or A. thaliana with functionally validated roles in flowering (SI Appendix, Dataset 7). All flowering QTL intervals contained at least one SNP significant in at least one mash run at a log10-transformed Bayes Factor \> 2, or in the 1% tail of significance, whichever was stricter (SI Appendix, Dataset 8). We also looked for enrichments of mash SNPs in the 1% tail of significance (the 'mash 1% tail') within each QTL interval. At the 5% level, five QTL had enrichments of SNPs in the mash 1% tail. Overall, there were eleven significant enrichments (p \< 0.05, hypergeometric test) of SNPs in the mash 1% tail in the QTL intervals. Our QTL intervals had more enrichments of SNPs in the mash 1% tail than were found for all but eleven of these sets of random genomic intervals (Fig. 4C, p = 0.011). Thus, we were able to experimentally support our mash intervals with a QTL mapping experiment using a separate mapping population.

# Discussion

As the climate and the natural environment change, it is increasingly critical to understand how patterns of gene-environment and plant-environment interactions will change in response. To do this, we must understand the current patterns of trait covariation across environments, the genetic underpinnings of these patterns, and the cases where this covariation can be altered. Here, we demonstrate that we can associate multiple patterns of GxE with specific genomic regions using a switchgrass diversity panel grown at eight common gardens, and also that we can assign specific SNP-associated patterns of GxE to both weather-based cues and to other, data-driven patterns. We use this approach to study GxE in both green-up and flowering phenological data in the deeply genetically diverged Gulf and Midwest populations of switchgrass.

Our analysis of green-up in the Gulf and in Both subpopulations revealed substantial antagonistic pleiotropy in effects between the Texas and North gardens (Figure 3A). This result supports theoretical models that local adaptation should involve antagonistic pleiotropy at the level of individual loci (22--25), and is the first experimental work using QTL mapping and GWAS across common gardens to find antagonistic pleiotropy to be common in small genomic regions (12, 36, 37).

Our analysis of flowering showed that the Gulf and Midwest subpopulations have two distinct photoperiod-related flowering responses: the Midwest subpopulation is day neutral, and flowering is cued primarily by a cumulative GDD threshold; in contrast, the Gulf subpopulation is photoperiod sensitive, and flowering is cued by the transition to shortening days. This result was supported by observations that expressing flowering date as a function of the day length at flowering increased its heritability in the Gulf subpopulation, while expressing flowering date as a function of cumulative GDD between green-up and flowering increased the heritability of flowering in the Midwest subpopulation (Fig. 1D). The genomic regions affecting flowering found in mash were also supported by QTL from an independent mapping population (Fig. 4C).

Identifying the environmental cues that are predictive of, or even correlated with, plant phenotypic responses remains a major challenge to studies interrogating gene action across many natural environments. Clearly, the photoperiod and cumulative GDD cues we identify here are functions of the genotypes measured, are not predictive, and capture only a minority of SNP effects on flowering. We know still less about the overwintering parameters that affect green-up, which is reflected in our lower ability to assign SNP effects on green-up to weather-based cues. More generally, it is difficult to predict the time scales over which individuals may integrate environmental cues, particularly in perennial species which may integrate these cues over longer time scales. Mash offers an opportunity to specify multiple environmental cues and compete them to explain patterns of SNP effects, allowing us to detect how important these cues are genome-wide, and how strongly each cue influences each SNP. This is a key development to further improve our understanding of genetic variation in GxE.

# Materials and Methods

Whenever possible, plant material will be shared upon request. Source data and code to replicate these analyses are available at: https://github.com/Alice-MacQueen/pvdiv-phenology-gxe.git. SNP data to replicate these analyses are available from the UT dataverse at https://doi.org/link.

### Genotype-by-environment effects on green-up and flowering as functions of weather-based cues

In 2019, we scored two phenological events every two days in two mapping populations of switchgrass, a diversity panel and a pseudo-F2 cross, planted at eight common garden locations (32, 34, 37). We scored green-up date as the day of the year when 50% of the tiller area of the crown of the plant cut the previous year had green growth. Flowering date was the day of the year when 50% of the plant tillers had panicles undergoing anthesis. We scored green-up and flowering as day of the year, then linked these dates to multiple weather-based environmental factors measured daily at each common garden (SI Appendix, Section S1, Table S1).

The formation and resequencing of the diversity panel has been described previously (32). The diversity panel contained 134 sequenced, clonally propagated individuals from the Midwest genetic subpopulation, and 229 from the Gulf genetic subpopulation. To allow for the possibility that different subpopulations had different strengths of connection between our phenotypes and genotypes (38), we conducted three sets of genetic analyses: on Gulf and Midwest genotypes separately, and on both subpopulations together ('Both'). Analyses to determine narrow-sense heritability (h2) for green-up and flowering were done using linear mixed models and followed (32). Details on these models can be found in (SI Appendix, Section S3,S4).

### Mapping major patterns of genotype-by-environment effects on green-up and flowering

To evaluate the prevalence and kinds of covariance patterns of SNP effects across our common gardens, we used multivariate adaptive shrinkage (mash) on SNP effect estimates from the diversity panel (35). Mash is a statistical method that allows estimation and comparison of many effects jointly across many different conditions; it improves on previous methods by allowing multiple, arbitrary correlations in effect sizes among conditions (SI Appendix, Section S5). To obtain SNP effect estimates, we first conducted univariate genome-wide association at each common garden for green-up and flowering date. We then analyzed SNP effects for the top 19K relatively unlinked (r2 \< 0.2) SNPs per condition using mash, as in (32). Details on these models can be found in (SI Appendix, Section S6).

We generated hypothesis-based covariance matrices derived from correlations in environmental cues in the green-up or flowering date windows for the three subpopulations (SI Appendix, Table S1, Section S1). These covariance matrices represent correlations between identical genotypes drawn from a specific population at pairs of common gardens; covariances near one mean that the population has a strong, positive linear relationship in individual responses at that pair of gardens, while covariances near zero mean that there is no relationship within the population for individual responses at that pair of gardens. Mash SNP effects will undergo strong shrinkage towards one another in the first case, and little shrinkage in the second case. Mash also generates data-driven covariance matrices corresponding to major patterns of SNP effects present in the data. We generated six data-driven matrices per mash run, five (denoted DD_PCA_1 through DD_PCA_5) produced by singular value decomposition (SVD) of an overall matrix, denoted 'DD_tPCA'. We used SVD to present vectors of garden-specific effects for each numbered DD_PCA matrix, because the first eigenvector of a SVD explains 100% of the variation in these matrices, and each value in this eigenvector represents one garden-specific effect.

Last, we characterized the overall patterns of antagonistic pleiotropy in the set of SNPs where there was pairwise significance of effects at pairs of gardens. To do this, we used the 'get_GxE' function of the switchgrassGWAS R package. First, this determines the set of SNPs with evidence of significant effects in both conditions for all pairs of conditions using local false sign rates (lfsr) as the significance criteria. Second, to determine antagonistic pleiotropy, this function determines if effects significant in both conditions are of opposite sign.

Using the lfsr rather than the local false discovery rate (lfdr) is a critical change in our ability to detect antagonistic pleiotropy. The lfdr, like other measures of FDR, focuses on if we have enough evidence to reject the null hypothesis that an effect j is 0 -- that there is a significant effect. Previous studies of antagonistic pleiotropy (e.g. (37)) have used the lfdr or equivalent statistical tests to detect antagonistic pleiotropy. These tests were conservative, in that they required two non-zero effects of different signs, while tests for differential sensitivity required only one non-zero effect. This previous work recognized that this testing bias could lead to undercounting occurrences of antagonistic pleiotropy (26, 27), and sought to reduce it by permutation (28). However, using the lfsr to test for antagonistic pleiotropy does not undercount occurrences of antagonistic pleiotropy, as this statistic answers a fundamentally different question. For each effect j, the lfsr_j is defined as the probability that we make an error in the sign of effect j if we were forced to declare the effect positive or negative (Stephens paper). Thus, rather than asking "Are these two effects different?" - as we reasonably expect two effects to be, even if this difference cannot be measured - the local false sign rate answers a more meaningful question: Can we be confident in the sign of this effect?

In addition, the get_GxE function also sets an arbitrary threshold to count an effect as changing in magnitude between environments, commonly known as differential sensitivity. For differential sensitivity, this function determines if effects significant in both conditions are of the same sign and of a magnitude (not tested for significance) that differs by a factor of 0.4 or more. The remaining effects that are significant in both conditions have the same effect sign and similar effect magnitudes and we denote these effects as having no GxE. The distinction between effects with different magnitudes is arbitrary but useful to fully characterize how effects vary across environments. and our use of the lfsr to determine significance and our specification that SNP effects must be significant in both conditions to be included means that our tests for antagonistic pleiotropy carry an equal statistical burden to those measuring differential sensitivity and effects without GxE.

### Confirmation of genotype-by-environment effects using an independent mapping population

To confirm candidate genomic regions and patterns of allelic effects found in the diversity panel, we analyzed flowering in an outbred pseudo-F2 cross between four individuals, two Midwest and two Gulf individuals. The formation of this mapping population has been described previously (34); additional details on QTL mapping can be found in SI Appendix, Section S6. To be directly comparable to the diversity panel data, only 2019 phenology data from the pseudo-F2 cross from the same eight common garden sites were used. To compare our QTL enrichments of significant mash associations to the null expectation, we used permutation to choose 1000 sets of 23 genomic regions of the same size randomly distributed throughout the genome, then calculated enrichments of the mash 1% tail in these random intervals.

## Author Affiliations {.unnumbered}

Include department, institution, and complete address, with the ZIP/postal code, for each author. Use lower case letters to match authors with institutions, as shown in the example. PNAS strongly encourages authors to supply an [ORCID identifier](https://orcid.org/) for each author. Individual authors must link their ORCID account to their PNAS account at [www.pnascentral.org](http://www.pnascentral.org/). For proper authentication, authors must provide their ORCID at submission and are not permitted to add ORCIDs on proofs.

## Manuscript Length {.unnumbered}

A standard 6-page article is approximately 4,000 words, 50 references, and 4 medium-size graphical elements (i.e., figures and tables). The preferred length of articles remains at 6 pages, but PNAS will allow articles up to a maximum of 12 pages.

## References {.unnumbered}

References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g. \cite{belkin2002using} and \cite{berard1994embedding,coifman2005geometric,phdthesis,masterthesis}. All references cited in the main text should be included in the main manuscript file.

## Data Archival {.unnumbered}

PNAS must be able to archive the data essential to a published article. Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the [Information for Authors](https://www.pnas.org/author-center/editorial-and-journal-policies#materials-and-data-availability), is acceptable.

## Figures and Tables {.unnumbered}

Figures can be included using Quarto's syntax, code chunks, or with LaTeX.

For simple figures, Quarto syntax is typically easiest. Full details are available in [the Quarto documentation](https://quarto.org/docs/authoring/figures.html).

![Placeholder image of a frog with a long example legend to show justification setting.](figs/frog.pdf){#fig-frog}

Tables can similarly be included using Quarto's syntax, code chunks, or with LaTeX. I tend to use code chunks, but LaTeX works very well for classic tables, like regression tables or summary statistics.

```{=tex}
\begin{table}[t!]
\centering
\caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
\begin{tabular}{lrrr}
Species & CBS & CV & G3 \\
\midrule
1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
\bottomrule
\end{tabular}

\addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
\end{table}
```
### Digital Figures {.unnumbered}

EPS, high-resolution PDF, and PowerPoint are preferred formats for figures that will be used in the main manuscript. Authors may submit PRC or U3D files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format. Color images must be in RGB (red, green, blue) mode. Include the font files for any text.

Images must be provided at final size, preferably 1 column width (8.7cm). Figures wider than 1 column should be sized to 11.4cm or 17.8cm wide. Numbers, letters, and symbols should be no smaller than 6 points (2mm) and no larger than 12 points (6mm) after reduction and must be consistent.

Figures and tables should be labelled and referenced in the standard way using the `\label{}` and `\ref{}` commands.

Figure @fig-frog shows an example of how to insert a column-wide figure. To insert a figure wider than one column, please use the `figure*` environment. Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide. Use `SCfigure*` environment for a wide figure with side legends, such as @fig-side.

In Quarto, we can do these by setting the `fig-env` command to `figure*` or `SCfigure*`

```{r}
#| label: fig-side
#| fig-cap: "This legend would be placed at the side of the figure, rather than below it."
#| fig-env: "SCfigure*"
#| echo: false
# tell it the options as comments with a | and a space, as above.
# set echo: false to avoid printing this text
knitr::include_graphics('figs/frog.pdf')
```

## Tables {.unnumbered}

Tables should be included in the main manuscript file and should not be uploaded separately.

## Single column equations {.unnumbered}

Authors may use 1- or 2-column equations in their article, according to their preference.

To allow an equation to span both columns, use the `\begin{figure*}...\end{figure*}` environment mentioned above for figures.

Note that the use of the `widetext` environment for equations is not recommended, and should not be used.

```{=tex}
\begin{figure*}[bt!]
\begin{align*}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
       &=x^3+3x^2y+3xy^3+x^3.
\end{align*}
\end{figure*}
```
# Material and Methods {.unnumbered}

## Supporting Information Appendix (SI) {.unnumbered}

Authors should submit SI as a single separate SI Appendix PDF file, combining all text, figures, tables, movie legends, and SI references. SI will be published as provided by the authors; it will not be edited or composed. Additional details can be found in the [PNAS Author Center](https://www.pnas.org/authors/submitting-your-manuscript#manuscript-formatting-guidelines). The PNAS Overleaf SI template can be found [here](https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd). Refer to the SI Appendix in the manuscript at an appropriate point in the text. Number supporting figures and tables starting with S1, S2, etc.

Authors who place detailed materials and methods in an SI Appendix must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the SI methods. If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.

### SI Datasets {.unnumbered}

Supply .xlsx, .csv, .txt, .rtf, or .pdf files. This file type will be published in raw format and will not be edited or composed.

### SI Movies {.unnumbered}

Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files. Movie legends should be included in the SI Appendix file. All movies should be submitted at the desired reproduction size and length. Movies should be no more than 10MB in size.

# References

\bibsplit[2]<!-- Use \bibsplit to split the references from the body of the text. Value "[2]" represents the number of reference in the left column (Note: Please avoid single column figures & tables on this page.) -->

::: {#refs}
:::

### 2023-12-02 manuscript/revision2/revision2.tex
Add revision Quarto document in PNAS format

--- 2023-12-02 manuscript/revision2/revision2.qmd
@@
----
-title: Quarto Template for PNAS Submissions
-format:
-  pnas-pdf:
-    keep-tex: true  
-author:
-  - name: Author One
-    affiliations:
-      - name: Harvard University
-        id: a
-        department: Department of Government
-        city: Cambridge
-        state: MA
-        postal-code: 02138
-      - name: Harvard University
-        id: c
-        department: Department of Statistics
-        city: Cambridge
-        state: MA
-        postal-code: 02138
-    attributes:
-      equal-contributor: true
-  - name: Author Two
-    affiliations:
-      - name: Yale University
-        id: b
-        department: Department of Political Science
-        city: New Haven
-        state: CT
-        postal-code: 06511
-    email: alicem@uchicago.edu
-    attributes:
-      equal-contributor: true
-      corresponding: true
-  - name: Author Three
-    affiliations:
-      - ref: a
-abstract: |
-  The timing of vegetative growth and flowering in plants ("phenological timings") multiple depend on environmental cues; thus, phenological traits can have important genotype-by-environment (GxE) interactions.
-Here, we map genotype-by-weather effects on phenological timings in multiple populations of two highly divergent switchgrass (Panicum virgatum) populations spanning the central United States.
-We distinguish GxWeather patterns covarying with interpretable weather-based cues and agnostic, empirically-inferred GxWeather patterns.
-Most GxWeather effects belong to the latter category and do not covary with weather-based cues.
-However, >26% of Gulf population effects on flowering time covary with photoperiod cues, while >__% of Midwest upland population SNPs affecting flowering had effects that covaried with __ cues.
-An independent pseudo-F2 cross of Gulf and Midwest individuals mapped 23 additive QTLs for flowering at the same common gardens, all with significant mash associations and ten with enrichment of highly significant mash associations.
-We demonstrate that we can identify QTL with GxE and assign them to specific weather-based cues and/or data-driven patterns.
-Breeding for particular alleles at these loci could change flowering responsiveness to photoperiod cues in switchgrass.
-More broadly, this approach could be used to identify genetic marker-environment interactions in any species with related populations phenotyped in multiple environments.
-runningauthor: "One, Two, and Three"
-significance: |
-  The timing of plant seasonal development (phenology) has major impacts on fitness because of the steep price of plant-environment mismatches.
-Here we map the genetic basis of two phenological events, the start of above-ground growth and flowering, in two genetically and phenologically distinct populations of switchgrass.
-We do this at eight field locations spanning the latitudinal range of both switchgrass populations.
-Our approach allows us to identify regions of the genome that covary with specific weather-related environmental features at every location.
-For flowering, these features differed by population: the Midwest population had genetic variation that primarily covaried with cumulative growing degree days, a temperature-related measure, while the Gulf population had genetic variation that primarily covaried with photoperiod, a day-length-related measure.
-author_contributions:  "T.E.J.
-designed research.
-D.B.L.
-contributed plant material and resources.
-J.B., D.B.L., and T.E.J.
-designed and executed field experiments.
-A.R.B., P.A.F., F.B.F., D.B.L., R.B.M., F.M.R., Y.W., and T.E.J.
-hosted field experiments.
-A.H.M.
-and L.Z.
-conducted statistical and computational analyses.
-The manuscript was written by A.H.M., L.Z., and T.E.J.
-with contributions from all authors."
-conflict_of_interest: "Please declare any competing interests here."
-equal-contributor: "A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable)."
-keywords:  [allele-by-environment effect variation, antagonistic pleiotropy, photoperiod, cumulative growing degree days, genetic variation]
-acknowledgements: | 
-  We thank the Brackenridge Field laboratory, the Ladybird Johnson Wildflower Center, and the Juenger laboratory for support with plant care and propagation.
-This material is based upon work supported in part by the Great Lakes Bioenergy Research Center, U.S.
-Department of Energy, Office of Science, Office of Biological and Environmental Research under Award Numbers DE-SC0018409 and DE-FC02-07ER64494, the US Department of Energy Awards DESC0014156 to T.E.J., DE-SC0017883 to D.B.L, National Science Foundation PGRP Awards IOS0922457 and IOS1444533 to T.E.J, and the Long-term Ecological Research Program (DEB 1832042) at the Kellogg Biological Station.
-bibliography: bibliography.bib  
-fontsize: "9pt"
-classoption:
- - twocolumn
- - twoside
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+%
+\documentclass[
+  9pt,
+  twocolumn,
+  twoside]{pnas-new}
 
-\firstpage[20]{2}
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-The timing of plant vegetative and reproductive development are major components of plant fitness affected by multiple external environmental cues (e.g.
-degree of winter chilling, day length, temperature, and water availability) that signal existing or upcoming growing conditions (1--3).
-Genetic responses to environmental cues determine the speed, timing, and energy apportioned to vegetative and reproductive growth and shape both the individual's lifespan and its lifetime production of viable seed.
-Day length (or photoperiod) is one of the most predictable environmental cues, and genetic sensitivity to photoperiod protects plants from potentially fatal consequences of phenological responses to temperature cues at the "wrong" time of year.
-However, the usefulness of specific environmental cues depends on both features of the environment and the species' adaptive strategies (4).
-Species with wide natural distributions can have multiple distinct environmentally-cued phenological responses: for example, populations of sunflower (Helianthus annuus) exhibit day-neutral, facultative short day, and facultative long-day flowering responses, which vary with their environments (5, 6).
-Distinct genetic responses in different environments are known as genotype by environment interactions, or GxE.
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+\makeatletter
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
+\makeatother
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+ {\begin{list}{}{%
+  \setlength{\itemindent}{0pt}
+  \setlength{\leftmargin}{0pt}
+  \setlength{\parsep}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+   \setlength{\leftmargin}{\cslhangindent}
+   \setlength{\itemindent}{-1\cslhangindent}
+  \fi
+  % set entry spacing
+  \setlength{\itemsep}{#2\baselineskip}}}
+ {\end{list}}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
-Flowering time, in particular, is a common subject of GxE research (5--11), a key output of selection driving adaptation to local environments (3, 12, 13), and a selection target for crop improvement to adapt crops to local or future environments (14).
-Changing flowering responsiveness to photoperiod cues has allowed geographic range expansion and increased yields in a number of cereal species (15--19) and other crops (20, 21).
-Recent statistical advances in studying phenological GxE have involved determining critical environmental indices before the phenological event occurs, such as photothermal time within a critical growth window (10).
-However, most studies of flowering GxE focus on finding a single, best fitting form of genotype-environment covariance, despite the key expectation that different genetic subpopulations, and even different genomic regions, have likely evolved distinct patterns of GxE.
-Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or sign-changing GxE, at the level of individual loci (22--25), previous work has found limited evidence of antagonistic pleiotropy (12, 26), but has been limited by a known statistical bias that reduced detection of SNP effects that differ in sign (26--28).
-Thus, despite substantial interest in the frequencies of various forms of GxE, the prevalence of antagonistic pleiotropy relative to other forms of GxE remains unknown.
+% TODO: Add custom LaTeX header directives here
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\AtBeginDocument{%
+\ifdefined\contentsname
+  \renewcommand*\contentsname{Table of contents}
+\else
+  \newcommand\contentsname{Table of contents}
+\fi
+\ifdefined\listfigurename
+  \renewcommand*\listfigurename{List of Figures}
+\else
+  \newcommand\listfigurename{List of Figures}
+\fi
+\ifdefined\listtablename
+  \renewcommand*\listtablename{List of Tables}
+\else
+  \newcommand\listtablename{List of Tables}
+\fi
+\ifdefined\figurename
+  \renewcommand*\figurename{Figure}
+\else
+  \newcommand\figurename{Figure}
+\fi
+\ifdefined\tablename
+  \renewcommand*\tablename{Table}
+\else
+  \newcommand\tablename{Table}
+\fi
+}
+\@ifpackageloaded{float}{}{\usepackage{float}}
+\floatstyle{ruled}
+\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
+\floatname{codelisting}{Listing}
+\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\makeatother
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={Quarto Template for PNAS Submissions},
+  pdfauthor={Author One; Author Two; Author Three},
+  pdfkeywords={allele-by-environment effect variation, antagonistic
+pleiotropy, photoperiod, cumulative growing degree days, genetic
+variation},
+  colorlinks=true,
+  linkcolor={blue},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={Blue},
+  pdfcreator={LaTeX via pandoc}}
 
-Switchgrass (Panicum virgatum) is considered a short-day plant with reproductive development strongly linked to day of the year (29).
-However, as part of its wide environmental adaptation across the eastern half of North America, its photoperiodicity has been predicted to differ by plant latitude of origin (30, 31).
-We previously found divergent Midwest and Gulf genetic subpopulations of switchgrass which segregate for distinct sets of environmental adaptations, based on the analysis of two measures of fitness (i.e.,biomass and survival) among 732 genotypes (32).
-The Midwest genetic subpopulation is primarily composed of individuals from the well-studied upland switchgrass ecotype (33, 34), while the Gulf subpopulation has individuals from the lowland ecotype and the phenotypically intermediate coastal ecotype (32).
+\templatetype{pnasresearcharticle}
 
-Here, we test how the Midwest and Gulf subpopulations differ in their phenological adaptations across environments, and hence their phenological GxE.
-We phenotype a diversity panel of hundreds of switchgrass genotypes from the Midwest and Gulf subpopulations for the timing of vegetative development ("green-up") and reproductive development (flowering) at eight common garden locations spanning 17 degrees of latitude.
-These gardens cover the majority of the latitudinal and climatic range of switchgrass and capture the most comprehensive picture to date of the environmental variation this species encounters.
-After single-site estimation of SNP effects for these phenotypes, we use multivariate adaptive shrinkage (mash) to jointly estimate SNP effects in each subpopulation and phenotype at all eight sites (35).
-Mash allows us to identify and specify multiple ways genetic effects linked to SNPs may covary with the environment, and does not have a statistical bias in detecting SNP effects with the same or opposite signs (39).
-To confirm our genetic mapping of GxE, we compare to mapping results from an outbred pseudo-F2 cross grown at the same sites.
-Taken together, our results allow us to describe the environmental cues and genetic variation affecting phenology in two divergent natural populations of switchgrass.
+\title{Quarto Template for PNAS Submissions}
 
-# Results
+\newcommand{\equalcont}{1}
 
-In our diversity panel of tetraploid switchgrass (32), genotypes from the Gulf and Midwest genetic subpopulations had distinct phenological timings and distinct patterns of phenological correlations across our eight common garden sites (Fig.
-1A,B).
-At the three Texas common gardens (hereafter 'Texas' gardens), located within the natural range of the Gulf subpopulation, Gulf green-up occurred before Midwestern green-up, and Gulf flowering occurred after Midwestern flowering (Fig.
-1A).
-At the four northernmost common gardens (hereafter 'North' gardens), located within the natural range of the Midwest subpopulation, both Gulf green-up and flowering occurred after Midwest green-up and flowering.
-At the Oklahoma common garden, located near the natural range limits of both the Gulf and the Midwest subpopulations, Gulf and Midwest green-up occurred over the same time period.
-These patterns led to strong negative phenotypic correlations for green-up between the North and Texas gardens, particularly in the Gulf and Both subpopulations, and contributed to positive phenotypic correlations for flowering time of larger magnitude at more northern gardens (Fig.
-1B).
+\newcommand{\correspond}{2}
 
-```         
+\author[a,c%
+,\equalcont%
+%
+]{Author One}
+\author[b%
+,\equalcont%
+,\correspond%
+]{Author Two}
+\author[a%
+%
+%
+]{Author Three}
+
+\affil[a]{Harvard University, Department of Government, Cambridge, 2138}
+\affil[c]{Harvard University, Department of Statistics, Cambridge, 2138}
+\affil[b]{Yale University, Department of Political Science, New
+Haven, 3401}
+
+\leadauthor{One, Two, and Three}
+
+\significancestatement{The timing of plant seasonal development
+(phenology) has major impacts on fitness because of the steep price of
+plant-environment mismatches. Here we map the genetic basis of two
+phenological events, the start of above-ground growth and flowering, in
+two genetically and phenologically distinct populations of switchgrass.
+We do this at eight field locations spanning the latitudinal range of
+both switchgrass populations. Our approach allows us to identify regions
+of the genome that covary with specific weather-related environmental
+features at every location. For flowering, these features differed by
+population: the Midwest population had genetic variation that primarily
+covaried with cumulative growing degree days, a temperature-related
+measure, while the Gulf population had genetic variation that primarily
+covaried with photoperiod, a day-length-related measure.}
+
+\authorcontributions{T.E.J. designed research. D.B.L. contributed plant
+material and resources. J.B., D.B.L., and T.E.J. designed and executed
+field experiments. A.R.B., P.A.F., F.B.F., D.B.L., R.B.M., F.M.R., Y.W.,
+and T.E.J. hosted field experiments. A.H.M. and L.Z. conducted
+statistical and computational analyses. The manuscript was written by
+A.H.M., L.Z., and T.E.J. with contributions from all authors.}
+
+\authordeclaration{Please declare any competing interests here.}
+
+\equalauthors{\textsuperscript{\equalcont} A.O.(Author One) contributed
+equally to this work with A.T. (Author Two) (remove if not applicable).}
+
+\correspondingauthor{\textsuperscript{\correspond}To whom correspondence should be addressed. E-mail: alicem@uchicago.edu}
+
+\keywords{allele-by-environment effect variation | antagonistic
+pleiotropy | photoperiod | cumulative growing degree days | genetic
+variation}
+
+\begin{document}
+\maketitle
+
+\begin{abstract}
+The timing of vegetative growth and flowering in plants (``phenological
+timings'') multiple depend on environmental cues; thus, phenological
+traits can have important genotype-by-environment (GxE) interactions.
+Here, we map genotype-by-weather effects on phenological timings in
+multiple populations of two highly divergent switchgrass (Panicum
+virgatum) populations spanning the central United States. We distinguish
+GxWeather patterns covarying with interpretable weather-based cues and
+agnostic, empirically-inferred GxWeather patterns. Most GxWeather
+effects belong to the latter category and do not covary with
+weather-based cues. However, \textgreater26\% of Gulf population effects
+on flowering time covary with photoperiod cues, while
+\textgreater{}\textbf{\% of Midwest upland population SNPs affecting
+flowering had effects that covaried with } cues. An independent
+pseudo-F2 cross of Gulf and Midwest individuals mapped 23 additive QTLs
+for flowering at the same common gardens, all with significant mash
+associations and ten with enrichment of highly significant mash
+associations. We demonstrate that we can identify QTL with GxE and
+assign them to specific weather-based cues and/or data-driven patterns.
+Breeding for particular alleles at these loci could change flowering
+responsiveness to photoperiod cues in switchgrass. More broadly, this
+approach could be used to identify genetic marker-environment
+interactions in any species with related populations phenotyped in
+multiple environments.
+\end{abstract}
+
+\dates{This manuscript was compiled on \today}
+\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
+
+\thispagestyle{firststyle}
+\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}\firstpage[20]{2}
+
+The timing of plant vegetative and reproductive development are major
+components of plant fitness affected by multiple external environmental
+cues (e.g.~degree of winter chilling, day length, temperature, and water
+availability) that signal existing or upcoming growing conditions
+(1--3). Genetic responses to environmental cues determine the speed,
+timing, and energy apportioned to vegetative and reproductive growth and
+shape both the individual's lifespan and its lifetime production of
+viable seed. Day length (or photoperiod) is one of the most predictable
+environmental cues, and genetic sensitivity to photoperiod protects
+plants from potentially fatal consequences of phenological responses to
+temperature cues at the ``wrong'' time of year. However, the usefulness
+of specific environmental cues depends on both features of the
+environment and the species' adaptive strategies (4). Species with wide
+natural distributions can have multiple distinct environmentally-cued
+phenological responses: for example, populations of sunflower
+(Helianthus annuus) exhibit day-neutral, facultative short day, and
+facultative long-day flowering responses, which vary with their
+environments (5, 6). Distinct genetic responses in different
+environments are known as genotype by environment interactions, or GxE.
+
+Flowering time, in particular, is a common subject of GxE research
+(5--11), a key output of selection driving adaptation to local
+environments (3, 12, 13), and a selection target for crop improvement to
+adapt crops to local or future environments (14). Changing flowering
+responsiveness to photoperiod cues has allowed geographic range
+expansion and increased yields in a number of cereal species (15--19)
+and other crops (20, 21). Recent statistical advances in studying
+phenological GxE have involved determining critical environmental
+indices before the phenological event occurs, such as photothermal time
+within a critical growth window (10). However, most studies of flowering
+GxE focus on finding a single, best fitting form of genotype-environment
+covariance, despite the key expectation that different genetic
+subpopulations, and even different genomic regions, have likely evolved
+distinct patterns of GxE. Additionally, despite theoretical predictions
+that local adaptation should involve antagonistic pleiotropy, or
+sign-changing GxE, at the level of individual loci (22--25), previous
+work has found limited evidence of antagonistic pleiotropy (12, 26), but
+has been limited by a known statistical bias that reduced detection of
+SNP effects that differ in sign (26--28). Thus, despite substantial
+interest in the frequencies of various forms of GxE, the prevalence of
+antagonistic pleiotropy relative to other forms of GxE remains unknown.
+
+Switchgrass (Panicum virgatum) is considered a short-day plant with
+reproductive development strongly linked to day of the year (29).
+However, as part of its wide environmental adaptation across the eastern
+half of North America, its photoperiodicity has been predicted to differ
+by plant latitude of origin (30, 31). We previously found divergent
+Midwest and Gulf genetic subpopulations of switchgrass which segregate
+for distinct sets of environmental adaptations, based on the analysis of
+two measures of fitness (i.e.,biomass and survival) among 732 genotypes
+(32). The Midwest genetic subpopulation is primarily composed of
+individuals from the well-studied upland switchgrass ecotype (33, 34),
+while the Gulf subpopulation has individuals from the lowland ecotype
+and the phenotypically intermediate coastal ecotype (32).
+
+Here, we test how the Midwest and Gulf subpopulations differ in their
+phenological adaptations across environments, and hence their
+phenological GxE. We phenotype a diversity panel of hundreds of
+switchgrass genotypes from the Midwest and Gulf subpopulations for the
+timing of vegetative development (``green-up'') and reproductive
+development (flowering) at eight common garden locations spanning 17
+degrees of latitude. These gardens cover the majority of the latitudinal
+and climatic range of switchgrass and capture the most comprehensive
+picture to date of the environmental variation this species encounters.
+After single-site estimation of SNP effects for these phenotypes, we use
+multivariate adaptive shrinkage (mash) to jointly estimate SNP effects
+in each subpopulation and phenotype at all eight sites (35). Mash allows
+us to identify and specify multiple ways genetic effects linked to SNPs
+may covary with the environment, and does not have a statistical bias in
+detecting SNP effects with the same or opposite signs (39). To confirm
+our genetic mapping of GxE, we compare to mapping results from an
+outbred pseudo-F2 cross grown at the same sites. Taken together, our
+results allow us to describe the environmental cues and genetic
+variation affecting phenology in two divergent natural populations of
+switchgrass.
+
+\section{Results}\label{results}
+
+In our diversity panel of tetraploid switchgrass (32), genotypes from
+the Gulf and Midwest genetic subpopulations had distinct phenological
+timings and distinct patterns of phenological correlations across our
+eight common garden sites (Fig. 1A,B). At the three Texas common gardens
+(hereafter `Texas' gardens), located within the natural range of the
+Gulf subpopulation, Gulf green-up occurred before Midwestern green-up,
+and Gulf flowering occurred after Midwestern flowering (Fig. 1A). At the
+four northernmost common gardens (hereafter `North' gardens), located
+within the natural range of the Midwest subpopulation, both Gulf
+green-up and flowering occurred after Midwest green-up and flowering. At
+the Oklahoma common garden, located near the natural range limits of
+both the Gulf and the Midwest subpopulations, Gulf and Midwest green-up
+occurred over the same time period. These patterns led to strong
+negative phenotypic correlations for green-up between the North and
+Texas gardens, particularly in the Gulf and Both subpopulations, and
+contributed to positive phenotypic correlations for flowering time of
+larger magnitude at more northern gardens (Fig. 1B).
+
+\begin{verbatim}
 Narrow-sense heritabilities (h2) indicated that rank-changing GxE for these phenotypes was present across the common gardens (Fig 1C).
@@
 These negative and small correlations undoubtedly contributed to the low h2 values for green-up and flowering date when estimated jointly at all eight gardens: h2 was 0.8% for green-up and 23.2% for flowering date.
-```
+\end{verbatim}
 
-### Mapping major patterns of genotype-by-environment effects on green-up and flowering
+\subsubsection{Mapping major patterns of genotype-by-environment effects
+on green-up and
+flowering}\label{mapping-major-patterns-of-genotype-by-environment-effects-on-green-up-and-flowering}
 
-```         
+\begin{verbatim}
 We next explored how genetic variation in phenology covaried with environmental cues by using mash to jointly discover trait-associated SNPs across all common gardens for the Midwest, Gulf, and ‘Both’ subpopulations (Fig 1D, SI Appendix, Datasets 1-6).
@@
 We then used mash on two subsets of these SNP effects, first to determine the covariance structures that significantly improved the mash model log likelihood when included, then to re-estimate the set of SNP effects with the largest effect estimates in univariate GWAS (hereafter “strong” effects) using the set of significant covariance matrices.
-```
+\end{verbatim}
 
-Three types of covariance matrices were tested for inclusion: "canonical" covariance matrices, with simple patterns of effects; "data-driven" matrices derived from common patterns of SNP effects in the data, and "hypothesis-based" covariance matrices that captured the covariance of weather-based phenological cues for cloned genotypes grown at multiple common gardens (Table 1, SI Appendix, Section S2).
-Mash first assigns mixture proportions for each SNP onto each provided covariance matrix using maximum likelihood.
-Then, in a second step, each SNP's effect estimate undergoes shrinkage such that it is more aligned with the covariance structure inferred in the first step.
-The hypothesis-based covariance matrices are an important advantage mash offers for studying patterns of GxE: they allow hypothesis testing of specific environmental drivers for each SNP effect, and loadings on these matrices provide information on genome-wide patterns of SNP-environment interaction.
-Say that a SNP in FLC controls flowering in a temperature-dependent manner.
-In that case, that SNP effect could have a high mixture proportion, or mass, on a covariance matrix created using a temperature-based environmental cue, such as average temperature on the day prior to flowering.
-In our data, we would infer that the effect of that SNP on flowering was caused by a response to that or a correlated environmental cue.
+Three types of covariance matrices were tested for inclusion:
+``canonical'' covariance matrices, with simple patterns of effects;
+``data-driven'' matrices derived from common patterns of SNP effects in
+the data, and ``hypothesis-based'' covariance matrices that captured the
+covariance of weather-based phenological cues for cloned genotypes grown
+at multiple common gardens (Table 1, SI Appendix, Section S2). Mash
+first assigns mixture proportions for each SNP onto each provided
+covariance matrix using maximum likelihood. Then, in a second step, each
+SNP's effect estimate undergoes shrinkage such that it is more aligned
+with the covariance structure inferred in the first step. The
+hypothesis-based covariance matrices are an important advantage mash
+offers for studying patterns of GxE: they allow hypothesis testing of
+specific environmental drivers for each SNP effect, and loadings on
+these matrices provide information on genome-wide patterns of
+SNP-environment interaction. Say that a SNP in FLC controls flowering in
+a temperature-dependent manner. In that case, that SNP effect could have
+a high mixture proportion, or mass, on a covariance matrix created using
+a temperature-based environmental cue, such as average temperature on
+the day prior to flowering. In our data, we would infer that the effect
+of that SNP on flowering was caused by a response to that or a
+correlated environmental cue.
 
-The phenotypic correlations for green up date had moderate negative correlations between the Texas and North gardens, particularly in the Gulf and in Both subpopulations (Fig 1B).
-If these phenotypic correlations have a genetic component, then they might be controlled by SNP effects that differed in sign across these regions, SNP effects that are large in one garden or region and non-significant in others, or a combination of these patterns.
-Five hypothesis-based covariance matrices were selected by the greedy mash algorithm, one to two per subpopulation.
-Of these five, three had mass on them in mash models of the strong effects (Fig 2A, 2B).
-Two of these three matrices had negative covariances between sites between Texas and North gardens (Fig 2A), while one had all positive or near-zero covariances.
-SNP-associated phenotypic effects covaried with different weather-based cues in the Gulf & in Both subpopulations (Fig 2B).
-In total, 65% of the posterior weight of strong SNP effects in the mash model of Gulf green-up fell on a covariance matrix of daylength 14 days prior to the date of green-up.
-The covariance matrix for this weather cue was very similar to the pattern of phenotypic correlation for green-up in the Gulf subpopulation (Fig 1B; Fig 2A).
-Mash models of Midwest and Both subpopulation green-up did not include this covariance matrix; the Midwest had no weight on any hypothesis-based matrix, while Both subpopulations had non-zero weights on two additional hypothesis-based matrices, average temperature one day prior to green-up, and the day length change in seconds in the day prior to green-up (Fig 2B).
-The average temperature covariance matrix had negative covariances between Texas and North gardens, though not as strong as the negative phenotypic correlations seen in Both subpopulations (Fig 1B; Fig 2A).
-Only the Gulf subpopulation and Both subpopulations had mass on any hypothesis-based covariance matrices (Fig 2C).
+The phenotypic correlations for green up date had moderate negative
+correlations between the Texas and North gardens, particularly in the
+Gulf and in Both subpopulations (Fig 1B). If these phenotypic
+correlations have a genetic component, then they might be controlled by
+SNP effects that differed in sign across these regions, SNP effects that
+are large in one garden or region and non-significant in others, or a
+combination of these patterns. Five hypothesis-based covariance matrices
+were selected by the greedy mash algorithm, one to two per
+subpopulation. Of these five, three had mass on them in mash models of
+the strong effects (Fig 2A, 2B). Two of these three matrices had
+negative covariances between sites between Texas and North gardens (Fig
+2A), while one had all positive or near-zero covariances. SNP-associated
+phenotypic effects covaried with different weather-based cues in the
+Gulf \& in Both subpopulations (Fig 2B). In total, 65\% of the posterior
+weight of strong SNP effects in the mash model of Gulf green-up fell on
+a covariance matrix of daylength 14 days prior to the date of green-up.
+The covariance matrix for this weather cue was very similar to the
+pattern of phenotypic correlation for green-up in the Gulf subpopulation
+(Fig 1B; Fig 2A). Mash models of Midwest and Both subpopulation green-up
+did not include this covariance matrix; the Midwest had no weight on any
+hypothesis-based matrix, while Both subpopulations had non-zero weights
+on two additional hypothesis-based matrices, average temperature one day
+prior to green-up, and the day length change in seconds in the day prior
+to green-up (Fig 2B). The average temperature covariance matrix had
+negative covariances between Texas and North gardens, though not as
+strong as the negative phenotypic correlations seen in Both
+subpopulations (Fig 1B; Fig 2A). Only the Gulf subpopulation and Both
+subpopulations had mass on any hypothesis-based covariance matrices (Fig
+2C).
 
-For flowering date, distinct weather-based cues captured SNP-associated effect patterns in the Gulf and Midwest subpopulations.
-33% of SNP effects on flowering in the Gulf subpopulation covaried with cumulative rainfall in the seven days prior to flowering (Fig 2D).
-22.6% of SNP effects on flowering in the Midwest subpopulation contrived with day length change in the two days prior to flowering (Fig 2D), which had negative covariances between Texas and North gardens.
-Neither covariance matrix was selected for in Both subpopulations, and no hypothesis-based covariance matrices had mass in Both subpopulations (Fig 2E).
-In five of the six mash models of strong effects, the hypothesis-based covariance matrices captured a minority of the posterior weights of the strong effects (Figure 2C,E); the majority of this mass was on various canonical covariance matrices.
-These matrices included simple heterozygosity, with intermediate, positive covariances between all gardens, and garden-specific effects present only at one garden.
+For flowering date, distinct weather-based cues captured SNP-associated
+effect patterns in the Gulf and Midwest subpopulations. 33\% of SNP
+effects on flowering in the Gulf subpopulation covaried with cumulative
+rainfall in the seven days prior to flowering (Fig 2D). 22.6\% of SNP
+effects on flowering in the Midwest subpopulation contrived with day
+length change in the two days prior to flowering (Fig 2D), which had
+negative covariances between Texas and North gardens. Neither covariance
+matrix was selected for in Both subpopulations, and no hypothesis-based
+covariance matrices had mass in Both subpopulations (Fig 2E). In five of
+the six mash models of strong effects, the hypothesis-based covariance
+matrices captured a minority of the posterior weights of the strong
+effects (Figure 2C,E); the majority of this mass was on various
+canonical covariance matrices. These matrices included simple
+heterozygosity, with intermediate, positive covariances between all
+gardens, and garden-specific effects present only at one garden.
 
-We next characterized the pairwise patterns of effects where we were confident in the sign of the effect at both gardens.
-We used the local false sign rate (lfsr), an analogue of the lfdr that establishes confidence in the effect sign, not the effect's difference from zero, to determine significance.
-We required significance (p \< 0.05) in both gardens to include effects.
-This means that our tests for antagonistic pleiotropy, or a sign change between conditions, carry an equal statistical burden to those for effects with the same sign.
-We also summarized the pairwise patterns of effect sign and magnitude within and between the Texas and North regions (Table 2).
+We next characterized the pairwise patterns of effects where we were
+confident in the sign of the effect at both gardens. We used the local
+false sign rate (lfsr), an analogue of the lfdr that establishes
+confidence in the effect sign, not the effect's difference from zero, to
+determine significance. We required significance (p \textless{} 0.05) in
+both gardens to include effects. This means that our tests for
+antagonistic pleiotropy, or a sign change between conditions, carry an
+equal statistical burden to those for effects with the same sign. We
+also summarized the pairwise patterns of effect sign and magnitude
+within and between the Texas and North regions (Table 2).
 
-For greenup date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, between pairs of Texas and North gardens (Fig.
-3A).
-78.7% of pairwise comparisons between North and Texas gardens had a difference in sign, while only 28.6% and 0.2% of North-North or Texas-Texas comparisons had a difference in sign, respectively (Table 2).
-The majority of pairwise effects for greenup for the Midwest (\>55%) and Both (\>85%) subpopulations were the same sign, and effects frequently differed in magnitude between the MO and OK garden and other gardens (Table 2; Fig 3A).
+For greenup date for the Gulf subpopulation, hundreds to thousands of
+pairwise effects exhibited antagonistic pleiotropy, or a difference in
+effect sign, between pairs of Texas and North gardens (Fig. 3A). 78.7\%
+of pairwise comparisons between North and Texas gardens had a difference
+in sign, while only 28.6\% and 0.2\% of North-North or Texas-Texas
+comparisons had a difference in sign, respectively (Table 2). The
+majority of pairwise effects for greenup for the Midwest
+(\textgreater55\%) and Both (\textgreater85\%) subpopulations were the
+same sign, and effects frequently differed in magnitude between the MO
+and OK garden and other gardens (Table 2; Fig 3A).
 
-For flowering date for the Gulf subpopulation, less than 2% of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, within or between regions (Table 2).
-More effects differed in magnitude between the Texas and North regions than within these regions (42.7% vs \<20%; Figure 3B).
-The Midwest population had relatively few significant effects for flowering, but a large proportion of these differed in sign between Texas and North regions (42.7%) or within the North region (65.4%).
-Finally, in Both subpopulations, less than 20% of pairwise effects differed in sign (Table 2).
-Most differences in sign were between TX1, the southernmost garden, and all other gardens (Fig 3B).
-Similarly, more effects that differed in magnitude included gardens in the Texas region (52.3-55.9%), and most effect pairs in the North region were not distinguishable (91.5%).
+For flowering date for the Gulf subpopulation, less than 2\% of pairwise
+effects exhibited antagonistic pleiotropy, or a difference in effect
+sign, within or between regions (Table 2). More effects differed in
+magnitude between the Texas and North regions than within these regions
+(42.7\% vs \textless20\%; Figure 3B). The Midwest population had
+relatively few significant effects for flowering, but a large proportion
+of these differed in sign between Texas and North regions (42.7\%) or
+within the North region (65.4\%). Finally, in Both subpopulations, less
+than 20\% of pairwise effects differed in sign (Table 2). Most
+differences in sign were between TX1, the southernmost garden, and all
+other gardens (Fig 3B). Similarly, more effects that differed in
+magnitude included gardens in the Texas region (52.3-55.9\%), and most
+effect pairs in the North region were not distinguishable (91.5\%).
 
-### Confirmation of genotype-by-environment effects using an independent mapping population
+\subsubsection{Confirmation of genotype-by-environment effects using an
+independent mapping
+population}\label{confirmation-of-genotype-by-environment-effects-using-an-independent-mapping-population}
 
-We sought additional experimental support for our mash intervals using an independent pseudo-F2 mapping population created from Gulf & Midwest individuals and grown at the same sites (Fig.
-4A,B).
-We conducted quantitative trait loci (QTL) mapping of flowering as functions of four environmental cues that we also used as covariance matrices in mash, and identified eight QTL for flowering date, six QTL for flowering GDD, ten QTL for flowering day length, and eight QTL flowering day length change, all of which showed QTL by environment interactions (SI Appendix, Fig.
-S3).
-All QTL for flowering overlapped one or more homologs from rice or A.
-thaliana with functionally validated roles in flowering (SI Appendix, Dataset 7).
-All flowering QTL intervals contained at least one SNP significant in at least one mash run at a log10-transformed Bayes Factor \> 2, or in the 1% tail of significance, whichever was stricter (SI Appendix, Dataset 8).
-We also looked for enrichments of mash SNPs in the 1% tail of significance (the 'mash 1% tail') within each QTL interval.
-At the 5% level, five QTL had enrichments of SNPs in the mash 1% tail.
-Overall, there were eleven significant enrichments (p \< 0.05, hypergeometric test) of SNPs in the mash 1% tail in the QTL intervals.
-Our QTL intervals had more enrichments of SNPs in the mash 1% tail than were found for all but eleven of these sets of random genomic intervals (Fig.
-4C, p = 0.011).
-Thus, we were able to experimentally support our mash intervals with a QTL mapping experiment using a separate mapping population.
+We sought additional experimental support for our mash intervals using
+an independent pseudo-F2 mapping population created from Gulf \& Midwest
+individuals and grown at the same sites (Fig. 4A,B). We conducted
+quantitative trait loci (QTL) mapping of flowering as functions of four
+environmental cues that we also used as covariance matrices in mash, and
+identified eight QTL for flowering date, six QTL for flowering GDD, ten
+QTL for flowering day length, and eight QTL flowering day length change,
+all of which showed QTL by environment interactions (SI Appendix, Fig.
+S3). All QTL for flowering overlapped one or more homologs from rice or
+A. thaliana with functionally validated roles in flowering (SI Appendix,
+Dataset 7). All flowering QTL intervals contained at least one SNP
+significant in at least one mash run at a log10-transformed Bayes Factor
+\textgreater{} 2, or in the 1\% tail of significance, whichever was
+stricter (SI Appendix, Dataset 8). We also looked for enrichments of
+mash SNPs in the 1\% tail of significance (the `mash 1\% tail') within
+each QTL interval. At the 5\% level, five QTL had enrichments of SNPs in
+the mash 1\% tail. Overall, there were eleven significant enrichments (p
+\textless{} 0.05, hypergeometric test) of SNPs in the mash 1\% tail in
+the QTL intervals. Our QTL intervals had more enrichments of SNPs in the
+mash 1\% tail than were found for all but eleven of these sets of random
+genomic intervals (Fig. 4C, p = 0.011). Thus, we were able to
+experimentally support our mash intervals with a QTL mapping experiment
+using a separate mapping population.
 
-# Discussion
+\section{Discussion}\label{discussion}
 
-As the climate and the natural environment change, it is increasingly critical to understand how patterns of gene-environment and plant-environment interactions will change in response.
-To do this, we must understand the current patterns of trait covariation across environments, the genetic underpinnings of these patterns, and the cases where this covariation can be altered.
-Here, we demonstrate that we can associate multiple patterns of GxE with specific genomic regions using a switchgrass diversity panel grown at eight common gardens, and also that we can assign specific SNP-associated patterns of GxE to both weather-based cues and to other, data-driven patterns.
-We use this approach to study GxE in both green-up and flowering phenological data in the deeply genetically diverged Gulf and Midwest populations of switchgrass.
+As the climate and the natural environment change, it is increasingly
+critical to understand how patterns of gene-environment and
+plant-environment interactions will change in response. To do this, we
+must understand the current patterns of trait covariation across
+environments, the genetic underpinnings of these patterns, and the cases
+where this covariation can be altered. Here, we demonstrate that we can
+associate multiple patterns of GxE with specific genomic regions using a
+switchgrass diversity panel grown at eight common gardens, and also that
+we can assign specific SNP-associated patterns of GxE to both
+weather-based cues and to other, data-driven patterns. We use this
+approach to study GxE in both green-up and flowering phenological data
+in the deeply genetically diverged Gulf and Midwest populations of
+switchgrass.
 
-Our analysis of green-up in the Gulf and in Both subpopulations revealed substantial antagonistic pleiotropy in effects between the Texas and North gardens (Figure 3A).
-This result supports theoretical models that local adaptation should involve antagonistic pleiotropy at the level of individual loci (22--25), and is the first experimental work using QTL mapping and GWAS across common gardens to find antagonistic pleiotropy to be common in small genomic regions (12, 36, 37).
+Our analysis of green-up in the Gulf and in Both subpopulations revealed
+substantial antagonistic pleiotropy in effects between the Texas and
+North gardens (Figure 3A). This result supports theoretical models that
+local adaptation should involve antagonistic pleiotropy at the level of
+individual loci (22--25), and is the first experimental work using QTL
+mapping and GWAS across common gardens to find antagonistic pleiotropy
+to be common in small genomic regions (12, 36, 37).
 
-Our analysis of flowering showed that the Gulf and Midwest subpopulations have two distinct photoperiod-related flowering responses: the Midwest subpopulation is day neutral, and flowering is cued primarily by a cumulative GDD threshold; in contrast, the Gulf subpopulation is photoperiod sensitive, and flowering is cued by the transition to shortening days.
-This result was supported by observations that expressing flowering date as a function of the day length at flowering increased its heritability in the Gulf subpopulation, while expressing flowering date as a function of cumulative GDD between green-up and flowering increased the heritability of flowering in the Midwest subpopulation (Fig.
-1D).
-The genomic regions affecting flowering found in mash were also supported by QTL from an independent mapping population (Fig.
-4C).
+Our analysis of flowering showed that the Gulf and Midwest
+subpopulations have two distinct photoperiod-related flowering
+responses: the Midwest subpopulation is day neutral, and flowering is
+cued primarily by a cumulative GDD threshold; in contrast, the Gulf
+subpopulation is photoperiod sensitive, and flowering is cued by the
+transition to shortening days. This result was supported by observations
+that expressing flowering date as a function of the day length at
+flowering increased its heritability in the Gulf subpopulation, while
+expressing flowering date as a function of cumulative GDD between
+green-up and flowering increased the heritability of flowering in the
+Midwest subpopulation (Fig. 1D). The genomic regions affecting flowering
+found in mash were also supported by QTL from an independent mapping
+population (Fig. 4C).
 
-Identifying the environmental cues that are predictive of, or even correlated with, plant phenotypic responses remains a major challenge to studies interrogating gene action across many natural environments.
-Clearly, the photoperiod and cumulative GDD cues we identify here are functions of the genotypes measured, are not predictive, and capture only a minority of SNP effects on flowering.
-We know still less about the overwintering parameters that affect green-up, which is reflected in our lower ability to assign SNP effects on green-up to weather-based cues.
-More generally, it is difficult to predict the time scales over which individuals may integrate environmental cues, particularly in perennial species which may integrate these cues over longer time scales.
-Mash offers an opportunity to specify multiple environmental cues and compete them to explain patterns of SNP effects, allowing us to detect how important these cues are genome-wide, and how strongly each cue influences each SNP.
-This is a key development to further improve our understanding of genetic variation in GxE.
+Identifying the environmental cues that are predictive of, or even
+correlated with, plant phenotypic responses remains a major challenge to
+studies interrogating gene action across many natural environments.
+Clearly, the photoperiod and cumulative GDD cues we identify here are
+functions of the genotypes measured, are not predictive, and capture
+only a minority of SNP effects on flowering. We know still less about
+the overwintering parameters that affect green-up, which is reflected in
+our lower ability to assign SNP effects on green-up to weather-based
+cues. More generally, it is difficult to predict the time scales over
+which individuals may integrate environmental cues, particularly in
+perennial species which may integrate these cues over longer time
+scales. Mash offers an opportunity to specify multiple environmental
+cues and compete them to explain patterns of SNP effects, allowing us to
+detect how important these cues are genome-wide, and how strongly each
+cue influences each SNP. This is a key development to further improve
+our understanding of genetic variation in GxE.
 
-# Materials and Methods
+\section{Materials and Methods}\label{materials-and-methods}
 
-Whenever possible, plant material will be shared upon request.
-Source data and code to replicate these analyses are available at: https://github.com/Alice-MacQueen/pvdiv-phenology-gxe.git.
-SNP data to replicate these analyses are available from the UT dataverse at https://doi.org/link.
+Whenever possible, plant material will be shared upon request. Source
+data and code to replicate these analyses are available at:
+https://github.com/Alice-MacQueen/pvdiv-phenology-gxe.git. SNP data to
+replicate these analyses are available from the UT dataverse at
+https://doi.org/link.
 
-### Genotype-by-environment effects on green-up and flowering as functions of weather-based cues
+\subsubsection{Genotype-by-environment effects on green-up and flowering
+as functions of weather-based
+cues}\label{genotype-by-environment-effects-on-green-up-and-flowering-as-functions-of-weather-based-cues}
 
-In 2019, we scored two phenological events every two days in two mapping populations of switchgrass, a diversity panel and a pseudo-F2 cross, planted at eight common garden locations (32, 34, 37).
-We scored green-up date as the day of the year when 50% of the tiller area of the crown of the plant cut the previous year had green growth.
-Flowering date was the day of the year when 50% of the plant tillers had panicles undergoing anthesis.
-We scored green-up and flowering as day of the year, then linked these dates to multiple weather-based environmental factors measured daily at each common garden (SI Appendix, Section S1, Table S1).
+In 2019, we scored two phenological events every two days in two mapping
+populations of switchgrass, a diversity panel and a pseudo-F2 cross,
+planted at eight common garden locations (32, 34, 37). We scored
+green-up date as the day of the year when 50\% of the tiller area of the
+crown of the plant cut the previous year had green growth. Flowering
+date was the day of the year when 50\% of the plant tillers had panicles
+undergoing anthesis. We scored green-up and flowering as day of the
+year, then linked these dates to multiple weather-based environmental
+factors measured daily at each common garden (SI Appendix, Section S1,
+Table S1).
 
-The formation and resequencing of the diversity panel has been described previously (32).
-The diversity panel contained 134 sequenced, clonally propagated individuals from the Midwest genetic subpopulation, and 229 from the Gulf genetic subpopulation.
-To allow for the possibility that different subpopulations had different strengths of connection between our phenotypes and genotypes (38), we conducted three sets of genetic analyses: on Gulf and Midwest genotypes separately, and on both subpopulations together ('Both').
-Analyses to determine narrow-sense heritability (h2) for green-up and flowering were done using linear mixed models and followed (32).
-Details on these models can be found in (SI Appendix, Section S3,S4).
+The formation and resequencing of the diversity panel has been described
+previously (32). The diversity panel contained 134 sequenced, clonally
+propagated individuals from the Midwest genetic subpopulation, and 229
+from the Gulf genetic subpopulation. To allow for the possibility that
+different subpopulations had different strengths of connection between
+our phenotypes and genotypes (38), we conducted three sets of genetic
+analyses: on Gulf and Midwest genotypes separately, and on both
+subpopulations together (`Both'). Analyses to determine narrow-sense
+heritability (h2) for green-up and flowering were done using linear
+mixed models and followed (32). Details on these models can be found in
+(SI Appendix, Section S3,S4).
 
-### Mapping major patterns of genotype-by-environment effects on green-up and flowering
+\subsubsection{Mapping major patterns of genotype-by-environment effects
+on green-up and
+flowering}\label{mapping-major-patterns-of-genotype-by-environment-effects-on-green-up-and-flowering-1}
 
-To evaluate the prevalence and kinds of covariance patterns of SNP effects across our common gardens, we used multivariate adaptive shrinkage (mash) on SNP effect estimates from the diversity panel (35).
-Mash is a statistical method that allows estimation and comparison of many effects jointly across many different conditions; it improves on previous methods by allowing multiple, arbitrary correlations in effect sizes among conditions (SI Appendix, Section S5).
-To obtain SNP effect estimates, we first conducted univariate genome-wide association at each common garden for green-up and flowering date.
-We then analyzed SNP effects for the top 19K relatively unlinked (r2 \< 0.2) SNPs per condition using mash, as in (32).
-Details on these models can be found in (SI Appendix, Section S6).
+To evaluate the prevalence and kinds of covariance patterns of SNP
+effects across our common gardens, we used multivariate adaptive
+shrinkage (mash) on SNP effect estimates from the diversity panel (35).
+Mash is a statistical method that allows estimation and comparison of
+many effects jointly across many different conditions; it improves on
+previous methods by allowing multiple, arbitrary correlations in effect
+sizes among conditions (SI Appendix, Section S5). To obtain SNP effect
+estimates, we first conducted univariate genome-wide association at each
+common garden for green-up and flowering date. We then analyzed SNP
+effects for the top 19K relatively unlinked (r2 \textless{} 0.2) SNPs
+per condition using mash, as in (32). Details on these models can be
+found in (SI Appendix, Section S6).
 
-We generated hypothesis-based covariance matrices derived from correlations in environmental cues in the green-up or flowering date windows for the three subpopulations (SI Appendix, Table S1, Section S1).
-These covariance matrices represent correlations between identical genotypes drawn from a specific population at pairs of common gardens; covariances near one mean that the population has a strong, positive linear relationship in individual responses at that pair of gardens, while covariances near zero mean that there is no relationship within the population for individual responses at that pair of gardens.
-Mash SNP effects will undergo strong shrinkage towards one another in the first case, and little shrinkage in the second case.
-Mash also generates data-driven covariance matrices corresponding to major patterns of SNP effects present in the data.
-We generated six data-driven matrices per mash run, five (denoted DD_PCA_1 through DD_PCA_5) produced by singular value decomposition (SVD) of an overall matrix, denoted 'DD_tPCA'.
-We used SVD to present vectors of garden-specific effects for each numbered DD_PCA matrix, because the first eigenvector of a SVD explains 100% of the variation in these matrices, and each value in this eigenvector represents one garden-specific effect.
+We generated hypothesis-based covariance matrices derived from
+correlations in environmental cues in the green-up or flowering date
+windows for the three subpopulations (SI Appendix, Table S1, Section
+S1). These covariance matrices represent correlations between identical
+genotypes drawn from a specific population at pairs of common gardens;
+covariances near one mean that the population has a strong, positive
+linear relationship in individual responses at that pair of gardens,
+while covariances near zero mean that there is no relationship within
+the population for individual responses at that pair of gardens. Mash
+SNP effects will undergo strong shrinkage towards one another in the
+first case, and little shrinkage in the second case. Mash also generates
+data-driven covariance matrices corresponding to major patterns of SNP
+effects present in the data. We generated six data-driven matrices per
+mash run, five (denoted DD\_PCA\_1 through DD\_PCA\_5) produced by
+singular value decomposition (SVD) of an overall matrix, denoted
+`DD\_tPCA'. We used SVD to present vectors of garden-specific effects
+for each numbered DD\_PCA matrix, because the first eigenvector of a SVD
+explains 100\% of the variation in these matrices, and each value in
+this eigenvector represents one garden-specific effect.
 
-Last, we characterized the overall patterns of antagonistic pleiotropy in the set of SNPs where there was pairwise significance of effects at pairs of gardens.
-To do this, we used the 'get_GxE' function of the switchgrassGWAS R package.
-First, this determines the set of SNPs with evidence of significant effects in both conditions for all pairs of conditions using local false sign rates (lfsr) as the significance criteria.
-Second, to determine antagonistic pleiotropy, this function determines if effects significant in both conditions are of opposite sign.
+Last, we characterized the overall patterns of antagonistic pleiotropy
+in the set of SNPs where there was pairwise significance of effects at
+pairs of gardens. To do this, we used the `get\_GxE' function of the
+switchgrassGWAS R package. First, this determines the set of SNPs with
+evidence of significant effects in both conditions for all pairs of
+conditions using local false sign rates (lfsr) as the significance
+criteria. Second, to determine antagonistic pleiotropy, this function
+determines if effects significant in both conditions are of opposite
+sign.
 
-Using the lfsr rather than the local false discovery rate (lfdr) is a critical change in our ability to detect antagonistic pleiotropy.
-The lfdr, like other measures of FDR, focuses on if we have enough evidence to reject the null hypothesis that an effect j is 0 -- that there is a significant effect.
-Previous studies of antagonistic pleiotropy (e.g.
-(37)) have used the lfdr or equivalent statistical tests to detect antagonistic pleiotropy.
-These tests were conservative, in that they required two non-zero effects of different signs, while tests for differential sensitivity required only one non-zero effect.
-This previous work recognized that this testing bias could lead to undercounting occurrences of antagonistic pleiotropy (26, 27), and sought to reduce it by permutation (28).
-However, using the lfsr to test for antagonistic pleiotropy does not undercount occurrences of antagonistic pleiotropy, as this statistic answers a fundamentally different question.
-For each effect j, the lfsr_j is defined as the probability that we make an error in the sign of effect j if we were forced to declare the effect positive or negative (Stephens paper).
-Thus, rather than asking "Are these two effects different?" - as we reasonably expect two effects to be, even if this difference cannot be measured - the local false sign rate answers a more meaningful question: Can we be confident in the sign of this effect?
+Using the lfsr rather than the local false discovery rate (lfdr) is a
+critical change in our ability to detect antagonistic pleiotropy. The
+lfdr, like other measures of FDR, focuses on if we have enough evidence
+to reject the null hypothesis that an effect j is 0 -- that there is a
+significant effect. Previous studies of antagonistic pleiotropy
+(e.g.~(37)) have used the lfdr or equivalent statistical tests to detect
+antagonistic pleiotropy. These tests were conservative, in that they
+required two non-zero effects of different signs, while tests for
+differential sensitivity required only one non-zero effect. This
+previous work recognized that this testing bias could lead to
+undercounting occurrences of antagonistic pleiotropy (26, 27), and
+sought to reduce it by permutation (28). However, using the lfsr to test
+for antagonistic pleiotropy does not undercount occurrences of
+antagonistic pleiotropy, as this statistic answers a fundamentally
+different question. For each effect j, the lfsr\_j is defined as the
+probability that we make an error in the sign of effect j if we were
+forced to declare the effect positive or negative (Stephens paper).
+Thus, rather than asking ``Are these two effects different?'' - as we
+reasonably expect two effects to be, even if this difference cannot be
+measured - the local false sign rate answers a more meaningful question:
+Can we be confident in the sign of this effect?
 
-In addition, the get_GxE function also sets an arbitrary threshold to count an effect as changing in magnitude between environments, commonly known as differential sensitivity.
-For differential sensitivity, this function determines if effects significant in both conditions are of the same sign and of a magnitude (not tested for significance) that differs by a factor of 0.4 or more.
-The remaining effects that are significant in both conditions have the same effect sign and similar effect magnitudes and we denote these effects as having no GxE.
-The distinction between effects with different magnitudes is arbitrary but useful to fully characterize how effects vary across environments.
-and our use of the lfsr to determine significance and our specification that SNP effects must be significant in both conditions to be included means that our tests for antagonistic pleiotropy carry an equal statistical burden to those measuring differential sensitivity and effects without GxE.
+In addition, the get\_GxE function also sets an arbitrary threshold to
+count an effect as changing in magnitude between environments, commonly
+known as differential sensitivity. For differential sensitivity, this
+function determines if effects significant in both conditions are of the
+same sign and of a magnitude (not tested for significance) that differs
+by a factor of 0.4 or more. The remaining effects that are significant
+in both conditions have the same effect sign and similar effect
+magnitudes and we denote these effects as having no GxE. The distinction
+between effects with different magnitudes is arbitrary but useful to
+fully characterize how effects vary across environments. and our use of
+the lfsr to determine significance and our specification that SNP
+effects must be significant in both conditions to be included means that
+our tests for antagonistic pleiotropy carry an equal statistical burden
+to those measuring differential sensitivity and effects without GxE.
 
-### Confirmation of genotype-by-environment effects using an independent mapping population
+\subsubsection{Confirmation of genotype-by-environment effects using an
+independent mapping
+population}\label{confirmation-of-genotype-by-environment-effects-using-an-independent-mapping-population-1}
 
-To confirm candidate genomic regions and patterns of allelic effects found in the diversity panel, we analyzed flowering in an outbred pseudo-F2 cross between four individuals, two Midwest and two Gulf individuals.
-The formation of this mapping population has been described previously (34); additional details on QTL mapping can be found in SI Appendix, Section S6.
-To be directly comparable to the diversity panel data, only 2019 phenology data from the pseudo-F2 cross from the same eight common garden sites were used.
-To compare our QTL enrichments of significant mash associations to the null expectation, we used permutation to choose 1000 sets of 23 genomic regions of the same size randomly distributed throughout the genome, then calculated enrichments of the mash 1% tail in these random intervals.
+To confirm candidate genomic regions and patterns of allelic effects
+found in the diversity panel, we analyzed flowering in an outbred
+pseudo-F2 cross between four individuals, two Midwest and two Gulf
+individuals. The formation of this mapping population has been described
+previously (34); additional details on QTL mapping can be found in SI
+Appendix, Section S6. To be directly comparable to the diversity panel
+data, only 2019 phenology data from the pseudo-F2 cross from the same
+eight common garden sites were used. To compare our QTL enrichments of
+significant mash associations to the null expectation, we used
+permutation to choose 1000 sets of 23 genomic regions of the same size
+randomly distributed throughout the genome, then calculated enrichments
+of the mash 1\% tail in these random intervals.
 
-## Author Affiliations {.unnumbered}
+\subsection*{Author Affiliations}\label{author-affiliations}
+\addcontentsline{toc}{subsection}{Author Affiliations}
 
-Include department, institution, and complete address, with the ZIP/postal code, for each author.
-Use lower case letters to match authors with institutions, as shown in the example.
-PNAS strongly encourages authors to supply an [ORCID identifier](https://orcid.org/) for each author.
-Individual authors must link their ORCID account to their PNAS account at [www.pnascentral.org](http://www.pnascentral.org/).
-For proper authentication, authors must provide their ORCID at submission and are not permitted to add ORCIDs on proofs.
+Include department, institution, and complete address, with the
+ZIP/postal code, for each author. Use lower case letters to match
+authors with institutions, as shown in the example. PNAS strongly
+encourages authors to supply an \href{https://orcid.org/}{ORCID
+identifier} for each author. Individual authors must link their ORCID
+account to their PNAS account at
+\href{http://www.pnascentral.org/}{www.pnascentral.org}. For proper
+authentication, authors must provide their ORCID at submission and are
+not permitted to add ORCIDs on proofs.
 
-## Manuscript Length {.unnumbered}
+\subsection*{Manuscript Length}\label{manuscript-length}
+\addcontentsline{toc}{subsection}{Manuscript Length}
 
-A standard 6-page article is approximately 4,000 words, 50 references, and 4 medium-size graphical elements (i.e., figures and tables).
-The preferred length of articles remains at 6 pages, but PNAS will allow articles up to a maximum of 12 pages.
+A standard 6-page article is approximately 4,000 words, 50 references,
+and 4 medium-size graphical elements (i.e., figures and tables). The
+preferred length of articles remains at 6 pages, but PNAS will allow
+articles up to a maximum of 12 pages.
 
-## References {.unnumbered}
+\subsection*{References}\label{references}
+\addcontentsline{toc}{subsection}{References}
 
-References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g.
-\cite{belkin2002using} and \cite{berard1994embedding,coifman2005geometric,phdthesis,masterthesis}.
-All references cited in the main text should be included in the main manuscript file.
+References should be cited in numerical order as they appear in text;
+this will be done automatically via bibtex, e.g.~\cite{belkin2002using}
+and
+\cite{berard1994embedding,coifman2005geometric,phdthesis,masterthesis}.
+All references cited in the main text should be included in the main
+manuscript file.
 
-## Data Archival {.unnumbered}
+\subsection*{Data Archival}\label{data-archival}
+\addcontentsline{toc}{subsection}{Data Archival}
 
 PNAS must be able to archive the data essential to a published article.
-Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the [Information for Authors](https://www.pnas.org/author-center/editorial-and-journal-policies#materials-and-data-availability), is acceptable.
+Where such archiving is not possible, deposition of data in public
+databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata,
+and others outlined in the
+\href{https://www.pnas.org/author-center/editorial-and-journal-policies\#materials-and-data-availability}{Information
+for Authors}, is acceptable.
 
-## Figures and Tables {.unnumbered}
+\subsection*{Figures and Tables}\label{figures-and-tables}
+\addcontentsline{toc}{subsection}{Figures and Tables}
 
-Figures can be included using Quarto's syntax, code chunks, or with LaTeX.
+Figures can be included using Quarto's syntax, code chunks, or with
+LaTeX.
 
-For simple figures, Quarto syntax is typically easiest. Full details are available in [the Quarto documentation](https://quarto.org/docs/authoring/figures.html).
+For simple figures, Quarto syntax is typically easiest. Full details are
+available in \href{https://quarto.org/docs/authoring/figures.html}{the
+Quarto documentation}.
 
-![Placeholder image of a frog with a long example legend to show justification setting.](figs/frog.pdf){#fig-frog}
+\begin{figure}
 
-Tables can similarly be included using Quarto's syntax, code chunks, or with LaTeX.
-I tend to use code chunks, but LaTeX works very well for classic tables, like regression tables or summary statistics.
+\centering{
 
-```{=tex}
+\includegraphics{figs/frog.pdf}
+
+}
+
+\caption{\label{fig-frog}Placeholder image of a frog with a long example
+legend to show justification setting.}
+
+\end{figure}%
+
+Tables can similarly be included using Quarto's syntax, code chunks, or
+with LaTeX. I tend to use code chunks, but LaTeX works very well for
+classic tables, like regression tables or summary statistics.
+
 \begin{table}[t!]
@@
 \end{table}
-```
-### Digital Figures {.unnumbered}
 
-EPS, high-resolution PDF, and PowerPoint are preferred formats for figures that will be used in the main manuscript.
-Authors may submit PRC or U3D files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.
-Color images must be in RGB (red, green, blue) mode.
-Include the font files for any text.
+\subsubsection*{Digital Figures}\label{digital-figures}
+\addcontentsline{toc}{subsubsection}{Digital Figures}
 
-Images must be provided at final size, preferably 1 column width (8.7cm).
-Figures wider than 1 column should be sized to 11.4cm or 17.8cm wide.
-Numbers, letters, and symbols should be no smaller than 6 points (2mm) and no larger than 12 points (6mm) after reduction and must be consistent.
+EPS, high-resolution PDF, and PowerPoint are preferred formats for
+figures that will be used in the main manuscript. Authors may submit PRC
+or U3D files for 3D images; these must be accompanied by 2D
+representations in TIFF, EPS, or high-resolution PDF format. Color
+images must be in RGB (red, green, blue) mode. Include the font files
+for any text.
 
-Figures and tables should be labelled and referenced in the standard way using the `\label{}` and `\ref{}` commands.
+Images must be provided at final size, preferably 1 column width
+(8.7cm). Figures wider than 1 column should be sized to 11.4cm or 17.8cm
+wide. Numbers, letters, and symbols should be no smaller than 6 points
+(2mm) and no larger than 12 points (6mm) after reduction and must be
+consistent.
 
-Figure @fig-frog shows an example of how to insert a column-wide figure.
-To insert a figure wider than one column, please use the `figure*` environment.
-Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide.
-Use `SCfigure*` environment for a wide figure with side legends, such as @fig-side.
+Figures and tables should be labelled and referenced in the standard way
+using the \texttt{\textbackslash{}label\{\}} and
+\texttt{\textbackslash{}ref\{\}} commands.
 
-In Quarto, we can do these by setting the `fig-env` command to `figure*` or `SCfigure*`
+Figure Figure~\ref{fig-frog} shows an example of how to insert a
+column-wide figure. To insert a figure wider than one column, please use
+the \texttt{figure*} environment. Figures wider than one column should
+be sized to 11.4 cm or 17.8 cm wide. Use \texttt{SCfigure*} environment
+for a wide figure with side legends, such as Figure~\ref{fig-side}.
 
-```{r}
-#| label: fig-side
-#| fig-cap: "This legend would be placed at the side of the figure, rather than below it."
-#| fig-env: "SCfigure*"
-#| echo: false
-# tell it the options as comments with a | and a space, as above.
-# set echo: false to avoid printing this text
-knitr::include_graphics('figs/frog.pdf')
-```
+In Quarto, we can do these by setting the \texttt{fig-env} command to
+\texttt{figure*} or \texttt{SCfigure*}
 
-## Tables {.unnumbered}
+\begin{SCfigure*}
 
-Tables should be included in the main manuscript file and should not be uploaded separately.
+\centering{
 
-## Single column equations {.unnumbered}
+\includegraphics{figs/frog.pdf}
 
-Authors may use 1- or 2-column equations in their article, according to their preference.
+}
 
-To allow an equation to span both columns, use the `\begin{figure*}...\end{figure*}` environment mentioned above for figures.
+\caption{\label{fig-side}This legend would be placed at the side of the
+figure, rather than below it.}
 
-Note that the use of the `widetext` environment for equations is not recommended, and should not be used.
+\end{SCfigure*}%
 
-```{=tex}
+\subsection*{Tables}\label{tables}
+\addcontentsline{toc}{subsection}{Tables}
+
+Tables should be included in the main manuscript file and should not be
+uploaded separately.
+
+\subsection*{Single column equations}\label{single-column-equations}
+\addcontentsline{toc}{subsection}{Single column equations}
+
+Authors may use 1- or 2-column equations in their article, according to
+their preference.
+
+To allow an equation to span both columns, use the
+\texttt{\textbackslash{}begin\{figure*\}...\textbackslash{}end\{figure*\}}
+environment mentioned above for figures.
+
+Note that the use of the \texttt{widetext} environment for equations is
+not recommended, and should not be used.
+
 \begin{figure*}[bt!]
@@
 \end{figure*}
-```
-# Material and Methods {.unnumbered}
 
-## Supporting Information Appendix (SI) {.unnumbered}
+\section*{Material and Methods}\label{material-and-methods}
+\addcontentsline{toc}{section}{Material and Methods}
 
-Authors should submit SI as a single separate SI Appendix PDF file, combining all text, figures, tables, movie legends, and SI references.
-SI will be published as provided by the authors; it will not be edited or composed.
-Additional details can be found in the [PNAS Author Center](https://www.pnas.org/authors/submitting-your-manuscript#manuscript-formatting-guidelines).
-The PNAS Overleaf SI template can be found [here](https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd).
-Refer to the SI Appendix in the manuscript at an appropriate point in the text.
-Number supporting figures and tables starting with S1, S2, etc.
+\subsection*{Supporting Information Appendix
+(SI)}\label{supporting-information-appendix-si}
+\addcontentsline{toc}{subsection}{Supporting Information Appendix (SI)}
 
-Authors who place detailed materials and methods in an SI Appendix must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the SI methods.
-If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.
+Authors should submit SI as a single separate SI Appendix PDF file,
+combining all text, figures, tables, movie legends, and SI references.
+SI will be published as provided by the authors; it will not be edited
+or composed. Additional details can be found in the
+\href{https://www.pnas.org/authors/submitting-your-manuscript\#manuscript-formatting-guidelines}{PNAS
+Author Center}. The PNAS Overleaf SI template can be found
+\href{https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd}{here}.
+Refer to the SI Appendix in the manuscript at an appropriate point in
+the text. Number supporting figures and tables starting with S1, S2,
+etc.
 
-### SI Datasets {.unnumbered}
+Authors who place detailed materials and methods in an SI Appendix must
+provide sufficient detail in the main text methods to enable a reader to
+follow the logic of the procedures and results and also must reference
+the SI methods. If a paper is fundamentally a study of a new method or
+technique, then the methods must be described completely in the main
+text.
 
-Supply .xlsx, .csv, .txt, .rtf, or .pdf files. This file type will be published in raw format and will not be edited or composed.
+\subsubsection*{SI Datasets}\label{si-datasets}
+\addcontentsline{toc}{subsubsection}{SI Datasets}
 
-### SI Movies {.unnumbered}
+Supply .xlsx, .csv, .txt, .rtf, or .pdf files. This file type will be
+published in raw format and will not be edited or composed.
 
-Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files.
-Movie legends should be included in the SI Appendix file.
-All movies should be submitted at the desired reproduction size and length.
-Movies should be no more than 10MB in size.
+\subsubsection*{SI Movies}\label{si-movies}
+\addcontentsline{toc}{subsubsection}{SI Movies}
 
-# References
+Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media
+(wmv), animated GIF (gif), or MPEG files. Movie legends should be
+included in the SI Appendix file. All movies should be submitted at the
+desired reproduction size and length. Movies should be no more than 10MB
+in size.
 
-\bibsplit[2]<!-- Use \bibsplit to split the references from the body of the text.
-Value "[2]" represents the number of reference in the left column (Note: Please avoid single column figures & tables on this page.) -->
+\section{References}\label{references-1}
 
-::: {#refs}
-:::
+\bibsplit[2]
+
+\phantomsection\label{refs}
+\begin{CSLReferences}{0}{1}
+\end{CSLReferences}
+
+\acknow{We thank the Brackenridge Field laboratory, the Ladybird Johnson
+Wildflower Center, and the Juenger laboratory for support with plant
+care and propagation. This material is based upon work supported in part
+by the Great Lakes Bioenergy Research Center, U.S. Department of Energy,
+Office of Science, Office of Biological and Environmental Research under
+Award Numbers DE-SC0018409 and DE-FC02-07ER64494, the US Department of
+Energy Awards DESC0014156 to T.E.J., DE-SC0017883 to D.B.L, National
+Science Foundation PGRP Awards IOS0922457 and IOS1444533 to T.E.J, and
+the Long-term Ecological Research Program (DEB 1832042) at the Kellogg
+Biological Station.}
+
+\showacknow{} % Display the acknowledgments section
+
+\end{document}

### 2023-12-02 manuscript/revision2/response_to_reviews.qmd
manuscript: SI Appendix and response to reviews for revision

---
title: "Response_to_Reviews"
format: pdf
editor: visual
---

**Reviewer comment**

Mash: A lot of the results interpretation is based on interpreting the SNP loadings on the specified covariance matrices. This is a secondary use of mash (the primary being the refinement of effect sizes), and while they do interpret these somewhat in the mash paper, these loadings need to be interpreted with caution. If these covariance matrices are similar, mash will somewhat arbitrarily assign weight to any of the matrices because all lead to the same fit to the effect-size data. The mash algorithm doesn\'t give the full posterior distribution on the loadings so you can\'t check for posterior correlations there. It seems likely to me that the hypothesis-driven and data-driven covariance matrices are somewhat correlated here, and the correlations may differ between green-up and flowering because of the better correlation between the environmental metrics and flowering than green-up. In the mash paper, they used cross-validation with the likelihood in the testing set as the evaluation metric to compare models. It might be safer to try dropping specific covariance matrices and comparing the model performance in a held-out testing set to evaluate the importance of the hypothesis-derived covariance matrices.

\

**Response**

We thank the reviewer for this comment and agree that the selection of covariance matrices to be included in a mash model is a nontrivial question. To clarify our interpretation of the posterior matrix weights provided by mash, we performed an extensive analysis of the performance of mash models when different covariance matrices are included. Specifically, we implemented a model selection approach that uses a greedy algorithm to evaluate the log likelihood of the mash model as additional covariance matrices were included (see Methods lines XXX-YYY and pseudo code below). This is a similar, but more computationally efficient design, than the leave-one-out approach recommended by the reviewer as the runtime of mash increases with the number of covariance matrices included. Additionally, while cross-validation is a powerful approach to model evaluation our limited sample size was prohibitive to the necessary partitioning of individuals included in our analysis.

\

In practice this greedy algorithm approach offers a fast way to identify the point where redundancy in the addition of a new covariance matrix results in no change to the likelihood of the mash model. Importantly, this removes the potential for arbitrary assignment of weights to the covariance matrices in the model and allows for the selection of matrices that most accurately capture the underlying biology captured by the site-specific effect sizes. We applied this algorithm to only the hypothesis matrices for each phenotype (Figures 1 and 2) using the previously used sets of significantly associated and randomly selected variants in the Gulf subpopulation individuals, Midwest subpopulation individuals, and the two subpopulations combined. The results indicate that only a small number of the original hypothesis matrices (between three and five) are necessary to reach the maximum likelihood model when using the same likelihood ratio test implemented in Urbut et al., indicating that correlation among hypothesis covariance matrices is high. This redundancy was confirmed through our application of the greedy algorithm to a combined set of canonical, data driven, and hypothesis covariance matrices (Figures 3 and 4)

Our primary conclusion from this analysis is in line with what the reviewer posited, that there is extensive redundancy among canonical, data driven, and hypothesis covariance matrices that is leading to biologically informative patterns of covariance between site effect sizes to be arbitrarily distributed among them. We additionally found evidence that while the correlations between data driven and environmentally informed (hypothesis) covariance matrices, there is extensive similarity in the matrices that are included in the maximum likelihood models across both phenotypes and in the analysis of the Gulf subpopulation, Midwest subpopulations, and the combined cohort.

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

Greedy algorithm for mash model selection.

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

\

matrices = \[all covariance matrices of interest\]

most_likely_matrices = \[\] #a list to store the most likely additional matrix at each iteration

maximum_likelihoods = \[\]#a list to store the maximum likelihood at each iteration

\

likelihood = -inf

most_likely_matrix = \'\'

for matrix in matrices:

	matrix_likelihood = mash(effects, std.errs, matrix)\[\'likelihood\'\] #obtain model likelihood

	if matrix_likelihood \> likelihood:

		most_likely_matrix = matrix

		likelihood = matrix_likelihood

\

matrices.remove(most_likely_matrix)

most_likely_matrices.append(most_likely_matrix)

maximum_likelihoods.append(likelihood)

\

lrt_pvalue = 0

iteration = 2

while lrt_pvalue \< 0.05:

	likelihood = -inf

most_likely_matrix = \'\'

	for matrix in matrices:

		model_matrices = most_likely_matrices + matrix

		model_likelihood = mash(effects, std.errs, model_matrices)\[\'likelihood\'\]

	if matrix_likelihood \> likelihood:

		most_likely_matrix = matrix

		likelihood = matrix_likelihood

\

matrices.remove(most_likely_matrix)

most_likely_matrices.append(most_likely_matrix)

maximum_likelihoods.append(likelihood)

\

#perform a likelihood ratio test of the most likely model from the previous iteration to the model in this step with one degree of freedom

lrt_pvalue = likelihood_ratio_test(likelihood, maximum_likelihoods\[iteration-1\], df = 1)

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

\

**Methods**

In order to better understand the behavior of the mash shrinkage algorithm in the presence of correlated covariance matrices, we implemented a \"greedy\" mash algorithm. Specifically, given n covariance matrices for each phenotype and corresponding effect sizes and standard errors across the k contexts of interest, we fit a set of mash models with each of the n matrices separately. We then identify the model with the maximum log likelihood estimates and select the corresponding matrix, leaving n-1 covariance matrices. We then fit a new set of mash models where the selected covariance matrix is paired with each of the remaining n-1 covariance matrices. The most likely pair of covariance matrices are then selected and the process is repeated for all possible n-2 matrix triplets. This process is repeated until one of two stop conditions are met: (i) all covariance matrices have been added to the mash model, resulting in a final model with all of the original n covariance matrices are included or (ii) a likelihood ratio test with one degree of freedom between the likelihood of the current model and the model from the previous iteration has a p-value \> 0.05. The result is a set of log likelihood estimates for each model and the stepwise \'path\' of covariance matrices that best explain the observed effect sizes and standard errors observed for the phenotype of interest.

\

We then applied the greedy algorithm to two sets of covariance matrices. First, we applied it to only the hypothesized covariance matrices for each phenotype as described in Supplementary Table 1. Briefly, the hypothesis covariance matrices included in our analysis of greenup time included: average temperature in the 5 days, 10 days, and 18 days preceding greenup and the number of cumulative growing degree days (over 12 degrees Celsius) in the 5 days, 10 days, and 18 days preceding greenup. Each of these six covariance matrices were calculated using individuals who were present at both sites. Additionally, each matrix was calculated using all individuals, only individuals from the Gulf subpopulation, and only individuals from the Midwest subpopulation.

\

For the flowering time phenotype, 12 hypothesized covariance matrices were included in the greedy algorithm. For each phenotype, we used the mash greedy algorithm to identify the most likely covariance matrix at each step using both a set of random independent variants and a set of independent variants that showed the strongest associations with the phenotype across all contexts. The resulting maximum likelihood paths for the selection of covariance matrices as determined by the greedy algorithm for each phenotype, subpopulation, and set of randomly selected or significantly associated variants are shown in Figures 1 and 2.

\

Our second analyses expanded the set of covariance matrices to all of the matrices included in our analyses of the empirical data. These include: the phenotype specific covariance matrices described above, the data-driven covariance matrices, and the suite of canonical covariance matrices representing context-specific effects, equal effects across contexts, and simple scenarios of heterogeneous effects across contexts. The resulting maximum likelihood paths for the selection of covariance matrices as determined by the greedy algorithm for each phenotype, subpopulation, and set of randomly selected or significantly associated variants are shown in Figures 3 and 4.

\

**Results**

We first applied the greedy algorithm to only the set of hypothesis matrices for each of the greenup and flowering phenotypes, respectively. Generally, we observed that only a small number of matrices significantly increased the likelihood of the model as the algorithm progressed. In our analysis of the set of variants that were significantly associated with greenup time in at least one context, 18 hypothesis covariance matrices were included in the initialization of the greedy algorithm. The stop condition, a likelihood ratio test p-value \> 0.05 between the most likely model at step n and most likely model at step n+1, was met after only four iterations in both the combined cohort (Figure 1A) as well as the analysis of the Gulf subpopulation alone (Figure 1B). In our analysis of the midwest subpopulation the stop condition was met after only three iterations, see Figure 1C. The relatively small number of necessary iterations, and correspondingly small number of covariance matrices, indicate that the hypothesis covariance matrices were highly correlated with one another and that inclusion of all the hypothesis covariance matrices introduced redundancy in the model. In the set of randomly selected independent variants, the stop condition was met in the combined cohort after five steps (Figure 1D). The stop condition was met after four and three iterations in the Gulf and Midwest subpopulations, respectively. The overlap between matrices included in the maximum likelihood model prior to reaching the stop condition in the analysis of significantly associated and randomly selected variants was high in the combined cohort; three of four matrices in the analysis of significantly associated variants were also included in the model for the randomly selected variants. Similarly, three of four matrices in the Gulf subpopulation and two of three matrices in the Midwest subpopulation that were included in the model for significantly associated variants were also included in the model for the randomly selected variants.

\

We then applied the greedy algorithm to the 12 hypothesis covariance matrices for the flowering phenotype. Similar to our analysis of the hypothesis covariance matrices for the greenup phenotype, we observed a similar pattern of correlation in hypothesis covariance matrices for the flowering phenotype. The stop condition for the greedy algorithm was met after five steps when analyzing the combined cohort, the Gulf subpopulation, and the Midwest subpopulation in analysis of variants that were significantly associated in at least one context (Figure 2A-C). When analyzing the set of randomly selected variants the stop condition was met after four steps in the combined cohort, five steps in the Gulf subpopulation, and six steps in the Midwest subpopulation. We also observed that many of the matrices included in the model for the significantly associated variants were also included in the corresponding model for randomly selected variants (Figure 2D-F).

\

Analysis of the hypothesis covariance matrices alone helped to establish the utility of the greedy algorithm implementation in identifying the maximum likelihood mash model for the observed summary statistics from our analysis of eight sites. However, due to substantial evidence that both canonical and data driven matrices explained a nontrivial proportion of the observed covariance between context specific summary statistics we applied the greedy algorithm with the all three classes of covariance matrices included in the initial set. In our analysis of the greenup phenotype, the stop condition was met after 16 iterations, resulting in a maximum likelihood model that included all 13 canonical covariance matrices, two data driven matrices (total PCA and PC 4), and the hypothesis matrix representing the covariance in average temperature ten days prior to greenup in the combined cohort (Figure 3A). When the Gulf subpopulation was analyzed alone the stop condition was met after five iterations. The five corresponding matrices were composed of four canonical matrices and the hypothesis matrix calculated as the covariance in average temperature five days prior to greenup in only individuals from the Gulf subpopulation (Figure 3B). Finally, the greedy algorithm met the stop condition after 13 iterations when applied to the Midwest subpopulation. Ten of the canonical covariance matrices, the total PCA data driven covariance matrix, and the two covariance matrices calculated using the five day temperature average in the five days prior to greenup and the number of growing degree days in the 18 days prior to greenup in the Midwest subpopulation individuals were all include in the mode (Figure 3C). Most notably, the canonical covariance matrix representing independent effects was the first matrix selected in all three applications of the greedy algorithm, indicating a large number of independent effects were present in our analysis of the greenup phenotype. We then applied the greedy algorithm to a set of randomly selected variants and observed that many of the same matrices included in the corresponding models from application to the significantly associated variants were observed in the most likely models once the stop condition was met (Figure 3D-F).

\

We then performed the same analysis of the entire suite of canonical, data driven, and hypothesis covariance matrices for the flowering phenotype. The stop condition was met after five iterations in the combined cohort and included four canonical covariance matrices representing independent effects, equal effects, and two simple models of effect heterogeneity among sites (Figure 4A). In application to the Gulf subpopulation, the stop condition was met after 11 iterations and included 10 canonical covariance matrices and the data driven covariance matrix derived from the third principal component (Figure 4B). The stop condition was met after 11 iterations in the analysis of Midwest subpopulation individuals and included nine canonical covariance matrices, the data driven matrix derived from the first principal component, and the hypothesis matrix calculated using the cumulative rainfall in the day prior to flowering tin the Midwest subpopulation (Figure 4C). Once again, we observed that the first matrix selected in each analysis of flowering time was the canonical matrix representing independent effects among conditions. Overlap between the models selected by the greedy algorithm when analyzing a set of randomly selected variants was high across all three analyses (Figure 4C-D).

\

**Figures**

\
\

![](https://lh7-us.googleusercontent.com/ae_Fi1djzcYo_ibeQWWjr2AXuwIX781VIa7wSug5K3H9T8F5ke1RYU8qKPRT-YGO_KvDJ5oK3pzzQX3G12zbJOo9i6_yLEV8NkpRHF4zT3dqrhXKfLISSV5N-iXjYIczVPSJcDEPx-W7vv7NvU1AQv8)

**Figure 1. Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the greenup phenotype using only hypothesis covariance matrices.** The matrix that resulted in the maximum likelihood model of covariance in summary statistics among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis. **(A)** The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals. The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05. **(B)** and **(C)** show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively. **(D)** The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals. The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05. \* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants. **(E)** and **(F)** show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.

\

![](https://lh7-us.googleusercontent.com/8AuFUNJTaIAgX7GVkOSnrBMMNqZr4mYdesj48zBZLSssUZ15epDhoC8FLp4gl_5kvZiR94cqb925c23NIRXakRL1dh2TVp_a1wtf-gw1w_QKx1hkaGvJ_CsQHwHube_7PFWBDq1zZgsaeZXQ65pON_8)

**Figure 2. Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the flowering phenotype using only hypothesis covariance matrices.** The matrix that resulted in the maximum likelihood model among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis. **(A)** The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals. The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05. **(B)** and **(C)** show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively. **(D)** The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals. The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05. \* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants. **(E)** and **(F)** show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.

\

![](https://lh7-us.googleusercontent.com/iO_JwhiNJbDDdV-UHqzlZOnGUemRBlDpw9qDXXHYfLUzRTgpU0w8fxQczs4NLhz5apnaD_x0CL1CBfPd0KlLtbprYsaC4Cub2-2nIyzkvMAcp3MkyzkB5x82-L3kJFfiBSJPxDg7XqZ6JIox4hG2e-A)

**Figure 3. Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the greenup phenotype using canonical, data drive, and hypothesis covariance matrices.** The matrix that resulted in the maximum likelihood model of covariance in summary statistics among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis. **(A)** The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals. The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05. Once this condition was met the greedy algorithm was halted. **(B)** and **(C)** show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively. **(D)** The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals. The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05. Once this condition was met the greedy algorithm was halted for computational considerations. \* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants. **(E)** and **(F)** show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.

\

![](https://lh7-us.googleusercontent.com/LnnT8NIjtolmYO5neLCaID6fOqbMyUBZalKFy9eiHmC4xTTSFoNoTVIxgIQpn3n5FDynzHwBdZmGaRv2ktSquMKcj5RsdqV453tRWtRzMvRk3va8YV5g-pItSMpPSQqTnW7iZ7l6xPR5Zfh-4tEorVI)

**Figure 4. Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the flowering phenotype using canonical, data drive, and hypothesis covariance matrices.** The matrix that resulted in the maximum likelihood model of covariance in summary statistics among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis. **(A)** The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals. The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05. Once this condition was met the greedy algorithm was halted. **(B)** and **(C)** show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively. **(D)** The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals. The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05. Once this condition was met the greedy algorithm was halted for computational considerations. \* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants. **(E)** and **(F)** show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.

### 2023-12-02 manuscript/revision2/response_to_reviews.tex
manuscript: SI Appendix and response to reviews for revision

--- 2023-12-02 manuscript/revision2/response_to_reviews.qmd
@@
----
-title: "Response_to_Reviews"
-format: pdf
-editor: visual
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+%
+\documentclass[
+  letterpaper,
+  DIV=11,
+  numbers=noendperiod]{scrartcl}
 
-**Reviewer comment**
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-Mash: A lot of the results interpretation is based on interpreting the SNP loadings on the specified covariance matrices.
-This is a secondary use of mash (the primary being the refinement of effect sizes), and while they do interpret these somewhat in the mash paper, these loadings need to be interpreted with caution.
-If these covariance matrices are similar, mash will somewhat arbitrarily assign weight to any of the matrices because all lead to the same fit to the effect-size data.
-The mash algorithm doesn\'t give the full posterior distribution on the loadings so you can\'t check for posterior correlations there.
-It seems likely to me that the hypothesis-driven and data-driven covariance matrices are somewhat correlated here, and the correlations may differ between green-up and flowering because of the better correlation between the environmental metrics and flowering than green-up.
-In the mash paper, they used cross-validation with the likelihood in the testing set as the evaluation metric to compare models.
-It might be safer to try dropping specific covariance matrices and comparing the model performance in a held-out testing set to evaluate the importance of the hypothesis-derived covariance matrices.
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 
-\
+\KOMAoption{captions}{tableheading}
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\AtBeginDocument{%
+\ifdefined\contentsname
+  \renewcommand*\contentsname{Table of contents}
+\else
+  \newcommand\contentsname{Table of contents}
+\fi
+\ifdefined\listfigurename
+  \renewcommand*\listfigurename{List of Figures}
+\else
+  \newcommand\listfigurename{List of Figures}
+\fi
+\ifdefined\listtablename
+  \renewcommand*\listtablename{List of Tables}
+\else
+  \newcommand\listtablename{List of Tables}
+\fi
+\ifdefined\figurename
+  \renewcommand*\figurename{Figure}
+\else
+  \newcommand\figurename{Figure}
+\fi
+\ifdefined\tablename
+  \renewcommand*\tablename{Table}
+\else
+  \newcommand\tablename{Table}
+\fi
+}
+\@ifpackageloaded{float}{}{\usepackage{float}}
+\floatstyle{ruled}
+\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
+\floatname{codelisting}{Listing}
+\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\makeatother
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={Response\_to\_Reviews},
+  colorlinks=true,
+  linkcolor={blue},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={Blue},
+  pdfcreator={LaTeX via pandoc}}
 
-**Response**
+\title{Response\_to\_Reviews}
+\author{}
+\date{}
 
-We thank the reviewer for this comment and agree that the selection of covariance matrices to be included in a mash model is a nontrivial question.
-To clarify our interpretation of the posterior matrix weights provided by mash, we performed an extensive analysis of the performance of mash models when different covariance matrices are included.
-Specifically, we implemented a model selection approach that uses a greedy algorithm to evaluate the log likelihood of the mash model as additional covariance matrices were included (see Methods lines XXX-YYY and pseudo code below).
-This is a similar, but more computationally efficient design, than the leave-one-out approach recommended by the reviewer as the runtime of mash increases with the number of covariance matrices included.
-Additionally, while cross-validation is a powerful approach to model evaluation our limited sample size was prohibitive to the necessary partitioning of individuals included in our analysis.
+\begin{document}
+\maketitle
+\textbf{Reviewer comment}
 
-\
+Mash: A lot of the results interpretation is based on interpreting the
+SNP loadings on the specified covariance matrices. This is a secondary
+use of mash (the primary being the refinement of effect sizes), and
+while they do interpret these somewhat in the mash paper, these loadings
+need to be interpreted with caution. If these covariance matrices are
+similar, mash will somewhat arbitrarily assign weight to any of the
+matrices because all lead to the same fit to the effect-size data. The
+mash algorithm doesn\textquotesingle t give the full posterior
+distribution on the loadings so you can\textquotesingle t check for
+posterior correlations there. It seems likely to me that the
+hypothesis-driven and data-driven covariance matrices are somewhat
+correlated here, and the correlations may differ between green-up and
+flowering because of the better correlation between the environmental
+metrics and flowering than green-up. In the mash paper, they used
+cross-validation with the likelihood in the testing set as the
+evaluation metric to compare models. It might be safer to try dropping
+specific covariance matrices and comparing the model performance in a
+held-out testing set to evaluate the importance of the
+hypothesis-derived covariance matrices.
 
-In practice this greedy algorithm approach offers a fast way to identify the point where redundancy in the addition of a new covariance matrix results in no change to the likelihood of the mash model.
-Importantly, this removes the potential for arbitrary assignment of weights to the covariance matrices in the model and allows for the selection of matrices that most accurately capture the underlying biology captured by the site-specific effect sizes.
-We applied this algorithm to only the hypothesis matrices for each phenotype (Figures 1 and 2) using the previously used sets of significantly associated and randomly selected variants in the Gulf subpopulation individuals, Midwest subpopulation individuals, and the two subpopulations combined.
-The results indicate that only a small number of the original hypothesis matrices (between three and five) are necessary to reach the maximum likelihood model when using the same likelihood ratio test implemented in Urbut et al., indicating that correlation among hypothesis covariance matrices is high.
-This redundancy was confirmed through our application of the greedy algorithm to a combined set of canonical, data driven, and hypothesis covariance matrices (Figures 3 and 4)
+\hfill\break
 
-Our primary conclusion from this analysis is in line with what the reviewer posited, that there is extensive redundancy among canonical, data driven, and hypothesis covariance matrices that is leading to biologically informative patterns of covariance between site effect sizes to be arbitrarily distributed among them.
-We additionally found evidence that while the correlations between data driven and environmentally informed (hypothesis) covariance matrices, there is extensive similarity in the matrices that are included in the maximum likelihood models across both phenotypes and in the analysis of the Gulf subpopulation, Midwest subpopulations, and the combined cohort.
+\textbf{Response}
+
+We thank the reviewer for this comment and agree that the selection of
+covariance matrices to be included in a mash model is a nontrivial
+question. To clarify our interpretation of the posterior matrix weights
+provided by mash, we performed an extensive analysis of the performance
+of mash models when different covariance matrices are included.
+Specifically, we implemented a model selection approach that uses a
+greedy algorithm to evaluate the log likelihood of the mash model as
+additional covariance matrices were included (see Methods lines XXX-YYY
+and pseudo code below). This is a similar, but more computationally
+efficient design, than the leave-one-out approach recommended by the
+reviewer as the runtime of mash increases with the number of covariance
+matrices included. Additionally, while cross-validation is a powerful
+approach to model evaluation our limited sample size was prohibitive to
+the necessary partitioning of individuals included in our analysis.
+
+\hfill\break
+
+In practice this greedy algorithm approach offers a fast way to identify
+the point where redundancy in the addition of a new covariance matrix
+results in no change to the likelihood of the mash model. Importantly,
+this removes the potential for arbitrary assignment of weights to the
+covariance matrices in the model and allows for the selection of
+matrices that most accurately capture the underlying biology captured by
+the site-specific effect sizes. We applied this algorithm to only the
+hypothesis matrices for each phenotype (Figures 1 and 2) using the
+previously used sets of significantly associated and randomly selected
+variants in the Gulf subpopulation individuals, Midwest subpopulation
+individuals, and the two subpopulations combined. The results indicate
+that only a small number of the original hypothesis matrices (between
+three and five) are necessary to reach the maximum likelihood model when
+using the same likelihood ratio test implemented in Urbut et al.,
+indicating that correlation among hypothesis covariance matrices is
+high. This redundancy was confirmed through our application of the
+greedy algorithm to a combined set of canonical, data driven, and
+hypothesis covariance matrices (Figures 3 and 4)
+
+Our primary conclusion from this analysis is in line with what the
+reviewer posited, that there is extensive redundancy among canonical,
+data driven, and hypothesis covariance matrices that is leading to
+biologically informative patterns of covariance between site effect
+sizes to be arbitrarily distributed among them. We additionally found
+evidence that while the correlations between data driven and
+environmentally informed (hypothesis) covariance matrices, there is
+extensive similarity in the matrices that are included in the maximum
+likelihood models across both phenotypes and in the analysis of the Gulf
+subpopulation, Midwest subpopulations, and the combined cohort.
+
+~
 
@@
 
-Greedy algorithm for mash model selection.
+Greedy algorithm for mash model selection.~
 
@@
 
-\
+\hfill\break
 
-matrices = \[all covariance matrices of interest\]
+matrices = {[}all covariance matrices of interest{]}
 
-most_likely_matrices = \[\] #a list to store the most likely additional matrix at each iteration
+most\_likely\_matrices = {[}{]} \#a list to store the most likely
+additional matrix at each iteration
 
-maximum_likelihoods = \[\]#a list to store the maximum likelihood at each iteration
+maximum\_likelihoods = {[}{]}\#a list to store the maximum likelihood at
+each iteration
 
-\
+\hfill\break
 
@@
 
-most_likely_matrix = \'\'
+most\_likely\_matrix = \textquotesingle\textquotesingle{}
 
@@
 
+\begin{verbatim}
 matrix_likelihood = mash(effects, std.errs, matrix)\[\'likelihood\'\] #obtain model likelihood
@@
     likelihood = matrix_likelihood
+\end{verbatim}
 
-\
+\hfill\break
 
-matrices.remove(most_likely_matrix)
+matrices.remove(most\_likely\_matrix)
 
-most_likely_matrices.append(most_likely_matrix)
+most\_likely\_matrices.append(most\_likely\_matrix)
 
-maximum_likelihoods.append(likelihood)
+maximum\_likelihoods.append(likelihood)
 
-\
+\hfill\break
 
-lrt_pvalue = 0
+lrt\_pvalue = 0
 
@@
 
-while lrt_pvalue \< 0.05:
+while lrt\_pvalue \textless{} 0.05:
 
+\begin{verbatim}
 likelihood = -inf
+\end{verbatim}
 
-most_likely_matrix = \'\'
+most\_likely\_matrix = \textquotesingle\textquotesingle{}
 
+\begin{verbatim}
 for matrix in matrices:
@@
     likelihood = matrix_likelihood
+\end{verbatim}
 
-\
+\hfill\break
 
-matrices.remove(most_likely_matrix)
+matrices.remove(most\_likely\_matrix)
 
-most_likely_matrices.append(most_likely_matrix)
+most\_likely\_matrices.append(most\_likely\_matrix)
 
-maximum_likelihoods.append(likelihood)
+maximum\_likelihoods.append(likelihood)
 
-\
+\hfill\break
 
-#perform a likelihood ratio test of the most likely model from the previous iteration to the model in this step with one degree of freedom
+\#perform a likelihood ratio test of the most likely model from the
+previous iteration to the model in this step with one degree of freedom
 
-lrt_pvalue = likelihood_ratio_test(likelihood, maximum_likelihoods\[iteration-1\], df = 1)
+lrt\_pvalue = likelihood\_ratio\_test(likelihood,
+maximum\_likelihoods{[}iteration-1{]}, df = 1)
 
@@
 
-\
+\hfill\break
 
-**Methods**
+\textbf{Methods}
 
-In order to better understand the behavior of the mash shrinkage algorithm in the presence of correlated covariance matrices, we implemented a \"greedy\" mash algorithm.
-Specifically, given n covariance matrices for each phenotype and corresponding effect sizes and standard errors across the k contexts of interest, we fit a set of mash models with each of the n matrices separately.
-We then identify the model with the maximum log likelihood estimates and select the corresponding matrix, leaving n-1 covariance matrices.
-We then fit a new set of mash models where the selected covariance matrix is paired with each of the remaining n-1 covariance matrices.
-The most likely pair of covariance matrices are then selected and the process is repeated for all possible n-2 matrix triplets.
-This process is repeated until one of two stop conditions are met: (i) all covariance matrices have been added to the mash model, resulting in a final model with all of the original n covariance matrices are included or (ii) a likelihood ratio test with one degree of freedom between the likelihood of the current model and the model from the previous iteration has a p-value \> 0.05.
-The result is a set of log likelihood estimates for each model and the stepwise \'path\' of covariance matrices that best explain the observed effect sizes and standard errors observed for the phenotype of interest.
+In order to better understand the behavior of the mash shrinkage
+algorithm in the presence of correlated covariance matrices, we
+implemented a "greedy" mash algorithm. Specifically, given n covariance
+matrices for each phenotype and corresponding effect sizes and standard
+errors across the k contexts of interest, we fit a set of mash models
+with each of the n matrices separately. We then identify the model with
+the maximum log likelihood estimates and select the corresponding
+matrix, leaving n-1 covariance matrices. We then fit a new set of mash
+models where the selected covariance matrix is paired with each of the
+remaining n-1 covariance matrices. The most likely pair of covariance
+matrices are then selected and the process is repeated for all possible
+n-2 matrix triplets. This process is repeated until one of two stop
+conditions are met: (i) all covariance matrices have been added to the
+mash model, resulting in a final model with all of the original n
+covariance matrices are included or (ii) a likelihood ratio test with
+one degree of freedom between the likelihood of the current model and
+the model from the previous iteration has a p-value \textgreater{} 0.05.
+The result is a set of log likelihood estimates for each model and the
+stepwise \textquotesingle path\textquotesingle{} of covariance matrices
+that best explain the observed effect sizes and standard errors observed
+for the phenotype of interest.~
 
-\
+\hfill\break
 
 We then applied the greedy algorithm to two sets of covariance matrices.
-First, we applied it to only the hypothesized covariance matrices for each phenotype as described in Supplementary Table 1.
-Briefly, the hypothesis covariance matrices included in our analysis of greenup time included: average temperature in the 5 days, 10 days, and 18 days preceding greenup and the number of cumulative growing degree days (over 12 degrees Celsius) in the 5 days, 10 days, and 18 days preceding greenup.
-Each of these six covariance matrices were calculated using individuals who were present at both sites.
-Additionally, each matrix was calculated using all individuals, only individuals from the Gulf subpopulation, and only individuals from the Midwest subpopulation.
+First, we applied it to only the hypothesized covariance matrices for
+each phenotype as described in Supplementary Table 1. Briefly, the
+hypothesis covariance matrices included in our analysis of greenup time
+included: average temperature in the 5 days, 10 days, and 18 days
+preceding greenup and the number of cumulative growing degree days (over
+12 degrees Celsius) in the 5 days, 10 days, and 18 days preceding
+greenup. Each of these six covariance matrices were calculated using
+individuals who were present at both sites. Additionally, each matrix
+was calculated using all individuals, only individuals from the Gulf
+subpopulation, and only individuals from the Midwest subpopulation.
 
-\
+\hfill\break
 
-For the flowering time phenotype, 12 hypothesized covariance matrices were included in the greedy algorithm.
-For each phenotype, we used the mash greedy algorithm to identify the most likely covariance matrix at each step using both a set of random independent variants and a set of independent variants that showed the strongest associations with the phenotype across all contexts.
-The resulting maximum likelihood paths for the selection of covariance matrices as determined by the greedy algorithm for each phenotype, subpopulation, and set of randomly selected or significantly associated variants are shown in Figures 1 and 2.
+For the flowering time phenotype, 12 hypothesized covariance matrices
+were included in the greedy algorithm. For each phenotype, we used the
+mash greedy algorithm to identify the most likely covariance matrix at
+each step using both a set of random independent variants and a set of
+independent variants that showed the strongest associations with the
+phenotype across all contexts. The resulting maximum likelihood paths
+for the selection of covariance matrices as determined by the greedy
+algorithm for each phenotype, subpopulation, and set of randomly
+selected or significantly associated variants are shown in Figures 1 and
+2.~
 
-\
+\hfill\break
 
-Our second analyses expanded the set of covariance matrices to all of the matrices included in our analyses of the empirical data.
-These include: the phenotype specific covariance matrices described above, the data-driven covariance matrices, and the suite of canonical covariance matrices representing context-specific effects, equal effects across contexts, and simple scenarios of heterogeneous effects across contexts.
-The resulting maximum likelihood paths for the selection of covariance matrices as determined by the greedy algorithm for each phenotype, subpopulation, and set of randomly selected or significantly associated variants are shown in Figures 3 and 4.
+Our second analyses expanded the set of covariance matrices to all of
+the matrices included in our analyses of the empirical data. These
+include: the phenotype specific covariance matrices described above, the
+data-driven covariance matrices, and the suite of canonical covariance
+matrices representing context-specific effects, equal effects across
+contexts, and simple scenarios of heterogeneous effects across contexts.
+The resulting maximum likelihood paths for the selection of covariance
+matrices as determined by the greedy algorithm for each phenotype,
+subpopulation, and set of randomly selected or significantly associated
+variants are shown in Figures 3 and 4.~
 
-\
+\hfill\break
 
-**Results**
+\textbf{Results}
 
-We first applied the greedy algorithm to only the set of hypothesis matrices for each of the greenup and flowering phenotypes, respectively.
-Generally, we observed that only a small number of matrices significantly increased the likelihood of the model as the algorithm progressed.
-In our analysis of the set of variants that were significantly associated with greenup time in at least one context, 18 hypothesis covariance matrices were included in the initialization of the greedy algorithm.
-The stop condition, a likelihood ratio test p-value \> 0.05 between the most likely model at step n and most likely model at step n+1, was met after only four iterations in both the combined cohort (Figure 1A) as well as the analysis of the Gulf subpopulation alone (Figure 1B).
-In our analysis of the midwest subpopulation the stop condition was met after only three iterations, see Figure 1C.
-The relatively small number of necessary iterations, and correspondingly small number of covariance matrices, indicate that the hypothesis covariance matrices were highly correlated with one another and that inclusion of all the hypothesis covariance matrices introduced redundancy in the model.
-In the set of randomly selected independent variants, the stop condition was met in the combined cohort after five steps (Figure 1D).
-The stop condition was met after four and three iterations in the Gulf and Midwest subpopulations, respectively.
-The overlap between matrices included in the maximum likelihood model prior to reaching the stop condition in the analysis of significantly associated and randomly selected variants was high in the combined cohort; three of four matrices in the analysis of significantly associated variants were also included in the model for the randomly selected variants.
-Similarly, three of four matrices in the Gulf subpopulation and two of three matrices in the Midwest subpopulation that were included in the model for significantly associated variants were also included in the model for the randomly selected variants.
+We first applied the greedy algorithm to only the set of hypothesis
+matrices for each of the greenup and flowering phenotypes, respectively.
+Generally, we observed that only a small number of matrices
+significantly increased the likelihood of the model as the algorithm
+progressed. In our analysis of the set of variants that were
+significantly associated with greenup time in at least one context, 18
+hypothesis covariance matrices were included in the initialization of
+the greedy algorithm. The stop condition, a likelihood ratio test
+p-value \textgreater{} 0.05 between the most likely model at step n and
+most likely model at step n+1, was met after only four iterations in
+both the combined cohort (Figure 1A) as well as the analysis of the Gulf
+subpopulation alone (Figure 1B). In our analysis of the midwest
+subpopulation the stop condition was met after only three iterations,
+see Figure 1C. The relatively small number of necessary iterations, and
+correspondingly small number of covariance matrices, indicate that the
+hypothesis covariance matrices were highly correlated with one another
+and that inclusion of all the hypothesis covariance matrices introduced
+redundancy in the model. In the set of randomly selected independent
+variants, the stop condition was met in the combined cohort after five
+steps (Figure 1D). The stop condition was met after four and three
+iterations in the Gulf and Midwest subpopulations, respectively. The
+overlap between matrices included in the maximum likelihood model prior
+to reaching the stop condition in the analysis of significantly
+associated and randomly selected variants was high in the combined
+cohort; three of four matrices in the analysis of significantly
+associated variants were also included in the model for the randomly
+selected variants. Similarly, three of four matrices in the Gulf
+subpopulation and two of three matrices in the Midwest subpopulation
+that were included in the model for significantly associated variants
+were also included in the model for the randomly selected variants.~
 
-\
+\hfill\break
 
-We then applied the greedy algorithm to the 12 hypothesis covariance matrices for the flowering phenotype.
-Similar to our analysis of the hypothesis covariance matrices for the greenup phenotype, we observed a similar pattern of correlation in hypothesis covariance matrices for the flowering phenotype.
-The stop condition for the greedy algorithm was met after five steps when analyzing the combined cohort, the Gulf subpopulation, and the Midwest subpopulation in analysis of variants that were significantly associated in at least one context (Figure 2A-C).
-When analyzing the set of randomly selected variants the stop condition was met after four steps in the combined cohort, five steps in the Gulf subpopulation, and six steps in the Midwest subpopulation.
-We also observed that many of the matrices included in the model for the significantly associated variants were also included in the corresponding model for randomly selected variants (Figure 2D-F).
+We then applied the greedy algorithm to the 12 hypothesis covariance
+matrices for the flowering phenotype. Similar to our analysis of the
+hypothesis covariance matrices for the greenup phenotype, we observed a
+similar pattern of correlation in hypothesis covariance matrices for the
+flowering phenotype. The stop condition for the greedy algorithm was met
+after five steps when analyzing the combined cohort, the Gulf
+subpopulation, and the Midwest subpopulation in analysis of variants
+that were significantly associated in at least one context (Figure
+2A-C). When analyzing the set of randomly selected variants the stop
+condition was met after four steps in the combined cohort, five steps in
+the Gulf subpopulation, and six steps in the Midwest subpopulation. We
+also observed that many of the matrices included in the model for the
+significantly associated variants were also included in the
+corresponding model for randomly selected variants (Figure 2D-F).~
 
-\
+\hfill\break
 
-Analysis of the hypothesis covariance matrices alone helped to establish the utility of the greedy algorithm implementation in identifying the maximum likelihood mash model for the observed summary statistics from our analysis of eight sites.
-However, due to substantial evidence that both canonical and data driven matrices explained a nontrivial proportion of the observed covariance between context specific summary statistics we applied the greedy algorithm with the all three classes of covariance matrices included in the initial set.
-In our analysis of the greenup phenotype, the stop condition was met after 16 iterations, resulting in a maximum likelihood model that included all 13 canonical covariance matrices, two data driven matrices (total PCA and PC 4), and the hypothesis matrix representing the covariance in average temperature ten days prior to greenup in the combined cohort (Figure 3A).
-When the Gulf subpopulation was analyzed alone the stop condition was met after five iterations.
-The five corresponding matrices were composed of four canonical matrices and the hypothesis matrix calculated as the covariance in average temperature five days prior to greenup in only individuals from the Gulf subpopulation (Figure 3B).
-Finally, the greedy algorithm met the stop condition after 13 iterations when applied to the Midwest subpopulation.
-Ten of the canonical covariance matrices, the total PCA data driven covariance matrix, and the two covariance matrices calculated using the five day temperature average in the five days prior to greenup and the number of growing degree days in the 18 days prior to greenup in the Midwest subpopulation individuals were all include in the mode (Figure 3C).
-Most notably, the canonical covariance matrix representing independent effects was the first matrix selected in all three applications of the greedy algorithm, indicating a large number of independent effects were present in our analysis of the greenup phenotype.
-We then applied the greedy algorithm to a set of randomly selected variants and observed that many of the same matrices included in the corresponding models from application to the significantly associated variants were observed in the most likely models once the stop condition was met (Figure 3D-F).
+Analysis of the hypothesis covariance matrices alone helped to establish
+the utility of the greedy algorithm implementation in identifying the
+maximum likelihood mash model for the observed summary statistics from
+our analysis of eight sites. However, due to substantial evidence that
+both canonical and data driven matrices explained a nontrivial
+proportion of the observed covariance between context specific summary
+statistics we applied the greedy algorithm with the all three classes of
+covariance matrices included in the initial set. In our analysis of the
+greenup phenotype, the stop condition was met after 16 iterations,
+resulting in a maximum likelihood model that included all 13 canonical
+covariance matrices, two data driven matrices (total PCA and PC 4), and
+the hypothesis matrix representing the covariance in average temperature
+ten days prior to greenup in the combined cohort (Figure 3A). When the
+Gulf subpopulation was analyzed alone the stop condition was met after
+five iterations. The five corresponding matrices were composed of four
+canonical matrices and the hypothesis matrix calculated as the
+covariance in average temperature five days prior to greenup in only
+individuals from the Gulf subpopulation (Figure 3B). Finally, the greedy
+algorithm met the stop condition after 13 iterations when applied to the
+Midwest subpopulation. Ten of the canonical covariance matrices, the
+total PCA data driven covariance matrix, and the two covariance matrices
+calculated using the five day temperature average in the five days prior
+to greenup and the number of growing degree days in the 18 days prior to
+greenup in the Midwest subpopulation individuals were all include in the
+mode (Figure 3C). Most notably, the canonical covariance matrix
+representing independent effects was the first matrix selected in all
+three applications of the greedy algorithm, indicating a large number of
+independent effects were present in our analysis of the greenup
+phenotype. We then applied the greedy algorithm to a set of randomly
+selected variants and observed that many of the same matrices included
+in the corresponding models from application to the significantly
+associated variants were observed in the most likely models once the
+stop condition was met (Figure 3D-F).~
 
-\
+\hfill\break
 
-We then performed the same analysis of the entire suite of canonical, data driven, and hypothesis covariance matrices for the flowering phenotype.
-The stop condition was met after five iterations in the combined cohort and included four canonical covariance matrices representing independent effects, equal effects, and two simple models of effect heterogeneity among sites (Figure 4A).
-In application to the Gulf subpopulation, the stop condition was met after 11 iterations and included 10 canonical covariance matrices and the data driven covariance matrix derived from the third principal component (Figure 4B).
-The stop condition was met after 11 iterations in the analysis of Midwest subpopulation individuals and included nine canonical covariance matrices, the data driven matrix derived from the first principal component, and the hypothesis matrix calculated using the cumulative rainfall in the day prior to flowering tin the Midwest subpopulation (Figure 4C).
-Once again, we observed that the first matrix selected in each analysis of flowering time was the canonical matrix representing independent effects among conditions.
-Overlap between the models selected by the greedy algorithm when analyzing a set of randomly selected variants was high across all three analyses (Figure 4C-D).
+We then performed the same analysis of the entire suite of canonical,
+data driven, and hypothesis covariance matrices for the flowering
+phenotype. The stop condition was met after five iterations in the
+combined cohort and included four canonical covariance matrices
+representing independent effects, equal effects, and two simple models
+of effect heterogeneity among sites (Figure 4A). In application to the
+Gulf subpopulation, the stop condition was met after 11 iterations and
+included 10 canonical covariance matrices and the data driven covariance
+matrix derived from the third principal component (Figure 4B). The stop
+condition was met after 11 iterations in the analysis of Midwest
+subpopulation individuals and included nine canonical covariance
+matrices, the data driven matrix derived from the first principal
+component, and the hypothesis matrix calculated using the cumulative
+rainfall in the day prior to flowering tin the Midwest subpopulation
+(Figure 4C). Once again, we observed that the first matrix selected in
+each analysis of flowering time was the canonical matrix representing
+independent effects among conditions. Overlap between the models
+selected by the greedy algorithm when analyzing a set of randomly
+selected variants was high across all three analyses (Figure 4C-D).
 
-\
+\hfill\break
 
-**Figures**
+\textbf{Figures}
 
-\
-\
+\hfill\break
+\hfill\break
 
-![](https://lh7-us.googleusercontent.com/ae_Fi1djzcYo_ibeQWWjr2AXuwIX781VIa7wSug5K3H9T8F5ke1RYU8qKPRT-YGO_KvDJ5oK3pzzQX3G12zbJOo9i6_yLEV8NkpRHF4zT3dqrhXKfLISSV5N-iXjYIczVPSJcDEPx-W7vv7NvU1AQv8)
+\includegraphics{response_to_reviews_files/mediabag/ae_Fi1djzcYo_ibeQWWj.pdf}
 
-**Figure 1.
-Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the greenup phenotype using only hypothesis covariance matrices.** The matrix that resulted in the maximum likelihood model of covariance in summary statistics among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis.
-**(A)** The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals.
-The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
-**(B)** and **(C)** show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.
-**(D)** The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals.
-The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
-\* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants.
-**(E)** and **(F)** show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.
+\textbf{Figure 1. Log-likelihood at each step in a greedy algorithm
+implementation of model selection for mash for the greenup phenotype
+using only hypothesis covariance matrices.} The matrix that resulted in
+the maximum likelihood model of covariance in summary statistics among
+sites is located at the base of the y-axis; matrices selected in the
+ensuing step are placed in ascending order along the y-axis.
+\textbf{(A)} The maximum likelihood path of covariance matrices in the
+greedy algorithm for the set of independent variants that were
+significant in at least one context when analyzing all individuals. The
+dashed black line represents the step in the algorithm where addition of
+the next matrix did not significantly increase the model likelihood,
+likelihood-ratio test p-value \textgreater{} 0.05. \textbf{(B)} and
+\textbf{(C)} show the maximum likelihood paths when the individuals
+analyzed are only from the Gulf and Midwest subpopulations,
+respectively. \textbf{(D)} The maximum likelihood path of covariance
+matrices in the greedy algorithm for the set of independent, randomly
+selected variants when analyzing all individuals. The dashed black line
+represents the step in the algorithm where addition of the next matrix
+did not significantly increase the model likelihood, likelihood-ratio
+test p-value \textgreater{} 0.05. * is used to denote those matrices
+that are identified by the greedy algorithm in the set of significantly
+associated variants and the randomly associated variants. \textbf{(E)}
+and \textbf{(F)} show the maximum likelihood paths when the individuals
+analyzed are only from the Gulf and Midwest subpopulations,
+respectively.
 
-\
+\hfill\break
 
-![](https://lh7-us.googleusercontent.com/8AuFUNJTaIAgX7GVkOSnrBMMNqZr4mYdesj48zBZLSssUZ15epDhoC8FLp4gl_5kvZiR94cqb925c23NIRXakRL1dh2TVp_a1wtf-gw1w_QKx1hkaGvJ_CsQHwHube_7PFWBDq1zZgsaeZXQ65pON_8)
+\includegraphics{response_to_reviews_files/mediabag/8AuFUNJTaIAgX7GVkOSn.pdf}
 
-**Figure 2.
-Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the flowering phenotype using only hypothesis covariance matrices.** The matrix that resulted in the maximum likelihood model among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis.
-**(A)** The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals.
-The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
-**(B)** and **(C)** show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.
-**(D)** The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals.
-The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
-\* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants.
-**(E)** and **(F)** show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.
+\textbf{Figure 2. Log-likelihood at each step in a greedy algorithm
+implementation of model selection for mash for the flowering phenotype
+using only hypothesis covariance matrices.} The matrix that resulted in
+the maximum likelihood model among sites is located at the base of the
+y-axis; matrices selected in the ensuing step are placed in ascending
+order along the y-axis. \textbf{(A)} The maximum likelihood path of
+covariance matrices in the greedy algorithm for the set of independent
+variants that were significant in at least one context when analyzing
+all individuals. The dashed black line represents the step in the
+algorithm where addition of the next matrix did not significantly
+increase the model likelihood, likelihood-ratio test p-value
+\textgreater{} 0.05. \textbf{(B)} and \textbf{(C)} show the maximum
+likelihood paths when the individuals analyzed are only from the Gulf
+and Midwest subpopulations, respectively. \textbf{(D)} The maximum
+likelihood path of covariance matrices in the greedy algorithm for the
+set of independent, randomly selected variants when analyzing all
+individuals. The dashed black line represents the step in the algorithm
+where addition of the next matrix did not significantly increase the
+model likelihood, likelihood-ratio test p-value \textgreater{} 0.05. *
+is used to denote those matrices that are identified by the greedy
+algorithm in the set of significantly associated variants and the
+randomly associated variants. \textbf{(E)} and \textbf{(F)} show the
+maximum likelihood paths when the individuals analyzed are only from the
+Gulf and Midwest subpopulations, respectively.
 
-\
+\hfill\break
 
-![](https://lh7-us.googleusercontent.com/iO_JwhiNJbDDdV-UHqzlZOnGUemRBlDpw9qDXXHYfLUzRTgpU0w8fxQczs4NLhz5apnaD_x0CL1CBfPd0KlLtbprYsaC4Cub2-2nIyzkvMAcp3MkyzkB5x82-L3kJFfiBSJPxDg7XqZ6JIox4hG2e-A)
+\includegraphics{response_to_reviews_files/mediabag/iO_JwhiNJbDDdV-UHqzl.pdf}
 
-**Figure 3.
-Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the greenup phenotype using canonical, data drive, and hypothesis covariance matrices.** The matrix that resulted in the maximum likelihood model of covariance in summary statistics among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis.
-**(A)** The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals.
-The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
+\textbf{Figure 3. Log-likelihood at each step in a greedy algorithm
+implementation of model selection for mash for the greenup phenotype
+using canonical, data drive, and hypothesis covariance matrices.} The
+matrix that resulted in the maximum likelihood model of covariance in
+summary statistics among sites is located at the base of the y-axis;
+matrices selected in the ensuing step are placed in ascending order
+along the y-axis. \textbf{(A)} The maximum likelihood path of covariance
+matrices in the greedy algorithm for the set of independent variants
+that were significant in at least one context when analyzing all
+individuals. The dashed black line represents the step in the algorithm
+where addition of the next matrix did not significantly increase the
+model likelihood, likelihood-ratio test p-value \textgreater{} 0.05.
 Once this condition was met the greedy algorithm was halted.
-**(B)** and **(C)** show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.
-**(D)** The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals.
-The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
-Once this condition was met the greedy algorithm was halted for computational considerations.
-\* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants.
-**(E)** and **(F)** show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.
+\textbf{(B)} and \textbf{(C)} show the maximum likelihood paths when the
+individuals analyzed are only from the Gulf and Midwest subpopulations,
+respectively. \textbf{(D)} The maximum likelihood path of covariance
+matrices in the greedy algorithm for the set of independent, randomly
+selected variants when analyzing all individuals. The dashed black line
+represents the step in the algorithm where addition of the next matrix
+did not significantly increase the model likelihood, likelihood-ratio
+test p-value \textgreater{} 0.05. Once this condition was met the greedy
+algorithm was halted for computational considerations. * is used to
+denote those matrices that are identified by the greedy algorithm in the
+set of significantly associated variants and the randomly associated
+variants. \textbf{(E)} and \textbf{(F)} show the maximum likelihood
+paths when the individuals analyzed are only from the Gulf and Midwest
+subpopulations, respectively.
 
-\
+\hfill\break
 
-![](https://lh7-us.googleusercontent.com/LnnT8NIjtolmYO5neLCaID6fOqbMyUBZalKFy9eiHmC4xTTSFoNoTVIxgIQpn3n5FDynzHwBdZmGaRv2ktSquMKcj5RsdqV453tRWtRzMvRk3va8YV5g-pItSMpPSQqTnW7iZ7l6xPR5Zfh-4tEorVI)
+\includegraphics{response_to_reviews_files/mediabag/LnnT8NIjtolmYO5neLCa.pdf}
 
-**Figure 4.
-Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the flowering phenotype using canonical, data drive, and hypothesis covariance matrices.** The matrix that resulted in the maximum likelihood model of covariance in summary statistics among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis.
-**(A)** The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals.
-The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
+\textbf{Figure 4. Log-likelihood at each step in a greedy algorithm
+implementation of model selection for mash for the flowering phenotype
+using canonical, data drive, and hypothesis covariance matrices.} The
+matrix that resulted in the maximum likelihood model of covariance in
+summary statistics among sites is located at the base of the y-axis;
+matrices selected in the ensuing step are placed in ascending order
+along the y-axis. \textbf{(A)} The maximum likelihood path of covariance
+matrices in the greedy algorithm for the set of independent variants
+that were significant in at least one context when analyzing all
+individuals. The dashed black line represents the step in the algorithm
+where addition of the next matrix did not significantly increase the
+model likelihood, likelihood-ratio test p-value \textgreater{} 0.05.
 Once this condition was met the greedy algorithm was halted.
-**(B)** and **(C)** show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.
-**(D)** The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals.
-The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
-Once this condition was met the greedy algorithm was halted for computational considerations.
-\* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants.
-**(E)** and **(F)** show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.
+\textbf{(B)} and \textbf{(C)} show the maximum likelihood paths when the
+individuals analyzed are only from the Gulf and Midwest subpopulations,
+respectively. \textbf{(D)} The maximum likelihood path of covariance
+matrices in the greedy algorithm for the set of independent, randomly
+selected variants when analyzing all individuals. The dashed black line
+represents the step in the algorithm where addition of the next matrix
+did not significantly increase the model likelihood, likelihood-ratio
+test p-value \textgreater{} 0.05. Once this condition was met the greedy
+algorithm was halted for computational considerations. * is used to
+denote those matrices that are identified by the greedy algorithm in the
+set of significantly associated variants and the randomly associated
+variants. \textbf{(E)} and \textbf{(F)} show the maximum likelihood
+paths when the individuals analyzed are only from the Gulf and Midwest
+subpopulations, respectively.
+
+\end{document}

### 2023-12-02 manuscript/revision2/SI_Appendix.qmd
manuscript: SI Appendix and response to reviews for revision

---
title: "SI Appendix"
format: pdf
editor: visual
---

**SI Appendix**

\

**Dataset 1.** SNP-associated effects and standard errors for green-up estimated using the Gulf genetic subpopulation.

**Dataset 2.** SNP-associated effects and standard errors for green-up estimated using the Midwest genetic subpopulation.

**Dataset 3.** SNP-associated effects and standard errors for green-up estimated using Both genetic subpopulations.

**Dataset 4.** SNP-associated effects and standard errors for flowering estimated using the Gulf genetic subpopulation.

**Dataset 5.** SNP-associated effects and standard errors for flowering estimated using the Midwest genetic subpopulation.

**Dataset 6.** SNP-associated effects and standard errors for flowering estimated using Both genetic subpopulations.

**Dataset 7.** Genes within QTL that have functionally validated roles in flowering in rice.

**Dataset 8.** Overlap between quantitative trait loci for flowering measured using an outbred pseudo-F2 mapping population, and significant mash effect estimates (with a log10Bayes Factor \> 2, or in the 1% tail of most significant mash results) found using three overlapping diversity panels. Genes within 20kb with a functionally validated role in flowering in rice are  indicated.

\

**Figure S1.** Variance components analysis of genetic (purple), genotype by environment (blue), environmental (green), and residual (yellow) terms in models of green-up time as functions of Julian date or weather, for the four northern common gardens (North), the three Texas common gardens (Texas), and for all eight common gardens (All). Dashed lines indicate the cumulative contribution of G and GxE for flowering as a function of day of the year. ![](https://lh7-us.googleusercontent.com/SPDH2Y9HThoK0pUZQxFuLPUDrIuXAcG73nUSk2ucYN3_CHXSG-c8G4b7Gq6crgo_1IapiGYFV66w_vhfyvxSIW8kPtTQA_tMeviTFamFdsXIPjIjTcaI0dwa7ZqpKN0P1RonX05vbxitm7vLYoYoZw)

**Figure S2.** Effect patterns exhibited by the major data-driven matrices for (a,d,e,j,m,n) Gulf, (b,f,g,k,o,p) Midwest, and (c,h,i,l,q,r) Both subpopulations from Figure 2. (a-i) Green-up date. (j-r) Flowering date. (a-c,j-l) Single-garden effect representations (eigenvectors) of the DD_tPCA data-derived matrices. The percent variation explained of DD_tPCA by each eigenvector is shown on the y-axis. Common gardens are arranged in latitudinal order along the x-axis. In addition, the first eigenvector corresponds to DD_PCA_1, and second corresponds to DD_PCA_2. (d,f,h,m,o,q) The number of SNPs with significant effects in both conditions that exhibit antagonistic pleiotropy between that pair of conditions. (e,g,i,n,p,r) Same as (d,f,h,m,o,q) except for differential sensitivity. Scales for (d,e), (f,g), (h,i), (m,n), (o,p), and (q,r) are identical and located on the first panel in alphabetical order.![](https://lh7-us.googleusercontent.com/JV_8jSfymcCRhEQsEuXD2WzHAkM7bOTpEws9j3_H4SbmaTArViZOZQyWld6aMI-xs8DPYLSZBqpxogUokZwU36l5NMMlJgLy-dY8GFwACkYAntFMqAvq9b1C3BsXnbacDYOF3i4kRqXIW4ISTu2Xzg)

\

\
![](https://lh7-us.googleusercontent.com/h-CoStrONeN2vAZcGvcrGRLg1RRQiAhfNYCT9xn9YF5uwInX02rCepO4NU1bRq1ykboFZcOjr1PeEiZXsXRDJ6v51BAzYPWKQ2wd8YhVg2LJA7IMkL5wvEEa8jOw0UDgrZgVwiYZlHQWSqGxT_2q0g)

**Figure S3.** Comparison of mash SNP effects from a diversity panel and effects from an outbred pseudo-F2 cross. (**a**) QTL mapping for four weather-related functions of flowering, and for green-up date, as indicated by the colors: pink, growing degree days (GDD) from green-up date to flowering date, yellow, day length change on the flowering date, green, day length on flowering date; purple, flowering date; gray, green-up date.Dotted lines indicate permutation-based significance thresholds for each weather-related function. Stars indicate QTL with significant enrichment for SNPs in the 1% mash tail; B, G, and M indicate which subpopulation had enrichment: B -  both subpopulations, G - Gulf subpopulation, M -Midwest subpopulation. Rug plots show genomic locations of SNPs in the 1% mash tail for flowering date for each subpopulation. (**b**) Schematic comparison of genotypes used for QTL mapping and genotypes used in mash. (**c**) Number of mash runs enriched for SNPs in the 1% mash tail in the 23 QTL intervals from (**a**) (dotted red line), compared to 1000 sets of 23 random QTL intervals of the same size (histogram).**(d**) SNP effects estimated using mash for SNPs with the highest Bayes factor for the three QTL with the highest LOD scores from (**a**) and enrichment of SNPs in the 1% mash tail. SNP genomic position and genetic subpopulation are indicated above each plot. Common gardens are arranged in latitudinal order. (**e**) QTL effect estimates for 23 non-overlapping and partially overlapping QTLs. Colors indicate which of the four weather-related functions of flowering for which the QTL was mapped. An additional five QTLs completely overlapped the QTL displayed; in this case, the QTL with the highest lod score is shown. Solid vertical lines separate QTLs that are not overlapping. Stars indicate QTL with significant enrichment for SNPs in the 1% mash tail; B, G, and M indicate which subpopulation had enrichment, as in (**b**):B -  both subpopulations, G - Gulf subpopulation, M -Midwest subpopulation. Common gardens are arranged in latitudinal order.

**Table S1.** Green-up and flowering dates phenotypes linked to environmental cues.

+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
| Phenotype | Name                              | Description                                                                                                                          |
+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
| green-up  | GDD, **n**d                       | Correlations in cumulative growing degree days for the five, ten, or eighteen days prior to green-up.                                |
|           |                                   |                                                                                                                                      |
|           | where \'**n**\' is in {5, 10, 18} |                                                                                                                                      |
+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
| green-up  | Temp Ave, **n**d                  | Correlations in average temperature for the five, ten, or eighteen days prior to green-up.                                           |
|           |                                   |                                                                                                                                      |
|           | where \'**n**\' is in {5, 10, 18} |                                                                                                                                      |
+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
| flowering | 50% green-up                      | Correlations in green-up.                                                                                                            |
+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
| flowering | daylength                         | Correlations in day length on the day of flowering.                                                                                  |
+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
| flowering | daylength change (s)              | Correlations in the change in daylength (in seconds) on the day of flowering.                                                        |
+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
| flowering | GDD, GR to FL                     | Correlations in cumulative GDD between green-up and flowering.                                                                       |
+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
| flowering | Rainfall, **n**d sum              | Correlations in cumulative rainfall in the one, three, or five days before flowering, or in the days between green-up and flowering. |
|           |                                   |                                                                                                                                      |
|           | where \'**n**\' is in {1, 3, 5}   |                                                                                                                                      |
+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+

\
\

**Section S1.** Weather-based environmental cues for green-up and flowering date.

\

	We scored green-up and flowering by day of the year, then linked these dates to multiple environmental factors (Table S1). We defined flowering as functions of eight variables: \'flowering date\' (day of the year), cumulative growing degree days (GDD) between green-up and flowering dates (\'flowering GDD\'), day length on the flowering date (\'flowering daylength\'), change in daylength relative to the previous day on the day of flowering (\'flowering daylength change\'), and four measures of cumulative rainfall: cumulative rainfall between green-up and flowering, and in the one day, three days, and five days before flowering. Cumulative GDD was calculated as GDD = GRFLmaxTmean−Tbase,0, where Tmean is the daily average temperature, defined as (Tmax+Tmin)/2, Tbase is the base temperature of 12 °C for switchgrass, Tmax is the maximum daily temperature, Tmin is the minimum daily temperature, GR is green-up date, and FL is flowering date (Kiniry et al. 2005; Behrman et al. 2013).  On a specific day, if Tmean is less than Tbase, the GDD for that day is 0; if Tmean is bigger than Tbase, the GDD for that day is the difference between Tmean and Tbase. We also defined green-up as functions of seven variables: \'green-up date\' (day of the year), cumulative GDD for the five, ten, or eighteen days prior to green-up, and the average air temperature for the five, ten, or eighteen days prior to green-up.

**Section S2.** Hypothesis-based covariance matrix definitions for mash.

	We generated hypothesis-based covariance matrices derived from correlations in environmental cues in the green-up or flowering date windows for the three subpopulations (Table S1). These covariance matrices were derived from the same set of phenotypes as environmental functions of green-up and flowering, and represent the correlations between genotypes for these phenotypes across our common gardens. For the diagonal of these matrices, we used the coefficient of variation in these phenotypes within the subpopulation at that garden. Code used to generate these matrices can be found at  [git\@github.com](mailto:git@github.com):Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd. We used distinct sets of user-specified, hypothesis-based covariance matrices for green-up and flowering, but the same set of matrices for all genetic subgroups. If a hypothesis-based matrix does not capture a common pattern of SNP effects, mash assigns small mixture proportions onto this matrix using maximum likelihood, giving that covariance matrix a low posterior weight summed across all SNP effects, which ultimately causes little to no change in the mash model. Alternatively, SNPs with high mixture proportions on particular environmental covariance matrices and large Bayes factors, which summarize the overall significance of a non-zero effect, represent small genomic intervals with strong evidence for a phenotypic effect correlated with an environmental driver.

\

**Section S3.** Narrow-sense heritability

	In the diversity panel, we determined narrow-sense heritabilities (h2) for green-up and flowering dates at single gardens using genomic relationship matrices calculated using the van Raden method (VanRaden et al. 2009). Genomic relationship matrices were calculated within each subpopulation (Midwest and Gulf) and for both genetic subpopulations (Both). We used ASReml (VSN International) to specify mixed models of the form:

**y** = 1 + Zu + e

Var(u) = Gσu2

Var(e) = Iσe2

in which the vector **y** represents the flowering date or green-up date values for that garden, Z the design matrix for random effects, u the whole genome additive genetic effect, and e the residual. Matrix G is the whole genomic relationship matrix based on all SNPs retained for subpopulation-specific analyses. I is the rank-y identity matrix. Phenotypic variance (σp2) is σu2 + σe2. Narrow-sense heritability is then h2 = (σu2/σp2).

	These models were run for each of the eight gardens, and across all gardens by adding an additional environmental effect of site without an interaction term. This resulted in 54 models: 3 sets of populations (the Gulf, Midwest, and Both subpopulations) for 9 garden sets (all eight gardens separately, and all eight gardens together) and two phenotypes (green-up date and flowering date).

\

**Section S4.** Variance components analysis

	In the diversity panel, to evaluate our environmental cues as genetic triggers of flowering, we defined green-up and flowering for individual genotypes as functions of seven and eight environmental cues, respectively, then used variance components analysis to partition variance attributed to genetic effects (G), genotype by environment interactions (GxE), environmental effects (E), and error for these phenology-related traits using linear mixed models. We used mixed models of the form:

**y** = 1 + Zuu + Zll + Zulul + e

Var(u) = Gσu2

Var(l) = IσE2

Var(ul) = GσuE2

Var(e) = Iσe2

in which the vector **y** represents the individual plant observations of a given trait and Z the design matrices associating trait observations with random factors. Random factors include u, the whole genome additive genetic (G) effect, l, the effect of each location (E), ul, the GxE effect, and e, the residual. Matrix G is the whole genomic relationship matrix. I are rank-y identity matrices.

	These models were run for each phenotype as a function of environmental cues. This resulted in 135 models: three sets of populations (the Gulf, Midwest, and Both subpopulations) for three garden sets (All, North, and Texas gardens) and 15 phenotypes (seven green-up functions, and eight flowering functions).

\

**Section S5.** Multivariate adaptive shrinkage

	To evaluate the prevalence and kinds of covariance patterns of SNP effects across our eight common gardens, we used multivariate adaptive shrinkage (mash) on SNP effect estimates from the diversity panel (Urbut et al. 2019). Mash is a statistical method that allows estimation and comparison of many effects jointly across many different conditions; it improves on previous methods by allowing arbitrary correlations in effect sizes among conditions. To this end, mash allows the user to include hypothesis-based covariance matrices, specifies some \'canonical\' covariance matrices with simple patterns of effects, and also generates \'data-driven\' covariance matrices from patterns of effects in the data. Then, mash assigns mixture proportions for each SNP onto each provided covariance matrix using maximum likelihood. Finally, mash uses Bayes\' theorem to shrink effects for each SNP towards the set of covariance matrices in accordance to their mixture proportions. For example, a SNP affecting flowering may have a high mixture proportion, or weight, on a covariance matrix created from a specific environmental cue. In that case, we can infer that the effect of that SNP on flowering is caused by a response to or interaction with that environmental cue. These user-specified and data-driven covariance matrices are an important advantage mash offers for studying patterns of GxE: the user-specified covariance matrices allow hypothesis testing of specific environmental drivers for each SNP, while the data-driven covariance matrices allow exploration of additional unexplained patterns of covariation.

	We first conducted univariate GWAS at each common garden for green-up and flowering date, then analyzed the allelic effects for the top 19K unlinked SNPs per univariate GWAS using mash, as in Lovell et al. (2021). To allow for the possibility that different subpopulations had different strengths of connection between our phenotypes and genotypes (Korte and Farlow 2013), we conducted three sets of genetic analyses: on Gulf and Midwest genotypes separately, and on both subpopulations together (hereafter \'Both\' subpopulations). Only SNPs with ≤ 20% missing data and minor allele frequencies \> 0.05 were used in univariate GWAS, resulting in 8.8 million SNPs retained for the Midwest subpopulation, 10.3 million SNPs retained for the Gulf subpopulation, and 12.3 million SNPs retained for Both subpopulations. Then, we created a set of \'random\' markers to account for correlation in effects using a set of 19K relatively unlinked (r2 \< 0.2) markers. Next, we created a set of \'strong\' markers by selecting 19K SNPs with the largest -log10(p-value) in any condition in the univariate GWAS, out of all markers unlinked at (r2 \< 0.2), which represented the SNP with the strongest effect in the 19K LD blocks with the largest univariate effects. We used the \'pvdiv_standard_run()\' function of the switchgrassGWAS R package (https://github.com/Alice-MacQueen/switchgrassGWAS) to conduct GWAS on genotypes from the Midwest, Gulf, and Both subpopulations, used the \'pvdiv_bigsnp2mashr()\' function to convert univariate GWAS output to the matrices of effects and standard errors needed for mash, then used the \'mash_standard_run()\' function to conduct mash, again specifying distinct sets of hypothesis-based covariance matrices for green-up and flowering, but the same set of matrices for all genetic subgroups.

\

**Section S6.** Outbred pseudo-F2 mapping population and Quantitative Trait Locus mapping

	To confirm candidate genomic regions and patterns of allelic effects found in the diversity panel, we analyzed flowering in an outbred pseudo-F2 cross between four individuals, two Midwest and two Gulf individuals. The formation of this mapping population has been described previously (Milano, Lowry, and Juenger 2016). The parents of this cross were DAC, an early flowering Midwest individual, VS16, a late flowering Midwest individual, AP13, an early flowering Gulf individual, and WBC, a late flowering Gulf individual. We made F1 crosses of the two early flowering genotypes, AP13xDAC, and the two late flowering genotypes, WBCxVS16. We then clonally propagated and planted the four parents, the two F1 genotypes (AP13xDAC, and VS16xWBC), and 801 F2 genotypes at eight field sites in May-July of 2015. To be directly comparable to the diversity panel data, only 2019 phenology data from the pseudo-F2 cross from the same eight common garden sites were used here.

	Details on the genetic map construction, map polishing and fine-scale reordering can be accessed on <https://datadryad.org/stash/dataset/doi:10.5061/dryad.ghx3ffbjv>. QTL mapping was conducted with R/qtl2 (Broman et al. 2019). We performed a genome scan with a linear mixed model that accounts for the relationships among individuals and for environmental covariates (i.e., field sites). The full model can be expressed as: phenotype = µ + QTL + E + QTLxE + kinship + e, where µ is the population mean, QTL is the marker genetic effect, E is the environmental effects (here, common garden), QTLxE is the interaction between marker genetic and environmental effects, kinship corresponds to the background polygenic variation, and e is the error term. The genome scan was accomplished with the \'scan1\' function. The statistical significance of the genome scan was established by performing a stratified (i.e., stratifying on common garden) permutation test (n=1000) using \'scan1perm\' function. The estimated QTL effect was obtained using \'scan1coef\' function in R/qtl2.

### 2023-12-02 manuscript/revision2/revision2.qmd
manuscript: SI Appendix and response to reviews for revision

@@
 
-```         
 Narrow-sense heritabilities (h2) indicated that rank-changing GxE for these phenotypes was present across the common gardens (Fig 1C).
 h2 were typically high at individual gardens: 59% on average for green-up date, and 87% for flowering date.
-However, h2 were variable across gardens, and green-up dates were uncorrelated (r2 < 0.2) or negatively correlated between pairs of gardens (Fig.
+However, h2 were variable across gardens, and green-up dates were uncorrelated (r2 \< 0.2) or negatively correlated between pairs of gardens (Fig.
 1B).
 These negative and small correlations undoubtedly contributed to the low h2 values for green-up and flowering date when estimated jointly at all eight gardens: h2 was 0.8% for green-up and 23.2% for flowering date.
-```
 
@@
 
-```         
-We next explored how genetic variation in phenology covaried with environmental cues by using mash to jointly discover trait-associated SNPs across all common gardens for the Midwest, Gulf, and ‘Both’ subpopulations (Fig 1D, SI Appendix, Datasets 1-6).
+We next explored how genetic variation in phenology covaried with environmental cues by using mash to jointly discover trait-associated SNPs across all common gardens for the Midwest, Gulf, and 'Both' subpopulations (Fig 1D, SI Appendix, Datasets 1-6).
 We first estimated SNP effects using GWAS on site-specific BLUPs for each phenotype.
-We then used mash on two subsets of these SNP effects, first to determine the covariance structures that significantly improved the mash model log likelihood when included, then to re-estimate the set of SNP effects with the largest effect estimates in univariate GWAS (hereafter “strong” effects) using the set of significant covariance matrices.
-```
+We then used mash on two subsets of these SNP effects, first to determine the covariance structures that significantly improved the mash model log likelihood when included, then to re-estimate the set of SNP effects with the largest effect estimates in univariate GWAS (hereafter "strong" effects) using the set of significant covariance matrices.
 
@@
 
+# Figures and Tables
+
+![Figure 1.
+Characterization of green-up and flowering dates from the switchgrass diversity panel.
+(a) Map and trait histograms of green-up and flowering dates across two genetically distinct switchgrass subpopulations and eight common gardens.
+Purple represents individuals from the Midwest genetic subpopulation, and pink individuals from the Gulf subpopulation.
+Vertical dashed lines indicate the summer solstice.
+Common gardens are arranged in latitudinal order.
+(b) Phenotypic correlations between clonal replicates planted at eight common gardens, within and between two genetic subpopulations.  (c) Narrow sense heritability of green-up and flowering within single common gardens (purple) and across all eight common gardens (green), within and between two genetic subpopulations.
+(d) Flow diagram of the methods applied to the green-up and flowering dates to jointly estimate SNP effects across all sites.
+Mash was fit to SNP effect data and used to find covariance matrices that improved the mash model likelihood using a large set of randomly selected, relatively unlinked SNP effects; this model was applied to a \"strong\" set of SNP effects with large effect sizes in the univariate GWAS.](images/Fig1_200dpi.png)
+
+![Figure 2.](images/Figure_2_Covariance_matrices_and_posterior_weights.png)
+
 ## Author Affiliations {.unnumbered}
@@
 ## Figures and Tables {.unnumbered}
+
+| **Weather variable**                                                      | **Time frame**                                               | **\# Covariance Matrices tested in greedy mash algorithm** |
+|---------------------------|-----------------------|----------------------|
+| cumulative GDD at 12C in the time frame                                   | 1, 2, 3, 4, 5, 6, 7, 14, 21, 28 days prior to green up date  | 60, 30 per phenotype and 10 per phenotype & genetic group  |
+| cumulative rainfall in the time frame                                     | 1, 2, 3, 4, 5, 6, 7, 14, 21, 28 days prior to green up date  | 60, 30 per phenotype and 10 per phenotype &  genetic group |
+| day length (hours) on a specific day indicated by the time frame          | 1, 2, 3, 4, 5, 6, 7, and 14 days prior to green up date      | 48, 24 per phenotype and 8 per phenotype &  genetic group  |
+| day length change (seconds) on a specific day indicated by the time frame | 1, 2, 3, 4, 5, 6, 7, and 14 days prior to green up date      | 48, 24 per phenotype and 8 per phenotype & genetic group   |
+| average temperature in the time frame                                     | 1, 2, 3, 4, 5, 6, 7, 14, 21, 28 days prior to green up date  | 60, 30 per phenotype and 10 per phenotype & genetic group  |
+
+: Weather variables and time frames around the green up and flowering date used to construct the hypothesis-based covariance matrices.
+The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices.
+Narrow-sense heritabilities for these values at each garden were used for the diagonal cells.
 

### 2023-12-02 manuscript/revision2/revision2.tex
manuscript: SI Appendix and response to reviews for revision

@@
 
-\begin{verbatim}
-Narrow-sense heritabilities (h2) indicated that rank-changing GxE for these phenotypes was present across the common gardens (Fig 1C).
-h2 were typically high at individual gardens: 59% on average for green-up date, and 87% for flowering date.
-However, h2 were variable across gardens, and green-up dates were uncorrelated (r2 < 0.2) or negatively correlated between pairs of gardens (Fig.
-1B).
-These negative and small correlations undoubtedly contributed to the low h2 values for green-up and flowering date when estimated jointly at all eight gardens: h2 was 0.8% for green-up and 23.2% for flowering date.
-\end{verbatim}
+Narrow-sense heritabilities (h2) indicated that rank-changing GxE for
+these phenotypes was present across the common gardens (Fig 1C). h2 were
+typically high at individual gardens: 59\% on average for green-up date,
+and 87\% for flowering date. However, h2 were variable across gardens,
+and green-up dates were uncorrelated (r2 \textless{} 0.2) or negatively
+correlated between pairs of gardens (Fig. 1B). These negative and small
+correlations undoubtedly contributed to the low h2 values for green-up
+and flowering date when estimated jointly at all eight gardens: h2 was
+0.8\% for green-up and 23.2\% for flowering date.
 
@@
 
-\begin{verbatim}
-We next explored how genetic variation in phenology covaried with environmental cues by using mash to jointly discover trait-associated SNPs across all common gardens for the Midwest, Gulf, and ‘Both’ subpopulations (Fig 1D, SI Appendix, Datasets 1-6).
-We first estimated SNP effects using GWAS on site-specific BLUPs for each phenotype.
-We then used mash on two subsets of these SNP effects, first to determine the covariance structures that significantly improved the mash model log likelihood when included, then to re-estimate the set of SNP effects with the largest effect estimates in univariate GWAS (hereafter “strong” effects) using the set of significant covariance matrices.
-\end{verbatim}
+We next explored how genetic variation in phenology covaried with
+environmental cues by using mash to jointly discover trait-associated
+SNPs across all common gardens for the Midwest, Gulf, and `Both'
+subpopulations (Fig 1D, SI Appendix, Datasets 1-6). We first estimated
+SNP effects using GWAS on site-specific BLUPs for each phenotype. We
+then used mash on two subsets of these SNP effects, first to determine
+the covariance structures that significantly improved the mash model log
+likelihood when included, then to re-estimate the set of SNP effects
+with the largest effect estimates in univariate GWAS (hereafter
+``strong'' effects) using the set of significant covariance matrices.
 
@@
 
+\section{Figures and Tables}\label{figures-and-tables}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{images/Fig1_200dpi.png}
+
+}
+
+\caption{Figure 1. Characterization of green-up and flowering dates from
+the switchgrass diversity panel. (a) Map and trait histograms of
+green-up and flowering dates across two genetically distinct switchgrass
+subpopulations and eight common gardens. Purple represents individuals
+from the Midwest genetic subpopulation, and pink individuals from the
+Gulf subpopulation. Vertical dashed lines indicate the summer solstice.
+Common gardens are arranged in latitudinal order. (b) Phenotypic
+correlations between clonal replicates planted at eight common gardens,
+within and between two genetic subpopulations.~ (c) Narrow sense
+heritability of green-up and flowering within single common gardens
+(purple) and across all eight common gardens (green), within and between
+two genetic subpopulations. (d) Flow diagram of the methods applied to
+the green-up and flowering dates to jointly estimate SNP effects across
+all sites. Mash was fit to SNP effect data and used to find covariance
+matrices that improved the mash model likelihood using a large set of
+randomly selected, relatively unlinked SNP effects; this model was
+applied to a "strong" set of SNP effects with large effect sizes in the
+univariate GWAS.}
+
+\end{figure}%
+
 \subsection*{Author Affiliations}\label{author-affiliations}
@@
 
-\subsection*{Figures and Tables}\label{figures-and-tables}
+\subsection*{Figures and Tables}\label{figures-and-tables-1}
 \addcontentsline{toc}{subsection}{Figures and Tables}
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3750}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3194}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3056}}@{}}
+\caption{Weather variables and time frames around the green up and
+flowering date used to construct the hypothesis-based covariance
+matrices. The correlations between values of these weather variables for
+genetically identical plants grown in different gardens were used to
+fill off-diagonal cells of the covariance matrices. Narrow-sense
+heritabilities for these values at each garden were used for the
+diagonal cells.}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Weather variable}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Time frame}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\textbf{\# Covariance Matrices tested in greedy mash algorithm}
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Weather variable}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Time frame}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\textbf{\# Covariance Matrices tested in greedy mash algorithm}
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+cumulative GDD at 12C in the time frame & 1, 2, 3, 4, 5, 6, 7, 14, 21,
+28 days prior to green up date~ & 60, 30 per phenotype and 10 per
+phenotype \& genetic group \\
+cumulative rainfall in the time frame & 1, 2, 3, 4, 5, 6, 7, 14, 21, 28
+days prior to green up date~ & 60, 30 per phenotype and 10 per phenotype
+\&~ genetic group \\
+day length (hours) on a specific day indicated by the time frame & 1, 2,
+3, 4, 5, 6, 7, and 14 days prior to green up date & 48, 24 per phenotype
+and 8 per phenotype \&~ genetic group \\
+day length change (seconds) on a specific day indicated by the time
+frame & 1, 2, 3, 4, 5, 6, 7, and 14 days prior to green up date & 48, 24
+per phenotype and 8 per phenotype \& genetic group \\
+average temperature in the time frame & 1, 2, 3, 4, 5, 6, 7, 14, 21, 28
+days prior to green up date & 60, 30 per phenotype and 10 per phenotype
+\& genetic group \\
+\end{longtable}
 

### 2023-12-28 manuscript/revision2/revision2.qmd
manuscript: revision 2 author and citation edits

@@
 ---
-title: Quarto Template for PNAS Submissions
+title: Switchgrass Phenology and Genotype-by-Environment Mapping
 format:
@@
     keep-tex: true  
+  docx:
+    highlight-style: github
+    fig-width: 7
+    fig-asp: 1.62
+    execute:
+      echo: false
+      cache: false
 author:
-  - name: Author One
+  - name: Alice H. MacQueen
     affiliations:
-      - name: Harvard University
+      - name: University of Texas at Austin
         id: a
-        department: Department of Government
-        city: Cambridge
-        state: MA
-        postal-code: 02138
-      - name: Harvard University
-        id: c
-        department: Department of Statistics
-        city: Cambridge
-        state: MA
-        postal-code: 02138
-    attributes:
-      equal-contributor: true
-  - name: Author Two
-    affiliations:
-      - name: Yale University
-        id: b
-        department: Department of Political Science
-        city: New Haven
-        state: CT
-        postal-code: 06511
+        department: Department of Integrative Biology
+        city: Austin
+        state: TX
+        postal-code: 78712
     email: alicem@uchicago.edu
@@
       corresponding: true
-  - name: Author Three
+  - name: Li Zhang
     affiliations:
       - ref: a
+    attributes:
+      equal-contributor: true
+  - name: Samuel A. Smith
+    affiliations:
+        - ref: a
+    attributes:
+      equal-contributor: true
+  - name: Jason Bonnette
+    affiliations:
+          - ref: a
+  - name: Arvid R. Boe
+    affiliations:
+      - name: South Dakota State University
+        id: b
+        department: Department of Agronomy
+        city: Brookings
+        state: SD
+        postal-code: 
+  - name: Phillip A. Fay
+    affiliations:
+      - name: USDA-ARS
+        id: c
+        department: Grassland, Soil and Water Research Laboratory
+        city: Temple
+        state: TX
+        postal-code: 
+  - name: Felix B. Fritschi
+    affiliations:
+      - name: University of Missouri
+        id: d
+        department: Division of Plant Sciences
+        city: Columbia
+        state: MO
+        postal-code: 
+  - name: David B. Lowry
+    affiliations:
+      - name: Michigan State University
+        id: e
+        department: Department of Plant Biology
+        city: East Lansing
+        state: MI
+        postal-code: 
+  - name: Robert B. Mitchell
+    affiliations:
+      - name: USDA-ARS
+        id: f
+        department: Wheat, Sorghum, and Forage Research Unit
+        city: Lincoln
+        state: NE
+        postal-code: 
+  - name: Francis M. Rouquette Jr
+    affiliations:
+      - name: Texas A&M University
+        id: g
+        department: Texas A&M AgriLife Research and Extension Center
+        city: Overton
+        state: TX
+        postal-code: 
+  - name: Yanqi Wu
+    affiliations:
+      - name: Oklahoma State University
+        id: h
+        department: Department of Plant and Soil Sciences
+        city: Stillwater
+        state: OK
+        postal-code: 
+  - name: Arbel Harpak
+    affiliations:
+      - ref: a
+  - name: Thomas E. Juenger
+    affiliations:
+      - ref: a
+    attributes:
+      corresponding: true
+    email: tjuenger@utexas.edu
 abstract: |
@@
 Most GxWeather effects belong to the latter category and do not covary with weather-based cues.
-However, >26% of Gulf population effects on flowering time covary with photoperiod cues, while >__% of Midwest upland population SNPs affecting flowering had effects that covaried with __ cues.
+However, >XX% of Gulf population effects on flowering time covary with photoperiod cues, while >XX% of Midwest upland population SNPs affecting flowering had effects that covaried with XX cues.
 An independent pseudo-F2 cross of Gulf and Midwest individuals mapped 23 additive QTLs for flowering at the same common gardens, all with significant mash associations and ten with enrichment of highly significant mash associations.
@@
 More broadly, this approach could be used to identify genetic marker-environment interactions in any species with related populations phenotyped in multiple environments.
-runningauthor: "One, Two, and Three"
+runningauthor: "MacQueen"
 significance: |
@@
 hosted field experiments.
-A.H.M.
-and L.Z.
+A.H.M., L.Z., and S.A.S.
 conducted statistical and computational analyses.
-The manuscript was written by A.H.M., L.Z., and T.E.J.
+The manuscript was written by A.H.M.
 with contributions from all authors."
-conflict_of_interest: "Please declare any competing interests here."
-equal-contributor: "A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable)."
+conflict_of_interest: "The authors declare no conflicts of interest."
+equal-contributor: "A.H.M. contributed equally to this work with L.Z. and S.A.S."
 keywords:  [allele-by-environment effect variation, antagonistic pleiotropy, photoperiod, cumulative growing degree days, genetic variation]
@@
 bibliography: bibliography.bib 
+csl: pnas.csl
 fontsize: "9pt"
@@
 
-\firstpage[20]{2}
+\firstpage[15]{2}
 
 The timing of plant vegetative and reproductive development are major components of plant fitness affected by multiple external environmental cues (e.g.
-degree of winter chilling, day length, temperature, and water availability) that signal existing or upcoming growing conditions (1--3).
+degree of winter chilling, day length, temperature, and water availability) that signal existing or upcoming growing conditions [@bauerle_photoperiodic_2012; @andres2012genetic; @korner2010phenology].
 Genetic responses to environmental cues determine the speed, timing, and energy apportioned to vegetative and reproductive growth and shape both the individual's lifespan and its lifetime production of viable seed.
 Day length (or photoperiod) is one of the most predictable environmental cues, and genetic sensitivity to photoperiod protects plants from potentially fatal consequences of phenological responses to temperature cues at the "wrong" time of year.
-However, the usefulness of specific environmental cues depends on both features of the environment and the species' adaptive strategies (4).
-Species with wide natural distributions can have multiple distinct environmentally-cued phenological responses: for example, populations of sunflower (Helianthus annuus) exhibit day-neutral, facultative short day, and facultative long-day flowering responses, which vary with their environments (5, 6).
+However, the usefulness of specific environmental cues depends on both features of the environment and the species' adaptive strategies @botero2015evolutionary.
+Species with wide natural distributions can have multiple distinct environmentally-cued phenological responses: for example, populations of sunflower (Helianthus annuus) exhibit day-neutral, facultative short day, and facultative long-day flowering responses, which vary with their environments [@blackman2013interacting; @henry2014transitions].
 Distinct genetic responses in different environments are known as genotype by environment interactions, or GxE.
 
-Flowering time, in particular, is a common subject of GxE research (5--11), a key output of selection driving adaptation to local environments (3, 12, 13), and a selection target for crop improvement to adapt crops to local or future environments (14).
-Changing flowering responsiveness to photoperiod cues has allowed geographic range expansion and increased yields in a number of cereal species (15--19) and other crops (20, 21).
-Recent statistical advances in studying phenological GxE have involved determining critical environmental indices before the phenological event occurs, such as photothermal time within a critical growth window (10).
+Flowering time, in particular, is a common subject of GxE research [@blackman2013interacting; @henry2014transitions; @aagren2017adaptive; @brachi2010linkage; @dittmar2014flowering; @li2018genomic; @romero2017study], a key output of selection driving adaptation to local environments [@andres2012genetic; @wadgymar2017identifying; @blumel2015flowering], and a selection target for crop improvement to adapt crops to local or future environments @jung2009flowering.
+Changing flowering responsiveness to photoperiod cues has allowed geographic range expansion and increased yields in a number of cereal species [@turner2005pseudo; @faure2012mutation; @hung2012zmcct; @zakhrabekova2012induced; @yang2013oself3] and other crops [@pin2012multifaceted; @weller2019parallel].
+Recent statistical advances in studying phenological GxE have involved determining critical environmental indices before the phenological event occurs, such as photothermal time within a critical growth window @li2018genomic.
 However, most studies of flowering GxE focus on finding a single, best fitting form of genotype-environment covariance, despite the key expectation that different genetic subpopulations, and even different genomic regions, have likely evolved distinct patterns of GxE.
-Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or sign-changing GxE, at the level of individual loci (22--25), previous work has found limited evidence of antagonistic pleiotropy (12, 26), but has been limited by a known statistical bias that reduced detection of SNP effects that differ in sign (26--28).
+Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or sign-changing GxE, at the level of individual loci [@levene1953genetic; @felsenstein1976theoretical; @kawecki2004conceptual; @hedrick1986genetic], previous work has found limited evidence of antagonistic pleiotropy [@des2013genotype; @wadgymar2017identifying], but has been limited by a known statistical bias that reduced detection of SNP effects that differ in sign [@des2013genotype; @anderson2013genetic; @anderson2011evolutionary].
 Thus, despite substantial interest in the frequencies of various forms of GxE, the prevalence of antagonistic pleiotropy relative to other forms of GxE remains unknown.
 
-Switchgrass (Panicum virgatum) is considered a short-day plant with reproductive development strongly linked to day of the year (29).
-However, as part of its wide environmental adaptation across the eastern half of North America, its photoperiodicity has been predicted to differ by plant latitude of origin (30, 31).
-We previously found divergent Midwest and Gulf genetic subpopulations of switchgrass which segregate for distinct sets of environmental adaptations, based on the analysis of two measures of fitness (i.e.,biomass and survival) among 732 genotypes (32).
-The Midwest genetic subpopulation is primarily composed of individuals from the well-studied upland switchgrass ecotype (33, 34), while the Gulf subpopulation has individuals from the lowland ecotype and the phenotypically intermediate coastal ecotype (32).
+Switchgrass (*Panicum virgatum*) is considered a short-day plant with reproductive development strongly linked to day of the year @mitchell1997predicting.
+However, as part of its wide environmental adaptation across the eastern half of North America, its photoperiodicity has been predicted to differ by plant latitude of origin [@parrish2005biology; @casler2004latitudinal].
+We previously found divergent Midwest and Gulf genetic subpopulations of switchgrass which segregate for distinct sets of environmental adaptations, based on the analysis of two measures of fitness (i.e.,biomass and survival) among 732 genotypes @lovell2021genomic.
+The Midwest genetic subpopulation is primarily composed of individuals from the well-studied upland switchgrass ecotype [@porter1966analysis; @milano2016genetic], while the Gulf subpopulation has individuals from the lowland ecotype and the phenotypically intermediate coastal ecotype @lovell2021genomic.
 
@@
 These gardens cover the majority of the latitudinal and climatic range of switchgrass and capture the most comprehensive picture to date of the environmental variation this species encounters.
-After single-site estimation of SNP effects for these phenotypes, we use multivariate adaptive shrinkage (mash) to jointly estimate SNP effects in each subpopulation and phenotype at all eight sites (35).
-Mash allows us to identify and specify multiple ways genetic effects linked to SNPs may covary with the environment, and does not have a statistical bias in detecting SNP effects with the same or opposite signs (39).
+After single-site estimation of SNP effects for these phenotypes, we use multivariate adaptive shrinkage (mash) to jointly estimate SNP effects in each subpopulation and phenotype at all eight sites @urbut2019flexible.
+Mash allows us to identify and specify multiple ways genetic effects linked to SNPs may covary with the environment, and does not have a statistical bias in detecting SNP effects with the same or opposite signs @10.1093/biostatistics/kxw041.
 To confirm our genetic mapping of GxE, we compare to mapping results from an outbred pseudo-F2 cross grown at the same sites.
@@
 
-In our diversity panel of tetraploid switchgrass (32), genotypes from the Gulf and Midwest genetic subpopulations had distinct phenological timings and distinct patterns of phenological correlations across our eight common garden sites (Fig.
-1A,B).
+In our diversity panel of tetraploid switchgrass [@lovell2021genomic], genotypes from the Gulf and Midwest genetic subpopulations had distinct phenological timings and distinct patterns of phenological correlations across our eight common garden sites (@fig-map).
 At the three Texas common gardens (hereafter 'Texas' gardens), located within the natural range of the Gulf subpopulation, Gulf green-up occurred before Midwestern green-up, and Gulf flowering occurred after Midwestern flowering (Fig.
@@
 
+# Figures and Tables
+
+```{=tex}
+\begin{table}[t!]
+\centering
+\caption{Weather variables and time frames around the green up and flowering date used to construct the hypothesis-based covariance matrices.
+The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices.
+Narrow-sense heritabilities for these values at each garden were used for the diagonal cells.}
+\begin{tabular}{lrr}
+Weather variable & Time Frame & Covariance Matrices tested \\
+\midrule
+1. cumulative GDD at 12C in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60 \\
+2. cumulative rainfall in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60  \\
+3. day length (hours) on a specific day indicated by the time frame & (1-7) and 14 days prior to green up date & 48 \\
+4. day length change (seconds) on a specific day indicated by the time frame & (1-7) and 14 days prior to green up date & 48 \\
+5. average temperature in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60 \\
+\bottomrule
+\end{tabular}
+
+\end{table}
+```
+![Figure 1.
+Characterization of green-up and flowering dates from the switchgrass diversity panel.
+(a) Map and trait histograms of green-up and flowering dates across two genetically distinct switchgrass subpopulations and eight common gardens.
+Purple represents individuals from the Midwest genetic subpopulation, and pink individuals from the Gulf subpopulation.
+Vertical dashed lines indicate the summer solstice.
+Common gardens are arranged in latitudinal order.
+(b) Phenotypic correlations between clonal replicates planted at eight common gardens, within and between two genetic subpopulations.
+(c) Narrow sense heritability of green-up and flowering within single common gardens (purple) and across all eight common gardens (green), within and between two genetic subpopulations.
+(d) Flow diagram of the methods applied to the green-up and flowering dates to jointly estimate SNP effects across all sites.
+Mash was fit to SNP effect data and used to find covariance matrices that improved the mash model likelihood using a large set of randomly selected, relatively unlinked SNP effects; this model was applied to a "strong" set of SNP effects with large effect sizes in the univariate GWAS.](images/Fig1_200dpi.png){#fig-map}
+
+![Figure 2.](images/Figure_2_Covariance_matrices_and_posterior_weights.png){#fig-covar}
+
 # Materials and Methods
@@
 
-# Figures and Tables
-
-![Figure 1.
-Characterization of green-up and flowering dates from the switchgrass diversity panel.
-(a) Map and trait histograms of green-up and flowering dates across two genetically distinct switchgrass subpopulations and eight common gardens.
-Purple represents individuals from the Midwest genetic subpopulation, and pink individuals from the Gulf subpopulation.
-Vertical dashed lines indicate the summer solstice.
-Common gardens are arranged in latitudinal order.
-(b) Phenotypic correlations between clonal replicates planted at eight common gardens, within and between two genetic subpopulations.  (c) Narrow sense heritability of green-up and flowering within single common gardens (purple) and across all eight common gardens (green), within and between two genetic subpopulations.
-(d) Flow diagram of the methods applied to the green-up and flowering dates to jointly estimate SNP effects across all sites.
-Mash was fit to SNP effect data and used to find covariance matrices that improved the mash model likelihood using a large set of randomly selected, relatively unlinked SNP effects; this model was applied to a \"strong\" set of SNP effects with large effect sizes in the univariate GWAS.](images/Fig1_200dpi.png)
-
-![Figure 2.](images/Figure_2_Covariance_matrices_and_posterior_weights.png)
-
-## Author Affiliations {.unnumbered}
-
-Include department, institution, and complete address, with the ZIP/postal code, for each author.
-Use lower case letters to match authors with institutions, as shown in the example.
-PNAS strongly encourages authors to supply an [ORCID identifier](https://orcid.org/) for each author.
-Individual authors must link their ORCID account to their PNAS account at [www.pnascentral.org](http://www.pnascentral.org/).
-For proper authentication, authors must provide their ORCID at submission and are not permitted to add ORCIDs on proofs.
-
-## Manuscript Length {.unnumbered}
-
-A standard 6-page article is approximately 4,000 words, 50 references, and 4 medium-size graphical elements (i.e., figures and tables).
-The preferred length of articles remains at 6 pages, but PNAS will allow articles up to a maximum of 12 pages.
-
-## References {.unnumbered}
-
-References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g.
-\cite{belkin2002using} and \cite{berard1994embedding,coifman2005geometric,phdthesis,masterthesis}.
-All references cited in the main text should be included in the main manuscript file.
-
-## Data Archival {.unnumbered}
-
-PNAS must be able to archive the data essential to a published article.
-Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the [Information for Authors](https://www.pnas.org/author-center/editorial-and-journal-policies#materials-and-data-availability), is acceptable.
-
-## Figures and Tables {.unnumbered}
-
-| **Weather variable**                                                      | **Time frame**                                               | **\# Covariance Matrices tested in greedy mash algorithm** |
-|---------------------------|-----------------------|----------------------|
-| cumulative GDD at 12C in the time frame                                   | 1, 2, 3, 4, 5, 6, 7, 14, 21, 28 days prior to green up date  | 60, 30 per phenotype and 10 per phenotype & genetic group  |
-| cumulative rainfall in the time frame                                     | 1, 2, 3, 4, 5, 6, 7, 14, 21, 28 days prior to green up date  | 60, 30 per phenotype and 10 per phenotype &  genetic group |
-| day length (hours) on a specific day indicated by the time frame          | 1, 2, 3, 4, 5, 6, 7, and 14 days prior to green up date      | 48, 24 per phenotype and 8 per phenotype &  genetic group  |
-| day length change (seconds) on a specific day indicated by the time frame | 1, 2, 3, 4, 5, 6, 7, and 14 days prior to green up date      | 48, 24 per phenotype and 8 per phenotype & genetic group   |
-| average temperature in the time frame                                     | 1, 2, 3, 4, 5, 6, 7, 14, 21, 28 days prior to green up date  | 60, 30 per phenotype and 10 per phenotype & genetic group  |
-
-: Weather variables and time frames around the green up and flowering date used to construct the hypothesis-based covariance matrices.
-The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices.
-Narrow-sense heritabilities for these values at each garden were used for the diagonal cells.
-
-Figures can be included using Quarto's syntax, code chunks, or with LaTeX.
-
-For simple figures, Quarto syntax is typically easiest. Full details are available in [the Quarto documentation](https://quarto.org/docs/authoring/figures.html).
-
-![Placeholder image of a frog with a long example legend to show justification setting.](figs/frog.pdf){#fig-frog}
-
-Tables can similarly be included using Quarto's syntax, code chunks, or with LaTeX.
-I tend to use code chunks, but LaTeX works very well for classic tables, like regression tables or summary statistics.
-
-```{=tex}
-\begin{table}[t!]
-\centering
-\caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
-\begin{tabular}{lrrr}
-Species & CBS & CV & G3 \\
-\midrule
-1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
-2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
-3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
-\bottomrule
-\end{tabular}
-
-\addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
-\end{table}
-```
-### Digital Figures {.unnumbered}
-
-EPS, high-resolution PDF, and PowerPoint are preferred formats for figures that will be used in the main manuscript.
-Authors may submit PRC or U3D files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.
-Color images must be in RGB (red, green, blue) mode.
-Include the font files for any text.
-
-Images must be provided at final size, preferably 1 column width (8.7cm).
-Figures wider than 1 column should be sized to 11.4cm or 17.8cm wide.
-Numbers, letters, and symbols should be no smaller than 6 points (2mm) and no larger than 12 points (6mm) after reduction and must be consistent.
-
-Figures and tables should be labelled and referenced in the standard way using the `\label{}` and `\ref{}` commands.
-
-Figure @fig-frog shows an example of how to insert a column-wide figure.
-To insert a figure wider than one column, please use the `figure*` environment.
-Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide.
-Use `SCfigure*` environment for a wide figure with side legends, such as @fig-side.
-
-In Quarto, we can do these by setting the `fig-env` command to `figure*` or `SCfigure*`
-
-```{r}
-#| label: fig-side
-#| fig-cap: "This legend would be placed at the side of the figure, rather than below it."
-#| fig-env: "SCfigure*"
-#| echo: false
-# tell it the options as comments with a | and a space, as above.
-# set echo: false to avoid printing this text
-knitr::include_graphics('figs/frog.pdf')
-```
-
-## Tables {.unnumbered}
-
-Tables should be included in the main manuscript file and should not be uploaded separately.
-
-## Single column equations {.unnumbered}
-
-Authors may use 1- or 2-column equations in their article, according to their preference.
-
-To allow an equation to span both columns, use the `\begin{figure*}...\end{figure*}` environment mentioned above for figures.
-
-Note that the use of the `widetext` environment for equations is not recommended, and should not be used.
-
-```{=tex}
-\begin{figure*}[bt!]
-\begin{align*}
-(x+y)^3&=(x+y)(x+y)^2\\
-       &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
-       &=x^3+3x^2y+3xy^3+x^3.
-\end{align*}
-\end{figure*}
-```
-# Material and Methods {.unnumbered}
-
-## Supporting Information Appendix (SI) {.unnumbered}
-
-Authors should submit SI as a single separate SI Appendix PDF file, combining all text, figures, tables, movie legends, and SI references.
-SI will be published as provided by the authors; it will not be edited or composed.
-Additional details can be found in the [PNAS Author Center](https://www.pnas.org/authors/submitting-your-manuscript#manuscript-formatting-guidelines).
-The PNAS Overleaf SI template can be found [here](https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd).
-Refer to the SI Appendix in the manuscript at an appropriate point in the text.
-Number supporting figures and tables starting with S1, S2, etc.
-
-Authors who place detailed materials and methods in an SI Appendix must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the SI methods.
-If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.
-
-### SI Datasets {.unnumbered}
-
-Supply .xlsx, .csv, .txt, .rtf, or .pdf files. This file type will be published in raw format and will not be edited or composed.
-
-### SI Movies {.unnumbered}
-
-Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files.
-Movie legends should be included in the SI Appendix file.
-All movies should be submitted at the desired reproduction size and length.
-Movies should be no more than 10MB in size.
-
 # References

### 2023-12-28 manuscript/revision2/revision2.tex
manuscript: revision 2 author and citation edits

@@
 \fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\usepackage{bookmark}
+
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 \hypersetup{
-  pdftitle={Quarto Template for PNAS Submissions},
-  pdfauthor={Author One; Author Two; Author Three},
+  pdftitle={Switchgrass Phenology and Genotype-by-Environment Mapping},
+  pdfauthor={Alice H.
+MacQueen; Li Zhang; Samuel A.
+Smith; Jason Bonnette; Arvid R.
+Boe; Phillip A.
+Fay; Felix B.
+Fritschi; David B.
+Lowry; Robert B.
+Mitchell; Francis M.
+Rouquette Jr; Yanqi Wu; Arbel Harpak; Thomas E.
+Juenger},
   pdfkeywords={allele-by-environment effect variation, antagonistic
@@
 
-\title{Quarto Template for PNAS Submissions}
+\title{Switchgrass Phenology and Genotype-by-Environment Mapping}
 
@@
 
-\author[a,c%
-,\equalcont%
-%
-]{Author One}
-\author[b%
+\author[a%
 ,\equalcont%
 ,\correspond%
-]{Author Two}
+]{Alice H. MacQueen}
 \author[a%
-%
-%
-]{Author Three}
-
-\affil[a]{Harvard University, Department of Government, Cambridge, 2138}
-\affil[c]{Harvard University, Department of Statistics, Cambridge, 2138}
-\affil[b]{Yale University, Department of Political Science, New
-Haven, 3401}
-
-\leadauthor{One, Two, and Three}
+,\equalcont%
+%
+]{Li Zhang}
+\author[a%
+,\equalcont%
+%
+]{Samuel A. Smith}
+\author[a%
+%
+%
+]{Jason Bonnette}
+\author[b%
+%
+%
+]{Arvid R. Boe}
+\author[c%
+%
+%
+]{Phillip A. Fay}
+\author[d%
+%
+%
+]{Felix B. Fritschi}
+\author[e%
+%
+%
+]{David B. Lowry}
+\author[f%
+%
+%
+]{Robert B. Mitchell}
+\author[g%
+%
+%
+]{Francis M. Rouquette Jr}
+\author[h%
+%
+%
+]{Yanqi Wu}
+\author[a%
+%
+%
+]{Arbel Harpak}
+\author[a%
+%
+,\correspond%
+]{Thomas E. Juenger}
+
+\affil[a]{University of Texas at Austin, Department of Integrative
+Biology, Austin, 78712}
+\affil[b]{South Dakota State University, Department of
+Agronomy, Brookings}
+\affil[c]{USDA-ARS, Grassland, Soil and Water Research
+Laboratory, Temple}
+\affil[d]{University of Missouri, Division of Plant Sciences, Columbia}
+\affil[e]{Michigan State University, Department of Plant Biology, East
+Lansing}
+\affil[f]{USDA-ARS, Wheat, Sorghum, and Forage Research Unit, Lincoln}
+\affil[g]{Texas A\&M University, Texas A\&M AgriLife Research and
+Extension Center, Overton}
+\affil[h]{Oklahoma State University, Department of Plant and Soil
+Sciences, Stillwater}
+
+\leadauthor{MacQueen}
 
@@
 field experiments. A.R.B., P.A.F., F.B.F., D.B.L., R.B.M., F.M.R., Y.W.,
-and T.E.J. hosted field experiments. A.H.M. and L.Z. conducted
+and T.E.J. hosted field experiments. A.H.M., L.Z., and S.A.S. conducted
 statistical and computational analyses. The manuscript was written by
-A.H.M., L.Z., and T.E.J. with contributions from all authors.}
-
-\authordeclaration{Please declare any competing interests here.}
-
-\equalauthors{\textsuperscript{\equalcont} A.O.(Author One) contributed
-equally to this work with A.T. (Author Two) (remove if not applicable).}
+A.H.M. with contributions from all authors.}
+
+\authordeclaration{The authors declare no conflicts of interest.}
+
+\equalauthors{\textsuperscript{\equalcont} A.H.M. contributed equally to
+this work with L.Z. and S.A.S.}
 
 \correspondingauthor{\textsuperscript{\correspond}To whom correspondence should be addressed. E-mail: alicem@uchicago.edu}
+\correspondingauthor{\textsuperscript{\correspond}To whom correspondence should be addressed. E-mail: tjuenger@utexas.edu}
 
@@
 effects belong to the latter category and do not covary with
-weather-based cues. However, \textgreater26\% of Gulf population effects
-on flowering time covary with photoperiod cues, while
-\textgreater{}\textbf{\% of Midwest upland population SNPs affecting
-flowering had effects that covaried with } cues. An independent
-pseudo-F2 cross of Gulf and Midwest individuals mapped 23 additive QTLs
-for flowering at the same common gardens, all with significant mash
-associations and ten with enrichment of highly significant mash
-associations. We demonstrate that we can identify QTL with GxE and
-assign them to specific weather-based cues and/or data-driven patterns.
-Breeding for particular alleles at these loci could change flowering
-responsiveness to photoperiod cues in switchgrass. More broadly, this
-approach could be used to identify genetic marker-environment
-interactions in any species with related populations phenotyped in
-multiple environments.
+weather-based cues. However, \textgreater XX\% of Gulf population
+effects on flowering time covary with photoperiod cues, while
+\textgreater XX\% of Midwest upland population SNPs affecting flowering
+had effects that covaried with XX cues. An independent pseudo-F2 cross
+of Gulf and Midwest individuals mapped 23 additive QTLs for flowering at
+the same common gardens, all with significant mash associations and ten
+with enrichment of highly significant mash associations. We demonstrate
+that we can identify QTL with GxE and assign them to specific
+weather-based cues and/or data-driven patterns. Breeding for particular
+alleles at these loci could change flowering responsiveness to
+photoperiod cues in switchgrass. More broadly, this approach could be
+used to identify genetic marker-environment interactions in any species
+with related populations phenotyped in multiple environments.
 \end{abstract}
@@
 \thispagestyle{firststyle}
-\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}\firstpage[20]{2}
+\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
+\firstpage[15]{2}
 
@@
 (5--11), a key output of selection driving adaptation to local
-environments (3, 12, 13), and a selection target for crop improvement to
+environments (2, 12, 13), and a selection target for crop improvement to
 adapt crops to local or future environments (14). Changing flowering
@@
 
-Switchgrass (Panicum virgatum) is considered a short-day plant with
-reproductive development strongly linked to day of the year (29).
+Switchgrass (\emph{Panicum virgatum}) is considered a short-day plant
+with reproductive development strongly linked to day of the year (29).
 However, as part of its wide environmental adaptation across the eastern
@@
 may covary with the environment, and does not have a statistical bias in
-detecting SNP effects with the same or opposite signs (39). To confirm
+detecting SNP effects with the same or opposite signs (36). To confirm
 our genetic mapping of GxE, we compare to mapping results from an
@@
 timings and distinct patterns of phenological correlations across our
-eight common garden sites (Fig. 1A,B). At the three Texas common gardens
-(hereafter `Texas' gardens), located within the natural range of the
-Gulf subpopulation, Gulf green-up occurred before Midwestern green-up,
-and Gulf flowering occurred after Midwestern flowering (Fig. 1A). At the
-four northernmost common gardens (hereafter `North' gardens), located
-within the natural range of the Midwest subpopulation, both Gulf
-green-up and flowering occurred after Midwest green-up and flowering. At
-the Oklahoma common garden, located near the natural range limits of
-both the Gulf and the Midwest subpopulations, Gulf and Midwest green-up
-occurred over the same time period. These patterns led to strong
-negative phenotypic correlations for green-up between the North and
-Texas gardens, particularly in the Gulf and Both subpopulations, and
-contributed to positive phenotypic correlations for flowering time of
-larger magnitude at more northern gardens (Fig. 1B).
+eight common garden sites (Figure~\ref{fig-map}). At the three Texas
+common gardens (hereafter `Texas' gardens), located within the natural
+range of the Gulf subpopulation, Gulf green-up occurred before
+Midwestern green-up, and Gulf flowering occurred after Midwestern
+flowering (Fig. 1A). At the four northernmost common gardens (hereafter
+`North' gardens), located within the natural range of the Midwest
+subpopulation, both Gulf green-up and flowering occurred after Midwest
+green-up and flowering. At the Oklahoma common garden, located near the
+natural range limits of both the Gulf and the Midwest subpopulations,
+Gulf and Midwest green-up occurred over the same time period. These
+patterns led to strong negative phenotypic correlations for green-up
+between the North and Texas gardens, particularly in the Gulf and Both
+subpopulations, and contributed to positive phenotypic correlations for
+flowering time of larger magnitude at more northern gardens (Fig. 1B).
 
@@
 
+\section{Figures and Tables}\label{figures-and-tables}
+
+\begin{table}[t!]
+\centering
+\caption{Weather variables and time frames around the green up and flowering date used to construct the hypothesis-based covariance matrices.
+The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices.
+Narrow-sense heritabilities for these values at each garden were used for the diagonal cells.}
+\begin{tabular}{lrr}
+Weather variable & Time Frame & Covariance Matrices tested \\
+\midrule
+1. cumulative GDD at 12C in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60 \\
+2. cumulative rainfall in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60  \\
+3. day length (hours) on a specific day indicated by the time frame & (1-7) and 14 days prior to green up date & 48 \\
+4. day length change (seconds) on a specific day indicated by the time frame & (1-7) and 14 days prior to green up date & 48 \\
+5. average temperature in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60 \\
+\bottomrule
+\end{tabular}
+
+\end{table}
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{images/Fig1_200dpi.png}
+
+}
+
+\caption{\label{fig-map}Figure 1. Characterization of green-up and
+flowering dates from the switchgrass diversity panel. (a) Map and trait
+histograms of green-up and flowering dates across two genetically
+distinct switchgrass subpopulations and eight common gardens. Purple
+represents individuals from the Midwest genetic subpopulation, and pink
+individuals from the Gulf subpopulation. Vertical dashed lines indicate
+the summer solstice. Common gardens are arranged in latitudinal order.
+(b) Phenotypic correlations between clonal replicates planted at eight
+common gardens, within and between two genetic subpopulations. (c)
+Narrow sense heritability of green-up and flowering within single common
+gardens (purple) and across all eight common gardens (green), within and
+between two genetic subpopulations. (d) Flow diagram of the methods
+applied to the green-up and flowering dates to jointly estimate SNP
+effects across all sites. Mash was fit to SNP effect data and used to
+find covariance matrices that improved the mash model likelihood using a
+large set of randomly selected, relatively unlinked SNP effects; this
+model was applied to a ``strong'' set of SNP effects with large effect
+sizes in the univariate GWAS.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{images/Figure_2_Covariance_matrices_and_posterior_weights.png}
+
+}
+
+\caption{\label{fig-covar}Figure 2.}
+
+\end{figure}%
+
 \section{Materials and Methods}\label{materials-and-methods}
@@
 
-\section{Figures and Tables}\label{figures-and-tables}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{images/Fig1_200dpi.png}
-
-}
-
-\caption{Figure 1. Characterization of green-up and flowering dates from
-the switchgrass diversity panel. (a) Map and trait histograms of
-green-up and flowering dates across two genetically distinct switchgrass
-subpopulations and eight common gardens. Purple represents individuals
-from the Midwest genetic subpopulation, and pink individuals from the
-Gulf subpopulation. Vertical dashed lines indicate the summer solstice.
-Common gardens are arranged in latitudinal order. (b) Phenotypic
-correlations between clonal replicates planted at eight common gardens,
-within and between two genetic subpopulations.~ (c) Narrow sense
-heritability of green-up and flowering within single common gardens
-(purple) and across all eight common gardens (green), within and between
-two genetic subpopulations. (d) Flow diagram of the methods applied to
-the green-up and flowering dates to jointly estimate SNP effects across
-all sites. Mash was fit to SNP effect data and used to find covariance
-matrices that improved the mash model likelihood using a large set of
-randomly selected, relatively unlinked SNP effects; this model was
-applied to a "strong" set of SNP effects with large effect sizes in the
-univariate GWAS.}
-
-\end{figure}%
-
-\subsection*{Author Affiliations}\label{author-affiliations}
-\addcontentsline{toc}{subsection}{Author Affiliations}
-
-Include department, institution, and complete address, with the
-ZIP/postal code, for each author. Use lower case letters to match
-authors with institutions, as shown in the example. PNAS strongly
-encourages authors to supply an \href{https://orcid.org/}{ORCID
-identifier} for each author. Individual authors must link their ORCID
-account to their PNAS account at
-\href{http://www.pnascentral.org/}{www.pnascentral.org}. For proper
-authentication, authors must provide their ORCID at submission and are
-not permitted to add ORCIDs on proofs.
-
-\subsection*{Manuscript Length}\label{manuscript-length}
-\addcontentsline{toc}{subsection}{Manuscript Length}
-
-A standard 6-page article is approximately 4,000 words, 50 references,
-and 4 medium-size graphical elements (i.e., figures and tables). The
-preferred length of articles remains at 6 pages, but PNAS will allow
-articles up to a maximum of 12 pages.
-
-\subsection*{References}\label{references}
-\addcontentsline{toc}{subsection}{References}
-
-References should be cited in numerical order as they appear in text;
-this will be done automatically via bibtex, e.g.~\cite{belkin2002using}
-and
-\cite{berard1994embedding,coifman2005geometric,phdthesis,masterthesis}.
-All references cited in the main text should be included in the main
-manuscript file.
-
-\subsection*{Data Archival}\label{data-archival}
-\addcontentsline{toc}{subsection}{Data Archival}
-
-PNAS must be able to archive the data essential to a published article.
-Where such archiving is not possible, deposition of data in public
-databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata,
-and others outlined in the
-\href{https://www.pnas.org/author-center/editorial-and-journal-policies\#materials-and-data-availability}{Information
-for Authors}, is acceptable.
-
-\subsection*{Figures and Tables}\label{figures-and-tables-1}
-\addcontentsline{toc}{subsection}{Figures and Tables}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3750}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3194}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3056}}@{}}
-\caption{Weather variables and time frames around the green up and
-flowering date used to construct the hypothesis-based covariance
-matrices. The correlations between values of these weather variables for
-genetically identical plants grown in different gardens were used to
-fill off-diagonal cells of the covariance matrices. Narrow-sense
-heritabilities for these values at each garden were used for the
-diagonal cells.}\tabularnewline
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Weather variable}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Time frame}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{\# Covariance Matrices tested in greedy mash algorithm}
-\end{minipage} \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Weather variable}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Time frame}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{\# Covariance Matrices tested in greedy mash algorithm}
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-cumulative GDD at 12C in the time frame & 1, 2, 3, 4, 5, 6, 7, 14, 21,
-28 days prior to green up date~ & 60, 30 per phenotype and 10 per
-phenotype \& genetic group \\
-cumulative rainfall in the time frame & 1, 2, 3, 4, 5, 6, 7, 14, 21, 28
-days prior to green up date~ & 60, 30 per phenotype and 10 per phenotype
-\&~ genetic group \\
-day length (hours) on a specific day indicated by the time frame & 1, 2,
-3, 4, 5, 6, 7, and 14 days prior to green up date & 48, 24 per phenotype
-and 8 per phenotype \&~ genetic group \\
-day length change (seconds) on a specific day indicated by the time
-frame & 1, 2, 3, 4, 5, 6, 7, and 14 days prior to green up date & 48, 24
-per phenotype and 8 per phenotype \& genetic group \\
-average temperature in the time frame & 1, 2, 3, 4, 5, 6, 7, 14, 21, 28
-days prior to green up date & 60, 30 per phenotype and 10 per phenotype
-\& genetic group \\
-\end{longtable}
-
-Figures can be included using Quarto's syntax, code chunks, or with
-LaTeX.
-
-For simple figures, Quarto syntax is typically easiest. Full details are
-available in \href{https://quarto.org/docs/authoring/figures.html}{the
-Quarto documentation}.
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{figs/frog.pdf}
-
-}
-
-\caption{\label{fig-frog}Placeholder image of a frog with a long example
-legend to show justification setting.}
-
-\end{figure}%
-
-Tables can similarly be included using Quarto's syntax, code chunks, or
-with LaTeX. I tend to use code chunks, but LaTeX works very well for
-classic tables, like regression tables or summary statistics.
-
-\begin{table}[t!]
-\centering
-\caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
-\begin{tabular}{lrrr}
-Species & CBS & CV & G3 \\
-\midrule
-1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
-2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
-3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
-\bottomrule
-\end{tabular}
-
-\addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
-\end{table}
-
-\subsubsection*{Digital Figures}\label{digital-figures}
-\addcontentsline{toc}{subsubsection}{Digital Figures}
-
-EPS, high-resolution PDF, and PowerPoint are preferred formats for
-figures that will be used in the main manuscript. Authors may submit PRC
-or U3D files for 3D images; these must be accompanied by 2D
-representations in TIFF, EPS, or high-resolution PDF format. Color
-images must be in RGB (red, green, blue) mode. Include the font files
-for any text.
-
-Images must be provided at final size, preferably 1 column width
-(8.7cm). Figures wider than 1 column should be sized to 11.4cm or 17.8cm
-wide. Numbers, letters, and symbols should be no smaller than 6 points
-(2mm) and no larger than 12 points (6mm) after reduction and must be
-consistent.
-
-Figures and tables should be labelled and referenced in the standard way
-using the \texttt{\textbackslash{}label\{\}} and
-\texttt{\textbackslash{}ref\{\}} commands.
-
-Figure Figure~\ref{fig-frog} shows an example of how to insert a
-column-wide figure. To insert a figure wider than one column, please use
-the \texttt{figure*} environment. Figures wider than one column should
-be sized to 11.4 cm or 17.8 cm wide. Use \texttt{SCfigure*} environment
-for a wide figure with side legends, such as Figure~\ref{fig-side}.
-
-In Quarto, we can do these by setting the \texttt{fig-env} command to
-\texttt{figure*} or \texttt{SCfigure*}
-
-\begin{SCfigure*}
-
-\centering{
-
-\includegraphics{figs/frog.pdf}
-
-}
-
-\caption{\label{fig-side}This legend would be placed at the side of the
-figure, rather than below it.}
-
-\end{SCfigure*}%
-
-\subsection*{Tables}\label{tables}
-\addcontentsline{toc}{subsection}{Tables}
-
-Tables should be included in the main manuscript file and should not be
-uploaded separately.
-
-\subsection*{Single column equations}\label{single-column-equations}
-\addcontentsline{toc}{subsection}{Single column equations}
-
-Authors may use 1- or 2-column equations in their article, according to
-their preference.
-
-To allow an equation to span both columns, use the
-\texttt{\textbackslash{}begin\{figure*\}...\textbackslash{}end\{figure*\}}
-environment mentioned above for figures.
-
-Note that the use of the \texttt{widetext} environment for equations is
-not recommended, and should not be used.
-
-\begin{figure*}[bt!]
-\begin{align*}
-(x+y)^3&=(x+y)(x+y)^2\\
-       &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
-       &=x^3+3x^2y+3xy^3+x^3.
-\end{align*}
-\end{figure*}
-
-\section*{Material and Methods}\label{material-and-methods}
-\addcontentsline{toc}{section}{Material and Methods}
-
-\subsection*{Supporting Information Appendix
-(SI)}\label{supporting-information-appendix-si}
-\addcontentsline{toc}{subsection}{Supporting Information Appendix (SI)}
-
-Authors should submit SI as a single separate SI Appendix PDF file,
-combining all text, figures, tables, movie legends, and SI references.
-SI will be published as provided by the authors; it will not be edited
-or composed. Additional details can be found in the
-\href{https://www.pnas.org/authors/submitting-your-manuscript\#manuscript-formatting-guidelines}{PNAS
-Author Center}. The PNAS Overleaf SI template can be found
-\href{https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd}{here}.
-Refer to the SI Appendix in the manuscript at an appropriate point in
-the text. Number supporting figures and tables starting with S1, S2,
-etc.
-
-Authors who place detailed materials and methods in an SI Appendix must
-provide sufficient detail in the main text methods to enable a reader to
-follow the logic of the procedures and results and also must reference
-the SI methods. If a paper is fundamentally a study of a new method or
-technique, then the methods must be described completely in the main
-text.
-
-\subsubsection*{SI Datasets}\label{si-datasets}
-\addcontentsline{toc}{subsubsection}{SI Datasets}
-
-Supply .xlsx, .csv, .txt, .rtf, or .pdf files. This file type will be
-published in raw format and will not be edited or composed.
-
-\subsubsection*{SI Movies}\label{si-movies}
-\addcontentsline{toc}{subsubsection}{SI Movies}
-
-Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media
-(wmv), animated GIF (gif), or MPEG files. Movie legends should be
-included in the SI Appendix file. All movies should be submitted at the
-desired reproduction size and length. Movies should be no more than 10MB
-in size.
-
-\section{References}\label{references-1}
+\section{References}\label{references}
 
@@
 \begin{CSLReferences}{0}{1}
+\bibitem[\citeproctext]{ref-bauerle_photoperiodic_2012}
+\CSLLeftMargin{1. }%
+\CSLRightInline{W. L. Bauerle, \emph{et al.},
+\href{https://doi.org/10.1073/pnas.1119131109}{Photoperiodic regulation
+of the seasonal pattern of photosynthetic capacity and the implications
+for carbon cycling}. \emph{Proceedings of the National Academy of
+Sciences} \textbf{109}, 8612--8617 (2012).}
+
+\bibitem[\citeproctext]{ref-andres2012genetic}
+\CSLLeftMargin{2. }%
+\CSLRightInline{F. Andrés, G. Coupland, The genetic basis of flowering
+responses to seasonal cues. \emph{Nature Reviews Genetics} \textbf{13},
+627--639 (2012).}
+
+\bibitem[\citeproctext]{ref-korner2010phenology}
+\CSLLeftMargin{3. }%
+\CSLRightInline{C. Körner, D. Basler, Phenology under global warming.
+\emph{Science} \textbf{327}, 1461--1462 (2010).}
+
+\bibitem[\citeproctext]{ref-botero2015evolutionary}
+\CSLLeftMargin{4. }%
+\CSLRightInline{C. A. Botero, F. J. Weissing, J. Wright, D. R.
+Rubenstein, Evolutionary tipping points in the capacity to adapt to
+environmental change. \emph{Proceedings of the National Academy of
+Sciences} \textbf{112}, 184--189 (2015).}
+
+\bibitem[\citeproctext]{ref-blackman2013interacting}
+\CSLLeftMargin{5. }%
+\CSLRightInline{B. K. Blackman, Interacting duplications, fluctuating
+selection, and convergence: The complex dynamics of flowering time
+evolution during sunflower domestication. \emph{Journal of experimental
+botany} \textbf{64}, 421--431 (2013).}
+
+\bibitem[\citeproctext]{ref-henry2014transitions}
+\CSLLeftMargin{6. }%
+\CSLRightInline{L. P. Henry, R. H. Watson, B. K. Blackman, Transitions
+in photoperiodic flowering are common and involve few loci in wild
+sunflowers (helianthus; asteraceae). \emph{American Journal of Botany}
+\textbf{101}, 1748--1758 (2014).}
+
+\bibitem[\citeproctext]{ref-aagren2017adaptive}
+\CSLLeftMargin{7. }%
+\CSLRightInline{J. Ågren, C. G. Oakley, S. Lundemo, D. W. Schemske,
+Adaptive divergence in flowering time among natural populations of
+arabidopsis thaliana: Estimates of selection and QTL mapping.
+\emph{Evolution} \textbf{71}, 550--564 (2017).}
+
+\bibitem[\citeproctext]{ref-brachi2010linkage}
+\CSLLeftMargin{8. }%
+\CSLRightInline{B. Brachi, \emph{et al.}, Linkage and association
+mapping of arabidopsis thaliana flowering time in nature. \emph{PLoS
+genetics} \textbf{6}, e1000940 (2010).}
+
+\bibitem[\citeproctext]{ref-dittmar2014flowering}
+\CSLLeftMargin{9. }%
+\CSLRightInline{E. L. Dittmar, C. G. Oakley, J. Ågren, D. W. Schemske,
+Flowering time QTL in natural populations of arabidopsis thaliana and
+implications for their adaptive value. \emph{Molecular ecology}
+\textbf{23}, 4291--4303 (2014).}
+
+\bibitem[\citeproctext]{ref-li2018genomic}
+\CSLLeftMargin{10. }%
+\CSLRightInline{X. Li, T. Guo, Q. Mu, X. Li, J. Yu, Genomic and
+environmental determinants and their interplay underlying phenotypic
+plasticity. \emph{Proceedings of the National Academy of Sciences}
+\textbf{115}, 6679--6684 (2018).}
+
+\bibitem[\citeproctext]{ref-romero2017study}
+\CSLLeftMargin{11. }%
+\CSLRightInline{J. A. Romero Navarro, \emph{et al.}, A study of allelic
+diversity underlying flowering-time adaptation in maize landraces.
+\emph{Nature genetics} \textbf{49}, 476--480 (2017).}
+
+\bibitem[\citeproctext]{ref-wadgymar2017identifying}
+\CSLLeftMargin{12. }%
+\CSLRightInline{S. M. Wadgymar, \emph{et al.}, Identifying targets and
+agents of selection: Innovative methods to evaluate the processes that
+contribute to local adaptation. \emph{Methods in Ecology and Evolution}
+\textbf{8}, 738--749 (2017).}
+
+\bibitem[\citeproctext]{ref-blumel2015flowering}
+\CSLLeftMargin{13. }%
+\CSLRightInline{M. Blümel, N. Dally, C. Jung, Flowering time regulation
+in crops---what did we learn from arabidopsis? \emph{Current opinion in
+biotechnology} \textbf{32}, 121--129 (2015).}
+
+\bibitem[\citeproctext]{ref-jung2009flowering}
+\CSLLeftMargin{14. }%
+\CSLRightInline{C. Jung, A. E. Müller, Flowering time control and
+applications in plant breeding. \emph{Trends in plant science}
+\textbf{14}, 563--573 (2009).}
+
+\bibitem[\citeproctext]{ref-turner2005pseudo}
+\CSLLeftMargin{15. }%
+\CSLRightInline{A. Turner, J. Beales, S. Faure, R. P. Dunford, D. A.
+Laurie, The pseudo-response regulator ppd-H1 provides adaptation to
+photoperiod in barley. \emph{Science} \textbf{310}, 1031--1034 (2005).}
+
+\bibitem[\citeproctext]{ref-faure2012mutation}
+\CSLLeftMargin{16. }%
+\CSLRightInline{S. Faure, \emph{et al.}, Mutation at the circadian clock
+gene EARLY MATURITY 8 adapts domesticated barley (hordeum vulgare) to
+short growing seasons. \emph{Proceedings of the National Academy of
+Sciences} \textbf{109}, 8328--8333 (2012).}
+
+\bibitem[\citeproctext]{ref-hung2012zmcct}
+\CSLLeftMargin{17. }%
+\CSLRightInline{H.-Y. Hung, \emph{et al.}, ZmCCT and the genetic basis
+of day-length adaptation underlying the postdomestication spread of
+maize. \emph{Proceedings of the National Academy of Sciences}
+\textbf{109}, E1913--E1921 (2012).}
+
+\bibitem[\citeproctext]{ref-zakhrabekova2012induced}
+\CSLLeftMargin{18. }%
+\CSLRightInline{S. Zakhrabekova, \emph{et al.}, Induced mutations in
+circadian clock regulator mat-a facilitated short-season adaptation and
+range extension in cultivated barley. \emph{Proceedings of the National
+Academy of Sciences} \textbf{109}, 4326--4331 (2012).}
+
+\bibitem[\citeproctext]{ref-yang2013oself3}
+\CSLLeftMargin{19. }%
+\CSLRightInline{Y. Yang, Q. Peng, G.-X. Chen, X.-H. Li, C.-Y. Wu, OsELF3
+is involved in circadian clock regulation for promoting flowering under
+long-day conditions in rice. \emph{Molecular Plant} \textbf{6}, 202--215
+(2013).}
+
+\bibitem[\citeproctext]{ref-pin2012multifaceted}
+\CSLLeftMargin{20. }%
+\CSLRightInline{P. Pin, O. Nilsson, The multifaceted roles of FLOWERING
+LOCUS t in plant development. \emph{Plant, cell \& environment}
+\textbf{35}, 1742--1755 (2012).}
+
+\bibitem[\citeproctext]{ref-weller2019parallel}
+\CSLLeftMargin{21. }%
+\CSLRightInline{J. L. Weller, \emph{et al.}, Parallel origins of
+photoperiod adaptation following dual domestications of common bean.
+\emph{Journal of Experimental Botany} \textbf{70}, 1209--1219 (2019).}
+
+\bibitem[\citeproctext]{ref-levene1953genetic}
+\CSLLeftMargin{22. }%
+\CSLRightInline{H. Levene, Genetic equilibrium when more than one
+ecological niche is available. \emph{The American Naturalist}
+\textbf{87}, 331--333 (1953).}
+
+\bibitem[\citeproctext]{ref-felsenstein1976theoretical}
+\CSLLeftMargin{23. }%
+\CSLRightInline{J. Felsenstein, The theoretical population genetics of
+variable selection and migration. \emph{Annual review of genetics}
+\textbf{10}, 253--280 (1976).}
+
+\bibitem[\citeproctext]{ref-kawecki2004conceptual}
+\CSLLeftMargin{24. }%
+\CSLRightInline{T. J. Kawecki, D. Ebert, Conceptual issues in local
+adaptation. \emph{Ecology letters} \textbf{7}, 1225--1241 (2004).}
+
+\bibitem[\citeproctext]{ref-hedrick1986genetic}
+\CSLLeftMargin{25. }%
+\CSLRightInline{P. W. Hedrick, Genetic polymorphism in heterogeneous
+environments: A decade later. \emph{Annual review of ecology and
+systematics} \textbf{17}, 535--566 (1986).}
+
+\bibitem[\citeproctext]{ref-des2013genotype}
+\CSLLeftMargin{26. }%
+\CSLRightInline{D. L. Des Marais, K. M. Hernandez, T. E. Juenger,
+Genotype-by-environment interaction and plasticity: Exploring genomic
+responses of plants to the abiotic environment. \emph{Annual Review of
+Ecology, Evolution, and Systematics} \textbf{44}, 5--29 (2013).}
+
+\bibitem[\citeproctext]{ref-anderson2013genetic}
+\CSLLeftMargin{27. }%
+\CSLRightInline{J. T. Anderson, C.-R. Lee, C. A. Rushworth, R. I.
+Colautti, T. Mitchell-Olds, Genetic trade-offs and conditional
+neutrality contribute to local adaptation. \emph{Molecular ecology}
+\textbf{22}, 699--708 (2013).}
+
+\bibitem[\citeproctext]{ref-anderson2011evolutionary}
+\CSLLeftMargin{28. }%
+\CSLRightInline{J. T. Anderson, J. H. Willis, T. Mitchell-Olds,
+Evolutionary genetics of plant adaptation. \emph{Trends in Genetics}
+\textbf{27}, 258--266 (2011).}
+
+\bibitem[\citeproctext]{ref-mitchell1997predicting}
+\CSLLeftMargin{29. }%
+\CSLRightInline{R. B. Mitchell, K. J. Moore, L. E. Moser, J. O. Fritz,
+D. D. Redfearn, Predicting developmental morphology in switchgrass and
+big bluestem. \emph{Agronomy Journal} \textbf{89}, 827--832 (1997).}
+
+\bibitem[\citeproctext]{ref-parrish2005biology}
+\CSLLeftMargin{30. }%
+\CSLRightInline{D. J. Parrish, J. H. Fike, The biology and agronomy of
+switchgrass for biofuels. \emph{BPTS} \textbf{24}, 423--459 (2005).}
+
+\bibitem[\citeproctext]{ref-casler2004latitudinal}
+\CSLLeftMargin{31. }%
+\CSLRightInline{M. Casler, K. P. Vogel, C. Taliaferro, R. Wynia,
+Latitudinal adaptation of switchgrass populations. \emph{Crop Science}
+\textbf{44}, 293--303 (2004).}
+
+\bibitem[\citeproctext]{ref-lovell2021genomic}
+\CSLLeftMargin{32. }%
+\CSLRightInline{J. T. Lovell, \emph{et al.}, Genomic mechanisms of
+climate adaptation in polyploid bioenergy switchgrass. \emph{Nature}
+\textbf{590}, 438--444 (2021).}
+
+\bibitem[\citeproctext]{ref-porter1966analysis}
+\CSLLeftMargin{33. }%
+\CSLRightInline{C. L. Porter Jr, An analysis of variation between upland
+and lowland switchgrass, panicum virgatum l., in central oklahoma.
+\emph{Ecology} \textbf{47}, 980--992 (1966).}
+
+\bibitem[\citeproctext]{ref-milano2016genetic}
+\CSLLeftMargin{34. }%
+\CSLRightInline{E. R. Milano, D. B. Lowry, T. E. Juenger, The genetic
+basis of upland/lowland ecotype divergence in switchgrass (panicum
+virgatum). \emph{G3: Genes, Genomes, Genetics} \textbf{6}, 3561--3570
+(2016).}
+
+\bibitem[\citeproctext]{ref-urbut2019flexible}
+\CSLLeftMargin{35. }%
+\CSLRightInline{S. M. Urbut, G. Wang, P. Carbonetto, M. Stephens,
+Flexible statistical methods for estimating and testing effects in
+genomic studies with multiple conditions. \emph{Nature genetics}
+\textbf{51}, 187--195 (2019).}
+
+\bibitem[\citeproctext]{ref-10.1093ux2fbiostatisticsux2fkxw041}
+\CSLLeftMargin{36. }%
+\CSLRightInline{M. Stephens,
+\href{https://doi.org/10.1093/biostatistics/kxw041}{{False discovery
+rates: a new deal}}. \emph{Biostatistics} \textbf{18}, 275--294 (2016).}
+
 \end{CSLReferences}

### 2023-12-28 manuscript/revision2/revision2.qmd
manuscript: Edit Abstract and Significance statement

@@
         state: SD
-        postal-code: 
+        postal-code: 57006
   - name: Phillip A. Fay
@@
         state: TX
-        postal-code: 
+        postal-code: 76502
   - name: Felix B. Fritschi
@@
         state: MO
-        postal-code: 
+        postal-code: 65211
   - name: David B. Lowry
@@
         state: MI
-        postal-code: 
+        postal-code: 48824
   - name: Robert B. Mitchell
@@
         state: NE
-        postal-code: 
+        postal-code: 68583
   - name: Francis M. Rouquette Jr
@@
         state: TX
-        postal-code: 
+        postal-code: 75684
   - name: Yanqi Wu
@@
         state: OK
-        postal-code: 
+        postal-code: 74078
   - name: Arbel Harpak
@@
 abstract: |
-  The timing of vegetative growth and flowering in plants ("phenological timings") multiple depend on environmental cues; thus, phenological traits can have important genotype-by-environment (GxE) interactions.
-Here, we map genotype-by-weather effects on phenological timings in multiple populations of two highly divergent switchgrass (Panicum virgatum) populations spanning the central United States.
-We distinguish GxWeather patterns covarying with interpretable weather-based cues and agnostic, empirically-inferred GxWeather patterns.
+  The timing of vegetative growth and flowering in plants ("phenological timings") depend on both genetic variation and environmental cues; thus, phenological traits can have important genotype-by-environment (GxE) interactions.
+In particular, understanding the mixture and prevalence of genotype-by-weather (GxWeather) interactions primed by weather prior to the phenological event should aid prediction and manipulation of phenological timings.
+Here, we map genotype-by-weather effects on phenological timings in two highly divergent switchgrass (*Panicum virgatum*) populations using repeated plantings of these populations at sites spanning the central United States.
+We distinguish GxWeather patterns covarying with interpretable weather-based cues from agnostic, site-based patterns.
 Most GxWeather effects belong to the latter category and do not covary with weather-based cues.
-However, >XX% of Gulf population effects on flowering time covary with photoperiod cues, while >XX% of Midwest upland population SNPs affecting flowering had effects that covaried with XX cues.
-An independent pseudo-F2 cross of Gulf and Midwest individuals mapped 23 additive QTLs for flowering at the same common gardens, all with significant mash associations and ten with enrichment of highly significant mash associations.
-We demonstrate that we can identify QTL with GxE and assign them to specific weather-based cues and/or data-driven patterns.
+However, 65% of Gulf population effects on the start of vegetative growth covary with daylength 14 days prior to green-up date.
+33% of Gulf population effects on flowering date covary with cumulative rainfall in the seven days prior to flowering, while 22% of Midwest population effects on flowering date covaried with day length change two days prior to flowering.
+An independent pseudo-F2 cross of Gulf and Midwest individuals mapped 23 additive QTLs for flowering at the same common gardens, all with significant mash associations and eleven with enrichment of highly significant mash associations.
+We demonstrate that we can identify QTL with GxWeather and assign them to specific weather-based cues or other patterns.
 Breeding for particular alleles at these loci could change flowering responsiveness to photoperiod cues in switchgrass.
@@
   The timing of plant seasonal development (phenology) has major impacts on fitness because of the steep price of plant-environment mismatches.
-Here we map the genetic basis of two phenological events, the start of above-ground growth and flowering, in two genetically and phenologically distinct populations of switchgrass.
-We do this at eight field locations spanning the latitudinal range of both switchgrass populations.
-Our approach allows us to identify regions of the genome that covary with specific weather-related environmental features at every location.
-For flowering, these features differed by population: the Midwest population had genetic variation that primarily covaried with cumulative growing degree days, a temperature-related measure, while the Gulf population had genetic variation that primarily covaried with photoperiod, a day-length-related measure.
+We infer the mixture of ways genetic effects on phenological traits depend on plant environment, particularly how effects covary with weather prior to the phenological event (GxWeather).
+Most effects do not have GxWeather, but a minority do.
+GxWeather is population-specific.
+The majority of effects on the start of vegetative growth in the Gulf population have GxWeather: these effects differ in sign across environments and covary with a photoperiod cue 14 days prior.
+A minority of effects on flowering date in the Gulf and Midwest populations had GxWeather, covarying with a cumulative rainfall and a photoperiod cue, respectively.
 author_contributions:  "T.E.J.
@@
 equal-contributor: "A.H.M. contributed equally to this work with L.Z. and S.A.S."
-keywords:  [allele-by-environment effect variation, antagonistic pleiotropy, photoperiod, cumulative growing degree days, genetic variation]
+keywords:  [allele-by-environment effect variation, antagonistic pleiotropy, photoperiod, cumulative rainfall, genetic variation]
 acknowledgements: | 
@@
 
-\firstpage[15]{2}
+\firstpage[10]{2}
 
@@
 However, most studies of flowering GxE focus on finding a single, best fitting form of genotype-environment covariance, despite the key expectation that different genetic subpopulations, and even different genomic regions, have likely evolved distinct patterns of GxE.
-Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or sign-changing GxE, at the level of individual loci [@levene1953genetic; @felsenstein1976theoretical; @kawecki2004conceptual; @hedrick1986genetic], previous work has found limited evidence of antagonistic pleiotropy [@des2013genotype; @wadgymar2017identifying], but has been limited by a known statistical bias that reduced detection of SNP effects that differ in sign [@des2013genotype; @anderson2013genetic; @anderson2011evolutionary].
+Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or sign-changing GxE, at the level of individual loci [@levene1953genetic; @felsenstein1976theoretical; @kawecki2004conceptual; @hedrick1986genetic], previous work has found limited evidence of antagonistic pleiotropy [@des2013genotype; @wadgymar2017identifying], but has been limited by a known statistical bias that reduced detection of genetic effects that differ in sign [@des2013genotype; @anderson2013genetic; @anderson2011evolutionary].
 Thus, despite substantial interest in the frequencies of various forms of GxE, the prevalence of antagonistic pleiotropy relative to other forms of GxE remains unknown.
@@
 
-In our diversity panel of tetraploid switchgrass [@lovell2021genomic], genotypes from the Gulf and Midwest genetic subpopulations had distinct phenological timings and distinct patterns of phenological correlations across our eight common garden sites (@fig-map).
-At the three Texas common gardens (hereafter 'Texas' gardens), located within the natural range of the Gulf subpopulation, Gulf green-up occurred before Midwestern green-up, and Gulf flowering occurred after Midwestern flowering (Fig.
-1A).
+In our diversity panel of tetraploid switchgrass @lovell2021genomic, genotypes from the Gulf and Midwest genetic subpopulations had distinct phenological timings and distinct patterns of phenological correlations across our eight common garden sites (@fig-map).
+At the three Texas common gardens (hereafter 'Texas' gardens), located within the natural range of the Gulf subpopulation, Gulf green-up occurred before Midwestern green-up, and Gulf flowering occurred after Midwestern flowering (@fig-map A).
 At the four northernmost common gardens (hereafter 'North' gardens), located within the natural range of the Midwest subpopulation, both Gulf green-up and flowering occurred after Midwest green-up and flowering.
 At the Oklahoma common garden, located near the natural range limits of both the Gulf and the Midwest subpopulations, Gulf and Midwest green-up occurred over the same time period.
-These patterns led to strong negative phenotypic correlations for green-up between the North and Texas gardens, particularly in the Gulf and Both subpopulations, and contributed to positive phenotypic correlations for flowering time of larger magnitude at more northern gardens (Fig.
-1B).
+These patterns led to strong negative phenotypic correlations for green-up between the North and Texas gardens, particularly in the Gulf and Both subpopulations, and contributed to positive phenotypic correlations for flowering time of larger magnitude at more northern gardens (@fig-map B).
 
-Narrow-sense heritabilities (h2) indicated that rank-changing GxE for these phenotypes was present across the common gardens (Fig 1C).
-h2 were typically high at individual gardens: 59% on average for green-up date, and 87% for flowering date.
-However, h2 were variable across gardens, and green-up dates were uncorrelated (r2 \< 0.2) or negatively correlated between pairs of gardens (Fig.
-1B).
-These negative and small correlations undoubtedly contributed to the low h2 values for green-up and flowering date when estimated jointly at all eight gardens: h2 was 0.8% for green-up and 23.2% for flowering date.
+Narrow-sense heritabilities (h^2^) indicated that rank-changing GxE for these phenotypes was present across the common gardens (@fig-map C).
+h^2^ were typically high at individual gardens: 59% on average for green-up date, and 87% for flowering date.
+However, h^2^ were variable across gardens, and green-up dates were uncorrelated (r^2^ \< 0.2) or negatively correlated between pairs of gardens (@fig-map B).
+These negative and small correlations undoubtedly contributed to the low h^2^ values for green-up and flowering date when estimated jointly at all eight gardens: h^2^ was 0.8% for green-up and 23.2% for flowering date.
 
@@
 
-We next explored how genetic variation in phenology covaried with environmental cues by using mash to jointly discover trait-associated SNPs across all common gardens for the Midwest, Gulf, and 'Both' subpopulations (Fig 1D, SI Appendix, Datasets 1-6).
+We next explored how genetic variation in phenology covaried with environmental cues by using mash to jointly discover trait-associated SNPs across all common gardens for the Midwest, Gulf, and 'Both' subpopulations (@fig-map D, SI Appendix, Datasets 1-6).
 We first estimated SNP effects using GWAS on site-specific BLUPs for each phenotype.

### 2023-12-28 manuscript/revision2/revision2.tex
manuscript: Edit Abstract and Significance statement

@@
   pdfkeywords={allele-by-environment effect variation, antagonistic
-pleiotropy, photoperiod, cumulative growing degree days, genetic
-variation},
+pleiotropy, photoperiod, cumulative rainfall, genetic variation},
   colorlinks=true,
@@
 \affil[b]{South Dakota State University, Department of
-Agronomy, Brookings}
+Agronomy, Brookings, 57006}
 \affil[c]{USDA-ARS, Grassland, Soil and Water Research
-Laboratory, Temple}
-\affil[d]{University of Missouri, Division of Plant Sciences, Columbia}
+Laboratory, Temple, 76502}
+\affil[d]{University of Missouri, Division of Plant
+Sciences, Columbia, 65211}
 \affil[e]{Michigan State University, Department of Plant Biology, East
-Lansing}
-\affil[f]{USDA-ARS, Wheat, Sorghum, and Forage Research Unit, Lincoln}
+Lansing, 48824}
+\affil[f]{USDA-ARS, Wheat, Sorghum, and Forage Research
+Unit, Lincoln, 68583}
 \affil[g]{Texas A\&M University, Texas A\&M AgriLife Research and
-Extension Center, Overton}
+Extension Center, Overton, 75684}
 \affil[h]{Oklahoma State University, Department of Plant and Soil
-Sciences, Stillwater}
+Sciences, Stillwater, 74078}
 
@@
 (phenology) has major impacts on fitness because of the steep price of
-plant-environment mismatches. Here we map the genetic basis of two
-phenological events, the start of above-ground growth and flowering, in
-two genetically and phenologically distinct populations of switchgrass.
-We do this at eight field locations spanning the latitudinal range of
-both switchgrass populations. Our approach allows us to identify regions
-of the genome that covary with specific weather-related environmental
-features at every location. For flowering, these features differed by
-population: the Midwest population had genetic variation that primarily
-covaried with cumulative growing degree days, a temperature-related
-measure, while the Gulf population had genetic variation that primarily
-covaried with photoperiod, a day-length-related measure.}
+plant-environment mismatches. We infer the mixture of ways genetic
+effects on phenological traits depend on plant environment, particularly
+how effects covary with weather prior to the phenological event
+(GxWeather). Most effects do not have GxWeather, but a minority do.
+GxWeather is population-specific. The majority of effects on the start
+of vegetative growth in the Gulf population have GxWeather: these
+effects differ in sign across environments and covary with a photoperiod
+cue 14 days prior. A minority of effects on flowering date in the Gulf
+and Midwest populations had GxWeather, covarying with a cumulative
+rainfall and a photoperiod cue, respectively.}
 
@@
 \keywords{allele-by-environment effect variation | antagonistic
-pleiotropy | photoperiod | cumulative growing degree days | genetic
-variation}
+pleiotropy | photoperiod | cumulative rainfall | genetic variation}
 
@@
 The timing of vegetative growth and flowering in plants (``phenological
-timings'') multiple depend on environmental cues; thus, phenological
-traits can have important genotype-by-environment (GxE) interactions.
-Here, we map genotype-by-weather effects on phenological timings in
-multiple populations of two highly divergent switchgrass (Panicum
-virgatum) populations spanning the central United States. We distinguish
-GxWeather patterns covarying with interpretable weather-based cues and
-agnostic, empirically-inferred GxWeather patterns. Most GxWeather
-effects belong to the latter category and do not covary with
-weather-based cues. However, \textgreater XX\% of Gulf population
-effects on flowering time covary with photoperiod cues, while
-\textgreater XX\% of Midwest upland population SNPs affecting flowering
-had effects that covaried with XX cues. An independent pseudo-F2 cross
-of Gulf and Midwest individuals mapped 23 additive QTLs for flowering at
-the same common gardens, all with significant mash associations and ten
-with enrichment of highly significant mash associations. We demonstrate
-that we can identify QTL with GxE and assign them to specific
-weather-based cues and/or data-driven patterns. Breeding for particular
-alleles at these loci could change flowering responsiveness to
-photoperiod cues in switchgrass. More broadly, this approach could be
-used to identify genetic marker-environment interactions in any species
-with related populations phenotyped in multiple environments.
+timings'') depend on both genetic variation and environmental cues;
+thus, phenological traits can have important genotype-by-environment
+(GxE) interactions. In particular, understanding the mixture and
+prevalence of genotype-by-weather (GxWeather) interactions primed by
+weather prior to the phenological event should aid prediction and
+manipulation of phenological timings. Here, we map genotype-by-weather
+effects on phenological timings in two highly divergent switchgrass
+(\emph{Panicum virgatum}) populations using repeated plantings of these
+populations at sites spanning the central United States. We distinguish
+GxWeather patterns covarying with interpretable weather-based cues from
+agnostic, site-based patterns. Most GxWeather effects belong to the
+latter category and do not covary with weather-based cues. However, 65\%
+of Gulf population effects on the start of vegetative growth covary with
+daylength 14 days prior to green-up date. 33\% of Gulf population
+effects on flowering date covary with cumulative rainfall in the seven
+days prior to flowering, while 22\% of Midwest population effects on
+flowering date covaried with day length change two days prior to
+flowering. An independent pseudo-F2 cross of Gulf and Midwest
+individuals mapped 23 additive QTLs for flowering at the same common
+gardens, all with significant mash associations and eleven with
+enrichment of highly significant mash associations. We demonstrate that
+we can identify QTL with GxWeather and assign them to specific
+weather-based cues or other patterns. Breeding for particular alleles at
+these loci could change flowering responsiveness to photoperiod cues in
+switchgrass. More broadly, this approach could be used to identify
+genetic marker-environment interactions in any species with related
+populations phenotyped in multiple environments.
 \end{abstract}
@@
 \ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
-\firstpage[15]{2}
+\firstpage[10]{2}
 
@@
 has been limited by a known statistical bias that reduced detection of
-SNP effects that differ in sign (26--28). Thus, despite substantial
+genetic effects that differ in sign (26--28). Thus, despite substantial
 interest in the frequencies of various forms of GxE, the prevalence of
@@
 Midwestern green-up, and Gulf flowering occurred after Midwestern
-flowering (Fig. 1A). At the four northernmost common gardens (hereafter
-`North' gardens), located within the natural range of the Midwest
-subpopulation, both Gulf green-up and flowering occurred after Midwest
-green-up and flowering. At the Oklahoma common garden, located near the
-natural range limits of both the Gulf and the Midwest subpopulations,
-Gulf and Midwest green-up occurred over the same time period. These
-patterns led to strong negative phenotypic correlations for green-up
-between the North and Texas gardens, particularly in the Gulf and Both
-subpopulations, and contributed to positive phenotypic correlations for
-flowering time of larger magnitude at more northern gardens (Fig. 1B).
-
-Narrow-sense heritabilities (h2) indicated that rank-changing GxE for
-these phenotypes was present across the common gardens (Fig 1C). h2 were
-typically high at individual gardens: 59\% on average for green-up date,
-and 87\% for flowering date. However, h2 were variable across gardens,
-and green-up dates were uncorrelated (r2 \textless{} 0.2) or negatively
-correlated between pairs of gardens (Fig. 1B). These negative and small
-correlations undoubtedly contributed to the low h2 values for green-up
-and flowering date when estimated jointly at all eight gardens: h2 was
-0.8\% for green-up and 23.2\% for flowering date.
+flowering (Figure~\ref{fig-map} A). At the four northernmost common
+gardens (hereafter `North' gardens), located within the natural range of
+the Midwest subpopulation, both Gulf green-up and flowering occurred
+after Midwest green-up and flowering. At the Oklahoma common garden,
+located near the natural range limits of both the Gulf and the Midwest
+subpopulations, Gulf and Midwest green-up occurred over the same time
+period. These patterns led to strong negative phenotypic correlations
+for green-up between the North and Texas gardens, particularly in the
+Gulf and Both subpopulations, and contributed to positive phenotypic
+correlations for flowering time of larger magnitude at more northern
+gardens (Figure~\ref{fig-map} B).
+
+Narrow-sense heritabilities (h\textsuperscript{2}) indicated that
+rank-changing GxE for these phenotypes was present across the common
+gardens (Fig 1C). h\textsuperscript{2} were typically high at individual
+gardens: 59\% on average for green-up date, and 87\% for flowering date.
+However, h2 were variable across gardens, and green-up dates were
+uncorrelated (r2 \textless{} 0.2) or negatively correlated between pairs
+of gardens (Fig. 1B). These negative and small correlations undoubtedly
+contributed to the low h2 values for green-up and flowering date when
+estimated jointly at all eight gardens: h2 was 0.8\% for green-up and
+23.2\% for flowering date.
 

### 2024-01-08 manuscript/revision2/revision2.qmd
manuscript: revision 2 text work Abstract through Discussion

@@
 ---
-title: Switchgrass Phenology and Genotype-by-Environment Mapping
+title: Mapping Mixtures of Genotype-by-Weather Interactions in Switchgrass (Panicum virgatum)
 format:
@@
 abstract: |
-  The timing of vegetative growth and flowering in plants ("phenological timings") depend on both genetic variation and environmental cues; thus, phenological traits can have important genotype-by-environment (GxE) interactions.
-In particular, understanding the mixture and prevalence of genotype-by-weather (GxWeather) interactions primed by weather prior to the phenological event should aid prediction and manipulation of phenological timings.
-Here, we map genotype-by-weather effects on phenological timings in two highly divergent switchgrass (*Panicum virgatum*) populations using repeated plantings of these populations at sites spanning the central United States.
+  The timing of vegetative and reproductive growth in plants ("phenological timings") depends on both genetic variation and environmental cues; thus, phenological traits can have important genotype-by-environment (GxE) interactions.
+Identifying the mixture and prevalence of GxE primed by weather prior to the phenological event (GxWeather) should aid prediction and manipulation of phenological timings.
+Here, we map GxWeather effects on phenological timings in two highly divergent switchgrass (*Panicum virgatum*) populations using repeated plantings of these populations at eight sites spanning the central United States.
 We distinguish GxWeather patterns covarying with interpretable weather-based cues from agnostic, site-based patterns.
 Most GxWeather effects belong to the latter category and do not covary with weather-based cues.
-However, 65% of Gulf population effects on the start of vegetative growth covary with daylength 14 days prior to green-up date.
-33% of Gulf population effects on flowering date covary with cumulative rainfall in the seven days prior to flowering, while 22% of Midwest population effects on flowering date covaried with day length change two days prior to flowering.
-An independent pseudo-F2 cross of Gulf and Midwest individuals mapped 23 additive QTLs for flowering at the same common gardens, all with significant mash associations and eleven with enrichment of highly significant mash associations.
+However, 65% of effects on the timing of vegetative growth in the Gulf population covary with daylength 14 days prior to green-up date.
+33% of effects on the timing of flowering in the Gulf population covary with cumulative rainfall in the seven days prior to flowering, while 22% of effects on flowering timing in the Midwest population covary with day length change two days prior to flowering.
+An independent pseudo-F2 cross of Gulf and Midwest individuals at the same sites mapped 23 additive QTLs for flowering, all with significant associations and eleven with enrichment of highly significant associations in our model of GxWeather effects.
 We demonstrate that we can identify QTL with GxWeather and assign them to specific weather-based cues or other patterns.
-Breeding for particular alleles at these loci could change flowering responsiveness to photoperiod cues in switchgrass.
+Breeding for particular alleles at these loci could change flowering responsiveness to photoperiod and rainfall cues in switchgrass.
 More broadly, this approach could be used to identify genetic marker-environment interactions in any species with related populations phenotyped in multiple environments.
@@
 significance: |
-  The timing of plant seasonal development (phenology) has major impacts on fitness because of the steep price of plant-environment mismatches.
-We infer the mixture of ways genetic effects on phenological traits depend on plant environment, particularly how effects covary with weather prior to the phenological event (GxWeather).
+  The timing of plant seasonal development (phenology) has major impacts on fitness because of the steep fitness cost of plant-environment mismatches.
+We infer the mixture of ways genetic effects on phenological traits depend on plant environment, focusing on how effects covary with weather prior to the phenological event (GxWeather).
 Most effects do not have GxWeather, but a minority do.
 GxWeather is population-specific.
-The majority of effects on the start of vegetative growth in the Gulf population have GxWeather: these effects differ in sign across environments and covary with a photoperiod cue 14 days prior.
+The majority of effects on the timing of vegetative growth in the Gulf population have GxWeather: these effects differ in sign across environments and covary with a photoperiod cue two weeks prior.
 A minority of effects on flowering date in the Gulf and Midwest populations had GxWeather, covarying with a cumulative rainfall and a photoperiod cue, respectively.
@@
 
-\firstpage[10]{2}
+\firstpage[9]{2}
 
-The timing of plant vegetative and reproductive development are major components of plant fitness affected by multiple external environmental cues (e.g.
+Plant phenological timings are major components of plant fitness affected by multiple external environmental cues (e.g.
 degree of winter chilling, day length, temperature, and water availability) that signal existing or upcoming growing conditions [@bauerle_photoperiodic_2012; @andres2012genetic; @korner2010phenology].
@@
 Day length (or photoperiod) is one of the most predictable environmental cues, and genetic sensitivity to photoperiod protects plants from potentially fatal consequences of phenological responses to temperature cues at the "wrong" time of year.
-However, the usefulness of specific environmental cues depends on both features of the environment and the species' adaptive strategies @botero2015evolutionary.
+However, the usefulness of specific environmental cues depends on both features of the environment, such as cue predictability and relevance, and the species' adaptive strategies @botero2015evolutionary.
 Species with wide natural distributions can have multiple distinct environmentally-cued phenological responses: for example, populations of sunflower (Helianthus annuus) exhibit day-neutral, facultative short day, and facultative long-day flowering responses, which vary with their environments [@blackman2013interacting; @henry2014transitions].
@@
 
-Flowering time, in particular, is a common subject of GxE research [@blackman2013interacting; @henry2014transitions; @aagren2017adaptive; @brachi2010linkage; @dittmar2014flowering; @li2018genomic; @romero2017study], a key output of selection driving adaptation to local environments [@andres2012genetic; @wadgymar2017identifying; @blumel2015flowering], and a selection target for crop improvement to adapt crops to local or future environments @jung2009flowering.
+Flowering time, or the transition from vegetative to reproductive growth, is a common subject of GxE research [@blackman2013interacting; @henry2014transitions; @aagren2017adaptive; @brachi2010linkage; @dittmar2014flowering; @li2018genomic; @romero2017study], a key output of selection driving adaptation to local environments [@andres2012genetic; @wadgymar2017identifying; @blumel2015flowering], and a selection target for crop improvement to adapt crops to local or future environments @jung2009flowering.
 Changing flowering responsiveness to photoperiod cues has allowed geographic range expansion and increased yields in a number of cereal species [@turner2005pseudo; @faure2012mutation; @hung2012zmcct; @zakhrabekova2012induced; @yang2013oself3] and other crops [@pin2012multifaceted; @weller2019parallel].
@@
 However, most studies of flowering GxE focus on finding a single, best fitting form of genotype-environment covariance, despite the key expectation that different genetic subpopulations, and even different genomic regions, have likely evolved distinct patterns of GxE.
-Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or sign-changing GxE, at the level of individual loci [@levene1953genetic; @felsenstein1976theoretical; @kawecki2004conceptual; @hedrick1986genetic], previous work has found limited evidence of antagonistic pleiotropy [@des2013genotype; @wadgymar2017identifying], but has been limited by a known statistical bias that reduced detection of genetic effects that differ in sign [@des2013genotype; @anderson2013genetic; @anderson2011evolutionary].
+Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or sign-changing GxE, at the level of individual loci [@levene1953genetic; @felsenstein1976theoretical; @kawecki2004conceptual; @hedrick1986genetic], previous work has found limited evidence of antagonistic pleiotropy [@des2013genotype; @wadgymar2017identifying].
+However, this work has been limited by a known statistical bias that reduced detection of genetic effects that differ in sign [@des2013genotype; @anderson2013genetic; @anderson2011evolutionary].
 Thus, despite substantial interest in the frequencies of various forms of GxE, the prevalence of antagonistic pleiotropy relative to other forms of GxE remains unknown.
@@
 However, as part of its wide environmental adaptation across the eastern half of North America, its photoperiodicity has been predicted to differ by plant latitude of origin [@parrish2005biology; @casler2004latitudinal].
-We previously found divergent Midwest and Gulf genetic subpopulations of switchgrass which segregate for distinct sets of environmental adaptations, based on the analysis of two measures of fitness (i.e.,biomass and survival) among 732 genotypes @lovell2021genomic.
-The Midwest genetic subpopulation is primarily composed of individuals from the well-studied upland switchgrass ecotype [@porter1966analysis; @milano2016genetic], while the Gulf subpopulation has individuals from the lowland ecotype and the phenotypically intermediate coastal ecotype @lovell2021genomic.
+We previously found divergent Midwest and Gulf genetic subpopulations of switchgrass with distinct sets of environmental adaptations, in that both populations had distinct genetic variation associated with each of two fitness proxies, biomass and survival @lovell2021genomic.
+The Midwest genetic subpopulation is primarily composed of individuals from the well-studied upland switchgrass ecotype [@porter1966analysis; @milano2016genetic], while the Gulf subpopulation has individuals from the well-studied lowland ecotype and the phenotypically intermediate coastal ecotype @lovell2021genomic.
 
-Here, we test how the Midwest and Gulf subpopulations differ in their phenological adaptations across environments, and hence their phenological GxE.
-We phenotype a diversity panel of hundreds of switchgrass genotypes from the Midwest and Gulf subpopulations for the timing of vegetative development ("green-up") and reproductive development (flowering) at eight common garden locations spanning 17 degrees of latitude.
+Here, we estimate the ways genetic effects on phenological timings depend on weather prior to the phenological event.
+To do this, we phenotype a diversity panel of hundreds of switchgrass genotypes from the Midwest and Gulf subpopulations for the timing of vegetative and reproductive development at eight common garden locations spanning 17 degrees of latitude.
 These gardens cover the majority of the latitudinal and climatic range of switchgrass and capture the most comprehensive picture to date of the environmental variation this species encounters.
-After single-site estimation of SNP effects for these phenotypes, we use multivariate adaptive shrinkage (mash) to jointly estimate SNP effects in each subpopulation and phenotype at all eight sites @urbut2019flexible.
-Mash allows us to identify and specify multiple ways genetic effects linked to SNPs may covary with the environment, and does not have a statistical bias in detecting SNP effects with the same or opposite signs @10.1093/biostatistics/kxw041.
-To confirm our genetic mapping of GxE, we compare to mapping results from an outbred pseudo-F2 cross grown at the same sites.
-Taken together, our results allow us to describe the environmental cues and genetic variation affecting phenology in two divergent natural populations of switchgrass.
+We define a number of ways phenological timings might covary with weather (\@tbl-covar) and additional ways phenological timings might vary by site (\@SI-info), then jointly re-estimate genetic effects on these timings at all eight sites using the set of these covariance types that significantly improved the modeled log-likelihood when included (\@SI-info) @urbut2019flexible.
+The method we use for effect re-estimation, mash, allows us to identify and specify multiple ways genetic effects may covary with the environment, and does not have a statistical bias in detecting genetic effects with the same or opposite signs @10.1093/biostatistics/kxw041.
+To confirm our genetic mapping of GxWeather, we compare our re-estimated genetic effects to mapping results from an outbred pseudo-F2 cross grown at the same sites.
+Our analysis allow us to describe the weather cues and genetic variation affecting phenology in two divergent natural populations of switchgrass.
 
@@
 
-In our diversity panel of tetraploid switchgrass @lovell2021genomic, genotypes from the Gulf and Midwest genetic subpopulations had distinct phenological timings and distinct patterns of phenological correlations across our eight common garden sites (@fig-map).
-At the three Texas common gardens (hereafter 'Texas' gardens), located within the natural range of the Gulf subpopulation, Gulf green-up occurred before Midwestern green-up, and Gulf flowering occurred after Midwestern flowering (@fig-map A).
-At the four northernmost common gardens (hereafter 'North' gardens), located within the natural range of the Midwest subpopulation, both Gulf green-up and flowering occurred after Midwest green-up and flowering.
-At the Oklahoma common garden, located near the natural range limits of both the Gulf and the Midwest subpopulations, Gulf and Midwest green-up occurred over the same time period.
-These patterns led to strong negative phenotypic correlations for green-up between the North and Texas gardens, particularly in the Gulf and Both subpopulations, and contributed to positive phenotypic correlations for flowering time of larger magnitude at more northern gardens (@fig-map B).
+Genotypes from the Gulf and Midwest subpopulations had distinct phenological timings and distinct patterns of phenological correlations across our eight common garden sites (@fig-map).
+At the three Texas common gardens (hereafter 'Texas' gardens), located within the natural range of the Gulf subpopulation, Gulf vegetative growth occurred before Midwestern vegetative growth, and Gulf flowering occurred after Midwestern flowering (@fig-map A).
+At the four northernmost common gardens (hereafter 'North' gardens), located within the natural range of the Midwest subpopulation, both Gulf vegetative growth and flowering occurred after Midwest vegetative growth and flowering.
+At the Oklahoma common garden, located near the natural range limits of both the Gulf and the Midwest subpopulations, Gulf and Midwest vegetative growth occurred over the same time period, and Gulf flowering occurred after Midwestern flowering (@fig-map A).
+These patterns led to strong negative phenotypic correlations for the timing of vegetative growth between the North and Texas gardens, particularly in the Gulf and across all individuals (hereafter, 'Both' subpopulations), and contributed to positive phenotypic correlations for the timing of flowering that had larger magnitudes at more northern gardens (@fig-map B).
 
@@
 
-### Mapping major patterns of genotype-by-environment effects on green-up and flowering
+### GxWeather mixtures for vegetative and reproductive timing
 
-We next explored how genetic variation in phenology covaried with environmental cues by using mash to jointly discover trait-associated SNPs across all common gardens for the Midwest, Gulf, and 'Both' subpopulations (@fig-map D, SI Appendix, Datasets 1-6).
-We first estimated SNP effects using GWAS on site-specific BLUPs for each phenotype.
-We then used mash on two subsets of these SNP effects, first to determine the covariance structures that significantly improved the mash model log likelihood when included, then to re-estimate the set of SNP effects with the largest effect estimates in univariate GWAS (hereafter "strong" effects) using the set of significant covariance matrices.
+We next looked for evidence of GxWeather by jointly re-estimating genetic effects across all eight common gardens using a model that allows effects to covary between gardens in many ways (@fig-map D, SI Appendix, Datasets 1-6).
+To do this, we used genome-wide association (GWAS) on site-specific BLUPs for the timing of vegetative growth and flowering to determine preliminary single nucleotide polymorphism (SNP) effect estimates at each garden.
+We then jointly modeled genetic effect estimates, and the mixture of ways these effects might covary, across all eight common gardens, using a two-part procedure.
+We first modeled genetic effects with different combinations of covariance types, using a subset of effects selected at random from all SNPs used in the GWAS.
+In these models, we specified many ways genetic effects could covary between gardens (as expanded on below) and used a greedy algorithm to iteratively select covariance structures that significantly improved the log likelihood of a model when included.
+Given the high correlation between some of the covariance patterns we tested, this approach allowed us to select only the subset of patterns of covariation that most improved the fit of the joint effect patterns present in the data.
+We then re-estimated the set of genetic effects that had the largest effect estimates in univariate GWAS (hereafter "strong" effects) using the set of significant covariance matrices that most improved the model fit on the random effects.
 
-Three types of covariance matrices were tested for inclusion: "canonical" covariance matrices, with simple patterns of effects; "data-driven" matrices derived from common patterns of SNP effects in the data, and "hypothesis-based" covariance matrices that captured the covariance of weather-based phenological cues for cloned genotypes grown at multiple common gardens (Table 1, SI Appendix, Section S2).
-Mash first assigns mixture proportions for each SNP onto each provided covariance matrix using maximum likelihood.
-Then, in a second step, each SNP's effect estimate undergoes shrinkage such that it is more aligned with the covariance structure inferred in the first step.
-The hypothesis-based covariance matrices are an important advantage mash offers for studying patterns of GxE: they allow hypothesis testing of specific environmental drivers for each SNP effect, and loadings on these matrices provide information on genome-wide patterns of SNP-environment interaction.
-Say that a SNP in FLC controls flowering in a temperature-dependent manner.
-In that case, that SNP effect could have a high mixture proportion, or mass, on a covariance matrix created using a temperature-based environmental cue, such as average temperature on the day prior to flowering.
+We created three categories of ways effects could covary across gardens: "canonical" covariance, with simple patterns of effects; "data-driven" covariances derived from common patterns of SNP effects in the data, and "GxWeather" covariances that captured the covariance of weather variables at specific times before the phenological event (Table 1, SI Appendix, Section S2).
+For example, GxWeather covariance between gardens in MI and MO, for rainfall in the seven days prior to flowering, captures the covariance in the amount of rainfall genotype A, B, and C receive in the seven days prior to their three flowering dates in MI and their three flowering dates in MO.
+If these six rainfall values are similar, covariance will be near one; if they are uncorrelated, covariance will be near zero; if there is sufficient rank changing in rainfall amounts between gardens for these genotypes, this covariance will be near negative one.
+
+The GxWeather covariances allow hypothesis testing of specific weather variables as cues for each genetic effect.
+Say that a SNP in FLC controls flowering in a photoperiod-dependent manner.
+In that case, that SNP effect could have a high mixture proportion, or mass, on a covariance matrix created using a photoperiod-based environmental cue, such as day length two days prior to flowering.
 In our data, we would infer that the effect of that SNP on flowering was caused by a response to that or a correlated environmental cue.
+Overall, loadings of genetic effects on these GxWeather matrices in the model with re-estimated effects provide information on genome-wide patterns of SNP-environment interaction.
 
-The phenotypic correlations for green up date had moderate negative correlations between the Texas and North gardens, particularly in the Gulf and in Both subpopulations (Fig 1B).
-If these phenotypic correlations have a genetic component, then they might be controlled by SNP effects that differed in sign across these regions, SNP effects that are large in one garden or region and non-significant in others, or a combination of these patterns.
-Five hypothesis-based covariance matrices were selected by the greedy mash algorithm, one to two per subpopulation.
-Of these five, three had mass on them in mash models of the strong effects (Fig 2A, 2B).
-Two of these three matrices had negative covariances between sites between Texas and North gardens (Fig 2A), while one had all positive or near-zero covariances.
-SNP-associated phenotypic effects covaried with different weather-based cues in the Gulf & in Both subpopulations (Fig 2B).
-In total, 65% of the posterior weight of strong SNP effects in the mash model of Gulf green-up fell on a covariance matrix of daylength 14 days prior to the date of green-up.
-The covariance matrix for this weather cue was very similar to the pattern of phenotypic correlation for green-up in the Gulf subpopulation (Fig 1B; Fig 2A).
-Mash models of Midwest and Both subpopulation green-up did not include this covariance matrix; the Midwest had no weight on any hypothesis-based matrix, while Both subpopulations had non-zero weights on two additional hypothesis-based matrices, average temperature one day prior to green-up, and the day length change in seconds in the day prior to green-up (Fig 2B).
-The average temperature covariance matrix had negative covariances between Texas and North gardens, though not as strong as the negative phenotypic correlations seen in Both subpopulations (Fig 1B; Fig 2A).
-Only the Gulf subpopulation and Both subpopulations had mass on any hypothesis-based covariance matrices (Fig 2C).
+The phenotypic correlations for the timing of vegetative growth had moderate negative correlations between the Texas and North gardens, particularly in the Gulf and in Both subpopulations (@fig-map B).
+If these phenotypic correlations have a genetic component, then they might be controlled by SNP effects that differed in sign across these regions (effects with antagonistic pleiotropy), SNP effects that are large in one garden or region and non-significant in others (effects with amplification), or a combination of these patterns.
+Five GxWeather covariance matrices were selected by the greedy mash algorithm, one to two per subpopulation.
+Of these five, three had mass on them in mash models of the strong effects (@fig-covar A-B).
+Two of these three matrices had negative covariances between sites between Texas and North gardens (@fig-covar A), while one had all positive or near-zero covariances.
+SNP-associated phenotypic effects covaried with different weather-based cues in the Gulf & in Both subpopulations (@fig-covar B).
+In total, 65% of the posterior weight of strong SNP effects in the mash model of Gulf vegetative growth fell on a covariance matrix of daylength 14 days prior to the date of vegetative growth.
+The covariance matrix for this weather cue was very similar to the pattern of phenotypic correlation for the timing of vegetative growth in the Gulf subpopulation (@fig-map B; @fig-covar A).
+Mash models of the timing of vegetative growth in Midwest and Both subpopulations did not include this GxWeather covariance; the Midwest had no weight on any GxWeather covariance, while Both subpopulations had non-zero weights on two additional GxWeather covariance types, average temperature one day prior to green-up, and the day length change in seconds in the day prior to green-up (@fig-covar B).
+The average temperature covariance matrix had negative covariances between Texas and North gardens, though not as strong as the negative phenotypic correlations seen in Both subpopulations (@fig-map B; @fig-covar A).
+Only the Gulf subpopulation and Both subpopulations had mass on any GxWeather covariance matrices (@fig-covar C).
 
-For flowering date, distinct weather-based cues captured SNP-associated effect patterns in the Gulf and Midwest subpopulations.
+For flowering date, distinct GxWeather covariances captured SNP-associated effect patterns in the Gulf and Midwest subpopulations.
 33% of SNP effects on flowering in the Gulf subpopulation covaried with cumulative rainfall in the seven days prior to flowering (Fig 2D).
-22.6% of SNP effects on flowering in the Midwest subpopulation contrived with day length change in the two days prior to flowering (Fig 2D), which had negative covariances between Texas and North gardens.
-Neither covariance matrix was selected for in Both subpopulations, and no hypothesis-based covariance matrices had mass in Both subpopulations (Fig 2E).
-In five of the six mash models of strong effects, the hypothesis-based covariance matrices captured a minority of the posterior weights of the strong effects (Figure 2C,E); the majority of this mass was on various canonical covariance matrices.
+22.6% of SNP effects on flowering in the Midwest subpopulation covaried with day length change in the two days prior to flowering (Fig 2D), which had negative covariances between Texas and North gardens.
+Neither covariance matrix was selected for in Both subpopulations, and no GxWeather covariance matrices had mass in Both subpopulations (Fig 2E).
+In five of the six mash models of strong effects, the GxWeather covariance matrices captured a minority of the posterior weights of the strong effects (Figure 2C,E); the majority of this mass was on various canonical covariance matrices.
 These matrices included simple heterozygosity, with intermediate, positive covariances between all gardens, and garden-specific effects present only at one garden.
@@
 
-### Confirmation of genotype-by-environment effects using an independent mapping population
+### Confirmation of effects on phenology using an independent mapping population
 
-We sought additional experimental support for our mash intervals using an independent pseudo-F2 mapping population created from Gulf & Midwest individuals and grown at the same sites (Fig.
+We sought additional experimental support for our significant effect re-estimates using an independent pseudo-F2 mapping population created from Gulf & Midwest individuals and grown at the same sites (Fig.
 4A,B).
@@
 S3).
-All QTL for flowering overlapped one or more homologs from rice or A.
-thaliana with functionally validated roles in flowering (SI Appendix, Dataset 7).
+All QTL for flowering overlapped one or more homologs from rice or *A.
+thaliana* with functionally validated roles in flowering (SI Appendix, Dataset 7).
 All flowering QTL intervals contained at least one SNP significant in at least one mash run at a log10-transformed Bayes Factor \> 2, or in the 1% tail of significance, whichever was stricter (SI Appendix, Dataset 8).
@@
 4C, p = 0.011).
-Thus, we were able to experimentally support our mash intervals with a QTL mapping experiment using a separate mapping population.
+Thus, we were able to experimentally support our significant effect re-estimates with a QTL mapping experiment using a separate mapping population.
 
@@
 To do this, we must understand the current patterns of trait covariation across environments, the genetic underpinnings of these patterns, and the cases where this covariation can be altered.
-Here, we demonstrate that we can associate multiple patterns of GxE with specific genomic regions using a switchgrass diversity panel grown at eight common gardens, and also that we can assign specific SNP-associated patterns of GxE to both weather-based cues and to other, data-driven patterns.
-We use this approach to study GxE in both green-up and flowering phenological data in the deeply genetically diverged Gulf and Midwest populations of switchgrass.
+Here, we demonstrate that we can associate multiple patterns of GxWeather with specific genomic regions using a switchgrass diversity panel grown at eight common gardens.
+We can assign genetic effects to both GxWeather patterns with interpretable weather-based cues, and to agnostic, site-based patterns.
+We use this approach to study GxWeather for the timings of vegetative and reproductive development in the deeply genetically diverged Gulf and Midwest subpopulations of switchgrass.
 
-Our analysis of green-up in the Gulf and in Both subpopulations revealed substantial antagonistic pleiotropy in effects between the Texas and North gardens (Figure 3A).
-This result supports theoretical models that local adaptation should involve antagonistic pleiotropy at the level of individual loci (22--25), and is the first experimental work using QTL mapping and GWAS across common gardens to find antagonistic pleiotropy to be common in small genomic regions (12, 36, 37).
+Our analysis of the timing of vegetative growth in the Gulf and in Both subpopulations revealed substantial antagonistic pleiotropy in effects between the Texas and North gardens (Figure 3A).
+This result supports theoretical models that local adaptation should involve antagonistic pleiotropy at the level of individual loci [@kawecki2004conceptual; @felsenstein1976theoretical; @levene1953genetic; @hedrick1986genetic], and is the first experimental work using QTL mapping and GWAS across common gardens to find antagonistic pleiotropy to be common in small genomic regions [@wadgymar2017identifying; @savolainen2013ecological; @lowry2019qtl].
 
-Our analysis of flowering showed that the Gulf and Midwest subpopulations have two distinct photoperiod-related flowering responses: the Midwest subpopulation is day neutral, and flowering is cued primarily by a cumulative GDD threshold; in contrast, the Gulf subpopulation is photoperiod sensitive, and flowering is cued by the transition to shortening days.
-This result was supported by observations that expressing flowering date as a function of the day length at flowering increased its heritability in the Gulf subpopulation, while expressing flowering date as a function of cumulative GDD between green-up and flowering increased the heritability of flowering in the Midwest subpopulation (Fig.
-1D).
-The genomic regions affecting flowering found in mash were also supported by QTL from an independent mapping population (Fig.
+Our analysis of the timing of flowering showed that the Gulf and Midwest subpopulations have distinct GxWeather: flowering timing in the Midwest subpopulation has photoperiod-related variation, in that flowering timing covaries with a day length change signal in the two days before flowering occurs.
+In contrast, the Gulf subpopulation does not have variation in flowering based on photoperiod.
+Instead, the Gulf subpopulation has variation in flowering that covaries with the rainfall that occurs in the week prior to flowering.
+The genomic regions affecting flowering that we re-estimated across all eight sites were also supported by QTL from an independent mapping population at these sites (Fig.
 4C).
@@
 ```
-![Figure 1.
-Characterization of green-up and flowering dates from the switchgrass diversity panel.
+![Characterization of green-up and flowering dates from the switchgrass diversity panel.
 (a) Map and trait histograms of green-up and flowering dates across two genetically distinct switchgrass subpopulations and eight common gardens.
@@
 
-![Figure 2.](images/Figure_2_Covariance_matrices_and_posterior_weights.png){#fig-covar}
+![Caption for this figure...](images/Figure_2_Covariance_matrices_and_posterior_weights.png){#fig-covar}
 

### 2024-01-08 manuscript/revision2/revision2.tex
manuscript: revision 2 text work Abstract through Discussion

@@
 \hypersetup{
-  pdftitle={Switchgrass Phenology and Genotype-by-Environment Mapping},
+  pdftitle={Mapping Mixtures of Genotype-by-Weather Interactions in Switchgrass (Panicum virgatum)},
   pdfauthor={Alice H.
@@
 
-\title{Switchgrass Phenology and Genotype-by-Environment Mapping}
+\title{Mapping Mixtures of Genotype-by-Weather Interactions in
+Switchgrass (Panicum virgatum)}
 
@@
 \significancestatement{The timing of plant seasonal development
-(phenology) has major impacts on fitness because of the steep price of
-plant-environment mismatches. We infer the mixture of ways genetic
-effects on phenological traits depend on plant environment, particularly
-how effects covary with weather prior to the phenological event
-(GxWeather). Most effects do not have GxWeather, but a minority do.
-GxWeather is population-specific. The majority of effects on the start
-of vegetative growth in the Gulf population have GxWeather: these
+(phenology) has major impacts on fitness because of the steep fitness
+cost of plant-environment mismatches. We infer the mixture of ways
+genetic effects on phenological traits depend on plant environment,
+focusing on how effects covary with weather prior to the phenological
+event (GxWeather). Most effects do not have GxWeather, but a minority
+do. GxWeather is population-specific. The majority of effects on the
+timing of vegetative growth in the Gulf population have GxWeather: these
 effects differ in sign across environments and covary with a photoperiod
-cue 14 days prior. A minority of effects on flowering date in the Gulf
+cue two weeks prior. A minority of effects on flowering date in the Gulf
 and Midwest populations had GxWeather, covarying with a cumulative
@@
 \begin{abstract}
-The timing of vegetative growth and flowering in plants (``phenological
-timings'') depend on both genetic variation and environmental cues;
-thus, phenological traits can have important genotype-by-environment
-(GxE) interactions. In particular, understanding the mixture and
-prevalence of genotype-by-weather (GxWeather) interactions primed by
-weather prior to the phenological event should aid prediction and
-manipulation of phenological timings. Here, we map genotype-by-weather
-effects on phenological timings in two highly divergent switchgrass
-(\emph{Panicum virgatum}) populations using repeated plantings of these
-populations at sites spanning the central United States. We distinguish
-GxWeather patterns covarying with interpretable weather-based cues from
-agnostic, site-based patterns. Most GxWeather effects belong to the
-latter category and do not covary with weather-based cues. However, 65\%
-of Gulf population effects on the start of vegetative growth covary with
-daylength 14 days prior to green-up date. 33\% of Gulf population
-effects on flowering date covary with cumulative rainfall in the seven
-days prior to flowering, while 22\% of Midwest population effects on
-flowering date covaried with day length change two days prior to
+The timing of vegetative and reproductive growth in plants
+(``phenological timings'') depends on both genetic variation and
+environmental cues; thus, phenological traits can have important
+genotype-by-environment (GxE) interactions. Identifying the mixture and
+prevalence of GxE primed by weather prior to the phenological event
+(GxWeather) should aid prediction and manipulation of phenological
+timings. Here, we map GxWeather effects on phenological timings in two
+highly divergent switchgrass (\emph{Panicum virgatum}) populations using
+repeated plantings of these populations at eight sites spanning the
+central United States. We distinguish GxWeather patterns covarying with
+interpretable weather-based cues from agnostic, site-based patterns.
+Most GxWeather effects belong to the latter category and do not covary
+with weather-based cues. However, 65\% of effects on the timing of
+vegetative growth in the Gulf population covary with daylength 14 days
+prior to green-up date. 33\% of effects on the timing of flowering in
+the Gulf population covary with cumulative rainfall in the seven days
+prior to flowering, while 22\% of effects on flowering timing in the
+Midwest population covary with day length change two days prior to
 flowering. An independent pseudo-F2 cross of Gulf and Midwest
-individuals mapped 23 additive QTLs for flowering at the same common
-gardens, all with significant mash associations and eleven with
-enrichment of highly significant mash associations. We demonstrate that
-we can identify QTL with GxWeather and assign them to specific
-weather-based cues or other patterns. Breeding for particular alleles at
-these loci could change flowering responsiveness to photoperiod cues in
-switchgrass. More broadly, this approach could be used to identify
-genetic marker-environment interactions in any species with related
-populations phenotyped in multiple environments.
+individuals at the same sites mapped 23 additive QTLs for flowering, all
+with significant associations and eleven with enrichment of highly
+significant associations in our model of GxWeather effects. We
+demonstrate that we can identify QTL with GxWeather and assign them to
+specific weather-based cues or other patterns. Breeding for particular
+alleles at these loci could change flowering responsiveness to
+photoperiod and rainfall cues in switchgrass. More broadly, this
+approach could be used to identify genetic marker-environment
+interactions in any species with related populations phenotyped in
+multiple environments.
 \end{abstract}
@@
 \ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
-\firstpage[10]{2}
-
-The timing of plant vegetative and reproductive development are major
-components of plant fitness affected by multiple external environmental
-cues (e.g.~degree of winter chilling, day length, temperature, and water
-availability) that signal existing or upcoming growing conditions
-(1--3). Genetic responses to environmental cues determine the speed,
-timing, and energy apportioned to vegetative and reproductive growth and
-shape both the individual's lifespan and its lifetime production of
-viable seed. Day length (or photoperiod) is one of the most predictable
-environmental cues, and genetic sensitivity to photoperiod protects
-plants from potentially fatal consequences of phenological responses to
-temperature cues at the ``wrong'' time of year. However, the usefulness
-of specific environmental cues depends on both features of the
-environment and the species' adaptive strategies (4). Species with wide
-natural distributions can have multiple distinct environmentally-cued
-phenological responses: for example, populations of sunflower
-(Helianthus annuus) exhibit day-neutral, facultative short day, and
-facultative long-day flowering responses, which vary with their
+\firstpage[9]{2}
+
+Plant phenological timings are major components of plant fitness
+affected by multiple external environmental cues (e.g.~degree of winter
+chilling, day length, temperature, and water availability) that signal
+existing or upcoming growing conditions (1--3). Genetic responses to
+environmental cues determine the speed, timing, and energy apportioned
+to vegetative and reproductive growth and shape both the individual's
+lifespan and its lifetime production of viable seed. Day length (or
+photoperiod) is one of the most predictable environmental cues, and
+genetic sensitivity to photoperiod protects plants from potentially
+fatal consequences of phenological responses to temperature cues at the
+``wrong'' time of year. However, the usefulness of specific
+environmental cues depends on both features of the environment, such as
+cue predictability and relevance, and the species' adaptive strategies
+(4). Species with wide natural distributions can have multiple distinct
+environmentally-cued phenological responses: for example, populations of
+sunflower (Helianthus annuus) exhibit day-neutral, facultative short
+day, and facultative long-day flowering responses, which vary with their
 environments (5, 6). Distinct genetic responses in different
@@
 
-Flowering time, in particular, is a common subject of GxE research
-(5--11), a key output of selection driving adaptation to local
-environments (2, 12, 13), and a selection target for crop improvement to
-adapt crops to local or future environments (14). Changing flowering
-responsiveness to photoperiod cues has allowed geographic range
-expansion and increased yields in a number of cereal species (15--19)
-and other crops (20, 21). Recent statistical advances in studying
-phenological GxE have involved determining critical environmental
-indices before the phenological event occurs, such as photothermal time
-within a critical growth window (10). However, most studies of flowering
-GxE focus on finding a single, best fitting form of genotype-environment
-covariance, despite the key expectation that different genetic
-subpopulations, and even different genomic regions, have likely evolved
-distinct patterns of GxE. Additionally, despite theoretical predictions
-that local adaptation should involve antagonistic pleiotropy, or
-sign-changing GxE, at the level of individual loci (22--25), previous
-work has found limited evidence of antagonistic pleiotropy (12, 26), but
-has been limited by a known statistical bias that reduced detection of
-genetic effects that differ in sign (26--28). Thus, despite substantial
-interest in the frequencies of various forms of GxE, the prevalence of
-antagonistic pleiotropy relative to other forms of GxE remains unknown.
+Flowering time, or the transition from vegetative to reproductive
+growth, is a common subject of GxE research (5--11), a key output of
+selection driving adaptation to local environments (2, 12, 13), and a
+selection target for crop improvement to adapt crops to local or future
+environments (14). Changing flowering responsiveness to photoperiod cues
+has allowed geographic range expansion and increased yields in a number
+of cereal species (15--19) and other crops (20, 21). Recent statistical
+advances in studying phenological GxE have involved determining critical
+environmental indices before the phenological event occurs, such as
+photothermal time within a critical growth window (10). However, most
+studies of flowering GxE focus on finding a single, best fitting form of
+genotype-environment covariance, despite the key expectation that
+different genetic subpopulations, and even different genomic regions,
+have likely evolved distinct patterns of GxE. Additionally, despite
+theoretical predictions that local adaptation should involve
+antagonistic pleiotropy, or sign-changing GxE, at the level of
+individual loci (22--25), previous work has found limited evidence of
+antagonistic pleiotropy (12, 26). However, this work has been limited by
+a known statistical bias that reduced detection of genetic effects that
+differ in sign (26--28). Thus, despite substantial interest in the
+frequencies of various forms of GxE, the prevalence of antagonistic
+pleiotropy relative to other forms of GxE remains unknown.
 
@@
 by plant latitude of origin (30, 31). We previously found divergent
-Midwest and Gulf genetic subpopulations of switchgrass which segregate
-for distinct sets of environmental adaptations, based on the analysis of
-two measures of fitness (i.e.,biomass and survival) among 732 genotypes
-(32). The Midwest genetic subpopulation is primarily composed of
-individuals from the well-studied upland switchgrass ecotype (33, 34),
-while the Gulf subpopulation has individuals from the lowland ecotype
-and the phenotypically intermediate coastal ecotype (32).
-
-Here, we test how the Midwest and Gulf subpopulations differ in their
-phenological adaptations across environments, and hence their
-phenological GxE. We phenotype a diversity panel of hundreds of
-switchgrass genotypes from the Midwest and Gulf subpopulations for the
-timing of vegetative development (``green-up'') and reproductive
-development (flowering) at eight common garden locations spanning 17
+Midwest and Gulf genetic subpopulations of switchgrass with distinct
+sets of environmental adaptations, in that both populations had distinct
+genetic variation associated with each of two fitness proxies, biomass
+and survival (32). The Midwest genetic subpopulation is primarily
+composed of individuals from the well-studied upland switchgrass ecotype
+(33, 34), while the Gulf subpopulation has individuals from the
+well-studied lowland ecotype and the phenotypically intermediate coastal
+ecotype (32).
+
+Here, we estimate the ways genetic effects on phenological timings
+depend on weather prior to the phenological event. To do this, we
+phenotype a diversity panel of hundreds of switchgrass genotypes from
+the Midwest and Gulf subpopulations for the timing of vegetative and
+reproductive development at eight common garden locations spanning 17
 degrees of latitude. These gardens cover the majority of the latitudinal
@@
 picture to date of the environmental variation this species encounters.
-After single-site estimation of SNP effects for these phenotypes, we use
-multivariate adaptive shrinkage (mash) to jointly estimate SNP effects
-in each subpopulation and phenotype at all eight sites (35). Mash allows
-us to identify and specify multiple ways genetic effects linked to SNPs
-may covary with the environment, and does not have a statistical bias in
-detecting SNP effects with the same or opposite signs (36). To confirm
-our genetic mapping of GxE, we compare to mapping results from an
-outbred pseudo-F2 cross grown at the same sites. Taken together, our
-results allow us to describe the environmental cues and genetic
-variation affecting phenology in two divergent natural populations of
-switchgrass.
+We define a number of ways phenological timings might covary with
+weather (@tbl-covar) and additional ways phenological timings might vary
+by site (@SI-info), then jointly re-estimate genetic effects on these
+timings at all eight sites using the set of these covariance types that
+significantly improved the modeled log-likelihood when included
+(@SI-info) (35). The method we use for effect re-estimation, mash,
+allows us to identify and specify multiple ways genetic effects may
+covary with the environment, and does not have a statistical bias in
+detecting genetic effects with the same or opposite signs (36). To
+confirm our genetic mapping of GxWeather, we compare our re-estimated
+genetic effects to mapping results from an outbred pseudo-F2 cross grown
+at the same sites. Our analysis allow us to describe the weather cues
+and genetic variation affecting phenology in two divergent natural
+populations of switchgrass.
 
@@
 
-In our diversity panel of tetraploid switchgrass (32), genotypes from
-the Gulf and Midwest genetic subpopulations had distinct phenological
-timings and distinct patterns of phenological correlations across our
-eight common garden sites (Figure~\ref{fig-map}). At the three Texas
-common gardens (hereafter `Texas' gardens), located within the natural
-range of the Gulf subpopulation, Gulf green-up occurred before
-Midwestern green-up, and Gulf flowering occurred after Midwestern
-flowering (Figure~\ref{fig-map} A). At the four northernmost common
-gardens (hereafter `North' gardens), located within the natural range of
-the Midwest subpopulation, both Gulf green-up and flowering occurred
-after Midwest green-up and flowering. At the Oklahoma common garden,
-located near the natural range limits of both the Gulf and the Midwest
-subpopulations, Gulf and Midwest green-up occurred over the same time
-period. These patterns led to strong negative phenotypic correlations
-for green-up between the North and Texas gardens, particularly in the
-Gulf and Both subpopulations, and contributed to positive phenotypic
-correlations for flowering time of larger magnitude at more northern
+Genotypes from the Gulf and Midwest subpopulations had distinct
+phenological timings and distinct patterns of phenological correlations
+across our eight common garden sites (Figure~\ref{fig-map}). At the
+three Texas common gardens (hereafter `Texas' gardens), located within
+the natural range of the Gulf subpopulation, Gulf vegetative growth
+occurred before Midwestern vegetative growth, and Gulf flowering
+occurred after Midwestern flowering (Figure~\ref{fig-map} A). At the
+four northernmost common gardens (hereafter `North' gardens), located
+within the natural range of the Midwest subpopulation, both Gulf
+vegetative growth and flowering occurred after Midwest vegetative growth
+and flowering. At the Oklahoma common garden, located near the natural
+range limits of both the Gulf and the Midwest subpopulations, Gulf and
+Midwest vegetative growth occurred over the same time period, and Gulf
+flowering occurred after Midwestern flowering (Figure~\ref{fig-map} A).
+These patterns led to strong negative phenotypic correlations for the
+timing of vegetative growth between the North and Texas gardens,
+particularly in the Gulf and across all individuals (hereafter, `Both'
+subpopulations), and contributed to positive phenotypic correlations for
+the timing of flowering that had larger magnitudes at more northern
 gardens (Figure~\ref{fig-map} B).
@@
 rank-changing GxE for these phenotypes was present across the common
-gardens (Fig 1C). h\textsuperscript{2} were typically high at individual
-gardens: 59\% on average for green-up date, and 87\% for flowering date.
-However, h2 were variable across gardens, and green-up dates were
-uncorrelated (r2 \textless{} 0.2) or negatively correlated between pairs
-of gardens (Fig. 1B). These negative and small correlations undoubtedly
-contributed to the low h2 values for green-up and flowering date when
-estimated jointly at all eight gardens: h2 was 0.8\% for green-up and
-23.2\% for flowering date.
-
-\subsubsection{Mapping major patterns of genotype-by-environment effects
-on green-up and
-flowering}\label{mapping-major-patterns-of-genotype-by-environment-effects-on-green-up-and-flowering}
-
-We next explored how genetic variation in phenology covaried with
-environmental cues by using mash to jointly discover trait-associated
-SNPs across all common gardens for the Midwest, Gulf, and `Both'
-subpopulations (Fig 1D, SI Appendix, Datasets 1-6). We first estimated
-SNP effects using GWAS on site-specific BLUPs for each phenotype. We
-then used mash on two subsets of these SNP effects, first to determine
-the covariance structures that significantly improved the mash model log
-likelihood when included, then to re-estimate the set of SNP effects
-with the largest effect estimates in univariate GWAS (hereafter
-``strong'' effects) using the set of significant covariance matrices.
-
-Three types of covariance matrices were tested for inclusion:
-``canonical'' covariance matrices, with simple patterns of effects;
-``data-driven'' matrices derived from common patterns of SNP effects in
-the data, and ``hypothesis-based'' covariance matrices that captured the
-covariance of weather-based phenological cues for cloned genotypes grown
-at multiple common gardens (Table 1, SI Appendix, Section S2). Mash
-first assigns mixture proportions for each SNP onto each provided
-covariance matrix using maximum likelihood. Then, in a second step, each
-SNP's effect estimate undergoes shrinkage such that it is more aligned
-with the covariance structure inferred in the first step. The
-hypothesis-based covariance matrices are an important advantage mash
-offers for studying patterns of GxE: they allow hypothesis testing of
-specific environmental drivers for each SNP effect, and loadings on
-these matrices provide information on genome-wide patterns of
-SNP-environment interaction. Say that a SNP in FLC controls flowering in
-a temperature-dependent manner. In that case, that SNP effect could have
-a high mixture proportion, or mass, on a covariance matrix created using
-a temperature-based environmental cue, such as average temperature on
-the day prior to flowering. In our data, we would infer that the effect
-of that SNP on flowering was caused by a response to that or a
-correlated environmental cue.
-
-The phenotypic correlations for green up date had moderate negative
-correlations between the Texas and North gardens, particularly in the
-Gulf and in Both subpopulations (Fig 1B). If these phenotypic
-correlations have a genetic component, then they might be controlled by
-SNP effects that differed in sign across these regions, SNP effects that
-are large in one garden or region and non-significant in others, or a
-combination of these patterns. Five hypothesis-based covariance matrices
-were selected by the greedy mash algorithm, one to two per
-subpopulation. Of these five, three had mass on them in mash models of
-the strong effects (Fig 2A, 2B). Two of these three matrices had
-negative covariances between sites between Texas and North gardens (Fig
-2A), while one had all positive or near-zero covariances. SNP-associated
-phenotypic effects covaried with different weather-based cues in the
-Gulf \& in Both subpopulations (Fig 2B). In total, 65\% of the posterior
-weight of strong SNP effects in the mash model of Gulf green-up fell on
-a covariance matrix of daylength 14 days prior to the date of green-up.
-The covariance matrix for this weather cue was very similar to the
-pattern of phenotypic correlation for green-up in the Gulf subpopulation
-(Fig 1B; Fig 2A). Mash models of Midwest and Both subpopulation green-up
-did not include this covariance matrix; the Midwest had no weight on any
-hypothesis-based matrix, while Both subpopulations had non-zero weights
-on two additional hypothesis-based matrices, average temperature one day
-prior to green-up, and the day length change in seconds in the day prior
-to green-up (Fig 2B). The average temperature covariance matrix had
-negative covariances between Texas and North gardens, though not as
+gardens (Figure~\ref{fig-map} C). h\textsuperscript{2} were typically
+high at individual gardens: 59\% on average for green-up date, and 87\%
+for flowering date. However, h\textsuperscript{2} were variable across
+gardens, and green-up dates were uncorrelated (r\textsuperscript{2}
+\textless{} 0.2) or negatively correlated between pairs of gardens
+(Figure~\ref{fig-map} B). These negative and small correlations
+undoubtedly contributed to the low h\textsuperscript{2} values for
+green-up and flowering date when estimated jointly at all eight gardens:
+h\textsuperscript{2} was 0.8\% for green-up and 23.2\% for flowering
+date.
+
+\subsubsection{GxWeather mixtures for vegetative and reproductive
+timing}\label{gxweather-mixtures-for-vegetative-and-reproductive-timing}
+
+We next looked for evidence of GxWeather by jointly re-estimating
+genetic effects across all eight common gardens using a model that
+allows effects to covary between gardens in many ways
+(Figure~\ref{fig-map} D, SI Appendix, Datasets 1-6). To do this, we used
+genome-wide association (GWAS) on site-specific BLUPs for the timing of
+vegetative growth and flowering to determine preliminary single
+nucleotide polymorphism (SNP) effect estimates at each garden. We then
+jointly modeled genetic effect estimates, and the mixture of ways these
+effects might covary, across all eight common gardens, using a two-part
+procedure. We first modeled genetic effects with different combinations
+of covariance types, using a subset of effects selected at random from
+all SNPs used in the GWAS. In these models, we specified many ways
+genetic effects could covary between gardens (as expanded on below) and
+used a greedy algorithm to iteratively select covariance structures that
+significantly improved the log likelihood of a model when included.
+Given the high correlation between some of the covariance patterns we
+tested, this approach allowed us to select only the subset of patterns
+of covariation that most improved the fit of the joint effect patterns
+present in the data. We then re-estimated the set of genetic effects
+that had the largest effect estimates in univariate GWAS (hereafter
+``strong'' effects) using the set of significant covariance matrices
+that most improved the model fit on the random effects.
+
+We created three categories of ways effects could covary across gardens:
+``canonical'' covariance, with simple patterns of effects;
+``data-driven'' covariances derived from common patterns of SNP effects
+in the data, and ``GxWeather'' covariances that captured the covariance
+of weather variables at specific times before the phenological event
+(Table 1, SI Appendix, Section S2). For example, GxWeather covariance
+between gardens in MI and MO, for rainfall in the seven days prior to
+flowering, captures the covariance in the amount of rainfall genotype A,
+B, and C receive in the seven days prior to their three flowering dates
+in MI and their three flowering dates in MO. If these six rainfall
+values are similar, covariance will be near one; if they are
+uncorrelated, covariance will be near zero; if there is sufficient rank
+changing in rainfall amounts between gardens for these genotypes, this
+covariance will be near negative one.
+
+The GxWeather covariances allow hypothesis testing of specific weather
+variables as cues for each genetic effect. Say that a SNP in FLC
+controls flowering in a photoperiod-dependent manner. In that case, that
+SNP effect could have a high mixture proportion, or mass, on a
+covariance matrix created using a photoperiod-based environmental cue,
+such as day length two days prior to flowering. In our data, we would
+infer that the effect of that SNP on flowering was caused by a response
+to that or a correlated environmental cue. Overall, loadings of genetic
+effects on these GxWeather matrices in the model with re-estimated
+effects provide information on genome-wide patterns of SNP-environment
+interaction.
+
+The phenotypic correlations for the timing of vegetative growth had
+moderate negative correlations between the Texas and North gardens,
+particularly in the Gulf and in Both subpopulations
+(Figure~\ref{fig-map} B). If these phenotypic correlations have a
+genetic component, then they might be controlled by SNP effects that
+differed in sign across these regions (effects with antagonistic
+pleiotropy), SNP effects that are large in one garden or region and
+non-significant in others (effects with amplification), or a combination
+of these patterns. Five GxWeather covariance matrices were selected by
+the greedy mash algorithm, one to two per subpopulation. Of these five,
+three had mass on them in mash models of the strong effects
+(Figure~\ref{fig-covar} A-B). Two of these three matrices had negative
+covariances between sites between Texas and North gardens
+(Figure~\ref{fig-covar} A), while one had all positive or near-zero
+covariances. SNP-associated phenotypic effects covaried with different
+weather-based cues in the Gulf \& in Both subpopulations
+(Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
+strong SNP effects in the mash model of Gulf vegetative growth fell on a
+covariance matrix of daylength 14 days prior to the date of vegetative
+growth. The covariance matrix for this weather cue was very similar to
+the pattern of phenotypic correlation for the timing of vegetative
+growth in the Gulf subpopulation (Figure~\ref{fig-map} B;
+Figure~\ref{fig-covar} A). Mash models of the timing of vegetative
+growth in Midwest and Both subpopulations did not include this GxWeather
+covariance; the Midwest had no weight on any GxWeather covariance, while
+Both subpopulations had non-zero weights on two additional GxWeather
+covariance types, average temperature one day prior to green-up, and the
+day length change in seconds in the day prior to green-up
+(Figure~\ref{fig-covar} B). The average temperature covariance matrix
+had negative covariances between Texas and North gardens, though not as
 strong as the negative phenotypic correlations seen in Both
-subpopulations (Fig 1B; Fig 2A). Only the Gulf subpopulation and Both
-subpopulations had mass on any hypothesis-based covariance matrices (Fig
-2C).
-
-For flowering date, distinct weather-based cues captured SNP-associated
-effect patterns in the Gulf and Midwest subpopulations. 33\% of SNP
-effects on flowering in the Gulf subpopulation covaried with cumulative
-rainfall in the seven days prior to flowering (Fig 2D). 22.6\% of SNP
-effects on flowering in the Midwest subpopulation contrived with day
-length change in the two days prior to flowering (Fig 2D), which had
-negative covariances between Texas and North gardens. Neither covariance
-matrix was selected for in Both subpopulations, and no hypothesis-based
-covariance matrices had mass in Both subpopulations (Fig 2E). In five of
-the six mash models of strong effects, the hypothesis-based covariance
-matrices captured a minority of the posterior weights of the strong
-effects (Figure 2C,E); the majority of this mass was on various
+subpopulations (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Only
+the Gulf subpopulation and Both subpopulations had mass on any GxWeather
+covariance matrices (Figure~\ref{fig-covar} C).
+
+For flowering date, distinct GxWeather covariances captured
+SNP-associated effect patterns in the Gulf and Midwest subpopulations.
+33\% of SNP effects on flowering in the Gulf subpopulation covaried with
+cumulative rainfall in the seven days prior to flowering (Fig 2D).
+22.6\% of SNP effects on flowering in the Midwest subpopulation covaried
+with day length change in the two days prior to flowering (Fig 2D),
+which had negative covariances between Texas and North gardens. Neither
+covariance matrix was selected for in Both subpopulations, and no
+GxWeather covariance matrices had mass in Both subpopulations (Fig 2E).
+In five of the six mash models of strong effects, the GxWeather
+covariance matrices captured a minority of the posterior weights of the
+strong effects (Figure 2C,E); the majority of this mass was on various
 canonical covariance matrices. These matrices included simple
@@
 
-\subsubsection{Confirmation of genotype-by-environment effects using an
-independent mapping
-population}\label{confirmation-of-genotype-by-environment-effects-using-an-independent-mapping-population}
-
-We sought additional experimental support for our mash intervals using
-an independent pseudo-F2 mapping population created from Gulf \& Midwest
-individuals and grown at the same sites (Fig. 4A,B). We conducted
-quantitative trait loci (QTL) mapping of flowering as functions of four
-environmental cues that we also used as covariance matrices in mash, and
-identified eight QTL for flowering date, six QTL for flowering GDD, ten
-QTL for flowering day length, and eight QTL flowering day length change,
-all of which showed QTL by environment interactions (SI Appendix, Fig.
-S3). All QTL for flowering overlapped one or more homologs from rice or
-A. thaliana with functionally validated roles in flowering (SI Appendix,
-Dataset 7). All flowering QTL intervals contained at least one SNP
-significant in at least one mash run at a log10-transformed Bayes Factor
-\textgreater{} 2, or in the 1\% tail of significance, whichever was
-stricter (SI Appendix, Dataset 8). We also looked for enrichments of
-mash SNPs in the 1\% tail of significance (the `mash 1\% tail') within
-each QTL interval. At the 5\% level, five QTL had enrichments of SNPs in
-the mash 1\% tail. Overall, there were eleven significant enrichments (p
-\textless{} 0.05, hypergeometric test) of SNPs in the mash 1\% tail in
-the QTL intervals. Our QTL intervals had more enrichments of SNPs in the
-mash 1\% tail than were found for all but eleven of these sets of random
-genomic intervals (Fig. 4C, p = 0.011). Thus, we were able to
-experimentally support our mash intervals with a QTL mapping experiment
-using a separate mapping population.
+\subsubsection{Confirmation of effects on phenology using an independent
+mapping
+population}\label{confirmation-of-effects-on-phenology-using-an-independent-mapping-population}
+
+We sought additional experimental support for our significant effect
+re-estimates using an independent pseudo-F2 mapping population created
+from Gulf \& Midwest individuals and grown at the same sites (Fig.
+4A,B). We conducted quantitative trait loci (QTL) mapping of flowering
+as functions of four environmental cues that we also used as covariance
+matrices in mash, and identified eight QTL for flowering date, six QTL
+for flowering GDD, ten QTL for flowering day length, and eight QTL
+flowering day length change, all of which showed QTL by environment
+interactions (SI Appendix, Fig. S3). All QTL for flowering overlapped
+one or more homologs from rice or \emph{A. thaliana} with functionally
+validated roles in flowering (SI Appendix, Dataset 7). All flowering QTL
+intervals contained at least one SNP significant in at least one mash
+run at a log10-transformed Bayes Factor \textgreater{} 2, or in the 1\%
+tail of significance, whichever was stricter (SI Appendix, Dataset 8).
+We also looked for enrichments of mash SNPs in the 1\% tail of
+significance (the `mash 1\% tail') within each QTL interval. At the 5\%
+level, five QTL had enrichments of SNPs in the mash 1\% tail. Overall,
+there were eleven significant enrichments (p \textless{} 0.05,
+hypergeometric test) of SNPs in the mash 1\% tail in the QTL intervals.
+Our QTL intervals had more enrichments of SNPs in the mash 1\% tail than
+were found for all but eleven of these sets of random genomic intervals
+(Fig. 4C, p = 0.011). Thus, we were able to experimentally support our
+significant effect re-estimates with a QTL mapping experiment using a
+separate mapping population.
 
@@
 where this covariation can be altered. Here, we demonstrate that we can
-associate multiple patterns of GxE with specific genomic regions using a
-switchgrass diversity panel grown at eight common gardens, and also that
-we can assign specific SNP-associated patterns of GxE to both
-weather-based cues and to other, data-driven patterns. We use this
-approach to study GxE in both green-up and flowering phenological data
-in the deeply genetically diverged Gulf and Midwest populations of
-switchgrass.
-
-Our analysis of green-up in the Gulf and in Both subpopulations revealed
-substantial antagonistic pleiotropy in effects between the Texas and
-North gardens (Figure 3A). This result supports theoretical models that
-local adaptation should involve antagonistic pleiotropy at the level of
-individual loci (22--25), and is the first experimental work using QTL
-mapping and GWAS across common gardens to find antagonistic pleiotropy
-to be common in small genomic regions (12, 36, 37).
-
-Our analysis of flowering showed that the Gulf and Midwest
-subpopulations have two distinct photoperiod-related flowering
-responses: the Midwest subpopulation is day neutral, and flowering is
-cued primarily by a cumulative GDD threshold; in contrast, the Gulf
-subpopulation is photoperiod sensitive, and flowering is cued by the
-transition to shortening days. This result was supported by observations
-that expressing flowering date as a function of the day length at
-flowering increased its heritability in the Gulf subpopulation, while
-expressing flowering date as a function of cumulative GDD between
-green-up and flowering increased the heritability of flowering in the
-Midwest subpopulation (Fig. 1D). The genomic regions affecting flowering
-found in mash were also supported by QTL from an independent mapping
-population (Fig. 4C).
+associate multiple patterns of GxWeather with specific genomic regions
+using a switchgrass diversity panel grown at eight common gardens. We
+can assign genetic effects to both GxWeather patterns with interpretable
+weather-based cues, and to agnostic, site-based patterns. We use this
+approach to study GxWeather for the timings of vegetative and
+reproductive development in the deeply genetically diverged Gulf and
+Midwest subpopulations of switchgrass.
+
+Our analysis of the timing of vegetative growth in the Gulf and in Both
+subpopulations revealed substantial antagonistic pleiotropy in effects
+between the Texas and North gardens (Figure 3A). This result supports
+theoretical models that local adaptation should involve antagonistic
+pleiotropy at the level of individual loci (22--25), and is the first
+experimental work using QTL mapping and GWAS across common gardens to
+find antagonistic pleiotropy to be common in small genomic regions (12,
+37, 38).
+
+Our analysis of the timing of flowering showed that the Gulf and Midwest
+subpopulations have distinct GxWeather: flowering timing in the Midwest
+subpopulation has photoperiod-related variation, in that flowering
+timing covaries with a day length change signal in the two days before
+flowering occurs. In contrast, the Gulf subpopulation does not have
+variation in flowering based on photoperiod. Instead, the Gulf
+subpopulation has variation in flowering that covaries with the rainfall
+that occurs in the week prior to flowering. The genomic regions
+affecting flowering that we re-estimated across all eight sites were
+also supported by QTL from an independent mapping population at these
+sites (Fig. 4C).
 
@@
 
-\caption{\label{fig-map}Figure 1. Characterization of green-up and
-flowering dates from the switchgrass diversity panel. (a) Map and trait
-histograms of green-up and flowering dates across two genetically
-distinct switchgrass subpopulations and eight common gardens. Purple
-represents individuals from the Midwest genetic subpopulation, and pink
-individuals from the Gulf subpopulation. Vertical dashed lines indicate
-the summer solstice. Common gardens are arranged in latitudinal order.
-(b) Phenotypic correlations between clonal replicates planted at eight
-common gardens, within and between two genetic subpopulations. (c)
-Narrow sense heritability of green-up and flowering within single common
-gardens (purple) and across all eight common gardens (green), within and
-between two genetic subpopulations. (d) Flow diagram of the methods
-applied to the green-up and flowering dates to jointly estimate SNP
-effects across all sites. Mash was fit to SNP effect data and used to
-find covariance matrices that improved the mash model likelihood using a
-large set of randomly selected, relatively unlinked SNP effects; this
-model was applied to a ``strong'' set of SNP effects with large effect
-sizes in the univariate GWAS.}
+\caption{\label{fig-map}Characterization of green-up and flowering dates
+from the switchgrass diversity panel. (a) Map and trait histograms of
+green-up and flowering dates across two genetically distinct switchgrass
+subpopulations and eight common gardens. Purple represents individuals
+from the Midwest genetic subpopulation, and pink individuals from the
+Gulf subpopulation. Vertical dashed lines indicate the summer solstice.
+Common gardens are arranged in latitudinal order. (b) Phenotypic
+correlations between clonal replicates planted at eight common gardens,
+within and between two genetic subpopulations. (c) Narrow sense
+heritability of green-up and flowering within single common gardens
+(purple) and across all eight common gardens (green), within and between
+two genetic subpopulations. (d) Flow diagram of the methods applied to
+the green-up and flowering dates to jointly estimate SNP effects across
+all sites. Mash was fit to SNP effect data and used to find covariance
+matrices that improved the mash model likelihood using a large set of
+randomly selected, relatively unlinked SNP effects; this model was
+applied to a ``strong'' set of SNP effects with large effect sizes in
+the univariate GWAS.}
 
@@
 
-\caption{\label{fig-covar}Figure 2.}
+\caption{\label{fig-covar}Caption for this figure\ldots{}}
 
@@
 on green-up and
-flowering}\label{mapping-major-patterns-of-genotype-by-environment-effects-on-green-up-and-flowering-1}
+flowering}\label{mapping-major-patterns-of-genotype-by-environment-effects-on-green-up-and-flowering}
 
@@
 independent mapping
-population}\label{confirmation-of-genotype-by-environment-effects-using-an-independent-mapping-population-1}
+population}\label{confirmation-of-genotype-by-environment-effects-using-an-independent-mapping-population}
 
@@
 
+\bibitem[\citeproctext]{ref-savolainen2013ecological}
+\CSLLeftMargin{37. }%
+\CSLRightInline{O. Savolainen, M. Lascoux, J. Merilä, Ecological
+genomics of local adaptation. \emph{Nature Reviews Genetics}
+\textbf{14}, 807--820 (2013).}
+
+\bibitem[\citeproctext]{ref-lowry2019qtl}
+\CSLLeftMargin{38. }%
+\CSLRightInline{D. B. Lowry, \emph{et al.}, QTL\(\times\) environment
+interactions underlie adaptive divergence in switchgrass across a large
+latitudinal gradient. \emph{Proceedings of the National Academy of
+Sciences} \textbf{116}, 12933--12941 (2019).}
+
 \end{CSLReferences}

### 2024-01-18 manuscript/revision2/revision2.qmd
manuscript: revisions to text from SAS from 20240110, some small additional revisions from AHM

@@
 However, the usefulness of specific environmental cues depends on both features of the environment, such as cue predictability and relevance, and the species' adaptive strategies @botero2015evolutionary.
-Species with wide natural distributions can have multiple distinct environmentally-cued phenological responses: for example, populations of sunflower (Helianthus annuus) exhibit day-neutral, facultative short day, and facultative long-day flowering responses, which vary with their environments [@blackman2013interacting; @henry2014transitions].
+Species with wide natural distributions can have multiple distinct environmentally cued phenological responses: for example, populations of sunflower (Helianthus annuus) exhibit day-neutral, facultative short day, and facultative long-day flowering responses, which vary with their environments [@blackman2013interacting; @henry2014transitions].
 Distinct genetic responses in different environments are known as genotype by environment interactions, or GxE.
@@
 Flowering time, or the transition from vegetative to reproductive growth, is a common subject of GxE research [@blackman2013interacting; @henry2014transitions; @aagren2017adaptive; @brachi2010linkage; @dittmar2014flowering; @li2018genomic; @romero2017study], a key output of selection driving adaptation to local environments [@andres2012genetic; @wadgymar2017identifying; @blumel2015flowering], and a selection target for crop improvement to adapt crops to local or future environments @jung2009flowering.
-Changing flowering responsiveness to photoperiod cues has allowed geographic range expansion and increased yields in a number of cereal species [@turner2005pseudo; @faure2012mutation; @hung2012zmcct; @zakhrabekova2012induced; @yang2013oself3] and other crops [@pin2012multifaceted; @weller2019parallel].
+Changing flowering responsiveness to photoperiod cues has allowed geographic range expansion and increased yields in several cereal species [@turner2005pseudo; @faure2012mutation; @hung2012zmcct; @zakhrabekova2012induced; @yang2013oself3] and other crops [@pin2012multifaceted; @weller2019parallel].
 Recent statistical advances in studying phenological GxE have involved determining critical environmental indices before the phenological event occurs, such as photothermal time within a critical growth window @li2018genomic.
@@
 These gardens cover the majority of the latitudinal and climatic range of switchgrass and capture the most comprehensive picture to date of the environmental variation this species encounters.
-We define a number of ways phenological timings might covary with weather (\@tbl-covar) and additional ways phenological timings might vary by site (\@SI-info), then jointly re-estimate genetic effects on these timings at all eight sites using the set of these covariance types that significantly improved the modeled log-likelihood when included (\@SI-info) @urbut2019flexible.
-The method we use for effect re-estimation, mash, allows us to identify and specify multiple ways genetic effects may covary with the environment, and does not have a statistical bias in detecting genetic effects with the same or opposite signs @10.1093/biostatistics/kxw041.
-To confirm our genetic mapping of GxWeather, we compare our re-estimated genetic effects to mapping results from an outbred pseudo-F2 cross grown at the same sites.
-Our analysis allow us to describe the weather cues and genetic variation affecting phenology in two divergent natural populations of switchgrass.
+We define multiple ways phenological timings might covary with weather (\@tbl-covar) and additional ways phenological timings might vary by site (\@SI-info), then jointly re-estimate genetic effects on these timings at all eight sites using the set of these covariance types that significantly improved the modeled log-likelihood when included (\@SI-info) @urbut2019flexible.
+We use the Bayesian framework *mash* (multivariate adaptive shrinkage), developed by @urbut2019flexible, to refine effect size estimates from genome-wide association (GWAS) conducted on individuals at each of the eight sites.
+*mash* allows us to identify and specify multiple covariance structures among genetic effect estimates across sites, including structures that represent covariance in weather variables of interest.
+Importantly, *mash* does not have a statistical bias in detecting genetic effects with the same or opposite signs @10.1093/biostatistics/kxw041.
+To confirm our genetic mapping of GxWeather, we compare our posterior genetic effect estimates from *mash* to mapping results from an outbred pseudo-F2 cross grown at the same sites.
+Our analyses allow us to describe the weather cues and genetic variation affecting phenology in two divergent natural populations of switchgrass.
 
@@
 At the four northernmost common gardens (hereafter 'North' gardens), located within the natural range of the Midwest subpopulation, both Gulf vegetative growth and flowering occurred after Midwest vegetative growth and flowering.
-At the Oklahoma common garden, located near the natural range limits of both the Gulf and the Midwest subpopulations, Gulf and Midwest vegetative growth occurred over the same time period, and Gulf flowering occurred after Midwestern flowering (@fig-map A).
+At the Oklahoma common garden, located near the natural range limits of both the Gulf and the Midwest subpopulations, Gulf and Midwest vegetative growth occurred over the same period, and Gulf flowering occurred after Midwestern flowering (@fig-map A).
 These patterns led to strong negative phenotypic correlations for the timing of vegetative growth between the North and Texas gardens, particularly in the Gulf and across all individuals (hereafter, 'Both' subpopulations), and contributed to positive phenotypic correlations for the timing of flowering that had larger magnitudes at more northern gardens (@fig-map B).
@@
 
-### GxWeather mixtures for vegetative and reproductive timing
+### Inference of GxWeather effects for vegetative and reproductive timing using *mash*
 
-We next looked for evidence of GxWeather by jointly re-estimating genetic effects across all eight common gardens using a model that allows effects to covary between gardens in many ways (@fig-map D, SI Appendix, Datasets 1-6).
-To do this, we used genome-wide association (GWAS) on site-specific BLUPs for the timing of vegetative growth and flowering to determine preliminary single nucleotide polymorphism (SNP) effect estimates at each garden.
+We next looked for evidence of GxWeather by jointly re-estimating genetic effects across all eight common gardens using the mash model, which allows effects to covary between gardens in many ways (@fig-map D, SI Appendix, Datasets 1-6).
+To do this, we used GWAS on site-specific BLUPs for the timing of vegetative growth and flowering to determine preliminary single nucleotide polymorphism (SNP) effect estimates for both traits at each garden.
 We then jointly modeled genetic effect estimates, and the mixture of ways these effects might covary, across all eight common gardens, using a two-part procedure.
-We first modeled genetic effects with different combinations of covariance types, using a subset of effects selected at random from all SNPs used in the GWAS.
-In these models, we specified many ways genetic effects could covary between gardens (as expanded on below) and used a greedy algorithm to iteratively select covariance structures that significantly improved the log likelihood of a model when included.
-Given the high correlation between some of the covariance patterns we tested, this approach allowed us to select only the subset of patterns of covariation that most improved the fit of the joint effect patterns present in the data.
-We then re-estimated the set of genetic effects that had the largest effect estimates in univariate GWAS (hereafter "strong" effects) using the set of significant covariance matrices that most improved the model fit on the random effects.
+We first modeled genetic effects with different combinations of covariance structures, using a subset of randomly selected SNPs used in the GWAS and their corresponding effect size estimates.
+In these models (hereafter referred to as "random"), we then specified a variety of potential covariance structures among genetic effect size estimates from each garden (as expanded on below) and used a greedy algorithm to iteratively select and add covariance structures that significantly improved the log likelihood of the *mash* model.
+Given the high correlation between some of the covariance patterns we tested, this approach allowed us to select only the subset of patterns of covariance that significantly improved the fit of the jointly inferred structure of genetic effect estimates present in the data.
+We then selected the set of SNPs and corresponding effect size estimates that had the largest effect estimates in univariate GWAS (hereafter "strong" effects) and jointly re-estimated these effect size estimates at all eight sites using the set of significant covariance matrices that most improved the model fit on the random effects.
 
-We created three categories of ways effects could covary across gardens: "canonical" covariance, with simple patterns of effects; "data-driven" covariances derived from common patterns of SNP effects in the data, and "GxWeather" covariances that captured the covariance of weather variables at specific times before the phenological event (Table 1, SI Appendix, Section S2).
-For example, GxWeather covariance between gardens in MI and MO, for rainfall in the seven days prior to flowering, captures the covariance in the amount of rainfall genotype A, B, and C receive in the seven days prior to their three flowering dates in MI and their three flowering dates in MO.
+We specified three categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event (Table 1, SI Appendix, Section S2).
+For example, GxWeather covariance between gardens in MI and MO, for rainfall in the seven days prior to flowering, captures the covariance in the amount of rainfall genotypes A, B, and C receive in the seven days prior to their three flowering dates in MI and their three flowering dates in MO.
 If these six rainfall values are similar, covariance will be near one; if they are uncorrelated, covariance will be near zero; if there is sufficient rank changing in rainfall amounts between gardens for these genotypes, this covariance will be near negative one.
+We specified multiple covariance structures for the canonical, data-driven, and GxWeather categories, each of which represented a specific pattern of covariance in genetic effect size estimates that could be included in the model.
 
-The GxWeather covariances allow hypothesis testing of specific weather variables as cues for each genetic effect.
+The GxWeather covariance structures allow hypothesis testing of specific weather variables as cues for the start of vegetative and reproductive growth.
 Say that a SNP in FLC controls flowering in a photoperiod-dependent manner.
-In that case, that SNP effect could have a high mixture proportion, or mass, on a covariance matrix created using a photoperiod-based environmental cue, such as day length two days prior to flowering.
-In our data, we would infer that the effect of that SNP on flowering was caused by a response to that or a correlated environmental cue.
-Overall, loadings of genetic effects on these GxWeather matrices in the model with re-estimated effects provide information on genome-wide patterns of SNP-environment interaction.
+In that case, the joint estimate of effects for that SNP could have a high mixture proportion, or mass, on a covariance matrix created using a photoperiod-based environmental cue, such as day length two days prior to flowering.
+In our data, we would infer that the effect of that SNP on flowering was caused by a response to the environmental cue used to construct the GxWeather covariance structure with the largest mass.
+Overall, loadings of genetic effects on these GxWeather matrices in the model, along with the posterior effect size estimates provide information on genome-wide patterns of SNP-environment interaction.
 
-The phenotypic correlations for the timing of vegetative growth had moderate negative correlations between the Texas and North gardens, particularly in the Gulf and in Both subpopulations (@fig-map B).
-If these phenotypic correlations have a genetic component, then they might be controlled by SNP effects that differed in sign across these regions (effects with antagonistic pleiotropy), SNP effects that are large in one garden or region and non-significant in others (effects with amplification), or a combination of these patterns.
-Five GxWeather covariance matrices were selected by the greedy mash algorithm, one to two per subpopulation.
+The phenotypic correlations for the timing of vegetative growth had moderate negative correlations between the Texas and North gardens, particularly in the Gulf subpopulation and Both subpopulations (@fig-map B).
+If these phenotypic correlations have a genetic component, then they may be controlled by SNP effects that differed in sign across these regions (effects with antagonistic pleiotropy), SNP effects that are large in one garden or region and non-significant in others (effects with amplification), or a combination of these patterns.
+Five GxWeather covariance structures were selected by the greedy algorithm, one to two per subpopulation.
 Of these five, three had mass on them in mash models of the strong effects (@fig-covar A-B).
@@
 SNP-associated phenotypic effects covaried with different weather-based cues in the Gulf & in Both subpopulations (@fig-covar B).
-In total, 65% of the posterior weight of strong SNP effects in the mash model of Gulf vegetative growth fell on a covariance matrix of daylength 14 days prior to the date of vegetative growth.
+In total, 65% of the posterior weight of strong SNP effects in the mash model of Gulf vegetative growth fell on a covariance matrix constructed using the covariance of daylength 14 days prior to the date of vegetative growth.
 The covariance matrix for this weather cue was very similar to the pattern of phenotypic correlation for the timing of vegetative growth in the Gulf subpopulation (@fig-map B; @fig-covar A).
-Mash models of the timing of vegetative growth in Midwest and Both subpopulations did not include this GxWeather covariance; the Midwest had no weight on any GxWeather covariance, while Both subpopulations had non-zero weights on two additional GxWeather covariance types, average temperature one day prior to green-up, and the day length change in seconds in the day prior to green-up (@fig-covar B).
+Mash models of the timing of vegetative growth in the Midwest subpopulation and Both subpopulations did not include this GxWeather covariance; the Midwest had no weight on any GxWeather covariance, while Both subpopulations had non-zero weights on two additional GxWeather covariance types, average temperature one day prior to green-up, and the day length change in seconds in the day prior to green-up (@fig-covar B).
 The average temperature covariance matrix had negative covariances between Texas and North gardens, though not as strong as the negative phenotypic correlations seen in Both subpopulations (@fig-map B; @fig-covar A).
@@
 
-For flowering date, distinct GxWeather covariances captured SNP-associated effect patterns in the Gulf and Midwest subpopulations.
-33% of SNP effects on flowering in the Gulf subpopulation covaried with cumulative rainfall in the seven days prior to flowering (Fig 2D).
-22.6% of SNP effects on flowering in the Midwest subpopulation covaried with day length change in the two days prior to flowering (Fig 2D), which had negative covariances between Texas and North gardens.
-Neither covariance matrix was selected for in Both subpopulations, and no GxWeather covariance matrices had mass in Both subpopulations (Fig 2E).
-In five of the six mash models of strong effects, the GxWeather covariance matrices captured a minority of the posterior weights of the strong effects (Figure 2C,E); the majority of this mass was on various canonical covariance matrices.
+For flowering date, distinct GxWeather covariance structures captured covariance in effect sizes for SNPs in the Gulf and Midwest subpopulations.
+33% of SNP effects on flowering in the Gulf subpopulation covaried with cumulative rainfall in the seven days prior to flowering (@fig-covar D).
+22.6% of SNP effects on flowering in the Midwest subpopulation covaried with day length change in the two days prior to flowering (@fig-covar D), which had negative covariances between Texas and North gardens.
+Neither covariance matrix was selected as significantly improving the log-likelihood of the *mash* model of Both subpopulations, and no GxWeather covariance matrices had a non-zero mass in models of Both subpopulations (@fig-covar E).
+In five of the six mash models of strong effects, the GxWeather covariance matrices captured a minority of the posterior weights of the strong effects (@fig-covar C,E); the majority of this mass was on various canonical covariance matrices.
 These matrices included simple heterozygosity, with intermediate, positive covariances between all gardens, and garden-specific effects present only at one garden.
@@
 We used the local false sign rate (lfsr), an analogue of the lfdr that establishes confidence in the effect sign, not the effect's difference from zero, to determine significance.
-We required significance (p \< 0.05) in both gardens to include effects.
+We required lfsr significance (p \< 0.05) in both gardens to include effects.
 This means that our tests for antagonistic pleiotropy, or a sign change between conditions, carry an equal statistical burden to those for effects with the same sign.
-We also summarized the pairwise patterns of effect sign and magnitude within and between the Texas and North regions (Table 2).
+We also summarized the pairwise patterns of effect sign and magnitude within and between the Texas and North regions, within the natural ranges of the Gulf and Midwest subpopulations (\@fig-effects).
 
-For greenup date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, between pairs of Texas and North gardens (Fig.
-3A).
-78.7% of pairwise comparisons between North and Texas gardens had a difference in sign, while only 28.6% and 0.2% of North-North or Texas-Texas comparisons had a difference in sign, respectively (Table 2).
-The majority of pairwise effects for greenup for the Midwest (\>55%) and Both (\>85%) subpopulations were the same sign, and effects frequently differed in magnitude between the MO and OK garden and other gardens (Table 2; Fig 3A).
+For greenup date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, between pairs of Texas and North gardens (\@fig-effects A).
+78.7% of pairwise comparisons between North and Texas gardens had a difference in sign, while only 28.6% and 0.2% of North-North or Texas-Texas comparisons had a difference in sign, respectively (\@fig-effects).
+The majority of pairwise effects for greenup for the Midwest (\>55%) and Both (\>85%) subpopulations were the same sign, and effects frequently differed in magnitude between the MO and OK garden and other gardens (\@fig-effects X).
 
-For flowering date for the Gulf subpopulation, less than 2% of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, within or between regions (Table 2).
-More effects differed in magnitude between the Texas and North regions than within these regions (42.7% vs \<20%; Figure 3B).
+For flowering date for the Gulf subpopulation, less than 2% of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, within or between regions (\@fig-effects).
+More effects differed in magnitude between the Texas and North regions than within these regions (42.7% vs \<20%; \@fig-effects X).
 The Midwest population had relatively few significant effects for flowering, but a large proportion of these differed in sign between Texas and North regions (42.7%) or within the North region (65.4%).
-Finally, in Both subpopulations, less than 20% of pairwise effects differed in sign (Table 2).
-Most differences in sign were between TX1, the southernmost garden, and all other gardens (Fig 3B).
+Finally, in Both subpopulations, less than 20% of pairwise effects differed in sign (\@fig-effects).
+Most differences in sign were between TX1, the southernmost garden, and all other gardens (\@fig-effects B).
 Similarly, more effects that differed in magnitude included gardens in the Texas region (52.3-55.9%), and most effect pairs in the North region were not distinguishable (91.5%).
@@
 
-We sought additional experimental support for our significant effect re-estimates using an independent pseudo-F2 mapping population created from Gulf & Midwest individuals and grown at the same sites (Fig.
-4A,B).
+We sought additional experimental support for our joint re-estimates of SNP effects using an independent pseudo-F2 mapping population created from Gulf & Midwest individuals and grown at the same sites (\@fig-qtl A,B).
 We conducted quantitative trait loci (QTL) mapping of flowering as functions of four environmental cues that we also used as covariance matrices in mash, and identified eight QTL for flowering date, six QTL for flowering GDD, ten QTL for flowering day length, and eight QTL flowering day length change, all of which showed QTL by environment interactions (SI Appendix, Fig.
@@
 Overall, there were eleven significant enrichments (p \< 0.05, hypergeometric test) of SNPs in the mash 1% tail in the QTL intervals.
-Our QTL intervals had more enrichments of SNPs in the mash 1% tail than were found for all but eleven of these sets of random genomic intervals (Fig.
-4C, p = 0.011).
-Thus, we were able to experimentally support our significant effect re-estimates with a QTL mapping experiment using a separate mapping population.
+Our QTL intervals had more enrichments of SNPs in the mash 1% tail than were found for all but eleven of these sets of random genomic intervals (\@fig-qtl C, p = 0.011).
+Thus, we were able to experimentally support our joint re-estimates of SNP effects with a QTL mapping experiment using a separate mapping population.
 
@@
 
-Our analysis of the timing of vegetative growth in the Gulf and in Both subpopulations revealed substantial antagonistic pleiotropy in effects between the Texas and North gardens (Figure 3A).
+Our analysis of the timing of vegetative growth in the Gulf and in Both subpopulations revealed substantial antagonistic pleiotropy in effects between the Texas and North gardens (\@fig-effects A).
 This result supports theoretical models that local adaptation should involve antagonistic pleiotropy at the level of individual loci [@kawecki2004conceptual; @felsenstein1976theoretical; @levene1953genetic; @hedrick1986genetic], and is the first experimental work using QTL mapping and GWAS across common gardens to find antagonistic pleiotropy to be common in small genomic regions [@wadgymar2017identifying; @savolainen2013ecological; @lowry2019qtl].
@@
 Instead, the Gulf subpopulation has variation in flowering that covaries with the rainfall that occurs in the week prior to flowering.
-The genomic regions affecting flowering that we re-estimated across all eight sites were also supported by QTL from an independent mapping population at these sites (Fig.
-4C).
+The genomic regions affecting flowering that we re-estimated across all eight sites were also supported by QTL from an independent mapping population at these sites (\@fig-qtl C).
 

### 2024-01-18 manuscript/revision2/revision2.tex
manuscript: revisions to text from SAS from 20240110, some small additional revisions from AHM

@@
 (4). Species with wide natural distributions can have multiple distinct
-environmentally-cued phenological responses: for example, populations of
+environmentally cued phenological responses: for example, populations of
 sunflower (Helianthus annuus) exhibit day-neutral, facultative short
@@
 environments (14). Changing flowering responsiveness to photoperiod cues
-has allowed geographic range expansion and increased yields in a number
-of cereal species (15--19) and other crops (20, 21). Recent statistical
+has allowed geographic range expansion and increased yields in several
+cereal species (15--19) and other crops (20, 21). Recent statistical
 advances in studying phenological GxE have involved determining critical
@@
 picture to date of the environmental variation this species encounters.
-We define a number of ways phenological timings might covary with
-weather (@tbl-covar) and additional ways phenological timings might vary
-by site (@SI-info), then jointly re-estimate genetic effects on these
-timings at all eight sites using the set of these covariance types that
+We define multiple ways phenological timings might covary with weather
+(@tbl-covar) and additional ways phenological timings might vary by site
+(@SI-info), then jointly re-estimate genetic effects on these timings at
+all eight sites using the set of these covariance types that
 significantly improved the modeled log-likelihood when included
-(@SI-info) (35). The method we use for effect re-estimation, mash,
-allows us to identify and specify multiple ways genetic effects may
-covary with the environment, and does not have a statistical bias in
+(@SI-info) (35). We use the Bayesian framework \emph{mash} (multivariate
+adaptive shrinkage), developed by (35), to refine effect size estimates
+from genome-wide association (GWAS) conducted on individuals at each of
+the eight sites. \emph{mash} allows us to identify and specify multiple
+covariance structures among genetic effect estimates across sites,
+including structures that represent covariance in weather variables of
+interest. Importantly, \emph{mash} does not have a statistical bias in
 detecting genetic effects with the same or opposite signs (36). To
-confirm our genetic mapping of GxWeather, we compare our re-estimated
-genetic effects to mapping results from an outbred pseudo-F2 cross grown
-at the same sites. Our analysis allow us to describe the weather cues
-and genetic variation affecting phenology in two divergent natural
-populations of switchgrass.
+confirm our genetic mapping of GxWeather, we compare our posterior
+genetic effect estimates from \emph{mash} to mapping results from an
+outbred pseudo-F2 cross grown at the same sites. Our analyses allow us
+to describe the weather cues and genetic variation affecting phenology
+in two divergent natural populations of switchgrass.
 
@@
 range limits of both the Gulf and the Midwest subpopulations, Gulf and
-Midwest vegetative growth occurred over the same time period, and Gulf
+Midwest vegetative growth occurred over the same period, and Gulf
 flowering occurred after Midwestern flowering (Figure~\ref{fig-map} A).
@@
 
-\subsubsection{GxWeather mixtures for vegetative and reproductive
-timing}\label{gxweather-mixtures-for-vegetative-and-reproductive-timing}
+\subsubsection{\texorpdfstring{Inference of GxWeather effects for
+vegetative and reproductive timing using
+\emph{mash}}{Inference of GxWeather effects for vegetative and reproductive timing using mash}}\label{inference-of-gxweather-effects-for-vegetative-and-reproductive-timing-using-mash}
 
 We next looked for evidence of GxWeather by jointly re-estimating
-genetic effects across all eight common gardens using a model that
-allows effects to covary between gardens in many ways
+genetic effects across all eight common gardens using the mash model,
+which allows effects to covary between gardens in many ways
 (Figure~\ref{fig-map} D, SI Appendix, Datasets 1-6). To do this, we used
-genome-wide association (GWAS) on site-specific BLUPs for the timing of
-vegetative growth and flowering to determine preliminary single
-nucleotide polymorphism (SNP) effect estimates at each garden. We then
-jointly modeled genetic effect estimates, and the mixture of ways these
-effects might covary, across all eight common gardens, using a two-part
-procedure. We first modeled genetic effects with different combinations
-of covariance types, using a subset of effects selected at random from
-all SNPs used in the GWAS. In these models, we specified many ways
-genetic effects could covary between gardens (as expanded on below) and
-used a greedy algorithm to iteratively select covariance structures that
-significantly improved the log likelihood of a model when included.
-Given the high correlation between some of the covariance patterns we
-tested, this approach allowed us to select only the subset of patterns
-of covariation that most improved the fit of the joint effect patterns
-present in the data. We then re-estimated the set of genetic effects
-that had the largest effect estimates in univariate GWAS (hereafter
-``strong'' effects) using the set of significant covariance matrices
-that most improved the model fit on the random effects.
-
-We created three categories of ways effects could covary across gardens:
-``canonical'' covariance, with simple patterns of effects;
-``data-driven'' covariances derived from common patterns of SNP effects
-in the data, and ``GxWeather'' covariances that captured the covariance
-of weather variables at specific times before the phenological event
-(Table 1, SI Appendix, Section S2). For example, GxWeather covariance
-between gardens in MI and MO, for rainfall in the seven days prior to
-flowering, captures the covariance in the amount of rainfall genotype A,
-B, and C receive in the seven days prior to their three flowering dates
-in MI and their three flowering dates in MO. If these six rainfall
-values are similar, covariance will be near one; if they are
-uncorrelated, covariance will be near zero; if there is sufficient rank
-changing in rainfall amounts between gardens for these genotypes, this
-covariance will be near negative one.
-
-The GxWeather covariances allow hypothesis testing of specific weather
-variables as cues for each genetic effect. Say that a SNP in FLC
-controls flowering in a photoperiod-dependent manner. In that case, that
-SNP effect could have a high mixture proportion, or mass, on a
+GWAS on site-specific BLUPs for the timing of vegetative growth and
+flowering to determine preliminary single nucleotide polymorphism (SNP)
+effect estimates for both traits at each garden. We then jointly modeled
+genetic effect estimates, and the mixture of ways these effects might
+covary, across all eight common gardens, using a two-part procedure. We
+first modeled genetic effects with different combinations of covariance
+structures, using a subset of randomly selected SNPs used in the GWAS
+and their corresponding effect size estimates. In these models
+(hereafter referred to as ``random''), we then specified a variety of
+potential covariance structures among genetic effect size estimates from
+each garden (as expanded on below) and used a greedy algorithm to
+iteratively select and add covariance structures that significantly
+improved the log likelihood of the \emph{mash} model. Given the high
+correlation between some of the covariance patterns we tested, this
+approach allowed us to select only the subset of patterns of covariance
+that significantly improved the fit of the jointly inferred structure of
+genetic effect estimates present in the data. We then selected the set
+of SNPs and corresponding effect size estimates that had the largest
+effect estimates in univariate GWAS (hereafter ``strong'' effects) and
+jointly re-estimated these effect size estimates at all eight sites
+using the set of significant covariance matrices that most improved the
+model fit on the random effects.
+
+We specified three categories of covariance structures across gardens:
+``canonical'' covariance, with simple patterns of effect size covariance
+introduced in the initial \emph{mash} manuscript; ``data-driven''
+covariances derived from common patterns of SNP effects observed in the
+data, and ``GxWeather'' covariance, estimated from the covariance of
+empirical weather patterns at each garden at specific times before the
+phenological event (Table 1, SI Appendix, Section S2). For example,
+GxWeather covariance between gardens in MI and MO, for rainfall in the
+seven days prior to flowering, captures the covariance in the amount of
+rainfall genotypes A, B, and C receive in the seven days prior to their
+three flowering dates in MI and their three flowering dates in MO. If
+these six rainfall values are similar, covariance will be near one; if
+they are uncorrelated, covariance will be near zero; if there is
+sufficient rank changing in rainfall amounts between gardens for these
+genotypes, this covariance will be near negative one. We specified
+multiple covariance structures for the canonical, data-driven, and
+GxWeather categories, each of which represented a specific pattern of
+covariance in genetic effect size estimates that could be included in
+the model.
+
+The GxWeather covariance structures allow hypothesis testing of specific
+weather variables as cues for the start of vegetative and reproductive
+growth. Say that a SNP in FLC controls flowering in a
+photoperiod-dependent manner. In that case, the joint estimate of
+effects for that SNP could have a high mixture proportion, or mass, on a
 covariance matrix created using a photoperiod-based environmental cue,
@@
 infer that the effect of that SNP on flowering was caused by a response
-to that or a correlated environmental cue. Overall, loadings of genetic
-effects on these GxWeather matrices in the model with re-estimated
-effects provide information on genome-wide patterns of SNP-environment
-interaction.
+to the environmental cue used to construct the GxWeather covariance
+structure with the largest mass. Overall, loadings of genetic effects on
+these GxWeather matrices in the model, along with the posterior effect
+size estimates provide information on genome-wide patterns of
+SNP-environment interaction.
 
@@
 moderate negative correlations between the Texas and North gardens,
-particularly in the Gulf and in Both subpopulations
+particularly in the Gulf subpopulation and Both subpopulations
 (Figure~\ref{fig-map} B). If these phenotypic correlations have a
-genetic component, then they might be controlled by SNP effects that
+genetic component, then they may be controlled by SNP effects that
 differed in sign across these regions (effects with antagonistic
@@
 non-significant in others (effects with amplification), or a combination
-of these patterns. Five GxWeather covariance matrices were selected by
-the greedy mash algorithm, one to two per subpopulation. Of these five,
-three had mass on them in mash models of the strong effects
+of these patterns. Five GxWeather covariance structures were selected by
+the greedy algorithm, one to two per subpopulation. Of these five, three
+had mass on them in mash models of the strong effects
 (Figure~\ref{fig-covar} A-B). Two of these three matrices had negative
@@
 strong SNP effects in the mash model of Gulf vegetative growth fell on a
-covariance matrix of daylength 14 days prior to the date of vegetative
-growth. The covariance matrix for this weather cue was very similar to
-the pattern of phenotypic correlation for the timing of vegetative
-growth in the Gulf subpopulation (Figure~\ref{fig-map} B;
-Figure~\ref{fig-covar} A). Mash models of the timing of vegetative
-growth in Midwest and Both subpopulations did not include this GxWeather
-covariance; the Midwest had no weight on any GxWeather covariance, while
-Both subpopulations had non-zero weights on two additional GxWeather
-covariance types, average temperature one day prior to green-up, and the
-day length change in seconds in the day prior to green-up
-(Figure~\ref{fig-covar} B). The average temperature covariance matrix
-had negative covariances between Texas and North gardens, though not as
-strong as the negative phenotypic correlations seen in Both
-subpopulations (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Only
-the Gulf subpopulation and Both subpopulations had mass on any GxWeather
-covariance matrices (Figure~\ref{fig-covar} C).
-
-For flowering date, distinct GxWeather covariances captured
-SNP-associated effect patterns in the Gulf and Midwest subpopulations.
-33\% of SNP effects on flowering in the Gulf subpopulation covaried with
-cumulative rainfall in the seven days prior to flowering (Fig 2D).
-22.6\% of SNP effects on flowering in the Midwest subpopulation covaried
-with day length change in the two days prior to flowering (Fig 2D),
-which had negative covariances between Texas and North gardens. Neither
-covariance matrix was selected for in Both subpopulations, and no
-GxWeather covariance matrices had mass in Both subpopulations (Fig 2E).
-In five of the six mash models of strong effects, the GxWeather
-covariance matrices captured a minority of the posterior weights of the
-strong effects (Figure 2C,E); the majority of this mass was on various
-canonical covariance matrices. These matrices included simple
-heterozygosity, with intermediate, positive covariances between all
-gardens, and garden-specific effects present only at one garden.
+covariance matrix constructed using the covariance of daylength 14 days
+prior to the date of vegetative growth. The covariance matrix for this
+weather cue was very similar to the pattern of phenotypic correlation
+for the timing of vegetative growth in the Gulf subpopulation
+(Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Mash models of the
+timing of vegetative growth in the Midwest subpopulation and Both
+subpopulations did not include this GxWeather covariance; the Midwest
+had no weight on any GxWeather covariance, while Both subpopulations had
+non-zero weights on two additional GxWeather covariance types, average
+temperature one day prior to green-up, and the day length change in
+seconds in the day prior to green-up (Figure~\ref{fig-covar} B). The
+average temperature covariance matrix had negative covariances between
+Texas and North gardens, though not as strong as the negative phenotypic
+correlations seen in Both subpopulations (Figure~\ref{fig-map} B;
+Figure~\ref{fig-covar} A). Only the Gulf subpopulation and Both
+subpopulations had mass on any GxWeather covariance matrices
+(Figure~\ref{fig-covar} C).
+
+For flowering date, distinct GxWeather covariance structures captured
+covariance in effect sizes for SNPs in the Gulf and Midwest
+subpopulations. 33\% of SNP effects on flowering in the Gulf
+subpopulation covaried with cumulative rainfall in the seven days prior
+to flowering (Figure~\ref{fig-covar} D). 22.6\% of SNP effects on
+flowering in the Midwest subpopulation covaried with day length change
+in the two days prior to flowering (Figure~\ref{fig-covar} D), which had
+negative covariances between Texas and North gardens. Neither covariance
+matrix was selected as significantly improving the log-likelihood of the
+\emph{mash} model of Both subpopulations, and no GxWeather covariance
+matrices had a non-zero mass in models of Both subpopulations
+(Figure~\ref{fig-covar} E). In five of the six mash models of strong
+effects, the GxWeather covariance matrices captured a minority of the
+posterior weights of the strong effects (Figure~\ref{fig-covar} C,E);
+the majority of this mass was on various canonical covariance matrices.
+These matrices included simple heterozygosity, with intermediate,
+positive covariances between all gardens, and garden-specific effects
+present only at one garden.
 
@@
 confidence in the effect sign, not the effect's difference from zero, to
-determine significance. We required significance (p \textless{} 0.05) in
-both gardens to include effects. This means that our tests for
+determine significance. We required lfsr significance (p \textless{}
+0.05) in both gardens to include effects. This means that our tests for
 antagonistic pleiotropy, or a sign change between conditions, carry an
@@
 also summarized the pairwise patterns of effect sign and magnitude
-within and between the Texas and North regions (Table 2).
+within and between the Texas and North regions, within the natural
+ranges of the Gulf and Midwest subpopulations (@fig-effects).
 
@@
 pairwise effects exhibited antagonistic pleiotropy, or a difference in
-effect sign, between pairs of Texas and North gardens (Fig. 3A). 78.7\%
-of pairwise comparisons between North and Texas gardens had a difference
-in sign, while only 28.6\% and 0.2\% of North-North or Texas-Texas
-comparisons had a difference in sign, respectively (Table 2). The
-majority of pairwise effects for greenup for the Midwest
-(\textgreater55\%) and Both (\textgreater85\%) subpopulations were the
-same sign, and effects frequently differed in magnitude between the MO
-and OK garden and other gardens (Table 2; Fig 3A).
+effect sign, between pairs of Texas and North gardens (@fig-effects A).
+78.7\% of pairwise comparisons between North and Texas gardens had a
+difference in sign, while only 28.6\% and 0.2\% of North-North or
+Texas-Texas comparisons had a difference in sign, respectively
+(@fig-effects). The majority of pairwise effects for greenup for the
+Midwest (\textgreater55\%) and Both (\textgreater85\%) subpopulations
+were the same sign, and effects frequently differed in magnitude between
+the MO and OK garden and other gardens (@fig-effects X).
 
@@
 effects exhibited antagonistic pleiotropy, or a difference in effect
-sign, within or between regions (Table 2). More effects differed in
+sign, within or between regions (@fig-effects). More effects differed in
 magnitude between the Texas and North regions than within these regions
-(42.7\% vs \textless20\%; Figure 3B). The Midwest population had
+(42.7\% vs \textless20\%; @fig-effects X). The Midwest population had
 relatively few significant effects for flowering, but a large proportion
@@
 within the North region (65.4\%). Finally, in Both subpopulations, less
-than 20\% of pairwise effects differed in sign (Table 2). Most
+than 20\% of pairwise effects differed in sign (@fig-effects). Most
 differences in sign were between TX1, the southernmost garden, and all
-other gardens (Fig 3B). Similarly, more effects that differed in
+other gardens (@fig-effects B). Similarly, more effects that differed in
 magnitude included gardens in the Texas region (52.3-55.9\%), and most
@@
 
-We sought additional experimental support for our significant effect
-re-estimates using an independent pseudo-F2 mapping population created
-from Gulf \& Midwest individuals and grown at the same sites (Fig.
-4A,B). We conducted quantitative trait loci (QTL) mapping of flowering
-as functions of four environmental cues that we also used as covariance
+We sought additional experimental support for our joint re-estimates of
+SNP effects using an independent pseudo-F2 mapping population created
+from Gulf \& Midwest individuals and grown at the same sites (@fig-qtl
+A,B). We conducted quantitative trait loci (QTL) mapping of flowering as
+functions of four environmental cues that we also used as covariance
 matrices in mash, and identified eight QTL for flowering date, six QTL
@@
 were found for all but eleven of these sets of random genomic intervals
-(Fig. 4C, p = 0.011). Thus, we were able to experimentally support our
-significant effect re-estimates with a QTL mapping experiment using a
-separate mapping population.
+(@fig-qtl C, p = 0.011). Thus, we were able to experimentally support
+our joint re-estimates of SNP effects with a QTL mapping experiment
+using a separate mapping population.
 
@@
 subpopulations revealed substantial antagonistic pleiotropy in effects
-between the Texas and North gardens (Figure 3A). This result supports
-theoretical models that local adaptation should involve antagonistic
-pleiotropy at the level of individual loci (22--25), and is the first
-experimental work using QTL mapping and GWAS across common gardens to
-find antagonistic pleiotropy to be common in small genomic regions (12,
-37, 38).
+between the Texas and North gardens (@fig-effects A). This result
+supports theoretical models that local adaptation should involve
+antagonistic pleiotropy at the level of individual loci (22--25), and is
+the first experimental work using QTL mapping and GWAS across common
+gardens to find antagonistic pleiotropy to be common in small genomic
+regions (12, 37, 38).
 
@@
 also supported by QTL from an independent mapping population at these
-sites (Fig. 4C).
+sites (@fig-qtl C).
 

### 2024-01-18 manuscript/revision2/revision2.qmd
manuscript: AHM revisions to Figure 3 and 4 and some M&M citation work

@@
 
-Our analysis of the timing of flowering showed that the Gulf and Midwest subpopulations have distinct GxWeather: flowering timing in the Midwest subpopulation has photoperiod-related variation, in that flowering timing covaries with a day length change signal in the two days before flowering occurs.
-In contrast, the Gulf subpopulation does not have variation in flowering based on photoperiod.
+Our analysis of the timing of flowering showed that the Gulf and Midwest subpopulations have distinct GxWeather: flowering timing in the Midwest subpopulation has photoperiod-related variation, in that flowering timing covaries with a day length change signal two days before flowering occurs.
+In contrast, the Gulf subpopulation does not have variation in flowering based on a photoperiod cue.
 Instead, the Gulf subpopulation has variation in flowering that covaries with the rainfall that occurs in the week prior to flowering.
@@
 Identifying the environmental cues that are predictive of, or even correlated with, plant phenotypic responses remains a major challenge to studies interrogating gene action across many natural environments.
-Clearly, the photoperiod and cumulative GDD cues we identify here are functions of the genotypes measured, are not predictive, and capture only a minority of SNP effects on flowering.
-We know still less about the overwintering parameters that affect green-up, which is reflected in our lower ability to assign SNP effects on green-up to weather-based cues.
+The GxWeather photoperiod and cumulative rainfall cues we identify here are functions of the genotypes measured and capture only a minority of SNP effects on flowering.
+We know still less about the overwintering parameters that cause variation in the start of vegetative.
+Though we could assign SNP effects to a GxWeather covariance structure in the Gulf subpopulation, we could not do this for the Midwest subpopulation.
 More generally, it is difficult to predict the time scales over which individuals may integrate environmental cues, particularly in perennial species which may integrate these cues over longer time scales.
-Mash offers an opportunity to specify multiple environmental cues and compete them to explain patterns of SNP effects, allowing us to detect how important these cues are genome-wide, and how strongly each cue influences each SNP.
+If this integration time itself varies between individuals, we cannot select a covariance structure that reflects this, though these structures would likely be highly correlated with GxWeather structures we did include.
+Mash offers an opportunity to specify multiple environmental cues and compete them to explain patterns of genetic effects, allowing us to detect how important these cues are genome-wide, and how strongly each cue influences each SNP.
 This is a key development to further improve our understanding of genetic variation in GxE.
@@
 
-![Caption for this figure...](images/Figure_2_Covariance_matrices_and_posterior_weights.png){#fig-covar}
+![Caption for this figure... blah blah blah.](images/Figure_2_Covariance_matrices_and_posterior_weights.png){#fig-covar}
+
+![Comparisons of pairs of jointly re-estimated effects across eight gardens, for effects with lfsr \< 0.05.
+AHM is considering moving this figure to the supplement and substantially simplifying the figure in the main text, perhaps just showing total effects in Texas, the North, and maybe Oklahoma.
+Essentially making Table 2 into a new, simplified figure.
+](images/Figure_3_Effect_Types_by_Site.png){#fig-effects}
+
+![SNP effects for SNPs within QTL.
+AHM to re-make this figure now LZ has re-run the QTL analysis, the assmption being the re-analysis will not change the figure or the text much.
+The bigger question for TEJ and AH is whether we want to keep this analysis in the paper, what does it add?](images/Figure_4_mash_SNP_effects.png){#fig-qtl}
 
@@
 In 2019, we scored two phenological events every two days in two mapping populations of switchgrass, a diversity panel and a pseudo-F2 cross, planted at eight common garden locations (32, 34, 37).
-We scored green-up date as the day of the year when 50% of the tiller area of the crown of the plant cut the previous year had green growth.
-Flowering date was the day of the year when 50% of the plant tillers had panicles undergoing anthesis.
+We scored the start of vegetative growth as the day of the year when 50% of the tiller area of the crown of the plant cut the previous year had green growth.
+The start of reproductive growth, or flowering date, was the day of the year when 50% of the plant tillers had panicles undergoing anthesis.
 We scored green-up and flowering as day of the year, then linked these dates to multiple weather-based environmental factors measured daily at each common garden (SI Appendix, Section S1, Table S1).
 
-The formation and resequencing of the diversity panel has been described previously (32).
+The formation and resequencing of the diversity panel has been described previously @lovell2021genomic.
 The diversity panel contained 134 sequenced, clonally propagated individuals from the Midwest genetic subpopulation, and 229 from the Gulf genetic subpopulation.
 To allow for the possibility that different subpopulations had different strengths of connection between our phenotypes and genotypes (38), we conducted three sets of genetic analyses: on Gulf and Midwest genotypes separately, and on both subpopulations together ('Both').
-Analyses to determine narrow-sense heritability (h2) for green-up and flowering were done using linear mixed models and followed (32).
+Analyses to determine narrow-sense heritability (h^2^) for green-up and flowering were done using linear mixed models and followed @lovell2021genomic.
 Details on these models can be found in (SI Appendix, Section S3,S4).
@@
 
-To evaluate the prevalence and kinds of covariance patterns of SNP effects across our common gardens, we used multivariate adaptive shrinkage (mash) on SNP effect estimates from the diversity panel (35).
+To evaluate the prevalence and kinds of covariance patterns of SNP effects across our common gardens, we used multivariate adaptive shrinkage (*mash*) on SNP effect estimates from the diversity panel (35).
 Mash is a statistical method that allows estimation and comparison of many effects jointly across many different conditions; it improves on previous methods by allowing multiple, arbitrary correlations in effect sizes among conditions (SI Appendix, Section S5).
 To obtain SNP effect estimates, we first conducted univariate genome-wide association at each common garden for green-up and flowering date.
-We then analyzed SNP effects for the top 19K relatively unlinked (r2 \< 0.2) SNPs per condition using mash, as in (32).
+We then analyzed SNP effects for the top 19K relatively unlinked (r^2^ \< 0.2) SNPs per condition using mash, as in @lovell2021genomic.
 Details on these models can be found in (SI Appendix, Section S6).
 
-We generated hypothesis-based covariance matrices derived from correlations in environmental cues in the green-up or flowering date windows for the three subpopulations (SI Appendix, Table S1, Section S1).
+We generated hypothesis-based covariance matrices derived from correlations in environmental cues in the green-up or flowering date windows for the three populations (SI Appendix, Table S1, Section S1).
 These covariance matrices represent correlations between identical genotypes drawn from a specific population at pairs of common gardens; covariances near one mean that the population has a strong, positive linear relationship in individual responses at that pair of gardens, while covariances near zero mean that there is no relationship within the population for individual responses at that pair of gardens.
@@
 Mash also generates data-driven covariance matrices corresponding to major patterns of SNP effects present in the data.
-We generated six data-driven matrices per mash run, five (denoted DD_PCA_1 through DD_PCA_5) produced by singular value decomposition (SVD) of an overall matrix, denoted 'DD_tPCA'.
-We used SVD to present vectors of garden-specific effects for each numbered DD_PCA matrix, because the first eigenvector of a SVD explains 100% of the variation in these matrices, and each value in this eigenvector represents one garden-specific effect.
+We generated six data-driven matrices per mash run, five produced by singular value decomposition (SVD) of an overall matrix.
 
@@
 Using the lfsr rather than the local false discovery rate (lfdr) is a critical change in our ability to detect antagonistic pleiotropy.
-The lfdr, like other measures of FDR, focuses on if we have enough evidence to reject the null hypothesis that an effect j is 0 -- that there is a significant effect.
+The lfdr, like other measures of FDR, focuses on if we have enough evidence to reject the null hypothesis that an effect j is 0, or that there is a significant effect.
 Previous studies of antagonistic pleiotropy (e.g.
@@
 However, using the lfsr to test for antagonistic pleiotropy does not undercount occurrences of antagonistic pleiotropy, as this statistic answers a fundamentally different question.
-For each effect j, the lfsr_j is defined as the probability that we make an error in the sign of effect j if we were forced to declare the effect positive or negative (Stephens paper).
+For each effect j, the lfsr~j~ is defined as the probability that we make an error in the sign of effect j if we were forced to declare the effect positive or negative @10.1093/biostatistics/kxw041.
 Thus, rather than asking "Are these two effects different?" - as we reasonably expect two effects to be, even if this difference cannot be measured - the local false sign rate answers a more meaningful question: Can we be confident in the sign of this effect?
 
-In addition, the get_GxE function also sets an arbitrary threshold to count an effect as changing in magnitude between environments, commonly known as differential sensitivity.
+In addition, the get_GxE function also sets an arbitrary threshold to count an effect as changing in magnitude between environments, commonly known as differential sensitivity or a change in amplitude of .
 For differential sensitivity, this function determines if effects significant in both conditions are of the same sign and of a magnitude (not tested for significance) that differs by a factor of 0.4 or more.

### 2024-01-18 manuscript/revision2/revision2.tex
manuscript: AHM revisions to Figure 3 and 4 and some M&M citation work

@@
 subpopulation has photoperiod-related variation, in that flowering
-timing covaries with a day length change signal in the two days before
+timing covaries with a day length change signal two days before
 flowering occurs. In contrast, the Gulf subpopulation does not have
-variation in flowering based on photoperiod. Instead, the Gulf
+variation in flowering based on a photoperiod cue. Instead, the Gulf
 subpopulation has variation in flowering that covaries with the rainfall
@@
 correlated with, plant phenotypic responses remains a major challenge to
-studies interrogating gene action across many natural environments.
-Clearly, the photoperiod and cumulative GDD cues we identify here are
-functions of the genotypes measured, are not predictive, and capture
-only a minority of SNP effects on flowering. We know still less about
-the overwintering parameters that affect green-up, which is reflected in
-our lower ability to assign SNP effects on green-up to weather-based
-cues. More generally, it is difficult to predict the time scales over
-which individuals may integrate environmental cues, particularly in
-perennial species which may integrate these cues over longer time
-scales. Mash offers an opportunity to specify multiple environmental
-cues and compete them to explain patterns of SNP effects, allowing us to
+studies interrogating gene action across many natural environments. The
+GxWeather photoperiod and cumulative rainfall cues we identify here are
+functions of the genotypes measured and capture only a minority of SNP
+effects on flowering. We know still less about the overwintering
+parameters that cause variation in the start of vegetative. Though we
+could assign SNP effects to a GxWeather covariance structure in the Gulf
+subpopulation, we could not do this for the Midwest subpopulation. More
+generally, it is difficult to predict the time scales over which
+individuals may integrate environmental cues, particularly in perennial
+species which may integrate these cues over longer time scales. If this
+integration time itself varies between individuals, we cannot select a
+covariance structure that reflects this, though these structures would
+likely be highly correlated with GxWeather structures we did include.
+Mash offers an opportunity to specify multiple environmental cues and
+compete them to explain patterns of genetic effects, allowing us to
 detect how important these cues are genome-wide, and how strongly each
@@
 
-\caption{\label{fig-covar}Caption for this figure\ldots{}}
+\caption{\label{fig-covar}Caption for this figure\ldots{} blah blah
+blah.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{images/Figure_3_Effect_Types_by_Site.png}
+
+}
+
+\caption{\label{fig-effects}Comparisons of pairs of jointly re-estimated
+effects across eight gardens, for effects with lfsr \textless{} 0.05.
+AHM is considering moving this figure to the supplement and
+substantially simplifying the figure in the main text, perhaps just
+showing total effects in Texas, the North, and maybe Oklahoma.
+Essentially making Table 2 into a new, simplified figure.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{images/Figure_4_mash_SNP_effects.png}
+
+}
+
+\caption{\label{fig-qtl}SNP effects for SNPs within QTL. AHM to re-make
+this figure now LZ has re-run the QTL analysis, the assmption being the
+re-analysis will not change the figure or the text much. The bigger
+question for TEJ and AH is whether we want to keep this analysis in the
+paper, what does it add?}
 
@@
 populations of switchgrass, a diversity panel and a pseudo-F2 cross,
-planted at eight common garden locations (32, 34, 37). We scored
-green-up date as the day of the year when 50\% of the tiller area of the
-crown of the plant cut the previous year had green growth. Flowering
-date was the day of the year when 50\% of the plant tillers had panicles
-undergoing anthesis. We scored green-up and flowering as day of the
-year, then linked these dates to multiple weather-based environmental
-factors measured daily at each common garden (SI Appendix, Section S1,
-Table S1).
+planted at eight common garden locations (32, 34, 37). We scored the
+start of vegetative growth as the day of the year when 50\% of the
+tiller area of the crown of the plant cut the previous year had green
+growth. The start of reproductive growth, or flowering date, was the day
+of the year when 50\% of the plant tillers had panicles undergoing
+anthesis. We scored green-up and flowering as day of the year, then
+linked these dates to multiple weather-based environmental factors
+measured daily at each common garden (SI Appendix, Section S1, Table
+S1).
 
@@
 subpopulations together (`Both'). Analyses to determine narrow-sense
-heritability (h2) for green-up and flowering were done using linear
-mixed models and followed (32). Details on these models can be found in
-(SI Appendix, Section S3,S4).
+heritability (h\textsuperscript{2}) for green-up and flowering were done
+using linear mixed models and followed (32). Details on these models can
+be found in (SI Appendix, Section S3,S4).
 
@@
 effects across our common gardens, we used multivariate adaptive
-shrinkage (mash) on SNP effect estimates from the diversity panel (35).
-Mash is a statistical method that allows estimation and comparison of
-many effects jointly across many different conditions; it improves on
+shrinkage (\emph{mash}) on SNP effect estimates from the diversity panel
+(35). Mash is a statistical method that allows estimation and comparison
+of many effects jointly across many different conditions; it improves on
 previous methods by allowing multiple, arbitrary correlations in effect
@@
 common garden for green-up and flowering date. We then analyzed SNP
-effects for the top 19K relatively unlinked (r2 \textless{} 0.2) SNPs
-per condition using mash, as in (32). Details on these models can be
-found in (SI Appendix, Section S6).
+effects for the top 19K relatively unlinked (r\textsuperscript{2}
+\textless{} 0.2) SNPs per condition using mash, as in (32). Details on
+these models can be found in (SI Appendix, Section S6).
 
@@
 correlations in environmental cues in the green-up or flowering date
-windows for the three subpopulations (SI Appendix, Table S1, Section
-S1). These covariance matrices represent correlations between identical
+windows for the three populations (SI Appendix, Table S1, Section S1).
+These covariance matrices represent correlations between identical
 genotypes drawn from a specific population at pairs of common gardens;
@@
 effects present in the data. We generated six data-driven matrices per
-mash run, five (denoted DD\_PCA\_1 through DD\_PCA\_5) produced by
-singular value decomposition (SVD) of an overall matrix, denoted
-`DD\_tPCA'. We used SVD to present vectors of garden-specific effects
-for each numbered DD\_PCA matrix, because the first eigenvector of a SVD
-explains 100\% of the variation in these matrices, and each value in
-this eigenvector represents one garden-specific effect.
+mash run, five produced by singular value decomposition (SVD) of an
+overall matrix.
 
@@
 lfdr, like other measures of FDR, focuses on if we have enough evidence
-to reject the null hypothesis that an effect j is 0 -- that there is a
+to reject the null hypothesis that an effect j is 0, or that there is a
 significant effect. Previous studies of antagonistic pleiotropy
@@
 antagonistic pleiotropy, as this statistic answers a fundamentally
-different question. For each effect j, the lfsr\_j is defined as the
-probability that we make an error in the sign of effect j if we were
-forced to declare the effect positive or negative (Stephens paper).
-Thus, rather than asking ``Are these two effects different?'' - as we
+different question. For each effect j, the lfsr\textsubscript{j} is
+defined as the probability that we make an error in the sign of effect j
+if we were forced to declare the effect positive or negative (36). Thus,
+rather than asking ``Are these two effects different?'' - as we
 reasonably expect two effects to be, even if this difference cannot be
@@
 count an effect as changing in magnitude between environments, commonly
-known as differential sensitivity. For differential sensitivity, this
-function determines if effects significant in both conditions are of the
-same sign and of a magnitude (not tested for significance) that differs
-by a factor of 0.4 or more. The remaining effects that are significant
-in both conditions have the same effect sign and similar effect
-magnitudes and we denote these effects as having no GxE. The distinction
-between effects with different magnitudes is arbitrary but useful to
-fully characterize how effects vary across environments. and our use of
-the lfsr to determine significance and our specification that SNP
-effects must be significant in both conditions to be included means that
-our tests for antagonistic pleiotropy carry an equal statistical burden
-to those measuring differential sensitivity and effects without GxE.
+known as differential sensitivity or a change in amplitude of . For
+differential sensitivity, this function determines if effects
+significant in both conditions are of the same sign and of a magnitude
+(not tested for significance) that differs by a factor of 0.4 or more.
+The remaining effects that are significant in both conditions have the
+same effect sign and similar effect magnitudes and we denote these
+effects as having no GxE. The distinction between effects with different
+magnitudes is arbitrary but useful to fully characterize how effects
+vary across environments. and our use of the lfsr to determine
+significance and our specification that SNP effects must be significant
+in both conditions to be included means that our tests for antagonistic
+pleiotropy carry an equal statistical burden to those measuring
+differential sensitivity and effects without GxE.
 

### 2024-02-21 R/Analysis_v0.6_QTL_GxE_and_mash_comparison.Rmd
feat: QTL analysis using phenotypes from new covariance matrices, incorporated into Figure 3

@@
 library(here)
+library(data.table)
 ```
@@
 qtlplot <- qtldf %>%
-  mutate(phe = case_when(lodcolumn == "FL50" ~ "50% flowering",
-                              lodcolumn == "dyln_fl50" ~ "Daylength",
-                              lodcolumn == "dyln_change_sec" ~ "daylength change (s)",
-                              lodcolumn == "CGDD_12C" ~ "GDD, GR to FL",
+  mutate(phe = case_when(lodcolumn == "FL50" ~ "flowering",
+                              lodcolumn == "dyln_change_sec_2d_prior" ~ "daylength change (s) 2d prior",
+                              lodcolumn == "dyln_change_sec" ~ "daylength change (s) 1d prior",
+                              lodcolumn == "GR50" ~ "greenup",
                               lodcolumn == "crain_gr2fl" ~ "Rainfall, GR to FL")) %>%
@@
 qtlplot$phe <- factor(qtlplot$phe, 
-                      levels = c("50% flowering", "Rainfall, GR to FL",
+                      levels = c("flowering", "daylength change (s) 2d prior",
                                  "GDD, GR to FL", "Daylength",
@@
 pw_3f <- pw_gf %>%
-  full_join(pw_mf) %>%
-  full_join(pw_gamf) %>%
+  filter(log10BayesFactor > 2.56) %>%
+  full_join(filter(pw_mf, log10BayesFactor > 2)) %>%
+  full_join(filter(pw_gamf, log10BayesFactor > 2.65)) %>%
   arrange(Chr, Pos) %>%
-  filter(log10BayesFactor > 3) %>%
-  select(Chr:Pos, log10BayesFactor, subpop, phenotype, everything())
+  #filter(log10BayesFactor > 3) %>%
+  dplyr::select(Chr:Pos, log10BayesFactor, subpop, phenotype, everything())
 
@@
   arrange(Chr, Pos) %>%
-  select(Chr:Pos, log10BayesFactor, subpop, phenotype, everything())
+  dplyr::select(Chr:Pos, log10BayesFactor, subpop, phenotype, everything())
 
@@
 summary(pw_3f$log10BayesFactor)
-quantile(pw_gf$log10BayesFactor, 0.96196895)
-quantile(pw_gf$log10BayesFactor, 0.96196895)
-quantile(pw_gamf$log10BayesFactor, 0.8006157)
+quantile(pw_gf$log10BayesFactor, 0.99)  # Gulf: Use 2.56
+quantile(pw_mf$log10BayesFactor, 0.99)  # Midwest: Use 2
+quantile(pw_gamf$log10BayesFactor, 0.99)  # GAM: Use 2.65
+quantile(pw_mf$log10BayesFactor, 0.995119)
+
+quantile(pw_gg$log10BayesFactor, 0.99)  # Gulf: Use 2.0989
+quantile(pw_mg$log10BayesFactor, 0.99)  # Midwest: Use 4.6831
+quantile(pw_gamg$log10BayesFactor, 0.99)  # GAM: Use 3.0865
+
 1-0.96196895
@@
 42/255 # GT 19.93843%? No 16.47%
+
+pw_3g <- pw_gg %>%
+  filter(log10BayesFactor > 2.0989) %>%
+  full_join(filter(pw_mg, log10BayesFactor > 4.6831)) %>%
+  full_join(filter(pw_gamg, log10BayesFactor > 3.0865)) %>%
+  arrange(Chr, Pos) %>%
+  #filter(log10BayesFactor > 3) %>%
+  dplyr::select(Chr:Pos, log10BayesFactor, subpop, phenotype, everything())
 ```
+## SNPs sig in 1% tail or BF>2, whichever is stricter:
+```{r}
+pw_3f %>%
+  arrange(desc(log10BayesFactor)) %>%
+  pivot_longer(cols = null:flowering.simple_het_1.13, names_to = "Covar Matrix", values_to = "Posterior Weight") %>%
+  filter(`Posterior Weight` > 0.01) %>%
+  arrange(desc("Posterior Weight"))
+
+pw_3g %>%
+  arrange(desc(log10BayesFactor)) %>%
+  pivot_longer(cols = null:greenup.singleton.MO.15, names_to = "Covar Matrix", values_to = "Posterior Weight") %>%
+  filter(`Posterior Weight` > 0.01) %>%
+  arrange(desc("Posterior Weight"))
+```
+3,4,6 >99% on gulf.crain_7d
+
+1,2,8 on singleton.TX3 9.3e-01
+5 on flowering.si
 
 ```{r}
+qtlplot_fl <- qtlplot %>% 
+  filter(lodcolumn %in% c("FL50", "dyln_change_sec_2d_prior")) 
+qtlplot_gr <- qtlplot %>% 
+  filter(lodcolumn %in% c("GR50", "dyln_change_sec"))
+
 i=1  # from 1 to 16
@@
   pw_df_all <- get_post_weights(m)
+  if("lod" %in% colnames(qtl_df)) {
+    qtl_df <- qtl_df %>%
+      arrange(desc(lod)) %>%
+      mutate(pos = as.integer(pos)) %>%
+      group_by(CHR, pos) %>%
+      dplyr::slice(1) %>%
+      arrange(CHR, pos)
+  }
 
@@
     filter(log10BayesFactor > BFthresh) %>%
-  select(Chr:Pos, log10BayesFactor, everything())
+    dplyr::select(Chr:Pos, log10BayesFactor, everything())
 
@@
 
-qtlplot <- qtlplot %>%
-  arrange(desc(lod)) %>%
-  mutate(pos = as.integer(pos)) %>%
-  group_by(CHR, pos) %>%
-  slice(1) %>%
-  arrange(CHR, pos)
-enrich_gf <- get_QTL_enrichment(m_gulf_fl_hyp, qtl_df = qtlplot, 0.99) %>%
+enrich_gf <- get_QTL_enrichment(m_gulf_fl_hyp, qtl_df = qtlplot_fl, 0.99) %>%
   mutate(subpop = "Gulf", phenotype = "flowering")
-enrich_mf <- get_QTL_enrichment(m_midw_fl_hyp, qtl_df = qtlplot, 0.99) %>%
+enrich_mf <- get_QTL_enrichment(m_midw_fl_hyp, qtl_df = qtlplot_fl, 0.99) %>%
   mutate(subpop = "Midwest", phenotype = "flowering")
-enrich_gamf <- get_QTL_enrichment(m_gam_fl_hyp, qtl_df = qtlplot, 0.99) %>%
+enrich_gamf <- get_QTL_enrichment(m_gam_fl_hyp, qtl_df = qtlplot_fl, 0.99) %>%
   mutate(subpop = "Both", phenotype = "flowering")
@@
 enrich_df <- enrich_df %>%
-  left_join(qtlplot, by = c("Chr" = "CHR", "Pos_lo" = "POS_lo", "Pos_hi" = "POS_hi")) %>%
-  select(Chr, Pos_lo, Pos_hi, inQTLandBF:subpop, lod, lodcolumn, pos) %>%
+  left_join(qtlplot_fl, by = c("Chr" = "CHR", "Pos_lo" = "POS_lo", "Pos_hi" = "POS_hi")) %>%
+  dplyr::select(Chr, Pos_lo, Pos_hi, inQTLandBF:subpop, lod, lodcolumn, pos) %>%
   unique() %>%
@@
 t.test(enrich_df[enrich_df$pvalue < 0.05,]$lod, enrich_df[enrich_df$pvalue >= 0.05,]$lod) 
-qtlplot %>% group_by(CHR, POS_lo) %>% tally() # 21-24 unique QTL by Pos lo/hi. 23 by QTL maximum. 12 of 24 have enriched mash hits.
+qtlplot_fl %>% group_by(CHR, POS_lo) %>% tally() # 21-24 unique QTL by Pos lo/hi. 23 by QTL maximum. 12 of 24 have enriched mash hits.
 
@@
 for(j in 1:1000){
-  qtl_size <- qtlplot %>%
-    select(CHR, POS_lo, POS_hi) %>%
-    unique() %>%   # don't doublecount completely overlapping QTL
+  qtl_size <- qtlplot_fl %>%
+    dplyr::select(CHR, POS_lo, POS_hi) %>%
+    dplyr::distinct() %>%   # don't doublecount completely overlapping QTL
     #(does doublecount partially overlapping QTL, however.)
@@
 
-  random_gf <- get_QTL_enrichment(m_gulf_fl_hyp, qtl_df = random_qtl, 0.99) %>%
+  random_gf <- get_QTL_enrichment(m = m_gulf_fl_hyp, qtl_df = random_qtl, 0.99) %>%
     mutate(subpop = "Gulf", phenotype = "flowering")
@@
 ```
+
+### Same process for greenup
+
+```{r}
+
+enrich_gg <- get_QTL_enrichment(m_gulf_gr_hyp, qtl_df = qtlplot_gr, 0.95) %>%
+  mutate(subpop = "Gulf", phenotype = "greenup")
+enrich_mg <- get_QTL_enrichment(m_midw_gr_hyp, qtl_df = qtlplot_gr, 0.9959) %>%
+  mutate(subpop = "Midwest", phenotype = "greenup")
+enrich_gamg <- get_QTL_enrichment(m_gam_gr_hyp, qtl_df = qtlplot_gr, 0.95) %>%
+  mutate(subpop = "Both", phenotype = "greenup")
+
+enrich_gr <- enrich_gg %>%
+  full_join(enrich_mg) %>%
+  full_join(enrich_gamg) %>%
+  arrange(Chr, Pos_lo)
+enrich_gr <- enrich_gr %>%
+  left_join(qtlplot_gr, by = c("Chr" = "CHR", "Pos_lo" = "POS_lo", "Pos_hi" = "POS_hi")) %>%
+  dplyr::select(Chr, Pos_lo, Pos_hi, inQTLandBF:subpop, lod, lodcolumn, pos) %>%
+  unique() %>%
+  arrange(Chr, Pos_lo, lodcolumn)
+
+enrich_gr %>%
+  filter(pvalue < 0.05)
+
+t.test(enrich_gr[enrich_gr$pvalue < 0.05,]$lod, enrich_gr[enrich_gr$pvalue >= 0.05,]$lod) 
+qtlplot_gr %>% group_by(CHR, POS_lo) %>% tally() # 21-24 unique QTL by Pos lo/hi. 23 by QTL maximum. 12 of 24 have enriched mash hits.
+
+```

### 2024-02-21 R/Analysis_v0.6_QTL_interval_plotting.Rmd
feat: QTL analysis using phenotypes from new covariance matrices, incorporated into Figure 3

@@
 ###load the original cross and reduce the markers from 4700 to 1185
-load('../phenology-fourway/5-QTLmappingForAlice/sexAveraged.cross.rda', verbose = T)
+#load('../phenology-fourway/5-QTLmappingForAlice/sexAveraged.cross.rda', verbose = T)
+load(file.path(workingdir, 'cross_reduced_marker_FL50_phenotypes_Alice_7_SITES.RData'))
 map  =  lapply(pull.map(cross), function(x) x[1,])
@@
 str(phenology) ###make sure your phenology traits are numeric, if not, change them into numeric
-phenos <- phenology %>% mutate_at(vars('ID'), list(as.character)) %>%
-  mutate_at(vars('ID'), list(as.numeric))%>% group_by(Covariate,ID ) %>%
-  summarize_all(mean, na.rm=T) %>% ungroup()%>% as.data.frame()
+phenos <- phenology %>% 
+  mutate_at(vars('ID'), list(as.character)) %>%
+  mutate_at(vars('ID'), list(as.numeric)) %>% 
+  group_by(Covariate,ID ) %>%
+  summarize_all(mean, na.rm=T) %>% 
+  ungroup() %>% 
+  as.data.frame()
 ```

### 2024-02-21 R/Analysis_v0.7_QTL_interval_lod_plot.Rmd
feat: QTL analysis using phenotypes from new covariance matrices, incorporated into Figure 3

--- 2021-05-04 R/Analysis_v0.7_QTL_interval_lod_plot.Rmd
@@
 load(file.path(workingdir, 'cross_reduced_marker_FL50_phenotypes_Alice_7_SITES.RData'))
-load(file.path(workingdir, 'scan1Results_FL50_phenotypes_Alice_7_SITES.RData.RData'))
+load(file.path(workingdir, 'scan1Results_FL50_phenotypes_Alice_7_SITES.RData'))
 
@@
 
+    flowering_cumulative_rainfall_7d    blue            1
+    flowering_daylength_change_s_2d     pink            4  
+    flowering                           purple          5
+
+    greenup_daylength_14d               green           2
+    greenup_daylength_change_s_1d       yellow          3
+    greenup                             grey            6
+    greenup_temp_ave_1d                 darkorange      7
+
 ```{r}
-threshold_df <- threshold_full %>% as_tibble(rownames = "Threshold") %>%
-  pivot_longer(GR50:dyln_change_sec, names_to = "Phenotype", values_to = "LOD")
+threshold_df <- threshold_full %>% 
+  as_tibble(rownames = "Threshold") %>%
+  pivot_longer(GR50:crain_7d, names_to = "Phenotype", values_to = "LOD") %>%
+  mutate(Phe = case_when(Phenotype %in% c("GR50", "tave_1d", "dyln_14d_prior", "dyln_change_sec") ~ "greenup",
+                         Phenotype %in% c("FL50", "dyln_change_sec_2d_prior", "crain_7d") ~ "flowering", 
+                         TRUE ~ NA_character_))
 
@@
   as_tibble(rownames = "Marker") %>%
-  pivot_longer(GR50:dyln_change_sec, names_to = "Phenotype", values_to = "LOD") %>%
+  pivot_longer(GR50:crain_7d, names_to = "Phenotype", values_to = "LOD") %>%
+  mutate(Phe = case_when(Phenotype %in% c("GR50", "tave_1d", "dyln_14d_prior", "dyln_change_sec") ~ "greenup",
+                         Phenotype %in% c("FL50", "dyln_change_sec_2d_prior", "crain_7d") ~ "flowering", 
+                         TRUE ~ NA_character_)) %>% 
   separate(Marker, into = c("Chr", "Pos"), sep = "_", remove = FALSE, convert = TRUE) %>%
@@
   geom_hline(data = threshold_df, aes(yintercept = LOD, color = Phenotype, linetype = Phenotype)) +
-  facet_wrap(~Chr, nrow = 1, scales = "free_x") +
+  facet_grid(Phe~Chr, scales = "free_x") +
   switchgrassGWAS::theme_oeco + 
-  scale_color_manual(values = c("#CC0066", "#FFCC00", "#006600", "#6600FF", "grey")) + 
-  scale_linetype_manual(values = c(4,5,4,4,4)) +
+  scale_color_manual(values = c("#000099", "#006600", "#FFCC00", "#CC0066", "#6600FF", "grey", "darkorange", "#000666")) + 
+  scale_linetype_manual(values = c(4,4,4,4,4,4,4)) +
   theme(legend.position = "top",
@@
 
-save_plot(filename = "../manuscript/plots/Figure_4_QTL_by_lod_score.svg", base_height = 3, base_width = 13.5, plot = last_plot())
+save_plot(filename = "../manuscript/revision2/Figure_4_QTL_by_lod_score.svg", base_height = 5, base_width = 13.5, plot = last_plot())
 ```
@@
          LOD = -0.5)
+
+test_rug4 <- pw_gg %>%
+  filter(log10BayesFactor > 2) %>% 
+  mutate(Pos = Pos/1000000, 
+         LOD = -0.5)
+test_rug5 <- pw_gamg %>%
+  filter(log10BayesFactor > 2) %>% 
+  mutate(Pos = Pos/1000000, 
+         LOD = -0.5)
+test_rug6 <- pw_mg %>%
+  filter(log10BayesFactor > 2) %>% 
+  mutate(Pos = Pos/1000000, 
+         LOD = -0.5)
 ```
@@
   as_tibble(rownames = "Marker") %>%
-  pivot_longer(GR50:dyln_change_sec, names_to = "Phenotype", values_to = "LOD") %>%
+  pivot_longer(GR50:crain_7d, names_to = "Phenotype", values_to = "LOD") %>%
+  filter(Phenotype %in% c("FL50", "dyln_change_sec_2d_prior", "crain_7d")) %>% 
   separate(Marker, into = c("Chr", "Pos"), sep = "_", remove = FALSE, convert = TRUE) %>%
@@
   geom_rug(data = test_rug3, outside = TRUE, sides = "b", alpha = 0.2) + 
-  geom_hline(data = threshold_df, aes(yintercept = LOD, color = Phenotype, linetype = Phenotype)) +
+  geom_hline(data = filter(threshold_df, 
+                           Phenotype %in% c("FL50", "dyln_change_sec_2d_prior", "crain_7d")),
+             aes(yintercept = LOD, color = Phenotype, linetype = Phenotype)) +
   facet_wrap(~Chr, nrow = 1, scales = "free_x") +
   switchgrassGWAS::theme_oeco + 
-  scale_color_manual(values = c("#CC0066", "#FFCC00", "#006600", "#6600FF", "grey")) + 
-  scale_linetype_manual(values = c(4,5,4,4,4)) +
+  scale_color_manual(values = c("#000099", "#CC0066", "darkgrey", "#000666")) + 
+  scale_linetype_manual(values = c(4,4,4,4,4)) +
   theme(legend.position = "top",
@@
         axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1),
-        panel.spacing.x = unit(-0., 'cm')
         ) + 
@@
   coord_cartesian(clip = "off")
-save_plot(filename = "../manuscript/plots/Figure_4_QTL_by_lod_score_w_rugs.svg", base_height = 3, base_width = 13.5, plot = last_plot())
+save_plot(filename = "../manuscript/revision2/Figure_4_FL50_QTL_by_lod_score_w_rugs.svg", base_height = 3, base_width = 13.5, plot = last_plot())
+
+s1full %>%
+  as_tibble(rownames = "Marker") %>%
+  pivot_longer(GR50:crain_7d, names_to = "Phenotype", values_to = "LOD") %>%
+  filter(Phenotype %in% c("GR50", "tave_1d", "dyln_14d_prior", "dyln_change_sec")) %>% 
+  separate(Marker, into = c("Chr", "Pos"), sep = "_", remove = FALSE, convert = TRUE) %>%
+  filter(!grepl("loc", Chr)) %>%
+  ggplot(aes(x = Pos, y = LOD)) + 
+  geom_line(aes(color = Phenotype)) +
+  geom_rug(data = test_rug5, sides = "t", alpha = 0.1) + 
+  geom_rug(data = test_rug4, sides = "b", alpha = 0.2) + 
+  geom_rug(data = test_rug6, outside = TRUE, sides = "b", alpha = 0.2) + 
+  geom_hline(data = filter(threshold_df, 
+                           Phenotype %in% c("GR50", "tave_1d", "dyln_14d_prior", 
+                                            "dyln_change_sec")), 
+             aes(yintercept = LOD, color = Phenotype, linetype = Phenotype)) +
+  facet_wrap(~Chr, nrow = 1, scales = "free_x") +
+  switchgrassGWAS::theme_oeco + 
+  scale_color_manual(values = c("#006600", "#FFCC00",  "grey", "darkorange", "#000666")) + 
+  scale_linetype_manual(values = c(4,4,4,4,4,4)) +
+  theme(legend.position = "top",
+        panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        text = element_text(size = 10, face="bold"),
+        strip.text = element_text(size = 10, face="bold"),
+        axis.title = element_text(size = 10,face="bold"),
+        axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1),
+        ) + 
+  xlab("Position (Mb)") + ylab("LOD score") +
+  scale_x_continuous(breaks = c(0, 20, 40, 60, 80)) + 
+  coord_cartesian(clip = "off")
+save_plot(filename = "../manuscript/revision2/Figure_4_GR50_QTL_by_lod_score_w_rugs.svg", base_height = 3, base_width = 13.5, plot = last_plot())
 ```

### 2024-02-21 R/Analysis_v1.2_mash_using_greedy_covar.qmd
feat: QTL analysis using phenotypes from new covariance matrices, incorporated into Figure 3

--- 2023-09-15 R/Analysis_v1.2_mash_using_greedy_covar.Rmd
@@
 ---
-title: "Analysis v1.2 mash"
+title: "Analysis for Revision 2"
 author: "Alice MacQueen"
-date: "2023-08-01"
-output: html_document
+format: 
+  html:
+    code-fold: true
+    code-annotations: hover
+    execute: 
+      echo: false
+      cache: true
+editor: visual
+toc: true
+number-sections: true
 ---
+
+::: callout-important
+## Analysis Version 1.2
+
+As of 2024-02-07, this file was a copy of the Rmd file of the same name. After 2024-02-07, this Quarto document will be modified and kept up to date for this analysis version.
+:::
 
@@
 
-# Installation reminders for any reinstalls
+::: callout-tip
+## Install reminders for reinstalls
+
 On debian: `apt install libgsl-dev` to install the GNU Scientific Library, zB.
@@
 For mashr: Sometimes install.packages("RcppGSL") is also needed.
-#
-
-Sam used a greedy mash algorithm to find the best set of covariance matrices to include, using 19K randomly associated variants.
-
-Now use the best set of covariance matrices on 19K strongly associated variants.
-
-Step 0. Set up.
-  0a. Look at phenotype distribution & heritability; run GWAS to get univariate SNP effect estimates & standard errors. Already done.
-  0b.
+:::
+
+## Overview
+
+Sam developed a greedy mash algorithm to find the best set of covariance matrices to include.
+Alice used this algorithm on 19K randomly associated variants using a large set of covariance matrices with h\^2 on their diagonals.
+
+Then, Alice used the best set of covariance matrices on 19K strongly associated variants.
+
+Outline of the current analysis:
+
+1.  Set up. Look at phenotype distribution & heritability; run GWAS to get univariate SNP effect estimates & standard errors. Done elsewhere.
+
+2.
 Make hypothesis covariance matrices based on weather data from 2018 - 2020.
-For cumulative weather variables, use a range of days, 1 - 20, say.
-Come back and add more if 20 is chosen by greedy mash.
-Step 1.
+For cumulative weather variables, use a range of days: every day for the first week, and every week until 28 days.
+Come back and add more if 28 is chosen by greedy mash.
+
+3.
 Run greedy mash to find best set of covariance matrices that explain SNP effect variation across environments, using 19K 'randomly associated variants'.
-Go back to 0b.
-if high number of day covar matrices tend to be chosen by greedy mash.
-Step 2. Use covariances from (1) to run mash on 19K 'strongly associated variants'. 
-3. Look at loadings on covariance matrices and meanings of covariance matrices as the writeup (Figure 2, 3; Table 1). Look at amount of antagonistic pleiotropy vs same sign significant effects.
-4. Repeat the QTL analysis... with Li.
-
-How much of a lift is it to re-run the greedy mash algorithm?
-If we, or if reviewers are thinking about this in a Jiaming Yu sense, I think the next logical question to ask is 'Why these covariance matrices'?
-The overly honest answer for the current set is that there were computational limits to running mash with so many covariance matrices, so I chose a few arbitrary time points for each weather variable we thought might be a environmental driver.
-But now we have introduced a method to choose between highly correlated covariance matrices, perhaps we could do better.
-2.
-
-# ----
-# Generate weather variables
-
-# Covar 1-28d weather
-As of 2023-09-11, want to remake these per reviewer revision to have h2 on the diagonal instead of the CV. Need to estimate h2 for these new covar matrices, probably using rrBLUP.
-
-## load phe & wea data
-```{r}
+Go back to (1) if high number of day covar matrices tend to be chosen by greedy mash.
+
+4.  Use covariances from (3) to run mash on 19K 'strongly associated variants'.
+
+5.  Look at loadings on covariance matrices and meanings of covariance matrices as the writeup (Figure 2, 3; Table 1). Look at amount of antagonistic pleiotropy vs same sign significant effects.
+
+6.  Have Li repeat the QTL analysis and incorporate her results.
+
+------------------------------------------------------------------------
+
+## Step 1: GWAS
+
+## Step 2: Covariance matrices
+
+### Generate weather variables
+
+### Covar 1-28d weather
+
+As of 2023-09-11, want to remake these per reviewer revision to have h\^2 on the diagonal instead of the CV. Estimate h\^2 for these new covar matrices using rrBLUP.
+
+### load phe & wea data
+
+```{r}
+#| eval: false
+
 phe_ten_sites <- readRDS(here("data", "Phenotypes_2018_to_2021_10_sites.rds"))
@@
 ```{r}
-
+#| eval: false
+#| 
 wea_ten_sites <- wea_ten_sites %>%
@@
 ```{r}
+#| eval: false
 wea_ten_sites
@@
 # Define phenotypes
-```{r}
+
+```{r}
+#| eval: false
+#| 
 Tbase <- 12
@@
 ## save weather-based phenotypes
-```{r}
+
+```{r}
+#| eval: false
 gr_ten_sites %>%
@@
 ```
+
 # --- can start here
+
 # Covar csv for greedy mash algorithm
 
-Make & save hypothesis matrices in a hypothesis_matrices directory in data. 
-Then run the greedy mash algorithm again.
-
-phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", 
-             "tave_5d", "tave_10d", "tave_18d")
-phe_hyp <- c("FL50", "GR50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
-             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")
+Make & save hypothesis matrices in a hypothesis_matrices directory in data. Then run the greedy mash algorithm again.
+
+phe_hyp \<- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", "tave_5d", "tave_10d", "tave_18d") phe_hyp \<- c("FL50", "GR50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")
 
 ### load weather phenotypes
+
 ```{r}
@@
 ## Calculate heritability for diagonal
+
 I don't have access to ASREML any more, so need a different package to do this.
@@
 Use the kinship matrix I've already calculated (using the van Raden method, using switchgrassGWAS::pvdiv_kinship()) using the SNPs as the K matrix.
-```{r}
+
+```{r}
+#| eval: false
+
 library(rrBLUP)
@@
 
-Another error: "Error in mash(data, covmats) : 
-  All U_scaled matrices should be positive semi-definite
+Another error: "Error in mash(data, covmats) : All U_scaled matrices should be positive semi-definite
 
@@
 ### Loop to calculate covar for flowering
-```{r}
+
+```{r}
+#| eval: false
+
 metadata <- read_rds(here("data", "metadata.rds"))
@@
 ```
-Another error: "Error in mash(data, covmats) : 
-  All U_scaled matrices should be positive semi-definite
+
+Another error: "Error in mash(data, covmats) : All U_scaled matrices should be positive semi-definite
 
 ### Loop to calculate covar for greenup
-```{r}
+
+```{r}
+#| eval: false
+#| 
 gr_ten_sites <- gr_ten_sites %>%
@@
 #### flowering: 6 sites
+
 MI, MO, SD, TX1, TX2, TX3
+
 ```{r flowering midwest}
+#| eval: false
 
@@
 #### greenup: 7 sites
+
 MI, MO, NE, OK, TX1, TX2, TX3
-```{r}
+
+```{r}
+#| eval: false
+
 phe_hyp <- colnames(gr_ten_sites)[8:50]
@@
 ```{r}
+#| eval: false
 workingdir <- here("analysis", "mash_greedy_algorithm", 
@@
 #### Visualize correlation matrices
-```{r}
-
+
+```{r}
+#| eval: false
 i=7
@@
 # ---------------
+
+## Step 3: Greedy Mash
+
 # Info to run greedy mash algo
 
-data/mash_switchgrass_models/mash/all_matrices/matrices hosts all covar mat (?)
-analysis/mash_greedy_algorithm/mash_switchgrass_models/mash/all_matrices/matrices has more? covar mat?
+data/mash_switchgrass_models/mash/all_matrices/matrices hosts all covar mat (?) analysis/mash_greedy_algorithm/mash_switchgrass_models/mash/all_matrices/matrices has more?
+covar mat?
 Can cp from these matrices dir to mash_h2_diag_covar/matrices the following regex, which run.mash.greedy.search.py looks for:
 
- '.singleton.*'
-'.identity.*'
-'.simple_het_*.*'
- '.equal_effects.*'
-'.*PCA*'
+'.singleton.*' '.identity.*' '.simple_het\_*.*' '.equal_effects.*' '.*PCA\*'
 
@@
 
-When I change to my inputs for effects & stderrs, I will need to adjust midwest covar & effects appropriately to reflect only 6 (flowering) and 7 (greenup) rows & columns used; the other GWAS were dropped as too low power or due to substantial phenotype data loss, & I think including them would be misleading, likely.
+When I change to my inputs for effects & stderrs, I will need to adjust midwest covar & effects appropriately to reflect only 6 (flowering) and 7 (greenup) rows & columns used; the other GWAS were dropped as too low power or due to substantial phenotype data loss, & including them would give misleading results.
 
 ### Next steps
+
 Redo Midwest: Remake Midwest covar matrices, including hypothesis, singleton, identity, simple het, equal effects, and data-driven matrices.
 Make mash_greedy.R variants for changed input dir & for different Midwest column numbers.
-Run two new run.mash.greedy.search.py commands:
-python run.mash.greedy.search.py -p greenup -c midwest -e random
-python run.mash.greedy.search.py -p flowering -c midwest -e random
-
-Try Gulf & All with my known inputs:
-python run.mash.greedy.search.py -p greenup -c gulf -e random
-python run.mash.greedy.search.py -p greenup -c all -e random
-python run.mash.greedy.search.py -p flowering -c gulf -e random
-python run.mash.greedy.search.py -p flowering -c all -e random
-
-Write inputs to mash_h2_diag_covar/All_inputs, Gulf_inputs, and Midwest_inputs, from the original mash work. Especially important for Midwest which has some poor GWAS at dropped sites.
+Run two new run.mash.greedy.search.py commands: python run.mash.greedy.search.py -p greenup -c midwest -e random python run.mash.greedy.search.py -p flowering -c midwest -e random
+
+Try Gulf & All with my known inputs: python run.mash.greedy.search.py -p greenup -c gulf -e random python run.mash.greedy.search.py -p greenup -c all -e random python run.mash.greedy.search.py -p flowering -c gulf -e random python run.mash.greedy.search.py -p flowering -c all -e random
+
+Write inputs to mash_h2_diag_covar/All_inputs, Gulf_inputs, and Midwest_inputs, from the original mash work.
+Especially important for Midwest which has some poor GWAS at dropped sites.
 The inputs Sam used have no column or rownames and look to be tab delimited.
 
-aka combined_mash_effects/greenup.all.betas.random.txt 
-and combined_mash_stderrs/flowering.gulf.betas.strong.txt
+aka combined_mash_effects/greenup.all.betas.random.txt and combined_mash_stderrs/flowering.gulf.betas.strong.txt
 
 I only really need the random ones here, run mash again on the strong effects below.
-```{r}
+
+```{r}
+#| eval: false
 popkey = list(all = "Gulf_and_Midwest", midwest = "Midwest", gulf = "Gulf")
@@
 ```{r}
+#| eval: false
 for (i in 1:35) {
@@
 # -----------------------------
-# Run mash
-
-```{r}
+
+# Step 4: Run mash
+
+```{r}
+#| eval: false
+#| 
+#|
+
 library(tidyverse)
@@
 
-Some univariate GWAS for the Midwest had population structure issues. Mostly for greenup - even GR50 at MO was quite bad. Drop MO for GR50.
-Midwest_GR50_SD have values of zero for Bhat and Shat. Drop this from Midwest analyses.
-
-# Remake mash figures
+Some univariate GWAS for the Midwest had population structure issues.
+Mostly for greenup - even GR50 at MO was quite bad.
+Drop MO for GR50.
+Midwest_GR50_SD have values of zero for Bhat and Shat.
+Drop this from Midwest analyses.
+
+# Step 5: Redo writeup & figures
 
 ## Load mash objects
+
 ```{r}
@@
 ## Figure 2
+
 ```{r}
@@
 get_cov_type(m_midw_fl_hyp, summarise = TRUE, subpop = TRUE)
-get_cov_type(m_gam_gr_hyp, summarise = TRUE, subpop = TRUE)
+get_cov_type(m_gam_gr_hyp, summarise = FALSE, subpop = TRUE)
 get_cov_type(m_gam_fl_hyp, summarise = TRUE, subpop = TRUE)
@@
 ```
+
 ### covar
@@
 legend <- get_legend(cov_gr1) 
-fig_eqeff <- cov_gr1$greenup.simple_het_2_df %>%
-  get_covar_fig2() +
+fig_eqeff <- plot_covar_Ulist_fig2(Ulist_mg[[2]], mg = TRUE) +
   labs(subtitle = "simple heterozygousity")
@@
 ```{r build figure 2}
+
 ggdraw() +
@@
 
-save_plot(filename = "../manuscript/Coauthors/PNAS/Revision/Figure_2_Covariance_matrices_and_posterior_weights.svg", plot = last_plot(), base_height = 6, base_width = 8.5, )
+```
+
+```{r}
+#| eval: false
+
+save_plot(filename = "../manuscript/Coauthors/PNAS/Revision/Figure_2_Covariance_matrices_and_posterior_weights.svg", plot = last_plot(), base_height = 6, base_width = 8.5)
 ```
@@
 Change these to look at S_AP, S_DS, S_2_no; S_1_row and S_1_col separately
+
 ```{r}
@@
   gxe_frac <- gxe_n %>%
-    mutate(North_North = round(North_North/sum(North_North), digits = 4),
-           North_Texas = round(North_Texas/sum(North_Texas), digits = 4),
-           Texas_Texas = round(Texas_Texas/sum(Texas_Texas), digits = 4))
+    mutate(North_North = round(North_North/sum(North_North), digits = 3),
+           North_Texas = round(North_Texas/sum(North_Texas), digits = 3),
+           Texas_Texas = round(Texas_Texas/sum(Texas_Texas), digits = 3))
   return(list(gxe_counts = gxe_n, gxe_fractions = gxe_frac))
@@
 get_gxe_summary(gxe_gg) 
+get_gxe_summary(gxe_mg, n_cond = 7) 
+get_gxe_summary(gxe_gamg, n_sep = 4) 
+print("flowering")
 get_gxe_summary(gxe_gf) 
-get_gxe_summary(gxe_mg, n_cond = 7) 
 get_gxe_summary(gxe_mf, n_cond = 6) 
-get_gxe_summary(gxe_gamg, n_sep = 4) 
 get_gxe_summary(gxe_gamf, n_sep = 4) 
@@
 ```{r}
-mash_plot_Ulist(m_midw_fl_hyp)
-mash_plot_manhattan_by_condition(m_midw_fl_hyp)
+ pvdiv_plot_cor(corrmatrix = gxe_gamf$S_AP, legend = TRUE) + 
+  labs(subtitle = " ")
 ```
@@
 
+```
+
+```{r}
+#| eval: false
+
 save_plot(filename = "../manuscript/Coauthors/PNAS/Revision/Figure_3_Effect_Types_by_Site.svg", plot = last_plot(), base_height = 10, base_width = 5.5)
@@
 ## Figure 4
+
+Comparison with QTL for the four-way mapping cross for green up (GR50), flowering (FL50), three functions of weather variables for green-up (from Figure 2; Both: Temp Ave, 1d; Both: Daylength change (s) 1d prior; Gulf: Daylength (14d prior), and two functions of weather variables for flowering (from Figure 2: Midwest: day length change (s) 2d prior; Gulf: cumulative rainfall, 7d).
+
+Li ran the QTL analysis 2024-02-14. Here Alice combines the QTL with the mash SNP effects to build the analysis for Figure 4.
+
+It's an open question how to combine these into an informative main text figure.
+
+The full QTL Lod plots for flowering & greenup phenotypes should go in the SI. Here, try plotting just the significant chromosomes, with a QTL above the permutation threshold.
+
+#### Figure 4A
+
+LOD QTL
+
+Adapted from Analysis_v0.7_QTL_interval_lod_plot.Rmd
+
+Make sure to Load mash objects at start of Step 5
+
+```{r}
+library(qtl)
+library(qtl2)
+# library(plyr)
+# library(tidyverse)
+library(qtl2convert)
+#library(cowplot)
+```
+
+```{r}
+workingdir <- "~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/"
+load(file.path(workingdir, 'cross_reduced_marker_FL50_phenotypes_Alice_7_SITES.RData'))
+load(file.path(workingdir, 'scan1Results_FL50_phenotypes_Alice_7_SITES.RData'))
+
+map  =  lapply(pull.map(cross), function(x) x[1,])
+###get the threshold value for each model and output the QTL intervals
+threshold_full = summary(s1permfull,alpha = 0.05)
+threshold_reduced = summary(s1permreduced,alpha = 0.05)
+```
+
+```{r}
+###  g x e
+gxe = s1full - s1reduced
+gxe_perm = threshold_full - threshold_reduced ###or gxe_perm = summary(gxe, alpha=0.05)
+
+###get the peak for full model and gxe model
+peaks_full = find_peaks(s1full, map, threshold = threshold_full, drop = 1.5, peakdrop = 5)
+peaks_gxe = find_peaks(gxe, map, threshold = gxe_perm, drop = 1.5)
+
+flank_marker_full = peaks_full %>% mutate(flank_lo = find_marker(map, chr, ci_lo), flank_hi=find_marker(map, chr, ci_hi))
+# write.csv(flank_marker_full,'QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv') ###change the name you want
+
+flank_marker_gxe = peaks_gxe %>% mutate(flank_lo = find_marker(map, chr, ci_lo), flank_hi=find_marker(map, chr, ci_hi))
+# write.csv(flank_marker_gxe,'QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_gxe_model.csv')
+
+###########
+```
+
+```{r}
+get_post_weights <- function(m){
+  log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
+      as.data.frame() %>%
+      rownames_to_column(var = "value") %>%
+      dplyr::mutate(value = as.integer(.data$value)) %>%
+      as_tibble() %>%
+      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
+      dplyr::rename(log10BayesFactor = .data$V1) %>%
+      dplyr::select(-.data$value)
+
+  pw_df <- as_tibble(m$posterior_weights)
+  pw_df <- switchgrassGWAS:::get_marker_df(m = m) %>%
+    bind_cols(pw_df)
+
+  ggman_df <- pw_df %>%
+    #left_join(m_meta) %>%
+    tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
+             extra = "merge", convert = TRUE) %>%
+    left_join(log10bf_df, by = "Marker") %>%
+    dplyr::arrange(.data$Chr, .data$Pos) 
+  return(ggman_df)
+}
+
+pw_gg <- get_post_weights(m_gulf_gr_hyp)
+pw_gf <- get_post_weights(m_gulf_fl_hyp) %>%
+  mutate(Subpop = "Gulf")
+
+pw_mg <- get_post_weights(m_midw_gr_hyp)
+pw_mf <- get_post_weights(m_midw_fl_hyp) %>%
+  mutate(Subpop = "Midwest")
+
+pw_gamg <- get_post_weights(m_gam_gr_hyp)
+pw_gamf <- get_post_weights(m_gam_fl_hyp) %>%
+  mutate(Subpop = "Both")
+```
+
+Decide on cutoff to use for rug plots and enrichment analysis: SNPs sig in 1% tail or BF>2, whichever is stricter
+
+1% tail is stricter in all cases except Midwest flowering, where BF>2 is stricter, equivalent to a 0.49% tail (0.995119 quantile).
+
+What is the Bayes Factor at the 99th quantile? If it's not above 2, what is the quantile where the BF > 2?
+```{r}
+quantile(pw_gf$log10BayesFactor, 0.99)  # Gulf: Use 2.56 BF
+quantile(pw_mf$log10BayesFactor, 0.99)  # Midwest: Use 2 BF
+quantile(pw_gamf$log10BayesFactor, 0.99)  # GAM: Use 2.65 BF
+quantile(pw_mf$log10BayesFactor, 0.995119)
+
+quantile(pw_gg$log10BayesFactor, 0.99)  # Gulf: Use 2.0989 BF
+quantile(pw_mg$log10BayesFactor, 0.99)  # Midwest: Use 4.6831 BF
+quantile(pw_gamg$log10BayesFactor, 0.99)  # GAM: Use 3.0865 BF
+```
+
+```{r}
+rug_gf <- pw_gf %>%
+  filter(log10BayesFactor > 2.559618) %>% 
+  mutate(Pos = Pos/1000000, 
+         LOD = -0.5)
+rug_gamf <- pw_gamf %>%
+  filter(log10BayesFactor > 2.64938) %>% 
+  mutate(Pos = Pos/1000000, 
+         LOD = -0.5)
+rug_mf <- pw_mf %>%
+  filter(log10BayesFactor > 2) %>% 
+  mutate(Pos = Pos/1000000, 
+         LOD = -0.5)
+
+rug_gg <- pw_gg %>%
+  filter(log10BayesFactor > 2.0989) %>% 
+  mutate(Pos = Pos/1000000, 
+         LOD = -0.5)
+rug_gamg <- pw_gamg %>%
+  filter(log10BayesFactor > 3.0865) %>% 
+  mutate(Pos = Pos/1000000, 
+         LOD = -0.5)
+rug_mg <- pw_mg %>%
+  filter(log10BayesFactor > 4.6831) %>% 
+  mutate(Pos = Pos/1000000, 
+         LOD = -0.5)
+```
+
+```{r}
+threshold_df <- threshold_full %>% 
+  as_tibble(rownames = "Threshold") %>%
+  pivot_longer(GR50:crain_7d, names_to = "Phenotype", values_to = "LOD") %>%
+  mutate(Phe = case_when(Phenotype %in% c("GR50", "tave_1d", "dyln_14d_prior", "dyln_change_sec") ~ "greenup",
+                         Phenotype %in% c("FL50", "dyln_change_sec_2d_prior", "crain_7d") ~ "flowering", 
+                         TRUE ~ NA_character_))
+
+s1full %>%
+  as_tibble(rownames = "Marker") %>%
+  pivot_longer(GR50:crain_7d, names_to = "Phenotype", values_to = "LOD") %>%
+  filter(Phenotype %in% c("FL50", "dyln_change_sec_2d_prior", "crain_7d")) %>% 
+  separate(Marker, into = c("Chr", "Pos"), sep = "_", remove = FALSE, convert = TRUE) %>%
+  filter(!grepl("loc", Chr)) %>%
+  ggplot(aes(x = Pos, y = LOD)) + 
+  geom_line(aes(color = Phenotype)) +
+  geom_rug(data = rug_gamf, sides = "t", alpha = 0.1) + 
+  geom_rug(data = rug_gf, sides = "b", alpha = 0.2) + 
+  geom_rug(data = rug_mf, outside = TRUE, sides = "b", alpha = 0.2) + 
+  geom_hline(data = filter(threshold_df, 
+                           Phenotype %in% c("FL50", "dyln_change_sec_2d_prior", "crain_7d")),
+             aes(yintercept = LOD, color = Phenotype, linetype = Phenotype)) +
+  facet_wrap(~Chr, nrow = 1, scales = "free_x") +
+  switchgrassGWAS::theme_oeco + 
+  scale_color_manual(values = c("#000099", "#CC0066", "darkgrey", "#000666")) + 
+  scale_linetype_manual(values = c(4,4,4,4,4)) +
+  theme(legend.position = "top",
+        panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        text = element_text(size = 10, face="bold"),
+        strip.text = element_text(size = 10, face="bold"),
+        axis.title = element_text(size = 10,face="bold"),
+        axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1),
+        ) + 
+  xlab("Position (Mb)") + ylab("LOD score") +
+  scale_x_continuous(breaks = c(0, 20, 40, 60, 80)) + 
+  coord_cartesian(clip = "off")
+save_plot(filename = "../manuscript/revision2/Figure_4_FL50_QTL_by_lod_score_w_rugs.svg", base_height = 3, base_width = 13.5, plot = last_plot())
+
+s1full %>%
+  as_tibble(rownames = "Marker") %>%
+  pivot_longer(GR50:crain_7d, names_to = "Phenotype", values_to = "LOD") %>%
+  filter(Phenotype %in% c("GR50", "tave_1d", "dyln_14d_prior", "dyln_change_sec")) %>% 
+  separate(Marker, into = c("Chr", "Pos"), sep = "_", remove = FALSE, convert = TRUE) %>%
+  filter(!grepl("loc", Chr)) %>%
+  ggplot(aes(x = Pos, y = LOD)) + 
+  geom_line(aes(color = Phenotype)) +
+  geom_rug(data = rug_gamg, sides = "t", alpha = 0.1) + 
+  geom_rug(data = rug_gg, sides = "b", alpha = 0.2) + 
+  geom_rug(data = rug_mg, outside = TRUE, sides = "b", alpha = 0.2) + 
+  geom_hline(data = filter(threshold_df, 
+                           Phenotype %in% c("GR50", "tave_1d", "dyln_14d_prior", 
+                                            "dyln_change_sec")), 
+             aes(yintercept = LOD, color = Phenotype, linetype = Phenotype)) +
+  facet_wrap(~Chr, nrow = 1, scales = "free_x") +
+  switchgrassGWAS::theme_oeco + 
+  scale_color_manual(values = c("#006600", "#FFCC00",  "grey", "darkorange", "#000666")) + 
+  scale_linetype_manual(values = c(4,4,4,4,4,4)) +
+  theme(legend.position = "top",
+        panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        text = element_text(size = 10, face="bold"),
+        strip.text = element_text(size = 10, face="bold"),
+        axis.title = element_text(size = 10,face="bold"),
+        axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1),
+        ) + 
+  xlab("Position (Mb)") + ylab("LOD score") +
+  scale_x_continuous(breaks = c(0, 20, 40, 60, 80)) + 
+  coord_cartesian(clip = "off")
+save_plot(filename = "../manuscript/revision2/Figure_4_GR50_QTL_by_lod_score_w_rugs.svg", base_height = 3, base_width = 13.5, plot = last_plot())
+```
+
+#### Figure 4B
+
+Same SVG line drawing that was used previously. Description of the two mapping panels used.
+
+#### Figure 4C
+
+Enrichment analysis
+
+Adapted from Analysis_v0.6_QTL_GxE_and_mash_comparison.Rmd
+
+Prepare the QTL dataframe
+```{r}
+qtldf <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv")
+
+qtldf <- qtldf %>%
+  separate(flank_lo, into = c("CHR", "Pos_Mb_lo"), sep = "_", convert = TRUE) %>%
+  separate(flank_hi, into = c("flank", "Pos_Mb_hi"), sep = "_", convert = TRUE) %>%
+  mutate(Pos_Mb_lo = round(Pos_Mb_lo, digits = 2),
+         Pos_Mb_hi = round(Pos_Mb_hi, digits = 2)) %>%
+  dplyr::select(CHR, Pos_Mb_lo, Pos_Mb_hi, lodcolumn, lod, everything()) %>%
+  arrange(CHR, Pos_Mb_lo)
+
+qtlplot <- qtldf %>%
+  mutate(phe = case_when(lodcolumn == "FL50" ~ "flowering",
+                              lodcolumn == "dyln_change_sec_2d_prior" ~ "daylength change (s) 2d prior",
+                              lodcolumn == "dyln_change_sec" ~ "daylength change (s) 1d prior",
+                              lodcolumn == "GR50" ~ "greenup",
+                              lodcolumn == "crain_gr2fl" ~ "Rainfall, GR to FL")) %>%
+  #filter(phe %in% c("Daylength", "GDD, GR to FL")) %>%
+  mutate(POS_lo = Pos_Mb_lo * 1000000,
+         POS_hi = Pos_Mb_hi * 1000000)
+
+qtlplot$phe <- factor(qtlplot$phe, 
+                      levels = c("flowering", "daylength change (s) 2d prior",
+                                 "GDD, GR to FL", "Daylength",
+                                          "daylength change (s)",
+                                          "Rainfall, 1 day",
+                                                  #"Rainfall, 2d sum",
+                                                  "Rainfall, 3d sum",
+                                                  "Rainfall, 5d sum"#,
+                                                  #"Rainfall, 7d sum"
+                                                  ))
+
+qtlplot_fl <- qtlplot %>% 
+  filter(lodcolumn %in% c("FL50", "dyln_change_sec_2d_prior")) 
+qtlplot_gr <- qtlplot %>% 
+  filter(lodcolumn %in% c("GR50", "dyln_change_sec")) 
+```
+Function to compute hypergeometric test for each row in a QTL df
+```{r}
+
+get_QTL_enrichment <- function(m, qtl_df, quantile = 0.99){
+  pw_df_all <- get_post_weights(m)
+  # do not filter for unique LOD intervals for the random QTL intervals 
+  # do this by not having lod in the random QTL intervals
+  if("lod" %in% colnames(qtl_df)) {
+    qtl_df <- qtl_df %>%
+      arrange(desc(lod)) %>%
+      mutate(pos = as.integer(pos)) %>%
+      group_by(CHR, pos) %>%
+      dplyr::slice(1) %>%
+      arrange(CHR, pos)
+  }
+
+  BFthresh <- quantile(pw_df_all$log10BayesFactor, quantile)
+  pw_df <- pw_df_all %>%
+    filter(log10BayesFactor > BFthresh) %>%
+    dplyr::select(Chr:Pos, log10BayesFactor, everything())
+
+  for(i in 1:nrow(qtl_df)){  # from 1 to 16
+  ## In block & > 2
+  QTLandBF <- pw_df %>%
+    filter(Chr %in% qtl_df$CHR[i] & between(Pos, qtl_df$POS_lo[i],
+                                             qtl_df$POS_hi[i])) %>% 
+    tally()
+  ## BF>2
+  inBF <- pw_df %>%
+    tally()
+  ## BF not > 2 (all minus BF > 2, all is below)
+  allwrtBF <- pw_df_all %>%
+    tally()
+  ## All in block
+  inQTL <- pw_df_all %>%
+    filter(Chr %in% qtl_df$CHR[i] & between(Pos, qtl_df$POS_lo[i],
+                                             qtl_df$POS_hi[i])) %>%
+    tally()
+  notinBF <- (allwrtBF$n[1]-inBF$n[1])
+  pval <- phyper(QTLandBF$n[1], inBF$n[1], allwrtBF$n[1]-inBF$n[1], 
+                 inQTL$n[1], lower.tail = FALSE)
+
+  if(i == 1){
+  outputdf <- tibble(Chr = qtl_df$CHR[i], Pos_lo = qtl_df$POS_lo[i], 
+                     Pos_hi = qtl_df$POS_hi[i], inQTLandBF = QTLandBF$n[1], 
+                     inBF = inBF$n[1], notinBF = notinBF, 
+                     inQTL = inQTL$n[1], pvalue = pval)
+  } else {
+    outputdf <- outputdf %>%
+      add_row(Chr = qtl_df$CHR[i], Pos_lo = qtl_df$POS_lo[i], 
+              Pos_hi = qtl_df$POS_hi[i],inQTLandBF = QTLandBF$n[1], 
+              inBF = inBF$n[1], notinBF = notinBF, 
+              inQTL = inQTL$n[1], 
+              pvalue = pval)
+    }
+  }
+  return(outputdf)
+}
+
+```
+
+```{r}
+enrich_gf <- get_QTL_enrichment(m_gulf_fl_hyp, qtl_df = qtlplot_fl, 0.99) %>%
+  mutate(subpop = "Gulf", phenotype = "flowering")
+enrich_mf <- get_QTL_enrichment(m_midw_fl_hyp, qtl_df = qtlplot_fl, 0.995119) %>%
+  mutate(subpop = "Midwest", phenotype = "flowering")
+enrich_gamf <- get_QTL_enrichment(m_gam_fl_hyp, qtl_df = qtlplot_fl, 0.99) %>%
+  mutate(subpop = "Both", phenotype = "flowering")
+
+enrich_df <- enrich_gf %>%
+  full_join(enrich_mf) %>%
+  full_join(enrich_gamf) %>%
+  arrange(Chr, Pos_lo)
+enrich_df <- enrich_df %>%
+  left_join(qtlplot_fl, by = c("Chr" = "CHR", "Pos_lo" = "POS_lo", "Pos_hi" = "POS_hi")) %>%
+  dplyr::select(Chr, Pos_lo, Pos_hi, inQTLandBF:subpop, lod, lodcolumn, pos) %>%
+  unique() %>%
+  arrange(Chr, Pos_lo, lodcolumn)
+
+enrich_df %>%
+  filter(pvalue < 0.05)
+
+# t.test(enrich_df[enrich_df$pvalue < 0.05,]$lod, enrich_df[enrich_df$pvalue >= 0.05,]$lod) 
+qtlplot_fl %>% group_by(CHR, POS_lo) %>% tally()
+# 10 unique QTL by Pos lo/hi. 5 of 10 have enriched mash hits.
+```
+
+```{r}
+enrich_gg <- get_QTL_enrichment(m_gulf_gr_hyp, qtl_df = qtlplot_gr, 0.99) %>%
+  mutate(subpop = "Gulf", phenotype = "greenup")
+enrich_mg <- get_QTL_enrichment(m_midw_gr_hyp, qtl_df = qtlplot_gr, 0.99) %>%
+  mutate(subpop = "Midwest", phenotype = "greenup")
+enrich_gamg <- get_QTL_enrichment(m_gam_gr_hyp, qtl_df = qtlplot_gr, 0.99) %>%
+  mutate(subpop = "Both", phenotype = "greenup")
+
+enrich_gr <- enrich_gg %>%
+  full_join(enrich_mg) %>%
+  full_join(enrich_gamg) %>%
+  arrange(Chr, Pos_lo)
+enrich_gr <- enrich_gr %>%
+  left_join(qtlplot_gr, by = c("Chr" = "CHR", "Pos_lo" = "POS_lo", "Pos_hi" = "POS_hi")) %>%
+  dplyr::select(Chr, Pos_lo, Pos_hi, inQTLandBF:subpop, lod, lodcolumn, pos) %>%
+  unique() %>%
+  arrange(Chr, Pos_lo, lodcolumn)
+
+enrich_gr  # 0 of 3 unique QTL have enriched mash hits for green-up.
+
+```
+
+```{r}
+#| eval: false
+
+pw_3f_all <- pw_gf %>%
+  full_join(pw_mf) %>%
+  full_join(pw_gamf) %>%
+  arrange(Chr, Pos)
+
+sig_random <- c()
+chr_size_mash <- pw_3f_all %>%
+  ungroup() %>%
+  group_by(Chr) %>%
+  dplyr::summarise(Pos_lo = min(Pos),
+                   Pos_hi = max(Pos))
+j = 1
+for(j in 1:1000){
+  qtl_size <- qtlplot_fl %>%
+    dplyr::select(CHR, POS_lo, POS_hi) %>%
+    dplyr::distinct() %>%   # don't doublecount completely overlapping QTL
+    #(does doublecount partially overlapping QTL, however.)
+    group_by(CHR, POS_lo) %>%
+    dplyr::mutate(Size = POS_hi - POS_lo) %>%
+    dplyr::select(CHR, POS_lo, Size)
+
+  rchr <- chr_size_mash$Chr[sample(1:18, nrow(qtl_size), replace = TRUE)]
+  rposlo <- c()
+  for(i in 1:length(rchr)){
+    rposlo[i] <- sample(chr_size_mash$Pos_lo[which(chr_size_mash$Chr == rchr[i])]:chr_size_mash$Pos_hi[which(chr_size_mash$Chr == rchr[i])], 1)
+  }
+
+  random_qtl <- tibble(CHR = rchr, POS_lo = rposlo, Size = qtl_size$Size) %>%
+    mutate(POS_hi = POS_lo + Size) %>%
+    arrange(CHR, POS_lo)
+
+  random_gf <- get_QTL_enrichment(m = m_gulf_fl_hyp, qtl_df = random_qtl, 0.99) %>%
+    mutate(subpop = "Gulf", phenotype = "flowering")
+  random_mf <- get_QTL_enrichment(m_midw_fl_hyp, qtl_df = random_qtl, 0.995119) %>%
+    mutate(subpop = "Midwest", phenotype = "flowering")
+  random_gamf <- get_QTL_enrichment(m_gam_fl_hyp, qtl_df = random_qtl, 0.99) %>%
+    mutate(subpop = "Both", phenotype = "flowering")
+
+  random_df <- random_gf %>%
+    full_join(random_mf) %>%
+    full_join(random_gamf) %>%
+    arrange(Chr, Pos_lo)
+
+  sig_random[j] <- nrow(filter(random_df, pvalue < 0.05))
+}
+saveRDS(sig_random, file = "../data/Random_QTL_df_FL50_significant_enrichments_of_mash_1per_tail_or_mash_BF2_tail_SNPs_1000runs.rds")
+
+```
+
+```{r}
+sig_random <- readRDS(file = "../data/Random_QTL_df_FL50_significant_enrichments_of_mash_1per_tail_or_mash_BF2_tail_SNPs_1000runs.rds")
+
+tibble(`mash 1% tail significant enrichments` = sig_random) %>%
+  ggplot(aes(x = `mash 1% tail significant enrichments` )) +
+  switchgrassGWAS::theme_oeco +
+  geom_histogram(binwidth = 1) + 
+  xlim(c(1, 24)) + ylab("") +
+  geom_vline(xintercept = 5, linetype = 2, color = "red")
+save_plot(filename = "../manuscript/revision2/Figure_4_QTL_overlaps_random_QTL_enrichments_with_mash_1000runs.png", plot = last_plot(), base_height = 1.8, base_asp = 1.8)
+```

### 2024-02-21 manuscript/revision2/revision2.qmd
feat: QTL analysis using phenotypes from new covariance matrices, incorporated into Figure 3

@@
 Identifying the mixture and prevalence of GxE primed by weather prior to the phenological event (GxWeather) should aid prediction and manipulation of phenological timings.
-Here, we map GxWeather effects on phenological timings in two highly divergent switchgrass (*Panicum virgatum*) populations using repeated plantings of these populations at eight sites spanning the central United States.
+Here, we map GxWeather effects on phenological timings in two highly divergent switchgrass (*Panicum virgatum*) populations using repeated plantings of cloned individuals from these populations at eight sites spanning the central United States.
 We distinguish GxWeather patterns covarying with interpretable weather-based cues from agnostic, site-based patterns.
@@
 33% of effects on the timing of flowering in the Gulf population covary with cumulative rainfall in the seven days prior to flowering, while 22% of effects on flowering timing in the Midwest population covary with day length change two days prior to flowering.
-An independent pseudo-F2 cross of Gulf and Midwest individuals at the same sites mapped 23 additive QTLs for flowering, all with significant associations and eleven with enrichment of highly significant associations in our model of GxWeather effects.
-We demonstrate that we can identify QTL with GxWeather and assign them to specific weather-based cues or other patterns.
+An independent pseudo-F2 cross of Gulf and Midwest individuals at the same sites mapped 10 additive QTLs for flowering and three for vegetative growth timing, all with significant associations and three with enrichment of highly significant associations in our model of GxWeather effects.
+We demonstrate that we can identify genetic variation with GxWeather and assign these loci to specific weather-based cues or other patterns.
 Breeding for particular alleles at these loci could change flowering responsiveness to photoperiod and rainfall cues in switchgrass.
@@
   The timing of plant seasonal development (phenology) has major impacts on fitness because of the steep fitness cost of plant-environment mismatches.
-We infer the mixture of ways genetic effects on phenological traits depend on plant environment, focusing on how effects covary with weather prior to the phenological event (GxWeather).
+We infer the mixture of ways that genetic effects on phenological traits depend on plant environment, focusing on how effects covary with weather prior to the phenological event (GxWeather).
 Most effects do not have GxWeather, but a minority do.
@@
 
-\firstpage[9]{2}
+\firstpage[8]{2}
 
@@
 However, the usefulness of specific environmental cues depends on both features of the environment, such as cue predictability and relevance, and the species' adaptive strategies @botero2015evolutionary.
-Species with wide natural distributions can have multiple distinct environmentally cued phenological responses: for example, populations of sunflower (Helianthus annuus) exhibit day-neutral, facultative short day, and facultative long-day flowering responses, which vary with their environments [@blackman2013interacting; @henry2014transitions].
+Species with wide natural distributions can have multiple distinct environmentally cued phenological responses: for example, populations of sunflower (*Helianthus annuus*) exhibit day-neutral, facultative short day, and facultative long-day flowering responses, which vary with their environments [@blackman2013interacting; @henry2014transitions].
 Distinct genetic responses in different environments are known as genotype by environment interactions, or GxE.
@@
 
-Switchgrass (*Panicum virgatum*) is considered a short-day plant with reproductive development strongly linked to day of the year @mitchell1997predicting.
+Previous research suggests that switchgrass phenological timings should have GxWeather and that these timings could differ by genetic subpopulation.
+Switchgrass is considered a short-day plant with reproductive development strongly linked to day of the year @mitchell1997predicting.
 However, as part of its wide environmental adaptation across the eastern half of North America, its photoperiodicity has been predicted to differ by plant latitude of origin [@parrish2005biology; @casler2004latitudinal].
@@
 
-Here, we estimate the ways genetic effects on phenological timings depend on weather prior to the phenological event.
+Here, we estimate GxWeather for two phenological timings in switchgrass by assigning patterns of genetic effects on phenology across gardens to many patterns of weather covariance at these gardens.
 To do this, we phenotype a diversity panel of hundreds of switchgrass genotypes from the Midwest and Gulf subpopulations for the timing of vegetative and reproductive development at eight common garden locations spanning 17 degrees of latitude.
 These gardens cover the majority of the latitudinal and climatic range of switchgrass and capture the most comprehensive picture to date of the environmental variation this species encounters.
-We define multiple ways phenological timings might covary with weather (\@tbl-covar) and additional ways phenological timings might vary by site (\@SI-info), then jointly re-estimate genetic effects on these timings at all eight sites using the set of these covariance types that significantly improved the modeled log-likelihood when included (\@SI-info) @urbut2019flexible.
+We define multiple ways phenological timings might covary with weather (\@tbl-covar) and additional ways phenological timings might vary by site (\@SI-info), then jointly re-estimate genetic effects on these timings at all eight sites using the set of these covariance matrices that significantly improved the modeled log-likelihood when included (\@SI-info) @urbut2019flexible.
 We use the Bayesian framework *mash* (multivariate adaptive shrinkage), developed by @urbut2019flexible, to refine effect size estimates from genome-wide association (GWAS) conducted on individuals at each of the eight sites.
@@
 
-![Caption for this figure... blah blah blah.](images/Figure_2_Covariance_matrices_and_posterior_weights.png){#fig-covar}
+![Example hypothesis-driven covariance matrices specified in mash and the posterior weights placed on all covariance matrices.
+(a) Left column: Five example hypothesized covariance matrices specified for the green-up date or flowering date phenotype; these matrices were created from environment-specific correlations across eight common gardens, and are described in Table XXX.
+Common gardens are arranged in latitudinal order within the matrices.
+Right column: Five example canonical covariance matrices.
+Canonical matrices (purple) have simple interpretations, such as equal effects across all common gardens, or effects specific to a single common garden.
+(b,d) Total posterior weight placed on each covariance matrix type specified for (b) green-up date and (d) flowering date mash models, within and between two genetic subpopulations.
+Hypothesized covariance matrices (green).
+Covariance matrices included in mash that had zero posterior weight in all three mash runs on the genetic subpopulations, such as the identity matrix, are not shown.
+(c,e) Total posterior weight placed on covariance matrices that were hypothesized or canonical, for the (c) green-up date phenotype and (e) flowering date phenotype.](images/Figure_2_Covariance_matrices_and_posterior_weights.png){#fig-covar}
 
@@
 Essentially making Table 2 into a new, simplified figure.
-](images/Figure_3_Effect_Types_by_Site.png){#fig-effects}
+Effect patterns for re-estimated SNP effects that are significant at pairs of sites by the local false sign rate test (lfsr), for (a) greenup date and (b) flowering date, within and between two genetic subpopulations.
+Common gardens are arranged in latitudinal order along the x- and y-axis.
+Effects in the first and fourth rows differ in sign at these pairs of gardens (p \< 0.05, lfsr).
+Effects in the second, third, fifth and sixth rows are identical in sign (p \< 0.05, lfsr).
+Effects in the second and fifth rows additionally differ in magnitude by a factor of \>0.4, while effects in the third and sixth row were not distinguishable by magnitude nor sign of the effect.](images/Figure_3_Effect_Types_by_Site.png){#fig-effects}
 

### 2024-02-21 manuscript/revision2/revision2.tex
feat: QTL analysis using phenotypes from new covariance matrices, incorporated into Figure 3

@@
 (phenology) has major impacts on fitness because of the steep fitness
-cost of plant-environment mismatches. We infer the mixture of ways
+cost of plant-environment mismatches. We infer the mixture of ways that
 genetic effects on phenological traits depend on plant environment,
@@
 highly divergent switchgrass (\emph{Panicum virgatum}) populations using
-repeated plantings of these populations at eight sites spanning the
-central United States. We distinguish GxWeather patterns covarying with
-interpretable weather-based cues from agnostic, site-based patterns.
-Most GxWeather effects belong to the latter category and do not covary
-with weather-based cues. However, 65\% of effects on the timing of
-vegetative growth in the Gulf population covary with daylength 14 days
-prior to green-up date. 33\% of effects on the timing of flowering in
-the Gulf population covary with cumulative rainfall in the seven days
-prior to flowering, while 22\% of effects on flowering timing in the
-Midwest population covary with day length change two days prior to
-flowering. An independent pseudo-F2 cross of Gulf and Midwest
-individuals at the same sites mapped 23 additive QTLs for flowering, all
-with significant associations and eleven with enrichment of highly
-significant associations in our model of GxWeather effects. We
-demonstrate that we can identify QTL with GxWeather and assign them to
+repeated plantings of cloned individuals from these populations at eight
+sites spanning the central United States. We distinguish GxWeather
+patterns covarying with interpretable weather-based cues from agnostic,
+site-based patterns. Most GxWeather effects belong to the latter
+category and do not covary with weather-based cues. However, 65\% of
+effects on the timing of vegetative growth in the Gulf population covary
+with daylength 14 days prior to green-up date. 33\% of effects on the
+timing of flowering in the Gulf population covary with cumulative
+rainfall in the seven days prior to flowering, while 22\% of effects on
+flowering timing in the Midwest population covary with day length change
+two days prior to flowering. An independent pseudo-F2 cross of Gulf and
+Midwest individuals at the same sites mapped 10 additive QTLs for
+flowering and three for vegetative growth timing, all with significant
+associations and three with enrichment of highly significant
+associations in our model of GxWeather effects. We demonstrate that we
+can identify genetic variation with GxWeather and assign these loci to
 specific weather-based cues or other patterns. Breeding for particular
@@
 \ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
-\firstpage[9]{2}
+\firstpage[8]{2}
 
@@
 environmentally cued phenological responses: for example, populations of
-sunflower (Helianthus annuus) exhibit day-neutral, facultative short
-day, and facultative long-day flowering responses, which vary with their
-environments (5, 6). Distinct genetic responses in different
+sunflower (\emph{Helianthus annuus}) exhibit day-neutral, facultative
+short day, and facultative long-day flowering responses, which vary with
+their environments (5, 6). Distinct genetic responses in different
 environments are known as genotype by environment interactions, or GxE.
@@
 
-Switchgrass (\emph{Panicum virgatum}) is considered a short-day plant
-with reproductive development strongly linked to day of the year (29).
+Previous research suggests that switchgrass phenological timings should
+have GxWeather and that these timings could differ by genetic
+subpopulation. Switchgrass is considered a short-day plant with
+reproductive development strongly linked to day of the year (29).
 However, as part of its wide environmental adaptation across the eastern
@@
 
-Here, we estimate the ways genetic effects on phenological timings
-depend on weather prior to the phenological event. To do this, we
+Here, we estimate GxWeather for two phenological timings in switchgrass
+by assigning patterns of genetic effects on phenology across gardens to
+many patterns of weather covariance at these gardens. To do this, we
 phenotype a diversity panel of hundreds of switchgrass genotypes from
@@
 (@SI-info), then jointly re-estimate genetic effects on these timings at
-all eight sites using the set of these covariance types that
+all eight sites using the set of these covariance matrices that
 significantly improved the modeled log-likelihood when included
@@
 
-\caption{\label{fig-covar}Caption for this figure\ldots{} blah blah
-blah.}
+\caption{\label{fig-covar}Example hypothesis-driven covariance matrices
+specified in mash and the posterior weights placed on all covariance
+matrices. (a) Left column: Five example hypothesized covariance matrices
+specified for the green-up date or flowering date phenotype; these
+matrices were created from environment-specific correlations across
+eight common gardens, and are described in Table XXX. Common gardens are
+arranged in latitudinal order within the matrices. Right column: Five
+example canonical covariance matrices. Canonical matrices (purple) have
+simple interpretations, such as equal effects across all common gardens,
+or effects specific to a single common garden. (b,d) Total posterior
+weight placed on each covariance matrix type specified for (b) green-up
+date and (d) flowering date mash models, within and between two genetic
+subpopulations. Hypothesized covariance matrices (green). Covariance
+matrices included in mash that had zero posterior weight in all three
+mash runs on the genetic subpopulations, such as the identity matrix,
+are not shown. (c,e) Total posterior weight placed on covariance
+matrices that were hypothesized or canonical, for the (c) green-up date
+phenotype and (e) flowering date phenotype.}
 
@@
 showing total effects in Texas, the North, and maybe Oklahoma.
-Essentially making Table 2 into a new, simplified figure.}
+Essentially making Table 2 into a new, simplified figure. Effect
+patterns for re-estimated SNP effects that are significant at pairs of
+sites by the local false sign rate test (lfsr), for (a) greenup date and
+(b) flowering date, within and between two genetic subpopulations.
+Common gardens are arranged in latitudinal order along the x- and
+y-axis. Effects in the first and fourth rows differ in sign at these
+pairs of gardens (p \textless{} 0.05, lfsr). Effects in the second,
+third, fifth and sixth rows are identical in sign (p \textless{} 0.05,
+lfsr). Effects in the second and fifth rows additionally differ in
+magnitude by a factor of \textgreater0.4, while effects in the third and
+sixth row were not distinguishable by magnitude nor sign of the effect.}
 

### 2024-02-21 manuscript/revision2/revision2.qmd
manuscript: M&M edits and some QTL incorporation

@@
 We next looked for evidence of GxWeather by jointly re-estimating genetic effects across all eight common gardens using the mash model, which allows effects to covary between gardens in many ways (@fig-map D, SI Appendix, Datasets 1-6).
-To do this, we used GWAS on site-specific BLUPs for the timing of vegetative growth and flowering to determine preliminary single nucleotide polymorphism (SNP) effect estimates for both traits at each garden.
-We then jointly modeled genetic effect estimates, and the mixture of ways these effects might covary, across all eight common gardens, using a two-part procedure.
-We first modeled genetic effects with different combinations of covariance structures, using a subset of randomly selected SNPs used in the GWAS and their corresponding effect size estimates.
-In these models (hereafter referred to as "random"), we then specified a variety of potential covariance structures among genetic effect size estimates from each garden (as expanded on below) and used a greedy algorithm to iteratively select and add covariance structures that significantly improved the log likelihood of the *mash* model.
-Given the high correlation between some of the covariance patterns we tested, this approach allowed us to select only the subset of patterns of covariance that significantly improved the fit of the jointly inferred structure of genetic effect estimates present in the data.
-We then selected the set of SNPs and corresponding effect size estimates that had the largest effect estimates in univariate GWAS (hereafter "strong" effects) and jointly re-estimated these effect size estimates at all eight sites using the set of significant covariance matrices that most improved the model fit on the random effects.
-
 We specified three categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event (Table 1, SI Appendix, Section S2).
-For example, GxWeather covariance between gardens in MI and MO, for rainfall in the seven days prior to flowering, captures the covariance in the amount of rainfall genotypes A, B, and C receive in the seven days prior to their three flowering dates in MI and their three flowering dates in MO.
-If these six rainfall values are similar, covariance will be near one; if they are uncorrelated, covariance will be near zero; if there is sufficient rank changing in rainfall amounts between gardens for these genotypes, this covariance will be near negative one.
+For example, the more similar that the amount of rainfall prior to flowering is for genotypes A and B between gardens in MI and MO, the nearer these covariances will be to one.
+If rainfalls are uncorrelated, covariance will be near zero; if there is rank changing (e.g.
+high rainfall before flowering of A at MO but not MI, and high rainfall before flowering for B at MI but not MO), this covariance will be near negative one.
 We specified multiple covariance structures for the canonical, data-driven, and GxWeather categories, each of which represented a specific pattern of covariance in genetic effect size estimates that could be included in the model.
@@
 Say that a SNP in FLC controls flowering in a photoperiod-dependent manner.
-In that case, the joint estimate of effects for that SNP could have a high mixture proportion, or mass, on a covariance matrix created using a photoperiod-based environmental cue, such as day length two days prior to flowering.
+In that case, the joint estimate of effects for that SNP could have a high mixture proportion, or mass, on a covariance matrix created using a photoperiod-based environmental cue, such as day length at some interval prior to flowering.
 In our data, we would infer that the effect of that SNP on flowering was caused by a response to the environmental cue used to construct the GxWeather covariance structure with the largest mass.
-Overall, loadings of genetic effects on these GxWeather matrices in the model, along with the posterior effect size estimates provide information on genome-wide patterns of SNP-environment interaction.
+Overall, loadings of genetic effects on these GxWeather matrices in the model, along with the posterior effect size estimates, provide information on genome-wide patterns of SNP-environment interaction.
 
 The phenotypic correlations for the timing of vegetative growth had moderate negative correlations between the Texas and North gardens, particularly in the Gulf subpopulation and Both subpopulations (@fig-map B).
-If these phenotypic correlations have a genetic component, then they may be controlled by SNP effects that differed in sign across these regions (effects with antagonistic pleiotropy), SNP effects that are large in one garden or region and non-significant in others (effects with amplification), or a combination of these patterns.
+If these phenotypic correlations have a genetic component, then they may be controlled by SNP effects that differ in sign across these regions (effects with antagonistic pleiotropy), SNP effects that are large in one garden or region and non-significant in others (effects with amplification), or a combination of these patterns.
 Five GxWeather covariance structures were selected by the greedy algorithm, one to two per subpopulation.
@@
 This means that our tests for antagonistic pleiotropy, or a sign change between conditions, carry an equal statistical burden to those for effects with the same sign.
-We also summarized the pairwise patterns of effect sign and magnitude within and between the Texas and North regions, within the natural ranges of the Gulf and Midwest subpopulations (\@fig-effects).
+We also summarized the pairwise patterns of effect sign and magnitude within and between the Texas and North regions, within the natural ranges of the Gulf and Midwest subpopulations (@fig-effects).
 
-For greenup date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, between pairs of Texas and North gardens (\@fig-effects A).
-78.7% of pairwise comparisons between North and Texas gardens had a difference in sign, while only 28.6% and 0.2% of North-North or Texas-Texas comparisons had a difference in sign, respectively (\@fig-effects).
-The majority of pairwise effects for greenup for the Midwest (\>55%) and Both (\>85%) subpopulations were the same sign, and effects frequently differed in magnitude between the MO and OK garden and other gardens (\@fig-effects X).
+For greenup date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, between pairs of Texas and North gardens (@fig-effects A).
+78.7% of pairwise comparisons between North and Texas gardens had a difference in sign, while only 28.6% and 0.2% of North-North or Texas-Texas comparisons had a difference in sign, respectively (@fig-effects).
+The majority of pairwise effects for greenup for the Midwest (\>55%) and Both (\>85%) subpopulations were the same sign, and effects frequently differed in magnitude between the MO and OK garden and other gardens (@fig-effects X).
 
-For flowering date for the Gulf subpopulation, less than 2% of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, within or between regions (\@fig-effects).
-More effects differed in magnitude between the Texas and North regions than within these regions (42.7% vs \<20%; \@fig-effects X).
+For flowering date for the Gulf subpopulation, less than 2% of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, within or between regions (@fig-effects).
+More effects differed in magnitude between the Texas and North regions than within these regions (42.7% vs \<20%; @fig-effects X).
 The Midwest population had relatively few significant effects for flowering, but a large proportion of these differed in sign between Texas and North regions (42.7%) or within the North region (65.4%).
-Finally, in Both subpopulations, less than 20% of pairwise effects differed in sign (\@fig-effects).
-Most differences in sign were between TX1, the southernmost garden, and all other gardens (\@fig-effects B).
+Finally, in Both subpopulations, less than 20% of pairwise effects differed in sign (@fig-effects).
+Most differences in sign were between TX1, the southernmost garden, and all other gardens (@fig-effects B).
 Similarly, more effects that differed in magnitude included gardens in the Texas region (52.3-55.9%), and most effect pairs in the North region were not distinguishable (91.5%).
@@
 
-We sought additional experimental support for our joint re-estimates of SNP effects using an independent pseudo-F2 mapping population created from Gulf & Midwest individuals and grown at the same sites (\@fig-qtl A,B).
-We conducted quantitative trait loci (QTL) mapping of flowering as functions of four environmental cues that we also used as covariance matrices in mash, and identified eight QTL for flowering date, six QTL for flowering GDD, ten QTL for flowering day length, and eight QTL flowering day length change, all of which showed QTL by environment interactions (SI Appendix, Fig.
+We sought additional experimental support for our joint re-estimates of SNP effects using an independent pseudo-F2 mapping population created from Gulf & Midwest individuals and grown at the same sites (@fig-qtl A,B).
+We conducted quantitative trait loci (QTL) mapping of flowering as functions of five environmental cues that we also used as covariance matrices in mash, and identified eight QTL for flowering date, eight QTL for flowering as a function of day length change two days prior, one QTL for the start of vegetative growth, and two QTL for vegetative growth as a function of daylength change one day prior, all of which showed QTL by environment interactions (SI Appendix, Fig.
 S3).
@@
 thaliana* with functionally validated roles in flowering (SI Appendix, Dataset 7).
-All flowering QTL intervals contained at least one SNP significant in at least one mash run at a log10-transformed Bayes Factor \> 2, or in the 1% tail of significance, whichever was stricter (SI Appendix, Dataset 8).
+All flowering and green-up QTL intervals contained at least one SNP significant in at least one mash run at a log10-transformed Bayes Factor \> 2, or in the 1% tail of significance, whichever was stricter (SI Appendix, Dataset 8).
 We also looked for enrichments of mash SNPs in the 1% tail of significance (the 'mash 1% tail') within each QTL interval.
-At the 5% level, five QTL had enrichments of SNPs in the mash 1% tail.
-Overall, there were eleven significant enrichments (p \< 0.05, hypergeometric test) of SNPs in the mash 1% tail in the QTL intervals.
-Our QTL intervals had more enrichments of SNPs in the mash 1% tail than were found for all but eleven of these sets of random genomic intervals (\@fig-qtl C, p = 0.011).
-Thus, we were able to experimentally support our joint re-estimates of SNP effects with a QTL mapping experiment using a separate mapping population.
+At the 5% level, three QTL had enrichments of SNPs in the mash 1% tail.
+Overall, there were five significant enrichments (p \< 0.05, hypergeometric test) of SNPs in the mash 1% tail in the QTL intervals; however, this number of enrichments was found in many random genomic intervals also (@fig-qtl C).
+Thus, we were able to experimentally support some of our re-estimates of SNP effects with a QTL mapping experiment using a separate mapping population.
 
@@
 Our analysis of the timing of vegetative growth in the Gulf and in Both subpopulations revealed substantial antagonistic pleiotropy in effects between the Texas and North gardens (\@fig-effects A).
-This result supports theoretical models that local adaptation should involve antagonistic pleiotropy at the level of individual loci [@kawecki2004conceptual; @felsenstein1976theoretical; @levene1953genetic; @hedrick1986genetic], and is the first experimental work using QTL mapping and GWAS across common gardens to find antagonistic pleiotropy to be common in small genomic regions [@wadgymar2017identifying; @savolainen2013ecological; @lowry2019qtl].
+This result supports theoretical models that local adaptation should involve antagonistic pleiotropy at the level of individual loci [@kawecki2004conceptual; @felsenstein1976theoretical; @levene1953genetic; @hedrick1986genetic], and is the first experimental work using GWAS across common gardens to find antagonistic pleiotropy to be common in small genomic regions [@wadgymar2017identifying; @savolainen2013ecological; @lowry2019qtl].
 
@@
 Instead, the Gulf subpopulation has variation in flowering that covaries with the rainfall that occurs in the week prior to flowering.
-The genomic regions affecting flowering that we re-estimated across all eight sites were also supported by QTL from an independent mapping population at these sites (\@fig-qtl C).
+Three genomic regions affecting flowering that we re-estimated across all eight sites were also supported by QTL from an independent mapping population at these sites (\@fig-qtl A).
 
@@
 The GxWeather photoperiod and cumulative rainfall cues we identify here are functions of the genotypes measured and capture only a minority of SNP effects on flowering.
-We know still less about the overwintering parameters that cause variation in the start of vegetative.
+We know still less about the overwintering parameters that cause variation in the start of vegetative growth.
 Though we could assign SNP effects to a GxWeather covariance structure in the Gulf subpopulation, we could not do this for the Midwest subpopulation.
@@
 
-![Comparisons of pairs of jointly re-estimated effects across eight gardens, for effects with lfsr \< 0.05.
-AHM is considering moving this figure to the supplement and substantially simplifying the figure in the main text, perhaps just showing total effects in Texas, the North, and maybe Oklahoma.
+![ AHM is considering moving this figure to the supplement and substantially simplifying the figure in the main text, perhaps just showing total effects in Texas, the North, and maybe Oklahoma.
+Comparisons of pairs of jointly re-estimated effects across eight gardens, for effects with lfsr \< 0.05.
 Essentially making Table 2 into a new, simplified figure.
@@
 To evaluate the prevalence and kinds of covariance patterns of SNP effects across our common gardens, we used multivariate adaptive shrinkage (*mash*) on SNP effect estimates from the diversity panel (35).
-Mash is a statistical method that allows estimation and comparison of many effects jointly across many different conditions; it improves on previous methods by allowing multiple, arbitrary correlations in effect sizes among conditions (SI Appendix, Section S5).
-To obtain SNP effect estimates, we first conducted univariate genome-wide association at each common garden for green-up and flowering date.
-We then analyzed SNP effects for the top 19K relatively unlinked (r^2^ \< 0.2) SNPs per condition using mash, as in @lovell2021genomic.
-Details on these models can be found in (SI Appendix, Section S6).
+To do this, we first conducted univarate genome-wide association (GWAS) on site-specific best linear unbiased predictors (BLUPs) for the timing of vegetative growth and flowering, using the switchgrassGWAS package and the methods in [@lovell2021].
+We used the effect estimates for single nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect estimates for both traits at each garden.
+We then jointly modeled genetic effect estimates, and the mixture of ways these effects might covary, across all eight common gardens, using a two-part procedure.
+We first modeled genetic effects with different combinations of covariance structures, using the effect size estimates from a subset of [19K]{.underline} relatively unlinked (r^2^ \< 0.2), randomly selected SNPs used in the GWAS.
+In these models (hereafter referred to as "random"), we then specified a variety of potential covariance structures among genetic effect size estimates from each garden (as expanded on below) and used a greedy algorithm to iteratively select and add covariance structures that significantly improved the log likelihood of the *mash* model.
+Given the high correlation between some of the covariance patterns we tested, this approach allowed us to select only the subset of patterns of covariance that significantly improved the fit of the jointly inferred structure of genetic effect estimates present in the data.
+We then selected 19K relatively unlinked (r^2^ \< 0.2) SNPs from each univariate GWAS that had the largest effect estimates in each GWAS (hereafter "strong" effects), which gave a set of [**48K - 80K**]{.underline} SNPs, and jointly re-estimated these effect size estimates at all eight sites using the set of significant covariance matrices that most improved the model fit on the random effects.
+Notably, the majority of these "strong" effects were not significant in univariate GWAS.
 

### 2024-02-21 manuscript/revision2/revision2.tex
manuscript: M&M edits and some QTL incorporation

@@
 \usepackage{xcolor}
+\ifLuaTeX
+  \usepackage{luacolor}
+  \usepackage[soul]{lua-ul}
+\else
+  \usepackage{soul}
+
+\fi
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 which allows effects to covary between gardens in many ways
-(Figure~\ref{fig-map} D, SI Appendix, Datasets 1-6). To do this, we used
-GWAS on site-specific BLUPs for the timing of vegetative growth and
-flowering to determine preliminary single nucleotide polymorphism (SNP)
-effect estimates for both traits at each garden. We then jointly modeled
-genetic effect estimates, and the mixture of ways these effects might
-covary, across all eight common gardens, using a two-part procedure. We
-first modeled genetic effects with different combinations of covariance
-structures, using a subset of randomly selected SNPs used in the GWAS
-and their corresponding effect size estimates. In these models
-(hereafter referred to as ``random''), we then specified a variety of
-potential covariance structures among genetic effect size estimates from
-each garden (as expanded on below) and used a greedy algorithm to
-iteratively select and add covariance structures that significantly
-improved the log likelihood of the \emph{mash} model. Given the high
-correlation between some of the covariance patterns we tested, this
-approach allowed us to select only the subset of patterns of covariance
-that significantly improved the fit of the jointly inferred structure of
-genetic effect estimates present in the data. We then selected the set
-of SNPs and corresponding effect size estimates that had the largest
-effect estimates in univariate GWAS (hereafter ``strong'' effects) and
-jointly re-estimated these effect size estimates at all eight sites
-using the set of significant covariance matrices that most improved the
-model fit on the random effects.
-
-We specified three categories of covariance structures across gardens:
-``canonical'' covariance, with simple patterns of effect size covariance
-introduced in the initial \emph{mash} manuscript; ``data-driven''
-covariances derived from common patterns of SNP effects observed in the
-data, and ``GxWeather'' covariance, estimated from the covariance of
-empirical weather patterns at each garden at specific times before the
-phenological event (Table 1, SI Appendix, Section S2). For example,
-GxWeather covariance between gardens in MI and MO, for rainfall in the
-seven days prior to flowering, captures the covariance in the amount of
-rainfall genotypes A, B, and C receive in the seven days prior to their
-three flowering dates in MI and their three flowering dates in MO. If
-these six rainfall values are similar, covariance will be near one; if
-they are uncorrelated, covariance will be near zero; if there is
-sufficient rank changing in rainfall amounts between gardens for these
-genotypes, this covariance will be near negative one. We specified
-multiple covariance structures for the canonical, data-driven, and
-GxWeather categories, each of which represented a specific pattern of
-covariance in genetic effect size estimates that could be included in
+(Figure~\ref{fig-map} D, SI Appendix, Datasets 1-6). We specified three
+categories of covariance structures across gardens: ``canonical''
+covariance, with simple patterns of effect size covariance introduced in
+the initial \emph{mash} manuscript; ``data-driven'' covariances derived
+from common patterns of SNP effects observed in the data, and
+``GxWeather'' covariance, estimated from the covariance of empirical
+weather patterns at each garden at specific times before the
+phenological event (Table 1, SI Appendix, Section S2). For example, the
+more similar that the amount of rainfall prior to flowering is for
+genotypes A and B between gardens in MI and MO, the nearer these
+covariances will be to one. If rainfalls are uncorrelated, covariance
+will be near zero; if there is rank changing (e.g.~high rainfall before
+flowering of A at MO but not MI, and high rainfall before flowering for
+B at MI but not MO), this covariance will be near negative one. We
+specified multiple covariance structures for the canonical, data-driven,
+and GxWeather categories, each of which represented a specific pattern
+of covariance in genetic effect size estimates that could be included in
 the model.
@@
 covariance matrix created using a photoperiod-based environmental cue,
-such as day length two days prior to flowering. In our data, we would
-infer that the effect of that SNP on flowering was caused by a response
-to the environmental cue used to construct the GxWeather covariance
-structure with the largest mass. Overall, loadings of genetic effects on
-these GxWeather matrices in the model, along with the posterior effect
-size estimates provide information on genome-wide patterns of
-SNP-environment interaction.
+such as day length at some interval prior to flowering. In our data, we
+would infer that the effect of that SNP on flowering was caused by a
+response to the environmental cue used to construct the GxWeather
+covariance structure with the largest mass. Overall, loadings of genetic
+effects on these GxWeather matrices in the model, along with the
+posterior effect size estimates, provide information on genome-wide
+patterns of SNP-environment interaction.
 
@@
 genetic component, then they may be controlled by SNP effects that
-differed in sign across these regions (effects with antagonistic
+differ in sign across these regions (effects with antagonistic
 pleiotropy), SNP effects that are large in one garden or region and
@@
 within and between the Texas and North regions, within the natural
-ranges of the Gulf and Midwest subpopulations (@fig-effects).
+ranges of the Gulf and Midwest subpopulations
+(Figure~\ref{fig-effects}).
 
@@
 pairwise effects exhibited antagonistic pleiotropy, or a difference in
-effect sign, between pairs of Texas and North gardens (@fig-effects A).
-78.7\% of pairwise comparisons between North and Texas gardens had a
-difference in sign, while only 28.6\% and 0.2\% of North-North or
-Texas-Texas comparisons had a difference in sign, respectively
-(@fig-effects). The majority of pairwise effects for greenup for the
-Midwest (\textgreater55\%) and Both (\textgreater85\%) subpopulations
-were the same sign, and effects frequently differed in magnitude between
-the MO and OK garden and other gardens (@fig-effects X).
+effect sign, between pairs of Texas and North gardens
+(Figure~\ref{fig-effects} A). 78.7\% of pairwise comparisons between
+North and Texas gardens had a difference in sign, while only 28.6\% and
+0.2\% of North-North or Texas-Texas comparisons had a difference in
+sign, respectively (Figure~\ref{fig-effects}). The majority of pairwise
+effects for greenup for the Midwest (\textgreater55\%) and Both
+(\textgreater85\%) subpopulations were the same sign, and effects
+frequently differed in magnitude between the MO and OK garden and other
+gardens (Figure~\ref{fig-effects} X).
 
@@
 effects exhibited antagonistic pleiotropy, or a difference in effect
-sign, within or between regions (@fig-effects). More effects differed in
-magnitude between the Texas and North regions than within these regions
-(42.7\% vs \textless20\%; @fig-effects X). The Midwest population had
-relatively few significant effects for flowering, but a large proportion
-of these differed in sign between Texas and North regions (42.7\%) or
-within the North region (65.4\%). Finally, in Both subpopulations, less
-than 20\% of pairwise effects differed in sign (@fig-effects). Most
-differences in sign were between TX1, the southernmost garden, and all
-other gardens (@fig-effects B). Similarly, more effects that differed in
-magnitude included gardens in the Texas region (52.3-55.9\%), and most
-effect pairs in the North region were not distinguishable (91.5\%).
+sign, within or between regions (Figure~\ref{fig-effects}). More effects
+differed in magnitude between the Texas and North regions than within
+these regions (42.7\% vs \textless20\%; Figure~\ref{fig-effects} X). The
+Midwest population had relatively few significant effects for flowering,
+but a large proportion of these differed in sign between Texas and North
+regions (42.7\%) or within the North region (65.4\%). Finally, in Both
+subpopulations, less than 20\% of pairwise effects differed in sign
+(Figure~\ref{fig-effects}). Most differences in sign were between TX1,
+the southernmost garden, and all other gardens (Figure~\ref{fig-effects}
+B). Similarly, more effects that differed in magnitude included gardens
+in the Texas region (52.3-55.9\%), and most effect pairs in the North
+region were not distinguishable (91.5\%).
 
@@
 SNP effects using an independent pseudo-F2 mapping population created
-from Gulf \& Midwest individuals and grown at the same sites (@fig-qtl
-A,B). We conducted quantitative trait loci (QTL) mapping of flowering as
-functions of four environmental cues that we also used as covariance
-matrices in mash, and identified eight QTL for flowering date, six QTL
-for flowering GDD, ten QTL for flowering day length, and eight QTL
-flowering day length change, all of which showed QTL by environment
-interactions (SI Appendix, Fig. S3). All QTL for flowering overlapped
-one or more homologs from rice or \emph{A. thaliana} with functionally
-validated roles in flowering (SI Appendix, Dataset 7). All flowering QTL
+from Gulf \& Midwest individuals and grown at the same sites
+(Figure~\ref{fig-qtl} A,B). We conducted quantitative trait loci (QTL)
+mapping of flowering as functions of five environmental cues that we
+also used as covariance matrices in mash, and identified eight QTL for
+flowering date, eight QTL for flowering as a function of day length
+change two days prior, one QTL for the start of vegetative growth, and
+two QTL for vegetative growth as a function of daylength change one day
+prior, all of which showed QTL by environment interactions (SI Appendix,
+Fig. S3). All QTL for flowering overlapped one or more homologs from
+rice or \emph{A. thaliana} with functionally validated roles in
+flowering (SI Appendix, Dataset 7). All flowering and green-up QTL
 intervals contained at least one SNP significant in at least one mash
@@
 significance (the `mash 1\% tail') within each QTL interval. At the 5\%
-level, five QTL had enrichments of SNPs in the mash 1\% tail. Overall,
-there were eleven significant enrichments (p \textless{} 0.05,
-hypergeometric test) of SNPs in the mash 1\% tail in the QTL intervals.
-Our QTL intervals had more enrichments of SNPs in the mash 1\% tail than
-were found for all but eleven of these sets of random genomic intervals
-(@fig-qtl C, p = 0.011). Thus, we were able to experimentally support
-our joint re-estimates of SNP effects with a QTL mapping experiment
-using a separate mapping population.
+level, three QTL had enrichments of SNPs in the mash 1\% tail. Overall,
+there were five significant enrichments (p \textless{} 0.05,
+hypergeometric test) of SNPs in the mash 1\% tail in the QTL intervals;
+however, this number of enrichments was found in many random genomic
+intervals also (Figure~\ref{fig-qtl} C). Thus, we were able to
+experimentally support some of our re-estimates of SNP effects with a
+QTL mapping experiment using a separate mapping population.
 
@@
 antagonistic pleiotropy at the level of individual loci (22--25), and is
-the first experimental work using QTL mapping and GWAS across common
-gardens to find antagonistic pleiotropy to be common in small genomic
-regions (12, 37, 38).
+the first experimental work using GWAS across common gardens to find
+antagonistic pleiotropy to be common in small genomic regions (12, 37,
+38).
 
@@
 subpopulation has variation in flowering that covaries with the rainfall
-that occurs in the week prior to flowering. The genomic regions
+that occurs in the week prior to flowering. Three genomic regions
 affecting flowering that we re-estimated across all eight sites were
 also supported by QTL from an independent mapping population at these
-sites (@fig-qtl C).
+sites (@fig-qtl A).
 
@@
 effects on flowering. We know still less about the overwintering
-parameters that cause variation in the start of vegetative. Though we
-could assign SNP effects to a GxWeather covariance structure in the Gulf
-subpopulation, we could not do this for the Midwest subpopulation. More
-generally, it is difficult to predict the time scales over which
-individuals may integrate environmental cues, particularly in perennial
-species which may integrate these cues over longer time scales. If this
-integration time itself varies between individuals, we cannot select a
-covariance structure that reflects this, though these structures would
-likely be highly correlated with GxWeather structures we did include.
-Mash offers an opportunity to specify multiple environmental cues and
-compete them to explain patterns of genetic effects, allowing us to
-detect how important these cues are genome-wide, and how strongly each
-cue influences each SNP. This is a key development to further improve
-our understanding of genetic variation in GxE.
+parameters that cause variation in the start of vegetative growth.
+Though we could assign SNP effects to a GxWeather covariance structure
+in the Gulf subpopulation, we could not do this for the Midwest
+subpopulation. More generally, it is difficult to predict the time
+scales over which individuals may integrate environmental cues,
+particularly in perennial species which may integrate these cues over
+longer time scales. If this integration time itself varies between
+individuals, we cannot select a covariance structure that reflects this,
+though these structures would likely be highly correlated with GxWeather
+structures we did include. Mash offers an opportunity to specify
+multiple environmental cues and compete them to explain patterns of
+genetic effects, allowing us to detect how important these cues are
+genome-wide, and how strongly each cue influences each SNP. This is a
+key development to further improve our understanding of genetic
+variation in GxE.
 
@@
 
-\caption{\label{fig-effects}Comparisons of pairs of jointly re-estimated
-effects across eight gardens, for effects with lfsr \textless{} 0.05.
-AHM is considering moving this figure to the supplement and
-substantially simplifying the figure in the main text, perhaps just
-showing total effects in Texas, the North, and maybe Oklahoma.
-Essentially making Table 2 into a new, simplified figure. Effect
-patterns for re-estimated SNP effects that are significant at pairs of
-sites by the local false sign rate test (lfsr), for (a) greenup date and
-(b) flowering date, within and between two genetic subpopulations.
-Common gardens are arranged in latitudinal order along the x- and
-y-axis. Effects in the first and fourth rows differ in sign at these
-pairs of gardens (p \textless{} 0.05, lfsr). Effects in the second,
-third, fifth and sixth rows are identical in sign (p \textless{} 0.05,
-lfsr). Effects in the second and fifth rows additionally differ in
-magnitude by a factor of \textgreater0.4, while effects in the third and
-sixth row were not distinguishable by magnitude nor sign of the effect.}
+\caption{\label{fig-effects}AHM is considering moving this figure to the
+supplement and substantially simplifying the figure in the main text,
+perhaps just showing total effects in Texas, the North, and maybe
+Oklahoma. Comparisons of pairs of jointly re-estimated effects across
+eight gardens, for effects with lfsr \textless{} 0.05. Essentially
+making Table 2 into a new, simplified figure. Effect patterns for
+re-estimated SNP effects that are significant at pairs of sites by the
+local false sign rate test (lfsr), for (a) greenup date and (b)
+flowering date, within and between two genetic subpopulations. Common
+gardens are arranged in latitudinal order along the x- and y-axis.
+Effects in the first and fourth rows differ in sign at these pairs of
+gardens (p \textless{} 0.05, lfsr). Effects in the second, third, fifth
+and sixth rows are identical in sign (p \textless{} 0.05, lfsr). Effects
+in the second and fifth rows additionally differ in magnitude by a
+factor of \textgreater0.4, while effects in the third and sixth row were
+not distinguishable by magnitude nor sign of the effect.}
 
@@
 shrinkage (\emph{mash}) on SNP effect estimates from the diversity panel
-(35). Mash is a statistical method that allows estimation and comparison
-of many effects jointly across many different conditions; it improves on
-previous methods by allowing multiple, arbitrary correlations in effect
-sizes among conditions (SI Appendix, Section S5). To obtain SNP effect
-estimates, we first conducted univariate genome-wide association at each
-common garden for green-up and flowering date. We then analyzed SNP
-effects for the top 19K relatively unlinked (r\textsuperscript{2}
-\textless{} 0.2) SNPs per condition using mash, as in (32). Details on
-these models can be found in (SI Appendix, Section S6).
+(35). To do this, we first conducted univarate genome-wide association
+(GWAS) on site-specific best linear unbiased predictors (BLUPs) for the
+timing of vegetative growth and flowering, using the switchgrassGWAS
+package and the methods in (39). We used the effect estimates for single
+nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect
+estimates for both traits at each garden. We then jointly modeled
+genetic effect estimates, and the mixture of ways these effects might
+covary, across all eight common gardens, using a two-part procedure. We
+first modeled genetic effects with different combinations of covariance
+structures, using the effect size estimates from a subset of \ul{19K}
+relatively unlinked (r\textsuperscript{2} \textless{} 0.2), randomly
+selected SNPs used in the GWAS. In these models (hereafter referred to
+as ``random''), we then specified a variety of potential covariance
+structures among genetic effect size estimates from each garden (as
+expanded on below) and used a greedy algorithm to iteratively select and
+add covariance structures that significantly improved the log likelihood
+of the \emph{mash} model. Given the high correlation between some of the
+covariance patterns we tested, this approach allowed us to select only
+the subset of patterns of covariance that significantly improved the fit
+of the jointly inferred structure of genetic effect estimates present in
+the data. We then selected 19K relatively unlinked (r\textsuperscript{2}
+\textless{} 0.2) SNPs from each univariate GWAS that had the largest
+effect estimates in each GWAS (hereafter ``strong'' effects), which gave
+a set of \ul{\textbf{48K - 80K}} SNPs, and jointly re-estimated these
+effect size estimates at all eight sites using the set of significant
+covariance matrices that most improved the model fit on the random
+effects. Notably, the majority of these ``strong'' effects were not
+significant in univariate GWAS.
 
@@
 
+\bibitem[\citeproctext]{ref-lovell2021}
+\CSLLeftMargin{39. }%
+\CSLRightInline{J. T. Lovell, \emph{et al.},
+\href{https://doi.org/10.1038/s41586-020-03127-1}{Genomic mechanisms of
+climate adaptation in polyploid bioenergy switchgrass}. \emph{Nature},
+1--7 (2021).}
+
 \end{CSLReferences}

### 2024-03-09 manuscript/revision2/revision2.qmd
manuscript: New Fig 3 and Fig 4 and revised text to reflect figure changes

@@
 33% of effects on the timing of flowering in the Gulf population covary with cumulative rainfall in the seven days prior to flowering, while 22% of effects on flowering timing in the Midwest population covary with day length change two days prior to flowering.
-An independent pseudo-F2 cross of Gulf and Midwest individuals at the same sites mapped 10 additive QTLs for flowering and three for vegetative growth timing, all with significant associations and three with enrichment of highly significant associations in our model of GxWeather effects.
+An independent pseudo-F2 cross of Gulf and Midwest individuals at the same sites mapped eight additive QTLs for flowering and three for vegetative growth timing, all with significant associations and three with enrichment of highly significant associations in our model of GxWeather effects.
 We demonstrate that we can identify genetic variation with GxWeather and assign these loci to specific weather-based cues or other patterns.
@@
 We specified multiple covariance structures for the canonical, data-driven, and GxWeather categories, each of which represented a specific pattern of covariance in genetic effect size estimates that could be included in the model.
+The phenotypic correlations for the timing of vegetative growth had moderate negative correlations between the Texas and North gardens, particularly in the Gulf subpopulation and Both subpopulations (@fig-map B).
+If these phenotypic correlations have a genetic component, they could be partially or completely captured by the covariance structures specified in the mash model.
 
@@
 
-The phenotypic correlations for the timing of vegetative growth had moderate negative correlations between the Texas and North gardens, particularly in the Gulf subpopulation and Both subpopulations (@fig-map B).
-If these phenotypic correlations have a genetic component, then they may be controlled by SNP effects that differ in sign across these regions (effects with antagonistic pleiotropy), SNP effects that are large in one garden or region and non-significant in others (effects with amplification), or a combination of these patterns.
-Five GxWeather covariance structures were selected by the greedy algorithm, one to two per subpopulation.
+Five GxWeather covariance structures were selected by the greedy algorithm, zero to two per subpopulation: two for the start of vegetative growth, and three for flowering date.
 Of these five, three had mass on them in mash models of the strong effects (@fig-covar A-B).
@@
 We next characterized the pairwise patterns of effects where we were confident in the sign of the effect at both gardens.
+If the overall pattern of phenotypic expression differs between a pair of gardens, this difference may be comprised of SNP effects that differ in sign between these gardens (effects with antagonistic pleiotropy), SNP effects that differ in magnitude - that are large in one garden or region and smaller in others, and SNP effects that are indistinguishable in the two gardens (@fig-effects A).
 We used the local false sign rate (lfsr), an analogue of the lfdr that establishes confidence in the effect sign, not the effect's difference from zero, to determine significance.
@@
 This means that our tests for antagonistic pleiotropy, or a sign change between conditions, carry an equal statistical burden to those for effects with the same sign.
-We also summarized the pairwise patterns of effect sign and magnitude within and between the Texas and North regions, within the natural ranges of the Gulf and Midwest subpopulations (@fig-effects).
 
-For greenup date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, between pairs of Texas and North gardens (@fig-effects A).
+For greenup date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, between pairs of Texas and North gardens (@fig-effects A; @SI-fig-gardens).
 78.7% of pairwise comparisons between North and Texas gardens had a difference in sign, while only 28.6% and 0.2% of North-North or Texas-Texas comparisons had a difference in sign, respectively (@fig-effects).
-The majority of pairwise effects for greenup for the Midwest (\>55%) and Both (\>85%) subpopulations were the same sign, and effects frequently differed in magnitude between the MO and OK garden and other gardens (@fig-effects X).
+The majority of pairwise effects for greenup for the Midwest (\>55%) and Both (\>85%) subpopulations were the same sign, and effects most often differed in magnitude between gardens within and between the regions (@fig-effects X).
 
 For flowering date for the Gulf subpopulation, less than 2% of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, within or between regions (@fig-effects).
-More effects differed in magnitude between the Texas and North regions than within these regions (42.7% vs \<20%; @fig-effects X).
+More effects differed in magnitude between the Texas and North regions than within these regions (42.7% vs \<20%; @fig-effects X), while the majority of effects had no GxE.
 The Midwest population had relatively few significant effects for flowering, but a large proportion of these differed in sign between Texas and North regions (42.7%) or within the North region (65.4%).
 Finally, in Both subpopulations, less than 20% of pairwise effects differed in sign (@fig-effects).
-Most differences in sign were between TX1, the southernmost garden, and all other gardens (@fig-effects B).
+Most differences in sign were between TX1, the southernmost garden, and all other gardens (@SI-fig-gardens).
 Similarly, more effects that differed in magnitude included gardens in the Texas region (52.3-55.9%), and most effect pairs in the North region were not distinguishable (91.5%).
@@
 At the 5% level, three QTL had enrichments of SNPs in the mash 1% tail.
-Overall, there were five significant enrichments (p \< 0.05, hypergeometric test) of SNPs in the mash 1% tail in the QTL intervals; however, this number of enrichments was found in many random genomic intervals also (@fig-qtl C).
+Overall, there were five significant enrichments (p \< 0.05, hypergeometric test) of SNPs in the mash 1% tail in the QTL intervals.
 Thus, we were able to experimentally support some of our re-estimates of SNP effects with a QTL mapping experiment using a separate mapping population.
@@
 
-Our analysis of the timing of vegetative growth in the Gulf and in Both subpopulations revealed substantial antagonistic pleiotropy in effects between the Texas and North gardens (\@fig-effects A).
+Our analysis of the timing of vegetative growth in the Gulf and in Both subpopulations revealed substantial antagonistic pleiotropy in effects between the Texas and North gardens (@fig-effects B).
 This result supports theoretical models that local adaptation should involve antagonistic pleiotropy at the level of individual loci [@kawecki2004conceptual; @felsenstein1976theoretical; @levene1953genetic; @hedrick1986genetic], and is the first experimental work using GWAS across common gardens to find antagonistic pleiotropy to be common in small genomic regions [@wadgymar2017identifying; @savolainen2013ecological; @lowry2019qtl].
@@
 Instead, the Gulf subpopulation has variation in flowering that covaries with the rainfall that occurs in the week prior to flowering.
-Three genomic regions affecting flowering that we re-estimated across all eight sites were also supported by QTL from an independent mapping population at these sites (\@fig-qtl A).
+Three genomic regions affecting flowering that we re-estimated across all eight sites were also supported by QTL from an independent mapping population at these sites (@fig-qtl B).
 
@@
 We know still less about the overwintering parameters that cause variation in the start of vegetative growth.
-Though we could assign SNP effects to a GxWeather covariance structure in the Gulf subpopulation, we could not do this for the Midwest subpopulation.
+We could only assign SNP effects to a GxWeather covariance structures in four of the six phenotype & genetic subpopulations we modeled.
 More generally, it is difficult to predict the time scales over which individuals may integrate environmental cues, particularly in perennial species which may integrate these cues over longer time scales.
@@
 
-![ AHM is considering moving this figure to the supplement and substantially simplifying the figure in the main text, perhaps just showing total effects in Texas, the North, and maybe Oklahoma.
-Comparisons of pairs of jointly re-estimated effects across eight gardens, for effects with lfsr \< 0.05.
-Essentially making Table 2 into a new, simplified figure.
-Effect patterns for re-estimated SNP effects that are significant at pairs of sites by the local false sign rate test (lfsr), for (a) greenup date and (b) flowering date, within and between two genetic subpopulations.
-Common gardens are arranged in latitudinal order along the x- and y-axis.
-Effects in the first and fourth rows differ in sign at these pairs of gardens (p \< 0.05, lfsr).
-Effects in the second, third, fifth and sixth rows are identical in sign (p \< 0.05, lfsr).
-Effects in the second and fifth rows additionally differ in magnitude by a factor of \>0.4, while effects in the third and sixth row were not distinguishable by magnitude nor sign of the effect.](images/Figure_3_Effect_Types_by_Site.png){#fig-effects}
+![Types of GxE present between of pairs of jointly re-estimated effects in eight common gardens, for effects with lfsr \< 0.05.
+a) Examples of effect patterns at three pairs of sites with three types of GxE.
+Sign: Effects that differ in sign at these pairs of gardens (p \< 0.05, lfsr).
+Magnitude: Effects identical in sign (p \< 0.05, lfsr) that differ in magnitude by a factor of \>0.4.
+None: Effects not distinguishable by magnitude nor sign of the effect.
+b) The fraction of effects with each GxE type for the start of vegetative growth (Greenup Date) and reproductive growth (Flowering Date), within and between two genetic subpopulations.
+Common gardens are grouped by the larger region they came from: North gardens are within the natural range of the Midwest subpopulation, while Texas gardens are within the natural range of the Gulf subpopulation.
+](Figure_3_GxE.png){#fig-effects}
 
-![SNP effects for SNPs within QTL.
-AHM to re-make this figure now LZ has re-run the QTL analysis, the assmption being the re-analysis will not change the figure or the text much.
-The bigger question for TEJ and AH is whether we want to keep this analysis in the paper, what does it add?](images/Figure_4_mash_SNP_effects.png){#fig-qtl}
+![Overlaps of QTL from an outbred pseudo-F2 cross and with jointly re-estimated SNP effects in the 1% tail of significance from a diversity panel.
+Dotted lines indicate permutation-based significance thresholds for each weather-related function.
+Stars indicate QTL with significant enrichment for SNPs in the 1% mash tail; G, M, and B indicate which subpopulation had enrichment: G - Gulf subpopulation, M -Midwest subpopulation, B - both subpopulations.
+Rug plots show genomic locations of SNPs in the 1% mash tail for flowering date for each subpopulation.
+a) QTL mapping for the start of vegetative growth (Greenup Date), and three weather-related functions of greenup date.
+b) QTL mapping for the start of reproductive growth (Flowering Date), and two weather-related functions of flowering date.
+](Figure_4_QTL_Overlaps.png){#fig-qtl}
 
@@
 Given the high correlation between some of the covariance patterns we tested, this approach allowed us to select only the subset of patterns of covariance that significantly improved the fit of the jointly inferred structure of genetic effect estimates present in the data.
-We then selected 19K relatively unlinked (r^2^ \< 0.2) SNPs from each univariate GWAS that had the largest effect estimates in each GWAS (hereafter "strong" effects), which gave a set of [**48K - 80K**]{.underline} SNPs, and jointly re-estimated these effect size estimates at all eight sites using the set of significant covariance matrices that most improved the model fit on the random effects.
+We then selected 19K relatively unlinked (r^2^ \< 0.2) SNPs from each univariate GWAS that had the largest effect estimates in each GWAS (hereafter "strong" effects), which gave sets of [**48K - 80K**]{.underline} SNPs, and jointly re-estimated these effect size estimates at all eight sites using the set of significant covariance matrices that most improved the model fit on the random effects.
 Notably, the majority of these "strong" effects were not significant in univariate GWAS.
@@
 
-In addition, the get_GxE function also sets an arbitrary threshold to count an effect as changing in magnitude between environments, commonly known as differential sensitivity or a change in amplitude of .
+In addition, the get_GxE function also sets an arbitrary threshold to count an effect as changing in magnitude between environments, commonly known as differential sensitivity or a change in amplitude of the effect.
 For differential sensitivity, this function determines if effects significant in both conditions are of the same sign and of a magnitude (not tested for significance) that differs by a factor of 0.4 or more.
 The remaining effects that are significant in both conditions have the same effect sign and similar effect magnitudes and we denote these effects as having no GxE.
-The distinction between effects with different magnitudes is arbitrary but useful to fully characterize how effects vary across environments.
-and our use of the lfsr to determine significance and our specification that SNP effects must be significant in both conditions to be included means that our tests for antagonistic pleiotropy carry an equal statistical burden to those measuring differential sensitivity and effects without GxE.
+The distinction between effects with different magnitudes is arbitrary but useful to fully characterize how effects vary across environments to ultimately influence phenotypes.
+Our use of the lfsr to determine significance and our specification that SNP effects must be significant in both conditions to be included means that our tests for antagonistic pleiotropy carry an equal statistical burden to those measuring differential sensitivity and effects without GxE.
 
@@
 To confirm candidate genomic regions and patterns of allelic effects found in the diversity panel, we analyzed flowering in an outbred pseudo-F2 cross between four individuals, two Midwest and two Gulf individuals.
-The formation of this mapping population has been described previously (34); additional details on QTL mapping can be found in SI Appendix, Section S6.
+The formation of this mapping population has been described previously [@milano2016genetic]; additional details on QTL mapping can be found in SI Appendix, Section S6.
 To be directly comparable to the diversity panel data, only 2019 phenology data from the pseudo-F2 cross from the same eight common garden sites were used.
-To compare our QTL enrichments of significant mash associations to the null expectation, we used permutation to choose 1000 sets of 23 genomic regions of the same size randomly distributed throughout the genome, then calculated enrichments of the mash 1% tail in these random intervals.
 

### 2024-03-09 manuscript/revision2/revision2.tex
manuscript: New Fig 3 and Fig 4 and revised text to reflect figure changes

@@
 two days prior to flowering. An independent pseudo-F2 cross of Gulf and
-Midwest individuals at the same sites mapped 10 additive QTLs for
+Midwest individuals at the same sites mapped eight additive QTLs for
 flowering and three for vegetative growth timing, all with significant
@@
 of covariance in genetic effect size estimates that could be included in
-the model.
+the model. The phenotypic correlations for the timing of vegetative
+growth had moderate negative correlations between the Texas and North
+gardens, particularly in the Gulf subpopulation and Both subpopulations
+(Figure~\ref{fig-map} B). If these phenotypic correlations have a
+genetic component, they could be partially or completely captured by the
+covariance structures specified in the mash model.
 
@@
 
-The phenotypic correlations for the timing of vegetative growth had
-moderate negative correlations between the Texas and North gardens,
-particularly in the Gulf subpopulation and Both subpopulations
-(Figure~\ref{fig-map} B). If these phenotypic correlations have a
-genetic component, then they may be controlled by SNP effects that
-differ in sign across these regions (effects with antagonistic
-pleiotropy), SNP effects that are large in one garden or region and
-non-significant in others (effects with amplification), or a combination
-of these patterns. Five GxWeather covariance structures were selected by
-the greedy algorithm, one to two per subpopulation. Of these five, three
+Five GxWeather covariance structures were selected by the greedy
+algorithm, zero to two per subpopulation: two for the start of
+vegetative growth, and three for flowering date. Of these five, three
 had mass on them in mash models of the strong effects
@@
 We next characterized the pairwise patterns of effects where we were
-confident in the sign of the effect at both gardens. We used the local
-false sign rate (lfsr), an analogue of the lfdr that establishes
-confidence in the effect sign, not the effect's difference from zero, to
-determine significance. We required lfsr significance (p \textless{}
-0.05) in both gardens to include effects. This means that our tests for
-antagonistic pleiotropy, or a sign change between conditions, carry an
-equal statistical burden to those for effects with the same sign. We
-also summarized the pairwise patterns of effect sign and magnitude
-within and between the Texas and North regions, within the natural
-ranges of the Gulf and Midwest subpopulations
-(Figure~\ref{fig-effects}).
+confident in the sign of the effect at both gardens. If the overall
+pattern of phenotypic expression differs between a pair of gardens, this
+difference may be comprised of SNP effects that differ in sign between
+these gardens (effects with antagonistic pleiotropy), SNP effects that
+differ in magnitude - that are large in one garden or region and smaller
+in others, and SNP effects that are indistinguishable in the two gardens
+(Figure~\ref{fig-effects} A). We used the local false sign rate (lfsr),
+an analogue of the lfdr that establishes confidence in the effect sign,
+not the effect's difference from zero, to determine significance. We
+required lfsr significance (p \textless{} 0.05) in both gardens to
+include effects. This means that our tests for antagonistic pleiotropy,
+or a sign change between conditions, carry an equal statistical burden
+to those for effects with the same sign.
 
@@
 effect sign, between pairs of Texas and North gardens
-(Figure~\ref{fig-effects} A). 78.7\% of pairwise comparisons between
-North and Texas gardens had a difference in sign, while only 28.6\% and
-0.2\% of North-North or Texas-Texas comparisons had a difference in
-sign, respectively (Figure~\ref{fig-effects}). The majority of pairwise
-effects for greenup for the Midwest (\textgreater55\%) and Both
-(\textgreater85\%) subpopulations were the same sign, and effects
-frequently differed in magnitude between the MO and OK garden and other
-gardens (Figure~\ref{fig-effects} X).
+(Figure~\ref{fig-effects} A; (\textbf{SI-fig-gardens?})). 78.7\% of
+pairwise comparisons between North and Texas gardens had a difference in
+sign, while only 28.6\% and 0.2\% of North-North or Texas-Texas
+comparisons had a difference in sign, respectively
+(Figure~\ref{fig-effects}). The majority of pairwise effects for greenup
+for the Midwest (\textgreater55\%) and Both (\textgreater85\%)
+subpopulations were the same sign, and effects most often differed in
+magnitude between gardens within and between the regions
+(Figure~\ref{fig-effects} X).
 
@@
 differed in magnitude between the Texas and North regions than within
-these regions (42.7\% vs \textless20\%; Figure~\ref{fig-effects} X). The
-Midwest population had relatively few significant effects for flowering,
-but a large proportion of these differed in sign between Texas and North
-regions (42.7\%) or within the North region (65.4\%). Finally, in Both
-subpopulations, less than 20\% of pairwise effects differed in sign
+these regions (42.7\% vs \textless20\%; Figure~\ref{fig-effects} X),
+while the majority of effects had no GxE. The Midwest population had
+relatively few significant effects for flowering, but a large proportion
+of these differed in sign between Texas and North regions (42.7\%) or
+within the North region (65.4\%). Finally, in Both subpopulations, less
+than 20\% of pairwise effects differed in sign
 (Figure~\ref{fig-effects}). Most differences in sign were between TX1,
-the southernmost garden, and all other gardens (Figure~\ref{fig-effects}
-B). Similarly, more effects that differed in magnitude included gardens
-in the Texas region (52.3-55.9\%), and most effect pairs in the North
-region were not distinguishable (91.5\%).
+the southernmost garden, and all other gardens
+((\textbf{SI-fig-gardens?})). Similarly, more effects that differed in
+magnitude included gardens in the Texas region (52.3-55.9\%), and most
+effect pairs in the North region were not distinguishable (91.5\%).
 
@@
 there were five significant enrichments (p \textless{} 0.05,
-hypergeometric test) of SNPs in the mash 1\% tail in the QTL intervals;
-however, this number of enrichments was found in many random genomic
-intervals also (Figure~\ref{fig-qtl} C). Thus, we were able to
-experimentally support some of our re-estimates of SNP effects with a
-QTL mapping experiment using a separate mapping population.
+hypergeometric test) of SNPs in the mash 1\% tail in the QTL intervals.
+Thus, we were able to experimentally support some of our re-estimates of
+SNP effects with a QTL mapping experiment using a separate mapping
+population.
 
@@
 subpopulations revealed substantial antagonistic pleiotropy in effects
-between the Texas and North gardens (@fig-effects A). This result
-supports theoretical models that local adaptation should involve
+between the Texas and North gardens (Figure~\ref{fig-effects} B). This
+result supports theoretical models that local adaptation should involve
 antagonistic pleiotropy at the level of individual loci (22--25), and is
@@
 also supported by QTL from an independent mapping population at these
-sites (@fig-qtl A).
+sites (Figure~\ref{fig-qtl} B).
 
@@
 effects on flowering. We know still less about the overwintering
-parameters that cause variation in the start of vegetative growth.
-Though we could assign SNP effects to a GxWeather covariance structure
-in the Gulf subpopulation, we could not do this for the Midwest
-subpopulation. More generally, it is difficult to predict the time
-scales over which individuals may integrate environmental cues,
-particularly in perennial species which may integrate these cues over
-longer time scales. If this integration time itself varies between
-individuals, we cannot select a covariance structure that reflects this,
-though these structures would likely be highly correlated with GxWeather
-structures we did include. Mash offers an opportunity to specify
-multiple environmental cues and compete them to explain patterns of
-genetic effects, allowing us to detect how important these cues are
-genome-wide, and how strongly each cue influences each SNP. This is a
-key development to further improve our understanding of genetic
-variation in GxE.
+parameters that cause variation in the start of vegetative growth. We
+could only assign SNP effects to a GxWeather covariance structures in
+four of the six phenotype \& genetic subpopulations we modeled. More
+generally, it is difficult to predict the time scales over which
+individuals may integrate environmental cues, particularly in perennial
+species which may integrate these cues over longer time scales. If this
+integration time itself varies between individuals, we cannot select a
+covariance structure that reflects this, though these structures would
+likely be highly correlated with GxWeather structures we did include.
+Mash offers an opportunity to specify multiple environmental cues and
+compete them to explain patterns of genetic effects, allowing us to
+detect how important these cues are genome-wide, and how strongly each
+cue influences each SNP. This is a key development to further improve
+our understanding of genetic variation in GxE.
 
@@
 
-\includegraphics{images/Figure_3_Effect_Types_by_Site.png}
+\includegraphics{Figure_3_GxE.png}
 
@@
 
-\caption{\label{fig-effects}AHM is considering moving this figure to the
-supplement and substantially simplifying the figure in the main text,
-perhaps just showing total effects in Texas, the North, and maybe
-Oklahoma. Comparisons of pairs of jointly re-estimated effects across
-eight gardens, for effects with lfsr \textless{} 0.05. Essentially
-making Table 2 into a new, simplified figure. Effect patterns for
-re-estimated SNP effects that are significant at pairs of sites by the
-local false sign rate test (lfsr), for (a) greenup date and (b)
-flowering date, within and between two genetic subpopulations. Common
-gardens are arranged in latitudinal order along the x- and y-axis.
-Effects in the first and fourth rows differ in sign at these pairs of
-gardens (p \textless{} 0.05, lfsr). Effects in the second, third, fifth
-and sixth rows are identical in sign (p \textless{} 0.05, lfsr). Effects
-in the second and fifth rows additionally differ in magnitude by a
-factor of \textgreater0.4, while effects in the third and sixth row were
-not distinguishable by magnitude nor sign of the effect.}
+\caption{\label{fig-effects}Types of GxE present between of pairs of
+jointly re-estimated effects in eight common gardens, for effects with
+lfsr \textless{} 0.05. a) Examples of effect patterns at three pairs of
+sites with three types of GxE. Sign: Effects that differ in sign at
+these pairs of gardens (p \textless{} 0.05, lfsr). Magnitude: Effects
+identical in sign (p \textless{} 0.05, lfsr) that differ in magnitude by
+a factor of \textgreater0.4. None: Effects not distinguishable by
+magnitude nor sign of the effect. b) The fraction of effects with each
+GxE type for the start of vegetative growth (Greenup Date) and
+reproductive growth (Flowering Date), within and between two genetic
+subpopulations. Common gardens are grouped by the larger region they
+came from: North gardens are within the natural range of the Midwest
+subpopulation, while Texas gardens are within the natural range of the
+Gulf subpopulation.}
 
@@
 
-\includegraphics{images/Figure_4_mash_SNP_effects.png}
+\includegraphics{Figure_4_QTL_Overlaps.png}
 
@@
 
-\caption{\label{fig-qtl}SNP effects for SNPs within QTL. AHM to re-make
-this figure now LZ has re-run the QTL analysis, the assmption being the
-re-analysis will not change the figure or the text much. The bigger
-question for TEJ and AH is whether we want to keep this analysis in the
-paper, what does it add?}
+\caption{\label{fig-qtl}Overlaps of QTL from an outbred pseudo-F2 cross
+and with jointly re-estimated SNP effects in the 1\% tail of
+significance from a diversity panel. Dotted lines indicate
+permutation-based significance thresholds for each weather-related
+function. Stars indicate QTL with significant enrichment for SNPs in the
+1\% mash tail; G, M, and B indicate which subpopulation had enrichment:
+G - Gulf subpopulation, M -Midwest subpopulation, B - both
+subpopulations. Rug plots show genomic locations of SNPs in the 1\% mash
+tail for flowering date for each subpopulation. a) QTL mapping for the
+start of vegetative growth (Greenup Date), and three weather-related
+functions of greenup date. b) QTL mapping for the start of reproductive
+growth (Flowering Date), and two weather-related functions of flowering
+date.}
 
@@
 effect estimates in each GWAS (hereafter ``strong'' effects), which gave
-a set of \ul{\textbf{48K - 80K}} SNPs, and jointly re-estimated these
+sets of \ul{\textbf{48K - 80K}} SNPs, and jointly re-estimated these
 effect size estimates at all eight sites using the set of significant
@@
 count an effect as changing in magnitude between environments, commonly
-known as differential sensitivity or a change in amplitude of . For
-differential sensitivity, this function determines if effects
-significant in both conditions are of the same sign and of a magnitude
-(not tested for significance) that differs by a factor of 0.4 or more.
-The remaining effects that are significant in both conditions have the
-same effect sign and similar effect magnitudes and we denote these
-effects as having no GxE. The distinction between effects with different
-magnitudes is arbitrary but useful to fully characterize how effects
-vary across environments. and our use of the lfsr to determine
-significance and our specification that SNP effects must be significant
-in both conditions to be included means that our tests for antagonistic
-pleiotropy carry an equal statistical burden to those measuring
-differential sensitivity and effects without GxE.
+known as differential sensitivity or a change in amplitude of the
+effect. For differential sensitivity, this function determines if
+effects significant in both conditions are of the same sign and of a
+magnitude (not tested for significance) that differs by a factor of 0.4
+or more. The remaining effects that are significant in both conditions
+have the same effect sign and similar effect magnitudes and we denote
+these effects as having no GxE. The distinction between effects with
+different magnitudes is arbitrary but useful to fully characterize how
+effects vary across environments to ultimately influence phenotypes. Our
+use of the lfsr to determine significance and our specification that SNP
+effects must be significant in both conditions to be included means that
+our tests for antagonistic pleiotropy carry an equal statistical burden
+to those measuring differential sensitivity and effects without GxE.
 
@@
 data, only 2019 phenology data from the pseudo-F2 cross from the same
-eight common garden sites were used. To compare our QTL enrichments of
-significant mash associations to the null expectation, we used
-permutation to choose 1000 sets of 23 genomic regions of the same size
-randomly distributed throughout the genome, then calculated enrichments
-of the mash 1\% tail in these random intervals.
+eight common garden sites were used.
 

### 2024-03-09 R/Functions_ggplot-theme-adjustments_2018-01-03.R

@@ -1,0 +1,43 @@
+# ------------------------------------------------------
+## 
+## ---- ggplot theme ----
+## 
+# ------------------------------------------------------
+## This script should run fine if you have run the 
+## Functions_Source-for-CDBN-Analyses R script first.
+require(tidyverse)
+require(gridExtra)
+
+## Many thanks to Rob Heckman for his ggplot theme script!
+
+theme_oeco <- theme_classic() +
+  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 10), 
+        axis.line.x = element_line(size = 0.35, colour = 'grey50'), axis.line.y = element_line(size = 0.35, colour = 'grey50'),
+        axis.ticks = element_line(size = 0.25, colour = 'grey50'), 
+        legend.justification = c(1, 0.75), legend.position = c(1, 0.9), legend.key.size = unit(0.35, 'cm'),
+        legend.title = element_blank(), legend.text = element_text(size = 9),
+        legend.text.align = 0, legend.background = element_blank(),
+        plot.subtitle = element_text(size = 10, vjust = 0), #plot.margin = unit(c(0.35, 0, 0.25, 0), 'cm'),
+        strip.background = element_blank(), strip.text = element_text(hjust = 0.5, size = 10 ,vjust = 0), 
+        strip.placement = 'outside', panel.spacing.x = unit(-0.5, 'cm'))
+  theme_set(theme_oeco)
+  
+theme_poster <- theme_classic() +
+    theme(axis.title = element_text(size = 18), axis.text = element_text(size = 14), 
+          axis.line.x = element_line(size = 0.35, colour = 'grey50'), axis.line.y = element_line(size = 0.35, colour = 'grey50'),
+          axis.ticks = element_line(size = 0.5, colour = 'grey50'), 
+          legend.justification = c(1, 0.75), legend.position = c(1, 0.9), legend.key.size = unit(0.5, 'cm'),
+          legend.title = element_blank(), legend.text = element_text(size = 18),
+          legend.text.align = 0, legend.background = element_blank(),
+          plot.subtitle = element_text(size = 14, vjust = 0), #plot.margin = unit(c(0.35, 0, 0.25, 0), 'cm'),
+          strip.background = element_blank(), strip.text = element_text(hjust = 0.5, size = 10 ,vjust = 0), 
+          strip.placement = 'outside', panel.spacing.x = unit(-0.5, 'cm'))
[…]

### 2024-03-12 R/Analysis_v1.2_mash_using_greedy_covar.qmd
manuscript: AH edits from 2024-03-07

@@
 ```{r}
-source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
 mashdir <- here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar")
@@
 
+New theme for ggplot2 v 3.5
+
+```{r}
+theme_oeco <- theme_classic() +
+  theme(axis.title = element_text(size = 10), 
+        axis.text = element_text(size = 10), 
+        axis.line.x = element_line(linewidth = 0.35, colour = 'grey50'),
+        axis.line.y = element_line(linewidth = 0.35, colour = 'grey50'),
+        axis.ticks = element_line(linewidth = 0.25, colour = 'grey50'), 
+        legend.justification = c(1, 0.75), 
+        legend.position = c(1, 0.9), 
+        legend.key.size = unit(0.35, 'cm'),
+        legend.title = element_blank(), 
+        legend.text = element_text(size = 9),
+        legend.background = element_blank(),
+        plot.subtitle = element_text(size = 10, vjust = 0), #plot.margin = unit(c(0.35, 0, 0.25, 0), 'cm'),
+        strip.background = element_blank(), 
+        strip.text = element_text(hjust = 0.5, size = 10 ,vjust = 0), 
+        strip.placement = 'outside', panel.spacing.x = unit(-0.5, 'cm'))
+theme_set(theme_oeco)
+```
+
 ```{r}
@@
 cov_gg$ggobject + 
+  theme_oeco +
   ggtitle("Gulf greenup", subtitle = "64.5% SNP mass")
 cov_gf$ggobject + 
+  theme_oeco +
   ggtitle("Gulf flowering", subtitle = "33.0% SNP mass")
 cov_mg$ggobject + 
+  theme_oeco +
   ggtitle("Midwest greenup", subtitle = "0% SNP mass")
 cov_mf$ggobject + 
+  theme_oeco +
   ggtitle("Midwest flowering", subtitle = "22.6% SNP mass")
 cov_gamg$ggobject + 
+  theme_oeco +
   ggtitle("Gulf and Midwest greenup", subtitle = "32.7% SNP mass")
 cov_gamf$ggobject + 
+  theme_oeco +
   ggtitle("Gulf and Midwest flowering", subtitle = "0% SNP mass")
@@
 ```{r pick covar matrices to plot fig 2}
-plot_covar_Ulist_fig2 <- function(df, mf = FALSE, mg = FALSE) {
+plot_covar_Ulist_fig2 <- function(df, mf = FALSE, mg = FALSE, legend = FALSE) {
   if (mg) {
@@
     geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
-    scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
-                                  "white", "#27AD81FF", "#5DC863FF",
-                                  "#FDE725FF"),
+    scale_fill_gradientn(colors = c("#30123BFF", "#4777EFFF", "#1BD0D5FF",
+                                  "white", "#FE9B2DFF", "#DB3A07FF",
+                                  "#7A0403FF"),
                          limits = c(-1, 1)) +
+    theme_oeco 
+  if (legend == FALSE) {
+    plot_covar <- plot_covar +
     theme(legend.position = "none",
@@
     labs(x = "", y = "")
+  } else if (legend == TRUE) {
+      plot_covar <- plot_covar +
+    theme(legend.position = "right",
+          axis.text = element_blank(),
+          axis.ticks = element_blank(),
+          axis.line.x = element_blank(),
+          axis.line.y = element_blank(),
+          axis.title.y.right = element_text()) +
+    labs(x = "", y = "")
+    }
+
   return(plot_covar)
 }
+library(scales)
+show_col(viridis_pal(option = "H")(8))
 ```
@@
   geom_bar(aes(fill = Type), stat = "identity") +
-  scale_fill_viridis_d(end = 0.9) +
+  scale_fill_viridis_d(end = 0.9, option = "H") +
+  theme_oeco + 
   theme(axis.title.x=element_blank(),
@@
   facet_wrap(~subpop) +
-  coord_flip() + scale_fill_viridis_d(end = 0.9) +
+  coord_flip() + scale_fill_viridis_d(end = 0.9, option = "H") +
   theme(legend.position = c(-0.4, 0.3),
@@
   facet_wrap(~subpop) +
-  coord_flip() + scale_fill_viridis_d(end = 0.9) +
+  coord_flip() + scale_fill_viridis_d(end = 0.9, option = "H") +
   theme(legend.position = "none",
@@
 
+pa_legend <- plot_covar_Ulist_fig2(Ulist_gg$`Daylength (14d prior)`, legend = TRUE) + 
+  labs(subtitle = "Daylength (14d prior)")
+legend <- get_legend(pa_legend) 
 ```
@@
 ggdraw() +
+  theme_oeco + 
   draw_plot(pa01, x = -0.02, y = 0.4, width = 0.1, height = 0.18) +
@@
   draw_plot(f2a, x = 0.3, y = 0.47, width = 0.55, height = 0.53) +
-  #draw_plot(legend, x = 0.15, y = 0.34, width = 0.1, height = 0.2) +
+  draw_plot(legend, x = 0.15, y = 0.34, width = 0.1, height = 0.2) +
   draw_plot(f2b, x = 0.85, y = 0.47, width = 0.15, height = 0.53) +
@@
 
-save_plot(filename = "../manuscript/Coauthors/PNAS/Revision/Figure_2_Covariance_matrices_and_posterior_weights.svg", plot = last_plot(), base_height = 6, base_width = 8.5)
+save_plot(filename = "../manuscript/revision2/Figure_2_recolor_covariance_matrices_and_posterior_weights.svg", plot = last_plot(), base_height = 5.5, base_width = 8)
 ```
@@
 ```{r}
-sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
+sites <- readRDS(here("data/sites.rds"))
 
@@
 ```{r}
-get_gxe_summary(gxe_gg) 
-get_gxe_summary(gxe_mg, n_cond = 7) 
-get_gxe_summary(gxe_gamg, n_sep = 4) 
+fig3_gg <- get_gxe_summary(gxe_gg) 
+fig3_mg <- get_gxe_summary(gxe_mg, n_cond = 7) 
+fig3_bg <- get_gxe_summary(gxe_gamg, n_sep = 4) 
 print("flowering")
-get_gxe_summary(gxe_gf) 
-get_gxe_summary(gxe_mf, n_cond = 6) 
-get_gxe_summary(gxe_gamf, n_sep = 4) 
+fig3_gf <- get_gxe_summary(gxe_gf) 
+fig3_mf <- get_gxe_summary(gxe_mf, n_cond = 6) 
+fig3_bf <- get_gxe_summary(gxe_gamf, n_sep = 4) 
+```
+
+```{r}
+fig3_fractions <- fig3_gg$gxe_fractions |> 
+  mutate(Phenotype = "Greenup Date",
+         Population = "Gulf") |> 
+  full_join(fig3_mg$gxe_fractions) |> 
+  mutate(Phenotype = case_when(is.na(Phenotype) ~ "Greenup Date",
+                               TRUE ~ Phenotype),
+         Population = case_when(is.na(Population) ~ "Midwest",
+                               TRUE ~ Population),
+         ) |> 
+  full_join(fig3_bg$gxe_fractions) |> 
+  mutate(Phenotype = case_when(is.na(Phenotype) ~ "Greenup Date",
+                               TRUE ~ Phenotype),
+         Population = case_when(is.na(Population) ~ "Both",
+                               TRUE ~ Population),
+         ) |> 
+  full_join(fig3_gf$gxe_fractions) |> 
+  mutate(Phenotype = case_when(is.na(Phenotype) ~ "Flowering Date",
+                               TRUE ~ Phenotype),
+         Population = case_when(is.na(Population) ~ "Gulf",
+                               TRUE ~ Population),
+         ) |> 
+  full_join(fig3_mf$gxe_fractions) |> 
+  mutate(Phenotype = case_when(is.na(Phenotype) ~ "Flowering Date",
+                               TRUE ~ Phenotype),
+         Population = case_when(is.na(Population) ~ "Midwest",
+                               TRUE ~ Population),
+         ) |> 
+  full_join(fig3_bf$gxe_fractions) |> 
+  mutate(Phenotype = case_when(is.na(Phenotype) ~ "Flowering Date",
+                               TRUE ~ Phenotype),
+         Population = case_when(is.na(Population) ~ "Both",
+                               TRUE ~ Population),
+         )
+```
+
+```{r}
+fig3_fractions <- fig3_fractions |> 
+  rename(`North vs North` = North_North,
+         `North vs Texas` = North_Texas,
+         `Texas vs Texas` = Texas_Texas) |> 
+  pivot_longer(cols = `North vs North`:`Texas vs Texas`, names_to = "Regions", values_to = "Fraction of Effects with this GxE Type")
+
+fig3_fractions$Type_of_GxE <- factor(fig3_fractions$Type_of_GxE, levels = c("Sign", "Magnitude", "None"))
+fig3_fractions$Phenotype <- factor(fig3_fractions$Phenotype, 
+                                   levels = c("Greenup Date", "Flowering Date"))
+fig3_fractions$Population <- factor(fig3_fractions$Population, 
+                                     levels = c("Gulf", "Midwest", "Both"))
+fig3_frac_plot <- fig3_fractions |> 
+  ggplot(aes(x = Regions, y = `Fraction of Effects with this GxE Type`)) + 
+  geom_bar(aes(fill = Type_of_GxE), position = "stack", stat = "identity", color = "black") + 
+  facet_grid(Phenotype ~ Population) +
+  scale_fill_manual(values = c("#440154FF", "#5DC863FF",
+                                  "white")) +
+  theme(legend.position = "top",
+        panel.spacing.x = unit(0.1, "mm"), 
+        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
+
+```
+
+```{r}
+#| eval: false
+
+ggdraw() +
+  theme_oeco +
+  draw_plot(fig3_frac_plot, x = 0, y = 0, width = 1, height = .67)
+save_plot(filename = "../manuscript/revision2/Figure_3_Fraction_GxE_by_Region.svg", base_height = 6, base_width = 4.5, plot = last_plot())
+```
+
+```{r}
+p1 <- mash_plot_effects(m_gulf_gr_hyp, n = 1)
+p1$ggobject + 
+  theme_oeco + 
+  coord_flip()
+
+save_plot(filename = "../manuscript/revision2/Figure_3_GxE_Types_Example.svg", base_height = 3, base_width = 2.5, plot = last_plot())
 ```
@@
     geom_tile(aes(fill = .data$covar)) +
-    scale_fill_viridis_c(option = "B", limits = c(0, upper)) +
+    scale_fill_viridis_c(option = "H", limits = c(0, upper)) +
     #geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
@@
  pvdiv_plot_cor(corrmatrix = gxe_gamf$S_AP, legend = TRUE) + 
+  theme_oeco +
   labs(subtitle = " ")
@@
 ```{r}
-p3legend <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_AP, legend = TRUE, upper = 4333) + 
+p3legend <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_AP, legend = TRUE, 
+                           upper = 4333) + 
+  theme_oeco +
   labs(title = "Gulf", subtitle = "Sign difference")
@@
 
-p3a1 <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_AP, legend = FALSE, upper = 4333) + 
+p3a1 <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_AP, legend = FALSE, 
+                       upper = 4333) + 
+  theme_oeco +
   labs(subtitle = " ")
-p3a2 <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_DS, legend = FALSE, upper = 4333) + 
+p3a2 <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_DS, legend = FALSE, 
+                       upper = 4333) + 
+  theme_oeco +
   labs(subtitle = " ")
-p3a3 <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_2_no, legend = FALSE, upper = 4333) + 
+p3a3 <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_2_no, legend = FALSE, 
+                       upper = 4333) + 
+  theme_oeco +
   labs(subtitle = " ")
-p3b1 <- pvdiv_plot_cor(corrmatrix = gxe_mg$S_AP, legend = FALSE, upper = 4333) + 
+p3b1 <- pvdiv_plot_cor(corrmatrix = gxe_mg$S_AP, legend = FALSE, 
+                       upper = 4333) + 
+  theme_oeco +
   labs(subtitle = "Sign difference")
-p3b2 <- pvdiv_plot_cor(corrmatrix = gxe_mg$S_DS, legend = FALSE, upper = 4333) + 
+p3b2 <- pvdiv_plot_cor(corrmatrix = gxe_mg$S_DS, legend = FALSE, 
+                       upper = 4333) + 
+  theme_oeco +
   labs(subtitle = "Magnitude difference")
-p3b3 <- pvdiv_plot_cor(corrmatrix = gxe_mg$S_2_no, legend = FALSE, upper = 4333) + 
+p3b3 <- pvdiv_plot_cor(corrmatrix = gxe_mg$S_2_no, legend = FALSE, 
+                       upper = 4333) + 
+  theme_oeco +
   labs(subtitle = "Not distinguishable")
-p3c1 <- pvdiv_plot_cor(corrmatrix = gxe_gamg$S_AP, legend = FALSE, upper = 4333) + 
+p3c1 <- pvdiv_plot_cor(corrmatrix = gxe_gamg$S_AP, legend = FALSE, 
+                       upper = 4333) + 
+  theme_oeco +
   labs(subtitle = " ")
-p3c2 <- pvdiv_plot_cor(corrmatrix = gxe_gamg$S_DS, legend = FALSE, upper = 4333) + 
+p3c2 <- pvdiv_plot_cor(corrmatrix = gxe_gamg$S_DS, legend = FALSE, 
+                       upper = 4333) + 
+  theme_oeco +
   labs(subtitle = " ")
-p3c3 <- pvdiv_plot_cor(corrmatrix = gxe_gamg$S_2_no, legend = FALSE, upper = 4333) + 
+p3c3 <- pvdiv_plot_cor(corrmatrix = gxe_gamg$S_2_no, legend = FALSE, 
+                       upper = 4333) + 
+  theme_oeco +
   labs(subtitle = " ")
 
-p3d1 <- pvdiv_plot_cor(corrmatrix = gxe_gf$S_AP, legend = FALSE, upper = 4333) + 
+p3d1 <- pvdiv_plot_cor(corrmatrix = gxe_gf$S_AP, legend = FALSE, 
+                       upper = 4333) + 
+  theme_oeco +
   labs(subtitle = " ")
-p3d2 <- pvdiv_plot_cor(corrmatrix = gxe_gf$S_DS, legend = FALSE, upper = 4333) + 
+p3d2 <- pvdiv_plot_cor(corrmatrix = gxe_gf$S_DS, legend = FALSE, 
+                       upper = 4333) + 
+  theme_oeco +
   labs(subtitle = " ")
-p3d3 <- pvdiv_plot_cor(corrmatrix = gxe_gf$S_2_no, legend = FALSE, upper = 4333) + 
+p3d3 <- pvdiv_plot_cor(corrmatrix = gxe_gf$S_2_no, legend = FALSE, 
+                       upper = 4333) + 
+  theme_oeco +
   labs(subtitle = " ")
-p3e1 <- pvdiv_plot_cor(corrmatrix = gxe_mf$S_AP, legend = FALSE, upper = 4333) + 
+p3e1 <- pvdiv_plot_cor(corrmatrix = gxe_mf$S_AP, legend = FALSE, 
+                       upper = 4333) + 
+  theme_oeco +
   labs(subtitle = "Sign difference")
-p3e2 <- pvdiv_plot_cor(corrmatrix = gxe_mf$S_DS, legend = FALSE, upper = 4333) + 
+p3e2 <- pvdiv_plot_cor(corrmatrix = gxe_mf$S_DS, legend = FALSE, 
+                       upper = 4333) + 
+  theme_oeco +
   labs(subtitle = "Magnitude difference")
-p3e3 <- pvdiv_plot_cor(corrmatrix = gxe_mf$S_2_no, legend = FALSE, upper = 4333) + 
+p3e3 <- pvdiv_plot_cor(corrmatrix = gxe_mf$S_2_no, legend = FALSE, 
+                       upper = 4333) + 
+  theme_oeco +
   labs(subtitle = "Not distinguishable")
-p3f1 <- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_AP, legend = FALSE, upper = 4333) + 
+p3f1 <- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_AP, legend = FALSE, 
+                       upper = 4333) + 
+  theme_oeco +
   labs(subtitle = " ")
-p3f2 <- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_DS, legend = FALSE, upper = 4333) + 
+p3f2 <- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_DS, legend = FALSE, 
+                       upper = 4333) + 
+  theme_oeco +
   labs(subtitle = " ")
-p3f3 <- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_2_no, legend = FALSE, upper = 4333) + 
+p3f3 <- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_2_no, legend = FALSE, 
+                       upper = 4333) + 
+  theme_oeco +
   labs(subtitle = " ")
@@
 ggdraw() +
+  theme_oeco +
   draw_plot(p3a1, x = 0, y = .82, width = 1/3, height = .18) +
@@
 
+2014-03-06: Change margins so entire QTL plots are shown. Panel a: Greenup QTL. Panel b: Flowering QTL. Indicate which QTL have 1% mash tail SNPs within.
+
 Comparison with QTL for the four-way mapping cross for green up (GR50), flowering (FL50), three functions of weather variables for green-up (from Figure 2; Both: Temp Ave, 1d; Both: Daylength change (s) 1d prior; Gulf: Daylength (14d prior), and two functions of weather variables for flowering (from Figure 2: Midwest: day length change (s) 2d prior; Gulf: cumulative rainfall, 7d).
@@
 ```{r}
-workingdir <- "~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/"
+workingdir <- here("phenology-fourway/5-QTLmappingForAlice/")
 load(file.path(workingdir, 'cross_reduced_marker_FL50_phenotypes_Alice_7_SITES.RData'))
@@
 
-Decide on cutoff to use for rug plots and enrichment analysis: SNPs sig in 1% tail or BF>2, whichever is stricter
-
-1% tail is stricter in all cases except Midwest flowering, where BF>2 is stricter, equivalent to a 0.49% tail (0.995119 quantile).
-
-What is the Bayes Factor at the 99th quantile? If it's not above 2, what is the quantile where the BF > 2?
+Decide on cutoff to use for rug plots and enrichment analysis: SNPs sig in 1% tail or BF\>2, whichever is stricter
+
+1% tail is stricter in all cases except Midwest flowering, where BF\>2 is stricter, equivalent to a 0.49% tail (0.995119 quantile).
+
+What is the Bayes Factor at the 99th quantile? If it's not above 2, what is the quantile where the BF \> 2?
+
 ```{r}
@@
   pivot_longer(GR50:crain_7d, names_to = "Phenotype", values_to = "LOD") %>%
-  mutate(Phe = case_when(Phenotype %in% c("GR50", "tave_1d", "dyln_14d_prior", "dyln_change_sec") ~ "greenup",
-                         Phenotype %in% c("FL50", "dyln_change_sec_2d_prior", "crain_7d") ~ "flowering", 
+  mutate(Phe = case_when(Phenotype %in% c("GR50", "tave_1d", "dyln_14d_prior",
+                                          "dyln_change_sec") ~ "greenup",
+                         Phenotype %in% c("FL50", "dyln_change_sec_2d_prior",
+                                          "crain_7d") ~ "flowering", 
+                         TRUE ~ NA_character_),
+         Phenotype_key = case_when(Phenotype == "GR50" ~ 
+                                     "Greenup Date",
+                                   Phenotype == "tave_1d" ~ 
+                                     "Temp Ave, 1d",
+                                   Phenotype == "dyln_14d_prior" ~ 
+                                     "Daylength (14d prior)",
+                                   Phenotype == "dyln_change_sec" ~ 
+                                     "Daylength change (s)",
+                                   Phenotype == "FL50" ~
+                                     "Flowering Date",
+                                   Phenotype == "dyln_change_sec_2d_prior" ~
+                                     "Daylength change (s), 2d prior",
+                                   Phenotype == "crain_7d" ~ 
+                                     "Cumulative rainfall, 7d",
                                    TRUE ~ NA_character_))
 
-s1full %>%
+fig4b <- s1full %>%
   as_tibble(rownames = "Marker") %>%
@@
   filter(Phenotype %in% c("FL50", "dyln_change_sec_2d_prior", "crain_7d")) %>% 
-  separate(Marker, into = c("Chr", "Pos"), sep = "_", remove = FALSE, convert = TRUE) %>%
+  separate(Marker, into = c("Chr", "Pos"), sep = "_", remove = FALSE, 
+           convert = TRUE) %>%
+  mutate(Phenotype_key = case_when(Phenotype == "GR50" ~ 
+                                     "Greenup Date",
+                                   Phenotype == "tave_1d" ~ 
+                                     "Temp Ave, 1d",
+                                   Phenotype == "dyln_14d_prior" ~ 
+                                     "Daylength (14d prior)",
+                                   Phenotype == "dyln_change_sec" ~ 
+                                     "Daylength change (s)",
+                                   Phenotype == "FL50" ~
+                                     "Flowering Date",
+                                   Phenotype == "dyln_change_sec_2d_prior" ~
+                                     "Daylength change (s), 2d prior",
+                                   Phenotype == "crain_7d" ~ 
+                                     "Cumulative rainfall, 7d",
+                                   TRUE ~ NA_character_)) |> 
   filter(!grepl("loc", Chr)) %>%
   ggplot(aes(x = Pos, y = LOD)) + 
-  geom_line(aes(color = Phenotype)) +
+  geom_line(aes(color = Phenotype_key)) +
   geom_rug(data = rug_gamf, sides = "t", alpha = 0.1) + 
@@
   geom_hline(data = filter(threshold_df, 
-                           Phenotype %in% c("FL50", "dyln_change_sec_2d_prior", "crain_7d")),
-             aes(yintercept = LOD, color = Phenotype, linetype = Phenotype)) +
-  facet_wrap(~Chr, nrow = 1, scales = "free_x") +
-  switchgrassGWAS::theme_oeco + 
+                           Phenotype_key %in% c("Flowering Date", 
+                                                "Daylength change (s), 2d prior",
+                                                "Cumulative rainfall, 7d")),
+             aes(yintercept = LOD, color = Phenotype_key, linetype = Phenotype_key)) +
+  facet_wrap(~Chr, nrow = 2, scales = "free_x") +
+  theme_oeco +
   scale_color_manual(values = c("#000099", "#CC0066", "darkgrey", "#000666")) + 
@@
         panel.grid.minor = element_blank(),
+        panel.spacing.x = unit(0.1, "mm"),
         text = element_text(size = 10, face="bold"),
@@
   coord_cartesian(clip = "off")
-save_plot(filename = "../manuscript/revision2/Figure_4_FL50_QTL_by_lod_score_w_rugs.svg", base_height = 3, base_width = 13.5, plot = last_plot())
-
-s1full %>%
+save_plot(filename = "../manuscript/revision2/Figure_4_FL50_QTL_by_lod_score_w_rugs.svg", base_height = 6, base_width = 6.5, plot = last_plot())
+```
+
+```{r}
+fig4a <- s1full |> 
   as_tibble(rownames = "Marker") %>%
   pivot_longer(GR50:crain_7d, names_to = "Phenotype", values_to = "LOD") %>%
-  filter(Phenotype %in% c("GR50", "tave_1d", "dyln_14d_prior", "dyln_change_sec")) %>% 
-  separate(Marker, into = c("Chr", "Pos"), sep = "_", remove = FALSE, convert = TRUE) %>%
+  filter(Phenotype %in% c("GR50", "tave_1d", "dyln_14d_prior", 
+                          "dyln_change_sec")) %>% 
+  mutate(Phenotype_key = case_when(Phenotype == "GR50" ~ 
+                                     "Greenup Date",
+                                   Phenotype == "tave_1d" ~ 
+                                     "Temp Ave, 1d",
+                                   Phenotype == "dyln_14d_prior" ~ 
+                                     "Daylength (14d prior)",
+                                   Phenotype == "dyln_change_sec" ~ 
+                                     "Daylength change (s)",
+                                   TRUE ~ NA_character_)) |> 
+  separate(Marker, into = c("Chr", "Pos"), sep = "_", remove = FALSE, 
+           convert = TRUE) %>%
   filter(!grepl("loc", Chr)) %>%
   ggplot(aes(x = Pos, y = LOD)) + 
-  geom_line(aes(color = Phenotype)) +
+  geom_line(aes(color = Phenotype_key)) +
   geom_rug(data = rug_gamg, sides = "t", alpha = 0.1) + 
@@
   geom_hline(data = filter(threshold_df, 
-                           Phenotype %in% c("GR50", "tave_1d", "dyln_14d_prior", 
-                                            "dyln_change_sec")), 
-             aes(yintercept = LOD, color = Phenotype, linetype = Phenotype)) +
-  facet_wrap(~Chr, nrow = 1, scales = "free_x") +
-  switchgrassGWAS::theme_oeco + 
-  scale_color_manual(values = c("#006600", "#FFCC00",  "grey", "darkorange", "#000666")) + 
+                           Phenotype_key %in% c("Greenup Date", "Temp Ave, 1d",
+                                            "Daylength (14d prior)", 
+                                            "Daylength change (s)")), 
+             aes(yintercept = LOD, color = Phenotype_key, 
+                 linetype = Phenotype_key)) +
+  facet_wrap(~Chr, nrow = 2, scales = "free_x") +
+  theme_oeco +
+  scale_color_manual(values = c("#006600", "#FFCC00",  "grey", "darkorange",
+                                "#000666")) + 
   scale_linetype_manual(values = c(4,4,4,4,4,4)) +
@@
   coord_cartesian(clip = "off")
-save_plot(filename = "../manuscript/revision2/Figure_4_GR50_QTL_by_lod_score_w_rugs.svg", base_height = 3, base_width = 13.5, plot = last_plot())
+save_plot(filename = "../manuscript/revision2/Figure_4_GR50_QTL_by_lod_score_w_rugs.svg", base_height = 6, base_width = 6.5, plot = last_plot())
 ```
@@
 Prepare the QTL dataframe
-```{r}
-qtldf <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv")
+
+```{r}
+qtldf <- read_csv(here("phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv"))
 
 qtldf <- qtldf %>%
-  separate(flank_lo, into = c("CHR", "Pos_Mb_lo"), sep = "_", convert = TRUE) %>%
-  separate(flank_hi, into = c("flank", "Pos_Mb_hi"), sep = "_", convert = TRUE) %>%
+  separate(flank_lo, into = c("CHR", "Pos_Mb_lo"), sep = "_", 
+           convert = TRUE) %>%
+  separate(flank_hi, into = c("flank", "Pos_Mb_hi"), sep = "_", 
+           convert = TRUE) %>%
   mutate(Pos_Mb_lo = round(Pos_Mb_lo, digits = 2),
@@
   mutate(phe = case_when(lodcolumn == "FL50" ~ "flowering",
-                              lodcolumn == "dyln_change_sec_2d_prior" ~ "daylength change (s) 2d prior",
-                              lodcolumn == "dyln_change_sec" ~ "daylength change (s) 1d prior",
+                              lodcolumn == "dyln_change_sec_2d_prior" ~
+                           "daylength change (s) 2d prior",
+                              lodcolumn == "dyln_change_sec" ~ 
+                           "daylength change (s) 1d prior",
                               lodcolumn == "GR50" ~ "greenup",
-                              lodcolumn == "crain_gr2fl" ~ "Rainfall, GR to FL")) %>%
+                              lodcolumn == "crain_gr2fl" ~ 
+                           "Rainfall, GR to FL")) %>%
   #filter(phe %in% c("Daylength", "GDD, GR to FL")) %>%
@@
 ```
+
 Function to compute hypergeometric test for each row in a QTL df
+
 ```{r}
@@
   filter(pvalue < 0.05)
+fl_enrichd <- enrich_df %>%
+  filter(inQTLandBF > 0) |> 
+  arrange(Chr, Pos_lo, lodcolumn) |> 
+  distinct(Chr, Pos_lo, subpop, .keep_all = TRUE)
 
@@
 # 10 unique QTL by Pos lo/hi. 5 of 10 have enriched mash hits.
+```
+
+```{r}
+fl_enrichd |> 
+  write_csv(here("data", "Enriched_flowering_QTL.csv"))
+fl_labels <- read_csv(here("data", "Enriched_flowering_QTL_labels.csv"))
+
+fig_4b <- fig4b +
+  geom_text_repel(data = fl_labels, aes(x = pos, y = lod, label = subpop), nudge_y = 2)
+library(ggrepel)
 ```
@@
 enrich_gr  # 0 of 3 unique QTL have enriched mash hits for green-up.
-
+enrich_gr |> write_csv(file = here("data", "Enriched_greenup_QTL.csv"))
+```
+
+```{r}
+gr_labels <- read_csv(here("data", "Enriched_greenup_QTL_labels.csv"))
+
+fig_4a <- fig4a +
+  geom_text_repel(data = gr_labels, aes(x = pos, y = lod, label = subpop), nudge_y = 2)
+library(ggrepel)
+```
+
+```{r}
+ggdraw() + 
+  theme_oeco +
+  draw_plot(fig_4a, x = 0, y = 0.5, width = 1, height = .5) +
+  draw_plot(fig_4b, x = 0, y = 0, width = 1, height = .5)
+
+save_plot(filename = "../manuscript/revision2/Figure_4_QTL_Overlaps.svg", base_height = 9, base_width = 6.5, plot = last_plot())
 ```
@@
 ```{r}
-sig_random <- readRDS(file = "../data/Random_QTL_df_FL50_significant_enrichments_of_mash_1per_tail_or_mash_BF2_tail_SNPs_1000runs.rds")
+sig_random <- readRDS(file = here( "data/Random_QTL_df_FL50_significant_enrichments_of_mash_1per_tail_or_mash_BF2_tail_SNPs_1000runs.rds"))
 
@@
   ggplot(aes(x = `mash 1% tail significant enrichments` )) +
-  switchgrassGWAS::theme_oeco +
+  theme_oeco +
   geom_histogram(binwidth = 1) + 

### 2024-03-12 manuscript/revision2/revision2.qmd
manuscript: AH edits from 2024-03-07

@@
 ---
-title: Mapping Mixtures of Genotype-by-Weather Interactions in Switchgrass (Panicum virgatum)
+title: Diverse Genotype-by-Weather Interactions in Switchgrass
 format:
@@
 abstract: |
-  The timing of vegetative and reproductive growth in plants ("phenological timings") depends on both genetic variation and environmental cues; thus, phenological traits can have important genotype-by-environment (GxE) interactions.
-Identifying the mixture and prevalence of GxE primed by weather prior to the phenological event (GxWeather) should aid prediction and manipulation of phenological timings.
-Here, we map GxWeather effects on phenological timings in two highly divergent switchgrass (*Panicum virgatum*) populations using repeated plantings of cloned individuals from these populations at eight sites spanning the central United States.
-We distinguish GxWeather patterns covarying with interpretable weather-based cues from agnostic, site-based patterns.
-Most GxWeather effects belong to the latter category and do not covary with weather-based cues.
-However, 65% of effects on the timing of vegetative growth in the Gulf population covary with daylength 14 days prior to green-up date.
-33% of effects on the timing of flowering in the Gulf population covary with cumulative rainfall in the seven days prior to flowering, while 22% of effects on flowering timing in the Midwest population covary with day length change two days prior to flowering.
-An independent pseudo-F2 cross of Gulf and Midwest individuals at the same sites mapped eight additive QTLs for flowering and three for vegetative growth timing, all with significant associations and three with enrichment of highly significant associations in our model of GxWeather effects.
+  The timing of vegetative and reproductive growth in plants ("phenological timings") depends on genetic effects, environmental (e.g., weather) cues, and possibly their interaction.
+Quantifying these GxWeather interactions can aid prediction and manipulation of phenological timings.
+Here, we map GxWeather effects on phenological timings in two highly divergent switchgrass (*Panicum virgatum*) populations using repeated plantings of cloned individuals from these populations at eight sites across the central United States.
+Some genetic effects covary with  interpretable weather-based cues.
+For example, in the Gulf population, 65% of genetic effects on the timing of vegetative growth covary with daylength 14 days prior to green-up date, and 33% of genetic effects on the timing of flowering covary with cumulative rainfall in the seven days prior to flowering.
+However, most variation in genetic effects across sites is site-specific and can not be attributed to variation in weather variables alone.
 We demonstrate that we can identify genetic variation with GxWeather and assign these loci to specific weather-based cues or other patterns.
-Breeding for particular alleles at these loci could change flowering responsiveness to photoperiod and rainfall cues in switchgrass.
-More broadly, this approach could be used to identify genetic marker-environment interactions in any species with related populations phenotyped in multiple environments.
+Breeding for particular alleles at the GxWeather loci we identified could change flowering responsiveness in a photoperiod and rainfall-specific way.
+More broadly, our approach proposes a refined, powerful characterization of genotype-by-environment interactions in any species phenotyped in multiple environments.
 runningauthor: "MacQueen"
@@
 
-\firstpage[8]{2}
+\firstpage[15]{2}
 
@@
 Species with wide natural distributions can have multiple distinct environmentally cued phenological responses: for example, populations of sunflower (*Helianthus annuus*) exhibit day-neutral, facultative short day, and facultative long-day flowering responses, which vary with their environments [@blackman2013interacting; @henry2014transitions].
-Distinct genetic responses in different environments are known as genotype by environment interactions, or GxE.
+Distinct genetic responses in different environments are known as genotype-by-environment interactions, or GxE.
 
@@
 Recent statistical advances in studying phenological GxE have involved determining critical environmental indices before the phenological event occurs, such as photothermal time within a critical growth window @li2018genomic.
-However, most studies of flowering GxE focus on finding a single, best fitting form of genotype-environment covariance, despite the key expectation that different genetic subpopulations, and even different genomic regions, have likely evolved distinct patterns of GxE.
-Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or sign-changing GxE, at the level of individual loci [@levene1953genetic; @felsenstein1976theoretical; @kawecki2004conceptual; @hedrick1986genetic], previous work has found limited evidence of antagonistic pleiotropy [@des2013genotype; @wadgymar2017identifying].
+However, most studies of flowering GxE focus on finding a single, best fitting form of genotype-environment covariance, despite the key expectation that different genetic subpopulations, and even different genomic regions, have likely evolved distinct patterns of GxE [@Weine2023.06.21.545998].
+Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or sign-changing GxE, at the level of individual loci [@levene1953genetic; @felsenstein1976theoretical; @kawecki2004conceptual; @hedrick1986genetic], previous empirical work has found limited evidence of antagonistic pleiotropy [@des2013genotype; @wadgymar2017identifying].
 However, this work has been limited by a known statistical bias that reduced detection of genetic effects that differ in sign [@des2013genotype; @anderson2013genetic; @anderson2011evolutionary].
@@
 We define multiple ways phenological timings might covary with weather (\@tbl-covar) and additional ways phenological timings might vary by site (\@SI-info), then jointly re-estimate genetic effects on these timings at all eight sites using the set of these covariance matrices that significantly improved the modeled log-likelihood when included (\@SI-info) @urbut2019flexible.
-We use the Bayesian framework *mash* (multivariate adaptive shrinkage), developed by @urbut2019flexible, to refine effect size estimates from genome-wide association (GWAS) conducted on individuals at each of the eight sites.
+We use the Bayesian framework *mash* (multivariate adaptive shrinkage) developed by @urbut2019flexible, to refine effect size estimates from genome-wide association (GWAS) conducted on individuals at each of the eight sites.
 *mash* allows us to identify and specify multiple covariance structures among genetic effect estimates across sites, including structures that represent covariance in weather variables of interest.
-Importantly, *mash* does not have a statistical bias in detecting genetic effects with the same or opposite signs @10.1093/biostatistics/kxw041.
+Importantly, this method circumvents statistical biases in detecting genetic effects with the same or opposite signs @10.1093/biostatistics/kxw041.
 To confirm our genetic mapping of GxWeather, we compare our posterior genetic effect estimates from *mash* to mapping results from an outbred pseudo-F2 cross grown at the same sites.
@@
 
-Narrow-sense heritabilities (h^2^) indicated that rank-changing GxE for these phenotypes was present across the common gardens (@fig-map C).
+Narrow-sense heritabilities (h^2^) suggested that rank-changing GxE for these phenotypes was present across the common gardens (@fig-map C).
 h^2^ were typically high at individual gardens: 59% on average for green-up date, and 87% for flowering date.
@@
 
-We next looked for evidence of GxWeather by jointly re-estimating genetic effects across all eight common gardens using the mash model, which allows effects to covary between gardens in many ways (@fig-map D, SI Appendix, Datasets 1-6).
-We specified three categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event (Table 1, SI Appendix, Section S2).
+We next looked for evidence of GxWeather by jointly re-estimating genetic effects across all eight common gardens using the mash model, which can flexibly capture many modes of covariance across gardens (@fig-map D, SI Appendix, Datasets 1-6).
+We specified three qualitative categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event (Table 1, SI Appendix, Section S2).
 For example, the more similar that the amount of rainfall prior to flowering is for genotypes A and B between gardens in MI and MO, the nearer these covariances will be to one.
@@
 
-As the climate and the natural environment change, it is increasingly critical to understand how patterns of gene-environment and plant-environment interactions will change in response.
+As the climate and the natural environment change, it is increasingly critical to understand how patterns of plant-environment interactions will change in response.
 To do this, we must understand the current patterns of trait covariation across environments, the genetic underpinnings of these patterns, and the cases where this covariation can be altered.
@@
 If this integration time itself varies between individuals, we cannot select a covariance structure that reflects this, though these structures would likely be highly correlated with GxWeather structures we did include.
-Mash offers an opportunity to specify multiple environmental cues and compete them to explain patterns of genetic effects, allowing us to detect how important these cues are genome-wide, and how strongly each cue influences each SNP.
+Our approach offers an opportunity to specify multiple environmental cues and compete them to explain patterns of genetic effects, allowing us to detect how important these cues are genome-wide, and how strongly each cue influences each SNP.
 This is a key development to further improve our understanding of genetic variation in GxE.
@@
 (a) Map and trait histograms of green-up and flowering dates across two genetically distinct switchgrass subpopulations and eight common gardens.
-Purple represents individuals from the Midwest genetic subpopulation, and pink individuals from the Gulf subpopulation.
+Purple represents individuals from the Midwest genetic subpopulation, and pink individuals from the Gulf subpopulation; map positions represent the original collection locations for the genotypes, and shapes represent the ecotype of the genotype.
 Vertical dashed lines indicate the summer solstice.
@@
 Covariance matrices included in mash that had zero posterior weight in all three mash runs on the genetic subpopulations, such as the identity matrix, are not shown.
-(c,e) Total posterior weight placed on covariance matrices that were hypothesized or canonical, for the (c) green-up date phenotype and (e) flowering date phenotype.](images/Figure_2_Covariance_matrices_and_posterior_weights.png){#fig-covar}
+(c,e) Total posterior weight placed on covariance matrices that were hypothesized or canonical, for the (c) green-up date phenotype and (e) flowering date phenotype.](Figure_2_recolor_covariance_matrices_and_posterior_weights.png){#fig-covar}
 

### 2024-03-12 manuscript/revision2/revision2.tex
manuscript: AH edits from 2024-03-07

@@
 \hypersetup{
-  pdftitle={Mapping Mixtures of Genotype-by-Weather Interactions in Switchgrass (Panicum virgatum)},
+  pdftitle={Diverse Genotype-by-Weather Interactions in Switchgrass},
   pdfauthor={Alice H.
@@
 
-\title{Mapping Mixtures of Genotype-by-Weather Interactions in
-Switchgrass (Panicum virgatum)}
+\title{Diverse Genotype-by-Weather Interactions in Switchgrass}
 
@@
 The timing of vegetative and reproductive growth in plants
-(``phenological timings'') depends on both genetic variation and
-environmental cues; thus, phenological traits can have important
-genotype-by-environment (GxE) interactions. Identifying the mixture and
-prevalence of GxE primed by weather prior to the phenological event
-(GxWeather) should aid prediction and manipulation of phenological
-timings. Here, we map GxWeather effects on phenological timings in two
-highly divergent switchgrass (\emph{Panicum virgatum}) populations using
-repeated plantings of cloned individuals from these populations at eight
-sites spanning the central United States. We distinguish GxWeather
-patterns covarying with interpretable weather-based cues from agnostic,
-site-based patterns. Most GxWeather effects belong to the latter
-category and do not covary with weather-based cues. However, 65\% of
-effects on the timing of vegetative growth in the Gulf population covary
-with daylength 14 days prior to green-up date. 33\% of effects on the
-timing of flowering in the Gulf population covary with cumulative
-rainfall in the seven days prior to flowering, while 22\% of effects on
-flowering timing in the Midwest population covary with day length change
-two days prior to flowering. An independent pseudo-F2 cross of Gulf and
-Midwest individuals at the same sites mapped eight additive QTLs for
-flowering and three for vegetative growth timing, all with significant
-associations and three with enrichment of highly significant
-associations in our model of GxWeather effects. We demonstrate that we
-can identify genetic variation with GxWeather and assign these loci to
-specific weather-based cues or other patterns. Breeding for particular
-alleles at these loci could change flowering responsiveness to
-photoperiod and rainfall cues in switchgrass. More broadly, this
-approach could be used to identify genetic marker-environment
-interactions in any species with related populations phenotyped in
+(``phenological timings'') depends on genetic effects, environmental
+(e.g., weather) cues, and possibly their interaction. Quantifying these
+GxWeather interactions can aid prediction and manipulation of
+phenological timings. Here, we map GxWeather effects on phenological
+timings in two highly divergent switchgrass (\emph{Panicum virgatum})
+populations using repeated plantings of cloned individuals from these
+populations at eight sites across the central United States. Some
+genetic effects covary with interpretable weather-based cues. For
+example, in the Gulf population, 65\% of genetic effects on the timing
+of vegetative growth covary with daylength 14 days prior to green-up
+date, and 33\% of genetic effects on the timing of flowering covary with
+cumulative rainfall in the seven days prior to flowering. However, most
+variation in genetic effects across sites is site-specific and can not
+be attributed to variation in weather variables alone. We demonstrate
+that we can identify genetic variation with GxWeather and assign these
+loci to specific weather-based cues or other patterns. Breeding for
+particular alleles at the GxWeather loci we identified could change
+flowering responsiveness in a photoperiod and rainfall-specific way.
+More broadly, our approach proposes a refined, powerful characterization
+of genotype-by-environment interactions in any species phenotyped in
 multiple environments.
@@
 \ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
-\firstpage[8]{2}
+\firstpage[15]{2}
 
@@
 their environments (5, 6). Distinct genetic responses in different
-environments are known as genotype by environment interactions, or GxE.
+environments are known as genotype-by-environment interactions, or GxE.
 
@@
 different genetic subpopulations, and even different genomic regions,
-have likely evolved distinct patterns of GxE. Additionally, despite
+have likely evolved distinct patterns of GxE (22). Additionally, despite
 theoretical predictions that local adaptation should involve
 antagonistic pleiotropy, or sign-changing GxE, at the level of
-individual loci (22--25), previous work has found limited evidence of
-antagonistic pleiotropy (12, 26). However, this work has been limited by
-a known statistical bias that reduced detection of genetic effects that
-differ in sign (26--28). Thus, despite substantial interest in the
-frequencies of various forms of GxE, the prevalence of antagonistic
-pleiotropy relative to other forms of GxE remains unknown.
+individual loci (23--26), previous empirical work has found limited
+evidence of antagonistic pleiotropy (12, 27). However, this work has
+been limited by a known statistical bias that reduced detection of
+genetic effects that differ in sign (27--29). Thus, despite substantial
+interest in the frequencies of various forms of GxE, the prevalence of
+antagonistic pleiotropy relative to other forms of GxE remains unknown.
 
@@
 subpopulation. Switchgrass is considered a short-day plant with
-reproductive development strongly linked to day of the year (29).
+reproductive development strongly linked to day of the year (30).
 However, as part of its wide environmental adaptation across the eastern
 half of North America, its photoperiodicity has been predicted to differ
-by plant latitude of origin (30, 31). We previously found divergent
+by plant latitude of origin (31, 32). We previously found divergent
 Midwest and Gulf genetic subpopulations of switchgrass with distinct
@@
 genetic variation associated with each of two fitness proxies, biomass
-and survival (32). The Midwest genetic subpopulation is primarily
+and survival (33). The Midwest genetic subpopulation is primarily
 composed of individuals from the well-studied upland switchgrass ecotype
-(33, 34), while the Gulf subpopulation has individuals from the
+(34, 35), while the Gulf subpopulation has individuals from the
 well-studied lowland ecotype and the phenotypically intermediate coastal
-ecotype (32).
+ecotype (33).
 
@@
 significantly improved the modeled log-likelihood when included
-(@SI-info) (35). We use the Bayesian framework \emph{mash} (multivariate
-adaptive shrinkage), developed by (35), to refine effect size estimates
+(@SI-info) (36). We use the Bayesian framework \emph{mash} (multivariate
+adaptive shrinkage) developed by (36), to refine effect size estimates
 from genome-wide association (GWAS) conducted on individuals at each of
@@
 including structures that represent covariance in weather variables of
-interest. Importantly, \emph{mash} does not have a statistical bias in
-detecting genetic effects with the same or opposite signs (36). To
+interest. Importantly, this method circumvents statistical biases in
+detecting genetic effects with the same or opposite signs (37). To
 confirm our genetic mapping of GxWeather, we compare our posterior
@@
 
-Narrow-sense heritabilities (h\textsuperscript{2}) indicated that
+Narrow-sense heritabilities (h\textsuperscript{2}) suggested that
 rank-changing GxE for these phenotypes was present across the common
@@
 genetic effects across all eight common gardens using the mash model,
-which allows effects to covary between gardens in many ways
+which can flexibly capture many modes of covariance across gardens
 (Figure~\ref{fig-map} D, SI Appendix, Datasets 1-6). We specified three
-categories of covariance structures across gardens: ``canonical''
-covariance, with simple patterns of effect size covariance introduced in
-the initial \emph{mash} manuscript; ``data-driven'' covariances derived
-from common patterns of SNP effects observed in the data, and
-``GxWeather'' covariance, estimated from the covariance of empirical
-weather patterns at each garden at specific times before the
+qualitative categories of covariance structures across gardens:
+``canonical'' covariance, with simple patterns of effect size covariance
+introduced in the initial \emph{mash} manuscript; ``data-driven''
+covariances derived from common patterns of SNP effects observed in the
+data, and ``GxWeather'' covariance, estimated from the covariance of
+empirical weather patterns at each garden at specific times before the
 phenological event (Table 1, SI Appendix, Section S2). For example, the
@@
 As the climate and the natural environment change, it is increasingly
-critical to understand how patterns of gene-environment and
-plant-environment interactions will change in response. To do this, we
-must understand the current patterns of trait covariation across
-environments, the genetic underpinnings of these patterns, and the cases
-where this covariation can be altered. Here, we demonstrate that we can
-associate multiple patterns of GxWeather with specific genomic regions
-using a switchgrass diversity panel grown at eight common gardens. We
-can assign genetic effects to both GxWeather patterns with interpretable
-weather-based cues, and to agnostic, site-based patterns. We use this
-approach to study GxWeather for the timings of vegetative and
-reproductive development in the deeply genetically diverged Gulf and
-Midwest subpopulations of switchgrass.
+critical to understand how patterns of plant-environment interactions
+will change in response. To do this, we must understand the current
+patterns of trait covariation across environments, the genetic
+underpinnings of these patterns, and the cases where this covariation
+can be altered. Here, we demonstrate that we can associate multiple
+patterns of GxWeather with specific genomic regions using a switchgrass
+diversity panel grown at eight common gardens. We can assign genetic
+effects to both GxWeather patterns with interpretable weather-based
+cues, and to agnostic, site-based patterns. We use this approach to
+study GxWeather for the timings of vegetative and reproductive
+development in the deeply genetically diverged Gulf and Midwest
+subpopulations of switchgrass.
 
@@
 result supports theoretical models that local adaptation should involve
-antagonistic pleiotropy at the level of individual loci (22--25), and is
+antagonistic pleiotropy at the level of individual loci (23--26), and is
 the first experimental work using GWAS across common gardens to find
-antagonistic pleiotropy to be common in small genomic regions (12, 37,
-38).
+antagonistic pleiotropy to be common in small genomic regions (12, 38,
+39).
 
@@
 likely be highly correlated with GxWeather structures we did include.
-Mash offers an opportunity to specify multiple environmental cues and
-compete them to explain patterns of genetic effects, allowing us to
-detect how important these cues are genome-wide, and how strongly each
-cue influences each SNP. This is a key development to further improve
-our understanding of genetic variation in GxE.
+Our approach offers an opportunity to specify multiple environmental
+cues and compete them to explain patterns of genetic effects, allowing
+us to detect how important these cues are genome-wide, and how strongly
+each cue influences each SNP. This is a key development to further
+improve our understanding of genetic variation in GxE.
 
@@
 from the Midwest genetic subpopulation, and pink individuals from the
-Gulf subpopulation. Vertical dashed lines indicate the summer solstice.
-Common gardens are arranged in latitudinal order. (b) Phenotypic
-correlations between clonal replicates planted at eight common gardens,
-within and between two genetic subpopulations. (c) Narrow sense
-heritability of green-up and flowering within single common gardens
-(purple) and across all eight common gardens (green), within and between
-two genetic subpopulations. (d) Flow diagram of the methods applied to
-the green-up and flowering dates to jointly estimate SNP effects across
-all sites. Mash was fit to SNP effect data and used to find covariance
-matrices that improved the mash model likelihood using a large set of
-randomly selected, relatively unlinked SNP effects; this model was
-applied to a ``strong'' set of SNP effects with large effect sizes in
-the univariate GWAS.}
+Gulf subpopulation; map positions represent the original collection
+locations for the genotypes, and shapes represent the ecotype of the
+genotype. Vertical dashed lines indicate the summer solstice. Common
+gardens are arranged in latitudinal order. (b) Phenotypic correlations
+between clonal replicates planted at eight common gardens, within and
+between two genetic subpopulations. (c) Narrow sense heritability of
+green-up and flowering within single common gardens (purple) and across
+all eight common gardens (green), within and between two genetic
+subpopulations. (d) Flow diagram of the methods applied to the green-up
+and flowering dates to jointly estimate SNP effects across all sites.
+Mash was fit to SNP effect data and used to find covariance matrices
+that improved the mash model likelihood using a large set of randomly
+selected, relatively unlinked SNP effects; this model was applied to a
+``strong'' set of SNP effects with large effect sizes in the univariate
+GWAS.}
 
@@
 
-\includegraphics{images/Figure_2_Covariance_matrices_and_posterior_weights.png}
+\includegraphics{Figure_2_recolor_covariance_matrices_and_posterior_weights.png}
 
@@
 The formation and resequencing of the diversity panel has been described
-previously (32). The diversity panel contained 134 sequenced, clonally
+previously (33). The diversity panel contained 134 sequenced, clonally
 propagated individuals from the Midwest genetic subpopulation, and 229
@@
 heritability (h\textsuperscript{2}) for green-up and flowering were done
-using linear mixed models and followed (32). Details on these models can
+using linear mixed models and followed (33). Details on these models can
 be found in (SI Appendix, Section S3,S4).
@@
 timing of vegetative growth and flowering, using the switchgrassGWAS
-package and the methods in (39). We used the effect estimates for single
+package and the methods in (40). We used the effect estimates for single
 nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect
@@
 defined as the probability that we make an error in the sign of effect j
-if we were forced to declare the effect positive or negative (36). Thus,
+if we were forced to declare the effect positive or negative (37). Thus,
 rather than asking ``Are these two effects different?'' - as we
@@
 individuals. The formation of this mapping population has been described
-previously (34); additional details on QTL mapping can be found in SI
+previously (35); additional details on QTL mapping can be found in SI
 Appendix, Section S6. To be directly comparable to the diversity panel
@@
 
+\bibitem[\citeproctext]{ref-Weine2023.06.21.545998}
+\CSLLeftMargin{22. }%
+\CSLRightInline{E. Weine, S. P. Smith, R. K. Knowlton, A. Harpak,
+Tradeoffs in modeling context dependency in complex trait genetics.
+\emph{bioRxiv} (2024)
+https:/doi.org/\href{https://doi.org/10.1101/2023.06.21.545998}{10.1101/2023.06.21.545998}.}
+
 \bibitem[\citeproctext]{ref-levene1953genetic}
-\CSLLeftMargin{22. }%
+\CSLLeftMargin{23. }%
 \CSLRightInline{H. Levene, Genetic equilibrium when more than one
@@
 \bibitem[\citeproctext]{ref-felsenstein1976theoretical}
-\CSLLeftMargin{23. }%
+\CSLLeftMargin{24. }%
 \CSLRightInline{J. Felsenstein, The theoretical population genetics of
@@
 \bibitem[\citeproctext]{ref-kawecki2004conceptual}
-\CSLLeftMargin{24. }%
+\CSLLeftMargin{25. }%
 \CSLRightInline{T. J. Kawecki, D. Ebert, Conceptual issues in local
@@
 \bibitem[\citeproctext]{ref-hedrick1986genetic}
-\CSLLeftMargin{25. }%
+\CSLLeftMargin{26. }%
 \CSLRightInline{P. W. Hedrick, Genetic polymorphism in heterogeneous
@@
 \bibitem[\citeproctext]{ref-des2013genotype}
-\CSLLeftMargin{26. }%
+\CSLLeftMargin{27. }%
 \CSLRightInline{D. L. Des Marais, K. M. Hernandez, T. E. Juenger,
@@
 \bibitem[\citeproctext]{ref-anderson2013genetic}
-\CSLLeftMargin{27. }%
+\CSLLeftMargin{28. }%
 \CSLRightInline{J. T. Anderson, C.-R. Lee, C. A. Rushworth, R. I.
@@
 \bibitem[\citeproctext]{ref-anderson2011evolutionary}
-\CSLLeftMargin{28. }%
+\CSLLeftMargin{29. }%
 \CSLRightInline{J. T. Anderson, J. H. Willis, T. Mitchell-Olds,
@@
 \bibitem[\citeproctext]{ref-mitchell1997predicting}
-\CSLLeftMargin{29. }%
+\CSLLeftMargin{30. }%
 \CSLRightInline{R. B. Mitchell, K. J. Moore, L. E. Moser, J. O. Fritz,
@@
 \bibitem[\citeproctext]{ref-parrish2005biology}
-\CSLLeftMargin{30. }%
+\CSLLeftMargin{31. }%
 \CSLRightInline{D. J. Parrish, J. H. Fike, The biology and agronomy of
@@
 \bibitem[\citeproctext]{ref-casler2004latitudinal}
-\CSLLeftMargin{31. }%
+\CSLLeftMargin{32. }%
 \CSLRightInline{M. Casler, K. P. Vogel, C. Taliaferro, R. Wynia,
@@
 \bibitem[\citeproctext]{ref-lovell2021genomic}
-\CSLLeftMargin{32. }%
+\CSLLeftMargin{33. }%
 \CSLRightInline{J. T. Lovell, \emph{et al.}, Genomic mechanisms of
@@
 \bibitem[\citeproctext]{ref-porter1966analysis}
-\CSLLeftMargin{33. }%
+\CSLLeftMargin{34. }%
 \CSLRightInline{C. L. Porter Jr, An analysis of variation between upland
@@
 \bibitem[\citeproctext]{ref-milano2016genetic}
-\CSLLeftMargin{34. }%
+\CSLLeftMargin{35. }%
 \CSLRightInline{E. R. Milano, D. B. Lowry, T. E. Juenger, The genetic
@@
 \bibitem[\citeproctext]{ref-urbut2019flexible}
-\CSLLeftMargin{35. }%
+\CSLLeftMargin{36. }%
 \CSLRightInline{S. M. Urbut, G. Wang, P. Carbonetto, M. Stephens,
@@
 \bibitem[\citeproctext]{ref-10.1093ux2fbiostatisticsux2fkxw041}
-\CSLLeftMargin{36. }%
+\CSLLeftMargin{37. }%
 \CSLRightInline{M. Stephens,
@@
 \bibitem[\citeproctext]{ref-savolainen2013ecological}
-\CSLLeftMargin{37. }%
+\CSLLeftMargin{38. }%
 \CSLRightInline{O. Savolainen, M. Lascoux, J. Merilä, Ecological
@@
 \bibitem[\citeproctext]{ref-lowry2019qtl}
-\CSLLeftMargin{38. }%
+\CSLLeftMargin{39. }%
 \CSLRightInline{D. B. Lowry, \emph{et al.}, QTL\(\times\) environment
@@
 \bibitem[\citeproctext]{ref-lovell2021}
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{40. }%
 \CSLRightInline{J. T. Lovell, \emph{et al.},

### 2024-05-09 manuscript/revision2/cover_letter/cover_letter.qmd
manuscript: first pass at revision2 cover letter

---
author:
  - name: Alice MacQueen* and Tom Juenger
    affiliation:
      - name: The University of Texas at Austin
        department: Department of Integrative Biology
        address: 2415 Speedway
        city: Austin, TX
        postal-code: 78712
address:
  - Editors
  - Proceedings of the National Academy of Sciences
subject: manuscript revision
subject-title: "Subject"
opening: "Dear Editors,"
closing: "Best,"
format: letter-pdf
bibliography: references.bib
---

Thank you for considering this work for publication in *PNAS*. Our manuscript, "Diverse Genotype-by-Weather Interactions in Switchgrass" has undergone substantial revision from our original submission in 2021. This timeframe is long - Dr. MacQueen took a new position in 2022. They continued working with the group as we developed a new algorithm to address major reviewer concerns and significantly altered the manuscript. Because of these alterations, the response to reviewers does not have track changes for each point, though we do point to line numbers for the associated revisions in our response.

Our manuscript is suitable for publication in *PNAS* for the following reasons.

Our work is a key methodological development in understanding genetic variation in genotype-by-environment interactions, or GxE. GxE interactions must vary across the genome; that is, different genes and genetic variants can exhibit different responses to the environment. Yet current statistical models to test for GxE don't have robust frameworks to specify and compete different models of GxE effects, instead focusing on finding a single, best fitting form of GxE covariance. We use multivariate adaptive shrinkage (mash [@urbut2018]) to determine the regions of the switchgrass genome with different types of GxE, specifically regions of the genome with effects on phenology that covary with several weather-based 'cues'. We use a greedy algorithm to compete and select types of GxE and GxWeather that improve our models, then use mash to identify which SNP effects are best modeled with each type of GxE or GxWeather.

Second, to our knowledge, this work is the first to use a statistically unbiased test to detect the relative frequencies of two important types of GxE; antagonistic pleiotropy, or a change in sign of effects between environments, and differential sensitivity, or a change in magnitude of effects between environments. Here, rather than using the local false discovery rate to ask "Are these two effects different?" - as we reasonably expect two effects to be, even if this difference cannot be measured - we use the local false sign rate to ask a more meaningful question: Can we be confident in the sign of these two effects? We use the lfsr to find thousands of genetic effects with antagonistic pleiotropy for the start of vegetative growth for switchgrass.

Finally, with current advances in high-throughput measurement, and the increased resources devoted to experiments in natural and semi-natural environments, we must consider how to share information across these experiments and how to associate natural variation in genomes with natural
variation in complex field environments. Advances in methods to find the environmental cues that matter should allow researchers to determine the adaptations important to responses to the natural environment, and how the importance of these genetic mechanisms changes as the environment itself changes. Our manuscript should be of broad interest to PNAS readership for the insights it provides into these important issues.

We suggest that the original editor, Editorial Board members, and anonymous reviewers for this manuscript would be good choices to assess this revision. In the event that is not practicable, we include some additional suggestions below.

We suggest that Douglas Schemske, \_\_\_, and Qifa Zhang would be good Editorial Board members for this manuscript. All would be able to provide strong insights into mapping the genetic basis of adaptive and ecologically relevant traits in a field experimental context.

We suggest that Johanna Schmitt, Trudy Mackay, and Detlef Weigel would be excellent editors for this manuscript. Dr. Schmitt has done extensive work in *Arabidopsis thaliana* and other plant species unravelling the genetic mechanisms involved in responses to the natural environment. Dr. Mackay has extensive experience mapping the genetics of complex traits and genotype-by-environment interactions. Dr. Weigel has done extensive work understanding genetic diversity in *A. thaliana* including extensive work on floral induction cues.

### 2024-05-09 manuscript/revision2/intermediate_files/response_to_reviews.qmd
manuscript: add response to reviews folder within manuscript/revision2/ and move intermediate files to intermediate_files folder

--- 2023-12-02 manuscript/revision2/response_to_reviews.qmd
@@
 If these covariance matrices are similar, mash will somewhat arbitrarily assign weight to any of the matrices because all lead to the same fit to the effect-size data.
-The mash algorithm doesn\'t give the full posterior distribution on the loadings so you can\'t check for posterior correlations there.
+The mash algorithm doesn't give the full posterior distribution on the loadings so you can't check for posterior correlations there.
 It seems likely to me that the hypothesis-driven and data-driven covariance matrices are somewhat correlated here, and the correlations may differ between green-up and flowering because of the better correlation between the environmental metrics and flowering than green-up.
@@
 
-most_likely_matrix = \'\'
+most_likely_matrix = ''
 
@@
 
+```         
 matrix_likelihood = mash(effects, std.errs, matrix)\[\'likelihood\'\] #obtain model likelihood
@@
     likelihood = matrix_likelihood
+```
 
@@
 
+```         
 likelihood = -inf
+```
 
-most_likely_matrix = \'\'
+most_likely_matrix = ''
 
+```         
 for matrix in matrices:
@@
 
-In order to better understand the behavior of the mash shrinkage algorithm in the presence of correlated covariance matrices, we implemented a \"greedy\" mash algorithm.
+In order to better understand the behavior of the mash shrinkage algorithm in the presence of correlated covariance matrices, we implemented a "greedy" mash algorithm.
 Specifically, given n covariance matrices for each phenotype and corresponding effect sizes and standard errors across the k contexts of interest, we fit a set of mash models with each of the n matrices separately.
@@
 This process is repeated until one of two stop conditions are met: (i) all covariance matrices have been added to the mash model, resulting in a final model with all of the original n covariance matrices are included or (ii) a likelihood ratio test with one degree of freedom between the likelihood of the current model and the model from the previous iteration has a p-value \> 0.05.
-The result is a set of log likelihood estimates for each model and the stepwise \'path\' of covariance matrices that best explain the observed effect sizes and standard errors observed for the phenotype of interest.
+The result is a set of log likelihood estimates for each model and the stepwise 'path' of covariance matrices that best explain the observed effect sizes and standard errors observed for the phenotype of interest.
 

### 2024-05-09 manuscript/revision2/intermediate_files/response_to_reviews.tex
manuscript: add response to reviews folder within manuscript/revision2/ and move intermediate files to intermediate_files folder

--- 2023-12-02 manuscript/revision2/response_to_reviews.tex
@@
 \maketitle
+
 \textbf{Reviewer comment}
@@
 matrices because all lead to the same fit to the effect-size data. The
-mash algorithm doesn\textquotesingle t give the full posterior
-distribution on the loadings so you can\textquotesingle t check for
-posterior correlations there. It seems likely to me that the
-hypothesis-driven and data-driven covariance matrices are somewhat
-correlated here, and the correlations may differ between green-up and
-flowering because of the better correlation between the environmental
-metrics and flowering than green-up. In the mash paper, they used
-cross-validation with the likelihood in the testing set as the
-evaluation metric to compare models. It might be safer to try dropping
-specific covariance matrices and comparing the model performance in a
-held-out testing set to evaluate the importance of the
+mash algorithm doesn't give the full posterior distribution on the
+loadings so you can't check for posterior correlations there. It seems
+likely to me that the hypothesis-driven and data-driven covariance
+matrices are somewhat correlated here, and the correlations may differ
+between green-up and flowering because of the better correlation between
+the environmental metrics and flowering than green-up. In the mash
+paper, they used cross-validation with the likelihood in the testing set
+as the evaluation metric to compare models. It might be safer to try
+dropping specific covariance matrices and comparing the model
+performance in a held-out testing set to evaluate the importance of the
 hypothesis-derived covariance matrices.
@@
 
-most\_likely\_matrix = \textquotesingle\textquotesingle{}
+most\_likely\_matrix = '\,'
 
@@
 algorithm in the presence of correlated covariance matrices, we
-implemented a "greedy" mash algorithm. Specifically, given n covariance
-matrices for each phenotype and corresponding effect sizes and standard
-errors across the k contexts of interest, we fit a set of mash models
-with each of the n matrices separately. We then identify the model with
-the maximum log likelihood estimates and select the corresponding
-matrix, leaving n-1 covariance matrices. We then fit a new set of mash
-models where the selected covariance matrix is paired with each of the
-remaining n-1 covariance matrices. The most likely pair of covariance
-matrices are then selected and the process is repeated for all possible
-n-2 matrix triplets. This process is repeated until one of two stop
-conditions are met: (i) all covariance matrices have been added to the
-mash model, resulting in a final model with all of the original n
+implemented a ``greedy'' mash algorithm. Specifically, given n
+covariance matrices for each phenotype and corresponding effect sizes
+and standard errors across the k contexts of interest, we fit a set of
+mash models with each of the n matrices separately. We then identify the
+model with the maximum log likelihood estimates and select the
+corresponding matrix, leaving n-1 covariance matrices. We then fit a new
+set of mash models where the selected covariance matrix is paired with
+each of the remaining n-1 covariance matrices. The most likely pair of
+covariance matrices are then selected and the process is repeated for
+all possible n-2 matrix triplets. This process is repeated until one of
+two stop conditions are met: (i) all covariance matrices have been added
+to the mash model, resulting in a final model with all of the original n
 covariance matrices are included or (ii) a likelihood ratio test with
@@
 The result is a set of log likelihood estimates for each model and the
-stepwise \textquotesingle path\textquotesingle{} of covariance matrices
-that best explain the observed effect sizes and standard errors observed
-for the phenotype of interest.~
+stepwise `path' of covariance matrices that best explain the observed
+effect sizes and standard errors observed for the phenotype of
+interest.~
 

### 2024-05-09 manuscript/revision2/response_to_reviews/response_to_reviews.qmd
manuscript: add response to reviews folder within manuscript/revision2/ and move intermediate files to intermediate_files folder

---
author:
  - name: Alice MacQueen* and Tom Juenger
    affiliation:
      - name: The University of Texas at Austin
        department: Department of Integrative Biology
        address: 2415 Speedway
        city: Austin, TX
        postal-code: 78712
address:
  - Editors
  - Proceedings of the National Academy of Sciences
subject: manuscript revision
subject-title: "Subject"
opening: "Dear Editors,"
closing: "Best,"
format: letter-pdf
---

Thank you for considering this work for publication in *PNAS*. Our manuscript, "Diverse Genotype-by-Weather Interactions in Switchgrass" has undergone substantial revision from our original submission in 2021. This timeframe is long - Dr. MacQueen took a new position in 2022. They continued working with the group as we developed a new algorithm to address major reviewer concerns and significantly altered the manuscript. Because of these alterations, the response to reviewers does not have track changes for each point, though we do point to line numbers for the associated revisions in our response.

We were pleased by the useful feedback given by the reviewers of our original submission. We have addressed all the editorial issues suggested by the two reviewers, and included in this letter are point-by-point responses to these issues.

The previous reviewers brought up two major concerns:

> Reviewer #1:\
> \
> Suitable Quality?: No\
> Sufficient General Interest?: No\
> Conclusions Justified?: Yes\
> Clearly Written?: No\
> Procedures Described?: Yes\
> \
> Comments on Significance Statement:\
> \
> Conclusions are very specific for the crop/population and traits. General conclusions are lacking.\
> \
> Comments:\
> \
> The paper by MacQueen et al. presents results of a study on two switchgrass populations grown in 8 garden experiments across a rane of latitudes. Traits are green-up date and flowering time. The data show strong GxE, effects of QTL changing in magnitude and sign. Results are partly validated in an independent segregating population.\
> \
> The topic is timely and highly relevant. A better understanding of GxE will be valuable for plant genetics, evolution and breeding. The data are complex and the authors have succeeded in making them accessible to the reader.\
> \
> \
> Relevant literature\
> The introduction and discussion are focused on findings in Arabidopsis and switchgrass. An excellent study from Arabidopsis with high relevance for this paper is\
> Fournier-Level et al. 2016 <https://doi.org/10.1073/pnas.1517456113>\
> \
> Major findings from other crop species have been ignored. To mention just two references:\
> \
> Bustos-Korts et al. 2019 The Plant Journal <https://doi.org/10.1111/tpj.14414>\
> Millet et al. 2016 Plant Physiology <https://doi.org/10.1104/pp.16.00621> and\
> There are many papers on crop modelling (eg work by Mark Cooper and others in maize) and the authors might want to consider linking their results to crop modelling to demonstrate the relevance of the results of this study

Comments on additional findings from other crop species

> Results\
> \
> There is a large overlap of data and findings with reference 32 Lovell et al. 2021 Nature. It has not become clear which additional insights can be gained from this study. One reason might be that objectives of the study have remained rather vague as well as the generic conclusions that can be drawn from the results.

We edited the Introduction substantially to differentiate this study from Lovell et al 2021.

> In the introduction the aim is given as „we test if these populations differ in their phenological adaptation and hence their phenological GxE". Given prior knowledge such as the origin of the populations and earlier results (Lovell et al.) what is expected?

We edit the Introduction to more clearly articulate the aims of this manuscript. Our key expectation is that different genetic subpopulations and genomic regions have likely evolved distinct patterns of GxE. Thus, we aim to identify the kinds of GxE present for each subpopulation and each trait.

> The results for these two populations did not show a general pattern leading to conclusions on the environmental cues. Associations trait/cue changed across populations, traits, gardens, analysis etc. , it was difficult to see the big picture.

We agree that we do not identify one general pattern or one environmental cue affecting any set of populations or phenological trait. Rather, we aim to identify the kinds of GxE present genome-wide and elaborate when these kinds of GxE differ (Figure 2), and demonstrate that individual loci have different types of GxE across environmental regions and subpopulations (Figure 3).

> The authors claim that the environmental cues in the hypothesis based models improved model fit. What inference is possible from this result? Unless results are validated in independent data or with cross validation the predictive power of the cues cannot be judged.

We edit the Results to a) more clearly articulate this claim and b) add an algorithm to more rigorously demonstrate this claim given the data that we have. We agree that given the data we have, the predictive power of these cues cannot be judged. *may need to look at QTL GxE to address this*. *It differs from our rewritten manuscript aims*.

> Results from the SNPs "mash model of Midwest green-up fell on a covariance matrix of average temperature in the 10 days prior to Midwest green-up". These results should be linked to the phenotypic GxE analysis.

We edit the results to directly compare the GxWeather matrices and the phenotypic correlations in Figure 1B. (lines 406-409; lines 417-420).

> Flowering posterior weights were higher than green-up weights. What does that mean? How does that link to the different heritabilities and levels of GxE?

We believe this comment refers to there being higher loadings on the GxWeather matrices for flowering date than there was for the date of the start of vegetative growth in the original paper (original Figure 2). Our revised analyses uses a selection process for canonical and GxWeather covariance matrices to select only the subset that significantly improves the model log-likelihood. Thus, the covariance matrices included in each model have changed in this revision, as have the relative loadings on GxWeather and canonical covariance matrices. *what do we see now?*

> What can you conclude from the pattern of antagonistic pleiotropy between Texas and Northern gardens that has not been seen in the GxE analysis?

No conclusions about individual loci effects can be drawn from the GxE analysis in Figure 2, as this plot does not show how shrinkage changes the jointly re-estimated SNP effect sizes. The GxE analysis in Figure 2 shows the types of GxE and GxWeather present genome-wide, from loadings on individual loci, but does not show the effects of these loci on traits. Rather, it is an overall characterization of GxE across all eight gardens.

The GxE analysis in Figure 3 shows the re-estimated effects of these loci on traits as contrasts between individual pairs of gardens. From Figure 3, you can conclude that many loci have effects with antagonistic pleiotropy at the site level.

> Discussion\
> "we must understand the current patterns of trait covariation across environments, .....". What is the understanding from your data? Are there any general patterns or do we need to assess them individually for different genetic material, environments, traits? It looks like the latter so what are the consequences?

We think that these patterns need to be assessed individually for different genetic material, environments, and traits (lines 575-578); thus, we develop an approach to specify multiple environmental cues and compete them to explain patterns of genetic effects (lines 591-595): we use a greedy algorithm to use with mash to select covariance patterns, and use mash to flexibly identify these patterns across populations and environments. (lines \_\_).

> "this is the first experimental work using QTL mapping and GWAS across ....". I disagree, see references above and others.

Edited to clarify our claim. NB: We do cite the first experimental work using both QTL mapping and GWAS (Brachi et al), and there are many others; we meant using these approaches to map GxE.

> "Gulf and Midwest subpopulations have two distinct photoperiod-related flowering responses....". This has been shown for other species, eg Unterseer et al. 2016 Genome Biology

*Try and add this reference in the introduction*

> Data/Methods\
> \
> Is one year data enough to make conclusions about the effect of environmental cues? It is well known that GxYear is more pronounced than GxLocation.

We agree that this would be an interesting topic for future work. Our aim was to demonstrate the value of this approach for mapping GxE, and we believe one year of data sufficient to do this.

> The reader should be able to understand what the hypothesis in the hypothesis-driven models is without looking at the supplement.

Agreed. We add Table 1 to better explain the covariance matrices we select from using the greedy mash algorithm. We include the matrices selected by this algorithm as visualizations in Figure 2.

> Reviewer #2:\
> \
> Suitable Quality?: No\
> Sufficient General Interest?: Yes\
> Conclusions Justified?: No\
> Clearly Written?: No\
> Procedures Described?: No\
> \
> Comments:\
> \
> The authors present a massive dataset on phenological variation among switchgrass cultivars across 8 common gardens. They use this dataset to describe the genetic architecture of gene-environment interactions in two phenological traits: the timing of green-up and flowering. They find that the genetic architecture of these traits varies considerably among two populations of switchgrass (called Gulf and Midwest) and among the latitudinal cline in common gardens, and that some of this variation seems to be related to variation among gardens in key environmental cues including temperature and daylength. They also identify genomic regions associated with this variation and replicate these regions in a separate set of F2 populations.\
> \
> While previous publications from this experiment have been published, I believe this is the first to focus on these two phenological traits which are of clear importance in switchgrass. The overall questions that they target are also of great importance - understanding the genetic basis of gene-environment interactions, and identifying the environmental drivers of these phenological traits - and the analytical approach that they use is creative and leverages powerful statistical methods. Given this, this study has the potential to be an important case study in the field for identifying GxE loci. However there are a number of issues with the statistical approach and also with the conceptual framework that I think make the current results uninterpretable. Also, because the methods are relatively novel and complex, it would be very helpful to provide more intuition behind the analytical approaches and more access to the raw results so that others can understand the approach more fully.

> First a note: It would be helpful if the authors would provide a pdf with line-numbers and with a font that can be copied directly.

Edited. We apologize for this omission.

> Major issues:\
> \
> Conceptual issues:\
> \
> - I like the idea of using environmental measures near the time of phenological events to try to identify environmental drivers and compare the plasticity functions across environments. However the choices here do not make physiological sense, especially for flowering. The environmental indices focus on either the day of flowering, or the 1-2 weeks prior to flowering. However, the developmental transition to flowering likely occurs well before this interval (I'm not sure how long in switchgrass, but it's likely much earlier). Photoperiod and temperature cues driving this developmental transition are irrelevant once the developmental commitment has been made. The length of time before flowers emerge and open may be dependent on these environmental factors as well, but this is physiological, not related mechanistically to the daylength or temperature requirements for flowering. The intervals used may be more relevant to green-up, I'm not familiar with the developmental basis of this trait. But as is, I don't think these metrics are interpretable for what they are designed for.

We agree that the choice of environmental measure is an extremely important consideration.

### 2024-05-09 manuscript/revision2/intermediate_files/SI_Appendix.qmd
manuscript: add response to reviews folder within manuscript/revision2/ and move intermediate files to intermediate_files folder

--- 2023-12-02 manuscript/revision2/SI_Appendix.qmd
@@
 ---
+
+## Manuscript Length {.unnumbered}
+
+A standard 6-page article is approximately 4,000 words, 50 references, and 4 medium-size graphical elements (i.e., figures and tables).
+The preferred length of articles remains at 6 pages, but PNAS will allow articles up to a maximum of 12 pages.
+
+### Digital Figures {.unnumbered}
+
+EPS, high-resolution PDF, and PowerPoint are preferred formats for figures that will be used in the main manuscript.
+Authors may submit PRC or U3D files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.
+Color images must be in RGB (red, green, blue) mode.
+Include the font files for any text.
+
+Images must be provided at final size, preferably 1 column width (8.7cm).
+Figures wider than 1 column should be sized to 11.4cm or 17.8cm wide.
+Numbers, letters, and symbols should be no smaller than 6 points (2mm) and no larger than 12 points (6mm) after reduction and must be consistent.
+
+Figures and tables should be labelled and referenced in the standard way using the `\label{}` and `\ref{}` commands.
+
+To insert a figure wider than one column, please use the `figure*` environment.
+Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide.
+Use `SCfigure*` environment for a wide figure with side legends, such as @fig-side.
+
+In Quarto, we can do these by setting the `fig-env` command to `figure*` or `SCfigure*`
+
+```{r}
+#| label: fig-side
+#| fig-cap: "This legend would be placed at the side of the figure, rather than below it."
+#| fig-env: "SCfigure*"
+#| echo: false
+# tell it the options as comments with a | and a space, as above.
+# set echo: false to avoid printing this text
+knitr::include_graphics('figs/frog.pdf')
+```
+
+## Supporting Information Appendix (SI)
+
+Authors should submit SI as a single separate SI Appendix PDF file, combining all text, figures, tables, movie legends, and SI references.
+SI will be published as provided by the authors; it will not be edited or composed.
+Additional details can be found in the [PNAS Author Center](https://www.pnas.org/authors/submitting-your-manuscript#manuscript-formatting-guidelines).
+The PNAS Overleaf SI template can be found [here](https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd).
+Refer to the SI Appendix in the manuscript at an appropriate point in the text.
+Number supporting figures and tables starting with S1, S2, etc.Authors who place detailed materials and methods in an SI Appendix must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the SI methods.
+If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.
+
+## Data Archival {.unnumbered}
+
+PNAS must be able to archive the data essential to a published article.
+Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the [Information for Authors](https://www.pnas.org/author-center/editorial-and-journal-policies#materials-and-data-availability), is acceptable.
 
@@
 
-+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
++-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
 | Phenotype | Name                            | Description                                                                                                                          |
-+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
++-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
 | green-up  | GDD, **n**d                     | Correlations in cumulative growing degree days for the five, ten, or eighteen days prior to green-up.                                |
 |           |                                 |                                                                                                                                      |
-|           | where \'**n**\' is in {5, 10, 18} |                                                                                                                                      |
-+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
+|           | where '**n**' is in {5, 10, 18} |                                                                                                                                      |
++-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
 | green-up  | Temp Ave, **n**d                | Correlations in average temperature for the five, ten, or eighteen days prior to green-up.                                           |
 |           |                                 |                                                                                                                                      |
-|           | where \'**n**\' is in {5, 10, 18} |                                                                                                                                      |
-+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
+|           | where '**n**' is in {5, 10, 18} |                                                                                                                                      |
++-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
 | flowering | 50% green-up                    | Correlations in green-up.                                                                                                            |
-+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
++-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
 | flowering | daylength                       | Correlations in day length on the day of flowering.                                                                                  |
-+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
++-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
 | flowering | daylength change (s)            | Correlations in the change in daylength (in seconds) on the day of flowering.                                                        |
-+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
++-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
 | flowering | GDD, GR to FL                   | Correlations in cumulative GDD between green-up and flowering.                                                                       |
-+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
++-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
 | flowering | Rainfall, **n**d sum            | Correlations in cumulative rainfall in the one, three, or five days before flowering, or in the days between green-up and flowering. |
 |           |                                 |                                                                                                                                      |
-|           | where \'**n**\' is in {1, 3, 5}   |                                                                                                                                      |
-+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
+|           | where '**n**' is in {1, 3, 5}   |                                                                                                                                      |
++-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
 
@@
 
+```         
 We scored green-up and flowering by day of the year, then linked these dates to multiple environmental factors (Table S1).
@@
 
+```
+
 **Section S2.** Hypothesis-based covariance matrix definitions for mash.
 
+```         
 We generated hypothesis-based covariance matrices derived from correlations in environmental cues in the green-up or flowering date windows for the three subpopulations (Table S1).
@@
 Alternatively, SNPs with high mixture proportions on particular environmental covariance matrices and large Bayes factors, which summarize the overall significance of a non-zero effect, represent small genomic intervals with strong evidence for a phenotypic effect correlated with an environmental driver.
+```
 
@@
 
+```         
 In the diversity panel, we determined narrow-sense heritabilities (h2) for green-up and flowering dates at single gardens using genomic relationship matrices calculated using the van Raden method (VanRaden et al.
@@
 We used ASReml (VSN International) to specify mixed models of the form:
+```
 
@@
 
+```         
 These models were run for each of the eight gardens, and across all gardens by adding an additional environmental effect of site without an interaction term.
 This resulted in 54 models: 3 sets of populations (the Gulf, Midwest, and Both subpopulations) for 9 garden sets (all eight gardens separately, and all eight gardens together) and two phenotypes (green-up date and flowering date).
+```
 
@@
 
+```         
 In the diversity panel, to evaluate our environmental cues as genetic triggers of flowering, we defined green-up and flowering for individual genotypes as functions of seven and eight environmental cues, respectively, then used variance components analysis to partition variance attributed to genetic effects (G), genotype by environment interactions (GxE), environmental effects (E), and error for these phenology-related traits using linear mixed models.
 We used mixed models of the form:
+```
 
@@
 
+```         
 These models were run for each phenotype as a function of environmental cues.
 This resulted in 135 models: three sets of populations (the Gulf, Midwest, and Both subpopulations) for three garden sets (All, North, and Texas gardens) and 15 phenotypes (seven green-up functions, and eight flowering functions).
+```
 
@@
 
+```         
 To evaluate the prevalence and kinds of covariance patterns of SNP effects across our eight common gardens, we used multivariate adaptive shrinkage (mash) on SNP effect estimates from the diversity panel (Urbut et al.
@@
 We used the \'pvdiv_standard_run()\' function of the switchgrassGWAS R package (https://github.com/Alice-MacQueen/switchgrassGWAS) to conduct GWAS on genotypes from the Midwest, Gulf, and Both subpopulations, used the \'pvdiv_bigsnp2mashr()\' function to convert univariate GWAS output to the matrices of effects and standard errors needed for mash, then used the \'mash_standard_run()\' function to conduct mash, again specifying distinct sets of hypothesis-based covariance matrices for green-up and flowering, but the same set of matrices for all genetic subgroups.
+```
 
@@
 
+```         
 To confirm candidate genomic regions and patterns of allelic effects found in the diversity panel, we analyzed flowering in an outbred pseudo-F2 cross between four individuals, two Midwest and two Gulf individuals.
@@
 The estimated QTL effect was obtained using \'scan1coef\' function in R/qtl2.
+```

### 2024-05-09 manuscript/revision2/intermediate_files/SI_Appendix.tex
manuscript: add response to reviews folder within manuscript/revision2/ and move intermediate files to intermediate_files folder

--- 2024-05-09 manuscript/revision2/intermediate_files/SI_Appendix.qmd
@@
----
-title: "SI Appendix"
-format: pdf
-editor: visual
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+%
+\documentclass[
+  letterpaper,
+  DIV=11,
+  numbers=noendperiod]{scrartcl}
 
-## Manuscript Length {.unnumbered}
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-A standard 6-page article is approximately 4,000 words, 50 references, and 4 medium-size graphical elements (i.e., figures and tables).
-The preferred length of articles remains at 6 pages, but PNAS will allow articles up to a maximum of 12 pages.
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 
-### Digital Figures {.unnumbered}
+\KOMAoption{captions}{tableheading}
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\AtBeginDocument{%
+\ifdefined\contentsname
+  \renewcommand*\contentsname{Table of contents}
+\else
+  \newcommand\contentsname{Table of contents}
+\fi
+\ifdefined\listfigurename
+  \renewcommand*\listfigurename{List of Figures}
+\else
+  \newcommand\listfigurename{List of Figures}
+\fi
+\ifdefined\listtablename
+  \renewcommand*\listtablename{List of Tables}
+\else
+  \newcommand\listtablename{List of Tables}
+\fi
+\ifdefined\figurename
+  \renewcommand*\figurename{Figure}
+\else
+  \newcommand\figurename{Figure}
+\fi
+\ifdefined\tablename
+  \renewcommand*\tablename{Table}
+\else
+  \newcommand\tablename{Table}
+\fi
+}
+\@ifpackageloaded{float}{}{\usepackage{float}}
+\floatstyle{ruled}
+\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
+\floatname{codelisting}{Listing}
+\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\makeatother
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\usepackage{bookmark}
 
-EPS, high-resolution PDF, and PowerPoint are preferred formats for figures that will be used in the main manuscript.
-Authors may submit PRC or U3D files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.
-Color images must be in RGB (red, green, blue) mode.
-Include the font files for any text.
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={SI Appendix},
+  colorlinks=true,
+  linkcolor={blue},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={Blue},
+  pdfcreator={LaTeX via pandoc}}
 
-Images must be provided at final size, preferably 1 column width (8.7cm).
-Figures wider than 1 column should be sized to 11.4cm or 17.8cm wide.
-Numbers, letters, and symbols should be no smaller than 6 points (2mm) and no larger than 12 points (6mm) after reduction and must be consistent.
+\title{SI Appendix}
+\author{}
+\date{}
 
-Figures and tables should be labelled and referenced in the standard way using the `\label{}` and `\ref{}` commands.
+\begin{document}
+\maketitle
 
-To insert a figure wider than one column, please use the `figure*` environment.
-Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide.
-Use `SCfigure*` environment for a wide figure with side legends, such as @fig-side.
+\subsection*{Manuscript Length}\label{manuscript-length}
+\addcontentsline{toc}{subsection}{Manuscript Length}
 
-In Quarto, we can do these by setting the `fig-env` command to `figure*` or `SCfigure*`
+A standard 6-page article is approximately 4,000 words, 50 references,
+and 4 medium-size graphical elements (i.e., figures and tables). The
+preferred length of articles remains at 6 pages, but PNAS will allow
+articles up to a maximum of 12 pages.
 
-```{r}
-#| label: fig-side
-#| fig-cap: "This legend would be placed at the side of the figure, rather than below it."
-#| fig-env: "SCfigure*"
-#| echo: false
-# tell it the options as comments with a | and a space, as above.
-# set echo: false to avoid printing this text
-knitr::include_graphics('figs/frog.pdf')
-```
+\subsubsection*{Digital Figures}\label{digital-figures}
+\addcontentsline{toc}{subsubsection}{Digital Figures}
 
-## Supporting Information Appendix (SI)
+EPS, high-resolution PDF, and PowerPoint are preferred formats for
+figures that will be used in the main manuscript. Authors may submit PRC
+or U3D files for 3D images; these must be accompanied by 2D
+representations in TIFF, EPS, or high-resolution PDF format. Color
+images must be in RGB (red, green, blue) mode. Include the font files
+for any text.
 
-Authors should submit SI as a single separate SI Appendix PDF file, combining all text, figures, tables, movie legends, and SI references.
-SI will be published as provided by the authors; it will not be edited or composed.
-Additional details can be found in the [PNAS Author Center](https://www.pnas.org/authors/submitting-your-manuscript#manuscript-formatting-guidelines).
-The PNAS Overleaf SI template can be found [here](https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd).
-Refer to the SI Appendix in the manuscript at an appropriate point in the text.
-Number supporting figures and tables starting with S1, S2, etc.Authors who place detailed materials and methods in an SI Appendix must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the SI methods.
-If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.
+Images must be provided at final size, preferably 1 column width
+(8.7cm). Figures wider than 1 column should be sized to 11.4cm or 17.8cm
+wide. Numbers, letters, and symbols should be no smaller than 6 points
+(2mm) and no larger than 12 points (6mm) after reduction and must be
+consistent.
 
-## Data Archival {.unnumbered}
+Figures and tables should be labelled and referenced in the standard way
+using the \texttt{\textbackslash{}label\{\}} and
+\texttt{\textbackslash{}ref\{\}} commands.
+
+To insert a figure wider than one column, please use the
+\texttt{figure*} environment. Figures wider than one column should be
+sized to 11.4 cm or 17.8 cm wide. Use \texttt{SCfigure*} environment for
+a wide figure with side legends, such as Figure~\ref{fig-side}.
+
+In Quarto, we can do these by setting the \texttt{fig-env} command to
+\texttt{figure*} or \texttt{SCfigure*}
+
+\begin{SCfigure*}
+
+\centering{
+
+\includegraphics{figs/frog.pdf}
+
+}
+
+\caption{\label{fig-side}This legend would be placed at the side of the
+figure, rather than below it.}
+
+\end{SCfigure*}%
+
+\subsection{Supporting Information Appendix
+(SI)}\label{supporting-information-appendix-si}
+
+Authors should submit SI as a single separate SI Appendix PDF file,
+combining all text, figures, tables, movie legends, and SI references.
+SI will be published as provided by the authors; it will not be edited
+or composed. Additional details can be found in the
+\href{https://www.pnas.org/authors/submitting-your-manuscript\#manuscript-formatting-guidelines}{PNAS
+Author Center}. The PNAS Overleaf SI template can be found
+\href{https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd}{here}.
+Refer to the SI Appendix in the manuscript at an appropriate point in
+the text. Number supporting figures and tables starting with S1, S2,
+etc.Authors who place detailed materials and methods in an SI Appendix
+must provide sufficient detail in the main text methods to enable a
+reader to follow the logic of the procedures and results and also must
+reference the SI methods. If a paper is fundamentally a study of a new
+method or technique, then the methods must be described completely in
+the main text.
+
+\subsection*{Data Archival}\label{data-archival}
+\addcontentsline{toc}{subsection}{Data Archival}
 
 PNAS must be able to archive the data essential to a published article.
-Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the [Information for Authors](https://www.pnas.org/author-center/editorial-and-journal-policies#materials-and-data-availability), is acceptable.
+Where such archiving is not possible, deposition of data in public
+databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata,
+and others outlined in the
+\href{https://www.pnas.org/author-center/editorial-and-journal-policies\#materials-and-data-availability}{Information
+for Authors}, is acceptable.
 
-**SI Appendix**
+\textbf{SI Appendix}
 
-\
+\hfill\break
 
-**Dataset 1.** SNP-associated effects and standard errors for green-up estimated using the Gulf genetic subpopulation.
+\textbf{Dataset 1.} SNP-associated effects and standard errors for
+green-up estimated using the Gulf genetic subpopulation.
 
-**Dataset 2.** SNP-associated effects and standard errors for green-up estimated using the Midwest genetic subpopulation.
+\textbf{Dataset 2.} SNP-associated effects and standard errors for
+green-up estimated using the Midwest genetic subpopulation.
 
-**Dataset 3.** SNP-associated effects and standard errors for green-up estimated using Both genetic subpopulations.
+\textbf{Dataset 3.} SNP-associated effects and standard errors for
+green-up estimated using Both genetic subpopulations.
 
-**Dataset 4.** SNP-associated effects and standard errors for flowering estimated using the Gulf genetic subpopulation.
+\textbf{Dataset 4.} SNP-associated effects and standard errors for
+flowering estimated using the Gulf genetic subpopulation.
 
-**Dataset 5.** SNP-associated effects and standard errors for flowering estimated using the Midwest genetic subpopulation.
+\textbf{Dataset 5.} SNP-associated effects and standard errors for
+flowering estimated using the Midwest genetic subpopulation.
 
-**Dataset 6.** SNP-associated effects and standard errors for flowering estimated using Both genetic subpopulations.
+\textbf{Dataset 6.} SNP-associated effects and standard errors for
+flowering estimated using Both genetic subpopulations.
 
-**Dataset 7.** Genes within QTL that have functionally validated roles in flowering in rice.
+\textbf{Dataset 7.} Genes within QTL that have functionally validated
+roles in flowering in rice.
 
-**Dataset 8.** Overlap between quantitative trait loci for flowering measured using an outbred pseudo-F2 mapping population, and significant mash effect estimates (with a log10Bayes Factor \> 2, or in the 1% tail of most significant mash results) found using three overlapping diversity panels.
-Genes within 20kb with a functionally validated role in flowering in rice are  indicated.
+\textbf{Dataset 8.} Overlap between quantitative trait loci for
+flowering measured using an outbred pseudo-F2 mapping population, and
+significant mash effect estimates (with a log10Bayes Factor
+\textgreater{} 2, or in the 1\% tail of most significant mash results)
+found using three overlapping diversity panels. Genes within 20kb with a
+functionally validated role in flowering in rice are~ indicated.~
 
-\
+\hfill\break
 
-**Figure S1.** Variance components analysis of genetic (purple), genotype by environment (blue), environmental (green), and residual (yellow) terms in models of green-up time as functions of Julian date or weather, for the four northern common gardens (North), the three Texas common gardens (Texas), and for all eight common gardens (All).
-Dashed lines indicate the cumulative contribution of G and GxE for flowering as a function of day of the year. ![](https://lh7-us.googleusercontent.com/SPDH2Y9HThoK0pUZQxFuLPUDrIuXAcG73nUSk2ucYN3_CHXSG-c8G4b7Gq6crgo_1IapiGYFV66w_vhfyvxSIW8kPtTQA_tMeviTFamFdsXIPjIjTcaI0dwa7ZqpKN0P1RonX05vbxitm7vLYoYoZw)
+\textbf{Figure S1.} Variance components analysis of genetic (purple),
+genotype by environment (blue), environmental (green), and residual
+(yellow) terms in models of green-up time as functions of Julian date or
+weather, for the four northern common gardens (North), the three Texas
+common gardens (Texas), and for all eight common gardens (All). Dashed
+lines indicate the cumulative contribution of G and GxE for flowering as
+a function of day of the
+year.~\includegraphics{SI_Appendix_files/mediabag/SPDH2Y9HThoK0pUZQxFu.pdf}
 
-**Figure S2.** Effect patterns exhibited by the major data-driven matrices for (a,d,e,j,m,n) Gulf, (b,f,g,k,o,p) Midwest, and (c,h,i,l,q,r) Both subpopulations from Figure 2.
-(a-i) Green-up date.
-(j-r) Flowering date.
-(a-c,j-l) Single-garden effect representations (eigenvectors) of the DD_tPCA data-derived matrices.
-The percent variation explained of DD_tPCA by each eigenvector is shown on the y-axis.
-Common gardens are arranged in latitudinal order along the x-axis.
-In addition, the first eigenvector corresponds to DD_PCA_1, and second corresponds to DD_PCA_2.
-(d,f,h,m,o,q) The number of SNPs with significant effects in both conditions that exhibit antagonistic pleiotropy between that pair of conditions.
-(e,g,i,n,p,r) Same as (d,f,h,m,o,q) except for differential sensitivity.
-Scales for (d,e), (f,g), (h,i), (m,n), (o,p), and (q,r) are identical and located on the first panel in alphabetical order.![](https://lh7-us.googleusercontent.com/JV_8jSfymcCRhEQsEuXD2WzHAkM7bOTpEws9j3_H4SbmaTArViZOZQyWld6aMI-xs8DPYLSZBqpxogUokZwU36l5NMMlJgLy-dY8GFwACkYAntFMqAvq9b1C3BsXnbacDYOF3i4kRqXIW4ISTu2Xzg)
+\textbf{Figure S2.} Effect patterns exhibited by the major data-driven
+matrices for (a,d,e,j,m,n) Gulf, (b,f,g,k,o,p) Midwest, and
+(c,h,i,l,q,r) Both subpopulations from Figure 2. (a-i) Green-up date.
+(j-r) Flowering date. (a-c,j-l) Single-garden effect representations
+(eigenvectors) of the DD\_tPCA data-derived matrices. The percent
+variation explained of DD\_tPCA by each eigenvector is shown on the
+y-axis. Common gardens are arranged in latitudinal order along the
+x-axis. In addition, the first eigenvector corresponds to DD\_PCA\_1,
+and second corresponds to DD\_PCA\_2. (d,f,h,m,o,q) The number of SNPs
+with significant effects in both conditions that exhibit antagonistic
+pleiotropy between that pair of conditions. (e,g,i,n,p,r) Same as
+(d,f,h,m,o,q) except for differential sensitivity. Scales for (d,e),
+(f,g), (h,i), (m,n), (o,p), and (q,r) are identical and located on the
+first panel in alphabetical
+order.\includegraphics{SI_Appendix_files/mediabag/JV_8jSfymcCRhEQsEuXD.pdf}
 
-\
+\hfill\break
 
-\
-![](https://lh7-us.googleusercontent.com/h-CoStrONeN2vAZcGvcrGRLg1RRQiAhfNYCT9xn9YF5uwInX02rCepO4NU1bRq1ykboFZcOjr1PeEiZXsXRDJ6v51BAzYPWKQ2wd8YhVg2LJA7IMkL5wvEEa8jOw0UDgrZgVwiYZlHQWSqGxT_2q0g)
+\hfill\break
+\includegraphics{SI_Appendix_files/mediabag/h-CoStrONeN2vAZcGvcr.pdf}
 
-**Figure S3.** Comparison of mash SNP effects from a diversity panel and effects from an outbred pseudo-F2 cross.
-(**a**) QTL mapping for four weather-related functions of flowering, and for green-up date, as indicated by the colors: pink, growing degree days (GDD) from green-up date to flowering date, yellow, day length change on the flowering date, green, day length on flowering date; purple, flowering date; gray, green-up date.Dotted lines indicate permutation-based significance thresholds for each weather-related function.
-Stars indicate QTL with significant enrichment for SNPs in the 1% mash tail; B, G, and M indicate which subpopulation had enrichment: B -  both subpopulations, G - Gulf subpopulation, M -Midwest subpopulation.
-Rug plots show genomic locations of SNPs in the 1% mash tail for flowering date for each subpopulation.
-(**b**) Schematic comparison of genotypes used for QTL mapping and genotypes used in mash.
-(**c**) Number of mash runs enriched for SNPs in the 1% mash tail in the 23 QTL intervals from (**a**) (dotted red line), compared to 1000 sets of 23 random QTL intervals of the same size (histogram).**(d**) SNP effects estimated using mash for SNPs with the highest Bayes factor for the three QTL with the highest LOD scores from (**a**) and enrichment of SNPs in the 1% mash tail.
-SNP genomic position and genetic subpopulation are indicated above each plot.
-Common gardens are arranged in latitudinal order.
-(**e**) QTL effect estimates for 23 non-overlapping and partially overlapping QTLs.
-Colors indicate which of the four weather-related functions of flowering for which the QTL was mapped.
-An additional five QTLs completely overlapped the QTL displayed; in this case, the QTL with the highest lod score is shown.
-Solid vertical lines separate QTLs that are not overlapping.
-Stars indicate QTL with significant enrichment for SNPs in the 1% mash tail; B, G, and M indicate which subpopulation had enrichment, as in (**b**):B -  both subpopulations, G - Gulf subpopulation, M -Midwest subpopulation.
-Common gardens are arranged in latitudinal order.
+\textbf{Figure S3.} Comparison of mash SNP effects from a diversity
+panel and effects from an outbred pseudo-F2 cross. (\textbf{a}) QTL
+mapping for four weather-related functions of flowering, and for
+green-up date, as indicated by the colors: pink, growing degree days
+(GDD) from green-up date to flowering date, yellow, day length change on
+the flowering date, green, day length on flowering date; purple,
+flowering date; gray, green-up date.Dotted lines indicate
+permutation-based significance thresholds for each weather-related
+function. Stars indicate QTL with significant enrichment for SNPs in the
+1\% mash tail; B, G, and M indicate which subpopulation had enrichment:
+B -~ both subpopulations, G - Gulf subpopulation, M -Midwest
+subpopulation. Rug plots show genomic locations of SNPs in the 1\% mash
+tail for flowering date for each subpopulation. (\textbf{b}) Schematic
+comparison of genotypes used for QTL mapping and genotypes used in mash.
+(\textbf{c}) Number of mash runs enriched for SNPs in the 1\% mash tail
+in the 23 QTL intervals from (\textbf{a}) (dotted red line), compared to
+1000 sets of 23 random QTL intervals of the same size
+(histogram).\textbf{(d}) SNP effects estimated using mash for SNPs with
+the highest Bayes factor for the three QTL with the highest LOD scores
+from (\textbf{a}) and enrichment of SNPs in the 1\% mash tail. SNP
+genomic position and genetic subpopulation are indicated above each
+plot. Common gardens are arranged in latitudinal order. (\textbf{e}) QTL
+effect estimates for 23 non-overlapping and partially overlapping QTLs.
+Colors indicate which of the four weather-related functions of flowering
+for which the QTL was mapped. An additional five QTLs completely
+overlapped the QTL displayed; in this case, the QTL with the highest lod
+score is shown. Solid vertical lines separate QTLs that are not
+overlapping. Stars indicate QTL with significant enrichment for SNPs in
+the 1\% mash tail; B, G, and M indicate which subpopulation had
+enrichment, as in (\textbf{b}):B -~ both subpopulations, G - Gulf
+subpopulation, M -Midwest subpopulation. Common gardens are arranged in
+latitudinal order.~
 
-**Table S1.** Green-up and flowering dates phenotypes linked to environmental cues.
+\textbf{Table S1.} Green-up and flowering dates phenotypes linked to
+environmental cues.
 
-+-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
-| Phenotype | Name                            | Description                                                                                                                          |
-+-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
-| green-up  | GDD, **n**d                     | Correlations in cumulative growing degree days for the five, ten, or eighteen days prior to green-up.                                |
-|           |                                 |                                                                                                                                      |
-|           | where '**n**' is in {5, 10, 18} |                                                                                                                                      |
-+-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
-| green-up  | Temp Ave, **n**d                | Correlations in average temperature for the five, ten, or eighteen days prior to green-up.                                           |
-|           |                                 |                                                                                                                                      |
-|           | where '**n**' is in {5, 10, 18} |                                                                                                                                      |
-+-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
-| flowering | 50% green-up                    | Correlations in green-up.                                                                                                            |
-+-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
-| flowering | daylength                       | Correlations in day length on the day of flowering.                                                                                  |
-+-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
-| flowering | daylength change (s)            | Correlations in the change in daylength (in seconds) on the day of flowering.                                                        |
-+-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
-| flowering | GDD, GR to FL                   | Correlations in cumulative GDD between green-up and flowering.                                                                       |
-+-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
-| flowering | Rainfall, **n**d sum            | Correlations in cumulative rainfall in the one, three, or five days before flowering, or in the days between green-up and flowering. |
-|           |                                 |                                                                                                                                      |
-|           | where '**n**' is in {1, 3, 5}   |                                                                                                                                      |
-+-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0659}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1868}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.7418}}@{}}
+\toprule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Phenotype & Name & Description \\
+green-up & GDD, \textbf{n}d
 
-\
-\
+where `\textbf{n}' is in \{5, 10, 18\} & Correlations in cumulative
+growing degree days for the five, ten, or eighteen days prior to
+green-up. \\
+green-up & Temp Ave, \textbf{n}d
 
-**Section S1.** Weather-based environmental cues for green-up and flowering date.
+where `\textbf{n}' is in \{5, 10, 18\} & Correlations in average
+temperature for the five, ten, or eighteen days prior to green-up. \\
+flowering & 50\% green-up & Correlations in green-up. \\
+flowering & daylength & Correlations in day length on the day of
+flowering. \\
+flowering & daylength change (s) & Correlations in the change in
+daylength (in seconds) on the day of flowering. \\
+flowering & GDD, GR to FL & Correlations in cumulative GDD between
+green-up and flowering. \\
+flowering & Rainfall, \textbf{n}d sum
 
-\
+where `\textbf{n}' is in \{1, 3, 5\} & Correlations in cumulative
+rainfall in the one, three, or five days before flowering, or in the
+days between green-up and flowering. \\
+\end{longtable}
 
-```         
+\hfill\break
+\hfill\break
+
+\textbf{Section S1.} Weather-based environmental cues for green-up and
+flowering date.
+
+\hfill\break
+
+\begin{verbatim}
 We scored green-up and flowering by day of the year, then linked these dates to multiple environmental factors (Table S1).
@@
 
-```
+\end{verbatim}
 
-**Section S2.** Hypothesis-based covariance matrix definitions for mash.
+\textbf{Section S2.} Hypothesis-based covariance matrix definitions for
+mash.
 
-```         
+\begin{verbatim}
 We generated hypothesis-based covariance matrices derived from correlations in environmental cues in the green-up or flowering date windows for the three subpopulations (Table S1).
@@
 Alternatively, SNPs with high mixture proportions on particular environmental covariance matrices and large Bayes factors, which summarize the overall significance of a non-zero effect, represent small genomic intervals with strong evidence for a phenotypic effect correlated with an environmental driver.
-```
+\end{verbatim}
 
-\
+\hfill\break
 
-**Section S3.** Narrow-sense heritability
+\textbf{Section S3.} Narrow-sense heritability~
 
-```         
+\begin{verbatim}
 In the diversity panel, we determined narrow-sense heritabilities (h2) for green-up and flowering dates at single gardens using genomic relationship matrices calculated using the van Raden method (VanRaden et al.
@@
 We used ASReml (VSN International) to specify mixed models of the form:
-```
+\end{verbatim}
 
-**y** = 1 + Zu + e
+\textbf{y} = 1 + Zu + e
 
@@
 
-in which the vector **y** represents the flowering date or green-up date values for that garden, Z the design matrix for random effects, u the whole genome additive genetic effect, and e the residual.
-Matrix G is the whole genomic relationship matrix based on all SNPs retained for subpopulation-specific analyses.
-I is the rank-y identity matrix.
-Phenotypic variance (σp2) is σu2 + σe2.
-Narrow-sense heritability is then h2 = (σu2/σp2).
+in which the vector \textbf{y} represents the flowering date or green-up
+date values for that garden, Z the design matrix for random effects, u
+the whole genome additive genetic effect, and e the residual. Matrix G
+is the whole genomic relationship matrix based on all SNPs retained for
+subpopulation-specific analyses. I is the rank-y identity matrix.
+Phenotypic variance (σp2) is σu2 + σe2. Narrow-sense heritability is
+then h2 = (σu2/σp2).
 
-```         
+\begin{verbatim}
 These models were run for each of the eight gardens, and across all gardens by adding an additional environmental effect of site without an interaction term.
 This resulted in 54 models: 3 sets of populations (the Gulf, Midwest, and Both subpopulations) for 9 garden sets (all eight gardens separately, and all eight gardens together) and two phenotypes (green-up date and flowering date).
-```
+\end{verbatim}
 
-\
+\hfill\break
 
-**Section S4.** Variance components analysis
+\textbf{Section S4.} Variance components analysis
 
-```         
+\begin{verbatim}
 In the diversity panel, to evaluate our environmental cues as genetic triggers of flowering, we defined green-up and flowering for individual genotypes as functions of seven and eight environmental cues, respectively, then used variance components analysis to partition variance attributed to genetic effects (G), genotype by environment interactions (GxE), environmental effects (E), and error for these phenology-related traits using linear mixed models.
 We used mixed models of the form:
-```
+\end{verbatim}
 
-**y** = 1 + Zuu + Zll + Zulul + e
+\textbf{y} = 1 + Zuu + Zll + Zulul + e
 
@@
 
-in which the vector **y** represents the individual plant observations of a given trait and Z the design matrices associating trait observations with random factors.
-Random factors include u, the whole genome additive genetic (G) effect, l, the effect of each location (E), ul, the GxE effect, and e, the residual.
-Matrix G is the whole genomic relationship matrix.
-I are rank-y identity matrices.
+in which the vector \textbf{y} represents the individual plant
+observations of a given trait and Z the design matrices associating
+trait observations with random factors. Random factors include u, the
+whole genome additive genetic (G) effect, l, the effect of each location
+(E), ul, the GxE effect, and e, the residual. Matrix G is the whole
+genomic relationship matrix. I are rank-y identity matrices.~
 
-```         
+\begin{verbatim}
 These models were run for each phenotype as a function of environmental cues.
 This resulted in 135 models: three sets of populations (the Gulf, Midwest, and Both subpopulations) for three garden sets (All, North, and Texas gardens) and 15 phenotypes (seven green-up functions, and eight flowering functions).
-```
+\end{verbatim}
 
-\
+\hfill\break
 
-**Section S5.** Multivariate adaptive shrinkage
+\textbf{Section S5.} Multivariate adaptive shrinkage
 
-```         
+\begin{verbatim}
 To evaluate the prevalence and kinds of covariance patterns of SNP effects across our eight common gardens, we used multivariate adaptive shrinkage (mash) on SNP effect estimates from the diversity panel (Urbut et al.
@@
 We used the \'pvdiv_standard_run()\' function of the switchgrassGWAS R package (https://github.com/Alice-MacQueen/switchgrassGWAS) to conduct GWAS on genotypes from the Midwest, Gulf, and Both subpopulations, used the \'pvdiv_bigsnp2mashr()\' function to convert univariate GWAS output to the matrices of effects and standard errors needed for mash, then used the \'mash_standard_run()\' function to conduct mash, again specifying distinct sets of hypothesis-based covariance matrices for green-up and flowering, but the same set of matrices for all genetic subgroups.
-```
+\end{verbatim}
 
-\
+\hfill\break
 
-**Section S6.** Outbred pseudo-F2 mapping population and Quantitative Trait Locus mapping
+\textbf{Section S6.} Outbred pseudo-F2 mapping population and
+Quantitative Trait Locus mapping
 
-```         
+\begin{verbatim}
 To confirm candidate genomic regions and patterns of allelic effects found in the diversity panel, we analyzed flowering in an outbred pseudo-F2 cross between four individuals, two Midwest and two Gulf individuals.
@@
 The estimated QTL effect was obtained using \'scan1coef\' function in R/qtl2.
-```
+\end{verbatim}
+
+\end{document}

### 2024-05-09 manuscript/revision2/revision2.qmd
manuscript: add response to reviews folder within manuscript/revision2/ and move intermediate files to intermediate_files folder

@@
 title: Diverse Genotype-by-Weather Interactions in Switchgrass
+pdf-engine: pdflatex
 format:
@@
 
-\firstpage[15]{2}
+\firstpage[17]{2}
 
@@
 At the Oklahoma common garden, located near the natural range limits of both the Gulf and the Midwest subpopulations, Gulf and Midwest vegetative growth occurred over the same period, and Gulf flowering occurred after Midwestern flowering (@fig-map A).
-These patterns led to strong negative phenotypic correlations for the timing of vegetative growth between the North and Texas gardens, particularly in the Gulf and across all individuals (hereafter, 'Both' subpopulations), and contributed to positive phenotypic correlations for the timing of flowering that had larger magnitudes at more northern gardens (@fig-map B).
+These patterns led to strong negative phenotypic correlations for the timing of the start of vegetative growth between the North and Texas gardens, particularly in the Gulf and across all individuals from the Gulf and Midwest subpopulations (hereafter, 'Both' subpopulations), and contributed to positive phenotypic correlations for the timing of flowering that had larger magnitudes at more northern gardens (@fig-map B).
 
@@
 For example, the more similar that the amount of rainfall prior to flowering is for genotypes A and B between gardens in MI and MO, the nearer these covariances will be to one.
-If rainfalls are uncorrelated, covariance will be near zero; if there is rank changing (e.g.
+If rainfall amounts are uncorrelated, covariance will be near zero; if there is rank changing (e.g.
 high rainfall before flowering of A at MO but not MI, and high rainfall before flowering for B at MI but not MO), this covariance will be near negative one.
@@
 
-Five GxWeather covariance structures were selected by the greedy algorithm, zero to two per subpopulation: two for the start of vegetative growth, and three for flowering date.
+We used a greedy algorithm to select covariance matrices from each model's set that significantly improved the model likelihood.
+Five GxWeather covariance structures were selected, zero to two per subpopulation: two for the start of vegetative growth, and three for flowering date.
 Of these five, three had mass on them in mash models of the strong effects (@fig-covar A-B).
@@
 In total, 65% of the posterior weight of strong SNP effects in the mash model of Gulf vegetative growth fell on a covariance matrix constructed using the covariance of daylength 14 days prior to the date of vegetative growth.
-The covariance matrix for this weather cue was very similar to the pattern of phenotypic correlation for the timing of vegetative growth in the Gulf subpopulation (@fig-map B; @fig-covar A).
+The covariance matrix for this weather cue was very similar to the observed pattern of phenotypic correlation for the timing of vegetative growth in the Gulf subpopulation (@fig-map B; @fig-covar A).
 Mash models of the timing of vegetative growth in the Midwest subpopulation and Both subpopulations did not include this GxWeather covariance; the Midwest had no weight on any GxWeather covariance, while Both subpopulations had non-zero weights on two additional GxWeather covariance types, average temperature one day prior to green-up, and the day length change in seconds in the day prior to green-up (@fig-covar B).
 The average temperature covariance matrix had negative covariances between Texas and North gardens, though not as strong as the negative phenotypic correlations seen in Both subpopulations (@fig-map B; @fig-covar A).
-Only the Gulf subpopulation and Both subpopulations had mass on any GxWeather covariance matrices (@fig-covar C).
+Only the Gulf subpopulation and Both subpopulations had mass on any GxWeather covariance matrices (@fig-covar C) for vegetative growth.
 
@@
 22.6% of SNP effects on flowering in the Midwest subpopulation covaried with day length change in the two days prior to flowering (@fig-covar D), which had negative covariances between Texas and North gardens.
-Neither covariance matrix was selected as significantly improving the log-likelihood of the *mash* model of Both subpopulations, and no GxWeather covariance matrices had a non-zero mass in models of Both subpopulations (@fig-covar E).
-In five of the six mash models of strong effects, the GxWeather covariance matrices captured a minority of the posterior weights of the strong effects (@fig-covar C,E); the majority of this mass was on various canonical covariance matrices.
+Neither of these covariance matrices significantly improved the log-likelihood of the *mash* model of Both subpopulations, and no GxWeather covariance matrices had non-zero mass in models of Both subpopulations (@fig-covar E).
+In five of the six mash models of strong effects, the GxWeather covariance matrices captured a minority of the posterior weights of the strong effects (@fig-covar C,E); in these five models, the majority of this mass was on various canonical covariance matrices.
 These matrices included simple heterozygosity, with intermediate, positive covariances between all gardens, and garden-specific effects present only at one garden.
@@
 We next characterized the pairwise patterns of effects where we were confident in the sign of the effect at both gardens.
-If the overall pattern of phenotypic expression differs between a pair of gardens, this difference may be comprised of SNP effects that differ in sign between these gardens (effects with antagonistic pleiotropy), SNP effects that differ in magnitude - that are large in one garden or region and smaller in others, and SNP effects that are indistinguishable in the two gardens (@fig-effects A).
+If the overall pattern of phenotypic expression differs between a pair of gardens, this difference may be comprised of SNP effects that differ in sign between these gardens (effects with antagonistic pleiotropy), SNP effects that differ in magnitude (effects that are large in one garden or region and smaller in others), and SNP effects that are indistinguishable in the two gardens (@fig-effects A).
 We used the local false sign rate (lfsr), an analogue of the lfdr that establishes confidence in the effect sign, not the effect's difference from zero, to determine significance.
@@
 The GxWeather photoperiod and cumulative rainfall cues we identify here are functions of the genotypes measured and capture only a minority of SNP effects on flowering.
-We know still less about the overwintering parameters that cause variation in the start of vegetative growth.
 We could only assign SNP effects to a GxWeather covariance structures in four of the six phenotype & genetic subpopulations we modeled.
+It's likely we did not include some GxWeather important to these phenological cues - for example, overwintering parameters that might cause variation in the start of vegetative growth.
 More generally, it is difficult to predict the time scales over which individuals may integrate environmental cues, particularly in perennial species which may integrate these cues over longer time scales.
-If this integration time itself varies between individuals, we cannot select a covariance structure that reflects this, though these structures would likely be highly correlated with GxWeather structures we did include.
+If this integration time itself varies between individuals, the covariance structures we modeled cannot reflect this, though these structures would likely be highly correlated with GxWeather structures we did include.
 Our approach offers an opportunity to specify multiple environmental cues and compete them to explain patterns of genetic effects, allowing us to detect how important these cues are genome-wide, and how strongly each cue influences each SNP.
@@
 b) The fraction of effects with each GxE type for the start of vegetative growth (Greenup Date) and reproductive growth (Flowering Date), within and between two genetic subpopulations.
-Common gardens are grouped by the larger region they came from: North gardens are within the natural range of the Midwest subpopulation, while Texas gardens are within the natural range of the Gulf subpopulation.
-](Figure_3_GxE.png){#fig-effects}
+Common gardens are grouped by the larger region they came from: North gardens are within the natural range of the Midwest subpopulation, while Texas gardens are within the natural range of the Gulf subpopulation.](Figure_3_GxE.png){#fig-effects}
 
@@
 a) QTL mapping for the start of vegetative growth (Greenup Date), and three weather-related functions of greenup date.
-b) QTL mapping for the start of reproductive growth (Flowering Date), and two weather-related functions of flowering date.
-](Figure_4_QTL_Overlaps.png){#fig-qtl}
+b) QTL mapping for the start of reproductive growth (Flowering Date), and two weather-related functions of flowering date.](Figure_4_QTL_Overlaps.png){#fig-qtl}
 
@@
 To do this, we first conducted univarate genome-wide association (GWAS) on site-specific best linear unbiased predictors (BLUPs) for the timing of vegetative growth and flowering, using the switchgrassGWAS package and the methods in [@lovell2021].
-We used the effect estimates for single nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect estimates for both traits at each garden.
+We used the effect estimates for single nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect estimates for the two traits at each garden.
 We then jointly modeled genetic effect estimates, and the mixture of ways these effects might covary, across all eight common gardens, using a two-part procedure.

### 2024-05-09 manuscript/revision2/revision2.tex
manuscript: add response to reviews folder within manuscript/revision2/ and move intermediate files to intermediate_files folder

@@
 \ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
-\firstpage[15]{2}
+\firstpage[17]{2}
 
@@
 These patterns led to strong negative phenotypic correlations for the
-timing of vegetative growth between the North and Texas gardens,
-particularly in the Gulf and across all individuals (hereafter, `Both'
-subpopulations), and contributed to positive phenotypic correlations for
-the timing of flowering that had larger magnitudes at more northern
-gardens (Figure~\ref{fig-map} B).
+timing of the start of vegetative growth between the North and Texas
+gardens, particularly in the Gulf and across all individuals from the
+Gulf and Midwest subpopulations (hereafter, `Both' subpopulations), and
+contributed to positive phenotypic correlations for the timing of
+flowering that had larger magnitudes at more northern gardens
+(Figure~\ref{fig-map} B).
 
@@
 genotypes A and B between gardens in MI and MO, the nearer these
-covariances will be to one. If rainfalls are uncorrelated, covariance
-will be near zero; if there is rank changing (e.g.~high rainfall before
-flowering of A at MO but not MI, and high rainfall before flowering for
-B at MI but not MO), this covariance will be near negative one. We
-specified multiple covariance structures for the canonical, data-driven,
-and GxWeather categories, each of which represented a specific pattern
-of covariance in genetic effect size estimates that could be included in
-the model. The phenotypic correlations for the timing of vegetative
-growth had moderate negative correlations between the Texas and North
-gardens, particularly in the Gulf subpopulation and Both subpopulations
-(Figure~\ref{fig-map} B). If these phenotypic correlations have a
-genetic component, they could be partially or completely captured by the
-covariance structures specified in the mash model.
+covariances will be to one. If rainfall amounts are uncorrelated,
+covariance will be near zero; if there is rank changing (e.g.~high
+rainfall before flowering of A at MO but not MI, and high rainfall
+before flowering for B at MI but not MO), this covariance will be near
+negative one. We specified multiple covariance structures for the
+canonical, data-driven, and GxWeather categories, each of which
+represented a specific pattern of covariance in genetic effect size
+estimates that could be included in the model. The phenotypic
+correlations for the timing of vegetative growth had moderate negative
+correlations between the Texas and North gardens, particularly in the
+Gulf subpopulation and Both subpopulations (Figure~\ref{fig-map} B). If
+these phenotypic correlations have a genetic component, they could be
+partially or completely captured by the covariance structures specified
+in the mash model.
 
@@
 
-Five GxWeather covariance structures were selected by the greedy
-algorithm, zero to two per subpopulation: two for the start of
-vegetative growth, and three for flowering date. Of these five, three
-had mass on them in mash models of the strong effects
-(Figure~\ref{fig-covar} A-B). Two of these three matrices had negative
-covariances between sites between Texas and North gardens
-(Figure~\ref{fig-covar} A), while one had all positive or near-zero
-covariances. SNP-associated phenotypic effects covaried with different
-weather-based cues in the Gulf \& in Both subpopulations
+We used a greedy algorithm to select covariance matrices from each
+model's set that significantly improved the model likelihood. Five
+GxWeather covariance structures were selected, zero to two per
+subpopulation: two for the start of vegetative growth, and three for
+flowering date. Of these five, three had mass on them in mash models of
+the strong effects (Figure~\ref{fig-covar} A-B). Two of these three
+matrices had negative covariances between sites between Texas and North
+gardens (Figure~\ref{fig-covar} A), while one had all positive or
+near-zero covariances. SNP-associated phenotypic effects covaried with
+different weather-based cues in the Gulf \& in Both subpopulations
 (Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
@@
 prior to the date of vegetative growth. The covariance matrix for this
-weather cue was very similar to the pattern of phenotypic correlation
-for the timing of vegetative growth in the Gulf subpopulation
-(Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Mash models of the
-timing of vegetative growth in the Midwest subpopulation and Both
-subpopulations did not include this GxWeather covariance; the Midwest
-had no weight on any GxWeather covariance, while Both subpopulations had
-non-zero weights on two additional GxWeather covariance types, average
-temperature one day prior to green-up, and the day length change in
-seconds in the day prior to green-up (Figure~\ref{fig-covar} B). The
-average temperature covariance matrix had negative covariances between
-Texas and North gardens, though not as strong as the negative phenotypic
-correlations seen in Both subpopulations (Figure~\ref{fig-map} B;
-Figure~\ref{fig-covar} A). Only the Gulf subpopulation and Both
-subpopulations had mass on any GxWeather covariance matrices
-(Figure~\ref{fig-covar} C).
+weather cue was very similar to the observed pattern of phenotypic
+correlation for the timing of vegetative growth in the Gulf
+subpopulation (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Mash
+models of the timing of vegetative growth in the Midwest subpopulation
+and Both subpopulations did not include this GxWeather covariance; the
+Midwest had no weight on any GxWeather covariance, while Both
+subpopulations had non-zero weights on two additional GxWeather
+covariance types, average temperature one day prior to green-up, and the
+day length change in seconds in the day prior to green-up
+(Figure~\ref{fig-covar} B). The average temperature covariance matrix
+had negative covariances between Texas and North gardens, though not as
+strong as the negative phenotypic correlations seen in Both
+subpopulations (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Only
+the Gulf subpopulation and Both subpopulations had mass on any GxWeather
+covariance matrices (Figure~\ref{fig-covar} C) for vegetative growth.
 
@@
 in the two days prior to flowering (Figure~\ref{fig-covar} D), which had
-negative covariances between Texas and North gardens. Neither covariance
-matrix was selected as significantly improving the log-likelihood of the
+negative covariances between Texas and North gardens. Neither of these
+covariance matrices significantly improved the log-likelihood of the
 \emph{mash} model of Both subpopulations, and no GxWeather covariance
-matrices had a non-zero mass in models of Both subpopulations
+matrices had non-zero mass in models of Both subpopulations
 (Figure~\ref{fig-covar} E). In five of the six mash models of strong
 effects, the GxWeather covariance matrices captured a minority of the
-posterior weights of the strong effects (Figure~\ref{fig-covar} C,E);
-the majority of this mass was on various canonical covariance matrices.
-These matrices included simple heterozygosity, with intermediate,
-positive covariances between all gardens, and garden-specific effects
-present only at one garden.
+posterior weights of the strong effects (Figure~\ref{fig-covar} C,E); in
+these five models, the majority of this mass was on various canonical
+covariance matrices. These matrices included simple heterozygosity, with
+intermediate, positive covariances between all gardens, and
+garden-specific effects present only at one garden.
 
@@
 these gardens (effects with antagonistic pleiotropy), SNP effects that
-differ in magnitude - that are large in one garden or region and smaller
-in others, and SNP effects that are indistinguishable in the two gardens
-(Figure~\ref{fig-effects} A). We used the local false sign rate (lfsr),
-an analogue of the lfdr that establishes confidence in the effect sign,
-not the effect's difference from zero, to determine significance. We
-required lfsr significance (p \textless{} 0.05) in both gardens to
-include effects. This means that our tests for antagonistic pleiotropy,
-or a sign change between conditions, carry an equal statistical burden
-to those for effects with the same sign.
+differ in magnitude (effects that are large in one garden or region and
+smaller in others), and SNP effects that are indistinguishable in the
+two gardens (Figure~\ref{fig-effects} A). We used the local false sign
+rate (lfsr), an analogue of the lfdr that establishes confidence in the
+effect sign, not the effect's difference from zero, to determine
+significance. We required lfsr significance (p \textless{} 0.05) in both
+gardens to include effects. This means that our tests for antagonistic
+pleiotropy, or a sign change between conditions, carry an equal
+statistical burden to those for effects with the same sign.
 
@@
 functions of the genotypes measured and capture only a minority of SNP
-effects on flowering. We know still less about the overwintering
-parameters that cause variation in the start of vegetative growth. We
-could only assign SNP effects to a GxWeather covariance structures in
-four of the six phenotype \& genetic subpopulations we modeled. More
-generally, it is difficult to predict the time scales over which
+effects on flowering. We could only assign SNP effects to a GxWeather
+covariance structures in four of the six phenotype \& genetic
+subpopulations we modeled. It's likely we did not include some GxWeather
+important to these phenological cues - for example, overwintering
+parameters that might cause variation in the start of vegetative growth.
+More generally, it is difficult to predict the time scales over which
 individuals may integrate environmental cues, particularly in perennial
 species which may integrate these cues over longer time scales. If this
-integration time itself varies between individuals, we cannot select a
-covariance structure that reflects this, though these structures would
+integration time itself varies between individuals, the covariance
+structures we modeled cannot reflect this, though these structures would
 likely be highly correlated with GxWeather structures we did include.
@@
 nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect
-estimates for both traits at each garden. We then jointly modeled
+estimates for the two traits at each garden. We then jointly modeled
 genetic effect estimates, and the mixture of ways these effects might

### 2024-05-09 manuscript/revision2/revision2.qmd
chores: ensure revision2 images render from images folder

@@
 Covariance matrices included in mash that had zero posterior weight in all three mash runs on the genetic subpopulations, such as the identity matrix, are not shown.
-(c,e) Total posterior weight placed on covariance matrices that were hypothesized or canonical, for the (c) green-up date phenotype and (e) flowering date phenotype.](Figure_2_recolor_covariance_matrices_and_posterior_weights.png){#fig-covar}
+(c,e) Total posterior weight placed on covariance matrices that were hypothesized or canonical, for the (c) green-up date phenotype and (e) flowering date phenotype.](images/Figure_2_recolor_covariance_matrices_and_posterior_weights.png){#fig-covar}
 
@@
 b) The fraction of effects with each GxE type for the start of vegetative growth (Greenup Date) and reproductive growth (Flowering Date), within and between two genetic subpopulations.
-Common gardens are grouped by the larger region they came from: North gardens are within the natural range of the Midwest subpopulation, while Texas gardens are within the natural range of the Gulf subpopulation.](Figure_3_GxE.png){#fig-effects}
+Common gardens are grouped by the larger region they came from: North gardens are within the natural range of the Midwest subpopulation, while Texas gardens are within the natural range of the Gulf subpopulation.](images/Figure_3_GxE.png){#fig-effects}
 
@@
 a) QTL mapping for the start of vegetative growth (Greenup Date), and three weather-related functions of greenup date.
-b) QTL mapping for the start of reproductive growth (Flowering Date), and two weather-related functions of flowering date.](Figure_4_QTL_Overlaps.png){#fig-qtl}
+b) QTL mapping for the start of reproductive growth (Flowering Date), and two weather-related functions of flowering date.](images/Figure_4_QTL_Overlaps.png){#fig-qtl}
 

### 2024-05-15 manuscript/revision2/response_to_reviews/response_to_reviews.qmd

@@
 format: letter-pdf
+code-annotations: hover
+
 ---
@@
 We agree that the choice of environmental measure is an extremely important consideration.
+
+> \
+> - I’ve struggled to understand the “hypothesis-based covariance matrix” idea.
+Since this idea is really key to the idea of this paper, I think the authors should work to make it much more clear and intuitive, perhaps a diagram could help.
+I think the idea is: if in two gardens the daylength-at-flowering metrics are correlated across accessions, this implies that the importance of daylength on flowering is similar in these two gardens.
+In a third environment that has little variation in daylength-at-flowering, the genetics of the daylength pathway are still the same, so the lack of observed variation must mean that this cue is less important here.
+If a SNP has a similar effect in these two gardens, but a smaller effect in gardens without much variation in this metric, this implies that this SNP may contribute to the signaling pathway that links daylength to flowering.
+This is a neat idea.
+However the way this is described in the text (eg “SNP effects on flowering ...
+covaried with daylength”) is confusing.
+This sounds like SNP effects were larger in gardens where the mean daylength at flowering was later.
+Or maybe that these SNPs were also GWAS hits for daylength at flowering itself.
+
+>Statistical issues:\
+>\
+> - Narrow-sense heritability: A key result is that the h^2 is low when measured across all trials, but high within trials.
+However the model used to estimate the global h^2 is missing a term for replication of lines across accessions.
+u only accounts for replication of additive effects of lines across gardens, but non-additive genetics can be persistent too.
+Also, this model doesn’t allow var(e) to vary across locations but it’s likely that it does.
+If the goal is to show rank-crossing GxE, actually fit a GxE model like was done for the environmental indices to directly show it.
+>\
+> - Variance components analysis: The specification of this model is incorrect because Var(u) and Var(ul) are assigned the same distribution.
+There are many ways of specifying GxE effects in a model like this, and it’s not clear how it is done here.
+If G is a nxn GRM for n accessions, it can’t be the covariance matrix for both Var(u) (dimension nxn) and Var(ul) (dimension nl x nl).
+Generally, Var(ul) would be G \otimes Psi where Psi is a lxl covariance matrix among gardens.
+Sometimes this is constrained to be diagonal (ie no covariance among gardens), and sometimes to be proportional to I (constant variance and no covariance among gardens), but these more restrictive models should be justified.
+
+> \
+> - GWAS: The GWAS model the authors use is not specified (nor was it specified in the 2021 Nature paper referenced).
+I believe looking at the source code that the model is a linear model with PCs to account for structure but no kinship matrix.
+It’s not clear how many PCs were used and since the raw GWAS results are not presented it’s impossible to tell if this was sufficient to account for the significant structure in these populations.
+I’m concerned about this because of the extremely unlikely number of significant markers (19K LD blocks).
+If each of these was a true positive, the average effect size of each block would be only 0.005% of the genetic variance.
+With only 350 accessions, the power to detect a locus that explains 10% of the genetic variance is only 50% with alpha = 10^-5 and MAF = 0.05.
+So if the model is detecting tons of loci it’s likely that the model is severely biased by population structure, and these biases are likely to be somewhat consistent across gardens because the structure is similar.
+mash doesn’t help when the individual models are biased, it’ll find patterns of correlation whether they are biological or not and use those to shrink effect sizes together.
+
+>\
+\- Mash: A lot of the results interpretation is based on interpreting the SNP loadings on the specified covariance matrices.
+This is a secondary use of mash (the primary being the refinement of effect sizes), and while they do interpret these somewhat in the mash paper, these loadings need to be interpreted with caution.
+If these covariance matrices are similar, mash will somewhat arbitrarily assign weight to any of the matrices because all lead to the same fit to the effect-size data.
+The mash algorithm doesn’t give the full posterior distribution on the loadings so you can’t check for posterior correlations there.
+It seems likely to me that the hypothesis-driven and data-driven covariance matrices are somewhat correlated here, and the correlations may differ between green-up and flowering because of the better correlation between the environmental metrics and flowering than green-up.
+In the mash paper, they used cross-validation with the likelihood in the testing set as the evaluation metric to compare models.
+It might be safer to try dropping specific covariance matrices and comparing the model performance in a held-out testing set to evaluate the importance of the hypothesis-derived covariance matrices.
+
+We thank the reviewer for this comment and agree that the selection of covariance matrices to be included in a mash model is a nontrivial question.
+To clarify our interpretation of the posterior matrix weights provided by mash, we performed an extensive analysis of the performance of mash models when different covariance matrices are included.
+Specifically, we implemented a model selection approach that uses a greedy algorithm to evaluate the log likelihood of the mash model as additional covariance matrices were included (see Methods lines \_-\_ and pseudo code below).
+This is a similar, but more computationally efficient design, than the leave-one-out approach recommended by the reviewer as the runtime of mash increases with the number of covariance matrices included.
+Additionally, while cross-validation is a powerful approach to model evaluation our limited sample size was prohibitive to the necessary partitioning of individuals included in our analysis.
+
+In practice this greedy algorithm approach offers a fast way to identify the point where redundancy in the addition of a new covariance matrix results in no change to the likelihood of the mash model.
+Importantly, this removes the potential for arbitrary assignment of weights to the covariance matrices in the model and allows for the selection of matrices that most accurately capture the underlying biology captured by the site-specific effect sizes.
+We applied this algorithm to only the hypothesis matrices for each phenotype (Figures 1 and 2) using the previously used sets of significantly associated and randomly selected variants in the Gulf subpopulation individuals, Midwest subpopulation individuals, and the two subpopulations combined.
+The results indicate that only a small number of the original hypothesis matrices (between three and five) are necessary to reach the maximum likelihood model when using the same likelihood ratio test implemented in Urbut et al., indicating that correlation among hypothesis covariance matrices is high.
+This redundancy was confirmed through our application of the greedy algorithm to a combined set of canonical, data driven, and hypothesis covariance matrices (Figures 3 and 4)
+Our primary conclusion from this analysis is in line with what the reviewer posited, that there is extensive redundancy among canonical, data driven, and hypothesis covariance matrices that is leading to biologically informative patterns of covariance between site effect sizes to be arbitrarily distributed among them.
+We additionally found evidence that while the correlations between data driven and environmentally informed (hypothesis) covariance matrices, there is extensive similarity in the matrices that are included in the maximum likelihood models across both phenotypes and in the analysis of the Gulf subpopulation, Midwest subpopulations, and the combined cohort.
+
+```r
+Greedy algorithm for mash model selection. 
+____________________________________________________________________________
+
+matrices = [] #<1>
+most_likely_matrices = [] #<2>
+maximum_likelihoods = []#<2>
+
+likelihood = -inf #<2>
+most_likely_matrix = ‘’ #<2>
+
+for matrix in matrices: #<3>
+	matrix_likelihood = mash(effects, std.errs, matrix) [‘likelihood’]
+	if matrix_likelihood > likelihood:
+		most_likely_matrix = matrix
+		likelihood = matrix_likelihood
+
+matrices.remove(most_likely_matrix) #<4>
+most_likely_matrices.append(most_likely_matrix)
+maximum_likelihoods.append(likelihood)
+
+lrt_pvalue = 0
+iteration = 2
+while lrt_pvalue < 0.05: #<5>
+	likelihood = -inf
+  most_likely_matrix = ‘’
+	for matrix in matrices:
+		model_matrices = most_likely_matrices + matrix 
+		model_likelihood = mash(effects, std.errs, model_matrices)[‘likelihood’]
+	if matrix_likelihood > likelihood:
+		most_likely_matrix = matrix
+		likelihood = matrix_likelihood
+
+  matrices.remove(most_likely_matrix)
+  most_likely_matrices.append(most_likely_matrix)
+  maximum_likelihoods.append(likelihood)
+
+  lrt_pvalue = likelihood_ratio_test(likelihood, maximum_likelihoods[iteration-1], df = 1)
+____________________________________________________________________________
+```
+1. Initialize covariance matrices of interest
+2. Initialize lists to store the most likely additional matrix and corresponding maximum likelihood
+3. Obtain model likelihood for a model fit with each matrix
+4. Remove the matrix with the greatest likelihood from the array and store it 
+5.
+Repeat the process of fitting mash models, finding the most likely, pair, trio, etc.
+(using the most likely matrices from the preceding iteration) until the likelihood ratio test is no longer significant
+
+>\
+- There appears to be an issue with the construction of the hypothesis-derived covariance matrices.
+I believe the goal here is to estimate the genetic covariance among gardens based on the phenotypic covariance (P) and the estimated residual covariance (E), where P = G + E.
+The described approach involves starting with the the phenotypic correlation matrix and then replacing the diagonal with the coefficient of variation.
+If the diagonal had been replaced with the h2 in each garden, that’d be a valid estimate of G.
+But as described, it’s likely that you’ll end up with an invalid covariance matrix, one that’s not positive-semi-definite.
+Instead the appropriate operation would be to both pre- and post-multiply the covariance matrix by a diagonal matrix with the square-root of the CV’s.
+This maintains a valid covariance-like structure.
+Or just mean-standardize the traits first and then calculate the covariance of this standardized matrix.
+
+>\
+\ - Antagonistic Pleiotropy: Maybe I’m missing the idea of this analysis, but it doesn’t seem to me that this strategy for partitioning loci between antagonistic pleiotropy and differential sensitivity would have equal power to detect each class.
+It seems that antagonistic pleiotropy should be a subset of differential sensitivity because the latter is defined as simply a change in magnitude.
+But even if it’s restricted to “different magnitude but same sign”, the practical definition here seems to be that antagonistic pleiotropy is detected when the signs are different and both lfsrs < threshold (0.05?), while differential sensitivity additionally requires a threshold on the difference in effect sizes (0.4) with no justification for why this size was chosen.
+This additional criterion will necessarily make the rate of detection different.
+Even without this, with differential sensitivity one effect size must be much larger than the other in absolute value (which isn’t the case with antagonistic pleiotropy), so differential sensitivity loci will be more likely to pass the lfsr loci than antagonistic pleiotropy loci.
+Note that previous literature here tried to separate antagonistic pleiotropy from conditional neutrality (ie where in one location the effect was zero), while here the contrast is with differential sensitivity where effects are non-zero in all locations.
+I’m not aware of previous discussions in the literature of trying to differentiate these two specific classes of GxE variants and not really sure theoretically what the importance would be.

### 2024-05-17 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
revision: Add SI Appendix folder/files for revision2

---
title: Diverse Genotype-by-Weather Interactions in Switchgrass
pdf-engine: pdflatex
format:
  pnas-si-pdf:
    keep-tex: false  
author:
  - name: Alice H. MacQueen
    affiliations:
      - name: University of Texas at Austin
        id: a
        department: Department of Integrative Biology
        city: Austin
        state: TX
        postal-code: 78712
    email: alicem@uchicago.edu
    attributes:
      equal-contributor: true
      corresponding: true
  - name: Li Zhang
    affiliations:
      - ref: a
    attributes:
      equal-contributor: true
  - name: Samuel A. Smith
    affiliations:
        - ref: a
    attributes:
      equal-contributor: true
  - name: Jason Bonnette
    affiliations:
          - ref: a
  - name: Arvid R. Boe
    affiliations:
      - name: South Dakota State University
        id: b
        department: Department of Agronomy
        city: Brookings
        state: SD
        postal-code: 57006
  - name: Phillip A. Fay
    affiliations:
      - name: USDA-ARS
        id: c
        department: Grassland, Soil and Water Research Laboratory
        city: Temple
        state: TX
        postal-code: 76502
  - name: Felix B. Fritschi
    affiliations:
      - name: University of Missouri
        id: d
        department: Division of Plant Sciences
        city: Columbia
        state: MO
        postal-code: 65211
  - name: David B. Lowry
    affiliations:
      - name: Michigan State University
        id: e
        department: Department of Plant Biology
        city: East Lansing
        state: MI
        postal-code: 48824
  - name: Robert B. Mitchell
    affiliations:
      - name: USDA-ARS
        id: f
        department: Wheat, Sorghum, and Forage Research Unit
        city: Lincoln
        state: NE
        postal-code: 68583
  - name: Francis M. Rouquette Jr
    affiliations:
      - name: Texas A&M University
        id: g
        department: Texas A&M AgriLife Research and Extension Center
        city: Overton
        state: TX
        postal-code: 75684
  - name: Yanqi Wu
    affiliations:
      - name: Oklahoma State University
        id: h
        department: Department of Plant and Soil Sciences
        city: Stillwater
        state: OK
        postal-code: 74078
  - name: Arbel Harpak
    affiliations:
      - ref: a
  - name: Thomas E. Juenger
    affiliations:
      - ref: a
    attributes:
      corresponding: true
    email: tjuenger@utexas.edu
bibliography: bibliography.bib  
fontsize: "9pt"
classoption:
 - twoside
---

\noindent This PNAS template for Supporting Information (SI) may be used to organize your supporting material. The template is intended to provide a clearly organized PDF file that will ensure readers can easily navigate to sections or specific figures and tables. Movie files or large datasets can be presented as separate files. Further information is available in the [PNAS Author Center](https://www.pnas.org/authors/submitting-your-manuscript).

## Using the template {.unnumbered}

Specify the title, author list, and corresponding authors with the same YAML specification as in the [corresponding PNAS Quarto template](https://github.com/christopherkenny/pnas). The cover page will be automatically generated with the relevant description of the SI.

Figures should be placed on separate pages with legends set immediately below each figure. Table titles should be set immediately above each table. Note that tables extending beyond the width of the page can be included in the PDF or provided as separate dataset files. Oversized/nonstandard page sizes are accepted as part of your SI Appendix file.

References cited in the SI text should be included in a separate reference list at the end of this SI file: @varga2016multilingual and @olsen1992optimal.

Supporting information for Brief Reports is limited to extended methods, essential supporting datasets, and videos (no additional tables or figures). Supporting figures and tables are not allowed for Brief Reports.

## Submitting SI {.unnumbered}

Save your completed SI file as a PDF for submission. Further submission instructions are available [here](https://www.pnas.org/page/authors/submission#preparation).

## Subhead {.unnumbered}

Type or paste text here. This should be additional explanatory text such as an extended technical description of results, full details of mathematical models, etc.

# Supplementary Methods

## Section S1. Covariance matrices generated to jointly re-estimate SNP effects across eight sites {.unnumbered}

Type or paste text here.

## Section S2. Greedy mash algorithm used to select covariance matrices that significantly improved the model log-likelihood {.unnumbered}

Type or paste text here. Break this section up into subheads as needed

## Section S3. Narrow-sense heritability estimation {.unnumbered}

In the diversity panel, we determined narrow-sense heritabilities (h^2^) for green-up and flowering dates at single gardens using genomic relationship matrices calculated using the van Raden method [@vanraden2008]. Genomic relationship matrices were calculated within each subpopulation (Midwest and Gulf) and for both genetic subpopulations (Both). We used ASReml (VSN International) to specify mixed models of the form:

$$
\mathbf{y} = 1 + Zu + e
$$

$$
Var(u) = G\sigma_u^2
$$

$$
Var(e) = I\sigma_e^2
$$

in which the vector **y** represents the flowering date or green-up date values for that garden, Z the design matrix for random effects, u the whole genome additive genetic effect, and e the residual. Matrix G is the whole genomic relationship matrix based on all SNPs retained for subpopulation-specific analyses. I is the rank-y identity matrix. Phenotypic variance $\sigma_p^2$ is $\sigma_u^2 + \sigma_e^2$. Narrow-sense heritability is then $h^2 = (\sigma_u^2 / \sigma_p^2)$.

These models were run for each of the eight gardens, and across all gardens by adding an additional environmental effect of site without an interaction term. This resulted in 54 models: 3 sets of populations (the Gulf, Midwest, and Both subpopulations) for 9 garden sets (all eight gardens separately, and all eight gardens together) and two phenotypes (green-up date and flowering date).

## Section S4. Outbred pseudo-F2 mapping population and Quantitative Trait Locus mapping {.unnumbered}

To confirm candidate genomic regions and patterns of allelic effects found in the diversity panel, we analyzed flowering in an outbred pseudo-F2 cross between four grandparents, two Midwest and two Gulf individuals. The formation of this mapping population has been described previously [@Milano2016]. The parents of this cross were DAC, an early flowering Midwest individual, VS16, a late flowering Midwest individual, AP13, an early flowering Gulf individual, and WBC, a late flowering Gulf individual. We made F1 crosses of the two early flowering genotypes, AP13xDAC, and the two late flowering genotypes, WBCxVS16. We then clonally propagated and planted the four parents, the two F1 genotypes (AP13xDAC, and VS16xWBC), and 801 F2 genotypes at eight field sites in May-July of 2015. To be directly comparable to the diversity panel data, only 2019 phenology data from the pseudo-F2 cross from the same eight common garden sites were used here.

Details on the genetic map construction, map polishing and fine-scale reordering can be accessed on [DataDryad](https://datadryad.org/stash/dataset/doi:10.5061/dryad.ghx3ffbjv). QTL mapping was conducted with R/qtl2 [@broman2019]. We performed a genome scan with a linear mixed model that accounts for the relationships among individuals and for environmental covariates (i.e., field sites). The full model can be expressed as:

$$
phenotype = \mu + QTL + E + QTL*E + kinship + e
$$

where $\mu$ is the population mean, QTL is the marker genetic effect, E is the environmental effects (here, common garden), QTL\*E is the interaction between marker genetic and environmental effects, kinship corresponds to the background polygenic variation, and e is the error term. The genome scan was accomplished with the ‘scan1’ function. The statistical significance of the genome scan was established by performing a stratified (i.e., stratifying on common garden) permutation test (n=1000) using ‘scan1perm’ function. The estimated QTL effect was obtained using ‘scan1coef’ function in R/qtl2.

Figures should be cross-referenced like @fig-frog. Each figure should be on its own page. You can control this by placing a pagebreak shortcode, with `{{< pagebreak >}}`.

This can also be done with LaTeX's `\newpage` command. You may find `\FloatBarrier` useful if your figures or tables are running off to the wrong pages.

{{< pagebreak >}}

![Figure](figs/frog.eps){#fig-frog}

\FloatBarrier <!-- always leave a blank line between this and figures --->

{{< pagebreak >}}

\FloatBarrier

<!-- use these FloatBarrier calls to avoid true-to-the-theme jumping ahead of the frog --->

| Species           |  CBS |   CV |   G3 |
|:------------------|-----:|-----:|-----:|
| Acetaldehyde      |  0.0 |  0.0 |  0.0 |
| Vinyl alcohol     |  9.1 |  9.6 | 13.5 |
| Hydroxyethylidene | 50.8 | 51.2 | 54.0 |

: Simple table example {#tbl-simple}

{{< pagebreak >}}

## Datasets {.unnumbered}

If your document relies on movies or datasets, please list them here with their captions. Use the `movie{your caption}` and `\dataset{file_name.ext}{your caption}` commands to do so.

\dataset{dataset_one.txt}{SNP-associated effects and standard errors for the start of vegetative growth jointly re-estimated in the Gulf genetic subpopulation.}

\dataset{dataset_two.txt}{SNP-associated effects and standard errors for the start of vegetative growth jointly re-estimated in the Midwest genetic subpopulation.}

\dataset{dataset_three.txt}{SNP-associated effects and standard errors for the start of vegetative growth jointly re-estimated in both Midwest and Gulf genetic subpopulations.}

\dataset{dataset_four.txt}{SNP-associated effects and standard errors for the start of reproductive growth jointly re-estimated in the Gulf genetic subpopulation.}

\dataset{dataset_five.txt}{SNP-associated effects and standard errors for the start of reproductive growth jointly re-estimated in the Midwest genetic subpopulation.}

\dataset{dataset_six.txt}{SNP-associated effects and standard errors for the start of reproductive growth jointly re-estimated in both the Midwest and Gulf genetic subpopulations.}

# References

<!-- references for the SI go here. The below div controls its printing --->

::: {#refs}
:::

### 2024-05-17 manuscript/revision2/revision2.qmd
revision: Add SI Appendix folder/files for revision2

@@
 However, as part of its wide environmental adaptation across the eastern half of North America, its photoperiodicity has been predicted to differ by plant latitude of origin [@parrish2005biology; @casler2004latitudinal].
-We previously found divergent Midwest and Gulf genetic subpopulations of switchgrass with distinct sets of environmental adaptations, in that both populations had distinct genetic variation associated with each of two fitness proxies, biomass and survival @lovell2021genomic.
+We previously found divergent Midwest and Gulf genetic subpopulations of switchgrass with distinct sets of environmental adaptations, in that both populations had distinct genetic variation associated with each of two fitness proxies, biomass and overwinter survival @lovell2021genomic.
 The Midwest genetic subpopulation is primarily composed of individuals from the well-studied upland switchgrass ecotype [@porter1966analysis; @milano2016genetic], while the Gulf subpopulation has individuals from the well-studied lowland ecotype and the phenotypically intermediate coastal ecotype @lovell2021genomic.
@@
 These gardens cover the majority of the latitudinal and climatic range of switchgrass and capture the most comprehensive picture to date of the environmental variation this species encounters.
-We define multiple ways phenological timings might covary with weather (\@tbl-covar) and additional ways phenological timings might vary by site (\@SI-info), then jointly re-estimate genetic effects on these timings at all eight sites using the set of these covariance matrices that significantly improved the modeled log-likelihood when included (\@SI-info) @urbut2019flexible.
+We define multiple ways phenological timings might covary with weather (Table 1) and additional ways phenological timings might vary by site (SI Appendix, Table S1, Section SX), then jointly re-estimate genetic effects on these timings at all eight sites using the set of these covariance matrices that significantly improved the modeled log-likelihood when included (SI Appendix, Section SX) @urbut2019flexible.
 We use the Bayesian framework *mash* (multivariate adaptive shrinkage) developed by @urbut2019flexible, to refine effect size estimates from genome-wide association (GWAS) conducted on individuals at each of the eight sites.
@@
 
-We next looked for evidence of GxWeather by jointly re-estimating genetic effects across all eight common gardens using the mash model, which can flexibly capture many modes of covariance across gardens (@fig-map D, SI Appendix, Datasets 1-6).
-We specified three qualitative categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event (Table 1, SI Appendix, Section S2).
+We next looked for evidence of GxWeather by jointly re-estimating genetic effects across all eight common gardens using the mash model, which can flexibly capture many modes of covariance across gardens (@fig-map D; SI Appendix, Datasets 1-6).
+We specified three qualitative categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event (Table 1; SI Appendix, Section S2).
 For example, the more similar that the amount of rainfall prior to flowering is for genotypes A and B between gardens in MI and MO, the nearer these covariances will be to one.
@@
 SNP-associated phenotypic effects covaried with different weather-based cues in the Gulf & in Both subpopulations (@fig-covar B).
-In total, 65% of the posterior weight of strong SNP effects in the mash model of Gulf vegetative growth fell on a covariance matrix constructed using the covariance of daylength 14 days prior to the date of vegetative growth.
-The covariance matrix for this weather cue was very similar to the observed pattern of phenotypic correlation for the timing of vegetative growth in the Gulf subpopulation (@fig-map B; @fig-covar A).
+In total, 65% of the posterior weight of strong SNP effects in the mash model of Gulf vegetative growth fell on a covariance matrix constructed using the covariance of day length 14 days prior to the date of vegetative growth.
+The covariance matrix for this weather cue was visually similar to the observed pattern of phenotypic correlation for the timing of vegetative growth in the Gulf subpopulation (@fig-map B; @fig-covar A).
 Mash models of the timing of vegetative growth in the Midwest subpopulation and Both subpopulations did not include this GxWeather covariance; the Midwest had no weight on any GxWeather covariance, while Both subpopulations had non-zero weights on two additional GxWeather covariance types, average temperature one day prior to green-up, and the day length change in seconds in the day prior to green-up (@fig-covar B).
@@
 
-For greenup date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, between pairs of Texas and North gardens (@fig-effects A; @SI-fig-gardens).
+For greenup date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, between pairs of Texas and North gardens (@fig-effects A; SI Appendix, Fig SX).
 78.7% of pairwise comparisons between North and Texas gardens had a difference in sign, while only 28.6% and 0.2% of North-North or Texas-Texas comparisons had a difference in sign, respectively (@fig-effects).
@@
 Finally, in Both subpopulations, less than 20% of pairwise effects differed in sign (@fig-effects).
-Most differences in sign were between TX1, the southernmost garden, and all other gardens (@SI-fig-gardens).
+Most differences in sign were between TX1, the southernmost garden, and all other gardens (SI Appendix, Fig SX).
 Similarly, more effects that differed in magnitude included gardens in the Texas region (52.3-55.9%), and most effect pairs in the North region were not distinguishable (91.5%).
@@
 ```{=tex}
+\clearpage
+\onecolumn
 \begin{table}[t!]
 \centering
-\caption{Weather variables and time frames around the green up and flowering date used to construct the hypothesis-based covariance matrices.
+\caption{Weather variables and time frames prior to the start of vegetative and reproductive growth used to construct the hypothesis-based covariance matrices.
+Covariance matrices were constructed for, and tested on, three subpopulations and two phenological traits.
 The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices.
@@
 \begin{tabular}{lrr}
-Weather variable & Time Frame & Covariance Matrices tested \\
+Weather variable & Time Frame & Covariance Matrices \\
 \midrule
@@
 \end{table}
+\clearpage
+\twocolumn
 ```
@@
 ![Example hypothesis-driven covariance matrices specified in mash and the posterior weights placed on all covariance matrices.
-(a) Left column: Five example hypothesized covariance matrices specified for the green-up date or flowering date phenotype; these matrices were created from environment-specific correlations across eight common gardens, and are described in Table XXX.
+(a) Left column: Five example hypothesized covariance matrices specified for the green-up date or flowering date phenotype; these matrices were created from environment-specific correlations across eight common gardens, and are described in Table 1.
 Common gardens are arranged in latitudinal order within the matrices.
@@
 
-In 2019, we scored two phenological events every two days in two mapping populations of switchgrass, a diversity panel and a pseudo-F2 cross, planted at eight common garden locations (32, 34, 37).
+In 2019, we scored two phenological events every two days in two mapping populations of switchgrass, a diversity panel and a pseudo-F2 cross, planted at eight common garden locations [@lovell2021; @milano2016genetic; @lowry2019qtl].
 We scored the start of vegetative growth as the day of the year when 50% of the tiller area of the crown of the plant cut the previous year had green growth.
@@
 The diversity panel contained 134 sequenced, clonally propagated individuals from the Midwest genetic subpopulation, and 229 from the Gulf genetic subpopulation.
-To allow for the possibility that different subpopulations had different strengths of connection between our phenotypes and genotypes (38), we conducted three sets of genetic analyses: on Gulf and Midwest genotypes separately, and on both subpopulations together ('Both').
+To allow for the possibility that different subpopulations had different strengths of connection between our phenotypes and genotypes [@kortefarlow2013], we conducted three sets of genetic analyses: on Gulf and Midwest genotypes separately, and on both subpopulations together ('Both').
 Analyses to determine narrow-sense heritability (h^2^) for green-up and flowering were done using linear mixed models and followed @lovell2021genomic.
@@
 
-To evaluate the prevalence and kinds of covariance patterns of SNP effects across our common gardens, we used multivariate adaptive shrinkage (*mash*) on SNP effect estimates from the diversity panel (35).
+To evaluate the prevalence and kinds of covariance patterns of SNP effects across our common gardens, we used multivariate adaptive shrinkage (*mash*) on SNP effect estimates from the diversity panel [@urbut2019flexible].
 To do this, we first conducted univarate genome-wide association (GWAS) on site-specific best linear unbiased predictors (BLUPs) for the timing of vegetative growth and flowering, using the switchgrassGWAS package and the methods in [@lovell2021].
@@
 
-We generated hypothesis-based covariance matrices derived from correlations in environmental cues in the green-up or flowering date windows for the three populations (SI Appendix, Table S1, Section S1).
+We generated hypothesis-based covariance matrices derived from correlations in environmental cues in the green-up or flowering date windows for the three populations (Table 1; SI Appendix, Table S1, Section S1).
 These covariance matrices represent correlations between identical genotypes drawn from a specific population at pairs of common gardens; covariances near one mean that the population has a strong, positive linear relationship in individual responses at that pair of gardens, while covariances near zero mean that there is no relationship within the population for individual responses at that pair of gardens.
@@
 Previous studies of antagonistic pleiotropy (e.g.
-(37)) have used the lfdr or equivalent statistical tests to detect antagonistic pleiotropy.
+[@lowry2019qtl]) have used the lfdr or equivalent statistical tests to detect antagonistic pleiotropy.
 These tests were conservative, in that they required two non-zero effects of different signs, while tests for differential sensitivity required only one non-zero effect.
-This previous work recognized that this testing bias could lead to undercounting occurrences of antagonistic pleiotropy (26, 27), and sought to reduce it by permutation (28).
+This previous work recognized that this testing bias could lead to undercounting occurrences of antagonistic pleiotropy [@anderson2011evolutionary; @des2013genotype], and sought to reduce it by permutation [@anderson2013genetic].
 However, using the lfsr to test for antagonistic pleiotropy does not undercount occurrences of antagonistic pleiotropy, as this statistic answers a fundamentally different question.

### 2024-05-17 manuscript/revision2/revision2.tex
chores: ensure revision2 images render from images folder
revision: Add SI Appendix folder/files for revision2

@@
 genetic variation associated with each of two fitness proxies, biomass
-and survival (33). The Midwest genetic subpopulation is primarily
-composed of individuals from the well-studied upland switchgrass ecotype
-(34, 35), while the Gulf subpopulation has individuals from the
-well-studied lowland ecotype and the phenotypically intermediate coastal
-ecotype (33).
+and overwinter survival (33). The Midwest genetic subpopulation is
+primarily composed of individuals from the well-studied upland
+switchgrass ecotype (34, 35), while the Gulf subpopulation has
+individuals from the well-studied lowland ecotype and the phenotypically
+intermediate coastal ecotype (33).
 
@@
 We define multiple ways phenological timings might covary with weather
-(@tbl-covar) and additional ways phenological timings might vary by site
-(@SI-info), then jointly re-estimate genetic effects on these timings at
-all eight sites using the set of these covariance matrices that
-significantly improved the modeled log-likelihood when included
-(@SI-info) (36). We use the Bayesian framework \emph{mash} (multivariate
-adaptive shrinkage) developed by (36), to refine effect size estimates
-from genome-wide association (GWAS) conducted on individuals at each of
-the eight sites. \emph{mash} allows us to identify and specify multiple
-covariance structures among genetic effect estimates across sites,
-including structures that represent covariance in weather variables of
-interest. Importantly, this method circumvents statistical biases in
-detecting genetic effects with the same or opposite signs (37). To
-confirm our genetic mapping of GxWeather, we compare our posterior
-genetic effect estimates from \emph{mash} to mapping results from an
-outbred pseudo-F2 cross grown at the same sites. Our analyses allow us
-to describe the weather cues and genetic variation affecting phenology
-in two divergent natural populations of switchgrass.
+(Table 1) and additional ways phenological timings might vary by site
+(SI Appendix, Table S1, Section SX), then jointly re-estimate genetic
+effects on these timings at all eight sites using the set of these
+covariance matrices that significantly improved the modeled
+log-likelihood when included (SI Appendix, Section SX) (36). We use the
+Bayesian framework \emph{mash} (multivariate adaptive shrinkage)
+developed by (36), to refine effect size estimates from genome-wide
+association (GWAS) conducted on individuals at each of the eight sites.
+\emph{mash} allows us to identify and specify multiple covariance
+structures among genetic effect estimates across sites, including
+structures that represent covariance in weather variables of interest.
+Importantly, this method circumvents statistical biases in detecting
+genetic effects with the same or opposite signs (37). To confirm our
+genetic mapping of GxWeather, we compare our posterior genetic effect
+estimates from \emph{mash} to mapping results from an outbred pseudo-F2
+cross grown at the same sites. Our analyses allow us to describe the
+weather cues and genetic variation affecting phenology in two divergent
+natural populations of switchgrass.
 
@@
 which can flexibly capture many modes of covariance across gardens
-(Figure~\ref{fig-map} D, SI Appendix, Datasets 1-6). We specified three
+(Figure~\ref{fig-map} D; SI Appendix, Datasets 1-6). We specified three
 qualitative categories of covariance structures across gardens:
@@
 empirical weather patterns at each garden at specific times before the
-phenological event (Table 1, SI Appendix, Section S2). For example, the
+phenological event (Table 1; SI Appendix, Section S2). For example, the
 more similar that the amount of rainfall prior to flowering is for
@@
 strong SNP effects in the mash model of Gulf vegetative growth fell on a
-covariance matrix constructed using the covariance of daylength 14 days
+covariance matrix constructed using the covariance of day length 14 days
 prior to the date of vegetative growth. The covariance matrix for this
-weather cue was very similar to the observed pattern of phenotypic
+weather cue was visually similar to the observed pattern of phenotypic
 correlation for the timing of vegetative growth in the Gulf
@@
 effect sign, between pairs of Texas and North gardens
-(Figure~\ref{fig-effects} A; (\textbf{SI-fig-gardens?})). 78.7\% of
-pairwise comparisons between North and Texas gardens had a difference in
-sign, while only 28.6\% and 0.2\% of North-North or Texas-Texas
-comparisons had a difference in sign, respectively
-(Figure~\ref{fig-effects}). The majority of pairwise effects for greenup
-for the Midwest (\textgreater55\%) and Both (\textgreater85\%)
-subpopulations were the same sign, and effects most often differed in
-magnitude between gardens within and between the regions
-(Figure~\ref{fig-effects} X).
+(Figure~\ref{fig-effects} A; SI Appendix, Fig SX). 78.7\% of pairwise
+comparisons between North and Texas gardens had a difference in sign,
+while only 28.6\% and 0.2\% of North-North or Texas-Texas comparisons
+had a difference in sign, respectively (Figure~\ref{fig-effects}). The
+majority of pairwise effects for greenup for the Midwest
+(\textgreater55\%) and Both (\textgreater85\%) subpopulations were the
+same sign, and effects most often differed in magnitude between gardens
+within and between the regions (Figure~\ref{fig-effects} X).
 
@@
 (Figure~\ref{fig-effects}). Most differences in sign were between TX1,
-the southernmost garden, and all other gardens
-((\textbf{SI-fig-gardens?})). Similarly, more effects that differed in
-magnitude included gardens in the Texas region (52.3-55.9\%), and most
-effect pairs in the North region were not distinguishable (91.5\%).
+the southernmost garden, and all other gardens (SI Appendix, Fig SX).
+Similarly, more effects that differed in magnitude included gardens in
+the Texas region (52.3-55.9\%), and most effect pairs in the North
+region were not distinguishable (91.5\%).
 
@@
 
+\clearpage
+\onecolumn
 \begin{table}[t!]
 \centering
-\caption{Weather variables and time frames around the green up and flowering date used to construct the hypothesis-based covariance matrices.
+\caption{Weather variables and time frames prior to the start of vegetative and reproductive growth used to construct the hypothesis-based covariance matrices.
+Covariance matrices were constructed for, and tested on, three subpopulations and two phenological traits.
 The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices.
@@
 \end{table}
+\clearpage
+\twocolumn
 
@@
 
-\includegraphics{Figure_2_recolor_covariance_matrices_and_posterior_weights.png}
+\includegraphics{images/Figure_2_recolor_covariance_matrices_and_posterior_weights.png}
 
@@
 matrices were created from environment-specific correlations across
-eight common gardens, and are described in Table XXX. Common gardens are
+eight common gardens, and are described in Table 1. Common gardens are
 arranged in latitudinal order within the matrices. Right column: Five
@@
 
-\includegraphics{Figure_3_GxE.png}
+\includegraphics{images/Figure_3_GxE.png}
 
@@
 
-\includegraphics{Figure_4_QTL_Overlaps.png}
+\includegraphics{images/Figure_4_QTL_Overlaps.png}
 
@@
 populations of switchgrass, a diversity panel and a pseudo-F2 cross,
-planted at eight common garden locations (32, 34, 37). We scored the
+planted at eight common garden locations (35, 39, 40). We scored the
 start of vegetative growth as the day of the year when 50\% of the
@@
 different subpopulations had different strengths of connection between
-our phenotypes and genotypes (38), we conducted three sets of genetic
+our phenotypes and genotypes (41), we conducted three sets of genetic
 analyses: on Gulf and Midwest genotypes separately, and on both
@@
 shrinkage (\emph{mash}) on SNP effect estimates from the diversity panel
-(35). To do this, we first conducted univarate genome-wide association
+(36). To do this, we first conducted univarate genome-wide association
 (GWAS) on site-specific best linear unbiased predictors (BLUPs) for the
@@
 correlations in environmental cues in the green-up or flowering date
-windows for the three populations (SI Appendix, Table S1, Section S1).
-These covariance matrices represent correlations between identical
-genotypes drawn from a specific population at pairs of common gardens;
-covariances near one mean that the population has a strong, positive
-linear relationship in individual responses at that pair of gardens,
-while covariances near zero mean that there is no relationship within
-the population for individual responses at that pair of gardens. Mash
-SNP effects will undergo strong shrinkage towards one another in the
-first case, and little shrinkage in the second case. Mash also generates
-data-driven covariance matrices corresponding to major patterns of SNP
-effects present in the data. We generated six data-driven matrices per
-mash run, five produced by singular value decomposition (SVD) of an
-overall matrix.
+windows for the three populations (Table 1; SI Appendix, Table S1,
+Section S1). These covariance matrices represent correlations between
+identical genotypes drawn from a specific population at pairs of common
+gardens; covariances near one mean that the population has a strong,
+positive linear relationship in individual responses at that pair of
+gardens, while covariances near zero mean that there is no relationship
+within the population for individual responses at that pair of gardens.
+Mash SNP effects will undergo strong shrinkage towards one another in
+the first case, and little shrinkage in the second case. Mash also
+generates data-driven covariance matrices corresponding to major
+patterns of SNP effects present in the data. We generated six
+data-driven matrices per mash run, five produced by singular value
+decomposition (SVD) of an overall matrix.
 
@@
 to reject the null hypothesis that an effect j is 0, or that there is a
-significant effect. Previous studies of antagonistic pleiotropy
-(e.g.~(37)) have used the lfdr or equivalent statistical tests to detect
+significant effect. Previous studies of antagonistic pleiotropy (e.g.
+(39)) have used the lfdr or equivalent statistical tests to detect
 antagonistic pleiotropy. These tests were conservative, in that they
@@
 previous work recognized that this testing bias could lead to
-undercounting occurrences of antagonistic pleiotropy (26, 27), and
+undercounting occurrences of antagonistic pleiotropy (27, 29), and
 sought to reduce it by permutation (28). However, using the lfsr to test
@@
 
+\bibitem[\citeproctext]{ref-kortefarlow2013}
+\CSLLeftMargin{41. }%
+\CSLRightInline{A. Korte, A. Farlow, The advantages and limitations of
+trait analysis with GWAS: A review. \emph{Plant methods} \textbf{9}, 29
+(2013).}
+
 \end{CSLReferences}

### 2024-05-17 R/Analysis_v1.2_mash_using_greedy_covar.qmd
revision2: Add new SI Datasets 1-6 to the SI Appendix

@@
 For cumulative weather variables, use a range of days: every day for the first week, and every week until 28 days.
-Come back and add more if 28 is chosen by greedy mash.
+Come back and add more weeks if 28 is chosen by greedy mash.
 
@@
 
-6.  Have Li repeat the QTL analysis and incorporate her results.
+6.  Have Li redo the QTL analysis with any new weather variables and incorporate QTL results.
 
@@
 
-So I should adjust these to make sure all are positive semi-definite, too, in addition to replacing the diagonals with h2.
+So I should adjust these to make sure all are positive semi-definite, too, in addition to replacing the diagonals with h^2^.
 
@@
 
-### Midwest covar fewer sites
+### Midwest covar with fewer sites
 
@@
 
-Given that the greedy algorithm, run on Sam's random inputs, however he generated or found those, gives many more covar matrices especially from Uhyp, I wonder if some repetition of this analysis would be valuable, before blowing these results all out of proportion in a paper.
-Either different random sets of effects, or effects from different years of data.
+Given that the greedy algorithm, run on Sam's random inputs, however he generated or found those, gives many more covar matrices especially from Uhyp, I wonder if some repetition of this analysis would be valuable - either different random sets of effects, or effects from different years of data.
 Think on this.

### 2024-05-17 R/SI_Appendix_Datasets.Rmd
revision2: Add new SI Datasets 1-6 to the SI Appendix

@@
 
+Edited 2024-05-17 for revision2.
+
 ```{r setup, include=FALSE}
@@
 library(tidyverse)
+library(here)
 library(cowplot)
@@
 
-Make all these datasets (except Dataset 7 is already made):
+Make all these datasets:
 
@@
 
-# mash objects
+# Dataset 1-6: mash objects
+
 ```{r}
-m_gulf_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_GR50_Gulf_Uhyp.rds")
-m_gulf_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_FL50_Gulf_Uhyp.rds")
+mashdir <- here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar")
 
-m_midw_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_24000_SNPs_GR50_Midwest_Uhyp.rds")
-m_midw_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_33000_SNPs_FL50_Midwest_Uhyp.rds")
-
-m_gam_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_GR50_Gulf_and_Midwest_Uhyp.rds")
-m_gam_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_FL50_Gulf_and_Midwest_Uhyp.rds")
+numSNPs_v <- c(19000, 19000, 19000, 19000, 24000, 33000)
+suffix_v <- c("FL50_Gulf_and_Midwest", "FL50_Gulf", "GR50_Gulf_and_Midwest",
+              "GR50_Gulf", "GR50_Midwest", "FL50_Midwest")
 ```
@@
 ```{r}
-m <- m_gam_gr_hyp
-dataset_creation_function <- function(m, filename, ncol = 27){
+m_gulf_gr_hyp <- readRDS(file.path(mashdir,  
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
+                                          numSNPs_v[4], "_SNPs_",
+                                          suffix_v[4], ".rds")))
+m_gulf_fl_hyp <- readRDS(file.path(mashdir,  
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
+                                          numSNPs_v[2], "_SNPs_",
+                                          suffix_v[2], ".rds")))
 
-  all_effects <- switchgrassGWAS:::get_marker_df(m) %>%
-    mutate(log10BF = switchgrassGWAS:::get_log10bf(m))
+m_midw_gr_hyp <- readRDS(file.path(mashdir,  
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
+                                          numSNPs_v[5], "_SNPs_",
+                                          suffix_v[5], ".rds")))
+m_midw_fl_hyp <- readRDS(file.path(mashdir,  
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
+                                          numSNPs_v[6], "_SNPs_",
+                                          suffix_v[6], ".rds")))
+
+m_gam_gr_hyp <- readRDS(file.path(mashdir,  
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
+                                          numSNPs_v[3], "_SNPs_",
+                                          suffix_v[3], ".rds")))
+m_gam_fl_hyp <- readRDS(file.path(mashdir,  
+                                   paste0("Strong_mash_posterior_",
+                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
+                                          numSNPs_v[1], "_SNPs_",
+                                          suffix_v[1], ".rds")))
+```
+
+```{r}
+m <- m_gulf_gr_hyp
+dataset_creation_function <- function(m, filename){
+
+  all_effects <- switchgrassGWAS:::get_marker_df(m)
+  BF <- switchgrassGWAS:::get_log10bf(m) |> 
+    as.vector()
+  all_effects$log10BF <- BF
 
@@
   lfsr <- get_lfsr(m)
+  # increase clarity of column names for Supplement data
   colnames(StdErr) <- colnames(StdErr) %>%
-    str_replace(., "Stand_Bhat", "Effect_SE_") %>%
-    str_replace(., "Gulf_and_Midwest", "Both")
+    str_replace(., "Stand_Bhat", "Effect_StandardError_") %>%
+    str_replace(., "Gulf_and_Midwest", "Both") %>%
+    str_replace(., "GR50", "GreenupDate") %>%
+    str_replace(., "FL50", "FloweringDate")
   colnames(Means) <- colnames(Means) %>%
     str_replace(., "Stand_Bhat", "Effect_Mean_") %>%
-    str_replace(., "Gulf_and_Midwest", "Both")
+    str_replace(., "Gulf_and_Midwest", "Both") %>%
+    str_replace(., "GR50", "GreenupDate") %>%
+    str_replace(., "FL50", "FloweringDate")
   colnames(lfsr) <- colnames(lfsr) %>%
     str_replace(., "Stand_Bhat", "Effect_lfsr_") %>%
-    str_replace(., "Gulf_and_Midwest", "Both")
+    str_replace(., "Gulf_and_Midwest", "Both") %>%
+    str_replace(., "GR50", "GreenupDate") %>%
+    str_replace(., "FL50", "FloweringDate")
 
-  all_effects <- cbind(all_effects, Means, StdErr, lfsr)
+  all_effects <- cbind(all_effects, Means, StdErr, lfsr) %>%
+    select(starts_with("Marker"), log10BF, ends_with("TX1"), ends_with("TX2"),
+           ends_with("TX3"), ends_with("OK"), ends_with("MO"), ends_with("NE"),
+           ends_with("MI"), ends_with("SD"))
 
-  effects_df <- all_effects %>%
-    pivot_longer(cols = 4:all_of(ncol), names_to = "Condition", values_to = "Value") %>%
-    separate(col = "Condition", into = c(NA, "Type", "Subpop", "Phe", "Site"), 
-             sep = "_") %>%
-    mutate(Phe = case_when(Phe == "GR50" ~ "green-up",
-                           Phe == "FL50" ~ "flowering")) %>%
-    arrange(Subpop, Phe, Site, Type) %>%
-    pivot_wider(names_from = c("Subpop", "Phe", "Site", "Type"), 
-                values_from = "Value")
+  all_effects |> 
+    write_csv(filename)
 
-  effects_df %>%
-    select(-value) %>%
-    write_csv(filename)
 }
 
-dataset_creation_function(m_gulf_gr_hyp, filename = "~/Github/pvdiv-phenology-gxe/manuscript/Coauthors/PNAS/Dataset_1_gulf_green-up.csv")
-dataset_creation_function(m_gulf_fl_hyp, filename = "~/Github/pvdiv-phenology-gxe/manuscript/Coauthors/PNAS/Dataset_4_gulf_flowering.csv")
+dataset_creation_function(m_gulf_gr_hyp, filename = here("manuscript", "revision2", "SI_Appendix", "Dataset_1_Gulf_vegetative_growth_date.csv"))
+dataset_creation_function(m_gulf_fl_hyp, filename = here("manuscript", "revision2", "SI_Appendix", "Dataset_4_Gulf_reproductive_growth_date.csv"))
 
-dataset_creation_function(m_gam_gr_hyp, filename = "~/Github/pvdiv-phenology-gxe/manuscript/Coauthors/PNAS/Dataset_3_both_subpopulations_green-up.csv")
-dataset_creation_function(m_gam_fl_hyp, filename = "~/Github/pvdiv-phenology-gxe/manuscript/Coauthors/PNAS/Dataset_6_both_subpopulations_flowering.csv")
+dataset_creation_function(m_gam_gr_hyp, filename = here("manuscript", "revision2", "SI_Appendix", "Dataset_3_Both_subpopulations_vegetative_growth_date.csv"))
+dataset_creation_function(m_gam_fl_hyp, filename = here("manuscript", "revision2", "SI_Appendix", "Dataset_6_Both_subpopulations_reproductive_growth_date.csv"))
 
-dataset_creation_function(m_midw_fl_hyp, filename = "~/Github/pvdiv-phenology-gxe/manuscript/Coauthors/PNAS/Dataset_5_midwest_flowering.csv",
-                          ncol = 21)
-dataset_creation_function(m_midw_gr_hyp, filename = "~/Github/pvdiv-phenology-gxe/manuscript/Coauthors/PNAS/Dataset_2_midwest_green-up.csv",
-                          ncol = 21)
+dataset_creation_function(m_midw_fl_hyp, filename = here("manuscript", "revision2", "SI_Appendix", "Dataset_5_Midwest_reproductive_growth_date.csv"))
+dataset_creation_function(m_midw_gr_hyp, filename = here("manuscript", "revision2", "SI_Appendix", "Dataset_2_Midwest_vegetative_growth_date.csv"))
 ```
+
+# Dataset 7
 

### 2024-05-17 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
revision2: Add new SI Datasets 1-6 to the SI Appendix

@@
 Genomic relationship matrices were calculated within each subpopulation (Midwest and Gulf) and for both genetic subpopulations (Both).
-We used ASReml (VSN International) to specify mixed models of the form:
+We used rrBLUP [@rrBLUP] to specify mixed models of the form:
 
@@
 
-\dataset{dataset_one.txt}{SNP-associated effects and standard errors for the start of vegetative growth jointly re-estimated in the Gulf genetic subpopulation.}
+Datasets 1 through 6 are csv files where the first two columns have the following definitions:
+
+-   Marker: The SNP marker in the format Chromosome_Position
+
+-   log10BF: log10(Bayes Factor) of the significance of the marker effect in the mash model
+
+The remaining column names follow the pattern Effect\_\[Mean/StandardError/lfsr\]\_\[Subpopulation\]\_\[Phenotype\]\_\[Garden\], where Mean and Standard Error are estimates of the effect mean and standard error, lfsr is the local false sign rate statistic for the effect, and \[Subpopulation\], \[Phenotype\], and \[Garden\] follow the conventions of Figure 1.
+
+\dataset{Dataset_1_Gulf_vegetative_growth_date.csv}{SNP-associated effects for the start of vegetative growth jointly re-estimated in the Gulf genetic subpopulation.}
 

### 2024-05-17 manuscript/revision2/revision2.qmd
revision2: Add new SI Datasets 1-6 to the SI Appendix

@@
 Genotypes from the Gulf and Midwest subpopulations had distinct phenological timings and distinct patterns of phenological correlations across our eight common garden sites (@fig-map).
-At the three Texas common gardens (hereafter 'Texas' gardens), located within the natural range of the Gulf subpopulation, Gulf vegetative growth occurred before Midwestern vegetative growth, and Gulf flowering occurred after Midwestern flowering (@fig-map A).
-At the four northernmost common gardens (hereafter 'North' gardens), located within the natural range of the Midwest subpopulation, both Gulf vegetative growth and flowering occurred after Midwest vegetative growth and flowering.
-At the Oklahoma common garden, located near the natural range limits of both the Gulf and the Midwest subpopulations, Gulf and Midwest vegetative growth occurred over the same period, and Gulf flowering occurred after Midwestern flowering (@fig-map A).
+At the three Texas common gardens (hereafter 'Texas' gardens), located within the natural range of the Gulf subpopulation, Gulf vegetative growth (hereafter 'green-up') occurred before Midwestern green-up, and Gulf reproductive growth (hereafter 'flowering') occurred after Midwestern flowering (@fig-map A).
+At the four northernmost common gardens (hereafter 'North' gardens), located within the natural range of the Midwest subpopulation, both Gulf green-up and flowering occurred after Midwest green-up and flowering.
+At the Oklahoma common garden, located near the natural range limits of both the Gulf and the Midwest subpopulations, Gulf and Midwest green-up occurred over the same period, and Gulf flowering occurred after Midwestern flowering (@fig-map A).
 These patterns led to strong negative phenotypic correlations for the timing of the start of vegetative growth between the North and Texas gardens, particularly in the Gulf and across all individuals from the Gulf and Midwest subpopulations (hereafter, 'Both' subpopulations), and contributed to positive phenotypic correlations for the timing of flowering that had larger magnitudes at more northern gardens (@fig-map B).
@@
 We used a greedy algorithm to select covariance matrices from each model's set that significantly improved the model likelihood.
-Five GxWeather covariance structures were selected, zero to two per subpopulation: two for the start of vegetative growth, and three for flowering date.
+Five GxWeather covariance structures were selected, zero to two per subpopulation: two for green-up date, and three for flowering date.
 Of these five, three had mass on them in mash models of the strong effects (@fig-covar A-B).
@@
 SNP-associated phenotypic effects covaried with different weather-based cues in the Gulf & in Both subpopulations (@fig-covar B).
-In total, 65% of the posterior weight of strong SNP effects in the mash model of Gulf vegetative growth fell on a covariance matrix constructed using the covariance of day length 14 days prior to the date of vegetative growth.
+In total, 65% of the posterior weight of strong SNP effects in the mash model of Gulf green-up fell on a covariance matrix constructed using the covariance of day length 14 days prior to the date of green-up.
 The covariance matrix for this weather cue was visually similar to the observed pattern of phenotypic correlation for the timing of vegetative growth in the Gulf subpopulation (@fig-map B; @fig-covar A).
@@
 
-For greenup date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, between pairs of Texas and North gardens (@fig-effects A; SI Appendix, Fig SX).
-78.7% of pairwise comparisons between North and Texas gardens had a difference in sign, while only 28.6% and 0.2% of North-North or Texas-Texas comparisons had a difference in sign, respectively (@fig-effects).
-The majority of pairwise effects for greenup for the Midwest (\>55%) and Both (\>85%) subpopulations were the same sign, and effects most often differed in magnitude between gardens within and between the regions (@fig-effects X).
+For green-up date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, between pairs of Texas and North gardens (@fig-effects B; SI Appendix, Fig SX).
+78.7% of pairwise comparisons between North and Texas gardens had a difference in sign, while only 28.6% and 0.2% of North-North or Texas-Texas comparisons had a difference in sign, respectively (@fig-effects B).
+The majority of pairwise effects for greenup for the Midwest (\>55%) and Both (\>85%) subpopulations were the same sign, and effects most often differed in magnitude between gardens within and between the regions (@fig-effects B).
 
-For flowering date for the Gulf subpopulation, less than 2% of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, within or between regions (@fig-effects).
-More effects differed in magnitude between the Texas and North regions than within these regions (42.7% vs \<20%; @fig-effects X), while the majority of effects had no GxE.
+For flowering date for the Gulf subpopulation, less than 2% of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, within or between regions (@fig-effects B).
+More effects differed in magnitude between the Texas and North regions than within these regions (42.7% vs \<20%; @fig-effects B), while the majority of effects had no GxE.
 The Midwest population had relatively few significant effects for flowering, but a large proportion of these differed in sign between Texas and North regions (42.7%) or within the North region (65.4%).
-Finally, in Both subpopulations, less than 20% of pairwise effects differed in sign (@fig-effects).
+Finally, in Both subpopulations, less than 20% of pairwise effects differed in sign (@fig-effects B).
 Most differences in sign were between TX1, the southernmost garden, and all other gardens (SI Appendix, Fig SX).
@@
 ```
-![Characterization of green-up and flowering dates from the switchgrass diversity panel.
+![Characterization of the timing of vegetative (green-up) and reproductive (flowering) growth in the switchgrass diversity panel.
 (a) Map and trait histograms of green-up and flowering dates across two genetically distinct switchgrass subpopulations and eight common gardens.
@@
 In 2019, we scored two phenological events every two days in two mapping populations of switchgrass, a diversity panel and a pseudo-F2 cross, planted at eight common garden locations [@lovell2021; @milano2016genetic; @lowry2019qtl].
-We scored the start of vegetative growth as the day of the year when 50% of the tiller area of the crown of the plant cut the previous year had green growth.
+We scored the start of vegetative growth, or green-up date, as the day of the year when 50% of the tiller area of the crown of the plant cut the previous year had green growth.
 The start of reproductive growth, or flowering date, was the day of the year when 50% of the plant tillers had panicles undergoing anthesis.

### 2024-05-17 manuscript/revision2/revision2.tex
revision2: Add new SI Datasets 1-6 to the SI Appendix

@@
 the natural range of the Gulf subpopulation, Gulf vegetative growth
-occurred before Midwestern vegetative growth, and Gulf flowering
-occurred after Midwestern flowering (Figure~\ref{fig-map} A). At the
-four northernmost common gardens (hereafter `North' gardens), located
-within the natural range of the Midwest subpopulation, both Gulf
-vegetative growth and flowering occurred after Midwest vegetative growth
-and flowering. At the Oklahoma common garden, located near the natural
-range limits of both the Gulf and the Midwest subpopulations, Gulf and
-Midwest vegetative growth occurred over the same period, and Gulf
-flowering occurred after Midwestern flowering (Figure~\ref{fig-map} A).
-These patterns led to strong negative phenotypic correlations for the
-timing of the start of vegetative growth between the North and Texas
-gardens, particularly in the Gulf and across all individuals from the
-Gulf and Midwest subpopulations (hereafter, `Both' subpopulations), and
-contributed to positive phenotypic correlations for the timing of
-flowering that had larger magnitudes at more northern gardens
-(Figure~\ref{fig-map} B).
+(hereafter `green-up') occurred before Midwestern green-up, and Gulf
+reproductive growth (hereafter `flowering') occurred after Midwestern
+flowering (Figure~\ref{fig-map} A). At the four northernmost common
+gardens (hereafter `North' gardens), located within the natural range of
+the Midwest subpopulation, both Gulf green-up and flowering occurred
+after Midwest green-up and flowering. At the Oklahoma common garden,
+located near the natural range limits of both the Gulf and the Midwest
+subpopulations, Gulf and Midwest green-up occurred over the same period,
+and Gulf flowering occurred after Midwestern flowering
+(Figure~\ref{fig-map} A). These patterns led to strong negative
+phenotypic correlations for the timing of the start of vegetative growth
+between the North and Texas gardens, particularly in the Gulf and across
+all individuals from the Gulf and Midwest subpopulations (hereafter,
+`Both' subpopulations), and contributed to positive phenotypic
+correlations for the timing of flowering that had larger magnitudes at
+more northern gardens (Figure~\ref{fig-map} B).
 
@@
 GxWeather covariance structures were selected, zero to two per
-subpopulation: two for the start of vegetative growth, and three for
-flowering date. Of these five, three had mass on them in mash models of
-the strong effects (Figure~\ref{fig-covar} A-B). Two of these three
-matrices had negative covariances between sites between Texas and North
-gardens (Figure~\ref{fig-covar} A), while one had all positive or
-near-zero covariances. SNP-associated phenotypic effects covaried with
-different weather-based cues in the Gulf \& in Both subpopulations
+subpopulation: two for green-up date, and three for flowering date. Of
+these five, three had mass on them in mash models of the strong effects
+(Figure~\ref{fig-covar} A-B). Two of these three matrices had negative
+covariances between sites between Texas and North gardens
+(Figure~\ref{fig-covar} A), while one had all positive or near-zero
+covariances. SNP-associated phenotypic effects covaried with different
+weather-based cues in the Gulf \& in Both subpopulations
 (Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
-strong SNP effects in the mash model of Gulf vegetative growth fell on a
+strong SNP effects in the mash model of Gulf green-up fell on a
 covariance matrix constructed using the covariance of day length 14 days
-prior to the date of vegetative growth. The covariance matrix for this
-weather cue was visually similar to the observed pattern of phenotypic
+prior to the date of green-up. The covariance matrix for this weather
+cue was visually similar to the observed pattern of phenotypic
 correlation for the timing of vegetative growth in the Gulf
@@
 
-For greenup date for the Gulf subpopulation, hundreds to thousands of
+For green-up date for the Gulf subpopulation, hundreds to thousands of
 pairwise effects exhibited antagonistic pleiotropy, or a difference in
 effect sign, between pairs of Texas and North gardens
-(Figure~\ref{fig-effects} A; SI Appendix, Fig SX). 78.7\% of pairwise
+(Figure~\ref{fig-effects} B; SI Appendix, Fig SX). 78.7\% of pairwise
 comparisons between North and Texas gardens had a difference in sign,
 while only 28.6\% and 0.2\% of North-North or Texas-Texas comparisons
-had a difference in sign, respectively (Figure~\ref{fig-effects}). The
+had a difference in sign, respectively (Figure~\ref{fig-effects} B). The
 majority of pairwise effects for greenup for the Midwest
@@
 same sign, and effects most often differed in magnitude between gardens
-within and between the regions (Figure~\ref{fig-effects} X).
+within and between the regions (Figure~\ref{fig-effects} B).
 
@@
 effects exhibited antagonistic pleiotropy, or a difference in effect
-sign, within or between regions (Figure~\ref{fig-effects}). More effects
-differed in magnitude between the Texas and North regions than within
-these regions (42.7\% vs \textless20\%; Figure~\ref{fig-effects} X),
-while the majority of effects had no GxE. The Midwest population had
+sign, within or between regions (Figure~\ref{fig-effects} B). More
+effects differed in magnitude between the Texas and North regions than
+within these regions (42.7\% vs \textless20\%; Figure~\ref{fig-effects}
+B), while the majority of effects had no GxE. The Midwest population had
 relatively few significant effects for flowering, but a large proportion
@@
 within the North region (65.4\%). Finally, in Both subpopulations, less
-than 20\% of pairwise effects differed in sign
-(Figure~\ref{fig-effects}). Most differences in sign were between TX1,
-the southernmost garden, and all other gardens (SI Appendix, Fig SX).
-Similarly, more effects that differed in magnitude included gardens in
-the Texas region (52.3-55.9\%), and most effect pairs in the North
-region were not distinguishable (91.5\%).
+than 20\% of pairwise effects differed in sign (Figure~\ref{fig-effects}
+B). Most differences in sign were between TX1, the southernmost garden,
+and all other gardens (SI Appendix, Fig SX). Similarly, more effects
+that differed in magnitude included gardens in the Texas region
+(52.3-55.9\%), and most effect pairs in the North region were not
+distinguishable (91.5\%).
 
@@
 
-\caption{\label{fig-map}Characterization of green-up and flowering dates
-from the switchgrass diversity panel. (a) Map and trait histograms of
-green-up and flowering dates across two genetically distinct switchgrass
-subpopulations and eight common gardens. Purple represents individuals
-from the Midwest genetic subpopulation, and pink individuals from the
-Gulf subpopulation; map positions represent the original collection
-locations for the genotypes, and shapes represent the ecotype of the
-genotype. Vertical dashed lines indicate the summer solstice. Common
-gardens are arranged in latitudinal order. (b) Phenotypic correlations
-between clonal replicates planted at eight common gardens, within and
-between two genetic subpopulations. (c) Narrow sense heritability of
-green-up and flowering within single common gardens (purple) and across
-all eight common gardens (green), within and between two genetic
-subpopulations. (d) Flow diagram of the methods applied to the green-up
-and flowering dates to jointly estimate SNP effects across all sites.
-Mash was fit to SNP effect data and used to find covariance matrices
-that improved the mash model likelihood using a large set of randomly
-selected, relatively unlinked SNP effects; this model was applied to a
-``strong'' set of SNP effects with large effect sizes in the univariate
-GWAS.}
+\caption{\label{fig-map}Characterization of the timing of vegetative
+(green-up) and reproductive (flowering) growth in the switchgrass
+diversity panel. (a) Map and trait histograms of green-up and flowering
+dates across two genetically distinct switchgrass subpopulations and
+eight common gardens. Purple represents individuals from the Midwest
+genetic subpopulation, and pink individuals from the Gulf subpopulation;
+map positions represent the original collection locations for the
+genotypes, and shapes represent the ecotype of the genotype. Vertical
+dashed lines indicate the summer solstice. Common gardens are arranged
+in latitudinal order. (b) Phenotypic correlations between clonal
+replicates planted at eight common gardens, within and between two
+genetic subpopulations. (c) Narrow sense heritability of green-up and
+flowering within single common gardens (purple) and across all eight
+common gardens (green), within and between two genetic subpopulations.
+(d) Flow diagram of the methods applied to the green-up and flowering
+dates to jointly estimate SNP effects across all sites. Mash was fit to
+SNP effect data and used to find covariance matrices that improved the
+mash model likelihood using a large set of randomly selected, relatively
+unlinked SNP effects; this model was applied to a ``strong'' set of SNP
+effects with large effect sizes in the univariate GWAS.}
 
@@
 planted at eight common garden locations (35, 39, 40). We scored the
-start of vegetative growth as the day of the year when 50\% of the
-tiller area of the crown of the plant cut the previous year had green
-growth. The start of reproductive growth, or flowering date, was the day
-of the year when 50\% of the plant tillers had panicles undergoing
-anthesis. We scored green-up and flowering as day of the year, then
-linked these dates to multiple weather-based environmental factors
-measured daily at each common garden (SI Appendix, Section S1, Table
-S1).
+start of vegetative growth, or green-up date, as the day of the year
+when 50\% of the tiller area of the crown of the plant cut the previous
+year had green growth. The start of reproductive growth, or flowering
+date, was the day of the year when 50\% of the plant tillers had
+panicles undergoing anthesis. We scored green-up and flowering as day of
+the year, then linked these dates to multiple weather-based
+environmental factors measured daily at each common garden (SI Appendix,
+Section S1, Table S1).
 

### 2024-05-21 R/Analysis_v1.2_mash_using_greedy_covar.qmd
revision: Add Section S1 to the SI Appendix

@@
 
-## Step 1: GWAS
+## Step 1: GWAS & mash input prep
+
+GWAS were run on BLUPs from models run at each garden individually.
+The GWAS were run in R/Plots_v0.3_manuscript_figures.Rmd, starting at line 1243.
+I copy-pasted that info here to have the full workflow in one qmd document; you may need to refer to it if you need additional information.
+
+### Run GWAS
+
+```{r}
+library(bigsnpr)
+library(tidyverse)
+library(switchgrassGWAS)
+library(AnnotationDbi)
+txdb <- loadDb(file = file.path("~", "Github", "pvdiv-genome",
+                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+inputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs")
+outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs")
+subpop_v <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", Gulf = "Gulf")
+snp_v <- c("12\\.1M", "8\\.8M", "10\\.3M")
+
+inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
+                                   "inputkinship.txt"),  delim = " ", 
+                                   col_names = "SNPfiles")
+set_v <- list(eight_sites = "eight_sites", tx_sites = "tx_sites", north_sites = "north_sites", single_sites = "!sites")
+
+for(j in 4){#seq_along(set_v)){
+  for(i in seq_along(subpop_v)){
+    dir.create(file.path(outputdir, names(set_v)[j]), showWarnings = FALSE)
+    snp <- snp_attach(inputfiles$SNPfiles[i])
+    phe_df <- read_csv(file.path(inputdir, paste0("BLUP_phenotypes_",
+                                                  "for_gwas_GR_FL_",
+                                                  names(subpop_v)[i],
+                                                  ".csv"))) 
+  if(names(set_v)[j] != "single_sites"){
+    phe_df <- phe_df %>%
+    dplyr::select(PLANT_ID, ends_with(names(set_v)[j]))
+  } else {
+    phe_df <- phe_df %>%
+    dplyr::select(PLANT_ID, !ends_with("sites"))
+    if(names(subpop_v)[i] != "Gulf"){
+      phe_df <- phe_df %>%
+    dplyr::select(PLANT_ID, ends_with("NE"), ends_with("MI"), ends_with("SD"))
+    }
+  }
+  pvdiv_standard_gwas(snp, df = phe_df, type = "linear", 
+                      outputdir = file.path(outputdir, names(set_v)[j]),
+                      savegwas = TRUE, saveannos = TRUE, txdb = txdb,
+                      minphe = 130)
+
+  gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir,
+                                                       names(set_v)[j]), 
+                                      pattern = paste0("GWAS_datatable*.*",
+                                                       snp_v[i]))
+  phenotypes <- str_sub(gwas_rds, start = 16, end = -42) %>%
+    str_replace(., "_$", "")
+  out1 <- pvdiv_bigsnp2mashr(path = file.path(outputdir, names(set_v)[j]), 
+                             snp = snp, gwas_rds = gwas_rds, 
+                             phenotypes = phenotypes, numSNPs = 1000, 
+                             clump = FALSE, scaled = FALSE, saveoutput = TRUE,
+                             suffix = paste0("FL_", names(set_v)[j], "_",
+                                             names(subpop_v)[i]))
+  }
+}
+
+```
+
+I made the mash inputs originally in the Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd file.
+I copy-pasted that info here to have the full workflow in one qmd document; some needs to be run on the Juenger lab server.
+
+### Make mashr inputs for 8 GR50 phenotypes & run mash.
+
+Some univariate GWAS for the Midwest had population structure issues. Mostly for greenup - even GR50 at MO was quite bad. Drop MO for GR50. Otherwise they look ok for FL50 and GR50.
+
+#### 2a. Gulf & Gulf&Midwest.
+
+NB: Do this on server, where the large GWAS_datatable objects were created.
+
+```{r mash inputs with gulf}
+#| eval: false
+# library()
+k = 3 # subpop: G&M, M, Gulf
+for(k in c(1, 3)){
+phe_suffix <- paste("GR50", subpop_v[k], 
+                    sep = "*.*")
+snp <- snp_attach(inputfiles$SNPfiles[k])
+gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
+                                      pattern = paste0("GWAS_datatable*.*",
+                                                       phe_suffix))
+phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
+    str_replace(., "_$", "")
+phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
+numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
+out1phe <- pvdiv_bigsnp2mashr(path = file.path(outputdir), 
+                           snp = snp, gwas_rds = gwas_rds, 
+                           phenotypes = phenotypes, numSNPs = numSNPs, 
+                           clump = TRUE, scaled = TRUE, saveoutput = TRUE,
+                           suffix = paste0("GR50_", names(subpop_v)[k]))
+}
+```
+
+#### 2b. Midwest, dropping gardens with zero values.
+
+Midwest_GR50_SD have values of zero for Bhat and Shat. Drop this from Midwest analyses.
+
+```{r check midwest}
+mwdf <- readRDS(file.path(outputdir,
+                          "B_hat_strong_df_19000topSNPs_GR50_Midwest.rds"))
+
+summary(mwdf)
+```
+
+```{r mash midwest}
+k=2
+phe_suffix <- paste("GR50", subpop_v[k], 
+                    sep = "*.*")
+snp <- snp_attach(inputfiles$SNPfiles[k])
+gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
+                                      pattern = paste0("GWAS_datatable*.*",
+                                                       phe_suffix))
+gwas_rds <- gwas_rds[c(1:4,6:8)]
+
+phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -41) %>%
+    str_replace(., "_$", "")
+phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
+numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
+out1phe <- pvdiv_bigsnp2mashr(path = file.path(outputdir), 
+                           snp = snp, gwas_rds = gwas_rds, 
+                           phenotypes = phenotypes, numSNPs = numSNPs, 
+                           clump = TRUE, scaled = TRUE, saveoutput = TRUE,
+                           suffix = paste0("GR50_", names(subpop_v)[k]))
+# Ran mash using a 'nohup_v0.6' R script.
+```
+
+Used scp to move these output files (which all have the same suffix) to laptop.
+
+e.g.
+`scp -i Github/jls_alice alice@popgen.biosci.utexas.edu:~/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/*19000topSNPs_GR50_Midwest*  /home/alice/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/Midwest/`
+
+`scp -i ~/Github/jls_alice alice@popgen.biosci.utexas.edu:~/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/Part-One-Output*  /home/alice/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/`
+
+Files with the six suffixes I have/want mash results for: B_hat_strong_df_19000topSNPs_FL50_Gulf_and_Midwest.rds B_hat_strong_df_19000topSNPs_FL50_Gulf.rds B_hat_strong_df_19000topSNPs_GR50_Gulf_and_Midwest.rds B_hat_strong_df_19000topSNPs_GR50_Gulf.rds B_hat_strong_df_24000topSNPs_GR50_Midwest.rds B_hat_strong_df_33000topSNPs_FL50_Midwest.rds
+
+# ----------------------
 
@@
 Think on this.
-I definitely want to use my known inputs, not Sam's, as mine have a traceable history (or anyways more traceable).
+I definitely want to use my known inputs, not Sam's, as mine have a traceable history (or anyways are currently traceable).
 

### 2024-05-21 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
revision: Add Section S1 to the SI Appendix

@@
 
-Type or paste text here.
+We specified three qualitative categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event.
+
+Canonical covariance matrices can be created by *mash* and fall into four groups: an identity matrix, an equal effects matrix, singleton matrices, and simple heterozygousity matrices.
+The identity matrix has 0 covariance between different gardens and correlations of 1 within gardens, and represents the situation in which the effects in different conditions are independent.
+The equal effects matrix has correlations of 1 for all matrix cells, and represents identical effects across all conditions.
+Each singleton matrix has a correlation of 1 for a single garden and 0's elsewhere, representing a garden-specific effect.
+Three simple heterozygousity matrices are generated, which have covariances of 0.25, 0.5, or 0.75 for all off-diagonal matrix cells, and correlations of 1 on the diagonal, representing moderately positively correlated effects across all sites.
+
+Data-driven covariance matrices can also be created by *mash* and fall into two groups: an overall pattern of covariance between all conditions, denoted as `tPCA`, and the first five principle components of a singular value decomposition of the tPCA, denoted as PCA_1 through PCA_5.
+
+To create the GxWeather covariance matrices, for each genotype's green-up and flowering date at each garden, we created new, date-related values that were summary functions for a weather variable across a specified date range.
+For example, the cumulative sum of the rainfall that occurred seven days prior to the flowering date was used to create one possible GxWeather cue variable for flowering date.
+The code for these calculations can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 208.
+
+We defined both phenological dates as functions of five weather variables and eight to ten time frames (Table 1): Cumulative growing degree days (GDD) variables for each time frame were calculated as $GDD = \sum^{DATE}_{DATE - i} max(T_{mean} - T_{base}, 0)$, where $T_{mean}$ is the daily average temperature, defined as $(T_{max} + T_{min}) / 2$, $T_{base}$ is the base temperature of 12°C for switchgrass, $T_{max}$ is the maximum daily temperature, $T_{min}$ is the minimum daily temperature, $DATE$ is the phenological date, and $i$ is the time frame, or the number of days prior to the phenological date that GDD is summed across.
+Second, cumulative rainfall values in each time frame were calculated as $C_{rain} = \sum^{DATE}_{DATE - i} PRCP$, where PRCP is the daily rainfall, measured in millimeters, $DATE$ is the phenological date, and $i$ is the time frame, or the number of days prior to the phenological date that rainfall is summed across.
+Third, day length in hours was determined for each time frame as a specific single day prior to the phenological date, as indicated by the time frame $i$, e.g.
+$DATE - i$.
+Day length was calculated as a function of latitude and day of the year as in [@forsythe1995].
+Fourth, day length change in seconds was determined for a specific single day prior to the phenological date as indicated by the time frame, as $DYLN_{change (s)} = (DYLN_{DATE - i} - DYLN_{DATE-i-1})*60*60$, where $DYLN$ in the day length in hours, $DATE$ is the phenological date, and $i$ is the number of days prior to the phenological date.
+Fifth, average temperature in each time frame was defined as $T_{ave} = \sum^{DATE}_{DATE - i} (T_{max} + T_{min}) / 2$, where $T_{max}$ is the maximum daily temperature, $T_{min}$ is the minimum daily temperature, $DATE$ is the phenological date, and $i$ is the time frame, or the number of days prior to the phenological date that average temperature is summed across.
+
+We then generated GxWeather covariance matrices derived from correlations in these GxWeather cue variables for the three subpopulations (Table S1, Figure S1).
+These covariance matrices represent the correlations between genotypes for these weather-related variables across our eight common gardens.
+To create these matrices, we determined the correlations between these weather-related cue variables for identical genotypes grown at different common gardens and used these correlations to fill the off-diagonals of a eight-by-eight common garden matrix.
+For the diagonal of these matrices, we used the narrow-sense heritability for each weather-related cue variable at each garden, calculated using the same methods as Section S3.
+Matrices need to be positive semi-definite to be used in *mash*, so when the resultant matrices were not positive semi-definite, the `Matrix::nearPD` function was used to compute the nearest positive definite matrix to the correlation matrix.
+Finally, all covariance matrices are rescaled so that the maximum diagonal element of the rescaled matrix is 1, as in [@Urbut2019].
+Code used to generate these matrices can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 536.
+
+We defined distinct sets of weather-based covariance matrices for the two phenological dates and for the three genetic subgroups.
+We used the same five weather variables for both green-up and flowering date, but the two sets of weather-based covariance matrices were defined relative to the two phenological dates.
+Weather-based covariance matrices were also defined separately for individual genotypes from the Gulf, Midwest, and Both subpopulations.
 

### 2024-05-21 manuscript/revision2/revision2.qmd
revision: Add Section S1 to the SI Appendix

@@
 
-We next looked for evidence of GxWeather by jointly re-estimating genetic effects across all eight common gardens using the mash model, which can flexibly capture many modes of covariance across gardens (@fig-map D; SI Appendix, Datasets 1-6).
+We next looked for evidence of GxWeather by jointly re-estimating genetic effects across all eight common gardens using the *mash* model, which can flexibly capture many modes of covariance across gardens (@fig-map D; SI Appendix, Datasets 1-6).
 We specified three qualitative categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event (Table 1; SI Appendix, Section S2).
@@
 The phenotypic correlations for the timing of vegetative growth had moderate negative correlations between the Texas and North gardens, particularly in the Gulf subpopulation and Both subpopulations (@fig-map B).
-If these phenotypic correlations have a genetic component, they could be partially or completely captured by the covariance structures specified in the mash model.
+If these phenotypic correlations have a genetic component, they could be partially or completely captured by the covariance structures specified in the *mash* model.
 
@@
 Five GxWeather covariance structures were selected, zero to two per subpopulation: two for green-up date, and three for flowering date.
-Of these five, three had mass on them in mash models of the strong effects (@fig-covar A-B).
+Of these five, three had mass on them in *mash* models of the strong effects (@fig-covar A-B).
 Two of these three matrices had negative covariances between sites between Texas and North gardens (@fig-covar A), while one had all positive or near-zero covariances.
 SNP-associated phenotypic effects covaried with different weather-based cues in the Gulf & in Both subpopulations (@fig-covar B).
-In total, 65% of the posterior weight of strong SNP effects in the mash model of Gulf green-up fell on a covariance matrix constructed using the covariance of day length 14 days prior to the date of green-up.
+In total, 65% of the posterior weight of strong SNP effects in the *mash* model of Gulf green-up fell on a covariance matrix constructed using the covariance of day length 14 days prior to the date of green-up.
 The covariance matrix for this weather cue was visually similar to the observed pattern of phenotypic correlation for the timing of vegetative growth in the Gulf subpopulation (@fig-map B; @fig-covar A).
-Mash models of the timing of vegetative growth in the Midwest subpopulation and Both subpopulations did not include this GxWeather covariance; the Midwest had no weight on any GxWeather covariance, while Both subpopulations had non-zero weights on two additional GxWeather covariance types, average temperature one day prior to green-up, and the day length change in seconds in the day prior to green-up (@fig-covar B).
+*Mash* models of the timing of vegetative growth in the Midwest subpopulation and Both subpopulations did not include this GxWeather covariance; the Midwest had no weight on any GxWeather covariance, while Both subpopulations had non-zero weights on two additional GxWeather covariance types, average temperature one day prior to green-up, and the day length change in seconds in the day prior to green-up (@fig-covar B).
 The average temperature covariance matrix had negative covariances between Texas and North gardens, though not as strong as the negative phenotypic correlations seen in Both subpopulations (@fig-map B; @fig-covar A).
@@
 Neither of these covariance matrices significantly improved the log-likelihood of the *mash* model of Both subpopulations, and no GxWeather covariance matrices had non-zero mass in models of Both subpopulations (@fig-covar E).
-In five of the six mash models of strong effects, the GxWeather covariance matrices captured a minority of the posterior weights of the strong effects (@fig-covar C,E); in these five models, the majority of this mass was on various canonical covariance matrices.
+In five of the six *mash* models of strong effects, the GxWeather covariance matrices captured a minority of the posterior weights of the strong effects (@fig-covar C,E); in these five models, the majority of this mass was on various canonical covariance matrices.
 These matrices included simple heterozygosity, with intermediate, positive covariances between all gardens, and garden-specific effects present only at one garden.
@@
 We sought additional experimental support for our joint re-estimates of SNP effects using an independent pseudo-F2 mapping population created from Gulf & Midwest individuals and grown at the same sites (@fig-qtl A,B).
-We conducted quantitative trait loci (QTL) mapping of flowering as functions of five environmental cues that we also used as covariance matrices in mash, and identified eight QTL for flowering date, eight QTL for flowering as a function of day length change two days prior, one QTL for the start of vegetative growth, and two QTL for vegetative growth as a function of daylength change one day prior, all of which showed QTL by environment interactions (SI Appendix, Fig.
+We conducted quantitative trait loci (QTL) mapping of flowering as functions of five environmental cues that we also used as covariance matrices in *mash*, and identified eight QTL for flowering date, eight QTL for flowering as a function of day length change two days prior, one QTL for the start of vegetative growth, and two QTL for vegetative growth as a function of daylength change one day prior, all of which showed QTL by environment interactions (SI Appendix, Fig.
 S3).
@@
 thaliana* with functionally validated roles in flowering (SI Appendix, Dataset 7).
-All flowering and green-up QTL intervals contained at least one SNP significant in at least one mash run at a log10-transformed Bayes Factor \> 2, or in the 1% tail of significance, whichever was stricter (SI Appendix, Dataset 8).
-We also looked for enrichments of mash SNPs in the 1% tail of significance (the 'mash 1% tail') within each QTL interval.
-At the 5% level, three QTL had enrichments of SNPs in the mash 1% tail.
-Overall, there were five significant enrichments (p \< 0.05, hypergeometric test) of SNPs in the mash 1% tail in the QTL intervals.
+All flowering and green-up QTL intervals contained at least one SNP significant in at least one *mash* run at a log10-transformed Bayes Factor \> 2, or in the 1% tail of significance, whichever was stricter (SI Appendix, Dataset 8).
+We also looked for enrichments of *mash* SNPs in the 1% tail of significance (the '*mash* 1% tail') within each QTL interval.
+At the 5% level, three QTL had enrichments of SNPs in the *mash* 1% tail.
+Overall, there were five significant enrichments (p \< 0.05, hypergeometric test) of SNPs in the *mash* 1% tail in the QTL intervals.
 Thus, we were able to experimentally support some of our re-estimates of SNP effects with a QTL mapping experiment using a separate mapping population.
@@
 \midrule
-1. cumulative GDD at 12C in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60 \\
-2. cumulative rainfall in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60  \\
-3. day length (hours) on a specific day indicated by the time frame & (1-7) and 14 days prior to green up date & 48 \\
-4. day length change (seconds) on a specific day indicated by the time frame & (1-7) and 14 days prior to green up date & 48 \\
-5. average temperature in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60 \\
+1. cumulative growing degree days at 12C in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60 \\
+2. cumulative rainfall in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60  \\
+3. day length (hours) on a specific day indicated by the time frame & (1-7) and 14 days prior to the phenological date & 48 \\
+4. day length change (seconds) on a specific day indicated by the time frame & (1-7) and 14 days prior to the phenological date & 48 \\
+5. average temperature in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60 \\
 \bottomrule
@@
 (d) Flow diagram of the methods applied to the green-up and flowering dates to jointly estimate SNP effects across all sites.
-Mash was fit to SNP effect data and used to find covariance matrices that improved the mash model likelihood using a large set of randomly selected, relatively unlinked SNP effects; this model was applied to a "strong" set of SNP effects with large effect sizes in the univariate GWAS.](images/Fig1_200dpi.png){#fig-map}
+*Mash* was fit to SNP effect data and used to find covariance matrices that improved the *mash* model likelihood using a large set of randomly selected, relatively unlinked SNP effects; this model was applied to a "strong" set of SNP effects with large effect sizes in the univariate GWAS.](images/Fig1_200dpi.png){#fig-map}
 
-![Example hypothesis-driven covariance matrices specified in mash and the posterior weights placed on all covariance matrices.
+![Example hypothesis-driven covariance matrices specified in *mash* and the posterior weights placed on all covariance matrices.
 (a) Left column: Five example hypothesized covariance matrices specified for the green-up date or flowering date phenotype; these matrices were created from environment-specific correlations across eight common gardens, and are described in Table 1.
@@
 Canonical matrices (purple) have simple interpretations, such as equal effects across all common gardens, or effects specific to a single common garden.
-(b,d) Total posterior weight placed on each covariance matrix type specified for (b) green-up date and (d) flowering date mash models, within and between two genetic subpopulations.
+(b,d) Total posterior weight placed on each covariance matrix type specified for (b) green-up date and (d) flowering date *mash* models, within and between two genetic subpopulations.
 Hypothesized covariance matrices (green).
-Covariance matrices included in mash that had zero posterior weight in all three mash runs on the genetic subpopulations, such as the identity matrix, are not shown.
+Covariance matrices included in *mash* that had zero posterior weight in all three *mash* runs on the genetic subpopulations, such as the identity matrix, are not shown.
 (c,e) Total posterior weight placed on covariance matrices that were hypothesized or canonical, for the (c) green-up date phenotype and (e) flowering date phenotype.](images/Figure_2_recolor_covariance_matrices_and_posterior_weights.png){#fig-covar}
@@
 Dotted lines indicate permutation-based significance thresholds for each weather-related function.
-Stars indicate QTL with significant enrichment for SNPs in the 1% mash tail; G, M, and B indicate which subpopulation had enrichment: G - Gulf subpopulation, M -Midwest subpopulation, B - both subpopulations.
-Rug plots show genomic locations of SNPs in the 1% mash tail for flowering date for each subpopulation.
+Stars indicate QTL with significant enrichment for SNPs in the 1% *mash* tail; G, M, and B indicate which subpopulation had enrichment: G - Gulf subpopulation, M -Midwest subpopulation, B - both subpopulations.
+Rug plots show genomic locations of SNPs in the 1% *mash* tail for flowering date for each subpopulation.
 a) QTL mapping for the start of vegetative growth (Greenup Date), and three weather-related functions of greenup date.
@@
 The diversity panel contained 134 sequenced, clonally propagated individuals from the Midwest genetic subpopulation, and 229 from the Gulf genetic subpopulation.
-To allow for the possibility that different subpopulations had different strengths of connection between our phenotypes and genotypes [@kortefarlow2013], we conducted three sets of genetic analyses: on Gulf and Midwest genotypes separately, and on both subpopulations together ('Both').
+To allow for the possibility that different subpopulations had different strengths of connection between our phenotypes and genotypes [@kortefarlow2013], we conducted three sets of genetic analyses: on Gulf and Midwest genotypes separately, and on both subpopulations together ('Both' subpopulations).
 Analyses to determine narrow-sense heritability (h^2^) for green-up and flowering were done using linear mixed models and followed @lovell2021genomic.
-Details on these models can be found in (SI Appendix, Section S3,S4).
+Details on these models can be found in (SI Appendix, Section S2).
 
@@
 In these models (hereafter referred to as "random"), we then specified a variety of potential covariance structures among genetic effect size estimates from each garden (as expanded on below) and used a greedy algorithm to iteratively select and add covariance structures that significantly improved the log likelihood of the *mash* model.
+Details on these models can be found in (SI Appendix, Section S3).
 Given the high correlation between some of the covariance patterns we tested, this approach allowed us to select only the subset of patterns of covariance that significantly improved the fit of the jointly inferred structure of genetic effect estimates present in the data.
@@
 These covariance matrices represent correlations between identical genotypes drawn from a specific population at pairs of common gardens; covariances near one mean that the population has a strong, positive linear relationship in individual responses at that pair of gardens, while covariances near zero mean that there is no relationship within the population for individual responses at that pair of gardens.
-Mash SNP effects will undergo strong shrinkage towards one another in the first case, and little shrinkage in the second case.
-Mash also generates data-driven covariance matrices corresponding to major patterns of SNP effects present in the data.
-We generated six data-driven matrices per mash run, five produced by singular value decomposition (SVD) of an overall matrix.
+*Mash* SNP effects will undergo strong shrinkage towards one another in the first case, and little shrinkage in the second case.
+*Mash* also generates data-driven covariance matrices corresponding to major patterns of SNP effects present in the data.
+We generated six data-driven matrices per *mash* run, five produced by singular value decomposition (SVD) of an overall matrix.
 
@@
 To confirm candidate genomic regions and patterns of allelic effects found in the diversity panel, we analyzed flowering in an outbred pseudo-F2 cross between four individuals, two Midwest and two Gulf individuals.
-The formation of this mapping population has been described previously [@milano2016genetic]; additional details on QTL mapping can be found in SI Appendix, Section S6.
+The formation of this mapping population has been described previously [@milano2016genetic]; additional details on QTL mapping can be found in SI Appendix, Section S4.
 To be directly comparable to the diversity panel data, only 2019 phenology data from the pseudo-F2 cross from the same eight common garden sites were used.

### 2024-05-21 manuscript/revision2/revision2.tex
revision: Add Section S1 to the SI Appendix

@@
 We next looked for evidence of GxWeather by jointly re-estimating
-genetic effects across all eight common gardens using the mash model,
-which can flexibly capture many modes of covariance across gardens
-(Figure~\ref{fig-map} D; SI Appendix, Datasets 1-6). We specified three
-qualitative categories of covariance structures across gardens:
-``canonical'' covariance, with simple patterns of effect size covariance
-introduced in the initial \emph{mash} manuscript; ``data-driven''
-covariances derived from common patterns of SNP effects observed in the
-data, and ``GxWeather'' covariance, estimated from the covariance of
-empirical weather patterns at each garden at specific times before the
-phenological event (Table 1; SI Appendix, Section S2). For example, the
-more similar that the amount of rainfall prior to flowering is for
-genotypes A and B between gardens in MI and MO, the nearer these
-covariances will be to one. If rainfall amounts are uncorrelated,
-covariance will be near zero; if there is rank changing (e.g.~high
-rainfall before flowering of A at MO but not MI, and high rainfall
-before flowering for B at MI but not MO), this covariance will be near
-negative one. We specified multiple covariance structures for the
-canonical, data-driven, and GxWeather categories, each of which
+genetic effects across all eight common gardens using the \emph{mash}
+model, which can flexibly capture many modes of covariance across
+gardens (Figure~\ref{fig-map} D; SI Appendix, Datasets 1-6). We
+specified three qualitative categories of covariance structures across
+gardens: ``canonical'' covariance, with simple patterns of effect size
+covariance introduced in the initial \emph{mash} manuscript;
+``data-driven'' covariances derived from common patterns of SNP effects
+observed in the data, and ``GxWeather'' covariance, estimated from the
+covariance of empirical weather patterns at each garden at specific
+times before the phenological event (Table 1; SI Appendix, Section S2).
+For example, the more similar that the amount of rainfall prior to
+flowering is for genotypes A and B between gardens in MI and MO, the
+nearer these covariances will be to one. If rainfall amounts are
+uncorrelated, covariance will be near zero; if there is rank changing
+(e.g.~high rainfall before flowering of A at MO but not MI, and high
+rainfall before flowering for B at MI but not MO), this covariance will
+be near negative one. We specified multiple covariance structures for
+the canonical, data-driven, and GxWeather categories, each of which
 represented a specific pattern of covariance in genetic effect size
@@
 partially or completely captured by the covariance structures specified
-in the mash model.
+in the \emph{mash} model.
 
@@
 subpopulation: two for green-up date, and three for flowering date. Of
-these five, three had mass on them in mash models of the strong effects
-(Figure~\ref{fig-covar} A-B). Two of these three matrices had negative
-covariances between sites between Texas and North gardens
+these five, three had mass on them in \emph{mash} models of the strong
+effects (Figure~\ref{fig-covar} A-B). Two of these three matrices had
+negative covariances between sites between Texas and North gardens
 (Figure~\ref{fig-covar} A), while one had all positive or near-zero
@@
 (Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
-strong SNP effects in the mash model of Gulf green-up fell on a
+strong SNP effects in the \emph{mash} model of Gulf green-up fell on a
 covariance matrix constructed using the covariance of day length 14 days
@@
 correlation for the timing of vegetative growth in the Gulf
-subpopulation (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Mash
-models of the timing of vegetative growth in the Midwest subpopulation
-and Both subpopulations did not include this GxWeather covariance; the
-Midwest had no weight on any GxWeather covariance, while Both
-subpopulations had non-zero weights on two additional GxWeather
+subpopulation (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A).
+\emph{Mash} models of the timing of vegetative growth in the Midwest
+subpopulation and Both subpopulations did not include this GxWeather
+covariance; the Midwest had no weight on any GxWeather covariance, while
+Both subpopulations had non-zero weights on two additional GxWeather
 covariance types, average temperature one day prior to green-up, and the
@@
 matrices had non-zero mass in models of Both subpopulations
-(Figure~\ref{fig-covar} E). In five of the six mash models of strong
-effects, the GxWeather covariance matrices captured a minority of the
-posterior weights of the strong effects (Figure~\ref{fig-covar} C,E); in
-these five models, the majority of this mass was on various canonical
-covariance matrices. These matrices included simple heterozygosity, with
-intermediate, positive covariances between all gardens, and
-garden-specific effects present only at one garden.
+(Figure~\ref{fig-covar} E). In five of the six \emph{mash} models of
+strong effects, the GxWeather covariance matrices captured a minority of
+the posterior weights of the strong effects (Figure~\ref{fig-covar}
+C,E); in these five models, the majority of this mass was on various
+canonical covariance matrices. These matrices included simple
+heterozygosity, with intermediate, positive covariances between all
+gardens, and garden-specific effects present only at one garden.
 
@@
 mapping of flowering as functions of five environmental cues that we
-also used as covariance matrices in mash, and identified eight QTL for
-flowering date, eight QTL for flowering as a function of day length
-change two days prior, one QTL for the start of vegetative growth, and
-two QTL for vegetative growth as a function of daylength change one day
-prior, all of which showed QTL by environment interactions (SI Appendix,
-Fig. S3). All QTL for flowering overlapped one or more homologs from
-rice or \emph{A. thaliana} with functionally validated roles in
-flowering (SI Appendix, Dataset 7). All flowering and green-up QTL
-intervals contained at least one SNP significant in at least one mash
-run at a log10-transformed Bayes Factor \textgreater{} 2, or in the 1\%
-tail of significance, whichever was stricter (SI Appendix, Dataset 8).
-We also looked for enrichments of mash SNPs in the 1\% tail of
-significance (the `mash 1\% tail') within each QTL interval. At the 5\%
-level, three QTL had enrichments of SNPs in the mash 1\% tail. Overall,
-there were five significant enrichments (p \textless{} 0.05,
-hypergeometric test) of SNPs in the mash 1\% tail in the QTL intervals.
-Thus, we were able to experimentally support some of our re-estimates of
-SNP effects with a QTL mapping experiment using a separate mapping
-population.
+also used as covariance matrices in \emph{mash}, and identified eight
+QTL for flowering date, eight QTL for flowering as a function of day
+length change two days prior, one QTL for the start of vegetative
+growth, and two QTL for vegetative growth as a function of daylength
+change one day prior, all of which showed QTL by environment
+interactions (SI Appendix, Fig. S3). All QTL for flowering overlapped
+one or more homologs from rice or \emph{A. thaliana} with functionally
+validated roles in flowering (SI Appendix, Dataset 7). All flowering and
+green-up QTL intervals contained at least one SNP significant in at
+least one \emph{mash} run at a log10-transformed Bayes Factor
+\textgreater{} 2, or in the 1\% tail of significance, whichever was
+stricter (SI Appendix, Dataset 8). We also looked for enrichments of
+\emph{mash} SNPs in the 1\% tail of significance (the `\emph{mash} 1\%
+tail') within each QTL interval. At the 5\% level, three QTL had
+enrichments of SNPs in the \emph{mash} 1\% tail. Overall, there were
+five significant enrichments (p \textless{} 0.05, hypergeometric test)
+of SNPs in the \emph{mash} 1\% tail in the QTL intervals. Thus, we were
+able to experimentally support some of our re-estimates of SNP effects
+with a QTL mapping experiment using a separate mapping population.
 
@@
 (d) Flow diagram of the methods applied to the green-up and flowering
-dates to jointly estimate SNP effects across all sites. Mash was fit to
-SNP effect data and used to find covariance matrices that improved the
-mash model likelihood using a large set of randomly selected, relatively
-unlinked SNP effects; this model was applied to a ``strong'' set of SNP
-effects with large effect sizes in the univariate GWAS.}
+dates to jointly estimate SNP effects across all sites. \emph{Mash} was
+fit to SNP effect data and used to find covariance matrices that
+improved the \emph{mash} model likelihood using a large set of randomly
+selected, relatively unlinked SNP effects; this model was applied to a
+``strong'' set of SNP effects with large effect sizes in the univariate
+GWAS.}
 
@@
 \caption{\label{fig-covar}Example hypothesis-driven covariance matrices
-specified in mash and the posterior weights placed on all covariance
-matrices. (a) Left column: Five example hypothesized covariance matrices
-specified for the green-up date or flowering date phenotype; these
-matrices were created from environment-specific correlations across
-eight common gardens, and are described in Table 1. Common gardens are
-arranged in latitudinal order within the matrices. Right column: Five
-example canonical covariance matrices. Canonical matrices (purple) have
-simple interpretations, such as equal effects across all common gardens,
-or effects specific to a single common garden. (b,d) Total posterior
-weight placed on each covariance matrix type specified for (b) green-up
-date and (d) flowering date mash models, within and between two genetic
-subpopulations. Hypothesized covariance matrices (green). Covariance
-matrices included in mash that had zero posterior weight in all three
-mash runs on the genetic subpopulations, such as the identity matrix,
-are not shown. (c,e) Total posterior weight placed on covariance
-matrices that were hypothesized or canonical, for the (c) green-up date
-phenotype and (e) flowering date phenotype.}
+specified in \emph{mash} and the posterior weights placed on all
+covariance matrices. (a) Left column: Five example hypothesized
+covariance matrices specified for the green-up date or flowering date
+phenotype; these matrices were created from environment-specific
+correlations across eight common gardens, and are described in Table 1.
+Common gardens are arranged in latitudinal order within the matrices.
+Right column: Five example canonical covariance matrices. Canonical
+matrices (purple) have simple interpretations, such as equal effects
+across all common gardens, or effects specific to a single common
+garden. (b,d) Total posterior weight placed on each covariance matrix
+type specified for (b) green-up date and (d) flowering date \emph{mash}
+models, within and between two genetic subpopulations. Hypothesized
+covariance matrices (green). Covariance matrices included in \emph{mash}
+that had zero posterior weight in all three \emph{mash} runs on the
+genetic subpopulations, such as the identity matrix, are not shown.
+(c,e) Total posterior weight placed on covariance matrices that were
+hypothesized or canonical, for the (c) green-up date phenotype and (e)
+flowering date phenotype.}
 
@@
 function. Stars indicate QTL with significant enrichment for SNPs in the
-1\% mash tail; G, M, and B indicate which subpopulation had enrichment:
-G - Gulf subpopulation, M -Midwest subpopulation, B - both
-subpopulations. Rug plots show genomic locations of SNPs in the 1\% mash
-tail for flowering date for each subpopulation. a) QTL mapping for the
-start of vegetative growth (Greenup Date), and three weather-related
-functions of greenup date. b) QTL mapping for the start of reproductive
-growth (Flowering Date), and two weather-related functions of flowering
-date.}
+1\% \emph{mash} tail; G, M, and B indicate which subpopulation had
+enrichment: G - Gulf subpopulation, M -Midwest subpopulation, B - both
+subpopulations. Rug plots show genomic locations of SNPs in the 1\%
+\emph{mash} tail for flowering date for each subpopulation. a) QTL
+mapping for the start of vegetative growth (Greenup Date), and three
+weather-related functions of greenup date. b) QTL mapping for the start
+of reproductive growth (Flowering Date), and two weather-related
+functions of flowering date.}
 
@@
 analyses: on Gulf and Midwest genotypes separately, and on both
-subpopulations together (`Both'). Analyses to determine narrow-sense
-heritability (h\textsuperscript{2}) for green-up and flowering were done
-using linear mixed models and followed (33). Details on these models can
-be found in (SI Appendix, Section S3,S4).
+subpopulations together (`Both' subpopulations). Analyses to determine
+narrow-sense heritability (h\textsuperscript{2}) for green-up and
+flowering were done using linear mixed models and followed (33). Details
+on these models can be found in (SI Appendix, Section S2).
 
@@
 add covariance structures that significantly improved the log likelihood
-of the \emph{mash} model. Given the high correlation between some of the
+of the \emph{mash} model. Details on these models can be found in (SI
+Appendix, Section S3). Given the high correlation between some of the
 covariance patterns we tested, this approach allowed us to select only
@@
 within the population for individual responses at that pair of gardens.
-Mash SNP effects will undergo strong shrinkage towards one another in
-the first case, and little shrinkage in the second case. Mash also
-generates data-driven covariance matrices corresponding to major
-patterns of SNP effects present in the data. We generated six
-data-driven matrices per mash run, five produced by singular value
-decomposition (SVD) of an overall matrix.
+\emph{Mash} SNP effects will undergo strong shrinkage towards one
+another in the first case, and little shrinkage in the second case.
+\emph{Mash} also generates data-driven covariance matrices corresponding
+to major patterns of SNP effects present in the data. We generated six
+data-driven matrices per \emph{mash} run, five produced by singular
+value decomposition (SVD) of an overall matrix.
 
@@
 previously (35); additional details on QTL mapping can be found in SI
-Appendix, Section S6. To be directly comparable to the diversity panel
+Appendix, Section S4. To be directly comparable to the diversity panel
 data, only 2019 phenology data from the pseudo-F2 cross from the same

### 2024-05-22 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
revision: small edits to Section S1 for clarity

--- 2024-05-17 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
@@
 
-Type or paste text here.
+We specified three qualitative categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data and introduced in the initial *mash* manuscript, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden over specific time frames before the phenological event.
+
+Canonical covariance matrices can be created by *mash* and fall into four groups: an identity matrix, an equal effects matrix, singleton matrices, and simple heterozygousity matrices.
+The identity matrix has values of zero for covariance between different gardens and correlations of one within gardens, and represents the situation in which the effects in different conditions are independent.
+The equal effects matrix has values of one for all matrix elements, and represents identical effects across all conditions.
+Each singleton matrix has a value of one for a single garden and zeros for all other elements, representing a garden-specific effect.
+Three simple heterozygousity matrices are generated, which have covariances of 0.25, 0.5, or 0.75 for all off-diagonal matrix elements, and correlations of one for diagonal entries, representing moderately positively correlated effects across all sites.
+
+Data-driven covariance matrices can also be created by *mash* and fall into two groups: an overall pattern of covariance between all conditions, denoted as $tPCA$, and the first five principle components of a singular value decomposition of the $tPCA$, denoted as `PCA_1`$PCA_1$ through $PCA_5$[@Urbut2019].
+
+To create the GxWeather covariance matrices we transformed each genotype's green-up and flowering date at each garden to create date-related values that were summary functions for a weather variable across a specified date range.
+For example, the cumulative sum of the rainfall that occurred seven days prior to the flowering date was used to create one possible GxWeather cue variable for flowering date.
+The code for these calculations can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 208.
+
+We defined both phenological dates as functions of five weather variables and eight to ten time frames (Table 1): Cumulative growing degree days (GDD) variables for each time frame were calculated as $GDD = \sum^{DATE}_{DATE - i} max(T_{mean} - T_{base}, 0)$, where $T_{mean}$ is the daily average temperature, defined as $(T_{max} + T_{min}) / 2$, $T_{base}$ is the base temperature of 12°C for switchgrass, $T_{max}$ is the maximum daily temperature, $T_{min}$ is the minimum daily temperature, $DATE$ is the phenological date, and $i$ is the time frame, or the number of days prior to the phenological date that GDD is summed across [@kiniry2005; @behrman2013].
+Second, cumulative rainfall values in each time frame were calculated as $C_{rain} = \sum^{DATE}_{DATE - i} PRCP$, where PRCP is the daily rainfall measured in millimeters, $DATE$ is the phenological date, and $i$ is the time frame, or the number of days prior to the phenological date that rainfall is summed across.
+Third, day length in hours was determined for each time frame as a specific single day prior to the phenological date, as indicated by the time frame $i$, e.g.
+$DATE - i$.
+Day length was calculated as a function of latitude and day of the year as in [@forsythe1995].
+Fourth, day length change in seconds was determined for a specific single day prior to the phenological date as indicated by the time frame, as $DYLN_{change (s)} = (DYLN_{DATE - i} - DYLN_{DATE-i-1})*60*60$, where $DYLN$ in the day length in hours, $DATE$ is the phenological date, and $i$ is the number of days prior to the phenological date.
+Fifth, average temperature in each time frame was defined as $T_{ave} = \sum^{DATE}_{DATE - i} (T_{max} + T_{min}) / 2$, where $T_{max}$ is the maximum daily temperature, $T_{min}$ is the minimum daily temperature, $DATE$ is the phenological date, and $i$ is the time frame, or the number of days prior to the phenological date that average temperature is summed across.
+
+We then generated GxWeather covariance matrices derived from correlations in these GxWeather cue variables for the three subpopulations.
+These covariance matrices represent the correlations between genotypes for these weather-related variables across our eight common gardens.
+To create these matrices, we determined the correlations between these weather-related cue variables for identical genotypes grown at different common gardens and used these correlations to fill the off-diagonals of a eight-by-eight common garden matrix.
+For the diagonal of these matrices, we used the narrow-sense heritability for each weather-related cue variable at each garden, calculated using the same methods as Section S3.
+Matrices need to be positive semi-definite to be used in *mash*, so when the resultant matrices were not positive semi-definite, the `Matrix::nearPD` function was used to compute the nearest positive definite matrix to the correlation matrix [@Matrix].
+Finally, all covariance matrices are rescaled so that the maximum diagonal element of the rescaled matrix is 1, as in [@Urbut2019].
+Code used to generate these matrices can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 536.
+
+We defined distinct sets of weather-based covariance matrices for the two phenological dates and for the three genetic subgroups.
+We used the same five weather variables to transform both green-up and flowering date, but defined the weather-based covariance matrices relative to each of the two phenological dates.
+Weather-based covariance matrices were also defined separately for individual genotypes from the Gulf, Midwest, and Both subpopulations.
 
@@
 
-\dataset{dataset_two.txt}{SNP-associated effects and standard errors for the start of vegetative growth jointly re-estimated in the Midwest genetic subpopulation.}
+\dataset{Dataset_2_Midwest_vegetative_growth_date.csv}{SNP-associated effects and standard errors for the start of vegetative growth jointly re-estimated in the Midwest genetic subpopulation.}
 
-\dataset{dataset_three.txt}{SNP-associated effects and standard errors for the start of vegetative growth jointly re-estimated in both Midwest and Gulf genetic subpopulations.}
+\dataset{Dataset_3_Both_subpopulations_vegetative_growth_date.csv}{SNP-associated effects and standard errors for the start of vegetative growth jointly re-estimated in both the Midwest and Gulf genetic subpopulations.}
 
-\dataset{dataset_four.txt}{SNP-associated effects and standard errors for the start of reproductive growth jointly re-estimated in the Gulf genetic subpopulation.}
+\dataset{Dataset_4_Gulf_reproductive_growth_date.csv}{SNP-associated effects and standard errors for the start of reproductive growth jointly re-estimated in the Gulf genetic subpopulation.}
 
-\dataset{dataset_five.txt}{SNP-associated effects and standard errors for the start of reproductive growth jointly re-estimated in the Midwest genetic subpopulation.}
+\dataset{Dataset_5_Midwest_reproductive_growth_date.csv}{SNP-associated effects and standard errors for the start of reproductive growth jointly re-estimated in the Midwest genetic subpopulation.}
 
-\dataset{dataset_six.txt}{SNP-associated effects and standard errors for the start of reproductive growth jointly re-estimated in both the Midwest and Gulf genetic subpopulations.}
+\dataset{Dataset_6_Both_subpopulations_reproductive_growth_date.csv}{SNP-associated effects and standard errors for the start of reproductive growth jointly re-estimated in both the Midwest and Gulf genetic subpopulations.}
 

### 2024-05-24 manuscript/revision2/revision2.qmd
revision: Address Tom's comment about sign-changing GxE vs antagonistic pleiotropy

@@
 For example, in the Gulf population, 65% of genetic effects on the timing of vegetative growth covary with daylength 14 days prior to green-up date, and 33% of genetic effects on the timing of flowering covary with cumulative rainfall in the seven days prior to flowering.
-However, most variation in genetic effects across sites is site-specific and can not be attributed to variation in weather variables alone.
+However, most variation in genetic effects is site-specific and can not be attributed to variation in the studied weather variables.
 We demonstrate that we can identify genetic variation with GxWeather and assign these loci to specific weather-based cues or other patterns.
-Breeding for particular alleles at the GxWeather loci we identified could change flowering responsiveness in a photoperiod and rainfall-specific way.
+Breeding for particular alleles at the GxWeather loci we identified could change flowering responsiveness in a photoperiod or rainfall-specific way.
 More broadly, our approach proposes a refined, powerful characterization of genotype-by-environment interactions in any species phenotyped in multiple environments.
@@
 equal-contributor: "A.H.M. contributed equally to this work with L.Z. and S.A.S."
-keywords:  [allele-by-environment effect variation, antagonistic pleiotropy, photoperiod, cumulative rainfall, genetic variation]
+keywords:  [allele-by-environment effect variation, rank-changing genotype-by-environment interaction, photoperiod, cumulative rainfall, genetic variation]
 acknowledgements: | 
@@
 However, most studies of flowering GxE focus on finding a single, best fitting form of genotype-environment covariance, despite the key expectation that different genetic subpopulations, and even different genomic regions, have likely evolved distinct patterns of GxE [@Weine2023.06.21.545998].
-Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or sign-changing GxE, at the level of individual loci [@levene1953genetic; @felsenstein1976theoretical; @kawecki2004conceptual; @hedrick1986genetic], previous empirical work has found limited evidence of antagonistic pleiotropy [@des2013genotype; @wadgymar2017identifying].
+Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or alleles with effects with opposing fitness outcomes [@levene1953genetic; @felsenstein1976theoretical; @kawecki2004conceptual; @hedrick1986genetic], previous empirical work has found limited evidence of antagonistic pleiotropy [@des2013genotype; @wadgymar2017identifying].
 However, this work has been limited by a known statistical bias that reduced detection of genetic effects that differ in sign [@des2013genotype; @anderson2013genetic; @anderson2011evolutionary].
-Thus, despite substantial interest in the frequencies of various forms of GxE, the prevalence of antagonistic pleiotropy relative to other forms of GxE remains unknown.
+Thus, despite substantial interest in the frequencies of various forms of GxE, particularly sign-changing GxE, relative to other forms of GxE remains unknown.
 
@@
 
-Here, we estimate GxWeather for two phenological timings in switchgrass by assigning patterns of genetic effects on phenology across gardens to many patterns of weather covariance at these gardens.
+Here, we estimate GxWeather for two phenological timings in switchgrass by assigning patterns of genetic effects on phenology for cloned individuals grown at eight common gardens to many patterns of weather covariance at these gardens.
 To do this, we phenotype a diversity panel of hundreds of switchgrass genotypes from the Midwest and Gulf subpopulations for the timing of vegetative and reproductive development at eight common garden locations spanning 17 degrees of latitude.
 These gardens cover the majority of the latitudinal and climatic range of switchgrass and capture the most comprehensive picture to date of the environmental variation this species encounters.
-We define multiple ways phenological timings might covary with weather (Table 1) and additional ways phenological timings might vary by site (SI Appendix, Table S1, Section SX), then jointly re-estimate genetic effects on these timings at all eight sites using the set of these covariance matrices that significantly improved the modeled log-likelihood when included (SI Appendix, Section SX) @urbut2019flexible.
-We use the Bayesian framework *mash* (multivariate adaptive shrinkage) developed by @urbut2019flexible, to refine effect size estimates from genome-wide association (GWAS) conducted on individuals at each of the eight sites.
+We define multiple ways phenological timings might covary with weather (Table 1) and additional ways phenological timings might vary by garden (SI Appendix, Section S1), then jointly re-estimate genetic effects on these timings at all eight common gardens using the set of these covariance matrices that significantly improved the modeled log-likelihood when included (SI Appendix, Section S2) @urbut2019flexible.
+We use the Bayesian framework *mash* (multivariate adaptive shrinkage) developed by @urbut2019flexible, to refine effect size estimates from genome-wide association (GWAS) conducted on individuals at each of the eight gardens.
 *mash* allows us to identify and specify multiple covariance structures among genetic effect estimates across sites, including structures that represent covariance in weather variables of interest.
 Importantly, this method circumvents statistical biases in detecting genetic effects with the same or opposite signs @10.1093/biostatistics/kxw041.
-To confirm our genetic mapping of GxWeather, we compare our posterior genetic effect estimates from *mash* to mapping results from an outbred pseudo-F2 cross grown at the same sites.
-Our analyses allow us to describe the weather cues and genetic variation affecting phenology in two divergent natural populations of switchgrass.
+To confirm our genetic mapping of GxWeather, we compare our significant posterior genetic effects locations from *mash* to mapping results from an outbred mapping population grown at the same sites.
+Our analyses allow us to describe the weather cues and genetic variation affecting phenology in two divergent natural populations of switchgrass across the species' latitudinal range.
 
@@
 
-We next looked for evidence of GxWeather by jointly re-estimating genetic effects across all eight common gardens using the *mash* model, which can flexibly capture many modes of covariance across gardens (@fig-map D; SI Appendix, Datasets 1-6).
+We next looked for evidence of GxWeather by jointly re-estimating genetic effects of SNPs across all eight common gardens using the *mash* model, which can flexibly capture many modes of covariance across gardens (@fig-map D; SI Appendix, Datasets 1-6).
 We specified three qualitative categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event (Table 1; SI Appendix, Section S2).
@@
 The GxWeather covariance structures allow hypothesis testing of specific weather variables as cues for the start of vegetative and reproductive growth.
-Say that a SNP in FLC controls flowering in a photoperiod-dependent manner.
+Say that a SNP in the gene *Flowering Locus C* (FLC), a well-known flowering time regulator, controls flowering in a photoperiod-dependent manner.
 In that case, the joint estimate of effects for that SNP could have a high mixture proportion, or mass, on a covariance matrix created using a photoperiod-based environmental cue, such as day length at some interval prior to flowering.
@@
 In five of the six *mash* models of strong effects, the GxWeather covariance matrices captured a minority of the posterior weights of the strong effects (@fig-covar C,E); in these five models, the majority of this mass was on various canonical covariance matrices.
-These matrices included simple heterozygosity, with intermediate, positive covariances between all gardens, and garden-specific effects present only at one garden.
+These matrices included simple heterozygosity, with intermediate, positive covariances between all gardens, and single effect matrices with garden-specific effects (SI Appendix, Section S1).
 
 We next characterized the pairwise patterns of effects where we were confident in the sign of the effect at both gardens.
-If the overall pattern of phenotypic expression differs between a pair of gardens, this difference may be comprised of SNP effects that differ in sign between these gardens (effects with antagonistic pleiotropy), SNP effects that differ in magnitude (effects that are large in one garden or region and smaller in others), and SNP effects that are indistinguishable in the two gardens (@fig-effects A).
+If the overall pattern of phenotypic expression differs between a pair of gardens, this difference may be comprised of SNP effects that differ in sign between these gardens (effects with rank-changing GxE between gardens), SNP effects that differ in magnitude (effects that are large in one garden or region and smaller in others), and SNP effects that are indistinguishable in the two gardens (@fig-effects A).
 We used the local false sign rate (lfsr), an analogue of the lfdr that establishes confidence in the effect sign, not the effect's difference from zero, to determine significance.
 We required lfsr significance (p \< 0.05) in both gardens to include effects.
-This means that our tests for antagonistic pleiotropy, or a sign change between conditions, carry an equal statistical burden to those for effects with the same sign.
+This means that our tests for a sign change between gardens carry an equal statistical burden to those for effects with the same sign.
 
-For green-up date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, between pairs of Texas and North gardens (@fig-effects B; SI Appendix, Fig SX).
+For green-up date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited a difference in effect sign, between pairs of Texas and North gardens (@fig-effects B; SI Appendix, Fig SX).
 78.7% of pairwise comparisons between North and Texas gardens had a difference in sign, while only 28.6% and 0.2% of North-North or Texas-Texas comparisons had a difference in sign, respectively (@fig-effects B).
@@
 
-For flowering date for the Gulf subpopulation, less than 2% of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, within or between regions (@fig-effects B).
+For flowering date for the Gulf subpopulation, less than 2% of pairwise effects exhibited a difference in effect sign within or between regions (@fig-effects B).
 More effects differed in magnitude between the Texas and North regions than within these regions (42.7% vs \<20%; @fig-effects B), while the majority of effects had no GxE.
@@
 
-We sought additional experimental support for our joint re-estimates of SNP effects using an independent pseudo-F2 mapping population created from Gulf & Midwest individuals and grown at the same sites (@fig-qtl A,B).
+We sought additional experimental support the occurrence and location of our significant SNP effect re-estimates using an independent pseudo-F2 mapping population created from Gulf & Midwest individuals and grown at the same sites (@fig-qtl A,B).
 We conducted quantitative trait loci (QTL) mapping of flowering as functions of five environmental cues that we also used as covariance matrices in *mash*, and identified eight QTL for flowering date, eight QTL for flowering as a function of day length change two days prior, one QTL for the start of vegetative growth, and two QTL for vegetative growth as a function of daylength change one day prior, all of which showed QTL by environment interactions (SI Appendix, Fig.
@@
 Here, we demonstrate that we can associate multiple patterns of GxWeather with specific genomic regions using a switchgrass diversity panel grown at eight common gardens.
-We can assign genetic effects to both GxWeather patterns with interpretable weather-based cues, and to agnostic, site-based patterns.
+We can assign genetic effects to both GxWeather patterns with interpretable weather-based cues, and to unmeasured, site-based patterns.
 We use this approach to study GxWeather for the timings of vegetative and reproductive development in the deeply genetically diverged Gulf and Midwest subpopulations of switchgrass.
 
-Our analysis of the timing of vegetative growth in the Gulf and in Both subpopulations revealed substantial antagonistic pleiotropy in effects between the Texas and North gardens (@fig-effects B).
-This result supports theoretical models that local adaptation should involve antagonistic pleiotropy at the level of individual loci [@kawecki2004conceptual; @felsenstein1976theoretical; @levene1953genetic; @hedrick1986genetic], and is the first experimental work using GWAS across common gardens to find antagonistic pleiotropy to be common in small genomic regions [@wadgymar2017identifying; @savolainen2013ecological; @lowry2019qtl].
+Our analysis of the timing of vegetative growth in the Gulf and in Both subpopulations revealed many alleles with sign changes in their effects between the Texas and North gardens (@fig-effects B).
+As phenological timings are major components of plant fitness, this result supports theoretical models that local adaptation should involve antagonistic pleiotropy at the level of individual loci [@kawecki2004conceptual; @felsenstein1976theoretical; @levene1953genetic; @hedrick1986genetic]; it expands the genetics of local adaptation research to more than two field sites and to a wider range of genetic variation by using effects from GWAS in common gardens located throughout the species' range to determine that rank-changing GxE for phenological timings is common in small genomic regions [@wadgymar2017identifying; @savolainen2013ecological; @lowry2019qtl].
 
-Our analysis of the timing of flowering showed that the Gulf and Midwest subpopulations have distinct GxWeather: flowering timing in the Midwest subpopulation has photoperiod-related variation, in that flowering timing covaries with a day length change signal two days before flowering occurs.
-In contrast, the Gulf subpopulation does not have variation in flowering based on a photoperiod cue.
-Instead, the Gulf subpopulation has variation in flowering that covaries with the rainfall that occurs in the week prior to flowering.
+Our analysis of the timing of flowering showed that the Gulf and Midwest subpopulations have distinct GxWeather: flowering timing in the Midwest subpopulation has photoperiod-related genetic variation, in that flowering timing covaries with a day length change signal two days before flowering occurs.
+In contrast, the Gulf subpopulation does not have genetic variation in flowering that covaries with a photoperiod cue.
+Instead, the Gulf subpopulation has genetic variation in flowering that covaries with the rainfall that occurs in the week prior to flowering.
 Three genomic regions affecting flowering that we re-estimated across all eight sites were also supported by QTL from an independent mapping population at these sites (@fig-qtl B).
@@
 
-Last, we characterized the overall patterns of antagonistic pleiotropy in the set of SNPs where there was pairwise significance of effects at pairs of gardens.
+Last, we characterized the overall patterns of GxE in the set of SNPs where there was pairwise significance of effects at pairs of gardens.
 To do this, we used the 'get_GxE' function of the switchgrassGWAS R package.
 First, this determines the set of SNPs with evidence of significant effects in both conditions for all pairs of conditions using local false sign rates (lfsr) as the significance criteria.
-Second, to determine antagonistic pleiotropy, this function determines if effects significant in both conditions are of opposite sign.
+Then, this function determines if effects significant in both conditions are of opposite sign.
 
-Using the lfsr rather than the local false discovery rate (lfdr) is a critical change in our ability to detect antagonistic pleiotropy.
+Using the lfsr rather than the local false discovery rate (lfdr) is a critical change in our ability to detect alleles directly contributing to rank-changing GxE between environments.
 The lfdr, like other measures of FDR, focuses on if we have enough evidence to reject the null hypothesis that an effect j is 0, or that there is a significant effect.
@@
 This previous work recognized that this testing bias could lead to undercounting occurrences of antagonistic pleiotropy [@anderson2011evolutionary; @des2013genotype], and sought to reduce it by permutation [@anderson2013genetic].
-However, using the lfsr to test for antagonistic pleiotropy does not undercount occurrences of antagonistic pleiotropy, as this statistic answers a fundamentally different question.
+However, using the lfsr to test for allelic effects that differ in sign does not undercount these occurences, as this statistic answers a fundamentally different question.
 For each effect j, the lfsr~j~ is defined as the probability that we make an error in the sign of effect j if we were forced to declare the effect positive or negative @10.1093/biostatistics/kxw041.
@@
 The distinction between effects with different magnitudes is arbitrary but useful to fully characterize how effects vary across environments to ultimately influence phenotypes.
-Our use of the lfsr to determine significance and our specification that SNP effects must be significant in both conditions to be included means that our tests for antagonistic pleiotropy carry an equal statistical burden to those measuring differential sensitivity and effects without GxE.
+Our use of the lfsr to determine significance and our specification that SNP effects must be significant in both conditions to be included means that our tests for alleles with rank-changing GxE carry an equal statistical burden to those measuring differential sensitivity and effects without GxE.
 

### 2024-05-24 manuscript/revision2/revision2.tex
revision: Address Tom's comment about sign-changing GxE vs antagonistic pleiotropy

@@
 Juenger},
-  pdfkeywords={allele-by-environment effect variation, antagonistic
-pleiotropy, photoperiod, cumulative rainfall, genetic variation},
+  pdfkeywords={allele-by-environment effect variation, rank-changing
+genotype-by-environment interaction, photoperiod, cumulative
+rainfall, genetic variation},
   colorlinks=true,
@@
 
-\keywords{allele-by-environment effect variation | antagonistic
-pleiotropy | photoperiod | cumulative rainfall | genetic variation}
+\keywords{allele-by-environment effect variation | rank-changing
+genotype-by-environment interaction | photoperiod | cumulative
+rainfall | genetic variation}
 
@@
 cumulative rainfall in the seven days prior to flowering. However, most
-variation in genetic effects across sites is site-specific and can not
-be attributed to variation in weather variables alone. We demonstrate
-that we can identify genetic variation with GxWeather and assign these
-loci to specific weather-based cues or other patterns. Breeding for
-particular alleles at the GxWeather loci we identified could change
-flowering responsiveness in a photoperiod and rainfall-specific way.
-More broadly, our approach proposes a refined, powerful characterization
-of genotype-by-environment interactions in any species phenotyped in
+variation in genetic effects is site-specific and can not be attributed
+to variation in the studied weather variables. We demonstrate that we
+can identify genetic variation with GxWeather and assign these loci to
+specific weather-based cues or other patterns. Breeding for particular
+alleles at the GxWeather loci we identified could change flowering
+responsiveness in a photoperiod or rainfall-specific way. More broadly,
+our approach proposes a refined, powerful characterization of
+genotype-by-environment interactions in any species phenotyped in
 multiple environments.
@@
 theoretical predictions that local adaptation should involve
-antagonistic pleiotropy, or sign-changing GxE, at the level of
-individual loci (23--26), previous empirical work has found limited
-evidence of antagonistic pleiotropy (12, 27). However, this work has
-been limited by a known statistical bias that reduced detection of
-genetic effects that differ in sign (27--29). Thus, despite substantial
-interest in the frequencies of various forms of GxE, the prevalence of
-antagonistic pleiotropy relative to other forms of GxE remains unknown.
+antagonistic pleiotropy, or alleles with effects with opposing fitness
+outcomes (23--26), previous empirical work has found limited evidence of
+antagonistic pleiotropy (12, 27). However, this work has been limited by
+a known statistical bias that reduced detection of genetic effects that
+differ in sign (27--29). Thus, despite substantial interest in the
+frequencies of various forms of GxE, particularly sign-changing GxE,
+relative to other forms of GxE remains unknown.
 
@@
 Here, we estimate GxWeather for two phenological timings in switchgrass
-by assigning patterns of genetic effects on phenology across gardens to
-many patterns of weather covariance at these gardens. To do this, we
-phenotype a diversity panel of hundreds of switchgrass genotypes from
-the Midwest and Gulf subpopulations for the timing of vegetative and
-reproductive development at eight common garden locations spanning 17
-degrees of latitude. These gardens cover the majority of the latitudinal
-and climatic range of switchgrass and capture the most comprehensive
-picture to date of the environmental variation this species encounters.
-We define multiple ways phenological timings might covary with weather
-(Table 1) and additional ways phenological timings might vary by site
-(SI Appendix, Table S1, Section SX), then jointly re-estimate genetic
-effects on these timings at all eight sites using the set of these
-covariance matrices that significantly improved the modeled
-log-likelihood when included (SI Appendix, Section SX) (36). We use the
-Bayesian framework \emph{mash} (multivariate adaptive shrinkage)
-developed by (36), to refine effect size estimates from genome-wide
-association (GWAS) conducted on individuals at each of the eight sites.
-\emph{mash} allows us to identify and specify multiple covariance
-structures among genetic effect estimates across sites, including
-structures that represent covariance in weather variables of interest.
-Importantly, this method circumvents statistical biases in detecting
-genetic effects with the same or opposite signs (37). To confirm our
-genetic mapping of GxWeather, we compare our posterior genetic effect
-estimates from \emph{mash} to mapping results from an outbred pseudo-F2
-cross grown at the same sites. Our analyses allow us to describe the
-weather cues and genetic variation affecting phenology in two divergent
-natural populations of switchgrass.
+by assigning patterns of genetic effects on phenology for cloned
+individuals grown at eight common gardens to many patterns of weather
+covariance at these gardens. To do this, we phenotype a diversity panel
+of hundreds of switchgrass genotypes from the Midwest and Gulf
+subpopulations for the timing of vegetative and reproductive development
+at eight common garden locations spanning 17 degrees of latitude. These
+gardens cover the majority of the latitudinal and climatic range of
+switchgrass and capture the most comprehensive picture to date of the
+environmental variation this species encounters. We define multiple ways
+phenological timings might covary with weather (Table 1) and additional
+ways phenological timings might vary by garden (SI Appendix, Section
+S1), then jointly re-estimate genetic effects on these timings at all
+eight common gardens using the set of these covariance matrices that
+significantly improved the modeled log-likelihood when included (SI
+Appendix, Section S2) (36). We use the Bayesian framework \emph{mash}
+(multivariate adaptive shrinkage) developed by (36), to refine effect
+size estimates from genome-wide association (GWAS) conducted on
+individuals at each of the eight gardens. \emph{mash} allows us to
+identify and specify multiple covariance structures among genetic effect
+estimates across sites, including structures that represent covariance
+in weather variables of interest. Importantly, this method circumvents
+statistical biases in detecting genetic effects with the same or
+opposite signs (37). To confirm our genetic mapping of GxWeather, we
+compare our significant posterior genetic effects locations from
+\emph{mash} to mapping results from an outbred mapping population grown
+at the same sites. Our analyses allow us to describe the weather cues
+and genetic variation affecting phenology in two divergent natural
+populations of switchgrass across the species' latitudinal range.
 
@@
 We next looked for evidence of GxWeather by jointly re-estimating
-genetic effects across all eight common gardens using the \emph{mash}
-model, which can flexibly capture many modes of covariance across
-gardens (Figure~\ref{fig-map} D; SI Appendix, Datasets 1-6). We
+genetic effects of SNPs across all eight common gardens using the
+\emph{mash} model, which can flexibly capture many modes of covariance
+across gardens (Figure~\ref{fig-map} D; SI Appendix, Datasets 1-6). We
 specified three qualitative categories of covariance structures across
@@
 weather variables as cues for the start of vegetative and reproductive
-growth. Say that a SNP in FLC controls flowering in a
+growth. Say that a SNP in the gene \emph{Flowering Locus C} (FLC), a
+well-known flowering time regulator, controls flowering in a
 photoperiod-dependent manner. In that case, the joint estimate of
@@
 heterozygosity, with intermediate, positive covariances between all
-gardens, and garden-specific effects present only at one garden.
+gardens, and single effect matrices with garden-specific effects (SI
+Appendix, Section S1).
 
@@
 difference may be comprised of SNP effects that differ in sign between
-these gardens (effects with antagonistic pleiotropy), SNP effects that
-differ in magnitude (effects that are large in one garden or region and
-smaller in others), and SNP effects that are indistinguishable in the
-two gardens (Figure~\ref{fig-effects} A). We used the local false sign
-rate (lfsr), an analogue of the lfdr that establishes confidence in the
-effect sign, not the effect's difference from zero, to determine
-significance. We required lfsr significance (p \textless{} 0.05) in both
-gardens to include effects. This means that our tests for antagonistic
-pleiotropy, or a sign change between conditions, carry an equal
-statistical burden to those for effects with the same sign.
+these gardens (effects with rank-changing GxE between gardens), SNP
+effects that differ in magnitude (effects that are large in one garden
+or region and smaller in others), and SNP effects that are
+indistinguishable in the two gardens (Figure~\ref{fig-effects} A). We
+used the local false sign rate (lfsr), an analogue of the lfdr that
+establishes confidence in the effect sign, not the effect's difference
+from zero, to determine significance. We required lfsr significance (p
+\textless{} 0.05) in both gardens to include effects. This means that
+our tests for a sign change between gardens carry an equal statistical
+burden to those for effects with the same sign.
 
 For green-up date for the Gulf subpopulation, hundreds to thousands of
-pairwise effects exhibited antagonistic pleiotropy, or a difference in
-effect sign, between pairs of Texas and North gardens
-(Figure~\ref{fig-effects} B; SI Appendix, Fig SX). 78.7\% of pairwise
-comparisons between North and Texas gardens had a difference in sign,
-while only 28.6\% and 0.2\% of North-North or Texas-Texas comparisons
-had a difference in sign, respectively (Figure~\ref{fig-effects} B). The
-majority of pairwise effects for greenup for the Midwest
-(\textgreater55\%) and Both (\textgreater85\%) subpopulations were the
-same sign, and effects most often differed in magnitude between gardens
-within and between the regions (Figure~\ref{fig-effects} B).
+pairwise effects exhibited a difference in effect sign, between pairs of
+Texas and North gardens (Figure~\ref{fig-effects} B; SI Appendix, Fig
+SX). 78.7\% of pairwise comparisons between North and Texas gardens had
+a difference in sign, while only 28.6\% and 0.2\% of North-North or
+Texas-Texas comparisons had a difference in sign, respectively
+(Figure~\ref{fig-effects} B). The majority of pairwise effects for
+greenup for the Midwest (\textgreater55\%) and Both (\textgreater85\%)
+subpopulations were the same sign, and effects most often differed in
+magnitude between gardens within and between the regions
+(Figure~\ref{fig-effects} B).
 
 For flowering date for the Gulf subpopulation, less than 2\% of pairwise
-effects exhibited antagonistic pleiotropy, or a difference in effect
-sign, within or between regions (Figure~\ref{fig-effects} B). More
-effects differed in magnitude between the Texas and North regions than
-within these regions (42.7\% vs \textless20\%; Figure~\ref{fig-effects}
-B), while the majority of effects had no GxE. The Midwest population had
-relatively few significant effects for flowering, but a large proportion
-of these differed in sign between Texas and North regions (42.7\%) or
-within the North region (65.4\%). Finally, in Both subpopulations, less
-than 20\% of pairwise effects differed in sign (Figure~\ref{fig-effects}
-B). Most differences in sign were between TX1, the southernmost garden,
-and all other gardens (SI Appendix, Fig SX). Similarly, more effects
-that differed in magnitude included gardens in the Texas region
+effects exhibited a difference in effect sign within or between regions
+(Figure~\ref{fig-effects} B). More effects differed in magnitude between
+the Texas and North regions than within these regions (42.7\% vs
+\textless20\%; Figure~\ref{fig-effects} B), while the majority of
+effects had no GxE. The Midwest population had relatively few
+significant effects for flowering, but a large proportion of these
+differed in sign between Texas and North regions (42.7\%) or within the
+North region (65.4\%). Finally, in Both subpopulations, less than 20\%
+of pairwise effects differed in sign (Figure~\ref{fig-effects} B). Most
+differences in sign were between TX1, the southernmost garden, and all
+other gardens (SI Appendix, Fig SX). Similarly, more effects that
+differed in magnitude included gardens in the Texas region
 (52.3-55.9\%), and most effect pairs in the North region were not
@@
 
-We sought additional experimental support for our joint re-estimates of
-SNP effects using an independent pseudo-F2 mapping population created
-from Gulf \& Midwest individuals and grown at the same sites
-(Figure~\ref{fig-qtl} A,B). We conducted quantitative trait loci (QTL)
-mapping of flowering as functions of five environmental cues that we
-also used as covariance matrices in \emph{mash}, and identified eight
-QTL for flowering date, eight QTL for flowering as a function of day
-length change two days prior, one QTL for the start of vegetative
-growth, and two QTL for vegetative growth as a function of daylength
-change one day prior, all of which showed QTL by environment
+We sought additional experimental support the occurrence and location of
+our significant SNP effect re-estimates using an independent pseudo-F2
+mapping population created from Gulf \& Midwest individuals and grown at
+the same sites (Figure~\ref{fig-qtl} A,B). We conducted quantitative
+trait loci (QTL) mapping of flowering as functions of five environmental
+cues that we also used as covariance matrices in \emph{mash}, and
+identified eight QTL for flowering date, eight QTL for flowering as a
+function of day length change two days prior, one QTL for the start of
+vegetative growth, and two QTL for vegetative growth as a function of
+daylength change one day prior, all of which showed QTL by environment
 interactions (SI Appendix, Fig. S3). All QTL for flowering overlapped
@@
 effects to both GxWeather patterns with interpretable weather-based
-cues, and to agnostic, site-based patterns. We use this approach to
+cues, and to unmeasured, site-based patterns. We use this approach to
 study GxWeather for the timings of vegetative and reproductive
@@
 Our analysis of the timing of vegetative growth in the Gulf and in Both
-subpopulations revealed substantial antagonistic pleiotropy in effects
-between the Texas and North gardens (Figure~\ref{fig-effects} B). This
-result supports theoretical models that local adaptation should involve
-antagonistic pleiotropy at the level of individual loci (23--26), and is
-the first experimental work using GWAS across common gardens to find
-antagonistic pleiotropy to be common in small genomic regions (12, 38,
-39).
+subpopulations revealed many alleles with sign changes in their effects
+between the Texas and North gardens (Figure~\ref{fig-effects} B). As
+phenological timings are major components of plant fitness, this result
+supports theoretical models that local adaptation should involve
+antagonistic pleiotropy at the level of individual loci (23--26); it
+expands the genetics of local adaptation research to more than two field
+sites and to a wider range of genetic variation by using effects from
+GWAS in common gardens located throughout the species' range to
+determine that rank-changing GxE for phenological timings is common in
+small genomic regions (12, 38, 39).
 
@@
 subpopulations have distinct GxWeather: flowering timing in the Midwest
-subpopulation has photoperiod-related variation, in that flowering
-timing covaries with a day length change signal two days before
-flowering occurs. In contrast, the Gulf subpopulation does not have
-variation in flowering based on a photoperiod cue. Instead, the Gulf
-subpopulation has variation in flowering that covaries with the rainfall
-that occurs in the week prior to flowering. Three genomic regions
-affecting flowering that we re-estimated across all eight sites were
-also supported by QTL from an independent mapping population at these
-sites (Figure~\ref{fig-qtl} B).
+subpopulation has photoperiod-related genetic variation, in that
+flowering timing covaries with a day length change signal two days
+before flowering occurs. In contrast, the Gulf subpopulation does not
+have genetic variation in flowering that covaries with a photoperiod
+cue. Instead, the Gulf subpopulation has genetic variation in flowering
+that covaries with the rainfall that occurs in the week prior to
+flowering. Three genomic regions affecting flowering that we
+re-estimated across all eight sites were also supported by QTL from an
+independent mapping population at these sites (Figure~\ref{fig-qtl} B).
 
@@
 
-Last, we characterized the overall patterns of antagonistic pleiotropy
-in the set of SNPs where there was pairwise significance of effects at
-pairs of gardens. To do this, we used the `get\_GxE' function of the
-switchgrassGWAS R package. First, this determines the set of SNPs with
-evidence of significant effects in both conditions for all pairs of
-conditions using local false sign rates (lfsr) as the significance
-criteria. Second, to determine antagonistic pleiotropy, this function
-determines if effects significant in both conditions are of opposite
-sign.
+Last, we characterized the overall patterns of GxE in the set of SNPs
+where there was pairwise significance of effects at pairs of gardens. To
+do this, we used the `get\_GxE' function of the switchgrassGWAS R
+package. First, this determines the set of SNPs with evidence of
+significant effects in both conditions for all pairs of conditions using
+local false sign rates (lfsr) as the significance criteria. Then, this
+function determines if effects significant in both conditions are of
+opposite sign.
 
 Using the lfsr rather than the local false discovery rate (lfdr) is a
-critical change in our ability to detect antagonistic pleiotropy. The
-lfdr, like other measures of FDR, focuses on if we have enough evidence
-to reject the null hypothesis that an effect j is 0, or that there is a
-significant effect. Previous studies of antagonistic pleiotropy (e.g.
-(39)) have used the lfdr or equivalent statistical tests to detect
-antagonistic pleiotropy. These tests were conservative, in that they
-required two non-zero effects of different signs, while tests for
-differential sensitivity required only one non-zero effect. This
-previous work recognized that this testing bias could lead to
-undercounting occurrences of antagonistic pleiotropy (27, 29), and
-sought to reduce it by permutation (28). However, using the lfsr to test
-for antagonistic pleiotropy does not undercount occurrences of
-antagonistic pleiotropy, as this statistic answers a fundamentally
-different question. For each effect j, the lfsr\textsubscript{j} is
-defined as the probability that we make an error in the sign of effect j
-if we were forced to declare the effect positive or negative (37). Thus,
-rather than asking ``Are these two effects different?'' - as we
-reasonably expect two effects to be, even if this difference cannot be
-measured - the local false sign rate answers a more meaningful question:
-Can we be confident in the sign of this effect?
+critical change in our ability to detect alleles directly contributing
+to rank-changing GxE between environments. The lfdr, like other measures
+of FDR, focuses on if we have enough evidence to reject the null
+hypothesis that an effect j is 0, or that there is a significant effect.
+Previous studies of antagonistic pleiotropy (e.g. (39)) have used the
+lfdr or equivalent statistical tests to detect antagonistic pleiotropy.
+These tests were conservative, in that they required two non-zero
+effects of different signs, while tests for differential sensitivity
+required only one non-zero effect. This previous work recognized that
+this testing bias could lead to undercounting occurrences of
+antagonistic pleiotropy (27, 29), and sought to reduce it by permutation
+(28). However, using the lfsr to test for allelic effects that differ in
+sign does not undercount these occurences, as this statistic answers a
+fundamentally different question. For each effect j, the
+lfsr\textsubscript{j} is defined as the probability that we make an
+error in the sign of effect j if we were forced to declare the effect
+positive or negative (37). Thus, rather than asking ``Are these two
+effects different?'' - as we reasonably expect two effects to be, even
+if this difference cannot be measured - the local false sign rate
+answers a more meaningful question: Can we be confident in the sign of
+this effect?
 
@@
 effects must be significant in both conditions to be included means that
-our tests for antagonistic pleiotropy carry an equal statistical burden
-to those measuring differential sensitivity and effects without GxE.
+our tests for alleles with rank-changing GxE carry an equal statistical
+burden to those measuring differential sensitivity and effects without
+GxE.
 

### 2024-06-05 R/SI_Appendix_Datasets.Rmd
revision2: Add Dataset 7 and 8 to SI Appendix material

@@
 
-# Dataset 8
+```{r}
+if (!require("BiocManager", quietly = TRUE))
+    install.packages("BiocManager")
 
-## QTL
+BiocManager::install("AnnotationDbi")
+BiocManager::install("GenomicFeatures")
+BiocManager::install("VariantAnnotation")
+```
+
+Find: `qtlplot` in Analysis_v0.6_QTL_interval_plotting. Really, find code to remake this with new QTL from revision2.
+
+Find this file on NAS: "~/Github/pvdiv-genome/Pvirgatum_516_v5.1.gene.txdb.sqlite"
+
+Couldn't find it on the NAS, scp'd it from the Juenger Lab Server:
+`scp -i ~/Documents/server_keys/jls_alice alice@popgen.biosci.utexas.edu:/home/alice/Github/pvdiv-genome/Pvirgatum_516_v5.1.gene.txdb.sqlite ~/Documents/Github/pvdiv-phenology-gxe/data-raw/ignore/`
+
+It's too big to commit; had to add it to data-raw/ignore files ignored by .gitignore.
+
 ```{r}
-qtldf <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv")
+qtldf <- read_csv(here("phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv"))
 
@@
 qtlplot <- qtldf %>%
-  mutate(phe = case_when(lodcolumn == "FL50" ~ "50% flowering",
+  mutate(phe = case_when(lodcolumn == "FL50" ~ "Flowering Date",
                          lodcolumn == "dyln_fl50" ~ "Daylength",
-                              lodcolumn == "dyln_change_sec" ~ "daylength change (s)",
-                              lodcolumn == "CGDD_12C" ~ "GDD, GR to FL",
+                         lodcolumn == "dyln_change_sec" ~ 
+                           "daylength change (s)",
+                         lodcolumn == "dyln_change_sec_2d_prior" ~ 
+                           "daylength change, 2 days prior (s)",
+                         lodcolumn == "GR50" ~ "Greenup Date",
                          lodcolumn == "crain_gr2fl" ~ "Rainfall, GR to FL")) %>%
@@
 qtlplot$phe <- factor(qtlplot$phe, 
-                      levels = c("50% flowering", "Rainfall, GR to FL",
+                      levels = c("Flowering Date", "Rainfall, GR to FL",
                                  "GDD, GR to FL", "Daylength",
                                  "daylength change (s)",
+                                 "daylength change, 2 days prior (s)",
                                  "Rainfall, 1 day",
+                                 "Greenup Date",
                                                   #"Rainfall, 2d sum",
@@
 ```{r}
+library(AnnotationDbi)
+library(GenomicFeatures)
+library(switchgrassGWAS)
+qtl_anno_df <- qtlplot %>%
+  dplyr::rename(start = POS_lo, end = POS_hi)
+not_all_na <- function(x) any(!is.na(x))
+txdb <- loadDb("~/Documents/Github/pvdiv-phenology-gxe/data-raw/ignore/Pvirgatum_516_v5.1.gene.txdb.sqlite")
+# source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+
+ftanno <- read_csv(here("data/Pvirgatum_v5.1.annotation_FTGeneHomologs.csv"))
+ftkey <- read_delim(here("data", "OsGene_keyword_table.txt"),
+                    delim = "\t")
+ftanno %>%
+  filter(!is.na(OsGene)) %>%
+  left_join(ftkey, by = c("OsGene" = "Symbol")) %>%
+  filter(Keyword %in% c("flowering time", "heading date")) %>% unique()
+
+osgene_keyword_df <- ftkey %>%
+  dplyr::select(-RAPdb, -MSU) %>%
+  group_by(Symbol, Keyword) %>%
+  dplyr::slice(1) %>%
+  pivot_wider(names_from = Keyword, values_from = Title)
+
+osgene_key <- osgene_keyword_df %>%
+  dplyr::select(Symbol, `flowering time`, `heading date`, anther, flower,
+                `floral meristem`, `floral meristem determinacy`, flowering,
+                `flower development`, `carpel differentiation`, stamen, 
+                `stamen number`, vegetative, `plant growth`, `chloroplast development`, chloroplast) %>%
+  pivot_longer(cols = -Symbol, names_to = "Keyword", values_to = "Title") %>%
+  filter(!is.na(Title)) %>%
+  pivot_wider(names_from = Keyword, values_from = Title)
+
+os_annos <- ftanno %>%
+  filter(OsGene %in% osgene_key$Symbol | !is.na(AtGene)) %>%
+  mutate(Pos_Mb = round(start/10000)/100,
+         #Pos_Mb_e = floor(start/10000)/100,
+         POS_bin = round(Pos_Mb*100/2)) %>%# ,
+         #POS_bin_hi = round(Pos_Mb_e*100/2))
+  dplyr::select(OsGene, AtGene, locusName, everything()) %>%
+  dplyr::rename(`Gene ID` = locusName)
+
+qtl_annos <- switchgrassGWAS::pvdiv_table_topsnps(df = qtl_anno_df, type = "table", txdb = txdb) %>%
+    inner_join(os_annos, by = c("Gene ID")) %>%
+    left_join(osgene_key, by = c("OsGene" = "Symbol")) %>%
+    dplyr::select(where(not_all_na)) %>%
+    group_by(CHR, pos, `Gene ID`) %>%
+    dplyr::slice(1)
+
+names(qtl_annos)
+all_na_gr_fl <- rowSums(is.na(qtl_annos[,59:70]))
+add_column(qtl_annos, GR_FL_Osgene_NA = all_na_gr_fl, .before = TRUE) %>%
+  group_by(CHR, pos, GR_FL_Osgene_NA < 12) %>%
+  tally()
+flowering_qtl_annos <- add_column(qtl_annos, GR_FL_Osgene_NA = all_na_gr_fl, .before = TRUE) %>%
+  filter(GR_FL_Osgene_NA < 12) %>%
+  dplyr::select(CHR, pos, phe, lod, Pos_Mb_lo, Pos_Mb_hi, ci_lo, ci_hi, `Gene ID`,  gene_start:gene_width, `Arabidopsis thaliana homolog`:`Rice gene annotation`, OsGene, anther:`floral meristem determinacy`) |>
+  mutate(
+    Phenology_type = case_when(phe %in% 
+                                 c("Flowering Date", 
+                                   "daylength change, 2 days prior (s)") ~
+                                 "Flowering Date",
+                               phe %in% c("Greenup Date", 
+                                          "daylength change (s)") ~ 
+                                 "Greenup Date",
+                               TRUE ~ NA_character_
+                               ),
+    .before = 3)
+
+flowering_qtl_annos |> 
+  filter(Phenology_type == "Greenup Date")
+
+flowering_qtl_annos |> 
+  arrange(Phenology_type) |> 
+  write_csv(here("manuscript/revision2/SI_Appendix/Dataset_7_QTL_intervals_with_functionally_validated_homologs.csv"))
+```
+
+# Dataset 8
+All flowering and green-up QTL intervals contained at least one SNP significant in at least one mash run at a log10-transformed Bayes Factor > 2, or in the 1% tail of significance, whichever was stricter.
+
+## Objects generated earlier
+Get `qtlplot` object by running code starting at about line 150 (for Dataset 7).
+Get `m_gulf_fl_hyp` and other objects by running code starting at about line 35 (for Datasets 1-6).
+
+## Thresholds
+Decide on cutoff to use for rug plots and enrichment analysis: SNPs sig in 1% tail or BF\>2, whichever is stricter
+
+1% tail is stricter in all cases except Midwest flowering, where BF\>2 is stricter, equivalent to a 0.49% tail (0.995119 quantile).
+
+What is the Bayes Factor at the 99th quantile? If it's not above 2, what is the quantile where the BF \> 2?
+
+```{r}
+#quantile(pw_gf$log10BayesFactor, 0.99)  # Gulf: Use 2.56 BF
+#quantile(pw_mf$log10BayesFactor, 0.99)  # Midwest: Use 2 BF
+#quantile(pw_gamf$log10BayesFactor, 0.99)  # GAM: Use 2.65 BF
+#quantile(pw_mf$log10BayesFactor, 0.995119)
+
+#quantile(pw_gg$log10BayesFactor, 0.99)  # Gulf: Use 2.0989 BF
+#quantile(pw_mg$log10BayesFactor, 0.99)  # Midwest: Use 4.6831 BF
+#quantile(pw_gamg$log10BayesFactor, 0.99)  # GAM: Use 3.0865 BF
+```
+
+Now there are significant greenup date QTL, separate QTL & mash into greenup and flowering associations, and look for overlaps just within each phenology type.
+
+```{r}
 m <- m_gulf_fl_hyp
-qtl_df <- qtlplot
+qtl_df <- qtlplot |>   # split by Phenology type. Define based on Figure 4.
+  mutate(
+    Phenology_type = case_when(phe %in% 
+                                 c("Flowering Date", 
+                                   "daylength change, 2 days prior (s)") ~
+                                 "Flowering Date",
+                               phe %in% c("Greenup Date", 
+                                          "daylength change (s)") ~ 
+                                 "Greenup Date",
+                               TRUE ~ NA_character_
+                               ),
+    .before = 3)
+qtl_fl <- qtl_df |> 
+  filter(Phenology_type == "Flowering Date")
+qtl_gr <- qtl_df |> 
+  filter(Phenology_type == "Greenup Date")
+
 subpop = "Gulf"
-phe = "flowering"
+phe = "Flowering Date"
 
-qtl_mash_overlap_function <- function(m, subpop, phe){
+qtl_mash_overlap_function <- function(qtl, m, subpop, phe, onepcttail = 2){
 
-  mash_df <- switchgrassGWAS:::get_marker_df(m) %>%
-    mutate(log10BF = switchgrassGWAS:::get_log10bf(m)) %>%
+  mash_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
+      as.data.frame() %>%
+      dplyr::mutate(value = row_number()) %>%
+      as_tibble() %>%
+      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
+      dplyr::rename(log10BayesFactor = .data$V1) %>%
+      dplyr::select(-.data$value) |> 
+  ## TODO: FIX list/matrix column here
     tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, 
                     sep = "_", extra = "merge", convert = TRUE) %>%
-    dplyr::arrange(.data$Chr, .data$Pos) %>%
-    select(-value)
+    dplyr::arrange(.data$Chr, .data$Pos)
 
-  for(i in 1:nrow(qtl_df)){  # from 1 to 16
+  for(i in 1:nrow(qtl)){  # from 1 to 16
     ## In block & > 2
     qtl_sub <- mash_df %>%
-      filter(Chr %in% qtl_df$CHR[i] & between(Pos, qtl_df$POS_lo[i],
-                                               qtl_df$POS_hi[i])) %>% 
-      filter(log10BF > 2)
+      filter(Chr %in% qtl$CHR[i] & between(Pos, qtl$POS_lo[i],
+                                               qtl$POS_hi[i])) %>% 
+      filter(log10BayesFactor > 2)
 
     if(i == 1){
-    outputdf <- tibble(Chr = qtl_df$CHR[i], QTL_Pos_lo = qtl_df$POS_lo[i], 
-                       QTL_Pos_hi = qtl_df$POS_hi[i], 
-                       QTL_phenotype = qtl_df$lodcolumn[i], 
-                       QTL_LOD = qtl_df$lod[i], mash_Pos = qtl_sub$Pos, 
-                       mash_log10BF = qtl_sub$log10BF, mash_subpop = subpop, 
+    outputdf <- tibble(Chr = qtl$CHR[i], QTL_Pos_lo = qtl$POS_lo[i], 
+                       QTL_Pos_hi = qtl$POS_hi[i], 
+                       QTL_phenotype = qtl$lodcolumn[i], 
+                       QTL_LOD = qtl$lod[i], mash_Pos = qtl_sub$Pos, 
+                       mash_log10BayesFactor = qtl_sub$log10BayesFactor, mash_subpop = subpop, 
                        mash_phenotype = phe)
     } else {
-      outputdf1 <- tibble(Chr = qtl_df$CHR[i], QTL_Pos_lo = qtl_df$POS_lo[i], 
-                          QTL_Pos_hi = qtl_df$POS_hi[i], 
-                          QTL_phenotype = qtl_df$lodcolumn[i], 
-                          QTL_LOD = qtl_df$lod[i], mash_Pos = qtl_sub$Pos, 
-                          mash_log10BF = qtl_sub$log10BF, mash_subpop = subpop, 
+      outputdf1 <- tibble(Chr = qtl$CHR[i], QTL_Pos_lo = qtl$POS_lo[i], 
+                          QTL_Pos_hi = qtl$POS_hi[i], 
+                          QTL_phenotype = qtl$lodcolumn[i], 
+                          QTL_LOD = qtl$lod[i], mash_Pos = qtl_sub$Pos, 
+                          mash_log10BayesFactor = qtl_sub$log10BayesFactor,
+                          mash_subpop = subpop, 
                           mash_phenotype = phe)
@@
   }
+  outputdf <- outputdf |> 
+    mutate(
+      Marker_in_mash_1pct_tail = 
+        case_when(mash_log10BayesFactor >= onepcttail ~ TRUE,
+                  mash_log10BayesFactor < onepcttail ~ FALSE,
+                  TRUE ~ NA)
+    )
   return(outputdf)
@@
 
-overlap1 <- qtl_mash_overlap_function(m_gulf_fl_hyp, subpop = "Gulf", 
-                                      phe = "flowering") 
-overlap2 <- qtl_mash_overlap_function(m_midw_fl_hyp, subpop = "Midwest", 
-                                      phe = "flowering") 
-overlap3 <- qtl_mash_overlap_function(m_gam_fl_hyp, subpop = "Both", 
-                                      phe = "flowering")
+overlap1 <- qtl_mash_overlap_function(qtl_fl, m_gulf_fl_hyp, 
+                                      subpop = "Gulf", 
+                                      phe = "Flowering Date",
+                                      onepcttail = 2.56) 
+overlap2 <- qtl_mash_overlap_function(qtl_fl, m_midw_fl_hyp, 
+                                      subpop = "Midwest", 
+                                      phe = "Flowering Date",
+                                      onepcttail = 2) 
+overlap3 <- qtl_mash_overlap_function(qtl_fl, m_gam_fl_hyp, 
+                                      subpop = "Both", 
+                                      phe = "Flowering Date",
+                                      onepcttail = 2.65) 
+overlap4 <- qtl_mash_overlap_function(qtl_gr, m_gulf_gr_hyp, 
+                                      subpop = "Gulf", 
+                                      phe = "Greenup Date",
+                                      onepcttail = 2.0989) 
+overlap5 <- qtl_mash_overlap_function(qtl_gr, m_midw_gr_hyp, 
+                                      subpop = "Midwest", 
+                                      phe = "Greenup Date",
+                                      onepcttail = 4.6831) 
+overlap6 <- qtl_mash_overlap_function(qtl_gr, m_gam_gr_hyp, 
+                                      subpop = "Both", 
+                                      phe = "Greenup Date",
+                                      onepcttail = 3.0865)
 
+overlap4 %>%
+  full_join(overlap5) %>%
+  full_join(overlap6) |> 
+  arrange(Chr, mash_Pos)
 dataset_8 <- overlap1 %>%
@@
   full_join(overlap3) %>%
-  arrange(Chr, mash_Pos)
+  full_join(overlap4) %>%
+  full_join(overlap5) %>%
+  full_join(overlap6) |> 
+  arrange(mash_phenotype, Chr, mash_Pos)
 
-write_csv(dataset_8, file = "~/Github/pvdiv-phenology-gxe/manuscript/Coauthors/PNAS/Dataset_8_QTL_mash_overlap.csv")
+write_csv(dataset_8, file = here("manuscript", "revision2", "SI_Appendix",
+                                 "Dataset_8_QTL_mash_overlap.csv"))
 ```
-
-Is there an enrichment of SNPs in the top 1% (5%? 10%?) of the distribution for each mash result in the QTL regions?
-1% is probably best out of these three arbitrary cutoffs.
-```{r}
-m <- m_gulf_fl_hyp
-qtl_df <- qtlplot
-
-get_QTL_enrichment <- function(m, qtl_df, quantile = 0.95){
-  pw_df_all <- get_post_weights(m)
-
-  BFthresh <- quantile(pw_df_all$log10BayesFactor, quantile)
-  pw_df <- pw_df_all %>%
-  filter(log10BayesFactor > BFthresh) %>%
-  select(Chr:Pos, log10BayesFactor, everything())
-
-  for(i in 1:nrow(qtl_df)){  # from 1 to 16
-  ## In block & > 2
-  QTLandBF <- pw_df %>%
-    filter(Chr %in% qtl_df$CHR[i] & between(Pos, qtl_df$POS_lo[i],
-                                             qtl_df$POS_hi[i])) %>% 
-    tally()
-  ## BF>2
-  inBF <- pw_df %>%
-    tally()
-  ## BF not > 2 (all minus BF > 2, all is below)
-  allwrtBF <- pw_df_all %>%
-    tally()
-  ## All in block
-  inQTL <- pw_df_all %>%
-    filter(Chr %in% qtl_df$CHR[i] & between(Pos, qtl_df$POS_lo[i],
-                                             qtl_df$POS_hi[i])) %>%
-    tally()
-  notinBF <- (allwrtBF$n[1]-inBF$n[1])
-  pval <- phyper(QTLandBF$n[1], inBF$n[1], allwrtBF$n[1]-inBF$n[1], 
-                 inQTL$n[1], lower.tail = FALSE)
-
-  if(i == 1){
-  outputdf <- tibble(Chr = qtl_df$CHR[i], Pos_lo = qtl_df$POS_lo[i], Pos_hi = qtl_df$POS_hi[i], inQTLandBF = QTLandBF$n[1], inBF = inBF$n[1], notinBF = notinBF, inQTL = inQTL$n[1], pvalue = pval)
-  } else {
-    outputdf <- outputdf %>%
-      add_row(Chr = qtl_df$CHR[i], Pos_lo = qtl_df$POS_lo[i], Pos_hi = qtl_df$POS_hi[i],inQTLandBF = QTLandBF$n[1], inBF = inBF$n[1], notinBF = notinBF, inQTL = inQTL$n[1], pvalue = pval)
-    }
-  }
-  return(outputdf)
-}
-
-get_post_weights <- function(m){
-  log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
-      as.data.frame() %>%
-      rownames_to_column(var = "value") %>%
-      dplyr::mutate(value = as.integer(.data$value)) %>%
-      as_tibble() %>%
-      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
-      dplyr::rename(log10BayesFactor = .data$V1) %>%
-      dplyr::select(-.data$value)
-
-  pw_df <- as_tibble(m$posterior_weights)
-  pw_df <- switchgrassGWAS:::get_marker_df(m = m) %>%
-    bind_cols(pw_df)
-
-  ggman_df <- pw_df %>%
-    #left_join(m_meta) %>%
-    tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
-             extra = "merge", convert = TRUE) %>%
-    dplyr::left_join(log10bf_df, by = "Marker") %>%
-    dplyr::arrange(.data$Chr, .data$Pos) 
-  return(ggman_df)
-}
-
-qtlplot <- qtlplot %>%
-  arrange(desc(lod)) %>%
-  mutate(pos = as.integer(pos)) %>%
-  group_by(CHR, pos) %>%
-  slice(1) %>%
-  arrange(CHR, pos)
-enrich_gf <- get_QTL_enrichment(m_gulf_fl_hyp, qtl_df = qtlplot, 0.99) %>%
-  mutate(subpop = "Gulf", phenotype = "flowering")
-enrich_mf <- get_QTL_enrichment(m_midw_fl_hyp, qtl_df = qtlplot, 0.99) %>%
-  mutate(subpop = "Midwest", phenotype = "flowering")
-enrich_gamf <- get_QTL_enrichment(m_gam_fl_hyp, qtl_df = qtlplot, 0.99) %>%
-  mutate(subpop = "Both", phenotype = "flowering")
-
-enrich_df <- enrich_gf %>%
-  full_join(enrich_mf) %>%
-  full_join(enrich_gamf) %>%
-  arrange(Chr, Pos_lo)
-enrich_df <- enrich_df %>%
-  left_join(qtlplot, by = c("Chr" = "CHR", "Pos_lo" = "POS_lo", "Pos_hi" = "POS_hi")) %>%
-  select(Chr, Pos_lo, Pos_hi, inQTLandBF:subpop, lod, lodcolumn, pos) %>%
-  unique() %>%
-  arrange(Chr, Pos_lo, lodcolumn)
-
-enrich_df %>%
-  filter(pvalue < 0.05)
-
-t.test(enrich_df[enrich_df$pvalue < 0.05,]$lod, enrich_df[enrich_df$pvalue >= 0.05,]$lod) 
-qtlplot %>% group_by(CHR, POS_lo) %>% tally() # 21-24 unique QTL by Pos lo/hi. 23 by QTL maximum. 12 of 24 have enriched mash hits.
-
-```
-
-16 QTL. In all cases there was at least one SNP in the QTL interval that was in the 1% tail of SNPs in one of the three flowering mash runs.
-We also looked for enrichment of SNPs in the 1% tail of significance in the mash distribution within each QTL interval.
-(in random genomic intervals of the same size do we see a similar level of enrichment?)
-Half of these QTL had a significant enrichment of significant results for mash compared to expectations given the number of SNPs in the QTL interval.
-
-2K: 2 QTL (lod 14 & 17), 1 enriched (in Midwest) (lod 17: CGDD_12C)
-2N: 4 QTL (3 non-overlapping) (lod 19, 39, 17, 15), 2 enriched (lod 39 dylnfl50 in Midwest, in Both & 15 CGDD_12C in all three)
-4K: 2 QTL, 2 enriched (lod 40 dyln in Midwest & 33 dyln in Gulf)
-5N: 4 QTL (3 non-overlapping) (lod 55, 25, 17, 33), 3 enriched (2 non-overlapping) (lod 25 dyln in Midwest, 17 CGDD in Gulf, in Midwest, 33 dyln in Gulf, in Both)
-7K: 2 QTL, 0 en (lod 14, 13)
-8N: 1 QTL, 0 en (lod 15)
-9N: 1 QTL, 0 en (lod 18)

### 2024-06-05 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
revision2: Add Dataset 7 and 8 to SI Appendix material

@@
 
+A trait measured in two environments can be considered two traits that are genetically correlated [@falconer1952].
+By extension, a phenological trait measured at eight common gardens can be modeled as eight genetically correlated traits represented by a covariance matrix, where diagonal entries represent narrow-sense heritability at a common garden, or the additive genetic variance present in the population at that garden, and non-diagonal entries represent the genetic correlation between the trait in a pair of environments.
+
+We were interested in the types of genetic correlation that could be expressed by different alleles across the genome and reasoned that different loci could have different patterns of genetic correlation.
+For example, one locus's effects could covary with a weather-based cues, if it had an allele that was responsive to this cue.
+Alternatively, a locus might have an effect at only one garden, for example if a herbivore or pathogen was only present at one garden.
+
 We specified three qualitative categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data and introduced in the initial *mash* manuscript, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden over specific time frames before the phenological event.
 
-Canonical covariance matrices can be created by *mash* and fall into four groups: an identity matrix, an equal effects matrix, singleton matrices, and simple heterozygousity matrices.
+Canonical covariance matrices were created by *mash* and fell into four groups: an identity matrix, an equal effects matrix, singleton matrices, and simple heterozygousity matrices.
 The identity matrix has values of zero for covariance between different gardens and correlations of one within gardens, and represents the situation in which the effects in different conditions are independent.
@@
 
-We defined distinct sets of weather-based covariance matrices for the two phenological dates and for the three genetic subgroups.
+We defined distinct sets of GxWeather covariance matrices for the two phenological dates and for the three genetic subgroups.
 We used the same five weather variables to transform both green-up and flowering date, but defined the weather-based covariance matrices relative to each of the two phenological dates.
-Weather-based covariance matrices were also defined separately for individual genotypes from the Gulf, Midwest, and Both subpopulations.
+GxWeather covariance matrices were also defined separately for individual genotypes from the Gulf, Midwest, and Both subpopulations.
 
 ## Section S2. Greedy mash algorithm used to select covariance matrices that significantly improved the model log-likelihood {.unnumbered}
+
+We generated 48 - 60 GxWeather covariance matrices for each of our five weather variables over different time frames. These GxWeather covariance matrices were in many cases highly correlated.
+
+selection of the GxWeather
+
+> describing that the weather-based cues are themselves correlated, and picking and choosing among weather cues is a model building problem – to address the issue, we used a heuristic greedy algorithm to pick the weather-based covariance structures that improved the likelihood of the mash model
+
+> A lot of the results interpretation is based on interpreting the SNP loadings on the specified covariance matrices.
+This is a secondary use of mash (the primary being the refinement of effect sizes), and while they do interpret these somewhat in the mash paper, these loadings need to be interpreted with caution.
+If these covariance matrices are similar, mash will somewhat arbitrarily assign weight to any of the matrices because all lead to the same fit to the effect-size data.
+The mash algorithm doesn’t give the full posterior distribution on the loadings so you can’t check for posterior correlations there.
+It seems likely to me that the hypothesis-driven and data-driven covariance matrices are somewhat correlated here, and the correlations may differ between green-up and flowering because of the better correlation between the environmental metrics and flowering than green-up.
 
@@
 
+\dataset{Dataset_7_QTL_intervals_with_functionally_validated_homologs.csv}{Genes in the quantitative trait loci (QTL) regions shown in Figure 4 that have functionally validated homologs in rice for vegetative or reproductive development.
+The first 9 columns of this table provide QTL information, the remainder provide annotation information for the switchgrass genes in the QTL interval with homologs in A.
+thaliana or rice that have functional validation in flowering or vegetative-growth related traits.
+Columns 20-30 provide the titles of papers providing functional validation of the rice homolog of the switchgrass gene.}
+
+\dataset{Dataset_8_blah.csv}{Overlap between quantitative trait loci (QTL) for the start of vegetative or reproductive growth with significant mash effect estimates (with a log10(Bayes Factor) > 2) found using diversity panels of the Midwest, Gulf, and Both genetic subpopulations.
+The first five columns of this table provide QTL information, and columns 6-10 provide mash Marker information for markers within the QTL regions.}
+
 # References

### 2024-06-06 R/Analysis_v1.2_mash_using_greedy_covar.qmd
revision2: Add SI Appendix Sections on BLUPs, GWAS, SNP subset selection for mash

@@
 The GWAS were run in R/Plots_v0.3_manuscript_figures.Rmd, starting at line 1243.
-I copy-pasted that info here to have the full workflow in one qmd document; you may need to refer to it if you need additional information.
+I copy-pasted that info here to have the full workflow in one qmd document; you may need to refer to the Rmd file if you need additional information.
+
+### BLUPs
+
+Copied this code from R/Plots_v0.2_manuscript_figures.Rmd, which is where the ASREML code to calculate BLUPs individually for each location x subpop x phenotype came from; you may need to refer to the Rmd file if you need additional information.
+
+ASReml h2 greenup analysis
+```{r}
+#| eval: false
+#| 
+library(tidyverse)
+library(asreml)
+library(switchgrassGWAS)
+library(cowplot)
+
+source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+outputdir <- file.path(workingdir, "analysis", "heritability")
+metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
+sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
+subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
+phe_gr_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
+phe_fl_df <- read_rds(file.path(workingdir, "data",
+                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
+k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
+
+```
+
+get BLUPs from ASReml
+```{r}
+site_v <- list(eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
+subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"))
+phe_fl <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl")
+phe_gr <- c("GR50", "tmin_10d", "cgdd_12c_10d")
+phe_dfs <- list(phe_gr_df, phe_fl_df)
+phe_lists <- list(phe_gr, phe_fl)
+h2_table <- tibble()
+
+i = 1
+j = 1
+k = 1
+p = 2
+
+for(p in seq_along(phe_dfs)){
+  phe_v <- phe_lists[[p]]
+  phe <- phe_dfs[[p]]
+
+  for(i in seq_along(site_v)){
+    for(j in seq_along(subpop_v)){
+      for(k in seq_along(phe_v)){
+        phe_single <- phe %>%
+    left_join(sites, by = "SITE") %>%
+    filter(manu_site %in% site_v[[i]] & SUBPOP %in% subpop_v[[j]]) %>%
+    rename(phenotype = eval(phe_v[k])) %>%
+    mutate(PLANT_ID = as_factor(PLANT_ID),
+           SITE = as_factor(SITE),
+           PLOT_GL = as_factor(PLOT_GL)) %>%
+    as.data.frame()
+
+  if(length(site_v[[i]]) > 1){  # need to add site as a random factor if 
+    # there is more than one site included in the dataset.
+    tryCatch({
+     asr_out <- asreml(phenotype ~ 1,
+                       random = ~vm(PLANT_ID, k_full) +
+                         ~idv(manu_site),
+                       residual = ~idv(units),
+                       data = phe_single, 
+                       workspace = "3gb")
+
+    h2_est <- vpredict(asr_out, h2 ~ V2/(V1+V2+V3)) %>% as_tibble() %>%
+      mutate(site = names(site_v)[i],
+             subpop = names(subpop_v)[j],
+             phe = phe_v[k],
+             bic = summary(asr_out)$bic[1],
+             loglik = summary(asr_out)$loglik)
+    blup <- summary(asr_out, coef=TRUE)$coef.random %>% as_tibble(rownames = "Effect")
+    write_csv(blup, file.path(outputdir, paste0("blups_", names(site_v)[i], "_",
+                                                 names(subpop_v)[j],
+                                                 "_", phe_v[k], ".csv")))
+    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+  } else {
+    tryCatch({
+    asr_out <- asreml(phenotype ~ 1,
+                       random = ~vm(PLANT_ID, k_full),
+                       residual = ~idv(units),
+                       data = phe_single, 
+                       workspace = "3gb")
+
+    h2_est <- vpredict(asr_out, h2 ~ V1/(V1+V2)) %>% as_tibble() %>%
+      mutate(site = names(site_v)[i],
+             subpop = names(subpop_v)[j],
+             phe = phe_v[k],
+             bic = summary(asr_out)$bic[1],
+             loglik = summary(asr_out)$loglik)
+  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+  }
+
+  if(i == 1 & j == 1 & k == 1){
+    h2_table <- h2_est
+  } else {
+    h2_table <- add_row(h2_table, h2_est)
+  }
+  tryCatch({
+      save_plot(filename=file.path(outputdir, paste0("ASReml_performance_",
+                                                 names(site_v)[i], "_",
+                                                 names(subpop_v)[j],
+                                                 "_", phe_v[k], ".png")), 
+            plot = plot(asr_out))
+    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+
+      }
+    }
+  }
+
+}
+
+```
+
+set up phenotype df for gwas
+```{r}
+#| eval: false 
+#| 
+site_v <- list(eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
+subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"))
+phe_fl <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl")
+phe_gr <- c("GR50", "tmin_10d", "cgdd_12c_10d")
+phe_dfs <- list(phe_gr_df, phe_fl_df)
+phe_lists <- list(phe_gr, phe_fl)
+
+i = 1
+j = 1
+k = 1
+p = 2
+phe_gwas <- tibble()
+
+for(j in seq_along(subpop_v)){
+  for(p in seq_along(phe_dfs)){
+    phe_v <- phe_lists[[p]]
+    phe <- phe_dfs[[p]]
+    for(i in seq_along(site_v)){
+      for(k in seq_along(phe_v)){
+        phe_raw <- read_csv(file.path(outputdir, paste0("blups_",
+                                                        names(site_v)[i], "_",
+                                                        names(subpop_v)[j],
+                                                        "_", phe_v[k],
+                                                        ".csv")))
+        names(phe_raw)[2] <- paste0(phe_v[k], "_G_BLUP")
+        phe_proc <- phe_raw %>%
+          separate(Effect, into = c("factor", "PLANT_ID"), sep = "\\)_") %>%
+          filter(!is.na(PLANT_ID)) %>%
+          left_join(metadata, by = "PLANT_ID") %>%
+          filter(SUBPOP %in% subpop_v[[j]]) %>%
+          select(PLANT_ID, 3)
+        if(i == 1 & k == 1 & p == 1){
+          phe_gwas <- phe_proc
+        } else {
+          phe_gwas <- phe_gwas %>%
+            left_join(phe_proc, by = "PLANT_ID")
+        }
+      }
+    }
+  }
+write_csv(phe_gwas, path = file.path(outputdir, paste0("G_BLUP_phenotypes_",
+                                                       "for_gwas_GR_FL_",
+                                                       names(subpop_v)[j],
+                                                       ".csv")))
+}
+```
 
@@
 
-```{r}
+NB: only the single_site BLUPs & GWAS that I generated were used in the subsequent analysis.
+```{r}
+#| eval: false
+#| 
 library(bigsnpr)
@@
 library(AnnotationDbi)
-txdb <- loadDb(file = file.path("~", "Github", "pvdiv-genome",
+library(here)
+txdb <- loadDb(file = here("data-raw", "ignore",
                            "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
@@
 #|
-#|
 
@@
 
-# ----------------
-
 Reasoning from Analysis_v0.6 script for dropping sites from the Midwest:
@@
 
+# ---
 # Step 5: Redo writeup & figures
@@
 legend <- get_legend(pa_legend)
+
 ```

### 2024-06-06 R/SI_Appendix_Datasets.Rmd
revision2: Add SI Appendix Sections on BLUPs, GWAS, SNP subset selection for mash

@@
 
+```         
 Dataset 1. SNP-associated effects and standard errors for green-up estimated using the Gulf genetic subpopulation.
@@
 Genes within 20kb with a functionally validated role in flowering in rice are  indicated.
+```
 
@@
                                           suffix_v[1], ".rds")))
+```
+
+## SI Appendix Table S1 Info
+
+#### random SNP inputs
+
+```{r}
+popkey = list(all = "Gulf_and_Midwest", midwest = "Midwest", gulf = "Gulf")
+phekey = list(greenup = "GR50", flowering = "FL50")
+# i=2; j=1
+
+for (i in seq_along(popkey)) {
+  for (j in seq_along(phekey)) {
+
+    if (i == 2 & j == 1) {
+      betas <- readRDS(file = here("analysis", "mash", popkey[[i]], "inputs", 
+                                 paste0("B_hat_random_df_24000topSNPs_",
+                                        phekey[[j]], "_", popkey[[i]], ".rds")))
+      stderr <- readRDS(file = here("analysis", "mash", popkey[[i]], "inputs", 
+                                 paste0("S_hat_random_df_24000topSNPs_",
+                                        phekey[[j]], "_", popkey[[i]], ".rds")))
+    } else if (i == 2 & j == 2) {
+      betas <- readRDS(file = here("analysis", "mash", popkey[[i]], "inputs", 
+                                 paste0("B_hat_random_df_33000topSNPs_",
+                                        phekey[[j]], "_", popkey[[i]], ".rds")))
+      stderr <- readRDS(file = here("analysis", "mash", popkey[[i]], "inputs", 
+                                 paste0("S_hat_random_df_33000topSNPs_",
+                                        phekey[[j]], "_", popkey[[i]], ".rds")))
+    } else {
+      betas <- readRDS(file = here("analysis", "mash", popkey[[i]], "inputs", 
+                                 paste0("B_hat_random_df_19000topSNPs_",
+                                        phekey[[j]], "_", popkey[[i]], ".rds")))
+      stderr <- readRDS(file = here("analysis", "mash", popkey[[i]], "inputs", 
+                                 paste0("S_hat_random_df_19000topSNPs_",
+                                        phekey[[j]], "_", popkey[[i]], ".rds")))
+    }
+    print(paste(names(popkey)[i], names(phekey)[j], "random SNP set (input into greedy mash algorithm) has", nrow(betas), "SNPs"))
+  }
+}
+
+```
+
+So random SNP sets have 71-127K SNPs.
+
+#### strong SNP inputs
+
+Numbers for this section: We then selected 19K relatively unlinked (r2 < 0.2) SNPs from each univariate GWAS that had the largest effect estimates in each GWAS (hereafter "strong" effects), which gave sets of 48K - 80K SNPs, and jointly re-estimated these effect size estimates at all eight sites using the set of significant covariance matrices that most improved the model fit on the random effects.
+Notably, the majority of these "strong" effects were not significant in univariate GWAS.
+
+```{r}
+switchgrassGWAS:::get_marker_df(m_gulf_gr_hyp) # 45931
+switchgrassGWAS:::get_marker_df(m_gulf_fl_hyp) # 50490
+switchgrassGWAS:::get_marker_df(m_midw_gr_hyp) # 49030
+switchgrassGWAS:::get_marker_df(m_midw_fl_hyp) # 63367
+switchgrassGWAS:::get_marker_df(m_gam_gr_hyp)  # 44765
+switchgrassGWAS:::get_marker_df(m_gam_fl_hyp)  # 35735
 ```
@@
 
-Find this file on NAS: "~/Github/pvdiv-genome/Pvirgatum_516_v5.1.gene.txdb.sqlite"
+Find this file on NAS: "\~/Github/pvdiv-genome/Pvirgatum_516_v5.1.gene.txdb.sqlite"
 
-Couldn't find it on the NAS, scp'd it from the Juenger Lab Server:
-`scp -i ~/Documents/server_keys/jls_alice alice@popgen.biosci.utexas.edu:/home/alice/Github/pvdiv-genome/Pvirgatum_516_v5.1.gene.txdb.sqlite ~/Documents/Github/pvdiv-phenology-gxe/data-raw/ignore/`
+Couldn't find it on the NAS, scp'd it from the Juenger Lab Server: `scp -i ~/Documents/server_keys/jls_alice alice@popgen.biosci.utexas.edu:/home/alice/Github/pvdiv-genome/Pvirgatum_516_v5.1.gene.txdb.sqlite ~/Documents/Github/pvdiv-phenology-gxe/data-raw/ignore/`
 
@@
 # Dataset 8
-All flowering and green-up QTL intervals contained at least one SNP significant in at least one mash run at a log10-transformed Bayes Factor > 2, or in the 1% tail of significance, whichever was stricter.
+
+All flowering and green-up QTL intervals contained at least one SNP significant in at least one mash run at a log10-transformed Bayes Factor \> 2, or in the 1% tail of significance, whichever was stricter.
 
 ## Objects generated earlier
-Get `qtlplot` object by running code starting at about line 150 (for Dataset 7).
-Get `m_gulf_fl_hyp` and other objects by running code starting at about line 35 (for Datasets 1-6).
+
+Get `qtlplot` object by running code starting at about line 150 (for Dataset 7). Get `m_gulf_fl_hyp` and other objects by running code starting at about line 35 (for Datasets 1-6).
 
 ## Thresholds
+
 Decide on cutoff to use for rug plots and enrichment analysis: SNPs sig in 1% tail or BF\>2, whichever is stricter

### 2024-06-06 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
revision2: Add SI Appendix Sections on BLUPs, GWAS, SNP subset selection for mash

@@
 
-References cited in the SI text should be included in a separate reference list at the end of this SI file: @varga2016multilingual and @olsen1992optimal.
-
-Supporting information for Brief Reports is limited to extended methods, essential supporting datasets, and videos (no additional tables or figures).
-Supporting figures and tables are not allowed for Brief Reports.
-
 ## Submitting SI {.unnumbered}
@@
 
-## Subhead {.unnumbered}
-
-Type or paste text here. This should be additional explanatory text such as an extended technical description of results, full details of mathematical models, etc.
-
 # Supplementary Methods
 
-## Section S1. Covariance matrices generated to jointly re-estimate SNP effects across eight sites {.unnumbered}
+## Section S1. Covariance matrices generated to provide multiple ways SNP effects could covary across common gardens {.unnumbered}
 
@@
 We were interested in the types of genetic correlation that could be expressed by different alleles across the genome and reasoned that different loci could have different patterns of genetic correlation.
-For example, one locus's effects could covary with a weather-based cues, if it had an allele that was responsive to this cue.
-Alternatively, a locus might have an effect at only one garden, for example if a herbivore or pathogen was only present at one garden.
+For example, the effects of a locus, $A$, could covary with a weather-based cue if $A$ had a common allele, $a$, that was responsive to this cue.
+In contrast, the effects of a different locus $B$ might have an effect at only one garden, for example if a pathogen was only present at one garden, and $B$ has a common allele $b$ that is resistant to that pathogen.
 
-We specified three qualitative categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data and introduced in the initial *mash* manuscript, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden over specific time frames before the phenological event.
+We specified three qualitative categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data and introduced in the initial *mash* manuscript, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden over specific time frames before each genotype underwent the phenological event.
 
@@
 
-Data-driven covariance matrices can also be created by *mash* and fall into two groups: an overall pattern of covariance between all conditions, denoted as $tPCA$, and the first five principle components of a singular value decomposition of the $tPCA$, denoted as `PCA_1`$PCA_1$ through $PCA_5$[@Urbut2019].
+Data-driven covariance matrices can also be created by *mash* and fall into two groups: an overall pattern of covariance between all conditions, denoted as $tPCA$, and the first five principle components of a singular value decomposition of the $tPCA$, denoted as $PCA_1$ through $PCA_5$ [@Urbut2019].
 
@@
 For example, the cumulative sum of the rainfall that occurred seven days prior to the flowering date was used to create one possible GxWeather cue variable for flowering date.
-The code for these calculations can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 208.
+The code for these calculations can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 390.
 
-We defined both phenological dates as functions of five weather variables and eight to ten time frames (Table 1): Cumulative growing degree days (GDD) variables for each time frame were calculated as $GDD = \sum^{DATE}_{DATE - i} max(T_{mean} - T_{base}, 0)$, where $T_{mean}$ is the daily average temperature, defined as $(T_{max} + T_{min}) / 2$, $T_{base}$ is the base temperature of 12°C for switchgrass, $T_{max}$ is the maximum daily temperature, $T_{min}$ is the minimum daily temperature, $DATE$ is the phenological date, and $i$ is the time frame, or the number of days prior to the phenological date that GDD is summed across [@kiniry2005; @behrman2013].
+We defined the two phenological dates as functions of five weather variables and eight to ten time frames (Table 1).
+First, cumulative growing degree days (GDD) variables for each time frame were calculated as $GDD = \sum^{DATE}_{DATE - i} max(T_{mean} - T_{base}, 0)$, where $T_{mean}$ is the daily average temperature, defined as $(T_{max} + T_{min}) / 2$, $T_{base}$ is the base temperature of 12°C for switchgrass, $T_{max}$ is the maximum daily temperature, $T_{min}$ is the minimum daily temperature, $DATE$ is the phenological date, and $i$ is the time frame, or the number of days prior to the phenological date that GDD is summed across [@kiniry2005; @behrman2013].
 Second, cumulative rainfall values in each time frame were calculated as $C_{rain} = \sum^{DATE}_{DATE - i} PRCP$, where PRCP is the daily rainfall measured in millimeters, $DATE$ is the phenological date, and $i$ is the time frame, or the number of days prior to the phenological date that rainfall is summed across.
@@
 These covariance matrices represent the correlations between genotypes for these weather-related variables across our eight common gardens.
-To create these matrices, we determined the correlations between these weather-related cue variables for identical genotypes grown at different common gardens and used these correlations to fill the off-diagonals of a eight-by-eight common garden matrix.
-For the diagonal of these matrices, we used the narrow-sense heritability for each weather-related cue variable at each garden, calculated using the same methods as Section S3.
+To create these matrices, we determined the correlations between these weather-related cue variables for identical genotypes grown at different common gardens and used these correlations to fill the off-diagonals of a eight-by-eight matrix.
+To fill the diagonal entries of these matrices, we used the narrow-sense heritability for each weather-related cue variable at each garden, calculated using the same methods as Section S3.
 Matrices need to be positive semi-definite to be used in *mash*, so when the resultant matrices were not positive semi-definite, the `Matrix::nearPD` function was used to compute the nearest positive definite matrix to the correlation matrix [@Matrix].
-Finally, all covariance matrices are rescaled so that the maximum diagonal element of the rescaled matrix is 1, as in [@Urbut2019].
-Code used to generate these matrices can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 536.
+Finally, all covariance matrices were rescaled so that the maximum diagonal element of the rescaled matrix is 1, as in [@Urbut2019].
+Code used to generate these matrices can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 716.
 
@@
 
-## Section S2. Greedy mash algorithm used to select covariance matrices that significantly improved the model log-likelihood {.unnumbered}
+## Section S2. Univariate genome-wide association to obtain initial effect estimates {.unnumbered}
 
-We generated 48 - 60 GxWeather covariance matrices for each of our five weather variables over different time frames. These GxWeather covariance matrices were in many cases highly correlated.
+To obtain initial effect estimates for each single nucleotide polymorphism (SNP) for each combination of phenological trait, genetic subpopulation, and common garden, we conducted univarate genome-wide association (GWAS) on site-specific best linear unbiased predictors (BLUPs) using the switchgrassGWAS package and the methods in [@Lovell2021].
+The full SNP diversity panel generated in [@Lovell2021] was subset within each subpopulation for all subsequenct analyses to retain only SNPs with $\leq$ 20% missing data and minor allele frequencies \> 0.05.
+This subsetting resulted in 8.8 million SNPs retained for the Midwest subpopulation, 10.3 million SNPs retained for the Gulf subpopulation, and 12.3 million SNPs retained for Both subpopulations.
 
-selection of the GxWeather
+BLUPs were calculated independently for each combination of trait, genetic subpopulation, and common garden. We used ASReml to estimate BLUPs using mixed models of the form:
 
-> describing that the weather-based cues are themselves correlated, and picking and choosing among weather cues is a model building problem – to address the issue, we used a heuristic greedy algorithm to pick the weather-based covariance structures that improved the likelihood of the mash model
+$$
+\mathbf{y} = 1 + Zu + e
+$$
 
-> A lot of the results interpretation is based on interpreting the SNP loadings on the specified covariance matrices.
-This is a secondary use of mash (the primary being the refinement of effect sizes), and while they do interpret these somewhat in the mash paper, these loadings need to be interpreted with caution.
-If these covariance matrices are similar, mash will somewhat arbitrarily assign weight to any of the matrices because all lead to the same fit to the effect-size data.
-The mash algorithm doesn’t give the full posterior distribution on the loadings so you can’t check for posterior correlations there.
-It seems likely to me that the hypothesis-driven and data-driven covariance matrices are somewhat correlated here, and the correlations may differ between green-up and flowering because of the better correlation between the environmental metrics and flowering than green-up.
+$$
+Var(u) = G\sigma^2_u
+$$
 
-Type or paste text here. Break this section up into subheads as needed
+$$
+Var(e) = I\sigma^2_e
+$$
 
-## Section S3. Narrow-sense heritability estimation {.unnumbered}
+where the vector $\mathbf{y}$ represents the flowering date or greenup date values for that garden, $Z$ the design matrix for random effects, $u$ the whole genome additive genetic effect, $e$ the residual, $G$ the genomic relationship matrix based on all SNPs retained for subpopulation-specific analyses, and $I$ the rank-$\mathbf{y}$ identity matrix [@clark2013].
+ASReml can specify considerably more complex variance structures; however, we were interested in specifying covariance structures using *mash*, which allows the specification of multiple covariance structures; hence, we used a relatively simple model to calculate BLUPs for use in univariate GWAS.
+The code for these calculations can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 74.
+
+Univariate GWAS were run independently for each combination of trait, genetic subpopulation, and common garden.
+The code used to run GWAS can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 240.
+We used [`pvdiv_standard_run()`](https://alice-macqueen.github.io/switchgrassGWAS/reference/pvdiv_standard_gwas.html), an opinionated function of the [switchgrassGWAS](https://github.com/Alice-MacQueen/switchgrassGWAS) R package, to conduct GWAS.
+To control for population structure and run GWAS, this function uses a three-part process.
+First, a fast truncated singular value decomposition (SVD) is used, which has initial pruning of regions in long-range linkage disequilibrium (LD) and iterative pruining and removal of regions in long-range LD [@privé2017a].
+By default, 15 vectors of singular values, $k$, are computed.
+Pruning and removal of long-range LD regions are recommended in human genetics to best control for population structure [@abdellaoui2013; @privé2017; @price2008].
+Second, linear regression models using 0-15 of these vectors of singular values were used to run GWAS and to calculate the genomic control coefficient, $\lambda_{GC}$, for each $k$ [@Devlin1999].
+Third, we chose the lowest value of $k$ that made $\lambda_{GC}$ closest to 1 to correct for population structure, and ran GWAS using a linear regression model with this $k$ for population structure correction.
+
+## Section S3. Selection of effect estimate subsets to model using *mash* {.unnumbered}
+
+To make the subsequent analyses using *mash* computationally feasible, we followed the author's [recommended analysis outline](https://stephenslab.github.io/mashr/articles/eQTL_outline.html) for large datasets using *mash* and considered subsets of the millions of effects at different stages of the mash analysis.
+We extracted two subsets of tests from our complete GWAS effect estimate datasets: 1) effects from a subset of "strong" tests corresponding to stronger effects on our traits; 2) results from a random subset of all tests to correspond to an unbiased representation of all effects.
+We also clumped effects before selection so that signal strength was standardized across different LD blocks in the genome, and not limited to closely linked associations.
+
+To prepare subsets of effect estimates and their standard errors from these GWAS for use in mash, we used [`pvdiv_bigsnp2mashr()`](https://alice-macqueen.github.io/switchgrassGWAS/reference/pvdiv_bigsnp2mashr.html) from the switchgrassGWAS R package.
+For each phenological trait and each genetic subpopulation, we created two subsets of effects and standard errors in paired $m \times n$ matrices, where $m$ were marker effects and $n$ were common gardens.
+The first subset of effects and standard errors, the strong effects, was created by selecting relatively unlinked (r^2^ \< 0.2) SNPs from each univariate GWAS that had the smallest *p*-values in those GWAS.
+Notably, the majority of these "strong" effects were not significant at a 5% level in univariate GWAS.
+The number of SNPs from each univariate GWAS was chosen to result in approximately million entry matrices for each trait & subpopulation.
+We selected 19K - 33K effects and standard errors for the strong effect subsets (@tbl-simple).
+In many cases, there was overlap in the markers with the largest effect estimates at different locations; thus, 19 - 33K SNPs selected for each of eight univariate GWAS ultimately gave sets of 35K - 63K SNPs (@tbl-simple).
+The second subset of effects and standard errors, the random subset, was created by selecting relatively unlinked (r^2^ \< 0.2) SNPs from each univariate GWAS after clumping using minor allele frequency.
+To account for the chance that true effects could be sparse, we doubled the size of the random subset relative to the strong subset.
+We selected 38 - 66K SNP effects and formed subsets of 71 - 127K SNP markers for the random subset.
+
+## Section S4. Greedy *mash* algorithm used to select covariance matrices that significantly improved the model log-likelihood {.unnumbered}
+
+Because we were particularly interested in determining whether genomic loci varied in their patterns of genetic covariance across gardens, specifying the optimal set of covariance matrices in *mash* was an important model-building problem.
+If very similar covariance matrices are included when fitting a model in *mash*, *mash* will arbitrarily assign effect loadings across these matrices, as all assignments will lead to the same fit to the effect-size data.
+Our GxWeather covariance matrices were in many cases highly correlated, as they represented covariation in the same weather variables in overlapping time frames.
+We thus evaluated the performance of *mash* models when different covariance matrices were included by implementing a "greedy" mash algorithm.
+
+## Section S5. *Mash* to jointly re-estimate SNP effects across eight common gardens {.unnumbered}
+
+## Section S6. Narrow-sense heritability estimation {.unnumbered}
 
@@
 
-## Section S4. Outbred pseudo-F2 mapping population and Quantitative Trait Locus mapping {.unnumbered}
+## Section S7. Outbred pseudo-F2 mapping population and Quantitative Trait Locus mapping {.unnumbered}
 
@@
 
-| Species           |  CBS |   CV |   G3 |
-|:------------------|-----:|-----:|-----:|
-| Acetaldehyde      |  0.0 |  0.0 |  0.0 |
-| Vinyl alcohol     |  9.1 |  9.6 | 13.5 |
-| Hydroxyethylidene | 50.8 | 51.2 | 54.0 |
+| Phenology type | Subpopulation |  Effects Selected per GWAS, Strong Subset |  Effects Selected per GWAS, Random Subset | Total Markers, Strong Subset | Total Markers, Random Subset |
+|:-----------|--------------:|-----------:|-----------:|------------|------------|
+| Greenup Date   |          Gulf |                                     19000 |                                     38000 | 45931                        | 91862                        |
+|                |       Midwest |                                     24000 |                                     48000 | 49030                        | 98060                        |
+|                |          Both |                                     19000 |                                     38000 | 44765                        | 89530                        |
+| Flowering Date |          Gulf |                                     19000 |                                     38000 | 50490                        | 100980                       |
+|                |       Midwest |                                     33000 |                                     66000 | 63367                        | 126734                       |
+|                |          Both |                                     19000 |                                     38000 | 35735                        | 71470                        |
 
-: Simple table example {#tbl-simple}
+: Effect estimates selected for joint re-estimation across eight common gardens using *mash* {#tbl-simple}
 
@@
 
-\dataset{Dataset_8_blah.csv}{Overlap between quantitative trait loci (QTL) for the start of vegetative or reproductive growth with significant mash effect estimates (with a log10(Bayes Factor) > 2) found using diversity panels of the Midwest, Gulf, and Both genetic subpopulations.
+\dataset{Dataset_8_QTL_mash_overlap.csv}{Overlap between quantitative trait loci (QTL) for the start of vegetative or reproductive growth with significant mash effect estimates (with a log10(Bayes Factor) > 2) found using diversity panels of the Midwest, Gulf, and Both genetic subpopulations.
 The first five columns of this table provide QTL information, and columns 6-10 provide mash Marker information for markers within the QTL regions.}

### 2024-06-06 manuscript/revision2/revision2.qmd
revision2: Add SI Appendix Sections on BLUPs, GWAS, SNP subset selection for mash

@@
 However, this work has been limited by a known statistical bias that reduced detection of genetic effects that differ in sign [@des2013genotype; @anderson2013genetic; @anderson2011evolutionary].
-Thus, despite substantial interest in the frequencies of various forms of GxE, particularly sign-changing GxE, relative to other forms of GxE remains unknown.
+Thus, despite substantial interest in the frequencies of various forms of GxE, the frequency of sign-changing GxE relative to other forms of GxE remains unknown.
 
@@
 These gardens cover the majority of the latitudinal and climatic range of switchgrass and capture the most comprehensive picture to date of the environmental variation this species encounters.
-We define multiple ways phenological timings might covary with weather (Table 1) and additional ways phenological timings might vary by garden (SI Appendix, Section S1), then jointly re-estimate genetic effects on these timings at all eight common gardens using the set of these covariance matrices that significantly improved the modeled log-likelihood when included (SI Appendix, Section S2) @urbut2019flexible.
+We define multiple ways phenological timings might covary with weather (Table 1) and additional ways phenological timings might vary by garden (SI Appendix, Section S1), then jointly re-estimate genetic effects on these timings at all eight common gardens using the set of these covariance matrices that significantly improved the modeled log-likelihood when included (SI Appendix, Section S2-S5) @urbut2019flexible.
 We use the Bayesian framework *mash* (multivariate adaptive shrinkage) developed by @urbut2019flexible, to refine effect size estimates from genome-wide association (GWAS) conducted on individuals at each of the eight gardens.
@@
 
-We next looked for evidence of GxWeather by jointly re-estimating genetic effects of SNPs across all eight common gardens using the *mash* model, which can flexibly capture many modes of covariance across gardens (@fig-map D; SI Appendix, Datasets 1-6).
-We specified three qualitative categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event (Table 1; SI Appendix, Section S2).
+We next looked for evidence of GxWeather by jointly re-estimating genetic effects of SNPs across all eight common gardens using the *mash* model, which can flexibly capture many modes of covariance across gardens (@fig-map D; SI Appendix, Sections S1-S5, Datasets 1-6).
+We specified three qualitative categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event (Table 1; SI Appendix, Section S1).
 For example, the more similar that the amount of rainfall prior to flowering is for genotypes A and B between gardens in MI and MO, the nearer these covariances will be to one.
@@
 
-For green-up date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited a difference in effect sign, between pairs of Texas and North gardens (@fig-effects B; SI Appendix, Fig SX).
+For green-up date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited a difference in effect sign, between pairs of Texas and North gardens (@fig-effects B; SI Appendix, Fig S1).
 78.7% of pairwise comparisons between North and Texas gardens had a difference in sign, while only 28.6% and 0.2% of North-North or Texas-Texas comparisons had a difference in sign, respectively (@fig-effects B).
@@
 Finally, in Both subpopulations, less than 20% of pairwise effects differed in sign (@fig-effects B).
-Most differences in sign were between TX1, the southernmost garden, and all other gardens (SI Appendix, Fig SX).
+Most differences in sign were between TX1, the southernmost garden, and all other gardens (SI Appendix, Figure S1).
 Similarly, more effects that differed in magnitude included gardens in the Texas region (52.3-55.9%), and most effect pairs in the North region were not distinguishable (91.5%).
@@
 We sought additional experimental support the occurrence and location of our significant SNP effect re-estimates using an independent pseudo-F2 mapping population created from Gulf & Midwest individuals and grown at the same sites (@fig-qtl A,B).
-We conducted quantitative trait loci (QTL) mapping of flowering as functions of five environmental cues that we also used as covariance matrices in *mash*, and identified eight QTL for flowering date, eight QTL for flowering as a function of day length change two days prior, one QTL for the start of vegetative growth, and two QTL for vegetative growth as a function of daylength change one day prior, all of which showed QTL by environment interactions (SI Appendix, Fig.
-S3).
+We conducted quantitative trait loci (QTL) mapping of flowering as functions of five environmental cues that we also used as covariance matrices in *mash*, and identified eight QTL for flowering date, eight QTL for flowering as a function of day length change two days prior, one QTL for the start of vegetative growth, and two QTL for vegetative growth as a function of daylength change one day prior, all of which showed QTL by environment interactions (SI Appendix, Figure S2).
 All QTL for flowering overlapped one or more homologs from rice or *A.
@@
 a) Examples of effect patterns at three pairs of sites with three types of GxE.
+All effects are for an alternate allele, with the reference allele effect defined to be zero at both gardens.
 Sign: Effects that differ in sign at these pairs of gardens (p \< 0.05, lfsr).
@@
 The start of reproductive growth, or flowering date, was the day of the year when 50% of the plant tillers had panicles undergoing anthesis.
-We scored green-up and flowering as day of the year, then linked these dates to multiple weather-based environmental factors measured daily at each common garden (SI Appendix, Section S1, Table S1).
+We scored green-up and flowering as day of the year, then linked these dates to multiple weather-based environmental factors measured daily at each common garden (Table 1, SI Appendix, Section S1).
 
@@
 Analyses to determine narrow-sense heritability (h^2^) for green-up and flowering were done using linear mixed models and followed @lovell2021genomic.
-Details on these models can be found in (SI Appendix, Section S2).
+Details on these models can be found in (SI Appendix, Section S6).
 
@@
 To evaluate the prevalence and kinds of covariance patterns of SNP effects across our common gardens, we used multivariate adaptive shrinkage (*mash*) on SNP effect estimates from the diversity panel [@urbut2019flexible].
-To do this, we first conducted univarate genome-wide association (GWAS) on site-specific best linear unbiased predictors (BLUPs) for the timing of vegetative growth and flowering, using the switchgrassGWAS package and the methods in [@lovell2021].
+To do this, we first conducted univarate genome-wide association (GWAS) on site-specific best linear unbiased predictors (BLUPs) for the timing of vegetative growth and flowering, using the switchgrassGWAS package and the methods in [@lovell2021] (SI Appendix, Section S2).
 We used the effect estimates for single nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect estimates for the two traits at each garden.
-We then jointly modeled genetic effect estimates, and the mixture of ways these effects might covary, across all eight common gardens, using a two-part procedure.
-We first modeled genetic effects with different combinations of covariance structures, using the effect size estimates from a subset of [19K]{.underline} relatively unlinked (r^2^ \< 0.2), randomly selected SNPs used in the GWAS.
+We then jointly modeled genetic effect estimates, and the mixture of ways these effects might covary, across all eight common gardens, using a two-part procedure: a greedy algorithm to choose covariance structures using a random SNP subset, followed by a final run of mash with the best covariance structures using a strong SNP subset (SI Appendix, Section S3-S5).
+In brief, we first modeled genetic effects with different combinations of covariance structures, using the effect size estimates from subsets of 71-127K relatively unlinked (r^2^ \< 0.2), randomly selected SNPs used in the GWAS.
+The number of SNPs was chosen to result in an approximately million entry matrix for each trait & subpopulation, where matrix rows were SNPs and columns were univariate GWAS.
 In these models (hereafter referred to as "random"), we then specified a variety of potential covariance structures among genetic effect size estimates from each garden (as expanded on below) and used a greedy algorithm to iteratively select and add covariance structures that significantly improved the log likelihood of the *mash* model.
-Details on these models can be found in (SI Appendix, Section S3).
 Given the high correlation between some of the covariance patterns we tested, this approach allowed us to select only the subset of patterns of covariance that significantly improved the fit of the jointly inferred structure of genetic effect estimates present in the data.
-We then selected 19K relatively unlinked (r^2^ \< 0.2) SNPs from each univariate GWAS that had the largest effect estimates in each GWAS (hereafter "strong" effects), which gave sets of [**48K - 80K**]{.underline} SNPs, and jointly re-estimated these effect size estimates at all eight sites using the set of significant covariance matrices that most improved the model fit on the random effects.
+We then selected 19K - 33K relatively unlinked (r^2^ \< 0.2) SNPs from each univariate GWAS that had the largest effect estimates in each GWAS (hereafter "strong" effects).
 Notably, the majority of these "strong" effects were not significant in univariate GWAS.
+In many cases, there was overlap in the markers with the largest effect estimates at different locations; thus, 19 - 33K SNPs selected for each of eight univariate GWAS ultimately gave sets of 35K - 63K SNPs.
+We jointly re-estimated these effect size estimates at all eight sites using the set of significant covariance matrices that most improved the model fit on the random effects (SI Appendix, Section S5, Datasets 1-6).
 
-We generated hypothesis-based covariance matrices derived from correlations in environmental cues in the green-up or flowering date windows for the three populations (Table 1; SI Appendix, Table S1, Section S1).
+We generated hypothesis-based covariance matrices derived from correlations in environmental cues in the green-up or flowering date windows for the three populations (Table 1; SI Appendix, Section S1).
 These covariance matrices represent correlations between identical genotypes drawn from a specific population at pairs of common gardens; covariances near one mean that the population has a strong, positive linear relationship in individual responses at that pair of gardens, while covariances near zero mean that there is no relationship within the population for individual responses at that pair of gardens.
@@
 To confirm candidate genomic regions and patterns of allelic effects found in the diversity panel, we analyzed flowering in an outbred pseudo-F2 cross between four individuals, two Midwest and two Gulf individuals.
-The formation of this mapping population has been described previously [@milano2016genetic]; additional details on QTL mapping can be found in SI Appendix, Section S4.
+The formation of this mapping population has been described previously [@milano2016genetic]; additional details on QTL mapping can be found in SI Appendix, Section S7.
 To be directly comparable to the diversity panel data, only 2019 phenology data from the pseudo-F2 cross from the same eight common garden sites were used.

### 2024-06-06 manuscript/revision2/revision2.tex
revision2: Add SI Appendix Sections on BLUPs, GWAS, SNP subset selection for mash

@@
 \usepackage{xcolor}
-\ifLuaTeX
-  \usepackage{luacolor}
-  \usepackage[soul]{lua-ul}
-\else
-  \usepackage{soul}
-
-\fi
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 differ in sign (27--29). Thus, despite substantial interest in the
-frequencies of various forms of GxE, particularly sign-changing GxE,
+frequencies of various forms of GxE, the frequency of sign-changing GxE
 relative to other forms of GxE remains unknown.
@@
 significantly improved the modeled log-likelihood when included (SI
-Appendix, Section S2) (36). We use the Bayesian framework \emph{mash}
+Appendix, Section S2-S5) (36). We use the Bayesian framework \emph{mash}
 (multivariate adaptive shrinkage) developed by (36), to refine effect
@@
 \emph{mash} model, which can flexibly capture many modes of covariance
-across gardens (Figure~\ref{fig-map} D; SI Appendix, Datasets 1-6). We
-specified three qualitative categories of covariance structures across
-gardens: ``canonical'' covariance, with simple patterns of effect size
-covariance introduced in the initial \emph{mash} manuscript;
-``data-driven'' covariances derived from common patterns of SNP effects
-observed in the data, and ``GxWeather'' covariance, estimated from the
-covariance of empirical weather patterns at each garden at specific
-times before the phenological event (Table 1; SI Appendix, Section S2).
-For example, the more similar that the amount of rainfall prior to
-flowering is for genotypes A and B between gardens in MI and MO, the
-nearer these covariances will be to one. If rainfall amounts are
-uncorrelated, covariance will be near zero; if there is rank changing
-(e.g.~high rainfall before flowering of A at MO but not MI, and high
-rainfall before flowering for B at MI but not MO), this covariance will
-be near negative one. We specified multiple covariance structures for
-the canonical, data-driven, and GxWeather categories, each of which
+across gardens (Figure~\ref{fig-map} D; SI Appendix, Sections S1-S5,
+Datasets 1-6). We specified three qualitative categories of covariance
+structures across gardens: ``canonical'' covariance, with simple
+patterns of effect size covariance introduced in the initial \emph{mash}
+manuscript; ``data-driven'' covariances derived from common patterns of
+SNP effects observed in the data, and ``GxWeather'' covariance,
+estimated from the covariance of empirical weather patterns at each
+garden at specific times before the phenological event (Table 1; SI
+Appendix, Section S1). For example, the more similar that the amount of
+rainfall prior to flowering is for genotypes A and B between gardens in
+MI and MO, the nearer these covariances will be to one. If rainfall
+amounts are uncorrelated, covariance will be near zero; if there is rank
+changing (e.g.~high rainfall before flowering of A at MO but not MI, and
+high rainfall before flowering for B at MI but not MO), this covariance
+will be near negative one. We specified multiple covariance structures
+for the canonical, data-driven, and GxWeather categories, each of which
 represented a specific pattern of covariance in genetic effect size
@@
 Texas and North gardens (Figure~\ref{fig-effects} B; SI Appendix, Fig
-SX). 78.7\% of pairwise comparisons between North and Texas gardens had
+S1). 78.7\% of pairwise comparisons between North and Texas gardens had
 a difference in sign, while only 28.6\% and 0.2\% of North-North or
@@
 differences in sign were between TX1, the southernmost garden, and all
-other gardens (SI Appendix, Fig SX). Similarly, more effects that
+other gardens (SI Appendix, Figure S1). Similarly, more effects that
 differed in magnitude included gardens in the Texas region
@@
 daylength change one day prior, all of which showed QTL by environment
-interactions (SI Appendix, Fig. S3). All QTL for flowering overlapped
+interactions (SI Appendix, Figure S2). All QTL for flowering overlapped
 one or more homologs from rice or \emph{A. thaliana} with functionally
@@
 lfsr \textless{} 0.05. a) Examples of effect patterns at three pairs of
-sites with three types of GxE. Sign: Effects that differ in sign at
-these pairs of gardens (p \textless{} 0.05, lfsr). Magnitude: Effects
-identical in sign (p \textless{} 0.05, lfsr) that differ in magnitude by
-a factor of \textgreater0.4. None: Effects not distinguishable by
-magnitude nor sign of the effect. b) The fraction of effects with each
-GxE type for the start of vegetative growth (Greenup Date) and
-reproductive growth (Flowering Date), within and between two genetic
-subpopulations. Common gardens are grouped by the larger region they
-came from: North gardens are within the natural range of the Midwest
-subpopulation, while Texas gardens are within the natural range of the
-Gulf subpopulation.}
+sites with three types of GxE. All effects are for an alternate allele,
+with the reference allele effect defined to be zero at both gardens.
+Sign: Effects that differ in sign at these pairs of gardens (p
+\textless{} 0.05, lfsr). Magnitude: Effects identical in sign (p
+\textless{} 0.05, lfsr) that differ in magnitude by a factor of
+\textgreater0.4. None: Effects not distinguishable by magnitude nor sign
+of the effect. b) The fraction of effects with each GxE type for the
+start of vegetative growth (Greenup Date) and reproductive growth
+(Flowering Date), within and between two genetic subpopulations. Common
+gardens are grouped by the larger region they came from: North gardens
+are within the natural range of the Midwest subpopulation, while Texas
+gardens are within the natural range of the Gulf subpopulation.}
 
@@
 the year, then linked these dates to multiple weather-based
-environmental factors measured daily at each common garden (SI Appendix,
-Section S1, Table S1).
+environmental factors measured daily at each common garden (Table 1, SI
+Appendix, Section S1).
 
@@
 flowering were done using linear mixed models and followed (33). Details
-on these models can be found in (SI Appendix, Section S2).
+on these models can be found in (SI Appendix, Section S6).
 
@@
 timing of vegetative growth and flowering, using the switchgrassGWAS
-package and the methods in (40). We used the effect estimates for single
-nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect
-estimates for the two traits at each garden. We then jointly modeled
-genetic effect estimates, and the mixture of ways these effects might
-covary, across all eight common gardens, using a two-part procedure. We
-first modeled genetic effects with different combinations of covariance
-structures, using the effect size estimates from a subset of \ul{19K}
-relatively unlinked (r\textsuperscript{2} \textless{} 0.2), randomly
-selected SNPs used in the GWAS. In these models (hereafter referred to
-as ``random''), we then specified a variety of potential covariance
+package and the methods in (40) (SI Appendix, Section S2). We used the
+effect estimates for single nucleotide polymorphisms (SNPs) from these
+GWAS as preliminary effect estimates for the two traits at each garden.
+We then jointly modeled genetic effect estimates, and the mixture of
+ways these effects might covary, across all eight common gardens, using
+a two-part procedure: a greedy algorithm to choose covariance structures
+using a random SNP subset, followed by a final run of mash with the best
+covariance structures using a strong SNP subset (SI Appendix, Section
+S3-S5). In brief, we first modeled genetic effects with different
+combinations of covariance structures, using the effect size estimates
+from subsets of 71-127K relatively unlinked (r\textsuperscript{2}
+\textless{} 0.2), randomly selected SNPs used in the GWAS. The number of
+SNPs was chosen to result in an approximately million entry matrix for
+each trait \& subpopulation, where matrix rows were SNPs and columns
+were univariate GWAS. In these models (hereafter referred to as
+``random''), we then specified a variety of potential covariance
 structures among genetic effect size estimates from each garden (as
@@
 add covariance structures that significantly improved the log likelihood
-of the \emph{mash} model. Details on these models can be found in (SI
-Appendix, Section S3). Given the high correlation between some of the
+of the \emph{mash} model. Given the high correlation between some of the
 covariance patterns we tested, this approach allowed us to select only
@@
 of the jointly inferred structure of genetic effect estimates present in
-the data. We then selected 19K relatively unlinked (r\textsuperscript{2}
-\textless{} 0.2) SNPs from each univariate GWAS that had the largest
-effect estimates in each GWAS (hereafter ``strong'' effects), which gave
-sets of \ul{\textbf{48K - 80K}} SNPs, and jointly re-estimated these
-effect size estimates at all eight sites using the set of significant
-covariance matrices that most improved the model fit on the random
-effects. Notably, the majority of these ``strong'' effects were not
-significant in univariate GWAS.
+the data. We then selected 19K - 33K relatively unlinked
+(r\textsuperscript{2} \textless{} 0.2) SNPs from each univariate GWAS
+that had the largest effect estimates in each GWAS (hereafter ``strong''
+effects). Notably, the majority of these ``strong'' effects were not
+significant in univariate GWAS. In many cases, there was overlap in the
+markers with the largest effect estimates at different locations; thus,
+19 - 33K SNPs selected for each of eight univariate GWAS ultimately gave
+sets of 35K - 63K SNPs. We jointly re-estimated these effect size
+estimates at all eight sites using the set of significant covariance
+matrices that most improved the model fit on the random effects (SI
+Appendix, Section S5, Datasets 1-6).
 
@@
 correlations in environmental cues in the green-up or flowering date
-windows for the three populations (Table 1; SI Appendix, Table S1,
-Section S1). These covariance matrices represent correlations between
-identical genotypes drawn from a specific population at pairs of common
-gardens; covariances near one mean that the population has a strong,
-positive linear relationship in individual responses at that pair of
-gardens, while covariances near zero mean that there is no relationship
-within the population for individual responses at that pair of gardens.
+windows for the three populations (Table 1; SI Appendix, Section S1).
+These covariance matrices represent correlations between identical
+genotypes drawn from a specific population at pairs of common gardens;
+covariances near one mean that the population has a strong, positive
+linear relationship in individual responses at that pair of gardens,
+while covariances near zero mean that there is no relationship within
+the population for individual responses at that pair of gardens.
 \emph{Mash} SNP effects will undergo strong shrinkage towards one
@@
 previously (35); additional details on QTL mapping can be found in SI
-Appendix, Section S4. To be directly comparable to the diversity panel
+Appendix, Section S7. To be directly comparable to the diversity panel
 data, only 2019 phenology data from the pseudo-F2 cross from the same

### 2024-06-06 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
revision2: Add greedy algorithm description to SI Appendix

@@
 
+Specifically, given $k$ covariance matrices for each phenotype, and $m$ corresponding effect sizes and standard errors across $n$ contexts of interest, we first fit a set of mash models with each of the $k$ matrices separately.
+We then identify the model with the maximum log likelihood estimates and select the corresponding matrix, leaving $k-1$ covariance matrices.
+We then fit a new set of mash models where the selected covariance matrix is paired with each of the remaining $k-1$ covariance matrices.
+The most likely pair of covariance matrices are then selected and the process is repeated for all possible $k-2$ matrix triplets.
+This process is repeated until one of two stop conditions are met: (i) all covariance matrices have been added to the mash model, resulting in a final model with all of the original $k$ covariance matrices are included or (ii) a likelihood ratio test with one degree of freedom between the likelihood of the current model and the model from the previous iteration has a *p*-value \> 0.05.
+The result is a set of log likelihood estimates for each model and the stepwise ‘path’ of covariance matrices that best explain the observed effect sizes and standard errors observed for the phenotype of interest.
+
+We applied the greedy *mash* algorithm to the entire set of canonical, data-driven, and GxWeather covariance matrices described in Section S1.
+We used the random subset of SNP effect estimates, described in Section S3, in the greedy *mash* algorithm in order to test covariance matrices on an unbiased representation of all effects across the genome.
+The code to replicate this analysis can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/analysis/mash_greedy_algorithm/mash_h2_diag_covar/ in four directories: all_inputs/, gulf_inputs/, midwest_flowering/, and midwest_greenup/.
+For our six models of two phenological traits in three genetic subpopulations, between eight and twelve iterations of the greedy algorithm were necessary before a stop condition was met (@tbl-greedy).
+The relatively small number of necessary iterations, and correspondingly small number of covariance matrices selected, indicate that the hypothesis covariance matrices were highly correlated with one another and that inclusion of all the hypothesis covariance matrices introduced redundancy in the model.
+Most notably, the canonical "Identity" covariance matrix, representing independent effects, was the first matrix selected in all six applications of the greedy algorithm (@tbl-greedy).
+This indicated that a large number of independent effects were present in our analyses of both phenological traits.
+In addition, most or all of the available canonical covariance matrices, which specifed garden-specific effects, were selected before any GxWeather covariance matrices or data-driven covariance matrices (@tbl-greedy).
+Finally, at least one GxWeather covariance matrix was selected in each greedy algorithm application; in contrast, only one data-driven covariance matrix was selected (@tbl-greedy).
+
 ## Section S5. *Mash* to jointly re-estimate SNP effects across eight common gardens {.unnumbered}
+
+To evaluate the prevalence and kinds of covariance patterns of SNP effects across our eight common gardens, we used multivariate adaptive shrinkage (*mash*) to jointly re-estimate SNP effect estimates for our strong effects subset using the set of covariance matrices selected by the greedy *mash* algorithm.
+Mash is a statistical method that allows estimation and comparison of many effects jointly across many different conditions; it improves on previous methods by allowing arbitrary correlations in effect sizes among conditions.
+Mash assigns mixture proportions for each SNP's effect onto each provided covariance matrix using maximum likelihood.
+Then, mash uses Bayes' theorem to shrink effects for each SNP towards the set of covariance matrices in accordance to their mixture proportions.
+The ability to specify multiple types of covariance matrix is an important advantage *mash* offers for studying patterns of GxE: user-specified covariance matrices allow hypothesis testing of specific environmental drivers for each SNP's effect, while the data-driven covariance matrices mash creates allow exploration of additional unexplained patterns of covariation.
+
+The code used to run *mash* can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 1325.
 
@@
 | Phenology type | Subpopulation | Effects Selected per GWAS, Strong Subset | Effects Selected per GWAS, Random Subset | Total Markers, Strong Subset | Total Markers, Random Subset |
-|:-----------|--------------:|-----------:|-----------:|------------|------------|
+|:-----------|-----------:|-----------:|-----------:|------------|------------|
 | Greenup Date   |          Gulf |                                    19000 |                                    38000 | 45931                        | 91862                        |
@@
 
+| Greedy Algorithm Iteration \# |     Gulf Greenup Date |           Midwest Greenup Date |    Both Greenup Date | Gulf Flowering Date     | Midwest Flowering Date         | Both Flowering Date     |
+|:----------|----------:|----------:|-----------:|------------|------------|------------------|
+| 1                             |              Identity |                       Identity |             Identity | Identity                | Identity                       | Identity                |
+| 2                             |         Simple Het. 2 |               Single Effect MO |     Single Effect MI | Simple Het. 1           | Simple Het. 1                  | Single Effect TX3       |
+| 3                             |      Single Effect MO |               Single Effect NE |     Single Effect SD | Single Effect MI        | Single Effect MI               | Single Effect TX1       |
+| 4                             |      Single Effect NE |              Single Effect TX2 |     Single Effect NE | Single Effect TX3       | Single Effect TX2              | Single Effect OK        |
+| 5                             | Daylength (14d prior) |              Single Effect TX1 |    Single Effect TX2 | Single Effect SD        | Single Effect MO               | Single Effect MI        |
+| 6                             |      Single Effect MI |               Single Effect MI |     Single Effect OK | Single Effect OK        | Single Effect TX3              | Single Effect NE        |
+| 7                             |      Single Effect SD |               Single Effect OK |    Single Effect TX3 | Single Effect NE        | Daylength change (s), 2d prior | Single Effect TX2       |
+| 8                             |     Single Effect TX2 |              Single Effect TX3 |        Simple Het. 1 | Single Effect TX1       | Single Effect SD               | Single Effect MO        |
+| 9                             |    Cumulative GDD, 6d |                  Simple Het. 1 |    Single Effect TX1 | Single Effect TX2       | \-                             | Single Effect SD        |
+| 10                            |     Single Effect TX3 | Daylength change (s), 7d prior |     Single Effect MO | Single Effect MO        | \-                             | Simple Het. 1           |
+| 11                            |   Data-driven $PCA_2$ |                             \- |         Temp Ave, 1d | Cumulative rainfall, 7d | \-                             | Cumulative rainfall, 1d |
+| 12                            |                    \- |                             \- | Daylength change (s) | \-                      | \-                             | \-                      |
+
+: Covariance matrices selected in each iteration of the greedy *mash* algorithm, for each phenological trait and genetic subpopulation.
+Naming matches the conventions in Figure 2 of the main text.
+{#tbl-greedy}
+
+{{< pagebreak >}}
+
 ## Datasets {.unnumbered}
-
-If your document relies on movies or datasets, please list them here with their captions. Use the `movie{your caption}` and `\dataset{file_name.ext}{your caption}` commands to do so.
 

### 2024-06-10 R/Analysis_v1.2_mash_using_greedy_covar.qmd
revision2: edit M&M based on Tom's March comments and new SI

@@
 ASReml h2 greenup analysis
+
 ```{r}
@@
 get BLUPs from ASReml
+
 ```{r}
@@
 set up phenotype df for gwas
+
 ```{r}
@@
 NB: only the single_site BLUPs & GWAS that I generated were used in the subsequent analysis.
+
 ```{r}
@@
 # ---
+
 # Step 5: Redo writeup & figures

### 2024-06-10 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
revision2: edit M&M based on Tom's March comments and new SI

@@
 
-## Section S1. Covariance matrices generated to provide multiple ways SNP effects could covary across common gardens {.unnumbered}
+## Section S1. Generation of covariance matrices to provide multiple ways SNP effects could covary across common gardens {.unnumbered}
 
 A trait measured in two environments can be considered two traits that are genetically correlated [@falconer1952].
-By extension, a phenological trait measured at eight common gardens can be modeled as eight genetically correlated traits represented by a covariance matrix, where diagonal entries represent narrow-sense heritability at a common garden, or the additive genetic variance present in the population at that garden, and non-diagonal entries represent the genetic correlation between the trait in a pair of environments.
+By extension, a phenological trait measured at eight common gardens can be modeled as eight genetically correlated traits represented by a covariance matrix, where diagonal entries represent the additive genetic variance present in the population at that garden, and non-diagonal entries represent the genetic correlation between the trait in a pair of environments.
 
 We were interested in the types of genetic correlation that could be expressed by different alleles across the genome and reasoned that different loci could have different patterns of genetic correlation.
-For example, the effects of a locus, $A$, could covary with a weather-based cue if $A$ had a common allele, $a$, that was responsive to this cue.
+For example, the effects of an allele of locus $A$ could covary with a weather-based cue if $A$ had a common allele, $a$, that was responsive to this cue.
 In contrast, the effects of a different locus $B$ might have an effect at only one garden, for example if a pathogen was only present at one garden, and $B$ has a common allele $b$ that is resistant to that pathogen.
@@
 For example, the cumulative sum of the rainfall that occurred seven days prior to the flowering date was used to create one possible GxWeather cue variable for flowering date.
-The code for these calculations can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 390.
+The code for these calculations can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 394.
 
@@
 Finally, all covariance matrices were rescaled so that the maximum diagonal element of the rescaled matrix is 1, as in [@Urbut2019].
-Code used to generate these matrices can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 716.
+Code used to generate these matrices can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 720.
 
@@
 To obtain initial effect estimates for each single nucleotide polymorphism (SNP) for each combination of phenological trait, genetic subpopulation, and common garden, we conducted univarate genome-wide association (GWAS) on site-specific best linear unbiased predictors (BLUPs) using the switchgrassGWAS package and the methods in [@Lovell2021].
-The full SNP diversity panel generated in [@Lovell2021] was subset within each subpopulation for all subsequenct analyses to retain only SNPs with $\leq$ 20% missing data and minor allele frequencies \> 0.05.
+The full SNP diversity panel generated in [@Lovell2021] was subset within each subpopulation for all analyses to retain only SNPs with $\leq$ 20% missing data and minor allele frequencies \> 0.05.
 This subsetting resulted in 8.8 million SNPs retained for the Midwest subpopulation, 10.3 million SNPs retained for the Gulf subpopulation, and 12.3 million SNPs retained for Both subpopulations.
@@
 ASReml can specify considerably more complex variance structures; however, we were interested in specifying covariance structures using *mash*, which allows the specification of multiple covariance structures; hence, we used a relatively simple model to calculate BLUPs for use in univariate GWAS.
-The code for these calculations can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 74.
+The code for these calculations can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 75.
 
@@
 To control for population structure and run GWAS, this function uses a three-part process.
-First, a fast truncated singular value decomposition (SVD) is used, which has initial pruning of regions in long-range linkage disequilibrium (LD) and iterative pruining and removal of regions in long-range LD [@privé2017a].
+First, a fast truncated singular value decomposition (SVD) is generated from the SNP data, which has initial pruning of regions in long-range linkage disequilibrium (LD) and iterative pruning and removal of regions in long-range LD [@privé2017a].
 By default, 15 vectors of singular values, $k$, are computed.
-Pruning and removal of long-range LD regions are recommended in human genetics to best control for population structure [@abdellaoui2013; @privé2017; @price2008].
+Pruning and removal of long-range LD regions are recommended in research in human genetics to best control for population structure [@abdellaoui2013; @privé2017; @price2008].
 Second, linear regression models using 0-15 of these vectors of singular values were used to run GWAS and to calculate the genomic control coefficient, $\lambda_{GC}$, for each $k$ [@Devlin1999].
-Third, we chose the lowest value of $k$ that made $\lambda_{GC}$ closest to 1 to correct for population structure, and ran GWAS using a linear regression model with this $k$ for population structure correction.
+Third, we chose the lowest value of $k$ that made $\lambda_{GC}$ closest to 1, and ran GWAS using a linear regression model with this $k$ for population structure correction.
 
@@
 
-To make the subsequent analyses using *mash* computationally feasible, we followed the author's [recommended analysis outline](https://stephenslab.github.io/mashr/articles/eQTL_outline.html) for large datasets using *mash* and considered subsets of the millions of effects at different stages of the mash analysis.
+To make the subsequent analyses using *mash* computationally feasible, we followed the author's [recommended analysis outline](https://stephenslab.github.io/mashr/articles/eQTL_outline.html) for large datasets using *mash* and considered subsets of effects at different stages of the mash analysis.
 We extracted two subsets of tests from our complete GWAS effect estimate datasets: 1) effects from a subset of "strong" tests corresponding to stronger effects on our traits; 2) results from a random subset of all tests to correspond to an unbiased representation of all effects.
-We also clumped effects before selection so that signal strength was standardized across different LD blocks in the genome, and not limited to closely linked associations.
+We also clumped effects before selection of both subsets so that signal strength was standardized across different LD blocks in the genome, and not limited to closely linked associations.
 
@@
 
-## Section S4. Greedy *mash* algorithm used to select covariance matrices that significantly improved the model log-likelihood {.unnumbered}
+## Section S4. Greedy *mash* algorithm to select covariance matrices that significantly improved the model log-likelihood {.unnumbered}
 
-Because we were particularly interested in determining whether genomic loci varied in their patterns of genetic covariance across gardens, specifying the optimal set of covariance matrices in *mash* was an important model-building problem.
+Because we were particularly interested in determining whether genomic loci varied in their patterns of genetic covariance across gardens, and thus loaded on to different covariance matrices in *mash*, specifying the optimal set of covariance matrices in *mash* was an important model-building problem.
 If very similar covariance matrices are included when fitting a model in *mash*, *mash* will arbitrarily assign effect loadings across these matrices, as all assignments will lead to the same fit to the effect-size data.
@@
 Specifically, given $k$ covariance matrices for each phenotype, and $m$ corresponding effect sizes and standard errors across $n$ contexts of interest, we first fit a set of mash models with each of the $k$ matrices separately.
-We then identify the model with the maximum log likelihood estimates and select the corresponding matrix, leaving $k-1$ covariance matrices.
+We then identified the model with the maximum log likelihood estimates and selected the corresponding matrix, leaving $k-1$ covariance matrices.
 We then fit a new set of mash models where the selected covariance matrix is paired with each of the remaining $k-1$ covariance matrices.
-The most likely pair of covariance matrices are then selected and the process is repeated for all possible $k-2$ matrix triplets.
-This process is repeated until one of two stop conditions are met: (i) all covariance matrices have been added to the mash model, resulting in a final model with all of the original $k$ covariance matrices are included or (ii) a likelihood ratio test with one degree of freedom between the likelihood of the current model and the model from the previous iteration has a *p*-value \> 0.05.
+The most likely pair of covariance matrices was then selected and the process was repeated for all possible $k-2$ matrix triplets.
+This process was repeated until one of two stop conditions was met: (i) all covariance matrices have been added to the mash model, resulting in a final model with all of the original $k$ covariance matrices are included or (ii) a likelihood ratio test with one degree of freedom between the likelihood of the current model and the model from the previous iteration has a *p*-value \> 0.05.
 The result is a set of log likelihood estimates for each model and the stepwise ‘path’ of covariance matrices that best explain the observed effect sizes and standard errors observed for the phenotype of interest.
@@
 For our six models of two phenological traits in three genetic subpopulations, between eight and twelve iterations of the greedy algorithm were necessary before a stop condition was met (@tbl-greedy).
-The relatively small number of necessary iterations, and correspondingly small number of covariance matrices selected, indicate that the hypothesis covariance matrices were highly correlated with one another and that inclusion of all the hypothesis covariance matrices introduced redundancy in the model.
+The relatively small number of necessary iterations, and correspondingly small number of covariance matrices selected, indicate that the hypothesis covariance matrices were highly correlated with one another and that inclusion of all the hypothesis covariance matrices would introduce redundancy into the model.
 Most notably, the canonical "Identity" covariance matrix, representing independent effects, was the first matrix selected in all six applications of the greedy algorithm (@tbl-greedy).
@@
 In addition, most or all of the available canonical covariance matrices, which specifed garden-specific effects, were selected before any GxWeather covariance matrices or data-driven covariance matrices (@tbl-greedy).
-Finally, at least one GxWeather covariance matrix was selected in each greedy algorithm application; in contrast, only one data-driven covariance matrix was selected (@tbl-greedy).
+Finally, at least one GxWeather covariance matrix was selected in each greedy algorithm application; in contrast, one data-driven covariance matrix was selected in just one of the six models (@tbl-greedy).
 
@@
 To evaluate the prevalence and kinds of covariance patterns of SNP effects across our eight common gardens, we used multivariate adaptive shrinkage (*mash*) to jointly re-estimate SNP effect estimates for our strong effects subset using the set of covariance matrices selected by the greedy *mash* algorithm.
-Mash is a statistical method that allows estimation and comparison of many effects jointly across many different conditions; it improves on previous methods by allowing arbitrary correlations in effect sizes among conditions.
-Mash assigns mixture proportions for each SNP's effect onto each provided covariance matrix using maximum likelihood.
-Then, mash uses Bayes' theorem to shrink effects for each SNP towards the set of covariance matrices in accordance to their mixture proportions.
-The ability to specify multiple types of covariance matrix is an important advantage *mash* offers for studying patterns of GxE: user-specified covariance matrices allow hypothesis testing of specific environmental drivers for each SNP's effect, while the data-driven covariance matrices mash creates allow exploration of additional unexplained patterns of covariation.
+*Mash* is a statistical method that allows estimation and comparison of many effects jointly across many different conditions; it improves on previous methods by allowing arbitrary correlations in effect sizes among conditions.
+*Mash* assigns mixture proportions for each SNP's effect onto each provided covariance matrix using maximum likelihood.
+Then, *mash* uses Bayes' theorem to shrink effects for each SNP towards the set of covariance matrices in accordance to their mixture proportions.
+The ability to specify multiple types of covariance matrix is an important advantage *mash* offers for studying patterns of GxE: user-specified covariance matrices allow hypothesis testing of specific drivers for each SNP's effect, such as weather-related cues, while the data-driven covariance matrices *mash* creates allow exploration of additional unexplained patterns of covariation.
 
-The code used to run *mash* can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 1325.
+The code used to run *mash* can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 1326.
 
@@
 | Phenology type | Subpopulation | Effects Selected per GWAS, Strong Subset | Effects Selected per GWAS, Random Subset | Total Markers, Strong Subset | Total Markers, Random Subset |
-|:-----------|-----------:|-----------:|-----------:|------------|------------|
+|:---------------|--------------:|-----------------------------------------:|-----------------------------------------:|------------------------------|------------------------------|
 | Greenup Date   |          Gulf |                                    19000 |                                    38000 | 45931                        | 91862                        |
@@
 | Greedy Algorithm Iteration \# |     Gulf Greenup Date |           Midwest Greenup Date |    Both Greenup Date | Gulf Flowering Date     | Midwest Flowering Date         | Both Flowering Date     |
-|:----------|----------:|----------:|-----------:|------------|------------|------------------|
+|:------------------------------|----------------------:|-------------------------------:|---------------------:|-------------------------|--------------------------------|-------------------------|
 | 1                             |              Identity |                       Identity |             Identity | Identity                | Identity                       | Identity                |

### 2024-06-10 manuscript/revision2/revision2.qmd
revision2: edit M&M based on Tom's March comments and new SI

@@
 
-Here, we estimate GxWeather for two phenological timings in switchgrass by assigning patterns of genetic effects on phenology for cloned individuals grown at eight common gardens to many patterns of weather covariance at these gardens.
-To do this, we phenotype a diversity panel of hundreds of switchgrass genotypes from the Midwest and Gulf subpopulations for the timing of vegetative and reproductive development at eight common garden locations spanning 17 degrees of latitude.
-These gardens cover the majority of the latitudinal and climatic range of switchgrass and capture the most comprehensive picture to date of the environmental variation this species encounters.
+Here, we test for GxWeather for two phenological timings by assigning patterns of genetic effects on phenology at eight common gardens to many patterns of weather covariance at these gardens.
+To do this, we phenotype a diversity panel of hundreds of switchgrass genotypes from the Midwest and Gulf subpopulations for the timing of vegetative and reproductive development.
+We do this at eight common garden locations spanning 17 degrees of latitude: these gardens cover the majority of the latitudinal and climatic range of switchgrass and capture the most comprehensive picture to date of the environmental variation this species encounters.
 We define multiple ways phenological timings might covary with weather (Table 1) and additional ways phenological timings might vary by garden (SI Appendix, Section S1), then jointly re-estimate genetic effects on these timings at all eight common gardens using the set of these covariance matrices that significantly improved the modeled log-likelihood when included (SI Appendix, Section S2-S5) @urbut2019flexible.
-We use the Bayesian framework *mash* (multivariate adaptive shrinkage) developed by @urbut2019flexible, to refine effect size estimates from genome-wide association (GWAS) conducted on individuals at each of the eight gardens.
-*mash* allows us to identify and specify multiple covariance structures among genetic effect estimates across sites, including structures that represent covariance in weather variables of interest.
+We use the Bayesian framework *mash* (multivariate adaptive shrinkage) developed by @urbut2019flexible, to refine effect size estimates from genome-wide association (GWAS).
+*Mash* allows us to identify and specify multiple covariance structures among genetic effect estimates across sites, including structures that represent covariance in weather variables of interest.
 Importantly, this method circumvents statistical biases in detecting genetic effects with the same or opposite signs @10.1093/biostatistics/kxw041.
-To confirm our genetic mapping of GxWeather, we compare our significant posterior genetic effects locations from *mash* to mapping results from an outbred mapping population grown at the same sites.
-Our analyses allow us to describe the weather cues and genetic variation affecting phenology in two divergent natural populations of switchgrass across the species' latitudinal range.
+To confirm our genetic mapping of GxWeather, we compare genomic locations of the significant posterior effect estimates from *mash* to mapping results from an outbred mapping population grown at the same sites.
+Our analyses allow us to describe the weather cues and types of GxE affecting phenology in two divergent natural populations of switchgrass across the species' latitudinal range.
 
@@
 Genotypes from the Gulf and Midwest subpopulations had distinct phenological timings and distinct patterns of phenological correlations across our eight common garden sites (@fig-map).
-At the three Texas common gardens (hereafter 'Texas' gardens), located within the natural range of the Gulf subpopulation, Gulf vegetative growth (hereafter 'green-up') occurred before Midwestern green-up, and Gulf reproductive growth (hereafter 'flowering') occurred after Midwestern flowering (@fig-map A).
-At the four northernmost common gardens (hereafter 'North' gardens), located within the natural range of the Midwest subpopulation, both Gulf green-up and flowering occurred after Midwest green-up and flowering.
-At the Oklahoma common garden, located near the natural range limits of both the Gulf and the Midwest subpopulations, Gulf and Midwest green-up occurred over the same period, and Gulf flowering occurred after Midwestern flowering (@fig-map A).
-These patterns led to strong negative phenotypic correlations for the timing of the start of vegetative growth between the North and Texas gardens, particularly in the Gulf and across all individuals from the Gulf and Midwest subpopulations (hereafter, 'Both' subpopulations), and contributed to positive phenotypic correlations for the timing of flowering that had larger magnitudes at more northern gardens (@fig-map B).
+At the three Texas common gardens (hereafter 'Texas' gardens), located within the natural range of the Gulf subpopulation, the onset of vegetative growth for Gulf genotypes (hereafter 'green-up') occurred before Midwestern green-up, and the onset of reproductive growth for Gulf genotypes (hereafter 'flowering') occurred after Midwestern flowering (@fig-map A).
+At the four northernmost common gardens (hereafter 'North' gardens), located within the natural range of the Midwest subpopulation, both green-up and flowering of Gulf genotypes occurred after that of Midwest genotypes.
+At the Oklahoma common garden, located near the natural range limits of both the Gulf and the Midwest subpopulations, Gulf and Midwest green-up occurred over the same period, and Gulf genotype flowering occurred after Midwestern flowering (@fig-map A).
+These patterns led to strong negative phenotypic correlations for the onset of vegetative growth between the North and Texas gardens, particularly in the Gulf and across all individuals from the Gulf and Midwest subpopulations (hereafter, 'Both' subpopulations), and contributed to positive phenotypic correlations for the onset of flowering that had larger magnitudes at more northern gardens (@fig-map B).
 
@@
 
-### Genotype-by-environment effects on green-up and flowering as functions of weather-based cues
+## Scoring onset of vegetative and reproductive development in two mapping panels
 
 In 2019, we scored two phenological events every two days in two mapping populations of switchgrass, a diversity panel and a pseudo-F2 cross, planted at eight common garden locations [@lovell2021; @milano2016genetic; @lowry2019qtl].
-We scored the start of vegetative growth, or green-up date, as the day of the year when 50% of the tiller area of the crown of the plant cut the previous year had green growth.
-The start of reproductive growth, or flowering date, was the day of the year when 50% of the plant tillers had panicles undergoing anthesis.
-We scored green-up and flowering as day of the year, then linked these dates to multiple weather-based environmental factors measured daily at each common garden (Table 1, SI Appendix, Section S1).
+We scored the onset of vegetative growth, or green-up date, as the day of the year when 50% of the tiller area of the crown of the plant cut the previous year had green growth.
+The onset of reproductive growth, or flowering date, was the day of the year when 50% of the plant tillers had panicles undergoing anthesis.
 
@@
 To allow for the possibility that different subpopulations had different strengths of connection between our phenotypes and genotypes [@kortefarlow2013], we conducted three sets of genetic analyses: on Gulf and Midwest genotypes separately, and on both subpopulations together ('Both' subpopulations).
-Analyses to determine narrow-sense heritability (h^2^) for green-up and flowering were done using linear mixed models and followed @lovell2021genomic.
-Details on these models can be found in (SI Appendix, Section S6).
+Analyses to determine narrow-sense heritability (h^2^) for green-up and flowering were done using linear mixed models and followed @lovell2021genomic (SI Appendix, Section S6).
 
-### Mapping major patterns of genotype-by-environment effects on green-up and flowering
+To confirm candidate genomic regions and patterns of allelic effects found in the diversity panel, we analyzed flowering in an outbred pseudo-F2 cross between four individuals, two Midwest and two Gulf individuals.
+The formation of this mapping population has been described previously [@milano2016genetic]; additional details on QTL mapping can be found in SI Appendix, Section S7.
+To be directly comparable to the diversity panel data, only 2019 phenology data from the pseudo-F2 cross from the same eight common garden sites were used.
 
-To evaluate the prevalence and kinds of covariance patterns of SNP effects across our common gardens, we used multivariate adaptive shrinkage (*mash*) on SNP effect estimates from the diversity panel [@urbut2019flexible].
-To do this, we first conducted univarate genome-wide association (GWAS) on site-specific best linear unbiased predictors (BLUPs) for the timing of vegetative growth and flowering, using the switchgrassGWAS package and the methods in [@lovell2021] (SI Appendix, Section S2).
-We used the effect estimates for single nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect estimates for the two traits at each garden.
-We then jointly modeled genetic effect estimates, and the mixture of ways these effects might covary, across all eight common gardens, using a two-part procedure: a greedy algorithm to choose covariance structures using a random SNP subset, followed by a final run of mash with the best covariance structures using a strong SNP subset (SI Appendix, Section S3-S5).
-In brief, we first modeled genetic effects with different combinations of covariance structures, using the effect size estimates from subsets of 71-127K relatively unlinked (r^2^ \< 0.2), randomly selected SNPs used in the GWAS.
-The number of SNPs was chosen to result in an approximately million entry matrix for each trait & subpopulation, where matrix rows were SNPs and columns were univariate GWAS.
-In these models (hereafter referred to as "random"), we then specified a variety of potential covariance structures among genetic effect size estimates from each garden (as expanded on below) and used a greedy algorithm to iteratively select and add covariance structures that significantly improved the log likelihood of the *mash* model.
-Given the high correlation between some of the covariance patterns we tested, this approach allowed us to select only the subset of patterns of covariance that significantly improved the fit of the jointly inferred structure of genetic effect estimates present in the data.
-We then selected 19K - 33K relatively unlinked (r^2^ \< 0.2) SNPs from each univariate GWAS that had the largest effect estimates in each GWAS (hereafter "strong" effects).
-Notably, the majority of these "strong" effects were not significant in univariate GWAS.
-In many cases, there was overlap in the markers with the largest effect estimates at different locations; thus, 19 - 33K SNPs selected for each of eight univariate GWAS ultimately gave sets of 35K - 63K SNPs.
-We jointly re-estimated these effect size estimates at all eight sites using the set of significant covariance matrices that most improved the model fit on the random effects (SI Appendix, Section S5, Datasets 1-6).
+## Joint re-estimation of SNP effects to assess the frequency of rank-changing GxE and assignment of genome-wide patterns of GxE and GxWeather
 
-We generated hypothesis-based covariance matrices derived from correlations in environmental cues in the green-up or flowering date windows for the three populations (Table 1; SI Appendix, Section S1).
-These covariance matrices represent correlations between identical genotypes drawn from a specific population at pairs of common gardens; covariances near one mean that the population has a strong, positive linear relationship in individual responses at that pair of gardens, while covariances near zero mean that there is no relationship within the population for individual responses at that pair of gardens.
-*Mash* SNP effects will undergo strong shrinkage towards one another in the first case, and little shrinkage in the second case.
-*Mash* also generates data-driven covariance matrices corresponding to major patterns of SNP effects present in the data.
-We generated six data-driven matrices per *mash* run, five produced by singular value decomposition (SVD) of an overall matrix.
+We were interested in specifying genetic models for trait variation that allowed more than one form of GxE, as we reasoned that different loci should display different forms of GxE.
+In addition, we were interested in an unbiased estimation of the frequency of rank-changing GxE relative to other forms of GxE, as the presence of rank-changing GxE at the level of individual loci is a key theoretical prediction of local adaptation.
+*Mash* allowed us to both specify multiple forms of GxE and GxWeather and conduct unbiased statistical tests for when SNP effects changed sign between common gardens.
 
-Last, we characterized the overall patterns of GxE in the set of SNPs where there was pairwise significance of effects at pairs of gardens.
+To use *mash* on our diversity panel, we had to specify both a relatively uncorrelated set of covariance matrices, which in our case defined types of GxE and GxWeather between gardens, and we had to specify subsets of SNP effect estimates and standard errors for our traits at each common garden.
+To specify a set of covariance matrices, we first defined many covariance matrices, including GxWeather matrices that represented the correlation in weather cues between gardenes before the phenological event (SI Appendix, Section S1), then implemented a model selection approach that used a greedy algorithm to evaluate if the log likelihood of the *mash* model was significantly improved as additional covariance matrices were included (SI Appendix, Section S4).
+To specify subsets of SNP effect estimates and standard errors to use for both the greedy algorithm and with the optimal set of covariance matrices, we first calculated best linear unbiased predictors (BLUPs) for each phenological trait in each genetic subpopulation and each common garden (SI Appendix, Section S2).
+Next, we determined effect estimates for 8.8 to 12.3 million SNPs per subpopulation by conducting garden-specific GWAS on these BLUPs using $k$ vectors of singular values to correct for population structure, where $k$ was the smallest integer value that made the genomic control coefficient closest to 1 (SI Appendix, Section S2).
+Singular values were computed using singular value decomposition of the matrix of all SNPs, with iterative SNP pruning and removal of regions in long-range linkage disequilibrium [@privé2017].
+Third, to make the *mash* models computationally feasible, we extracted two subsets of SNP effect estimates and standard errors from our GWAS effect estimates: 1) effects from a subset of "strong" tests corresponding to stronger effects on our traits; 2) results from a random subset of all tests to correspond to an unbiased representation of all effects (SI Appendix, Section S3).
+We used the subset of random effects in our greedy *mash* algorithm (SI Appendix, Section S4) and used the subset of strong effects in our *mash* models with the optimal set of covariance matrices (SI Appendix, Section S5).
+
+The loadings of genetic effects onto the multiple covariance structures specified in our *mash* models provided information on genome-wide patterns of SNP-environment interaction.
+In addition, the the GxWeather covariance structures allowed hypothesis testing of specific weather variables as cues for the start of vegetative and reproductive growth.
+Say that a SNP in the gene Flowering Locus C (FLC), a well-known flowering time regulator, controls flowering in a photoperiod-dependent manner.
+In that case, the joint estimate of effects for that SNP could have a high mixture proportion, or mass, on a covariance matrix created using a photoperiod-based environmental cue, such as day length at some interval prior to flowering.
+In our data, we would infer that the effect of that SNP on flowering was caused by a response to the environmental cue used to construct the GxWeather covariance structure with the largest mass.
+
+Our joint re-estimation of SNP effects also allowed us to characterize the overall patterns of GxE in the set of SNPs where there was pairwise significance of effects at pairs of gardens.
 To do this, we used the 'get_GxE' function of the switchgrassGWAS R package.
-First, this determines the set of SNPs with evidence of significant effects in both conditions for all pairs of conditions using local false sign rates (lfsr) as the significance criteria.
+First, this function determines the set of SNPs with evidence of significant effects in both conditions for all pairs of conditions using local false sign rates (lfsr) as the significance criteria.
 Then, this function determines if effects significant in both conditions are of opposite sign.
@@
 However, using the lfsr to test for allelic effects that differ in sign does not undercount these occurences, as this statistic answers a fundamentally different question.
-For each effect j, the lfsr~j~ is defined as the probability that we make an error in the sign of effect j if we were forced to declare the effect positive or negative @10.1093/biostatistics/kxw041.
+For each effect $j$, the $lfsr_j$ is defined as the probability that we make an error in the sign of effect $j$ if we were forced to declare the effect positive or negative @10.1093/biostatistics/kxw041.
 Thus, rather than asking "Are these two effects different?" - as we reasonably expect two effects to be, even if this difference cannot be measured - the local false sign rate answers a more meaningful question: Can we be confident in the sign of this effect?
@@
 
-### Confirmation of genotype-by-environment effects using an independent mapping population
-
-To confirm candidate genomic regions and patterns of allelic effects found in the diversity panel, we analyzed flowering in an outbred pseudo-F2 cross between four individuals, two Midwest and two Gulf individuals.
-The formation of this mapping population has been described previously [@milano2016genetic]; additional details on QTL mapping can be found in SI Appendix, Section S7.
-To be directly comparable to the diversity panel data, only 2019 phenology data from the pseudo-F2 cross from the same eight common garden sites were used.
-
 # References

### 2024-06-10 manuscript/revision2/revision2.tex
revision2: edit M&M based on Tom's March comments and new SI

@@
 
-Here, we estimate GxWeather for two phenological timings in switchgrass
-by assigning patterns of genetic effects on phenology for cloned
-individuals grown at eight common gardens to many patterns of weather
-covariance at these gardens. To do this, we phenotype a diversity panel
-of hundreds of switchgrass genotypes from the Midwest and Gulf
-subpopulations for the timing of vegetative and reproductive development
-at eight common garden locations spanning 17 degrees of latitude. These
-gardens cover the majority of the latitudinal and climatic range of
-switchgrass and capture the most comprehensive picture to date of the
-environmental variation this species encounters. We define multiple ways
-phenological timings might covary with weather (Table 1) and additional
-ways phenological timings might vary by garden (SI Appendix, Section
-S1), then jointly re-estimate genetic effects on these timings at all
-eight common gardens using the set of these covariance matrices that
-significantly improved the modeled log-likelihood when included (SI
-Appendix, Section S2-S5) (36). We use the Bayesian framework \emph{mash}
-(multivariate adaptive shrinkage) developed by (36), to refine effect
-size estimates from genome-wide association (GWAS) conducted on
-individuals at each of the eight gardens. \emph{mash} allows us to
-identify and specify multiple covariance structures among genetic effect
-estimates across sites, including structures that represent covariance
-in weather variables of interest. Importantly, this method circumvents
-statistical biases in detecting genetic effects with the same or
-opposite signs (37). To confirm our genetic mapping of GxWeather, we
-compare our significant posterior genetic effects locations from
-\emph{mash} to mapping results from an outbred mapping population grown
-at the same sites. Our analyses allow us to describe the weather cues
-and genetic variation affecting phenology in two divergent natural
-populations of switchgrass across the species' latitudinal range.
+Here, we test for GxWeather for two phenological timings by assigning
+patterns of genetic effects on phenology at eight common gardens to many
+patterns of weather covariance at these gardens. To do this, we
+phenotype a diversity panel of hundreds of switchgrass genotypes from
+the Midwest and Gulf subpopulations for the timing of vegetative and
+reproductive development. We do this at eight common garden locations
+spanning 17 degrees of latitude: these gardens cover the majority of the
+latitudinal and climatic range of switchgrass and capture the most
+comprehensive picture to date of the environmental variation this
+species encounters. We define multiple ways phenological timings might
+covary with weather (Table 1) and additional ways phenological timings
+might vary by garden (SI Appendix, Section S1), then jointly re-estimate
+genetic effects on these timings at all eight common gardens using the
+set of these covariance matrices that significantly improved the modeled
+log-likelihood when included (SI Appendix, Section S2-S5) (36). We use
+the Bayesian framework \emph{mash} (multivariate adaptive shrinkage)
+developed by (36), to refine effect size estimates from genome-wide
+association (GWAS). \emph{Mash} allows us to identify and specify
+multiple covariance structures among genetic effect estimates across
+sites, including structures that represent covariance in weather
+variables of interest. Importantly, this method circumvents statistical
+biases in detecting genetic effects with the same or opposite signs
+(37). To confirm our genetic mapping of GxWeather, we compare genomic
+locations of the significant posterior effect estimates from \emph{mash}
+to mapping results from an outbred mapping population grown at the same
+sites. Our analyses allow us to describe the weather cues and types of
+GxE affecting phenology in two divergent natural populations of
+switchgrass across the species' latitudinal range.
 
@@
 three Texas common gardens (hereafter `Texas' gardens), located within
-the natural range of the Gulf subpopulation, Gulf vegetative growth
-(hereafter `green-up') occurred before Midwestern green-up, and Gulf
-reproductive growth (hereafter `flowering') occurred after Midwestern
-flowering (Figure~\ref{fig-map} A). At the four northernmost common
-gardens (hereafter `North' gardens), located within the natural range of
-the Midwest subpopulation, both Gulf green-up and flowering occurred
-after Midwest green-up and flowering. At the Oklahoma common garden,
+the natural range of the Gulf subpopulation, the onset of vegetative
+growth for Gulf genotypes (hereafter `green-up') occurred before
+Midwestern green-up, and the onset of reproductive growth for Gulf
+genotypes (hereafter `flowering') occurred after Midwestern flowering
+(Figure~\ref{fig-map} A). At the four northernmost common gardens
+(hereafter `North' gardens), located within the natural range of the
+Midwest subpopulation, both green-up and flowering of Gulf genotypes
+occurred after that of Midwest genotypes. At the Oklahoma common garden,
 located near the natural range limits of both the Gulf and the Midwest
 subpopulations, Gulf and Midwest green-up occurred over the same period,
-and Gulf flowering occurred after Midwestern flowering
+and Gulf genotype flowering occurred after Midwestern flowering
 (Figure~\ref{fig-map} A). These patterns led to strong negative
-phenotypic correlations for the timing of the start of vegetative growth
-between the North and Texas gardens, particularly in the Gulf and across
-all individuals from the Gulf and Midwest subpopulations (hereafter,
-`Both' subpopulations), and contributed to positive phenotypic
-correlations for the timing of flowering that had larger magnitudes at
-more northern gardens (Figure~\ref{fig-map} B).
+phenotypic correlations for the onset of vegetative growth between the
+North and Texas gardens, particularly in the Gulf and across all
+individuals from the Gulf and Midwest subpopulations (hereafter, `Both'
+subpopulations), and contributed to positive phenotypic correlations for
+the onset of flowering that had larger magnitudes at more northern
+gardens (Figure~\ref{fig-map} B).
 
@@
 
-\subsubsection{Genotype-by-environment effects on green-up and flowering
-as functions of weather-based
-cues}\label{genotype-by-environment-effects-on-green-up-and-flowering-as-functions-of-weather-based-cues}
+\subsection{Scoring onset of vegetative and reproductive development in
+two mapping
+panels}\label{scoring-onset-of-vegetative-and-reproductive-development-in-two-mapping-panels}
 
@@
 planted at eight common garden locations (35, 39, 40). We scored the
-start of vegetative growth, or green-up date, as the day of the year
+onset of vegetative growth, or green-up date, as the day of the year
 when 50\% of the tiller area of the crown of the plant cut the previous
-year had green growth. The start of reproductive growth, or flowering
+year had green growth. The onset of reproductive growth, or flowering
 date, was the day of the year when 50\% of the plant tillers had
-panicles undergoing anthesis. We scored green-up and flowering as day of
-the year, then linked these dates to multiple weather-based
-environmental factors measured daily at each common garden (Table 1, SI
-Appendix, Section S1).
+panicles undergoing anthesis.
 
@@
 narrow-sense heritability (h\textsuperscript{2}) for green-up and
-flowering were done using linear mixed models and followed (33). Details
-on these models can be found in (SI Appendix, Section S6).
-
-\subsubsection{Mapping major patterns of genotype-by-environment effects
-on green-up and
-flowering}\label{mapping-major-patterns-of-genotype-by-environment-effects-on-green-up-and-flowering}
-
-To evaluate the prevalence and kinds of covariance patterns of SNP
-effects across our common gardens, we used multivariate adaptive
-shrinkage (\emph{mash}) on SNP effect estimates from the diversity panel
-(36). To do this, we first conducted univarate genome-wide association
-(GWAS) on site-specific best linear unbiased predictors (BLUPs) for the
-timing of vegetative growth and flowering, using the switchgrassGWAS
-package and the methods in (40) (SI Appendix, Section S2). We used the
-effect estimates for single nucleotide polymorphisms (SNPs) from these
-GWAS as preliminary effect estimates for the two traits at each garden.
-We then jointly modeled genetic effect estimates, and the mixture of
-ways these effects might covary, across all eight common gardens, using
-a two-part procedure: a greedy algorithm to choose covariance structures
-using a random SNP subset, followed by a final run of mash with the best
-covariance structures using a strong SNP subset (SI Appendix, Section
-S3-S5). In brief, we first modeled genetic effects with different
-combinations of covariance structures, using the effect size estimates
-from subsets of 71-127K relatively unlinked (r\textsuperscript{2}
-\textless{} 0.2), randomly selected SNPs used in the GWAS. The number of
-SNPs was chosen to result in an approximately million entry matrix for
-each trait \& subpopulation, where matrix rows were SNPs and columns
-were univariate GWAS. In these models (hereafter referred to as
-``random''), we then specified a variety of potential covariance
-structures among genetic effect size estimates from each garden (as
-expanded on below) and used a greedy algorithm to iteratively select and
-add covariance structures that significantly improved the log likelihood
-of the \emph{mash} model. Given the high correlation between some of the
-covariance patterns we tested, this approach allowed us to select only
-the subset of patterns of covariance that significantly improved the fit
-of the jointly inferred structure of genetic effect estimates present in
-the data. We then selected 19K - 33K relatively unlinked
-(r\textsuperscript{2} \textless{} 0.2) SNPs from each univariate GWAS
-that had the largest effect estimates in each GWAS (hereafter ``strong''
-effects). Notably, the majority of these ``strong'' effects were not
-significant in univariate GWAS. In many cases, there was overlap in the
-markers with the largest effect estimates at different locations; thus,
-19 - 33K SNPs selected for each of eight univariate GWAS ultimately gave
-sets of 35K - 63K SNPs. We jointly re-estimated these effect size
-estimates at all eight sites using the set of significant covariance
-matrices that most improved the model fit on the random effects (SI
-Appendix, Section S5, Datasets 1-6).
-
-We generated hypothesis-based covariance matrices derived from
-correlations in environmental cues in the green-up or flowering date
-windows for the three populations (Table 1; SI Appendix, Section S1).
-These covariance matrices represent correlations between identical
-genotypes drawn from a specific population at pairs of common gardens;
-covariances near one mean that the population has a strong, positive
-linear relationship in individual responses at that pair of gardens,
-while covariances near zero mean that there is no relationship within
-the population for individual responses at that pair of gardens.
-\emph{Mash} SNP effects will undergo strong shrinkage towards one
-another in the first case, and little shrinkage in the second case.
-\emph{Mash} also generates data-driven covariance matrices corresponding
-to major patterns of SNP effects present in the data. We generated six
-data-driven matrices per \emph{mash} run, five produced by singular
-value decomposition (SVD) of an overall matrix.
-
-Last, we characterized the overall patterns of GxE in the set of SNPs
-where there was pairwise significance of effects at pairs of gardens. To
-do this, we used the `get\_GxE' function of the switchgrassGWAS R
-package. First, this determines the set of SNPs with evidence of
-significant effects in both conditions for all pairs of conditions using
-local false sign rates (lfsr) as the significance criteria. Then, this
-function determines if effects significant in both conditions are of
-opposite sign.
+flowering were done using linear mixed models and followed (33) (SI
+Appendix, Section S6).
+
+To confirm candidate genomic regions and patterns of allelic effects
+found in the diversity panel, we analyzed flowering in an outbred
+pseudo-F2 cross between four individuals, two Midwest and two Gulf
+individuals. The formation of this mapping population has been described
+previously (35); additional details on QTL mapping can be found in SI
+Appendix, Section S7. To be directly comparable to the diversity panel
+data, only 2019 phenology data from the pseudo-F2 cross from the same
+eight common garden sites were used.
+
+\subsection{Joint re-estimation of SNP effects to assess the frequency
+of rank-changing GxE and assignment of genome-wide patterns of GxE and
+GxWeather}\label{joint-re-estimation-of-snp-effects-to-assess-the-frequency-of-rank-changing-gxe-and-assignment-of-genome-wide-patterns-of-gxe-and-gxweather}
+
+We were interested in specifying genetic models for trait variation that
+allowed more than one form of GxE, as we reasoned that different loci
+should display different forms of GxE. In addition, we were interested
+in an unbiased estimation of the frequency of rank-changing GxE relative
+to other forms of GxE, as the presence of rank-changing GxE at the level
+of individual loci is a key theoretical prediction of local adaptation.
+\emph{Mash} allowed us to both specify multiple forms of GxE and
+GxWeather and conduct unbiased statistical tests for when SNP effects
+changed sign between common gardens.
+
+To use \emph{mash} on our diversity panel, we had to specify both a
+relatively uncorrelated set of covariance matrices, which in our case
+defined types of GxE and GxWeather between gardens, and we had to
+specify subsets of SNP effect estimates and standard errors for our
+traits at each common garden. To specify a set of covariance matrices,
+we first defined many covariance matrices, including GxWeather matrices
+that represented the correlation in weather cues between gardenes before
+the phenological event (SI Appendix, Section S1), then implemented a
+model selection approach that used a greedy algorithm to evaluate if the
+log likelihood of the \emph{mash} model was significantly improved as
+additional covariance matrices were included (SI Appendix, Section S4).
+To specify subsets of SNP effect estimates and standard errors to use
+for both the greedy algorithm and with the optimal set of covariance
+matrices, we first calculated best linear unbiased predictors (BLUPs)
+for each phenological trait in each genetic subpopulation and each
+common garden (SI Appendix, Section S2). Next, we determined effect
+estimates for 8.8 to 12.3 million SNPs per subpopulation by conducting
+garden-specific GWAS on these BLUPs using \(k\) vectors of singular
+values to correct for population structure, where \(k\) was the smallest
+integer value that made the genomic control coefficient closest to 1 (SI
+Appendix, Section S2). Singular values were computed using singular
+value decomposition of the matrix of all SNPs, with iterative SNP
+pruning and removal of regions in long-range linkage disequilibrium
+(42). Third, to make the \emph{mash} models computationally feasible, we
+extracted two subsets of SNP effect estimates and standard errors from
+our GWAS effect estimates: 1) effects from a subset of ``strong'' tests
+corresponding to stronger effects on our traits; 2) results from a
+random subset of all tests to correspond to an unbiased representation
+of all effects (SI Appendix, Section S3). We used the subset of random
+effects in our greedy \emph{mash} algorithm (SI Appendix, Section S4)
+and used the subset of strong effects in our \emph{mash} models with the
+optimal set of covariance matrices (SI Appendix, Section S5).
+
+The loadings of genetic effects onto the multiple covariance structures
+specified in our \emph{mash} models provided information on genome-wide
+patterns of SNP-environment interaction. In addition, the the GxWeather
+covariance structures allowed hypothesis testing of specific weather
+variables as cues for the start of vegetative and reproductive growth.
+Say that a SNP in the gene Flowering Locus C (FLC), a well-known
+flowering time regulator, controls flowering in a photoperiod-dependent
+manner. In that case, the joint estimate of effects for that SNP could
+have a high mixture proportion, or mass, on a covariance matrix created
+using a photoperiod-based environmental cue, such as day length at some
+interval prior to flowering. In our data, we would infer that the effect
+of that SNP on flowering was caused by a response to the environmental
+cue used to construct the GxWeather covariance structure with the
+largest mass.
+
+Our joint re-estimation of SNP effects also allowed us to characterize
+the overall patterns of GxE in the set of SNPs where there was pairwise
+significance of effects at pairs of gardens. To do this, we used the
+`get\_GxE' function of the switchgrassGWAS R package. First, this
+function determines the set of SNPs with evidence of significant effects
+in both conditions for all pairs of conditions using local false sign
+rates (lfsr) as the significance criteria. Then, this function
+determines if effects significant in both conditions are of opposite
+sign.
 
@@
 sign does not undercount these occurences, as this statistic answers a
-fundamentally different question. For each effect j, the
-lfsr\textsubscript{j} is defined as the probability that we make an
-error in the sign of effect j if we were forced to declare the effect
-positive or negative (37). Thus, rather than asking ``Are these two
-effects different?'' - as we reasonably expect two effects to be, even
-if this difference cannot be measured - the local false sign rate
-answers a more meaningful question: Can we be confident in the sign of
-this effect?
+fundamentally different question. For each effect \(j\), the \(lfsr_j\)
+is defined as the probability that we make an error in the sign of
+effect \(j\) if we were forced to declare the effect positive or
+negative (37). Thus, rather than asking ``Are these two effects
+different?'' - as we reasonably expect two effects to be, even if this
+difference cannot be measured - the local false sign rate answers a more
+meaningful question: Can we be confident in the sign of this effect?
 
@@
 
-\subsubsection{Confirmation of genotype-by-environment effects using an
-independent mapping
-population}\label{confirmation-of-genotype-by-environment-effects-using-an-independent-mapping-population}
-
-To confirm candidate genomic regions and patterns of allelic effects
-found in the diversity panel, we analyzed flowering in an outbred
-pseudo-F2 cross between four individuals, two Midwest and two Gulf
-individuals. The formation of this mapping population has been described
-previously (35); additional details on QTL mapping can be found in SI
-Appendix, Section S7. To be directly comparable to the diversity panel
-data, only 2019 phenology data from the pseudo-F2 cross from the same
-eight common garden sites were used.
-
 \section{References}\label{references}
@@
 (2013).}
+
+\bibitem[\citeproctext]{ref-privuxe92017}
+\CSLLeftMargin{42. }%
+\CSLRightInline{F. Privé, H. Aschard, M. G. B. Blum,
+\href{http://dx.doi.org/10.1101/190926}{Efficient management and
+analysis of large-scale genome-wide data with two r packages: Bigstatsr
+and bigsnpr} (2017).}
 

### 2024-06-11 R/SI_Appendix_Datasets.Rmd
revision2: Coherent M&M between Results, M&M, & SI Appendix

@@
 ```
+
+### Overlap with Hd1 from switchgrass functional validation study
+
+```{r}
+df_gam_fl <- read_csv(here("manuscript", "revision2", "SI_Appendix", "Dataset_6_Both_subpopulations_reproductive_growth_date.csv"))
+df_gam_fl <- read_csv(here("manuscript", "revision2", "SI_Appendix", "Dataset_4_Gulf_reproductive_growth_date.csv"))
+df_gam_fl <- read_csv(here("manuscript", "revision2", "SI_Appendix", "Dataset_5_Midwest_reproductive_growth_date.csv"))
+
+df_gam_fl |> 
+  separate(Marker, into = c("Chr", "Pos"), sep = "_", 
+           remove = FALSE, convert = TRUE) |> 
+  filter(Chr == "Chr04K", between(Pos, 23100000, 23200000)) |> 
+  pivot_longer(cols = matches("lfsr"), names_to = "Garden", values_to = "lfsr")
+
+get_significant_results(m_gam_fl_hyp) |> 
+  as.data.frame() |> 
+  rownames_to_column(var = "Marker") |> 
+  separate(Marker, into = c("Chr", "Pos"), sep = "_", 
+           remove = FALSE, convert = TRUE) |> 
+  filter(Chr == "Chr04K", between(Pos, 23100000, 23200000))
+
+mash_plot_effects(m = m_gam_fl_hyp, i = 13819)
+```
+Chr04K_23164108 in Both has log10BF ~ 2.3
+Chr04K_23179015 in Gulf has log10BF ~ 1.86
+Nearby SNPs _not_ significant in Midwest. log10BF ~ 0.1
 
@@
 ```
+
+## Figure S1
+
+Figure 3 is North x Texas GxE effects. Figure S1 is garden x garden GxE effects of individual loci.
+
+Most differences in sign were between TX1, the southernmost garden, and all other gardens (SI Appendix, Figure S1).
+Similarly, more effects that differed in magnitude included gardens in the Texas region (52.3-55.9%), and most effect pairs in the North region were not distinguishable (91.5%).
+
+## Figure S2
+
+QTL x Env effects
+
+We conducted quantitative trait loci (QTL) mapping of flowering as functions of five environmental cues that we also used as covariance matrices in mash, and identified eight QTL for flowering date, eight QTL for flowering as a function of day length change two days prior, one QTL for the start of vegetative growth, and two QTL for vegetative growth as a function of daylength change one day prior, all of which showed QTL by environment interactions (SI Appendix, Figure S2).

### 2024-06-11 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
revision2: Coherent M&M between Results, M&M, & SI Appendix

@@
 
-Figures should be cross-referenced like @fig-frog. Each figure should be on its own page. You can control this by placing a pagebreak shortcode, with `{{< pagebreak >}}`.
+Figures should be cross-referenced like @fig-qtl-by-env. Each figure should be on its own page. You can control this by placing a pagebreak shortcode, with `{{< pagebreak >}}`.
 
@@
 
-![Figure](figs/frog.eps){#fig-frog}
+![Figure](figs/frog.eps){#fig-garden-by-garden-gxe}
+
+\FloatBarrier <!-- always leave a blank line between this and figures --->
+
+{{< pagebreak >}}
+
+![Figure](figs/frog.eps){#fig-qtl-by-env}
 
@@
 | Phenology type | Subpopulation | Effects Selected per GWAS, Strong Subset | Effects Selected per GWAS, Random Subset | Total Markers, Strong Subset | Total Markers, Random Subset |
-|:---------------|--------------:|-----------------------------------------:|-----------------------------------------:|------------------------------|------------------------------|
+|:-----------|-----------:|-----------:|-----------:|------------|------------|
 | Greenup Date   |          Gulf |                                    19000 |                                    38000 | 45931                        | 91862                        |
@@
 | Greedy Algorithm Iteration \# |     Gulf Greenup Date |           Midwest Greenup Date |    Both Greenup Date | Gulf Flowering Date     | Midwest Flowering Date         | Both Flowering Date     |
-|:------------------------------|----------------------:|-------------------------------:|---------------------:|-------------------------|--------------------------------|-------------------------|
+|:----------|----------:|----------:|----------:|-----------|-----------|-----------|
 | 1                             |              Identity |                       Identity |             Identity | Identity                | Identity                       | Identity                |

### 2024-06-11 manuscript/revision2/revision2.qmd
revision2: Coherent M&M between Results, M&M, & SI Appendix

@@
 Genotypes from the Gulf and Midwest subpopulations had distinct phenological timings and distinct patterns of phenological correlations across our eight common garden sites (@fig-map).
-At the three Texas common gardens (hereafter 'Texas' gardens), located within the natural range of the Gulf subpopulation, the onset of vegetative growth for Gulf genotypes (hereafter 'green-up') occurred before Midwestern green-up, and the onset of reproductive growth for Gulf genotypes (hereafter 'flowering') occurred after Midwestern flowering (@fig-map A).
+At the three Texas common gardens (hereafter 'Texas' gardens), located within the natural range of the Gulf subpopulation, the onset of vegetative growth, or 'green-up date', for Gulf genotypes occurred before Midwestern green-up, and the onset of reproductive growth, or 'flowering date', for Gulf genotypes occurred after Midwestern flowering (@fig-map A).
 At the four northernmost common gardens (hereafter 'North' gardens), located within the natural range of the Midwest subpopulation, both green-up and flowering of Gulf genotypes occurred after that of Midwest genotypes.
@@
 
-### Inference of GxWeather effects for vegetative and reproductive timing using *mash*
+### Inference of genome-wide patterns of GxE and GxWeather for vegetative and reproductive timing
 
-We next looked for evidence of GxWeather by jointly re-estimating genetic effects of SNPs across all eight common gardens using the *mash* model, which can flexibly capture many modes of covariance across gardens (@fig-map D; SI Appendix, Sections S1-S5, Datasets 1-6).
-We specified three qualitative categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event (Table 1; SI Appendix, Section S1).
-For example, the more similar that the amount of rainfall prior to flowering is for genotypes A and B between gardens in MI and MO, the nearer these covariances will be to one.
-If rainfall amounts are uncorrelated, covariance will be near zero; if there is rank changing (e.g.
-high rainfall before flowering of A at MO but not MI, and high rainfall before flowering for B at MI but not MO), this covariance will be near negative one.
-We specified multiple covariance structures for the canonical, data-driven, and GxWeather categories, each of which represented a specific pattern of covariance in genetic effect size estimates that could be included in the model.
-The phenotypic correlations for the timing of vegetative growth had moderate negative correlations between the Texas and North gardens, particularly in the Gulf subpopulation and Both subpopulations (@fig-map B).
+We expected patterns of GxE and GxWeather to vary at the locus, subpopulation, and trait level; thus we were interested in modeling the different ways effects might covary across our eight common gardens.
+We jointly re-estimated genetic effects of a subset of SNPs across all eight common gardens using *mash* models, which can flexibly capture many modes of covariance across gardens (@fig-map D; SI Appendix, Sections S1-S5, Datasets 1-6).
+To obtain genetic effects, we used subsets of effects from garden- and subpopulation GWAS (SI Appendix, Section S2-S3; Table S1).
+To obtain covariance structures, we specified both covariance structures introduced in the initial *mash* manuscript, and GxWeather covariance matrices estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event (Table 1; SI Appendix, Section S1).
+Then, we used a greedy algorithm to select covariance matrices from each model's set that significantly improved the model likelihood (SI Appendix, Section S4; Table S2).
+Finally, we fit *mash* models using the selected set of covariance matrices on a subset relatively unlinked (r^2^ \< 0.2), 'strong' genetic effects with low *p*-values (SI Appendix, Section S5).
+
+The loadings of genetic effects onto the GxE and GxWeather covariance matrices included in each model provide information on genome-wide patterns of SNP-environment interaction.
+For example, the phenotypic correlations for the onset of vegetative growth had moderate negative correlations between the Texas and North gardens, particularly in the Gulf subpopulation and Both subpopulations (@fig-map B).
 If these phenotypic correlations have a genetic component, they could be partially or completely captured by the covariance structures specified in the *mash* model.
+In this case, the joint estimate of effects for many SNPs could have high mixture proportions, or mass, on a covariance matrix that captures the negative correlation between Texas and North gardens.
+In our data, daylength 14 days prior to green-up date has negative correlations between Texas and North gardens (@fig-covar A); if many SNP effects have mass on this GxWeather matrix, we would infer that the effects of those SNPs have that form of GxWeather covariation.
 
-The GxWeather covariance structures allow hypothesis testing of specific weather variables as cues for the start of vegetative and reproductive growth.
-Say that a SNP in the gene *Flowering Locus C* (FLC), a well-known flowering time regulator, controls flowering in a photoperiod-dependent manner.
-In that case, the joint estimate of effects for that SNP could have a high mixture proportion, or mass, on a covariance matrix created using a photoperiod-based environmental cue, such as day length at some interval prior to flowering.
-In our data, we would infer that the effect of that SNP on flowering was caused by a response to the environmental cue used to construct the GxWeather covariance structure with the largest mass.
-Overall, loadings of genetic effects on these GxWeather matrices in the model, along with the posterior effect size estimates, provide information on genome-wide patterns of SNP-environment interaction.
+Eight GxWeather covariance structures were selected by the greedy algorithm, one to two per subpopulation: five for green-up date, and three for flowering date (SI Appendix, Table S2).
+Of these eight, five had mass on them in *mash* models of the strong effects (@fig-covar A-B).
+Three of these five matrices had negative covariances between sites between Texas and North gardens (@fig-covar A), while two had all positive or near-zero covariances.
 
-We used a greedy algorithm to select covariance matrices from each model's set that significantly improved the model likelihood.
-Five GxWeather covariance structures were selected, zero to two per subpopulation: two for green-up date, and three for flowering date.
-Of these five, three had mass on them in *mash* models of the strong effects (@fig-covar A-B).
-Two of these three matrices had negative covariances between sites between Texas and North gardens (@fig-covar A), while one had all positive or near-zero covariances.
-SNP-associated phenotypic effects covaried with different weather-based cues in the Gulf & in Both subpopulations (@fig-covar B).
+For green-up date, SNP-associated phenotypic effects covaried with different weather-based cues in the Gulf & in Both subpopulations (@fig-covar B).
 In total, 65% of the posterior weight of strong SNP effects in the *mash* model of Gulf green-up fell on a covariance matrix constructed using the covariance of day length 14 days prior to the date of green-up.
@@
 In five of the six *mash* models of strong effects, the GxWeather covariance matrices captured a minority of the posterior weights of the strong effects (@fig-covar C,E); in these five models, the majority of this mass was on various canonical covariance matrices.
-These matrices included simple heterozygosity, with intermediate, positive covariances between all gardens, and single effect matrices with garden-specific effects (SI Appendix, Section S1).
+These matrices included simple heterozygosity, with intermediate, positive covariances between all gardens, and single effect matrices with garden-specific effects.
 
-We next characterized the pairwise patterns of effects where we were confident in the sign of the effect at both gardens.
-If the overall pattern of phenotypic expression differs between a pair of gardens, this difference may be comprised of SNP effects that differ in sign between these gardens (effects with rank-changing GxE between gardens), SNP effects that differ in magnitude (effects that are large in one garden or region and smaller in others), and SNP effects that are indistinguishable in the two gardens (@fig-effects A).
-We used the local false sign rate (lfsr), an analogue of the lfdr that establishes confidence in the effect sign, not the effect's difference from zero, to determine significance.
+### Frequency of rank-changing GxE in significant SNP effects
+
+We expected to observe common rank-changing GxE at the level of individual loci as this is a key theoretical prediction of local adaptation [@felsenstein1976theoretical; @levene1953genetic; @kawecki2004conceptual; @hedrick1986genetic].
+To determine the frequency of rank-changing GxE, we used the local false sign rate (lfsr), an analogue of the lfdr that establishes confidence in the effect sign, not the effect's difference from zero, to determine significance.
 We required lfsr significance (p \< 0.05) in both gardens to include effects.
 This means that our tests for a sign change between gardens carry an equal statistical burden to those for effects with the same sign.
+We separated kinds of effects at the level of individual loci into SNP effects that differ in sign between gardens (effects with rank-changing GxE between gardens), SNP effects that differ in magnitude (effects that are large in one garden or region and smaller in others), and SNP effects that are indistinguishable in two gardens (@fig-effects A).
 
-For green-up date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited a difference in effect sign, between pairs of Texas and North gardens (@fig-effects B; SI Appendix, Fig S1).
+For green-up date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited rank-changing GxE, or a difference in sign between pairs of Texas and North gardens (@fig-effects B; SI Appendix, Fig S1).
 78.7% of pairwise comparisons between North and Texas gardens had a difference in sign, while only 28.6% and 0.2% of North-North or Texas-Texas comparisons had a difference in sign, respectively (@fig-effects B).
@@
 
-For flowering date for the Gulf subpopulation, less than 2% of pairwise effects exhibited a difference in effect sign within or between regions (@fig-effects B).
+For flowering date for the Gulf subpopulation, less than 2% of pairwise effects exhibited rank-changing GxE, or a difference in effect sign within or between regions (@fig-effects B).
 More effects differed in magnitude between the Texas and North regions than within these regions (42.7% vs \<20%; @fig-effects B), while the majority of effects had no GxE.
@@
 All QTL for flowering overlapped one or more homologs from rice or *A.
-thaliana* with functionally validated roles in flowering (SI Appendix, Dataset 7).
+thaliana* with functionally validated roles in flowering (SI Appendix, Dataset 7); the QTL on Chr04K overlapped a gene recently functionally validated for flowering in switchgrass [@choi2023].
 All flowering and green-up QTL intervals contained at least one SNP significant in at least one *mash* run at a log10-transformed Bayes Factor \> 2, or in the 1% tail of significance, whichever was stricter (SI Appendix, Dataset 8).
@@
 Overall, there were five significant enrichments (p \< 0.05, hypergeometric test) of SNPs in the *mash* 1% tail in the QTL intervals.
-Thus, we were able to experimentally support some of our re-estimates of SNP effects with a QTL mapping experiment using a separate mapping population.
+Thus, we were able to experimentally support the genomic windows of some re-estimates of significant SNP effects with a QTL mapping experiment using a separate mapping population.
 
@@
 
-Our analysis of the timing of vegetative growth in the Gulf and in Both subpopulations revealed many alleles with sign changes in their effects between the Texas and North gardens (@fig-effects B).
-As phenological timings are major components of plant fitness, this result supports theoretical models that local adaptation should involve antagonistic pleiotropy at the level of individual loci [@kawecki2004conceptual; @felsenstein1976theoretical; @levene1953genetic; @hedrick1986genetic]; it expands the genetics of local adaptation research to more than two field sites and to a wider range of genetic variation by using effects from GWAS in common gardens located throughout the species' range to determine that rank-changing GxE for phenological timings is common in small genomic regions [@wadgymar2017identifying; @savolainen2013ecological; @lowry2019qtl].
+Our analysis of the timing of vegetative growth in the Gulf and in Both subpopulations revealed many alleles with rank-changing GxE, or sign changes in their effects between the Texas and North gardens (@fig-effects B).
+As phenological timings are major components of plant fitness, this result supports theoretical models that local adaptation should involve antagonistic pleiotropy at the level of individual loci [@kawecki2004conceptual; @felsenstein1976theoretical; @levene1953genetic; @hedrick1986genetic].
+In addition, it expands the genetics of local adaptation research to more than two field sites and to a wider range of genetic variation by re-estimating effects from GWAS to determine that rank-changing GxE for phenological timings is common in small genomic regions [@wadgymar2017identifying; @savolainen2013ecological; @lowry2019qtl].
 
@@
 Source data and code to replicate these analyses are available at: https://github.com/Alice-MacQueen/pvdiv-phenology-gxe.git.
-SNP data to replicate these analyses are available from the UT dataverse at https://doi.org/link.
+SNP data to replicate these analyses are available from the UT dataverse at <https://doi.org/10.18738/T8/A604BU>.
 
@@
 
-To confirm candidate genomic regions and patterns of allelic effects found in the diversity panel, we analyzed flowering in an outbred pseudo-F2 cross between four individuals, two Midwest and two Gulf individuals.
+To confirm candidate genomic regions found using *mash* on the diversity panel, we analyzed flowering in an outbred pseudo-F2 cross between four individuals, two Midwest and two Gulf individuals.
 The formation of this mapping population has been described previously [@milano2016genetic]; additional details on QTL mapping can be found in SI Appendix, Section S7.
@@
 Singular values were computed using singular value decomposition of the matrix of all SNPs, with iterative SNP pruning and removal of regions in long-range linkage disequilibrium [@privé2017].
-Third, to make the *mash* models computationally feasible, we extracted two subsets of SNP effect estimates and standard errors from our GWAS effect estimates: 1) effects from a subset of "strong" tests corresponding to stronger effects on our traits; 2) results from a random subset of all tests to correspond to an unbiased representation of all effects (SI Appendix, Section S3).
+Third, to make the *mash* models computationally feasible, we extracted two subsets of SNP effect estimates and standard errors from our GWAS effect estimates: (i) effects from a subset of "strong" tests corresponding to stronger effects on our traits; (ii) results from a random subset of all tests to correspond to an unbiased representation of all effects (SI Appendix, Section S3).
 We used the subset of random effects in our greedy *mash* algorithm (SI Appendix, Section S4) and used the subset of strong effects in our *mash* models with the optimal set of covariance matrices (SI Appendix, Section S5).

### 2024-06-11 manuscript/revision2/revision2.tex
revision2: Coherent M&M between Results, M&M, & SI Appendix

@@
 the natural range of the Gulf subpopulation, the onset of vegetative
-growth for Gulf genotypes (hereafter `green-up') occurred before
-Midwestern green-up, and the onset of reproductive growth for Gulf
-genotypes (hereafter `flowering') occurred after Midwestern flowering
+growth, or `green-up date', for Gulf genotypes occurred before
+Midwestern green-up, and the onset of reproductive growth, or `flowering
+date', for Gulf genotypes occurred after Midwestern flowering
 (Figure~\ref{fig-map} A). At the four northernmost common gardens
@@
 
-\subsubsection{\texorpdfstring{Inference of GxWeather effects for
-vegetative and reproductive timing using
-\emph{mash}}{Inference of GxWeather effects for vegetative and reproductive timing using mash}}\label{inference-of-gxweather-effects-for-vegetative-and-reproductive-timing-using-mash}
-
-We next looked for evidence of GxWeather by jointly re-estimating
-genetic effects of SNPs across all eight common gardens using the
-\emph{mash} model, which can flexibly capture many modes of covariance
-across gardens (Figure~\ref{fig-map} D; SI Appendix, Sections S1-S5,
-Datasets 1-6). We specified three qualitative categories of covariance
-structures across gardens: ``canonical'' covariance, with simple
-patterns of effect size covariance introduced in the initial \emph{mash}
-manuscript; ``data-driven'' covariances derived from common patterns of
-SNP effects observed in the data, and ``GxWeather'' covariance,
-estimated from the covariance of empirical weather patterns at each
-garden at specific times before the phenological event (Table 1; SI
-Appendix, Section S1). For example, the more similar that the amount of
-rainfall prior to flowering is for genotypes A and B between gardens in
-MI and MO, the nearer these covariances will be to one. If rainfall
-amounts are uncorrelated, covariance will be near zero; if there is rank
-changing (e.g.~high rainfall before flowering of A at MO but not MI, and
-high rainfall before flowering for B at MI but not MO), this covariance
-will be near negative one. We specified multiple covariance structures
-for the canonical, data-driven, and GxWeather categories, each of which
-represented a specific pattern of covariance in genetic effect size
-estimates that could be included in the model. The phenotypic
-correlations for the timing of vegetative growth had moderate negative
+\subsubsection{Inference of genome-wide patterns of GxE and GxWeather
+for vegetative and reproductive
+timing}\label{inference-of-genome-wide-patterns-of-gxe-and-gxweather-for-vegetative-and-reproductive-timing}
+
+We expected patterns of GxE and GxWeather to vary at the locus,
+subpopulation, and trait level; thus we were interested in modeling the
+different ways effects might covary across our eight common gardens. We
+jointly re-estimated genetic effects of a subset of SNPs across all
+eight common gardens using \emph{mash} models, which can flexibly
+capture many modes of covariance across gardens (Figure~\ref{fig-map} D;
+SI Appendix, Sections S1-S5, Datasets 1-6). To obtain genetic effects,
+we used subsets of effects from garden- and subpopulation GWAS (SI
+Appendix, Section S2-S3; Table S1). To obtain covariance structures, we
+specified both covariance structures introduced in the initial
+\emph{mash} manuscript, and GxWeather covariance matrices estimated from
+the covariance of empirical weather patterns at each garden at specific
+times before the phenological event (Table 1; SI Appendix, Section S1).
+Then, we used a greedy algorithm to select covariance matrices from each
+model's set that significantly improved the model likelihood (SI
+Appendix, Section S4; Table S2). Finally, we fit \emph{mash} models
+using the selected set of covariance matrices on a subset relatively
+unlinked (r\textsuperscript{2} \textless{} 0.2), `strong' genetic
+effects with low \emph{p}-values (SI Appendix, Section S5).
+
+The loadings of genetic effects onto the GxE and GxWeather covariance
+matrices included in each model provide information on genome-wide
+patterns of SNP-environment interaction. For example, the phenotypic
+correlations for the onset of vegetative growth had moderate negative
 correlations between the Texas and North gardens, particularly in the
@@
 partially or completely captured by the covariance structures specified
-in the \emph{mash} model.
-
-The GxWeather covariance structures allow hypothesis testing of specific
-weather variables as cues for the start of vegetative and reproductive
-growth. Say that a SNP in the gene \emph{Flowering Locus C} (FLC), a
-well-known flowering time regulator, controls flowering in a
-photoperiod-dependent manner. In that case, the joint estimate of
-effects for that SNP could have a high mixture proportion, or mass, on a
-covariance matrix created using a photoperiod-based environmental cue,
-such as day length at some interval prior to flowering. In our data, we
-would infer that the effect of that SNP on flowering was caused by a
-response to the environmental cue used to construct the GxWeather
-covariance structure with the largest mass. Overall, loadings of genetic
-effects on these GxWeather matrices in the model, along with the
-posterior effect size estimates, provide information on genome-wide
-patterns of SNP-environment interaction.
-
-We used a greedy algorithm to select covariance matrices from each
-model's set that significantly improved the model likelihood. Five
-GxWeather covariance structures were selected, zero to two per
-subpopulation: two for green-up date, and three for flowering date. Of
-these five, three had mass on them in \emph{mash} models of the strong
-effects (Figure~\ref{fig-covar} A-B). Two of these three matrices had
-negative covariances between sites between Texas and North gardens
-(Figure~\ref{fig-covar} A), while one had all positive or near-zero
-covariances. SNP-associated phenotypic effects covaried with different
-weather-based cues in the Gulf \& in Both subpopulations
+in the \emph{mash} model. In this case, the joint estimate of effects
+for many SNPs could have high mixture proportions, or mass, on a
+covariance matrix that captures the negative correlation between Texas
+and North gardens. In our data, daylength 14 days prior to green-up date
+has negative correlations between Texas and North gardens
+(Figure~\ref{fig-covar} A); if many SNP effects have mass on this
+GxWeather matrix, we would infer that the effects of those SNPs have
+that form of GxWeather covariation.
+
+Eight GxWeather covariance structures were selected by the greedy
+algorithm, one to two per subpopulation: five for green-up date, and
+three for flowering date (SI Appendix, Table S2). Of these eight, five
+had mass on them in \emph{mash} models of the strong effects
+(Figure~\ref{fig-covar} A-B). Three of these five matrices had negative
+covariances between sites between Texas and North gardens
+(Figure~\ref{fig-covar} A), while two had all positive or near-zero
+covariances.
+
+For green-up date, SNP-associated phenotypic effects covaried with
+different weather-based cues in the Gulf \& in Both subpopulations
 (Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
@@
 heterozygosity, with intermediate, positive covariances between all
-gardens, and single effect matrices with garden-specific effects (SI
-Appendix, Section S1).
-
-We next characterized the pairwise patterns of effects where we were
-confident in the sign of the effect at both gardens. If the overall
-pattern of phenotypic expression differs between a pair of gardens, this
-difference may be comprised of SNP effects that differ in sign between
-these gardens (effects with rank-changing GxE between gardens), SNP
-effects that differ in magnitude (effects that are large in one garden
-or region and smaller in others), and SNP effects that are
-indistinguishable in the two gardens (Figure~\ref{fig-effects} A). We
+gardens, and single effect matrices with garden-specific effects.
+
+\subsubsection{Frequency of rank-changing GxE in significant SNP
+effects}\label{frequency-of-rank-changing-gxe-in-significant-snp-effects}
+
+We expected to observe common rank-changing GxE at the level of
+individual loci as this is a key theoretical prediction of local
+adaptation (23--26). To determine the frequency of rank-changing GxE, we
 used the local false sign rate (lfsr), an analogue of the lfdr that
@@
 our tests for a sign change between gardens carry an equal statistical
-burden to those for effects with the same sign.
+burden to those for effects with the same sign. We separated kinds of
+effects at the level of individual loci into SNP effects that differ in
+sign between gardens (effects with rank-changing GxE between gardens),
+SNP effects that differ in magnitude (effects that are large in one
+garden or region and smaller in others), and SNP effects that are
+indistinguishable in two gardens (Figure~\ref{fig-effects} A).
 
 For green-up date for the Gulf subpopulation, hundreds to thousands of
-pairwise effects exhibited a difference in effect sign, between pairs of
-Texas and North gardens (Figure~\ref{fig-effects} B; SI Appendix, Fig
-S1). 78.7\% of pairwise comparisons between North and Texas gardens had
-a difference in sign, while only 28.6\% and 0.2\% of North-North or
-Texas-Texas comparisons had a difference in sign, respectively
-(Figure~\ref{fig-effects} B). The majority of pairwise effects for
-greenup for the Midwest (\textgreater55\%) and Both (\textgreater85\%)
-subpopulations were the same sign, and effects most often differed in
-magnitude between gardens within and between the regions
-(Figure~\ref{fig-effects} B).
+pairwise effects exhibited rank-changing GxE, or a difference in sign
+between pairs of Texas and North gardens (Figure~\ref{fig-effects} B; SI
+Appendix, Fig S1). 78.7\% of pairwise comparisons between North and
+Texas gardens had a difference in sign, while only 28.6\% and 0.2\% of
+North-North or Texas-Texas comparisons had a difference in sign,
+respectively (Figure~\ref{fig-effects} B). The majority of pairwise
+effects for greenup for the Midwest (\textgreater55\%) and Both
+(\textgreater85\%) subpopulations were the same sign, and effects most
+often differed in magnitude between gardens within and between the
+regions (Figure~\ref{fig-effects} B).
 
 For flowering date for the Gulf subpopulation, less than 2\% of pairwise
-effects exhibited a difference in effect sign within or between regions
-(Figure~\ref{fig-effects} B). More effects differed in magnitude between
-the Texas and North regions than within these regions (42.7\% vs
-\textless20\%; Figure~\ref{fig-effects} B), while the majority of
-effects had no GxE. The Midwest population had relatively few
-significant effects for flowering, but a large proportion of these
-differed in sign between Texas and North regions (42.7\%) or within the
-North region (65.4\%). Finally, in Both subpopulations, less than 20\%
-of pairwise effects differed in sign (Figure~\ref{fig-effects} B). Most
-differences in sign were between TX1, the southernmost garden, and all
-other gardens (SI Appendix, Figure S1). Similarly, more effects that
-differed in magnitude included gardens in the Texas region
+effects exhibited rank-changing GxE, or a difference in effect sign
+within or between regions (Figure~\ref{fig-effects} B). More effects
+differed in magnitude between the Texas and North regions than within
+these regions (42.7\% vs \textless20\%; Figure~\ref{fig-effects} B),
+while the majority of effects had no GxE. The Midwest population had
+relatively few significant effects for flowering, but a large proportion
+of these differed in sign between Texas and North regions (42.7\%) or
+within the North region (65.4\%). Finally, in Both subpopulations, less
+than 20\% of pairwise effects differed in sign (Figure~\ref{fig-effects}
+B). Most differences in sign were between TX1, the southernmost garden,
+and all other gardens (SI Appendix, Figure S1). Similarly, more effects
+that differed in magnitude included gardens in the Texas region
 (52.3-55.9\%), and most effect pairs in the North region were not
@@
 one or more homologs from rice or \emph{A. thaliana} with functionally
-validated roles in flowering (SI Appendix, Dataset 7). All flowering and
-green-up QTL intervals contained at least one SNP significant in at
-least one \emph{mash} run at a log10-transformed Bayes Factor
-\textgreater{} 2, or in the 1\% tail of significance, whichever was
-stricter (SI Appendix, Dataset 8). We also looked for enrichments of
-\emph{mash} SNPs in the 1\% tail of significance (the `\emph{mash} 1\%
-tail') within each QTL interval. At the 5\% level, three QTL had
-enrichments of SNPs in the \emph{mash} 1\% tail. Overall, there were
-five significant enrichments (p \textless{} 0.05, hypergeometric test)
-of SNPs in the \emph{mash} 1\% tail in the QTL intervals. Thus, we were
-able to experimentally support some of our re-estimates of SNP effects
-with a QTL mapping experiment using a separate mapping population.
+validated roles in flowering (SI Appendix, Dataset 7); the QTL on Chr04K
+overlapped a gene recently functionally validated for flowering in
+switchgrass (38). All flowering and green-up QTL intervals contained at
+least one SNP significant in at least one \emph{mash} run at a
+log10-transformed Bayes Factor \textgreater{} 2, or in the 1\% tail of
+significance, whichever was stricter (SI Appendix, Dataset 8). We also
+looked for enrichments of \emph{mash} SNPs in the 1\% tail of
+significance (the `\emph{mash} 1\% tail') within each QTL interval. At
+the 5\% level, three QTL had enrichments of SNPs in the \emph{mash} 1\%
+tail. Overall, there were five significant enrichments (p \textless{}
+0.05, hypergeometric test) of SNPs in the \emph{mash} 1\% tail in the
+QTL intervals. Thus, we were able to experimentally support the genomic
+windows of some re-estimates of significant SNP effects with a QTL
+mapping experiment using a separate mapping population.
 
@@
 Our analysis of the timing of vegetative growth in the Gulf and in Both
-subpopulations revealed many alleles with sign changes in their effects
-between the Texas and North gardens (Figure~\ref{fig-effects} B). As
-phenological timings are major components of plant fitness, this result
-supports theoretical models that local adaptation should involve
-antagonistic pleiotropy at the level of individual loci (23--26); it
-expands the genetics of local adaptation research to more than two field
-sites and to a wider range of genetic variation by using effects from
-GWAS in common gardens located throughout the species' range to
+subpopulations revealed many alleles with rank-changing GxE, or sign
+changes in their effects between the Texas and North gardens
+(Figure~\ref{fig-effects} B). As phenological timings are major
+components of plant fitness, this result supports theoretical models
+that local adaptation should involve antagonistic pleiotropy at the
+level of individual loci (23--26). In addition, it expands the genetics
+of local adaptation research to more than two field sites and to a wider
+range of genetic variation by re-estimating effects from GWAS to
 determine that rank-changing GxE for phenological timings is common in
-small genomic regions (12, 38, 39).
+small genomic regions (12, 39, 40).
 
@@
 replicate these analyses are available from the UT dataverse at
-https://doi.org/link.
+\url{https://doi.org/10.18738/T8/A604BU}.
 
@@
 populations of switchgrass, a diversity panel and a pseudo-F2 cross,
-planted at eight common garden locations (35, 39, 40). We scored the
+planted at eight common garden locations (35, 40, 41). We scored the
 onset of vegetative growth, or green-up date, as the day of the year
@@
 different subpopulations had different strengths of connection between
-our phenotypes and genotypes (41), we conducted three sets of genetic
+our phenotypes and genotypes (42), we conducted three sets of genetic
 analyses: on Gulf and Midwest genotypes separately, and on both
@@
 
-To confirm candidate genomic regions and patterns of allelic effects
-found in the diversity panel, we analyzed flowering in an outbred
-pseudo-F2 cross between four individuals, two Midwest and two Gulf
-individuals. The formation of this mapping population has been described
-previously (35); additional details on QTL mapping can be found in SI
-Appendix, Section S7. To be directly comparable to the diversity panel
-data, only 2019 phenology data from the pseudo-F2 cross from the same
-eight common garden sites were used.
+To confirm candidate genomic regions found using \emph{mash} on the
+diversity panel, we analyzed flowering in an outbred pseudo-F2 cross
+between four individuals, two Midwest and two Gulf individuals. The
+formation of this mapping population has been described previously (35);
+additional details on QTL mapping can be found in SI Appendix, Section
+S7. To be directly comparable to the diversity panel data, only 2019
+phenology data from the pseudo-F2 cross from the same eight common
+garden sites were used.
 
@@
 pruning and removal of regions in long-range linkage disequilibrium
-(42). Third, to make the \emph{mash} models computationally feasible, we
+(43). Third, to make the \emph{mash} models computationally feasible, we
 extracted two subsets of SNP effect estimates and standard errors from
-our GWAS effect estimates: 1) effects from a subset of ``strong'' tests
-corresponding to stronger effects on our traits; 2) results from a
+our GWAS effect estimates: (i) effects from a subset of ``strong'' tests
+corresponding to stronger effects on our traits; (ii) results from a
 random subset of all tests to correspond to an unbiased representation
@@
 hypothesis that an effect j is 0, or that there is a significant effect.
-Previous studies of antagonistic pleiotropy (e.g. (39)) have used the
+Previous studies of antagonistic pleiotropy (e.g. (40)) have used the
 lfdr or equivalent statistical tests to detect antagonistic pleiotropy.
@@
 
+\bibitem[\citeproctext]{ref-choi2023}
+\CSLLeftMargin{38. }%
+\CSLRightInline{S. Choi, \emph{et al.},
+\href{https://doi.org/10.1093/jxb/erad255}{A single amino acid change
+led to structural and functional differentiation of {\emph{PvHd1}} to
+control flowering in switchgrass}. \emph{Journal of Experimental Botany}
+\textbf{74}, 5532--5546 (2023).}
+
 \bibitem[\citeproctext]{ref-savolainen2013ecological}
-\CSLLeftMargin{38. }%
+\CSLLeftMargin{39. }%
 \CSLRightInline{O. Savolainen, M. Lascoux, J. Merilä, Ecological
@@
 \bibitem[\citeproctext]{ref-lowry2019qtl}
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{40. }%
 \CSLRightInline{D. B. Lowry, \emph{et al.}, QTL\(\times\) environment
@@
 \bibitem[\citeproctext]{ref-lovell2021}
-\CSLLeftMargin{40. }%
+\CSLLeftMargin{41. }%
 \CSLRightInline{J. T. Lovell, \emph{et al.},
@@
 \bibitem[\citeproctext]{ref-kortefarlow2013}
-\CSLLeftMargin{41. }%
+\CSLLeftMargin{42. }%
 \CSLRightInline{A. Korte, A. Farlow, The advantages and limitations of
@@
 \bibitem[\citeproctext]{ref-privuxe92017}
-\CSLLeftMargin{42. }%
+\CSLLeftMargin{43. }%
 \CSLRightInline{F. Privé, H. Aschard, M. G. B. Blum,

### 2024-06-12 R/Analysis_v1.2_mash_using_greedy_covar.qmd
revision2: generate revised figures 1-4 and SI figures 1-4

@@
 
+NB: Figure 1 is in Plots_v0.3_manuscript_figures.Rmd
+
 ## Load mash objects
@@
     geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
-    scale_fill_gradientn(colors = c("#30123BFF", "#4777EFFF", "#1BD0D5FF",
+    scale_fill_gradientn(colors = c("#30123BFF", "#4777EFFF",
+                                    "#1BD0D5FF",
                                   "white", "#FE9B2DFF", "#DB3A07FF",
@@
 
+Build Figure 2 with covariance matrices as first two rows over covar barplots
+
+```{r}
+
+ggdraw() +
+  theme_oeco + 
+  draw_plot(pa01, x = 0, y = 0.66, width = 0.1, height = 0.18) +
+  draw_plot(pa02, x = 0.2, y = 0.66, width = 0.1, height = 0.18) +
+  #draw_plot(pa03, x = -0.02, y = 0.6, width = 0.1, height = 0.18) +
+  #draw_plot(pa04, x = 0.1, y = 0.6, width = 0.1, height = 0.18) +
+  draw_plot(pa05, x = 0.4, y = 0.66, width = 0.1, height = 0.18) +
+  draw_plot(pa06, x = 0.6, y = 0.66, width = 0.1, height = 0.18) +
+  draw_plot(pa07, x = 0.8, y = 0.66, width = 0.1, height = 0.18) +
+  #draw_plot(pa08, x = 0.2, y = 0, width = 0.1, height = 0.18) +
+  draw_plot(pa09, x = 0, y = 0.83, width = 0.1, height = 0.18) +
+  draw_plot(pa10, x = 0.2, y = 0.83, width = 0.1, height = 0.18) +
+  draw_plot(pa11, x = 0.4, y = 0.83, width = 0.1, height = 0.18) +
+  draw_plot(pa12, x = 0.6, y = 0.83, width = 0.1, height = 0.18) +
+  draw_plot(pa13, x = 0.8, y = 0.83, width = 0.1, height = 0.18) +
+  draw_plot(f2a, x = 0, y = 0.33, width = 0.66, height = 0.38) +
+  draw_plot(legend, x = 0.1, y = 0.77, width = 0.1, height = 0.2) +
+  draw_plot(f2b, x = 0.66, y = 0.33, width = 0.33, height = 0.38) +
+  draw_plot(f2c, x = 0, y = 0, width = 0.66, height = 0.33) +
+  draw_plot(f2d, x = 0.66, y = 0, width = 0.33, height = 0.33)
+```
+
 ```{r}
@@
 
-save_plot(filename = "../manuscript/revision2/Figure_2_recolor_covariance_matrices_and_posterior_weights.svg", plot = last_plot(), base_height = 5.5, base_width = 8)
+save_plot(filename = "../manuscript/revision2/Figure_2_recolor_resize_covariance_matrices_and_posterior_weights.svg", plot = last_plot(), base_height = 7.5, base_width = 6.5)
+
 ```

### 2024-06-12 R/Plots_v0.3_manuscript_figures.Rmd
revision2: generate revised figures 1-4 and SI figures 1-4

@@
 usa <- map_data("usa")
-source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+# source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+```
+
+```{r}
+theme_oeco <- theme_classic() +
+  theme(axis.title = element_text(size = 10), 
+        axis.text = element_text(size = 10), 
+        axis.line.x = element_line(linewidth = 0.35, colour = 'grey50'),
+        axis.line.y = element_line(linewidth = 0.35, colour = 'grey50'),
+        axis.ticks = element_line(linewidth = 0.25, colour = 'grey50'), 
+        legend.justification = c(1, 0.75), 
+        legend.position = c(1, 0.9), 
+        legend.key.size = unit(0.35, 'cm'),
+        legend.title = element_blank(), 
+        legend.text = element_text(size = 9),
+        legend.background = element_blank(),
+        plot.subtitle = element_text(size = 10, vjust = 0), #plot.margin = unit(c(0.35, 0, 0.25, 0), 'cm'),
+        strip.background = element_blank(), 
+        strip.text = element_text(hjust = 0.5, size = 10 ,vjust = 0), 
+        strip.placement = 'outside', panel.spacing.x = unit(-0.5, 'cm'))
+theme_set(theme_oeco)
 ```
@@
 ```{r}
-workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
-metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
-sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
-subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
-phe_gr_df <- read_rds(file.path(workingdir, "data",
+# workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+metadata <- read_rds(here("data", "metadata.rds"))
+sites <- read_rds(here("data", "sites.rds"))
+subpops <- read_rds(here("data", "subpop_color_coding.rds"))
+phe_gr_df <- read_rds(here("data",
                                 "Weather_related_phe_for_asreml_h2_GR50.rds"))
-phe_fl_df <- read_rds(file.path(workingdir, "data",
+phe_fl_df <- read_rds(here("data",
                                 "Weather_related_phe_for_asreml_h2_FL50.rds"))
-phenotypes <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
-h2gr <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "h2_df_GR50_related_functions_of_weather_v1.csv"))
-h2fl <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "h2_df_FL50_related_functions_of_weather_v1.csv"))
-h2gr_gxe <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "gxe_and_h2_df_GR50_related_functions_of_weather_v2.csv"))
-h2fl_gxe <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "gxe_and_h2_df_FL50_related_functions_of_weather_v1.csv"))
-
-h2gr <- h2gr %>%
-  full_join(h2gr_gxe)
+phenotypes <- readRDS(here("data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds"))
+h2gr <- read_csv(here("analysis", "heritability", "kinship_subpop", "h2_df_GR50_related_functions_of_weather_v1.csv"))
+h2fl <- read_csv(here("analysis", "heritability", "kinship_subpop", "h2_df_FL50_related_functions_of_weather_v1.csv"))
+h2gr_gxe <- read_csv(here("analysis", "heritability", "kinship_subpop", "gxe_and_h2_df_GR50_related_functions_of_weather_v2.csv"))
+h2fl_gxe <- read_csv(here("analysis", "heritability", "kinship_subpop", "gxe_and_h2_df_FL50_related_functions_of_weather_v1.csv"))
+
+#h2gr <- h2gr %>%
+#  full_join(h2gr_gxe)
 h2gr <- h2gr_gxe
@@
   #scale_alpha_manual(values = c(0.7, 0.7, 0.7, 0.7), guide = FALSE)
-save_plot(filename = "../manuscript/plots/Map_subpops_and_Ecotypes.png", plot = last_plot(), base_height = 3)
+save_plot(filename = here("manuscript", "revision2", "intermediate_files", "Map_subpops_and_Ecotypes.png"), plot = last_plot(), base_height = 3)
 ?geom_point
@@
   ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
-  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
-                                      "white", "#27AD81FF", "#5DC863FF",
-                                      "#FDE725FF"), limits = c(-1,1)) +
+  scale_fill_gradientn(colors = c("#30123BFF", "#4777EFFF", "#1BD0D5FF",
+                                  "white", "#FE9B2DFF", "#DB3A07FF",
+                                  "#7A0403FF"), limits = c(-1,1)) +
   theme(legend.position = "none")
@@
   ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
-  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
-                                      "white", "#27AD81FF", "#5DC863FF",
-                                      "#FDE725FF"), limits = c(-1,1)) +
+  scale_fill_gradientn(colors = c("#30123BFF", "#4777EFFF", "#1BD0D5FF",
+                                  "white", "#FE9B2DFF", "#DB3A07FF",
+                                  "#7A0403FF"), limits = c(-1,1)) +
   theme(legend.position = c(0.4, 0.85))
@@
   ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
-  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
-                                      "white", "#27AD81FF", "#5DC863FF",
-                                      "#FDE725FF"), limits = c(-1,1)) +
+  scale_fill_gradientn(colors = c("#30123BFF", "#4777EFFF", "#1BD0D5FF",
+                                  "white", "#FE9B2DFF", "#DB3A07FF",
+                                  "#7A0403FF"), limits = c(-1,1)) +
   theme(legend.position = "none")
 
-pheno_corr <- plot_grid(pana, panb, panc, pand, pane, panf, nrow = 2)
-save_plot(filename = "../manuscript/plots/Phenotypic_correlations_GR50_FL50_2.svg", plot = pheno_corr, base_width = 6.5)
+pheno_corr <- plot_grid(panc, panf, panb, pane, pana, pand, nrow = 3)
+save_plot(filename = here("manuscript", "revision2", "intermediate_files", "Phenotypic_correlations_GR50_FL50_2.svg"), plot = pheno_corr, base_width = 3.5, base_height = 4.5)
 ```

### 2024-06-12 R/SI_Appendix_Datasets.Rmd
revision2: generate revised figures 1-4 and SI figures 1-4

@@
 
-Edited 2024-05-17 for revision2.
+Edited 2024-05-17+ for revision2.
 
@@
 
-Figure 3 is North x Texas GxE effects. Figure S1 is garden x garden GxE effects of individual loci.
+h^2 analyses that was originally figure 1C. Try moving to the supplement.
+
+## Figure S2 & S3
+
+Figure 3 is North x Texas GxE effects. Figure S2 is garden x garden GxE effects of individual loci for green-up; Figure S3 is for flowering date.
 
@@
 
-## Figure S2
+## Figure S4
 
@@
 
-We conducted quantitative trait loci (QTL) mapping of flowering as functions of five environmental cues that we also used as covariance matrices in mash, and identified eight QTL for flowering date, eight QTL for flowering as a function of day length change two days prior, one QTL for the start of vegetative growth, and two QTL for vegetative growth as a function of daylength change one day prior, all of which showed QTL by environment interactions (SI Appendix, Figure S2).
+We conducted quantitative trait loci (QTL) mapping of flowering as functions of five environmental cues that we also used as covariance matrices in mash, and identified eight QTL for flowering date, eight QTL for flowering as a function of day length change two days prior, one QTL for the start of vegetative growth, and two QTL for vegetative growth as a function of daylength change one day prior, all of which showed QTL by environment interactions (SI Appendix, Figure S4).
+
+```{r}
+library(qtl)
+library(qtl2)
+library(qtl2convert)
+```
+
+#
+
+```{r prep QTL_Eff df}
+###QTL Effect plot
+###load the original cross and reduce the markers from 4700 to 1185
+#load('../phenology-fourway/5-QTLmappingForAlice/sexAveraged.cross.rda', verbose = T)
+workingdir <- here("phenology-fourway/5-QTLmappingForAlice/")
+load(file.path(workingdir, 'cross_reduced_marker_FL50_phenotypes_Alice_7_SITES.RData'))
+map  =  lapply(pull.map(cross), function(x) x[1,])
+# reduced_markers  =  unlist(
+#   lapply(
+#     reduce_markers(
+#       lapply(pull.map(cross), function(x)
+#         x[1,]), min_distance = 1), names))
+#
+# mars2drop  =  markernames(cross)[!markernames(cross) %in% reduced_markers]
+# cross  =  drop.markers(cross, mars2drop)
+cross = calc.genoprob(cross, step=1, error.prob = 1e-3,map.function = "kosambi",stepwidth = "max")
+######### Converting probalibity from qtl1 to qtl2
+probs  =  probs_qtl_to_qtl2(cross)$probs
+##calculate the kinship matrix for single site
+kinship_loco_Single = calc_kinship(probs,use_allele_probs = F,type = "loco")
+
+###extract the genotype ID from the original cross file, so we can align the phenology file with genotype ID
+ID <- data.frame(ID=cross$pheno$id)
+ID <- ID %>% 
+  mutate_at(vars('ID'), list(as.character)) %>%
+  mutate_at(vars('ID'), list(as.numeric))
+
+######get the phenotype at each site
+phenology = read.csv('../phenology-fourway/5-QTLmappingForAlice/FL50_phenotypes_Alice.csv', header = TRUE)
+colnames(phenology)[1:2] = c('Covariate','ID')
+str(phenology) ###make sure your phenology traits are numeric, if not, change them into numeric
+phenos <- phenology %>% 
+  mutate_at(vars('Covariate'), list(as.character)) %>%
+  mutate(ID = as.numeric(ID)) %>% 
+  group_by(Covariate,ID) %>%
+  summarize_all(mean, na.rm=T) %>% 
+  ungroup() %>% 
+  as.data.frame()
+```
+
+```{r setup QTL effects for all traits}
+###QTL effect for each trait
+EffectMatrix = vector('list', length=0)
+t=0
+for (envi in unique(phenos$Covariate)){
+    t=t+1
+    pheno = phenos %>% 
+      filter(Covariate==envi) %>% 
+      dplyr::select(-Covariate) %>%
+      select(ID, FL50) %>% 
+      right_join(ID) %>%  column_to_rownames('ID') %>% as.matrix()
+
+    ##ran effect scan for each chromosome
+    eff_out = vector('list',length = 0)
+    se_out = vector('list',length = 0)
+    add_contrasts=cbind(mu=c(1,1,1,1), add1=c(-1, 1, -1, 1),add2=c(-1, -1, 1, 1), dom=c(-1, 1, 1, -1))
+
+    z = 0
+    # chr = "1K"
+    for (chr in chrnames(cross)){
+      print(c(chr, envi,t))
+      z= z+1
+      eff = scan1coef(probs[,chr],pheno,
+                      kinship = kinship_loco_Single[chr], 
+                      se=T, contrasts = add_contrasts)
+      se = attr(eff, 'SE')
+      eff_out[[z]] =  eff %>% as.data.frame() %>% rownames_to_column('MARKER')
+      se_out[[z]] =  se %>% as.data.frame() %>% rownames_to_column('MARKER')
+    }
+    eff_out = bind_rows(eff_out)
+    se_out = bind_rows(se_out)
+
+    se_out2 = se_out %>% dplyr::select(-mu) %>% gather('CROSS','SE',-MARKER)
+    eff_out2 = eff_out %>% dplyr::select(-mu) %>% gather('CROSS','EFF',-MARKER) %>% mutate(SE=se_out2$SE, Covariate=envi)
+
+    EffectMatrix[[t]] = eff_out2
+  }
+
+EffectMatrix2 = bind_rows(EffectMatrix)
+
+###unlist nmap and merge with Effect file
+nmap = data.frame(pos = unlist(map))
+nnmap = nmap %>% rownames_to_column('MARKER') %>% rowwise() %>%
+  mutate(chr = str_sub(MARKER, 1,2), MARKER=str_sub(MARKER, 4,nchar(MARKER)))%>%
+  ungroup() %>% mutate(pos=round(pos,4))
+
+flank_marker_full = read.csv('../phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv')
+flank_marker_full = flank_marker_full %>% mutate_at(vars(chr),list(as.character))
+
+for (i in 1:nrow(flank_marker_full)){
+  tmp = nnmap %>% filter(chr==flank_marker_full$chr[i] & pos==round(flank_marker_full$pos[i],4))
+  flank_marker_full$MARKER[i] = tmp$MARKER
+}
+
+QTL_full_model = flank_marker_full %>% dplyr::rename(trait=lodcolumn)
+
+```
+
+```{r ggplots of QTL effects}
+QTLdf <- QTL_full_model %>% left_join(EffectMatrix2)%>%
+  mutate(CROSS = str_replace(CROSS,'add2','C x D')) %>%
+  mutate(CROSS = str_replace(CROSS,'add1','A x B'),
+         Cross = case_when(CROSS == "A x B" ~ "AP13 x DAC",
+                           CROSS == "C x D" ~ "VS16 x WBC",
+                           CROSS == "dom" ~ "Dominance",
+                           TRUE ~ "Other"),
+         Trait = case_when(trait == "dyln_change_sec" ~ "Daylength change (s), on Green-up Date",
+                           trait == "dyln_change_sec_2d_prior" ~ "Daylength change (s), 2d prior to Flowering Date",
+                           trait == "FL50" ~ "Flowering Date",
+                           trait == "GR50" ~ "Green-up Date",
+                           TRUE ~ "Other")
+         ) %>%
+  separate(MARKER, into = c("Chr", "Pos"), sep = "_", convert = TRUE, 
+           remove = FALSE) %>%
+  mutate(Pos_plot = paste0(Chr, "\n", round(Pos, 1), " Mb"))
+
+QTLdf$Covariate = factor(QTLdf$Covariate, levels = c("TX1", "TX2", "OK", "MO", 
+                                                   "NE", "MI", "SD"))
+
+QTLdf %>% 
+  #filter(Cross != "Dominance") %>% # trait == 'FL50' &
+  ggplot(aes(x = Covariate, y = EFF, group = Cross)) +
+  geom_point(size = 2) + facet_grid(Cross + trait ~ Pos_plot) +
+  geom_hline(yintercept = 0, linetype=2) + 
+  geom_errorbar(aes(ymin=EFF-SE, ymax=EFF+SE), size=0.6, width=0.5) +
+  xlab("") +
+  ylab('QTL Effect') + 
+  coord_flip() + 
+  theme_bw()+
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        text=element_text(size=12, face="bold"),
+        axis.title=element_text(size=12,face="bold"),
+        axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1))
+# shared QTL for more than one trait differ in LOD score, but not meaningfully in effect patterns across gardens. Useful to know.
+
+QTLplot <- QTLdf %>%
+  arrange(desc(lod)) %>%
+  group_by(Chr, Pos, Covariate, Cross) %>%
+  slice(1) %>% 
+  filter(Cross != "Dominance") # trait == 'FL50' &
+
+QTLplot %>%
+  ggplot(aes(x = Covariate, y = EFF, group = Cross)) +
+  geom_col(size = 1, aes(fill = Trait)) + 
+  facet_grid(Cross ~ Pos_plot) +
+  geom_hline(yintercept = 0, linetype=2) + 
+  geom_errorbar(aes(ymin=EFF-SE, ymax=EFF+SE, color = Trait), size=0.6, width=0.5) +
+  xlab("") + scale_y_continuous(limits = c(-2.5, 4), breaks = c(-2, 0, 2)) +
+  ylab('QTL Effect') + 
+  coord_flip() + 
+  switchgrassGWAS::theme_oeco + 
+  scale_color_manual(values = c("#CC0066", "#FFCC00", "darkgrey", "darkgreen")) +
+  scale_fill_manual(values = c("#CC0066", "#FFCC00", "darkgrey", "darkgreen")) +
+  theme(legend.position = "top",
+        panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        text = element_text(size = 10, face="bold"),
+        strip.text = element_text(size = 10, face="bold"),
+        axis.title = element_text(size = 10,face="bold"),
+        axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1),
+        )# +
+    #scale_y_continuous(expand = c(0.2, 0.2)) 
+save_plot(filename = here("manuscript/revision2/intermediate_files/ Figure_S4_QTL_effects_by_trait.svg"), base_height = 4.5, base_width = 9, plot = last_plot())
+
+QTLdf %>%
+  filter(trait == 'dyln_change_sec_2d_prior') 
+QTLdf %>% 
+  filter(trait == 'FL50')
+QTLdf %>% 
+  filter(trait == 'dyln_change_sec')
+
+unique(QTLdf$trait)
+```

### 2024-06-12 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
revision2: generate revised figures 1-4 and SI figures 1-4

@@
 
-\noindent This PNAS template for Supporting Information (SI) may be used to organize your supporting material.
-The template is intended to provide a clearly organized PDF file that will ensure readers can easily navigate to sections or specific figures and tables.
-Movie files or large datasets can be presented as separate files.
-Further information is available in the [PNAS Author Center](https://www.pnas.org/authors/submitting-your-manuscript).
-
-## Using the template {.unnumbered}
-
-Specify the title, author list, and corresponding authors with the same YAML specification as in the [corresponding PNAS Quarto template](https://github.com/christopherkenny/pnas).
-The cover page will be automatically generated with the relevant description of the SI.
-
-Figures should be placed on separate pages with legends set immediately below each figure.
-Table titles should be set immediately above each table.
-Note that tables extending beyond the width of the page can be included in the PDF or provided as separate dataset files.
-Oversized/nonstandard page sizes are accepted as part of your SI Appendix file.
-
-## Submitting SI {.unnumbered}
-
-Save your completed SI file as a PDF for submission. Further submission instructions are available [here](https://www.pnas.org/page/authors/submission#preparation).
-
 # Supplementary Methods
@@
 Third, we chose the lowest value of $k$ that made $\lambda_{GC}$ closest to 1, and ran GWAS using a linear regression model with this $k$ for population structure correction.
+
+At this stage, we reviewed the univariate GWAS results, both Manhattan and QQ-plots, for population structure issues that were not resolved using the $k$ that minimized $\lambda_{GC}$.
+These issues were apparent in the data in two ways: (i) univariate GWAS with many significant SNPs with extremely similar *p*-values, usually indicative of genetic effects that, by chance or, more likely, because of confounding effects of population structure, were identical across many SNPs and thus not resolvable to a single genomic location; (ii) univariate GWAS with many effect estimates and standard errors of zero, indicating an inability to estimate effects given the constraints of small population size & population structure.
+We only noticed these issues in our smallest subpopulation, the Midwest subpopulation, in the green-up date trait at the SD garden, and in the flowering date trait at the OK and NE gardens.
+Due to these issues, we dropped the SD garden from subsequent green-up analyses for the Midwest subpopulation, and we dropped the OK and NE gardens from subsequent Midwest flowering date analyses.
 
@@
 The statistical significance of the genome scan was established by performing a stratified (i.e., stratifying on common garden) permutation test (n=1000) using ‘scan1perm’ function.
-The estimated QTL effect was obtained using ‘scan1coef’ function in R/qtl2.
-
-Figures should be cross-referenced like @fig-qtl-by-env. Each figure should be on its own page. You can control this by placing a pagebreak shortcode, with `{{< pagebreak >}}`.
-
-This can also be done with LaTeX's `\newpage` command. You may find `\FloatBarrier` useful if your figures or tables are running off to the wrong pages.
+The estimated QTL effects (@fig-qtl-by-env) were obtained using ‘scan1coef’ function in R/qtl2.
 
@@
 
-![Figure](figs/frog.eps){#fig-garden-by-garden-gxe}
+![Narrow sense heritability of green-up date and flowering date within single common gardens (purple) and across all eight common gardens (green), within and between two genetic subpopulations.](images/Figure_S1_h2.png){#fig-garden-by-garden-gxe}
 
@@
 
-![Figure](figs/frog.eps){#fig-qtl-by-env}
+![Types of GxE present between of pairs of jointly re-estimated effects in eight common gardens, for effects with lfsr \< 0.05 at both gardens in each pair.
+Effect patterns: Sign: Effects that differ in sign at these pairs of gardens (p \< 0.05, lfsr).
+Magnitude: Effects identical in sign (p \< 0.05, lfsr) that differ in magnitude by a factor of \>0.4.
+Not Distinguishable: Effects not distinguishable by magnitude nor sign of the effect, with no measurable GxE.
+b) The number of effects with each GxE type for the onset of vegetative growth (green-up date) within and between two genetic subpopulations.](images/S2_Greenup.png)
+
+\FloatBarrier <!-- always leave a blank line between this and figures --->
+
+{{< pagebreak >}}
+
+![Types of GxE present between of pairs of jointly re-estimated effects in eight common gardens, for effects with lfsr \< 0.05 at both gardens in each pair.
+Effect patterns: Sign: Effects that differ in sign at these pairs of gardens (p \< 0.05, lfsr).
+Magnitude: Effects identical in sign (p \< 0.05, lfsr) that differ in magnitude by a factor of \>0.4.
+Not Distinguishable: Effects not distinguishable by magnitude nor sign of the effect, with no measurable GxE.
+b) The number of effects with each GxE type for the onset of reproductive growth (green-up date) within and between two genetic subpopulations.](images/S3_Flowering.png)
+
+\FloatBarrier <!-- always leave a blank line between this and figures --->
+
+{{< pagebreak >}}
+
+![All QTL identified in the outbred mapping population displayed GxE, in that all QTL had effects that differed in at least two gardens.
+Gardens are on the y-axis in latitudinal order; the effects of the significant QTL in Figure 4 are displayed as differences between allelic effects for the two sets of grandparent lines in the outbred mapping panel.
+](images/Figure_S4_QTL_effects_by_trait.png){#fig-qtl-by-env}
 
@@
 
+{{< pagebreak >}}
+
 # References

### 2024-06-12 manuscript/revision2/revision2.qmd
revision2: generate revised figures 1-4 and SI figures 1-4

@@
 
-Narrow-sense heritabilities (h^2^) suggested that rank-changing GxE for these phenotypes was present across the common gardens (@fig-map C).
+Narrow-sense heritabilities (h^2^) suggested that rank-changing GxE for these phenotypes was present across the common gardens (SI Appendix, Figure S1).
 h^2^ were typically high at individual gardens: 59% on average for green-up date, and 87% for flowering date.
 However, h^2^ were variable across gardens, and green-up dates were uncorrelated (r^2^ \< 0.2) or negatively correlated between pairs of gardens (@fig-map B).
-These negative and small correlations undoubtedly contributed to the low h^2^ values for green-up and flowering date when estimated jointly at all eight gardens: h^2^ was 0.8% for green-up and 23.2% for flowering date.
+These negative and small correlations undoubtedly contributed to the low h^2^ values for green-up and flowering date when estimated jointly at all eight gardens: h^2^ was 0.8% for green-up and 23.2% for flowering date (SI Appendix, Figure S1).
 
@@
 We expected patterns of GxE and GxWeather to vary at the locus, subpopulation, and trait level; thus we were interested in modeling the different ways effects might covary across our eight common gardens.
-We jointly re-estimated genetic effects of a subset of SNPs across all eight common gardens using *mash* models, which can flexibly capture many modes of covariance across gardens (@fig-map D; SI Appendix, Sections S1-S5, Datasets 1-6).
+We jointly re-estimated genetic effects of a subset of SNPs across all eight common gardens using *mash* models, which can flexibly capture many modes of covariance across gardens (SI Appendix, Sections S1-S5, Datasets 1-6).
 To obtain genetic effects, we used subsets of effects from garden- and subpopulation GWAS (SI Appendix, Section S2-S3; Table S1).
@@
 Eight GxWeather covariance structures were selected by the greedy algorithm, one to two per subpopulation: five for green-up date, and three for flowering date (SI Appendix, Table S2).
-Of these eight, five had mass on them in *mash* models of the strong effects (@fig-covar A-B).
+Of these eight, five had mass on them in *mash* models of the strong effects (@fig-covar B,D).
 Three of these five matrices had negative covariances between sites between Texas and North gardens (@fig-covar A), while two had all positive or near-zero covariances.
@@
 
-For green-up date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited rank-changing GxE, or a difference in sign between pairs of Texas and North gardens (@fig-effects B; SI Appendix, Fig S1).
+For green-up date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited rank-changing GxE, or a difference in sign between pairs of Texas and North gardens (@fig-effects B; SI Appendix, Fig S2).
 78.7% of pairwise comparisons between North and Texas gardens had a difference in sign, while only 28.6% and 0.2% of North-North or Texas-Texas comparisons had a difference in sign, respectively (@fig-effects B).
-The majority of pairwise effects for greenup for the Midwest (\>55%) and Both (\>85%) subpopulations were the same sign, and effects most often differed in magnitude between gardens within and between the regions (@fig-effects B).
+The majority of pairwise effects for greenup for the Midwest (\>55%) and Both (\>85%) subpopulations were the same sign, and effects most often differed in magnitude between gardens within and between the regions (@fig-effects B; SI Appendix, Fig S2).
 
@@
 Finally, in Both subpopulations, less than 20% of pairwise effects differed in sign (@fig-effects B).
-Most differences in sign were between TX1, the southernmost garden, and all other gardens (SI Appendix, Figure S1).
+Most differences in sign were between TX1, the southernmost garden, and all other gardens (SI Appendix, Figure S3).
 Similarly, more effects that differed in magnitude included gardens in the Texas region (52.3-55.9%), and most effect pairs in the North region were not distinguishable (91.5%).
@@
 We sought additional experimental support the occurrence and location of our significant SNP effect re-estimates using an independent pseudo-F2 mapping population created from Gulf & Midwest individuals and grown at the same sites (@fig-qtl A,B).
-We conducted quantitative trait loci (QTL) mapping of flowering as functions of five environmental cues that we also used as covariance matrices in *mash*, and identified eight QTL for flowering date, eight QTL for flowering as a function of day length change two days prior, one QTL for the start of vegetative growth, and two QTL for vegetative growth as a function of daylength change one day prior, all of which showed QTL by environment interactions (SI Appendix, Figure S2).
+We conducted quantitative trait loci (QTL) mapping of flowering as functions of five environmental cues that we also used as covariance matrices in *mash*, and identified eight QTL for flowering date, eight QTL for flowering as a function of day length change two days prior, one QTL for the start of vegetative growth, and two QTL for vegetative growth as a function of daylength change one day prior, all of which showed QTL by environment interactions (SI Appendix, Figure S4).
 All QTL for flowering overlapped one or more homologs from rice or *A.
@@
 ```
-![Characterization of the timing of vegetative (green-up) and reproductive (flowering) growth in the switchgrass diversity panel.
+![Characterization of the timing of the onset of vegetative (green-up) and reproductive (flowering) growth in the switchgrass diversity panel.
 (a) Map and trait histograms of green-up and flowering dates across two genetically distinct switchgrass subpopulations and eight common gardens.
 Purple represents individuals from the Midwest genetic subpopulation, and pink individuals from the Gulf subpopulation; map positions represent the original collection locations for the genotypes, and shapes represent the ecotype of the genotype.
-Vertical dashed lines indicate the summer solstice.
+Histogram vertical dashed lines indicate the summer solstice.
 Common gardens are arranged in latitudinal order.
-(b) Phenotypic correlations between clonal replicates planted at eight common gardens, within and between two genetic subpopulations.
-(c) Narrow sense heritability of green-up and flowering within single common gardens (purple) and across all eight common gardens (green), within and between two genetic subpopulations.
-(d) Flow diagram of the methods applied to the green-up and flowering dates to jointly estimate SNP effects across all sites.
-*Mash* was fit to SNP effect data and used to find covariance matrices that improved the *mash* model likelihood using a large set of randomly selected, relatively unlinked SNP effects; this model was applied to a "strong" set of SNP effects with large effect sizes in the univariate GWAS.](images/Fig1_200dpi.png){#fig-map}
+(b) Phenotypic correlations between clonal replicates planted at eight common gardens, within and between two genetic subpopulations.](images/Figure_1_AB.png){#fig-map}
 
 ![Example hypothesis-driven covariance matrices specified in *mash* and the posterior weights placed on all covariance matrices.
-(a) Left column: Five example hypothesized covariance matrices specified for the green-up date or flowering date phenotype; these matrices were created from environment-specific correlations across eight common gardens, and are described in Table 1.
-Common gardens are arranged in latitudinal order within the matrices.
-Right column: Five example canonical covariance matrices.
+(a) Common gardens are arranged in latitudinal order within the matrices.
+Top row: Four example canonical covariance matrices.
 Canonical matrices (purple) have simple interpretations, such as equal effects across all common gardens, or effects specific to a single common garden.
+Bottom row: Five example GxWeather covariance matrices specified for the green-up date or flowering date phenotype; these matrices were created from environment-specific correlations across eight common gardens, and are described in Table 1.
 (b,d) Total posterior weight placed on each covariance matrix type specified for (b) green-up date and (d) flowering date *mash* models, within and between two genetic subpopulations.
-Hypothesized covariance matrices (green).
 Covariance matrices included in *mash* that had zero posterior weight in all three *mash* runs on the genetic subpopulations, such as the identity matrix, are not shown.
-(c,e) Total posterior weight placed on covariance matrices that were hypothesized or canonical, for the (c) green-up date phenotype and (e) flowering date phenotype.](images/Figure_2_recolor_covariance_matrices_and_posterior_weights.png){#fig-covar}
+(c,e) Total posterior weight placed on covariance matrices that were hypothesized or canonical, for the (c) green-up date phenotype and (e) flowering date phenotype.](images/Figure_2.png){#fig-covar}
 
@@
 a) Examples of effect patterns at three pairs of sites with three types of GxE.
-All effects are for an alternate allele, with the reference allele effect defined to be zero at both gardens.
+All effects are for an alternate allele, with the reference allele effect defined to be zero at both gardens and represented by the dashed vertical line.
 Sign: Effects that differ in sign at these pairs of gardens (p \< 0.05, lfsr).
 Magnitude: Effects identical in sign (p \< 0.05, lfsr) that differ in magnitude by a factor of \>0.4.
-None: Effects not distinguishable by magnitude nor sign of the effect.
-b) The fraction of effects with each GxE type for the start of vegetative growth (Greenup Date) and reproductive growth (Flowering Date), within and between two genetic subpopulations.
-Common gardens are grouped by the larger region they came from: North gardens are within the natural range of the Midwest subpopulation, while Texas gardens are within the natural range of the Gulf subpopulation.](images/Figure_3_GxE.png){#fig-effects}
+Not Distinguishable: Effects not distinguishable by magnitude nor sign of the effect, with no measurable GxE.
+b) The fraction of effects with each GxE type for the onset of vegetative growth (green-up date) and reproductive growth (flowering date), within and between two genetic subpopulations.
+Common gardens are grouped by the larger region they came from: North gardens are within the natural range of the Midwest subpopulation, and include MO, NE, MI, and SD, while Texas gardens are within the natural range of the Gulf subpopulation, and include TX1, TX2, and TX3.](images/Figure_3_GxE.png){#fig-effects}
 
@@
 Stars indicate QTL with significant enrichment for SNPs in the 1% *mash* tail; G, M, and B indicate which subpopulation had enrichment: G - Gulf subpopulation, M -Midwest subpopulation, B - both subpopulations.
-Rug plots show genomic locations of SNPs in the 1% *mash* tail for flowering date for each subpopulation.
-a) QTL mapping for the start of vegetative growth (Greenup Date), and three weather-related functions of greenup date.
-b) QTL mapping for the start of reproductive growth (Flowering Date), and two weather-related functions of flowering date.](images/Figure_4_QTL_Overlaps.png){#fig-qtl}
+Rug plots show genomic locations of SNPs in the 1% *mash* tail for flowering date for each subpopulation: Both subpopulations are above the plot panel, the Gulf subpopulation is above the x-axis, and the Midwest subpopulation is below the x-axis.
+a) QTL mapping for the onset of vegetative growth (green-up date), and three weather-related functions of green-up date.
+b) QTL mapping for the onset of reproductive growth (flowering date), and two weather-related functions of flowering date.](images/Figure_4_QTL_Overlaps.png){#fig-qtl}
 

### 2024-06-12 manuscript/revision2/revision2.tex
revision2: generate revised figures 1-4 and SI figures 1-4

@@
 rank-changing GxE for these phenotypes was present across the common
-gardens (Figure~\ref{fig-map} C). h\textsuperscript{2} were typically
+gardens (SI Appendix, Figure S1). h\textsuperscript{2} were typically
 high at individual gardens: 59\% on average for green-up date, and 87\%
@@
 h\textsuperscript{2} was 0.8\% for green-up and 23.2\% for flowering
-date.
+date (SI Appendix, Figure S1).
 
@@
 eight common gardens using \emph{mash} models, which can flexibly
-capture many modes of covariance across gardens (Figure~\ref{fig-map} D;
-SI Appendix, Sections S1-S5, Datasets 1-6). To obtain genetic effects,
-we used subsets of effects from garden- and subpopulation GWAS (SI
-Appendix, Section S2-S3; Table S1). To obtain covariance structures, we
-specified both covariance structures introduced in the initial
-\emph{mash} manuscript, and GxWeather covariance matrices estimated from
-the covariance of empirical weather patterns at each garden at specific
-times before the phenological event (Table 1; SI Appendix, Section S1).
-Then, we used a greedy algorithm to select covariance matrices from each
-model's set that significantly improved the model likelihood (SI
-Appendix, Section S4; Table S2). Finally, we fit \emph{mash} models
-using the selected set of covariance matrices on a subset relatively
-unlinked (r\textsuperscript{2} \textless{} 0.2), `strong' genetic
-effects with low \emph{p}-values (SI Appendix, Section S5).
+capture many modes of covariance across gardens (SI Appendix, Sections
+S1-S5, Datasets 1-6). To obtain genetic effects, we used subsets of
+effects from garden- and subpopulation GWAS (SI Appendix, Section S2-S3;
+Table S1). To obtain covariance structures, we specified both covariance
+structures introduced in the initial \emph{mash} manuscript, and
+GxWeather covariance matrices estimated from the covariance of empirical
+weather patterns at each garden at specific times before the
+phenological event (Table 1; SI Appendix, Section S1). Then, we used a
+greedy algorithm to select covariance matrices from each model's set
+that significantly improved the model likelihood (SI Appendix, Section
+S4; Table S2). Finally, we fit \emph{mash} models using the selected set
+of covariance matrices on a subset relatively unlinked
+(r\textsuperscript{2} \textless{} 0.2), `strong' genetic effects with
+low \emph{p}-values (SI Appendix, Section S5).
 
@@
 had mass on them in \emph{mash} models of the strong effects
-(Figure~\ref{fig-covar} A-B). Three of these five matrices had negative
+(Figure~\ref{fig-covar} B,D). Three of these five matrices had negative
 covariances between sites between Texas and North gardens
@@
 between pairs of Texas and North gardens (Figure~\ref{fig-effects} B; SI
-Appendix, Fig S1). 78.7\% of pairwise comparisons between North and
+Appendix, Fig S2). 78.7\% of pairwise comparisons between North and
 Texas gardens had a difference in sign, while only 28.6\% and 0.2\% of
@@
 often differed in magnitude between gardens within and between the
-regions (Figure~\ref{fig-effects} B).
+regions (Figure~\ref{fig-effects} B; SI Appendix, Fig S2).
 
@@
 B). Most differences in sign were between TX1, the southernmost garden,
-and all other gardens (SI Appendix, Figure S1). Similarly, more effects
+and all other gardens (SI Appendix, Figure S3). Similarly, more effects
 that differed in magnitude included gardens in the Texas region
@@
 daylength change one day prior, all of which showed QTL by environment
-interactions (SI Appendix, Figure S2). All QTL for flowering overlapped
+interactions (SI Appendix, Figure S4). All QTL for flowering overlapped
 one or more homologs from rice or \emph{A. thaliana} with functionally
@@
 
-\includegraphics{images/Fig1_200dpi.png}
+\includegraphics{images/Figure_1_AB.png}
 
@@
 
-\caption{\label{fig-map}Characterization of the timing of vegetative
-(green-up) and reproductive (flowering) growth in the switchgrass
-diversity panel. (a) Map and trait histograms of green-up and flowering
-dates across two genetically distinct switchgrass subpopulations and
-eight common gardens. Purple represents individuals from the Midwest
-genetic subpopulation, and pink individuals from the Gulf subpopulation;
-map positions represent the original collection locations for the
-genotypes, and shapes represent the ecotype of the genotype. Vertical
-dashed lines indicate the summer solstice. Common gardens are arranged
-in latitudinal order. (b) Phenotypic correlations between clonal
-replicates planted at eight common gardens, within and between two
-genetic subpopulations. (c) Narrow sense heritability of green-up and
-flowering within single common gardens (purple) and across all eight
-common gardens (green), within and between two genetic subpopulations.
-(d) Flow diagram of the methods applied to the green-up and flowering
-dates to jointly estimate SNP effects across all sites. \emph{Mash} was
-fit to SNP effect data and used to find covariance matrices that
-improved the \emph{mash} model likelihood using a large set of randomly
-selected, relatively unlinked SNP effects; this model was applied to a
-``strong'' set of SNP effects with large effect sizes in the univariate
-GWAS.}
+\caption{\label{fig-map}Characterization of the timing of the onset of
+vegetative (green-up) and reproductive (flowering) growth in the
+switchgrass diversity panel. (a) Map and trait histograms of green-up
+and flowering dates across two genetically distinct switchgrass
+subpopulations and eight common gardens. Purple represents individuals
+from the Midwest genetic subpopulation, and pink individuals from the
+Gulf subpopulation; map positions represent the original collection
+locations for the genotypes, and shapes represent the ecotype of the
+genotype. Histogram vertical dashed lines indicate the summer solstice.
+Common gardens are arranged in latitudinal order. (b) Phenotypic
+correlations between clonal replicates planted at eight common gardens,
+within and between two genetic subpopulations.}
 
@@
 
-\includegraphics{images/Figure_2_recolor_covariance_matrices_and_posterior_weights.png}
+\includegraphics{images/Figure_2.png}
 
@@
 specified in \emph{mash} and the posterior weights placed on all
-covariance matrices. (a) Left column: Five example hypothesized
-covariance matrices specified for the green-up date or flowering date
-phenotype; these matrices were created from environment-specific
-correlations across eight common gardens, and are described in Table 1.
-Common gardens are arranged in latitudinal order within the matrices.
-Right column: Five example canonical covariance matrices. Canonical
-matrices (purple) have simple interpretations, such as equal effects
-across all common gardens, or effects specific to a single common
-garden. (b,d) Total posterior weight placed on each covariance matrix
-type specified for (b) green-up date and (d) flowering date \emph{mash}
-models, within and between two genetic subpopulations. Hypothesized
-covariance matrices (green). Covariance matrices included in \emph{mash}
-that had zero posterior weight in all three \emph{mash} runs on the
-genetic subpopulations, such as the identity matrix, are not shown.
-(c,e) Total posterior weight placed on covariance matrices that were
-hypothesized or canonical, for the (c) green-up date phenotype and (e)
-flowering date phenotype.}
+covariance matrices. (a) Common gardens are arranged in latitudinal
+order within the matrices. Top row: Four example canonical covariance
+matrices. Canonical matrices (purple) have simple interpretations, such
+as equal effects across all common gardens, or effects specific to a
+single common garden. Bottom row: Five example GxWeather covariance
+matrices specified for the green-up date or flowering date phenotype;
+these matrices were created from environment-specific correlations
+across eight common gardens, and are described in Table 1. (b,d) Total
+posterior weight placed on each covariance matrix type specified for (b)
+green-up date and (d) flowering date \emph{mash} models, within and
+between two genetic subpopulations. Covariance matrices included in
+\emph{mash} that had zero posterior weight in all three \emph{mash} runs
+on the genetic subpopulations, such as the identity matrix, are not
+shown. (c,e) Total posterior weight placed on covariance matrices that
+were hypothesized or canonical, for the (c) green-up date phenotype and
+(e) flowering date phenotype.}
 
@@
 sites with three types of GxE. All effects are for an alternate allele,
-with the reference allele effect defined to be zero at both gardens.
-Sign: Effects that differ in sign at these pairs of gardens (p
-\textless{} 0.05, lfsr). Magnitude: Effects identical in sign (p
-\textless{} 0.05, lfsr) that differ in magnitude by a factor of
-\textgreater0.4. None: Effects not distinguishable by magnitude nor sign
-of the effect. b) The fraction of effects with each GxE type for the
-start of vegetative growth (Greenup Date) and reproductive growth
-(Flowering Date), within and between two genetic subpopulations. Common
+with the reference allele effect defined to be zero at both gardens and
+represented by the dashed vertical line. Sign: Effects that differ in
+sign at these pairs of gardens (p \textless{} 0.05, lfsr). Magnitude:
+Effects identical in sign (p \textless{} 0.05, lfsr) that differ in
+magnitude by a factor of \textgreater0.4. Not Distinguishable: Effects
+not distinguishable by magnitude nor sign of the effect, with no
+measurable GxE. b) The fraction of effects with each GxE type for the
+onset of vegetative growth (green-up date) and reproductive growth
+(flowering date), within and between two genetic subpopulations. Common
 gardens are grouped by the larger region they came from: North gardens
-are within the natural range of the Midwest subpopulation, while Texas
-gardens are within the natural range of the Gulf subpopulation.}
+are within the natural range of the Midwest subpopulation, and include
+MO, NE, MI, and SD, while Texas gardens are within the natural range of
+the Gulf subpopulation, and include TX1, TX2, and TX3.}
 
@@
 subpopulations. Rug plots show genomic locations of SNPs in the 1\%
-\emph{mash} tail for flowering date for each subpopulation. a) QTL
-mapping for the start of vegetative growth (Greenup Date), and three
-weather-related functions of greenup date. b) QTL mapping for the start
-of reproductive growth (Flowering Date), and two weather-related
+\emph{mash} tail for flowering date for each subpopulation: Both
+subpopulations are above the plot panel, the Gulf subpopulation is above
+the x-axis, and the Midwest subpopulation is below the x-axis. a) QTL
+mapping for the onset of vegetative growth (green-up date), and three
+weather-related functions of green-up date. b) QTL mapping for the onset
+of reproductive growth (flowering date), and two weather-related
 functions of flowering date.}

### 2024-06-16 R/SI_Appendix_Datasets.Rmd
plots: examine GxE in QTL and mash SNP overlaps

@@
 
-Get `qtlplot` object by running code starting at about line 150 (for Dataset 7). Get `m_gulf_fl_hyp` and other objects by running code starting at about line 35 (for Datasets 1-6).
+Get `qtlplot` object by running code starting at about line 233 (for Dataset 7). Get `m_gulf_fl_hyp` and other objects by running code starting at about line 35 (for Datasets 1-6).
 
@@
       dplyr::rename(log10BayesFactor = .data$V1) %>%
-      dplyr::select(-.data$value) |> 
+      #dplyr::select(-.data$value) |> 
   ## TODO: FIX list/matrix column here
@@
                        QTL_LOD = qtl$lod[i], mash_Pos = qtl_sub$Pos, 
+                       mash_idx = qtl_sub$value,
                        mash_log10BayesFactor = qtl_sub$log10BayesFactor, mash_subpop = subpop, 
@@
                           QTL_LOD = qtl$lod[i], mash_Pos = qtl_sub$Pos, 
+                          mash_idx = qtl_sub$value,
                           mash_log10BayesFactor = qtl_sub$log10BayesFactor,
@@
                                  "Dataset_8_QTL_mash_overlap.csv"))
+```
+
+### Exploratory effect plots for mash QTL overlaps
+
+2N 59.5Mb: biggest effect at TX2, mostly opposite to other gardens or other gardens have NSE
+2N 60.9Mb: biggest effect at TX2, AP between TX1 and TX2
+2 Gulf effects I would say look nothing like that.
+```{r}
+overlap <- dataset_8 %>%
+  select(mash_idx, mash_subpop, everything()) |> 
+  distinct(mash_idx, .keep_all = TRUE)
+overlap
+```
+
+```{r}
+refactor_conditions <- function(phe, subpop, idx) {
+  if (phe == "Flowering Date" & subpop == "Both") {
+    plot <- mash_plot_effects(m_gam_fl_hyp, i = idx)
+    plot$effect_df$Conditions <- 
+      factor(plot$effect_df$Conditions, 
+             levels = c(paste0("Gulf_and_Midwest_FL50_", c("TX1", "TX2", "TX3",
+                                                           "OK", "MO", "NE",
+                                                           "MI", "SD"))))
+  } else if (phe == "Flowering Date" & subpop == "Gulf") {
+    plot <- mash_plot_effects(m_gulf_fl_hyp, i = idx)
+    plot$effect_df$Conditions <- 
+      factor(plot$effect_df$Conditions, 
+             levels = c(paste0("Gulf_FL50_", c("TX1", "TX2", "TX3",
+                                               "OK", "MO", "NE",
+                                               "MI", "SD"))))
+  } else if (phe == "Flowering Date" & subpop == "Midwest") {
+    plot <- mash_plot_effects(m_midw_fl_hyp, i = idx)
+    plot$effect_df$Conditions <- 
+      factor(plot$effect_df$Conditions, 
+             levels = c(paste0("Midwest_FL50_", c("TX1", "TX2", "TX3",
+                                               "MO", "M", "SD"))))
+    } else if (phe == "Greenup Date" & subpop == "Both") {
+    plot <- mash_plot_effects(m_gam_gr_hyp, i = idx)
+    plot$effect_df$Conditions <- 
+      factor(plot$effect_df$Conditions, 
+             levels = c(paste0("Gulf_and_Midwest_GR50_", c("TX1", "TX2", "TX3",
+                                                           "OK", "MO", "NE",
+                                                           "MI", "SD"))))
+  } else if (phe == "Greenup Date" & subpop == "Gulf") {
+    plot <- mash_plot_effects(m_gulf_gr_hyp, i = idx)
+    plot$effect_df$Conditions <- 
+      factor(plot$effect_df$Conditions, 
+             levels = c(paste0("Gulf_GR50_", c("TX1", "TX2", "TX3",
+                                               "OK", "MO", "NE",
+                                               "MI", "SD"))))
+  } else if (phe == "Greenup Date" & subpop == "Midwest") {
+    plot <- mash_plot_effects(m_midw_gr_hyp, i = idx)
+    plot$effect_df$Conditions <- 
+      factor(plot$effect_df$Conditions, 
+             levels = c(paste0("Midwest_GR50_", c("TX1", "TX2", "TX3", "OK",
+                                               "MO", "NE", "MI", "SD"))))
+  } else stop("not built in yet")
+  plot$ggobject <- ggplot(data = plot$effect_df) + 
+      geom_point(mapping = aes(x = as.factor(.data$Conditions), 
+                               y = .data$mn)) + 
+      switchgrassGWAS::theme_oeco + 
+      geom_errorbar(mapping = aes(ymin = .data$mn - 
+                                    .data$se, ymax = .data$mn + .data$se, 
+                                  x = .data$Conditions), 
+                    width = 0.3) + geom_hline(yintercept = 0, lty = 2) + 
+        labs(x = "Conditions", y = "Effect Size") + 
+      theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+      coord_flip() +
+      ggtitle(plot$marker)
+  return(plot)
+}
+```
+
+```{r}
+gr_qtl_overlap <- overlap |> 
+  mutate(row_n = row_number(),
+         .before = 1) |> 
+  filter(mash_phenotype == "Greenup Date") |> 
+  arrange(Chr, desc(mash_log10BayesFactor)) |> 
+  select(row_n, mash_subpop, mash_log10BayesFactor, Chr, QTL_Pos_lo,
+         Marker_in_mash_1pct_tail, everything())
+
+fl_qtl_overlap <- overlap |> 
+  mutate(row_n = row_number(),
+         .before = 1) |> 
+  filter(mash_phenotype == "Flowering Date") |> 
+  arrange(Chr, desc(mash_log10BayesFactor)) |> 
+  select(row_n, mash_subpop, mash_log10BayesFactor, Chr, QTL_Pos_lo,
+         Marker_in_mash_1pct_tail, everything())
+```
+row_n 117: No match to QTL effects, just large Midwest mash effect at MO
+
+9N QTL: (-) effects at OK & MI.
+
+For greenup, all but 3 significant mash SNPs (in the Gulf subpop) have just a single significant effect at MO.
+No QTL are like this, the only strong effect at MO is on Chr3K in the QTL.
+The 3 Gulf QTL don't really look like the 9N QTL either.
+
+```{r}
+i = 142 # 152
+for (i in 1:nrow(gr_qtl_overlap)) {
+  i <- gr_qtl_overlap$row_n[i]
+  replot <- refactor_conditions(phe = overlap$mash_phenotype[i],
+                    subpop = overlap$mash_subpop[i],
+                    idx = overlap$mash_idx[i])
+  save_plot(filename = here("manuscript", "revision2", "intermediate_files", "mash_effect_plots", "greenup", paste0(str_replace_all(replot$marker, pattern = " ", "_"), "_effect.png")), plot = replot$ggobject)
+}
+
+for (i in 1:nrow(fl_qtl_overlap)) {
+  i <- fl_qtl_overlap$row_n[i]
+  replot <- refactor_conditions(phe = overlap$mash_phenotype[i],
+                    subpop = overlap$mash_subpop[i],
+                    idx = overlap$mash_idx[i])
+  save_plot(filename = here("manuscript", "revision2", "intermediate_files", "mash_effect_plots", "flowering", paste0(str_replace_all(replot$marker, pattern = " ", "_"), "_effect.png")), plot = replot$ggobject)
+}
 ```

### 2024-06-22 manuscript/revision2/revision2.qmd
revision2: Add additional recent GxE analyses as references

@@
 Plant phenological timings are major components of plant fitness affected by multiple external environmental cues (e.g.
-degree of winter chilling, day length, temperature, and water availability) that signal existing or upcoming growing conditions [@bauerle_photoperiodic_2012; @andres2012genetic; @korner2010phenology].
+degree of winter chilling, day length, temperature, and water availability) that signal existing or upcoming growing conditions [@bauerle_photoperiodic_2012; @andres2012genetic; @korner2010phenology; @maple2024].
 Genetic responses to environmental cues determine the speed, timing, and energy apportioned to vegetative and reproductive growth and shape both the individual's lifespan and its lifetime production of viable seed.
@@
 
-Flowering time, or the transition from vegetative to reproductive growth, is a common subject of GxE research [@blackman2013interacting; @henry2014transitions; @aagren2017adaptive; @brachi2010linkage; @dittmar2014flowering; @li2018genomic; @romero2017study], a key output of selection driving adaptation to local environments [@andres2012genetic; @wadgymar2017identifying; @blumel2015flowering], and a selection target for crop improvement to adapt crops to local or future environments @jung2009flowering.
+Flowering time, or the transition from vegetative to reproductive growth, is a common subject of GxE research [@blackman2013interacting; @henry2014transitions; @aagren2017adaptive; @brachi2010linkage; @dittmar2014flowering; @li2018genomic; @romero2017study; @bustos-korts2019], a key output of selection driving adaptation to local environments [@andres2012genetic; @wadgymar2017identifying; @blumel2015flowering; @fournier-level2016], and a selection target for crop improvement to adapt crops to local or future environments @jung2009flowering.
 Changing flowering responsiveness to photoperiod cues has allowed geographic range expansion and increased yields in several cereal species [@turner2005pseudo; @faure2012mutation; @hung2012zmcct; @zakhrabekova2012induced; @yang2013oself3] and other crops [@pin2012multifaceted; @weller2019parallel].
@@
 However, most studies of flowering GxE focus on finding a single, best fitting form of genotype-environment covariance, despite the key expectation that different genetic subpopulations, and even different genomic regions, have likely evolved distinct patterns of GxE [@Weine2023.06.21.545998].
-Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or alleles with effects with opposing fitness outcomes [@levene1953genetic; @felsenstein1976theoretical; @kawecki2004conceptual; @hedrick1986genetic], previous empirical work has found limited evidence of antagonistic pleiotropy [@des2013genotype; @wadgymar2017identifying].
+Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or alleles with effects with opposing fitness outcomes [@levene1953genetic; @felsenstein1976theoretical; @kawecki2004conceptual; @hedrick1986genetic], previous empirical work has found limited evidence of antagonistic pleiotropy [@des2013genotype; @wadgymar2017identifying; @millet2016].
 However, this work has been limited by a known statistical bias that reduced detection of genetic effects that differ in sign [@des2013genotype; @anderson2013genetic; @anderson2011evolutionary].
@@
 As phenological timings are major components of plant fitness, this result supports theoretical models that local adaptation should involve antagonistic pleiotropy at the level of individual loci [@kawecki2004conceptual; @felsenstein1976theoretical; @levene1953genetic; @hedrick1986genetic].
-In addition, it expands the genetics of local adaptation research to more than two field sites and to a wider range of genetic variation by re-estimating effects from GWAS to determine that rank-changing GxE for phenological timings is common in small genomic regions [@wadgymar2017identifying; @savolainen2013ecological; @lowry2019qtl].
+In addition, it advances local adaptation research by using more than two field sites and a wide range of genetic variation to determine that rank-changing GxE for phenological timings is common in small genomic regions [@wadgymar2017identifying; @savolainen2013ecological; @lowry2019qtl].
 
@@
 In addition, the the GxWeather covariance structures allowed hypothesis testing of specific weather variables as cues for the start of vegetative and reproductive growth.
-Say that a SNP in the gene Flowering Locus C (FLC), a well-known flowering time regulator, controls flowering in a photoperiod-dependent manner.
+Say that our diversity panel contains a SNP in the gene CONSTANS (CO), a well-known flowering time regulator, and that only one of the alleles affects the promotion of flowering in a photoperiod-dependent manner.
 In that case, the joint estimate of effects for that SNP could have a high mixture proportion, or mass, on a covariance matrix created using a photoperiod-based environmental cue, such as day length at some interval prior to flowering.
-In our data, we would infer that the effect of that SNP on flowering was caused by a response to the environmental cue used to construct the GxWeather covariance structure with the largest mass.
+In our data, we would infer that the effect of that SNP on flowering was cued by the weather variable used to create the GxWeather covariance structure.
 

### 2024-06-22 manuscript/revision2/revision2.tex
revision2: Add additional recent GxE analyses as references

@@
 chilling, day length, temperature, and water availability) that signal
-existing or upcoming growing conditions (1--3). Genetic responses to
+existing or upcoming growing conditions (1--4). Genetic responses to
 environmental cues determine the speed, timing, and energy apportioned
@@
 cue predictability and relevance, and the species' adaptive strategies
-(4). Species with wide natural distributions can have multiple distinct
+(5). Species with wide natural distributions can have multiple distinct
 environmentally cued phenological responses: for example, populations of
@@
 short day, and facultative long-day flowering responses, which vary with
-their environments (5, 6). Distinct genetic responses in different
+their environments (6, 7). Distinct genetic responses in different
 environments are known as genotype-by-environment interactions, or GxE.
@@
 Flowering time, or the transition from vegetative to reproductive
-growth, is a common subject of GxE research (5--11), a key output of
-selection driving adaptation to local environments (2, 12, 13), and a
+growth, is a common subject of GxE research (6--13), a key output of
+selection driving adaptation to local environments (2, 14--16), and a
 selection target for crop improvement to adapt crops to local or future
-environments (14). Changing flowering responsiveness to photoperiod cues
+environments (17). Changing flowering responsiveness to photoperiod cues
 has allowed geographic range expansion and increased yields in several
-cereal species (15--19) and other crops (20, 21). Recent statistical
+cereal species (18--22) and other crops (23, 24). Recent statistical
 advances in studying phenological GxE have involved determining critical
 environmental indices before the phenological event occurs, such as
-photothermal time within a critical growth window (10). However, most
+photothermal time within a critical growth window (11). However, most
 studies of flowering GxE focus on finding a single, best fitting form of
@@
 different genetic subpopulations, and even different genomic regions,
-have likely evolved distinct patterns of GxE (22). Additionally, despite
+have likely evolved distinct patterns of GxE (25). Additionally, despite
 theoretical predictions that local adaptation should involve
 antagonistic pleiotropy, or alleles with effects with opposing fitness
-outcomes (23--26), previous empirical work has found limited evidence of
-antagonistic pleiotropy (12, 27). However, this work has been limited by
-a known statistical bias that reduced detection of genetic effects that
-differ in sign (27--29). Thus, despite substantial interest in the
-frequencies of various forms of GxE, the frequency of sign-changing GxE
-relative to other forms of GxE remains unknown.
+outcomes (26--29), previous empirical work has found limited evidence of
+antagonistic pleiotropy (14, 30, 31). However, this work has been
+limited by a known statistical bias that reduced detection of genetic
+effects that differ in sign (30, 32, 33). Thus, despite substantial
+interest in the frequencies of various forms of GxE, the frequency of
+sign-changing GxE relative to other forms of GxE remains unknown.
 
@@
 subpopulation. Switchgrass is considered a short-day plant with
-reproductive development strongly linked to day of the year (30).
+reproductive development strongly linked to day of the year (34).
 However, as part of its wide environmental adaptation across the eastern
 half of North America, its photoperiodicity has been predicted to differ
-by plant latitude of origin (31, 32). We previously found divergent
+by plant latitude of origin (35, 36). We previously found divergent
 Midwest and Gulf genetic subpopulations of switchgrass with distinct
@@
 genetic variation associated with each of two fitness proxies, biomass
-and overwinter survival (33). The Midwest genetic subpopulation is
+and overwinter survival (37). The Midwest genetic subpopulation is
 primarily composed of individuals from the well-studied upland
-switchgrass ecotype (34, 35), while the Gulf subpopulation has
+switchgrass ecotype (38, 39), while the Gulf subpopulation has
 individuals from the well-studied lowland ecotype and the phenotypically
-intermediate coastal ecotype (33).
+intermediate coastal ecotype (37).
 
@@
 set of these covariance matrices that significantly improved the modeled
-log-likelihood when included (SI Appendix, Section S2-S5) (36). We use
+log-likelihood when included (SI Appendix, Section S2-S5) (40). We use
 the Bayesian framework \emph{mash} (multivariate adaptive shrinkage)
-developed by (36), to refine effect size estimates from genome-wide
+developed by (40), to refine effect size estimates from genome-wide
 association (GWAS). \emph{Mash} allows us to identify and specify
@@
 biases in detecting genetic effects with the same or opposite signs
-(37). To confirm our genetic mapping of GxWeather, we compare genomic
+(41). To confirm our genetic mapping of GxWeather, we compare genomic
 locations of the significant posterior effect estimates from \emph{mash}
@@
 individual loci as this is a key theoretical prediction of local
-adaptation (23--26). To determine the frequency of rank-changing GxE, we
+adaptation (26--29). To determine the frequency of rank-changing GxE, we
 used the local false sign rate (lfsr), an analogue of the lfdr that
@@
 overlapped a gene recently functionally validated for flowering in
-switchgrass (38). All flowering and green-up QTL intervals contained at
+switchgrass (42). All flowering and green-up QTL intervals contained at
 least one SNP significant in at least one \emph{mash} run at a
@@
 that local adaptation should involve antagonistic pleiotropy at the
-level of individual loci (23--26). In addition, it expands the genetics
-of local adaptation research to more than two field sites and to a wider
-range of genetic variation by re-estimating effects from GWAS to
-determine that rank-changing GxE for phenological timings is common in
-small genomic regions (12, 39, 40).
+level of individual loci (26--29). In addition, it advances local
+adaptation research by using more than two field sites and a wide range
+of genetic variation to determine that rank-changing GxE for
+phenological timings is common in small genomic regions (14, 43, 44).
 
@@
 populations of switchgrass, a diversity panel and a pseudo-F2 cross,
-planted at eight common garden locations (35, 40, 41). We scored the
+planted at eight common garden locations (39, 44, 45). We scored the
 onset of vegetative growth, or green-up date, as the day of the year
@@
 The formation and resequencing of the diversity panel has been described
-previously (33). The diversity panel contained 134 sequenced, clonally
+previously (37). The diversity panel contained 134 sequenced, clonally
 propagated individuals from the Midwest genetic subpopulation, and 229
@@
 different subpopulations had different strengths of connection between
-our phenotypes and genotypes (42), we conducted three sets of genetic
+our phenotypes and genotypes (46), we conducted three sets of genetic
 analyses: on Gulf and Midwest genotypes separately, and on both
@@
 narrow-sense heritability (h\textsuperscript{2}) for green-up and
-flowering were done using linear mixed models and followed (33) (SI
+flowering were done using linear mixed models and followed (37) (SI
 Appendix, Section S6).
@@
 between four individuals, two Midwest and two Gulf individuals. The
-formation of this mapping population has been described previously (35);
+formation of this mapping population has been described previously (39);
 additional details on QTL mapping can be found in SI Appendix, Section
@@
 pruning and removal of regions in long-range linkage disequilibrium
-(43). Third, to make the \emph{mash} models computationally feasible, we
+(47). Third, to make the \emph{mash} models computationally feasible, we
 extracted two subsets of SNP effect estimates and standard errors from
@@
 variables as cues for the start of vegetative and reproductive growth.
-Say that a SNP in the gene Flowering Locus C (FLC), a well-known
-flowering time regulator, controls flowering in a photoperiod-dependent
-manner. In that case, the joint estimate of effects for that SNP could
-have a high mixture proportion, or mass, on a covariance matrix created
-using a photoperiod-based environmental cue, such as day length at some
-interval prior to flowering. In our data, we would infer that the effect
-of that SNP on flowering was caused by a response to the environmental
-cue used to construct the GxWeather covariance structure with the
-largest mass.
+Say that our diversity panel contains a SNP in the gene CONSTANS (CO), a
+well-known flowering time regulator, and that only one of the alleles
+affects the promotion of flowering in a photoperiod-dependent manner. In
+that case, the joint estimate of effects for that SNP could have a high
+mixture proportion, or mass, on a covariance matrix created using a
+photoperiod-based environmental cue, such as day length at some interval
+prior to flowering. In our data, we would infer that the effect of that
+SNP on flowering was cued by the weather variable used to create the
+GxWeather covariance structure.
 
@@
 hypothesis that an effect j is 0, or that there is a significant effect.
-Previous studies of antagonistic pleiotropy (e.g. (40)) have used the
+Previous studies of antagonistic pleiotropy (e.g. (44)) have used the
 lfdr or equivalent statistical tests to detect antagonistic pleiotropy.
@@
 this testing bias could lead to undercounting occurrences of
-antagonistic pleiotropy (27, 29), and sought to reduce it by permutation
-(28). However, using the lfsr to test for allelic effects that differ in
+antagonistic pleiotropy (30, 33), and sought to reduce it by permutation
+(32). However, using the lfsr to test for allelic effects that differ in
 sign does not undercount these occurences, as this statistic answers a
@@
 effect \(j\) if we were forced to declare the effect positive or
-negative (37). Thus, rather than asking ``Are these two effects
+negative (41). Thus, rather than asking ``Are these two effects
 different?'' - as we reasonably expect two effects to be, even if this
@@
 
+\bibitem[\citeproctext]{ref-maple2024}
+\CSLLeftMargin{4. }%
+\CSLRightInline{R. Maple, P. Zhu, J. Hepworth, J.-W. Wang, C. Dean,
+\href{https://doi.org/10.1093/plphys/kiae109}{Flowering time: From
+physiology, through genetics to mechanism}. \emph{Plant Physiology}
+\textbf{195}, 190--212 (2024).}
+
 \bibitem[\citeproctext]{ref-botero2015evolutionary}
-\CSLLeftMargin{4. }%
+\CSLLeftMargin{5. }%
 \CSLRightInline{C. A. Botero, F. J. Weissing, J. Wright, D. R.
@@
 \bibitem[\citeproctext]{ref-blackman2013interacting}
-\CSLLeftMargin{5. }%
+\CSLLeftMargin{6. }%
 \CSLRightInline{B. K. Blackman, Interacting duplications, fluctuating
@@
 \bibitem[\citeproctext]{ref-henry2014transitions}
-\CSLLeftMargin{6. }%
+\CSLLeftMargin{7. }%
 \CSLRightInline{L. P. Henry, R. H. Watson, B. K. Blackman, Transitions
@@
 \bibitem[\citeproctext]{ref-aagren2017adaptive}
-\CSLLeftMargin{7. }%
+\CSLLeftMargin{8. }%
 \CSLRightInline{J. Ågren, C. G. Oakley, S. Lundemo, D. W. Schemske,
@@
 \bibitem[\citeproctext]{ref-brachi2010linkage}
-\CSLLeftMargin{8. }%
+\CSLLeftMargin{9. }%
 \CSLRightInline{B. Brachi, \emph{et al.}, Linkage and association
@@
 \bibitem[\citeproctext]{ref-dittmar2014flowering}
-\CSLLeftMargin{9. }%
+\CSLLeftMargin{10. }%
 \CSLRightInline{E. L. Dittmar, C. G. Oakley, J. Ågren, D. W. Schemske,
@@
 \bibitem[\citeproctext]{ref-li2018genomic}
-\CSLLeftMargin{10. }%
+\CSLLeftMargin{11. }%
 \CSLRightInline{X. Li, T. Guo, Q. Mu, X. Li, J. Yu, Genomic and
@@
 \bibitem[\citeproctext]{ref-romero2017study}
-\CSLLeftMargin{11. }%
+\CSLLeftMargin{12. }%
 \CSLRightInline{J. A. Romero Navarro, \emph{et al.}, A study of allelic
@@
 
+\bibitem[\citeproctext]{ref-bustos-korts2019}
+\CSLLeftMargin{13. }%
+\CSLRightInline{D. Bustos-Korts, \emph{et al.},
+\href{https://doi.org/10.1111/tpj.14414}{Exome sequences and
+multi{-}environment field trials elucidate the genetic basis of
+adaptation in barley}. \emph{The Plant Journal} \textbf{99}, 1172--1191
+(2019).}
+
 \bibitem[\citeproctext]{ref-wadgymar2017identifying}
-\CSLLeftMargin{12. }%
+\CSLLeftMargin{14. }%
 \CSLRightInline{S. M. Wadgymar, \emph{et al.}, Identifying targets and
@@
 \bibitem[\citeproctext]{ref-blumel2015flowering}
-\CSLLeftMargin{13. }%
+\CSLLeftMargin{15. }%
 \CSLRightInline{M. Blümel, N. Dally, C. Jung, Flowering time regulation
@@
 
+\bibitem[\citeproctext]{ref-fournier-level2016}
+\CSLLeftMargin{16. }%
+\CSLRightInline{A. Fournier-Level, \emph{et al.},
+\href{https://doi.org/10.1073/pnas.1517456113}{Predicting the
+evolutionary dynamics of seasonal adaptation to novel climates in
+{\emph{Arabidopsis thaliana}}}. \emph{Proceedings of the National
+Academy of Sciences} \textbf{113} (2016).}
+
 \bibitem[\citeproctext]{ref-jung2009flowering}
-\CSLLeftMargin{14. }%
+\CSLLeftMargin{17. }%
 \CSLRightInline{C. Jung, A. E. Müller, Flowering time control and
@@
 \bibitem[\citeproctext]{ref-turner2005pseudo}
-\CSLLeftMargin{15. }%
+\CSLLeftMargin{18. }%
 \CSLRightInline{A. Turner, J. Beales, S. Faure, R. P. Dunford, D. A.
@@
 \bibitem[\citeproctext]{ref-faure2012mutation}
-\CSLLeftMargin{16. }%
+\CSLLeftMargin{19. }%
 \CSLRightInline{S. Faure, \emph{et al.}, Mutation at the circadian clock
@@
 \bibitem[\citeproctext]{ref-hung2012zmcct}
-\CSLLeftMargin{17. }%
+\CSLLeftMargin{20. }%
 \CSLRightInline{H.-Y. Hung, \emph{et al.}, ZmCCT and the genetic basis
@@
 \bibitem[\citeproctext]{ref-zakhrabekova2012induced}
-\CSLLeftMargin{18. }%
+\CSLLeftMargin{21. }%
 \CSLRightInline{S. Zakhrabekova, \emph{et al.}, Induced mutations in
@@
 \bibitem[\citeproctext]{ref-yang2013oself3}
-\CSLLeftMargin{19. }%
+\CSLLeftMargin{22. }%
 \CSLRightInline{Y. Yang, Q. Peng, G.-X. Chen, X.-H. Li, C.-Y. Wu, OsELF3
@@
 \bibitem[\citeproctext]{ref-pin2012multifaceted}
-\CSLLeftMargin{20. }%
+\CSLLeftMargin{23. }%
 \CSLRightInline{P. Pin, O. Nilsson, The multifaceted roles of FLOWERING
@@
 \bibitem[\citeproctext]{ref-weller2019parallel}
-\CSLLeftMargin{21. }%
+\CSLLeftMargin{24. }%
 \CSLRightInline{J. L. Weller, \emph{et al.}, Parallel origins of
@@
 \bibitem[\citeproctext]{ref-Weine2023.06.21.545998}
-\CSLLeftMargin{22. }%
+\CSLLeftMargin{25. }%
 \CSLRightInline{E. Weine, S. P. Smith, R. K. Knowlton, A. Harpak,
@@
 \bibitem[\citeproctext]{ref-levene1953genetic}
-\CSLLeftMargin{23. }%
+\CSLLeftMargin{26. }%
 \CSLRightInline{H. Levene, Genetic equilibrium when more than one
@@
 \bibitem[\citeproctext]{ref-felsenstein1976theoretical}
-\CSLLeftMargin{24. }%
+\CSLLeftMargin{27. }%
 \CSLRightInline{J. Felsenstein, The theoretical population genetics of
@@
 \bibitem[\citeproctext]{ref-kawecki2004conceptual}
-\CSLLeftMargin{25. }%
+\CSLLeftMargin{28. }%
 \CSLRightInline{T. J. Kawecki, D. Ebert, Conceptual issues in local
@@
 \bibitem[\citeproctext]{ref-hedrick1986genetic}
-\CSLLeftMargin{26. }%
+\CSLLeftMargin{29. }%
 \CSLRightInline{P. W. Hedrick, Genetic polymorphism in heterogeneous
@@
 \bibitem[\citeproctext]{ref-des2013genotype}
-\CSLLeftMargin{27. }%
+\CSLLeftMargin{30. }%
 \CSLRightInline{D. L. Des Marais, K. M. Hernandez, T. E. Juenger,
@@
 
+\bibitem[\citeproctext]{ref-millet2016}
+\CSLLeftMargin{31. }%
+\CSLRightInline{E. Millet, \emph{et al.},
+\href{https://doi.org/10.1104/pp.16.00621}{Genome-wide analysis of yield
+in Europe: allelic effects as functions of drought and heat scenarios}.
+\emph{Plant Physiology}, pp.00621.2016 (2016).}
+
 \bibitem[\citeproctext]{ref-anderson2013genetic}
-\CSLLeftMargin{28. }%
+\CSLLeftMargin{32. }%
 \CSLRightInline{J. T. Anderson, C.-R. Lee, C. A. Rushworth, R. I.
@@
 \bibitem[\citeproctext]{ref-anderson2011evolutionary}
-\CSLLeftMargin{29. }%
+\CSLLeftMargin{33. }%
 \CSLRightInline{J. T. Anderson, J. H. Willis, T. Mitchell-Olds,
@@
 \bibitem[\citeproctext]{ref-mitchell1997predicting}
-\CSLLeftMargin{30. }%
+\CSLLeftMargin{34. }%
 \CSLRightInline{R. B. Mitchell, K. J. Moore, L. E. Moser, J. O. Fritz,
@@
 \bibitem[\citeproctext]{ref-parrish2005biology}
-\CSLLeftMargin{31. }%
+\CSLLeftMargin{35. }%
 \CSLRightInline{D. J. Parrish, J. H. Fike, The biology and agronomy of
@@
 \bibitem[\citeproctext]{ref-casler2004latitudinal}
-\CSLLeftMargin{32. }%
+\CSLLeftMargin{36. }%
 \CSLRightInline{M. Casler, K. P. Vogel, C. Taliaferro, R. Wynia,
@@
 \bibitem[\citeproctext]{ref-lovell2021genomic}
-\CSLLeftMargin{33. }%
+\CSLLeftMargin{37. }%
 \CSLRightInline{J. T. Lovell, \emph{et al.}, Genomic mechanisms of
@@
 \bibitem[\citeproctext]{ref-porter1966analysis}
-\CSLLeftMargin{34. }%
+\CSLLeftMargin{38. }%
 \CSLRightInline{C. L. Porter Jr, An analysis of variation between upland
@@
 \bibitem[\citeproctext]{ref-milano2016genetic}
-\CSLLeftMargin{35. }%
+\CSLLeftMargin{39. }%
 \CSLRightInline{E. R. Milano, D. B. Lowry, T. E. Juenger, The genetic
@@
 \bibitem[\citeproctext]{ref-urbut2019flexible}
-\CSLLeftMargin{36. }%
+\CSLLeftMargin{40. }%
 \CSLRightInline{S. M. Urbut, G. Wang, P. Carbonetto, M. Stephens,
@@
 \bibitem[\citeproctext]{ref-10.1093ux2fbiostatisticsux2fkxw041}
-\CSLLeftMargin{37. }%
+\CSLLeftMargin{41. }%
 \CSLRightInline{M. Stephens,
@@
 \bibitem[\citeproctext]{ref-choi2023}
-\CSLLeftMargin{38. }%
+\CSLLeftMargin{42. }%
 \CSLRightInline{S. Choi, \emph{et al.},
@@
 \bibitem[\citeproctext]{ref-savolainen2013ecological}
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{43. }%
 \CSLRightInline{O. Savolainen, M. Lascoux, J. Merilä, Ecological
@@
 \bibitem[\citeproctext]{ref-lowry2019qtl}
-\CSLLeftMargin{40. }%
+\CSLLeftMargin{44. }%
 \CSLRightInline{D. B. Lowry, \emph{et al.}, QTL\(\times\) environment
@@
 \bibitem[\citeproctext]{ref-lovell2021}
-\CSLLeftMargin{41. }%
+\CSLLeftMargin{45. }%
 \CSLRightInline{J. T. Lovell, \emph{et al.},
@@
 \bibitem[\citeproctext]{ref-kortefarlow2013}
-\CSLLeftMargin{42. }%
+\CSLLeftMargin{46. }%
 \CSLRightInline{A. Korte, A. Farlow, The advantages and limitations of
@@
 \bibitem[\citeproctext]{ref-privuxe92017}
-\CSLLeftMargin{43. }%
+\CSLLeftMargin{47. }%
 \CSLRightInline{F. Privé, H. Aschard, M. G. B. Blum,

### 2024-06-25 manuscript/revision2/response_to_reviews/response_to_reviews.qmd
revision2: Add first pass of response to reviews

@@
 author:
-  - name: Alice MacQueen* and Tom Juenger
+  - name: Alice MacQueen and Tom Juenger
     affiliation:
@@
 code-annotations: hover
-
+bibliography: references.bib
 ---
@@
 They continued working with the group as we developed a new algorithm to address major reviewer concerns and significantly altered the manuscript.
-Because of these alterations, the response to reviewers does not have track changes for each point, though we do point to line numbers for the associated revisions in our response.
+Because of the major alterations over this time frame, the response to reviewers does not have track changes for each point, though we do point to line numbers for the associated revisions in our response.
 
-We were pleased by the useful feedback given by the reviewers of our original submission.
+We were pleased by the useful feedback given by the reviewers of our original submission and believe that the greedy mash algorithm we implemented addresses the major statistical concern of reviewer 2, as well as substantially improving our use and interpretation of covariance matrices in *mash*.
 We have addressed all the editorial issues suggested by the two reviewers, and included in this letter are point-by-point responses to these issues.
 
-The previous reviewers brought up two major concerns:
-
+> ::: {.callout-warning appearance="minimal"}
 > Reviewer #1:\
@@
 > There are many papers on crop modelling (eg work by Mark Cooper and others in maize) and the authors might want to consider linking their results to crop modelling to demonstrate the relevance of the results of this study
+> :::
 
-Comments on additional findings from other crop species
+*I'm still reading through these to determine which to add. May need to make additional edits to the Intro/Discussion to add these citations.*
 
+> ::: {.callout-warning appearance="minimal"}
 > Results\
@@
 One reason might be that objectives of the study have remained rather vague as well as the generic conclusions that can be drawn from the results.
+> :::
 
-We edited the Introduction substantially to differentiate this study from Lovell et al 2021.
+While we do rely on the same genotypic data and common gardens used in [Lovell *et al.,* 2021](https://doi.org/10.1038/s41586-020-03127-1), and make use of the population structure findings of [Lovell *et al.,* 2021](https://doi.org/10.1038/s41586-020-03127-1) to divide our genotypes into subpopulations for GWAS, the previous work considered fitness proxies and this work considers phenological traits.
+This work has two additional, key advances over the previous work: it can assign loci genome-wide to multiple kinds of GxE and GxWeather, and has an unbiased statistical method to identify loci with and without rank-changing GxE.
+We now summarize the findings from [Lovell *et al.,* 2021](https://doi.org/10.1038/s41586-020-03127-1) that we use as a springboard for this work on lines 136-145, and outline the additional advances of this study on lines 156-163, 169-171, and 883-893.
 
+> ::: {.callout-warning appearance="minimal"}
 > In the introduction the aim is given as „we test if these populations differ in their phenological adaptation and hence their phenological GxE".
 Given prior knowledge such as the origin of the populations and earlier results (Lovell et al.) what is expected?
+> :::
 
-We edit the Introduction to more clearly articulate the aims of this manuscript.
-Our key expectation is that different genetic subpopulations and genomic regions have likely evolved distinct patterns of GxE.
+We now edit the Introduction to more clearly articulate the aims of this manuscript.
+Our key expectation is that different genetic subpopulations and genomic regions have likely evolved distinct patterns of GxE (lines 55-58; lines 219-223).
 Thus, we aim to identify the kinds of GxE present for each subpopulation and each trait.
+(lines 129-131).
 
+> ::: {.callout-warning appearance="minimal"}
 > The results for these two populations did not show a general pattern leading to conclusions on the environmental cues.
@@
 , it was difficult to see the big picture.
+> :::
 
 We agree that we do not identify one general pattern or one environmental cue affecting any set of populations or phenological trait.
-Rather, we aim to identify the kinds of GxE present genome-wide and elaborate when these kinds of GxE differ (Figure 2), and demonstrate that individual loci have different types of GxE across environmental regions and subpopulations (Figure 3).
+Instead, our key expectation is that different populations and different genetic loci will have evolved distinct patterns of GxE (lines 55-58; lines 219-223; SI Appendix Section S1).
+Thus, we aim to identify the kinds of GxE present genome-wide and elaborate when these kinds of GxE differ (Figure 2), and demonstrate that individual loci have different types of GxE across environmental regions and subpopulations (Figure 3).
 
+> ::: {.callout-warning appearance="minimal"}
 > The authors claim that the environmental cues in the hypothesis based models improved model fit.
@@
 Unless results are validated in independent data or with cross validation the predictive power of the cues cannot be judged.
+> :::
 
-We edit the Results to a) more clearly articulate this claim and b) add an algorithm to more rigorously demonstrate this claim given the data that we have.
-We agree that given the data we have, the predictive power of these cues cannot be judged.
-*may need to look at QTL GxE to address this*.
-*It differs from our rewritten manuscript aims*.
+We edit the Results to (i) change our analysis add an algorithm to more rigorously demonstrate this claim given the data that we have, and (ii) more clearly articulate this claim.
+We now employ a greedy mash algorithm to iteratively add covariance matrices to the *mash* model, only adding matrices that significantly improve the mash model log-likelihood, as we elaborate on further below in our [response to reviewer 2](#fig-greedy).
 
+We agree that given the data we have, the predictive power of these cues cannot be judged - in fact, it is hard to imagine how our GxWeather covariances could be used predictively, as we define the GxWeather covariance matrices retroactively from each genotype's green-up date and flowering date (lines 231 - 235; SI Appendix, Section S1).
+
+> ::: {.callout-warning appearance="minimal"}
 > Results from the SNPs "mash model of Midwest green-up fell on a covariance matrix of average temperature in the 10 days prior to Midwest green-up".
 These results should be linked to the phenotypic GxE analysis.
+> :::
 
-We edit the results to directly compare the GxWeather matrices and the phenotypic correlations in Figure 1B. (lines 406-409; lines 417-420).
+We edit the results to directly compare the GxWeather matrices and the phenotypic correlations in Figure 1B. (lines 241-258; lines 269-276; lines 283-287).
 
+> ::: {.callout-warning appearance="minimal"}
 > Flowering posterior weights were higher than green-up weights. What does that mean? How does that link to the different heritabilities and levels of GxE?
+> :::
 
-We believe this comment refers to there being higher loadings on the GxWeather matrices for flowering date than there was for the date of the start of vegetative growth in the original paper (original Figure 2).
-Our revised analyses uses a selection process for canonical and GxWeather covariance matrices to select only the subset that significantly improves the model log-likelihood.
+We believe this comment refers to there being higher loadings on the GxWeather matrices for flowering date than there were for the date of the start of vegetative growth in the original paper (original Figure 2).
+Our revised analyses uses a selection process for canonical and GxWeather covariance matrices to select only the subset that significantly improves the model log-likelihood, using a [greedy mash algorithm to iteratively select matrices](#fig-greedy).
 Thus, the covariance matrices included in each model have changed in this revision, as have the relative loadings on GxWeather and canonical covariance matrices.
-*what do we see now?*
 
+Now, posterior weights on the GxWeather matrices vary between 0 to 0.6, but we do not see clear patterns between posterior weights between phenotypes or subpopulations, save that for most models, a large fraction of effects, 0.2 to 0.6, load onto a matrix of 'No Effects' that shrinks effects towards zero.
+We don't believe it's appropriate to make direct comparisons between heritability or a variance components analysis and these mash loadings, so we do not do so in the paper; instead of the variance due to additive genetic variation and GxE, these show the additive genetic effects that most closely matched different patterns of GxE and GxWeather, regardless of the proportion of additive variance each locus has.
+
+> ::: {.callout-warning appearance="minimal"}
 > What can you conclude from the pattern of antagonistic pleiotropy between Texas and Northern gardens that has not been seen in the GxE analysis?
+> :::
 
@@
 
+> ::: {.callout-warning appearance="minimal"}
 > Discussion\
@@
 It looks like the latter so what are the consequences?
+> :::
 
-We think that these patterns need to be assessed individually for different genetic material, environments, and traits (lines 575-578); thus, we develop an approach to specify multiple environmental cues and compete them to explain patterns of genetic effects (lines 591-595): we use a greedy algorithm to use with mash to select covariance patterns, and use mash to flexibly identify these patterns across populations and environments.
-(lines \_\_).
+We think that these patterns need to be assessed individually for different genetic material, environments, and traits (lines 220-223); thus, we develop an approach to specify multiple environmental cues and compete them to explain patterns of genetic effects (lines 230-237): we use a greedy algorithm to use with mash to select covariance patterns, and use mash to flexibly identify these patterns across populations and environments.
+(lines 238-240).
 
+*Need to add more here to fully address this point*
+
+> ::: {.callout-warning appearance="minimal"}
 > "this is the first experimental work using QTL mapping and GWAS across ....". I disagree, see references above and others.
+> :::
 
-Edited to clarify our claim. NB: We do cite the first experimental work using both QTL mapping and GWAS (Brachi et al), and there are many others; we meant using these approaches to map GxE.
+We edit this to clarify our claim.
+NB: We do cite the first experimental work using both QTL mapping and GWAS ([Brachi *et al.,* 2010](https://doi.org/10.1371/journal.pgen.1000940)), and there are many others; we meant using these approaches to map multiple types of GxE.
 
+> ::: {.callout-warning appearance="minimal"}
 > "Gulf and Midwest subpopulations have two distinct photoperiod-related flowering responses....". This has been shown for other species, eg Unterseer et al. 2016 Genome Biology
+> :::
 
@@
 
+> ::: {.callout-warning appearance="minimal"}
 > Data/Methods\
@@
 > Is one year data enough to make conclusions about the effect of environmental cues? It is well known that GxYear is more pronounced than GxLocation.
+> :::
 
@@
 
+> ::: {.callout-warning appearance="minimal"}
 > The reader should be able to understand what the hypothesis in the hypothesis-driven models is without looking at the supplement.
+> :::
 
@@
 
+> ::: {.callout-warning appearance="minimal"}
 > Reviewer #2:\
@@
 Also, because the methods are relatively novel and complex, it would be very helpful to provide more intuition behind the analytical approaches and more access to the raw results so that others can understand the approach more fully.
-
+>
 > First a note: It would be helpful if the authors would provide a pdf with line-numbers and with a font that can be copied directly.
+> :::
 
@@
 
+> ::: {.callout-warning appearance="minimal"}
 > Major issues:\
@@
 But as is, I don't think these metrics are interpretable for what they are designed for.
+> :::
 
-We agree that the choice of environmental measure is an extremely important consideration.
+We agree that the choice of environmental measure is an extremely important consideration, and we recognize that the time window the weather signal is integrated over could be long.
+We wanted our model to inform us on reasonable time windows, rather than having us assert these (probably incorrectly) in our model.
+Initially, we thought computational feasibility constrained the number of covariance matrices we could add to the mash model, as the runtime of mash increases with the number of covariance matrices included.
+To address this comment and the mash statistical issue the reviewer points out below, we now specify many more GxWeather covariance matrices with more time frames - 1-7 days prior to the phenological event, and 14, 21, and 28 days prior to the phenological event, generating 48-60 matrices per weather variable.
+Then, we used a greedy mash algorithm, [explained below](#fig-greedy), to select GxWeather covariance matrices from this set that significantly improve the model likelihood.
+We thus extended the time frame for GxWeather covariance that our models could capture; however, no 21 or 28 day matrices were selected by the greedy algorithm.
 
-> \
-> - I’ve struggled to understand the “hypothesis-based covariance matrix” idea.
+We also now try and explicitly describe the two phenotypes more clearly in the text - both of the traits we map were measured on individuals, but specifically when approximately half of the ramets of the genotype had open flowers (for flowering) or green leaves (for green-up).
+Thus we are looking for cues driving vegetative and reproductive transitions for a majority of ramets, not for heading (flower emergence on the panicle).
+While there may be physiological mechanisms driving flower opening once the transition to reproductive development has been made, our results show that there is also genetic variation underlying differences in timing of the physiological mechanisms underlying flower opening.
+
+> ::: {.callout-warning appearance="minimal"}
+> \- I’ve struggled to understand the “hypothesis-based covariance matrix” idea.
 Since this idea is really key to the idea of this paper, I think the authors should work to make it much more clear and intuitive, perhaps a diagram could help.
@@
 Or maybe that these SNPs were also GWAS hits for daylength at flowering itself.
+> :::
 
->Statistical issues:\
->\
-> - Narrow-sense heritability: A key result is that the h^2 is low when measured across all trials, but high within trials.
-However the model used to estimate the global h^2 is missing a term for replication of lines across accessions.
+We thank the reviewer for this insight; we have reframed the way we talk about these covariance matrices in the paper (starting with the title, abstract, and significance statement; continuing throughout).
+Instead of hypothesis-based covariance matrices, we emphasize that the matrices we create are looking at an interaction between genetics and weather - they are GxWeather matrices, and we look for effects with different GxWeather interactions that are based on specific weather cues.
+
+> ::: {.callout-warning appearance="minimal"}
+> Statistical issues:\
+> \
+> - Narrow-sense heritability: A key result is that the h\^2 is low when measured across all trials, but high within trials.
+However the model used to estimate the global h\^2 is missing a term for replication of lines across accessions.
 u only accounts for replication of additive effects of lines across gardens, but non-additive genetics can be persistent too.
 Also, this model doesn’t allow var(e) to vary across locations but it’s likely that it does.
-If the goal is to show rank-crossing GxE, actually fit a GxE model like was done for the environmental indices to directly show it.
->\
+If the goal is to show rank-crossing GxE, actually fit a GxE model like was done for the environmental indices to directly show it.\
 > - Variance components analysis: The specification of this model is incorrect because Var(u) and Var(ul) are assigned the same distribution.
@@
 Sometimes this is constrained to be diagonal (ie no covariance among gardens), and sometimes to be proportional to I (constant variance and no covariance among gardens), but these more restrictive models should be justified.
+> :::
 
-> \
-> - GWAS: The GWAS model the authors use is not specified (nor was it specified in the 2021 Nature paper referenced).
+We agree that correctly specifying the type of GxE in our models of additive variation across gardens is important for all models in this manuscript.
+In many cases the complexity of the models that we can specify using this data is limited, because of the low sample size - we typically have one replicate per genotype in each common garden, and the low number of genotypes present at each garden in each strongly stratified subpopulation means that we cannot fit complicated GxE structures using linear mixed models, often needing to fit very simple/restrictive variance-covariance structures.
+The entire focus on mash was to allow us to fit less structured variance-covariance structures, as well as more of them.
+
+Thus we thought hard about the minimum number of linear mixed models we could specify to justify our modeling approach with *mash*.
+We think that the presence of negative phenotypic covariation and additive genetic variation at each garden is sufficient to motivate a search for the sign- or rank-changing GxE that could potentially underlie strong negative covariation between the North and Texas gardens.
+Clearly, additive genetic variation is necessary to conduct a genetic mapping analysis; in addition, we find many more loci with rank-changing GxE for green-up than for flowering, and green-up has negative phenotypic correlations which flowering does not.
+
+We could not find a similar compelling reason to include the variance components analysis of the phenological traits or weather-derived traits based on these phenological dates.
+Because we construct the GxWeather matrices from covariation in these weather-derived cues, then mash loads loci onto these GxWeather matrices, we felt it did not add value to the paper to include a variance components analysis of these as additional traits; thus, we remove the variance components analysis from this revision.
+
+Additionally, we move the narrow-sense heritability analysis to the SI Appendix (Figure S1).
+Our goal was to show suitability for further analysis of additive genetic variation, not directly show rank-crossing GxE with this mixed linear model, so we do not change the environmental effect of site, which does not have an interaction term, in the narrow-sense heritability models.
+
+> ::: {.callout-warning appearance="minimal"}
+> \- GWAS: The GWAS model the authors use is not specified (nor was it specified in the 2021 Nature paper referenced).
 I believe looking at the source code that the model is a linear model with PCs to account for structure but no kinship matrix.
@@
 If each of these was a true positive, the average effect size of each block would be only 0.005% of the genetic variance.
-With only 350 accessions, the power to detect a locus that explains 10% of the genetic variance is only 50% with alpha = 10^-5 and MAF = 0.05.
+With only 350 accessions, the power to detect a locus that explains 10% of the genetic variance is only 50% with alpha = 10\^-5 and MAF = 0.05.
 So if the model is detecting tons of loci it’s likely that the model is severely biased by population structure, and these biases are likely to be somewhat consistent across gardens because the structure is similar.
 mash doesn’t help when the individual models are biased, it’ll find patterns of correlation whether they are biological or not and use those to shrink effect sizes together.
+> :::
 
->\
-\- Mash: A lot of the results interpretation is based on interpreting the SNP loadings on the specified covariance matrices.
+First, we now better document our GWAS methodology for this manuscript in Section S2 of the SI Appendix.
+We also include additional information on GWAS, including the number of PCs used and the Manhattan and QQ plots and associated data tables in the Github repository associated with the paper.
+
+Second, though we selected 19K LD blocks in our set of 'strong' effects for *mash*, very few of these effects were significant (\<500 relatively unlinked LD blocks per set of eight gardens; this still represents some genomic inflation, no doubt, but not nearly as bad as 19K significant blocks would be).
+We now clarify this important detail in the main manuscript and in the SI Appendix.
+
+Third, population structure could certainly bias the resultant mash models.
+We describe the steps we take to prevent "garbage in" to our mash models, including removing conditions where GWAS had significant population structure.
+However, it is our intuition that, should biases remain in our models, as they undoubtedly do, mash would not load these biased effects onto our GxWeather covariance matrices - instead, we these effects could perhaps load onto canonical matrices, such as garden-specific effects, or perhaps onto data-driven matrices.
+Thus, we expect our assessment of the proportion of the genome that shows GxWeather effects to be conservative if there is residual population structure.
+
+> ::: {#comment-cov-correlations .callout-warning appearance="minimal"}
+> \- Mash: A lot of the results interpretation is based on interpreting the SNP loadings on the specified covariance matrices.
 This is a secondary use of mash (the primary being the refinement of effect sizes), and while they do interpret these somewhat in the mash paper, these loadings need to be interpreted with caution.
@@
 It might be safer to try dropping specific covariance matrices and comparing the model performance in a held-out testing set to evaluate the importance of the hypothesis-derived covariance matrices.
+> :::
 
@@
 The results indicate that only a small number of the original hypothesis matrices (between three and five) are necessary to reach the maximum likelihood model when using the same likelihood ratio test implemented in Urbut et al., indicating that correlation among hypothesis covariance matrices is high.
-This redundancy was confirmed through our application of the greedy algorithm to a combined set of canonical, data driven, and hypothesis covariance matrices (Figures 3 and 4)
-Our primary conclusion from this analysis is in line with what the reviewer posited, that there is extensive redundancy among canonical, data driven, and hypothesis covariance matrices that is leading to biologically informative patterns of covariance between site effect sizes to be arbitrarily distributed among them.
+This redundancy was confirmed through our application of the greedy algorithm to a combined set of canonical, data driven, and hypothesis covariance matrices (Figures 3 and 4) Our primary conclusion from this analysis is in line with what the reviewer posited, that there is extensive redundancy among canonical, data driven, and hypothesis covariance matrices that is leading to biologically informative patterns of covariance between site effect sizes to be arbitrarily distributed among them.
 We additionally found evidence that while the correlations between data driven and environmentally informed (hypothesis) covariance matrices, there is extensive similarity in the matrices that are included in the maximum likelihood models across both phenotypes and in the analysis of the Gulf subpopulation, Midwest subpopulations, and the combined cohort.
 
-```r
+``` {#fig-greedy .r}
 Greedy algorithm for mash model selection. 
@@
 ```
+
 1.  Initialize covariance matrices of interest
@@
 
->\
-- There appears to be an issue with the construction of the hypothesis-derived covariance matrices.
+> ::: {.callout-warning appearance="minimal"}
+> \- There appears to be an issue with the construction of the hypothesis-derived covariance matrices.
 I believe the goal here is to estimate the genetic covariance among gardens based on the phenotypic covariance (P) and the estimated residual covariance (E), where P = G + E.
@@
 Or just mean-standardize the traits first and then calculate the covariance of this standardized matrix.
+> :::
 
->\
-\ - Antagonistic Pleiotropy: Maybe I’m missing the idea of this analysis, but it doesn’t seem to me that this strategy for partitioning loci between antagonistic pleiotropy and differential sensitivity would have equal power to detect each class.
+We now use the narrow-sense heritability as the diagonal for each GxWeather matrix, and clarify this in the Supplementary Methods (SI Appendix, Section S1, first and eighth paragraph - first and last paragraph on page two).
+
+> ::: {.callout-warning appearance="minimal"}
+> \- Antagonistic Pleiotropy: Maybe I’m missing the idea of this analysis, but it doesn’t seem to me that this strategy for partitioning loci between antagonistic pleiotropy and differential sensitivity would have equal power to detect each class.
 It seems that antagonistic pleiotropy should be a subset of differential sensitivity because the latter is defined as simply a change in magnitude.
-But even if it’s restricted to “different magnitude but same sign”, the practical definition here seems to be that antagonistic pleiotropy is detected when the signs are different and both lfsrs < threshold (0.05?), while differential sensitivity additionally requires a threshold on the difference in effect sizes (0.4) with no justification for why this size was chosen.
+But even if it’s restricted to “different magnitude but same sign”, the practical definition here seems to be that antagonistic pleiotropy is detected when the signs are different and both lfsrs \< threshold (0.05?), while differential sensitivity additionally requires a threshold on the difference in effect sizes (0.4) with no justification for why this size was chosen.
 This additional criterion will necessarily make the rate of detection different.
@@
 I’m not aware of previous discussions in the literature of trying to differentiate these two specific classes of GxE variants and not really sure theoretically what the importance would be.
+> :::
+
+We believe that the use of the lfsr to detect effects with rank-changing GxE is a key advance in our manuscript, and so we have made changes to the Introduction, Results, and Materials & Methods (lines 957-997; lines 1055-1056), as well as the SI Appendix, to explain and motivate this statistical change.
+
+Our expectation is that in nature, effects will always differ.
+Statisically, this difference will not always be detectable.
+Thus, rather than ask "Are these two effects different?" - as we reasonably expect two effects to be, even if this difference cannot be measured - the local false sign rate answers a more meaningful question: Can we be confident in the sign of this effect?
+
+We think that if evolutionary geneticists are interested in antagonistic pleiotropy and rank-changing GxE, then they should use the lfsr to measure their confidence in the sign of the effect, rather than the lfdr to measure their confidence that the effect is different than zero.
+In the lfsr case, comparisons between antagonistic pleiotropy and conditional neutrality are no longer sensible, as we are no longer doing a statistical test that can robustly detect conditional neutrality (just as the lfdr does not robustly detect antagonistic pleiotropy).
+
+It's true that the threshold for differential sensitivity is arbitrary.
+We have reworded the manuscript to stress that there is equal power to detect effects of different sign as there are effects of the same sign (lines 994-997; lines 1055-1056).
+Another class of effects we have in our analysis, that differ from classic antagonistic pleiotropy & conditional neutrality comparisons are effects which are not distinguishable by sign or magnitude.
+It is not accurate that differentially sensitive loci will be more likely to pass the lfsr, because the loci are first selected by significant lfsr, then divided further into effects that can & cannot be distinguished by magnitude.
+Thus effects that are significant, but have small magnitude in both conditions will be labeled as not being distinguishable - essentially, these effects do not have detectable GxE.
+We propose that, if our intent is to detect sign-changing GxE, then we should use an unbiased statistical test to look at loci with and without sign-changing GxE, the lfsr.

### 2024-06-26 R/Analysis_v1.2_mash_using_greedy_covar.qmd
revision2: Drop MO from Midwest GR50 mash analysis due to GWAS pop structure issues

--- 2024-06-10 R/Analysis_v1.2_mash_using_greedy_covar.qmd
@@
                            "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
-workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
+workingdir <- here()
 inputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs")
@@
 inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
-                                   "inputkinship.txt"),  delim = " ", 
+                                   "inputkinship.txt"),  
+                         delim = " ", comment = "#",
                          col_names = "SNPfiles")
@@
 for(k in c(1, 3)){
-phe_suffix <- paste("GR50", subpop_v[k], 
+phe_suffix <- paste("GR50", snp_v[k], 
                     sep = "*.*")
 snp <- snp_attach(inputfiles$SNPfiles[k])
-gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
+gwas_rds <- pvdiv_results_in_folder(path = here("analysis", "gwas", "BLUPs", "single_sites", "Datatables_single_sites"), 
                                       pattern = paste0("GWAS_datatable*.*",
                                                        phe_suffix))
-phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
-    str_replace(., "_$", "")
-phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
-numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
-out1phe <- pvdiv_bigsnp2mashr(path = file.path(outputdir), 
-                           snp = snp, gwas_rds = gwas_rds, 
-                           phenotypes = phenotypes, numSNPs = numSNPs, 
-                           clump = TRUE, scaled = TRUE, saveoutput = TRUE,
-                           suffix = paste0("GR50_", names(subpop_v)[k]))
-}
-```
-
-#### 2b. Midwest, dropping gardens with zero values.
-
-Midwest_GR50_SD have values of zero for Bhat and Shat. Drop this from Midwest analyses.
-
-```{r check midwest}
-mwdf <- readRDS(file.path(outputdir,
-                          "B_hat_strong_df_19000topSNPs_GR50_Midwest.rds"))
-
-summary(mwdf)
-```
-
-```{r mash midwest}
-k=2
-phe_suffix <- paste("GR50", subpop_v[k], 
-                    sep = "*.*")
-snp <- snp_attach(inputfiles$SNPfiles[k])
-gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
-                                      pattern = paste0("GWAS_datatable*.*",
-                                                       phe_suffix))
-gwas_rds <- gwas_rds[c(1:4,6:8)]
-
 phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -41) %>%
@@
                            suffix = paste0("GR50_", names(subpop_v)[k]))
-# Ran mash using a 'nohup_v0.6' R script.
+}
+```
+
+#### 2b. Midwest, dropping gardens with zero values.
+
+Midwest_GR50_SD have values of zero for Bhat and Shat. Drop this from Midwest analyses.
+
+```{r check midwest}
+mwdf <- readRDS(file.path(outputdir,
+                          "B_hat_strong_df_19000topSNPs_GR50_Midwest.rds"))
+
+summary(mwdf)
+```
+
+```{r mash midwest}
+k=2
+phe_suffix <- paste("GR50", snp_v[k], 
+                    sep = "*.*")
+snp <- snp_attach(inputfiles$SNPfiles[k])
+gwas_rds <- pvdiv_results_in_folder(path = here("analysis", "gwas", "BLUPs", "single_sites", "Datatables_single_sites"), 
+                                      pattern = paste0("GWAS_datatable*.*",
+                                                       phe_suffix))
+gwas_rds <- gwas_rds[c(1, 3:4,6:8)]
+
+phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -41) %>%
+    str_replace(., "_$", "")
+phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
+numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
+out1phe <- pvdiv_bigsnp2mashr(path = here("analysis", "gwas", "BLUPs",
+                                          "single_sites",
+                                          "Datatables_single_sites"), 
+                           snp = snp, gwas_rds = gwas_rds, 
+                           phenotypes = phenotypes, numSNPs = numSNPs, 
+                           clump = TRUE, scaled = TRUE, saveoutput = TRUE,
+                           suffix = paste0("GR50_", names(subpop_v)[k]))
+
 ```
@@
 
-Files with the six suffixes I have/want mash results for: B_hat_strong_df_19000topSNPs_FL50_Gulf_and_Midwest.rds B_hat_strong_df_19000topSNPs_FL50_Gulf.rds B_hat_strong_df_19000topSNPs_GR50_Gulf_and_Midwest.rds B_hat_strong_df_19000topSNPs_GR50_Gulf.rds B_hat_strong_df_24000topSNPs_GR50_Midwest.rds B_hat_strong_df_33000topSNPs_FL50_Midwest.rds
+Files with the six suffixes I have/want mash results for: B_hat_strong_df_19000topSNPs_FL50_Gulf_and_Midwest.rds B_hat_strong_df_19000topSNPs_FL50_Gulf.rds B_hat_strong_df_19000topSNPs_GR50_Gulf_and_Midwest.rds B_hat_strong_df_19000topSNPs_GR50_Gulf.rds B_hat_strong_df_33000topSNPs_GR50_Midwest.rds B_hat_strong_df_33000topSNPs_FL50_Midwest.rds
 
@@
 
-### sketch out how h2 is calculated in rrBLUP
+### Calculate h2 in rrBLUP
 
@@
 library(rrBLUP)
-k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
+k_full <- read_rds(here("data/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds"))
 fl_ten_sites
@@
 
-#### greenup: 7 sites
-
-MI, MO, NE, OK, TX1, TX2, TX3
+#### greenup: 6 sites
+
+MO has clear population structure bias in the Manhattan & QQplot
+
+SD dropped because of many 0 Bhat and 0 Shat values
+
+MI, NE, OK, TX1, TX2, TX3
 
@@
              IND = as.factor(IND)) %>%
-      select(PLANT_ID, SUBPOP, IND, MI, MO, NE, OK, TX1, TX2, TX3) #####  IL,
+      select(PLANT_ID, SUBPOP, IND, MI, NE, OK, TX1, TX2, TX3) #####  IL,
 
@@
     h2_diag <- c()
-    site_names <- c("MI", "MO", "NE", "OK", "TX1", "TX2", "TX3")
+    site_names <- c("MI", "NE", "OK", "TX1", "TX2", "TX3")
     for(j in seq_along(site_names)) {
@@
   if (is.matrix(U_hyp_2019[[i]])) {
-    matrixout = matrix(unlist(U_hyp_2019[[i]]), ncol = 7, byrow = TRUE)
+    matrixout = matrix(unlist(U_hyp_2019[[i]]), ncol = 6, byrow = TRUE)
   } else {
-    matrixout = matrix(unlist(U_hyp_2019[[i]]$mat), ncol = 7, byrow = TRUE)
+    matrixout = matrix(unlist(U_hyp_2019[[i]]$mat), ncol = 6, byrow = TRUE)
   }
@@
 
+## June 2024
+
 ```{r}
@@
 workingdir <- here("analysis", "mash_greedy_algorithm", 
-                                   "mash_h2_diag_covar", "midwest_greenup")
+                                   "mash_h2_diag_covar", "midwest_inputs")
 pheno = "greenup"
@@
 
-mat_betas = matrix(unlist(betas),ncol=7,byrow = TRUE)
-mat_stderrs = matrix(unlist(stderrs),ncol=7,byrow = TRUE)
-
-sites = c('MI','MO','NE', 'OK','TX1','TX2','TX3')
+mat_betas = matrix(unlist(betas),ncol=6,byrow = TRUE)
+mat_stderrs = matrix(unlist(stderrs),ncol=6,byrow = TRUE)
+
+sites = c('MI','NE', 'OK','TX1','TX2','TX3')
 data=mash_set_data(mat_betas,mat_stderrs)
@@
   if (matrix %in% c('equal_effects','identity','simple_het_1','simple_het_2','simple_het_3')) {
-  matrixout = matrix(unlist(canonical[matrix]), ncol = 7, byrow = TRUE)
+  matrixout = matrix(unlist(canonical[matrix]), ncol = 6, byrow = TRUE)
   row.names(matrixout) = sites
@@
 print(canonical)
-for (i in 1:7) {
+for (i in 1:6) {
   singleton = paste0('singletons_',i)
@@
 
-  singmat = matrix(unlist(canonical[singleton]),ncol = 7, byrow = TRUE)
+  singmat = matrix(unlist(canonical[singleton]),ncol = 6, byrow = TRUE)
   row.names(singmat) = sites
@@
 for (mat in matnames) {
-  matrixout = matrix(unlist(canonical[mat]),ncol = 7,byrow=TRUE)
+  matrixout = matrix(unlist(canonical[mat]),ncol = 6,byrow=TRUE)
   row.names(matrixout) = sites
@@
   write.csv(matrixout, file = here(workingdir, "matrices",
-                                   paste0(pheno,'.', cohort,'.', effect,'.', mat, '.csv')))
+                                   paste0(pheno,'.', cohort,'.', effect,'.',
+                                          mat, '.csv')))
 }
@@
 
-Set up new analysis/mash_greedy_algorithm/mash_1_to_28d_covar/ or subdirectory to have all the subdir necessary to run `run.mash.greedy.search.py` python script.
+Set up new analysis/mash_greedy_algorithm/mash_h2_diag_covar/ or subdirectory to have all the subdir necessary to run `run.mash.greedy.search.py` python script.
 I think these may have to be pre-generated for the script to work, and I'll make sure they're empty to keep there from being I/O issues I haven't tested for.
@@
 
-When I change to my inputs for effects & stderrs, I will need to adjust midwest covar & effects appropriately to reflect only 6 (flowering) and 7 (greenup) rows & columns used; the other GWAS were dropped as too low power or due to substantial phenotype data loss, & including them would give misleading results.
+When I change to my inputs for effects & stderrs, I will need to adjust midwest covar & effects appropriately to reflect only 6 (flowering) and 6 (greenup) rows & columns used; the other GWAS were dropped as too low power, due to population structure issues in the GWAS, or due to substantial phenotype data loss, & including them would give misleading results, even if it would just result in loadings onto the single garden effect matrices.
 
@@
 
-Try Gulf & All with my known inputs: python run.mash.greedy.search.py -p greenup -c gulf -e random python run.mash.greedy.search.py -p greenup -c all -e random python run.mash.greedy.search.py -p flowering -c gulf -e random python run.mash.greedy.search.py -p flowering -c all -e random
+Try Gulf & All with my known inputs: python run.mash.greedy.search.py -p greenup -c gulf -e random python run.mash.greedy.search.py -p greenup -c all -e random python run.mash.greedy.search.py -p flowering -c gulf -e random
+
+python run.mash.greedy.search.py -p flowering -c all -e random
 
@@
     if (i == 2 & j == 1) {
-      betas <- readRDS(file = here("analysis", "mash", popkey[[i]], "inputs", 
-                                 paste0("B_hat_random_df_24000topSNPs_",
+      betas <- readRDS(file = here("analysis", "gwas", "BLUPs", "single_sites",
+                                   "Datatables_single_sites", 
+                                 paste0("B_hat_random_df_33000topSNPs_",
                                         phekey[[j]], "_", popkey[[i]], ".rds")))
-      stderr <- readRDS(file = here("analysis", "mash", popkey[[i]], "inputs", 
-                                 paste0("S_hat_random_df_24000topSNPs_",
+      stderr <- readRDS(file = here("analysis", "gwas", "BLUPs", "single_sites",
+                                   "Datatables_single_sites",  
+                                 paste0("S_hat_random_df_33000topSNPs_",
                                         phekey[[j]], "_", popkey[[i]], ".rds")))
@@
 
-source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
+# source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
 mashdir <- here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar")
 
-numSNPs_v <- c(19000, 19000, 19000, 19000, 24000, 33000)
+numSNPs_v <- c(19000, 19000, 19000, 19000, 33000, 33000)
 suffix_v <- c("FL50_Gulf_and_Midwest", "FL50_Gulf", "GR50_Gulf_and_Midwest",
@@
             here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar",
-                 "midwest_greenup"),
+                 "midwest_inputs"),
             here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar",
@@
 
-#   i = 4
+#   i = 5
 for (i in 1:6) {  # change this for different scripts
@@
       U_single <- U_single %>%
-        filter(`...1` %in% c("MI", "MO", "NE", "OK", "TX1", "TX2", "TX3")) %>%
-        select(`...1`, .data$MI, .data$MO, .data$NE, .data$OK, .data$TX1,
+        filter(`...1` %in% c("MI", "NE", "OK", "TX1", "TX2", "TX3")) %>%
+        select(`...1`, .data$MI, .data$NE, .data$OK, .data$TX1,
                .data$TX2, .data$TX3)
@@
 
-numSNPs_v <- c(19000, 19000, 19000, 19000, 24000, 33000)
+numSNPs_v <- c(19000, 19000, 19000, 19000, 33000, 33000)
 suffix_v <- c("FL50_Gulf_and_Midwest", "FL50_Gulf", "GR50_Gulf_and_Midwest",
@@
 
+```{r}
+get_significant_results(m2)
+print(length(get_significant_results(m2)))
+get_significant_results(m_midw_gr_hyp)
+print(length(get_significant_results(m_midw_gr_hyp)))
+print(length(get_significant_results(m_midw_fl_hyp)))
+print(length(get_significant_results(m_gulf_gr_hyp)))
+print(length(get_significant_results(m_gulf_fl_hyp)))
+print(length(get_significant_results(m_gam_gr_hyp)))
+print(length(get_significant_results(m_gam_fl_hyp)))
+```
+
 ## Figure 2
@@
                             grepl("singleton", `Covariance Matrix`) ~
-                              "canonical",
+                              "Canonical",
                             grepl("simple_het", `Covariance Matrix`) ~
-                              "canonical",
+                              "Canonical",
                             grepl("no_effects", `Covariance Matrix`) ~
-                              "canonical",
+                              "Canonical",
                             grepl("identity", `Covariance Matrix`) ~
-                              "canonical",
+                              "Canonical",
                             grepl("equal_effects", `Covariance Matrix`) ~
-                              "canonical",
+                              "Canonical",
                             grepl("U.mat$", `Covariance Matrix`) ~
-                              "hypothesized",
+                              "GxWeather",
                             TRUE ~ "other"
@@
   pop_sc <- c("all", "gulf", "all", "gulf", "midwest", "midwest")
-  phe_sc <- c("flowering", "flowering", "greenup", "greenup", "greenup",
-              "flowering")
+  phe_sc <- c("flowering", "flowering", "greenup", "greenup", 
+              "greenup", "flowering")
   Upaths <- c(here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar",
@@
               here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar",
-                   "midwest_greenup"),
+                   "midwest_inputs"),
               here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar",
@@
          `Covariance Matrix` = str_replace(`Covariance Matrix`, "no_effects", "No Effects"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "equal_effects", "Equal Effects"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "cgdd_12c_28d.U.mat", "Cumulative GDD (28d)"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "cgdd_12c_1d.U.mat", "Cumulative GDD (1d)"),
          `Covariance Matrix` = str_replace(`Covariance Matrix`, "tave_1d.U.mat", "Temp Ave, 1d"),
@@
     df <- df %>%
-      select(`...1`, TX1, TX2, TX3, OK, MO, NE, MI) 
+      select(`...1`, TX1, TX2, TX3, OK, NE, MI) 
   } else if (mf) {
@@
 Ulist_gf$`Cumulative rainfall, 7d`
-Ulist_mg$dyln_change_sec_7d_prior.U.mat
+Ulist_mg$`Cumulative GDD (28d)`
 ```
@@
 
-pa13 <- plot_covar_Ulist_fig2(Ulist_gf[[1]]) + 
-  labs(subtitle = "identity (Mass <1%)") 
+pa13 <- plot_covar_Ulist_fig2(Ulist_mg[[3]], mg = TRUE) + 
+  labs(subtitle = "Equal Effects") 
+pa14 <- plot_covar_Ulist_fig2(Ulist_mg[[7]], mg = TRUE) + 
+  labs(subtitle = "Cumulative GDD (28d)") 
+pa15 <- plot_covar_Ulist_fig2(Ulist_gf[[1]]) + 
+  labs(subtitle = "Identity (Mass <1%)") 
 pa09 <- plot_covar_Ulist_fig2(Ulist_gf[[5]]) + 
@@
 legend <- get_legend(cov_gr1) 
-fig_eqeff <- plot_covar_Ulist_fig2(Ulist_mg[[2]], mg = TRUE) +
-  labs(subtitle = "simple heterozygousity")
+#fig_eqeff <- plot_covar_Ulist_fig2(Ulist_mg[[2]], mg = TRUE) +
+#  labs(subtitle = "simple heterozygousity")
 
@@
          `Covariance Matrix` = str_replace(`Covariance Matrix`, "no_effects", "No Effects"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "equal_effects", "Equal Effects"),
          `Covariance Matrix` = str_replace(`Covariance Matrix`, "tave_1d.U.mat", " Temp Ave, 1d"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "cgdd_12c_28d.U.mat", "  Cumulative GDD, 28d"),
+         `Covariance Matrix` = str_replace(`Covariance Matrix`, "cgdd_12c_1d.U.mat", "  Cumulative GDD, 1d"),
          `Covariance Matrix` = str_replace(`Covariance Matrix`, "dyln_14d_prior.U.mat", " Daylength (14d prior)"),
@@
 df_2fl_a$subpop <- factor(df_2fl_a$subpop, levels = c("Gulf", "Midwest", "Both"))
-df_2fl_a$Type <- factor(df_2fl_a$Type, levels = c("canonical", "hypothesized"))
-df_2fl_a$Type <- factor(df_2fl_a$Type, levels = c("canonical", "hypothesized"))
+df_2fl_a$Type <- factor(df_2fl_a$Type, levels = c("Canonical", "GxWeather"))
+df_2fl_b$Type <- factor(df_2fl_b$Type, levels = c("Canonical", "GxWeather"))
 
@@
         axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45))
-f2a <- df_2gr_a %>%
-  filter(Mass > 1e-06) %>%
+(f2a <- df_2gr_a %>%
+  filter(Mass > 0.001) %>%
   ggplot(aes(x = `Covariance Matrix`, y = Mass)) + 
@@
         axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45), 
-        panel.spacing.x = unit(-0, 'cm')) #+ labs(x = "")
-f2c <- df_2fl_a %>%
-  filter(Mass > 1e-06) %>%
+        panel.spacing.x = unit(-0, 'cm'))) #+ labs(x = "")
+(f2c <- df_2fl_a %>%
+  filter(Mass > 0.001) %>%
   ggplot(aes(x = `Covariance Matrix`, y = Mass)) + 
@@
         axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45), 
-        panel.spacing.x = unit(-0, 'cm')) #+ labs(x = "")
+        panel.spacing.x = unit(-0, 'cm'))) #+ labs(x = "")
 
@@
 
+Build Figure 2 with covariance matrices as first two rows over covar barplots
+
+```{r}
+
+ggdraw() +
+  theme_oeco + 
+  draw_plot(pa01, x = 0, y = 0.66, width = 0.1, height = 0.18) +
+  draw_plot(pa14, x = 1/6, y = 0.66, width = 0.1, height = 0.18) +
+  #draw_plot(pa03, x = -0.02, y = 0.6, width = 0.1, height = 0.18) +
+  #draw_plot(pa04, x = 0.1, y = 0.6, width = 0.1, height = 0.18) +
+  draw_plot(pa05, x = 2/6, y = 0.66, width = 0.1, height = 0.18) +
+  draw_plot(pa02, x = 0.5, y = 0.66, width = 0.1, height = 0.18) +
+  draw_plot(pa06, x = 4/6, y = 0.66, width = 0.1, height = 0.18) +
+  draw_plot(pa07, x = 5/6, y = 0.66, width = 0.1, height = 0.18) +
+  #draw_plot(pa08, x = 0.2, y = 0, width = 0.1, height = 0.18) +
+  #draw_plot(pa09, x = 0, y = 0.83, width = 0.1, height = 0.18) +
+  draw_plot(pa10, x = 1/6, y = 0.83, width = 0.1, height = 0.18) +
+  draw_plot(pa12, x = 2/6, y = 0.83, width = 0.1, height = 0.18) +
+  draw_plot(pa13, x = 0.5, y = 0.83, width = 0.1, height = 0.18) +
+  draw_plot(pa11, x = 4/6, y = 0.83, width = 0.1, height = 0.18) +
+  draw_plot(pa09, x = 5/6, y = 0.83, width = 0.1, height = 0.18) +
+  draw_plot(f2a, x = 0, y = 0.33, width = 0.66, height = 0.38) +
+  draw_plot(legend, x = 0.1, y = 0.77, width = 0.1, height = 0.2) +
+  draw_plot(f2b, x = 0.66, y = 0.33, width = 0.33, height = 0.38) +
+  draw_plot(f2c, x = 0, y = 0, width = 0.66, height = 0.33) +
+  draw_plot(f2d, x = 0.66, y = 0, width = 0.33, height = 0.33)
+```
+
 ```{r}
@@
 
-save_plot(filename = "../manuscript/revision2/Figure_2_recolor_covariance_matrices_and_posterior_weights.svg", plot = last_plot(), base_height = 5.5, base_width = 8)
+save_plot(filename = "../manuscript/revision2/intermediate_files/Figure_2_six_sites_Midwest.svg", plot = last_plot(), base_height = 7.5, base_width = 6.5)
+
 ```
@@
     left_join(site_regions, by = c("Garden_1" = "Site")) %>%
-    rename("Region_1" = "Region") %>%
+    dplyr::rename("Region_1" = "Region") %>%
     left_join(site_regions, by = c("Garden_2" = "Site")) %>%
-    rename("Region_2" = "Region") %>%
+    dplyr::rename("Region_2" = "Region") %>%
     filter(Garden_1 != Garden_2) %>%
@@
     ungroup() %>%
-    select(-Region_1, -Region_2) %>%
+    dplyr::select(-Region_1, -Region_2) %>%
     group_by(Region_comparison) %>%
@@
   AP <- one_gxe_summary(gxe_list$S_AP, n_cond = n_cond, n_sep = n_sep) %>%
-    rename(Sign = n)
+    dplyr::rename(Sign = n)
   DS <- one_gxe_summary(gxe_list$S_DS, n_cond = n_cond, n_sep = n_sep) %>%
-    rename(Magnitude = n)
+    dplyr::rename(Magnitude = n)
   ND <- one_gxe_summary(gxe_list$S_2_no, n_cond = n_cond, n_sep = n_sep) %>%
-    rename(None = n)
+    dplyr::rename(None = n)
   gxe_n <- AP %>%
@@
 
-get_gxe_summary(gxe_gg) 
+get_gxe_summary(gxe_mf, n_cond = 6) 
 ```
@@
 fig3_gg <- get_gxe_summary(gxe_gg) 
-fig3_mg <- get_gxe_summary(gxe_mg, n_cond = 7) 
+fig3_mg <- get_gxe_summary(gxe_mg, n_cond = 6) 
 fig3_bg <- get_gxe_summary(gxe_gamg, n_sep = 4) 
@@
   draw_plot(fig3_frac_plot, x = 0, y = 0, width = 1, height = .67)
-save_plot(filename = "../manuscript/revision2/Figure_3_Fraction_GxE_by_Region.svg", base_height = 6, base_width = 4.5, plot = last_plot())
+save_plot(filename = "../manuscript/revision2/intermediate_files/Figure_3_Fraction_GxE_by_Region_six_midwest_sites.svg", base_height = 6, base_width = 4.5, plot = last_plot())
 ```
@@
     geom_tile(aes(fill = .data$covar)) +
-    scale_fill_viridis_c(option = "H", limits = c(0, upper)) +
+    scale_fill_viridis_c(option = "A", limits = c(0, upper)) +
     #geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
+    theme_oeco +
     theme(legend.position = "none",
@@
     xlab("") + ylab("") 
-  if(legend == TRUE){
+  if (legend == TRUE) {
     U1_covar <- U1_covar +
-      theme(legend.position = c(0.2, 0.9)) +
+      theme(legend.position = c(0.05, 0.9),
+            axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
       geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey")
@@
 ```{r}
- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_AP, legend = TRUE) + 
-  theme_oeco +
+ pvdiv_plot_cor(corrmatrix = gxe_mf$S_AP, legend = TRUE) + 
   labs(subtitle = " ")
@@
                            upper = 4333) + 
-  theme_oeco +
   labs(title = "Gulf", subtitle = "Sign difference")
 legend <- get_legend(p3legend)
+p3legend_fl <- pvdiv_plot_cor(corrmatrix = gxe_gg$S_AP, legend = TRUE, 
+                           upper = 2500) + 
+  labs(title = "Gulf", subtitle = "Sign difference")
+legend_fl <- get_legend(p3legend_fl)
 
@@
                        upper = 4333) + 
-  theme_oeco +
+  theme(legend.position = c(0.45, 0.9)) +
   labs(subtitle = " ")
@@
                        upper = 4333) + 
-  theme_oeco +
   labs(subtitle = " ")
@@
                        upper = 4333) + 
-  theme_oeco +
   labs(subtitle = "Sign difference")
@@
                        upper = 4333) + 
-  theme_oeco +
   labs(subtitle = "Magnitude difference")
@@
                        upper = 4333) + 
-  theme_oeco +
   labs(subtitle = "Not distinguishable")
@@
 p3d1 <- pvdiv_plot_cor(corrmatrix = gxe_gf$S_AP, legend = FALSE, 
-                       upper = 4333) + 
-  theme_oeco +
+                       upper = 2500) + 
+  theme(legend.position = c(0.45, 0.9))
   labs(subtitle = " ")
 p3d2 <- pvdiv_plot_cor(corrmatrix = gxe_gf$S_DS, legend = FALSE, 
-                       upper = 4333) + 
-  theme_oeco +
+                       upper = 2500) + 
   labs(subtitle = " ")
 p3d3 <- pvdiv_plot_cor(corrmatrix = gxe_gf$S_2_no, legend = FALSE, 
-                       upper = 4333) + 
-  theme_oeco +
+                       upper = 2500) + 
   labs(subtitle = " ")
 p3e1 <- pvdiv_plot_cor(corrmatrix = gxe_mf$S_AP, legend = FALSE, 
-                       upper = 4333) + 
-  theme_oeco +
+                       upper = 2500) + 
   labs(subtitle = "Sign difference")
 p3e2 <- pvdiv_plot_cor(corrmatrix = gxe_mf$S_DS, legend = FALSE, 
-                       upper = 4333) + 
-  theme_oeco +
+                       upper = 2500) + 
   labs(subtitle = "Magnitude difference")
 p3e3 <- pvdiv_plot_cor(corrmatrix = gxe_mf$S_2_no, legend = FALSE, 
-                       upper = 4333) + 
-  theme_oeco +
+                       upper = 2500) + 
   labs(subtitle = "Not distinguishable")
 p3f1 <- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_AP, legend = FALSE, 
-                       upper = 4333) + 
-  theme_oeco +
+                       upper = 2500) + 
   labs(subtitle = " ")
 p3f2 <- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_DS, legend = FALSE, 
-                       upper = 4333) + 
-  theme_oeco +
+                       upper = 2500) + 
   labs(subtitle = " ")
 p3f3 <- pvdiv_plot_cor(corrmatrix = gxe_gamf$S_2_no, legend = FALSE, 
-                       upper = 4333) + 
-  theme_oeco +
+                       upper = 2500) + 
   labs(subtitle = " ")
@@
   draw_plot(p3f3, x = 0.66, y = 0/6, width = 1/3, height = .18) +
-  draw_plot(legend, x = 0.37, y = .38, width = 1/3, height = .18)
+  draw_plot(legend, x = 0.67, y = .68, width = 1/3, height = .18) +
+  draw_plot(legend_fl, x = 0.37, y = .38, width = 1/3, height = .18)
 
@@
 
-save_plot(filename = "../manuscript/Coauthors/PNAS/Revision/Figure_3_Effect_Types_by_Site.svg", plot = last_plot(), base_height = 10, base_width = 5.5)
+save_plot(filename = "../manuscript/revision2/intermediate_files/Figure_S3_S4_Effect_Types_by_Garden.svg", plot = last_plot(), base_height = 10, base_width = 5.5)
 ```
@@
 quantile(pw_gg$log10BayesFactor, 0.99)  # Gulf: Use 2.0989 BF
-quantile(pw_mg$log10BayesFactor, 0.99)  # Midwest: Use 4.6831 BF
+quantile(pw_mg$log10BayesFactor, 0.9998446)  # Midwest: Use 2 BF
 quantile(pw_gamg$log10BayesFactor, 0.99)  # GAM: Use 3.0865 BF
@@
 rug_mg <- pw_mg %>%
-  filter(log10BayesFactor > 4.6831) %>% 
+  filter(log10BayesFactor > 2) %>% 
   mutate(Pos = Pos/1000000, 
@@
 
+library(ggrepel)
 fig_4b <- fig4b +
   geom_text_repel(data = fl_labels, aes(x = pos, y = lod, label = subpop), nudge_y = 2)
-library(ggrepel)
+
 ```
@@
   mutate(subpop = "Gulf", phenotype = "greenup")
-enrich_mg <- get_QTL_enrichment(m_midw_gr_hyp, qtl_df = qtlplot_gr, 0.99) %>%
+enrich_mg <- get_QTL_enrichment(m_midw_gr_hyp, qtl_df = qtlplot_gr, 0.9998446) %>%
   mutate(subpop = "Midwest", phenotype = "greenup")
@@
 
-save_plot(filename = "../manuscript/revision2/Figure_4_QTL_Overlaps.svg", base_height = 9, base_width = 6.5, plot = last_plot())
+save_plot(filename = "../manuscript/revision2/intermediate_files/Figure_4_QTL_Overlaps_6_Midwest_Gardens.svg", base_height = 9, base_width = 6.5, plot = last_plot())
 ```

### 2024-06-26 R/SI_Appendix_Datasets.Rmd
revision2: Drop MO from Midwest GR50 mash analysis due to GWAS pop structure issues

@@
       betas <- readRDS(file = here("analysis", "mash", popkey[[i]], "inputs", 
-                                 paste0("B_hat_random_df_24000topSNPs_",
+                                 paste0("B_hat_random_df_33000topSNPs_",
                                         phekey[[j]], "_", popkey[[i]], ".rds")))
       stderr <- readRDS(file = here("analysis", "mash", popkey[[i]], "inputs", 
-                                 paste0("S_hat_random_df_24000topSNPs_",
+                                 paste0("S_hat_random_df_33000topSNPs_",
                                         phekey[[j]], "_", popkey[[i]], ".rds")))
@@
 switchgrassGWAS:::get_marker_df(m_gulf_fl_hyp) # 50490
-switchgrassGWAS:::get_marker_df(m_midw_gr_hyp) # 49030
+switchgrassGWAS:::get_marker_df(m_midw_gr_hyp) # 57729
 switchgrassGWAS:::get_marker_df(m_midw_fl_hyp) # 63367
@@
 ```{r}
-m <- m_gulf_gr_hyp
+m <- m_midw_gr_hyp
 dataset_creation_function <- function(m, filename){
@@
 overlap <- dataset_8 %>%
-  select(mash_idx, mash_subpop, everything()) |> 
+  dplyr::select(mash_idx, mash_subpop, everything()) |> 
   distinct(mash_idx, .keep_all = TRUE)

### 2024-06-26 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
revision2: Drop MO from Midwest GR50 mash analysis due to GWAS pop structure issues

@@
 Not Distinguishable: Effects not distinguishable by magnitude nor sign of the effect, with no measurable GxE.
-b) The number of effects with each GxE type for the onset of vegetative growth (green-up date) within and between two genetic subpopulations.](images/S2_Greenup.png)
+b) The number of effects with each GxE type for the onset of vegetative growth (green-up date) within and between two genetic subpopulations.](images/Fig_S2.png)
 
@@
 Not Distinguishable: Effects not distinguishable by magnitude nor sign of the effect, with no measurable GxE.
-b) The number of effects with each GxE type for the onset of reproductive growth (green-up date) within and between two genetic subpopulations.](images/S3_Flowering.png)
+b) The number of effects with each GxE type for the onset of reproductive growth (green-up date) within and between two genetic subpopulations.](images/Fig_S3.png)
 
@@
 ![All QTL identified in the outbred mapping population displayed GxE, in that all QTL had effects that differed in at least two gardens.
-Gardens are on the y-axis in latitudinal order; the effects of the significant QTL in Figure 4 are displayed as differences between allelic effects for the two sets of grandparent lines in the outbred mapping panel.
-](images/Figure_S4_QTL_effects_by_trait.png){#fig-qtl-by-env}
+Gardens are on the y-axis in latitudinal order; the effects of the significant QTL in Figure 4 are displayed as differences between allelic effects for the two sets of grandparent lines in the outbred mapping panel.](images/Figure_S4_QTL_effects_by_trait.png){#fig-qtl-by-env}
 
@@
 | Greenup Date   |          Gulf |                                    19000 |                                    38000 | 45931                        | 91862                        |
-|                |       Midwest |                                    24000 |                                    48000 | 49030                        | 98060                        |
+|                |       Midwest |                                    33000 |                                    66000 | 57729                        | 115458                       |
 |                |          Both |                                    19000 |                                    38000 | 44765                        | 89530                        |
@@
 | 1                             |              Identity |                 Identity |             Identity | Identity                | Identity                       | Identity                |
-| 2                             |         Simple Het. 2 |               Single Effect MO |     Single Effect MI | Simple Het. 1           | Simple Het. 1                  | Single Effect TX3       |
-| 3                             |      Single Effect MO |               Single Effect NE |     Single Effect SD | Single Effect MI        | Single Effect MI               | Single Effect TX1       |
-| 4                             |      Single Effect NE |              Single Effect TX2 |     Single Effect NE | Single Effect TX3       | Single Effect TX2              | Single Effect OK        |
-| 5                             | Daylength (14d prior) |              Single Effect TX1 |    Single Effect TX2 | Single Effect SD        | Single Effect MO               | Single Effect MI        |
-| 6                             |      Single Effect MI |               Single Effect MI |     Single Effect OK | Single Effect OK        | Single Effect TX3              | Single Effect NE        |
-| 7                             |      Single Effect SD |               Single Effect OK |    Single Effect TX3 | Single Effect NE        | Daylength change (s), 2d prior | Single Effect TX2       |
-| 8                             |     Single Effect TX2 |              Single Effect TX3 |        Simple Het. 1 | Single Effect TX1       | Single Effect SD               | Single Effect MO        |
-| 9                             |    Cumulative GDD, 6d |                  Simple Het. 1 |    Single Effect TX1 | Single Effect TX2       | \-                             | Single Effect SD        |
-| 10                            |     Single Effect TX3 | Daylength change (s), 7d prior |     Single Effect MO | Single Effect MO        | \-                             | Simple Het. 1           |
+| 2                             |         Simple Het. 2 |         Single Effect OK |     Single Effect MI | Simple Het. 1           | Simple Het. 1                  | Single Effect TX3       |
+| 3                             |      Single Effect MO |            Equal Effects |     Single Effect SD | Single Effect MI        | Single Effect MI               | Single Effect TX1       |
+| 4                             |      Single Effect NE |        Single Effect TX3 |     Single Effect NE | Single Effect TX3       | Single Effect TX2              | Single Effect OK        |
+| 5                             | Daylength (14d prior) |        Single Effect TX2 |    Single Effect TX2 | Single Effect SD        | Single Effect MO               | Single Effect MI        |
+| 6                             |      Single Effect MI |  Cumulative GDD, 12C, 1d |     Single Effect OK | Single Effect OK        | Single Effect TX3              | Single Effect NE        |
+| 7                             |      Single Effect SD | Cumulative GDD, 12C, 28d |    Single Effect TX3 | Single Effect NE        | Daylength change (s), 2d prior | Single Effect TX2       |
+| 8                             |     Single Effect TX2 |                       \- |        Simple Het. 1 | Single Effect TX1       | Single Effect SD               | Single Effect MO        |
+| 9                             |    Cumulative GDD, 6d |                       \- |    Single Effect TX1 | Single Effect TX2       | \-                             | Single Effect SD        |
+| 10                            |     Single Effect TX3 |                       \- |     Single Effect MO | Single Effect MO        | \-                             | Simple Het. 1           |
 | 11                            |   Data-driven $PCA_2$ |                       \- |         Temp Ave, 1d | Cumulative rainfall, 7d | \-                             | Cumulative rainfall, 1d |

### 2024-06-26 manuscript/revision2/revision2.qmd
revision2: Drop MO from Midwest GR50 mash analysis due to GWAS pop structure issues

@@
 
-Eight GxWeather covariance structures were selected by the greedy algorithm, one to two per subpopulation: five for green-up date, and three for flowering date (SI Appendix, Table S2).
-Of these eight, five had mass on them in *mash* models of the strong effects (@fig-covar B,D).
-Three of these five matrices had negative covariances between sites between Texas and North gardens (@fig-covar A), while two had all positive or near-zero covariances.
+Nine GxWeather covariance structures were selected by the greedy algorithm, one to two per subpopulation: six for green-up date, and three for flowering date (SI Appendix, Table S2).
+Of these nine, six had mass (\>0.1%) on them in *mash* models of the strong effects (@fig-covar B,D).
+Five of these six matrices had negative covariances between gardens in the Texas and North regions (@fig-covar A), while one had all positive or near-zero covariances.
 
@@
 The covariance matrix for this weather cue was visually similar to the observed pattern of phenotypic correlation for the timing of vegetative growth in the Gulf subpopulation (@fig-map B; @fig-covar A).
-*Mash* models of the timing of vegetative growth in the Midwest subpopulation and Both subpopulations did not include this GxWeather covariance; the Midwest had no weight on any GxWeather covariance, while Both subpopulations had non-zero weights on two additional GxWeather covariance types, average temperature one day prior to green-up, and the day length change in seconds in the day prior to green-up (@fig-covar B).
+*Mash* models of the timing of vegetative growth in the Midwest subpopulation and Both subpopulations did not include this GxWeather covariance.
+The Midwest had 0.15% weight on a GxWeather covariance matrix of cumulative GDD for the 28d prior to green-up, a GxWeather matrix that was visually similar to the Midwest's observed pattern of phenotypic correlation (@fig-map B).
+Both subpopulations had non-zero weights on two additional GxWeather covariance types, average temperature one day prior to green-up, and the day length change in seconds in the day prior to green-up (@fig-covar B).
 The average temperature covariance matrix had negative covariances between Texas and North gardens, though not as strong as the negative phenotypic correlations seen in Both subpopulations (@fig-map B; @fig-covar A).
-Only the Gulf subpopulation and Both subpopulations had mass on any GxWeather covariance matrices (@fig-covar C) for vegetative growth.
+The Gulf subpopulation and Both subpopulations had substantially more mass on GxWeather covariance matrices than the Midwest population (@fig-covar C) for the onset of vegetative growth.
 
@@
 This means that our tests for a sign change between gardens carry an equal statistical burden to those for effects with the same sign.
-We separated kinds of effects at the level of individual loci into SNP effects that differ in sign between gardens (effects with rank-changing GxE between gardens), SNP effects that differ in magnitude (effects that are large in one garden or region and smaller in others), and SNP effects that are indistinguishable in two gardens (@fig-effects A).
+We separated kinds of effects at the level of individual loci into SNP effects that differ in sign between gardens (effects with rank-changing GxE between gardens), SNP effects that differ in magnitude (effects that are large in one garden or region and smaller in others), and SNP effects that are indistinguishable in two gardens (similarly large or small in comparisons between gardens) (@fig-effects A).
 
@@
 78.7% of pairwise comparisons between North and Texas gardens had a difference in sign, while only 28.6% and 0.2% of North-North or Texas-Texas comparisons had a difference in sign, respectively (@fig-effects B).
-The majority of pairwise effects for greenup for the Midwest (\>55%) and Both (\>85%) subpopulations were the same sign, and effects most often differed in magnitude between gardens within and between the regions (@fig-effects B; SI Appendix, Fig S2).
+The majority of pairwise effects for greenup for the Midwest (\>73%; 68-342 effects) were indistinguishable, or had no GxE.
+The majority of pairwise effects for Both (\>85%) subpopulations were the same sign, and effects most often differed in magnitude between gardens within and between the regions; however, these differences in effect sign and magnitude were mostly between the MO site and other sites (@fig-effects B; SI Appendix, Fig S2).
 
@@
 More effects differed in magnitude between the Texas and North regions than within these regions (42.7% vs \<20%; @fig-effects B), while the majority of effects had no GxE.
-The Midwest population had relatively few significant effects for flowering, but a large proportion of these differed in sign between Texas and North regions (42.7%) or within the North region (65.4%).
+The Midwest population had relatively few significant effects for flowering (0-174 per pairwise comparison), but a large proportion of these differed in sign between Texas and North regions (42.7%) or within the North region (65.4%).
 Finally, in Both subpopulations, less than 20% of pairwise effects differed in sign (@fig-effects B).
@@
 
-Our analysis of the timing of vegetative growth in the Gulf and in Both subpopulations revealed many alleles with rank-changing GxE, or sign changes in their effects between the Texas and North gardens (@fig-effects B).
+Our analysis of the timing of vegetative and reproductive growth revealed many alleles with rank-changing GxE, or sign changes in their effects between the Texas and North gardens (@fig-effects B).
+The Gulf & Both subpopulations had rank-changing GxE for green-up date, while the Midwest subpopulation had rank-changing GxE for flowering date.
 As phenological timings are major components of plant fitness, this result supports theoretical models that local adaptation should involve antagonistic pleiotropy at the level of individual loci [@kawecki2004conceptual; @felsenstein1976theoretical; @levene1953genetic; @hedrick1986genetic].
@@
 Covariance matrices included in *mash* that had zero posterior weight in all three *mash* runs on the genetic subpopulations, such as the identity matrix, are not shown.
-(c,e) Total posterior weight placed on covariance matrices that were hypothesized or canonical, for the (c) green-up date phenotype and (e) flowering date phenotype.](images/Figure_2.png){#fig-covar}
+(c,e) Total posterior weight placed on covariance matrices that were hypothesized or canonical, for the (c) green-up date phenotype and (e) flowering date phenotype.](images/Figure_2_six_sites_Midwest.png){#fig-covar}
 
@@
 b) The fraction of effects with each GxE type for the onset of vegetative growth (green-up date) and reproductive growth (flowering date), within and between two genetic subpopulations.
-Common gardens are grouped by the larger region they came from: North gardens are within the natural range of the Midwest subpopulation, and include MO, NE, MI, and SD, while Texas gardens are within the natural range of the Gulf subpopulation, and include TX1, TX2, and TX3.](images/Figure_3_GxE.png){#fig-effects}
+Common gardens are grouped by the larger region they came from: North gardens are within the natural range of the Midwest subpopulation, and include MO, NE, MI, and SD, while Texas gardens are within the natural range of the Gulf subpopulation, and include TX1, TX2, and TX3.](images/Figure_3_Fraction_GxE_by_Region_six_midwest_sites.png){#fig-effects}
 
@@
 a) QTL mapping for the onset of vegetative growth (green-up date), and three weather-related functions of green-up date.
-b) QTL mapping for the onset of reproductive growth (flowering date), and two weather-related functions of flowering date.](images/Figure_4_QTL_Overlaps.png){#fig-qtl}
+b) QTL mapping for the onset of reproductive growth (flowering date), and two weather-related functions of flowering date.](images/Figure_4_QTL_Overlaps_6_Midwest_Gardens.png){#fig-qtl}
 

### 2024-06-26 manuscript/revision2/revision2.tex
revision2: Drop MO from Midwest GR50 mash analysis due to GWAS pop structure issues

@@
 
-Eight GxWeather covariance structures were selected by the greedy
-algorithm, one to two per subpopulation: five for green-up date, and
-three for flowering date (SI Appendix, Table S2). Of these eight, five
-had mass on them in \emph{mash} models of the strong effects
-(Figure~\ref{fig-covar} B,D). Three of these five matrices had negative
-covariances between sites between Texas and North gardens
-(Figure~\ref{fig-covar} A), while two had all positive or near-zero
+Nine GxWeather covariance structures were selected by the greedy
+algorithm, one to two per subpopulation: six for green-up date, and
+three for flowering date (SI Appendix, Table S2). Of these nine, six had
+mass (\textgreater0.1\%) on them in \emph{mash} models of the strong
+effects (Figure~\ref{fig-covar} B,D). Five of these six matrices had
+negative covariances between gardens in the Texas and North regions
+(Figure~\ref{fig-covar} A), while one had all positive or near-zero
 covariances.
@@
 subpopulation and Both subpopulations did not include this GxWeather
-covariance; the Midwest had no weight on any GxWeather covariance, while
-Both subpopulations had non-zero weights on two additional GxWeather
-covariance types, average temperature one day prior to green-up, and the
-day length change in seconds in the day prior to green-up
-(Figure~\ref{fig-covar} B). The average temperature covariance matrix
-had negative covariances between Texas and North gardens, though not as
-strong as the negative phenotypic correlations seen in Both
-subpopulations (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Only
-the Gulf subpopulation and Both subpopulations had mass on any GxWeather
-covariance matrices (Figure~\ref{fig-covar} C) for vegetative growth.
+covariance. The Midwest had 0.15\% weight on a GxWeather covariance
+matrix of cumulative GDD for the 28d prior to green-up, a GxWeather
+matrix that was visually similar to the Midwest's observed pattern of
+phenotypic correlation (Figure~\ref{fig-map} B). Both subpopulations had
+non-zero weights on two additional GxWeather covariance types, average
+temperature one day prior to green-up, and the day length change in
+seconds in the day prior to green-up (Figure~\ref{fig-covar} B). The
+average temperature covariance matrix had negative covariances between
+Texas and North gardens, though not as strong as the negative phenotypic
+correlations seen in Both subpopulations (Figure~\ref{fig-map} B;
+Figure~\ref{fig-covar} A). The Gulf subpopulation and Both
+subpopulations had substantially more mass on GxWeather covariance
+matrices than the Midwest population (Figure~\ref{fig-covar} C) for the
+onset of vegetative growth.
 
@@
 garden or region and smaller in others), and SNP effects that are
-indistinguishable in two gardens (Figure~\ref{fig-effects} A).
+indistinguishable in two gardens (similarly large or small in
+comparisons between gardens) (Figure~\ref{fig-effects} A).
 
@@
 respectively (Figure~\ref{fig-effects} B). The majority of pairwise
-effects for greenup for the Midwest (\textgreater55\%) and Both
-(\textgreater85\%) subpopulations were the same sign, and effects most
-often differed in magnitude between gardens within and between the
-regions (Figure~\ref{fig-effects} B; SI Appendix, Fig S2).
+effects for greenup for the Midwest (\textgreater73\%; 68-342 effects)
+were indistinguishable, or had no GxE. The majority of pairwise effects
+for Both (\textgreater85\%) subpopulations were the same sign, and
+effects most often differed in magnitude between gardens within and
+between the regions; however, these differences in effect sign and
+magnitude were mostly between the MO site and other sites
+(Figure~\ref{fig-effects} B; SI Appendix, Fig S2).
 
@@
 while the majority of effects had no GxE. The Midwest population had
-relatively few significant effects for flowering, but a large proportion
-of these differed in sign between Texas and North regions (42.7\%) or
-within the North region (65.4\%). Finally, in Both subpopulations, less
-than 20\% of pairwise effects differed in sign (Figure~\ref{fig-effects}
-B). Most differences in sign were between TX1, the southernmost garden,
-and all other gardens (SI Appendix, Figure S3). Similarly, more effects
-that differed in magnitude included gardens in the Texas region
-(52.3-55.9\%), and most effect pairs in the North region were not
-distinguishable (91.5\%).
+relatively few significant effects for flowering (0-174 per pairwise
+comparison), but a large proportion of these differed in sign between
+Texas and North regions (42.7\%) or within the North region (65.4\%).
+Finally, in Both subpopulations, less than 20\% of pairwise effects
+differed in sign (Figure~\ref{fig-effects} B). Most differences in sign
+were between TX1, the southernmost garden, and all other gardens (SI
+Appendix, Figure S3). Similarly, more effects that differed in magnitude
+included gardens in the Texas region (52.3-55.9\%), and most effect
+pairs in the North region were not distinguishable (91.5\%).
 
@@
 
-Our analysis of the timing of vegetative growth in the Gulf and in Both
-subpopulations revealed many alleles with rank-changing GxE, or sign
-changes in their effects between the Texas and North gardens
-(Figure~\ref{fig-effects} B). As phenological timings are major
-components of plant fitness, this result supports theoretical models
-that local adaptation should involve antagonistic pleiotropy at the
-level of individual loci (26--29). In addition, it advances local
-adaptation research by using more than two field sites and a wide range
-of genetic variation to determine that rank-changing GxE for
-phenological timings is common in small genomic regions (14, 43, 44).
+Our analysis of the timing of vegetative and reproductive growth
+revealed many alleles with rank-changing GxE, or sign changes in their
+effects between the Texas and North gardens (Figure~\ref{fig-effects}
+B). The Gulf \& Both subpopulations had rank-changing GxE for green-up
+date, while the Midwest subpopulation had rank-changing GxE for
+flowering date. As phenological timings are major components of plant
+fitness, this result supports theoretical models that local adaptation
+should involve antagonistic pleiotropy at the level of individual loci
+(26--29). In addition, it advances local adaptation research by using
+more than two field sites and a wide range of genetic variation to
+determine that rank-changing GxE for phenological timings is common in
+small genomic regions (14, 43, 44).
 
@@
 
-\includegraphics{images/Figure_2.png}
+\includegraphics{images/Figure_2_six_sites_Midwest.png}
 
@@
 
-\includegraphics{images/Figure_3_GxE.png}
+\includegraphics{images/Figure_3_Fraction_GxE_by_Region_six_midwest_sites.png}
 
@@
 
-\includegraphics{images/Figure_4_QTL_Overlaps.png}
+\includegraphics{images/Figure_4_QTL_Overlaps_6_Midwest_Gardens.png}
 

### 2024-06-26 analysis/mash_greedy_algorithm/mash_h2_diag_covar/midwest_inputs/mash_greedy.R

@@ -1,0 +1,64 @@
+## Run GWAS 
+suppressMessages(library(dplyr))
+suppressMessages(library(bigsnpr))
+suppressMessages(library(readr))
+suppressMessages(library(mashr))
+suppressMessages(library(AnnotationDbi))
+suppressMessages(library(optparse))
+suppressMessages(library(collections))
+suppressMessages(library(bigparallelr))
+suppressMessages(library('matrixcalc'))
+assert_cores(1)
+
+option_list = list(
+  make_option(c("-p","--pheno"), type="character", default=NULL,help="which site to use",metavar="character"),
+  make_option(c("-c","--cohort"), type="character", default=NULL,help="gulf, midwest, or all",metavar="character"),
+  make_option(c("-m","--matrices"), type="character", default=NULL,help="gulf, midwest, or all",metavar="character"),
+  make_option(c("-s","--step"), type="character", default=NULL,help="step in the greedy algorithm",metavar="character"),
+  make_option(c("-e","--effects"), type="character", default=NULL,help="gulf, midwest, or all",metavar="character")
+)
+
+opt_parser = OptionParser(option_list=option_list)
+opt = parse_args(opt_parser)
+
+if (is.null(opt$pheno)) {
+  print_help(opt_parser)
+  stop("Missing argument for phenotype code", call.=FALSE)
+}
+
+pheno = opt$pheno
+cohort = opt$cohort
+matrices = unlist(strsplit(opt$matrices,','))
+effect = opt$effects
+step = opt$step
+
+betas = read.table(paste0('combined_mash_effects/',pheno,'.',cohort,'.betas.',effect,'.txt'))
+stderrs = read.table(paste0('combined_mash_stderrs/',pheno,'.',cohort,'.stderrs.',effect,'.txt'))
+
+mat_betas = matrix(unlist(betas), ncol = 6, byrow = TRUE)
+mat_stderrs = matrix(unlist(stderrs), ncol = 6, byrow = TRUE)
+
+data = mash_set_data(mat_betas, mat_stderrs) 
+
+
+covmats = list()
+labels = list()
+for (matrix in matrices) {
+
+  label=gsub('.csv','',matrix)
+  label = gsub('matrices/','',label)
+  temp = read.csv(matrix,header = TRUE, row.names = 1)
+  temp = matrix(unlist(temp),ncol=6,byrow = TRUE)
+  covmats[[label]] = temp
[…]

### 2024-06-26 analysis/mash_greedy_algorithm/mash_h2_diag_covar/midwest_inputs/pod.greedy.sh

@@ -1,0 +1,1 @@
+python run.mash.greedy.search.py -p greenup -c midwest -e random

### 2024-07-17 manuscript/revision2/response_to_reviews/response_to_reviews.qmd
revision2: Sync cross-references between revision, SI Appendix, and response letter.
Also, add Figures to response letter and change templating to allow this

@@
 ---
-author:
-  - name: Alice MacQueen and Tom Juenger
-    affiliation:
-      - name: The University of Texas at Austin
-        department: Department of Integrative Biology
-        address: 2415 Speedway
-        city: Austin, TX
-        postal-code: 78712
-address:
-  - Editors
-  - Proceedings of the National Academy of Sciences
-subject: manuscript revision
-subject-title: "Subject"
-opening: "Dear Editors,"
-closing: "Best,"
-format: letter-pdf
+format:
+  pdf:
+    keep-tex: true
+    pdfengine: xelatex
+    geometry: 
+      - "top=1in"
+      - "bottom=1in"
+      - "left=1in"
+      - "right=1in"
     code-annotations: hover
-bibliography: references.bib
 ---
+
+::: {.flushright data-latex="1" linestretch="1"}
+The University of Texas at Austin\
+Department of Integrative Biology\
+2415 Speedway\
+Austin, TX 78712
+:::
+
+::: {.flushright data-latex="1" linestretch="1"}
+July 17th, 2024
+:::
+
+::: {.flushleft data-latex="1" linestretch="1"}
+Editors\
+Proceedings of the National Academy of Sciences\
+
+Subject: Manuscript revision
+:::
+
+Dear Editors,
 
@@
 We have addressed all the editorial issues suggested by the two reviewers, and included in this letter are point-by-point responses to these issues.
+
+We suggest that the original editor, Editorial Board members, and anonymous reviewers for this manuscript would be good choices to assess this revision.
+In the event that is not practicable, Douglas Schemske, Stephen P.
+Long, and Qifa Zhang would be good Editorial Board members for this manuscript.
+All would be able to provide strong insights into mapping the genetic basis of adaptive and ecologically relevant traits in a field experimental context.
+Johanna Schmitt, Trudy Mackay, and Detlef Weigel would be excellent editors for this manuscript.
+Dr.
+Schmitt has done extensive work in *Arabidopsis thaliana* and other plant species unravelling the genetic mechanisms involved in responses to the natural environment.
+Dr.
+Mackay has extensive experience mapping the genetics of complex traits and genotype-by-environment interactions.
+Dr.
+Weigel has done extensive work understanding genetic diversity in *A.
+thaliana* including extensive work on floral induction cues.
 
@@
 
-*I'm still reading through these to determine which to add. May need to make additional edits to the Intro/Discussion to add these citations.*
+We now add these citations (lines 47-50; lines 58-61).
+We regret we can't comprehensively reference the impressive crop modeling literature.
+Our goal was to reference work on natural populations that focused the genetics on phenology and/or local adaptation, and to touch on how these ideas have interacted with crop improvement.
+This was because our work uses a natural population of switchgrass with potential relevance for breeding efforts in this species.
 
@@
 
-We agree that given the data we have, the predictive power of these cues cannot be judged - in fact, it is hard to imagine how our GxWeather covariances could be used predictively, as we define the GxWeather covariance matrices retroactively from each genotype's green-up date and flowering date (lines 231 - 235; SI Appendix, Section S1).
+We do see ways we could ultimately use this for prediction - shrinking effects so that they have more accurate GxE would improve prediction, as in ([Zhu *et al.*, 2023](https://doi.org/10.1016/j.xgen.2023.100297)).
+However, prediction is beyond the scope of this manuscript; here we are introducing these ideas and using them to do two specific things: (i) find the GxE structures and (ii) look at the genome-wide prevalence of these structures.
+We feel that using mash to improve predictions is beyond the scope of this paper.
 
@@
 
-We edit the results to directly compare the GxWeather matrices and the phenotypic correlations in Figure 1B. (lines 241-258; lines 269-276; lines 283-287).
+We edit the results to directly compare the GxWeather matrices and the phenotypic correlations in Figure 1B. (lines 241-258; lines 269-276; lines 283-290).
 
@@
 We believe this comment refers to there being higher loadings on the GxWeather matrices for flowering date than there were for the date of the start of vegetative growth in the original paper (original Figure 2).
-Our revised analyses uses a selection process for canonical and GxWeather covariance matrices to select only the subset that significantly improves the model log-likelihood, using a [greedy mash algorithm to iteratively select matrices](#fig-greedy).
+Our revised analyses [uses a greedy algorithm on canonical and GxWeather covariance matrices](#fig-greedy) to iteratively select only the subset that significantly improves the model log-likelihood.
 Thus, the covariance matrices included in each model have changed in this revision, as have the relative loadings on GxWeather and canonical covariance matrices.
 
-Now, posterior weights on the GxWeather matrices vary between 0 to 0.6, but we do not see clear patterns between posterior weights between phenotypes or subpopulations, save that for most models, a large fraction of effects, 0.2 to 0.6, load onto a matrix of 'No Effects' that shrinks effects towards zero.
+We expected patterns of GxE and GxWeather to vary at the locus, subpopulation, and trait level, and we did not see a consistently higher fraction of green-up date mass than flowering date mass on the GxWeather matrices, nor vice versa.
 We don't believe it's appropriate to make direct comparisons between heritability or a variance components analysis and these mash loadings, so we do not do so in the paper; instead of the variance due to additive genetic variation and GxE, these show the additive genetic effects that most closely matched different patterns of GxE and GxWeather, regardless of the proportion of additive variance each locus has.
@@
 
-We think that these patterns need to be assessed individually for different genetic material, environments, and traits (lines 220-223); thus, we develop an approach to specify multiple environmental cues and compete them to explain patterns of genetic effects (lines 230-237): we use a greedy algorithm to use with mash to select covariance patterns, and use mash to flexibly identify these patterns across populations and environments.
+As we state in the SI Appendix (Section S1, paragraph two), we were interested in the types of genetic correlation that could be expressed by different alleles across the genome.
+We reasoned that different loci could have different patterns of genetic correlation.
+For example, the effects of an allele of locus $A$ could covary with a weather-based cue if $A$ had a common allele, $a$, that was responsive to this cue.
+In contrast, the effects of a different locus $B$ might have an effect at only one garden, for example if a pathogen was only present at one garden, and $B$ has a common allele $b$ that is resistant to that pathogen.
+
+As a consequence, we develop an approach to specify multiple environmental cues and compete them to explain patterns of genetic effects (lines 230-237): we use a greedy algorithm to use with mash to select covariance patterns, and use mash to flexibly identify these patterns across populations and environments.
 (lines 238-240).
-
-*Need to add more here to fully address this point*
+With this approach, we are able to demonstrate that we can associate multiple patterns of GxWeather with specific genomic regions (lines 407-410).
+We can also assign genetic effects to both GxWeather patterns derived from weather variables and to other, site-based patterns (lines 410-412).
 
@@
 
-We edit this to clarify our claim.
-NB: We do cite the first experimental work using both QTL mapping and GWAS ([Brachi *et al.,* 2010](https://doi.org/10.1371/journal.pgen.1000940)), and there are many others; we meant using these approaches to map multiple types of GxE.
+We edit the manuscript to remove this unclear claim.
+NB: We do cite the first experimental work using both QTL mapping and GWAS ([Brachi *et al.,* 2010](https://doi.org/10.1371/journal.pgen.1000940)), and there are many others; we meant, more specifically, using these approaches to map multiple types of GxE.
 
@@
 
-*Try and add this reference in the introduction*
+We now reference this paper in the introduction (lines 47-52).
 
@@
 
-Agreed. We add Table 1 to better explain the covariance matrices we select from using the greedy mash algorithm. We include the matrices selected by this algorithm as visualizations in Figure 2.
+Agreed.
+We add Table 1 to better explain the covariance matrices we select using the greedy mash algorithm.
+We include the matrices selected by this algorithm that have \>0.1% weight (lines 262-263) in the mash model as visualizations in Figure 2.
 
@@
 Initially, we thought computational feasibility constrained the number of covariance matrices we could add to the mash model, as the runtime of mash increases with the number of covariance matrices included.
-To address this comment and the mash statistical issue the reviewer points out below, we now specify many more GxWeather covariance matrices with more time frames - 1-7 days prior to the phenological event, and 14, 21, and 28 days prior to the phenological event, generating 48-60 matrices per weather variable.
+To address this comment and the mash statistical issue the reviewer points out below, we now specify many more GxWeather covariance matrices with more time frames - 1-7 days prior to the phenological event, and 14, 21, and 28 days prior to the phenological event, generating 48-60 matrices per weather variable (Table 1).
 Then, we used a greedy mash algorithm, [explained below](#fig-greedy), to select GxWeather covariance matrices from this set that significantly improve the model likelihood.
-We thus extended the time frame for GxWeather covariance that our models could capture; however, no 21 or 28 day matrices were selected by the greedy algorithm.
+We thus extended the time frame for GxWeather covariance that our models could capture; however, only one 28 day matrix was selected by the greedy algorithm and had \>0.1% weight in our *mash* models.
 
-We also now try and explicitly describe the two phenotypes more clearly in the text - both of the traits we map were measured on individuals, but specifically when approximately half of the ramets of the genotype had open flowers (for flowering) or green leaves (for green-up).
+We now also explicitly describe the two phenological traits more clearly in the text - both of the traits we map were measured on individuals, but specifically when approximately half of the ramets of the genotype had open flowers (for flowering) or green leaves (for green-up).
 Thus we are looking for cues driving vegetative and reproductive transitions for a majority of ramets, not for heading (flower emergence on the panicle).
-While there may be physiological mechanisms driving flower opening once the transition to reproductive development has been made, our results show that there is also genetic variation underlying differences in timing of the physiological mechanisms underlying flower opening.
+While there may be physiological mechanisms driving flower opening once the transition to reproductive development has been made, our results show that genetic variation underlies these differences in the timing of the physiological mechanisms underlying flower opening.
 
@@
 Third, population structure could certainly bias the resultant mash models.
-We describe the steps we take to prevent "garbage in" to our mash models, including removing conditions where GWAS had significant population structure.
-However, it is our intuition that, should biases remain in our models, as they undoubtedly do, mash would not load these biased effects onto our GxWeather covariance matrices - instead, we these effects could perhaps load onto canonical matrices, such as garden-specific effects, or perhaps onto data-driven matrices.
-Thus, we expect our assessment of the proportion of the genome that shows GxWeather effects to be conservative if there is residual population structure.
+We describe the steps we take to prevent "garbage in" to our mash models, including removing conditions where GWAS had significant population structure (SI Appendix, Section S2, last paragraph).
+However, it is our intuition that, should biases remain in our models, as they undoubtedly do, mash would not load these biased effects onto our GxWeather covariance matrices - instead, these effects likely load onto canonical matrices, such as garden-specific effects, or potentially onto data-driven matrices.
+Thus, we expect our assessment of the proportion of the genome that shows GxWeather effects to be conservative if our models include residual population structure.
 
@@
 To clarify our interpretation of the posterior matrix weights provided by mash, we performed an extensive analysis of the performance of mash models when different covariance matrices are included.
-Specifically, we implemented a model selection approach that uses a greedy algorithm to evaluate the log likelihood of the mash model as additional covariance matrices were included (see Methods lines \_-\_ and pseudo code below).
+Specifically, we implemented a model selection approach that uses a greedy algorithm to evaluate the log likelihood of the mash model as additional covariance matrices were included (see Methods lines 899-906, SI Appendix Section S4, and pseudo code below).
 This is a similar, but more computationally efficient design, than the leave-one-out approach recommended by the reviewer as the runtime of mash increases with the number of covariance matrices included.
@@
 Importantly, this removes the potential for arbitrary assignment of weights to the covariance matrices in the model and allows for the selection of matrices that most accurately capture the underlying biology captured by the site-specific effect sizes.
-We applied this algorithm to only the hypothesis matrices for each phenotype (Figures 1 and 2) using the previously used sets of significantly associated and randomly selected variants in the Gulf subpopulation individuals, Midwest subpopulation individuals, and the two subpopulations combined.
-The results indicate that only a small number of the original hypothesis matrices (between three and five) are necessary to reach the maximum likelihood model when using the same likelihood ratio test implemented in Urbut et al., indicating that correlation among hypothesis covariance matrices is high.
-This redundancy was confirmed through our application of the greedy algorithm to a combined set of canonical, data driven, and hypothesis covariance matrices (Figures 3 and 4) Our primary conclusion from this analysis is in line with what the reviewer posited, that there is extensive redundancy among canonical, data driven, and hypothesis covariance matrices that is leading to biologically informative patterns of covariance between site effect sizes to be arbitrarily distributed among them.
+We applied this algorithm to only the GxWeather covariance matrices for each phenotype (Figures 1 and 2) using the previously used sets of significantly associated and randomly selected variants in the Gulf subpopulation individuals, Midwest subpopulation individuals, and the two subpopulations combined.
+The results indicate that only a small number of the original hypothesis matrices (between three and five) are necessary to reach the maximum likelihood model when using the same likelihood ratio test implemented in Urbut *et al*., indicating that correlation among hypothesis covariance matrices is high.
+This redundancy was confirmed through our application of the greedy algorithm to a combined set of canonical, data driven, and GxWeather covariance matrices (Figures 3 and 4) Our primary conclusion from this analysis is in line with what the reviewer posited, that there is extensive redundancy among canonical, data driven, and hypothesis covariance matrices that is leading to biologically informative patterns of covariance between site effect sizes to be arbitrarily distributed among them.
 We additionally found evidence that while the correlations between data driven and environmentally informed (hypothesis) covariance matrices, there is extensive similarity in the matrices that are included in the maximum likelihood models across both phenotypes and in the analysis of the Gulf subpopulation, Midwest subpopulations, and the combined cohort.
@@
 
+![**Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the green-up date phenotype using only GxWeather covariance matrices.** The matrix that resulted in the maximum likelihood model of covariance in summary statistics among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis.
+(A) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals.
+The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
+(B) and (C) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.
+(D) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals.
+The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
+\* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants.
+(E) and (F) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.](images/RR_Figure1.png)
+
+![**Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the flowering date phenotype using only GxWeather covariance matrices.** The matrix that resulted in the maximum likelihood model among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis.
+(A) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals.
+The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
+(B) and (C) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.
+(D) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals.
+The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
+\* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants.
+(E) and (F) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.](images/RR_Figure2.png)
+
+![**Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the greenup date phenotype using canonical, data driven, and GxWeather covariance matrices.** The matrix that resulted in the maximum likelihood model of covariance in summary statistics among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis.
+(A) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals.
+The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
+Once this condition was met the greedy algorithm was halted.
+(B) and (C) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.
+(D) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals.
+The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
+Once this condition was met the greedy algorithm was halted for computational considerations.
+\* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants.
+(E) and (F) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.](images/RR_Figure3.png)
+
+![**Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the flowering phenotype using canonical, data driven, and GxWeather covariance matrices.** The matrix that resulted in the maximum likelihood model of covariance in summary statistics among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis.
+(A) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals.
+The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
+Once this condition was met the greedy algorithm was halted.
+(B) and (C) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.
+(D) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals.
+The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
+Once this condition was met the greedy algorithm was halted for computational considerations.
+\* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants.
+(E) and (F) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.](images/RR_Figure4.png)
+
 > ::: {.callout-warning appearance="minimal"}
@@
 
-We believe that the use of the lfsr to detect effects with rank-changing GxE is a key advance in our manuscript, and so we have made changes to the Introduction, Results, and Materials & Methods (lines 957-997; lines 1055-1056), as well as the SI Appendix, to explain and motivate this statistical change.
+We believe that the use of the lfsr to detect effects with rank-changing GxE is a key advance in our manuscript, and so we have made changes to the Introduction (lines 169-171), Results (lines 319-326), and Materials & Methods (lines 957-997; lines 1055-1056) to explain and motivate this statistical change.
 
 Our expectation is that in nature, effects will always differ.
-Statisically, this difference will not always be detectable.
+Statistically, this difference will not always be detectable.
 Thus, rather than ask "Are these two effects different?" - as we reasonably expect two effects to be, even if this difference cannot be measured - the local false sign rate answers a more meaningful question: Can we be confident in the sign of this effect?
 
-We think that if evolutionary geneticists are interested in antagonistic pleiotropy and rank-changing GxE, then they should use the lfsr to measure their confidence in the sign of the effect, rather than the lfdr to measure their confidence that the effect is different than zero.
-In the lfsr case, comparisons between antagonistic pleiotropy and conditional neutrality are no longer sensible, as we are no longer doing a statistical test that can robustly detect conditional neutrality (just as the lfdr does not robustly detect antagonistic pleiotropy).
+We think that if evolutionary geneticists are interested in antagonistic pleiotropy and rank-changing GxE, then they should use the lfsr to measure their confidence in the sign of the effect, rather than using statistical tests where the null hypothesis is that the effect is different than zero.
+However, if we use the lfsr, comparisons between antagonistic pleiotropy and conditional neutrality are no longer sensible, as we are no longer doing a statistical test that can robustly detect conditional neutrality (just as the FDR does not robustly detect antagonistic pleiotropy).
+This is justified because detection of antagonistic pleiotropy is more important than detection of conditional neutrality for models of local adaptation.
 
@@
 We have reworded the manuscript to stress that there is equal power to detect effects of different sign as there are effects of the same sign (lines 994-997; lines 1055-1056).
-Another class of effects we have in our analysis, that differ from classic antagonistic pleiotropy & conditional neutrality comparisons are effects which are not distinguishable by sign or magnitude.
+Another class of effects we have in our analysis that differ from classic antagonistic pleiotropy & conditional neutrality comparisons are effects which are not distinguishable by sign or magnitude.
 It is not accurate that differentially sensitive loci will be more likely to pass the lfsr, because the loci are first selected by significant lfsr, then divided further into effects that can & cannot be distinguished by magnitude.
 Thus effects that are significant, but have small magnitude in both conditions will be labeled as not being distinguishable - essentially, these effects do not have detectable GxE.
+Effects without GxE may be even more common and should also be quantified.
 We propose that, if our intent is to detect sign-changing GxE, then we should use an unbiased statistical test to look at loci with and without sign-changing GxE, the lfsr.
+
+::: {.center data-latex="1"}
+Best,
+:::
+
+::: {.flushright data-latex="1"}
+Alice MacQueen and Tom Juenger
+:::

### 2024-07-17 manuscript/revision2/response_to_reviews/response_to_reviews.tex
revision2: Sync cross-references between revision, SI Appendix, and response letter.
Also, add Figures to response letter and change templating to allow this

--- 2024-07-17 manuscript/revision2/response_to_reviews/response_to_reviews.qmd
@@
----
-format:
-  pdf:
-    keep-tex: true
-    pdfengine: xelatex
-    geometry: 
-      - "top=1in"
-      - "bottom=1in"
-      - "left=1in"
-      - "right=1in"
-    code-annotations: hover
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+%
+\documentclass[
+  letterpaper,
+  DIV=11,
+  numbers=noendperiod]{scrartcl}
 
-::: {.flushright data-latex="1" linestretch="1"}
-The University of Texas at Austin\
-Department of Integrative Biology\
-2415 Speedway\
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage[top=1in,bottom=1in,left=1in,right=1in]{geometry}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{241,243,245}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
+\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+
+\KOMAoption{captions}{tableheading}
+\makeatletter
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
+\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
+\definecolor{quarto-callout-color}{HTML}{909090}
+\definecolor{quarto-callout-note-color}{HTML}{0758E5}
+\definecolor{quarto-callout-important-color}{HTML}{CC1914}
+\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
+\definecolor{quarto-callout-tip-color}{HTML}{00A047}
+\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
+\definecolor{quarto-callout-color-frame}{HTML}{acacac}
+\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
+\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
+\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
+\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
+\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\AtBeginDocument{%
+\ifdefined\contentsname
+  \renewcommand*\contentsname{Table of contents}
+\else
+  \newcommand\contentsname{Table of contents}
+\fi
+\ifdefined\listfigurename
+  \renewcommand*\listfigurename{List of Figures}
+\else
+  \newcommand\listfigurename{List of Figures}
+\fi
+\ifdefined\listtablename
+  \renewcommand*\listtablename{List of Tables}
+\else
+  \newcommand\listtablename{List of Tables}
+\fi
+\ifdefined\figurename
+  \renewcommand*\figurename{Figure}
+\else
+  \newcommand\figurename{Figure}
+\fi
+\ifdefined\tablename
+  \renewcommand*\tablename{Table}
+\else
+  \newcommand\tablename{Table}
+\fi
+}
+\@ifpackageloaded{float}{}{\usepackage{float}}
+\floatstyle{ruled}
+\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
+\floatname{codelisting}{Listing}
+\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\makeatother
+\makeatletter
+\@ifpackageloaded{tikz}{}{\usepackage{tikz}}
+\makeatother
+        \newcommand*\circled[1]{\tikz[baseline=(char.base)]{
+          \node[shape=circle,draw,inner sep=1pt] (char) {{\scriptsize#1}};}}
+
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\usepackage{bookmark}
+
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  colorlinks=true,
+  linkcolor={blue},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={Blue},
+  pdfcreator={LaTeX via pandoc}}
+
+\author{}
+\date{}
+
+\begin{document}
+
+\begin{flushright}
+The University of Texas at Austin\\
+Department of Integrative Biology\\
+2415 Speedway\\
 Austin, TX 78712
-:::
 
-::: {.flushright data-latex="1" linestretch="1"}
+\end{flushright}
+
+~
+
+\begin{flushright}
 July 17th, 2024
-:::
 
-::: {.flushleft data-latex="1" linestretch="1"}
-Editors\
-Proceedings of the National Academy of Sciences\
+\end{flushright}
+
+~
+
+\begin{flushleft}
+Editors\\
+Proceedings of the National Academy of Sciences\\
 
 Subject: Manuscript revision
-:::
+
+\end{flushleft}
+
+~
 
@@
 
-Thank you for considering this work for publication in *PNAS*.
-Our manuscript, "Diverse Genotype-by-Weather Interactions in Switchgrass" has undergone substantial revision from our original submission in 2021.
-This timeframe is long - Dr.
-MacQueen took a new position in 2022.
-They continued working with the group as we developed a new algorithm to address major reviewer concerns and significantly altered the manuscript.
-Because of the major alterations over this time frame, the response to reviewers does not have track changes for each point, though we do point to line numbers for the associated revisions in our response.
+Thank you for considering this work for publication in \emph{PNAS}. Our
+manuscript, ``Diverse Genotype-by-Weather Interactions in Switchgrass''
+has undergone substantial revision from our original submission in 2021.
+This timeframe is long - Dr.~MacQueen took a new position in 2022. They
+continued working with the group as we developed a new algorithm to
+address major reviewer concerns and significantly altered the
+manuscript. Because of the major alterations over this time frame, the
+response to reviewers does not have track changes for each point, though
+we do point to line numbers for the associated revisions in our
+response.
 
-We were pleased by the useful feedback given by the reviewers of our original submission and believe that the greedy mash algorithm we implemented addresses the major statistical concern of reviewer 2, as well as substantially improving our use and interpretation of covariance matrices in *mash*.
-We have addressed all the editorial issues suggested by the two reviewers, and included in this letter are point-by-point responses to these issues.
+We were pleased by the useful feedback given by the reviewers of our
+original submission and believe that the greedy mash algorithm we
+implemented addresses the major statistical concern of reviewer 2, as
+well as substantially improving our use and interpretation of covariance
+matrices in \emph{mash}. We have addressed all the editorial issues
+suggested by the two reviewers, and included in this letter are
+point-by-point responses to these issues.
 
-We suggest that the original editor, Editorial Board members, and anonymous reviewers for this manuscript would be good choices to assess this revision.
-In the event that is not practicable, Douglas Schemske, Stephen P.
-Long, and Qifa Zhang would be good Editorial Board members for this manuscript.
-All would be able to provide strong insights into mapping the genetic basis of adaptive and ecologically relevant traits in a field experimental context.
-Johanna Schmitt, Trudy Mackay, and Detlef Weigel would be excellent editors for this manuscript.
-Dr.
-Schmitt has done extensive work in *Arabidopsis thaliana* and other plant species unravelling the genetic mechanisms involved in responses to the natural environment.
-Dr.
-Mackay has extensive experience mapping the genetics of complex traits and genotype-by-environment interactions.
-Dr.
-Weigel has done extensive work understanding genetic diversity in *A.
-thaliana* including extensive work on floral induction cues.
+We suggest that the original editor, Editorial Board members, and
+anonymous reviewers for this manuscript would be good choices to assess
+this revision. In the event that is not practicable, Douglas Schemske,
+Stephen P. Long, and Qifa Zhang would be good Editorial Board members
+for this manuscript. All would be able to provide strong insights into
+mapping the genetic basis of adaptive and ecologically relevant traits
+in a field experimental context. Johanna Schmitt, Trudy Mackay, and
+Detlef Weigel would be excellent editors for this manuscript.
+Dr.~Schmitt has done extensive work in \emph{Arabidopsis thaliana} and
+other plant species unravelling the genetic mechanisms involved in
+responses to the natural environment. Dr.~Mackay has extensive
+experience mapping the genetics of complex traits and
+genotype-by-environment interactions. Dr.~Weigel has done extensive work
+understanding genetic diversity in \emph{A. thaliana} including
+extensive work on floral induction cues.
 
-> ::: {.callout-warning appearance="minimal"}
-> Reviewer #1:\
-> \
-> Suitable Quality?: No\
-> Sufficient General Interest?: No\
-> Conclusions Justified?: Yes\
-> Clearly Written?: No\
-> Procedures Described?: Yes\
-> \
-> Comments on Significance Statement:\
-> \
-> Conclusions are very specific for the crop/population and traits. General conclusions are lacking.\
-> \
-> Comments:\
-> \
-> The paper by MacQueen et al.
-presents results of a study on two switchgrass populations grown in 8 garden experiments across a rane of latitudes.
-Traits are green-up date and flowering time.
-The data show strong GxE, effects of QTL changing in magnitude and sign.
-Results are partly validated in an independent segregating population.\
-> \
-> The topic is timely and highly relevant.
-A better understanding of GxE will be valuable for plant genetics, evolution and breeding.
-The data are complex and the authors have succeeded in making them accessible to the reader.\
-> \
-> \
-> Relevant literature\
-> The introduction and discussion are focused on findings in Arabidopsis and switchgrass. An excellent study from Arabidopsis with high relevance for this paper is\
-> Fournier-Level et al. 2016 <https://doi.org/10.1073/pnas.1517456113>\
-> \
-> Major findings from other crop species have been ignored. To mention just two references:\
-> \
-> Bustos-Korts et al. 2019 The Plant Journal <https://doi.org/10.1111/tpj.14414>\
-> Millet et al. 2016 Plant Physiology <https://doi.org/10.1104/pp.16.00621> and\
-> There are many papers on crop modelling (eg work by Mark Cooper and others in maize) and the authors might want to consider linking their results to crop modelling to demonstrate the relevance of the results of this study
-> :::
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
 
-We now add these citations (lines 47-50; lines 58-61).
-We regret we can't comprehensively reference the impressive crop modeling literature.
-Our goal was to reference work on natural populations that focused the genetics on phenology and/or local adaptation, and to touch on how these ideas have interacted with crop improvement.
-This was because our work uses a natural population of switchgrass with potential relevance for breeding efforts in this species.
+Reviewer \#1:\\
+\strut \\
+Suitable Quality?: No\\
+Sufficient General Interest?: No\\
+Conclusions Justified?: Yes\\
+Clearly Written?: No\\
+Procedures Described?: Yes\\
+\strut \\
+Comments on Significance Statement:\\
+\strut \\
+Conclusions are very specific for the crop/population and traits.
+General conclusions are lacking.\\
+\strut \\
+Comments:\\
+\strut \\
+The paper by MacQueen et al.~presents results of a study on two
+switchgrass populations grown in 8 garden experiments across a rane of
+latitudes. Traits are green-up date and flowering time. The data show
+strong GxE, effects of QTL changing in magnitude and sign. Results are
+partly validated in an independent segregating population.\\
+\strut \\
+The topic is timely and highly relevant. A better understanding of GxE
+will be valuable for plant genetics, evolution and breeding. The data
+are complex and the authors have succeeded in making them accessible to
+the reader.\\
+\strut \\
+\strut \\
+Relevant literature\\
+The introduction and discussion are focused on findings in Arabidopsis
+and switchgrass. An excellent study from Arabidopsis with high relevance
+for this paper is\\
+Fournier-Level et al.~2016
+\url{https://doi.org/10.1073/pnas.1517456113}\\
+\strut \\
+Major findings from other crop species have been ignored. To mention
+just two references:\\
+\strut \\
+Bustos-Korts et al.~2019 The Plant Journal
+\url{https://doi.org/10.1111/tpj.14414}\\
+Millet et al.~2016 Plant Physiology
+\url{https://doi.org/10.1104/pp.16.00621} and\\
+There are many papers on crop modelling (eg work by Mark Cooper and
+others in maize) and the authors might want to consider linking their
+results to crop modelling to demonstrate the relevance of the results of
+this study
 
-> ::: {.callout-warning appearance="minimal"}
-> Results\
-> \
-> There is a large overlap of data and findings with reference 32 Lovell et al.
-2021 Nature.
-It has not become clear which additional insights can be gained from this study.
-One reason might be that objectives of the study have remained rather vague as well as the generic conclusions that can be drawn from the results.
-> :::
+\end{tcolorbox}
+\end{quote}
 
-While we do rely on the same genotypic data and common gardens used in [Lovell *et al.,* 2021](https://doi.org/10.1038/s41586-020-03127-1), and make use of the population structure findings of [Lovell *et al.,* 2021](https://doi.org/10.1038/s41586-020-03127-1) to divide our genotypes into subpopulations for GWAS, the previous work considered fitness proxies and this work considers phenological traits.
-This work has two additional, key advances over the previous work: it can assign loci genome-wide to multiple kinds of GxE and GxWeather, and has an unbiased statistical method to identify loci with and without rank-changing GxE.
-We now summarize the findings from [Lovell *et al.,* 2021](https://doi.org/10.1038/s41586-020-03127-1) that we use as a springboard for this work on lines 136-145, and outline the additional advances of this study on lines 156-163, 169-171, and 883-893.
+We now add these citations (lines 47-50; lines 58-61). We regret we
+can't comprehensively reference the impressive crop modeling literature.
+Our goal was to reference work on natural populations that focused the
+genetics on phenology and/or local adaptation, and to touch on how these
+ideas have interacted with crop improvement. This was because our work
+uses a natural population of switchgrass with potential relevance for
+breeding efforts in this species.
 
-> ::: {.callout-warning appearance="minimal"}
-> In the introduction the aim is given as „we test if these populations differ in their phenological adaptation and hence their phenological GxE".
-Given prior knowledge such as the origin of the populations and earlier results (Lovell et al.) what is expected?
-> :::
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
 
-We now edit the Introduction to more clearly articulate the aims of this manuscript.
-Our key expectation is that different genetic subpopulations and genomic regions have likely evolved distinct patterns of GxE (lines 55-58; lines 219-223).
-Thus, we aim to identify the kinds of GxE present for each subpopulation and each trait.
-(lines 129-131).
+Results\\
+\strut \\
+There is a large overlap of data and findings with reference 32 Lovell
+et al.~2021 Nature. It has not become clear which additional insights
+can be gained from this study. One reason might be that objectives of
+the study have remained rather vague as well as the generic conclusions
+that can be drawn from the results.
 
-> ::: {.callout-warning appearance="minimal"}
-> The results for these two populations did not show a general pattern leading to conclusions on the environmental cues.
-Associations trait/cue changed across populations, traits, gardens, analysis etc.
-, it was difficult to see the big picture.
-> :::
+\end{tcolorbox}
+\end{quote}
 
-We agree that we do not identify one general pattern or one environmental cue affecting any set of populations or phenological trait.
-Instead, our key expectation is that different populations and different genetic loci will have evolved distinct patterns of GxE (lines 55-58; lines 219-223; SI Appendix Section S1).
-Thus, we aim to identify the kinds of GxE present genome-wide and elaborate when these kinds of GxE differ (Figure 2), and demonstrate that individual loci have different types of GxE across environmental regions and subpopulations (Figure 3).
+While we do rely on the same genotypic data and common gardens used in
+\href{https://doi.org/10.1038/s41586-020-03127-1}{Lovell \emph{et al.,}
+2021}, and make use of the population structure findings of
+\href{https://doi.org/10.1038/s41586-020-03127-1}{Lovell \emph{et al.,}
+2021} to divide our genotypes into subpopulations for GWAS, the previous
+work considered fitness proxies and this work considers phenological
+traits. This work has two additional, key advances over the previous
+work: it can assign loci genome-wide to multiple kinds of GxE and
+GxWeather, and has an unbiased statistical method to identify loci with
+and without rank-changing GxE. We now summarize the findings from
+\href{https://doi.org/10.1038/s41586-020-03127-1}{Lovell \emph{et al.,}
+2021} that we use as a springboard for this work on lines 136-145, and
+outline the additional advances of this study on lines 156-163, 169-171,
+and 883-893.
 
-> ::: {.callout-warning appearance="minimal"}
-> The authors claim that the environmental cues in the hypothesis based models improved model fit.
-What inference is possible from this result?
-Unless results are validated in independent data or with cross validation the predictive power of the cues cannot be judged.
-> :::
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
 
-We edit the Results to (i) change our analysis add an algorithm to more rigorously demonstrate this claim given the data that we have, and (ii) more clearly articulate this claim.
-We now employ a greedy mash algorithm to iteratively add covariance matrices to the *mash* model, only adding matrices that significantly improve the mash model log-likelihood, as we elaborate on further below in our [response to reviewer 2](#fig-greedy).
+In the introduction the aim is given as „we test if these populations
+differ in their phenological adaptation and hence their phenological
+GxE''. Given prior knowledge such as the origin of the populations and
+earlier results (Lovell et al.) what is expected?
 
-We do see ways we could ultimately use this for prediction - shrinking effects so that they have more accurate GxE would improve prediction, as in ([Zhu *et al.*, 2023](https://doi.org/10.1016/j.xgen.2023.100297)).
-However, prediction is beyond the scope of this manuscript; here we are introducing these ideas and using them to do two specific things: (i) find the GxE structures and (ii) look at the genome-wide prevalence of these structures.
-We feel that using mash to improve predictions is beyond the scope of this paper.
+\end{tcolorbox}
+\end{quote}
 
-> ::: {.callout-warning appearance="minimal"}
-> Results from the SNPs "mash model of Midwest green-up fell on a covariance matrix of average temperature in the 10 days prior to Midwest green-up".
-These results should be linked to the phenotypic GxE analysis.
-> :::
+We now edit the Introduction to more clearly articulate the aims of this
+manuscript. Our key expectation is that different genetic subpopulations
+and genomic regions have likely evolved distinct patterns of GxE (lines
+55-58; lines 219-223). Thus, we aim to identify the kinds of GxE present
+for each subpopulation and each trait. (lines 129-131).
 
-We edit the results to directly compare the GxWeather matrices and the phenotypic correlations in Figure 1B. (lines 241-258; lines 269-276; lines 283-290).
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
 
-> ::: {.callout-warning appearance="minimal"}
-> Flowering posterior weights were higher than green-up weights. What does that mean? How does that link to the different heritabilities and levels of GxE?
-> :::
+The results for these two populations did not show a general pattern
+leading to conclusions on the environmental cues. Associations trait/cue
+changed across populations, traits, gardens, analysis etc. , it was
+difficult to see the big picture.
 
-We believe this comment refers to there being higher loadings on the GxWeather matrices for flowering date than there were for the date of the start of vegetative growth in the original paper (original Figure 2).
-Our revised analyses [uses a greedy algorithm on canonical and GxWeather covariance matrices](#fig-greedy) to iteratively select only the subset that significantly improves the model log-likelihood.
-Thus, the covariance matrices included in each model have changed in this revision, as have the relative loadings on GxWeather and canonical covariance matrices.
+\end{tcolorbox}
+\end{quote}
 
-We expected patterns of GxE and GxWeather to vary at the locus, subpopulation, and trait level, and we did not see a consistently higher fraction of green-up date mass than flowering date mass on the GxWeather matrices, nor vice versa.
-We don't believe it's appropriate to make direct comparisons between heritability or a variance components analysis and these mash loadings, so we do not do so in the paper; instead of the variance due to additive genetic variation and GxE, these show the additive genetic effects that most closely matched different patterns of GxE and GxWeather, regardless of the proportion of additive variance each locus has.
+We agree that we do not identify one general pattern or one
+environmental cue affecting any set of populations or phenological
+trait. Instead, our key expectation is that different populations and
+different genetic loci will have evolved distinct patterns of GxE (lines
+55-58; lines 219-223; SI Appendix Section S1). Thus, we aim to identify
+the kinds of GxE present genome-wide and elaborate when these kinds of
+GxE differ (Figure 2), and demonstrate that individual loci have
+different types of GxE across environmental regions and subpopulations
+(Figure 3).
 
-> ::: {.callout-warning appearance="minimal"}
-> What can you conclude from the pattern of antagonistic pleiotropy between Texas and Northern gardens that has not been seen in the GxE analysis?
-> :::
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
 
-No conclusions about individual loci effects can be drawn from the GxE analysis in Figure 2, as this plot does not show how shrinkage changes the jointly re-estimated SNP effect sizes.
-The GxE analysis in Figure 2 shows the types of GxE and GxWeather present genome-wide, from loadings on individual loci, but does not show the effects of these loci on traits.
-Rather, it is an overall characterization of GxE across all eight gardens.
+The authors claim that the environmental cues in the hypothesis based
+models improved model fit. What inference is possible from this result?
+Unless results are validated in independent data or with cross
+validation the predictive power of the cues cannot be judged.
 
-The GxE analysis in Figure 3 shows the re-estimated effects of these loci on traits as contrasts between individual pairs of gardens.
-From Figure 3, you can conclude that many loci have effects with antagonistic pleiotropy at the site level.
+\end{tcolorbox}
+\end{quote}
 
-> ::: {.callout-warning appearance="minimal"}
-> Discussion\
-> "we must understand the current patterns of trait covariation across environments, .....".
-What is the understanding from your data?
-Are there any general patterns or do we need to assess them individually for different genetic material, environments, traits?
-It looks like the latter so what are the consequences?
-> :::
+We edit the Results to (i) change our analysis add an algorithm to more
+rigorously demonstrate this claim given the data that we have, and (ii)
+more clearly articulate this claim. We now employ a greedy mash
+algorithm to iteratively add covariance matrices to the \emph{mash}
+model, only adding matrices that significantly improve the mash model
+log-likelihood, as we elaborate on further below in our
+\hyperref[fig-greedy]{response to reviewer 2}.
 
-As we state in the SI Appendix (Section S1, paragraph two), we were interested in the types of genetic correlation that could be expressed by different alleles across the genome.
-We reasoned that different loci could have different patterns of genetic correlation.
-For example, the effects of an allele of locus $A$ could covary with a weather-based cue if $A$ had a common allele, $a$, that was responsive to this cue.
-In contrast, the effects of a different locus $B$ might have an effect at only one garden, for example if a pathogen was only present at one garden, and $B$ has a common allele $b$ that is resistant to that pathogen.
+We do see ways we could ultimately use this for prediction - shrinking
+effects so that they have more accurate GxE would improve prediction, as
+in (\href{https://doi.org/10.1016/j.xgen.2023.100297}{Zhu \emph{et al.},
+2023}). However, prediction is beyond the scope of this manuscript; here
+we are introducing these ideas and using them to do two specific things:
+(i) find the GxE structures and (ii) look at the genome-wide prevalence
+of these structures. We feel that using mash to improve predictions is
+beyond the scope of this paper.
 
-As a consequence, we develop an approach to specify multiple environmental cues and compete them to explain patterns of genetic effects (lines 230-237): we use a greedy algorithm to use with mash to select covariance patterns, and use mash to flexibly identify these patterns across populations and environments.
-(lines 238-240).
-With this approach, we are able to demonstrate that we can associate multiple patterns of GxWeather with specific genomic regions (lines 407-410).
-We can also assign genetic effects to both GxWeather patterns derived from weather variables and to other, site-based patterns (lines 410-412).
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
 
-> ::: {.callout-warning appearance="minimal"}
-> "this is the first experimental work using QTL mapping and GWAS across ....". I disagree, see references above and others.
-> :::
+Results from the SNPs ``mash model of Midwest green-up fell on a
+covariance matrix of average temperature in the 10 days prior to Midwest
+green-up''. These results should be linked to the phenotypic GxE
+analysis.
 
-We edit the manuscript to remove this unclear claim.
-NB: We do cite the first experimental work using both QTL mapping and GWAS ([Brachi *et al.,* 2010](https://doi.org/10.1371/journal.pgen.1000940)), and there are many others; we meant, more specifically, using these approaches to map multiple types of GxE.
+\end{tcolorbox}
+\end{quote}
 
-> ::: {.callout-warning appearance="minimal"}
-> "Gulf and Midwest subpopulations have two distinct photoperiod-related flowering responses....". This has been shown for other species, eg Unterseer et al. 2016 Genome Biology
-> :::
+We edit the results to directly compare the GxWeather matrices and the
+phenotypic correlations in Figure 1B. (lines 241-258; lines 269-276;
+lines 283-290).
+
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
+
+Flowering posterior weights were higher than green-up weights. What does
+that mean? How does that link to the different heritabilities and levels
+of GxE?
+
+\end{tcolorbox}
+\end{quote}
+
+We believe this comment refers to there being higher loadings on the
+GxWeather matrices for flowering date than there were for the date of
+the start of vegetative growth in the original paper (original Figure
+2). Our revised analyses \hyperref[fig-greedy]{uses a greedy algorithm
+on canonical and GxWeather covariance matrices} to iteratively select
+only the subset that significantly improves the model log-likelihood.
+Thus, the covariance matrices included in each model have changed in
+this revision, as have the relative loadings on GxWeather and canonical
+covariance matrices.
+
+We expected patterns of GxE and GxWeather to vary at the locus,
+subpopulation, and trait level, and we did not see a consistently higher
+fraction of green-up date mass than flowering date mass on the GxWeather
+matrices, nor vice versa. We don't believe it's appropriate to make
+direct comparisons between heritability or a variance components
+analysis and these mash loadings, so we do not do so in the paper;
+instead of the variance due to additive genetic variation and GxE, these
+show the additive genetic effects that most closely matched different
+patterns of GxE and GxWeather, regardless of the proportion of additive
+variance each locus has.
+
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
+
+What can you conclude from the pattern of antagonistic pleiotropy
+between Texas and Northern gardens that has not been seen in the GxE
+analysis?
+
+\end{tcolorbox}
+\end{quote}
+
+No conclusions about individual loci effects can be drawn from the GxE
+analysis in Figure 2, as this plot does not show how shrinkage changes
+the jointly re-estimated SNP effect sizes. The GxE analysis in Figure 2
+shows the types of GxE and GxWeather present genome-wide, from loadings
+on individual loci, but does not show the effects of these loci on
+traits. Rather, it is an overall characterization of GxE across all
+eight gardens.
+
+The GxE analysis in Figure 3 shows the re-estimated effects of these
+loci on traits as contrasts between individual pairs of gardens. From
+Figure 3, you can conclude that many loci have effects with antagonistic
+pleiotropy at the site level.
+
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
+
+Discussion\\
+``we must understand the current patterns of trait covariation across
+environments, \ldots..''. What is the understanding from your data? Are
+there any general patterns or do we need to assess them individually for
+different genetic material, environments, traits? It looks like the
+latter so what are the consequences?
+
+\end{tcolorbox}
+\end{quote}
+
+As we state in the SI Appendix (Section S1, paragraph two), we were
+interested in the types of genetic correlation that could be expressed
+by different alleles across the genome. We reasoned that different loci
+could have different patterns of genetic correlation. For example, the
+effects of an allele of locus \(A\) could covary with a weather-based
+cue if \(A\) had a common allele, \(a\), that was responsive to this
+cue. In contrast, the effects of a different locus \(B\) might have an
+effect at only one garden, for example if a pathogen was only present at
+one garden, and \(B\) has a common allele \(b\) that is resistant to
+that pathogen.
+
+As a consequence, we develop an approach to specify multiple
+environmental cues and compete them to explain patterns of genetic
+effects (lines 230-237): we use a greedy algorithm to use with mash to
+select covariance patterns, and use mash to flexibly identify these
+patterns across populations and environments. (lines 238-240). With this
+approach, we are able to demonstrate that we can associate multiple
+patterns of GxWeather with specific genomic regions (lines 407-410). We
+can also assign genetic effects to both GxWeather patterns derived from
+weather variables and to other, site-based patterns (lines 410-412).
+
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
+
+``this is the first experimental work using QTL mapping and GWAS across
+\ldots.''. I disagree, see references above and others.
+
+\end{tcolorbox}
+\end{quote}
+
+We edit the manuscript to remove this unclear claim. NB: We do cite the
+first experimental work using both QTL mapping and GWAS
+(\href{https://doi.org/10.1371/journal.pgen.1000940}{Brachi \emph{et
+al.,} 2010}), and there are many others; we meant, more specifically,
+using these approaches to map multiple types of GxE.
+
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
+
+``Gulf and Midwest subpopulations have two distinct photoperiod-related
+flowering responses\ldots.''. This has been shown for other species, eg
+Unterseer et al.~2016 Genome Biology
+
+\end{tcolorbox}
+\end{quote}
 
@@
 
-> ::: {.callout-warning appearance="minimal"}
-> Data/Methods\
-> \
-> Is one year data enough to make conclusions about the effect of environmental cues? It is well known that GxYear is more pronounced than GxLocation.
-> :::
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
 
-We agree that this would be an interesting topic for future work. Our aim was to demonstrate the value of this approach for mapping GxE, and we believe one year of data sufficient to do this.
+Data/Methods\\
+\strut \\
+Is one year data enough to make conclusions about the effect of
+environmental cues? It is well known that GxYear is more pronounced than
+GxLocation.
 
-> ::: {.callout-warning appearance="minimal"}
-> The reader should be able to understand what the hypothesis in the hypothesis-driven models is without looking at the supplement.
-> :::
+\end{tcolorbox}
+\end{quote}
 
-Agreed.
-We add Table 1 to better explain the covariance matrices we select using the greedy mash algorithm.
-We include the matrices selected by this algorithm that have \>0.1% weight (lines 262-263) in the mash model as visualizations in Figure 2.
+We agree that this would be an interesting topic for future work. Our
+aim was to demonstrate the value of this approach for mapping GxE, and
+we believe one year of data sufficient to do this.
 
-> ::: {.callout-warning appearance="minimal"}
-> Reviewer #2:\
-> \
-> Suitable Quality?: No\
-> Sufficient General Interest?: Yes\
-> Conclusions Justified?: No\
-> Clearly Written?: No\
-> Procedures Described?: No\
-> \
-> Comments:\
-> \
-> The authors present a massive dataset on phenological variation among switchgrass cultivars across 8 common gardens.
-They use this dataset to describe the genetic architecture of gene-environment interactions in two phenological traits: the timing of green-up and flowering.
-They find that the genetic architecture of these traits varies considerably among two populations of switchgrass (called Gulf and Midwest) and among the latitudinal cline in common gardens, and that some of this variation seems to be related to variation among gardens in key environmental cues including temperature and daylength.
-They also identify genomic regions associated with this variation and replicate these regions in a separate set of F2 populations.\
-> \
-> While previous publications from this experiment have been published, I believe this is the first to focus on these two phenological traits which are of clear importance in switchgrass.
-The overall questions that they target are also of great importance - understanding the genetic basis of gene-environment interactions, and identifying the environmental drivers of these phenological traits - and the analytical approach that they use is creative and leverages powerful statistical methods.
-Given this, this study has the potential to be an important case study in the field for identifying GxE loci.
-However there are a number of issues with the statistical approach and also with the conceptual framework that I think make the current results uninterpretable.
-Also, because the methods are relatively novel and complex, it would be very helpful to provide more intuition behind the analytical approaches and more access to the raw results so that others can understand the approach more fully.
->
-> First a note: It would be helpful if the authors would provide a pdf with line-numbers and with a font that can be copied directly.
-> :::
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
+
+The reader should be able to understand what the hypothesis in the
+hypothesis-driven models is without looking at the supplement.
+
+\end{tcolorbox}
+\end{quote}
+
+Agreed. We add Table 1 to better explain the covariance matrices we
+select using the greedy mash algorithm. We include the matrices selected
+by this algorithm that have \textgreater0.1\% weight (lines 262-263) in
+the mash model as visualizations in Figure 2.
+
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
+
+Reviewer \#2:\\
+\strut \\
+Suitable Quality?: No\\
+Sufficient General Interest?: Yes\\
+Conclusions Justified?: No\\
+Clearly Written?: No\\
+Procedures Described?: No\\
+\strut \\
+Comments:\\
+\strut \\
+The authors present a massive dataset on phenological variation among
+switchgrass cultivars across 8 common gardens. They use this dataset to
+describe the genetic architecture of gene-environment interactions in
+two phenological traits: the timing of green-up and flowering. They find
+that the genetic architecture of these traits varies considerably among
+two populations of switchgrass (called Gulf and Midwest) and among the
+latitudinal cline in common gardens, and that some of this variation
+seems to be related to variation among gardens in key environmental cues
+including temperature and daylength. They also identify genomic regions
+associated with this variation and replicate these regions in a separate
+set of F2 populations.\\
+\strut \\
+While previous publications from this experiment have been published, I
+believe this is the first to focus on these two phenological traits
+which are of clear importance in switchgrass. The overall questions that
+they target are also of great importance - understanding the genetic
+basis of gene-environment interactions, and identifying the
+environmental drivers of these phenological traits - and the analytical
+approach that they use is creative and leverages powerful statistical
+methods. Given this, this study has the potential to be an important
+case study in the field for identifying GxE loci. However there are a
+number of issues with the statistical approach and also with the
+conceptual framework that I think make the current results
+uninterpretable. Also, because the methods are relatively novel and
+complex, it would be very helpful to provide more intuition behind the
+analytical approaches and more access to the raw results so that others
+can understand the approach more fully.
+
+First a note: It would be helpful if the authors would provide a pdf
+with line-numbers and with a font that can be copied directly.
+
+\end{tcolorbox}
+\end{quote}
 
@@
 
-> ::: {.callout-warning appearance="minimal"}
-> Major issues:\
-> \
-> Conceptual issues:\
-> \
-> - I like the idea of using environmental measures near the time of phenological events to try to identify environmental drivers and compare the plasticity functions across environments.
-However the choices here do not make physiological sense, especially for flowering.
-The environmental indices focus on either the day of flowering, or the 1-2 weeks prior to flowering.
-However, the developmental transition to flowering likely occurs well before this interval (I'm not sure how long in switchgrass, but it's likely much earlier).
-Photoperiod and temperature cues driving this developmental transition are irrelevant once the developmental commitment has been made.
-The length of time before flowers emerge and open may be dependent on these environmental factors as well, but this is physiological, not related mechanistically to the daylength or temperature requirements for flowering.
-The intervals used may be more relevant to green-up, I'm not familiar with the developmental basis of this trait.
-But as is, I don't think these metrics are interpretable for what they are designed for.
-> :::
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
 
-We agree that the choice of environmental measure is an extremely important consideration, and we recognize that the time window the weather signal is integrated over could be long.
-We wanted our model to inform us on reasonable time windows, rather than having us assert these (probably incorrectly) in our model.
-Initially, we thought computational feasibility constrained the number of covariance matrices we could add to the mash model, as the runtime of mash increases with the number of covariance matrices included.
-To address this comment and the mash statistical issue the reviewer points out below, we now specify many more GxWeather covariance matrices with more time frames - 1-7 days prior to the phenological event, and 14, 21, and 28 days prior to the phenological event, generating 48-60 matrices per weather variable (Table 1).
-Then, we used a greedy mash algorithm, [explained below](#fig-greedy), to select GxWeather covariance matrices from this set that significantly improve the model likelihood.
-We thus extended the time frame for GxWeather covariance that our models could capture; however, only one 28 day matrix was selected by the greedy algorithm and had \>0.1% weight in our *mash* models.
+Major issues:\\
+\strut \\
+Conceptual issues:\\
+\strut \\
+- I like the idea of using environmental measures near the time of
+phenological events to try to identify environmental drivers and compare
+the plasticity functions across environments. However the choices here
+do not make physiological sense, especially for flowering. The
+environmental indices focus on either the day of flowering, or the 1-2
+weeks prior to flowering. However, the developmental transition to
+flowering likely occurs well before this interval (I'm not sure how long
+in switchgrass, but it's likely much earlier). Photoperiod and
+temperature cues driving this developmental transition are irrelevant
+once the developmental commitment has been made. The length of time
+before flowers emerge and open may be dependent on these environmental
+factors as well, but this is physiological, not related mechanistically
+to the daylength or temperature requirements for flowering. The
+intervals used may be more relevant to green-up, I'm not familiar with
+the developmental basis of this trait. But as is, I don't think these
+metrics are interpretable for what they are designed for.
 
-We now also explicitly describe the two phenological traits more clearly in the text - both of the traits we map were measured on individuals, but specifically when approximately half of the ramets of the genotype had open flowers (for flowering) or green leaves (for green-up).
-Thus we are looking for cues driving vegetative and reproductive transitions for a majority of ramets, not for heading (flower emergence on the panicle).
-While there may be physiological mechanisms driving flower opening once the transition to reproductive development has been made, our results show that genetic variation underlies these differences in the timing of the physiological mechanisms underlying flower opening.
+\end{tcolorbox}
+\end{quote}
 
-> ::: {.callout-warning appearance="minimal"}
-> \- I’ve struggled to understand the “hypothesis-based covariance matrix” idea.
-Since this idea is really key to the idea of this paper, I think the authors should work to make it much more clear and intuitive, perhaps a diagram could help.
-I think the idea is: if in two gardens the daylength-at-flowering metrics are correlated across accessions, this implies that the importance of daylength on flowering is similar in these two gardens.
-In a third environment that has little variation in daylength-at-flowering, the genetics of the daylength pathway are still the same, so the lack of observed variation must mean that this cue is less important here.
-If a SNP has a similar effect in these two gardens, but a smaller effect in gardens without much variation in this metric, this implies that this SNP may contribute to the signaling pathway that links daylength to flowering.
-This is a neat idea.
-However the way this is described in the text (eg “SNP effects on flowering ...
-covaried with daylength”) is confusing.
-This sounds like SNP effects were larger in gardens where the mean daylength at flowering was later.
-Or maybe that these SNPs were also GWAS hits for daylength at flowering itself.
-> :::
+We agree that the choice of environmental measure is an extremely
+important consideration, and we recognize that the time window the
+weather signal is integrated over could be long. We wanted our model to
+inform us on reasonable time windows, rather than having us assert these
+(probably incorrectly) in our model. Initially, we thought computational
+feasibility constrained the number of covariance matrices we could add
+to the mash model, as the runtime of mash increases with the number of
+covariance matrices included. To address this comment and the mash
+statistical issue the reviewer points out below, we now specify many
+more GxWeather covariance matrices with more time frames - 1-7 days
+prior to the phenological event, and 14, 21, and 28 days prior to the
+phenological event, generating 48-60 matrices per weather variable
+(Table 1). Then, we used a greedy mash algorithm,
+\hyperref[fig-greedy]{explained below}, to select GxWeather covariance
+matrices from this set that significantly improve the model likelihood.
+We thus extended the time frame for GxWeather covariance that our models
+could capture; however, only one 28 day matrix was selected by the
+greedy algorithm and had \textgreater0.1\% weight in our \emph{mash}
+models.
 
-We thank the reviewer for this insight; we have reframed the way we talk about these covariance matrices in the paper (starting with the title, abstract, and significance statement; continuing throughout).
-Instead of hypothesis-based covariance matrices, we emphasize that the matrices we create are looking at an interaction between genetics and weather - they are GxWeather matrices, and we look for effects with different GxWeather interactions that are based on specific weather cues.
+We now also explicitly describe the two phenological traits more clearly
+in the text - both of the traits we map were measured on individuals,
+but specifically when approximately half of the ramets of the genotype
+had open flowers (for flowering) or green leaves (for green-up). Thus we
+are looking for cues driving vegetative and reproductive transitions for
+a majority of ramets, not for heading (flower emergence on the panicle).
+While there may be physiological mechanisms driving flower opening once
+the transition to reproductive development has been made, our results
+show that genetic variation underlies these differences in the timing of
+the physiological mechanisms underlying flower opening.
 
-> ::: {.callout-warning appearance="minimal"}
-> Statistical issues:\
-> \
-> - Narrow-sense heritability: A key result is that the h\^2 is low when measured across all trials, but high within trials.
-However the model used to estimate the global h\^2 is missing a term for replication of lines across accessions.
-u only accounts for replication of additive effects of lines across gardens, but non-additive genetics can be persistent too.
-Also, this model doesn’t allow var(e) to vary across locations but it’s likely that it does.
-If the goal is to show rank-crossing GxE, actually fit a GxE model like was done for the environmental indices to directly show it.\
-> - Variance components analysis: The specification of this model is incorrect because Var(u) and Var(ul) are assigned the same distribution.
-There are many ways of specifying GxE effects in a model like this, and it’s not clear how it is done here.
-If G is a nxn GRM for n accessions, it can’t be the covariance matrix for both Var(u) (dimension nxn) and Var(ul) (dimension nl x nl).
-Generally, Var(ul) would be G \otimes Psi where Psi is a lxl covariance matrix among gardens.
-Sometimes this is constrained to be diagonal (ie no covariance among gardens), and sometimes to be proportional to I (constant variance and no covariance among gardens), but these more restrictive models should be justified.
-> :::
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
 
-We agree that correctly specifying the type of GxE in our models of additive variation across gardens is important for all models in this manuscript.
-In many cases the complexity of the models that we can specify using this data is limited, because of the low sample size - we typically have one replicate per genotype in each common garden, and the low number of genotypes present at each garden in each strongly stratified subpopulation means that we cannot fit complicated GxE structures using linear mixed models, often needing to fit very simple/restrictive variance-covariance structures.
-The entire focus on mash was to allow us to fit less structured variance-covariance structures, as well as more of them.
+- I've struggled to understand the ``hypothesis-based covariance
+matrix'' idea. Since this idea is really key to the idea of this paper,
+I think the authors should work to make it much more clear and
+intuitive, perhaps a diagram could help. I think the idea is: if in two
+gardens the daylength-at-flowering metrics are correlated across
+accessions, this implies that the importance of daylength on flowering
+is similar in these two gardens. In a third environment that has little
+variation in daylength-at-flowering, the genetics of the daylength
+pathway are still the same, so the lack of observed variation must mean
+that this cue is less important here. If a SNP has a similar effect in
+these two gardens, but a smaller effect in gardens without much
+variation in this metric, this implies that this SNP may contribute to
+the signaling pathway that links daylength to flowering. This is a neat
+idea. However the way this is described in the text (eg ``SNP effects on
+flowering \ldots{} covaried with daylength'') is confusing. This sounds
+like SNP effects were larger in gardens where the mean daylength at
+flowering was later. Or maybe that these SNPs were also GWAS hits for
+daylength at flowering itself.
 
-Thus we thought hard about the minimum number of linear mixed models we could specify to justify our modeling approach with *mash*.
-We think that the presence of negative phenotypic covariation and additive genetic variation at each garden is sufficient to motivate a search for the sign- or rank-changing GxE that could potentially underlie strong negative covariation between the North and Texas gardens.
-Clearly, additive genetic variation is necessary to conduct a genetic mapping analysis; in addition, we find many more loci with rank-changing GxE for green-up than for flowering, and green-up has negative phenotypic correlations which flowering does not.
+\end{tcolorbox}
+\end{quote}
 
-We could not find a similar compelling reason to include the variance components analysis of the phenological traits or weather-derived traits based on these phenological dates.
-Because we construct the GxWeather matrices from covariation in these weather-derived cues, then mash loads loci onto these GxWeather matrices, we felt it did not add value to the paper to include a variance components analysis of these as additional traits; thus, we remove the variance components analysis from this revision.
+We thank the reviewer for this insight; we have reframed the way we talk
+about these covariance matrices in the paper (starting with the title,
+abstract, and significance statement; continuing throughout). Instead of
+hypothesis-based covariance matrices, we emphasize that the matrices we
+create are looking at an interaction between genetics and weather - they
+are GxWeather matrices, and we look for effects with different GxWeather
+interactions that are based on specific weather cues.
 
-Additionally, we move the narrow-sense heritability analysis to the SI Appendix (Figure S1).
-Our goal was to show suitability for further analysis of additive genetic variation, not directly show rank-crossing GxE with this mixed linear model, so we do not change the environmental effect of site, which does not have an interaction term, in the narrow-sense heritability models.
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
 
-> ::: {.callout-warning appearance="minimal"}
-> \- GWAS: The GWAS model the authors use is not specified (nor was it specified in the 2021 Nature paper referenced).
-I believe looking at the source code that the model is a linear model with PCs to account for structure but no kinship matrix.
-It’s not clear how many PCs were used and since the raw GWAS results are not presented it’s impossible to tell if this was sufficient to account for the significant structure in these populations.
-I’m concerned about this because of the extremely unlikely number of significant markers (19K LD blocks).
-If each of these was a true positive, the average effect size of each block would be only 0.005% of the genetic variance.
-With only 350 accessions, the power to detect a locus that explains 10% of the genetic variance is only 50% with alpha = 10\^-5 and MAF = 0.05.
-So if the model is detecting tons of loci it’s likely that the model is severely biased by population structure, and these biases are likely to be somewhat consistent across gardens because the structure is similar.
-mash doesn’t help when the individual models are biased, it’ll find patterns of correlation whether they are biological or not and use those to shrink effect sizes together.
-> :::
+Statistical issues:\\
+\strut \\
+- Narrow-sense heritability: A key result is that the h\^{}2 is low when
+measured across all trials, but high within trials. However the model
+used to estimate the global h\^{}2 is missing a term for replication of
+lines across accessions. u only accounts for replication of additive
+effects of lines across gardens, but non-additive genetics can be
+persistent too. Also, this model doesn't allow var(e) to vary across
+locations but it's likely that it does. If the goal is to show
+rank-crossing GxE, actually fit a GxE model like was done for the
+environmental indices to directly show it.\\
+- Variance components analysis: The specification of this model is
+incorrect because Var(u) and Var(ul) are assigned the same distribution.
+There are many ways of specifying GxE effects in a model like this, and
+it's not clear how it is done here. If G is a nxn GRM for n accessions,
+it can't be the covariance matrix for both Var(u) (dimension nxn) and
+Var(ul) (dimension nl x nl). Generally, Var(ul) would be G \otimes Psi
+where Psi is a lxl covariance matrix among gardens. Sometimes this is
+constrained to be diagonal (ie no covariance among gardens), and
+sometimes to be proportional to I (constant variance and no covariance
+among gardens), but these more restrictive models should be justified.
 
-First, we now better document our GWAS methodology for this manuscript in Section S2 of the SI Appendix.
-We also include additional information on GWAS, including the number of PCs used and the Manhattan and QQ plots and associated data tables in the Github repository associated with the paper.
+\end{tcolorbox}
+\end{quote}
 
-Second, though we selected 19K LD blocks in our set of 'strong' effects for *mash*, very few of these effects were significant (\<500 relatively unlinked LD blocks per set of eight gardens; this still represents some genomic inflation, no doubt, but not nearly as bad as 19K significant blocks would be).
-We now clarify this important detail in the main manuscript and in the SI Appendix.
+We agree that correctly specifying the type of GxE in our models of
+additive variation across gardens is important for all models in this
+manuscript. In many cases the complexity of the models that we can
+specify using this data is limited, because of the low sample size - we
+typically have one replicate per genotype in each common garden, and the
+low number of genotypes present at each garden in each strongly
+stratified subpopulation means that we cannot fit complicated GxE
+structures using linear mixed models, often needing to fit very
+simple/restrictive variance-covariance structures. The entire focus on
+mash was to allow us to fit less structured variance-covariance
+structures, as well as more of them.
 
-Third, population structure could certainly bias the resultant mash models.
-We describe the steps we take to prevent "garbage in" to our mash models, including removing conditions where GWAS had significant population structure (SI Appendix, Section S2, last paragraph).
-However, it is our intuition that, should biases remain in our models, as they undoubtedly do, mash would not load these biased effects onto our GxWeather covariance matrices - instead, these effects likely load onto canonical matrices, such as garden-specific effects, or potentially onto data-driven matrices.
-Thus, we expect our assessment of the proportion of the genome that shows GxWeather effects to be conservative if our models include residual population structure.
+Thus we thought hard about the minimum number of linear mixed models we
+could specify to justify our modeling approach with \emph{mash}. We
+think that the presence of negative phenotypic covariation and additive
+genetic variation at each garden is sufficient to motivate a search for
+the sign- or rank-changing GxE that could potentially underlie strong
+negative covariation between the North and Texas gardens. Clearly,
+additive genetic variation is necessary to conduct a genetic mapping
+analysis; in addition, we find many more loci with rank-changing GxE for
+green-up than for flowering, and green-up has negative phenotypic
+correlations which flowering does not.
 
-> ::: {#comment-cov-correlations .callout-warning appearance="minimal"}
-> \- Mash: A lot of the results interpretation is based on interpreting the SNP loadings on the specified covariance matrices.
-This is a secondary use of mash (the primary being the refinement of effect sizes), and while they do interpret these somewhat in the mash paper, these loadings need to be interpreted with caution.
-If these covariance matrices are similar, mash will somewhat arbitrarily assign weight to any of the matrices because all lead to the same fit to the effect-size data.
-The mash algorithm doesn’t give the full posterior distribution on the loadings so you can’t check for posterior correlations there.
-It seems likely to me that the hypothesis-driven and data-driven covariance matrices are somewhat correlated here, and the correlations may differ between green-up and flowering because of the better correlation between the environmental metrics and flowering than green-up.
-In the mash paper, they used cross-validation with the likelihood in the testing set as the evaluation metric to compare models.
-It might be safer to try dropping specific covariance matrices and comparing the model performance in a held-out testing set to evaluate the importance of the hypothesis-derived covariance matrices.
-> :::
+We could not find a similar compelling reason to include the variance
+components analysis of the phenological traits or weather-derived traits
+based on these phenological dates. Because we construct the GxWeather
+matrices from covariation in these weather-derived cues, then mash loads
+loci onto these GxWeather matrices, we felt it did not add value to the
+paper to include a variance components analysis of these as additional
+traits; thus, we remove the variance components analysis from this
+revision.
 
-We thank the reviewer for this comment and agree that the selection of covariance matrices to be included in a mash model is a nontrivial question.
-To clarify our interpretation of the posterior matrix weights provided by mash, we performed an extensive analysis of the performance of mash models when different covariance matrices are included.
-Specifically, we implemented a model selection approach that uses a greedy algorithm to evaluate the log likelihood of the mash model as additional covariance matrices were included (see Methods lines 899-906, SI Appendix Section S4, and pseudo code below).
-This is a similar, but more computationally efficient design, than the leave-one-out approach recommended by the reviewer as the runtime of mash increases with the number of covariance matrices included.
-Additionally, while cross-validation is a powerful approach to model evaluation our limited sample size was prohibitive to the necessary partitioning of individuals included in our analysis.
+Additionally, we move the narrow-sense heritability analysis to the SI
+Appendix (Figure S1). Our goal was to show suitability for further
+analysis of additive genetic variation, not directly show rank-crossing
+GxE with this mixed linear model, so we do not change the environmental
+effect of site, which does not have an interaction term, in the
+narrow-sense heritability models.
 
-In practice this greedy algorithm approach offers a fast way to identify the point where redundancy in the addition of a new covariance matrix results in no change to the likelihood of the mash model.
-Importantly, this removes the potential for arbitrary assignment of weights to the covariance matrices in the model and allows for the selection of matrices that most accurately capture the underlying biology captured by the site-specific effect sizes.
-We applied this algorithm to only the GxWeather covariance matrices for each phenotype (Figures 1 and 2) using the previously used sets of significantly associated and randomly selected variants in the Gulf subpopulation individuals, Midwest subpopulation individuals, and the two subpopulations combined.
-The results indicate that only a small number of the original hypothesis matrices (between three and five) are necessary to reach the maximum likelihood model when using the same likelihood ratio test implemented in Urbut *et al*., indicating that correlation among hypothesis covariance matrices is high.
-This redundancy was confirmed through our application of the greedy algorithm to a combined set of canonical, data driven, and GxWeather covariance matrices (Figures 3 and 4) Our primary conclusion from this analysis is in line with what the reviewer posited, that there is extensive redundancy among canonical, data driven, and hypothesis covariance matrices that is leading to biologically informative patterns of covariance between site effect sizes to be arbitrarily distributed among them.
-We additionally found evidence that while the correlations between data driven and environmentally informed (hypothesis) covariance matrices, there is extensive similarity in the matrices that are included in the maximum likelihood models across both phenotypes and in the analysis of the Gulf subpopulation, Midwest subpopulations, and the combined cohort.
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
 
-``` {#fig-greedy .r}
-Greedy algorithm for mash model selection. 
-____________________________________________________________________________
+- GWAS: The GWAS model the authors use is not specified (nor was it
+specified in the 2021 Nature paper referenced). I believe looking at the
+source code that the model is a linear model with PCs to account for
+structure but no kinship matrix. It's not clear how many PCs were used
+and since the raw GWAS results are not presented it's impossible to tell
+if this was sufficient to account for the significant structure in these
+populations. I'm concerned about this because of the extremely unlikely
+number of significant markers (19K LD blocks). If each of these was a
+true positive, the average effect size of each block would be only
+0.005\% of the genetic variance. With only 350 accessions, the power to
+detect a locus that explains 10\% of the genetic variance is only 50\%
+with alpha = 10\^{}-5 and MAF = 0.05. So if the model is detecting tons
+of loci it's likely that the model is severely biased by population
+structure, and these biases are likely to be somewhat consistent across
+gardens because the structure is similar. mash doesn't help when the
+individual models are biased, it'll find patterns of correlation whether
+they are biological or not and use those to shrink effect sizes
+together.
 
-matrices = [] #<1>
-most_likely_matrices = [] #<2>
-maximum_likelihoods = []#<2>
+\end{tcolorbox}
+\end{quote}
 
-likelihood = -inf #<2>
-most_likely_matrix = ‘’ #<2>
+First, we now better document our GWAS methodology for this manuscript
+in Section S2 of the SI Appendix. We also include additional information
+on GWAS, including the number of PCs used and the Manhattan and QQ plots
+and associated data tables in the Github repository associated with the
+paper.
 
-for matrix in matrices: #<3>
-    matrix_likelihood = mash(effects, std.errs, matrix) [‘likelihood’]
-    if matrix_likelihood > likelihood:
-        most_likely_matrix = matrix
-        likelihood = matrix_likelihood
+Second, though we selected 19K LD blocks in our set of `strong' effects
+for \emph{mash}, very few of these effects were significant
+(\textless500 relatively unlinked LD blocks per set of eight gardens;
+this still represents some genomic inflation, no doubt, but not nearly
+as bad as 19K significant blocks would be). We now clarify this
+important detail in the main manuscript and in the SI Appendix.
 
-matrices.remove(most_likely_matrix) #<4>
-most_likely_matrices.append(most_likely_matrix)
-maximum_likelihoods.append(likelihood)
+Third, population structure could certainly bias the resultant mash
+models. We describe the steps we take to prevent ``garbage in'' to our
+mash models, including removing conditions where GWAS had significant
+population structure (SI Appendix, Section S2, last paragraph). However,
+it is our intuition that, should biases remain in our models, as they
+undoubtedly do, mash would not load these biased effects onto our
+GxWeather covariance matrices - instead, these effects likely load onto
+canonical matrices, such as garden-specific effects, or potentially onto
+data-driven matrices. Thus, we expect our assessment of the proportion
+of the genome that shows GxWeather effects to be conservative if our
+models include residual population structure.
 
-lrt_pvalue = 0
-iteration = 2
-while lrt_pvalue < 0.05: #<5>
-    likelihood = -inf
-  most_likely_matrix = ‘’
-    for matrix in matrices:
-        model_matrices = most_likely_matrices + matrix 
-        model_likelihood = mash(effects, std.errs, model_matrices)[‘likelihood’]
-    if matrix_likelihood > likelihood:
-        most_likely_matrix = matrix
-        likelihood = matrix_likelihood
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
 
-  matrices.remove(most_likely_matrix)
-  most_likely_matrices.append(most_likely_matrix)
-  maximum_likelihoods.append(likelihood)
+- Mash: A lot of the results interpretation is based on interpreting the
+SNP loadings on the specified covariance matrices. This is a secondary
+use of mash (the primary being the refinement of effect sizes), and
+while they do interpret these somewhat in the mash paper, these loadings
+need to be interpreted with caution. If these covariance matrices are
+similar, mash will somewhat arbitrarily assign weight to any of the
+matrices because all lead to the same fit to the effect-size data. The
+mash algorithm doesn't give the full posterior distribution on the
+loadings so you can't check for posterior correlations there. It seems
+likely to me that the hypothesis-driven and data-driven covariance
+matrices are somewhat correlated here, and the correlations may differ
+between green-up and flowering because of the better correlation between
+the environmental metrics and flowering than green-up. In the mash
+paper, they used cross-validation with the likelihood in the testing set
+as the evaluation metric to compare models. It might be safer to try
+dropping specific covariance matrices and comparing the model
+performance in a held-out testing set to evaluate the importance of the
+hypothesis-derived covariance matrices.
 
-  lrt_pvalue = likelihood_ratio_test(likelihood, maximum_likelihoods[iteration-1], df = 1)
-____________________________________________________________________________
-```
+\end{tcolorbox}
+\end{quote}
 
-1.  Initialize covariance matrices of interest
-2.  Initialize lists to store the most likely additional matrix and corresponding maximum likelihood
-3.  Obtain model likelihood for a model fit with each matrix
-4.  Remove the matrix with the greatest likelihood from the array and store it
-5.
-Repeat the process of fitting mash models, finding the most likely, pair, trio, etc.
-(using the most likely matrices from the preceding iteration) until the likelihood ratio test is no longer significant
+We thank the reviewer for this comment and agree that the selection of
+covariance matrices to be included in a mash model is a nontrivial
+question. To clarify our interpretation of the posterior matrix weights
+provided by mash, we performed an extensive analysis of the performance
+of mash models when different covariance matrices are included.
+Specifically, we implemented a model selection approach that uses a
+greedy algorithm to evaluate the log likelihood of the mash model as
+additional covariance matrices were included (see Methods lines 899-906,
+SI Appendix Section S4, and pseudo code below). This is a similar, but
+more computationally efficient design, than the leave-one-out approach
+recommended by the reviewer as the runtime of mash increases with the
+number of covariance matrices included. Additionally, while
+cross-validation is a powerful approach to model evaluation our limited
+sample size was prohibitive to the necessary partitioning of individuals
+included in our analysis.
 
-![**Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the green-up date phenotype using only GxWeather covariance matrices.** The matrix that resulted in the maximum likelihood model of covariance in summary statistics among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis.
-(A) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals.
-The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
-(B) and (C) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.
-(D) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals.
-The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
-\* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants.
-(E) and (F) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.](images/RR_Figure1.png)
+In practice this greedy algorithm approach offers a fast way to identify
+the point where redundancy in the addition of a new covariance matrix
+results in no change to the likelihood of the mash model. Importantly,
+this removes the potential for arbitrary assignment of weights to the
+covariance matrices in the model and allows for the selection of
+matrices that most accurately capture the underlying biology captured by
+the site-specific effect sizes. We applied this algorithm to only the
+GxWeather covariance matrices for each phenotype (Figures 1 and 2) using
+the previously used sets of significantly associated and randomly
+selected variants in the Gulf subpopulation individuals, Midwest
+subpopulation individuals, and the two subpopulations combined. The
+results indicate that only a small number of the original hypothesis
+matrices (between three and five) are necessary to reach the maximum
+likelihood model when using the same likelihood ratio test implemented
+in Urbut \emph{et al}., indicating that correlation among hypothesis
+covariance matrices is high. This redundancy was confirmed through our
+application of the greedy algorithm to a combined set of canonical, data
+driven, and GxWeather covariance matrices (Figures 3 and 4) Our primary
+conclusion from this analysis is in line with what the reviewer posited,
+that there is extensive redundancy among canonical, data driven, and
+hypothesis covariance matrices that is leading to biologically
+informative patterns of covariance between site effect sizes to be
+arbitrarily distributed among them. We additionally found evidence that
+while the correlations between data driven and environmentally informed
+(hypothesis) covariance matrices, there is extensive similarity in the
+matrices that are included in the maximum likelihood models across both
+phenotypes and in the analysis of the Gulf subpopulation, Midwest
+subpopulations, and the combined cohort.
 
-![**Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the flowering date phenotype using only GxWeather covariance matrices.** The matrix that resulted in the maximum likelihood model among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis.
-(A) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals.
-The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
-(B) and (C) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.
-(D) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals.
-The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
-\* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants.
-(E) and (F) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.](images/RR_Figure2.png)
+\phantomsection\label{fig-greedy}%
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{Greedy algorithm }\ControlFlowTok{for}\NormalTok{ mash model selection. }
+\NormalTok{\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_}
 
-![**Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the greenup date phenotype using canonical, data driven, and GxWeather covariance matrices.** The matrix that resulted in the maximum likelihood model of covariance in summary statistics among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis.
-(A) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals.
-The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
-Once this condition was met the greedy algorithm was halted.
-(B) and (C) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.
-(D) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals.
-The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
-Once this condition was met the greedy algorithm was halted for computational considerations.
-\* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants.
-(E) and (F) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.](images/RR_Figure3.png)
+\NormalTok{matrices }\OtherTok{=}\NormalTok{ [] }\hspace*{\fill}\NormalTok{\circled{1}}
+\NormalTok{most\_likely\_matrices }\OtherTok{=}\NormalTok{ [] }\hspace*{\fill}\NormalTok{\circled{2}}
+\NormalTok{maximum\_likelihoods }\OtherTok{=}\NormalTok{ [] }
 
-![**Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the flowering phenotype using canonical, data driven, and GxWeather covariance matrices.** The matrix that resulted in the maximum likelihood model of covariance in summary statistics among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis.
-(A) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals.
-The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
-Once this condition was met the greedy algorithm was halted.
-(B) and (C) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.
-(D) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals.
-The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05.
-Once this condition was met the greedy algorithm was halted for computational considerations.
-\* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants.
-(E) and (F) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.](images/RR_Figure4.png)
+\NormalTok{likelihood }\OtherTok{=} \SpecialCharTok{{-}}\NormalTok{inf }
+\NormalTok{most\_likely\_matrix }\OtherTok{=}\NormalTok{ ‘’ }
 
-> ::: {.callout-warning appearance="minimal"}
-> \- There appears to be an issue with the construction of the hypothesis-derived covariance matrices.
-I believe the goal here is to estimate the genetic covariance among gardens based on the phenotypic covariance (P) and the estimated residual covariance (E), where P = G + E.
-The described approach involves starting with the the phenotypic correlation matrix and then replacing the diagonal with the coefficient of variation.
-If the diagonal had been replaced with the h2 in each garden, that’d be a valid estimate of G.
-But as described, it’s likely that you’ll end up with an invalid covariance matrix, one that’s not positive-semi-definite.
-Instead the appropriate operation would be to both pre- and post-multiply the covariance matrix by a diagonal matrix with the square-root of the CV’s.
-This maintains a valid covariance-like structure.
-Or just mean-standardize the traits first and then calculate the covariance of this standardized matrix.
-> :::
+\ControlFlowTok{for}\NormalTok{ matrix }\ControlFlowTok{in}\NormalTok{ matrices}\SpecialCharTok{:} \hspace*{\fill}\NormalTok{\circled{3}}
+\NormalTok{    matrix\_likelihood }\OtherTok{=} \FunctionTok{mash}\NormalTok{(effects, std.errs, matrix) [‘likelihood’]}
+    \ControlFlowTok{if}\NormalTok{ matrix\_likelihood }\SpecialCharTok{\textgreater{}}\NormalTok{ likelihood}\SpecialCharTok{:}
+\NormalTok{        most\_likely\_matrix }\OtherTok{=}\NormalTok{ matrix}
+\NormalTok{        likelihood }\OtherTok{=}\NormalTok{ matrix\_likelihood}
 
-We now use the narrow-sense heritability as the diagonal for each GxWeather matrix, and clarify this in the Supplementary Methods (SI Appendix, Section S1, first and eighth paragraph - first and last paragraph on page two).
+\FunctionTok{matrices.remove}\NormalTok{(most\_likely\_matrix) }\hspace*{\fill}\NormalTok{\circled{4}}
+\FunctionTok{most\_likely\_matrices.append}\NormalTok{(most\_likely\_matrix)}
+\FunctionTok{maximum\_likelihoods.append}\NormalTok{(likelihood)}
 
-> ::: {.callout-warning appearance="minimal"}
-> \- Antagonistic Pleiotropy: Maybe I’m missing the idea of this analysis, but it doesn’t seem to me that this strategy for partitioning loci between antagonistic pleiotropy and differential sensitivity would have equal power to detect each class.
-It seems that antagonistic pleiotropy should be a subset of differential sensitivity because the latter is defined as simply a change in magnitude.
-But even if it’s restricted to “different magnitude but same sign”, the practical definition here seems to be that antagonistic pleiotropy is detected when the signs are different and both lfsrs \< threshold (0.05?), while differential sensitivity additionally requires a threshold on the difference in effect sizes (0.4) with no justification for why this size was chosen.
-This additional criterion will necessarily make the rate of detection different.
-Even without this, with differential sensitivity one effect size must be much larger than the other in absolute value (which isn’t the case with antagonistic pleiotropy), so differential sensitivity loci will be more likely to pass the lfsr loci than antagonistic pleiotropy loci.
-Note that previous literature here tried to separate antagonistic pleiotropy from conditional neutrality (ie where in one location the effect was zero), while here the contrast is with differential sensitivity where effects are non-zero in all locations.
-I’m not aware of previous discussions in the literature of trying to differentiate these two specific classes of GxE variants and not really sure theoretically what the importance would be.
-> :::
+\NormalTok{lrt\_pvalue }\OtherTok{=} \DecValTok{0}
+\NormalTok{iteration }\OtherTok{=} \DecValTok{2}
+\ControlFlowTok{while}\NormalTok{ lrt\_pvalue }\SpecialCharTok{\textless{}} \FloatTok{0.05}\SpecialCharTok{:} \hspace*{\fill}\NormalTok{\circled{5}}
+\NormalTok{    likelihood }\OtherTok{=} \SpecialCharTok{{-}}\NormalTok{inf}
+\NormalTok{  most\_likely\_matrix }\OtherTok{=}\NormalTok{ ‘’}
+    \ControlFlowTok{for}\NormalTok{ matrix }\ControlFlowTok{in}\NormalTok{ matrices}\SpecialCharTok{:}
+\NormalTok{        model\_matrices }\OtherTok{=}\NormalTok{ most\_likely\_matrices }\SpecialCharTok{+}\NormalTok{ matrix }
+\NormalTok{        model\_likelihood }\OtherTok{=} \FunctionTok{mash}\NormalTok{(effects, std.errs, model\_matrices)[‘likelihood’]}
+    \ControlFlowTok{if}\NormalTok{ matrix\_likelihood }\SpecialCharTok{\textgreater{}}\NormalTok{ likelihood}\SpecialCharTok{:}
+\NormalTok{        most\_likely\_matrix }\OtherTok{=}\NormalTok{ matrix}
+\NormalTok{        likelihood }\OtherTok{=}\NormalTok{ matrix\_likelihood}
 
-We believe that the use of the lfsr to detect effects with rank-changing GxE is a key advance in our manuscript, and so we have made changes to the Introduction (lines 169-171), Results (lines 319-326), and Materials & Methods (lines 957-997; lines 1055-1056) to explain and motivate this statistical change.
+  \FunctionTok{matrices.remove}\NormalTok{(most\_likely\_matrix)}
+  \FunctionTok{most\_likely\_matrices.append}\NormalTok{(most\_likely\_matrix)}
+  \FunctionTok{maximum\_likelihoods.append}\NormalTok{(likelihood)}
+
+\NormalTok{  lrt\_pvalue }\OtherTok{=} \FunctionTok{likelihood\_ratio\_test}\NormalTok{(likelihood, maximum\_likelihoods[iteration}\DecValTok{{-}1}\NormalTok{], }\AttributeTok{df =} \DecValTok{1}\NormalTok{)}
+\NormalTok{\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{description}
+\tightlist
+\item[\circled{1}]
+Initialize covariance matrices of interest
+\item[\circled{2}]
+Initialize lists to store the most likely additional matrix and
+corresponding maximum likelihood
+\item[\circled{3}]
+Obtain model likelihood for a model fit with each matrix
+\item[\circled{4}]
+Remove the matrix with the greatest likelihood from the array and store
+it
+\item[\circled{5}]
+Repeat the process of fitting mash models, finding the most likely,
+pair, trio, etc. (using the most likely matrices from the preceding
+iteration) until the likelihood ratio test is no longer significant
+\end{description}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{images/RR_Figure1.png}
+
+}
+
+\caption{\textbf{Log-likelihood at each step in a greedy algorithm
+implementation of model selection for mash for the green-up date
+phenotype using only GxWeather covariance matrices.} The matrix that
+resulted in the maximum likelihood model of covariance in summary
+statistics among sites is located at the base of the y-axis; matrices
+selected in the ensuing step are placed in ascending order along the
+y-axis. (A) The maximum likelihood path of covariance matrices in the
+greedy algorithm for the set of independent variants that were
+significant in at least one context when analyzing all individuals. The
+dashed black line represents the step in the algorithm where addition of
+the next matrix did not significantly increase the model likelihood,
+likelihood-ratio test p-value \textgreater{} 0.05. (B) and (C) show the
+maximum likelihood paths when the individuals analyzed are only from the
+Gulf and Midwest subpopulations, respectively. (D) The maximum
+likelihood path of covariance matrices in the greedy algorithm for the
+set of independent, randomly selected variants when analyzing all
+individuals. The dashed black line represents the step in the algorithm
+where addition of the next matrix did not significantly increase the
+model likelihood, likelihood-ratio test p-value \textgreater{} 0.05. *
+is used to denote those matrices that are identified by the greedy
+algorithm in the set of significantly associated variants and the
+randomly associated variants. (E) and (F) show the maximum likelihood
+paths when the individuals analyzed are only from the Gulf and Midwest
+subpopulations, respectively.}
+
+\end{figure}%%
+\begin{figure}[H]
+
+{\centering \includegraphics{images/RR_Figure2.png}
+
+}
+
+\caption{\textbf{Log-likelihood at each step in a greedy algorithm
+implementation of model selection for mash for the flowering date
+phenotype using only GxWeather covariance matrices.} The matrix that
+resulted in the maximum likelihood model among sites is located at the
+base of the y-axis; matrices selected in the ensuing step are placed in
+ascending order along the y-axis. (A) The maximum likelihood path of
+covariance matrices in the greedy algorithm for the set of independent
+variants that were significant in at least one context when analyzing
+all individuals. The dashed black line represents the step in the
+algorithm where addition of the next matrix did not significantly
+increase the model likelihood, likelihood-ratio test p-value
+\textgreater{} 0.05. (B) and (C) show the maximum likelihood paths when
+the individuals analyzed are only from the Gulf and Midwest
+subpopulations, respectively. (D) The maximum likelihood path of
+covariance matrices in the greedy algorithm for the set of independent,
+randomly selected variants when analyzing all individuals. The dashed
+black line represents the step in the algorithm where addition of the
+next matrix did not significantly increase the model likelihood,
+likelihood-ratio test p-value \textgreater{} 0.05. * is used to denote
+those matrices that are identified by the greedy algorithm in the set of
+significantly associated variants and the randomly associated variants.
+(E) and (F) show the maximum likelihood paths when the individuals
+analyzed are only from the Gulf and Midwest subpopulations,
+respectively.}
+
+\end{figure}%%
+\begin{figure}[H]
+
+{\centering \includegraphics{images/RR_Figure3.png}
+
+}
+
+\caption{\textbf{Log-likelihood at each step in a greedy algorithm
+implementation of model selection for mash for the greenup date
+phenotype using canonical, data driven, and GxWeather covariance
+matrices.} The matrix that resulted in the maximum likelihood model of
+covariance in summary statistics among sites is located at the base of
+the y-axis; matrices selected in the ensuing step are placed in
+ascending order along the y-axis. (A) The maximum likelihood path of
+covariance matrices in the greedy algorithm for the set of independent
+variants that were significant in at least one context when analyzing
+all individuals. The dashed black line represents the step in the
+algorithm where addition of the next matrix did not significantly
+increase the model likelihood, likelihood-ratio test p-value
+\textgreater{} 0.05. Once this condition was met the greedy algorithm
+was halted. (B) and (C) show the maximum likelihood paths when the
+individuals analyzed are only from the Gulf and Midwest subpopulations,
+respectively. (D) The maximum likelihood path of covariance matrices in
+the greedy algorithm for the set of independent, randomly selected
+variants when analyzing all individuals. The dashed black line
+represents the step in the algorithm where addition of the next matrix
+did not significantly increase the model likelihood, likelihood-ratio
+test p-value \textgreater{} 0.05. Once this condition was met the greedy
+algorithm was halted for computational considerations. * is used to
+denote those matrices that are identified by the greedy algorithm in the
+set of significantly associated variants and the randomly associated
+variants. (E) and (F) show the maximum likelihood paths when the
+individuals analyzed are only from the Gulf and Midwest subpopulations,
+respectively.}
+
+\end{figure}%%
+\begin{figure}[H]
+
+{\centering \includegraphics{images/RR_Figure4.png}
+
+}
+
+\caption{\textbf{Log-likelihood at each step in a greedy algorithm
+implementation of model selection for mash for the flowering phenotype
+using canonical, data driven, and GxWeather covariance matrices.} The
+matrix that resulted in the maximum likelihood model of covariance in
+summary statistics among sites is located at the base of the y-axis;
+matrices selected in the ensuing step are placed in ascending order
+along the y-axis. (A) The maximum likelihood path of covariance matrices
+in the greedy algorithm for the set of independent variants that were
+significant in at least one context when analyzing all individuals. The
+dashed black line represents the step in the algorithm where addition of
+the next matrix did not significantly increase the model likelihood,
+likelihood-ratio test p-value \textgreater{} 0.05. Once this condition
+was met the greedy algorithm was halted. (B) and (C) show the maximum
+likelihood paths when the individuals analyzed are only from the Gulf
+and Midwest subpopulations, respectively. (D) The maximum likelihood
+path of covariance matrices in the greedy algorithm for the set of
+independent, randomly selected variants when analyzing all individuals.
+The dashed black line represents the step in the algorithm where
+addition of the next matrix did not significantly increase the model
+likelihood, likelihood-ratio test p-value \textgreater{} 0.05. Once this
+condition was met the greedy algorithm was halted for computational
+considerations. * is used to denote those matrices that are identified
+by the greedy algorithm in the set of significantly associated variants
+and the randomly associated variants. (E) and (F) show the maximum
+likelihood paths when the individuals analyzed are only from the Gulf
+and Midwest subpopulations, respectively.}
+
+\end{figure}%
+
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
+
+- There appears to be an issue with the construction of the
+hypothesis-derived covariance matrices. I believe the goal here is to
+estimate the genetic covariance among gardens based on the phenotypic
+covariance (P) and the estimated residual covariance (E), where P = G +
+E. The described approach involves starting with the the phenotypic
+correlation matrix and then replacing the diagonal with the coefficient
+of variation. If the diagonal had been replaced with the h2 in each
+garden, that'd be a valid estimate of G. But as described, it's likely
+that you'll end up with an invalid covariance matrix, one that's not
+positive-semi-definite. Instead the appropriate operation would be to
+both pre- and post-multiply the covariance matrix by a diagonal matrix
+with the square-root of the CV's. This maintains a valid covariance-like
+structure. Or just mean-standardize the traits first and then calculate
+the covariance of this standardized matrix.
+
+\end{tcolorbox}
+\end{quote}
+
+We now use the narrow-sense heritability as the diagonal for each
+GxWeather matrix, and clarify this in the Supplementary Methods (SI
+Appendix, Section S1, first and eighth paragraph - first and last
+paragraph on page two).
+
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
+
+- Antagonistic Pleiotropy: Maybe I'm missing the idea of this analysis,
+but it doesn't seem to me that this strategy for partitioning loci
+between antagonistic pleiotropy and differential sensitivity would have
+equal power to detect each class. It seems that antagonistic pleiotropy
+should be a subset of differential sensitivity because the latter is
+defined as simply a change in magnitude. But even if it's restricted to
+``different magnitude but same sign'', the practical definition here
+seems to be that antagonistic pleiotropy is detected when the signs are
+different and both lfsrs \textless{} threshold (0.05?), while
+differential sensitivity additionally requires a threshold on the
+difference in effect sizes (0.4) with no justification for why this size
+was chosen. This additional criterion will necessarily make the rate of
+detection different. Even without this, with differential sensitivity
+one effect size must be much larger than the other in absolute value
+(which isn't the case with antagonistic pleiotropy), so differential
+sensitivity loci will be more likely to pass the lfsr loci than
+antagonistic pleiotropy loci. Note that previous literature here tried
+to separate antagonistic pleiotropy from conditional neutrality (ie
+where in one location the effect was zero), while here the contrast is
+with differential sensitivity where effects are non-zero in all
+locations. I'm not aware of previous discussions in the literature of
+trying to differentiate these two specific classes of GxE variants and
+not really sure theoretically what the importance would be.
+
+\end{tcolorbox}
+\end{quote}
+
+We believe that the use of the lfsr to detect effects with rank-changing
+GxE is a key advance in our manuscript, and so we have made changes to
+the Introduction (lines 169-171), Results (lines 319-326), and Materials
+\& Methods (lines 957-997; lines 1055-1056) to explain and motivate this
+statistical change.
 
 Our expectation is that in nature, effects will always differ.
-Statistically, this difference will not always be detectable.
-Thus, rather than ask "Are these two effects different?" - as we reasonably expect two effects to be, even if this difference cannot be measured - the local false sign rate answers a more meaningful question: Can we be confident in the sign of this effect?
+Statistically, this difference will not always be detectable. Thus,
+rather than ask ``Are these two effects different?'' - as we reasonably
+expect two effects to be, even if this difference cannot be measured -
+the local false sign rate answers a more meaningful question: Can we be
+confident in the sign of this effect?
 
-We think that if evolutionary geneticists are interested in antagonistic pleiotropy and rank-changing GxE, then they should use the lfsr to measure their confidence in the sign of the effect, rather than using statistical tests where the null hypothesis is that the effect is different than zero.
-However, if we use the lfsr, comparisons between antagonistic pleiotropy and conditional neutrality are no longer sensible, as we are no longer doing a statistical test that can robustly detect conditional neutrality (just as the FDR does not robustly detect antagonistic pleiotropy).
-This is justified because detection of antagonistic pleiotropy is more important than detection of conditional neutrality for models of local adaptation.
+We think that if evolutionary geneticists are interested in antagonistic
+pleiotropy and rank-changing GxE, then they should use the lfsr to
+measure their confidence in the sign of the effect, rather than using
+statistical tests where the null hypothesis is that the effect is
+different than zero. However, if we use the lfsr, comparisons between
+antagonistic pleiotropy and conditional neutrality are no longer
+sensible, as we are no longer doing a statistical test that can robustly
+detect conditional neutrality (just as the FDR does not robustly detect
+antagonistic pleiotropy). This is justified because detection of
+antagonistic pleiotropy is more important than detection of conditional
+neutrality for models of local adaptation.
 
 It's true that the threshold for differential sensitivity is arbitrary.
-We have reworded the manuscript to stress that there is equal power to detect effects of different sign as there are effects of the same sign (lines 994-997; lines 1055-1056).
-Another class of effects we have in our analysis that differ from classic antagonistic pleiotropy & conditional neutrality comparisons are effects which are not distinguishable by sign or magnitude.
-It is not accurate that differentially sensitive loci will be more likely to pass the lfsr, because the loci are first selected by significant lfsr, then divided further into effects that can & cannot be distinguished by magnitude.
-Thus effects that are significant, but have small magnitude in both conditions will be labeled as not being distinguishable - essentially, these effects do not have detectable GxE.
-Effects without GxE may be even more common and should also be quantified.
-We propose that, if our intent is to detect sign-changing GxE, then we should use an unbiased statistical test to look at loci with and without sign-changing GxE, the lfsr.
+We have reworded the manuscript to stress that there is equal power to
+detect effects of different sign as there are effects of the same sign
+(lines 994-997; lines 1055-1056). Another class of effects we have in
+our analysis that differ from classic antagonistic pleiotropy \&
+conditional neutrality comparisons are effects which are not
+distinguishable by sign or magnitude. It is not accurate that
+differentially sensitive loci will be more likely to pass the lfsr,
+because the loci are first selected by significant lfsr, then divided
+further into effects that can \& cannot be distinguished by magnitude.
+Thus effects that are significant, but have small magnitude in both
+conditions will be labeled as not being distinguishable - essentially,
+these effects do not have detectable GxE. Effects without GxE may be
+even more common and should also be quantified. We propose that, if our
+intent is to detect sign-changing GxE, then we should use an unbiased
+statistical test to look at loci with and without sign-changing GxE, the
+lfsr.
 
-::: {.center data-latex="1"}
+~
+
+~
+
+\begin{center}
 Best,
-:::
 
-::: {.flushright data-latex="1"}
+\end{center}
+
+\begin{flushright}
 Alice MacQueen and Tom Juenger
-:::
+
+\end{flushright}
+
+\end{document}

### 2024-07-17 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
revision2: Sync cross-references between revision, SI Appendix, and response letter.
Also, add Figures to response letter and change templating to allow this

@@
 For example, the cumulative sum of the rainfall that occurred seven days prior to the flowering date was used to create one possible GxWeather cue variable for flowering date.
-The code for these calculations can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 394.
+The code for these calculations can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 387.
 
@@
 Finally, all covariance matrices were rescaled so that the maximum diagonal element of the rescaled matrix is 1, as in [@Urbut2019].
-Code used to generate these matrices can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 720.
+Code used to generate these matrices can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 715.
 
@@
 ASReml can specify considerably more complex variance structures; however, we were interested in specifying covariance structures using *mash*, which allows the specification of multiple covariance structures; hence, we used a relatively simple model to calculate BLUPs for use in univariate GWAS.
-The code for these calculations can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 75.
+The code for these calculations can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 69.
 
 Univariate GWAS were run independently for each combination of trait, genetic subpopulation, and common garden.
-The code used to run GWAS can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 240.
-We used [`pvdiv_standard_run()`](https://alice-macqueen.github.io/switchgrassGWAS/reference/pvdiv_standard_gwas.html), an opinionated function of the [switchgrassGWAS](https://github.com/Alice-MacQueen/switchgrassGWAS) R package, to conduct GWAS.
+The code used to run GWAS can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using\_ greedy_covar.qmd, starting at line 240.
+We used [`pvdiv_standard_run()`](#0), an opinionated function of the [switchgrassGWAS](#0) R package, to conduct GWAS.
 To control for population structure and run GWAS, this function uses a three-part process.
-First, a fast truncated singular value decomposition (SVD) is generated from the SNP data, which has initial pruning of regions in long-range linkage disequilibrium (LD) and iterative pruning and removal of regions in long-range LD [@privé2017a].
+First, a fast truncated singular value decomposition (SVD) is generated from the SNP data, which has initial pruning of regions in long-range linkage disequilibrium (LD) and iterative pruning and removal of regions in long-range LD [@privé2018].
 By default, 15 vectors of singular values, $k$, are computed.
-Pruning and removal of long-range LD regions are recommended in research in human genetics to best control for population structure [@abdellaoui2013; @privé2017; @price2008].
+Pruning and removal of long-range LD regions are recommended in research in human genetics to best control for population structure [@abdellaoui2013; @privé2018; @price2008].
 Second, linear regression models using 0-15 of these vectors of singular values were used to run GWAS and to calculate the genomic control coefficient, $\lambda_{GC}$, for each $k$ [@Devlin1999].
@@
 These issues were apparent in the data in two ways: (i) univariate GWAS with many significant SNPs with extremely similar *p*-values, usually indicative of genetic effects that, by chance or, more likely, because of confounding effects of population structure, were identical across many SNPs and thus not resolvable to a single genomic location; (ii) univariate GWAS with many effect estimates and standard errors of zero, indicating an inability to estimate effects given the constraints of small population size & population structure.
-We only noticed these issues in our smallest subpopulation, the Midwest subpopulation, in the green-up date trait at the SD garden, and in the flowering date trait at the OK and NE gardens.
-Due to these issues, we dropped the SD garden from subsequent green-up analyses for the Midwest subpopulation, and we dropped the OK and NE gardens from subsequent Midwest flowering date analyses.
+We only noticed these issues in our smallest subpopulation, the Midwest subpopulation, in the green-up date trait at the MO and SD garden, and in the flowering date trait at the OK and NE gardens.
+Due to these issues, we dropped the MO and SD garden from subsequent green-up analyses for the Midwest subpopulation, and we dropped the OK and NE gardens from subsequent Midwest flowering date analyses.
 
@@
 We also clumped effects before selection of both subsets so that signal strength was standardized across different LD blocks in the genome, and not limited to closely linked associations.
+
+The intuition behind taking a subset of relatively unlinked, randomly selected effects is that this will capture a realistic fraction of the null and non-null effects in the full dataset, and thus be an unbiased representation of the amount and types of signal in the full dataset.
+Mash then uses this subset of random effects to estimate the number of null effects in the dataset and shrink some estimates towards null effects, effectively, towards a covariance matrix with no effects.
+The intuition behind using a subset of relatively unlinked effects with strong effects, or low *p*-values, is that these effects are the most likely to end up with significant jointly re-estimated SNP effects, and every SNP in a linkage block is likely to have effects re-estimated similarly.
 
@@
 
-The code used to run *mash* can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash_using_greedy_covar.qmd, starting at line 1326.
+The code used to run *mash* can be found at git\@github.com:Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v1.2_mash\_ using_greedy_covar.qmd, starting at line 1340.
 

### 2024-07-17 manuscript/revision2/revision2.qmd
revision2: Sync cross-references between revision, SI Appendix, and response letter.
Also, add Figures to response letter and change templating to allow this

@@
 degree of winter chilling, day length, temperature, and water availability) that signal existing or upcoming growing conditions [@bauerle_photoperiodic_2012; @andres2012genetic; @korner2010phenology; @maple2024].
-Genetic responses to environmental cues determine the speed, timing, and energy apportioned to vegetative and reproductive growth and shape both the individual's lifespan and its lifetime production of viable seed.
+Genetic responses to environmental cues determine the speed, timing, and energy apportioned to vegetative and reproductive growth and shape the individual's physiological responses, lifespan, and lifetime production of viable seed.
 Day length (or photoperiod) is one of the most predictable environmental cues, and genetic sensitivity to photoperiod protects plants from potentially fatal consequences of phenological responses to temperature cues at the "wrong" time of year.
@@
 
-Flowering time, or the transition from vegetative to reproductive growth, is a common subject of GxE research [@blackman2013interacting; @henry2014transitions; @aagren2017adaptive; @brachi2010linkage; @dittmar2014flowering; @li2018genomic; @romero2017study; @bustos-korts2019], a key output of selection driving adaptation to local environments [@andres2012genetic; @wadgymar2017identifying; @blumel2015flowering; @fournier-level2016], and a selection target for crop improvement to adapt crops to local or future environments @jung2009flowering.
-Changing flowering responsiveness to photoperiod cues has allowed geographic range expansion and increased yields in several cereal species [@turner2005pseudo; @faure2012mutation; @hung2012zmcct; @zakhrabekova2012induced; @yang2013oself3] and other crops [@pin2012multifaceted; @weller2019parallel].
+Flowering time, or the transition from vegetative to reproductive growth, is a common subject of GxE research [@blackman2013interacting; @henry2014transitions; @aagren2017adaptive; @brachi2010linkage; @dittmar2014flowering; @li2018genomic; @romero2017study; @bustos-korts2019], a key output of selection driving adaptation to local environments [@unterseer2016; @andres2012genetic; @wadgymar2017identifying; @blumel2015flowering; @fournier-level2016], and a selection target for crop improvement to adapt crops to local or future environments @jung2009flowering.
+Changing flowering responsiveness to photoperiod cues has allowed geographic range expansion and increased yields in several cereal species [@unterseer2016; @turner2005pseudo; @faure2012mutation; @hung2012zmcct; @zakhrabekova2012induced; @yang2013oself3] and other crops [@pin2012multifaceted; @weller2019parallel].
 Recent statistical advances in studying phenological GxE have involved determining critical environmental indices before the phenological event occurs, such as photothermal time within a critical growth window @li2018genomic.
@@
 However, as part of its wide environmental adaptation across the eastern half of North America, its photoperiodicity has been predicted to differ by plant latitude of origin [@parrish2005biology; @casler2004latitudinal].
-We previously found divergent Midwest and Gulf genetic subpopulations of switchgrass with distinct sets of environmental adaptations, in that both populations had distinct genetic variation associated with each of two fitness proxies, biomass and overwinter survival @lovell2021genomic.
-The Midwest genetic subpopulation is primarily composed of individuals from the well-studied upland switchgrass ecotype [@porter1966analysis; @milano2016genetic], while the Gulf subpopulation has individuals from the well-studied lowland ecotype and the phenotypically intermediate coastal ecotype @lovell2021genomic.
+We previously found divergent Midwest and Gulf genetic subpopulations of switchgrass with distinct sets of environmental adaptations, in that both populations had distinct genetic variation associated with each of two fitness proxies, biomass and overwinter survival @lovell2021.
+The Midwest genetic subpopulation is primarily composed of individuals from the well-studied upland switchgrass ecotype [@porter1966analysis; @milano2016genetic], while the Gulf subpopulation has individuals from the well-studied lowland ecotype and the phenotypically intermediate coastal ecotype @lovell2021.
 
@@
 We expected to observe common rank-changing GxE at the level of individual loci as this is a key theoretical prediction of local adaptation [@felsenstein1976theoretical; @levene1953genetic; @kawecki2004conceptual; @hedrick1986genetic].
-To determine the frequency of rank-changing GxE, we used the local false sign rate (lfsr), an analogue of the lfdr that establishes confidence in the effect sign, not the effect's difference from zero, to determine significance.
+To determine the frequency of rank-changing GxE, we used the local false sign rate (lfsr), an analogue of the local false discovery rate that establishes confidence in the effect sign, not the effect's difference from zero, to determine significance.
 We required lfsr significance (p \< 0.05) in both gardens to include effects.
@@
 
-The formation and resequencing of the diversity panel has been described previously @lovell2021genomic.
+The formation and resequencing of the diversity panel has been described previously @lovell2021.
 The diversity panel contained 134 sequenced, clonally propagated individuals from the Midwest genetic subpopulation, and 229 from the Gulf genetic subpopulation.
 To allow for the possibility that different subpopulations had different strengths of connection between our phenotypes and genotypes [@kortefarlow2013], we conducted three sets of genetic analyses: on Gulf and Midwest genotypes separately, and on both subpopulations together ('Both' subpopulations).
-Analyses to determine narrow-sense heritability (h^2^) for green-up and flowering were done using linear mixed models and followed @lovell2021genomic (SI Appendix, Section S6).
+Analyses to determine narrow-sense heritability (h^2^) for green-up and flowering were done using linear mixed models and followed @lovell2021 (SI Appendix, Section S6).
 
@@
 
-Using the lfsr rather than the local false discovery rate (lfdr) is a critical change in our ability to detect alleles directly contributing to rank-changing GxE between environments.
-The lfdr, like other measures of FDR, focuses on if we have enough evidence to reject the null hypothesis that an effect j is 0, or that there is a significant effect.
+Using the lfsr rather than the local false discovery rate (local FDR) is a critical change in our ability to detect alleles directly contributing to rank-changing GxE between environments.
+The local FDR, like other measures of FDR, focuses on if we have enough evidence to reject the null hypothesis that an effect j is 0, or that there is a significant effect.
 Previous studies of antagonistic pleiotropy (e.g.
-[@lowry2019qtl]) have used the lfdr or equivalent statistical tests to detect antagonistic pleiotropy.
+[@lowry2019qtl]) have used the local FDR or equivalent statistical tests to detect antagonistic pleiotropy.
 These tests were conservative, in that they required two non-zero effects of different signs, while tests for differential sensitivity required only one non-zero effect.

### 2024-07-17 manuscript/revision2/revision2.tex
revision2: Sync cross-references between revision, SI Appendix, and response letter.
Also, add Figures to response letter and change templating to allow this

@@
 environmental cues determine the speed, timing, and energy apportioned
-to vegetative and reproductive growth and shape both the individual's
-lifespan and its lifetime production of viable seed. Day length (or
-photoperiod) is one of the most predictable environmental cues, and
-genetic sensitivity to photoperiod protects plants from potentially
-fatal consequences of phenological responses to temperature cues at the
-``wrong'' time of year. However, the usefulness of specific
-environmental cues depends on both features of the environment, such as
-cue predictability and relevance, and the species' adaptive strategies
-(5). Species with wide natural distributions can have multiple distinct
-environmentally cued phenological responses: for example, populations of
-sunflower (\emph{Helianthus annuus}) exhibit day-neutral, facultative
-short day, and facultative long-day flowering responses, which vary with
-their environments (6, 7). Distinct genetic responses in different
-environments are known as genotype-by-environment interactions, or GxE.
+to vegetative and reproductive growth and shape the individual's
+physiological responses, lifespan, and lifetime production of viable
+seed. Day length (or photoperiod) is one of the most predictable
+environmental cues, and genetic sensitivity to photoperiod protects
+plants from potentially fatal consequences of phenological responses to
+temperature cues at the ``wrong'' time of year. However, the usefulness
+of specific environmental cues depends on both features of the
+environment, such as cue predictability and relevance, and the species'
+adaptive strategies (5). Species with wide natural distributions can
+have multiple distinct environmentally cued phenological responses: for
+example, populations of sunflower (\emph{Helianthus annuus}) exhibit
+day-neutral, facultative short day, and facultative long-day flowering
+responses, which vary with their environments (6, 7). Distinct genetic
+responses in different environments are known as genotype-by-environment
+interactions, or GxE.
 
@@
 growth, is a common subject of GxE research (6--13), a key output of
-selection driving adaptation to local environments (2, 14--16), and a
+selection driving adaptation to local environments (2, 14--17), and a
 selection target for crop improvement to adapt crops to local or future
-environments (17). Changing flowering responsiveness to photoperiod cues
+environments (18). Changing flowering responsiveness to photoperiod cues
 has allowed geographic range expansion and increased yields in several
-cereal species (18--22) and other crops (23, 24). Recent statistical
+cereal species (14, 19--23) and other crops (24, 25). Recent statistical
 advances in studying phenological GxE have involved determining critical
@@
 different genetic subpopulations, and even different genomic regions,
-have likely evolved distinct patterns of GxE (25). Additionally, despite
+have likely evolved distinct patterns of GxE (26). Additionally, despite
 theoretical predictions that local adaptation should involve
 antagonistic pleiotropy, or alleles with effects with opposing fitness
-outcomes (26--29), previous empirical work has found limited evidence of
-antagonistic pleiotropy (14, 30, 31). However, this work has been
+outcomes (27--30), previous empirical work has found limited evidence of
+antagonistic pleiotropy (15, 31, 32). However, this work has been
 limited by a known statistical bias that reduced detection of genetic
-effects that differ in sign (30, 32, 33). Thus, despite substantial
+effects that differ in sign (31, 33, 34). Thus, despite substantial
 interest in the frequencies of various forms of GxE, the frequency of
@@
 subpopulation. Switchgrass is considered a short-day plant with
-reproductive development strongly linked to day of the year (34).
+reproductive development strongly linked to day of the year (35).
 However, as part of its wide environmental adaptation across the eastern
 half of North America, its photoperiodicity has been predicted to differ
-by plant latitude of origin (35, 36). We previously found divergent
+by plant latitude of origin (36, 37). We previously found divergent
 Midwest and Gulf genetic subpopulations of switchgrass with distinct
@@
 genetic variation associated with each of two fitness proxies, biomass
-and overwinter survival (37). The Midwest genetic subpopulation is
+and overwinter survival (38). The Midwest genetic subpopulation is
 primarily composed of individuals from the well-studied upland
-switchgrass ecotype (38, 39), while the Gulf subpopulation has
+switchgrass ecotype (39, 40), while the Gulf subpopulation has
 individuals from the well-studied lowland ecotype and the phenotypically
-intermediate coastal ecotype (37).
+intermediate coastal ecotype (38).
 
@@
 set of these covariance matrices that significantly improved the modeled
-log-likelihood when included (SI Appendix, Section S2-S5) (40). We use
+log-likelihood when included (SI Appendix, Section S2-S5) (41). We use
 the Bayesian framework \emph{mash} (multivariate adaptive shrinkage)
-developed by (40), to refine effect size estimates from genome-wide
+developed by (41), to refine effect size estimates from genome-wide
 association (GWAS). \emph{Mash} allows us to identify and specify
@@
 biases in detecting genetic effects with the same or opposite signs
-(41). To confirm our genetic mapping of GxWeather, we compare genomic
+(42). To confirm our genetic mapping of GxWeather, we compare genomic
 locations of the significant posterior effect estimates from \emph{mash}
@@
 individual loci as this is a key theoretical prediction of local
-adaptation (26--29). To determine the frequency of rank-changing GxE, we
-used the local false sign rate (lfsr), an analogue of the lfdr that
-establishes confidence in the effect sign, not the effect's difference
-from zero, to determine significance. We required lfsr significance (p
-\textless{} 0.05) in both gardens to include effects. This means that
-our tests for a sign change between gardens carry an equal statistical
-burden to those for effects with the same sign. We separated kinds of
-effects at the level of individual loci into SNP effects that differ in
-sign between gardens (effects with rank-changing GxE between gardens),
-SNP effects that differ in magnitude (effects that are large in one
-garden or region and smaller in others), and SNP effects that are
-indistinguishable in two gardens (similarly large or small in
-comparisons between gardens) (Figure~\ref{fig-effects} A).
+adaptation (27--30). To determine the frequency of rank-changing GxE, we
+used the local false sign rate (lfsr), an analogue of the local false
+discovery rate that establishes confidence in the effect sign, not the
+effect's difference from zero, to determine significance. We required
+lfsr significance (p \textless{} 0.05) in both gardens to include
+effects. This means that our tests for a sign change between gardens
+carry an equal statistical burden to those for effects with the same
+sign. We separated kinds of effects at the level of individual loci into
+SNP effects that differ in sign between gardens (effects with
+rank-changing GxE between gardens), SNP effects that differ in magnitude
+(effects that are large in one garden or region and smaller in others),
+and SNP effects that are indistinguishable in two gardens (similarly
+large or small in comparisons between gardens) (Figure~\ref{fig-effects}
+A).
 
@@
 overlapped a gene recently functionally validated for flowering in
-switchgrass (42). All flowering and green-up QTL intervals contained at
+switchgrass (43). All flowering and green-up QTL intervals contained at
 least one SNP significant in at least one \emph{mash} run at a
@@
 should involve antagonistic pleiotropy at the level of individual loci
-(26--29). In addition, it advances local adaptation research by using
+(27--30). In addition, it advances local adaptation research by using
 more than two field sites and a wide range of genetic variation to
 determine that rank-changing GxE for phenological timings is common in
-small genomic regions (14, 43, 44).
+small genomic regions (15, 44, 45).
 
@@
 populations of switchgrass, a diversity panel and a pseudo-F2 cross,
-planted at eight common garden locations (39, 44, 45). We scored the
+planted at eight common garden locations (38, 40, 45). We scored the
 onset of vegetative growth, or green-up date, as the day of the year
@@
 The formation and resequencing of the diversity panel has been described
-previously (37). The diversity panel contained 134 sequenced, clonally
+previously (38). The diversity panel contained 134 sequenced, clonally
 propagated individuals from the Midwest genetic subpopulation, and 229
@@
 narrow-sense heritability (h\textsuperscript{2}) for green-up and
-flowering were done using linear mixed models and followed (37) (SI
+flowering were done using linear mixed models and followed (38) (SI
 Appendix, Section S6).
@@
 between four individuals, two Midwest and two Gulf individuals. The
-formation of this mapping population has been described previously (39);
+formation of this mapping population has been described previously (40);
 additional details on QTL mapping can be found in SI Appendix, Section
@@
 
-Using the lfsr rather than the local false discovery rate (lfdr) is a
-critical change in our ability to detect alleles directly contributing
-to rank-changing GxE between environments. The lfdr, like other measures
-of FDR, focuses on if we have enough evidence to reject the null
-hypothesis that an effect j is 0, or that there is a significant effect.
-Previous studies of antagonistic pleiotropy (e.g. (44)) have used the
-lfdr or equivalent statistical tests to detect antagonistic pleiotropy.
-These tests were conservative, in that they required two non-zero
-effects of different signs, while tests for differential sensitivity
-required only one non-zero effect. This previous work recognized that
-this testing bias could lead to undercounting occurrences of
-antagonistic pleiotropy (30, 33), and sought to reduce it by permutation
-(32). However, using the lfsr to test for allelic effects that differ in
-sign does not undercount these occurences, as this statistic answers a
-fundamentally different question. For each effect \(j\), the \(lfsr_j\)
-is defined as the probability that we make an error in the sign of
-effect \(j\) if we were forced to declare the effect positive or
-negative (41). Thus, rather than asking ``Are these two effects
-different?'' - as we reasonably expect two effects to be, even if this
-difference cannot be measured - the local false sign rate answers a more
-meaningful question: Can we be confident in the sign of this effect?
+Using the lfsr rather than the local false discovery rate (local FDR) is
+a critical change in our ability to detect alleles directly contributing
+to rank-changing GxE between environments. The local FDR, like other
+measures of FDR, focuses on if we have enough evidence to reject the
+null hypothesis that an effect j is 0, or that there is a significant
+effect. Previous studies of antagonistic pleiotropy (e.g. (45)) have
+used the local FDR or equivalent statistical tests to detect
+antagonistic pleiotropy. These tests were conservative, in that they
+required two non-zero effects of different signs, while tests for
+differential sensitivity required only one non-zero effect. This
+previous work recognized that this testing bias could lead to
+undercounting occurrences of antagonistic pleiotropy (31, 34), and
+sought to reduce it by permutation (33). However, using the lfsr to test
+for allelic effects that differ in sign does not undercount these
+occurences, as this statistic answers a fundamentally different
+question. For each effect \(j\), the \(lfsr_j\) is defined as the
+probability that we make an error in the sign of effect \(j\) if we were
+forced to declare the effect positive or negative (42). Thus, rather
+than asking ``Are these two effects different?'' - as we reasonably
+expect two effects to be, even if this difference cannot be measured -
+the local false sign rate answers a more meaningful question: Can we be
+confident in the sign of this effect?
 
@@
 
+\bibitem[\citeproctext]{ref-unterseer2016}
+\CSLLeftMargin{14. }%
+\CSLRightInline{S. Unterseer, \emph{et al.},
+\href{https://doi.org/10.1186/s13059-016-1009-x}{A comprehensive study
+of the genomic differentiation between temperate Dent and Flint maize}.
+\emph{Genome Biology} \textbf{17} (2016).}
+
 \bibitem[\citeproctext]{ref-wadgymar2017identifying}
-\CSLLeftMargin{14. }%
+\CSLLeftMargin{15. }%
 \CSLRightInline{S. M. Wadgymar, \emph{et al.}, Identifying targets and
@@
 \bibitem[\citeproctext]{ref-blumel2015flowering}
-\CSLLeftMargin{15. }%
+\CSLLeftMargin{16. }%
 \CSLRightInline{M. Blümel, N. Dally, C. Jung, Flowering time regulation
@@
 \bibitem[\citeproctext]{ref-fournier-level2016}
-\CSLLeftMargin{16. }%
+\CSLLeftMargin{17. }%
 \CSLRightInline{A. Fournier-Level, \emph{et al.},
@@
 \bibitem[\citeproctext]{ref-jung2009flowering}
-\CSLLeftMargin{17. }%
+\CSLLeftMargin{18. }%
 \CSLRightInline{C. Jung, A. E. Müller, Flowering time control and
@@
 \bibitem[\citeproctext]{ref-turner2005pseudo}
-\CSLLeftMargin{18. }%
+\CSLLeftMargin{19. }%
 \CSLRightInline{A. Turner, J. Beales, S. Faure, R. P. Dunford, D. A.
@@
 \bibitem[\citeproctext]{ref-faure2012mutation}
-\CSLLeftMargin{19. }%
+\CSLLeftMargin{20. }%
 \CSLRightInline{S. Faure, \emph{et al.}, Mutation at the circadian clock
@@
 \bibitem[\citeproctext]{ref-hung2012zmcct}
-\CSLLeftMargin{20. }%
+\CSLLeftMargin{21. }%
 \CSLRightInline{H.-Y. Hung, \emph{et al.}, ZmCCT and the genetic basis
@@
 \bibitem[\citeproctext]{ref-zakhrabekova2012induced}
-\CSLLeftMargin{21. }%
+\CSLLeftMargin{22. }%
 \CSLRightInline{S. Zakhrabekova, \emph{et al.}, Induced mutations in
@@
 \bibitem[\citeproctext]{ref-yang2013oself3}
-\CSLLeftMargin{22. }%
+\CSLLeftMargin{23. }%
 \CSLRightInline{Y. Yang, Q. Peng, G.-X. Chen, X.-H. Li, C.-Y. Wu, OsELF3
@@
 \bibitem[\citeproctext]{ref-pin2012multifaceted}
-\CSLLeftMargin{23. }%
+\CSLLeftMargin{24. }%
 \CSLRightInline{P. Pin, O. Nilsson, The multifaceted roles of FLOWERING
@@
 \bibitem[\citeproctext]{ref-weller2019parallel}
-\CSLLeftMargin{24. }%
+\CSLLeftMargin{25. }%
 \CSLRightInline{J. L. Weller, \emph{et al.}, Parallel origins of
@@
 \bibitem[\citeproctext]{ref-Weine2023.06.21.545998}
-\CSLLeftMargin{25. }%
+\CSLLeftMargin{26. }%
 \CSLRightInline{E. Weine, S. P. Smith, R. K. Knowlton, A. Harpak,
@@
 \bibitem[\citeproctext]{ref-levene1953genetic}
-\CSLLeftMargin{26. }%
+\CSLLeftMargin{27. }%
 \CSLRightInline{H. Levene, Genetic equilibrium when more than one
@@
 \bibitem[\citeproctext]{ref-felsenstein1976theoretical}
-\CSLLeftMargin{27. }%
+\CSLLeftMargin{28. }%
 \CSLRightInline{J. Felsenstein, The theoretical population genetics of
@@
 \bibitem[\citeproctext]{ref-kawecki2004conceptual}
-\CSLLeftMargin{28. }%
+\CSLLeftMargin{29. }%
 \CSLRightInline{T. J. Kawecki, D. Ebert, Conceptual issues in local
@@
 \bibitem[\citeproctext]{ref-hedrick1986genetic}
-\CSLLeftMargin{29. }%
+\CSLLeftMargin{30. }%
 \CSLRightInline{P. W. Hedrick, Genetic polymorphism in heterogeneous
@@
 \bibitem[\citeproctext]{ref-des2013genotype}
-\CSLLeftMargin{30. }%
+\CSLLeftMargin{31. }%
 \CSLRightInline{D. L. Des Marais, K. M. Hernandez, T. E. Juenger,
@@
 \bibitem[\citeproctext]{ref-millet2016}
-\CSLLeftMargin{31. }%
+\CSLLeftMargin{32. }%
 \CSLRightInline{E. Millet, \emph{et al.},
@@
 \bibitem[\citeproctext]{ref-anderson2013genetic}
-\CSLLeftMargin{32. }%
+\CSLLeftMargin{33. }%
 \CSLRightInline{J. T. Anderson, C.-R. Lee, C. A. Rushworth, R. I.
@@
 \bibitem[\citeproctext]{ref-anderson2011evolutionary}
-\CSLLeftMargin{33. }%
+\CSLLeftMargin{34. }%
 \CSLRightInline{J. T. Anderson, J. H. Willis, T. Mitchell-Olds,
@@
 \bibitem[\citeproctext]{ref-mitchell1997predicting}
-\CSLLeftMargin{34. }%
+\CSLLeftMargin{35. }%
 \CSLRightInline{R. B. Mitchell, K. J. Moore, L. E. Moser, J. O. Fritz,
@@
 \bibitem[\citeproctext]{ref-parrish2005biology}
-\CSLLeftMargin{35. }%
+\CSLLeftMargin{36. }%
 \CSLRightInline{D. J. Parrish, J. H. Fike, The biology and agronomy of
@@
 \bibitem[\citeproctext]{ref-casler2004latitudinal}
-\CSLLeftMargin{36. }%
+\CSLLeftMargin{37. }%
 \CSLRightInline{M. Casler, K. P. Vogel, C. Taliaferro, R. Wynia,
@@
 
-\bibitem[\citeproctext]{ref-lovell2021genomic}
-\CSLLeftMargin{37. }%
-\CSLRightInline{J. T. Lovell, \emph{et al.}, Genomic mechanisms of
-climate adaptation in polyploid bioenergy switchgrass. \emph{Nature}
-\textbf{590}, 438--444 (2021).}
+\bibitem[\citeproctext]{ref-lovell2021}
+\CSLLeftMargin{38. }%
+\CSLRightInline{J. T. Lovell, \emph{et al.},
+\href{https://doi.org/10.1038/s41586-020-03127-1}{Genomic mechanisms of
+climate adaptation in polyploid bioenergy switchgrass}. \emph{Nature},
+1--7 (2021).}
 
 \bibitem[\citeproctext]{ref-porter1966analysis}
-\CSLLeftMargin{38. }%
+\CSLLeftMargin{39. }%
 \CSLRightInline{C. L. Porter Jr, An analysis of variation between upland
@@
 \bibitem[\citeproctext]{ref-milano2016genetic}
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{40. }%
 \CSLRightInline{E. R. Milano, D. B. Lowry, T. E. Juenger, The genetic
@@
 \bibitem[\citeproctext]{ref-urbut2019flexible}
-\CSLLeftMargin{40. }%
+\CSLLeftMargin{41. }%
 \CSLRightInline{S. M. Urbut, G. Wang, P. Carbonetto, M. Stephens,
@@
 \bibitem[\citeproctext]{ref-10.1093ux2fbiostatisticsux2fkxw041}
-\CSLLeftMargin{41. }%
+\CSLLeftMargin{42. }%
 \CSLRightInline{M. Stephens,
@@
 \bibitem[\citeproctext]{ref-choi2023}
-\CSLLeftMargin{42. }%
+\CSLLeftMargin{43. }%
 \CSLRightInline{S. Choi, \emph{et al.},
@@
 \bibitem[\citeproctext]{ref-savolainen2013ecological}
-\CSLLeftMargin{43. }%
+\CSLLeftMargin{44. }%
 \CSLRightInline{O. Savolainen, M. Lascoux, J. Merilä, Ecological
@@
 \bibitem[\citeproctext]{ref-lowry2019qtl}
-\CSLLeftMargin{44. }%
+\CSLLeftMargin{45. }%
 \CSLRightInline{D. B. Lowry, \emph{et al.}, QTL\(\times\) environment
@@
 Sciences} \textbf{116}, 12933--12941 (2019).}
-
-\bibitem[\citeproctext]{ref-lovell2021}
-\CSLLeftMargin{45. }%
-\CSLRightInline{J. T. Lovell, \emph{et al.},
-\href{https://doi.org/10.1038/s41586-020-03127-1}{Genomic mechanisms of
-climate adaptation in polyploid bioenergy switchgrass}. \emph{Nature},
-1--7 (2021).}
 

### 2024-07-30 manuscript/revision2/revision2.qmd
revision2: Coauthor line edits from SS+FM+RM+DL+AH+FF

@@
 abstract: |
-  The timing of vegetative and reproductive growth in plants ("phenological timings") depends on genetic effects, environmental (e.g., weather) cues, and possibly their interaction.
-Quantifying these GxWeather interactions can aid prediction and manipulation of phenological timings.
+  The timing of vegetative and reproductive growth in plants ("phenological timings") depends on genetic effects (G), environmental (e.g., weather) cues, and their interaction.
+Measuring GxWeather interactions can aid prediction and manipulation of phenological timings.
 Here, we map GxWeather effects on phenological timings in two highly divergent switchgrass (*Panicum virgatum*) populations using repeated plantings of cloned individuals from these populations at eight sites across the central United States.
@@
 For example, in the Gulf population, 65% of genetic effects on the timing of vegetative growth covary with daylength 14 days prior to green-up date, and 33% of genetic effects on the timing of flowering covary with cumulative rainfall in the seven days prior to flowering.
-However, most variation in genetic effects is site-specific and can not be attributed to variation in the studied weather variables.
+However, most variation in genetic effects is site-specific and cannot be attributed to variation in weather variables.
 We demonstrate that we can identify genetic variation with GxWeather and assign these loci to specific weather-based cues or other patterns.
-Breeding for particular alleles at the GxWeather loci we identified could change flowering responsiveness in a photoperiod or rainfall-specific way.
-More broadly, our approach proposes a refined, powerful characterization of genotype-by-environment interactions in any species phenotyped in multiple environments.
+We predict that selective breeding for particular alleles at the GxWeather loci could alter flowering responsiveness in a photoperiod or rainfall-specific way.
+More broadly, our approach proposes a refined characterization of genotype-by-environment interactions in any species phenotyped in multiple environments.
 runningauthor: "MacQueen"
@@
 A.H.M., L.Z., and S.A.S.
-conducted statistical and computational analyses.
+conducted statistical and computational analyses with input from T.E.J.
+and A.H.
 The manuscript was written by A.H.M.
@@
 This material is based upon work supported in part by the Great Lakes Bioenergy Research Center, U.S.
-Department of Energy, Office of Science, Office of Biological and Environmental Research under Award Numbers DE-SC0018409 and DE-FC02-07ER64494, the US Department of Energy Awards DESC0014156 to T.E.J., DE-SC0017883 to D.B.L, National Science Foundation PGRP Awards IOS0922457 and IOS1444533 to T.E.J, and the Long-term Ecological Research Program (DEB 1832042) at the Kellogg Biological Station.
+Department of Energy, Office of Science, Office of Biological and Environmental Research under Award Numbers DE-SC0018409 and DE-FC02-07ER64494, the US Department of Energy Awards DESC0014156 to T.E.J., DE-SC0017883 to D.B.L, NIH grant R35GM151108 and a Pew Scholarship to A.H., National Science Foundation PGRP Awards IOS0922457 and IOS1444533 to T.E.J, and the Long-term Ecological Research Program (DEB 1832042) at the Kellogg Biological Station.
 bibliography: bibliography.bib 
@@
 
-Plant phenological timings are major components of plant fitness affected by multiple external environmental cues (e.g.
-degree of winter chilling, day length, temperature, and water availability) that signal existing or upcoming growing conditions [@bauerle_photoperiodic_2012; @andres2012genetic; @korner2010phenology; @maple2024].
-Genetic responses to environmental cues determine the speed, timing, and energy apportioned to vegetative and reproductive growth and shape the individual's physiological responses, lifespan, and lifetime production of viable seed.
+Plant phenological traits are important components of plant fitness that are affected by multiple external environmental cues (e.g.
+degree of winter chilling, day length, temperature, soil fertility, and water availability), which signal existing or upcoming growing conditions [@bauerle_photoperiodic_2012; @andres2012genetic; @korner2010phenology; @maple2024].
+Genetic responses to environmental cues determine the speed, timing, and energy apportioned to vegetative and reproductive growth and shape plant physiological responses, lifespan, and lifetime production of viable seed.
 Day length (or photoperiod) is one of the most predictable environmental cues, and genetic sensitivity to photoperiod protects plants from potentially fatal consequences of phenological responses to temperature cues at the "wrong" time of year.
-However, the usefulness of specific environmental cues depends on both features of the environment, such as cue predictability and relevance, and the species' adaptive strategies @botero2015evolutionary.
-Species with wide natural distributions can have multiple distinct environmentally cued phenological responses: for example, populations of sunflower (*Helianthus annuus*) exhibit day-neutral, facultative short day, and facultative long-day flowering responses, which vary with their environments [@blackman2013interacting; @henry2014transitions].
-Distinct genetic responses in different environments are known as genotype-by-environment interactions, or GxE.
+However, the utility of specific environmental cues depends on both features of the environment, such as cue predictability and relevance, and the species' adaptive strategies @botero2015evolutionary.
+Species with wide natural distributions can have multiple distinct environmentally cued phenological responses.
+For example, populations of sunflower (*Helianthus annuus*) exhibit day-neutral, facultative short day, and facultative long-day flowering responses, which vary with their environments [@blackman2013interacting; @henry2014transitions].
+These distinct genetic responses in different environments are known as genotype-by-environment interactions, or GxE.
 
@@
 However, as part of its wide environmental adaptation across the eastern half of North America, its photoperiodicity has been predicted to differ by plant latitude of origin [@parrish2005biology; @casler2004latitudinal].
-We previously found divergent Midwest and Gulf genetic subpopulations of switchgrass with distinct sets of environmental adaptations, in that both populations had distinct genetic variation associated with each of two fitness proxies, biomass and overwinter survival @lovell2021.
+We previously found that divergent Midwest and Gulf genetic subpopulations of switchgrass have distinct sets of environmental adaptations associated with each of two fitness proxies, biomass and overwinter survival @lovell2021.
 The Midwest genetic subpopulation is primarily composed of individuals from the well-studied upland switchgrass ecotype [@porter1966analysis; @milano2016genetic], while the Gulf subpopulation has individuals from the well-studied lowland ecotype and the phenotypically intermediate coastal ecotype @lovell2021.
 
-Here, we test for GxWeather for two phenological timings by assigning patterns of genetic effects on phenology at eight common gardens to many patterns of weather covariance at these gardens.
-To do this, we phenotype a diversity panel of hundreds of switchgrass genotypes from the Midwest and Gulf subpopulations for the timing of vegetative and reproductive development.
-We do this at eight common garden locations spanning 17 degrees of latitude: these gardens cover the majority of the latitudinal and climatic range of switchgrass and capture the most comprehensive picture to date of the environmental variation this species encounters.
-We define multiple ways phenological timings might covary with weather (Table 1) and additional ways phenological timings might vary by garden (SI Appendix, Section S1), then jointly re-estimate genetic effects on these timings at all eight common gardens using the set of these covariance matrices that significantly improved the modeled log-likelihood when included (SI Appendix, Section S2-S5) @urbut2019flexible.
-We use the Bayesian framework *mash* (multivariate adaptive shrinkage) developed by @urbut2019flexible, to refine effect size estimates from genome-wide association (GWAS).
-*Mash* allows us to identify and specify multiple covariance structures among genetic effect estimates across sites, including structures that represent covariance in weather variables of interest.
-Importantly, this method circumvents statistical biases in detecting genetic effects with the same or opposite signs @10.1093/biostatistics/kxw041.
-To confirm our genetic mapping of GxWeather, we compare genomic locations of the significant posterior effect estimates from *mash* to mapping results from an outbred mapping population grown at the same sites.
-Our analyses allow us to describe the weather cues and types of GxE affecting phenology in two divergent natural populations of switchgrass across the species' latitudinal range.
+Here, we test for GxWeather for the timing of two phenological traits by loading patterns of genetic effects on phenology at eight common gardens onto many patterns of weather covariance at these gardens.
+To do this, we phenotyped a diversity panel of hundreds of switchgrass genotypes from the Midwest and Gulf subpopulations for the timing of vegetative and reproductive development.
+We did this at eight common garden locations spanning 17 degrees of latitude: these gardens covered the majority of the latitudinal and climatic range of switchgrass and captured the most comprehensive picture to date of the environmental variation this species encounters.
+We defined multiple ways phenological traits might covary with weather (Table 1) and additional ways phenological traits might vary by garden (SI Appendix, Section S1), then jointly re-estimated genetic effects on these timings at all eight common gardens using the set of these covariance matrices that significantly improved the modeled log-likelihood when included (SI Appendix, Section S2-S5) @urbut2019flexible.
+We used the Bayesian framework *mash* (multivariate adaptive shrinkage) developed by @urbut2019flexible, to refine effect size estimates from genome-wide association (GWAS).
+*Mash* allowed us to identify and specify multiple covariance structures among genetic effect estimates across sites, including structures that represent covariance in weather variables of interest.
+Importantly, this method circumvented statistical biases in detecting genetic effects with the same or opposite signs @10.1093/biostatistics/kxw041.
+To confirm our genetic mapping of GxWeather, we compared genomic locations of the significant posterior effect estimates from *mash* to mapping results from an outbred mapping population grown at the same sites.
+Our analyses allowed us to describe the weather cues and types of GxE affecting phenology in two divergent natural populations of switchgrass across the species' latitudinal range.
 
@@
 
-Genotypes from the Gulf and Midwest subpopulations had distinct phenological timings and distinct patterns of phenological correlations across our eight common garden sites (@fig-map).
+Genotypes from the Gulf and Midwest subpopulations had distinct phenological trait timings and distinct patterns of phenological trait correlations across our eight common garden sites (@fig-map).
 At the three Texas common gardens (hereafter 'Texas' gardens), located within the natural range of the Gulf subpopulation, the onset of vegetative growth, or 'green-up date', for Gulf genotypes occurred before Midwestern green-up, and the onset of reproductive growth, or 'flowering date', for Gulf genotypes occurred after Midwestern flowering (@fig-map A).
@@
 At the Oklahoma common garden, located near the natural range limits of both the Gulf and the Midwest subpopulations, Gulf and Midwest green-up occurred over the same period, and Gulf genotype flowering occurred after Midwestern flowering (@fig-map A).
-These patterns led to strong negative phenotypic correlations for the onset of vegetative growth between the North and Texas gardens, particularly in the Gulf and across all individuals from the Gulf and Midwest subpopulations (hereafter, 'Both' subpopulations), and contributed to positive phenotypic correlations for the onset of flowering that had larger magnitudes at more northern gardens (@fig-map B).
+These patterns led to strong negative phenotypic correlations for the onset of vegetative growth between the North and Texas gardens, particularly in the Gulf and in the population containing all individuals from the Gulf and Midwest subpopulations (hereafter, 'Both' subpopulations) and contributed to positive phenotypic correlations for the onset of flowering that had larger magnitudes at more northern gardens (@fig-map B).
 
@@
 h^2^ were typically high at individual gardens: 59% on average for green-up date, and 87% for flowering date.
-However, h^2^ were variable across gardens, and green-up dates were uncorrelated (r^2^ \< 0.2) or negatively correlated between pairs of gardens (@fig-map B).
+However, h^2^ was variable across gardens, and green-up dates were uncorrelated (r^2^ \< 0.2) or negatively correlated between pairs of gardens (@fig-map B).
 These negative and small correlations undoubtedly contributed to the low h^2^ values for green-up and flowering date when estimated jointly at all eight gardens: h^2^ was 0.8% for green-up and 23.2% for flowering date (SI Appendix, Figure S1).
@@
 
-We expected patterns of GxE and GxWeather to vary at the locus, subpopulation, and trait level; thus we were interested in modeling the different ways effects might covary across our eight common gardens.
-We jointly re-estimated genetic effects of a subset of SNPs across all eight common gardens using *mash* models, which can flexibly capture many modes of covariance across gardens (SI Appendix, Sections S1-S5, Datasets 1-6).
+We expected patterns of GxE and GxWeather to vary at the locus, subpopulation, and trait level.
+Thus, we were interested in modeling the different ways effects might covary across our eight common gardens.
+We jointly re-estimated the genetic effects of a subset of SNPs across all eight common gardens using *mash* models, which can flexibly capture many modes of covariance across gardens (SI Appendix, Sections S1-S5, Datasets 1-6).
 To obtain genetic effects, we used subsets of effects from garden- and subpopulation GWAS (SI Appendix, Section S2-S3; Table S1).
@@
 If these phenotypic correlations have a genetic component, they could be partially or completely captured by the covariance structures specified in the *mash* model.
-In this case, the joint estimate of effects for many SNPs could have high mixture proportions, or mass, on a covariance matrix that captures the negative correlation between Texas and North gardens.
-In our data, daylength 14 days prior to green-up date has negative correlations between Texas and North gardens (@fig-covar A); if many SNP effects have mass on this GxWeather matrix, we would infer that the effects of those SNPs have that form of GxWeather covariation.
+In this case, the joint estimate of effects for many SNPs could have high mixture proportions, or mass, on a covariance matrix that captures the negative correlation between Texas (TX1, TX2, and TX3) and North (MO, NE, MI, and SD) gardens.
+In our data, daylength 14 days prior to green-up date was negatively correlated between Texas and North gardens (@fig-covar A); if many SNP effects have mass on this GxWeather matrix, we would infer that the effects of those SNPs have that form of GxWeather covariation.
 
@@
 Both subpopulations had non-zero weights on two additional GxWeather covariance types, average temperature one day prior to green-up, and the day length change in seconds in the day prior to green-up (@fig-covar B).
-The average temperature covariance matrix had negative covariances between Texas and North gardens, though not as strong as the negative phenotypic correlations seen in Both subpopulations (@fig-map B; @fig-covar A).
+The average temperature covariance matrix had negative covariances between Texas and North gardens, though not as strong as the negative phenotypic correlations observed in Both subpopulations (@fig-map B; @fig-covar A).
 The Gulf subpopulation and Both subpopulations had substantially more mass on GxWeather covariance matrices than the Midwest population (@fig-covar C) for the onset of vegetative growth.
@@
 The majority of pairwise effects for greenup for the Midwest (\>73%; 68-342 effects) were indistinguishable, or had no GxE.
-The majority of pairwise effects for Both (\>85%) subpopulations were the same sign, and effects most often differed in magnitude between gardens within and between the regions; however, these differences in effect sign and magnitude were mostly between the MO site and other sites (@fig-effects B; SI Appendix, Fig S2).
+The majority of pairwise effects for Both subpopulations (\>85%) were the same sign, and effects most often differed in magnitude between gardens within and between the regions; however, these differences in effect sign and magnitude were mostly between the MO site and other sites (@fig-effects B; SI Appendix, Fig S2).
 
@@
 We conducted quantitative trait loci (QTL) mapping of flowering as functions of five environmental cues that we also used as covariance matrices in *mash*, and identified eight QTL for flowering date, eight QTL for flowering as a function of day length change two days prior, one QTL for the start of vegetative growth, and two QTL for vegetative growth as a function of daylength change one day prior, all of which showed QTL by environment interactions (SI Appendix, Figure S4).
-All QTL for flowering overlapped one or more homologs from rice or *A.
+All QTL for flowering overlapped with one or more homologs from rice or *A.
 thaliana* with functionally validated roles in flowering (SI Appendix, Dataset 7); the QTL on Chr04K overlapped a gene recently functionally validated for flowering in switchgrass [@choi2023].
@@
 Here, we demonstrate that we can associate multiple patterns of GxWeather with specific genomic regions using a switchgrass diversity panel grown at eight common gardens.
-We can assign genetic effects to both GxWeather patterns with interpretable weather-based cues, and to unmeasured, site-based patterns.
-We use this approach to study GxWeather for the timings of vegetative and reproductive development in the deeply genetically diverged Gulf and Midwest subpopulations of switchgrass.
+We assigned genetic effects to both GxWeather patterns with interpretable weather-based cues, and to unmeasured, site-based patterns.
+We used this approach to study GxWeather for the timings of vegetative and reproductive development in the deeply genetically diverged Gulf and Midwest subpopulations of switchgrass.
 
@@
 The Gulf & Both subpopulations had rank-changing GxE for green-up date, while the Midwest subpopulation had rank-changing GxE for flowering date.
-As phenological timings are major components of plant fitness, this result supports theoretical models that local adaptation should involve antagonistic pleiotropy at the level of individual loci [@kawecki2004conceptual; @felsenstein1976theoretical; @levene1953genetic; @hedrick1986genetic].
-In addition, it advances local adaptation research by using more than two field sites and a wide range of genetic variation to determine that rank-changing GxE for phenological timings is common in small genomic regions [@wadgymar2017identifying; @savolainen2013ecological; @lowry2019qtl].
+As phenological timings are major components of plant fitness, this result supports theoretical models that local adaptation should involve trade-offs at the level of individual loci [@kawecki2004conceptual; @felsenstein1976theoretical; @levene1953genetic; @hedrick1986genetic].
+In addition, it advances local adaptation research by using more than two field sites and a wide range of genetic variation to determine that rank-changing GxE for phenological traits is common in small genomic regions [@wadgymar2017identifying; @savolainen2013ecological; @lowry2019qtl].
 
@@
 Common gardens are arranged in latitudinal order.
-(b) Phenotypic correlations between clonal replicates planted at eight common gardens, within and between two genetic subpopulations.](images/Figure_1_AB.png){#fig-map}
+(b) Phenotypic correlations between clonal replicates planted at eight common gardens, within and between two genetic subpopulations.](images/Fig1_AB.png){#fig-map}
 
@@
 
-![Types of GxE present between of pairs of jointly re-estimated effects in eight common gardens, for effects with lfsr \< 0.05.
+![Types of GxE present between of pairs of jointly re-estimated effects in eight common gardens, for effects with lfsr \< 0.05 at each pair of sites contrasted.
 a) Examples of effect patterns at three pairs of sites with three types of GxE.

### 2024-07-30 manuscript/revision2/revision2.tex
revision2: Coauthor line edits from SS+FM+RM+DL+AH+FF

@@
 and T.E.J. hosted field experiments. A.H.M., L.Z., and S.A.S. conducted
-statistical and computational analyses. The manuscript was written by
-A.H.M. with contributions from all authors.}
+statistical and computational analyses with input from T.E.J. and A.H.
+The manuscript was written by A.H.M. with contributions from all
+authors.}
 
@@
 The timing of vegetative and reproductive growth in plants
-(``phenological timings'') depends on genetic effects, environmental
-(e.g., weather) cues, and possibly their interaction. Quantifying these
-GxWeather interactions can aid prediction and manipulation of
-phenological timings. Here, we map GxWeather effects on phenological
-timings in two highly divergent switchgrass (\emph{Panicum virgatum})
-populations using repeated plantings of cloned individuals from these
-populations at eight sites across the central United States. Some
-genetic effects covary with interpretable weather-based cues. For
-example, in the Gulf population, 65\% of genetic effects on the timing
-of vegetative growth covary with daylength 14 days prior to green-up
-date, and 33\% of genetic effects on the timing of flowering covary with
-cumulative rainfall in the seven days prior to flowering. However, most
-variation in genetic effects is site-specific and can not be attributed
-to variation in the studied weather variables. We demonstrate that we
-can identify genetic variation with GxWeather and assign these loci to
-specific weather-based cues or other patterns. Breeding for particular
-alleles at the GxWeather loci we identified could change flowering
-responsiveness in a photoperiod or rainfall-specific way. More broadly,
-our approach proposes a refined, powerful characterization of
-genotype-by-environment interactions in any species phenotyped in
-multiple environments.
+(``phenological timings'') depends on genetic effects (G), environmental
+(e.g., weather) cues, and their interaction. Measuring GxWeather
+interactions can aid prediction and manipulation of phenological
+timings. Here, we map GxWeather effects on phenological timings in two
+highly divergent switchgrass (\emph{Panicum virgatum}) populations using
+repeated plantings of cloned individuals from these populations at eight
+sites across the central United States. Some genetic effects covary with
+interpretable weather-based cues. For example, in the Gulf population,
+65\% of genetic effects on the timing of vegetative growth covary with
+daylength 14 days prior to green-up date, and 33\% of genetic effects on
+the timing of flowering covary with cumulative rainfall in the seven
+days prior to flowering. However, most variation in genetic effects is
+site-specific and cannot be attributed to variation in weather
+variables. We demonstrate that we can identify genetic variation with
+GxWeather and assign these loci to specific weather-based cues or other
+patterns. We predict that selective breeding for particular alleles at
+the GxWeather loci could alter flowering responsiveness in a photoperiod
+or rainfall-specific way. More broadly, our approach proposes a refined
+characterization of genotype-by-environment interactions in any species
+phenotyped in multiple environments.
 \end{abstract}
@@
 
-Plant phenological timings are major components of plant fitness
-affected by multiple external environmental cues (e.g.~degree of winter
-chilling, day length, temperature, and water availability) that signal
-existing or upcoming growing conditions (1--4). Genetic responses to
-environmental cues determine the speed, timing, and energy apportioned
-to vegetative and reproductive growth and shape the individual's
-physiological responses, lifespan, and lifetime production of viable
-seed. Day length (or photoperiod) is one of the most predictable
-environmental cues, and genetic sensitivity to photoperiod protects
-plants from potentially fatal consequences of phenological responses to
-temperature cues at the ``wrong'' time of year. However, the usefulness
-of specific environmental cues depends on both features of the
-environment, such as cue predictability and relevance, and the species'
-adaptive strategies (5). Species with wide natural distributions can
-have multiple distinct environmentally cued phenological responses: for
-example, populations of sunflower (\emph{Helianthus annuus}) exhibit
-day-neutral, facultative short day, and facultative long-day flowering
-responses, which vary with their environments (6, 7). Distinct genetic
-responses in different environments are known as genotype-by-environment
-interactions, or GxE.
+Plant phenological traits are important components of plant fitness that
+are affected by multiple external environmental cues (e.g.~degree of
+winter chilling, day length, temperature, soil fertility, and water
+availability), which signal existing or upcoming growing conditions
+(1--4). Genetic responses to environmental cues determine the speed,
+timing, and energy apportioned to vegetative and reproductive growth and
+shape plant physiological responses, lifespan, and lifetime production
+of viable seed. Day length (or photoperiod) is one of the most
+predictable environmental cues, and genetic sensitivity to photoperiod
+protects plants from potentially fatal consequences of phenological
+responses to temperature cues at the ``wrong'' time of year. However,
+the utility of specific environmental cues depends on both features of
+the environment, such as cue predictability and relevance, and the
+species' adaptive strategies (5). Species with wide natural
+distributions can have multiple distinct environmentally cued
+phenological responses. For example, populations of sunflower
+(\emph{Helianthus annuus}) exhibit day-neutral, facultative short day,
+and facultative long-day flowering responses, which vary with their
+environments (6, 7). These distinct genetic responses in different
+environments are known as genotype-by-environment interactions, or GxE.
 
@@
 half of North America, its photoperiodicity has been predicted to differ
-by plant latitude of origin (36, 37). We previously found divergent
-Midwest and Gulf genetic subpopulations of switchgrass with distinct
-sets of environmental adaptations, in that both populations had distinct
-genetic variation associated with each of two fitness proxies, biomass
-and overwinter survival (38). The Midwest genetic subpopulation is
-primarily composed of individuals from the well-studied upland
-switchgrass ecotype (39, 40), while the Gulf subpopulation has
+by plant latitude of origin (36, 37). We previously found that divergent
+Midwest and Gulf genetic subpopulations of switchgrass have distinct
+sets of environmental adaptations associated with each of two fitness
+proxies, biomass and overwinter survival (38). The Midwest genetic
+subpopulation is primarily composed of individuals from the well-studied
+upland switchgrass ecotype (39, 40), while the Gulf subpopulation has
 individuals from the well-studied lowland ecotype and the phenotypically
@@
 
-Here, we test for GxWeather for two phenological timings by assigning
-patterns of genetic effects on phenology at eight common gardens to many
-patterns of weather covariance at these gardens. To do this, we
-phenotype a diversity panel of hundreds of switchgrass genotypes from
-the Midwest and Gulf subpopulations for the timing of vegetative and
-reproductive development. We do this at eight common garden locations
-spanning 17 degrees of latitude: these gardens cover the majority of the
-latitudinal and climatic range of switchgrass and capture the most
-comprehensive picture to date of the environmental variation this
-species encounters. We define multiple ways phenological timings might
-covary with weather (Table 1) and additional ways phenological timings
-might vary by garden (SI Appendix, Section S1), then jointly re-estimate
-genetic effects on these timings at all eight common gardens using the
-set of these covariance matrices that significantly improved the modeled
-log-likelihood when included (SI Appendix, Section S2-S5) (41). We use
-the Bayesian framework \emph{mash} (multivariate adaptive shrinkage)
-developed by (41), to refine effect size estimates from genome-wide
-association (GWAS). \emph{Mash} allows us to identify and specify
-multiple covariance structures among genetic effect estimates across
-sites, including structures that represent covariance in weather
-variables of interest. Importantly, this method circumvents statistical
-biases in detecting genetic effects with the same or opposite signs
-(42). To confirm our genetic mapping of GxWeather, we compare genomic
-locations of the significant posterior effect estimates from \emph{mash}
-to mapping results from an outbred mapping population grown at the same
-sites. Our analyses allow us to describe the weather cues and types of
-GxE affecting phenology in two divergent natural populations of
-switchgrass across the species' latitudinal range.
+Here, we test for GxWeather for the timing of two phenological traits by
+loading patterns of genetic effects on phenology at eight common gardens
+onto many patterns of weather covariance at these gardens. To do this,
+we phenotyped a diversity panel of hundreds of switchgrass genotypes
+from the Midwest and Gulf subpopulations for the timing of vegetative
+and reproductive development. We did this at eight common garden
+locations spanning 17 degrees of latitude: these gardens covered the
+majority of the latitudinal and climatic range of switchgrass and
+captured the most comprehensive picture to date of the environmental
+variation this species encounters. We defined multiple ways phenological
+traits might covary with weather (Table 1) and additional ways
+phenological traits might vary by garden (SI Appendix, Section S1), then
+jointly re-estimated genetic effects on these timings at all eight
+common gardens using the set of these covariance matrices that
+significantly improved the modeled log-likelihood when included (SI
+Appendix, Section S2-S5) (41). We used the Bayesian framework
+\emph{mash} (multivariate adaptive shrinkage) developed by (41), to
+refine effect size estimates from genome-wide association (GWAS).
+\emph{Mash} allowed us to identify and specify multiple covariance
+structures among genetic effect estimates across sites, including
+structures that represent covariance in weather variables of interest.
+Importantly, this method circumvented statistical biases in detecting
+genetic effects with the same or opposite signs (42). To confirm our
+genetic mapping of GxWeather, we compared genomic locations of the
+significant posterior effect estimates from \emph{mash} to mapping
+results from an outbred mapping population grown at the same sites. Our
+analyses allowed us to describe the weather cues and types of GxE
+affecting phenology in two divergent natural populations of switchgrass
+across the species' latitudinal range.
 
@@
 Genotypes from the Gulf and Midwest subpopulations had distinct
-phenological timings and distinct patterns of phenological correlations
-across our eight common garden sites (Figure~\ref{fig-map}). At the
-three Texas common gardens (hereafter `Texas' gardens), located within
-the natural range of the Gulf subpopulation, the onset of vegetative
-growth, or `green-up date', for Gulf genotypes occurred before
-Midwestern green-up, and the onset of reproductive growth, or `flowering
-date', for Gulf genotypes occurred after Midwestern flowering
-(Figure~\ref{fig-map} A). At the four northernmost common gardens
-(hereafter `North' gardens), located within the natural range of the
-Midwest subpopulation, both green-up and flowering of Gulf genotypes
-occurred after that of Midwest genotypes. At the Oklahoma common garden,
-located near the natural range limits of both the Gulf and the Midwest
-subpopulations, Gulf and Midwest green-up occurred over the same period,
-and Gulf genotype flowering occurred after Midwestern flowering
-(Figure~\ref{fig-map} A). These patterns led to strong negative
-phenotypic correlations for the onset of vegetative growth between the
-North and Texas gardens, particularly in the Gulf and across all
-individuals from the Gulf and Midwest subpopulations (hereafter, `Both'
-subpopulations), and contributed to positive phenotypic correlations for
-the onset of flowering that had larger magnitudes at more northern
-gardens (Figure~\ref{fig-map} B).
+phenological trait timings and distinct patterns of phenological trait
+correlations across our eight common garden sites
+(Figure~\ref{fig-map}). At the three Texas common gardens (hereafter
+`Texas' gardens), located within the natural range of the Gulf
+subpopulation, the onset of vegetative growth, or `green-up date', for
+Gulf genotypes occurred before Midwestern green-up, and the onset of
+reproductive growth, or `flowering date', for Gulf genotypes occurred
+after Midwestern flowering (Figure~\ref{fig-map} A). At the four
+northernmost common gardens (hereafter `North' gardens), located within
+the natural range of the Midwest subpopulation, both green-up and
+flowering of Gulf genotypes occurred after that of Midwest genotypes. At
+the Oklahoma common garden, located near the natural range limits of
+both the Gulf and the Midwest subpopulations, Gulf and Midwest green-up
+occurred over the same period, and Gulf genotype flowering occurred
+after Midwestern flowering (Figure~\ref{fig-map} A). These patterns led
+to strong negative phenotypic correlations for the onset of vegetative
+growth between the North and Texas gardens, particularly in the Gulf and
+in the population containing all individuals from the Gulf and Midwest
+subpopulations (hereafter, `Both' subpopulations) and contributed to
+positive phenotypic correlations for the onset of flowering that had
+larger magnitudes at more northern gardens (Figure~\ref{fig-map} B).
 
@@
 high at individual gardens: 59\% on average for green-up date, and 87\%
-for flowering date. However, h\textsuperscript{2} were variable across
+for flowering date. However, h\textsuperscript{2} was variable across
 gardens, and green-up dates were uncorrelated (r\textsuperscript{2}
@@
 We expected patterns of GxE and GxWeather to vary at the locus,
-subpopulation, and trait level; thus we were interested in modeling the
+subpopulation, and trait level. Thus, we were interested in modeling the
 different ways effects might covary across our eight common gardens. We
-jointly re-estimated genetic effects of a subset of SNPs across all
+jointly re-estimated the genetic effects of a subset of SNPs across all
 eight common gardens using \emph{mash} models, which can flexibly
@@
 covariance matrix that captures the negative correlation between Texas
-and North gardens. In our data, daylength 14 days prior to green-up date
-has negative correlations between Texas and North gardens
-(Figure~\ref{fig-covar} A); if many SNP effects have mass on this
-GxWeather matrix, we would infer that the effects of those SNPs have
-that form of GxWeather covariation.
+(TX1, TX2, and TX3) and North (MO, NE, MI, and SD) gardens. In our data,
+daylength 14 days prior to green-up date was negatively correlated
+between Texas and North gardens (Figure~\ref{fig-covar} A); if many SNP
+effects have mass on this GxWeather matrix, we would infer that the
+effects of those SNPs have that form of GxWeather covariation.
 
@@
 Texas and North gardens, though not as strong as the negative phenotypic
-correlations seen in Both subpopulations (Figure~\ref{fig-map} B;
+correlations observed in Both subpopulations (Figure~\ref{fig-map} B;
 Figure~\ref{fig-covar} A). The Gulf subpopulation and Both
@@
 were indistinguishable, or had no GxE. The majority of pairwise effects
-for Both (\textgreater85\%) subpopulations were the same sign, and
+for Both subpopulations (\textgreater85\%) were the same sign, and
 effects most often differed in magnitude between gardens within and
@@
 interactions (SI Appendix, Figure S4). All QTL for flowering overlapped
-one or more homologs from rice or \emph{A. thaliana} with functionally
-validated roles in flowering (SI Appendix, Dataset 7); the QTL on Chr04K
-overlapped a gene recently functionally validated for flowering in
-switchgrass (43). All flowering and green-up QTL intervals contained at
-least one SNP significant in at least one \emph{mash} run at a
-log10-transformed Bayes Factor \textgreater{} 2, or in the 1\% tail of
-significance, whichever was stricter (SI Appendix, Dataset 8). We also
-looked for enrichments of \emph{mash} SNPs in the 1\% tail of
+with one or more homologs from rice or \emph{A. thaliana} with
+functionally validated roles in flowering (SI Appendix, Dataset 7); the
+QTL on Chr04K overlapped a gene recently functionally validated for
+flowering in switchgrass (43). All flowering and green-up QTL intervals
+contained at least one SNP significant in at least one \emph{mash} run
+at a log10-transformed Bayes Factor \textgreater{} 2, or in the 1\% tail
+of significance, whichever was stricter (SI Appendix, Dataset 8). We
+also looked for enrichments of \emph{mash} SNPs in the 1\% tail of
 significance (the `\emph{mash} 1\% tail') within each QTL interval. At
@@
 patterns of GxWeather with specific genomic regions using a switchgrass
-diversity panel grown at eight common gardens. We can assign genetic
+diversity panel grown at eight common gardens. We assigned genetic
 effects to both GxWeather patterns with interpretable weather-based
-cues, and to unmeasured, site-based patterns. We use this approach to
+cues, and to unmeasured, site-based patterns. We used this approach to
 study GxWeather for the timings of vegetative and reproductive
@@
 fitness, this result supports theoretical models that local adaptation
-should involve antagonistic pleiotropy at the level of individual loci
-(27--30). In addition, it advances local adaptation research by using
-more than two field sites and a wide range of genetic variation to
-determine that rank-changing GxE for phenological timings is common in
-small genomic regions (15, 44, 45).
+should involve trade-offs at the level of individual loci (27--30). In
+addition, it advances local adaptation research by using more than two
+field sites and a wide range of genetic variation to determine that
+rank-changing GxE for phenological traits is common in small genomic
+regions (15, 44, 45).
 
@@
 
-\includegraphics{images/Figure_1_AB.png}
+\includegraphics{images/Fig1_AB.png}
 
@@
 jointly re-estimated effects in eight common gardens, for effects with
-lfsr \textless{} 0.05. a) Examples of effect patterns at three pairs of
-sites with three types of GxE. All effects are for an alternate allele,
-with the reference allele effect defined to be zero at both gardens and
-represented by the dashed vertical line. Sign: Effects that differ in
-sign at these pairs of gardens (p \textless{} 0.05, lfsr). Magnitude:
-Effects identical in sign (p \textless{} 0.05, lfsr) that differ in
-magnitude by a factor of \textgreater0.4. Not Distinguishable: Effects
-not distinguishable by magnitude nor sign of the effect, with no
-measurable GxE. b) The fraction of effects with each GxE type for the
-onset of vegetative growth (green-up date) and reproductive growth
-(flowering date), within and between two genetic subpopulations. Common
-gardens are grouped by the larger region they came from: North gardens
-are within the natural range of the Midwest subpopulation, and include
-MO, NE, MI, and SD, while Texas gardens are within the natural range of
-the Gulf subpopulation, and include TX1, TX2, and TX3.}
+lfsr \textless{} 0.05 at each pair of sites contrasted. a) Examples of
+effect patterns at three pairs of sites with three types of GxE. All
+effects are for an alternate allele, with the reference allele effect
+defined to be zero at both gardens and represented by the dashed
+vertical line. Sign: Effects that differ in sign at these pairs of
+gardens (p \textless{} 0.05, lfsr). Magnitude: Effects identical in sign
+(p \textless{} 0.05, lfsr) that differ in magnitude by a factor of
+\textgreater0.4. Not Distinguishable: Effects not distinguishable by
+magnitude nor sign of the effect, with no measurable GxE. b) The
+fraction of effects with each GxE type for the onset of vegetative
+growth (green-up date) and reproductive growth (flowering date), within
+and between two genetic subpopulations. Common gardens are grouped by
+the larger region they came from: North gardens are within the natural
+range of the Midwest subpopulation, and include MO, NE, MI, and SD,
+while Texas gardens are within the natural range of the Gulf
+subpopulation, and include TX1, TX2, and TX3.}
 
@@
 Award Numbers DE-SC0018409 and DE-FC02-07ER64494, the US Department of
-Energy Awards DESC0014156 to T.E.J., DE-SC0017883 to D.B.L, National
-Science Foundation PGRP Awards IOS0922457 and IOS1444533 to T.E.J, and
-the Long-term Ecological Research Program (DEB 1832042) at the Kellogg
-Biological Station.}
+Energy Awards DESC0014156 to T.E.J., DE-SC0017883 to D.B.L, NIH grant
+R35GM151108 and a Pew Scholarship to A.H., National Science Foundation
+PGRP Awards IOS0922457 and IOS1444533 to T.E.J, and the Long-term
+Ecological Research Program (DEB 1832042) at the Kellogg Biological
+Station.}
 

### 2024-08-17 manuscript/revision2/response_to_reviews/response_to_reviews.qmd
revison2: finalize manuscript and cover letter for resubmission

@@
 ::: {.flushright data-latex="1" linestretch="1"}
-July 17th, 2024
+August 17th, 2024
 :::
@@
 
-We now add these citations (lines 47-50; lines 58-61).
+We now add these citations (lines 51-54; lines 62, 125-131).
 We regret we can't comprehensively reference the impressive crop modeling literature.
@@
 This work has two additional, key advances over the previous work: it can assign loci genome-wide to multiple kinds of GxE and GxWeather, and has an unbiased statistical method to identify loci with and without rank-changing GxE.
-We now summarize the findings from [Lovell *et al.,* 2021](https://doi.org/10.1038/s41586-020-03127-1) that we use as a springboard for this work on lines 136-145, and outline the additional advances of this study on lines 156-163, 169-171, and 883-893.
+We now summarize the findings from [Lovell *et al.,* 2021](https://doi.org/10.1038/s41586-020-03127-1) that we use as a springboard for this work on lines 142-150, and outline the additional advances of this study on lines 162-169, 175-177, and 884-895.
 
@@
 We now edit the Introduction to more clearly articulate the aims of this manuscript.
-Our key expectation is that different genetic subpopulations and genomic regions have likely evolved distinct patterns of GxE (lines 55-58; lines 219-223).
+Our key expectation is that different genetic subpopulations and genomic regions have likely evolved distinct patterns of GxE (lines 59-62; lines 228-232).
 Thus, we aim to identify the kinds of GxE present for each subpopulation and each trait.
-(lines 129-131).
+(lines 135-137).
 
@@
 We agree that we do not identify one general pattern or one environmental cue affecting any set of populations or phenological trait.
-Instead, our key expectation is that different populations and different genetic loci will have evolved distinct patterns of GxE (lines 55-58; lines 219-223; SI Appendix Section S1).
+Instead, our key expectation is that different populations and different genetic loci will have evolved distinct patterns of GxE (lines 59-62; lines 228-232; SI Appendix Section S1).
 Thus, we aim to identify the kinds of GxE present genome-wide and elaborate when these kinds of GxE differ (Figure 2), and demonstrate that individual loci have different types of GxE across environmental regions and subpopulations (Figure 3).
@@
 
-We edit the results to directly compare the GxWeather matrices and the phenotypic correlations in Figure 1B. (lines 241-258; lines 269-276; lines 283-290).
+We edit the results to directly compare the GxWeather matrices and the phenotypic correlations in Figure 1B. (lines 251-269; lines 284-287; lines 290-293; lines 297-301).
 
@@
 
-As a consequence, we develop an approach to specify multiple environmental cues and compete them to explain patterns of genetic effects (lines 230-237): we use a greedy algorithm to use with mash to select covariance patterns, and use mash to flexibly identify these patterns across populations and environments.
-(lines 238-240).
-With this approach, we are able to demonstrate that we can associate multiple patterns of GxWeather with specific genomic regions (lines 407-410).
-We can also assign genetic effects to both GxWeather patterns derived from weather variables and to other, site-based patterns (lines 410-412).
+As a consequence, we develop an approach to specify multiple environmental cues and compete them to explain patterns of genetic effects (lines 238-244): we use a greedy algorithm to use with mash to select covariance patterns, and use mash to flexibly identify these patterns across populations and environments (lines 244-250).
+With this approach, we are able to demonstrate that we can associate multiple patterns of GxWeather with specific genomic regions (lines 427-430).
+We can also assign genetic effects to both GxWeather patterns derived from weather variables and to other, site-based patterns (lines 430-432).
 
@@
 
-We now reference this paper in the introduction (lines 47-52).
+We now reference this paper in the introduction (lines 51-56).
 
@@
 We add Table 1 to better explain the covariance matrices we select using the greedy mash algorithm.
-We include the matrices selected by this algorithm that have \>0.1% weight (lines 262-263) in the mash model as visualizations in Figure 2.
+We include the matrices selected by this algorithm that have \>0.1% weight (lines 273-274) in the mash model as visualizations in Figure 2.
 
@@
 We agree that correctly specifying the type of GxE in our models of additive variation across gardens is important for all models in this manuscript.
-In many cases the complexity of the models that we can specify using this data is limited, because of the low sample size - we typically have one replicate per genotype in each common garden, and the low number of genotypes present at each garden in each strongly stratified subpopulation means that we cannot fit complicated GxE structures using linear mixed models, often needing to fit very simple/restrictive variance-covariance structures.
-The entire focus on mash was to allow us to fit less structured variance-covariance structures, as well as more of them.
+In many cases the complexity of the genetic variance-covariance structures we wish to fit are not be possible to fit using standard approaches.
+Eight gardens give eight variance components and 8 x 8 covariances - most datasets are not large enough to give solutions for these models.
+The standard approach is to either do data reduction (e.g.
+factor analysis) or impose unrealistic constraints (e.g., equal variances, or only positive covariances) in order to solve these models.
+Our approach and our focus on mash was to allow us to fit less structured variance-covariance structures, as well as more of them.
 
@@
 To clarify our interpretation of the posterior matrix weights provided by mash, we performed an extensive analysis of the performance of mash models when different covariance matrices are included.
-Specifically, we implemented a model selection approach that uses a greedy algorithm to evaluate the log likelihood of the mash model as additional covariance matrices were included (see Methods lines 899-906, SI Appendix Section S4, and pseudo code below).
+Specifically, we implemented a model selection approach that uses a greedy algorithm to evaluate the log likelihood of the mash model as additional covariance matrices were included (see Methods lines 901-908, SI Appendix Section S4, and pseudo code below).
 This is a similar, but more computationally efficient design, than the leave-one-out approach recommended by the reviewer as the runtime of mash increases with the number of covariance matrices included.
@@
 
-We believe that the use of the lfsr to detect effects with rank-changing GxE is a key advance in our manuscript, and so we have made changes to the Introduction (lines 169-171), Results (lines 319-326), and Materials & Methods (lines 957-997; lines 1055-1056) to explain and motivate this statistical change.
+We believe that the use of the lfsr to detect effects with rank-changing GxE is a key advance in our manuscript, and so we have made changes to the Introduction (lines 175-177), Results (lines 338-343), and Materials & Methods (lines 959-982; lines 996-998; lines 1055-1057) to explain and motivate this statistical change.
 
@@
 It's true that the threshold for differential sensitivity is arbitrary.
-We have reworded the manuscript to stress that there is equal power to detect effects of different sign as there are effects of the same sign (lines 994-997; lines 1055-1056).
+We have reworded the manuscript to stress that there is equal power to detect effects of different sign as there are effects of the same sign (lines 996-998; lines 1055-1057).
 Another class of effects we have in our analysis that differ from classic antagonistic pleiotropy & conditional neutrality comparisons are effects which are not distinguishable by sign or magnitude.

### 2024-08-17 manuscript/revision2/response_to_reviews/response_to_reviews.tex
revison2: finalize manuscript and cover letter for resubmission

@@
 \begin{flushright}
-July 17th, 2024
+August 17th, 2024
 
@@
 \begin{quote}
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]
 
@@
 
-We now add these citations (lines 47-50; lines 58-61). We regret we
-can't comprehensively reference the impressive crop modeling literature.
-Our goal was to reference work on natural populations that focused the
-genetics on phenology and/or local adaptation, and to touch on how these
-ideas have interacted with crop improvement. This was because our work
-uses a natural population of switchgrass with potential relevance for
-breeding efforts in this species.
-
-\begin{quote}
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
+We now add these citations (lines 51-54; lines 62, 125-131). We regret
+we can't comprehensively reference the impressive crop modeling
+literature. Our goal was to reference work on natural populations that
+focused the genetics on phenology and/or local adaptation, and to touch
+on how these ideas have interacted with crop improvement. This was
+because our work uses a natural population of switchgrass with potential
+relevance for breeding efforts in this species.
+
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]
 
@@
 \href{https://doi.org/10.1038/s41586-020-03127-1}{Lovell \emph{et al.,}
-2021} that we use as a springboard for this work on lines 136-145, and
-outline the additional advances of this study on lines 156-163, 169-171,
-and 883-893.
-
-\begin{quote}
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
+2021} that we use as a springboard for this work on lines 142-150, and
+outline the additional advances of this study on lines 162-169, 175-177,
+and 884-895.
+
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]
 
@@
 and genomic regions have likely evolved distinct patterns of GxE (lines
-55-58; lines 219-223). Thus, we aim to identify the kinds of GxE present
-for each subpopulation and each trait. (lines 129-131).
-
-\begin{quote}
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
+59-62; lines 228-232). Thus, we aim to identify the kinds of GxE present
+for each subpopulation and each trait. (lines 135-137).
+
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]
 
@@
 different genetic loci will have evolved distinct patterns of GxE (lines
-55-58; lines 219-223; SI Appendix Section S1). Thus, we aim to identify
+59-62; lines 228-232; SI Appendix Section S1). Thus, we aim to identify
 the kinds of GxE present genome-wide and elaborate when these kinds of
@@
 We edit the results to directly compare the GxWeather matrices and the
-phenotypic correlations in Figure 1B. (lines 241-258; lines 269-276;
-lines 283-290).
-
-\begin{quote}
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
+phenotypic correlations in Figure 1B. (lines 251-269; lines 284-287;
+lines 290-293; lines 297-301).
+
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]
 
@@
 environmental cues and compete them to explain patterns of genetic
-effects (lines 230-237): we use a greedy algorithm to use with mash to
+effects (lines 238-244): we use a greedy algorithm to use with mash to
 select covariance patterns, and use mash to flexibly identify these
-patterns across populations and environments. (lines 238-240). With this
+patterns across populations and environments (lines 244-250). With this
 approach, we are able to demonstrate that we can associate multiple
-patterns of GxWeather with specific genomic regions (lines 407-410). We
+patterns of GxWeather with specific genomic regions (lines 427-430). We
 can also assign genetic effects to both GxWeather patterns derived from
-weather variables and to other, site-based patterns (lines 410-412).
-
-\begin{quote}
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
+weather variables and to other, site-based patterns (lines 430-432).
+
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]
 
@@
 
-We now reference this paper in the introduction (lines 47-52).
-
-\begin{quote}
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]
+We now reference this paper in the introduction (lines 51-56).
+
+\begin{quote}
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, arc=.35mm, colback=white, opacityback=0, left=2mm, breakable, toprule=.15mm, bottomrule=.15mm]
 
@@
 select using the greedy mash algorithm. We include the matrices selected
-by this algorithm that have \textgreater0.1\% weight (lines 262-263) in
+by this algorithm that have \textgreater0.1\% weight (lines 273-274) in
 the mash model as visualizations in Figure 2.
@@
 additive variation across gardens is important for all models in this
-manuscript. In many cases the complexity of the models that we can
-specify using this data is limited, because of the low sample size - we
-typically have one replicate per genotype in each common garden, and the
-low number of genotypes present at each garden in each strongly
-stratified subpopulation means that we cannot fit complicated GxE
-structures using linear mixed models, often needing to fit very
-simple/restrictive variance-covariance structures. The entire focus on
-mash was to allow us to fit less structured variance-covariance
-structures, as well as more of them.
+manuscript. In many cases the complexity of the genetic
+variance-covariance structures we wish to fit are not be possible to fit
+using standard approaches. Eight gardens give eight variance components
+and 8 x 8 covariances - most datasets are not large enough to give
+solutions for these models. The standard approach is to either do data
+reduction (e.g.~factor analysis) or impose unrealistic constraints
+(e.g., equal variances, or only positive covariances) in order to solve
+these models. Our approach and our focus on mash was to allow us to fit
+less structured variance-covariance structures, as well as more of them.
 
@@
 greedy algorithm to evaluate the log likelihood of the mash model as
-additional covariance matrices were included (see Methods lines 899-906,
+additional covariance matrices were included (see Methods lines 901-908,
 SI Appendix Section S4, and pseudo code below). This is a similar, but
@@
 GxE is a key advance in our manuscript, and so we have made changes to
-the Introduction (lines 169-171), Results (lines 319-326), and Materials
-\& Methods (lines 957-997; lines 1055-1056) to explain and motivate this
-statistical change.
+the Introduction (lines 175-177), Results (lines 338-343), and Materials
+\& Methods (lines 959-982; lines 996-998; lines 1055-1057) to explain
+and motivate this statistical change.
 
@@
 detect effects of different sign as there are effects of the same sign
-(lines 994-997; lines 1055-1056). Another class of effects we have in
+(lines 996-998; lines 1055-1057). Another class of effects we have in
 our analysis that differ from classic antagonistic pleiotropy \&

### 2024-08-17 manuscript/revision2/revision2.qmd
revison2: finalize manuscript and cover letter for resubmission

@@
         postal-code: 57006
-  - name: Phillip A. Fay
+  - name: Philip A. Fay
     affiliations:
@@
 abstract: |
-  The timing of vegetative and reproductive growth in plants ("phenological timings") depends on genetic effects (G), environmental (e.g., weather) cues, and their interaction.
-Measuring GxWeather interactions can aid prediction and manipulation of phenological timings.
-Here, we map GxWeather effects on phenological timings in two highly divergent switchgrass (*Panicum virgatum*) populations using repeated plantings of cloned individuals from these populations at eight sites across the central United States.
-Some genetic effects covary with  interpretable weather-based cues.
-For example, in the Gulf population, 65% of genetic effects on the timing of vegetative growth covary with daylength 14 days prior to green-up date, and 33% of genetic effects on the timing of flowering covary with cumulative rainfall in the seven days prior to flowering.
-However, most variation in genetic effects is site-specific and cannot be attributed to variation in weather variables.
-We demonstrate that we can identify genetic variation with GxWeather and assign these loci to specific weather-based cues or other patterns.
-We predict that selective breeding for particular alleles at the GxWeather loci could alter flowering responsiveness in a photoperiod or rainfall-specific way.
-More broadly, our approach proposes a refined characterization of genotype-by-environment interactions in any species phenotyped in multiple environments.
+  The timing of vegetative and reproductive growth in plants ("phenological timings") depend on genetic effects (G), environmental (e.g., weather) cues, and their interaction.
+Here, we measure phenological timings in two highly divergent switchgrass (*Panicum virgatum*) subpopulations using repeated plantings of cloned individuals at eight sites across the central United States.
+The timing of vegetative growth for the two subpopulations reversed between their two natural ranges and had strong negative correlations between these regions; in contrast, the timing of flowering was positively correlated between gardens.
+We expect that these phenotypic correlations consist of polygenic effects on phenology which have distinct patterns of GxE segregating at different mapped loci.
+Thus, we infer the mixture of ways genetic effects impact phenological timings, such as across common gardens (GxE) or with weather cues (GxWeather).
+We demonstrate that we can identify genetic variation with GxWeather and assign genetic loci to specific weather-based cues or other patterns.
+For example, in the Gulf subpopulation, 65% of genetic effects on the timing of vegetative growth covary with daylength 14 days prior to green-up date, and 33% of genetic effects on the timing of flowering covary with cumulative rainfall in the week prior to flowering.
+However, most variation in genetic effects cannot be attributed to variation in weather variables.
+Selective breeding for particular alleles at GxWeather loci could alter flowering responsiveness in a photoperiod or rainfall-specific way.
+More broadly, our approach refines the characterization of genotype-by-environment interactions and can be implemented in any species phenotyped in multiple environments.
 runningauthor: "MacQueen"
@@
 This material is based upon work supported in part by the Great Lakes Bioenergy Research Center, U.S.
-Department of Energy, Office of Science, Office of Biological and Environmental Research under Award Numbers DE-SC0018409 and DE-FC02-07ER64494, the US Department of Energy Awards DESC0014156 to T.E.J., DE-SC0017883 to D.B.L, NIH grant R35GM151108 and a Pew Scholarship to A.H., National Science Foundation PGRP Awards IOS0922457 and IOS1444533 to T.E.J, and the Long-term Ecological Research Program (DEB 1832042) at the Kellogg Biological Station.
+Department of Energy, Office of Science, Office of Biological and Environmental Research under Award Numbers DE-SC0018409 and DE-FC02-07ER64494, the US Department of Energy Awards DE-SC0021126 and DESC0014156 to T.E.J., DE-SC0017883 to D.B.L, NIH grant R35GM151108 and a Pew Scholarship to A.H., National Science Foundation PGRP Awards IOS0922457 and IOS1444533 to T.E.J, and the Long-term Ecological Research Program (DEB 1832042) at the Kellogg Biological Station.
 bibliography: bibliography.bib 
@@
 
-\firstpage[17]{2}
+\firstpage[13]{2}
 
@@
 However, most studies of flowering GxE focus on finding a single, best fitting form of genotype-environment covariance, despite the key expectation that different genetic subpopulations, and even different genomic regions, have likely evolved distinct patterns of GxE [@Weine2023.06.21.545998].
-Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or alleles with effects with opposing fitness outcomes [@levene1953genetic; @felsenstein1976theoretical; @kawecki2004conceptual; @hedrick1986genetic], previous empirical work has found limited evidence of antagonistic pleiotropy [@des2013genotype; @wadgymar2017identifying; @millet2016].
+Additionally, despite theoretical predictions that local adaptation should involve trade-offs caused by antagonistic pleiotropy, or alleles with effects with opposing fitness outcomes [@levene1953genetic; @felsenstein1976theoretical; @kawecki2004conceptual; @hedrick1986genetic], previous empirical work has found limited evidence of trade-offs caused by this form of GxE [@des2013genotype; @wadgymar2017identifying; @millet2016].
 However, this work has been limited by a known statistical bias that reduced detection of genetic effects that differ in sign [@des2013genotype; @anderson2013genetic; @anderson2011evolutionary].
@@
 We expected patterns of GxE and GxWeather to vary at the locus, subpopulation, and trait level.
-Thus, we were interested in modeling the different ways effects might covary across our eight common gardens.
+Thus, we were interested in modeling the different ways polygenic effects might covary across our eight common gardens[@weine2024].
 We jointly re-estimated the genetic effects of a subset of SNPs across all eight common gardens using *mash* models, which can flexibly capture many modes of covariance across gardens (SI Appendix, Sections S1-S5, Datasets 1-6).
 To obtain genetic effects, we used subsets of effects from garden- and subpopulation GWAS (SI Appendix, Section S2-S3; Table S1).
-To obtain covariance structures, we specified both covariance structures introduced in the initial *mash* manuscript, and GxWeather covariance matrices estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event (Table 1; SI Appendix, Section S1).
+To obtain covariance structures, we specified both covariance structures introduced in the initial *mash* manuscript (e.g., garden-specific effects; equal effects at all gardens, @fig-covar A), and GxWeather covariance matrices estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event (Table 1; SI Appendix, Section S1).
 Then, we used a greedy algorithm to select covariance matrices from each model's set that significantly improved the model likelihood (SI Appendix, Section S4; Table S2).
-Finally, we fit *mash* models using the selected set of covariance matrices on a subset relatively unlinked (r^2^ \< 0.2), 'strong' genetic effects with low *p*-values (SI Appendix, Section S5).
+Finally, we fit *mash* models using the selected set of covariance matrices on a subset of relatively unlinked (r^2^ \< 0.2), 'strong' genetic effects with low *p*-values (SI Appendix, Section S5).
 
@@
 The Midwest had 0.15% weight on a GxWeather covariance matrix of cumulative GDD for the 28d prior to green-up, a GxWeather matrix that was visually similar to the Midwest's observed pattern of phenotypic correlation (@fig-map B).
-Both subpopulations had non-zero weights on two additional GxWeather covariance types, average temperature one day prior to green-up, and the day length change in seconds in the day prior to green-up (@fig-covar B).
+Both subpopulations had non-zero weights on two additional GxWeather covariance types, average temperature one day prior to green-up, and the day length change in the day prior to green-up (@fig-covar B).
 The average temperature covariance matrix had negative covariances between Texas and North gardens, though not as strong as the negative phenotypic correlations observed in Both subpopulations (@fig-map B; @fig-covar A).
@@
 These matrices included simple heterozygosity, with intermediate, positive covariances between all gardens, and single effect matrices with garden-specific effects.
+Re-estimation of genetic effects with *mash* allowed us to identify and quantify the fraction of loci exhibiting GxWeather patterns; a minority of GxE was GxWeather covariance, and the majority of GxE covariance in these models is driven by other, unknown drivers of GxE.
 
@@
 We expected to observe common rank-changing GxE at the level of individual loci as this is a key theoretical prediction of local adaptation [@felsenstein1976theoretical; @levene1953genetic; @kawecki2004conceptual; @hedrick1986genetic].
+Previous empirical work has found limited evidence of trade-offs caused by antagonistic pleiotropy; however, this work had a known statistical bias reducing the detection of effects that differed in sign [@wadgymar2017identifying; @des2013genotype; @millet2016].
 To determine the frequency of rank-changing GxE, we used the local false sign rate (lfsr), an analogue of the local false discovery rate that establishes confidence in the effect sign, not the effect's difference from zero, to determine significance.
@@
 As the climate and the natural environment change, it is increasingly critical to understand how patterns of plant-environment interactions will change in response.
-To do this, we must understand the current patterns of trait covariation across environments, the genetic underpinnings of these patterns, and the cases where this covariation can be altered.
+To do this, we must understand the current patterns of trait covariation across environments, the genetic underpinnings of these patterns, and the cases where this covariation can be altered through selection at individual loci.
 Here, we demonstrate that we can associate multiple patterns of GxWeather with specific genomic regions using a switchgrass diversity panel grown at eight common gardens.
@@
 The Gulf & Both subpopulations had rank-changing GxE for green-up date, while the Midwest subpopulation had rank-changing GxE for flowering date.
-As phenological timings are major components of plant fitness, this result supports theoretical models that local adaptation should involve trade-offs at the level of individual loci [@kawecki2004conceptual; @felsenstein1976theoretical; @levene1953genetic; @hedrick1986genetic].
-In addition, it advances local adaptation research by using more than two field sites and a wide range of genetic variation to determine that rank-changing GxE for phenological traits is common in small genomic regions [@wadgymar2017identifying; @savolainen2013ecological; @lowry2019qtl].
+As phenological timings are major components of plant fitness, this result supports theoretical models that local adaptation should involve trade-offs due to antagonistic pleiotropy at the level of individual loci [@kawecki2004conceptual; @felsenstein1976theoretical; @levene1953genetic; @hedrick1986genetic].
+Experimental designs in local adaptation research now often use more than two field sites and a wide range of genetic variation; *mash* and the local false sign rate facilitate the analysis of these experimental designs.
+We use *mash* to determine that rank-changing GxE for phenological traits is common in small genomic regions [@wadgymar2017identifying; @savolainen2013ecological; @lowry2019qtl].
 
@@
 Three genomic regions affecting flowering that we re-estimated across all eight sites were also supported by QTL from an independent mapping population at these sites (@fig-qtl B).
+Models combining both subpopulations showed less signal, perhaps due to the distinctness of cues in both subpopulations or to confounding caused by population structure.
 
@@
 
-![Example hypothesis-driven covariance matrices specified in *mash* and the posterior weights placed on all covariance matrices.
+![Example Canonical and GxWeather covariance matrices specified in *mash* and the posterior weights placed on all covariance matrices.
 (a) Common gardens are arranged in latitudinal order within the matrices.
@@
 Covariance matrices included in *mash* that had zero posterior weight in all three *mash* runs on the genetic subpopulations, such as the identity matrix, are not shown.
-(c,e) Total posterior weight placed on covariance matrices that were hypothesized or canonical, for the (c) green-up date phenotype and (e) flowering date phenotype.](images/Figure_2_six_sites_Midwest.png){#fig-covar}
+(c,e) Total posterior weight placed on covariance matrices that were GxWeather or Canonical, for the (c) green-up date phenotype and (e) flowering date phenotype.](images/Figure_2_six_sites_Midwest.png){#fig-covar}
 
-![Types of GxE present between of pairs of jointly re-estimated effects in eight common gardens, for effects with lfsr \< 0.05 at each pair of sites contrasted.
+![Types of GxE present between of pairs of jointly re-estimated SNP effects in eight common gardens, for effects with lfsr \< 0.05 at both gardens for each pair of gardens contrasted.
 a) Examples of effect patterns at three pairs of sites with three types of GxE.
@@
 Not Distinguishable: Effects not distinguishable by magnitude nor sign of the effect, with no measurable GxE.
+Confidence intervals are illustrative that the effect estimate does not overlap zero.
 b) The fraction of effects with each GxE type for the onset of vegetative growth (green-up date) and reproductive growth (flowering date), within and between two genetic subpopulations.
-Common gardens are grouped by the larger region they came from: North gardens are within the natural range of the Midwest subpopulation, and include MO, NE, MI, and SD, while Texas gardens are within the natural range of the Gulf subpopulation, and include TX1, TX2, and TX3.](images/Figure_3_Fraction_GxE_by_Region_six_midwest_sites.png){#fig-effects}
+Common gardens are grouped by the larger region they came from: North gardens are within the natural range of the Midwest subpopulation, and include MO, NE, MI, and SD, while Texas gardens are within the natural range of the Gulf subpopulation, and include TX1, TX2, and TX3.](images/Figure_3_AH_comments.png){#fig-effects}
 
@@
 
-We were interested in specifying genetic models for trait variation that allowed more than one form of GxE, as we reasoned that different loci should display different forms of GxE.
+We were interested in specifying genetic models for trait variation that allowed more than one form of GxE, as we reasoned that different loci should display different forms of GxE (e.g., changes in effect magnitude vs sign).
 In addition, we were interested in an unbiased estimation of the frequency of rank-changing GxE relative to other forms of GxE, as the presence of rank-changing GxE at the level of individual loci is a key theoretical prediction of local adaptation.
@@
 To use *mash* on our diversity panel, we had to specify both a relatively uncorrelated set of covariance matrices, which in our case defined types of GxE and GxWeather between gardens, and we had to specify subsets of SNP effect estimates and standard errors for our traits at each common garden.
-To specify a set of covariance matrices, we first defined many covariance matrices, including GxWeather matrices that represented the correlation in weather cues between gardenes before the phenological event (SI Appendix, Section S1), then implemented a model selection approach that used a greedy algorithm to evaluate if the log likelihood of the *mash* model was significantly improved as additional covariance matrices were included (SI Appendix, Section S4).
+To specify a set of covariance matrices, we first defined many covariance matrices, including GxWeather matrices that represented the correlation in weather cues between gardens before the phenological event (SI Appendix, Section S1), then implemented a model selection approach that used a greedy algorithm to evaluate if the log likelihood of the *mash* model was significantly improved as additional covariance matrices were included (SI Appendix, Section S4).
 To specify subsets of SNP effect estimates and standard errors to use for both the greedy algorithm and with the optimal set of covariance matrices, we first calculated best linear unbiased predictors (BLUPs) for each phenological trait in each genetic subpopulation and each common garden (SI Appendix, Section S2).
@@
 This previous work recognized that this testing bias could lead to undercounting occurrences of antagonistic pleiotropy [@anderson2011evolutionary; @des2013genotype], and sought to reduce it by permutation [@anderson2013genetic].
-However, using the lfsr to test for allelic effects that differ in sign does not undercount these occurences, as this statistic answers a fundamentally different question.
+However, using the lfsr to test for allelic effects that differ in sign does not undercount these occurrences, as this statistic answers a fundamentally different question.
 For each effect $j$, the $lfsr_j$ is defined as the probability that we make an error in the sign of effect $j$ if we were forced to declare the effect positive or negative @10.1093/biostatistics/kxw041.

### 2024-08-17 manuscript/revision2/revision2.tex
revison2: finalize manuscript and cover letter for resubmission

@@
 Smith; Jason Bonnette; Arvid R.
-Boe; Phillip A.
+Boe; Philip A.
 Fay; Felix B.
@@
 %
-]{Phillip A. Fay}
+]{Philip A. Fay}
 \author[d%
@@
 The timing of vegetative and reproductive growth in plants
-(``phenological timings'') depends on genetic effects (G), environmental
-(e.g., weather) cues, and their interaction. Measuring GxWeather
-interactions can aid prediction and manipulation of phenological
-timings. Here, we map GxWeather effects on phenological timings in two
-highly divergent switchgrass (\emph{Panicum virgatum}) populations using
-repeated plantings of cloned individuals from these populations at eight
-sites across the central United States. Some genetic effects covary with
-interpretable weather-based cues. For example, in the Gulf population,
-65\% of genetic effects on the timing of vegetative growth covary with
-daylength 14 days prior to green-up date, and 33\% of genetic effects on
-the timing of flowering covary with cumulative rainfall in the seven
-days prior to flowering. However, most variation in genetic effects is
-site-specific and cannot be attributed to variation in weather
-variables. We demonstrate that we can identify genetic variation with
-GxWeather and assign these loci to specific weather-based cues or other
-patterns. We predict that selective breeding for particular alleles at
-the GxWeather loci could alter flowering responsiveness in a photoperiod
-or rainfall-specific way. More broadly, our approach proposes a refined
-characterization of genotype-by-environment interactions in any species
-phenotyped in multiple environments.
+(``phenological timings'') depend on genetic effects (G), environmental
+(e.g., weather) cues, and their interaction. Here, we measure
+phenological timings in two highly divergent switchgrass (\emph{Panicum
+virgatum}) subpopulations using repeated plantings of cloned individuals
+at eight sites across the central United States. The timing of
+vegetative growth for the two subpopulations reversed between their two
+natural ranges and had strong negative correlations between these
+regions; in contrast, the timing of flowering was positively~correlated
+between gardens. We expect that these phenotypic correlations consist of
+polygenic effects on phenology which have distinct patterns of GxE
+segregating at different mapped loci. Thus, we infer the mixture of ways
+genetic effects impact phenological timings, such as across common
+gardens (GxE) or with weather cues (GxWeather). We demonstrate that we
+can identify genetic variation with GxWeather and assign genetic loci to
+specific weather-based cues or other patterns. For example, in the Gulf
+subpopulation, 65\% of genetic effects on the timing of vegetative
+growth covary with daylength 14 days prior to green-up date, and 33\% of
+genetic effects on the timing of flowering covary with cumulative
+rainfall in the week prior to flowering. However, most variation in
+genetic effects cannot be attributed to variation in weather variables.
+Selective breeding for particular alleles at GxWeather loci could alter
+flowering responsiveness in a photoperiod or rainfall-specific way. More
+broadly, our approach refines the characterization of
+genotype-by-environment interactions and can be implemented in any
+species phenotyped in multiple environments.
 \end{abstract}
@@
 \ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
-\firstpage[17]{2}
+\firstpage[13]{2}
 
@@
 have likely evolved distinct patterns of GxE (26). Additionally, despite
-theoretical predictions that local adaptation should involve
-antagonistic pleiotropy, or alleles with effects with opposing fitness
-outcomes (27--30), previous empirical work has found limited evidence of
-antagonistic pleiotropy (15, 31, 32). However, this work has been
-limited by a known statistical bias that reduced detection of genetic
-effects that differ in sign (31, 33, 34). Thus, despite substantial
-interest in the frequencies of various forms of GxE, the frequency of
-sign-changing GxE relative to other forms of GxE remains unknown.
+theoretical predictions that local adaptation should involve trade-offs
+caused by antagonistic pleiotropy, or alleles with effects with opposing
+fitness outcomes (27--30), previous empirical work has found limited
+evidence of trade-offs caused by this form of GxE (15, 31, 32). However,
+this work has been limited by a known statistical bias that reduced
+detection of genetic effects that differ in sign (31, 33, 34). Thus,
+despite substantial interest in the frequencies of various forms of GxE,
+the frequency of sign-changing GxE relative to other forms of GxE
+remains unknown.
 
@@
 subpopulation, and trait level. Thus, we were interested in modeling the
-different ways effects might covary across our eight common gardens. We
-jointly re-estimated the genetic effects of a subset of SNPs across all
-eight common gardens using \emph{mash} models, which can flexibly
-capture many modes of covariance across gardens (SI Appendix, Sections
-S1-S5, Datasets 1-6). To obtain genetic effects, we used subsets of
-effects from garden- and subpopulation GWAS (SI Appendix, Section S2-S3;
-Table S1). To obtain covariance structures, we specified both covariance
-structures introduced in the initial \emph{mash} manuscript, and
-GxWeather covariance matrices estimated from the covariance of empirical
-weather patterns at each garden at specific times before the
-phenological event (Table 1; SI Appendix, Section S1). Then, we used a
-greedy algorithm to select covariance matrices from each model's set
-that significantly improved the model likelihood (SI Appendix, Section
-S4; Table S2). Finally, we fit \emph{mash} models using the selected set
-of covariance matrices on a subset relatively unlinked
-(r\textsuperscript{2} \textless{} 0.2), `strong' genetic effects with
-low \emph{p}-values (SI Appendix, Section S5).
+different ways polygenic effects might covary across our eight common
+gardens(43). We jointly re-estimated the genetic effects of a subset of
+SNPs across all eight common gardens using \emph{mash} models, which can
+flexibly capture many modes of covariance across gardens (SI Appendix,
+Sections S1-S5, Datasets 1-6). To obtain genetic effects, we used
+subsets of effects from garden- and subpopulation GWAS (SI Appendix,
+Section S2-S3; Table S1). To obtain covariance structures, we specified
+both covariance structures introduced in the initial \emph{mash}
+manuscript (e.g., garden-specific effects; equal effects at all gardens,
+Figure~\ref{fig-covar} A), and GxWeather covariance matrices estimated
+from the covariance of empirical weather patterns at each garden at
+specific times before the phenological event (Table 1; SI Appendix,
+Section S1). Then, we used a greedy algorithm to select covariance
+matrices from each model's set that significantly improved the model
+likelihood (SI Appendix, Section S4; Table S2). Finally, we fit
+\emph{mash} models using the selected set of covariance matrices on a
+subset of relatively unlinked (r\textsuperscript{2} \textless{} 0.2),
+`strong' genetic effects with low \emph{p}-values (SI Appendix, Section
+S5).
 
@@
 non-zero weights on two additional GxWeather covariance types, average
-temperature one day prior to green-up, and the day length change in
-seconds in the day prior to green-up (Figure~\ref{fig-covar} B). The
-average temperature covariance matrix had negative covariances between
-Texas and North gardens, though not as strong as the negative phenotypic
+temperature one day prior to green-up, and the day length change in the
+day prior to green-up (Figure~\ref{fig-covar} B). The average
+temperature covariance matrix had negative covariances between Texas and
+North gardens, though not as strong as the negative phenotypic
 correlations observed in Both subpopulations (Figure~\ref{fig-map} B;
@@
 gardens, and single effect matrices with garden-specific effects.
+Re-estimation of genetic effects with \emph{mash} allowed us to identify
+and quantify the fraction of loci exhibiting GxWeather patterns; a
+minority of GxE was GxWeather covariance, and the majority of GxE
+covariance in these models is driven by other, unknown drivers of GxE.
 
@@
 individual loci as this is a key theoretical prediction of local
-adaptation (27--30). To determine the frequency of rank-changing GxE, we
-used the local false sign rate (lfsr), an analogue of the local false
+adaptation (27--30). Previous empirical work has found limited evidence
+of trade-offs caused by antagonistic pleiotropy; however, this work had
+a known statistical bias reducing the detection of effects that differed
+in sign (15, 31, 32). To determine the frequency of rank-changing GxE,
+we used the local false sign rate (lfsr), an analogue of the local false
 discovery rate that establishes confidence in the effect sign, not the
@@
 QTL on Chr04K overlapped a gene recently functionally validated for
-flowering in switchgrass (43). All flowering and green-up QTL intervals
+flowering in switchgrass (44). All flowering and green-up QTL intervals
 contained at least one SNP significant in at least one \emph{mash} run
@@
 underpinnings of these patterns, and the cases where this covariation
-can be altered. Here, we demonstrate that we can associate multiple
-patterns of GxWeather with specific genomic regions using a switchgrass
-diversity panel grown at eight common gardens. We assigned genetic
-effects to both GxWeather patterns with interpretable weather-based
-cues, and to unmeasured, site-based patterns. We used this approach to
-study GxWeather for the timings of vegetative and reproductive
-development in the deeply genetically diverged Gulf and Midwest
-subpopulations of switchgrass.
+can be altered through selection at individual loci. Here, we
+demonstrate that we can associate multiple patterns of GxWeather with
+specific genomic regions using a switchgrass diversity panel grown at
+eight common gardens. We assigned genetic effects to both GxWeather
+patterns with interpretable weather-based cues, and to unmeasured,
+site-based patterns. We used this approach to study GxWeather for the
+timings of vegetative and reproductive development in the deeply
+genetically diverged Gulf and Midwest subpopulations of switchgrass.
 
@@
 fitness, this result supports theoretical models that local adaptation
-should involve trade-offs at the level of individual loci (27--30). In
-addition, it advances local adaptation research by using more than two
-field sites and a wide range of genetic variation to determine that
-rank-changing GxE for phenological traits is common in small genomic
-regions (15, 44, 45).
+should involve trade-offs due to antagonistic pleiotropy at the level of
+individual loci (27--30). Experimental designs in local adaptation
+research now often use more than two field sites and a wide range of
+genetic variation; \emph{mash} and the local false sign rate facilitate
+the analysis of these experimental designs. We use \emph{mash} to
+determine that rank-changing GxE for phenological traits is common in
+small genomic regions (15, 45, 46).
 
@@
 independent mapping population at these sites (Figure~\ref{fig-qtl} B).
+Models combining both subpopulations showed less signal, perhaps due to
+the distinctness of cues in both subpopulations or to confounding caused
+by population structure.
 
@@
 
-\caption{\label{fig-covar}Example hypothesis-driven covariance matrices
-specified in \emph{mash} and the posterior weights placed on all
-covariance matrices. (a) Common gardens are arranged in latitudinal
+\caption{\label{fig-covar}Example Canonical and GxWeather covariance
+matrices specified in \emph{mash} and the posterior weights placed on
+all covariance matrices. (a) Common gardens are arranged in latitudinal
 order within the matrices. Top row: Four example canonical covariance
@@
 shown. (c,e) Total posterior weight placed on covariance matrices that
-were hypothesized or canonical, for the (c) green-up date phenotype and
-(e) flowering date phenotype.}
+were GxWeather or Canonical, for the (c) green-up date phenotype and (e)
+flowering date phenotype.}
 
@@
 
-\includegraphics{images/Figure_3_Fraction_GxE_by_Region_six_midwest_sites.png}
+\includegraphics{images/Figure_3_AH_comments.png}
 
@@
 \caption{\label{fig-effects}Types of GxE present between of pairs of
-jointly re-estimated effects in eight common gardens, for effects with
-lfsr \textless{} 0.05 at each pair of sites contrasted. a) Examples of
-effect patterns at three pairs of sites with three types of GxE. All
-effects are for an alternate allele, with the reference allele effect
-defined to be zero at both gardens and represented by the dashed
-vertical line. Sign: Effects that differ in sign at these pairs of
-gardens (p \textless{} 0.05, lfsr). Magnitude: Effects identical in sign
-(p \textless{} 0.05, lfsr) that differ in magnitude by a factor of
-\textgreater0.4. Not Distinguishable: Effects not distinguishable by
-magnitude nor sign of the effect, with no measurable GxE. b) The
-fraction of effects with each GxE type for the onset of vegetative
-growth (green-up date) and reproductive growth (flowering date), within
-and between two genetic subpopulations. Common gardens are grouped by
-the larger region they came from: North gardens are within the natural
-range of the Midwest subpopulation, and include MO, NE, MI, and SD,
-while Texas gardens are within the natural range of the Gulf
-subpopulation, and include TX1, TX2, and TX3.}
+jointly re-estimated SNP effects in eight common gardens, for effects
+with lfsr \textless{} 0.05 at both gardens for each pair of gardens
+contrasted. a) Examples of effect patterns at three pairs of sites with
+three types of GxE. All effects are for an alternate allele, with the
+reference allele effect defined to be zero at both gardens and
+represented by the dashed vertical line. Sign: Effects that differ in
+sign at these pairs of gardens (p \textless{} 0.05, lfsr). Magnitude:
+Effects identical in sign (p \textless{} 0.05, lfsr) that differ in
+magnitude by a factor of \textgreater0.4. Not Distinguishable: Effects
+not distinguishable by magnitude nor sign of the effect, with no
+measurable GxE. Confidence intervals are illustrative that the effect
+estimate does not overlap zero. b) The fraction of effects with each GxE
+type for the onset of vegetative growth (green-up date) and reproductive
+growth (flowering date), within and between two genetic subpopulations.
+Common gardens are grouped by the larger region they came from: North
+gardens are within the natural range of the Midwest subpopulation, and
+include MO, NE, MI, and SD, while Texas gardens are within the natural
+range of the Gulf subpopulation, and include TX1, TX2, and TX3.}
 
@@
 populations of switchgrass, a diversity panel and a pseudo-F2 cross,
-planted at eight common garden locations (38, 40, 45). We scored the
+planted at eight common garden locations (38, 40, 46). We scored the
 onset of vegetative growth, or green-up date, as the day of the year
@@
 different subpopulations had different strengths of connection between
-our phenotypes and genotypes (46), we conducted three sets of genetic
+our phenotypes and genotypes (47), we conducted three sets of genetic
 analyses: on Gulf and Midwest genotypes separately, and on both
@@
 allowed more than one form of GxE, as we reasoned that different loci
-should display different forms of GxE. In addition, we were interested
-in an unbiased estimation of the frequency of rank-changing GxE relative
-to other forms of GxE, as the presence of rank-changing GxE at the level
-of individual loci is a key theoretical prediction of local adaptation.
-\emph{Mash} allowed us to both specify multiple forms of GxE and
-GxWeather and conduct unbiased statistical tests for when SNP effects
-changed sign between common gardens.
+should display different forms of GxE (e.g., changes in effect magnitude
+vs sign). In addition, we were interested in an unbiased estimation of
+the frequency of rank-changing GxE relative to other forms of GxE, as
+the presence of rank-changing GxE at the level of individual loci is a
+key theoretical prediction of local adaptation. \emph{Mash} allowed us
+to both specify multiple forms of GxE and GxWeather and conduct unbiased
+statistical tests for when SNP effects changed sign between common
+gardens.
 
@@
 we first defined many covariance matrices, including GxWeather matrices
-that represented the correlation in weather cues between gardenes before
+that represented the correlation in weather cues between gardens before
 the phenological event (SI Appendix, Section S1), then implemented a
@@
 pruning and removal of regions in long-range linkage disequilibrium
-(47). Third, to make the \emph{mash} models computationally feasible, we
+(48). Third, to make the \emph{mash} models computationally feasible, we
 extracted two subsets of SNP effect estimates and standard errors from
@@
 null hypothesis that an effect j is 0, or that there is a significant
-effect. Previous studies of antagonistic pleiotropy (e.g. (45)) have
+effect. Previous studies of antagonistic pleiotropy (e.g. (46)) have
 used the local FDR or equivalent statistical tests to detect
@@
 for allelic effects that differ in sign does not undercount these
-occurences, as this statistic answers a fundamentally different
+occurrences, as this statistic answers a fundamentally different
 question. For each effect \(j\), the \(lfsr_j\) is defined as the
@@
 
+\bibitem[\citeproctext]{ref-weine2024}
+\CSLLeftMargin{43. }%
+\CSLRightInline{E. Weine, S. P. Smith, R. K. Knowlton, A. Harpak,
+\href{http://dx.doi.org/10.7554/eLife.99210.1}{Tradeoffs in modeling
+context dependency in complex trait genetics} (2024).}
+
 \bibitem[\citeproctext]{ref-choi2023}
-\CSLLeftMargin{43. }%
+\CSLLeftMargin{44. }%
 \CSLRightInline{S. Choi, \emph{et al.},
@@
 \bibitem[\citeproctext]{ref-savolainen2013ecological}
-\CSLLeftMargin{44. }%
+\CSLLeftMargin{45. }%
 \CSLRightInline{O. Savolainen, M. Lascoux, J. Merilä, Ecological
@@
 \bibitem[\citeproctext]{ref-lowry2019qtl}
-\CSLLeftMargin{45. }%
+\CSLLeftMargin{46. }%
 \CSLRightInline{D. B. Lowry, \emph{et al.}, QTL\(\times\) environment
@@
 \bibitem[\citeproctext]{ref-kortefarlow2013}
-\CSLLeftMargin{46. }%
+\CSLLeftMargin{47. }%
 \CSLRightInline{A. Korte, A. Farlow, The advantages and limitations of
@@
 \bibitem[\citeproctext]{ref-privuxe92017}
-\CSLLeftMargin{47. }%
+\CSLLeftMargin{48. }%
 \CSLRightInline{F. Privé, H. Aschard, M. G. B. Blum,
@@
 Award Numbers DE-SC0018409 and DE-FC02-07ER64494, the US Department of
-Energy Awards DESC0014156 to T.E.J., DE-SC0017883 to D.B.L, NIH grant
-R35GM151108 and a Pew Scholarship to A.H., National Science Foundation
-PGRP Awards IOS0922457 and IOS1444533 to T.E.J, and the Long-term
-Ecological Research Program (DEB 1832042) at the Kellogg Biological
-Station.}
+Energy Awards DE-SC0021126 and DESC0014156 to T.E.J., DE-SC0017883 to
+D.B.L, NIH grant R35GM151108 and a Pew Scholarship to A.H., National
+Science Foundation PGRP Awards IOS0922457 and IOS1444533 to T.E.J, and
+the Long-term Ecological Research Program (DEB 1832042) at the Kellogg
+Biological Station.}
 

### 2024-08-18 manuscript/revision2/revision_biorxiv.qmd
revision2: journal neutral format for bioRxiv revision

--- 2024-08-17 manuscript/revision2/revision2.qmd
@@
 format: 
-  pnas-pdf:
-    keep-tex: true  
   docx:
@@
       cache: false
+  simple-article-pdf:
+    keep-tex: true  
 author:
@@
       equal-contributor: true
-  - name: Samuel A. Smith
+  - name: Samuel P. Smith
     affiliations:
@@
 runningauthor: "MacQueen"
-significance: |
-  The timing of plant seasonal development (phenology) has major impacts on fitness because of the steep fitness cost of plant-environment mismatches.
-We infer the mixture of ways that genetic effects on phenological traits depend on plant environment, focusing on how effects covary with weather prior to the phenological event (GxWeather).
-Most effects do not have GxWeather, but a minority do.
-GxWeather is population-specific.
-The majority of effects on the timing of vegetative growth in the Gulf population have GxWeather: these effects differ in sign across environments and covary with a photoperiod cue two weeks prior.
-A minority of effects on flowering date in the Gulf and Midwest populations had GxWeather, covarying with a cumulative rainfall and a photoperiod cue, respectively.
-author_contributions:  "T.E.J.
-designed research.
-D.B.L.
-contributed plant material and resources.
-J.B., D.B.L., and T.E.J.
-designed and executed field experiments.
-A.R.B., P.A.F., F.B.F., D.B.L., R.B.M., F.M.R., Y.W., and T.E.J.
-hosted field experiments.
-A.H.M., L.Z., and S.A.S.
-conducted statistical and computational analyses with input from T.E.J.
-and A.H.
-The manuscript was written by A.H.M.
-with contributions from all authors."
-conflict_of_interest: "The authors declare no conflicts of interest."
-equal-contributor: "A.H.M. contributed equally to this work with L.Z. and S.A.S."
-keywords:  [allele-by-environment effect variation, rank-changing genotype-by-environment interaction, photoperiod, cumulative rainfall, genetic variation]
-acknowledgements: | 
-  We thank the Brackenridge Field laboratory, the Ladybird Johnson Wildflower Center, and the Juenger laboratory for support with plant care and propagation.
-This material is based upon work supported in part by the Great Lakes Bioenergy Research Center, U.S.
-Department of Energy, Office of Science, Office of Biological and Environmental Research under Award Numbers DE-SC0018409 and DE-FC02-07ER64494, the US Department of Energy Awards DE-SC0021126 and DESC0014156 to T.E.J., DE-SC0017883 to D.B.L, NIH grant R35GM151108 and a Pew Scholarship to A.H., National Science Foundation PGRP Awards IOS0922457 and IOS1444533 to T.E.J, and the Long-term Ecological Research Program (DEB 1832042) at the Kellogg Biological Station.
 bibliography: bibliography.bib 
@@
 
-\firstpage[13]{2}
+# Significance Statement
+
+The timing of plant seasonal development (phenology) has major impacts on fitness because of the steep fitness cost of plant-environment mismatches.
+We infer the mixture of ways that genetic effects on phenological traits depend on plant environment, focusing on how effects covary with weather prior to the phenological event (GxWeather).
+Most effects do not have GxWeather, but a minority do.
+GxWeather is population-specific.
+The majority of effects on the timing of vegetative growth in the Gulf population have GxWeather: these effects differ in sign across environments and covary with a photoperiod cue two weeks prior.
+A minority of effects on flowering date in the Gulf and Midwest populations had GxWeather, covarying with a cumulative rainfall and a photoperiod cue, respectively.
+
+# Introduction
 
@@
 
-# Figures and Tables
+# Author Contributions
 
-```{=tex}
-\clearpage
-\onecolumn
-\begin{table}[t!]
-\centering
-\caption{Weather variables and time frames prior to the start of vegetative and reproductive growth used to construct the hypothesis-based covariance matrices.
-Covariance matrices were constructed for, and tested on, three subpopulations and two phenological traits.
-The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices.
-Narrow-sense heritabilities for these values at each garden were used for the diagonal cells.}
-\begin{tabular}{lrr}
-Weather variable & Time Frame & Covariance Matrices \\
-\midrule
-1. cumulative growing degree days at 12C in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60 \\
-2. cumulative rainfall in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60  \\
-3. day length (hours) on a specific day indicated by the time frame & (1-7) and 14 days prior to the phenological date & 48 \\
-4. day length change (seconds) on a specific day indicated by the time frame & (1-7) and 14 days prior to the phenological date & 48 \\
-5. average temperature in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60 \\
-\bottomrule
-\end{tabular}
+T.E.J.
+designed research.
+D.B.L.
+contributed plant material and resources.
+J.B., D.B.L., and T.E.J.
+designed and executed field experiments.
+A.R.B., P.A.F., F.B.F., D.B.L., R.B.M., F.M.R., Y.W., and T.E.J.
+hosted field experiments.
+A.H.M., L.Z., and S.P.S.
+conducted statistical and computational analyses with input from T.E.J.
+and A.H.
+The manuscript was written by A.H.M.
+with contributions from all authors.
 
-\end{table}
-\clearpage
-\twocolumn
-```
-![Characterization of the timing of the onset of vegetative (green-up) and reproductive (flowering) growth in the switchgrass diversity panel.
-(a) Map and trait histograms of green-up and flowering dates across two genetically distinct switchgrass subpopulations and eight common gardens.
-Purple represents individuals from the Midwest genetic subpopulation, and pink individuals from the Gulf subpopulation; map positions represent the original collection locations for the genotypes, and shapes represent the ecotype of the genotype.
-Histogram vertical dashed lines indicate the summer solstice.
-Common gardens are arranged in latitudinal order.
-(b) Phenotypic correlations between clonal replicates planted at eight common gardens, within and between two genetic subpopulations.](images/Fig1_AB.png){#fig-map}
+# Conflict of Interest
 
-![Example Canonical and GxWeather covariance matrices specified in *mash* and the posterior weights placed on all covariance matrices.
-(a) Common gardens are arranged in latitudinal order within the matrices.
-Top row: Four example canonical covariance matrices.
-Canonical matrices (purple) have simple interpretations, such as equal effects across all common gardens, or effects specific to a single common garden.
-Bottom row: Five example GxWeather covariance matrices specified for the green-up date or flowering date phenotype; these matrices were created from environment-specific correlations across eight common gardens, and are described in Table 1.
-(b,d) Total posterior weight placed on each covariance matrix type specified for (b) green-up date and (d) flowering date *mash* models, within and between two genetic subpopulations.
-Covariance matrices included in *mash* that had zero posterior weight in all three *mash* runs on the genetic subpopulations, such as the identity matrix, are not shown.
-(c,e) Total posterior weight placed on covariance matrices that were GxWeather or Canonical, for the (c) green-up date phenotype and (e) flowering date phenotype.](images/Figure_2_six_sites_Midwest.png){#fig-covar}
+The authors declare no conflicts of interest.
 
-![Types of GxE present between of pairs of jointly re-estimated SNP effects in eight common gardens, for effects with lfsr \< 0.05 at both gardens for each pair of gardens contrasted.
-a) Examples of effect patterns at three pairs of sites with three types of GxE.
-All effects are for an alternate allele, with the reference allele effect defined to be zero at both gardens and represented by the dashed vertical line.
-Sign: Effects that differ in sign at these pairs of gardens (p \< 0.05, lfsr).
-Magnitude: Effects identical in sign (p \< 0.05, lfsr) that differ in magnitude by a factor of \>0.4.
-Not Distinguishable: Effects not distinguishable by magnitude nor sign of the effect, with no measurable GxE.
-Confidence intervals are illustrative that the effect estimate does not overlap zero.
-b) The fraction of effects with each GxE type for the onset of vegetative growth (green-up date) and reproductive growth (flowering date), within and between two genetic subpopulations.
-Common gardens are grouped by the larger region they came from: North gardens are within the natural range of the Midwest subpopulation, and include MO, NE, MI, and SD, while Texas gardens are within the natural range of the Gulf subpopulation, and include TX1, TX2, and TX3.](images/Figure_3_AH_comments.png){#fig-effects}
+# Equal Contributor
 
-![Overlaps of QTL from an outbred pseudo-F2 cross and with jointly re-estimated SNP effects in the 1% tail of significance from a diversity panel.
-Dotted lines indicate permutation-based significance thresholds for each weather-related function.
-Stars indicate QTL with significant enrichment for SNPs in the 1% *mash* tail; G, M, and B indicate which subpopulation had enrichment: G - Gulf subpopulation, M -Midwest subpopulation, B - both subpopulations.
-Rug plots show genomic locations of SNPs in the 1% *mash* tail for flowering date for each subpopulation: Both subpopulations are above the plot panel, the Gulf subpopulation is above the x-axis, and the Midwest subpopulation is below the x-axis.
-a) QTL mapping for the onset of vegetative growth (green-up date), and three weather-related functions of green-up date.
-b) QTL mapping for the onset of reproductive growth (flowering date), and two weather-related functions of flowering date.](images/Figure_4_QTL_Overlaps_6_Midwest_Gardens.png){#fig-qtl}
+A.H.M. contributed equally to this work with L.Z. and S.P.S.
+
+# Acknowledgements
+
+We thank the Brackenridge Field laboratory, the Ladybird Johnson Wildflower Center, and the Juenger laboratory for support with plant care and propagation.
+The authors acknowledge the Texas Advanced Computing Center (TACC) at The University of Texas at Austin for providing HPC storage resources that have contributed to the research results reported within this paper.
+This material is based upon work supported in part by the Great Lakes Bioenergy Research Center, U.S.
+Department of Energy, Office of Science, Office of Biological and Environmental Research under Award Numbers DE-SC0018409 and DE-FC02-07ER64494, the US Department of Energy Awards DE-SC0021126 and DESC0014156 to T.E.J., DE-SC0017883 to D.B.L, NIH grant R35GM151108 and a Pew Scholarship to A.H., National Science Foundation PGRP Awards IOS0922457 and IOS1444533 to T.E.J, and the Long-term Ecological Research Program (DEB 1832042) at the Kellogg Biological Station.
 
@@
 
+# Figures and Tables
+
+| Weather variable                                                          | Time Frame                                            | Covariance Matrices |
+|---------------------------------|------------------------|----------------|
+| cumulative growing degree days at 12C in the time frame                   | (1-7), 14, 21, 28 days prior to the phenological date | 60                  |
+| cumulative rainfall in the time frame                                     | (1-7), 14, 21, 28 days prior to the phenological date | 60                  |
+| day length (hours) on a specific day indicated by the time frame          | (1-7) and 14 days prior to the phenological date      | 48                  |
+| day length change (seconds) on a specific day indicated by the time frame | (1-7) and 14 days prior to the phenological date      | 48                  |
+| average temperature in the time frame                                     | (1-7), 14, 21, 28 days prior to the phenological date | 60                  |
+
+: Weather variables and time frames prior to the start of vegetative and reproductive growth used to construct the hypothesis-based covariance matrices.
+Covariance matrices were constructed for, and tested on, three subpopulations and two phenological traits.
+The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices.
+Narrow-sense heritabilities for these values at each garden were used for the diagonal cells.
+
+```{=tex}
+\clearpage
+\onecolumn
+\begin{table}[t!]
+\centering
+\caption{Weather variables and time frames prior to the start of vegetative and reproductive growth used to construct the hypothesis-based covariance matrices.
+Covariance matrices were constructed for, and tested on, three subpopulations and two phenological traits.
+The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices.
+Narrow-sense heritabilities for these values at each garden were used for the diagonal cells.}
+\begin{tabular}{lrr}
+Weather variable & Time Frame & Covariance Matrices \\
+\midrule
+1. cumulative growing degree days at 12C in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60 \\
+2. cumulative rainfall in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60  \\
+3. day length (hours) on a specific day indicated by the time frame & (1-7) and 14 days prior to the phenological date & 48 \\
+4. day length change (seconds) on a specific day indicated by the time frame & (1-7) and 14 days prior to the phenological date & 48 \\
+5. average temperature in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60 \\
+\bottomrule
+\end{tabular}
+
+\end{table}
+\clearpage
+\twocolumn
+```
+![Characterization of the timing of the onset of vegetative (green-up) and reproductive (flowering) growth in the switchgrass diversity panel.
+(a) Map and trait histograms of green-up and flowering dates across two genetically distinct switchgrass subpopulations and eight common gardens.
+Purple represents individuals from the Midwest genetic subpopulation, and pink individuals from the Gulf subpopulation; map positions represent the original collection locations for the genotypes, and shapes represent the ecotype of the genotype.
+Histogram vertical dashed lines indicate the summer solstice.
+Common gardens are arranged in latitudinal order.
+(b) Phenotypic correlations between clonal replicates planted at eight common gardens, within and between two genetic subpopulations.](images/Fig1_AB.png){#fig-map}
+
+![Example Canonical and GxWeather covariance matrices specified in *mash* and the posterior weights placed on all covariance matrices.
+(a) Common gardens are arranged in latitudinal order within the matrices.
+Top row: Four example canonical covariance matrices.
+Canonical matrices (purple) have simple interpretations, such as equal effects across all common gardens, or effects specific to a single common garden.
+Bottom row: Five example GxWeather covariance matrices specified for the green-up date or flowering date phenotype; these matrices were created from environment-specific correlations across eight common gardens, and are described in Table 1.
+(b,d) Total posterior weight placed on each covariance matrix type specified for (b) green-up date and (d) flowering date *mash* models, within and between two genetic subpopulations.
+Covariance matrices included in *mash* that had zero posterior weight in all three *mash* runs on the genetic subpopulations, such as the identity matrix, are not shown.
+(c,e) Total posterior weight placed on covariance matrices that were GxWeather or Canonical, for the (c) green-up date phenotype and (e) flowering date phenotype.](images/Figure_2_six_sites_Midwest.png){#fig-covar}
+
+![Types of GxE present between of pairs of jointly re-estimated SNP effects in eight common gardens, for effects with lfsr \< 0.05 at both gardens for each pair of gardens contrasted.
+a) Examples of effect patterns at three pairs of sites with three types of GxE.
+All effects are for an alternate allele, with the reference allele effect defined to be zero at both gardens and represented by the dashed vertical line.
+Sign: Effects that differ in sign at these pairs of gardens (p \< 0.05, lfsr).
+Magnitude: Effects identical in sign (p \< 0.05, lfsr) that differ in magnitude by a factor of \>0.4.
+Not Distinguishable: Effects not distinguishable by magnitude nor sign of the effect, with no measurable GxE.
+Confidence intervals are illustrative that the effect estimate does not overlap zero.
+b) The fraction of effects with each GxE type for the onset of vegetative growth (green-up date) and reproductive growth (flowering date), within and between two genetic subpopulations.
+Common gardens are grouped by the larger region they came from: North gardens are within the natural range of the Midwest subpopulation, and include MO, NE, MI, and SD, while Texas gardens are within the natural range of the Gulf subpopulation, and include TX1, TX2, and TX3.](images/Figure_3_AH_comments.png){#fig-effects}
+
+![Overlaps of QTL from an outbred pseudo-F2 cross and with jointly re-estimated SNP effects in the 1% tail of significance from a diversity panel.
+Dotted lines indicate permutation-based significance thresholds for each weather-related function.
+Stars indicate QTL with significant enrichment for SNPs in the 1% *mash* tail; G, M, and B indicate which subpopulation had enrichment: G - Gulf subpopulation, M -Midwest subpopulation, B - both subpopulations.
+Rug plots show genomic locations of SNPs in the 1% *mash* tail for flowering date for each subpopulation: Both subpopulations are above the plot panel, the Gulf subpopulation is above the x-axis, and the Midwest subpopulation is below the x-axis.
+a) QTL mapping for the onset of vegetative growth (green-up date), and three weather-related functions of green-up date.
+b) QTL mapping for the onset of reproductive growth (flowering date), and two weather-related functions of flowering date.](images/Figure_4_QTL_Overlaps_6_Midwest_Gardens.png){#fig-qtl}
+
+\
+
 # References
 
-\bibsplit[2]<!-- Use \bibsplit to split the references from the body of the text.
+<!-- \bibsplit[2] Use \bibsplit to split the references from the body of the text.
 Value "[2]" represents the number of reference in the left column (Note: Please avoid single column figures & tables on this page.) -->

### 2024-08-18 manuscript/revision2/revision_biorxiv.tex
revision2: journal neutral format for bioRxiv revision

--- 2024-08-17 manuscript/revision2/revision2.tex
@@
 \PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-%
 \documentclass[
@@
   twocolumn,
-  twoside]{pnas-new}
-
-\usepackage{amsmath,amssymb}
-\usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-\usepackage{lmodern}
-\ifPDFTeX\else  
-    % xetex/luatex font selection
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
+  twoside]{simple-article}% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
+\IfFileExists{flushend.sty}{\usepackage{flushend}}{}
+
 \makeatletter
@@
 \makeatother
-\usepackage{xcolor}
+
 \setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\setcounter{secnumdepth}{5}
 % Make \paragraph and \subparagraph free-standing
@@
 \makeatother
-% definitions for citeproc citations
-\NewDocumentCommand\citeproctext{}{}
-\NewDocumentCommand\citeproc{mm}{%
-  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
-\makeatletter
- % allow citations to break across lines
- \let\@cite@ofmt\@firstofone
- % avoid brackets around text for \cite:
- \def\@biblabel#1{}
- \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
-\makeatother
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
- {\begin{list}{}{%
-  \setlength{\itemindent}{0pt}
-  \setlength{\leftmargin}{0pt}
-  \setlength{\parsep}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-   \setlength{\leftmargin}{\cslhangindent}
-   \setlength{\itemindent}{-1\cslhangindent}
-  \fi
-  % set entry spacing
-  \setlength{\itemsep}{#2\baselineskip}}}
- {\end{list}}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-
-% TODO: Add custom LaTeX header directives here
+
+\usepackage{authblk}
+\usepackage{orcidlink}
+\definecolor{mypink}{RGB}{219, 48, 122}
 \makeatletter
@@
 \fi
-\usepackage{bookmark}
-
+\usepackage[]{natbib}
+\bibliographystyle{simple-article}
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
 \urlstyle{same} % disable monospaced font for URLs
+
 \hypersetup{
@@
   pdfauthor={Alice H.
-MacQueen; Li Zhang; Samuel A.
+MacQueen; Li Zhang; Samuel P.
 Smith; Jason Bonnette; Arvid R.
@@
 Juenger},
-  pdfkeywords={allele-by-environment effect variation, rank-changing
-genotype-by-environment interaction, photoperiod, cumulative
-rainfall, genetic variation},
   colorlinks=true,
-  linkcolor={blue},
-  filecolor={Maroon},
+  linkcolor={Blue},
+  filecolor={Blue},
   citecolor={Blue},
-  urlcolor={Blue},
+  urlcolor={Maroon},
   pdfcreator={LaTeX via pandoc}}
 
-\templatetype{pnasresearcharticle}
-
 \title{Diverse Genotype-by-Weather Interactions in Switchgrass}
 
-\newcommand{\equalcont}{1}
-
-\newcommand{\correspond}{2}
-
-\author[a%
-,\equalcont%
-,\correspond%
-]{Alice H. MacQueen}
-\author[a%
-,\equalcont%
-%
-]{Li Zhang}
-\author[a%
-,\equalcont%
-%
-]{Samuel A. Smith}
-\author[a%
-%
-%
-]{Jason Bonnette}
-\author[b%
-%
-%
-]{Arvid R. Boe}
-\author[c%
-%
-%
-]{Philip A. Fay}
-\author[d%
-%
-%
-]{Felix B. Fritschi}
-\author[e%
-%
-%
-]{David B. Lowry}
-\author[f%
-%
-%
-]{Robert B. Mitchell}
-\author[g%
-%
-%
-]{Francis M. Rouquette Jr}
-\author[h%
-%
-%
-]{Yanqi Wu}
-\author[a%
-%
-%
-]{Arbel Harpak}
-\author[a%
-%
-,\correspond%
-]{Thomas E. Juenger}
-
-\affil[a]{University of Texas at Austin, Department of Integrative
-Biology, Austin, 78712}
-\affil[b]{South Dakota State University, Department of
-Agronomy, Brookings, 57006}
-\affil[c]{USDA-ARS, Grassland, Soil and Water Research
-Laboratory, Temple, 76502}
-\affil[d]{University of Missouri, Division of Plant
-Sciences, Columbia, 65211}
-\affil[e]{Michigan State University, Department of Plant Biology, East
-Lansing, 48824}
-\affil[f]{USDA-ARS, Wheat, Sorghum, and Forage Research
-Unit, Lincoln, 68583}
-\affil[g]{Texas A\&M University, Texas A\&M AgriLife Research and
-Extension Center, Overton, 75684}
-\affil[h]{Oklahoma State University, Department of Plant and Soil
-Sciences, Stillwater, 74078}
-
-\leadauthor{MacQueen}
-
-\significancestatement{The timing of plant seasonal development
-(phenology) has major impacts on fitness because of the steep fitness
-cost of plant-environment mismatches. We infer the mixture of ways that
-genetic effects on phenological traits depend on plant environment,
-focusing on how effects covary with weather prior to the phenological
-event (GxWeather). Most effects do not have GxWeather, but a minority
-do. GxWeather is population-specific. The majority of effects on the
-timing of vegetative growth in the Gulf population have GxWeather: these
-effects differ in sign across environments and covary with a photoperiod
-cue two weeks prior. A minority of effects on flowering date in the Gulf
-and Midwest populations had GxWeather, covarying with a cumulative
-rainfall and a photoperiod cue, respectively.}
-
-\authorcontributions{T.E.J. designed research. D.B.L. contributed plant
-material and resources. J.B., D.B.L., and T.E.J. designed and executed
-field experiments. A.R.B., P.A.F., F.B.F., D.B.L., R.B.M., F.M.R., Y.W.,
-and T.E.J. hosted field experiments. A.H.M., L.Z., and S.A.S. conducted
-statistical and computational analyses with input from T.E.J. and A.H.
-The manuscript was written by A.H.M. with contributions from all
-authors.}
-
-\authordeclaration{The authors declare no conflicts of interest.}
-
-\equalauthors{\textsuperscript{\equalcont} A.H.M. contributed equally to
-this work with L.Z. and S.A.S.}
-
-\correspondingauthor{\textsuperscript{\correspond}To whom correspondence should be addressed. E-mail: alicem@uchicago.edu}
-\correspondingauthor{\textsuperscript{\correspond}To whom correspondence should be addressed. E-mail: tjuenger@utexas.edu}
-
-\keywords{allele-by-environment effect variation | rank-changing
-genotype-by-environment interaction | photoperiod | cumulative
-rainfall | genetic variation}
+\newif\ifnoAffil\noAffilfalse
+\newif\ifoneAffil\oneAffilfalse
+
+\newcommand{\setNumAffil}[1]{%
+    \ifnum#1=0
+        \noAffiltrue
+    \else
+        \ifnum#1=1
+            \oneAffiltrue
+        \fi
+    \fi
+}
+
+\setNumAffil{8}
+
+\ifnoAffil
+%% No affiliation given....
+\author{Alice H. MacQueen, Li Zhang, Samuel P. Smith, Jason
+Bonnette, Arvid R. Boe, Philip A. Fay, Felix B. Fritschi, David B.
+Lowry, Robert B. Mitchell, Francis M. Rouquette Jr, Yanqi Wu, Arbel
+Harpak, Thomas E. Juenger}
+\else
+\ifoneAffil
+%% Single affiliation given...
+\author{Alice H. MacQueen}
+\author{Li Zhang}
+\author{Samuel P. Smith}
+\author{Jason Bonnette}
+\author{Arvid R. Boe}
+\author{Philip A. Fay}
+\author{Felix B. Fritschi}
+\author{David B. Lowry}
+\author{Robert B. Mitchell}
+\author{Francis M. Rouquette Jr}
+\author{Yanqi Wu}
+\author{Arbel Harpak}
+\author{Thomas E. Juenger}
+\affil{University of Texas at Austin, Austin}
+\else
+%% Multiple affiliations given
+\author[1]{Alice H. MacQueen}
+\author[1]{Li Zhang}
+\author[1]{Samuel P. Smith}
+\author[1]{Jason Bonnette}
+\author[2]{Arvid R. Boe}
+\author[3]{Philip A. Fay}
+\author[4]{Felix B. Fritschi}
+\author[5]{David B. Lowry}
+\author[6]{Robert B. Mitchell}
+\author[7]{Francis M. Rouquette Jr}
+\author[8]{Yanqi Wu}
+\author[1]{Arbel Harpak}
+\author[1]{Thomas E. Juenger}
+\affil[1]{University of Texas at Austin, Austin}
+\affil[2]{South Dakota State University, Brookings}
+\affil[3]{USDA-ARS, Temple}
+\affil[4]{University of Missouri, Columbia}
+\affil[5]{Michigan State University, East Lansing}
+\affil[6]{USDA-ARS, Lincoln}
+\affil[7]{Texas A\&M University, Overton}
+\affil[8]{Oklahoma State University, Stillwater}
+\fi
+\fi
+
+\date{}
+
+\fancyhead[C]{Alice H. MacQueen $\bullet$ Diverse Genotype-by-Weather
+Interactions in Switchgrass}
 
 \begin{document}
-\maketitle
-
+
+\renewcommand{\maketitlehookd}{%
 \begin{abstract}
-The timing of vegetative and reproductive growth in plants
+\noindent{}The timing of vegetative and reproductive growth in plants
 (``phenological timings'') depend on genetic effects (G), environmental
@@
 \end{abstract}
-
-\dates{This manuscript was compiled on \today}
-\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
-
-\thispagestyle{firststyle}
-\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
-\firstpage[13]{2}
+}
+
+\maketitle
+
+\section{Significance Statement}\label{significance-statement}
+
+The timing of plant seasonal development (phenology) has major impacts
+on fitness because of the steep fitness cost of plant-environment
+mismatches. We infer the mixture of ways that genetic effects on
+phenological traits depend on plant environment, focusing on how effects
+covary with weather prior to the phenological event (GxWeather). Most
+effects do not have GxWeather, but a minority do. GxWeather is
+population-specific. The majority of effects on the timing of vegetative
+growth in the Gulf population have GxWeather: these effects differ in
+sign across environments and covary with a photoperiod cue two weeks
+prior. A minority of effects on flowering date in the Gulf and Midwest
+populations had GxWeather, covarying with a cumulative rainfall and a
+photoperiod cue, respectively.
+
+\section{Introduction}\label{introduction}
 
@@
 availability), which signal existing or upcoming growing conditions
-(1--4). Genetic responses to environmental cues determine the speed,
-timing, and energy apportioned to vegetative and reproductive growth and
-shape plant physiological responses, lifespan, and lifetime production
-of viable seed. Day length (or photoperiod) is one of the most
-predictable environmental cues, and genetic sensitivity to photoperiod
-protects plants from potentially fatal consequences of phenological
-responses to temperature cues at the ``wrong'' time of year. However,
-the utility of specific environmental cues depends on both features of
-the environment, such as cue predictability and relevance, and the
-species' adaptive strategies (5). Species with wide natural
+\citep{bauerle_photoperiodic_2012, andres2012genetic, korner2010phenology, maple2024}.
+Genetic responses to environmental cues determine the speed, timing, and
+energy apportioned to vegetative and reproductive growth and shape plant
+physiological responses, lifespan, and lifetime production of viable
+seed. Day length (or photoperiod) is one of the most predictable
+environmental cues, and genetic sensitivity to photoperiod protects
+plants from potentially fatal consequences of phenological responses to
+temperature cues at the ``wrong'' time of year. However, the utility of
+specific environmental cues depends on both features of the environment,
+such as cue predictability and relevance, and the species' adaptive
+strategies \citet{botero2015evolutionary}. Species with wide natural
 distributions can have multiple distinct environmentally cued
@@
 and facultative long-day flowering responses, which vary with their
-environments (6, 7). These distinct genetic responses in different
-environments are known as genotype-by-environment interactions, or GxE.
+environments \citep{blackman2013interacting, henry2014transitions}.
+These distinct genetic responses in different environments are known as
+genotype-by-environment interactions, or GxE.
 
 Flowering time, or the transition from vegetative to reproductive
-growth, is a common subject of GxE research (6--13), a key output of
-selection driving adaptation to local environments (2, 14--17), and a
-selection target for crop improvement to adapt crops to local or future
-environments (18). Changing flowering responsiveness to photoperiod cues
-has allowed geographic range expansion and increased yields in several
-cereal species (14, 19--23) and other crops (24, 25). Recent statistical
-advances in studying phenological GxE have involved determining critical
-environmental indices before the phenological event occurs, such as
-photothermal time within a critical growth window (11). However, most
-studies of flowering GxE focus on finding a single, best fitting form of
-genotype-environment covariance, despite the key expectation that
-different genetic subpopulations, and even different genomic regions,
-have likely evolved distinct patterns of GxE (26). Additionally, despite
-theoretical predictions that local adaptation should involve trade-offs
-caused by antagonistic pleiotropy, or alleles with effects with opposing
-fitness outcomes (27--30), previous empirical work has found limited
-evidence of trade-offs caused by this form of GxE (15, 31, 32). However,
+growth, is a common subject of GxE research
+\citep{blackman2013interacting, henry2014transitions, aagren2017adaptive, brachi2010linkage, dittmar2014flowering, li2018genomic, romero2017study, bustos-korts2019},
+a key output of selection driving adaptation to local environments
+\citep{unterseer2016, andres2012genetic, wadgymar2017identifying, blumel2015flowering, fournier-level2016},
+and a selection target for crop improvement to adapt crops to local or
+future environments \citet{jung2009flowering}. Changing flowering
+responsiveness to photoperiod cues has allowed geographic range
+expansion and increased yields in several cereal species
+\citep{unterseer2016, turner2005pseudo, faure2012mutation, hung2012zmcct, zakhrabekova2012induced, yang2013oself3}
+and other crops \citep{pin2012multifaceted, weller2019parallel}. Recent
+statistical advances in studying phenological GxE have involved
+determining critical environmental indices before the phenological event
+occurs, such as photothermal time within a critical growth window
+\citet{li2018genomic}. However, most studies of flowering GxE focus on
+finding a single, best fitting form of genotype-environment covariance,
+despite the key expectation that different genetic subpopulations, and
+even different genomic regions, have likely evolved distinct patterns of
+GxE \citep{Weine2023.06.21.545998}. Additionally, despite theoretical
+predictions that local adaptation should involve trade-offs caused by
+antagonistic pleiotropy, or alleles with effects with opposing fitness
+outcomes
+\citep{levene1953genetic, felsenstein1976theoretical, kawecki2004conceptual, hedrick1986genetic},
+previous empirical work has found limited evidence of trade-offs caused
+by this form of GxE
+\citep{des2013genotype, wadgymar2017identifying, millet2016}. However,
 this work has been limited by a known statistical bias that reduced
-detection of genetic effects that differ in sign (31, 33, 34). Thus,
-despite substantial interest in the frequencies of various forms of GxE,
-the frequency of sign-changing GxE relative to other forms of GxE
-remains unknown.
+detection of genetic effects that differ in sign
+\citep{des2013genotype, anderson2013genetic, anderson2011evolutionary}.
+Thus, despite substantial interest in the frequencies of various forms
+of GxE, the frequency of sign-changing GxE relative to other forms of
+GxE remains unknown.
 
@@
 subpopulation. Switchgrass is considered a short-day plant with
-reproductive development strongly linked to day of the year (35).
-However, as part of its wide environmental adaptation across the eastern
-half of North America, its photoperiodicity has been predicted to differ
-by plant latitude of origin (36, 37). We previously found that divergent
-Midwest and Gulf genetic subpopulations of switchgrass have distinct
-sets of environmental adaptations associated with each of two fitness
-proxies, biomass and overwinter survival (38). The Midwest genetic
-subpopulation is primarily composed of individuals from the well-studied
-upland switchgrass ecotype (39, 40), while the Gulf subpopulation has
-individuals from the well-studied lowland ecotype and the phenotypically
-intermediate coastal ecotype (38).
+reproductive development strongly linked to day of the year
+\citet{mitchell1997predicting}. However, as part of its wide
+environmental adaptation across the eastern half of North America, its
+photoperiodicity has been predicted to differ by plant latitude of
+origin \citep{parrish2005biology, casler2004latitudinal}. We previously
+found that divergent Midwest and Gulf genetic subpopulations of
+switchgrass have distinct sets of environmental adaptations associated
+with each of two fitness proxies, biomass and overwinter survival
+\citet{lovell2021}. The Midwest genetic subpopulation is primarily
+composed of individuals from the well-studied upland switchgrass ecotype
+\citep{porter1966analysis, milano2016genetic}, while the Gulf
+subpopulation has individuals from the well-studied lowland ecotype and
+the phenotypically intermediate coastal ecotype \citet{lovell2021}.
 
@@
 significantly improved the modeled log-likelihood when included (SI
-Appendix, Section S2-S5) (41). We used the Bayesian framework
-\emph{mash} (multivariate adaptive shrinkage) developed by (41), to
-refine effect size estimates from genome-wide association (GWAS).
-\emph{Mash} allowed us to identify and specify multiple covariance
-structures among genetic effect estimates across sites, including
-structures that represent covariance in weather variables of interest.
-Importantly, this method circumvented statistical biases in detecting
-genetic effects with the same or opposite signs (42). To confirm our
-genetic mapping of GxWeather, we compared genomic locations of the
-significant posterior effect estimates from \emph{mash} to mapping
-results from an outbred mapping population grown at the same sites. Our
-analyses allowed us to describe the weather cues and types of GxE
-affecting phenology in two divergent natural populations of switchgrass
-across the species' latitudinal range.
+Appendix, Section S2-S5) \citet{urbut2019flexible}. We used the Bayesian
+framework \emph{mash} (multivariate adaptive shrinkage) developed by
+\citet{urbut2019flexible}, to refine effect size estimates from
+genome-wide association (GWAS). \emph{Mash} allowed us to identify and
+specify multiple covariance structures among genetic effect estimates
+across sites, including structures that represent covariance in weather
+variables of interest. Importantly, this method circumvented statistical
+biases in detecting genetic effects with the same or opposite signs
+\citet{10.1093/biostatistics/kxw041}. To confirm our genetic mapping of
+GxWeather, we compared genomic locations of the significant posterior
+effect estimates from \emph{mash} to mapping results from an outbred
+mapping population grown at the same sites. Our analyses allowed us to
+describe the weather cues and types of GxE affecting phenology in two
+divergent natural populations of switchgrass across the species'
+latitudinal range.
 
@@
 different ways polygenic effects might covary across our eight common
-gardens(43). We jointly re-estimated the genetic effects of a subset of
-SNPs across all eight common gardens using \emph{mash} models, which can
-flexibly capture many modes of covariance across gardens (SI Appendix,
-Sections S1-S5, Datasets 1-6). To obtain genetic effects, we used
-subsets of effects from garden- and subpopulation GWAS (SI Appendix,
-Section S2-S3; Table S1). To obtain covariance structures, we specified
-both covariance structures introduced in the initial \emph{mash}
-manuscript (e.g., garden-specific effects; equal effects at all gardens,
-Figure~\ref{fig-covar} A), and GxWeather covariance matrices estimated
-from the covariance of empirical weather patterns at each garden at
-specific times before the phenological event (Table 1; SI Appendix,
-Section S1). Then, we used a greedy algorithm to select covariance
-matrices from each model's set that significantly improved the model
-likelihood (SI Appendix, Section S4; Table S2). Finally, we fit
-\emph{mash} models using the selected set of covariance matrices on a
-subset of relatively unlinked (r\textsuperscript{2} \textless{} 0.2),
+gardens\citep{weine2024}. We jointly re-estimated the genetic effects of
+a subset of SNPs across all eight common gardens using \emph{mash}
+models, which can flexibly capture many modes of covariance across
+gardens (SI Appendix, Sections S1-S5, Datasets 1-6). To obtain genetic
+effects, we used subsets of effects from garden- and subpopulation GWAS
+(SI Appendix, Section S2-S3; Table S1). To obtain covariance structures,
+we specified both covariance structures introduced in the initial
+\emph{mash} manuscript (e.g., garden-specific effects; equal effects at
+all gardens, Figure~\ref{fig-covar} A), and GxWeather covariance
+matrices estimated from the covariance of empirical weather patterns at
+each garden at specific times before the phenological event (Table 1; SI
+Appendix, Section S1). Then, we used a greedy algorithm to select
+covariance matrices from each model's set that significantly improved
+the model likelihood (SI Appendix, Section S4; Table S2). Finally, we
+fit \emph{mash} models using the selected set of covariance matrices on
+a subset of relatively unlinked (r\textsuperscript{2} \textless{} 0.2),
 `strong' genetic effects with low \emph{p}-values (SI Appendix, Section
@@
 individual loci as this is a key theoretical prediction of local
-adaptation (27--30). Previous empirical work has found limited evidence
-of trade-offs caused by antagonistic pleiotropy; however, this work had
-a known statistical bias reducing the detection of effects that differed
-in sign (15, 31, 32). To determine the frequency of rank-changing GxE,
-we used the local false sign rate (lfsr), an analogue of the local false
-discovery rate that establishes confidence in the effect sign, not the
-effect's difference from zero, to determine significance. We required
-lfsr significance (p \textless{} 0.05) in both gardens to include
-effects. This means that our tests for a sign change between gardens
-carry an equal statistical burden to those for effects with the same
-sign. We separated kinds of effects at the level of individual loci into
-SNP effects that differ in sign between gardens (effects with
-rank-changing GxE between gardens), SNP effects that differ in magnitude
-(effects that are large in one garden or region and smaller in others),
-and SNP effects that are indistinguishable in two gardens (similarly
-large or small in comparisons between gardens) (Figure~\ref{fig-effects}
-A).
+adaptation
+\citep{felsenstein1976theoretical, levene1953genetic, kawecki2004conceptual, hedrick1986genetic}.
+Previous empirical work has found limited evidence of trade-offs caused
+by antagonistic pleiotropy; however, this work had a known statistical
+bias reducing the detection of effects that differed in sign
+\citep{wadgymar2017identifying, des2013genotype, millet2016}. To
+determine the frequency of rank-changing GxE, we used the local false
+sign rate (lfsr), an analogue of the local false discovery rate that
+establishes confidence in the effect sign, not the effect's difference
+from zero, to determine significance. We required lfsr significance (p
+\textless{} 0.05) in both gardens to include effects. This means that
+our tests for a sign change between gardens carry an equal statistical
+burden to those for effects with the same sign. We separated kinds of
+effects at the level of individual loci into SNP effects that differ in
+sign between gardens (effects with rank-changing GxE between gardens),
+SNP effects that differ in magnitude (effects that are large in one
+garden or region and smaller in others), and SNP effects that are
+indistinguishable in two gardens (similarly large or small in
+comparisons between gardens) (Figure~\ref{fig-effects} A).
 
@@
 QTL on Chr04K overlapped a gene recently functionally validated for
-flowering in switchgrass (44). All flowering and green-up QTL intervals
-contained at least one SNP significant in at least one \emph{mash} run
-at a log10-transformed Bayes Factor \textgreater{} 2, or in the 1\% tail
-of significance, whichever was stricter (SI Appendix, Dataset 8). We
-also looked for enrichments of \emph{mash} SNPs in the 1\% tail of
-significance (the `\emph{mash} 1\% tail') within each QTL interval. At
-the 5\% level, three QTL had enrichments of SNPs in the \emph{mash} 1\%
-tail. Overall, there were five significant enrichments (p \textless{}
-0.05, hypergeometric test) of SNPs in the \emph{mash} 1\% tail in the
-QTL intervals. Thus, we were able to experimentally support the genomic
-windows of some re-estimates of significant SNP effects with a QTL
-mapping experiment using a separate mapping population.
+flowering in switchgrass \citep{choi2023}. All flowering and green-up
+QTL intervals contained at least one SNP significant in at least one
+\emph{mash} run at a log10-transformed Bayes Factor \textgreater{} 2, or
+in the 1\% tail of significance, whichever was stricter (SI Appendix,
+Dataset 8). We also looked for enrichments of \emph{mash} SNPs in the
+1\% tail of significance (the `\emph{mash} 1\% tail') within each QTL
+interval. At the 5\% level, three QTL had enrichments of SNPs in the
+\emph{mash} 1\% tail. Overall, there were five significant enrichments
+(p \textless{} 0.05, hypergeometric test) of SNPs in the \emph{mash} 1\%
+tail in the QTL intervals. Thus, we were able to experimentally support
+the genomic windows of some re-estimates of significant SNP effects with
+a QTL mapping experiment using a separate mapping population.
 
@@
 should involve trade-offs due to antagonistic pleiotropy at the level of
-individual loci (27--30). Experimental designs in local adaptation
-research now often use more than two field sites and a wide range of
-genetic variation; \emph{mash} and the local false sign rate facilitate
-the analysis of these experimental designs. We use \emph{mash} to
-determine that rank-changing GxE for phenological traits is common in
-small genomic regions (15, 45, 46).
+individual loci
+\citep{kawecki2004conceptual, felsenstein1976theoretical, levene1953genetic, hedrick1986genetic}.
+Experimental designs in local adaptation research now often use more
+than two field sites and a wide range of genetic variation; \emph{mash}
+and the local false sign rate facilitate the analysis of these
+experimental designs. We use \emph{mash} to determine that rank-changing
+GxE for phenological traits is common in small genomic regions
+\citep{wadgymar2017identifying, savolainen2013ecological, lowry2019qtl}.
 
@@
 
-\section{Figures and Tables}\label{figures-and-tables}
-
-\clearpage
-\onecolumn
-\begin{table}[t!]
-\centering
-\caption{Weather variables and time frames prior to the start of vegetative and reproductive growth used to construct the hypothesis-based covariance matrices.
-Covariance matrices were constructed for, and tested on, three subpopulations and two phenological traits.
-The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices.
-Narrow-sense heritabilities for these values at each garden were used for the diagonal cells.}
-\begin{tabular}{lrr}
-Weather variable & Time Frame & Covariance Matrices \\
-\midrule
-1. cumulative growing degree days at 12C in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60 \\
-2. cumulative rainfall in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60  \\
-3. day length (hours) on a specific day indicated by the time frame & (1-7) and 14 days prior to the phenological date & 48 \\
-4. day length change (seconds) on a specific day indicated by the time frame & (1-7) and 14 days prior to the phenological date & 48 \\
-5. average temperature in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60 \\
-\bottomrule
-\end{tabular}
-
-\end{table}
-\clearpage
-\twocolumn
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{images/Fig1_AB.png}
-
-}
-
-\caption{\label{fig-map}Characterization of the timing of the onset of
-vegetative (green-up) and reproductive (flowering) growth in the
-switchgrass diversity panel. (a) Map and trait histograms of green-up
-and flowering dates across two genetically distinct switchgrass
-subpopulations and eight common gardens. Purple represents individuals
-from the Midwest genetic subpopulation, and pink individuals from the
-Gulf subpopulation; map positions represent the original collection
-locations for the genotypes, and shapes represent the ecotype of the
-genotype. Histogram vertical dashed lines indicate the summer solstice.
-Common gardens are arranged in latitudinal order. (b) Phenotypic
-correlations between clonal replicates planted at eight common gardens,
-within and between two genetic subpopulations.}
-
-\end{figure}%
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{images/Figure_2_six_sites_Midwest.png}
-
-}
-
-\caption{\label{fig-covar}Example Canonical and GxWeather covariance
-matrices specified in \emph{mash} and the posterior weights placed on
-all covariance matrices. (a) Common gardens are arranged in latitudinal
-order within the matrices. Top row: Four example canonical covariance
-matrices. Canonical matrices (purple) have simple interpretations, such
-as equal effects across all common gardens, or effects specific to a
-single common garden. Bottom row: Five example GxWeather covariance
-matrices specified for the green-up date or flowering date phenotype;
-these matrices were created from environment-specific correlations
-across eight common gardens, and are described in Table 1. (b,d) Total
-posterior weight placed on each covariance matrix type specified for (b)
-green-up date and (d) flowering date \emph{mash} models, within and
-between two genetic subpopulations. Covariance matrices included in
-\emph{mash} that had zero posterior weight in all three \emph{mash} runs
-on the genetic subpopulations, such as the identity matrix, are not
-shown. (c,e) Total posterior weight placed on covariance matrices that
-were GxWeather or Canonical, for the (c) green-up date phenotype and (e)
-flowering date phenotype.}
-
-\end{figure}%
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{images/Figure_3_AH_comments.png}
-
-}
-
-\caption{\label{fig-effects}Types of GxE present between of pairs of
-jointly re-estimated SNP effects in eight common gardens, for effects
-with lfsr \textless{} 0.05 at both gardens for each pair of gardens
-contrasted. a) Examples of effect patterns at three pairs of sites with
-three types of GxE. All effects are for an alternate allele, with the
-reference allele effect defined to be zero at both gardens and
-represented by the dashed vertical line. Sign: Effects that differ in
-sign at these pairs of gardens (p \textless{} 0.05, lfsr). Magnitude:
-Effects identical in sign (p \textless{} 0.05, lfsr) that differ in
-magnitude by a factor of \textgreater0.4. Not Distinguishable: Effects
-not distinguishable by magnitude nor sign of the effect, with no
-measurable GxE. Confidence intervals are illustrative that the effect
-estimate does not overlap zero. b) The fraction of effects with each GxE
-type for the onset of vegetative growth (green-up date) and reproductive
-growth (flowering date), within and between two genetic subpopulations.
-Common gardens are grouped by the larger region they came from: North
-gardens are within the natural range of the Midwest subpopulation, and
-include MO, NE, MI, and SD, while Texas gardens are within the natural
-range of the Gulf subpopulation, and include TX1, TX2, and TX3.}
-
-\end{figure}%
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{images/Figure_4_QTL_Overlaps_6_Midwest_Gardens.png}
-
-}
-
-\caption{\label{fig-qtl}Overlaps of QTL from an outbred pseudo-F2 cross
-and with jointly re-estimated SNP effects in the 1\% tail of
-significance from a diversity panel. Dotted lines indicate
-permutation-based significance thresholds for each weather-related
-function. Stars indicate QTL with significant enrichment for SNPs in the
-1\% \emph{mash} tail; G, M, and B indicate which subpopulation had
-enrichment: G - Gulf subpopulation, M -Midwest subpopulation, B - both
-subpopulations. Rug plots show genomic locations of SNPs in the 1\%
-\emph{mash} tail for flowering date for each subpopulation: Both
-subpopulations are above the plot panel, the Gulf subpopulation is above
-the x-axis, and the Midwest subpopulation is below the x-axis. a) QTL
-mapping for the onset of vegetative growth (green-up date), and three
-weather-related functions of green-up date. b) QTL mapping for the onset
-of reproductive growth (flowering date), and two weather-related
-functions of flowering date.}
-
-\end{figure}%
+\section{Author Contributions}\label{author-contributions}
+
+T.E.J. designed research. D.B.L. contributed plant material and
+resources. J.B., D.B.L., and T.E.J. designed and executed field
+experiments. A.R.B., P.A.F., F.B.F., D.B.L., R.B.M., F.M.R., Y.W., and
+T.E.J. hosted field experiments. A.H.M., L.Z., and S.P.S. conducted
+statistical and computational analyses with input from T.E.J. and A.H.
+The manuscript was written by A.H.M. with contributions from all
+authors.
+
+\section{Conflict of Interest}\label{conflict-of-interest}
+
+The authors declare no conflicts of interest.
+
+\section{Equal Contributor}\label{equal-contributor}
+
+A.H.M. contributed equally to this work with L.Z. and S.P.S.
+
+\section{Acknowledgements}\label{acknowledgements}
+
+We thank the Brackenridge Field laboratory, the Ladybird Johnson
+Wildflower Center, and the Juenger laboratory for support with plant
+care and propagation. The authors acknowledge the Texas Advanced
+Computing Center (TACC) at The University of Texas at Austin for
+providing HPC storage resources that have contributed to the research
+results reported within this paper.This material is based upon work
+supported in part by the Great Lakes Bioenergy Research Center, U.S.
+Department of Energy, Office of Science, Office of Biological and
+Environmental Research under Award Numbers DE-SC0018409 and
+DE-FC02-07ER64494, the US Department of Energy Awards DE-SC0021126 and
+DESC0014156 to T.E.J., DE-SC0017883 to D.B.L, NIH grant R35GM151108 and
+a Pew Scholarship to A.H., National Science Foundation PGRP Awards
+IOS0922457 and IOS1444533 to T.E.J, and the Long-term Ecological
+Research Program (DEB 1832042) at the Kellogg Biological Station.
 
@@
 populations of switchgrass, a diversity panel and a pseudo-F2 cross,
-planted at eight common garden locations (38, 40, 46). We scored the
-onset of vegetative growth, or green-up date, as the day of the year
-when 50\% of the tiller area of the crown of the plant cut the previous
-year had green growth. The onset of reproductive growth, or flowering
-date, was the day of the year when 50\% of the plant tillers had
-panicles undergoing anthesis.
+planted at eight common garden locations
+\citep{lovell2021, milano2016genetic, lowry2019qtl}. We scored the onset
+of vegetative growth, or green-up date, as the day of the year when 50\%
+of the tiller area of the crown of the plant cut the previous year had
+green growth. The onset of reproductive growth, or flowering date, was
+the day of the year when 50\% of the plant tillers had panicles
+undergoing anthesis.
 
 The formation and resequencing of the diversity panel has been described
-previously (38). The diversity panel contained 134 sequenced, clonally
-propagated individuals from the Midwest genetic subpopulation, and 229
-from the Gulf genetic subpopulation. To allow for the possibility that
-different subpopulations had different strengths of connection between
-our phenotypes and genotypes (47), we conducted three sets of genetic
-analyses: on Gulf and Midwest genotypes separately, and on both
-subpopulations together (`Both' subpopulations). Analyses to determine
-narrow-sense heritability (h\textsuperscript{2}) for green-up and
-flowering were done using linear mixed models and followed (38) (SI
-Appendix, Section S6).
+previously \citet{lovell2021}. The diversity panel contained 134
+sequenced, clonally propagated individuals from the Midwest genetic
+subpopulation, and 229 from the Gulf genetic subpopulation. To allow for
+the possibility that different subpopulations had different strengths of
+connection between our phenotypes and genotypes \citep{kortefarlow2013},
+we conducted three sets of genetic analyses: on Gulf and Midwest
+genotypes separately, and on both subpopulations together (`Both'
+subpopulations). Analyses to determine narrow-sense heritability
+(h\textsuperscript{2}) for green-up and flowering were done using linear
+mixed models and followed \citet{lovell2021} (SI Appendix, Section S6).
 
@@
 between four individuals, two Midwest and two Gulf individuals. The
-formation of this mapping population has been described previously (40);
-additional details on QTL mapping can be found in SI Appendix, Section
-S7. To be directly comparable to the diversity panel data, only 2019
-phenology data from the pseudo-F2 cross from the same eight common
-garden sites were used.
+formation of this mapping population has been described previously
+\citep{milano2016genetic}; additional details on QTL mapping can be
+found in SI Appendix, Section S7. To be directly comparable to the
+diversity panel data, only 2019 phenology data from the pseudo-F2 cross
+from the same eight common garden sites were used.
 
@@
 pruning and removal of regions in long-range linkage disequilibrium
-(48). Third, to make the \emph{mash} models computationally feasible, we
-extracted two subsets of SNP effect estimates and standard errors from
-our GWAS effect estimates: (i) effects from a subset of ``strong'' tests
-corresponding to stronger effects on our traits; (ii) results from a
-random subset of all tests to correspond to an unbiased representation
-of all effects (SI Appendix, Section S3). We used the subset of random
-effects in our greedy \emph{mash} algorithm (SI Appendix, Section S4)
-and used the subset of strong effects in our \emph{mash} models with the
-optimal set of covariance matrices (SI Appendix, Section S5).
+\citep{privé2017}. Third, to make the \emph{mash} models computationally
+feasible, we extracted two subsets of SNP effect estimates and standard
+errors from our GWAS effect estimates: (i) effects from a subset of
+``strong'' tests corresponding to stronger effects on our traits; (ii)
+results from a random subset of all tests to correspond to an unbiased
+representation of all effects (SI Appendix, Section S3). We used the
+subset of random effects in our greedy \emph{mash} algorithm (SI
+Appendix, Section S4) and used the subset of strong effects in our
+\emph{mash} models with the optimal set of covariance matrices (SI
+Appendix, Section S5).
 
@@
 null hypothesis that an effect j is 0, or that there is a significant
-effect. Previous studies of antagonistic pleiotropy (e.g. (46)) have
-used the local FDR or equivalent statistical tests to detect
-antagonistic pleiotropy. These tests were conservative, in that they
-required two non-zero effects of different signs, while tests for
-differential sensitivity required only one non-zero effect. This
-previous work recognized that this testing bias could lead to
-undercounting occurrences of antagonistic pleiotropy (31, 34), and
-sought to reduce it by permutation (33). However, using the lfsr to test
-for allelic effects that differ in sign does not undercount these
-occurrences, as this statistic answers a fundamentally different
+effect. Previous studies of antagonistic pleiotropy (e.g.
+\citep{lowry2019qtl}) have used the local FDR or equivalent statistical
+tests to detect antagonistic pleiotropy. These tests were conservative,
+in that they required two non-zero effects of different signs, while
+tests for differential sensitivity required only one non-zero effect.
+This previous work recognized that this testing bias could lead to
+undercounting occurrences of antagonistic pleiotropy
+\citep{anderson2011evolutionary, des2013genotype}, and sought to reduce
+it by permutation \citep{anderson2013genetic}. However, using the lfsr
+to test for allelic effects that differ in sign does not undercount
+these occurrences, as this statistic answers a fundamentally different
 question. For each effect \(j\), the \(lfsr_j\) is defined as the
 probability that we make an error in the sign of effect \(j\) if we were
-forced to declare the effect positive or negative (42). Thus, rather
-than asking ``Are these two effects different?'' - as we reasonably
-expect two effects to be, even if this difference cannot be measured -
-the local false sign rate answers a more meaningful question: Can we be
-confident in the sign of this effect?
+forced to declare the effect positive or negative
+\citet{10.1093/biostatistics/kxw041}. Thus, rather than asking ``Are
+these two effects different?'' - as we reasonably expect two effects to
+be, even if this difference cannot be measured - the local false sign
+rate answers a more meaningful question: Can we be confident in the sign
+of this effect?
 
@@
 
+\section{Figures and Tables}\label{figures-and-tables}
+
+\clearpage
+\onecolumn
+\begin{table}[t!]
+\centering
+\caption{Weather variables and time frames prior to the start of vegetative and reproductive growth used to construct the hypothesis-based covariance matrices.
+Covariance matrices were constructed for, and tested on, three subpopulations and two phenological traits.
+The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices.
+Narrow-sense heritabilities for these values at each garden were used for the diagonal cells.}
+\begin{tabular}{lrr}
+Weather variable & Time Frame & Covariance Matrices \\
+\midrule
+1. cumulative growing degree days at 12C in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60 \\
+2. cumulative rainfall in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60  \\
+3. day length (hours) on a specific day indicated by the time frame & (1-7) and 14 days prior to the phenological date & 48 \\
+4. day length change (seconds) on a specific day indicated by the time frame & (1-7) and 14 days prior to the phenological date & 48 \\
+5. average temperature in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60 \\
+\bottomrule
+\end{tabular}
+
+\end{table}
+\clearpage
+\twocolumn
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{images/Fig1_AB.png}
+
+}
+
+\caption{\label{fig-map}Characterization of the timing of the onset of
+vegetative (green-up) and reproductive (flowering) growth in the
+switchgrass diversity panel. (a) Map and trait histograms of green-up
+and flowering dates across two genetically distinct switchgrass
+subpopulations and eight common gardens. Purple represents individuals
+from the Midwest genetic subpopulation, and pink individuals from the
+Gulf subpopulation; map positions represent the original collection
+locations for the genotypes, and shapes represent the ecotype of the
+genotype. Histogram vertical dashed lines indicate the summer solstice.
+Common gardens are arranged in latitudinal order. (b) Phenotypic
+correlations between clonal replicates planted at eight common gardens,
+within and between two genetic subpopulations.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{images/Figure_2_six_sites_Midwest.png}
+
+}
+
+\caption{\label{fig-covar}Example Canonical and GxWeather covariance
+matrices specified in \emph{mash} and the posterior weights placed on
+all covariance matrices. (a) Common gardens are arranged in latitudinal
+order within the matrices. Top row: Four example canonical covariance
+matrices. Canonical matrices (purple) have simple interpretations, such
+as equal effects across all common gardens, or effects specific to a
+single common garden. Bottom row: Five example GxWeather covariance
+matrices specified for the green-up date or flowering date phenotype;
+these matrices were created from environment-specific correlations
+across eight common gardens, and are described in Table 1. (b,d) Total
+posterior weight placed on each covariance matrix type specified for (b)
+green-up date and (d) flowering date \emph{mash} models, within and
+between two genetic subpopulations. Covariance matrices included in
+\emph{mash} that had zero posterior weight in all three \emph{mash} runs
+on the genetic subpopulations, such as the identity matrix, are not
+shown. (c,e) Total posterior weight placed on covariance matrices that
+were GxWeather or Canonical, for the (c) green-up date phenotype and (e)
+flowering date phenotype.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{images/Figure_3_AH_comments.png}
+
+}
+
+\caption{\label{fig-effects}Types of GxE present between of pairs of
+jointly re-estimated SNP effects in eight common gardens, for effects
+with lfsr \textless{} 0.05 at both gardens for each pair of gardens
+contrasted. a) Examples of effect patterns at three pairs of sites with
+three types of GxE. All effects are for an alternate allele, with the
+reference allele effect defined to be zero at both gardens and
+represented by the dashed vertical line. Sign: Effects that differ in
+sign at these pairs of gardens (p \textless{} 0.05, lfsr). Magnitude:
+Effects identical in sign (p \textless{} 0.05, lfsr) that differ in
+magnitude by a factor of \textgreater0.4. Not Distinguishable: Effects
+not distinguishable by magnitude nor sign of the effect, with no
+measurable GxE. Confidence intervals are illustrative that the effect
+estimate does not overlap zero. b) The fraction of effects with each GxE
+type for the onset of vegetative growth (green-up date) and reproductive
+growth (flowering date), within and between two genetic subpopulations.
+Common gardens are grouped by the larger region they came from: North
+gardens are within the natural range of the Midwest subpopulation, and
+include MO, NE, MI, and SD, while Texas gardens are within the natural
+range of the Gulf subpopulation, and include TX1, TX2, and TX3.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{images/Figure_4_QTL_Overlaps_6_Midwest_Gardens.png}
+
+}
+
+\caption{\label{fig-qtl}Overlaps of QTL from an outbred pseudo-F2 cross
+and with jointly re-estimated SNP effects in the 1\% tail of
+significance from a diversity panel. Dotted lines indicate
+permutation-based significance thresholds for each weather-related
+function. Stars indicate QTL with significant enrichment for SNPs in the
+1\% \emph{mash} tail; G, M, and B indicate which subpopulation had
+enrichment: G - Gulf subpopulation, M -Midwest subpopulation, B - both
+subpopulations. Rug plots show genomic locations of SNPs in the 1\%
+\emph{mash} tail for flowering date for each subpopulation: Both
+subpopulations are above the plot panel, the Gulf subpopulation is above
+the x-axis, and the Midwest subpopulation is below the x-axis. a) QTL
+mapping for the onset of vegetative growth (green-up date), and three
+weather-related functions of green-up date. b) QTL mapping for the onset
+of reproductive growth (flowering date), and two weather-related
+functions of flowering date.}
+
+\end{figure}%
+
+\hfill\break
+~
+
 \section{References}\label{references}
 
-\bibsplit[2]
-
-\phantomsection\label{refs}
-\begin{CSLReferences}{0}{1}
-\bibitem[\citeproctext]{ref-bauerle_photoperiodic_2012}
-\CSLLeftMargin{1. }%
-\CSLRightInline{W. L. Bauerle, \emph{et al.},
-\href{https://doi.org/10.1073/pnas.1119131109}{Photoperiodic regulation
-of the seasonal pattern of photosynthetic capacity and the implications
-for carbon cycling}. \emph{Proceedings of the National Academy of
-Sciences} \textbf{109}, 8612--8617 (2012).}
-
-\bibitem[\citeproctext]{ref-andres2012genetic}
-\CSLLeftMargin{2. }%
-\CSLRightInline{F. Andrés, G. Coupland, The genetic basis of flowering
-responses to seasonal cues. \emph{Nature Reviews Genetics} \textbf{13},
-627--639 (2012).}
-
-\bibitem[\citeproctext]{ref-korner2010phenology}
-\CSLLeftMargin{3. }%
-\CSLRightInline{C. Körner, D. Basler, Phenology under global warming.
-\emph{Science} \textbf{327}, 1461--1462 (2010).}
-
-\bibitem[\citeproctext]{ref-maple2024}
-\CSLLeftMargin{4. }%
-\CSLRightInline{R. Maple, P. Zhu, J. Hepworth, J.-W. Wang, C. Dean,
-\href{https://doi.org/10.1093/plphys/kiae109}{Flowering time: From
-physiology, through genetics to mechanism}. \emph{Plant Physiology}
-\textbf{195}, 190--212 (2024).}
-
-\bibitem[\citeproctext]{ref-botero2015evolutionary}
-\CSLLeftMargin{5. }%
-\CSLRightInline{C. A. Botero, F. J. Weissing, J. Wright, D. R.
-Rubenstein, Evolutionary tipping points in the capacity to adapt to
-environmental change. \emph{Proceedings of the National Academy of
-Sciences} \textbf{112}, 184--189 (2015).}
-
-\bibitem[\citeproctext]{ref-blackman2013interacting}
-\CSLLeftMargin{6. }%
-\CSLRightInline{B. K. Blackman, Interacting duplications, fluctuating
-selection, and convergence: The complex dynamics of flowering time
-evolution during sunflower domestication. \emph{Journal of experimental
-botany} \textbf{64}, 421--431 (2013).}
-
-\bibitem[\citeproctext]{ref-henry2014transitions}
-\CSLLeftMargin{7. }%
-\CSLRightInline{L. P. Henry, R. H. Watson, B. K. Blackman, Transitions
-in photoperiodic flowering are common and involve few loci in wild
-sunflowers (helianthus; asteraceae). \emph{American Journal of Botany}
-\textbf{101}, 1748--1758 (2014).}
-
-\bibitem[\citeproctext]{ref-aagren2017adaptive}
-\CSLLeftMargin{8. }%
-\CSLRightInline{J. Ågren, C. G. Oakley, S. Lundemo, D. W. Schemske,
-Adaptive divergence in flowering time among natural populations of
-arabidopsis thaliana: Estimates of selection and QTL mapping.
-\emph{Evolution} \textbf{71}, 550--564 (2017).}
-
-\bibitem[\citeproctext]{ref-brachi2010linkage}
-\CSLLeftMargin{9. }%
-\CSLRightInline{B. Brachi, \emph{et al.}, Linkage and association
-mapping of arabidopsis thaliana flowering time in nature. \emph{PLoS
-genetics} \textbf{6}, e1000940 (2010).}
-
-\bibitem[\citeproctext]{ref-dittmar2014flowering}
-\CSLLeftMargin{10. }%
-\CSLRightInline{E. L. Dittmar, C. G. Oakley, J. Ågren, D. W. Schemske,
-Flowering time QTL in natural populations of arabidopsis thaliana and
-implications for their adaptive value. \emph{Molecular ecology}
-\textbf{23}, 4291--4303 (2014).}
-
-\bibitem[\citeproctext]{ref-li2018genomic}
-\CSLLeftMargin{11. }%
-\CSLRightInline{X. Li, T. Guo, Q. Mu, X. Li, J. Yu, Genomic and
-environmental determinants and their interplay underlying phenotypic
-plasticity. \emph{Proceedings of the National Academy of Sciences}
-\textbf{115}, 6679--6684 (2018).}
-
-\bibitem[\citeproctext]{ref-romero2017study}
-\CSLLeftMargin{12. }%
-\CSLRightInline{J. A. Romero Navarro, \emph{et al.}, A study of allelic
-diversity underlying flowering-time adaptation in maize landraces.
-\emph{Nature genetics} \textbf{49}, 476--480 (2017).}
-
-\bibitem[\citeproctext]{ref-bustos-korts2019}
-\CSLLeftMargin{13. }%
-\CSLRightInline{D. Bustos-Korts, \emph{et al.},
-\href{https://doi.org/10.1111/tpj.14414}{Exome sequences and
-multi{-}environment field trials elucidate the genetic basis of
-adaptation in barley}. \emph{The Plant Journal} \textbf{99}, 1172--1191
-(2019).}
-
-\bibitem[\citeproctext]{ref-unterseer2016}
-\CSLLeftMargin{14. }%
-\CSLRightInline{S. Unterseer, \emph{et al.},
-\href{https://doi.org/10.1186/s13059-016-1009-x}{A comprehensive study
-of the genomic differentiation between temperate Dent and Flint maize}.
-\emph{Genome Biology} \textbf{17} (2016).}
-
-\bibitem[\citeproctext]{ref-wadgymar2017identifying}
-\CSLLeftMargin{15. }%
-\CSLRightInline{S. M. Wadgymar, \emph{et al.}, Identifying targets and
-agents of selection: Innovative methods to evaluate the processes that
-contribute to local adaptation. \emph{Methods in Ecology and Evolution}
-\textbf{8}, 738--749 (2017).}
-
-\bibitem[\citeproctext]{ref-blumel2015flowering}
-\CSLLeftMargin{16. }%
-\CSLRightInline{M. Blümel, N. Dally, C. Jung, Flowering time regulation
-in crops---what did we learn from arabidopsis? \emph{Current opinion in
-biotechnology} \textbf{32}, 121--129 (2015).}
-
-\bibitem[\citeproctext]{ref-fournier-level2016}
-\CSLLeftMargin{17. }%
-\CSLRightInline{A. Fournier-Level, \emph{et al.},
-\href{https://doi.org/10.1073/pnas.1517456113}{Predicting the
-evolutionary dynamics of seasonal adaptation to novel climates in
-{\emph{Arabidopsis thaliana}}}. \emph{Proceedings of the National
-Academy of Sciences} \textbf{113} (2016).}
-
-\bibitem[\citeproctext]{ref-jung2009flowering}
-\CSLLeftMargin{18. }%
-\CSLRightInline{C. Jung, A. E. Müller, Flowering time control and
-applications in plant breeding. \emph{Trends in plant science}
-\textbf{14}, 563--573 (2009).}
-
-\bibitem[\citeproctext]{ref-turner2005pseudo}
-\CSLLeftMargin{19. }%
-\CSLRightInline{A. Turner, J. Beales, S. Faure, R. P. Dunford, D. A.
-Laurie, The pseudo-response regulator ppd-H1 provides adaptation to
-photoperiod in barley. \emph{Science} \textbf{310}, 1031--1034 (2005).}
-
-\bibitem[\citeproctext]{ref-faure2012mutation}
-\CSLLeftMargin{20. }%
-\CSLRightInline{S. Faure, \emph{et al.}, Mutation at the circadian clock
-gene EARLY MATURITY 8 adapts domesticated barley (hordeum vulgare) to
-short growing seasons. \emph{Proceedings of the National Academy of
-Sciences} \textbf{109}, 8328--8333 (2012).}
-
-\bibitem[\citeproctext]{ref-hung2012zmcct}
-\CSLLeftMargin{21. }%
-\CSLRightInline{H.-Y. Hung, \emph{et al.}, ZmCCT and the genetic basis
-of day-length adaptation underlying the postdomestication spread of
-maize. \emph{Proceedings of the National Academy of Sciences}
-\textbf{109}, E1913--E1921 (2012).}
-
-\bibitem[\citeproctext]{ref-zakhrabekova2012induced}
-\CSLLeftMargin{22. }%
-\CSLRightInline{S. Zakhrabekova, \emph{et al.}, Induced mutations in
-circadian clock regulator mat-a facilitated short-season adaptation and
-range extension in cultivated barley. \emph{Proceedings of the National
-Academy of Sciences} \textbf{109}, 4326--4331 (2012).}
-
-\bibitem[\citeproctext]{ref-yang2013oself3}
-\CSLLeftMargin{23. }%
-\CSLRightInline{Y. Yang, Q. Peng, G.-X. Chen, X.-H. Li, C.-Y. Wu, OsELF3
-is involved in circadian clock regulation for promoting flowering under
-long-day conditions in rice. \emph{Molecular Plant} \textbf{6}, 202--215
-(2013).}
-
-\bibitem[\citeproctext]{ref-pin2012multifaceted}
-\CSLLeftMargin{24. }%
-\CSLRightInline{P. Pin, O. Nilsson, The multifaceted roles of FLOWERING
-LOCUS t in plant development. \emph{Plant, cell \& environment}
-\textbf{35}, 1742--1755 (2012).}
-
-\bibitem[\citeproctext]{ref-weller2019parallel}
-\CSLLeftMargin{25. }%
-\CSLRightInline{J. L. Weller, \emph{et al.}, Parallel origins of
-photoperiod adaptation following dual domestications of common bean.
-\emph{Journal of Experimental Botany} \textbf{70}, 1209--1219 (2019).}
-
-\bibitem[\citeproctext]{ref-Weine2023.06.21.545998}
-\CSLLeftMargin{26. }%
-\CSLRightInline{E. Weine, S. P. Smith, R. K. Knowlton, A. Harpak,
-Tradeoffs in modeling context dependency in complex trait genetics.
-\emph{bioRxiv} (2024)
-https:/doi.org/\href{https://doi.org/10.1101/2023.06.21.545998}{10.1101/2023.06.21.545998}.}
-
-\bibitem[\citeproctext]{ref-levene1953genetic}
-\CSLLeftMargin{27. }%
-\CSLRightInline{H. Levene, Genetic equilibrium when more than one
-ecological niche is available. \emph{The American Naturalist}
-\textbf{87}, 331--333 (1953).}
-
-\bibitem[\citeproctext]{ref-felsenstein1976theoretical}
-\CSLLeftMargin{28. }%
-\CSLRightInline{J. Felsenstein, The theoretical population genetics of
-variable selection and migration. \emph{Annual review of genetics}
-\textbf{10}, 253--280 (1976).}
-
-\bibitem[\citeproctext]{ref-kawecki2004conceptual}
-\CSLLeftMargin{29. }%
-\CSLRightInline{T. J. Kawecki, D. Ebert, Conceptual issues in local
-adaptation. \emph{Ecology letters} \textbf{7}, 1225--1241 (2004).}
-
-\bibitem[\citeproctext]{ref-hedrick1986genetic}
-\CSLLeftMargin{30. }%
-\CSLRightInline{P. W. Hedrick, Genetic polymorphism in heterogeneous
-environments: A decade later. \emph{Annual review of ecology and
-systematics} \textbf{17}, 535--566 (1986).}
-
-\bibitem[\citeproctext]{ref-des2013genotype}
-\CSLLeftMargin{31. }%
-\CSLRightInline{D. L. Des Marais, K. M. Hernandez, T. E. Juenger,
-Genotype-by-environment interaction and plasticity: Exploring genomic
-responses of plants to the abiotic environment. \emph{Annual Review of
-Ecology, Evolution, and Systematics} \textbf{44}, 5--29 (2013).}
-
-\bibitem[\citeproctext]{ref-millet2016}
-\CSLLeftMargin{32. }%
-\CSLRightInline{E. Millet, \emph{et al.},
-\href{https://doi.org/10.1104/pp.16.00621}{Genome-wide analysis of yield
-in Europe: allelic effects as functions of drought and heat scenarios}.
-\emph{Plant Physiology}, pp.00621.2016 (2016).}
-
-\bibitem[\citeproctext]{ref-anderson2013genetic}
-\CSLLeftMargin{33. }%
-\CSLRightInline{J. T. Anderson, C.-R. Lee, C. A. Rushworth, R. I.
-Colautti, T. Mitchell-Olds, Genetic trade-offs and conditional
-neutrality contribute to local adaptation. \emph{Molecular ecology}
-\textbf{22}, 699--708 (2013).}
-
-\bibitem[\citeproctext]{ref-anderson2011evolutionary}
-\CSLLeftMargin{34. }%
-\CSLRightInline{J. T. Anderson, J. H. Willis, T. Mitchell-Olds,
-Evolutionary genetics of plant adaptation. \emph{Trends in Genetics}
-\textbf{27}, 258--266 (2011).}
-
-\bibitem[\citeproctext]{ref-mitchell1997predicting}
-\CSLLeftMargin{35. }%
-\CSLRightInline{R. B. Mitchell, K. J. Moore, L. E. Moser, J. O. Fritz,
-D. D. Redfearn, Predicting developmental morphology in switchgrass and
-big bluestem. \emph{Agronomy Journal} \textbf{89}, 827--832 (1997).}
-
-\bibitem[\citeproctext]{ref-parrish2005biology}
-\CSLLeftMargin{36. }%
-\CSLRightInline{D. J. Parrish, J. H. Fike, The biology and agronomy of
-switchgrass for biofuels. \emph{BPTS} \textbf{24}, 423--459 (2005).}
-
-\bibitem[\citeproctext]{ref-casler2004latitudinal}
-\CSLLeftMargin{37. }%
-\CSLRightInline{M. Casler, K. P. Vogel, C. Taliaferro, R. Wynia,
-Latitudinal adaptation of switchgrass populations. \emph{Crop Science}
-\textbf{44}, 293--303 (2004).}
-
-\bibitem[\citeproctext]{ref-lovell2021}
-\CSLLeftMargin{38. }%
-\CSLRightInline{J. T. Lovell, \emph{et al.},
-\href{https://doi.org/10.1038/s41586-020-03127-1}{Genomic mechanisms of
-climate adaptation in polyploid bioenergy switchgrass}. \emph{Nature},
-1--7 (2021).}
-
-\bibitem[\citeproctext]{ref-porter1966analysis}
-\CSLLeftMargin{39. }%
-\CSLRightInline{C. L. Porter Jr, An analysis of variation between upland
-and lowland switchgrass, panicum virgatum l., in central oklahoma.
-\emph{Ecology} \textbf{47}, 980--992 (1966).}
-
-\bibitem[\citeproctext]{ref-milano2016genetic}
-\CSLLeftMargin{40. }%
-\CSLRightInline{E. R. Milano, D. B. Lowry, T. E. Juenger, The genetic
-basis of upland/lowland ecotype divergence in switchgrass (panicum
-virgatum). \emph{G3: Genes, Genomes, Genetics} \textbf{6}, 3561--3570
-(2016).}
-
-\bibitem[\citeproctext]{ref-urbut2019flexible}
-\CSLLeftMargin{41. }%
-\CSLRightInline{S. M. Urbut, G. Wang, P. Carbonetto, M. Stephens,
-Flexible statistical methods for estimating and testing effects in
-genomic studies with multiple conditions. \emph{Nature genetics}
-\textbf{51}, 187--195 (2019).}
-
-\bibitem[\citeproctext]{ref-10.1093ux2fbiostatisticsux2fkxw041}
-\CSLLeftMargin{42. }%
-\CSLRightInline{M. Stephens,
-\href{https://doi.org/10.1093/biostatistics/kxw041}{{False discovery
-rates: a new deal}}. \emph{Biostatistics} \textbf{18}, 275--294 (2016).}
-
-\bibitem[\citeproctext]{ref-weine2024}
-\CSLLeftMargin{43. }%
-\CSLRightInline{E. Weine, S. P. Smith, R. K. Knowlton, A. Harpak,
-\href{http://dx.doi.org/10.7554/eLife.99210.1}{Tradeoffs in modeling
-context dependency in complex trait genetics} (2024).}
-
-\bibitem[\citeproctext]{ref-choi2023}
-\CSLLeftMargin{44. }%
-\CSLRightInline{S. Choi, \emph{et al.},
-\href{https://doi.org/10.1093/jxb/erad255}{A single amino acid change
-led to structural and functional differentiation of {\emph{PvHd1}} to
-control flowering in switchgrass}. \emph{Journal of Experimental Botany}
-\textbf{74}, 5532--5546 (2023).}
-
-\bibitem[\citeproctext]{ref-savolainen2013ecological}
-\CSLLeftMargin{45. }%
-\CSLRightInline{O. Savolainen, M. Lascoux, J. Merilä, Ecological
-genomics of local adaptation. \emph{Nature Reviews Genetics}
-\textbf{14}, 807--820 (2013).}
-
-\bibitem[\citeproctext]{ref-lowry2019qtl}
-\CSLLeftMargin{46. }%
-\CSLRightInline{D. B. Lowry, \emph{et al.}, QTL\(\times\) environment
-interactions underlie adaptive divergence in switchgrass across a large
-latitudinal gradient. \emph{Proceedings of the National Academy of
-Sciences} \textbf{116}, 12933--12941 (2019).}
-
-\bibitem[\citeproctext]{ref-kortefarlow2013}
-\CSLLeftMargin{47. }%
-\CSLRightInline{A. Korte, A. Farlow, The advantages and limitations of
-trait analysis with GWAS: A review. \emph{Plant methods} \textbf{9}, 29
-(2013).}
-
-\bibitem[\citeproctext]{ref-privuxe92017}
-\CSLLeftMargin{48. }%
-\CSLRightInline{F. Privé, H. Aschard, M. G. B. Blum,
-\href{http://dx.doi.org/10.1101/190926}{Efficient management and
-analysis of large-scale genome-wide data with two r packages: Bigstatsr
-and bigsnpr} (2017).}
-
-\end{CSLReferences}
-
-\acknow{We thank the Brackenridge Field laboratory, the Ladybird Johnson
-Wildflower Center, and the Juenger laboratory for support with plant
-care and propagation. This material is based upon work supported in part
-by the Great Lakes Bioenergy Research Center, U.S. Department of Energy,
-Office of Science, Office of Biological and Environmental Research under
-Award Numbers DE-SC0018409 and DE-FC02-07ER64494, the US Department of
-Energy Awards DE-SC0021126 and DESC0014156 to T.E.J., DE-SC0017883 to
-D.B.L, NIH grant R35GM151108 and a Pew Scholarship to A.H., National
-Science Foundation PGRP Awards IOS0922457 and IOS1444533 to T.E.J, and
-the Long-term Ecological Research Program (DEB 1832042) at the Kellogg
-Biological Station.}
-
-\showacknow{} % Display the acknowledgments section
+\renewcommand{\bibsection}{}
+\bibliography{bibliography.bib}
 

### 2024-08-18 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
revision2: Edit author middle initial in SI Appendix

@@
       equal-contributor: true
-  - name: Jason Bonnette
+  - name: Jason E. Bonnette
     affiliations:

### 2024-08-18 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
revision2: Edit author middle name in SI Appendix and main manuscript

@@
       equal-contributor: true
-  - name: Samuel P. Smith
+  - name: Samuel Pattillo Smith
     affiliations:

### 2024-08-19 README.Rmd
Add DOI link in README to genomic data
revision2: Modify Readme to point to most recent code and analyses for this project

@@
 
-pvdiv-phenology-gxe contains the non-genomic data and code used to analyze the genetics of GxE in the phenology data collected for the switchgrass diversity panel in 2019, including green-up dates and flowering dates in 2019.
-This repository will be private until the associated manuscript is published.
+pvdiv-phenology-gxe contains the non-genomic data and code used to analyze the genetics of GxE in the phenology data collected for the switchgrass diversity panel in 2019, including green-up dates (the timing of vegetative growth) and flowering dates (the timing of reproductive growth) in 2019.
+
+The **most recent submission** is on bioRxiv: https://doi.org/10.1101/2021.08.19.456975
+
+**Code** used in the 2024 submission is located here: https://github.com/Alice-MacQueen/pvdiv-phenology-gxe/blob/main/R/Analysis_v1.2_mash_using_greedy_covar.qmd.
+
+You can find the **genomic data** at: https://doi.org/10.18738/T8/A604BU.
 
 In general, if you have any questions about how to locate an analysis folder or code used for an analysis, look at this README first.
-Then, if you still have questions, feel free to open a pull request with your issue and I will be happy to help.
+Then, if you still have questions, feel free to open a pull request with your issue or contact me directly at alice.macqueen[at]gmail.com and I will be happy to help.
 
@@
 
-    - Data objects and draft figures created as part of the analysis of heritability (including variance components analysis, done in ASReml), genome-wide association (on phenotypic BLUPs, done using the switchgrassGWAS R package), and mash (on results from the univariate GWAS, done using the switchgrassGWAS R package).
+Data objects and draft figures created as part of different discrete analyses, including:
+
+    - heritability/ contains the analysis of heritability (including variance components analysis, done in ASReml);
+    - gwas/BLUPs/  contains genome-wide association results (on phenotypic BLUPs, done using the switchgrassGWAS R package);
+    - hypothesis_matrices/ contains unpublished analyses of correlation data between GxWeather matrices, used as exploratory plots for a revision;
+    - mash/ contains mash models (on results from the univariate GWAS, done using the switchgrassGWAS R package). 
+    - mash_greedy_algorithm/mash_h2_diag_covar/ contains runs of the greedy mash algorithm used to select covariance matrices for inclusion in the mash model for the current revision.
+mash_greedy_algorithm/ also contains other runs of the greedy mash algorithm used as exploratory analyses for suitability for the revision.
+
 Generally, the code to replicate these analyses can be found in the R/ directory.
@@
 
-    - Contains the majority of the code used in the manuscript (some additional code used to prepare data is in data-raw/.) 
-    - Files that start with "Plots" were used to make manuscript figures
-    - Files that start with "nohup" were used to run mash
+R/ contains the majority of the code used in the manuscript (some additional code used to prepare data is in data-raw/.)
+
+    - The file titled "Analysis_v1.2_mash_using_greedy_covar.qmd" contains all of the code used to run the analyses and generate the figures for the 2024 revision (in manuscript/revision2/). 
+    - THe file titled "SI_Appendix_Datasets.Rmd" contains the code used to generate the SI Appendix datasets and the SI Appendix figures for the 2024 revision.
+    - The file titled "UT_Dataverse_files.Rmd" was used to generate the SNP datasets uploaded to the UT Dataverse for replication of this analysis. 
+    - Files that start with "Plots" were used to make manuscript figures for a 2021 submission
+    - Files that start with "nohup" were used to run mash for a 2021 submission
     - Files that start with "Analysis" were used to generate files in the analysis/ folder.
-I tried to name these descriptively but there are a lot of them; please open a pull request if you want help locating specific code.
+I tried to name these descriptively but there are a lot of them; please open a pull request if you want help locating specific code from earlier versions of this analysis.
+The most recent version is in the file titled "Analysis_v1.2_mash_using_greedy_covar.qmd".
 
@@
 
-    - Contains completed drafts of a manuscript on the genetics of GxE in phenology in the switchgrass diversity panel.
+This folder contains completed drafts of a manuscript on the genetics of GxE in phenology in the switchgrass diversity panel.
+
+    - revision2/ contains the most recent, 2024 revision.
+The main text is called revision2.pdf.
+There are also folders containing pdfs of the response to reviews (in response_to_reviews/), the SI Appendix data (in SI_Appendix) and images used in the main manuscript (in images/).
+    - Coauthors/ contains the manuscript from a submission in 2021.

### 2024-10-16 manuscript/revision2/cover_letter/cover_letter.qmd
revision2: Edit cover letter

@@
 ---
-author:
-  - name: Alice MacQueen* and Tom Juenger
-    affiliation:
-      - name: The University of Texas at Austin
-        department: Department of Integrative Biology
-        address: 2415 Speedway
-        city: Austin, TX
-        postal-code: 78712
-address:
-  - Editors
-  - Proceedings of the National Academy of Sciences
-subject: manuscript revision
-subject-title: "Subject"
-opening: "Dear Editors,"
-closing: "Best,"
-format: letter-pdf
-bibliography: references.bib
+format:
+  pdf:
+    keep-tex: true
+    pdfengine: xelatex
+    geometry: 
+      - "top=1in"
+      - "bottom=1in"
+      - "left=1in"
+      - "right=1in"
+    code-annotations: hover
 ---
 
-Thank you for considering this work for publication in *PNAS*.
-Our manuscript, "Diverse Genotype-by-Weather Interactions in Switchgrass" has undergone substantial revision from our original submission in 2021.
-This timeframe is long - Dr.
-MacQueen took a new position in 2022.
-They continued working with the group as we developed a new algorithm to address major reviewer concerns and significantly altered the manuscript.
-Because of these alterations, the response to reviewers does not have track changes for each point, though we do point to line numbers for the associated revisions in our response.
+::: {.flushright data-latex="1" linestretch="1"}
+The University of Texas at Austin\
+Department of Integrative Biology\
+2415 Speedway\
+Austin, TX 78712
+:::
 
-Our manuscript is suitable for publication in *PNAS* for the following reasons.
+::: {.flushright data-latex="1" linestretch="1"}
+October 16th, 2024
+:::
+
+Dear Editors,
+
+Thank you for considering this work for publication in *eLife*.
 
@@
 
-Second, to our knowledge, this work is the first to use a statistically unbiased test to detect the relative frequencies of two important types of GxE; antagonistic pleiotropy, or a change in sign of effects between environments, and differential sensitivity, or a change in magnitude of effects between environments.
-Here, rather than using the local false discovery rate to ask "Are these two effects different?" - as we reasonably expect two effects to be, even if this difference cannot be measured - we use the local false sign rate to ask a more meaningful question: Can we be confident in the sign of these two effects?
-We use the lfsr to find thousands of genetic effects with antagonistic pleiotropy for the start of vegetative growth for switchgrass.
+Second, to our knowledge, this work is the first to use a statistically unbiased test to detect the relative frequencies of loci with and without sign-changing GxE.
+Here, rather than using a standard statistical test such as the false discovery rate to ask "Are these two effects different?" - as we reasonably expect two effects to be, even if this difference cannot be measured - we use the local false sign rate to ask a more meaningful question: Can we be confident in the sign of these two effects?
+In contrast with previous empirical work which finds few genetic effects that differ in sign, we use the local false sign rate to find thousands of genetic effects with antagonistic pleiotropy for the start of vegetative growth for switchgrass.
 
-Finally, with current advances in high-throughput measurement, and the increased resources devoted to experiments in natural and semi-natural environments, we must consider how to share information across these experiments and how to associate natural variation in genomes with natural
-variation in complex field environments.
+Finally, with current advances in high-throughput measurement, and the increased resources devoted to experiments in natural and semi-natural environments, we must consider how to share information across these experiments and how to associate natural variation in genomes with natural variation in complex field environments.
 Advances in methods to find the environmental cues that matter should allow researchers to determine the adaptations important to responses to the natural environment, and how the importance of these genetic mechanisms changes as the environment itself changes.
-Our manuscript should be of broad interest to PNAS readership for the insights it provides into these important issues.
+Our manuscript should be of broad interest to *eLife* readership for the insights it provides into these important issues.
 
-We suggest that the original editor, Editorial Board members, and anonymous reviewers for this manuscript would be good choices to assess this revision.
-In the event that is not practicable, we include some additional suggestions below.
+::: {.center data-latex="1"}
+Best,
+:::
 
-We suggest that Douglas Schemske, \_\_\_, and Qifa Zhang would be good Editorial Board members for this manuscript.
-All would be able to provide strong insights into mapping the genetic basis of adaptive and ecologically relevant traits in a field experimental context.
-
-We suggest that Johanna Schmitt, Trudy Mackay, and Detlef Weigel would be excellent editors for this manuscript.
-Dr.
-Schmitt has done extensive work in *Arabidopsis thaliana* and other plant species unravelling the genetic mechanisms involved in responses to the natural environment.
-Dr.
-Mackay has extensive experience mapping the genetics of complex traits and genotype-by-environment interactions.
-Dr.
-Weigel has done extensive work understanding genetic diversity in *A.
-thaliana* including extensive work on floral induction cues.
+::: {.flushright data-latex="1"}
+Alice MacQueen and Tom Juenger
+:::

### 2024-10-16 manuscript/revision2/cover_letter/cover_letter.tex
revision2: Edit cover letter

--- 2023-12-02 manuscript/revision2/SI_Appendix.qmd
@@
----
-title: "SI Appendix"
-format: pdf
-editor: visual
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+%
+\documentclass[
+  letterpaper,
+  DIV=11,
+  numbers=noendperiod]{scrartcl}
 
-**SI Appendix**
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage[top=1in,bottom=1in,left=1in,right=1in]{geometry}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-\
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 
-**Dataset 1.** SNP-associated effects and standard errors for green-up estimated using the Gulf genetic subpopulation.
+\KOMAoption{captions}{tableheading}
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\AtBeginDocument{%
+\ifdefined\contentsname
+  \renewcommand*\contentsname{Table of contents}
+\else
+  \newcommand\contentsname{Table of contents}
+\fi
+\ifdefined\listfigurename
+  \renewcommand*\listfigurename{List of Figures}
+\else
+  \newcommand\listfigurename{List of Figures}
+\fi
+\ifdefined\listtablename
+  \renewcommand*\listtablename{List of Tables}
+\else
+  \newcommand\listtablename{List of Tables}
+\fi
+\ifdefined\figurename
+  \renewcommand*\figurename{Figure}
+\else
+  \newcommand\figurename{Figure}
+\fi
+\ifdefined\tablename
+  \renewcommand*\tablename{Table}
+\else
+  \newcommand\tablename{Table}
+\fi
+}
+\@ifpackageloaded{float}{}{\usepackage{float}}
+\floatstyle{ruled}
+\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
+\floatname{codelisting}{Listing}
+\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\makeatother
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\usepackage{bookmark}
 
-**Dataset 2.** SNP-associated effects and standard errors for green-up estimated using the Midwest genetic subpopulation.
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  colorlinks=true,
+  linkcolor={blue},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={Blue},
+  pdfcreator={LaTeX via pandoc}}
 
-**Dataset 3.** SNP-associated effects and standard errors for green-up estimated using Both genetic subpopulations.
+\author{}
+\date{}
 
-**Dataset 4.** SNP-associated effects and standard errors for flowering estimated using the Gulf genetic subpopulation.
+\begin{document}
 
-**Dataset 5.** SNP-associated effects and standard errors for flowering estimated using the Midwest genetic subpopulation.
+\begin{flushright}
+The University of Texas at Austin\\
+Department of Integrative Biology\\
+2415 Speedway\\
+Austin, TX 78712
 
-**Dataset 6.** SNP-associated effects and standard errors for flowering estimated using Both genetic subpopulations.
+\end{flushright}
 
-**Dataset 7.** Genes within QTL that have functionally validated roles in flowering in rice.
+~
 
-**Dataset 8.** Overlap between quantitative trait loci for flowering measured using an outbred pseudo-F2 mapping population, and significant mash effect estimates (with a log10Bayes Factor \> 2, or in the 1% tail of most significant mash results) found using three overlapping diversity panels.
-Genes within 20kb with a functionally validated role in flowering in rice are  indicated.
+\begin{flushright}
+October 16th, 2024
 
-\
+\end{flushright}
 
-**Figure S1.** Variance components analysis of genetic (purple), genotype by environment (blue), environmental (green), and residual (yellow) terms in models of green-up time as functions of Julian date or weather, for the four northern common gardens (North), the three Texas common gardens (Texas), and for all eight common gardens (All).
-Dashed lines indicate the cumulative contribution of G and GxE for flowering as a function of day of the year. ![](https://lh7-us.googleusercontent.com/SPDH2Y9HThoK0pUZQxFuLPUDrIuXAcG73nUSk2ucYN3_CHXSG-c8G4b7Gq6crgo_1IapiGYFV66w_vhfyvxSIW8kPtTQA_tMeviTFamFdsXIPjIjTcaI0dwa7ZqpKN0P1RonX05vbxitm7vLYoYoZw)
+~
 
-**Figure S2.** Effect patterns exhibited by the major data-driven matrices for (a,d,e,j,m,n) Gulf, (b,f,g,k,o,p) Midwest, and (c,h,i,l,q,r) Both subpopulations from Figure 2.
-(a-i) Green-up date.
-(j-r) Flowering date.
-(a-c,j-l) Single-garden effect representations (eigenvectors) of the DD_tPCA data-derived matrices.
-The percent variation explained of DD_tPCA by each eigenvector is shown on the y-axis.
-Common gardens are arranged in latitudinal order along the x-axis.
-In addition, the first eigenvector corresponds to DD_PCA_1, and second corresponds to DD_PCA_2.
-(d,f,h,m,o,q) The number of SNPs with significant effects in both conditions that exhibit antagonistic pleiotropy between that pair of conditions.
-(e,g,i,n,p,r) Same as (d,f,h,m,o,q) except for differential sensitivity.
-Scales for (d,e), (f,g), (h,i), (m,n), (o,p), and (q,r) are identical and located on the first panel in alphabetical order.![](https://lh7-us.googleusercontent.com/JV_8jSfymcCRhEQsEuXD2WzHAkM7bOTpEws9j3_H4SbmaTArViZOZQyWld6aMI-xs8DPYLSZBqpxogUokZwU36l5NMMlJgLy-dY8GFwACkYAntFMqAvq9b1C3BsXnbacDYOF3i4kRqXIW4ISTu2Xzg)
+Dear Editors,
 
-\
+Thank you for considering this work for publication in \emph{eLife}.
 
-\
-![](https://lh7-us.googleusercontent.com/h-CoStrONeN2vAZcGvcrGRLg1RRQiAhfNYCT9xn9YF5uwInX02rCepO4NU1bRq1ykboFZcOjr1PeEiZXsXRDJ6v51BAzYPWKQ2wd8YhVg2LJA7IMkL5wvEEa8jOw0UDgrZgVwiYZlHQWSqGxT_2q0g)
+Our work is a key methodological development in understanding genetic
+variation in genotype-by-environment interactions, or GxE. GxE
+interactions must vary across the genome; that is, different genes and
+genetic variants can exhibit different responses to the environment. Yet
+current statistical models to test for GxE don't have robust frameworks
+to specify and compete different models of GxE effects, instead focusing
+on finding a single, best fitting form of GxE covariance. We use
+multivariate adaptive shrinkage (mash {[}@urbut2018{]}) to determine the
+regions of the switchgrass genome with different types of GxE,
+specifically regions of the genome with effects on phenology that covary
+with several weather-based `cues'. We use a greedy algorithm to compete
+and select types of GxE and GxWeather that improve our models, then use
+mash to identify which SNP effects are best modeled with each type of
+GxE or GxWeather.
 
-**Figure S3.** Comparison of mash SNP effects from a diversity panel and effects from an outbred pseudo-F2 cross.
-(**a**) QTL mapping for four weather-related functions of flowering, and for green-up date, as indicated by the colors: pink, growing degree days (GDD) from green-up date to flowering date, yellow, day length change on the flowering date, green, day length on flowering date; purple, flowering date; gray, green-up date.Dotted lines indicate permutation-based significance thresholds for each weather-related function.
-Stars indicate QTL with significant enrichment for SNPs in the 1% mash tail; B, G, and M indicate which subpopulation had enrichment: B -  both subpopulations, G - Gulf subpopulation, M -Midwest subpopulation.
-Rug plots show genomic locations of SNPs in the 1% mash tail for flowering date for each subpopulation.
-(**b**) Schematic comparison of genotypes used for QTL mapping and genotypes used in mash.
-(**c**) Number of mash runs enriched for SNPs in the 1% mash tail in the 23 QTL intervals from (**a**) (dotted red line), compared to 1000 sets of 23 random QTL intervals of the same size (histogram).**(d**) SNP effects estimated using mash for SNPs with the highest Bayes factor for the three QTL with the highest LOD scores from (**a**) and enrichment of SNPs in the 1% mash tail.
-SNP genomic position and genetic subpopulation are indicated above each plot.
-Common gardens are arranged in latitudinal order.
-(**e**) QTL effect estimates for 23 non-overlapping and partially overlapping QTLs.
-Colors indicate which of the four weather-related functions of flowering for which the QTL was mapped.
-An additional five QTLs completely overlapped the QTL displayed; in this case, the QTL with the highest lod score is shown.
-Solid vertical lines separate QTLs that are not overlapping.
-Stars indicate QTL with significant enrichment for SNPs in the 1% mash tail; B, G, and M indicate which subpopulation had enrichment, as in (**b**):B -  both subpopulations, G - Gulf subpopulation, M -Midwest subpopulation.
-Common gardens are arranged in latitudinal order.
+Second, to our knowledge, this work is the first to use a statistically
+unbiased test to detect the relative frequencies of loci with and
+without sign-changing GxE. Here, rather than using a standard
+statistical test such as the false discovery rate to ask ``Are these two
+effects different?'' - as we reasonably expect two effects to be, even
+if this difference cannot be measured - we use the local false sign rate
+to ask a more meaningful question: Can we be confident in the sign of
+these two effects? In contrast with previous empirical work which finds
+few genetic effects that differ in sign, we use the local false sign
+rate to find thousands of genetic effects with antagonistic pleiotropy
+for the start of vegetative growth for switchgrass.
 
-**Table S1.** Green-up and flowering dates phenotypes linked to environmental cues.
+Finally, with current advances in high-throughput measurement, and the
+increased resources devoted to experiments in natural and semi-natural
+environments, we must consider how to share information across these
+experiments and how to associate natural variation in genomes with
+natural variation in complex field environments. Advances in methods to
+find the environmental cues that matter should allow researchers to
+determine the adaptations important to responses to the natural
+environment, and how the importance of these genetic mechanisms changes
+as the environment itself changes. Our manuscript should be of broad
+interest to \emph{eLife} readership for the insights it provides into
+these important issues.
 
-+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
-| Phenotype | Name                              | Description                                                                                                                          |
-+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
-| green-up  | GDD, **n**d                       | Correlations in cumulative growing degree days for the five, ten, or eighteen days prior to green-up.                                |
-|           |                                   |                                                                                                                                      |
-|           | where \'**n**\' is in {5, 10, 18} |                                                                                                                                      |
-+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
-| green-up  | Temp Ave, **n**d                  | Correlations in average temperature for the five, ten, or eighteen days prior to green-up.                                           |
-|           |                                   |                                                                                                                                      |
-|           | where \'**n**\' is in {5, 10, 18} |                                                                                                                                      |
-+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
-| flowering | 50% green-up                      | Correlations in green-up.                                                                                                            |
-+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
-| flowering | daylength                         | Correlations in day length on the day of flowering.                                                                                  |
-+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
-| flowering | daylength change (s)              | Correlations in the change in daylength (in seconds) on the day of flowering.                                                        |
-+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
-| flowering | GDD, GR to FL                     | Correlations in cumulative GDD between green-up and flowering.                                                                       |
-+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
-| flowering | Rainfall, **n**d sum              | Correlations in cumulative rainfall in the one, three, or five days before flowering, or in the days between green-up and flowering. |
-|           |                                   |                                                                                                                                      |
-|           | where \'**n**\' is in {1, 3, 5}   |                                                                                                                                      |
-+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+
+~
 
-\
-\
+~
 
-**Section S1.** Weather-based environmental cues for green-up and flowering date.
+\begin{center}
+Best,
 
-\
+\end{center}
 
-	We scored green-up and flowering by day of the year, then linked these dates to multiple environmental factors (Table S1).
-We defined flowering as functions of eight variables: \'flowering date\' (day of the year), cumulative growing degree days (GDD) between green-up and flowering dates (\'flowering GDD\'), day length on the flowering date (\'flowering daylength\'), change in daylength relative to the previous day on the day of flowering (\'flowering daylength change\'), and four measures of cumulative rainfall: cumulative rainfall between green-up and flowering, and in the one day, three days, and five days before flowering.
-Cumulative GDD was calculated as GDD = GRFLmaxTmean−Tbase,0, where Tmean is the daily average temperature, defined as (Tmax+Tmin)/2, Tbase is the base temperature of 12 °C for switchgrass, Tmax is the maximum daily temperature, Tmin is the minimum daily temperature, GR is green-up date, and FL is flowering date (Kiniry et al.
-2005; Behrman et al.
-2013).  On a specific day, if Tmean is less than Tbase, the GDD for that day is 0; if Tmean is bigger than Tbase, the GDD for that day is the difference between Tmean and Tbase.
-We also defined green-up as functions of seven variables: \'green-up date\' (day of the year), cumulative GDD for the five, ten, or eighteen days prior to green-up, and the average air temperature for the five, ten, or eighteen days prior to green-up.
+\begin{flushright}
+Alice MacQueen and Tom Juenger
 
-**Section S2.** Hypothesis-based covariance matrix definitions for mash.
+\end{flushright}
 
-	We generated hypothesis-based covariance matrices derived from correlations in environmental cues in the green-up or flowering date windows for the three subpopulations (Table S1).
-These covariance matrices were derived from the same set of phenotypes as environmental functions of green-up and flowering, and represent the correlations between genotypes for these phenotypes across our common gardens.
-For the diagonal of these matrices, we used the coefficient of variation in these phenotypes within the subpopulation at that garden.
-Code used to generate these matrices can be found at  [git\@github.com](mailto:git@github.com):Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd.
-We used distinct sets of user-specified, hypothesis-based covariance matrices for green-up and flowering, but the same set of matrices for all genetic subgroups.
-If a hypothesis-based matrix does not capture a common pattern of SNP effects, mash assigns small mixture proportions onto this matrix using maximum likelihood, giving that covariance matrix a low posterior weight summed across all SNP effects, which ultimately causes little to no change in the mash model.
-Alternatively, SNPs with high mixture proportions on particular environmental covariance matrices and large Bayes factors, which summarize the overall significance of a non-zero effect, represent small genomic intervals with strong evidence for a phenotypic effect correlated with an environmental driver.
-
-\
-
-**Section S3.** Narrow-sense heritability
-
-	In the diversity panel, we determined narrow-sense heritabilities (h2) for green-up and flowering dates at single gardens using genomic relationship matrices calculated using the van Raden method (VanRaden et al.
-2009).
-Genomic relationship matrices were calculated within each subpopulation (Midwest and Gulf) and for both genetic subpopulations (Both).
-We used ASReml (VSN International) to specify mixed models of the form:
-
-**y** = 1 + Zu + e
-
-Var(u) = Gσu2
-
-Var(e) = Iσe2
-
-in which the vector **y** represents the flowering date or green-up date values for that garden, Z the design matrix for random effects, u the whole genome additive genetic effect, and e the residual.
-Matrix G is the whole genomic relationship matrix based on all SNPs retained for subpopulation-specific analyses.
-I is the rank-y identity matrix.
-Phenotypic variance (σp2) is σu2 + σe2.
-Narrow-sense heritability is then h2 = (σu2/σp2).
-
-	These models were run for each of the eight gardens, and across all gardens by adding an additional environmental effect of site without an interaction term.
-This resulted in 54 models: 3 sets of populations (the Gulf, Midwest, and Both subpopulations) for 9 garden sets (all eight gardens separately, and all eight gardens together) and two phenotypes (green-up date and flowering date).
-
-\
-
-**Section S4.** Variance components analysis
-
-	In the diversity panel, to evaluate our environmental cues as genetic triggers of flowering, we defined green-up and flowering for individual genotypes as functions of seven and eight environmental cues, respectively, then used variance components analysis to partition variance attributed to genetic effects (G), genotype by environment interactions (GxE), environmental effects (E), and error for these phenology-related traits using linear mixed models.
-We used mixed models of the form:
-
-**y** = 1 + Zuu + Zll + Zulul + e
-
-Var(u) = Gσu2
-
-Var(l) = IσE2
-
-Var(ul) = GσuE2
-
-Var(e) = Iσe2
-
-in which the vector **y** represents the individual plant observations of a given trait and Z the design matrices associating trait observations with random factors.
-Random factors include u, the whole genome additive genetic (G) effect, l, the effect of each location (E), ul, the GxE effect, and e, the residual.
-Matrix G is the whole genomic relationship matrix.
-I are rank-y identity matrices.
-
-	These models were run for each phenotype as a function of environmental cues.
-This resulted in 135 models: three sets of populations (the Gulf, Midwest, and Both subpopulations) for three garden sets (All, North, and Texas gardens) and 15 phenotypes (seven green-up functions, and eight flowering functions).
-
-\
-
-**Section S5.** Multivariate adaptive shrinkage
-
-	To evaluate the prevalence and kinds of covariance patterns of SNP effects across our eight common gardens, we used multivariate adaptive shrinkage (mash) on SNP effect estimates from the diversity panel (Urbut et al.
-2019).
-Mash is a statistical method that allows estimation and comparison of many effects jointly across many different conditions; it improves on previous methods by allowing arbitrary correlations in effect sizes among conditions.
-To this end, mash allows the user to include hypothesis-based covariance matrices, specifies some \'canonical\' covariance matrices with simple patterns of effects, and also generates \'data-driven\' covariance matrices from patterns of effects in the data.
-Then, mash assigns mixture proportions for each SNP onto each provided covariance matrix using maximum likelihood.
-Finally, mash uses Bayes\' theorem to shrink effects for each SNP towards the set of covariance matrices in accordance to their mixture proportions.
-For example, a SNP affecting flowering may have a high mixture proportion, or weight, on a covariance matrix created from a specific environmental cue.
-In that case, we can infer that the effect of that SNP on flowering is caused by a response to or interaction with that environmental cue.
-These user-specified and data-driven covariance matrices are an important advantage mash offers for studying patterns of GxE: the user-specified covariance matrices allow hypothesis testing of specific environmental drivers for each SNP, while the data-driven covariance matrices allow exploration of additional unexplained patterns of covariation.
-
-	We first conducted univariate GWAS at each common garden for green-up and flowering date, then analyzed the allelic effects for the top 19K unlinked SNPs per univariate GWAS using mash, as in Lovell et al.
-(2021).
-To allow for the possibility that different subpopulations had different strengths of connection between our phenotypes and genotypes (Korte and Farlow 2013), we conducted three sets of genetic analyses: on Gulf and Midwest genotypes separately, and on both subpopulations together (hereafter \'Both\' subpopulations).
-Only SNPs with ≤ 20% missing data and minor allele frequencies \> 0.05 were used in univariate GWAS, resulting in 8.8 million SNPs retained for the Midwest subpopulation, 10.3 million SNPs retained for the Gulf subpopulation, and 12.3 million SNPs retained for Both subpopulations.
-Then, we created a set of \'random\' markers to account for correlation in effects using a set of 19K relatively unlinked (r2 \< 0.2) markers.
-Next, we created a set of \'strong\' markers by selecting 19K SNPs with the largest -log10(p-value) in any condition in the univariate GWAS, out of all markers unlinked at (r2 \< 0.2), which represented the SNP with the strongest effect in the 19K LD blocks with the largest univariate effects.
-We used the \'pvdiv_standard_run()\' function of the switchgrassGWAS R package (https://github.com/Alice-MacQueen/switchgrassGWAS) to conduct GWAS on genotypes from the Midwest, Gulf, and Both subpopulations, used the \'pvdiv_bigsnp2mashr()\' function to convert univariate GWAS output to the matrices of effects and standard errors needed for mash, then used the \'mash_standard_run()\' function to conduct mash, again specifying distinct sets of hypothesis-based covariance matrices for green-up and flowering, but the same set of matrices for all genetic subgroups.
-
-\
-
-**Section S6.** Outbred pseudo-F2 mapping population and Quantitative Trait Locus mapping
-
-	To confirm candidate genomic regions and patterns of allelic effects found in the diversity panel, we analyzed flowering in an outbred pseudo-F2 cross between four individuals, two Midwest and two Gulf individuals.
-The formation of this mapping population has been described previously (Milano, Lowry, and Juenger 2016).
-The parents of this cross were DAC, an early flowering Midwest individual, VS16, a late flowering Midwest individual, AP13, an early flowering Gulf individual, and WBC, a late flowering Gulf individual.
-We made F1 crosses of the two early flowering genotypes, AP13xDAC, and the two late flowering genotypes, WBCxVS16.
-We then clonally propagated and planted the four parents, the two F1 genotypes (AP13xDAC, and VS16xWBC), and 801 F2 genotypes at eight field sites in May-July of 2015.
-To be directly comparable to the diversity panel data, only 2019 phenology data from the pseudo-F2 cross from the same eight common garden sites were used here.
-
-	Details on the genetic map construction, map polishing and fine-scale reordering can be accessed on <https://datadryad.org/stash/dataset/doi:10.5061/dryad.ghx3ffbjv>.
-QTL mapping was conducted with R/qtl2 (Broman et al.
-2019).
-We performed a genome scan with a linear mixed model that accounts for the relationships among individuals and for environmental covariates (i.e., field sites).
-The full model can be expressed as: phenotype = µ + QTL + E + QTLxE + kinship + e, where µ is the population mean, QTL is the marker genetic effect, E is the environmental effects (here, common garden), QTLxE is the interaction between marker genetic and environmental effects, kinship corresponds to the background polygenic variation, and e is the error term.
-The genome scan was accomplished with the \'scan1\' function.
-The statistical significance of the genome scan was established by performing a stratified (i.e., stratifying on common garden) permutation test (n=1000) using \'scan1perm\' function.
-The estimated QTL effect was obtained using \'scan1coef\' function in R/qtl2.
+\end{document}

### 2024-10-16 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
revision2: minor edits to cover letter and SI Appendix

@@
 | Greedy Algorithm Iteration \# |     Gulf Greenup Date |     Midwest Greenup Date |    Both Greenup Date | Gulf Flowering Date     | Midwest Flowering Date         | Both Flowering Date     |
-|:----------|----------:|----------:|----------:|-----------|-----------|-----------|
+|:------------------------------|----------------------:|-------------------------:|---------------------:|-------------------------|--------------------------------|-------------------------|
 | 1                             |              Identity |                 Identity |             Identity | Identity                | Identity                       | Identity                |

### 2024-10-16 manuscript/revision2/cover_letter/cover_letter.qmd
revision2: minor edits to cover letter and SI Appendix

@@
 Yet current statistical models to test for GxE don't have robust frameworks to specify and compete different models of GxE effects, instead focusing on finding a single, best fitting form of GxE covariance.
-We use multivariate adaptive shrinkage (mash [@urbut2018]) to determine the regions of the switchgrass genome with different types of GxE, specifically regions of the genome with effects on phenology that covary with several weather-based 'cues'.
+We use multivariate adaptive shrinkage (mash, [Urbut *et al*., 2018](https://doi.org/10.1038/s41588-018-0268-8)) to determine the regions of the switchgrass genome with different types of GxE, specifically regions of the genome with effects on phenology that covary with several weather-based 'cues'.
 We use a greedy algorithm to compete and select types of GxE and GxWeather that improve our models, then use mash to identify which SNP effects are best modeled with each type of GxE or GxWeather.

### 2024-10-16 manuscript/revision2/cover_letter/cover_letter.tex
revision2: minor edits to cover letter and SI Appendix

@@
 on finding a single, best fitting form of GxE covariance. We use
-multivariate adaptive shrinkage (mash {[}@urbut2018{]}) to determine the
-regions of the switchgrass genome with different types of GxE,
-specifically regions of the genome with effects on phenology that covary
-with several weather-based `cues'. We use a greedy algorithm to compete
-and select types of GxE and GxWeather that improve our models, then use
-mash to identify which SNP effects are best modeled with each type of
-GxE or GxWeather.
+multivariate adaptive shrinkage (mash,
+\href{https://doi.org/10.1038/s41588-018-0268-8}{Urbut \emph{et al}.,
+2018}) to determine the regions of the switchgrass genome with different
+types of GxE, specifically regions of the genome with effects on
+phenology that covary with several weather-based `cues'. We use a greedy
+algorithm to compete and select types of GxE and GxWeather that improve
+our models, then use mash to identify which SNP effects are best modeled
+with each type of GxE or GxWeather.
 
